---
title: "DECODE-KIDNEY"
author: "Ye Ji Choi"
date: "`r lubridate::today()`"
format: 
  html:
    toc: true
    toc-depth: 5
    toc-location: left
    code-fold: true
    embed-resources: true
---

Ye Ji and Hailey present the groups characteristics (age, sex, HbA1c, BMI, SBP, DBP, eGFR, UACR), 

including medication use (SGLT2i, GLP1RA, MRA)

Stratify according to NWC, OC, T1D, T2D, and PKD

Indicate which which studies include kidney biopsies, 
and which we already have scRNAseq data (PB90+ATTEMPT), 
and others where we have biopsies but haven't performed done tissue interrogation (e.g., PANDA)

```{r echo = F}
library(dplyr)
library(arsenal)
library(ggplot2)
library(readxl)
library(purrr)
```

```{r echo = F}
# load "decode" df
load("/Users/choiyej/Library/CloudStorage/OneDrive-SharedLibraries-UW/Laura Pyle - Bjornstad/Biostatistics Core Shared Drive/DECODE/Data Clean/decode_harmonized.RData")

# List of studies:
# PANDA, RH, RH2, CASPER, CROC, 
# PENGUIN, ATTEMPT, TODAY/TOODAY2,
# TEEN-LABS, T1-DISCO, REMODEL-T1D, RPC2

decode_baseline <- decode %>%
  filter((study == "TEEN-LABS" & visit == "Month 1")|visit == "baseline") %>%
  select(record_id, visit, study, group, age, sex, hba1c, bmi, sbp, dbp, eGFR_CKD_epi, acr_u, sglti_timepoint, glp1_agonist_timepoint, mra_med) %>% # select just these vars because otherwise the next code raises too many issues
  dplyr::summarise(across(where(negate(is.numeric)), ~ ifelse(all(is.na(.x)), NA_character_, last(na.omit(.x)))),
                   across(where(is.numeric), ~ ifelse(all(is.na(.x)), NA_real_, mean(.x, na.rm = TRUE))),
                   .by = record_id)

# read in biopsy master spreadsheet (updated real time)
biopsy_master <- read_excel("/Users/choiyej/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Bjornstad Lab/Biopsy master tracker.xlsx", skip = 2) 
colnames(biopsy_master) <- gsub(" ", "_", colnames(biopsy_master))  # Replace spaces with underscores
colnames(biopsy_master) <- gsub("/", "_", colnames(biopsy_master))  # Replace slashes with underscores
colnames(biopsy_master) <- gsub("\\(", "_", colnames(biopsy_master))  # Replace opening parentheses with underscores
colnames(biopsy_master) <- gsub("\\)", "", colnames(biopsy_master))  # Remove closing parentheses
biopsy_master <- biopsy_master %>%
  rename(record_id=Study_ID,
         visit=Visit_ID) %>%
  mutate(visit = case_when(Study == "ATTEMPT" ~ visit,
                           Study == "REMODEL" ~ visit, 
                           Study == "IMPROVE" & visit == "4M" ~ "3_months_post_surgery",
                           visit == "12M" ~ "12_months_post_surgery",
                           visit == "Follow-up" ~ "follow-up",
                           T ~ "baseline"),
         record_id = case_when(startsWith(record_id, "RPC") & !is.na(Coenroll_ID__Same_visit_only) ~ Coenroll_ID__Same_visit_only,
                               T ~ record_id)) # RH2-60-T and RH2-48-T coenrolled into RPC2 as of 02/03/25 data pull

biopsy_master %>%
  select(record_id, Study, visit, Shipped_Y_N, scRNA_status) %>%
  filter(!is.na(record_id)) %>%
  filter(Shipped_Y_N == "Yes") %>%
  filter(scRNA_status != "No sample") %>%
  mutate(sequenced = case_when(scRNA_status == "Complete" ~ "Yes", T~ "No")) %>%
  summarise(n(), .by = c(Study, visit, sequenced)) %>%
  arrange(Study, visit, sequenced)

attempt_biopsy_ids <- read.csv("/Volumes/Peds Endo/Petter Bjornstad/scRNA/data_clean/ATTEMPT_record_ids.csv") 

attempt_biopsy_ids %>%
  summarise(n(), .by = c(visit))

decode_baseline$group <- factor(decode_baseline$group, levels = c("Lean Control",
                                                                     "Obese Control",
                                                                     "Type 1 Diabetes",
                                                                     "Type 2 Diabetes",
                                                                     "PKD"),
                                labels = c("NWC", "OBC", "T1D", "T2D", "PKD"))
```


```{r echo = F}
summary(arsenal::tableby(group ~ study + age + sex + hba1c + bmi + sbp + dbp + eGFR_CKD_epi + kwt(acr_u,"Nmiss", "medianq1q3", "range")  + sglti_timepoint + glp1_agonist_timepoint + mra_med, data = decode_baseline), test = F)
```

```{r echo = F}
# summary(arsenal::tableby(study ~ group + age + sex + hba1c + bmi + sbp + dbp + eGFR_CKD_epi + kwt(acr_u,"Nmiss", "medianq1q3", "range") + sglti_timepoint + glp1_agonist_timepoint, data = decode_baseline), test = F)
```

## visuals
```{r echo = F, warning=F}
## ggplots
decode_baseline %>%
  ggplot(aes(x = group, y = age, color = group)) + 
  geom_boxplot(size = 1.1) + 
  theme_bw() + 
  theme(panel.background = element_blank(),
        panel.grid = element_blank(),
        legend.position = "none",
        text = element_text(size = 18)) +
  scale_color_manual(values=c("NWC" = "#e76f51",
                              "OBC" = "#f4a261",
                              "T1D" = "#2a9d8f",
                              "T2D" = "#264653",
                              "PKD" = "#e9c46a")) +
  labs(x = NULL,
       y = "Age")
ggsave("/Users/choiyej/Library/CloudStorage/OneDrive-SharedLibraries-UW/Laura Pyle - Bjornstad/Biostatistics Core Shared Drive/DECODE/Figures/harmonized_age.jpeg", width = 7, height = 7)


decode_baseline %>%
  ggplot(aes(x = group, y = hba1c, color = group)) + 
  geom_boxplot(size = 1.1) + 
  theme_bw() + 
  theme(panel.background = element_blank(),
        panel.grid = element_blank(),
        legend.position = "none",
        text = element_text(size = 18)) +
  scale_color_manual(values=c("NWC" = "#e76f51",
                              "OBC" = "#f4a261",
                              "T1D" = "#2a9d8f",
                              "T2D" = "#264653",
                              "PKD" = "#e9c46a")) +
  labs(x = NULL,
       y = "HbA1c")
ggsave("/Users/choiyej/Library/CloudStorage/OneDrive-SharedLibraries-UW/Laura Pyle - Bjornstad/Biostatistics Core Shared Drive/DECODE/Figures/harmonized_hba1c.jpeg", width = 7, height = 7)


decode_baseline %>%
  ggplot(aes(x = group, y = bmi, color = group)) + 
  geom_boxplot(size = 1.1) + 
  theme_bw() + 
  theme(panel.background = element_blank(),
        panel.grid = element_blank(),
        legend.position = "none",
        text = element_text(size = 18)) +
  scale_color_manual(values=c("NWC" = "#e76f51",
                              "OBC" = "#f4a261",
                              "T1D" = "#2a9d8f",
                              "T2D" = "#264653",
                              "PKD" = "#e9c46a")) +
  labs(x = NULL,
       y = "BMI")
ggsave("/Users/choiyej/Library/CloudStorage/OneDrive-SharedLibraries-UW/Laura Pyle - Bjornstad/Biostatistics Core Shared Drive/DECODE/Figures/harmonized_bmi.jpeg", width = 7, height = 7)


decode_baseline %>%
  ggplot(aes(x = group, y = sbp, color = group)) + 
  geom_boxplot(size = 1.1) + 
  theme_bw() + 
  theme(panel.background = element_blank(),
        panel.grid = element_blank(),
        legend.position = "none",
        text = element_text(size = 18)) +
  scale_color_manual(values=c("NWC" = "#e76f51",
                              "OBC" = "#f4a261",
                              "T1D" = "#2a9d8f",
                              "T2D" = "#264653",
                              "PKD" = "#e9c46a")) +
  labs(x = NULL,
       y = "SBP")
ggsave("/Users/choiyej/Library/CloudStorage/OneDrive-SharedLibraries-UW/Laura Pyle - Bjornstad/Biostatistics Core Shared Drive/DECODE/Figures/harmonized_sbp.jpeg", width = 7, height = 7)


decode_baseline %>%
  ggplot(aes(x = group, y = dbp, color = group)) + 
  geom_boxplot(size = 1.1) + 
  theme_bw() + 
  theme(panel.background = element_blank(),
        panel.grid = element_blank(),
        legend.position = "none",
        text = element_text(size = 18)) +
  scale_color_manual(values=c("NWC" = "#e76f51",
                              "OBC" = "#f4a261",
                              "T1D" = "#2a9d8f",
                              "T2D" = "#264653",
                              "PKD" = "#e9c46a")) +
  labs(x = NULL,
       y = "DBP")
ggsave("/Users/choiyej/Library/CloudStorage/OneDrive-SharedLibraries-UW/Laura Pyle - Bjornstad/Biostatistics Core Shared Drive/DECODE/Figures/harmonized_dbp.jpeg", width = 7, height = 7)


decode_baseline %>%
  ggplot(aes(x = group, y = eGFR_CKD_epi, color = group)) + 
  geom_boxplot(size = 1.1) + 
  theme_bw() + 
  theme(panel.background = element_blank(),
        panel.grid = element_blank(),
        legend.position = "none",
        text = element_text(size = 18)) +
  scale_color_manual(values=c("NWC" = "#e76f51",
                              "OBC" = "#f4a261",
                              "T1D" = "#2a9d8f",
                              "T2D" = "#264653",
                              "PKD" = "#e9c46a")) +
  labs(x = NULL,
       y = "eGFR (CKD EPI)")
ggsave("/Users/choiyej/Library/CloudStorage/OneDrive-SharedLibraries-UW/Laura Pyle - Bjornstad/Biostatistics Core Shared Drive/DECODE/Figures/harmonized_egfr_ckdepi.jpeg", width = 7, height = 7)


decode_baseline %>%
  ggplot(aes(x = group, y = log(acr_u), color = group)) + 
  geom_boxplot(size = 1.1) + 
  theme_bw() + 
  theme(panel.background = element_blank(),
        panel.grid = element_blank(),
        legend.position = "none",
        text = element_text(size = 18)) +
  scale_color_manual(values=c("NWC" = "#e76f51",
                              "OBC" = "#f4a261",
                              "T1D" = "#2a9d8f",
                              "T2D" = "#264653",
                              "PKD" = "#e9c46a")) +
  labs(x = NULL,
       y = "log(UACR)")
ggsave("/Users/choiyej/Library/CloudStorage/OneDrive-SharedLibraries-UW/Laura Pyle - Bjornstad/Biostatistics Core Shared Drive/DECODE/Figures/harmonized_log_uacr.jpeg", width = 7, height = 7)
```