---
title: "TODAY Somalogic Pathway Analyses: Linear models for insulin sensitivity variables"
author: "Tim Vigers"
date: "`r format(Sys.time(), '%d %B %Y')`"
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    self-contained: true
    code-fold: true
    fig-width: 8
    fig-height: 10
---

```{r include=FALSE}
library(readxl)
library(clusterProfiler)
library(ReactomePA)
library(enrichplot)
library(EnhancedVolcano)
```

# Enrichment

Enrichment analyses were performed on nominally significant proteins using the R package `clusterProfiler` and a gene set ranked by log FC. Entrez gene IDs were used for all annotation. All genes in the `org.Hs.eg.db` were used as background for GO enrichment analysis. All duplicate gene/protein names were removed from the differentially expressed gene list.

```{r}
# Make a funciton for pathway analysis
# Need a genelist in the format used by clusterProfiler, with names in UniProt form
pathway_analysis = function(de_genes){
  # Remove duplicates and fix names
  names(de_genes) = sub("\\|.*","",names(de_genes))
  de_genes = de_genes[!is.na(names(de_genes))]
  de_genes = de_genes[which(!duplicated(names(de_genes)))]
  # GSEA
  gsea = gsePathway(geneList = de_genes[unique(names(de_genes))],pvalueCutoff = 1)
  gsea_upset = upsetplot(gsea) + ggtitle("GSEA Upset Plot")
  ridge = ridgeplot(gsea) + ggtitle("GSEA Ridge Plot")
  gseap = gseaplot2(gsea,geneSetID = which(gsea@result$p.adjust<=0.05),
                    pvalue_table = TRUE,ES_geom = "dot") + ggtitle("GSEA Plot")
  gsea_heat = heatplot(gsea,foldChange=de_genes) + ggtitle("GSEA Heatmap")
  # GO enrichment
  ego = enrichGO(names(de_genes),OrgDb = 'org.Hs.eg.db')
  bar = barplot(ego,showCategory=20) + ggtitle("GO Enrichment Bar Plot")
  dot = dotplot(ego,showCategory=20) + ggtitle("GO Enrichment Dot Plot")
  # Convert to gene symbols
  egox = setReadable(ego, 'org.Hs.eg.db', 'ENTREZID')
  # More plots!
  cnet = cnetplot(egox, foldChange=de_genes) + ggtitle("GO Enrichment Network Plot")
  circos = cnetplot(egox, foldChange=de_genes, circular = TRUE, colorEdge = TRUE) + 
    ggtitle("GO Enrichment Circos Plot")
  heat = heatplot(egox, foldChange=de_genes) + ggtitle("GO Enrichment Heatmap")
  egop = pairwise_termsim(ego)
  emap = emapplot(egop) + ggtitle("GO Enrichment Map")
  upset = upsetplot(egop) + ggtitle("GO Enrichment Upset Plot")
  # Return those beautiful plots
  return(list("gsea_upset" = gsea_upset,"ridge" = ridge,"gseap" = gseap,
              "gsea_heat" = gsea_heat,"bar" = bar,"dot" = dot,"cnet" = cnet,
              "circos" = circos,"heat" = heat,"emap" = emap,"upset" = upset))
}
# Function for comparing two gene lists
# de_gene_list is a list of multiple gene sets
pathway_comparison = function(de_gene_list){
  # Format names
  de_gene_list = lapply(de_gene_list, function(l){
    n = names(l)
    n = sub("\\|.*","",n)
    eg = bitr(n, fromType="UNIPROT", toType="ENTREZID", OrgDb="org.Hs.eg.db")
    n = eg$ENTREZID[match(n,eg$UNIPROT)]
    n = n[!is.na(n)]
    dups = n[duplicated(n)]
    n = n[!n %in% dups]
    return(n)
  })
  # Compare with GO enrichment
  ck = compareCluster(geneClusters = de_gene_list,
                      fun='enrichGO', OrgDb='org.Hs.eg.db')
  # Gene names
  ckx = setReadable(ck, OrgDb = "org.Hs.eg.db", keyType="ENTREZID")
  # Plots
  dot = dotplot(ckx) + ggtitle("Cluster Comparison Dot Plot")
  net = cnetplot(ckx) + ggtitle("Cluster Comparison Network Plot")
  xx = pairwise_termsim(ckx)
  emap = emapplot(xx) + ggtitle("Cluster Comparison Enrichment Map")
  # Return
  return(list("dot" = dot,"net" = net,"emap" = emap))
}
# Volcano plot
volcano = function(df,lab = "Target",xcol = "logFC",ycol = "P.Value",top = 15,
                   xlimit = c(-1, 1),ylimit = c(0, -log10(10e-7)),
                   xlabel = bquote(~Log ~ "fold change"),
                   pCutoff = 0.05){
  df = as.data.frame(df)
  df = df[order(df[,ycol]),]
  df$logp = -log10(df[,ycol])
  df$fc = df[,xcol]
  df$sig = df[,ycol] <= pCutoff
  p = ggplot(data = df,aes(x = fc,y = logp,color = sig)) + 
    geom_hline(yintercept = -log10(pCutoff),linetype = "dashed") +
    geom_point(size=2) + 
    geom_label_repel(data=df[1:top,],aes(label=Target),color = "black",
                     max.overlaps = 15,size=6) +
    scale_color_manual(values = c("grey","#3e6dbf")) +
    xlab(xlabel) + ylab(bquote(~-Log[10] ~ italic(P))) +
    theme_bw() +
    theme(legend.position = "none",
          axis.text=element_text(size=15),
          axis.title=element_text(size=20)) 
  return(p)
}
```

# Baseline (moderated adjusted linear models)

```{r}
#f = "/Volumes/Peds Endo/Petter Bjornstad/TODAY subaward/Results/Linear and Cox models/TODAY somalogic Cox models scaled baseline adjusted.xlsx"
f <- "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/TODAY subaward/Results/Linear and Cox models/TODAY somalogic limma baseline unadjusted insulin sensitivity.xlsx"
```

## Baseline BMI

```{r warning=FALSE,message=FALSE}
# changed "estimate" to "logFC" and "p.value" to "P.value" but not sure why that would be needed?
# might be best to change code block below back and then create new variables with original names

# Prepare gene list
bmi = read_excel(f,sheet = "BMI0_moderated_FDR")
volcano(bmi,xcol = "logFC",ycol = "P.Value",xlabel = "LogFC")
bmi = bmi[bmi$P.Value <= 0.05,]
de_genes_bmi = bmi$logFC
names(de_genes_bmi) = bmi$EntrezGeneID
de_genes_bmi = sort(de_genes_bmi, decreasing = TRUE)
# Pathway analysis
bmi_p = pathway_analysis(de_genes_bmi)
```

### Gene ontology (GO) enrichment analysis

```{r echo=FALSE,warning=FALSE,message=FALSE}
bmi_p$bar
bmi_p$dot
bmi_p$cnet
bmi_p$circos
bmi_p$heat
bmi_p$emap
bmi_p$upset
```

### Gene set enrichment analysis (GSEA) with KEGG

```{r echo=FALSE,warning=FALSE,message=FALSE}
bmi_p$gsea_upset
bmi_p$ridge
bmi_p$gseap
bmi_p$gsea_heat
```

## Baseline eIS

```{r warning=FALSE,message=FALSE}
# Prepare gene list
eis = read_excel(f,sheet = "si_1_ins00_moderated_FDR")
volcano(eis,xcol = "logFC",ycol = "P.Value",xlabel = "logFC")
eis = eis[eis$P.Value <= 0.05,]
de_genes_eis = eis$logFC
names(de_genes_eis) = eis$EntrezGeneID
de_genes_eis = sort(de_genes_eis, decreasing = TRUE)
# Pathway analysis
eis_p = pathway_analysis(de_genes_eis)
```

### Gene ontology (GO) enrichment analysis

```{r echo=FALSE,warning=FALSE,message=FALSE}
eis_p$bar
eis_p$dot
eis_p$cnet
eis_p$circos
eis_p$heat
eis_p$emap
eis_p$upset
```

### Gene set enrichment analysis (GSEA) with KEGG

```{r echo=FALSE,warning=FALSE,message=FALSE}
eis_p$gsea_upset
eis_p$ridge
eis_p$gseap
eis_p$gsea_heat
```

## Baseline C-peptide oDI

```{r warning=FALSE,message=FALSE}
# Prepare gene list
codi = read_excel(f,sheet = "codi0_moderated_FDR")
volcano(codi,xcol = "logFC",ycol = "P.Value",xlabel = "logFC")
codi = codi[codi$P.Value <= 0.05,]
de_genes_codi = codi$logFC
names(de_genes_codi) = codi$EntrezGeneID
de_genes_codi = sort(de_genes_codi, decreasing = TRUE)
# Pathway analysis
codi_p = pathway_analysis(de_genes_codi)
```

### Gene ontology (GO) enrichment analysis

```{r echo=FALSE,warning=FALSE,message=FALSE}
codi_p$bar
codi_p$dot
codi_p$cnet
codi_p$circos
codi_p$heat
codi_p$emap
codi_p$upset
```

### Gene set enrichment analysis (GSEA) with KEGG

```{r echo=FALSE,warning=FALSE,message=FALSE}
codi_p$gsea_upset
codi_p$ridge
# codi_p$gseap
codi_p$gsea_heat
```


change below


## Change in BMI

```{r warning=FALSE,message=FALSE}
# Prepare gene list
bmislope = read_excel(f,sheet = "slopebmi_moderated_FDR")
volcano(bmislope,xcol = "logFC",ycol = "P.Value",xlabel = "logFC")
bmislope = bmislope[bmislope$P.Value <= 0.05,]
de_genes_bmislope = bmislope$logFC
names(de_genes_bmislope) = bmislope$EntrezGeneID
de_genes_bmislope = sort(de_genes_bmislope, decreasing = TRUE)
# Pathway analysis
bmislope_p = pathway_analysis(de_genes_bmislope)
```

### Gene ontology (GO) enrichment analysis

```{r echo=FALSE,warning=FALSE,message=FALSE}
bmislope_p$bar
bmislope_p$dot
bmislope_p$cnet
bmislope_p$circos
bmislope_p$heat
bmislope_p$emap
bmislope_p$upset
```

### Gene set enrichment analysis (GSEA) with KEGG

```{r echo=FALSE,warning=FALSE,message=FALSE}
bmislope_p$gsea_upset
bmislope_p$ridge
# bmislope_p$gseap
bmislope_p$gsea_heat
```

## Change in eIS

```{r warning=FALSE,message=FALSE}
# Prepare gene list
eisslope = read_excel(f,sheet = "slopesi_1_ins0_moderated_FDR")
volcano(eisslope,xcol = "logFC",ycol = "P.Value",xlabel = "logFC")
eisslope = eisslope[eisslope$P.Value <= 0.05,]
de_genes_eisslope = eisslope$logFC
names(de_genes_eisslope) = eisslope$EntrezGeneID
de_genes_eisslope = sort(de_genes_eisslope, decreasing = TRUE)
# Pathway analysis
eisslope_p = pathway_analysis(de_genes_eisslope)
```

### Gene ontology (GO) enrichment analysis

```{r echo=FALSE,warning=FALSE,message=FALSE}
eisslope_p$bar
eisslope_p$dot
eisslope_p$cnet
eisslope_p$circos
eisslope_p$heat
eisslope_p$emap
eisslope_p$upset
```

### Gene set enrichment analysis (GSEA) with KEGG

```{r echo=FALSE,warning=FALSE,message=FALSE}
eisslope_p$gsea_upset
eisslope_p$ridge
#eisslope_p$gseap
eisslope_p$gsea_heat
```

## Change in C-peptide oDI

```{r warning=FALSE,message=FALSE}
# Prepare gene list
codislope = read_excel(f,sheet = "slopecodi_moderated_FDR")
volcano(codislope,xcol = "logFC",ycol = "P.Value",xlabel = "logFC")
codislope = codislope[codislope$P.Value <= 0.05,]
de_genes_codislope = codislope$logFC
names(de_genes_codislope) = codislope$EntrezGeneID
de_genes_codislope = sort(de_genes_codislope, decreasing = TRUE)
# Pathway analysis
codislope_p = pathway_analysis(de_genes_codislope)
```

### Gene ontology (GO) enrichment analysis

```{r echo=FALSE,warning=FALSE,message=FALSE}
codislope_p$bar
codislope_p$dot
codislope_p$cnet
codislope_p$circos
codislope_p$heat
codislope_p$emap
codislope_p$upset
```

### Gene set enrichment analysis (GSEA) with KEGG

```{r echo=FALSE,warning=FALSE,message=FALSE}
codislope_p$gsea_upset
codislope_p$ridge
# codislope_p$gseap
codislope_p$gsea_heat
```
