---
title: "TODAY Somalogic Pathway Analyses: Moderated Adjusted Linear Models"
author: "Tim Vigers"
date: "`r format(Sys.time(), '%d %B %Y')`"
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    self-contained: true
    code-fold: true
    fig-width: 8
    fig-height: 10
---

```{r include=FALSE}
library(readxl)
library(clusterProfiler)
library(ReactomePA)
library(enrichplot)
library(EnhancedVolcano)
```

# Enrichment

Enrichment analyses were performed on nominally significant proteins using the R package `clusterProfiler` and a gene set ranked by log FC. Entrez gene IDs were used for all annotation. All genes in the `org.Hs.eg.db` were used as background for GO enrichment analysis. All duplicate gene/protein names were removed from the differentially expressed gene list.

```{r}
# Make a function for pathway analysis
# Need a genelist in the format used by clusterProfiler, with names in UniProt form
pathway_analysis = function(de_genes){
  # Remove duplicates and fix names
  de_genes = de_genes[!is.na(names(de_genes))]
  names(de_genes) = sub("\\|.*","",names(de_genes))
  de_genes = de_genes[which(!duplicated(names(de_genes)))]
  # GSEA with KEGG
  gseaKEGG = gseKEGG(geneList = de_genes[unique(names(de_genes))],
                     keyType = 'uniprot',pvalueCutoff = 1)
  gsea_upset = upsetplot(gseaKEGG) + ggtitle("GSEA Upset Plot")
  ridge = ridgeplot(gseaKEGG) + ggtitle("GSEA Ridge Plot")
  gseap = gseaplot2(gseaKEGG,geneSetID = 1:nrow(gseaKEGG@result)) + 
    ggtitle("GSEA Plot")
  gsea_heat = heatplot(gseaKEGG,foldChange=de_genes) + ggtitle("GSEA Heatmap")
  # Convert to gene names (some won't match unfortunately)
  eg = bitr(names(de_genes), fromType="UNIPROT", toType="ENTREZID", OrgDb="org.Hs.eg.db")
  de_genes = de_genes[which(names(de_genes) %in% eg$UNIPROT)]
  names(de_genes) = eg$ENTREZID[match(names(de_genes),eg$UNIPROT)]
  # GO enrichment
  ego = enrichGO(names(de_genes),OrgDb = 'org.Hs.eg.db')
  bar = barplot(ego,showCategory=20) + ggtitle("GO Enrichment Bar Plot")
  dot = dotplot(ego,showCategory=20) + ggtitle("GO Enrichment Dot Plot")
  # Convert to gene symbols
  egox = setReadable(ego, 'org.Hs.eg.db', 'ENTREZID')
  # More plots!
  cnet = cnetplot(egox, foldChange=de_genes) + ggtitle("GO Enrichment Network Plot")
  circos = cnetplot(egox, foldChange=de_genes, circular = TRUE, colorEdge = TRUE) + 
    ggtitle("GO Enrichment Circos Plot")
  heat = heatplot(egox, foldChange=de_genes) + ggtitle("GO Enrichment Heatmap")
  egop = pairwise_termsim(ego)
  emap = emapplot(egop) + ggtitle("GO Enrichment Map")
  upset = upsetplot(egop) + ggtitle("GO Enrichment Upset Plot")
  # Return those beautiful plots
  return(list("gsea_upset" = gsea_upset,"ridge" = ridge,"gseap" = gseap,
              "gsea_heat" = gsea_heat,"bar" = bar,"dot" = dot,"cnet" = cnet,
              "circos" = circos,"heat" = heat,"emap" = emap,"upset" = upset))
}
# Function for comparing two gene lists
# de_gene_list is a list of multiple gene sets
pathway_comparison = function(de_gene_list){
  # Format names
  de_gene_list = lapply(de_gene_list, function(l){
    n = names(l)
    n = sub("\\|.*","",n)
    eg = bitr(n, fromType="UNIPROT", toType="ENTREZID", OrgDb="org.Hs.eg.db")
    n = eg$ENTREZID[match(n,eg$UNIPROT)]
    n = n[!is.na(n)]
    dups = n[duplicated(n)]
    n = n[!n %in% dups]
    return(n)
  })
  # Compare with GO enrichment
  ck = compareCluster(geneClusters = de_gene_list,
                      fun='enrichGO', OrgDb='org.Hs.eg.db')
  # Gene names
  ckx = setReadable(ck, OrgDb = "org.Hs.eg.db", keyType="ENTREZID")
  # Plots
  dot = dotplot(ckx) + ggtitle("Cluster Comparison Dot Plot")
  net = cnetplot(ckx) + ggtitle("Cluster Comparison Network Plot")
  xx = pairwise_termsim(ckx)
  emap = emapplot(xx) + ggtitle("Cluster Comparison Enrichment Map")
  # Return
  return(list("dot" = dot,"net" = net,"emap" = emap))
}
# Volcano plot
volcano = function(df,lab = "Target",xcol = "logFC",ycol = "P.Value",top = 10,
                           xlimit = c(-1, 1),ylimit = c(0, -log10(10e-7)),
                           pCutoff = 0.05){
  df = as.data.frame(df)
  df = df[order(df[,ycol]),]
  df$logp = -log10(df[,ycol])
  df$fc = df[,xcol]
  df$sig = df$P.Value <= pCutoff
  p = ggplot(data = df,aes(x = fc,y = logp,color = sig)) + 
    geom_hline(yintercept = -log10(pCutoff),linetype = "dashed") +
    geom_point() + 
    geom_label_repel(data=df[1:top,],aes(label=Target),color = "black") +
    scale_color_manual(values = c("grey","#00BFC4")) +
    xlab(bquote(~Log ~ "fold change")) + ylab(bquote(~-Log[10] ~ italic(P))) +
    theme_bw() +
    theme(legend.position = "none") 
  return(p)
}
```

# Baseline (moderated adjusted linear models)

```{r}
#f = "/Users/timvigers/Documents/Work/CHCO/Petter Bjornstad/TODAY subaward/Results/TODAY proteomics moderated adjusted linear models 2022-08-31.xlsx"
f = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/TODAY subaward/Results/Linear and Cox models/TODAY somalogic limma baseline adjusted.xlsx"
```

## MAC

```{r warning=FALSE,message=FALSE}
# Prepare gene list
mac = read_excel(f,sheet = "MAC_moderated_FDR")
volcano(mac)
mac = mac[mac$P.Value <= 0.05,]
de_genes_mac = mac$logFC
names(de_genes_mac) = mac$UniProt
de_genes_mac = sort(de_genes_mac, decreasing = TRUE)
# Pathway analysis
mac_p = pathway_analysis(de_genes_mac)
```

### Gene ontology (GO) enrichment analysis

```{r echo=FALSE,warning=FALSE,message=FALSE}
mac_p$bar
mac_p$dot
mac_p$cnet
mac_p$circos
mac_p$heat
mac_p$emap
mac_p$upset
```

### Gene set enrichment analysis (GSEA) with KEGG

```{r echo=FALSE,warning=FALSE,message=FALSE}
mac_p$gsea_upset
mac_p$ridge
mac_p$gseap
mac_p$gsea_heat
```

## MIC

```{r warning=FALSE,message=FALSE}
# Prepare gene list
mic = read_excel(f,sheet = "MIC_moderated_FDR")
volcano(mic)
mic = mic[mic$P.Value <= 0.05,]
de_genes_mic = mic$logFC
names(de_genes_mic) = mic$UniProt
de_genes_mic = sort(de_genes_mic, decreasing = TRUE)
# Pathway analysis
mic_p = pathway_analysis(de_genes_mic)
```

### Gene ontology (GO) enrichment analysis

```{r echo=FALSE,warning=FALSE,message=FALSE}
mic_p$bar
mic_p$dot
mic_p$cnet
mic_p$circos
mic_p$heat
mic_p$emap
mic_p$upset
```

### Gene set enrichment analysis (GSEA) with KEGG

```{r echo=FALSE,warning=FALSE,message=FALSE}
mic_p$gsea_upset
mic_p$ridge
mic_p$gseap
mac_p$gsea_heat
```

## MAC vs. MIC

```{r warning=FALSE,message=FALSE}
mac_vs_mic = list("MAC" = de_genes_mac,"MIC" = de_genes_mic)
mac_vs_mic_p = pathway_comparison(mac_vs_mic)
```

```{r echo=FALSE,warning=FALSE,message=FALSE}
mac_vs_mic_p$dot
mac_vs_mic_p$net
mac_vs_mic_p$emap
```

## Hyperfiltration

```{r warning=FALSE,message=FALSE}
# Prepare gene list
hyper = read_excel(f,sheet = "hyperfilt_moderated_FDR")
volcano(hyper)
hyper = hyper[hyper$P.Value <= 0.05,]
de_genes_hyper = hyper$logFC
names(de_genes_hyper) = hyper$UniProt
de_genes_hyper = sort(de_genes_hyper, decreasing = TRUE)
# Pathway analysis
hyper_p = pathway_analysis(de_genes_hyper)
```

### Gene ontology (GO) enrichment analysis

```{r echo=FALSE,warning=FALSE,message=FALSE}
hyper_p$bar
hyper_p$dot
hyper_p$cnet
hyper_p$circos
hyper_p$heat
hyper_p$emap
hyper_p$upset
```

### Gene set enrichment analysis (GSEA) with KEGG

```{r echo=FALSE,warning=FALSE,message=FALSE}
hyper_p$gsea_upset
hyper_p$ridge
hyper_p$gseap
hyper_p$gsea_heat
```

## Rapid eGFR decline

```{r warning=FALSE,message=FALSE}
# Prepare gene list
rapid = read_excel(f,sheet = "rapid_moderated_FDR")
volcano(rapid)
rapid = rapid[rapid$P.Value <= 0.05,]
de_genes_rapid = rapid$logFC
names(de_genes_rapid) = rapid$UniProt
de_genes_rapid = sort(de_genes_rapid, decreasing = TRUE)
# Pathway analysis
rapid_p = pathway_analysis(de_genes_rapid)
```

### Gene ontology (GO) enrichment analysis

```{r echo=FALSE,warning=FALSE,message=FALSE}
rapid_p$bar
rapid_p$dot
rapid_p$cnet
rapid_p$circos
rapid_p$heat
rapid_p$emap
rapid_p$upset
```

### Gene set enrichment analysis (GSEA) with KEGG

```{r echo=FALSE,warning=FALSE,message=FALSE}
rapid_p$gsea_upset
rapid_p$ridge
rapid_p$gseap
rapid_p$gsea_heat
```

## Hypertension

```{r warning=FALSE,message=FALSE}
# Prepare gene list
htn = read_excel(f,sheet = "htn_moderated_FDR")
volcano(htn)
htn = htn[htn$P.Value <= 0.05,]
de_genes_htn = htn$logFC
names(de_genes_htn) = htn$UniProt
de_genes_htn = sort(de_genes_htn, decreasing = TRUE)
# Pathway analysis
htn_p = pathway_analysis(de_genes_htn)
```

### Gene ontology (GO) enrichment analysis

```{r echo=FALSE,warning=FALSE,message=FALSE}
htn_p$bar
htn_p$dot
htn_p$cnet
htn_p$circos
htn_p$heat
htn_p$emap
htn_p$upset
```

### Gene set enrichment analysis (GSEA) with KEGG

```{r echo=FALSE,warning=FALSE,message=FALSE}
htn_p$gsea_upset
htn_p$ridge
htn_p$gseap
htn_p$gsea_heat
```

## Glycemic failure

```{r warning=FALSE,message=FALSE}
# Prepare gene list
glyc = read_excel(f,sheet = "glyc_moderated_FDR")
volcano(glyc)
glyc = glyc[glyc$P.Value <= 0.05,]
de_genes_glyc = glyc$logFC
names(de_genes_glyc) = glyc$UniProt
de_genes_glyc = sort(de_genes_glyc, decreasing = TRUE)
# Pathway analysis
glyc_p = pathway_analysis(de_genes_glyc)
```

### Gene ontology (GO) enrichment analysis

```{r echo=FALSE,warning=FALSE,message=FALSE}
glyc_p$bar
glyc_p$dot
glyc_p$cnet
glyc_p$circos
glyc_p$heat
glyc_p$emap
glyc_p$upset
```

### Gene set enrichment analysis (GSEA) with KEGG

```{r echo=FALSE,warning=FALSE,message=FALSE}
glyc_p$gsea_upset
glyc_p$ridge
glyc_p$gseap
glyc_p$gsea_heat
```

# 10-year follow-up (adjusted)

```{r}
#f = "/Users/timvigers/Documents/Work/CHCO/Petter Bjornstad/TODAY subaward/Results/TODAY proteomics moderated t-tests 10 year cross-sectional adjusted 2022-08-31.xlsx"
f = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/TODAY subaward/Results/Linear and Cox models/TODAY somalogic limma yr10 adjusted.xlsx"
```

## MAC

```{r warning=FALSE,message=FALSE}
# Prepare gene list
mac_10y = read_excel(f,sheet = "MAC_moderated_FDR")
volcano(mac_10y)
mac_10y = mac_10y[mac_10y$P.Value <= 0.05,]
de_genes_mac_10y = mac_10y$logFC
names(de_genes_mac_10y) = mac_10y$UniProt
de_genes_mac_10y = sort(de_genes_mac_10y, decreasing = TRUE)
# Pathway analysis
mac_10y_p = pathway_analysis(de_genes_mac_10y)
```

### Gene ontology (GO) enrichment analysis

```{r echo=FALSE,warning=FALSE,message=FALSE}
mac_10y_p$bar
mac_10y_p$dot
mac_10y_p$cnet
mac_10y_p$circos
mac_10y_p$heat
mac_10y_p$emap
mac_10y_p$upset
```

### Gene set enrichment analysis (GSEA) with KEGG

```{r echo=FALSE,warning=FALSE,message=FALSE}
mac_10y_p$gsea_upset
mac_10y_p$ridge
mac_10y_p$gseap
mac_10y_p$gsea_heat
```

## MIC

```{r warning=FALSE,message=FALSE}
# Prepare gene list
mic_10y = read_excel(f,sheet = "MIC_moderated_FDR")
volcano(mic_10y)
mic_10y = mic_10y[mic_10y$P.Value <= 0.05,]
de_genes_mic_10y = mic_10y$logFC
names(de_genes_mic_10y) = mic_10y$UniProt
de_genes_mic_10y = sort(de_genes_mic_10y, decreasing = TRUE)
# Pathway analysis
mic_10y_p = pathway_analysis(de_genes_mic_10y)
```

### Gene ontology (GO) enrichment analysis

```{r echo=FALSE,warning=FALSE,message=FALSE}
mic_10y_p$bar
mic_10y_p$dot
mic_10y_p$cnet
mic_10y_p$circos
mic_10y_p$heat
mic_10y_p$emap
mic_10y_p$upset
```

### Gene set enrichment analysis (GSEA) with KEGG

```{r echo=FALSE,warning=FALSE,message=FALSE}
mic_10y_p$gsea_upset
mic_10y_p$ridge
mic_10y_p$gseap
mac_p$gsea_heat
```

## MAC vs. MIC

```{r warning=FALSE,message=FALSE}
mac_vs_mic_10y = list("MAC" = de_genes_mac_10y,"MIC" = de_genes_mic_10y)
mac_vs_mic_p_10y = pathway_comparison(mac_vs_mic_10y)
```

```{r echo=FALSE,warning=FALSE,message=FALSE}
mac_vs_mic_p_10y$dot
mac_vs_mic_p_10y$net
mac_vs_mic_p_10y$emap
```

## Hyperfiltration

```{r warning=FALSE,message=FALSE}
# Prepare gene list
hyper_10y = read_excel(f,sheet = "hyperfilt_moderated_FDR")
volcano(hyper_10y)
hyper_10y = hyper_10y[hyper_10y$P.Value <= 0.05,]
de_genes_hyper_10y = hyper_10y$logFC
names(de_genes_hyper_10y) = hyper_10y$UniProt
de_genes_hyper_10y = sort(de_genes_hyper_10y, decreasing = TRUE)
# Pathway analysis
hyper_10y_p = pathway_analysis(de_genes_hyper_10y)
```

### Gene ontology (GO) enrichment analysis

```{r echo=FALSE,warning=FALSE,message=FALSE}
hyper_10y_p$bar
hyper_10y_p$dot
hyper_10y_p$cnet
hyper_10y_p$circos
hyper_10y_p$heat
hyper_10y_p$emap
hyper_10y_p$upset
```

### Gene set enrichment analysis (GSEA) with KEGG

```{r echo=FALSE,warning=FALSE,message=FALSE}
hyper_10y_p$gsea_upset
hyper_10y_p$ridge
hyper_10y_p$gseap
hyper_10y_p$gsea_heat
```

## Rapid eGFR decline

```{r warning=FALSE,message=FALSE}
# Prepare gene list
rapid_10y = read_excel(f,sheet = "rapid_moderated_FDR")
volcano(rapid_10y)
rapid_10y = rapid_10y[rapid_10y$P.Value <= 0.05,]
de_genes_rapid_10y = rapid_10y$logFC
names(de_genes_rapid_10y) = rapid_10y$UniProt
de_genes_rapid_10y = sort(de_genes_rapid_10y, decreasing = TRUE)
# Pathway analysis
rapid_10y_p = pathway_analysis(de_genes_rapid_10y)
```

### Gene ontology (GO) enrichment analysis

```{r echo=FALSE,warning=FALSE,message=FALSE}
rapid_10y_p$bar
rapid_10y_p$dot
rapid_10y_p$cnet
rapid_10y_p$circos
rapid_10y_p$heat
rapid_10y_p$emap
rapid_10y_p$upset
```

### Gene set enrichment analysis (GSEA) with KEGG

```{r echo=FALSE,warning=FALSE,message=FALSE}
rapid_10y_p$gsea_upset
rapid_10y_p$ridge
rapid_10y_p$gseap
rapid_10y_p$gsea_heat
```

## Hypertension

```{r warning=FALSE,message=FALSE}
# Prepare gene list
htn_10y = read_excel(f,sheet = "htn_moderated_FDR")
labs = htn_10y$Target[order(htn_10y$P.Value)[1:10]]
volcano(htn_10y)
htn_10y = htn_10y[htn_10y$P.Value <= 0.05,]
de_genes_htn_10y = htn_10y$logFC
names(de_genes_htn_10y) = htn_10y$UniProt
de_genes_htn_10y = sort(de_genes_htn_10y, decreasing = TRUE)
# Pathway analysis
htn_10y_p = pathway_analysis(de_genes_htn_10y)
```

### Gene ontology (GO) enrichment analysis

```{r echo=FALSE,warning=FALSE,message=FALSE}
htn_10y_p$bar
htn_10y_p$dot
htn_10y_p$cnet
htn_10y_p$circos
htn_10y_p$heat
htn_10y_p$emap
htn_10y_p$upset
```

### Gene set enrichment analysis (GSEA) with KEGG

```{r echo=FALSE,warning=FALSE,message=FALSE}
htn_10y_p$gsea_upset
htn_10y_p$ridge
htn_10y_p$gseap
htn_10y_p$gsea_heat
```

## Glycemic failure

```{r warning=FALSE,message=FALSE}
# Prepare gene list
glyc_10y = read_excel(f,sheet = "glyc_moderated_FDR")
volcano(glyc_10y)
glyc_10y = glyc_10y[glyc_10y$P.Value <= 0.05,]
de_genes_glyc_10y = glyc_10y$logFC
names(de_genes_glyc_10y) = glyc_10y$UniProt
de_genes_glyc_10y = sort(de_genes_glyc_10y, decreasing = TRUE)
# Pathway analysis
glyc_10y_p = pathway_analysis(de_genes_glyc_10y)
```

### Gene ontology (GO) enrichment analysis

```{r echo=FALSE,warning=FALSE,message=FALSE}
glyc_10y_p$bar
glyc_10y_p$dot
glyc_10y_p$cnet
glyc_10y_p$circos
glyc_10y_p$heat
glyc_10y_p$emap
glyc_10y_p$upset
```

### Gene set enrichment analysis (GSEA) with KEGG

```{r echo=FALSE,warning=FALSE,message=FALSE}
glyc_10y_p$gsea_upset
glyc_10y_p$ridge
glyc_10y_p$gseap
glyc_10y_p$gsea_heat
```

# Baseline vs. 10-year follow-up

## MAC

```{r warning=FALSE,message=FALSE}
mac_base_follow = list("Baseline" = de_genes_mac,"Follow-up" = de_genes_mac_10y)
mac_base_follow_p = pathway_comparison(mac_base_follow)
```

```{r echo=FALSE,warning=FALSE,message=FALSE}
mac_base_follow_p$dot
mac_base_follow_p$net
mac_base_follow_p$emap
```

## MIC

```{r warning=FALSE,message=FALSE}
mic_base_follow = list("Baseline" = de_genes_mic,"Follow-up" = de_genes_mic_10y)
mic_base_follow_p = pathway_comparison(mic_base_follow)
```

```{r echo=FALSE,warning=FALSE,message=FALSE}
mic_base_follow_p$dot
mic_base_follow_p$net
mic_base_follow_p$emap
```

## Hyperfiltration

```{r warning=FALSE,message=FALSE}
hyper_base_follow = list("Baseline" = de_genes_hyper,"Follow-up" = de_genes_hyper_10y)
hyper_base_follow_p = pathway_comparison(hyper_base_follow)
```

```{r echo=FALSE,warning=FALSE,message=FALSE}
hyper_base_follow_p$dot
hyper_base_follow_p$net
hyper_base_follow_p$emap
```

## Rapid eGFR decline

```{r warning=FALSE,message=FALSE}
rapid_base_follow = list("Baseline" = de_genes_rapid,"Follow-up" = de_genes_rapid_10y)
rapid_base_follow_p = pathway_comparison(rapid_base_follow)
```

```{r echo=FALSE,warning=FALSE,message=FALSE}
rapid_base_follow_p$dot
rapid_base_follow_p$net
rapid_base_follow_p$emap
```

## Hypertension

```{r warning=FALSE,message=FALSE}
htn_base_follow = list("Baseline" = de_genes_htn,"Follow-up" = de_genes_htn_10y)
htn_base_follow_p = pathway_comparison(htn_base_follow)
```

```{r echo=FALSE,warning=FALSE,message=FALSE}
htn_base_follow_p$dot
htn_base_follow_p$net
htn_base_follow_p$emap
```

## Glycemic failure

```{r warning=FALSE,message=FALSE}
glyc_base_follow = list("Baseline" = de_genes_glyc,"Follow-up" = de_genes_glyc_10y)
glyc_base_follow_p = pathway_comparison(glyc_base_follow)
```

```{r echo=FALSE,warning=FALSE,message=FALSE}
glyc_base_follow_p$dot
glyc_base_follow_p$net
glyc_base_follow_p$emap
```
