---
title: "EMERALD echo electro analysis"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include=FALSE}

library(knitr)

knitr::opts_chunk$set(echo = FALSE,cache = FALSE)
library(dplyr)
library(tableone)
library(reshape2)
library(car)
library(Hmisc)

# read in data
alldata <- read.csv("E:\\Nadeau\\Nadeau EMERALD\\Echo\\Michal\\Raw data\\EMERALG_Echo_Electro_final.csv")

# correct randomizations
alldata$randomization <- ifelse(alldata$subject_id %in% c("6377-02","6377-24"),1,alldata$randomization)

# drop unneeded variables
alldata$X <- NULL
alldata$first_name <- NULL
alldata$last_name <- NULL
alldata$screening_date <- NULL

# code randomization group
alldata$randgroup <- ifelse(alldata$randomization==1,"Metformin","Placebo")

# make sure everything is numeric
notecho <- c("subject_id","randomization","Visit","randgroup")
echovars <- colnames(alldata[,!colnames(alldata) %in% notecho])
alldata[,echovars] <- apply(alldata[,echovars],2,as.numeric)

# calculate deltas
wide <- reshape(alldata,idvar = "subject_id",timevar = "Visit",direction = "wide") 
wide$randomization.2 <- NULL
wide$randomization <- wide$randomization.1
wide$randomization.1 <- NULL
wide$randgroup.2 <- NULL
wide$randgroup <- wide$randgroup.1
wide$randgroup.1 <- NULL

for(i in 1:length(echovars)) {
  wide[,paste0(echovars[i],"_delta")] <-  wide[,paste0(echovars[i],".2")] - wide[,paste0(echovars[i],".1")]
}

# compare changes in the treatment groups
t1_vars <- colnames(wide[,grep("delta",colnames(wide))])
t1 <- CreateTableOne(data=wide,vars=t1_vars,strata = "randgroup")
t1 <- print(t1,printToggle = F,varLabels=TRUE,missing=T,showAllLevels = T,test = T,nonnorm=c("SSF_delta","SSF.1_delta","SSF.2_delta","SSF.3_delta",
                                                                                             "SSF.4_delta","SSF.5_delta","SSF.6_delta",
                                                                                             "DRF.1_delta","DRF.2_delta","DRF.3_delta",
                                                                                             "DRF.4_delta","DRF.5_delta","DRF.6_delta"))
t1 <- t1[,1:4]

p <- NULL
p_met <- NULL
for (i in 1:length(t1_vars)) {
  p <- round(t.test(wide[wide$randgroup=="Metformin",paste0(t1_vars[i])],mu=0)$p.value,4)
  p_met <- c(p_met,p)
}
p_met <- c("",p_met)
t1 <- cbind(t1,p_met)

p <- NULL
p_pla <- NULL
for (i in 1:length(t1_vars)) {
  p <- round(t.test(wide[wide$randgroup=="Placebo",paste0(t1_vars[i])],mu=0)$p.value,4)
  p_pla <- c(p_pla,p)
}
p_pla <- c("",p_pla)
t1 <- cbind(t1,p_pla)

# get baseline characteristics
emerald <- read.csv("E:\\Nadeau\\Nadeau EMERALD\\EMERALD_DATA_2017-08-24.csv",na.strings = c("","."))
keep <- emerald[,c("subject_id","age","gender","a1c","diabetes_duration","bmi_percentile","bmi_baseline_clamp","bmi_final_clamp",
                   "dexa_fat_perc","dexa_fat_perc_final","mg_kg_min_80_gluc_corrected","mg_kg_min_80_gluc_corrected_final",
                   "mg_kglean__min_80_gluc_corrected","mg_kglean__min_80_gluc_corrected_final")]
keep$bmi_delta <- keep$bmi_final_clamp - keep$bmi_baseline_clamp
keep$dexa_fat_perc_delta <- keep$dexa_fat_perc_final - keep$dexa_fat_perc
keep$mg_kg_min_80_gluc_corrected_delta <- keep$mg_kg_min_80_gluc_corrected_final - keep$mg_kg_min_80_gluc_corrected
keep$mg_kglean__min_80_gluc_corrected_delta <- keep$mg_kglean__min_80_gluc_corrected_final - keep$mg_kglean__min_80_gluc_corrected
wide <- merge(wide,keep,by="subject_id",all.x = T,all.y = F)

# table of demographic and clinical characteristics
wide$genderfactor <- as.factor(ifelse(wide$gender==1,"Female","Male"))
demovars <- c("age","genderfactor","a1c","diabetes_duration","bmi_percentile","bmi_baseline_clamp","bmi_final_clamp",
                   "dexa_fat_perc","dexa_fat_perc_final","mg_kg_min_80_gluc_corrected","mg_kg_min_80_gluc_corrected_final",
                   "mg_kglean__min_80_gluc_corrected","mg_kglean__min_80_gluc_corrected_final","bmi_delta","dexa_fat_perc_delta",
                    "mg_kg_min_80_gluc_corrected_delta","mg_kglean__min_80_gluc_corrected_delta")
demotable <- CreateTableOne(data=wide,vars=demovars,strata = "randgroup")
demotable <- print(demotable,printToggle = F,varLabels=TRUE,missing=T,showAllLevels = T,test = T)
demotable <- demotable[,1:4]

# table of baseline factors by sex
sexvars <- c("age","a1c","diabetes_duration","bmi_percentile","bmi_baseline_clamp","bmi_final_clamp",
                   "dexa_fat_perc","dexa_fat_perc_final","mg_kg_min_80_gluc_corrected","mg_kg_min_80_gluc_corrected_final",
                   "mg_kglean__min_80_gluc_corrected","mg_kglean__min_80_gluc_corrected_final","bmi_delta","dexa_fat_perc_delta",
                    "mg_kg_min_80_gluc_corrected_delta","mg_kglean__min_80_gluc_corrected_delta")
sextable <- CreateTableOne(data=wide,vars=sexvars,strata = "genderfactor")
sextable <- print(sextable,printToggle = F,varLabels=TRUE,missing=T,showAllLevels = T,test = T)
sextable <- sextable[,1:4]


 # for(i in 1:length(t1_vars)) {
 #     hist(wide[,noquote(t1_vars[i])],main=noquote(t1_vars[i]))
 # }



```

```{r model_adj function}
model_adj = function(outcome_name,df){
  # Convert to numeric
  #df[,outcome_name] = as.numeric(df[,outcome_name])
  # Fit model
  mod = lm(as.formula(paste0(outcome_name,"~randgroup+diabetes_duration+a1c")),data=df)
  # Print model results
  #print(kable(Anova(mod),digits = 3,caption = "Type III ANOVA"))
  print(kable(summary(mod)$coefficients,digits = 3,caption = "Fixed Effects"))
}
```

```{r corrmatrix}
# ++++++++++++++++++++++++++++
# flattenCorrMatrix
# ++++++++++++++++++++++++++++
# cormat : matrix of the correlation coefficients
# pmat : matrix of the correlation p-values
flattenCorrMatrix <- function(cormat, pmat) {
  ut <- upper.tri(cormat)
  data.frame(
    row = rownames(cormat)[row(cormat)[ut]],
    column = rownames(cormat)[col(cormat)[ut]],
    cor  =(cormat)[ut],
    p = pmat[ut]
  )
}

deltavars <- tidyselect::vars_select(names(wide), matches('delta'))
corrm <- rcorr(as.matrix(wide[, deltavars]),type = "spearman")
corrm <- flattenCorrMatrix(round(corrm$r,2), round(corrm$P,4))
corrm <- corrm[corrm$column %in% c("dexa_fat_perc_delta","bmi_delta","mg_kg_min_80_gluc_corrected_delta","mg_kglean__min_80_gluc_corrected_delta"),]
```

# Background

The goal of this analysis was to examine changes in echo measures during the EMERALD study, and to test whether those changes were associated with changes in insulin sensitivity, BMI, and fat mass.

# Methods

Groups were compared using t-tests for normally distributed variables and the Mann-Whitney tests for non-normally distributed variables, or the chi-square test for binary variables.  Linear models were used to compare groups while adjusting for diabetes duration and baseline A1c.  Spearman correlation coefficients were calculated between the changes in echo measures and the change in insulin sensitivity, BMI, and fat mass.

# Results

## Descriptive statistics

```{r, echo=FALSE}
kable(demotable,caption="Table 1a.  Descriptive statistics for demographic and clinical factors by treatment group.  Statistics presented are mean (SD) or median [25th %ile, 75th %ile].")
```

```{r, echo=FALSE}
kable(sextable,caption="Table 1b.  Descriptive statistics for baseline demographic and clinical factors by sex.  Statistics presented are mean (SD) or median [25th %ile, 75th %ile].")
```

## Changes in echo parameters

```{r, echo=FALSE}
kable(t1,caption="Table 2.  Changes in echo parameters.    Statistics presented are mean (SD) or median [25th %ile, 75th %ile].  The column labeled 'p' is the p-value comparing the change in the metformin group to the change in the placebo group.  The column labeled 'p_met' tests whether the change in the metformin group is different from 0, and and the column labeled 'p_pla' tests whether the change in the placebo group is different from 0.")
```

## Models comparing changes in echo measures between the groups, adjusted for diabetes duration and A1c

### How to interpret the model results

To interpret these models: the p-value in each row represents the test of whether each term is significant, so the p-value for "randgroupPlacebo" is the test of whether the difference in treatment groups is significant.  The column labeled "Estimate" is the mean difference for a one-unit change in the variable, so the estimate in the "randgroupPlacebo" row is the mean in the Placebo group (group 2) minus the mean in the Metformin group (group 1).

### AHA.1_delta

```{r results='asis'}
model_adj(outcome_name = "AHA.1_delta", df=wide)
```

### AHA.2_delta

```{r results='asis'}
model_adj(outcome_name = "AHA.2_delta", df=wide)
```

### AHA.3_delta

```{r results='asis'}
model_adj(outcome_name = "AHA.3_delta", df=wide)
```

### AHA.4_delta

```{r results='asis'}
model_adj(outcome_name = "AHA.4_delta", df=wide)
```

### AHA.5_delta

```{r results='asis'}
model_adj(outcome_name = "AHA.5_delta", df=wide)
```

### AHA.6_delta

```{r results='asis'}
model_adj(outcome_name = "AHA.6_delta", df=wide)
```

### Mdys_delta

```{r results='asis'}
model_adj(outcome_name = "Mdys_delta", df=wide)
```

### HR_delta

```{r results='asis'}
model_adj(outcome_name = "HR_delta", df=wide)
```

### GCS_delta

```{r results='asis'}
model_adj(outcome_name = "GCS_delta", df=wide)
```

### TTP.1_delta

```{r results='asis'}
model_adj(outcome_name = "TTP.1_delta", df=wide)
```

### TTP.2_delta

```{r results='asis'}
model_adj(outcome_name = "TTP.2_delta", df=wide)
```

### TTP.3_delta

```{r results='asis'}
model_adj(outcome_name = "TTP.3_delta", df=wide)
```

### TTP.4_delta

```{r results='asis'}
model_adj(outcome_name = "TTP.4_delta", df=wide)
```

### TTP.5_delta

```{r results='asis'}
model_adj(outcome_name = "TTP.5_delta", df=wide)
```

### TTP.6_delta

```{r results='asis'}
model_adj(outcome_name = "TTP.6_delta", df=wide)
```

### SSR.1_delta

```{r results='asis'}
model_adj(outcome_name = "SSR.1_delta", df=wide)
```

### SSR.2_delta

```{r results='asis'}
model_adj(outcome_name = "SSR.2_delta", df=wide)
```

### SSR.3_delta

```{r results='asis'}
model_adj(outcome_name = "SSR.3_delta", df=wide)
```

### SSR.4_delta

```{r results='asis'}
model_adj(outcome_name = "SSR.4_delta", df=wide)
```

### SSR.5_delta

```{r results='asis'}
model_adj(outcome_name = "SSR.5_delta", df=wide)
```

### SSR.6_delta

```{r results='asis'}
model_adj(outcome_name = "SSR.6_delta", df=wide)
```

### DSR.1_delta

```{r results='asis'}
model_adj(outcome_name = "DSR.1_delta", df=wide)
```

### DSR.2_delta

```{r results='asis'}
model_adj(outcome_name = "DSR.2_delta", df=wide)
```

### DSR.3_delta

```{r results='asis'}
model_adj(outcome_name = "DSR.3_delta", df=wide)
```

### DSR.4_delta

```{r results='asis'}
model_adj(outcome_name = "DSR.4_delta", df=wide)
```

### DSR.5_delta

```{r results='asis'}
model_adj(outcome_name = "DSR.5_delta", df=wide)
```

### DSR.6_delta

```{r results='asis'}
model_adj(outcome_name = "DSR.6_delta", df=wide)
```

### DRF_delta

```{r results='asis'}
model_adj(outcome_name = "DRF_delta", df=wide)
```

### SSF_delta

```{r results='asis'}
model_adj(outcome_name = "SSF_delta", df=wide)
```

### SSF.1_delta

```{r results='asis'}
model_adj(outcome_name = "SSF.1_delta", df=wide)
```

### SSF.2_delta

```{r results='asis'}
model_adj(outcome_name = "SSF.2_delta", df=wide)
```

### SSF.3_delta

```{r results='asis'}
model_adj(outcome_name = "SSF.3_delta", df=wide)
```

### SSF.4_delta

```{r results='asis'}
model_adj(outcome_name = "SSF.4_delta", df=wide)
```

### SSF.5_delta

```{r results='asis'}
model_adj(outcome_name = "SSF.5_delta", df=wide)
```

### SSF.6_delta

```{r results='asis'}
model_adj(outcome_name = "SSF.6_delta", df=wide)
```

### DRF.1_delta

```{r results='asis'}
model_adj(outcome_name = "DRF.1_delta", df=wide)
```

### DRF.2_delta

```{r results='asis'}
model_adj(outcome_name = "DRF.2_delta", df=wide)
```

### DRF.3_delta

```{r results='asis'}
model_adj(outcome_name = "DRF.3_delta", df=wide)
```

### DRF.4_delta

```{r results='asis'}
model_adj(outcome_name = "DRF.4_delta", df=wide)
```

### DRF.5_delta

```{r results='asis'}
model_adj(outcome_name = "DRF.5_delta", df=wide)
```

### DRF.6_delta

```{r results='asis'}
model_adj(outcome_name = "DRF.6_delta", df=wide)
```

### onset.1_delta

```{r results='asis'}
model_adj(outcome_name = "onset.1_delta", df=wide)
```

### onset.2_delta

```{r results='asis'}
model_adj(outcome_name = "onset.2_delta", df=wide)
```

### onset.3_delta

```{r results='asis'}
model_adj(outcome_name = "onset.3_delta", df=wide)
```

### onset.4_delta

```{r results='asis'}
model_adj(outcome_name = "onset.4_delta", df=wide)
```

### onset.5_delta

```{r results='asis'}
model_adj(outcome_name = "onset.5_delta", df=wide)
```

### onset.5.1_delta

```{r results='asis'}
model_adj(outcome_name = "onset.5.1_delta", df=wide)
```

## Spearman correlation coefficients between changes in echo measures and changes in BMI, Si, and DEXA % fat

```{r results='asis'}
kable(corrm,row.names = F)
```
