---
title: "SGLT2i manuscript - clinical characteristics"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 5
    toc_float: yes
---

```{r setup, include=FALSE}
library(tableone)
library(dplyr)
library(knitr)
library(questionr)
library(skimr)
library(readxl)
library(arsenal)
knitr::opts_chunk$set(echo = FALSE,warning = FALSE)
knitr::opts_knit$set(root.dir = "~/Dropbox/Work/CHCO/Petter Bjornstad/T2D scRNA SGLT2i")
```

```{r include=FALSE}
# read in CROCODILE
source("./Data raw/CROCODILE-SGLT2iDescriptiveSta_R_2021-12-02_1719.r")
croc <- data
data <- NULL

# read in IMPROVET2D
source("./Data raw/IMPROVET2D-SGLT2iManuscriptDesc_R_2021-12-02_1719.r")
improve <- data
data <- NULL

# read in RENALHEIR
source("./Data raw/RENALHEIR-SGLT2iManuscriptDesc_R_2021-12-02_1718.r")
heir <- data
data <- NULL

# read in Michigan data
# michigan <- read.csv("./Data raw/data_set_without_hpi.csv",na.strings = c("NA",""," "))
michigan <- read.csv("./Data raw/clinical_from_RDS.csv",na.strings = c("NA",""," "))

# correct one value
michigan$total_protein <- ifelse(michigan$ID=="CRC-09",7.5,michigan$total_protein)

# calculate pglo_abs
michigan$TotProtgdL <- michigan$total_protein
michigan$HCt <- michigan$hematocrit
michigan$MAP_clamp <- michigan$clamp_map
michigan$HCt_clamp <- michigan$HCt/100
michigan$GFRabsmLs <- michigan$gfr/60
michigan$ERPFabsmLs <- michigan$rpf/60
michigan$FFabs <- michigan$gfr/michigan$rpf
michigan$RBFabs <- michigan$rpf/(1-michigan$HCt_clamp)
michigan$RBFGomabs <- michigan$ERPFabsmLs/(1-michigan$HCt_clamp)
michigan$RVRAbs <- michigan$MAP_clamp/michigan$RBFabs
michigan$Kfg <- ifelse(michigan$Group %in% c("T1D","T2D"),0.1012,0.1733)
michigan$deltaPFabs <- michigan$GFRabsmLs/michigan$Kfg
michigan$Cm_abs <- (michigan$TotProtgdL/michigan$FFabs)*log(1/(1-michigan$FFabs))
michigan$Pg_abs <- 5*(michigan$Cm_abs-2)
michigan$Pglo_abs <- michigan$Pg_abs+michigan$deltaPFabs+10
michigan$Ra_abs <- ((michigan$MAP_clamp - michigan$Pglo_abs)/michigan$RBFGomabs)*1328
michigan$Re_abs <- (michigan$GFRabsmLs/(michigan$Kfg*(michigan$RBFGomabs-michigan$GFRabsmLs)))*1328

write.csv(michigan,"./Data clean/data_set_without_hpi_with_pgloabs.csv",na = "",row.names = F)

michigan$clamp_height_m <- michigan$clamp_height/100
michigan$clamp_bmi <- michigan$clamp_weight/(michigan$clamp_height_m^2)
michigan$screen_elevated_albuminuria <- as.factor(ifelse(michigan$screen_urine_mab>=30,1,
                                                         ifelse(is.na(michigan$screen_urine_mab),NA,0)))
michigan$clamp_elevated_albuminuria <- as.factor(ifelse(michigan$clamp_urine_mab_baseline>=30,1,
                                                        ifelse(is.na(michigan$clamp_urine_mab_baseline),NA,0)))
michigan$sex <- as.factor(michigan$sex)
michigan$Glomerular.Nuclear.Count <- NULL
michigan$Fractional.Interstitial.Area <- NULL

# merge in morphometrics
morph <- read.csv("./Data raw/PB Morpho 2-2022.csv")
morph$KL.ID <- morph$Study.ID
morph <- morph %>% select(-c("Study.ID","Group","ID","Priority.24.."))
morph$KL.ID <- ifelse(morph$KL.ID=="KL-0010097","KL-0019097",morph$KL.ID)
michigan <- merge(michigan, morph, by="KL.ID", all.x = T, all.y = F)

# use Tim's code to calculate eGFR by FAS
# source("C:/Users/pylell/Documents/GitHub/shared-resources/Data Cleaning/Calculated Variables/eGFR.r")
source("~/GitHub/shared-resources/Data Cleaning/Calculated Variables/eGFR.R")
x <- egfr_calc(age=michigan$Age,serum_creatinine = michigan$serum_creatinine, cystatin_c = michigan$cystatin_c,
               bun=michigan$labs_bun, height = michigan$clamp_height_m*100, sex = michigan$sex,
               female="0", male="1")
michigan$eGFR_fas_cr <- x$eGFR_fas_cr
michigan$eGFR_fas_cr_cysc <- x$eGFR_fas_cr_cysc

# exclude T1D for this paper
t1data <- michigan[!michigan$Group=="T1D",]
t1data$three_groups = paste(t1data$Group,t1data$SGLT2i,sep = "_")
t1data$three_groups = factor(t1data$three_groups,
                             levels = c("HC_NO","T2D_NO","T2D_YES"),
                             labels = c("Healthy Controls","T2D No SGLT2i","T2D SGLT2i"))
# Get MRNs for Petter
croc$croc_id = paste0("CRC-",stringr::str_pad(croc$record_id,2,"left","0"))
heir$croc_id = heir$subject_id
heir$mrn = heir$mr_number
improve$croc_id = improve$subject_id
improve$mrn = improve$mr_number
mrns = do.call(rbind,list(croc[,c("croc_id","mrn")],heir[,c("croc_id","mrn")],
                          improve[,c("croc_id","mrn")]))
colnames(mrns) = c("ID","MRN")
mrns = mrns[mrns$ID %in% t1data$ID,]
mrns = mrns[complete.cases(mrns),]
# Add ACE inhibitor and Angiotensin Receptor Blocker info
croc$ACEi = croc$htn_med___1.factor
levels(croc$ACEi) = c("No","Yes")
croc$ARB = croc$htn_med___2.factor
levels(croc$ARB) = c("No","Yes")
heir$ACEi = heir$htn_med_type___1.factor
levels(heir$ACEi) = c("No","Yes")
heir$ARB = heir$htn_med_type___2.factor
levels(heir$ARB) = c("No","Yes")
improve$ACEi = improve$htn_med_type___1.factor
levels(improve$ACEi) = c("No","Yes")
improve$ARB = improve$htn_med_type___2.factor
levels(improve$ARB) = c("No","Yes")
ace_arb = do.call(rbind,list(croc[,c("croc_id","ACEi","ARB")],heir[,c("croc_id","ACEi","ARB")],
                             improve[,c("croc_id","ACEi","ARB")]))
ace_arb = ace_arb[complete.cases(ace_arb),]
ace_arb = ace_arb[!duplicated(ace_arb$croc_id),]
t1data = left_join(t1data,ace_arb,by = c("ID" = "croc_id"))
# Electrolytes
electro = read.csv("./Data clean/SGLT2i_electrolytes.csv")
electro$MRN = NULL
t1data = left_join(t1data,electro,by = "ID")
# SGLT2i details
meds = read.csv("./Data clean/IT2D_RH_medications.csv")
meds$Dose = factor(meds$Dose)
t1data = left_join(t1data,meds)
# Tables
vars <- c("Age","sex","clamp_bmi","dxa_body_fat","cholesterol","ldl","hdl",
          "triglycerides","clamp_sbp","clamp_dbp","clamp_map","clamp_acr_baseline",
          "screen_elevated_albuminuria","clamp_elevated_albuminuria","hba1c",
          "gfr","gfr_bsa","eGFR_fas_cr",
          "serum_creatinine","GIR","labs_bun","Metformin","ACEi","ARB",
          "Insulin","TZD","GLPYES",
          "Na","K","Cl","HCO3","BUN",
          "Fractional.Interstitial.Area",
          "Glomerular.Tuft.Area..um.2.",
          "Glomerular.Volume...Weibel..10.6.um.3.",
          "Glomerular.Volume...Wiggins..10.6.um.3.",
          "Glomerular.Volume...Consensus..10.6.um.3.",
          "Mesangial.Matrix.Area..um.2.",
          "Mesangial.Index....",
          "Mesangial.Volume...Weibel..10.6.um.3.",
          "Mesangial.Volume...Wiggins..10.6.um.3.",
          "Mesangial.Volume...Consensus..10.6.um.3.",
          "Mesangial.Nuclear.Count",
          "Glomerular.Nuclear.Count",
          "SGLT2i_duration_in_all",
          "SGLT2i_duration_in_exposed",
          "Type","Dose")
nonnorm <- c("Na","K","Cl","HCO3","BUN",
             "Glomerular.Tuft.Area..um.2.",
             "Glomerular.Volume...Weibel..10.6.um.3.",
             "Glomerular.Volume...Wiggins..10.6.um.3.",
             "Glomerular.Volume...Consensus..10.6.um.3.",
             "Mesangial.Matrix.Area..um.2.",
             "Mesangial.Volume...Weibel..10.6.um.3.",
             "Mesangial.Volume...Wiggins..10.6.um.3.",
             "Mesangial.Volume...Consensus..10.6.um.3.",
             "Mesangial.Nuclear.Count",
             "Glomerular.Nuclear.Count",
             "SGLT2i_duration_in_all",
             "SGLT2i_duration_in_exposed","clamp_acr_baseline")
exact <- c("screen_elevated_albuminuria","clamp_elevated_albuminuria","Metformin","Insulin","TZD","GLPYES","ACEi","ARB","Type","Dose")
t1 <- CreateTableOne(vars = vars, data = t1data, strata = "Group", test = TRUE)
t1 <- print(t1,nonnormal = nonnorm, exact=exact)
freq.na(t1data[,vars])

# exclude controls now
t1data_nocontrol <- t1data[!t1data$three_groups == "Healthy Controls",]
t1data_nocontrol$three_groups = droplevels(t1data_nocontrol$three_groups)
t2 <- CreateTableOne(vars = vars, data = t1data_nocontrol, strata = "three_groups", test = TRUE)
t2 <- print(t2,nonnormal = nonnorm, exact=exact)
```

# Data summary

```{r echo=FALSE}
skim(t1data[,vars])
```

# Table 1a - healthy controls vs T2D (T2D SGLT2i + T2D no SGLT2i).  

P-values are from Fisher's exact test (if test="exact") or chi-square test for categorical variables,Mann-Whitney test (if test="nonnorm") or t-tests for continuous variables.

```{r echo=FALSE}
kable(t1)
```

# Table 1b - T2D SGLT2i vs. T2D no SGLT2i.  

P-values are from Fisher's exact test (if test="exact") or chi-square test for categorical variables, Mann-Whitney test (if test="nonnorm") or t-tests for continuous variables.

```{r echo=FALSE}
kable(t2)
```

# Tim's tables

## eGFR, etc.

### All groups

```{r results='asis'}
t1 <- tableby(three_groups ~ Age + sex + clamp_bmi + dxa_body_fat + cholesterol +
                ldl + hdl + triglycerides + clamp_sbp + clamp_dbp +
                kwt(clamp_acr_baseline,"Nmiss","median","q1q3") +
                fe(screen_elevated_albuminuria) + fe(clamp_elevated_albuminuria) +
                hba1c + serum_creatinine + eGFR_fas_cr + SGLT2i_duration_in_exposed +
                fe(GLPYES) + fe(Metformin) + fe(TZD) + fe(Insulin) + fe(ACEi) + fe(ARB) +
                fe(Type) + fe(Dose) +
                kwt(Na,"Nmiss","median","q1q3") + 
                kwt(K,"Nmiss","median","q1q3") + 
                kwt(Cl,"Nmiss","median","q1q3") + 
                kwt(HCO3,"Nmiss","median","q1q3") + 
                kwt(BUN,"Nmiss","median","q1q3") +
                Fractional.Interstitial.Area +
                kwt(Glomerular.Tuft.Area..um.2.,"Nmiss","median","q1q3") +
                kwt(Glomerular.Volume...Weibel..10.6.um.3.,"Nmiss","median","q1q3") +
                kwt(Glomerular.Volume...Wiggins..10.6.um.3.,"Nmiss","median","q1q3") +
                kwt(Glomerular.Volume...Consensus..10.6.um.3.,"Nmiss","median","q1q3") +
                kwt(Mesangial.Matrix.Area..um.2.,"Nmiss","median","q1q3") +
                kwt(Mesangial.Index....,"Nmiss","median","q1q3") +
                kwt(Mesangial.Volume...Weibel..10.6.um.3.,"Nmiss","median","q1q3") +
                kwt(Mesangial.Volume...Wiggins..10.6.um.3.,"Nmiss","median","q1q3") +
                kwt(Mesangial.Volume...Consensus..10.6.um.3.,"Nmiss","median","q1q3") +
                kwt(Mesangial.Nuclear.Count,"Nmiss","median","q1q3") +
                kwt(Glomerular.Nuclear.Count,"Nmiss","median","q1q3"),
              data = t1data, control = list(digits = 2)
)
summary(t1,pfootnote = T)
```

### T2D only

```{r results='asis'}
t1 <- tableby(three_groups ~ Age + sex + clamp_bmi + dxa_body_fat + cholesterol +
                ldl + hdl + triglycerides + clamp_sbp + clamp_dbp +
                kwt(clamp_acr_baseline,"Nmiss","median","q1q3") +
                fe(screen_elevated_albuminuria) + fe(clamp_elevated_albuminuria) +
                hba1c + serum_creatinine + eGFR_fas_cr + SGLT2i_duration_in_exposed +
                fe(GLPYES) + fe(Metformin) + fe(TZD) + fe(Insulin) + fe(ACEi) + fe(ARB) +
                fe(Type) + fe(Dose) +
                kwt(Na,"Nmiss","median","q1q3") + 
                kwt(K,"Nmiss","median","q1q3") + 
                kwt(Cl,"Nmiss","median","q1q3") + 
                kwt(HCO3,"Nmiss","median","q1q3") + 
                kwt(BUN,"Nmiss","median","q1q3") +
                Fractional.Interstitial.Area +
                kwt(Glomerular.Tuft.Area..um.2.,"Nmiss","median","q1q3") +
                kwt(Glomerular.Volume...Weibel..10.6.um.3.,"Nmiss","median","q1q3") +
                kwt(Glomerular.Volume...Wiggins..10.6.um.3.,"Nmiss","median","q1q3") +
                kwt(Glomerular.Volume...Consensus..10.6.um.3.,"Nmiss","median","q1q3") +
                kwt(Mesangial.Matrix.Area..um.2.,"Nmiss","median","q1q3") +
                kwt(Mesangial.Index....,"Nmiss","median","q1q3") +
                kwt(Mesangial.Volume...Weibel..10.6.um.3.,"Nmiss","median","q1q3") +
                kwt(Mesangial.Volume...Wiggins..10.6.um.3.,"Nmiss","median","q1q3") +
                kwt(Mesangial.Volume...Consensus..10.6.um.3.,"Nmiss","median","q1q3") +
                kwt(Mesangial.Nuclear.Count,"Nmiss","median","q1q3") +
                kwt(Glomerular.Nuclear.Count,"Nmiss","median","q1q3"),
              data = t1data_nocontrol, control = list(digits = 2)
)
summary(t1,pfootnote = T)
```

### T2D vs. HC

```{r results='asis'}
t1 <- tableby(Group ~ Age + sex + clamp_bmi + dxa_body_fat + cholesterol +
                ldl + hdl + triglycerides + clamp_sbp + clamp_dbp + 
                kwt(clamp_acr_baseline,"Nmiss","median","q1q3") +
                fe(screen_elevated_albuminuria) + fe(clamp_elevated_albuminuria) +
                hba1c + serum_creatinine + eGFR_fas_cr + SGLT2i_duration_in_exposed +
                fe(GLPYES) + fe(Metformin) + fe(TZD) + fe(Insulin) + fe(ACEi) + fe(ARB) +
                fe(Type) + fe(Dose) +
                kwt(Na,"Nmiss","median","q1q3") + 
                kwt(K,"Nmiss","median","q1q3") + 
                kwt(Cl,"Nmiss","median","q1q3") + 
                kwt(HCO3,"Nmiss","median","q1q3") + 
                kwt(BUN,"Nmiss","median","q1q3") +
                Fractional.Interstitial.Area +
                kwt(Glomerular.Tuft.Area..um.2.,"Nmiss","median","q1q3") +
                kwt(Glomerular.Volume...Weibel..10.6.um.3.,"Nmiss","median","q1q3") +
                kwt(Glomerular.Volume...Wiggins..10.6.um.3.,"Nmiss","median","q1q3") +
                kwt(Glomerular.Volume...Consensus..10.6.um.3.,"Nmiss","median","q1q3") +
                kwt(Mesangial.Matrix.Area..um.2.,"Nmiss","median","q1q3") +
                kwt(Mesangial.Index....,"Nmiss","median","q1q3") +
                kwt(Mesangial.Volume...Weibel..10.6.um.3.,"Nmiss","median","q1q3") +
                kwt(Mesangial.Volume...Wiggins..10.6.um.3.,"Nmiss","median","q1q3") +
                kwt(Mesangial.Volume...Consensus..10.6.um.3.,"Nmiss","median","q1q3") +
                kwt(Mesangial.Nuclear.Count,"Nmiss","median","q1q3") +
                kwt(Glomerular.Nuclear.Count,"Nmiss","median","q1q3"),
              data = t1data, control = list(digits = 2)
)
summary(t1,pfootnote = T)
```
