---
title: "PCOS and Age Preliminary Analysis Results"
author: "Callie Rountree-Jablin"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include=FALSE}
library(tidyverse)
library(nlme)
library(lme4)
library(emmeans)
library(kableExtra)
library(gtsummary)
pacman::p_load(naniar)
library(fields)
library(corrplot)
options(scipen=999)

home_dir = "B:/Melanie Green/PCOS and age"
knitr::opts_knit$set(echo=FALSE, root.dir = home_dir)
```

```{r read, include=FALSE}
raw_dat <- readxl::read_xlsx("./Raw Data/PCOS_Baseline_8.14.2024.xlsx")
```

```{r data cleaning, include=FALSE}
## clean up raw data, create dfs for each assay 
clean_dat <- raw_dat %>% rename(pid = 'young: age <16; old: age >=16',
                                age_group = 'Code 1=Young, 2=Old') %>% 
  mutate(age_group = factor(age_group, levels = c(1, 2), labels = c('young', 'old'))) %>% 
  mutate(study = case_when(str_detect(pid, 'TEAL') ~ 'TEAL',
                           str_detect(pid, '18-0803') ~ 'ORANGE',
                           str_detect(pid, '6530') ~ 'APPLE',
                           TRUE ~ 'PLUM'))

glucose <- clean_dat %>% select(pid, study, age_group, glucose_0, glucose_10, glucose_20, glucose_30, glucose_45, glucose_60,
                                     glucose_90, glucose_120, glucose_150, glucose_180, glucose_210, glucose_240)
insulin <- clean_dat %>% select(pid, study, age_group, insulin_0, insulin_10, insulin_20, insulin_30, insulin_45, insulin_60,
                                     insulin_90, insulin_120, insulin_150, insulin_180, insulin_210, insulin_240)
cpep <- clean_dat %>% select(pid, study, age_group, cpeptide_0, cpeptide_10, cpeptide_20, cpeptide_30)
glp1 <- clean_dat %>%  select(pid, study, age_group, glp1_0, glp1_10, glp1_30, glp1_60) %>% mutate(glp1_0 = as.numeric(glp1_0))
glucagon <- clean_dat %>% select(pid, study, age_group, glucagon_0, glucagon_10, glucagon_30, glucagon_60, glucagon_90, glucagon_240)
ffa <- clean_dat %>% select(pid, study, age_group, ffa_0, ffa_30, ffa_60, ffa_120, ffa_180, ffa_210, ffa_240)

## switch to long format
# define function for converting to long format
pivot_to_long <- function(df, assay_name){
  df %>% pivot_longer(cols = starts_with(assay_name),
                      names_to = c(".value", "time"),
                      names_sep = "_",
                      values_to = assay_name) %>% 
    mutate(time = as.factor(as.numeric(time)))
}

# convert dfs to long
glucose_long <- pivot_to_long(glucose, "glucose")
insulin_long <- pivot_to_long(insulin, "insulin")
cpep_long <- pivot_to_long(cpep, "cpeptide")
glp1_long <- pivot_to_long(glp1, "glp1")
glucagon_long <- pivot_to_long(glucagon, "glucagon")
ffa_long <- pivot_to_long(ffa, "ffa")
```

# Exploratory Data Analysis

## Spaghetti Plots

```{r spaghetti plot function, echo=FALSE, message=FALSE, warning=FALSE}
## function for spaghetti plots with mean line and coloring by study
spaghetti_plot <- function(df, assay_name){
  df %>% 
    ggplot(aes_string(x = 'time', y = assay_name, group = 'pid', color = 'study')) +
    geom_line(alpha = 0.4) +
    stat_summary(fun = mean, geom = 'line', aes(group = 1), color = 'black') +
    stat_summary(fun = mean, geom = 'point', aes(group = 1), color = 'black') +
    facet_wrap( ~ age_group) +
    scale_color_manual(values = c('red3','orange','green4','royalblue')) +
    theme_bw()
}
```

### Glucose
```{r glucose spaghetti plot, echo=FALSE, message=FALSE, warning=FALSE}
spaghetti_plot(glucose_long, 'glucose')
```

### Insulin
```{r insulin spaghetti plot, echo=FALSE, message=FALSE, warning=FALSE}
spaghetti_plot(insulin_long, 'insulin')
```

### C-Peptide
```{r cpep spaghetti plot, echo=FALSE, message=FALSE, warning=FALSE}
spaghetti_plot(cpep_long, 'cpeptide')
```

### GLP-1
```{r glp1 spaghetti plot, echo=FALSE, message=FALSE, warning=FALSE}
spaghetti_plot(glp1_long, 'glp1')
```

### Glucagon
```{r glucagon spaghetti plot, echo=FALSE, message=FALSE, warning=FALSE}
spaghetti_plot(glucagon_long, 'glucagon')
```

### FFA
```{r ffa spaghetti plot, echo=FALSE, message=FALSE, warning=FALSE}
spaghetti_plot(ffa_long, 'ffa')
```


```{r outlier search based on EDA, eval=FALSE, include=FALSE}
## find outliers

# glp1 subject with notably large baseline value
glp1 %>% filter(glp1_0 > 50)

# glucagon old subject with slightly large values across the board
glucagon %>% filter(glucagon_0 > 150)

# glucagon young subject with large values
glucagon %>%  filter(glucagon_90 > 150)

# ffa with low baseline starting point - are entries for this subject shifted?
ffa %>% filter(ffa_0 < 300) 
```

## Missingness Summary Plots

### Glucose
```{r glucose missingness exploration, echo=FALSE, warning=FALSE}
gg_miss_var(glucose[,-c(1,2)], facet = age_group)
gg_miss_upset(glucose, nset = 10)
```

### Insulin
```{r insulin missingness exploration, echo=FALSE, warning=FALSE}
gg_miss_var(insulin[,-c(1,2)], facet = age_group)
gg_miss_upset(insulin, nset = 10)
```

### C-Peptide
```{r cpep missingness exploration, echo=FALSE, warning=FALSE}
gg_miss_var(cpep[,-c(1,2)], facet = age_group)
gg_miss_upset(cpep, nset = 10)
```

### GLP-1
```{r glp1 missingness exploration, echo=FALSE, warning=FALSE}
gg_miss_var(glp1[,-c(1,2)], facet = age_group)
gg_miss_upset(glp1, nset = 10)
```

### Glucagon
```{r glucagon missingness exploration, echo=FALSE, warning=FALSE}
gg_miss_var(glucagon[,-c(1,2)], facet = age_group)
gg_miss_upset(glucagon, nset = 10)
```

### FFA
```{r ffa missingness exploration, echo=FALSE, warning=FALSE}
# ffa at 210 not collected for PLUM or APPLE
gg_miss_var(ffa[,-c(1,2)], facet = age_group)
gg_miss_upset(ffa, nset = 10)
```


```{r missingness summary, include=FALSE}
# summarize missing vars quantitatively
miss_var_summary(glucose[-c(1,2),])
miss_var_summary(insulin[-c(1,2),])
miss_var_summary(cpep[-c(1,2),])
miss_var_summary(glp1[-c(1,2),])
miss_var_summary(glucagon[-c(1,2),])
miss_var_summary(ffa[-c(1,2),])

# look at specific subjects with missing values 
summarize_missingness <- function(df){
  df %>% 
    group_by(pid) %>% 
    miss_var_summary() %>% 
    filter(pct_miss > 0) %>% 
    arrange(desc(pct_miss))
}

summarize_missingness(glucose)
summarize_missingness(insulin)
summarize_missingness(cpep)
summarize_missingness(glp1)
summarize_missingness(glucagon)
summarize_missingness(ffa)
```


<!-- ## Correlation -->

```{r correlation function, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
## define function for visualizing correlation using heat map and mixed plot
correlation_plots <- function(df){
  cor_matrix <- df %>% 
    select(-c(pid, study, age_group)) %>% 
    cor(use = 'pairwise.complete.obs')
  
  fields::image.plot(cor_matrix, main = 'Correlation Heatmap')
  corrplot::corrplot.mixed(cor_matrix, main = 'Mixed Correlation Plot')
}
```

<!-- ### Glucose -->
```{r glucose correlation, eval=FALSE, include=FALSE}
correlation_plots(glucose)
```

<!-- ### Insulin -->
```{r insulin correlation, eval=FALSE, include=FALSE}
correlation_plots(insulin)
```

<!-- ### C-Peptide -->
```{r cpep correlation, eval=FALSE, include=FALSE}
correlation_plots(cpep)
```

<!-- ### GLP-1 -->
```{r glp1 correlation, eval=FALSE, include=FALSE}
correlation_plots(glp1)
```

<!-- ### Glucagon -->
```{r glucagon correlation, eval=FALSE, include=FALSE}
correlation_plots(glucagon)
```

<!-- ### FFA -->
```{r ffa correlation, eval=FALSE, include=FALSE}
correlation_plots(ffa)
```


# Methods

Mixed effects models were used to evaluate differences in labs between age groups (young: age < 16, old: age $\ge$ 16), while adjusting for the correlation of repeated measures within a participant. Furthermore, a first order autoregressive (AR(1)) correlation structure was assumed when accounting for the correlated repeated measures, based on visual inspection of within-participant correlation patterns. Estimated marginal means were calculated for each timepoint of the lab within a visit, and pairwise comparisons were used to test whether each timepoint differed between age groups. 

Analyses for each metric are based on the following timepoints:

-   Glucose, Insulin: 0, 10, 20, 30, 45, 60, 90, 120, 150, 180, 210, 240 min
-   C-Peptide: 0, 10, 20, 30
-   GLP-1: 0, 10, 30, 60
-   Glucagon: 0, 10, 30, 60, 90, 240
-   FFA: 0, 30, 60, 120, 180, 210, 240

# Mixed Model Results

How to interpret the output:

* ANOVA Results: This table provides the parameter estimates for each term in the model. The `age_group:time` term tests whether the pattern of change of the lab measure across time differed by age group. For example, based on the C-peptide results, we can see that the trajectory of C-peptide across time did differ between age groups (p < 0.05).

* Estimated Marginal Means: This table reports the estimated marginal means for each age group at each timepoint. `lower.CL` and `upper.CL` correspond to the lower and upper bounds of the 95% confidence interval.

* Pairwise Comparisons: This table includes p-values for testing whether there is a notable difference in the lab measure between age group for each of the timepoints. For example, in the C-peptide pairwise comparisons table, a significant difference in the lab measure between age groups was only observed at 30 min (p < 0.001).

```{r mixed functions, include=FALSE}
## define function that fits LMM with random intercept for pid then outputs a list of the ANOVA, marginal mean, and pairwise comparisons results (inputted dat is the doubly long df for the assay of interest)
mixed <- function(dat, outcome){
  form = paste0(outcome,"~ age_group * time")
  mod <- lme(as.formula(form), random = ~1|pid, data = dat, na.action = na.omit,
             correlation = corAR1(form = ~1 | pid)) # specifies within-subject variability as AR(1)
  anova <- anova.lme(mod, type="marginal") # use ANOVA to get significance of interaction
  m <- emmeans(mod, c("age_group", "time")) # compute estimated marginal means to obtain effect at each timepoint
  prs <-  pairs(m, adjust = "tukey", by = "time") # perform pairwise comparisons by timepoint

  return(list(anova, m, prs))
}
```

## Glucose
<span style="color: blue;">Estimated marginal means and pairwise comparisons may not be meaningful to interpret due to non-significant findings.</span>
```{r glucose analysis, echo=FALSE}
glucose_results <- mixed(dat = glucose_long, outcome = "glucose")

glucose_results[[1]] %>%
  mutate(`p-value` = ifelse(`p-value` < 0.001, format.pval(`p-value`, digits = 3, eps = 0.001), sprintf('%.3f',`p-value`))) %>%
  kable(caption = 'ANOVA Results:', booktabs = TRUE, digits = 3) %>%
  kable_styling(latex_options='basic', html_font='Arial') %>%
      htmltools::HTML()
glucose_results[[2]] %>%
  kable(caption = 'Estimated Marginal Means:', booktabs = TRUE, digits =3) %>%
  kable_styling(latex_options='basic', html_font='Arial') %>%
      htmltools::HTML()
glucose_results[[3]] %>%
  kable(caption = 'Pairwise Comparisons:', booktabs = TRUE, digits =3) %>%
  kable_styling(latex_options='basic', html_font='Arial') %>%
      htmltools::HTML()
```

## Insulin
```{r insulin analysis, echo=FALSE}
insulin_results <- mixed(dat = insulin_long, outcome = "insulin")

insulin_results[[1]] %>%
  as.data.frame() %>% 
  mutate(`p-value` = ifelse(`p-value` < 0.001, format.pval(`p-value`, digits = 3, eps = 0.001), sprintf('%.3f',`p-value`))) %>% 
  kable(caption = 'ANOVA Results:', booktabs = TRUE, digits = 3) %>%
  kable_styling(latex_options='basic', html_font='Arial') %>%
      htmltools::HTML()
insulin_results[[2]] %>%
  kable(caption = 'Estimated Marginal Means:', booktabs = TRUE, digits =3) %>%
  kable_styling(latex_options='basic', html_font='Arial') %>%
      htmltools::HTML()
insulin_results[[3]] %>%
  kable(caption = 'Pairwise Comparisons:', booktabs = TRUE, digits =3) %>%
  kable_styling(latex_options='basic', html_font='Arial') %>%
      htmltools::HTML()
```

## C-Peptide
<!-- <span style="color: blue;">Estimated marginal means and pairwise comparisons were not computed due to non-significant findings.</span> -->
```{r cpep analysis, echo=FALSE}
cpep_results <- mixed(dat = cpep_long, outcome = "cpeptide")

cpep_results[[1]] %>%
  mutate(`p-value` = ifelse(`p-value` < 0.001, format.pval(`p-value`, digits = 3, eps = 0.001), sprintf('%.3f',`p-value`))) %>% 
  kable(caption = 'ANOVA Results:', booktabs = TRUE, digits = 3) %>%
  kable_styling(latex_options='basic', html_font='Arial') %>%
      htmltools::HTML()
cpep_results[[2]] %>%
  kable(caption = 'Estimated Marginal Means:', booktabs = TRUE, digits =3) %>%
  kable_styling(latex_options='basic', html_font='Arial') %>%
      htmltools::HTML()
cpep_results[[3]] %>%
  as.data.frame() %>% 
  mutate(p.value = ifelse(p.value < 0.001, format.pval(p.value, digits = 3, eps = 0.001), sprintf('%.3f',p.value))) %>% 
  kable(caption = 'Pairwise Comparisons:', booktabs = TRUE, digits = 3) %>%
  kable_styling(latex_options='basic', html_font='Arial') %>%
      htmltools::HTML()
```

## GLP-1
<span style="color: blue;">Estimated marginal means and pairwise comparisons may not be meaningful to interpret due to non-significant findings.</span>
```{r glp1 analysis, echo=FALSE}
glp1_results <- mixed(dat = glp1_long, outcome = "glp1")

glp1_results[[1]] %>%
  mutate(`p-value` = ifelse(`p-value` < 0.001, format.pval(`p-value`, digits = 3, eps = 0.001), sprintf('%.3f',`p-value`))) %>% 
  kable(caption = 'ANOVA Results:', booktabs = TRUE, digits = 3) %>%
  kable_styling(latex_options='basic', html_font='Arial') %>%
      htmltools::HTML()
glp1_results[[2]] %>%
  kable(caption = 'Estimated Marginal Means:', booktabs = TRUE, digits =3) %>%
  kable_styling(latex_options='basic', html_font='Arial') %>%
      htmltools::HTML()
glp1_results[[3]] %>%
  kable(caption = 'Pairwise Comparisons:', booktabs = TRUE, digits =3) %>%
  kable_styling(latex_options='basic', html_font='Arial') %>%
      htmltools::HTML()
```

## Glucagon
<span style="color: blue;">Estimated marginal means and pairwise comparisons may not be meaningful to interpret due to non-significant findings.</span>
```{r glucagon analysis, echo=FALSE}
glucagon_results <- mixed(dat = glucagon_long, outcome = "glucagon")

glucagon_results[[1]] %>%
  mutate(`p-value` = ifelse(`p-value` < 0.001, format.pval(`p-value`, digits = 3, eps = 0.001), sprintf('%.3f',`p-value`))) %>% 
  kable(caption = 'ANOVA Results:', booktabs = TRUE, digits = 3) %>%
  kable_styling(latex_options='basic', html_font='Arial') %>%
      htmltools::HTML()
glucagon_results[[2]] %>%
  kable(caption = 'Estimated Marginal Means:', booktabs = TRUE, digits =3) %>%
  kable_styling(latex_options='basic', html_font='Arial') %>%
      htmltools::HTML()
glucagon_results[[3]] %>%
  kable(caption = 'Pairwise Comparisons:', booktabs = TRUE, digits =3) %>%
  kable_styling(latex_options='basic', html_font='Arial') %>%
      htmltools::HTML()
```

## FFA
<span style="color: blue;">Estimated marginal means and pairwise comparisons may not be meaningful to interpret due to non-significant findings.</span>
```{r ffa analysis, echo=FALSE}
ffa_results <- mixed(dat = ffa_long, outcome = "ffa")

ffa_results[[1]] %>%
  mutate(`p-value` = ifelse(`p-value` < 0.001, format.pval(`p-value`, digits = 3, eps = 0.001), sprintf('%.3f',`p-value`))) %>% 
  kable(caption = 'ANOVA Results:', booktabs = TRUE, digits = 3) %>%
  kable_styling(latex_options='basic', html_font='Arial') %>%
      htmltools::HTML()
ffa_results[[2]] %>%
  kable(caption = 'Estimated Marginal Means:', booktabs = TRUE, digits =3) %>%
  kable_styling(latex_options='basic', html_font='Arial') %>%
      htmltools::HTML()
ffa_results[[3]] %>%
  kable(caption = 'Pairwise Comparisons:', booktabs = TRUE, digits =3) %>%
  kable_styling(latex_options='basic', html_font='Arial') %>%
      htmltools::HTML()
```