---
title: "Gourgari - Analysis for PAS"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = FALSE)

library(tableone)
library(Hmisc)
library(childsds)
library(tableone)
library(mixOmics)

source("C:/Users/pylell/Documents/GitHub/CHCO-Code/Jenny Gourgari/make final dataset.R")

source("c:/Users/pylell/Documents/GitHub/General-code/temp_table1.r")

source("c:/Users/pylell/Documents/GitHub/General-code/01_functions.r")

# need to combine race/ethnicity 
alldata$raceeth[alldata$Ethnicity=="Hispanic or Latino"] <- "Hispanic"
alldata$raceeth[alldata$Ethnicity !="Hispanic or Latino" & alldata$Race=="White"] <- "Non-hispanic White"
alldata$raceeth[is.na(alldata$raceeth)] <- "Other"
label(alldata$raceeth) = "Race/ethnicity"

# calculate BMI-Z
alldata$bmiz = sds(alldata$bmi,
              age = alldata$Age,
              sex = alldata$Gender, male = "M", female =  "F",
              ref = cdc.ref,
              item = "bmi",
              type = "SDS")
alldata$bmiz <- round(alldata$bmiz,2)
label(alldata$bmiz)="BMI-Z"

# calculate units/kg
alldata$units_kg <- alldata$UnitsInsTotal / alldata$weight
label(alldata$units_kg)="Units insulin/kg"

# for subjects with protein data, make a wide dataset and calculate deltas for protein and clinical variables
hasproteins <- alldata[!is.na(alldata$subjectid),]
# make a wide dataset for those with protein dataset
hasproteins_keep <- select(hasproteins,-c("subjectid","IL.1b","IL.12.IL.23p40","IL.17A","IL.18","IL.6","IL.8","il17F","MCP.1..Human.","TNF.a",
                                          "Value.ADIP","Value.CRPM","Value.EIA-ESelectin","Value.EIA-IL-6","Value.EIA-SICAM","Value.EIA-TNF-A", 
                                          "Value.SVCAM","BldPrSys","BldPrDia","DEXATotMass","DEXAFat","DEXALean","DEXABMC",
                                          "UnitsInsTotalPumpOrLog","UnitsInsBasalOrLongAct","InsDeliveryMethod",
                                          "height","units_per_kg","adipo_ugml","date","IFN.g"))
deltavars <- c("efflux.value","sp|P01024|CO3_HUMAN","sp|P02768|ALBU_HUMAN", "sp|P02671|FIBA_HUMAN",  "sp|P02751|FINC_HUMAN",
                                        "sp|P0C0L4|CO4A_HUMAN",  "sp|P02675|FIBB_HUMAN", "sp|P01023|A2MG_HUMAN",  "sp|P04114|APOB_HUMAN",
                                        "sp|P02679|FIBG_HUMAN",  "sp|P02647|APOA1_HUMAN", "sp|P06727|APOA4_HUMAN", "sp|P02787|TRFE_HUMAN",
                                        "sp|P00738|HPT_HUMAN",   "sp|P00751|CFAB_HUMAN", "sp|P00747|PLMN_HUMAN",  "sp|P02790|HEMO_HUMAN", 
                                        "sp|P01031|CO5_HUMAN",   "sp|P02774|VTDB_HUMAN",  "sp|P08603|CFAH_HUMAN",  "sp|P02748|CO9_HUMAN",
                                        "sp|P01871|IGHM_HUMAN",  "sp|P00450|CERU_HUMAN",  "sp|P02649|APOE_HUMAN",  "sp|P19823|ITIH2_HUMAN",
                                        "sp|P00736|C1R_HUMAN",   "sp|P10909|CLUS_HUMAN", "sp|P00734|THRB_HUMAN",  "sp|Q14624|ITIH4_HUMAN",
                                        "sp|P06396|GELS_HUMAN",  "sp|P01008|ANT3_HUMAN",  "sp|P03952|KLKB1_HUMAN", "sp|P0CG05|LAC2_HUMAN",
                                        "sp|P36955|PEDF_HUMAN",  "sp|P04004|VTNC_HUMAN",  "sp|P12259|FA5_HUMAN",   "sp|P04217|A1BG_HUMAN", 
                                        "sp|P07358|CO8B_HUMAN",  "sp|P08697|A2AP_HUMAN", "sp|P02765|FETUA_HUMAN", "sp|P10643|CO7_HUMAN",  
                                        "sp|P01834|IGKC_HUMAN",  "sp|P35858|ALS_HUMAN",   "sp|P27169|PON1_HUMAN",  "sp|P09871|C1S_HUMAN",  
                                        "sp|P02749|APOH_HUMAN",  "sp|P51884|LUM_HUMAN",   "sp|P01876|IGHA1_HUMAN", "sp|P07357|CO8A_HUMAN", 
                                        "sp|P68871|HBB_HUMAN",   "sp|P02654|APOC1_HUMAN", "sp|P01764|HV303_HUMAN", "sp|P36980|FHR2_HUMAN", 
                                        "sp|P05452|TETN_HUMAN",  "sp|P05156|CFAI_HUMAN",  "sp|O95445|APOM_HUMAN",  "sp|Q96PD5|PGRP2_HUMAN",
                                        "sp|P18135|KV312_HUMAN", "sp|P43652|AFAM_HUMAN",  "sp|P01779|HV318_HUMAN", "sp|P69905|HBA_HUMAN",  
                                        "sp|P01781|HV320_HUMAN", "sp|P63261|ACTG_HUMAN", "sp|P80748|LV302_HUMAN", "sp|P01743|HV102_HUMAN",
                                        "sp|Q96IY4|CBPB2_HUMAN", "sp|P01019|ANGT_HUMAN",  "sp|P02753|RET4_HUMAN",  "sp|P00748|FA12_HUMAN", 
                                        "sp|P01591|IGJ_HUMAN",   "sp|O14791|APOL1_HUMAN", "sp|Q9UK55|ZPI_HUMAN",   "sp|P02655|APOC2_HUMAN",
                                        "sp|P02766|TTHY_HUMAN",  "sp|Q03591|FHR1_HUMAN", "sp|P04433|KV309_HUMAN", "sp|P04434|KV310_HUMAN",
                                        "sp|P01707|LV204_HUMAN", "sp|B9A064|IGLL5_HUMAN", "sp|P07360|CO8G_HUMAN",  "sp|P01861|IGHG4_HUMAN",
                                        "sp|P02746|C1QB_HUMAN",  "sp|P00740|FA9_HUMAN",   "sp|P00739|HPTR_HUMAN",  "sp|P04208|LV106_HUMAN", 
                                        "HbA1c","Value.LDL","Value.VLDL","Value.HDL","Value.TG-NET","Value.CHOL","DEXAPercFat",
                                        "bmi","bmiperc","eis","bmiz","WaistCircum","weight","UnitsInsTotal","units_kg")
hasproteins_wide <- reshape(hasproteins_keep,idvar = c("analyticid","Age","Gender","raceeth","TannerPubicH","TannerBreGen"),
                            timevar = "visit",v.names = deltavars,direction="wide")

analyticid <- as.matrix(hasproteins_wide$analyticid)
for (x in deltavars)  {
  del.x <- paste0("delta.",x) 
  mo6.x <- paste0(x,".6 month")
  base.x <- paste0(x,".Baseline")
  temp <- select(hasproteins_wide,mo6.x) - select(hasproteins_wide,base.x)
  colnames(temp) <- del.x
  analyticid <- as.data.frame(cbind(analyticid,temp))
}
hasproteins_wide <- merge(hasproteins_wide,analyticid,by="analyticid")
# calculate percent change
analyticid <- as.matrix(hasproteins_wide$analyticid)
for (x in deltavars)  {
  delper.x <- paste0("dp.",x) 
  mo6.x <- paste0(x,".6 month")
  base.x <- paste0(x,".Baseline")
  temp <- ((select(hasproteins_wide,mo6.x) - select(hasproteins_wide,base.x))/select(hasproteins_wide,base.x))*100
  colnames(temp) <- delper.x
  analyticid <- as.data.frame(cbind(analyticid,temp))
}
hasproteins_wide <- merge(hasproteins_wide,analyticid,by="analyticid")
# get rid of participants with only one time point
hasproteins_wide <- hasproteins_wide[!is.na(hasproteins_wide$`delta.sp|P00740|FA9_HUMAN`),]
# check
#View(hasproteins_wide[,c("deltaper.sp|P01024|CO3_HUMAN","delta.sp|P01024|CO3_HUMAN","sp|P01024|CO3_HUMAN.Baseline","sp|P01024|CO3_HUMAN.6 month")])

# calculate change in BMI-Z and waist and insulin 
hasproteins_wide$bmiz_delta <- hasproteins_wide$`bmiz.6 month` - hasproteins_wide$bmiz.Baseline
hasproteins_wide$waist_delta <- hasproteins_wide$`WaistCircum.6 month` - hasproteins_wide$WaistCircum.Baseline
hasproteins_wide$ins_delta <- hasproteins_wide$`units_kg.6 month` - hasproteins_wide$units_kg.Baseline
label(hasproteins_wide$bmiz_delta)="Change in BMI-Z"
label(hasproteins_wide$waist_delta)="Change in waist circ"
label(hasproteins_wide$ins_delta)="Change in units of insulin/kg"

# categorize people whether they have reduction in BMI of 10% and reduction in insulin/kg of 25%
hasproteins_wide$bmiz10p <- ifelse(is.na(hasproteins_wide$dp.bmiz),NA,
                                   ifelse(hasproteins_wide$dp.bmiz<= -10,1,0))
hasproteins_wide$ins25p <- ifelse(is.na(hasproteins_wide$dp.units_kg),NA,
                                   ifelse(hasproteins_wide$dp.units_kg<= -25,1,0))
hasproteins_wide$bmiz10p <- as.factor(hasproteins_wide$bmiz10p)
hasproteins_wide$ins25p <- as.factor(hasproteins_wide$ins25p)
      
# table 1 for participants in protein dataset
t1_prot <- final_table(data=hasproteins_wide,variables=c("Age","Gender","raceeth","TannerPubicH","TannerBreGen"),group=hasproteins_wide$Treatment.Group,
                       margin=2,single=0,ron=1)
nprot <- nrow(hasproteins_wide)

# create table of deltas by treatment group
#for (i in 1:length(deltavars)) {
#  deltavars[i] <- paste0("delta.",deltavars[i])
#}
#tab_delta_prot <- final_table(data=hasproteins_wide,variables=deltavars,group=hasproteins_wide$Treatment.Group,margin=2,single=0,ron=3,summary.stat = "median")

# create table of percent deltas by treatment group
deltapervars <- NULL
for (i in 1:length(deltavars)) {
  deltapervars[i] <- paste0("dp.",deltavars[i])
}
tab_deltaper_prot <- final_table(data=hasproteins_wide,variables=deltapervars,group=hasproteins_wide$Treatment.Group,margin=2,single=0,ron=3,summary.stat = "median")
tab_deltaper_prot$adjp <- NA
tab_deltaper_prot$adjp <- p.adjust(tab_deltaper_prot$`P Value`,method="BH")

# create a table of clinical vars at each time point
clinvars <- c("HbA1c.Baseline","Value.LDL.Baseline","Value.VLDL.Baseline","Value.HDL.Baseline","Value.TG-NET.Baseline",
              "Value.CHOL.Baseline","DEXAPercFat.Baseline","bmi.Baseline","bmiperc.Baseline","bmiz.Baseline","eis.Baseline","WaistCircum.Baseline",
              "units_kg.Baseline","HbA1c.6 month","Value.LDL.6 month","Value.VLDL.6 month","Value.HDL.6 month","Value.TG-NET.6 month",
              "Value.CHOL.6 month","DEXAPercFat.6 month","bmi.6 month","bmiperc.6 month","bmiz.6 month","eis.6 month",
              "WaistCircum.6 month","units_kg.6 month","bmiz_delta","waist_delta","ins_delta")
tab_clinvars <- final_table(data=hasproteins_wide,variables=clinvars,group=hasproteins_wide$Treatment.Group,margin=2,single=0,ron=3,summary.stat = "median")
tab_clinvars[,1] <- clinvars

tab_clinvars_cat <- CreateTableOne(data=hasproteins_wide,vars=c("ins25p","bmiz10p"),strata = "Treatment.Group")
tab_clinvars_cat <- print(tab_clinvars_cat,showAllLevels = T,exact=c("ins25p","bmiz10p"))

# correlations of changes in proteins with changes in clinical variables
protdelt <- grep("sp|",deltapervars,value=T,fixed=T)
otherdelt <- deltapervars[!(deltapervars %in% protdelt)]
# ++++++++++++++++++++++++++++
# flattenCorrMatrix
# ++++++++++++++++++++++++++++
# cormat : matrix of the correlation coefficients
# pmat : matrix of the correlation p-values
flattenCorrMatrix <- function(cormat, pmat) {
  ut <- upper.tri(cormat)
  data.frame(
    row = rownames(cormat)[row(cormat)[ut]],
    column = rownames(cormat)[col(cormat)[ut]],
    cor  =(cormat)[ut],
    p = pmat[ut]
    )
}
corrout <- rcorr(as.matrix(hasproteins_wide[,deltapervars]),type = "spearman")
corrout <- flattenCorrMatrix(corrout$r, corrout$P)
corrout <- corrout[corrout$column %in% otherdelt,]
corrout <- corrout[corrout$row %in% protdelt,]
corrout$adjp <- p.adjust(corrout$p)

# correlations with change in efflux
corrout_efflux <- rcorr(as.matrix(hasproteins_wide[,c(protdelt,"dp.efflux.value")]),type = "spearman")
corrout_efflux <- flattenCorrMatrix(corrout_efflux$r, corrout_efflux$P)
corrout_efflux <- corrout_efflux[corrout_efflux$column=="dp.efflux.value",]
corrout_efflux$adjp <- p.adjust(corrout_efflux$p)

# who has protein data and cytokine data
x <- alldata[!is.na(alldata$subjectid) & !is.na(alldata$IL.1b),]

# correlations between cytokines and proteins at baseline
cytokinevars <- c("IFN.g","IL.1b","IL.12.IL.23p40","IL.17A","IL.18","IL.6","IL.8","MCP.1..Human.","TNF.a","il17F")
protvars <- c("sp|P01024|CO3_HUMAN","sp|P02768|ALBU_HUMAN", "sp|P02671|FIBA_HUMAN",  "sp|P02751|FINC_HUMAN",
                                        "sp|P0C0L4|CO4A_HUMAN",  "sp|P02675|FIBB_HUMAN", "sp|P01023|A2MG_HUMAN",  "sp|P04114|APOB_HUMAN",
                                        "sp|P02679|FIBG_HUMAN",  "sp|P02647|APOA1_HUMAN", "sp|P06727|APOA4_HUMAN", "sp|P02787|TRFE_HUMAN",
                                        "sp|P00738|HPT_HUMAN",   "sp|P00751|CFAB_HUMAN", "sp|P00747|PLMN_HUMAN",  "sp|P02790|HEMO_HUMAN", 
                                        "sp|P01031|CO5_HUMAN",   "sp|P02774|VTDB_HUMAN",  "sp|P08603|CFAH_HUMAN",  "sp|P02748|CO9_HUMAN",
                                        "sp|P01871|IGHM_HUMAN",  "sp|P00450|CERU_HUMAN",  "sp|P02649|APOE_HUMAN",  "sp|P19823|ITIH2_HUMAN",
                                        "sp|P00736|C1R_HUMAN",   "sp|P10909|CLUS_HUMAN", "sp|P00734|THRB_HUMAN",  "sp|Q14624|ITIH4_HUMAN",
                                        "sp|P06396|GELS_HUMAN",  "sp|P01008|ANT3_HUMAN",  "sp|P03952|KLKB1_HUMAN", "sp|P0CG05|LAC2_HUMAN",
                                        "sp|P36955|PEDF_HUMAN",  "sp|P04004|VTNC_HUMAN",  "sp|P12259|FA5_HUMAN",   "sp|P04217|A1BG_HUMAN", 
                                        "sp|P07358|CO8B_HUMAN",  "sp|P08697|A2AP_HUMAN", "sp|P02765|FETUA_HUMAN", "sp|P10643|CO7_HUMAN",  
                                        "sp|P01834|IGKC_HUMAN",  "sp|P35858|ALS_HUMAN",   "sp|P27169|PON1_HUMAN",  "sp|P09871|C1S_HUMAN",  
                                        "sp|P02749|APOH_HUMAN",  "sp|P51884|LUM_HUMAN",   "sp|P01876|IGHA1_HUMAN", "sp|P07357|CO8A_HUMAN", 
                                        "sp|P68871|HBB_HUMAN",   "sp|P02654|APOC1_HUMAN", "sp|P01764|HV303_HUMAN", "sp|P36980|FHR2_HUMAN", 
                                        "sp|P05452|TETN_HUMAN",  "sp|P05156|CFAI_HUMAN",  "sp|O95445|APOM_HUMAN",  "sp|Q96PD5|PGRP2_HUMAN",
                                        "sp|P18135|KV312_HUMAN", "sp|P43652|AFAM_HUMAN",  "sp|P01779|HV318_HUMAN", "sp|P69905|HBA_HUMAN",  
                                        "sp|P01781|HV320_HUMAN", "sp|P63261|ACTG_HUMAN", "sp|P80748|LV302_HUMAN", "sp|P01743|HV102_HUMAN",
                                        "sp|Q96IY4|CBPB2_HUMAN", "sp|P01019|ANGT_HUMAN",  "sp|P02753|RET4_HUMAN",  "sp|P00748|FA12_HUMAN", 
                                        "sp|P01591|IGJ_HUMAN",   "sp|O14791|APOL1_HUMAN", "sp|Q9UK55|ZPI_HUMAN",   "sp|P02655|APOC2_HUMAN",
                                        "sp|P02766|TTHY_HUMAN",  "sp|Q03591|FHR1_HUMAN", "sp|P04433|KV309_HUMAN", "sp|P04434|KV310_HUMAN",
                                        "sp|P01707|LV204_HUMAN", "sp|B9A064|IGLL5_HUMAN", "sp|P07360|CO8G_HUMAN",  "sp|P01861|IGHG4_HUMAN",
                                        "sp|P02746|C1QB_HUMAN",  "sp|P00740|FA9_HUMAN",   "sp|P00739|HPTR_HUMAN",  "sp|P04208|LV106_HUMAN")

corrbase <- rcorr(as.matrix(x[x$visit=="Baseline",c(cytokinevars,protvars)]),type = "spearman")
corrbase <- flattenCorrMatrix(corrbase$r, corrbase$P)
corrbase <- corrbase[corrbase$column %in% protvars,]
corrbase <- corrbase[corrbase$row %in% cytokinevars,]

# correlations between changes in proteins and changes in cytokines
x_keep <-select(x,-c("subjectid","Value.ADIP","Value.CRPM","Value.EIA-ESelectin","Value.EIA-IL-6","Value.EIA-SICAM","Value.EIA-TNF-A", 
                                          "Value.SVCAM","WaistCircum","BldPrSys","BldPrDia","DEXATotMass","DEXAFat","DEXALean","DEXABMC",
                                          "UnitsInsTotal","UnitsInsTotalPumpOrLog","UnitsInsBasalOrLongAct","InsDeliveryMethod","weight",
                                          "height","units_per_kg","adipo_ugml","date","efflux.value","HbA1c","Value.LDL",
                                          "Value.VLDL","Value.HDL","Value.TG-NET","Value.CHOL","DEXAPercFat","bmi","bmiperc","eis"))
proteinscytokines_wide <- reshape(x_keep,idvar = c("analyticid","Age","Gender","raceeth","TannerPubicH","TannerBreGen"),
                            timevar = "visit",v.names = c(protvars,cytokinevars),direction="wide")
analyticid <- as.matrix(proteinscytokines_wide$analyticid)
for (x in c(protvars,cytokinevars))  {
  del.x <- paste0("delta.",x) 
  mo6.x <- paste0(x,".6 month")
  base.x <- paste0(x,".Baseline")
  temp <- select(proteinscytokines_wide,mo6.x) - select(proteinscytokines_wide,base.x)
  colnames(temp) <- del.x
  analyticid <- as.data.frame(cbind(analyticid,temp))
}
proteinscytokines_wide <- merge(proteinscytokines_wide,analyticid,by="analyticid")
# get rid of participants with only one time point
proteinscytokines_wide <- proteinscytokines_wide[!is.na(proteinscytokines_wide$`delta.sp|P00740|FA9_HUMAN`),]
proteinscytokines_wide$dummy <- rep(1,nrow(proteinscytokines_wide))
deltavars2 <- c(protvars,cytokinevars)
for (i in 1:length(deltavars2)) {
  deltavars2[i] <- paste0("delta.",deltavars2[i])
}
cytodelt <- deltavars2[!(deltavars2 %in% protdelt)]
corrcyto <- rcorr(as.matrix(proteinscytokines_wide[,deltavars2]),type = "spearman")
corrcyto <- flattenCorrMatrix(corrcyto$r, corrcyto$P)
corrcyto <- corrcyto[corrcyto$column %in% cytodelt,]
corrcyto <- corrcyto[corrcyto$row %in% protdelt,]


# table 1 for participants in protein dataset
t1b <- final_table(data=proteinscytokines_wide,variables=c("Age","Gender","raceeth","TannerPubicH","TannerBreGen"),group=proteinscytokines_wide$dummy,
                   margin=2,single=0,ron=1)
ncytoprot <- nrow(proteinscytokines_wide)

# efflux analysis
hasefflux <- alldata[!is.na(alldata$efflux.value),]
hasefflux_keep <- select(hasefflux,c("analyticid","visit","Age","Gender","raceeth","TannerPubicH","TannerBreGen","efflux.value","Value.HDL",
                                     "Treatment.Group"))
hasefflux_wide <- reshape(hasefflux_keep,idvar = c("analyticid","Age","Gender","raceeth","TannerPubicH","TannerBreGen","Treatment.Group"),
                            timevar = "visit",v.names = c("efflux.value","Value.HDL"),direction="wide")
analyticid <- as.matrix(hasefflux_wide$analyticid)
for (x in c("efflux.value","Value.HDL"))  {
  del.x <- paste0("delta.",x) 
  mo6.x <- paste0(x,".6 month")
  base.x <- paste0(x,".Baseline")
  temp <- select(hasefflux_wide,mo6.x) - select(hasefflux_wide,base.x)
  colnames(temp) <- del.x
  analyticid <- as.data.frame(cbind(analyticid,temp))
}
hasefflux_wide <- merge(hasefflux_wide,analyticid,by="analyticid")

# table 1 for participants in efflux dataset
t1 <- final_table(data=hasefflux_wide,variables=c("Age","Gender","raceeth","TannerPubicH","TannerBreGen"),
                  group=hasefflux_wide$Treatment.Group,margin=2,single=0,ron=1)
nefflux <- nrow(hasefflux_wide)

# raw changes in efflux
label(hasefflux_wide$delta.efflux.value) = "Change in efflux"
effluxtab <- final_table(data=hasefflux_wide,variables="delta.efflux.value",
                  group=hasefflux_wide$Treatment.Group,margin=2,single=0,ron=3)

# change in efflux by treatment group, adjusted for baseline HDL
mod <- lm(data=hasefflux_wide, delta.efflux.value ~ Treatment.Group + Value.HDL.Baseline )
mod <- mod_tab(mod)

# repeat efflux analysis for people with proteomics data
effluxtab_prot <- final_table(data=hasefflux_wide[hasefflux_wide$analyticid %in% hasproteins_wide$analyticid,],variables="delta.efflux.value",
                  group=hasefflux_wide[hasefflux_wide$analyticid %in% hasproteins_wide$analyticid,]$Treatment.Group,margin=2,single=0,ron=3)
mod_prot <- lm(data=hasefflux_wide[hasefflux_wide$analyticid %in% hasproteins_wide$analyticid,], delta.efflux.value ~ Treatment.Group + Value.HDL.Baseline )
mod_prot <- mod_tab(mod_prot)


# plots - save files
jpeg("E:\\Nadeau\\T1D Exchange metformin and lipids\\PAS 2019 abstract\\PGRP2_eis.jpg")
plot(hasproteins_wide$dp.eis,hasproteins_wide$`dp.sp|Q96PD5|PGRP2_HUMAN`,xlab="% change in eIS",ylab="% change in PGRP2")
abline(lm(`dp.sp|Q96PD5|PGRP2_HUMAN`~dp.eis,data=hasproteins_wide))
dev.off()

jpeg("E:\\Nadeau\\T1D Exchange metformin and lipids\\PAS 2019 abstract\\PGRP2_tg.jpg")
plot(hasproteins_wide$`dp.Value.TG-NET`,hasproteins_wide$`dp.sp|Q96PD5|PGRP2_HUMAN`,xlab="% change in TG",ylab="% change in PGRP2")
abline(lm(`dp.sp|Q96PD5|PGRP2_HUMAN`~`dp.Value.TG-NET`,data=hasproteins_wide))
dev.off()

jpeg("E:\\Nadeau\\T1D Exchange metformin and lipids\\PAS 2019 abstract\\PGRP2_VLDL.jpg")
plot(hasproteins_wide$dp.Value.VLDL~hasproteins_wide$`dp.sp|Q96PD5|PGRP2_HUMAN`,xlab="% change in VLDL",ylab="% change in PGRP2")
abline(lm(`dp.sp|Q96PD5|PGRP2_HUMAN`~dp.Value.VLDL,data=hasproteins_wide))
dev.off()


#a <- boxplot(hasproteins_wide$`dp.sp|P01023|A2MG_HUMAN`,hasproteins_wide$Treatment.Group,names=c('Metformin','Placebo'),
 #       outline=FALSE)

# plsda of changes in proteins vs. treatmentgroup
plsda_tx <- splsda(hasproteins_wide[,protdelt],Y=as.factor(hasproteins_wide$Treatment.Group))

```

# Background

The purpose of this analysis is to compare changes in HDL efflux in the metformin and placebo groups in the T1D Exchange metformin and lipids cohort.  For participants with proteomic data, changes in proteins by treatment group were also compared, and correlations between changes in proteins and changes in clinical variables (BMI percentile, estimated insulin sensitivity, HbA1c, % fat by DXA, and lipids) were examined.  For participants with proteomic data and cytokine data, correlations between changes in proteins and changes in cytokines were examined.

# Methods

Some samples had two sets of efflux measurments from assays run on different days.  Results from the "copy of Jenny manifest" file were used if available.  If there were no results for a sample in that file, then results from the "Jenny proteome efflux for me" file were used.

BMI percentile was calculated using the childsds package in r.  Estimated insulin sensitivity was calculated using the CACTI equation for people with T1D, using the version of the equation with adiponectin.  If adiponectin was missing at a visit, then the equation without adiponectin was used.

The distributions of all variables were examined prior to analysis.  Descriptive statistics reported include frequencies and percentages for categorical variables, means and standard deviations for normally distributed continuous variables, and medians and percentiles for non-normally distributed continous variables.  To compare groups, either the chi-square test or Fisher's Exact test was used for categorical variables, and either the Mann-Whitney test or t-tests were used for continuous variables.  Spearman's correlation coefficient was used to assess correlation between continuous variables.  A linear model was used to test the difference in change in efflux between the two treatment groups while adjusting for baseline HDL.  

All analyses were performed using R, version 3.5.1.

# Results

There were `r nprot` partcipants with protein data at baseline and 6 months.  There were `r ncytoprot` participants with both protein and cytokine data at baseline and 6 months.  There were `r nefflux` participants with efflux data at baseline and 6 months.  


```{r echo=FALSE}
kable(t1_prot,caption = "Table 1a.  Descriptive statistics for participants with proteomics data, by treatment group.  Statistics given are N(%), mean?SD, or median (25th %ile, 75th %ile).  * denotes that the Mann-Whitney test was used, and ** denotes that Fisher's exact test was used.")
```
<br>


```{r echo=FALSE}
kable(t1,caption = "Table 1b.  Descriptive statistics for participants with efflux data, by treatment group.  Statistics given are N(%), mean?SD, or median (25th %ile, 75th %ile). * denotes that the Mann-Whitney test was used, and ** denotes that Fisher's exact test was used.")
```
<br>


```{r echo=FALSE}
kable(tab_deltaper_prot,caption = "Table 2a.  Descriptive statistics for percent changes in proteins and other variables, for participants with proteomics data, by treatment group.  Statistics given are median (25th %ile, 75th %ile).  P-values are from the Mann-Whitney test.")
```
<br>


```{r echo=FALSE}
kable(tab_clinvars,caption = "Table 2b.  Descriptive statistics for clinical variables, for participants with proteomics data, by treatment group.  Statistics given are median (25th %ile, 75th %ile).  P-values are from the Mann-Whitney test.")
```
<br>


```{r echo=FALSE}
kable(tab_clinvars_cat, caption="Table 2c.  Percentage of participants with at least a 10% reduction in BMI-Z and at least a 25% reduction in insulin/kg.")
```
<br>
      

```{r echo=FALSE}
kable(corrout,caption = "Table 3a.  Spearman correlation coefficients for percent changes in proteins with percent changes in clinical variables.")
```
<br>

```{r echo=FALSE}
kable(corrout_efflux,caption = "Table 3b.  Spearman correlation coefficients for percent changes in proteins with percent change in efflux.")
```
<br>

```{r echo=FALSE}
kable(effluxtab,caption = "Table 4.  Unadjusted changes in efflux, by treatment group.  Statistics given are mean +/- SD.")
```
<br>

```{r echo=FALSE}
kable(mod,caption = "Table 5.  Linear model of change in efflux by treatment group, adjusted for baseline HDL.  Statistics given are beta coefficients and 95% confidence intervals.")
```
<br>

```{r echo=FALSE}
kable(effluxtab_prot,caption = "Table 6.  Unadjusted changes in efflux for participants with proteomics data, by treatment group.  Statistics given are mean +/- SD.")
```
<br>

```{r echo=FALSE}
kable(mod_prot,caption = "Table 7.  Linear model of change in efflux by treatment group for participants with proteomics data, adjusted for baseline HDL.  Statistics given are beta coefficients and 95% confidence intervals.")
```

/newpage

# Individual plot from sPLS-DA of change in proteins by treatment group

```{r, echo=FALSE}
plotIndiv(plsda_tx, comp = c(1,2),
          ind.names = FALSE, 
          ellipse = TRUE, legend = TRUE, title="")
```

```{r, echo=FALSE}
top_1 <- selectVar(plsda_tx,comp=1)
top_2 <- selectVar(plsda_tx,comp=2)
```

# Ranking of proteins in terms of discrimination of treatment group, for component 1

```{r, echo=FALSE}
top_1$value
```

# Ranking of proteins in terms of discrimination of treatment group, for component 2

```{r, echo=FALSE}
top_2$value
```

```{r, echo=FALSE}
plsda.perf = perf(plsda_tx, validation = 'Mfold', folds = 5, 
                           progressBar = FALSE, nrepeat = 10, dist = 'max.dist',auc=TRUE)
auc_save <- plsda.perf$auc$`comp1`[1]    
auc_true <- as.numeric(plsda.perf$auc$comp1["AUC.mean"])
```

# Evaluation of sPLS-DA model 

Leave-one-out cross-validation was used to examine the performance of the sPLS-DA model.  The AUC indicates that the model performed just slightly better than randomly chance (i.e., a coin flip).

 Overall Error rate : `r plsda.perf$error.rate$overall[1,1]`
 AUC: `r auc_save[1]`

```{r plsda,cache=TRUE, warning=FALSE}
# Permutation testing - takes forever, make sure to cache
n_perm <- 1000
aucs <- vector(mode="numeric", length=n_perm)
for (i in 1:n_perm) {
  Y <- sample(as.factor(hasproteins_wide$Treatment.Group),replace = F)
  plsda_res <- plsda(hasproteins_wide[,protdelt],Y,ncomp = 1)
  perf_plsda <- perf(plsda_res,progressBar=FALSE, auc=TRUE,
                   validation = "loo")
  aucs[i] <- as.numeric(perf_plsda$auc$comp1["AUC.mean"])
}
ggplot(as.data.frame(aucs),aes(x=aucs)) + 
  geom_histogram(binwidth = 0.01) + 
  geom_vline(aes(xintercept=auc_true),color="red") + 
  theme_bw() + xlab("AUC")
```

Group labels were permuted `r n_perm` times. One component PLS-DA was run for each permutation and AUC was calculated using leave one out cross-validation. Red line indicates the AUC calculated for non-permuted data.

