---
title: "Preliminary CVD cytokine analysis"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(knitr)
library(dplyr)
library(data.table)
library(tidyr)
library(Hmisc)
library(rmarkdown)

####################################################################################################
#Data Manipulation #
####################################################################################################

# read main data sheet
all <- read.csv("H:/Endocrinology/Nadeau/T1D Exchange metformin and lipids/Preliminary CVD cytokine data/Jenny manifest v4 GRP2 9th Jan 2019.csv",na.strings = c("","NA"," ","error"))
# delete records with missing data
all <- all[!is.na(all$IFN.g),]

source("C:/Users/pylell/Documents/GitHub/General-code/temp_table1.r")

# read in randomization
rand <- read.csv("H:/Endocrinology/Nadeau/T1D Exchange metformin and lipids/Preliminary CVD cytokine data/metformin vs placebo.csv")

# merge datasets
alldata <- merge(all,rand,by=c("ANALYTICID"))

# get rid of unneeded variables
alldata <- select(alldata,-c("RN","PARENT_ID","SAMPLE_ID","STORAGETYPE","SAMPLEGROUP","BOX.ID","NO.","ROW","COL","pg.ml","X"))

# assign variable for pre vs. post treatment
alldata <- alldata[order(alldata$ANALYTICID,alldata$COLLECTIONDT),]
alldata <- alldata %>% group_by(ANALYTICID) %>% mutate(id=row_number())
alldata$prepost[alldata$id==1] <- "PRE"
alldata$prepost[alldata$id==2] <- "POST"

# Tim deltas
nums <- unlist(lapply(alldata, is.numeric)) 
deltas <- alldata[alldata$prepost == "POST",nums] - alldata[alldata$prepost == "PRE",nums]
deltas$ANALYTICID <- alldata$ANALYTICID[alldata$prepost == "PRE"]
# merge datasets
deltas <- merge(deltas,rand,by=c("ANALYTICID"))

# labels
label(deltas$IFN.g)="Change in IFN-g"
label(deltas$IL.1b)="Change in IL.1b"
label(deltas$IL.12.IL.23p40)="Change in IL.12.IL.23p40"
label(deltas$IL.17A)="Change in IL.17A"
label(deltas$IL.18)="Change in IL.18"
label(deltas$IL.6)="Change in IL.6"
label(deltas$IL.8)="Change in IL.8"
label(deltas$MCP.1..Human.)="Change in MCP.1..Human."
label(deltas$TNF.a)="Change in TNF.a"
label(deltas$il17F)="Change in IL17F"

# compare deltas in the treatment groups
table <- final_table(deltas,c("IFN.g","IL.1b","IL.12.IL.23p40","IL.17A","IL.18","IL.6","IL.8","MCP.1..Human.","TNF.a","il17F"),
                     group=as.factor(deltas$Treatment.Group),margin=2,single=0,ron=2)

N <- nrow(deltas)

# adjust for baseline
base <- alldata[alldata$prepost=="PRE",]
base <- select(base,-c("COLLECTIONDT","Treatment.Group","id","prepost"))
setnames(base,"IFN.g","IFN.g_base")
setnames(base,"IL.1b","IL.1b_base")
setnames(base,"IL.12.IL.23p40","IL.12.IL.23p40_base")
setnames(base,"IL.17A","IL.17A_base")
setnames(base,"IL.18","IL.18_base")
setnames(base,"IL.6","IL.6_base")
setnames(base,"IL.8","IL.8_base")
setnames(base,"il17F","il17F_base")
setnames(base,"MCP.1..Human.","MCP.1..Human._base")
setnames(base,"TNF.a","TNF.a_base")
deltas <- merge(deltas,base,by="ANALYTICID")
mod1 <- lm(data=deltas,IFN.g ~ as.factor(Treatment.Group) + IFN.g_base)
mod2 <- lm(data=deltas,IL.1b ~ as.factor(Treatment.Group) + IL.1b_base)
mod3 <- lm(data=deltas,IL.12.IL.23p40 ~ as.factor(Treatment.Group) + IL.12.IL.23p40_base)
mod4 <- lm(data=deltas, IL.17A ~ as.factor(Treatment.Group) + IL.17A_base)
mod5 <- lm(data=deltas, IL.18~ as.factor(Treatment.Group) + IL.18_base)
mod6 <- lm(data=deltas, IL.6 ~ as.factor(Treatment.Group) + IL.6_base)
mod7 <- lm(data=deltas, IL.8 ~ as.factor(Treatment.Group) + IL.8_base)
mod8 <- lm(data=deltas, il17F ~ as.factor(Treatment.Group) + il17F_base)
mod9 <- lm(data=deltas, MCP.1..Human. ~ as.factor(Treatment.Group) + MCP.1..Human._base)
mod10 <- lm(data=deltas, TNF.a ~ as.factor(Treatment.Group) + TNF.a_base)


```

# Background

The purpose of this analysis is to compare the change in CVD inflammatory markers in participants on metformin to participants on placebo in the T1D Exchange metformin study.

# Methods

The change in each inflammatory marker was calculated as the post-treatment value minus the pre-treatment value.  The distributions of these changes were examined prior to analysis.  For normally distributed variables, the mean and standard deviation were calculated, and groups were compared using t-tests.  For variables that were not normally distributed, the median, 25th and 75th percentiles were calculated, and the groups were compared using the Mann Whitney test.

Linear models adjusted for the baseline value of the marker were also used to compare changes in the two groups.

# Results

There were `r N` subjects included in this analysis.  Changes in inflammatory CVD markers are shown in Table 1.

```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(table,caption="Table 1.  Change in CVD inflammatory markers in metformin and placebo groups.  Descriptive statistics are mean ± standard deviation for normally distributed variables, or median (25th percentile, 75th percentile for variables that were not normally distributed.  Normally distributed variables were compared using t-tests; variables that were not normally distributed were compared using the Mann Whitney test.")
```

Tables 2-11 show the results of the linear models adjusted for the baseline values of the markers.

```{r, results='asis',tidy=TRUE, echo=FALSE}
print(kable(summary(mod1)$coefficients,caption="Table 2.  Change in IFN.g adjusted for baseline."))

```
 <br/>
 
```{r, results='asis',tidy=TRUE, echo=FALSE}
print(kable(summary(mod2)$coefficients,caption="Table 3.  Change in IL.1b adjusted for baseline."))
```
 <br/>
 
``` {r, results='asis',tidy=TRUE, echo=FALSE}
kable(summary(mod3)$coefficients,caption="Table 4.  Change in IL.12.IL.23p40 adjusted for baseline.")
```
 <br/>
 
```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(summary(mod4)$coefficients,caption="Table 5.  Change in IL.17A adjusted for baseline.")
```
 <br/>
 
```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(summary(mod5)$coefficients,caption="Table 6.  Change in IL.18 adjusted for baseline.")
```
 <br/>
 
```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(summary(mod6)$coefficients,caption="Table 7.  Change in IL.6 adjusted for baseline.")
```
 <br/>
 
```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(summary(mod7)$coefficients,caption="Table 8.  Change in IL.8 adjusted for baseline.")
```
 <br/>
 
```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(summary(mod8)$coefficients,caption="Table 9.  Change in il17F adjusted for baseline.")
```
 <br/>
 
```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(summary(mod9)$coefficients,caption="Table 10.  Change in MCP.1..Human. adjusted for baseline.")
```
 <br/>
 
```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(summary(mod10)$coefficients,caption="Table 11.  Change in TNF.a adjusted for baseline.")

```




