---
title: "Pima Trajectory Analysis"
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
---

```{r setup, include=FALSE}
library(readxl)
library(DT)
library(tidyverse)
library(plotly)
knitr::opts_chunk$set(echo = FALSE)
home_dir = ifelse(.Platform$OS.type != "unix",
                  "C:/Users/timvigers/Dropbox/Work/Pima/",
                  "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/")
knitr::opts_knit$set(root.dir = home_dir)
```

```{r import}
combined = read_excel("Combined Kidney Data.xls")
combined$NIH = factor(combined$NIH)
combined$PROTOCOL = factor(combined$PROTOCOL)
```

# Check that DOB is the same across visits for everyone

```{r}
dob = combined %>% group_by(NIH) %>% summarize(n_dobs = length(unique(DOB)))
multiples = unique(dob$NIH[which(dob$n_dobs != 1)])
if(length(multiples) > 0){
  combined %>% filter(NIH %in% multiples) %>% group_by(NIH) %>% 
    count(DOB,PROTOCOL) %>% select(-n) %>% 
    datatable(.,extensions = "Buttons",
              options = list(dom = 'lBfrtip',
                             buttons = c('copy', 'csv', 'excel')))
}
```

There were `r length(multiples)` participants with multiple DOB values.

# Check that date of diagnosis is the same across visits for everyone

```{r}
dx = combined %>% group_by(NIH) %>% summarize(n_dx_dates = length(unique(DATE_DX)))
multiples = unique(dx$NIH[which(dx$n_dx_dates != 1)])
if(length(multiples) > 0){
  combined %>% filter(NIH %in% multiples) %>% group_by(NIH) %>% 
    count(DATE_DX,PROTOCOL) %>% select(-n) %>% 
    datatable(.,extensions = "Buttons",
              options = list(dom = 'lBfrtip',
                             buttons = c('copy', 'csv', 'excel')))
}
```

There were `r length(multiples)` participants with multiple date of diagnosis values.

# Check that sex is the same across visits for everyone

```{r}
sex = combined %>% group_by(NIH) %>% summarize(n_sex = length(unique(SEX)))
multiples = unique(sex$NIH[which(sex$n_sex != 1)])
if(length(multiples) > 0){
  combined %>% filter(NIH %in% multiples) %>% group_by(NIH) %>% 
    count(SEX,PROTOCOL) %>% select(-n) %>% 
    datatable(.,extensions = "Buttons",
              options = list(dom = 'lBfrtip',
                             buttons = c('copy', 'csv', 'excel')))
}
```

There were `r length(multiples)` participants with multiple sex values.

# Check that height is mostly the same across visits for everyone

```{r}
ht = combined %>% group_by(NIH) %>% summarize(n_ht = length(unique(HT_CM)))
multiples = unique(ht$NIH[which(ht$n_ht != 1)])
if(length(multiples) > 0){
  combined %>% filter(NIH %in% multiples) %>% group_by(NIH) %>% 
    count(HT_CM,PROTOCOL) %>% select(-n) %>% 
    datatable(.,extensions = "Buttons",
              options = list(dom = 'lBfrtip',
                             buttons = c('copy', 'csv', 'excel')))
}
```

There were `r length(multiples)` participants with multiple height values.

# Check for big fluctuations in WT_KG, SBP, DBP, and PULSE 

We calculated change in each variable and change in each variable per day between all visits for each participant. The tables show the 30 visits with highest absolute change. Information about the previous visit is in columns with the "prev_" prefix. 

```{r functions}
# Plots
traj_plots = function(df,var){
  plots = lapply(levels(df$PROTOCOL), function(p){
    plot = ggplot(df[df$PROTOCOL == p,],
                  aes(x = TESTINT,y = !!sym(var),group = NIH)) + 
      geom_line(alpha = 0.1) + ggtitle(p) +
      theme_bw() + theme(legend.position = "none")
    plot
  })
  return(plots)
}
# Check differences between visits
traj_check = function(df,var,row_limit = 30){
  # Calculate differences
  d = df %>% arrange(NIH,DATE) %>% group_by(NIH) %>% 
    mutate(diff = !!sym(var) - lag(!!sym(var)),
           date_diff = as.numeric(DATE - lag(DATE),units = "days"),
           diff_per_day = diff / date_diff,
           prev_date = lag(DATE),
           prev = lag(!!sym(var)),
           prev_protocol = lag(PROTOCOL)) %>%
    select(NIH,PROTOCOL,prev_protocol,DATE,prev_date,date_diff,
           !!sym(var),prev,diff,diff_per_day) %>% ungroup() %>%
    slice_max(order_by = abs(diff),n = row_limit)
  datatable(d,extensions = "Buttons",
            options = list(dom = 'lBfrtip',
                           buttons = c('copy', 'csv', 'excel'),
                           pageLength = row_limit)) %>%
    formatSignif(c('diff','diff_per_day'),digits = 2)
}
```

## WT_KG

### Plots

```{r warning=FALSE}
# Plot by protocol (facet wrap is too small)
plots = traj_plots(combined,var = "WT_KG")
ggplotly(plots[[1]])
ggplotly(plots[[2]])
ggplotly(plots[[3]])
ggplotly(plots[[4]])
ggplotly(plots[[5]])
```

### Table

```{r}
traj_check(combined,var = "WT_KG")
```

## SBP

### Plots

```{r warning=FALSE}
# Plot by protocol (facet wrap is too small)
plots = traj_plots(combined,var = "SBP")
ggplotly(plots[[1]])
ggplotly(plots[[2]])
ggplotly(plots[[3]])
ggplotly(plots[[4]])
ggplotly(plots[[5]])
```

### Table

```{r}
traj_check(combined,var = "SBP")
```

## DBP

### Plots

```{r warning=FALSE}
# Plot by protocol (facet wrap is too small)
plots = traj_plots(combined,var = "DBP")
ggplotly(plots[[1]])
ggplotly(plots[[2]])
ggplotly(plots[[3]])
ggplotly(plots[[4]])
ggplotly(plots[[5]])
```

### Table

```{r}
traj_check(combined,var = "DBP")
```

## PULSE

### Plots

```{r warning=FALSE}
# Plot by protocol (facet wrap is too small)
plots = traj_plots(combined,var = "PULSE")
ggplotly(plots[[1]])
ggplotly(plots[[2]])
ggplotly(plots[[3]])
ggplotly(plots[[4]])
ggplotly(plots[[5]])
```

### Table

```{r}
traj_check(combined,var = "PULSE")
```

# Check for big fluctuations in SMAC variables

## SMAC_GLU

```{r}
traj_check(combined,var = "SMAC_GLU")
```

## SMAC_BUN

```{r}
traj_check(combined,var = "SMAC_BUN")
```

## SMAC_SCR

```{r}
traj_check(combined,var = "SMAC_SCR")
```

## SMAC_NA

```{r}
traj_check(combined,var = "SMAC_NA")
```

## SMAC_K

```{r}
traj_check(combined,var = "SMAC_K")
```

## SMAC_CL

```{r}
traj_check(combined,var = "SMAC_CL")
```

## SMAC_CA

```{r}
traj_check(combined,var = "SMAC_CA")
```

## SMAC_PHOS

```{r}
traj_check(combined,var = "SMAC_PHOS")
```

## SMAC_TP

```{r}
traj_check(combined,var = "SMAC_TP")
```

## SMAC_ALB

```{r}
traj_check(combined,var = "SMAC_ALB")
```

## SMAC_C02

```{r}
traj_check(combined,var = "SMAC_C02")
```

## SMAC_CHOL

```{r}
traj_check(combined,var = "SMAC_CHOL")
```

## SMAC_TRI

```{r}
traj_check(combined,var = "SMAC_TRI")
```

## SMAC_HDL

```{r}
traj_check(combined,var = "SMAC_HDL")
```

## SMAC_LDL

```{r}
traj_check(combined,var = "SMAC_LDL")
```

## SMAC_TBIL

```{r}
traj_check(combined,var = "SMAC_TBIL")
```

## SMAC_ALK

```{r}
traj_check(combined,var = "SMAC_ALK")
```

## SMAC_AST

```{r}
traj_check(combined,var = "SMAC_AST")
```