---
title: "Pima Trajectory Analysis"
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
---

```{r setup, include=FALSE}
library(readxl)
library(DT)
library(tidyverse)
library(plotly)
knitr::opts_chunk$set(echo = FALSE)
home_dir = ifelse(.Platform$OS.type != "unix",
                  "C:/Users/timvigers/Dropbox/Work/Pima/",
                  "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects/")
knitr::opts_knit$set(root.dir = home_dir)
```

```{r import}
combined = read_excel("Combined Kidney Data.xls")
combined$NIH = factor(combined$NIH)
combined$PROTOCOL = factor(combined$PROTOCOL)
```

# Check that DOB is the same across visits for everyone

```{r}
dob = combined %>% group_by(NIH) %>% summarize(n_dobs = length(unique(DOB)))
multiples = unique(dob$NIH[which(dob$n_dobs != 1)])
if(length(multiples) > 0){
  combined %>% filter(NIH %in% multiples) %>% group_by(NIH) %>% 
    count(DOB,PROTOCOL) %>% select(-n) %>% 
    datatable(.,extensions = "Buttons",
              options = list(dom = 'lBfrtip',
                             buttons = c('copy', 'csv', 'excel')))
}
```

There were `r length(multiples)` participants with multiple DOB values.

# Check that date of diagnosis is the same across visits for everyone

```{r}
dx = combined %>% group_by(NIH) %>% summarize(n_dx_dates = length(unique(DATE_DX)))
multiples = unique(dx$NIH[which(dx$n_dx_dates != 1)])
if(length(multiples) > 0){
  combined %>% filter(NIH %in% multiples) %>% group_by(NIH) %>% 
    count(DATE_DX,PROTOCOL) %>% select(-n) %>% 
    datatable(.,extensions = "Buttons",
              options = list(dom = 'lBfrtip',
                             buttons = c('copy', 'csv', 'excel')))
}
```

There were `r length(multiples)` participants with multiple date of diagnosis values.

# Check that sex is the same across visits for everyone

```{r}
sex = combined %>% group_by(NIH) %>% summarize(n_sex = length(unique(SEX)))
multiples = unique(sex$NIH[which(sex$n_sex != 1)])
if(length(multiples) > 0){
  combined %>% filter(NIH %in% multiples) %>% group_by(NIH) %>% 
    count(SEX,PROTOCOL) %>% select(-n) %>% 
    datatable(.,extensions = "Buttons",
              options = list(dom = 'lBfrtip',
                             buttons = c('copy', 'csv', 'excel')))
}
```

There were `r length(multiples)` participants with multiple sex values.

# Check that height is mostly the same across visits for everyone

```{r}
ht = combined %>% group_by(NIH) %>% summarize(n_ht = length(unique(HT_CM)))
multiples = unique(ht$NIH[which(ht$n_ht != 1)])
if(length(multiples) > 0){
  combined %>% filter(NIH %in% multiples) %>% group_by(NIH) %>% 
    count(HT_CM,PROTOCOL) %>% select(-n) %>% 
    datatable(.,extensions = "Buttons",
              options = list(dom = 'lBfrtip',
                             buttons = c('copy', 'csv', 'excel')))
}
```

There were `r length(multiples)` participants with multiple height values.

# Check for big fluctuations in WT_KG, SBP, DBP, and PULSE 

## WT_KG

```{r warning=FALSE}
# Plot by protocol (facet wrap is too small)
plots = lapply(levels(combined$PROTOCOL), function(p){
  wt_plot = ggplot(combined[combined$PROTOCOL == p,],
                   aes(x = TESTINT,y = WT_KG,group = NIH)) + 
    geom_line(alpha = 0.1) + ggtitle(p) +
    theme_bw() + theme(legend.position = "none")
  wt_plot
})
ggplotly(plots[[1]])
ggplotly(plots[[2]])
ggplotly(plots[[3]])
ggplotly(plots[[4]])
ggplotly(plots[[5]])
```

```{r}
# Check maximum difference between visits by person
wt = combined %>% arrange(NIH,DATE) %>% group_by(NIH) %>% 
  mutate(wt_diff = WT_KG - lag(WT_KG),
         date_diff = as.numeric(DATE - lag(DATE),units = "days"),
         wt_per_day = wt_diff / date_diff,
         prev_date = lag(DATE),
         prev_wt = lag(WT_KG),
         prev_protocol = lag(PROTOCOL)) %>%
  filter(abs(wt_per_day) > 0.1) %>%
  select(NIH,PROTOCOL,prev_protocol,DATE,prev_date,date_diff,
         WT_KG,prev_wt,wt_diff,wt_per_day) %>%
  arrange(abs(wt_per_day))
datatable(wt,extensions = "Buttons",
          options = list(dom = 'lBfrtip',
                         buttons = c('copy', 'csv', 'excel'))) %>%
  formatSignif(c('wt_diff','wt_per_day'),digits = 2)
```

We calculated change in weight and change in weight per day between all visits for each participant. The above table shows visits with absolute change in weight $\geq$ 0.1 kg/day. Information about the previous visit is in columns with the "prev_" prefix. 
