---
title: "MISS Study - Correlations of Microbiome With Insulin Sensitivity"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
# Load libraries and functions
library(microbiome)
library(vegan)
library(psych)
library(stringr)
library(stringi)
library(dplyr)
library(knitr)
library(propr)
library(phyloseq)

source('C:/Users/pylell/Documents/GitHub/Tim-and-Laura/tim_R_functions.R')
# Check OS.
pathstart <- os.check()
```

```{r echo=FALSE,include=FALSE}
# Data management
# Read microbiome data.
# data are relative abundance, in proportions, not percentages
filename <- "H:/Endocrinology/Kelsey/Kelsey microbiome/Raw data/MISS_Alltaxa_OTU_12_17_18.csv"
# read relative abundance data 
ra <- read.csv(filename,row.names = 1)
ra <- ra[,-1]

# Make OTU table.
otumat <- ra
otumat <- as.matrix(otumat)
OTU <- otu_table(otumat,taxa_are_rows = T)
# Make taxa table
taxmat <- str_split_fixed(rownames(ra),pattern = "/", n = 7)
rownames(taxmat) <- row.names(ra)
colnames(taxmat) <-   c("Domain", "Phylum", "Class", "Order", "Family", "Genus", "Species")
taxmat[taxmat == ""] <- "Unknown"
TAX <- tax_table(taxmat)

# Read in sample data.
source("C:/Users/pylell/Documents/GitHub/Kelsey-MISS/MISSENROLLEDPTS-MinimalModelResults_R_2019-02-15_1812.r")
sammat <- data
sammat$miss_id <- gsub("-","",sammat$miss_id)
sammat$miss_id <- gsub("_","",sammat$miss_id)
sammat$miss_id <- paste0(stri_sub(sammat$miss_id,from=1,to=4),stri_sub(sammat$miss_id,from=6,to=7))
rownames(sammat) <- sammat$miss_id
SAM <- sample_data(sammat)
# Combine into physeq object
pseq <- phyloseq(OTU, TAX, SAM)
```

```{r echo=FALSE,include=FALSE}
# Relative abundance.
relative <- pseq
otus <- as.data.frame(otu_table(relative))
# Transpose.
otus <- as.data.frame(t(otus))
otus$subject.id <- rownames(otus)
# Free testosterone
temp <- c("insulin_sensitivity","insulin_secretion_mm","disposition_index")
si <- sammat[temp]
si$subject.id <- rownames(si)
final <- merge(otus,si,by="subject.id",all.x = TRUE,all.y=TRUE)

# remove people who are missing all SI variables
final <- final[!is.na(final$disposition_index) | !is.na(final$insulin_secretion_mm) | !is.na(final$insulin_sensitivity),]

# Move columns
final <- final %>% select("subject.id",everything())

# delete columns with all 0 counts
final <- final[,-(which(colSums(final[,-1])==0))]
N <- nrow(final)

# export final dataset for Megan
write.csv(final,"H:\\Endocrinology\\Kelsey\\Kelsey microbiome\\Clean data\\export.csv")

# Correlations with insulin sensitivity
corrs <- corr.test(final[,2:(ncol(final)-3)],final$insulin_sensitivity,method = "spearman",adjust = "fdr")
# Table
corr.table <- as.data.frame(cbind(corrs$r,corrs$n,corrs$p))
corr.table <- round(corr.table,3)
colnames(corr.table) <- c("R","n","P Value")
corr.table$Taxa <- rownames(corr.table)
corr.table <- select(corr.table,"Taxa",everything())
corr.table <- corr.table[order(corr.table$`P Value`),]

# read in the phylum level data
phylum <- read.csv('H:/Endocrinology/Kelsey/Kelsey microbiome/Raw data/MISS_Phylum_OTU_12_18_18.csv',row.names = 1)
phylum <- phylum[,-1]
# Make OTU table.
phylummat <- phylum
phylummat <- as.matrix(phylummat)
PHY <- otu_table(phylummat,taxa_are_rows = T)
# Transpose.
phylummat <- as.data.frame(t(phylummat))
phylummat$subject.id <- rownames(phylummat)
phylummat$fb <- phylummat$`Bacteria/Firmicutes` / phylummat$`Bacteria/Bacteroidetes`
sammat$subject.id <- sammat$miss_id
phylummat <- merge(phylummat,sammat,by="subject.id",all.x = TRUE,all.y = FALSE)
# remove people who are missing all SI variables
phylummat <- phylummat[!is.na(phylummat$disposition_index) | !is.na(phylummat$insulin_secretion_mm) | !is.na(phylummat$insulin_sensitivity),]
# Move columns
phylummat <- phylummat %>% select("subject.id",everything())

# Correlations with insulin sensitivity
corrs_phy <- corr.test(phylummat[c("insulin_sensitivity","insulin_secretion_mm","disposition_index")],phylummat$fb,method = "spearman",adjust = "none")
corr.table.phy <- as.data.frame(cbind(corrs_phy$r,corrs_phy$n,corrs_phy$p))
corr.table.phy <- round(corr.table.phy,3)
colnames(corr.table.phy) <- c("R","n","P Value")
corr.table.phy$Variable <- rownames(corr.table.phy)

```

# Background

The purpose of this analysis is to examine correlations between relative abundance of microbiota in the MISS study samples with insulin sensitivity, insulin secretion, and disposition index.

# Methods

The following data cleaning and manipulation steps were performed:

1) Patients with missing data from the minimal model were excluded.

Spearman's correlation was used to examine correlation between microbiota relative abundance and insulin sensitivity and secretion variables.  Both unadjusted p-values and p-values with FDR adjustment for multiple testing are reported.  All analyses were performed using R version 3.5.2.

# Results

There were `r N` subjects included in this analysis.

Table 1: Relative Abundance Spearman Correlation with Insulin Sensitivity, FDR Adjusted P-Values
```{r echo=FALSE}
kable(corr.table,row.names = F)
```

```{r echo=FALSE,include=FALSE}
# Correlations with insulin secretion
corrs <- corr.test(final[,2:(ncol(final)-3)],final$insulin_secretion_mm,method = "spearman",adjust = "fdr")
# Table
corr.table <- as.data.frame(cbind(corrs$r,corrs$n,corrs$p))
corr.table <- round(corr.table,3)
colnames(corr.table) <- c("R","n","P Value")
corr.table$Taxa <- rownames(corr.table)
corr.table <- select(corr.table,"Taxa",everything())
corr.table <- corr.table[order(corr.table$`P Value`),]
```

Table 2: Relative Abundance Spearman Correlation with Insulin Secretion, FDR Adjusted P-Values
```{r echo=FALSE}
kable(corr.table,row.names = F)
```

```{r echo=FALSE,include=FALSE}
# Correlations with disposition index
corrs <- corr.test(final[,2:(ncol(final)-3)],final$disposition_index,method = "spearman",adjust = "fdr")
# Table
corr.table <- as.data.frame(cbind(corrs$r,corrs$n,corrs$p))
corr.table <- round(corr.table,3)
colnames(corr.table) <- c("R","n","P Value")
corr.table$Taxa <- rownames(corr.table)
corr.table <- select(corr.table,"Taxa",everything())
corr.table <- corr.table[order(corr.table$`P Value`),]
```

Table 3: Relative Abundance Spearman Correlation with Disposition Index, FDR Adjusted P-Values
```{r echo=FALSE}
kable(corr.table,row.names = F)
```


```{r echo=FALSE,include=FALSE}
# Correlations with insulin sensitivity
corrs <- corr.test(final[,2:(ncol(final)-3)],final$insulin_sensitivity,method = "spearman",adjust = "none")
# Table
corr.table <- as.data.frame(cbind(corrs$r,corrs$n,corrs$p))
corr.table <- round(corr.table,3)
colnames(corr.table) <- c("R","n","P Value")
corr.table$Taxa <- rownames(corr.table)
corr.table <- select(corr.table,"Taxa",everything())
corr.table <- corr.table[order(corr.table$`P Value`),]
```

Table 4: Relative Abundance Spearman Correlation with Insulin Sensitivity, Unadjusted P-Values
```{r echo=FALSE}
kable(corr.table,row.names = F)
```

```{r echo=FALSE,include=FALSE}
# Correlations with insulin secretion
corrs <- corr.test(final[,2:(ncol(final)-3)],final$insulin_secretion_mm,method = "spearman",adjust = "none")
# Table
corr.table <- as.data.frame(cbind(corrs$r,corrs$n,corrs$p))
corr.table <- round(corr.table,3)
colnames(corr.table) <- c("R","n","P Value")
corr.table$Taxa <- rownames(corr.table)
corr.table <- select(corr.table,"Taxa",everything())
corr.table <- corr.table[order(corr.table$`P Value`),]
```

Table 5: Relative Abundance Spearman Correlation with Insulin Secretion, Unadjusted P-Values
```{r echo=FALSE}
kable(corr.table,row.names = F)
```

```{r echo=FALSE,include=FALSE}
# Correlations with disposition index
corrs <- corr.test(final[,2:(ncol(final)-3)],final$disposition_index,method = "spearman",adjust = "none")
# Table
corr.table <- as.data.frame(cbind(corrs$r,corrs$n,corrs$p))
corr.table <- round(corr.table,3)
colnames(corr.table) <- c("R","n","P Value")
corr.table$Taxa <- rownames(corr.table)
corr.table <- select(corr.table,"Taxa",everything())
corr.table <- corr.table[order(corr.table$`P Value`),]
```

Table 6: Relative Abundance Spearman Correlation with Disposition Index, Unadjusted P-Values
```{r echo=FALSE}
kable(corr.table,row.names = F)
```

Table 7: Correlation of Insulin Sensitivity, Insulin Secretion, and DI with F:B ratio, Unadjusted P-Values
```{r echo=FALSE}
kable(corr.table.phy,row.names = F)
```
