---
title: "Multi-omics Integration Analysis with Brain Biomarkers & Diabetes Duration"
author: "Hailey Hampson"
date: "2024-04-02"
output: html_document
---

#1. Set Up R Markdown 
```{r setup, include=FALSE, echo=FALSE}
#Load Libraries 
source("Libraries.R")

#Load Functions
source("Functions.R")

#Directories
dir.dat <- c("/Volumes/Peds Endo/Petter Bjornstad")
dir.results <- c("/Users/hhampson/Documents/UW/1_Ongoing Projects/Brain_Biomarkers/Results")
```

#2. Load & Clean Data
```{r load clean data, include=FALSE, echo=FALSE}
#Load Data
dat_harm <- read.csv(fs::path(dir.dat,"Data Harmonization","Data Clean","soma_harmonized_dataset.csv"))
# dat_harm2 <- read.csv(fs::path(dir.dat,"Data Harmonization","Data Clean","harmonized_dataset.csv"))

#Filter to CROCODILE & Diabetes - heatmap for beta 
dat_harm <- dat_harm %>% 
  filter(study=="CROCODILE")  
#& penguine plasma, and 
#renal heritage = serum

#Metabolomics & proteomics
#Select metabolomics data
met.names <- c("record_id",colnames(dat_harm)[which(startsWith(colnames(dat_harm),"c0") | startsWith(colnames(dat_harm),"c1"))])
met <- dat_harm %>% 
  filter(procedure == "metabolomics_blood") %>% #Filter to metabolomics measurement
  dplyr::select(all_of(met.names)) #Select only metabolite features
met <- met %>% #Remove metabolites with missing data
  dplyr::select(where(~ !all(is.na(.))))
metabolites <- colnames(met)[-1]
#Select proteomics data
prot.names <- c("record_id",colnames(dat_harm)[which(grepl("seq.",colnames(dat_harm)))])
prot <- dat_harm %>% 
  filter(procedure == "metabolomics_blood") %>% #Proteomics available for all measurements, select one arbitrarily)
  dplyr::select(all_of(prot.names)) #Select only protein features
#CRC-53 & CRC-60 missing proteomics data
proteins <- prot.names[-1]
rm(prot.names)
rm(met.names)

#Brain biomarkers
brain <- dat_harm %>% 
  filter(procedure=="neuro_markers") %>% 
  dplyr::select(record_id,ab40_avg_conc,ab42_avg_conc,
                tau_avg_conc,nfl_avg_conc,gfap_avg_conc,
                ptau_181_avg_conc,ptau_217_avg_conc) %>% 
  filter(!is.na(ab40_avg_conc))

#Select covariates
cov <- dat_harm %>% 
  filter(procedure=="physical_exam") %>%
  dplyr::select(record_id,procedure,diabetes_duration,group,age,sex,bmi,race_ethnicity) %>% 
  filter(record_id %in% brain$record_id) %>%  #filter to individuals with proteomics & metabolomics & brain biomarkers
  mutate(diabetes_duration=ifelse(group=="Lean Control",0,diabetes_duration)) %>% 
  dplyr::select(-procedure)
  
#Combine datasets
dat <- tidylog::left_join(cov,brain,by="record_id")
dat <- tidylog::left_join(dat,prot,by="record_id")
dat <- tidylog::left_join(dat,met,by="record_id")
  
rm(brain,cov,dat_harm,met,prot)

#Make Covariates factors
dat <- dat %>% 
  mutate(diabetes_status=as.factor(ifelse(group=="Lean Control",0,1))) %>% #0 = Lean Control, 1 = Type 1
  mutate(sex=as.factor(sex),
         race_ethnicity=as.factor(race_ethnicity))
```

# 3. Annotate Feature Tables 
```{r annotation, include=FALSE,echo=FALSE}
#Match Confirmed Features to feature tables: match on mz and rt (mz rounded to 5 digits, rt to 1 digit)
## C18 Negative Features - 117 confirmed features ----
c18neg_annotated <- tidylog::left_join(c18neg_conf_ftrs,c18neg_ftrs,by=c("mz","rt"))

## C18 Positive Features - 115 confirmed features ----
c18pos_annotated <- tidylog::left_join(c18pos_conf_ftrs,c18pos_ftrs,by=c("mz","rt"))

## Hilic Negative Features - 125 confirmed features ----
hilicneg_annotated <- tidylog::left_join(hilicneg_conf_ftrs,hilicneg_ftrs,by=c("mz","rt"))

## Hilic Positive Features - 97 confirmed features ----
hilicpos_annotated <- tidylog::left_join(hilicpos_conf_ftrs,hilicpos_ftrs,by=c("mz","rt"))

#Combine Annoted Features 
annotated_ftrs <- rbind(c18neg_annotated,c18pos_annotated)
annotated_ftrs <- rbind(annotated_ftrs,hilicneg_annotated)
annotated_ftrs <- rbind(annotated_ftrs,hilicpos_annotated)

#Duplicate metabolites - select based on smaller delta mz and rt
dups <- annotated_ftrs$name[which(duplicated(annotated_ftrs$name))]
dup_metab <- annotated_ftrs %>% 
  filter(name %in% dups)

#Select smaller delta mz of duplicates
# dup_metab_test2 <- dup_metab %>%
#   filter((name=="5-Hydroxy-L-Tryptophan" & delta_ppm==0.35) |
#            (name=="D-Sedoheptulose" & delta_ppm==0.61) |
#            (name=="Hypoxanthine" & delta_ppm==0.13) |
#            (name=="Inosine" & delta_ppm==0.32) |
#            (name=="Kynurenine" & delta_ppm==0.03) |
#            (name=="Mevalolactone" & delta_ppm==0.35) |
#            (name=="N-Acetyltryptophan" & delta_ppm==0.21) |
#            (name=="NNN-Trimethyllysine" & delta_ppm==0.6) |
#            (name=="Palmitoylcarnitine" & delta_ppm==0.11) |
#            (name=="Tryptophan" & delta_ppm==0.12))
dup_metab <- dup_metab %>% 
  group_by(name) %>% 
  slice_min(delta_ppm) %>% 
  ungroup()

#Remove duplicates from annotated features
annotated_ftrs_filt <- annotated_ftrs %>% 
  filter(!name %in% dups)

#Merge back in duplicates 
annotated_ftrs_filt <- rbind(annotated_ftrs_filt,dup_metab)

metabolites <- annotated_ftrs_filt %>%
  dplyr::select(mz_rt,name,mz,rt)

# #Merge confirmed features with feature tables
c18neg_ftrs <- data.frame(ftrs$c18neg)
c18pos_ftrs <- data.frame(ftrs$c18pos)
hilicneg_ftrs <- data.frame(ftrs$hilicneg)
hilicpos_ftrs <- data.frame(ftrs$hilicpos)
colnames(c18pos_ftrs) <- str_remove(colnames(c18pos_ftrs),"X")
colnames(c18neg_ftrs) <- str_remove(colnames(c18neg_ftrs),"X")
colnames(hilicpos_ftrs) <- str_remove(colnames(hilicpos_ftrs),"X")
colnames(hilicneg_ftrs) <- str_remove(colnames(hilicneg_ftrs),"X")

#Combine
metab.data <- tidylog::full_join(c18neg_ftrs,c18pos_ftrs,by="id")
metab.data <- tidylog::full_join(metab.data,hilicneg_ftrs,by="id")
metab.data <- tidylog::full_join(metab.data,hilicpos_ftrs,by="id")

#Filter to the 454 annotated metabolites
confirmed.metabolites <- metabolites$mz_rt
metab.data <- metab.data %>% 
  dplyr::select(id,all_of(confirmed.metabolites))

#Create easier metabolomics names for analysis
metab.data <- metab.data %>% 
  mutate(across(everything(),~as.vector(.)))
metabolites$key <- paste0("metabolite",seq(1:length(metabolites$mz_rt)))        

colnames(metab.data) <- c("id",metabolites$key)
```

#4. Load & Clean Meta data
##a. Meta Data
```{r Metadata, include=FALSE, echo=FALSE}
#Load PFAS & covariates data metaair 
PFAS <- readRDS(fs::path(dir.dat %>% dirname(),"CHS MetaAir MetaChem Cleaned Redcap and Exposures Outcome Data long V3.rds")) 
PFAS <- PFAS %>%   
  filter(study=="MetaAir") %>% 
  dplyr::select(id,ma_pfda:ma_pfpes,ma_age,ma_sex,ma_bmi,ma_race_eth,ma_parental_edu) %>% 
  mutate(id=as.numeric(id))

#Load eGFR metachem 
eGFR <- read.csv(fs::path(dir.dat %>% dirname(),"archive_versions","chs_complete_data.csv"))
eGFR <- eGFR %>% 
  dplyr::select(id,mc_creatinine) 
# filter(!is.na(mc_creatinine))

#Load baseline creatinine-like protein metaair
cst3 <- read.csv(fs::path(dir.dat %>% dirname(),"archive_versions","chs_complete_data.csv")) %>%  
  dplyr::select(id,ma_CST3,mc_CST3)

#Load cystatin-C 
cystatin_c <- readRDS("/Users/hhampson/USC Lab Dropbox/Chatzi Active Projects/MOVED Chatzi Projects (Active)/Env Chem SOL-CHS/Analysis/1_analysis_ready_data/CHS MetaAir MetaChem Cleaned Redcap and Exposures Outcome Data long V3.rds") %>%
  filter(study=="MetaChem") %>%
  dplyr::select(id,cystatin_c) %>%
  mutate(id=as.numeric(id))

#Load baseline fiber & energy
nutrients <- readRDS("/Users/hhampson/USC Lab Dropbox/Chatzi Active Projects/MOVED Chatzi Projects (Active)/Env Chem SOL-CHS/Analysis/0_data_processing/SOL CHS Diet Data Processing/CHS/2_cleaned_data/temp/Nutrients_Data_Formatted.rds") %>%
  dplyr::select(id,m_energy_kcal,m_total_dietary_fiber_g)

#Load microbiome data
#Reduced Microbiome Features
# micro <- readRDS(fs::path(dir.dat %>% dirname() %>% dirname(),"2_ongoing/CHS PFAS Microbiome (Hailey)/BHRM_microbiome/Formatted Data/SignificantGenusPFASData.rds")) %>%
#   mutate(id = as.numeric(id))
# Full Mirobiome Features
micro <- readRDS(fs::path(dir.dat %>% dirname() %>% dirname(),"2_ongoing/CHS PFAS Microbiome (Hailey)/BHRM_microbiome/Formatted Data/GenusPFASData.rds")) %>%
  mutate(id = as.numeric(id))

colnames(micro) <- str_replace(colnames(micro),".*g__","")
colnames(micro) <- str_replace(colnames(micro), "UCG_002","Oscillospiraceae_g__UCG_002")
colnames(micro) <- str_replace(colnames(micro), "UCG_005","Oscillospiraceae_g__UCG_005")
colnames(micro) <- str_replace(colnames(micro), "unclassified887","Lachnospiraceae_g__unclassified887")
colnames(micro) <- str_replace(colnames(micro), "unclassified498","Lachnospiraceae_g__unclassified498")
colnames(micro) <- str_replace(colnames(micro), "GCA_900066575","Lachnospiraceae_g__GCA_900066575")

#Calculate Relative Abundance Data
rownames(micro) <- micro$id
micro <- micro %>%
  dplyr::select(-id) 

micro_rel <- micro %>% 
  mutate(across(everything(),~./sum(.)))

micro_rel$id <- as.numeric(rownames(micro_rel))
rownames(micro_rel) <- NULL


#Load Proteomics Outcomes 
prot <- read.csv(fs::path(dir.dat %>% dirname(),"proteomics","MetaChem_screened_proteomics.csv")) %>% 
  dplyr::select(id,TINAGL1, LCN2)

#Baseline creatinine
ftrs <- readRDS(fs::path(dir.dat, "sol_chs_batch_cor_scaled_untargeted_fts.rds"))
ftrs <- ftrs$chs
c18neg_ftrs_table <- data.frame(ftrs$c18neg)
colnames(c18neg_ftrs_table) <- str_remove(colnames(c18neg_ftrs_table),"X")
c18neg_ftrs_table <- c18neg_ftrs_table %>% 
  dplyr::select(id,`112.051649_23.21579379`) %>% 
  dplyr::rename(ma_creatinine=`112.051649_23.21579379`)
c18neg_ftrs_table$id <- as.numeric(c18neg_ftrs_table$id)

#Merge together
meta <- tidylog::full_join(PFAS,eGFR,by="id")
meta <- tidylog::full_join(meta,cst3,by="id")
# meta <- tidylog::full_join(meta,nutrients,by="id")
meta <- tidylog::full_join(meta,micro_rel,by="id")
meta <- tidylog::full_join(meta,prot,by="id")
meta <- tidylog::full_join(meta,c18neg_ftrs_table,by="id")
meta <- tidylog::left_join(meta,cystatin_c,by="id") #ng/mL
meta <- tidylog::left_join(meta,nutrients,by="id") 
meta$cystatin_c2 <- meta$cystatin_c/1000 #Convert from ng/mL to mg/L for eGFR calculation

meta <- meta %>% 
  rename(mc_CST3_log=mc_CST3,
         ma_CST3_log=ma_CST3,
         ma_creatinine_scaled=ma_creatinine,
         LCN2_log=LCN2,
         TINAGL1_log=TINAGL1)

#Filter to complete pfas & eGFR data 
meta <- meta %>%
  filter(!is.na(mc_creatinine)) %>%
  filter(!is.na(ma_pfda)) #78 remain

#Imput energy and fiber for missing individual 
meta$ma_sex <- as.factor(meta$ma_sex)
meta$ma_race_eth <- as.factor(meta$ma_race_eth)
meta$ma_parental_edu <- as.factor(meta$ma_parental_edu)
meta$id <- as.numeric(meta$id)
# meta$ma_sex <- ifelse(meta$ma_sex=="female",2,1)

#Standardize creatinine
# meta <- meta %>% 
#   mutate(mc_creatinine_scaled=((mc_creatinine-mean(mc_creatinine))/sd(mc_creatinine)))

#Calculate eGFR ----
#Calculate egfr based on creatinine, age and gender
calculate_eGFR <- function(creatinine, age, gender) {
  kappa <- ifelse(gender == "female", 0.7, 0.9)
  alpha <- ifelse(gender == "female", -0.329, -0.411)
  min_creatinine_kappa <- pmin(creatinine / kappa, 1)
  max_creatinine_kappa <- pmax(creatinine / kappa, 1)
  gfr <- 141 * min_creatinine_kappa^alpha * max_creatinine_kappa^(-1.209) * 0.993^age
  return(gfr)
}

eGFR <- calculate_eGFR(creatinine=meta$mc_creatinine,age=meta$ma_age,gender=meta$ma_sex)
meta$eGFR <- eGFR
meta$kidney_binary1<-ifelse(meta$eGFR<116,1,0) #average eGFR number
meta$kidney_binary2<-ifelse(meta$eGFR<90,1,0) #early stage kidney disease


#Calculate egfr based on cystatin c and creatinine, age and gender
calculate_eGFR2 <- function(creatinine, cystatin_c, age, gender) {
  kappa <- ifelse(gender == "female", 0.7, 0.9)
  alpha <- ifelse(gender == "female", -0.219, -0.144)
  min_creatinine_kappa <- pmin(creatinine / kappa, 1)
  max_creatinine_kappa <- pmax(creatinine / kappa, 1)
  min_cys <- pmin(cystatin_c/0.8, 1)
  max_cys <- pmax(cystatin_c/0.8, 1)
  female <- ifelse(gender=="female",0.963,1)
  gfr2 <- 135 * min_creatinine_kappa^alpha * max_creatinine_kappa^(-0.544) * min_cys^(-0.323) * max_cys^(-0.778) * 0.9961^age * female
  return(gfr2)
}

eGFR2 <- calculate_eGFR2(creatinine=meta$mc_creatinine,cystatin_c=meta$cystatin_c2,age=meta$ma_age,gender=meta$ma_sex)
meta$eGFR2 <- eGFR2

#Calculate egfr based on creatinine, age and gender with kidney.org equation
calculate_eGFR3 <- function(creatinine, age, gender) {
  kappa <- ifelse(gender == "female", 0.7, 0.9)
  alpha <- ifelse(gender == "female", -0.241, -0.302)
  min_creatinine_kappa <- pmin(creatinine / kappa, 1)
  max_creatinine_kappa <- pmax(creatinine / kappa, 1)
  female <- ifelse(gender=="female",1.012,1)
  gfr3 <- 142 * min_creatinine_kappa^alpha * max_creatinine_kappa^(-1.200) * 0.9938^age
  return(gfr3)
}

eGFR3 <- calculate_eGFR3(creatinine=meta$mc_creatinine,age=meta$ma_age,gender=meta$ma_sex)
meta$eGFR3 <- eGFR3
# meta$kidney_binary1<-ifelse(meta$eGFR3<116,1,0) #average eGFR number
# meta$kidney_binary2<-ifelse(meta$eGFR3<90,1,0) #early stage kidney disease

# meta$ma_sex <- ifelse(meta$ma_sex==2,"female","male")
# 
# ggplot(data=meta,aes(x=eGFR1,y=eGFR2))+
#   geom_point()+
#   geom_smooth(method="lm")
# 
# ggplot(data=meta,aes(x=eGFR1,y=eGFR3))+
#   geom_point()+
#   geom_smooth(method="lm")

#Scale exposure and outcome
exposure <- c("ma_pfos","ma_pfoa","ma_pfda","ma_pfna","ma_pfhxs","ma_pfhps","ma_pfpes")
outcome <- c("eGFR","eGFR2","eGFR3","mc_creatinine","cystatin_c")
# outcome  <- c("eGFR","mc_creatinine")

#Log transform
meta <- meta %>%
  mutate(across(all_of(exposure), ~log(.),.names = "{col}_log")) %>% 
  mutate(across(all_of(outcome), ~log(.),.names = "{col}_log"))
#Scale
meta <- meta %>%
  mutate(across(all_of(exposure),~((. - mean(.)) / sd(.)),.names = "{col}_scaled")) %>%
  mutate(across(all_of(outcome),~((. - mean(.)) / sd(.)),.names = "{col}_scaled"))
exposure_scaled <- c("ma_pfos_scaled","ma_pfoa_scaled","ma_pfda_scaled","ma_pfna_scaled","ma_pfhxs_scaled","ma_pfhps_scaled","ma_pfpes_scaled")

#Scale the log transformed
exposure <- c("ma_pfos_log","ma_pfoa_log","ma_pfda_log","ma_pfna_log","ma_pfhxs_log","ma_pfhps_log","ma_pfpes_log")
# outcome  <- c("eGFR1_log","eGFR2_log","eGFR3_log","mc_creatinine_log","LCN2_log","mc_CST3_log")
outcome  <- c("eGFR_log","eGFR2_log","eGFR3_log","mc_creatinine_log","LCN2_log","mc_CST3_log","cystatin_c_log")
microbiome <- colnames(micro_rel)[-93]
meta <- meta %>%
  mutate(across(all_of(exposure),~((. - mean(.)) / sd(.)),.names = "{col}_scaled")) %>%
  mutate(across(all_of(outcome),~((. - mean(.)) / sd(.)),.names = "{col}_scaled")) %>% 
  mutate(across(all_of(microbiome),~((. - mean(.,na.rm=T)) / sd(.,na.rm=T)),.names = "{col}_scaled")) 
exposure_log_scaled <- c("ma_pfos_log_scaled","ma_pfoa_log_scaled","ma_pfda_log_scaled",
                         "ma_pfna_log_scaled","ma_pfhxs_log_scaled","ma_pfhps_log_scaled",
                         "ma_pfpes_log_scaled")
# outcome_log_scaled  <- c("eGFR1_log_scaled","eGFR2_log_scaled","eGFR3_log_scaled","mc_creatinine_log_scaled","LCN2_log_scaled","mc_CST3_log_scaled")
outcome_log_scaled  <- c("eGFR_log_scaled","eGFR2_log_scaled","eGFR3_log_scaled","mc_creatinine_log_scaled","LCN2_log_scaled","mc_CST3_log_scaled","cystatin_c_log_scaled")
microbiome_scaled <- paste0(microbiome,"_scaled")
quantile(meta$ma_pfna, probs = seq(0, 1, 1/4))
meta$ma_pfna_tert <- cut(meta$ma_pfna,
                         breaks=c(0,0.4116666667,0.4903333333,0.9),
                         labels=c("Low","Medium","High"))
meta$ma_pfhxs_tert <- cut(meta$ma_pfhxs,
                          breaks=c(0,0.6651666667,1.3080000000,11),
                          labels=c("Low","Medium","High"))

meta$ma_pfhxs_quartile <- cut(meta$ma_pfhxs,
                              breaks=c(0,0.59575,0.94200,1.66100,11),
                              labels=c("Q1","Q2","Q3","Q4"))
meta$ma_pfna_quartile <- cut(meta$ma_pfna,
                             breaks=c(0,0.4025,0.4515,0.5265,0.9),
                             labels=c("Q1","Q2","Q3","Q4"))

#Merge metabolites into metadata
metab.data$id <- as.numeric(metab.data$id)
meta <- tidylog::full_join(meta,metab.data,by="id") #137 total, 78 matched

#Filter to 78 with pfas data
meta_filt <- meta %>% 
  filter(!is.na(ma_pfda))

meta_filt$ma_sex
meta_filt$ma_parental_edu
meta_filt$ma_race_eth

meta_filt$m_total_dietary_fiber_g

# impute_dat <- data.frame(meta_filt[c("id",exposure_log_scaled,outcome_log_scaled,omics,omics2)])
# # impute_dat <- data.frame(meta_filt[c("id",exposure_log_scaled[4],outcome_log_scaled[1],omics2)])
# rownames(impute_dat) <- impute_dat$id
# impute_dat <- impute_dat %>%
#   dplyr::select(-id)
covars <- c("ma_age","ma_sex","ma_parental_edu","ma_race_eth","ma_creatinine_scaled")
meta_filt$ma_creatinine_scaled2 <- as.numeric(meta_filt$ma_creatinine_scaled)
impute_dat <- meta_filt[c(covars[-5],"ma_creatinine_scaled2","m_total_dietary_fiber_g")]
impute_dat_temp <- mice(impute_dat,m=3,seed=500,method="sample") #rf, sample
# summary(impute_dat_temp)
covs_complete <- complete(impute_dat_temp,1)
covs_complete$m_total_dietary_fiber_g
# completedData$id <- rownames(completedData)
# saveRDS(completedData,fs::path(here::here() %>% dirname() %>% dirname(),"0_Data","Imputed_Data_rel_abundance_reduced_v1.rds"))

# meta_filt <- meta %>%
#   filter(!is.na(m_energy_kcal)) #77

# ggplot(data=meta_filt,aes(x=eGFR1,y=eGFR2))+
#   geom_point()+
#   geom_smooth(method="lm")
# 
# ggplot(data=meta_filt,aes(x=eGFR1,y=eGFR3))+
#   geom_point()+
#   geom_smooth(method="lm")
# 
# ggplot(data=meta_filt,aes(x=eGFR2,y=eGFR3))+
#   geom_point()+
#   geom_smooth(method="lm")


# meta_filt$ma_sex <- as.factor(meta_filt$ma_sex)


# mean(meta_filt$ma_age)
# sd(meta_filt$ma_age)
# 
# table(meta_filt$ma_sex)
# (41/78)*100
# (37/78)*100
# mean(meta_filt$ma_creatinine_scaled)
# original_values <- c((meta_filt$ma_creatinine_scaled * 268.2032636) + 1078.121404)
# mean(original_values)
# sd(original_values)
# 
# mean(meta_filt$eGFR)
# sd(meta_filt$eGFR)
# 
# mean(meta_filt$mc_creatinine)*1000
# sd(meta_filt$mc_creatinine)*1000
# 
# table(meta_filt$ma_race_eth)
# 6/78
# 44/78
# 28/78
# 
# table(meta_filt$ma_parental_edu)
# (2/78)*100
# (13/78)*100
# (10/78)*100
# (53/78)*100
# 
# mean(meta_filt$ma_bmi)
# sd(meta_filt$ma_bmi)
# 
# geometric.mean(meta_filt$ma_pfpes)
# geometric_mean <- prod(meta_filt$ma_pfpes)^(1/length(meta_filt$ma_pfpes))
# n <- length(meta_filt$ma_pfpes)
# 
# # Calculate the sum of squared differences
# sum_sq_diff <- sum((meta_filt$ma_pfpes - geometric_mean)^2)
# 
# 
# # Calculate the standard error
# SE <- sqrt(sum_sq_diff /(n * (n - 1)))
# 
# # Z-value for 95% confidence level
# z <- qnorm(1 - 0.05/2)
# 
# ci_lower <- geometric_mean * exp(-z * SE)
# ci_upper <- geometric_mean * exp(z * SE)
# geometric_mean
# ci_lower
# ci_upper
# 
# 
# (meta_filt$ma_pfos<0.43)
# (meta_filt$ma_pfoa<0.01)
# (meta_filt$ma_pfhxs<0.01)
# (meta_filt$ma_pfna<0.01)
# (meta_filt$ma_pfhps<0.05)
# (meta_filt$ma_pfda<0.01)
# length(which((meta_filt$ma_pfpes<0.01)))
# # (14/78)*100
# sd(meta_filt$ma_pfos)
# sd(meta_filt$ma_pfoa)
# sd(meta_filt$ma_pfhxs)
# sd(meta_filt$ma_pfna)
# sd(meta_filt$ma_pfda)
# sd(meta_filt$ma_pfhps)
# sd(meta_filt$ma_pfpes)
# mean(meta_filt$eap.pfas)
# sd(meta_filt$eap.pfas)
# se <- sd(meta_filt$eap.pfas)/sqrt(78)
# mean(meta_filt$eap.pfas)-(1.96*se)
# mean(meta_filt$eap.pfas)+(1.96*se)
# 



#Calculate PFAS burden scores --------------
## A. Create categorical PFAS which are needed for the score ------
pfas_names_all <- c("ma_pfos","ma_pfoa","ma_pfna","ma_pfhxs","ma_pfpes","ma_pfhps","ma_pfda")
data_analysis <- meta_filt %>%
  mutate_at(.vars = vars(all_of(pfas_names_all[c(1:7)])), 
            .funs = list(
              quartile = ~as.integer(cut(., quantile(., probs = seq(0, 1, 0.25)), 
                                         include.lowest = TRUE
              ))))
# mutate(pfas_pfda_detected = ifelse(pfas_pfda<0.05|pfas_pfda==0.2/sqrt(2), 1, 2)) %>%
# mutate(pfas_pfuna_detected = ifelse(pfas_pfuna<0.05|pfas_pfuna==0.1/sqrt(2), 1, 2)) %>%
# mutate_at(.vars = vars(c("pfas_pfhpa","pfas_pfuna")),
#           .funs = list(detected = ~ifelse(.<0.05|.==0.1/sqrt(2), 1, 2))) %>%
# tidylog::mutate_at(.vars = vars(c("pfas_pfhps", "pfas_pfhpa")), #, 
#                    .funs = list(tert = ~cut(., 
#                                                    quantile(., probs = seq(0, 1, 1/3)), 
#                                                    include.lowest = TRUE) %>% 
# as.integer()))

# Select just the categorical PFAS
burden_score_pfas <- data_analysis %>% 
  dplyr::select(contains("detected"), 
                contains("quartile"), 
                contains("tert")) %>% 
  as.data.frame()

eap.pfas <- ltm::factor.scores(grm(burden_score_pfas), 
                               method="EAP", 
                               resp.patterns = burden_score_pfas)$score.dat$z1
meta_filt$eap.pfas <- eap.pfas
meta_filt$burden_scaled <- scale(meta_filt$eap.pfas)

burden_score_pfcas <- burden_score_pfas |> 
  dplyr::select(ma_pfoa_quartile, 
                ma_pfna_quartile, 
                ma_pfda_quartile) 
# pfas_pfuna_detected, 
# pfas_pfhpa_tert)

# Select PFSAs
burden_score_pfsas <- burden_score_pfas |> 
  dplyr::select(ma_pfos_quartile, 
                ma_pfhxs_quartile, 
                ma_pfhps_quartile)


eap.pfcas <- ltm::factor.scores(grm(burden_score_pfcas), 
                                method="EAP", 
                                resp.patterns = burden_score_pfcas)$score.dat$z1
eap.pfsas <- ltm::factor.scores(grm(burden_score_pfsas), 
                                method="EAP", 
                                resp.patterns = burden_score_pfsas)$score.dat$z1

meta_filt$eap.pfcas <- eap.pfcas
meta_filt$burden_scaled_pfcas <- scale(meta_filt$eap.pfcas)

meta_filt$eap.pfsas <- eap.pfsas
meta_filt$burden_scaled_pfsas <- scale(meta_filt$eap.pfsas)


burden_score <- c("burden_scaled","burden_scaled_pfcas","burden_scaled_pfsas")

```

#5. Integration Analysis
##a. Prepare Datasets for Analysis 
```{r Prepare Datasets for Analysis , echo=F, include=F}
# Define exposure and outcome names
exposure <- c("diabetes_duration","diabetes_status")
exposure_dat <- dat[exposure]
#Scale continuous exposure
exposure_dat <- exposure_dat %>% 
  mutate(diabetes_duration=(diabetes_duration-mean(diabetes_duration))/sd(diabetes_duration))

outcome  <- c("ab40_avg_conc","ab42_avg_conc","tau_avg_conc",
              "nfl_avg_conc","gfap_avg_conc","ptau_181_avg_conc","ptau_217_avg_conc")
outcome_dat  <- dat[outcome]
#Scale outcome variables
outcome_dat <- outcome_dat %>% 
  mutate(across(all_of(outcome),~((.-mean(.,na.rm=T))/sd(.,na.rm=T))))

# Get numeric matrix of covariates 
covars <- c("age","sex","bmi","race_ethnicity")
covs <- dat[covars]
covs <- covs %>% 
  mutate(sex=as.numeric(sex),
         race_ethnicity=as.numeric(race_ethnicity))
covs <- as.matrix(covs)
covs[,c(1,3)] <- scale(covs[,c(1,3)]) #Scale continuous covariates


#Create list of omics data 
metabolomics <- dat[metabolites]
metabolomics <- metabolomics %>% 
  mutate(across(all_of(metabolites),~((.-mean(.,na.rm=T))/sd(.,na.rm=T))))

proteomics <- dat[proteins]
proteomics <- proteomics %>% 
  mutate(across(all_of(proteins),~((.-mean(.,na.rm=T))/sd(.,na.rm=T))))

omics_lst <- list(proteomics=proteomics,metabolomics = metabolomics)
```

##b. Exposure Outcome Analysis 
```{r Analysis, include=FALSE,echo=FALSE}

#1. Exposure - Outcome Analysis 
# covars.reduced <- c("ma_age","ma_sex","ma_parental_edu","ma_race_eth","ma_CST3","fiber.adj","m_energy_kcal")
exposure_log <- c("ma_pfos_log","ma_pfoa_log","ma_pfda_log","ma_pfna_log","ma_pfhxs_log",
                  "ma_pfhps_log","ma_pfpes_log")
exposure_log_scaled
exposure <- str_remove(exposure_log,"_log")
outcome_log <- c("eGFR3_log","mc_creatinine_log","cystatin_c_log")
# "LCN2_log","mc_CST3_log",
# outcome_log <- c("eGFR_log","mc_creatinine_log","LCN2_log")
outcome_log_scaled <- outcome_log_scaled[-c(1,6)]
covars <- c("ma_age","ma_sex","ma_race_eth","ma_parental_edu","ma_creatinine_scaled2") #"ma_bmi"
#"ma_creatinine_scaled"
burden_score <- c("burden_scaled","burden_scaled_pfcas","burden_scaled_pfsas")
kidney <- c("kidney_binary1")
pfna <- c("ma_pfna")
# egfr <- c("eGFR1","eGFR2","eGFR3")
egfr <- c("eGFR")
exposures <- c(exposure_log_scaled,"burden_scaled")
# outcome <- c("eGFR3","mc_creatinine","cystatin_c")
exposure_scaled <- c(exposure_scaled[-9])
exposure <- str_remove(exposure_log,"_log")
exposure <- c(exposure,"eap.pfas")
outcome <- c("eGFR2_log","cystatin_c_log","mc_creatinine_log")
# meta_filt <- meta_filt %>%
#   mutate(across(all_of(burden_score), ~log(.),.names = "{col}_log")) 

meta_filt$kidney_binary1 <- as.factor(meta_filt$kidney_binary1)
# outcome <- outcome_log_scaled[c(1,3,5)]
covars2 <- covars[-c(6)]
covariates <- paste0(covars,collapse=" + ")
# covariates <- paste0(covars2,collapse=" + ")
results <- data.frame()
# pdf(fs::path(dir.results,"PFAS_Kidney_Outcomes_scatterplots_04_04.pdf"))
for(x in exposures){
  for (y in outcome){
    # for(x in burden_score){
    # for (y in outcome_log_scaled){
    m1 <- as.formula(paste0(y, "~ ",x," + ",covariates))
    lm_res <- lm(m1,data = meta_filt)
    # lm_res <- glm(m1,data = meta_filt,family="binomial")
    beta <- summary(lm_res)$coef[2,1]
    pval <- round(summary(lm_res)$coef[2,4],3)
    SE <- summary(lm_res)$coef[2,2]
    ll <- (beta - (1.96*SE))
    ul <- (beta + (1.96*SE))
    beta <- round(exp(beta),3)
    ll <- round(exp(ll),3)
    ul <- round(exp(ul),3)
    # p <- ggplot(data = meta, 
    #             aes(x = .data[[x]], y = .data[[y]])) + 
    #   geom_point() +
    #   stat_smooth(method = "lm",
    #               formula = y ~ x ,
    #               geom = "smooth") +
    #   xlab(paste0(x)) +
    #   ylab(paste0(y))+
    #   labs(subtitle=paste0("Beta = ",beta,", P-value = ",pval))+
    #   theme_bw()
    # plot(p)
    
    results.data <- data.frame(exposure=x,outcome=y,beta=beta,ll=ll,ul=ul,pval=pval)
    results <- rbind(results,results.data)
  }
}
sig <- results %>% 
  filter(pval<0.05)
# burdenscore.results <- results
#Coefficient plot
results$exposure <- str_remove(results$exposure,"ma_")
# results$exposure <- str_remove(results$exposure,"_log_scaled")
results$exposure <- str_remove(results$exposure,"_scaled")
results$exposure <- str_to_upper(results$exposure)
results$exposure <- str_replace(results$exposure,"PFHPS","PFHpS")
results$exposure <- str_replace(results$exposure,"PFHXS","PFHxS")
results$exposure <- str_replace(results$exposure,"PFPES","PFPeS")
results$exposure <- str_replace(results$exposure,"BURDEN","Burden Score")

results$outcome <- str_remove(results$outcome,"_log")
results$outcome <- str_replace(results$outcome,"mc_creatinine","Creatinine")
# results$outcome <- str_replace(results$outcome,"mc_CST3","CST3")
results$outcome <- str_replace(results$outcome,"cystatin_c","Cystatin-C")
results$outcome <- str_replace(results$outcome,"eGFR2","eGFR")
# results$outcome <- str_replace(results$outcome,"Creatinine","Follow-up Creatinine")
# results$outcome <- factor(results$outcome,levels=c("eGFR"))
# results <- results %>% 
#   filter(outcome=="eGFR")
results$outcome <- factor(results$outcome,levels=unique(c(results$outcome)))

results <- results[order(results$beta),]
results$exposure <- factor(results$exposure,levels=c("Burden Score","PFHpS","PFNA","PFOS","PFOA","PFHxS","PFPeS","PFDA"))


ggplot(aes(beta, exposure), data=results) +
  geom_errorbarh(aes(xmin = ll, xmax = ul),height=0.4) +
  geom_point() +
  # add in a dotted line at zero
  geom_vline(xintercept = 0, lty = 2) +
  facet_wrap("outcome",scale="free_x")+
  labs(
    x = "Beta Coefficient (95% CI)"
  )+
  theme_bw()

results <- results %>% 
  mutate(beta_perc = ifelse(beta<1,-((1-beta)*100),((beta-1)*100))) %>% 
  mutate(ll_perc = ifelse(ll<1,-((1-ll)*100),((ll-1)*100))) %>% 
  mutate(ul_perc = ifelse(ul<1,-((1-ul)*100),((ul-1)*100)))

ggplot(aes(beta_perc, exposure), data=results) +
  geom_errorbarh(aes(xmin = ll_perc, xmax = ul_perc),height=0.4) +
  geom_point() +
  # add in a dotted line at zero
  geom_vline(xintercept = 0, lty = 2) +
  facet_wrap("outcome",scale="free_x")+
  labs(
    x = "Beta Coefficient (95% CI)",
    y="Baseline PFAS"
  )+
  theme_bw()

alpha = log((100 + 10)/100) 
beta=-0.165
exp(alpha*beta)
1-0.984396831
0.015603169*100
1-0.9703649347
0.0296350653*100
-(1-exp(alpha*-0.165))*100
#1.6%
ll=-0.311
exp(alpha*ll)
1-0.9448755964
0.0551244036*100

ul=-0.019
exp(alpha*ul)
1-0.9965418835
0.0034581165*100

-(1-exp(alpha*-0.019))*100


1-0.9950525
0.0049475*100
1-.954
0.046*100
# dev.off()
# write.csv(results,fs::path(dir.results,"Results_04_04.csv"))


0.243
beta=0.243
(exp(alpha*beta)-1)*100
1.045300222-1
0.045300222*100 #4.53% increase

#Calculate 10% increase
alpha = log((100 + 10)/100) 
results_wide <- results %>% 
  mutate(beta_est=(ifelse(beta>0,round((exp(alpha*beta)-1)*100,3),round(-(1-exp(alpha*beta))*100,3)))) %>% 
  mutate(ll_est=(ifelse(beta>0,round((exp(alpha*ll)-1)*100,3),round(-(1-exp(alpha*ll))*100,3)))) %>% 
  mutate(ul_est=(ifelse(beta>0,round((exp(alpha*ul)-1)*100,3),round(-(1-exp(alpha*ul))*100,3)))) %>% 
  mutate(sig=ifelse(pval<0.05,"*","")) %>% 
  mutate(Estimate=paste(beta_est,"% [",ll_est,"%,",ul_est,"%]",sig)) %>% 
  dplyr::select(-c(beta:sig)) %>% 
  pivot_wider(names_from=outcome,values_from=Estimate)
write.csv(results_wide,fs::path(dir.results,"Table_S1.csv"))


#PFNA/PFHxS & LCN2
m1 <- as.formula(paste0("LCN2_log ~ ma_pfna_tert +", covariates))
lm_res <- lm(m1,data = meta)
summary(lm_res)

m1 <- as.formula(paste0("LCN2_log ~ ma_pfhxs_tert +", covariates))
lm_res <- lm(m1,data = meta)
summary(lm_res)


#Mixture Analysis 
#Exposure mixture
exposure_names = exposure_log
colnames_omic_fts <- outcome
# colnames_omic_fts <- kidney

#OWAS Function - PFAS Mixtures vs. Microbiome Taxa
out <- owas_qgcomp(df = meta_filt,
                   expnms = exposure_names,
                   omics = colnames_omic_fts,
                   q = 7, 
                   covars = c("ma_age","ma_sex","ma_parental_edu","ma_race_eth","ma_creatinine_scaled"), 
                   confidence_level = 0.95,
                   family="gaussian",
                   run.qgcomp.boot = FALSE) 
# write.csv(out,fs::path(dir.results,"Log Mixture PFAS Kidney q_null_scaled.csv"))

#BKMR
y <- meta_filt$LCN2_log
X_vars <- meta_filt[covars]
X_vars$ma_sex <- as.character(X_vars$ma_sex)
X_vars$ma_sex <- ifelse(X_vars$ma_sex=="male",1,0)
X_vars$ma_parental_edu <- as.numeric(X_vars$ma_parental_edu)
X_vars$ma_race_eth <- as.numeric(X_vars$ma_race_eth)
X <- as.matrix(X_vars)
Z <- as.matrix(meta_filt[exposure])
temp <- kmbayes(y = y, Z = Z, X = X, iter = 10000, verbose = FALSE, varsel = TRUE)
summary(temp)
ExtractPIPs(temp)

sel<-seq(100,1000,by=1)

#Visualize the overall mixture effect
pred.resp.univar <- PredictorResponseUnivar(fit=temp, sel=sel,
                                            method="approx")

pred.resp.bivar  <- PredictorResponseBivar(fit=temp,  min.plot.dist = 1,
                                           sel=sel,  method="approx")

pred.resp.bivar.levels <- PredictorResponseBivarLevels(pred.resp.df=
                                                         pred.resp.bivar,   Z = Z, both_pairs=TRUE,
                                                       qs = c(0.1, 0.5, 0.9))

risks.overall <- OverallRiskSummaries(fit=temp, qs=seq(0.1, 0.9, by=0.05),
                                      q.fixed = 0.5, method = "approx",
                                      sel=sel)

risks.singvar <- SingVarRiskSummaries(fit=temp, qs.diff = c(0.1, 0.9),
                                      q.fixed = c(0.1, 0.50, 0.9),
                                      method = "approx")

risks.int <- SingVarIntSummaries(fit=temp, qs.diff = c(0.1, 0.9),
                                 qs.fixed = c(0.1, 0.9))

ggplot(risks.overall, aes(quantile, est, ymin = est - 1.96*sd,
                          ymax = est + 1.96*sd)) +
  geom_hline(yintercept=00, linetype="dashed", color="gray") +
  geom_pointrange() + scale_y_continuous(name="estimate")+
  theme_bw()


# Make exposure outcome associaiton plots for eGFR and LCN2 - x axis mixture effects, y axis pfas


```


##c. Integrated/Quasi-Mediation 
```{r Integrated/Quasi-Mediation,echo=F,include=T}

#Convert metabolomics matrices to vector variables
omics <- colnames(omics_df)

omics2 <- colnames(microbiome)
#Remove unclassified microbiome features
# omics2 <- omics2[which(!grepl("unclassified",omics2))] #80 features
# omics2 <- microbiome_features

# missing <- meta_filt[c(omics,omics2)]

# Impute missing microbiome data before analysis - 42% missingness
# impute_dat <- data.frame(meta_filt[c("id",exposure_log_scaled,outcome_log_scaled,omics,omics2)])
# # impute_dat <- data.frame(meta_filt[c("id",exposure_log_scaled[4],outcome_log_scaled[1],omics2)])
# rownames(impute_dat) <- impute_dat$id
# impute_dat <- impute_dat %>%
#   dplyr::select(-id)

# impute_dat_temp <- mice(impute_dat,m=3,seed=500,method="sample") #rf, sample
# summary(impute_dat_temp)
# completedData <- complete(impute_dat_temp,1)
# completedData$id <- rownames(completedData)
# saveRDS(completedData,fs::path(here::here() %>% dirname() %>% dirname(),"0_Data","Imputed_Data_rel_abundance_reduced_v1.rds"))
# 
# completedData <- readRDS(fs::path(here::here() %>% dirname() %>% dirname(),"0_Data","Imputed_Data_rel_abundance_reduced_v1.rds"))
completedData <- readRDS(fs::path(here::here() %>% dirname() %>% dirname(),"0_Data","Imputed_Data_rel_abundance_v2.rds"))
microbiome_unscaled <- str_remove(microbiome_scaled,"_scaled")
#Scale microbime features
completedData <- completedData %>% 
  mutate(across(all_of(microbiome_unscaled),~((. - mean(.,na.rm=T)) / sd(.,na.rm=T)),.names = "{col}_scaled")) 
eGFR_vals <- meta_filt[c("id","eGFR","eGFR2","eGFR3","eGFR_log","eGFR2_log","eGFR3_log","eGFR_log_scaled","eGFR2_log_scaled","eGFR3_log_scaled","burden_scaled")]
completedData$id <- as.numeric(completedData$id)
completedData <- tidylog::full_join(completedData,eGFR_vals,by="id")
completedData <- tidylog::left_join(completedData,nutrients,by="id")
# completedData <- completedData %>%
#   filter(!is.na(m_energy_kcal))
# completedData <- completedData %>%
#   dplyr::select(-m_energy_kcal)
```

###i. Sensitivity Analysis
```{r sensitivity, echo = F}
omics <- colnames(omics_df)

omics2 <- colnames(microbiome)

meta_sens <- meta_filt %>% 
  filter(!is.na(Prevotella_9))
microbiome_unscaled <- str_remove(microbiome_scaled,"_scaled")

full_data <- meta_sens
# full_data$id <- rownames(full_data)
# covariates <- meta_filt[c("id",covars)]
# covariates$id <- as.character(covariates$id)
# full_data <- tidylog::left_join(full_data,covariates,by="id")
#PFAS with microbime
microbiome_scaled
exposure_log_scaled
outcome_log_scaled


lm_fxn1 <- function(x,y) {
  m0 <- as.formula(paste0(y,"~",x,"+ma_age+ma_sex+ma_race_eth+ma_parental_edu"))
  # m1 <- lm(m0,data=meta_filt)
  m1 <- lm(m0,data=full_data)
  beta <- round(summary(m1)$coef[2,1],5)
  pval <- summary(m1)$coef[2,4]
  output1 <- data.frame(PFAS=x,Microbiome=y,Beta=beta,P_Val=pval)
  return(output1)
}

#PFAS with Microbime
combinations <- crossing(x = exposure_log_scaled, y = microbiome_scaled)
# combinations <- crossing(x = burden_score[1], y = microbiome_scaled)

# Apply lm_fxn to each combination
results1 <- combinations %>%
  pmap_dfr(~ lm_fxn1(..1, ..2))
results1 <- results1 %>% 
  mutate(Direction=ifelse(Beta>0,"Positive","Negative"))

#Microbime with eGFR
lm_fxn2 <- function(x,y) {
  m0 <- as.formula(paste0(y,"~",x,"+ma_age+ma_sex+ma_race_eth+ma_parental_edu"))
  m1 <- lm(m0,data=full_data)
  beta <- round(summary(m1)$coef[2,1],5)
  pval <- summary(m1)$coef[2,4]
  output1 <- data.frame(Microbiome=x,Outcome=y,Beta=beta,P_Val=pval)
  return(output1)
}

combinations <- crossing(x = microbiome_scaled, y=outcome_log_scaled[3])
# combinations <- crossing(x = microbiome_scaled, y=outcome_log_scaled[2])
# Apply lm_fxn to each combination
results2 <- combinations %>%
  pmap_dfr(~ lm_fxn2(..1, ..2))
results2 <- results2 %>% 
  mutate(Direction=ifelse(Beta>0,"Positive","Negative"))
# results2.sig <- results2 %>%
#   filter(P_Val<0.05)
# microbiome_features2 <- unique(results2.sig$Microbiome) #5
# microbiome_features <- unique(c(microbiome_features1,microbiome_features2))#19

microbiome_results_full <- tidylog::full_join(results1,results2,by=c("Microbiome"))
microbiome_results_reduced <- microbiome_results_full %>% 
  mutate(Dir=ifelse((Direction.x=="Positive" & Direction.y=="Negative") | (Direction.x=="Negative" & Direction.y=="Positive"),"Diff","Same"))
microbiome_results_reduced <- microbiome_results_reduced %>% 
  filter((Dir=="Diff" & P_Val.x<0.05)|(Dir=="Diff"& P_Val.y<0.05))
microbiome_features <- unique(microbiome_results_reduced$Microbiome) #18 (15) (3) (17) (15) (10)

#PFAS with Metabolites 
metabolite <- metabolites$key
lm_fxn3 <- function(x,y) {
  m0 <- as.formula(paste0(y,"~",x,"+ma_age+ma_sex+ma_race_eth+ma_parental_edu"))
  m1 <- lm(m0,data=full_data)
  beta <- round(summary(m1)$coef[2,1],5)
  pval <- summary(m1)$coef[2,4]
  output1 <- data.frame(PFAS=x,Metabolites=y,Beta=beta,P_Val=pval)
  return(output1)
}
combinations <- crossing(x = exposure_log_scaled, y = metabolite)
# combinations <- crossing(x = burden_score[1], y = metabolite)
# combinations <- crossing(x = microbiome, y = metabolite)
results3 <- combinations %>%
  pmap_dfr(~ lm_fxn3(..1, ..2)) 
results3 <- results3 %>% 
  mutate(Direction=ifelse(Beta>0,"Positive","Negative"))
# results3.sig <- results3 %>%
#   filter(P_Val<0.05)
# metabolite_features1 <- unique(results3.sig$Metabolites) #105 metabolites

#Metabolites with eGFR
lm_fxn4 <- function(x,y) {
  m0 <- as.formula(paste0(y,"~",x,"+ma_age+ma_sex+ma_race_eth+ma_parental_edu"))
  m1 <- lm(m0,data=full_data)
  beta <- round(summary(m1)$coef[2,1],5)
  pval <- summary(m1)$coef[2,4]
  output1 <- data.frame(Metabolites=x,Outcome=y,Beta=beta,P_Val=pval)
  return(output1)
}

# combinations <- crossing(x = metabolite, y = outcome_log_scaled[1])
# combinations <- crossing(x = metabolite, y = outcome_log_scaled[2])
combinations <- crossing(x = metabolite, y = outcome_log_scaled[3])

# Apply lm_fxn to each combination
results4 <- combinations %>%
  pmap_dfr(~ lm_fxn4(..1, ..2))
results4 <- results4 %>% 
  mutate(Direction=ifelse(Beta>0,"Positive","Negative"))
# results4.sig <- results4 %>%
#   filter(P_Val<0.05)
# metabolite_features2 <- unique(results4.sig$Metabolites) #15 metabolites
# metabolite_features <- unique(c(metabolite_features1,metabolite_features2)) #105
# metabolite_features <- metabolite_features2

metabolite_results_full <- tidylog::full_join(results3,results4,by=c("Metabolites"))
metabolite_results_reduced <- metabolite_results_full %>% 
  mutate(Dir=ifelse((Direction.x=="Positive" & Direction.y=="Negative") | (Direction.x=="Negative" & Direction.y=="Positive"),"Diff","Same"))
metabolite_results_reduced <- metabolite_results_reduced %>% 
  filter((Dir=="Diff" & P_Val.x<0.05)|(Dir=="Diff"& P_Val.y<0.05))
metabolite_features <- unique(metabolite_results_reduced$Metabolites) #100 (97) (44) (92) (97) (94)


omics2 <- microbiome_features
microbiome_features_complete <- full_data[,omics2]
microbiome_features_complete$id <- rownames(microbiome_features_complete)
# microbiome_features_complete <- meta_filt[c("id",omics2)]

#Metabolomics
omics <- metabolite_features
metabolomics_features_complete <- full_data[c("id",omics)]
# metabolomics_features_complete <- meta_filt[c("id",omics)]


#Exposure-PFAS
pfas_features_complete <- full_data[c(exposure_log_scaled)]
# pfas_features_complete <- completedData[c(burden_score[1])]
# pfas_features_complete <- completedData[c(exposure_log_scaled[4])]
# pfas_features_complete <- completedData[c("eap.pfas","eap.pfcas","eap.pfsas")]
rownames(pfas_features_complete) <- NULL

#Outcome-eGFR
outcome_outcome <- full_data[c(outcome_log_scaled[3])]
rownames(outcome_outcome) <- NULL

#Create Omics Dataframe
microbiome_features_complete$id <- as.numeric(microbiome_features_complete$id)
omics_complete <- tidylog::full_join(microbiome_features_complete,metabolomics_features_complete,by="id")

# omics_complete <- microbiome_features_complete
omics_complete <- omics_complete %>% 
  dplyr::select(-id)
rownames(omics_complete) <- NULL

#Create Omics List
# omics_lst <- list(microbiome=data.frame(microbiome_features_complete[-19]),metabolomics = data.frame(scale(metabolomics_features_complete[-1])))
omics_lst <- list(microbiome=data.frame(microbiome_features_complete[-11]),metabolomics = data.frame(scale(metabolomics_features_complete[-c(1,2)])))
# omics

# Define exposure and outcome name
exposure <- as.vector(full_data$ma_pfna_log_scaled)
exposure <- as.vector(full_data$ma_pfhps_log_scaled)
exposure <- as.vector(full_data$burden_scaled)

# exposure <- meta_filt[exposure_log_scaled]|> as.matrix()
outcome  <- as.vector(full_data$eGFR3_log_scaled)

# Get numeric matrix of covariates 
covs <- meta_sens[colnames(covs)]

# create list of omics data 
# omics_lst <- simulated_data[-which(names(simulated_data) == "phenotype")]
omics_lst <- omics_lst


# Create data frame of omics data
omics_df <- omics_lst |> 
  purrr::map(~as_tibble(.x, rownames = "name")) |>
  purrr::reduce(left_join, by = "name") |>
  column_to_rownames("name")

# # Transpose list of matrices
omics_t <- lapply(omics_lst, t)

# Rename omics datasets
names(omics_t) = names(omics_lst)

#Data reduction step
jive1 <- jive(omics_t, rankJ = 5, rankA = c(5,5),
              method = "given",
              # conv = 1e-04,
              maxiter = 100,
              showProgress = FALSE)
# #
summary(jive1)
showVarExplained(jive1)
```

###ii. Sensitivity Analysis 2 
```{r Microbiome and Metabolomcs Data Reduction, echo=F, include=F}
covars <- c("ma_age","ma_sex","ma_race_eth","ma_parental_edu","ma_creatinine_scaled","m_energy_kcal")
#FUll Data
full_data <- completedData
# full_data <- meta_filt %>%
# filter(!is.na(Prevotella_9))
# full_data$id <- rownames(full_data)
covariates <- meta_filt[c("id",covars)]
# covariates$id <- as.character(covariates$id)
full_data <- tidylog::left_join(full_data,covariates,by="id")
#PFAS with microbime
microbiome_scaled
exposure_log_scaled
outcome_log_scaled
# pfna_exposure <- exposure_log_scaled[4]

lm_fxn1 <- function(x,y) {
  m0 <- as.formula(paste0(y,"~",x,"+ma_age+ma_sex+ma_race_eth+ma_parental_edu+m_energy_kcal"))
  # m1 <- lm(m0,data=meta_filt)
  m1 <- lm(m0,data=full_data)
  beta <- round(summary(m1)$coef[2,1],5)
  pval <- summary(m1)$coef[2,4]
  output1 <- data.frame(PFAS=x,Microbiome=y,Beta=beta,P_Val=pval)
  return(output1)
}

#PFAS with Microbime
combinations <- crossing(x = exposure_log_scaled, y = microbiome_scaled)
# combinations <- crossing(x = burden_score[1], y = microbiome_scaled)

# Apply lm_fxn to each combination
results1 <- combinations %>%
  pmap_dfr(~ lm_fxn1(..1, ..2))
results1 <- results1 %>% 
  mutate(Direction=ifelse(Beta>0,"Positive","Negative"))
results1.sig <- results1 %>%
  filter(P_Val<0.05)
# microbiome_features1 <- unique(results1.sig$Microbiome) #13
results1_l <- results1 %>% 
  mutate(P_Val=ifelse(P_Val<0.05,paste0(round(P_Val,3),"*"),paste0(round(P_Val,3)))) %>% 
  mutate(Estimate=paste(round(Beta,3)," (p = ",P_Val,")",sep=""))
results1_l <- results1_l %>% 
  dplyr::select(PFAS,Microbiome,Estimate)
results1_l <- results1_l %>%
  pivot_wider(names_from = PFAS,values_from = Estimate)
# write.csv(results1,fs::path(dir.results,"PFAS_Microbiome_Sensitivity_Results.csv"))

#Microbime with eGFR
lm_fxn2 <- function(x,y) {
  m0 <- as.formula(paste0(y,"~",x,"+ma_age+ma_sex+ma_race_eth+ma_parental_edu+m_energy_kcal"))
  m1 <- lm(m0,data=full_data)
  beta <- round(summary(m1)$coef[2,1],5)
  pval <- summary(m1)$coef[2,4]
  output1 <- data.frame(Microbiome=x,Outcome=y,Beta=beta,P_Val=pval)
  return(output1)
}

combinations <- crossing(x = microbiome_scaled, y=outcome_log_scaled[2])
# combinations <- crossing(x = microbiome_scaled, y=outcome_log_scaled[2])
# Apply lm_fxn to each combination
results2 <- combinations %>%
  pmap_dfr(~ lm_fxn2(..1, ..2))
results2 <- results2 %>% 
  mutate(Direction=ifelse(Beta>0,"Positive","Negative"))
results2.sig <- results2 %>%
  filter(P_Val<0.05)
# results1_l <- results1 %>% 
#   mutate(P_Val=ifelse(P_Val<0.05,paste0(round(P_Val,3),"*"),paste0(round(P_Val,3)))) %>% 
#   mutate(Estimate=paste(round(Beta,3)," (p = ",P_Val,")",sep=""))
# results1_l <- results1_l %>% 
#   dplyr::select(PFAS,Microbiome,Estimate)
# results1_l <- results1_l %>%
#   pivot_wider(names_from = PFAS,values_from = Estimate)
# write.csv(results1,fs::path(dir.results,"PFAS_Microbiome_Sensitivity_Results.csv"))
# write.csv(results2.sig,fs::path(dir.results,"Sig_Mic_eGFR_Sens.csv"))

# microbiome_features2 <- unique(results2.sig$Microbiome) #5
# microbiome_features <- unique(c(microbiome_features1,microbiome_features2))#19

microbiome_results_full <- tidylog::full_join(results1,results2,by=c("Microbiome"))
microbiome_results_reduced <- microbiome_results_full %>% 
  mutate(Dir=ifelse((Direction.x=="Positive" & Direction.y=="Negative") | (Direction.x=="Negative" & Direction.y=="Positive"),"Diff","Same"))
microbiome_results_reduced <- microbiome_results_reduced %>% 
  filter((Dir=="Diff" & P_Val.x<0.05)|(Dir=="Diff"& P_Val.y<0.05))
microbiome_features <- unique(microbiome_results_reduced$Microbiome) #18 (15) (3) (17) (15)
# write.csv(microbiome_results_reduced,fs::path(dir.results,"Microbiome_Feature_Selection_Sensitivity.csv"))
#PFAS with Metabolites 
metabolite <- metabolites$key
lm_fxn3 <- function(x,y) {
  m0 <- as.formula(paste0(y,"~",x,"+ma_age+ma_sex+ma_race_eth+ma_parental_edu+m_energy_kcal"))
  m1 <- lm(m0,data=full_data)
  beta <- round(summary(m1)$coef[2,1],5)
  pval <- summary(m1)$coef[2,4]
  output1 <- data.frame(PFAS=x,Metabolites=y,Beta=beta,P_Val=pval)
  return(output1)
}
combinations <- crossing(x = exposure_log_scaled, y = metabolite)
# combinations <- crossing(x = burden_score[1], y = metabolite)
# combinations <- crossing(x = microbiome, y = metabolite)
results3 <- combinations %>%
  pmap_dfr(~ lm_fxn3(..1, ..2)) 
results3 <- results3 %>% 
  mutate(Direction=ifelse(Beta>0,"Positive","Negative"))
# results3.sig <- results3 %>%
#   filter(P_Val<0.05)
# metabolite_features1 <- unique(results3.sig$Metabolites) #105 metabolites

#Metabolites with eGFR
lm_fxn4 <- function(x,y) {
  m0 <- as.formula(paste0(y,"~",x,"+ma_age+ma_sex+ma_race_eth+ma_parental_edu+m_energy_kcal"))
  m1 <- lm(m0,data=full_data)
  beta <- round(summary(m1)$coef[2,1],5)
  pval <- summary(m1)$coef[2,4]
  output1 <- data.frame(Metabolites=x,Outcome=y,Beta=beta,P_Val=pval)
  return(output1)
}

# combinations <- crossing(x = metabolite, y = outcome_log_scaled[1])
# combinations <- crossing(x = metabolite, y = outcome_log_scaled[2])
combinations <- crossing(x = metabolite, y = outcome_log_scaled[2])

# Apply lm_fxn to each combination
results4 <- combinations %>%
  pmap_dfr(~ lm_fxn4(..1, ..2))
results4 <- results4 %>% 
  mutate(Direction=ifelse(Beta>0,"Positive","Negative"))
# results4.sig <- results4 %>%
#   filter(P_Val<0.05)
# metabolite_features2 <- unique(results4.sig$Metabolites) #15 metabolites
# metabolite_features <- unique(c(metabolite_features1,metabolite_features2)) #105
# metabolite_features <- metabolite_features2

metabolite_results_full <- tidylog::full_join(results3,results4,by=c("Metabolites"))
metabolite_results_reduced <- metabolite_results_full %>% 
  mutate(Dir=ifelse((Direction.x=="Positive" & Direction.y=="Negative") | (Direction.x=="Negative" & Direction.y=="Positive"),"Diff","Same"))
metabolite_results_reduced <- metabolite_results_reduced %>% 
  filter((Dir=="Diff" & P_Val.x<0.05)|(Dir=="Diff"& P_Val.y<0.05))
metabolite_features <- unique(metabolite_results_reduced$Metabolites) #100 (97) (44) (92) (97)


sig.met1 <- metabolite_results_reduced %>% 
  filter(P_Val.x<0.05)

unique(sig.met1$Metabolites)

sig.met2 <- metabolite_results_reduced %>% 
  filter(P_Val.y<0.05)
unique(sig.met2$Metabolites)
```


##d. Microbiome & Metabolomics Data Reduction 
```{r Microbiome and Metabolomcs Data Reduction, echo=F, include=F}

#FUll Data
full_data <- completedData
# full_data <- meta_filt %>%
# filter(!is.na(Prevotella_9))
# full_data$id <- rownames(full_data)

covariates <- meta_filt[c("id",covars)]
# covariates$id <- as.character(covariates$id)
full_data <- tidylog::left_join(full_data,covariates,by="id")
#PFAS with microbime
microbiome_scaled
exposure_log_scaled
outcome_log_scaled
# pfna_exposure <- exposure_log_scaled[4]

lm_fxn1 <- function(x,y) {
  m0 <- as.formula(paste0(y,"~",x,"+ma_age+ma_sex+ma_race_eth+ma_parental_edu"))
  # m1 <- lm(m0,data=meta_filt)
  m1 <- lm(m0,data=full_data)
  beta <- round(summary(m1)$coef[2,1],5)
  pval <- summary(m1)$coef[2,4]
  output1 <- data.frame(PFAS=x,Microbiome=y,Beta=beta,P_Val=pval)
  return(output1)
}

#PFAS with Microbime
combinations <- crossing(x = exposure_log_scaled, y = microbiome_scaled)
# combinations <- crossing(x = burden_score[1], y = microbiome_scaled)

# Apply lm_fxn to each combination
results1 <- combinations %>%
  pmap_dfr(~ lm_fxn1(..1, ..2))
results1 <- results1 %>% 
  mutate(Direction=ifelse(Beta>0,"Positive","Negative"))
results1.sig <- results1 %>%
  filter(P_Val<0.05)
# microbiome_features1 <- unique(results1.sig$Microbiome) #13
# write.csv(results1.sig,fs::path(dir.results,"Sig_Mic_PFAS_Sens.csv"))

#Microbime with eGFR
lm_fxn2 <- function(x,y) {
  m0 <- as.formula(paste0(y,"~",x,"+ma_age+ma_sex+ma_race_eth+ma_parental_edu"))
  m1 <- lm(m0,data=full_data)
  beta <- round(summary(m1)$coef[2,1],5)
  pval <- summary(m1)$coef[2,4]
  output1 <- data.frame(Microbiome=x,Outcome=y,Beta=beta,P_Val=pval)
  return(output1)
}

combinations <- crossing(x = microbiome_scaled, y=outcome_log_scaled[2])
# combinations <- crossing(x = microbiome_scaled, y=outcome_log_scaled[2])
# Apply lm_fxn to each combination
results2 <- combinations %>%
  pmap_dfr(~ lm_fxn2(..1, ..2))
results2 <- results2 %>% 
  mutate(Direction=ifelse(Beta>0,"Positive","Negative"))
results2.sig <- results2 %>%
  filter(P_Val<0.05)
# write.csv(results2.sig,fs::path(dir.results,"Sig_Mic_eGFR_Sens.csv"))

# microbiome_features2 <- unique(results2.sig$Microbiome) #5
# microbiome_features <- unique(c(microbiome_features1,microbiome_features2))#19

microbiome_results_full <- tidylog::full_join(results1,results2,by=c("Microbiome"))
microbiome_results_reduced <- microbiome_results_full %>% 
  mutate(Dir=ifelse((Direction.x=="Positive" & Direction.y=="Negative") | (Direction.x=="Negative" & Direction.y=="Positive"),"Diff","Same"))
microbiome_results_reduced <- microbiome_results_reduced %>% 
  filter((Dir=="Diff" & P_Val.x<0.05)|(Dir=="Diff"& P_Val.y<0.05))
microbiome_features <- unique(microbiome_results_reduced$Microbiome) #18 (15) (3) (17) (15)
# write.csv(microbiome_results_reduced,fs::path(dir.results,"Microbiome_Feature_Selection_Sensitivity.csv"))
#PFAS with Metabolites 
metabolite <- metabolites$key
lm_fxn3 <- function(x,y) {
  m0 <- as.formula(paste0(y,"~",x,"+ma_age+ma_sex+ma_race_eth+ma_parental_edu"))
  m1 <- lm(m0,data=full_data)
  beta <- round(summary(m1)$coef[2,1],5)
  pval <- summary(m1)$coef[2,4]
  output1 <- data.frame(PFAS=x,Metabolites=y,Beta=beta,P_Val=pval)
  return(output1)
}
combinations <- crossing(x = exposure_log_scaled, y = metabolite)
# combinations <- crossing(x = burden_score[1], y = metabolite)
# combinations <- crossing(x = microbiome, y = metabolite)
results3 <- combinations %>%
  pmap_dfr(~ lm_fxn3(..1, ..2)) 
results3 <- results3 %>% 
  mutate(Direction=ifelse(Beta>0,"Positive","Negative"))
# results3.sig <- results3 %>%
#   filter(P_Val<0.05)
# metabolite_features1 <- unique(results3.sig$Metabolites) #105 metabolites

#Metabolites with eGFR
lm_fxn4 <- function(x,y) {
  m0 <- as.formula(paste0(y,"~",x,"+ma_age+ma_sex+ma_race_eth+ma_parental_edu"))
  m1 <- lm(m0,data=full_data)
  beta <- round(summary(m1)$coef[2,1],5)
  pval <- summary(m1)$coef[2,4]
  output1 <- data.frame(Metabolites=x,Outcome=y,Beta=beta,P_Val=pval)
  return(output1)
}

# combinations <- crossing(x = metabolite, y = outcome_log_scaled[1])
# combinations <- crossing(x = metabolite, y = outcome_log_scaled[2])
combinations <- crossing(x = metabolite, y = outcome_log_scaled[2])

# Apply lm_fxn to each combination
results4 <- combinations %>%
  pmap_dfr(~ lm_fxn4(..1, ..2))
results4 <- results4 %>% 
  mutate(Direction=ifelse(Beta>0,"Positive","Negative"))
# results4.sig <- results4 %>%
#   filter(P_Val<0.05)
# metabolite_features2 <- unique(results4.sig$Metabolites) #15 metabolites
# metabolite_features <- unique(c(metabolite_features1,metabolite_features2)) #105
# metabolite_features <- metabolite_features2

metabolite_results_full <- tidylog::full_join(results3,results4,by=c("Metabolites"))
metabolite_results_reduced <- metabolite_results_full %>% 
  mutate(Dir=ifelse((Direction.x=="Positive" & Direction.y=="Negative") | (Direction.x=="Negative" & Direction.y=="Positive"),"Diff","Same"))
metabolite_results_reduced <- metabolite_results_reduced %>% 
  filter((Dir=="Diff" & P_Val.x<0.05)|(Dir=="Diff"& P_Val.y<0.05))
metabolite_features <- unique(metabolite_results_reduced$Metabolites) #100 (97) (44) (92) (97)


sig.met1 <- metabolite_results_reduced %>% 
  filter(P_Val.x<0.05)

unique(sig.met1$Metabolites)

sig.met2 <- metabolite_results_reduced %>% 
  filter(P_Val.y<0.05)
unique(sig.met2$Metabolites)
```


#Finalize before analysis
```{r ,echo=FALSE}
#Create dataframes/matrices for LUCID anlaysis 
#Microbiome
omics2 <- microbiome_features
microbiome_features_complete <- completedData[,omics2]
microbiome_features_complete$id <- rownames(microbiome_features_complete)
# microbiome_features_complete <- meta_filt[c("id",omics2)]

#Metabolomics
omics <- metabolite_features
metabolomics_features_complete <- completedData[c("id",omics)]
# metabolomics_features_complete <- meta_filt[c("id",omics)]


#Exposure-PFAS
pfas_features_complete <- completedData[c(exposure_log_scaled)]
# pfas_features_complete <- completedData[c(burden_score[1])]
# pfas_features_complete <- completedData[c(exposure_log_scaled[4])]
# pfas_features_complete <- completedData[c("eap.pfas","eap.pfcas","eap.pfsas")]
rownames(pfas_features_complete) <- NULL

#Outcome-eGFR
outcome_outcome <- completedData[c(outcome_log_scaled[1])]
rownames(outcome_outcome) <- NULL

#Create Omics Dataframe
microbiome_features_complete$id <- as.numeric(microbiome_features_complete$id)
omics_complete <- tidylog::full_join(microbiome_features_complete,metabolomics_features_complete,by="id")

# omics_complete <- microbiome_features_complete
omics_complete <- omics_complete %>% 
  dplyr::select(-id)
rownames(omics_complete) <- NULL

#Create Omics List
# omics_lst <- list(microbiome=data.frame(microbiome_features_complete[-19]),metabolomics = data.frame(scale(metabolomics_features_complete[-1])))
# omics_lst <- list(microbiome=data.frame(microbiome_features_complete[-18]),metabolomics = data.frame(scale(metabolomics_features_complete[-c(1,2)])))
# omics_lst <- list(microbiome=data.frame(scale(microbiome_features_complete[-1])))

omics_lst <- list(microbiome=data.frame(microbiome_features_complete[-16]),metabolomics = data.frame(scale(metabolomics_features_complete[-c(1,2)])))
# omics_lst <- list(microbiome=data.frame(microbiome_features_complete[-17]),metabolomics = data.frame(scale(metabolomics_features_complete[-c(1,2)])))

```

###i. Early Integration 
```{r Early Integration,echo=F,include=T}
#Outcome - eGFR
#pfos, pfoa, pfna, pfhxs, pfpes
# G = exposure_dat[c("ma_pfos_scaled","ma_pfoa_scaled","ma_pfna_scaled","ma_pfhxs_scaled","ma_pfpes_scaled")] |> as.matrix()
# G = exposure_dat |> as.matrix()
G = pfas_features_complete|> as.matrix()
G = scale(G)
# Z = omics_complete[1:24] |> as.matrix()
Z = omics_complete |> as.matrix()
Z <- scale(Z)
Y <- eGFR_complete[1]|> as.matrix()
# Y <- rnorm(78,0,1)|> as.matrix()

fit1 <- estimate_lucid(lucid_model = "early",
                       G = G,
                       Z = Z,
                       Y = Y,
                       K = 2,
                       CoY = covs,
                       CoG = NULL,
                       Rho_Z_Mu = 1,
                       Rho_Z_Cov = 0.1,
                       useY = TRUE,
                       init_par = "mclust",
                       family = "normal",
                       init_omic.data.model = NULL)

fit2 <- reorder_lucid(fit1, "early",2)

# reorder_lucid <- function(lucidus_fit,lucid_model,
#                                 order = NULL)

# #Outcome - LCN2
# #pfda, pfna, pfhps, pfpes
# G = exposure_dat[c("ma_pfda_scaled","ma_pfna_scaled","ma_pfhps_scaled","ma_pfpes_scaled")] |> as.matrix()
# Z = omics_complete |> as.matrix()
# Z <- scale(Z)
# Y <- meta_filt$LCN2_log_scaled
# 
# 
# fit2 <- estimate_lucid(lucid_model = "early",
#                        G = G,
#                        Z = Z,
#                        Y = Y,
#                        K = 2,
#                        CoY = covs,
#                        CoG = NULL,
#                        Rho_Z_Mu = 0.1,
#                        Rho_Z_Cov = 0.005,
#                        useY = TRUE,
#                        init_par = "mclust",
#                        family = "normal",
#                        init_omic.data.model = NULL)


#Get PIPs from fit1 
meta_filt$pips <- fit1$inclusion.p[,2]


```


###i. Intermediate (Parallel) Integration 
```{r Intermediate Integration,echo=F,include=T}
#eGFR
# G = exposure_dat %>% as.matrix()
G = pfas_features_complete|> as.matrix()
# G = exposure_dat[c("ma_pfna_scaled")] |> as.matrix()
# Z = list(as.matrix(omics2_data)[-1],as.matrix(omics_data)[-1])
Z = list(as.matrix(microbiome_features_complete[-1]),as.matrix(metabolomics_features_complete[-1]))
Y = eGFR_complete

fit3 <- estimate_lucid(lucid_model = "parallel",
                       G = G, 
                       Z = Z, 
                       Y = Y,
                       K = rep(2, length(Z)), 
                       family = "normal", 
                       init_par="mclust",
                       max_itr = 200, 
                       useY = TRUE, 
                       init_omic.data.model  = "EII",
                       Rho_Z_Mu = 0.1,
                       Rho_Z_Cov = 0.0001) 

# Reorder the clusters
fit_reordered <- reorder_lucid_parallel(fit3, reference = c(2,2))

# #LCN2
# G = exposure_dat %>% as.matrix()
# Z = omics_lst[c(1:2)]
# Y = outcome_dat[2]
# 
# fit4 <- estimate_lucid(lucid_model = "parallel",
#                        G = G, 
#                        Z = Z, 
#                        Y = Y,
#                        K = rep(2, length(Z)), 
#                        family = "normal", 
#                        init_par="mclust",
#                        max_itr = 200, 
#                        useY = TRUE, 
#                        init_omic.data.model  = "EII",
#                        Rho_Z_Mu = 0.1,
#                        Rho_Z_Cov = 0.0001) 
# 
# # Reorder the clusters
# fit_reordered2 <- reorder_lucid_parallel(fit4, reference = c(2,2))
```


###i. Late (Serial) Integration 
```{r Late Integration,echo=F,include=T}
#eGFR
set.seed(100)
# covs[,1] <- scale(covs[,1])
# covs1 <- covs[,1]
# G = completedData[exposure] %>% as.matrix()
# G = exposure_dat[c("ma_pfos_scaled","ma_pfoa_scaled","ma_pfna_scaled","ma_pfhxs_scaled","ma_pfpes_scaled")] |> as.matrix()
# G = pfas_features_complete|> as.matrix()
# colnames(G) <- c("PFOS","PFOA","PFNA","PFHxS","PFPeS")
G = pfas_features_complete[4] |> as.matrix()
# colnames(G) <- c("PFOS","PFOA","PFNA","PFHxS","PFPeS")
# G <- G[,c(1,3)]
G <- scale(G)
# metabolomics_features_complete2 <- metabolomics_features_complete[sig.metabolites2]
Z = list(scale(as.matrix(omics_complete[omics2])),scale(as.matrix(omics_complete[omics])))
# Z = omics_lst
Y = completedData$eGFR_log_scaled
Y <- scale(Y)

# Run the LUCID Model  
fit <- estimate_lucid(lucid_model = "serial",
                      G = G,
                      Z = Z,
                      Y = Y ,
                      # CoY = NULL,
                      CoY = covs,
                      CoG = NULL,
                      K = list(2,2),
                      useY = TRUE,
                      family = "normal", 
                      init_par="mclust",
                      # Rho_Z_Mu = 5,
                      # Rho_Z_Cov = 0.3,
                      init_omic.data.model = NULL) 

#extract selected Z features to to refit lUCID in serial
selected_Z = vector(mode = "list", length = length(Z))
for (i in (1:length(Z))){
  fiti_select_Z = fit$submodel[[i]]$select$selectZ
  selected_Z[[i]] = Z[[i]][,fiti_select_Z]
}

#Refit ther LUCID in serial model with selected Z
fit <- estimate_lucid(lucid_model = "serial",
                      G = G,
                      Z = selected_Z,
                      Y = Y, 
                      # CoY = NULL,
                      CoY = covs,
                      CoG = NULL,
                      K = list(2,2),
                      useY = TRUE,
                      init_par = "mclust",
                      family = "normal",
                      init_omic.data.model = NULL)

# # Rename Exposure
fit1 <- fit$submodel[[1]]
# fit1$var.names$Gnames[1] <- "PFAS"
fit2 <- fit$submodel[[2]]
fit2$var.names$Gnames[1] <- "<b>Microbiome\nProfile 1</b>"
# fit3 <- fit$submodel[[3]]
# fit3$var.names$Gnames[1] <- "<b>Microbiome\nProfile 1</b>"

# fit1$res_Beta #pfas to microbiome
# fit2$res_Gamma$beta #metabolites to eGFR
# fit2$res_Beta #microbiome to metabolites
# 
fit1$res_Beta #1.874438517 -0.3702220566 -0.02550096442 -0.1810410163 0.2978761659 -0.1169606285
fit2$res_Gamma$beta #-1.3858956193  0.4320394754 -> -0.9538561439 -0.4320394754
fit2$res_Beta #1.410956514 -0.2826748594 -> -1.410956514 0.2826748594

fit1 = fit$submodel[[1]]
fit1 = reorder_lucid(fit1, "early", 2)
fit2 = fit2 <- fit$submodel[[2]]
fit2 = reorder_lucid(fit2, "serial_sub_with_sub_before_it_reordered", 2)
fit2 <- reorder_lucid(fit2, "early", 2)

fit$submodel[[1]] = fit1
fit$submodel[[2]] = fit2

rm(fit)
rm(fit1)  
rm(fit2)

meta_filt_microbiome <- meta_filt_microbiome %>% 
  filter(!is.na(Dorea_scaled))
#Just eGFR and microbiome
# meta_filt_microbiome["eGFR_log_scaled"]
G = meta_filt_microbiome[exposure_log_scaled]|> as.matrix()
# G = pfas_features_complete[1] |> as.matrix()
# colnames(G) <- c("PFOS","PFOA","PFNA","PFHxS","PFPeS")
# G <- G[,c(1,3)]
# G <- scale(G)

# Z = list(scale(as.matrix(omics_complete[omics2])),scale(as.matrix(omics_complete[omics])))
# Z = list(as.matrix(microbiome_features_complete[-1]),as.matrix(metabolomics_features_complete[-1]))
Z = list(as.matrix(meta_filt_microbiome[microbiome_scaled]))
# Y = eGFR_complete[1]
Y = meta_filt_microbiome["eGFR_log_scaled"]
# Y <- scale(Y)

# Run the LUCID Model  
fit <- estimate_lucid(lucid_model = "serial",
                      G = G,
                      Z = Z,
                      Y = FALSE ,
                      # CoY = NULL,
                      CoY = covs,
                      CoG = NULL,
                      K = list(2),
                      useY = TRUE,
                      family = "normal", 
                      init_par="mclust",
                      # Rho_Z_Mu = 5,
                      # Rho_Z_Cov = 0.3,
                      init_omic.data.model = NULL) 

#extract selected Z features to to refit lUCID in serial
selected_Z = vector(mode = "list", length = length(Z))
for (i in (1:length(Z))){
  fiti_select_Z = fit$submodel[[i]]$select$selectZ
  selected_Z[[i]] = Z[[i]][,fiti_select_Z]
}

#Refit ther LUCID in serial model with selected Z
fit <- estimate_lucid(lucid_model = "serial",
                      G = G,
                      Z = selected_Z,
                      Y = Y, 
                      # CoY = NULL,
                      CoY = covs,
                      CoG = NULL,
                      K = list(2,2),
                      useY = TRUE,
                      init_par = "mclust",
                      family = "normal",
                      init_omic.data.model = NULL)

# # Rename Exposure
fit1 <- fit$submodel[[1]]
# fit1$var.names$Gnames[1] <- "PFAS"
fit2 <- fit$submodel[[2]]
fit2$var.names$Gnames[1] <- "<b>Microbiome\nProfile 1</b>"
```


##d. JIVE and Mediation Analysis
```{r JIVE and Mediation Analysis,echo=F,include=T}
# Define exposure and outcome name
# exposure <- as.vector(completedData$ma_pfna_log_scaled)
# exposure <- as.vector(completedData$ma_pfhps_log_scaled)
exposure <- as.vector(completedData$burden_scaled)

# exposure <- meta_filt[exposure_log_scaled]|> as.matrix()
outcome  <- as.vector(completedData$eGFR3_log_scaled)

# Get numeric matrix of covariates 
# covs <- covs[,-6]
covs <- covs

# create list of omics data 
# omics_lst <- simulated_data[-which(names(simulated_data) == "phenotype")]
omics_lst <- omics_lst


# Create data frame of omics data
omics_df <- omics_lst |> 
  purrr::map(~as_tibble(.x, rownames = "name")) |>
  purrr::reduce(left_join, by = "name") |>
  column_to_rownames("name")



#Rename metabolites 
full_metabolites <- metabolites %>% 
  dplyr::select(name,key)
full_metabolites <- tidylog::left_join(full_metabolites,conf_ftrs,by="name")
#Select lower diff rt
get_min_name <- function(rt_diff, name) {
  rt_diff_values <- as.numeric(unlist(strsplit(rt_diff, "; ")))
  name_values <- unlist(strsplit(name, "; "))
  
  min_index <- which.min(rt_diff_values)
  min_name <- name_values[min_index]
  
  return(min_name)
}

# Apply the function to each row and create the new column
new_names <- mapply(get_min_name, full_metabolites$rt_diff, full_metabolites$name)
names(new_names) <- NULL
full_metabolites$new_names <- new_names
key <- full_metabolites %>% 
  dplyr::select(new_names,key)  
key <- key %>% 
  dplyr::distinct(new_names,.keep_all = T)
key <- key %>% 
  rename(feature=key)
panel_b_dat_top_ft_named <- tidylog::left_join(panel_b_dat_top_ft,key,by="feature")
panel_b_dat_top_ft_named <- panel_b_dat_top_ft_named %>% 
  mutate(feature=ifelse(!is.na(new_names),new_names,feature))
panel_b_dat_top_ft_named <- panel_b_dat_top_ft_named %>% 
  dplyr::select(-new_names)
panel_b_dat_top_ft <-panel_b_dat_top_ft_named

# # # Transpose list of matrices
omics_t <- lapply(omics_lst, t)
# 
# # Rename omics datasets
names(omics_t) = names(omics_lst)
# 
# #Data reduction step
jive1 <- jive(omics_t, rankJ = 5, rankA = c(5,5),
              method = "given",
              # conv = 1e-04,
              maxiter = 100,
              showProgress = FALSE)
# #
summary(jive1)
showVarExplained(jive1)

summary(result_jive2)
showVarExplained(result_jive2)


# # # Run Analysis
# result_med_with_latent_fctrs_JIVE <-
#   lafamum(exposure,
#           outcome,
#           omics_lst,
#           covs = covs,
#           Y.family = "gaussian",
#           fdr.level = 0.05,
#           integration = "intermediate",
#           jive.rankJ = 5,
#           jive.rankA = c(5,5))
# # 
# result_med_with_latent_fctrs_JIVE <- res
# plot <- (plot_lafamum(result_med_with_latent_fctrs_JIVE))
# plot
# significant_features_cor <- c(microbiome_features,metabolite_features)
# ggsave(plot=plot,fs::path(dir.results,"Mediation_Ranks5_40_features.jpeg"))


#Ranks=2, NULL
#Ranks=3, Metabolomics_3
#Ranks=4, Metabolomics_3 & Joint_2, N.S. FDR, 8 metabolites (294,291,288,287,279,277,275,139), 3 taxa(unclassified356, streptococcus, unclassified887), joint alpha = negative, beta = positive, var %>40%, metabolite pc alpha=negative, beta=positive, var>20%

#Ranks=5, Metabolomics_2 & Joint_2, 6 metabolites (88,313,312,214,199,172), 3 taxa (Romboutsia, Oscillibacter, Eubacterium halli), N.S. FDR,alpha=positive, beta=negative joint, alpha = positive, beta=negative metabolite pc,Joint var explains >40%, metabolite >20% 

#Ranks=6, Metabolomics_3 & Joint_2, 4 metabolites, 2 taxa, N.S. FDR, 4 metabolites(88,313,312,139), 2 taxa(unclassified356, streptococcus), joint pc alpha=negative,beta=positive, var explained >40%, met pc alpha=positive, beta=negative, varexplained>20%

#Ranks=7, metabolomics_3, joint_2, 4 metabolites(88,313,312,139), 2 taxa(unclassified356,streptococcus), joint pc alpha=negative, beta=psoitive, var>40%, metab alpha=positive,beta=negative,var>20%, N.S. FDR

#Ranks=8, microbiome_8, metabolomics_8, metabolomics_7, N.S. FDR, alphas and betas all in same directions

#Ranks=9, metabolomics_9

#Ranks=10, metabolomics_10

# 
# #Select top 30 loadings 
# head(sort(abs(result_med_with_latent_fctrs_JIVE[["result_ftr_cor_sig_pcs_jive"]]$Joint_2)),20)
# 
# #Joint
# top_40 <- result_med_with_latent_fctrs_JIVE[["result_ftr_cor_sig_pcs_jive"]][order(abs(result_med_with_latent_fctrs_JIVE[["result_ftr_cor_sig_pcs_jive"]]$Joint_2), decreasing = TRUE), ][1:40, ]
# features <- top_40$feature
# 
# #Metabolite
# top_20_metab <- result_med_with_latent_fctrs_JIVE[["result_ftr_cor_sig_pcs_jive"]][order(abs(result_med_with_latent_fctrs_JIVE[["result_ftr_cor_sig_pcs_jive"]]$metabolomics_2), decreasing = TRUE), ][1:20, ]
# features_metab <- top_20_metab$feature
# 

# 
# selected_metabolites <- top_40 %>% 
#   filter(grepl("metabolite",feature))
# selected_metabolites <- unique(selected_metabolites$feature)



#Select only metabolites & bacteria from jive components
joint_features <- panel_b_dat_top_ft %>% 
  filter(lf_num=="Joint_1")
microbiome_features <- joint_features %>% 
  filter(omic_layer=="microbiome")
mic_features <- microbiome_features$feature
metabolite_features_joint <- joint_features %>% 
  filter(omic_layer=="metabolomics")
joint_met_features <- metabolite_features_joint$feature #56

#Pathway analysis:
#Joint Factor
joint_factor<- factors_jive$Joint_1
# joint_factor_B <- factors_jive$Joint_2
completedData <- completedData %>% 
  dplyr::select(-metabolite10)

# microbiome_features[microbiome_features %in% panel_b_dat_top_ft$feature]
# selected_metabolites_matrix <- as.matrix(completedData[selected_metabolites])
# selected_metabolites_matrix <- as.matrix(completedData[metabolite_features[-1][metabolite_features[-1] %in% panel_b_dat_top_ft$feature]])
# selected_microbiome_matrix <- as.matrix(completedData[microbiome_features[microbiome_features %in% panel_b_dat_top_ft$feature]])


selected_metabolites_matrix <- as.matrix(completedData[joint_met_features])
selected_microbiome_matrix <- as.matrix(completedData[mic_features])


rownames(selected_metabolites_matrix) <- NULL
# selected_metabolites %in% omics
#Correlation
# cor(x, y, method = c("pearson", "kendall", "spearman"))
# cor.test(x, y, method=c("pearson", "kendall", "spearman"))

results_cor <- data.frame()
# for (i in 1:length(selected_metabolites)){
for (i in 1:length(joint_met_features)){
  correlation <- cor.test(selected_metabolites_matrix[,i],joint_factor,method="pearson")
  r <- round(correlation$estimate,3)
  t <- round(correlation$statistic,3)
  p <- round(correlation$p.value,5)
  # met <- selected_metabolites[i]
  met <- joint_met_features[i]
  corr.results <- data.frame(key=met,r=r,T_stat=t,PVal=p)
  results_cor <- rbind(results_cor,corr.results)
}

sig.met1 <- results_cor %>% 
  filter(PVal<0.05)
sig_metabolite_features1 <- sig.met1$key
#Signficiant Metabolites: "metabolite199" "metabolite172" "metabolite214" "metabolite139" "metabolite254" "metabolite96"  "metabolite232" "metabolite251"

# ipa_metabolites <- metabolites %>%
#   filter(key %in% results_cor$key)
# ipa_metabolites <- ipa_metabolites[,c(2,5)]
# metabolite.names <- ipa_metabolites$name
# ipa_data <- conf_ftrs %>%
#   filter(name %in% metabolite.names)
# ipa_data <- ipa_data[,c(12,20,30)] #Select the names and pubchem ids
# ipa_data <- tidylog::left_join(ipa_data,ipa_metabolites,by="name")
# ipa_data <- tidylog::left_join(ipa_data,results_cor,by="key")

#Microbiome features with joint facto
results_cor2 <- data.frame()
# for (i in 1:length(selected_metabolites)){
for (i in 1:length(mic_features)){
  correlation <- cor.test(selected_microbiome_matrix[,i],joint_factor,method="pearson")
  # r <- round(correlation$estimate,3)
  t <- round(correlation$statistic,3)
  p <- round(correlation$p.value,5)
  # met <- selected_metabolites[i]
  mic <- mic_features[i]
  corr.results2 <- data.frame(Taxa=mic,T_stat=t,PVal=p)
  results_cor2 <- rbind(results_cor2,corr.results2)
}

sig.mic <- results_cor2 %>% 
  filter(PVal<0.05)
sig_microbiome_features <- sig.mic$Taxa

joint_factor_features <- c(sig_metabolite_features1,sig_microbiome_features)
met_factor_features <- joint_met_features[!joint_met_features %in% joint_factor_features]
#Metabolomics Factor
met_factor <- factors_jive$metabolomics_5
# selected_metabolites_matrix <- as.matrix(completedData[selected_metabolites])
# selected_metabolites_matrix2 <- as.matrix(completedData[metabolite_features[-1]])
selected_metabolites_matrix2 <- as.matrix(completedData[met_factor_features])

rownames(selected_metabolites_matrix2) <- NULL
# selected_metabolites %in% omics
#Correlation
# cor(x, y, method = c("pearson", "kendall", "spearman"))
# cor.test(x, y, method=c("pearson", "kendall", "spearman"))
results_cor3 <- data.frame()
# for (i in 1:length(selected_metabolites)){
for (i in 1:length(met_factor_features)){
  correlation <- cor.test(selected_metabolites_matrix[,i],met_factor,method="pearson")
  r <- round(correlation$estimate,3)
  t <- round(correlation$statistic,3)
  p <- round(correlation$p.value,5)
  # met <- selected_metabolites[i]
  met <- met_factor_features[i]
  corr.results <- data.frame(key=met,r=r,T_stat2=t,PVal2=p)
  results_cor3 <- rbind(results_cor3,corr.results)
}

sig.met2 <- results_cor3 %>%
  filter(PVal2<0.05)
sig_metabolite_features2 <- sig.met2$key
sig_metabolites <- c(sig_metabolite_features1,sig_metabolite_features2)
key2 <-key %>% 
  filter(feature %in% sig_metabolites)
sig_metabolite_names <- key2$new_names
# significant_features_cor <- c(,sig_metabolite_features2)

sig_met_key <- metabolites %>%
  filter(key %in% sig_metabolites) %>% 
  dplyr::select(name,key)

significant_features_cor <- c(sig_microbiome_features,sig_met_key$name)
significant_features_cor <- str_replace(significant_features_cor,"Succinate; Succinic acid; Methylmalonate; Methylmalonic acid","Succinate")
significant_features_cor <- str_replace(significant_features_cor,"Ursocholic acid; Allocholic acid; Gama-Muricholic acid","Ursocholic acid")




#Coef plots
#Joint
joint_factor_features
key3 <-key %>% 
  filter(feature %in% joint_factor_features)
sig_metabolite_joint <- key3$new_names
joint_factor_features2 <- c(sig_microbiome_features,sig_metabolite_joint)
joint_factor_features2  <- str_remove(joint_factor_features2 ,"_scaled")
joint_factor_features2  <- str_replace(joint_factor_features2 ,"unclassified569","Lachnospiraceae g.569")
joint_factor_features2  <- str_replace(joint_factor_features2 ,"unclassified356","Oscillospiraceae g.356")


joint_results <- panel_b_dat_top_ft %>% 
  filter(lf_num=="Joint_1") %>% 
  filter(feature %in% joint_factor_features2) %>% 
  mutate(Type=ifelse(omic_layer=="microbiome","Microbiome","Metabolome"))
joint_results$feature <- factor(joint_results$feature,levels=joint_results$feature)
joint_results$Type <- factor(joint_results$Type,levels=c("Microbiome","Metabolome"))

plot1 <- ggplot(joint_results,aes(x=feature,y=Correlation))+
  geom_bar(stat = "identity")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  facet_grid(~Type,scale="free_x",space="free_x")+
  labs(y="Factor Loadings",
       x=NULL)


#Met
sig_metabolite_features2
key4 <-key %>% 
  filter(feature %in% sig_metabolite_features2)
sig_metabolite_met<- key4$new_names
# met_factor_features2 <- c(sig_microbiome_features,sig_metabolite_joint)
# met_factor_features2  <- str_remove(joint_factor_features2 ,"_scaled")
# met_factor_features2  <- str_replace(joint_factor_features2 ,"unclassified569","Lachnospiraceae g.569")
# met_factor_features2  <- str_replace(joint_factor_features2 ,"unclassified356","Oscillospiraceae g.356")


met_results <- panel_b_dat_top_ft %>% 
  filter(lf_num!="Joint_1") %>% 
  filter(feature %in% sig_metabolite_met) %>% 
  mutate(Type=ifelse(omic_layer=="microbiome","Microbiome","Metabolome"))
met_results$feature <- factor(met_results$feature,levels=met_results$feature)
met_results$Type <- factor(met_results$Type,levels=c("Microbiome","Metabolome"))

plot2 <- ggplot(met_results,aes(x=feature,y=Correlation))+
  geom_bar(stat = "identity")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  facet_grid(~Type,scale="free_x",space="free_x")+
  labs(y="Factor Loadings",
       x=NULL)


#Get pubchem ids
sig_met_key$name
key5 <- metabolites %>% 
  filter(key %in% sig_metabolite_features2)
key6 <- conf_ftrs %>% 
  filter(name %in% key5$name)
key6$pub_chem_cid

key7 <- metabolites %>% 
  filter(key %in% sig_metabolite_features1)
key8 <- conf_ftrs %>% 
  filter(name %in% key7$name)
key8$pub_chem_cid


ipa_data2 <- conf_ftrs %>%
  filter(name %in% metabolite.names2)


sig_met <- c(sig_metabolite_features1)
sig_met_key <- metabolites %>%
  filter(key %in% sig_met)

ipa_metabolites2 <- metabolites %>%
  filter(key %in% results_cor$key)
ipa_metabolites2 <- sig_met_key[,c(2,5)]
metabolite.names2 <- ipa_metabolites2$name
ipa_data2 <- conf_ftrs %>%
  filter(name %in% metabolite.names2)
ipa_data2 <- ipa_data2[,c(12,20,30)] #Select the names and pubchem ids
ipa_data2 <- tidylog::left_join(ipa_data2,ipa_metabolites2,by="name")
ipa_data2 <- tidylog::left_join(ipa_data2,results_cor2,by="key")
ipa_formatted_data <- tidylog::full_join(ipa_data,ipa_data2,by=c("name","pub_chem_cid","key","rt_diff"))

#Select smaller delta rt for multiples
ipa_formatted_data2 <- ipa_formatted_data %>% 
  separate(rt_diff, c("diff1", "diff2","diff3","diff4","diff5","diff6","diff7"), sep=";",fill="right")
diff <- c("diff1", "diff2","diff3","diff4","diff5","diff6","diff7")
ipa_formatted_data2 <- ipa_formatted_data2 %>% 
  dplyr::mutate(across(diff,~as.numeric(.)))

diff <- ipa_formatted_data2 %>% 
  dplyr::select(diff1:diff7) 

find_min <- function(row) {
  min_value <- min(row, na.rm = TRUE)
  min_name <- names(row)[which.min(ifelse(is.na(row), Inf, row))]
  return(c(min_value, min_name))
}

# Apply function to find the minimum value and column name for each row
result <- t(apply(diff, 1, find_min))
result_df <- data.frame(min_diff = as.numeric(result[, 1]), diff_name = result[, 2])

# Add the result to the original dataframe
ipa_formatted_data3 <- cbind(ipa_formatted_data2, result_df)  

ipa_formatted_data3$index <- as.numeric(str_remove(ipa_formatted_data3$diff_name,"diff"))
ipa_formatted_data4 <- ipa_formatted_data3 %>% 
  dplyr::select(name) %>% 
  mutate(new_name=name) %>% 
  separate(new_name, c("name1", "name2","name3","name4","name5","name6","name7"), sep=";",fill="right",)
index <- ipa_formatted_data3$index
ipa_formatted_data4 <- cbind(ipa_formatted_data4,index)
ipa_formatted_data5 <- ipa_formatted_data4 %>% 
  mutate(selected_name=case_when(index==1~name1,
                                 index==2~name2,
                                 index==3~name3,
                                 index==4~name4,
                                 index==5~name5,
                                 index==6~name6,
                                 index==7~name7))
selected_name <- ipa_formatted_data5[,c(1,10)]
ipa_formatted_data <- tidylog::full_join(ipa_formatted_data,selected_name,by="name")
ipa_formatted_data <- ipa_formatted_data %>% 
  dplyr::select(selected_name,pub_chem_cid,T_stat:PVal2)

#Pubchem ids
ipa_formatted_data6 <- ipa_formatted_data3 %>% 
  dplyr::select(pub_chem_cid) %>% 
  mutate(new_pub_chem_cid=pub_chem_cid) %>% 
  separate(new_pub_chem_cid, c("name1", "name2","name3","name4","name5","name6","name7"), sep=";",fill="right",)
index <- ipa_formatted_data3$index
ipa_formatted_data7 <- cbind(ipa_formatted_data6,index)
ipa_formatted_data8 <- ipa_formatted_data7 %>% 
  mutate(selected_name=case_when(index==1~name1,
                                 index==2~name2,
                                 index==3~name3,
                                 index==4~name4,
                                 index==5~name5,
                                 index==6~name6,
                                 index==7~name7))

ipa_formatted_data8 <- ipa_formatted_data8 %>% 
  rename(select_pub_chem_id=selected_name)
selected_pubchem <- ipa_formatted_data8 %>% 
  dplyr::select(pub_chem_cid,select_pub_chem_id)
ipa_formatted_data <- tidylog::full_join(ipa_formatted_data,selected_pubchem,by="pub_chem_cid")
ipa_formatted_data <- ipa_formatted_data %>% 
  dplyr::select(selected_name,select_pub_chem_id,T_stat:PVal2)

ipa_formatted_data <- ipa_formatted_data %>% 
  rename(Metabolite=selected_name) %>% 
  rename(pub_chem_cid=select_pub_chem_id)
# write.csv(ipa_formatted_data,fs::path(dir.results,"IPA_Formatted_Data_v2.csv"))
sig.met <- ipa_formatted_data %>% 
  filter(PVal<0.05 | PVal2<0.05)
sig_metabolite_features <- sig.met$Metabolite

# ipa_data <- ipa_data[-8,]
# hydroxy1 <- data.frame(name="7-Hydroxycholesterol",pub_chem_cid=473141)
# hydroxy2 <- data.frame(name="7a-Hydroxycholesterol",pub_chem_cid=53477732)
# ipa_data<- rbind(ipa_data,hydroxy1)
# ipa_data <- rbind(ipa_data,hydroxy2)
# write.csv(ipa_data,fs::path(dir.results,"IPA_Formatted_Data_v1.csv"))


result_med_with_latent_fctrs_JIVE <- res
plot <- (plot_lafamum(result_med_with_latent_fctrs_JIVE))
plot

# significant_features_cor<- c(microbiome_features,metabolite_features)
```


##e. Pathway analysis 
```{r pathway analysis, echo=F}
metabolite.data <- KEGG_Id_Data %>% 
  janitor::clean_names() 
# na.omit()
corr.data <- IPA_Formatted_Data_v2 %>% 
  janitor::clean_names()
corr.data <- corr.data %>% 
  dplyr::select(pubchem_id,hmdb:p_value2)
metabolite.data <- tidylog::left_join(corr.data,metabolite.data,by=c("hmdb","pubchem_id"))
metabolite.data <- metabolite.data %>% 
  na.omit() 
kegg_ids <- metabolite.data %>% 
  dplyr::select(p_value1,p_value2)
kegg_ids <- as.matrix(kegg_ids)
rownames(kegg_ids) <- metabolite.data$kegg


# kegg_ids <- kegg_ids %>% 
#   dplyr::select(-kegg)
# kegg_ids <- c(metabolite.data$p_value1)
# names(kegg_ids) <- metabolite.data$kegg

pathways <- keggList("pathway", "hsa")
metabolic_pathways <- pathways[grep("Metabolic pathways", pathways, ignore.case = TRUE)]
pathway_id <- "hsa01100"
#Pathway analysis
pv.out <- pathview(gene.data = NULL, cpd.data = kegg_ids, pathway.id = pathway_id, species = "hsa",cpd.idtype = "kegg")
data(demo.paths)
keggview.graph(plot.data.gene=pv.out$plot.data.gene)

demo.paths$sel.paths[1]
data(gse16873.d)

pv.out <- pathview(gene.data = gse16873.d[, 1], pathway.id = demo.paths$sel.paths[1],
                   species = "hsa", out.suffix = "gse16873", kegg.native = T)
head(pv.out$plot.data.gene)

pathway_info <- pv.out$plot.data.gene

# Use keggview.graph on pathway_info
keggview.graph(pathway_info,path.graph = pathway_info)
```

##f. High Dimensional Multiomic Mediation Analysis
```{r High Dimensional Multiomic Mediation, echo=F}
# Define exposure and outcome name
exposure <- as.vector(completedData$ma_pfna_log_scaled)
# exposure <- meta_filt[exposure_log_scaled]|> as.matrix()
outcome  <- as.vector(completedData$eGFR_log_scaled)

# Get numeric matrix of covariates 
covs <- covs

# create list of omics data 
# omics_lst <- simulated_data[-which(names(simulated_data) == "phenotype")]
omics_lst <- omics_lst

# Run Analysis
result_hima_intermediate <- hidimum(omics_lst = omics_lst,
                                    covs = covs,
                                    outcome = outcome,
                                    exposure = exposure,
                                    n_boot = 12,
                                    Y.family = "gaussian", 
                                    integration = "intermediate")

# plot
plot_hidimum(result_hima_intermediate)

```


#g. Mediation Analysis 
```{r Mediation ANalysis, echo=F}
# install.packages("mediation")
library(mediation)

full_data$joint_factor <- factors_jive$Joint_1
full_data$met_factor <- factors_jive$metabolomics_5
sd(full_data$joint_factor)
sd(full_data$met_factor)
# joint_factor_A <- factors_jive$Joint_2
# joint_factor_B <- factors_jive$Joint_2
full_data$ma_creatinine_scaled
full_data$burden <- meta_filt$eap.pfas
full_data$ma_pfhps_scaled <- meta_filt$ma_pfhps_scaled
full_data$ma_pfna_scaled <- meta_filt$ma_pfna_scaled


# full_data$burden_scaled <- as.numeric(full_data$burden_scaled)
# full_data$burden_scaled <- as.numeric(full_data$burden_scaled)
model.0 <- lm(eGFR3_log_scaled ~ ma_pfna_log_scaled, full_data)
summary(model.0)

model.M <- lm(joint_factor ~ ma_pfna_log_scaled, full_data)
summary(model.M)

model.Y <- lm(eGFR3_log_scaled ~ ma_pfna_log_scaled + joint_factor, full_data)
summary(model.Y)

results <- mediate(model.M, model.Y, treat='ma_pfna_log_scaled', mediator='joint_factor',
                   boot=F, sims=500)
plot(results)
summary(results)

full_data$burden_scaled <- as.numeric(full_data$burden_scaled)
# devtools::install_github("BS1125/CMAverse",force=T)
library(CMAverse)
# full_data$burden_scaled
# full_data$burden <- meta_filt$eap.pfas


full_data$metabolite254
set.seed(500)

med <- cmest(data=full_data,model="rb",outcome="eGFR2_log",exposure="burden_scaled",a=2,astar=-2,basec = c("ma_age", "ma_sex", "ma_race_eth","ma_parental_edu","ma_creatinine_scaled"),
             mediator="metabolite254",EMint=F,mval=list(1),mreg = list("linear"), yreg = "linear",
             estimation = "paramfunc", inference = "bootstrap", nboot =500)

summary(med)


med <- cmest(data=full_data,model="rb",outcome="eGFR2_log",exposure="burden_scaled",a=2,astar=-2,basec = c("ma_age", "ma_sex", "ma_race_eth","ma_parental_edu","ma_creatinine_scaled","joint_factor"),
             mediator="met_factor",EMint=F,mval=list(1),mreg = list("linear"), yreg = "linear",
             estimation = "paramfunc", inference = "bootstrap", nboot =500)

summary(med)

med <- cmest(data=full_data,model="rb",outcome="eGFR2_log",exposure="burden_scaled",a=2,astar=-2,basec = c("ma_age", "ma_sex", "ma_race_eth","ma_parental_edu","ma_creatinine_scaled","met_factor"),
             mediator="joint_factor",EMint=F,mval=list(1),mreg = list("linear"), yreg = "linear",
             estimation = "paramfunc", inference = "bootstrap", nboot =500)

summary(med)


#PFHpS
med <- cmest(data=full_data,model="rb",outcome="eGFR2_log",exposure="ma_pfhps_scaled",a=2,astar=-2,basec = c("ma_age", "ma_sex", "ma_race_eth","ma_parental_edu","ma_creatinine_scaled","joint_factor"),
             mediator="met_factor",EMint=F,mval=list(1),mreg = list("linear"), yreg = "linear",
             estimation = "paramfunc", inference = "bootstrap", nboot =500)

summary(med)

med <- cmest(data=full_data,model="rb",outcome="eGFR2_log",exposure="ma_pfhps_scaled",a=2,astar=-2,basec = c("ma_age", "ma_sex", "ma_race_eth","ma_parental_edu","ma_creatinine_scaled","met_factor"),
             mediator="joint_factor",EMint=F,mval=list(1),mreg = list("linear"), yreg = "linear",
             estimation = "paramfunc", inference = "bootstrap", nboot =500)
summary(med)

#PFNA
med <- cmest(data=full_data,model="rb",outcome="eGFR2_log",exposure="ma_pfna_scaled",a=2,astar=-2,basec = c("ma_age", "ma_sex", "ma_race_eth","ma_parental_edu","ma_creatinine_scaled","joint_factor"),
             mediator="met_factor",EMint=F,mval=list(1),mreg = list("linear"), yreg = "linear",
             estimation = "paramfunc", inference = "bootstrap", nboot =500)

summary(med)
med <- cmest(data=full_data,model="rb",outcome="eGFR2_log",exposure="ma_pfna_scaled",a=2,astar=-2,basec = c("ma_age", "ma_sex", "ma_race_eth","ma_parental_edu","ma_creatinine_scaled","met_factor"),
             mediator="joint_factor",EMint=F,mval=list(1),mreg = list("linear"), yreg = "linear",
             estimation = "paramfunc", inference = "bootstrap", nboot =500)
summary(med)

#Fiber
full_data$m_total_dietary_fiber_g2 <- covs_complete$m_total_dietary_fiber_g

med <- cmest(data=full_data,model="rb",outcome="eGFR2_log",exposure="burden_scaled",a=2,astar=-2,basec = c("ma_age", "ma_sex", "ma_race_eth","ma_parental_edu","ma_creatinine_scaled","joint_factor","m_total_dietary_fiber_g2"),
             mediator="met_factor",EMint=F,mval=list(1),mreg = list("linear"), yreg = "linear",
             estimation = "paramfunc", inference = "bootstrap", nboot =500)

summary(med)


med <- cmest(data=full_data,model="rb",outcome="eGFR2_log",exposure="burden_scaled",a=2,astar=-2,basec = c("ma_age", "ma_sex", "ma_race_eth","ma_parental_edu","ma_creatinine_scaled","met_factor","m_total_dietary_fiber_g2"),
             mediator="joint_factor",EMint=F,mval=list(1),mreg = list("linear"), yreg = "linear",
             estimation = "paramfunc", inference = "bootstrap", nboot =500)

summary(med)


# Create the data frame with the updated data for specific terms
coefficients <- data.frame(
  Term = c("te", "tnie", "pnde"),
  Estimate = c(-4.58004, -0.96006, -3.61998),
  # Std.error = c(1.82677, 0.67183, 1.94587),
  CIL = c(-8.14668, -2.34884, -7.67258),
  CIU = c(-1.035, 0.133, 0.148),
  P_val=c(0.012,0.076,0.076)
)

# Ensure the terms are factors for ordered plotting
coef_data$Term <- factor(coef_data$Term, levels = coef_data$Term)
coef_data$Significance <- ifelse(coef_data$P.val < 0.05, "Significant", "Not Significant")

# Create the coefficient plot
ggplot(coef_data, aes(x = Term, y = Estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = CI_Lower, ymax = CI_Upper), width = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed",size=0.05) +
  labs(title = "Coefficient Plot",
       x = "Terms",
       y = "Estimate") +
  theme_bw() +
  coord_flip()  # Flip coordinates for better readability

p3 <- ggplot(coef_data, aes(x = Term, y = Estimate, color = Significance)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed",color="lightgray") +
  geom_errorbar(aes(ymin = CI_Lower, ymax = CI_Upper), width = 0.2) +
  scale_color_manual(values = c("Significant" = "red", "Not Significant" = "black")) +
  labs(title = "Coefficient Plot for Selected Terms",
       x = "Effects",
       y = "Estimate",
       color = "Significance") +
  theme_bw() +
  coord_flip() 


summary(lm(eGFR3~burden_scaled+ma_age+ma_sex+ma_parental_edu+ma_race_eth,data=full_data))

```


###h. Name metabolites
```{r Key Metabolties,echo=F}
joint_factor <- factors_jive$Joint_2
completedData <- completedData %>% 
  dplyr::select(-metabolite10)
selected_metabolites_matrix <- as.matrix(completedData[metabolite_features[-1]])
selected_microbiome_matrix <- as.matrix(completedData[microbiome_features])

rownames(selected_metabolites_matrix) <- NULL

results_cor <- data.frame()
for (i in 1:length(metabolite_features[-1])){
  correlation <- cor.test(selected_metabolites_matrix[,i],joint_factor,method="pearson")
  r <- round(correlation$estimate,3)
  t <- round(correlation$statistic,3)
  p <- round(correlation$p.value,5)
  # met <- selected_metabolites[i]
  met <- metabolite_features[-1][i]
  corr.results <- data.frame(key=met,r=r,T_stat=t,PVal=p)
  results_cor <- rbind(results_cor,corr.results)
}

sig.met1 <- results_cor %>% 
  filter(PVal<0.05)
sig_metabolite_features1 <- sig.met1$key


ipa_metabolites <- metabolites %>%
  filter(key %in% results_cor$key)
ipa_metabolites <- ipa_metabolites[,c(2,5)]
metabolite.names <- ipa_metabolites$name
ipa_data <- conf_ftrs %>%
  filter(name %in% metabolite.names)
ipa_data <- ipa_data[,c(12,20,30)] #Select the names and pubchem ids
ipa_data <- tidylog::left_join(ipa_data,ipa_metabolites,by="name")
ipa_data <- tidylog::left_join(ipa_data,results_cor,by="key")

#Microbiome features with joint facto
results_cor2 <- data.frame()
new_features <- microbiome_features[which(microbiome_features %in% panel_b_dat_top_ft$feature)]
# for (i in 1:length(selected_metabolites)){
for (i in 1:length(new_features)){
  correlation <- cor.test(selected_microbiome_matrix[,i],joint_factor,method="pearson")
  r <- round(correlation$estimate,3)
  t <- round(correlation$statistic,3)
  p <- round(correlation$p.value,5)
  # met <- selected_metabolites[i]
  mic <-new_features[i]
  corr.results2 <- data.frame(Taxa=mic,r=r,T_stat=t,PVal=p)
  results_cor2 <- rbind(results_cor2,corr.results2)
}

sig.mic <- results_cor2 %>% 
  filter(PVal<0.05)
# sig_microbiome_features <- sig.mic$Taxa[c(4,5)]
# sig_metabolite_features1

sig_met <- c(sig_metabolite_features1)
sig_met_key <- metabolites %>% 
  filter(key %in% sig_met) %>% 
  mutate(name=str_replace(name,"7-Hydroxycholesterol; 7a-Hydroxycholesterol","7a-Hydroxycholesterol"))
sig_met_key <- sig_met_key %>% 
  dplyr::select(name,key)
Signficiant_Metabolite_results <- tidylog::left_join(sig.met1,sig_met_key,by="key")

Signficiant_Metabolite_results$r <- sort(Signficiant_Metabolite_results$r)
order <- Signficiant_Metabolite_results$name
Signficiant_Metabolite_results$name <- factor(Signficiant_Metabolite_results$name,levels=order)
Signficiant_Metabolite_results <- Signficiant_Metabolite_results %>% 
  mutate(Type=ifelse(grepl("scaled",name),"Microbiome","Metabolome"))
results_plot <- Signficiant_Metabolite_results %>% 
  dplyr::select(-key) %>% 
  dplyr::select(name,r:PVal)
sig.mic <- sig.mic %>% 
  rename(name=Taxa)
rownames(sig.mic) <- NULL
results_plot <- rbind(sig.mic,results_plot)
results_plot <- results_plot %>% 
  mutate(Type=ifelse(grepl("scaled",name),"Microbiome","Metabolome")) %>% 
  mutate(name=str_replace(name,"_scaled",""))
sig.mic <- sig.mic %>%
  mutate(Type=ifelse(grepl("scaled",name),"Microbiome","Metabolome"))
Signficiant_Metabolite_results <- Signficiant_Metabolite_results %>% 
  mutate(Type=ifelse(grepl("scaled",name),"Microbiome","Metabolome"))

results_plot$Type <- factor(results_plot$Type, levels = c("Microbiome", "Metabolome"))
# results_plot$sign <- ifelse(results_plot$r > 0, "positive", "negative")
results_plot <- results_plot %>% 
  mutate(name=str_replace(name,"Eubacterium_siraeum_group","Eubacterium siraeum group"),
         name=str_replace(name,"Fusicatenibacter","Fusicatenibacter"),
         name=str_replace(name,"Dorea","Dorea"),
         name=str_replace(name,"Lachnospiraceae_g__GCA_900066575","GCA 900066575"))
results_plot$name <- factor(results_plot$name,levels=results_plot$name)

plot1 <- ggplot(results_plot,aes(x=name,y=r))+
  geom_bar(stat = "identity")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  facet_grid(~Type,scale="free_x",space="free_x")+
  labs(y="Factor Loadings",
       x=NULL)
# scale_fill_manual(values = c("positive" = "orange", "negative" = "blue"))

a <- ggplot(Signficiant_Metabolite_results,aes(x=name,y=r))+
  geom_bar(stat = "identity")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        axis.ticks.y=element_blank(),
        axis.text.y=element_blank())+
  facet_wrap("Type")+
  labs(y=NULL,
       x=NULL)

sig.mic <- sig.mic %>% 
  mutate(name=str_replace(name,"Eubacterium_siraeum_group_scaled","Eubacterium siraeum group"),
         name=str_replace(name,"Fusicatenibacter_scaled","Fusicatenibacter"),
         name=str_replace(name,"Dorea_scaled","Dorea"),
         name=str_replace(name,"Lachnospiraceae_g__GCA_900066575_scaled","GCA 900066575"))
b <-ggplot(sig.mic,aes(x=name,y=r))+
  geom_bar(stat = "identity",width=0.5)+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  facet_wrap("Type")+
  labs(y="Factor Loadings",
       x=NULL)
# install.packages("ggpubr")
library(ggpubr)
# install.packages("patchwork")
library(patchwork)

combined_plot <- ggarrange(b,a, ncol = 2, nrow = 1)
gridExtra::grid.arrange(b,a,ncol=2)
combined_plot <- plot_grid(b,a,rel_widths = c(1,2),rel_heights = c(1.5,1))








#Match metabolites to IDs for pathway analysis
key <- metabolites %>% 
  dplyr::select(name,key)
metabolites_matching <- tidylog::left_join(results_cor,key,by="key")
metabolites_matching <- tidylog::left_join(metabolites_matching,conf_ftrs,by="name") 
#Select lower diff rt
get_min_name <- function(rt_diff, name) {
  rt_diff_values <- as.numeric(unlist(strsplit(rt_diff, "; ")))
  name_values <- unlist(strsplit(name, "; "))
  
  min_index <- which.min(rt_diff_values)
  min_name <- name_values[min_index]
  
  return(min_name)
}

# Apply the function to each row and create the new column
new_names <- mapply(get_min_name, metabolites_matching$rt_diff, metabolites_matching$name)
names(new_names) <- NULL
metabolites_matching$new_names <- new_names

new_ids <- mapply(get_min_name, metabolites_matching$rt_diff, metabolites_matching$pub_chem_cid)
names(new_ids) <- NULL
metabolites_matching$new_pub_chem_cids <- new_ids


metabolites_matching <- metabolites_matching %>% 
  dplyr::select(r:name,pub_chem_cid,new_names,new_pub_chem_cids)
metabolites_matching <- metabolites_matching %>% 
  rename(pubchem_id=new_pub_chem_cids,
         metabolites=new_names) %>% 
  dplyr::select(metabolites,pubchem_id,r:PVal)
write.csv(metabolites_matching,fs::path(dir.results,"Pathway_Metabolites.csv"))

```

#New results
```{r new results,echo=F}
joint_factor <- factors_jive$Joint_2
completedData <- completedData %>% 
  dplyr::select(-metabolite10)
selected_metabolites_matrix <- as.matrix(completedData[metabolite_features[-1]])
selected_microbiome_matrix <- as.matrix(completedData[microbiome_features])

rownames(selected_metabolites_matrix) <- NULL

results_cor <- data.frame()
for (i in 1:length(metabolite_features[-1])){
  correlation <- cor.test(selected_metabolites_matrix[,i],joint_factor,method="pearson")
  r <- round(correlation$estimate,3)
  t <- round(correlation$statistic,3)
  p <- round(correlation$p.value,5)
  # met <- selected_metabolites[i]
  met <- metabolite_features[-1][i]
  corr.results <- data.frame(key=met,r=r,T_stat=t,PVal=p)
  results_cor <- rbind(results_cor,corr.results)
}

sig.met1 <- results_cor %>% 
  filter(PVal<0.05)
sig_metabolite_features1 <- sig.met1$key


ipa_metabolites <- metabolites %>%
  filter(key %in% results_cor$key)
ipa_metabolites <- ipa_metabolites[,c(2,5)]
metabolite.names <- ipa_metabolites$name
ipa_data <- conf_ftrs %>%
  filter(name %in% metabolite.names)
ipa_data <- ipa_data[,c(12,20,30)] #Select the names and pubchem ids
ipa_data <- tidylog::left_join(ipa_data,ipa_metabolites,by="name")
ipa_data <- tidylog::left_join(ipa_data,results_cor,by="key")

#Microbiome features with joint factor
results_cor2 <- data.frame()
new_features <- microbiome_features[which(microbiome_features %in% panel_b_dat_top_ft$feature)]
# for (i in 1:length(selected_metabolites)){
for (i in 1:length(new_features)){
  correlation <- cor.test(selected_microbiome_matrix[,i],joint_factor,method="pearson")
  r <- round(correlation$estimate,3)
  t <- round(correlation$statistic,3)
  p <- round(correlation$p.value,5)
  # met <- selected_metabolites[i]
  mic <-new_features[i]
  corr.results2 <- data.frame(Taxa=mic,r=r,T_stat=t,PVal=p)
  results_cor2 <- rbind(results_cor2,corr.results2)
}

sig.mic <- results_cor2 %>% 
  filter(PVal<0.05)
sig_microbiome_features <- sig.mic$Taxa
# sig_metabolite_features1

sig_met <- c(sig_metabolite_features1)
sig_met_key <- metabolites %>% 
  filter(key %in% sig_met) %>% 
  mutate(name=str_replace(name,"7-Hydroxycholesterol; 7a-Hydroxycholesterol","7a-Hydroxycholesterol"))
sig_met_key <- sig_met_key %>% 
  dplyr::select(name,key)
Signficiant_Metabolite_results <- tidylog::left_join(sig.met1,sig_met_key,by="key")


met_factor <- factors_jive$metabolomics_1
# completedData <- completedData %>% 
#   dplyr::select(-metabolite10)
# selected_metabolites_matrix <- as.matrix(completedData[metabolite_features[-1]])
# selected_microbiome_matrix <- as.matrix(completedData[microbiome_features])

# rownames(selected_metabolites_matrix) <- NULL

results_cor3 <- data.frame()
for (i in 1:length(metabolite_features[-1])){
  correlation <- cor.test(selected_metabolites_matrix[,i],met_factor,method="pearson")
  r <- round(correlation$estimate,3)
  t <- round(correlation$statistic,3)
  p <- round(correlation$p.value,5)
  # met <- selected_metabolites[i]
  met <- metabolite_features[-1][i]
  corr.results3 <- data.frame(key=met,r=r,T_stat=t,PVal=p)
  results_cor3 <- rbind(results_cor3,corr.results3)
}

sig.met2 <- results_cor3 %>% 
  filter(PVal<0.05)
sig_metabolite_features2 <- sig.met2$key

ipa_metabolites2 <- metabolites %>%
  filter(key %in% results_cor3$key)
ipa_metabolites2 <- ipa_metabolites2[,c(2,5)]
metabolite.names2 <- ipa_metabolites2$name
ipa_data2 <- conf_ftrs %>%
  filter(name %in% metabolite.names2)
ipa_data2 <- ipa_data2[,c(12,20,30)] #Select the names and pubchem ids
ipa_data2 <- tidylog::left_join(ipa_data2,ipa_metabolites2,by="name")
ipa_data2 <- tidylog::left_join(ipa_data2,results_cor3,by="key")

sig_met2 <- c(sig_metabolite_features2)
sig_met_key2 <- metabolites %>% 
  filter(key %in% sig_met2) 
# mutate(name=str_replace(name,"7-Hydroxycholesterol; 7a-Hydroxycholesterol","7a-Hydroxycholesterol"))
sig_met_key2 <- sig_met_key2 %>% 
  dplyr::select(name,key)
Signficiant_Metabolite_results2 <- tidylog::left_join(sig.met2,sig_met_key2,by="key")
Signficiant_Metabolite_results_all <- rbind(Signficiant_Metabolite_results,Signficiant_Metabolite_results2)

sig.mic <- sig.mic %>% 
  rename(name=Taxa)
Signficiant_Metabolite_results <- Signficiant_Metabolite_results %>% 
  dplyr::select(name,r:PVal)
joint_results <- rbind(sig.mic,Signficiant_Metabolite_results)
rownames(joint_results) <-NULL 
joint_results$Type <- ifelse(grepl("scaled",joint_results$name),"Microbiome","Metabolome")
joint_results$r <- sort(joint_results$r)
order <- joint_results$name
joint_results$name <- factor(joint_results$name,levels=order)
joint_results$Type <- factor(joint_results$Type, levels = c("Microbiome", "Metabolome"))

panel_b_dat_top_ft_joint <- panel_b_dat_top_ft %>% 
  filter(lf_num=="Joint_2")
panel_b_dat_top_ft_joint$key <- panel_b_dat_top_ft_joint$feature
key <- metabolites %>% 
  dplyr::select(name,key)
panel_b_dat_top_ft_joint <- tidylog::left_join(panel_b_dat_top_ft_joint,key,by="key")

#Coef plot metabolomics component
key <- metabolites %>% 
  dplyr::select(name,key)
# metabolites_matching <- tidylog::left_join(panel_b_dat_top_ft_joint,key,by=c("key"))
metabolites_matching <- tidylog::left_join(panel_b_dat_top_ft_joint,conf_ftrs,by="name") 
metabolites_matching <- metabolites_matching %>% 
  filter(!is.na(name))
#Select lower diff rt
get_min_name <- function(rt_diff, name) {
  rt_diff_values <- as.numeric(unlist(strsplit(rt_diff, "; ")))
  name_values <- unlist(strsplit(name, "; "))
  
  min_index <- which.min(rt_diff_values)
  min_name <- name_values[min_index]
  
  return(min_name)
}

# Apply the function to each row and create the new column
new_names <- mapply(get_min_name, metabolites_matching$rt_diff, metabolites_matching$name)
names(new_names) <- NULL
metabolites_matching$new_names <- new_names

# new_ids <- mapply(get_min_name, metabolites_matching$rt_diff, metabolites_matching$pub_chem_cid)
# names(new_ids) <- NULL
# metabolites_matching$new_pub_chem_cids <- new_ids
new <- metabolites_matching %>% 
  dplyr::select(feature,new_names)
panel_b_dat_top_ft_joint <- tidylog::left_join(panel_b_dat_top_ft_joint,new,by="feature")
panel_b_dat_top_ft_joint <- panel_b_dat_top_ft_joint %>% 
  mutate(feature=ifelse(!is.na(new_names),new_names,feature))
panel_b_dat_top_ft_joint <- panel_b_dat_top_ft_joint %>% 
  dplyr::select(-c(key,name,new_names))



get_names <- metabolites %>% 
  filter(key %in% sig_metabolite_features1 )
names <- get_names$name
names <- str_replace(names,"7-Hydroxycholesterol; 7a-Hydroxycholesterol","7a-Hydroxycholesterol")

filt <- c(sig_microbiome_features,names)
#Remove non-significant from correlations
joint_results <- panel_b_dat_top_ft_joint %>% 
  filter(feature %in% filt)
joint_results$Type <- ifelse(grepl("scale",joint_results$feature),"Microbiome","Metabolome")

joint_results$Correlation <- sort(joint_results$Correlation)
order <- joint_results$feature
joint_results$feature <- factor(joint_results$feature,levels=order)
joint_results$Type <- factor(joint_results$Type, levels = c("Microbiome", "Metabolome"))

# joint_results <- panel_b_dat_top_ft_joint
#Coef plot jiont component
plot1 <- ggplot(joint_results,aes(x=feature,y=Correlation))+
  geom_bar(stat = "identity")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  facet_grid(~Type,scale="free_x",space="free_x")+
  labs(y="Factor Loadings",
       x=NULL)








# Signficiant_Metabolite_results2 <- Signficiant_Metabolite_results2 %>% 
#   dplyr::select(name,r:PVal)
met_results <- Signficiant_Metabolite_results2 
rownames(met_results) <-NULL 
met_results$Type <- ifelse(grepl("scaled",met_results$new_name),"Microbiome","Metabolome")
met_results$r <- sort(met_results$r)
order <- met_results$new_name
met_results$new_name <- factor(met_results$new_name,levels=order)
met_results$Type <- factor(met_results$Type, levels = c("Microbiome", "Metabolome"))

#Coef plot jiont component
plot2 <- ggplot(met_results,aes(x=new_name,y=r))+
  geom_bar(stat = "identity")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  facet_grid(~Type,scale="free_x",space="free_x")+
  labs(y="Factor Loadings",
       x=NULL)










#Match metabolites to IDs for pathway analysis
key <- metabolites %>% 
  dplyr::select(name,key)
metabolites_matching <- tidylog::left_join(results_cor,key,by="key")
metabolites_matching <- tidylog::left_join(metabolites_matching,conf_ftrs,by="name") 
#Select lower diff rt
get_min_name <- function(rt_diff, name) {
  rt_diff_values <- as.numeric(unlist(strsplit(rt_diff, "; ")))
  name_values <- unlist(strsplit(name, "; "))
  
  min_index <- which.min(rt_diff_values)
  min_name <- name_values[min_index]
  
  return(min_name)
}

# Apply the function to each row and create the new column
new_names <- mapply(get_min_name, metabolites_matching$rt_diff, metabolites_matching$name)
names(new_names) <- NULL
metabolites_matching$new_names <- new_names

new_ids <- mapply(get_min_name, metabolites_matching$rt_diff, metabolites_matching$pub_chem_cid)
names(new_ids) <- NULL
metabolites_matching$new_pub_chem_cids <- new_ids


metabolites_matching <- metabolites_matching %>% 
  dplyr::select(r:name,pub_chem_cid,new_names,new_pub_chem_cids)
metabolites_matching <- metabolites_matching %>% 
  rename(pubchem_id=new_pub_chem_cids,
         metabolites=new_names) %>% 
  dplyr::select(metabolites,pubchem_id,r:PVal)
write.csv(metabolites_matching,fs::path(dir.results,"Pathway_Metabolites_Joint_Component.csv"))

#Match metabolites to IDs for pathway analysis
key <- metabolites %>% 
  dplyr::select(name,key)
metabolites_matching <- tidylog::left_join(results_cor3,key,by="key")
metabolites_matching <- tidylog::left_join(metabolites_matching,conf_ftrs,by="name") 
#Select lower diff rt
get_min_name <- function(rt_diff, name) {
  rt_diff_values <- as.numeric(unlist(strsplit(rt_diff, "; ")))
  name_values <- unlist(strsplit(name, "; "))
  
  min_index <- which.min(rt_diff_values)
  min_name <- name_values[min_index]
  
  return(min_name)
}

# Apply the function to each row and create the new column
new_names <- mapply(get_min_name, metabolites_matching$rt_diff, metabolites_matching$name)
names(new_names) <- NULL
metabolites_matching$new_names <- new_names

new_ids <- mapply(get_min_name, metabolites_matching$rt_diff, metabolites_matching$pub_chem_cid)
names(new_ids) <- NULL
metabolites_matching$new_pub_chem_cids <- new_ids


metabolites_matching <- metabolites_matching %>% 
  dplyr::select(r:name,pub_chem_cid,new_names,new_pub_chem_cids)
metabolites_matching <- metabolites_matching %>% 
  rename(pubchem_id=new_pub_chem_cids,
         metabolites=new_names) %>% 
  dplyr::select(metabolites,pubchem_id,r:PVal)
write.csv(metabolites_matching,fs::path(dir.results,"Pathway_Metabolites_Metabolite_Component.csv"))



sig_met <- c(sig_metabolite_features1)
sig_met_key <- key %>%
  filter(feature %in% sig_met)
features <- sig_met_key$new_names
features <- c(features,sig.mic$Taxa)

joint_results <- panel_b_dat_top_ft %>% 
  filter(feature %in% features) %>% 
  filter(lf_num=="Joint_1")

joint_results$Correlation <- sort(joint_results$Correlation)
order <- joint_results$feature
joint_results$feature <- factor(joint_results$feature,levels=order)
joint_results$omic_layer <- factor(joint_results$omic_layer, levels = c("microbiome", "metabolomics"))

# joint_results <- panel_b_dat_top_ft_joint
#Coef plot jiont component
plot1 <- ggplot(joint_results,aes(x=feature,y=Correlation))+
  geom_bar(stat = "identity")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  facet_grid(~omic_layer,scale="free_x",space="free_x")+
  labs(y="Factor Loadings",
       x=NULL)

sig_met <- c(sig_metabolite_features2)
sig_met_key <- key %>%
  filter(feature %in% sig_met)
features <- sig_met_key$new_names
features <- c(features)

met_results <- panel_b_dat_top_ft %>% 
  filter(feature %in% features) %>% 
  filter(lf_num=="metabolomics_5")

met_results$Correlation <- sort(met_results$Correlation)
order <- met_results$feature
met_results$feature <- factor(met_results$feature,levels=order)
met_results$omic_layer <- factor(met_results$omic_layer, levels = c("microbiome", "metabolomics"))

# joint_results <- panel_b_dat_top_ft_joint
#Coef plot jiont component
plot2 <- ggplot(met_results,aes(x=feature,y=Correlation))+
  geom_bar(stat = "identity")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  facet_grid(~omic_layer,scale="free_x",space="free_x")+
  labs(y="Factor Loadings",
       x=NULL)

# key <- key %>% 
#   rename(key=feature)
# results_cor <- tidylog::left_join(results_cor,key,by="key")
# results_cpr <- tidylog::left_join(results_cor,conf_ftrs,by="name")

#Rename metabolites 
full_metabolites <- metabolites %>% 
  dplyr::select(name,key)
full_metabolites <- tidylog::left_join(full_metabolites,conf_ftrs,by="name")
#Select lower diff rt
get_min_name <- function(rt_diff, name) {
  rt_diff_values <- as.numeric(unlist(strsplit(rt_diff, "; ")))
  name_values <- unlist(strsplit(name, "; "))
  
  min_index <- which.min(rt_diff_values)
  min_name <- name_values[min_index]
  
  return(min_name)
}

# Apply the function to each row and create the new column
new_names <- mapply(get_min_name, full_metabolites$rt_diff, full_metabolites$name)
names(new_names) <- NULL
full_metabolites$new_names <- new_names

new_ids <- mapply(get_min_name, full_metabolites$rt_diff, full_metabolites$pub_chem_cid)
names(new_ids) <- NULL
full_metabolites$new_ids <- new_ids

key <- full_metabolites %>% 
  dplyr::select(new_names,new_ids,key)  
key <- key %>% 
  dplyr::distinct(new_names,.keep_all = T)
key 
results_cor <- tidylog::left_join(results_cor,key,by="key")
write.csv(results_cor,fs::path(dir.results,"Joint_Component_Correlations.csv"))

results_cor3 <- tidylog::left_join(results_cor3,key,by="key")
write.csv(results_cor3,fs::path(dir.results,"Metabolite_Component_Correlations.csv"))

```

#6. Missing At Random Test 
```{r MCAR Test, include=F}
library(naniar)
# covs.dat <- meta_filt[c("id",covars)]
mic <- meta_filt[c(2:108,255:621)]
mic <- as.data.frame(mic)
mic <- mic %>% 
  dplyr::mutate(across(everything(),~as.numeric(.)))
# dat <- tidylog::full_join(mic,covs.dat,by="id")
# mcar_test(mic)

mic.name <- colnames(mic[16:107])
mic <- mic %>% 
  mutate(across(all_of(mic.name),~ifelse(!is.na(.),1,0)))
mic2 <- mic[,c(1:16,108:474)]
mic3 <- mic[,c(1:16)]

mic2 <- mic2 %>%
  rename(missing=Bifidobacterium)
mic2$missing <- ifelse(mic2$missing==0,NA,1)
mic2 <- as.data.frame(mic2)
# mcar_test(mic2)
mic3 <- mic3 %>% 
  rename(missing=Bifidobacterium)
mic3$missing <- ifelse(mic3$missing==0,NA,1)
mic3 <- as.data.frame(mic3)
mcar_test(mic3)

vars <- colnames(mic2)[-16]
pvals <- data.frame()
for (x in vars){
  formula <- as.formula(paste0("missing ~ ",x))
  m1 <- glm(formula, data = mic2, family = "binomial")
  p <- summary(m1)$coef[2,4]
  pvalues <- data.frame(Value = x,P=p)
  pvals <- rbind(pvals,pvalues)
}

pvals$fdr <- p.adjust(pvals$P, method = "fdr")

vars <- colnames(mic3)[-9]
pvals <- data.frame()
for (x in vars){
  formula <- as.formula(paste0("missing ~ ",x))
  m1 <- glm(formula, data = mic3, family = "binomial")
  p <- summary(m1)$coef[2,4]
  pvalues <- data.frame(Value = x,P=p)
  pvals <- rbind(pvals,pvalues)
}

pvals$fdr <- p.adjust(pvals$P, method = "fdr")

```

#7. Determine Group Differences
```{r group differences, echo=F}
# meta_filt <- meta_filt %>% 
#   mutate(across(everything(),~ifelse(!is.na(.),1,0)))
meta_filt$mis.group <- ifelse(is.na(meta_filt$Bifidobacterium),0,1)

#Determine if group differences of covariates
t.test(meta_filt$ma_age~meta_filt$mis.group) #Age is slightly lower among missing
stu_data <- table(meta_filt$ma_sex, meta_filt$mis.group)
chisq.test(stu_data) #No difference in sex distribution by missing groups
stu_data2 <- table(meta_filt$ma_race_eth, meta_filt$mis.group)
chisq.test(stu_data2) #No difference in race/ethnicity by missing groups
stu_data3 <- table(meta_filt$ma_parental_edu, meta_filt$mis.group)
chisq.test(stu_data3) #No difference in parental education by missing groups
t.test(meta_filt$ma_creatinine_scaled~meta_filt$mis.group) #No difference in creatinine by group
t.test(meta_filt$m_total_dietary_fiber_g~meta_filt$mis.group) #No difference in dietary fiber/energy by group
t.test(meta_filt$ma_bmi~meta_filt$mis.group) #No differences in bmi by group




```

