---
title: "Shanlee Davis - XXY GXT testing"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 2
    toc_float: yes
---

```{r setup, include=FALSE}
library(knitr)
library(nlme)
library(emmeans)
library(ggplot2)

knitr::opts_chunk$set(echo = FALSE,warning = FALSE)

# read in longitudinal data
alldata <- read.csv("E:\\Davis\\XXY exercise testing\\GXT variables (2).csv")

# drop AT and Peak rows
data <- alldata[!alldata$stage %in% c("AT","Peak"),]
# recode baseline as 0
data$stage <- ifelse(data$stage=="Baseline",0,data$stage)
# clean up
data$notes <- NULL
data$X <- NULL
data$X.1 <- NULL
data$X.2 <- NULL
data$X.3 <- NULL
data$stage <- as.factor(data$stage)
data$Case <- as.factor(data$Case)

# drop stages 11 and 12 - few observations
data <- data[data$stage != 12,]
data <- data[data$stage != 11,]
data$stage <- droplevels(data$stage)
# need to order the stage factor correctly
data$stage <- ordered(data$stage,levels=c("0","1","2","3","4","5","6","7","8","9","10"))

# VO2 
vo2_mod <- lme(VO2 ~ stage*Case,random=~1|ID,data = data,na.action = na.omit)
vo2_mod_anova <- anova.lme(vo2_mod, type="marginal")
# use emmeans 
vo2_mod_means <- emmeans(vo2_mod,c("stage","Case"))
vo2_mod_pairs <-  pairs(vo2_mod_means,adjust="tukey",simple="Case")
# plot the data
p1 <- ggplot(data,aes(x = stage,y = VO2)) +
  geom_smooth(aes(group = Case, linetype=Case), se = T, color="black") + 
  theme_bw() + xlab("Stage") + ylab("VO2") + labs(linetype ='Case/Control')

# HR 
HR_mod <- lme(HR ~ stage*Case,random=~1|ID,data = data,na.action = na.omit)
HR_mod_anova <- anova.lme(HR_mod, type="marginal")
# use emmeans 
HR_mod_means <- emmeans(HR_mod,c("stage","Case"))
HR_mod_pairs <-  pairs(HR_mod_means,adjust="tukey",simple="Case")
# plot the data
p2 <- ggplot(data,aes(x = stage,y = HR)) +
  geom_smooth(aes(group = Case, linetype=Case), se = T, color="black") + 
  theme_bw() + xlab("Stage") + ylab("HR") + labs(linetype ='Case/Control')

```

# Background

The purpose of this analysis is to compare longitudinal graded exercise testing in XXY cases and controls.

# Methods

Stage "Baseline" was recoded as 0.  Stages 11 and 12 were excluded from analysis because of small numbers.  A mixed-effects model with an interaction between case/control status and stage was used to test whether trajectories of VO2 differed in the two groups.

# Results

## VO2 across stages in cases and controls

Plot

```{r echo=FALSE,comment=""}
p1
```
<br>

ANOVA table

```{r echo=FALSE,comment=""}
kable(vo2_mod_anova)
```
<br>

Least-squares means

```{r echo=FALSE,comment=""}
kable(vo2_mod_means)
```
<br>

Comparison at each stage

```{r echo=FALSE,comment=""}
vo2_mod_pairs
```
<br>

## HR across stages in cases and controls

Plot

```{r echo=FALSE,comment=""}
p2
```
<br>

ANOVA table

```{r echo=FALSE,comment=""}
kable(HR_mod_anova)
```
<br>

Least-squares means

```{r echo=FALSE,comment=""}
kable(HR_mod_means)
```
<br>

Comparison at each stage

```{r echo=FALSE,comment=""}
HR_mod_pairs
```
<br>