---
title: "PCSK9, IDO1, PLXNB2 for Pierre"
author: "Ye Ji Choi"
date: "`r lubridate::today()`"
format: 
  html:
    toc: true
    toc-depth: 5
    toc-location: left
    code-fold: true
    embed-resources: true
---

```{r, include = F}
library(dplyr)
library(tidyr)
library(jsonlite)
library(reticulate)
library(Seurat)
library(glmmTMB)
library(arsenal)
library(purrr)
```

```{r include = F}
reticulate::use_python("/mmfs1/gscratch/scrubbed/yejichoi/mypy/bin/python", required = T)
reticulate::repl_python()
reticulate::py_config()
# set up Kopah connection
## Load boto3 and pandas
boto3 <- reticulate::import("boto3")
pd <- reticulate::import("pandas")

## Create an S3 client
keys <- fromJSON("/mmfs1/home/yejichoi/keys.json")
session <- boto3$session$Session(
  aws_access_key_id = keys$MY_ACCESS_KEY,
  aws_secret_access_key = keys$MY_SECRET_KEY
)

## Create an S3 client with the session
s3 <- session$client("s3", endpoint_url = "https://s3.kopah.uw.edu")
```

```{r echo = F, include=F}
# read PB90 Seurat file
temp_file <- tempfile(fileext = ".rds")
s3$download_file("scrna", "Kidney transcriptomics/Single cell RNA seq/PB_90_RPCAFix_Metadata_LR_RM_kpmpV1labelled.rds", temp_file)
so90 <- readRDS(temp_file)

temp_file <- tempfile(fileext = ".csv")
s3$download_file("harmonized.dataset", "harmonized_dataset.csv", temp_file)
harm_dat <- read.csv(temp_file)

dat <- harm_dat %>%
  dplyr::mutate(across(where(is.character), ~ na_if(., ""))) %>%
  dplyr::summarise(across(where(negate(is.numeric)), ~ ifelse(all(is.na(.x)), NA_character_, last(na.omit(.x)))),
                   across(where(is.numeric), ~ ifelse(all(is.na(.x)), NA_real_, mean(.x, na.rm = TRUE))),
                   .by = record_id) %>%
  mutate(kit_id = gsub("^K.*-", "KL-", kit_id),
         albuminuria_yn = case_when(acr_u < 30 ~ "No", acr_u >= 30 ~ "Yes"),
         group_t2ob = case_when(group == "Lean Control" ~ "LC",
                                group == "Type 2 Diabetes" | group == "Obese Control" ~ "T2D/OB"),
         group_sglt2i = case_when(group == "Type 2 Diabetes" & epic_ever_sglt2i_1 == "Yes" ~ "T2D(+)",
                                  group == "Type 2 Diabetes" & epic_ever_sglt2i_1 == "No" ~ "T2D(-)",
                                  group == "Obese Control" & epic_ever_sglt2i_1 == "Yes" ~ "OB(+)",
                                  group == "Obese Control" & epic_ever_sglt2i_1 == "No" ~ "OB(-)",
                                  T ~ "LC"),
         group_t2ob_sglt2i = case_when(epic_ever_sglt2i_1 == "Yes" ~ "T2D/OB(+)",
                                       group != "Lean Control" & epic_ever_sglt2i_1 == "No" ~ "T2D/OB(-)",
                                       T~ "LC"),
         group_t2d = case_when(group == "Type 2 Diabetes" ~ "T2D", T ~ "Non-T2D")) %>%
  dplyr::select(-c(cryostor_id)) %>%
  filter(!is.na(kit_id))

so90_meta <- so90@meta.data %>%
  dplyr::mutate(kit_id = gsub("(?i)^k.*-", "KL-", kit_id))
so90_meta <- left_join(so90_meta, dat, by = c("kit_id", "record_id"))
rownames(so90_meta) <- so90_meta$barcode
so90 <- AddMetaData(so90, so90_meta)
so90 <- subset(so90, record_id != "CRC-55") ## remove control with IgAN
```

```{r echo = F}
markers <- c("PCSK9", "IDO1", "PLXNB2")
```


# Descriptive Table

```{r echo = F, results='asis'}
so90@meta.data <- so90@meta.data %>%
  mutate(renal_function = case_when(group %in% c("Type_1_Diabetes", "Type_2_Diabetes") & 
                                      eGFR_CKD_epi >= 60 & eGFR_CKD_epi <= 120 | acr_u < 30 ~ "Early Stage",
                                    group %in% c("Type_1_Diabetes", "Type_2_Diabetes") & 
                                      (eGFR_CKD_epi >= 30 & eGFR_CKD_epi < 60 | acr_u >= 30) ~ "Advanced Stage"),
         group_diab = case_when(group %in% c("Type_1_Diabetes", "Type_2_Diabetes") ~ "Diabetes",
                                group == "Lean_Control" ~ "HC"),
         noglp1_sglt2i = case_when((group == "Type_2_Diabetes" & glp1_agonist_timepoint != "Yes") & 
                                     sglt2i_timepoint == "Yes" ~ "SGLT2i+",
                                   (group == "Type_2_Diabetes" & glp1_agonist_timepoint != "Yes") & 
                                     sglt2i_timepoint == "No" ~ "SGLT2i-"),
         nosglt2i_glp1 = case_when((group == "Type_2_Diabetes" & sglt2i_timepoint != "Yes") & 
                                     glp1_agonist_timepoint == "Yes" ~ "GLP1RA+",
                                   (group == "Type_2_Diabetes" & sglt2i_timepoint != "Yes") & 
                                     glp1_agonist_timepoint == "No" ~ "GLP1RA-"))

desc_dat <- so90@meta.data %>%
  distinct(kit_id, .keep_all = T) %>%
  filter(group != "Obese_Control") %>%
  mutate(eGFR_CKD_epi = case_when(eGFR_CKD_epi == "." ~ NA, T ~ eGFR_CKD_epi),
         eGFR_CKD_epi = as.numeric(eGFR_CKD_epi))
tab <- arsenal::tableby(group ~ age + sex + bmi + eGFR_CKD_epi + kwt(acr_u, "Nmiss", "medianq1q3", "range") + renal_function + glp1_agonist_timepoint + sglt2i_timepoint + noglp1_sglt2i + nosglt2i_glp1 , desc_dat)
summary(tab, test = F, total = F)
```


# Overall scRNA analysis (All cell types)

```{r echo = F}
FeaturePlot(so90, reduction = "umap.harmony", features = markers, cols = c("lightgray", "blue"), raster = F)
FeaturePlot(so90, reduction = "umap.harmony", features = "PLXNB2", cols = c("lightgray", "blue"), raster = F)

DimPlot(so90, reduction = "umap.harmony", group.by = "celltype_harmony", raster = F, alpha = 0.2, label = T)
```

## Impact of chronic hyperglycemia

### Comparison of T1D/T2D vs. HC

```{r echo = F}
# T1D + T2D vs. HC
DimPlot(so90, reduction = "umap.harmony", group.by = "group_diab", raster = F, alpha = 0.2)
```


Group 1: Diabetes (T1D + T2D)

Group 2: Healthy Controls (No obese controls)

* +: higher expression in diabetes group than HC
 
* -: lower expression in diabetes group than HC


```{r echo = F}
FindMarkers(subset(so90, features = markers), group.by = "group_diab", ident.1 = "Diabetes", ident.2 = "HC", features = markers, 
            logfc.threshold = 0, min.pct = 0.05)
```

## Impact of renal function loss

### Patients with diabetes only (T1D + T2D): normoalbuminuric normal GFR and normal albuminuria ([G1 and G2= ie GFR 60 to 120] AND A1[=ACR<30 mg/g] ) vs advanced stage (G3 [=GFR 30 to 59] AND/OR A2/A3 [=ACR >=30 mg/g]) 

```{r echo = F}
DimPlot(so90, reduction = "umap.harmony", group.by = "renal_function", raster = F, alpha = 0.2)
```


Group 1: Advanced stage (patients with T1D or T2D, G3 [=GFR 30 to 59] AND/OR A2/A3 [=ACR >=30 mg/g])

Group 2: Early stage (patients with T1D or T2D, [G1 and G2= ie GFR 60 to 120] AND/OR A1[=ACR<30 mg/g])

* +: higher expression in advanced stage group than early stage

* -: lower expression in advanced stage group than early stage


```{r echo = F}
FindMarkers(subset(so90, features = markers), group.by = "renal_function", ident.1 = "Advanced Stage", ident.2 = "Early Stage", features = markers, 
            logfc.threshold = 0, min.pct = 0.05)
```

## Impact of drugs  of chronic hyperglycemia

### Patients with T2D only: drugs vs. no drugs

#### within those with no GLP1RA exposure : SGLt2i+ vs  SGLt2i-

```{r echo = F}
# at visit drug use
DimPlot(subset(so90, group == "Type_2_Diabetes"), reduction = "umap.harmony", group.by = "noglp1_sglt2i", raster = F, alpha = 0.2)
```

Group 1: SGLT2i+ in no GLP1RA exposure with T2D

Group 2: SGLT2i- in no GLP1RA exposure with T2D


* +: higher expression in SGLT2i+ than SGLT2i-

* -: lower expression in SGLT2i+ than SGLT2i-


```{r echo = F}
FindMarkers(subset(so90, features = markers), group.by = "noglp1_sglt2i", ident.1 = "SGLT2i+", ident.2 = "SGLT2i-", features = markers, 
            logfc.threshold = 0, min.pct = 0.05)
```

#### within those with no SGLt2i exposure : GLP1RA+ vs  GLP1RA-

```{r echo = F}
# at visit drug use
DimPlot(subset(so90, group == "Type_2_Diabetes"), reduction = "umap.harmony", group.by = "nosglt2i_glp1", raster = F, alpha = 0.2)
```

Group 1: GLP1RA+ in no SGLT2i exposure with T2D

Group 2: GLP1RA- in no SGLT2i exposure with T2D

* +: higher expression in GLP1RA+ than GLP1RA-

* -: lower expression in GLP1RA+ than GLP1RA-


```{r echo = F}
FindMarkers(subset(so90, features = markers), group.by = "nosglt2i_glp1", ident.1 = "GLP1RA+", ident.2 = "GLP1RA-", features = markers, 
            logfc.threshold = 0, min.pct = 0.05)
```

# PC cells

```{r echo = F, include = F}
so90$celltype_pc <- ifelse(grepl("^PC-\\d+$", so90$celltype_harmony),
                                 "PC", as.character(so90$celltype_harmony))
so90_pc <- subset(so90, celltype_pc == "PC")
so90_pc <- NormalizeData(so90_pc)
so90_pc <- ScaleData(so90_pc)
```

```{r echo = F}
FeaturePlot(so90_pc, reduction = "umap.harmony", features = markers, cols = c("lightgray", "blue"), raster = F)
FeaturePlot(so90_pc, reduction = "umap.harmony", features = "PLXNB2", cols = c("lightgray", "blue"), raster = F)

DimPlot(so90_pc, reduction = "umap.harmony", group.by = "celltype_harmony", raster = F, alpha = 0.2, label = T)
```

## Impact of chronic hyperglycemia

### Comparison of T1D/T2D vs. HC

```{r echo = F}
# T1D + T2D vs. HC
DimPlot(so90_pc, reduction = "umap.harmony", group.by = "group_diab", raster = F, alpha = 0.2)
```


Group 1: Diabetes (T1D + T2D)

Group 2: Healthy Controls (No obese controls)

* +: higher expression in diabetes group than HC
 
* -: lower expression in diabetes group than HC


```{r echo = F}
FindMarkers(subset(so90_pc, features = markers), group.by = "group_diab", ident.1 = "Diabetes", ident.2 = "HC", features = markers, 
            logfc.threshold = 0, min.pct = 0.05)
```

## Impact of renal function loss

### Patients with diabetes only (T1D + T2D): normoalbuminuric normal GFR and normal albuminuria ([G1 and G2= ie GFR 60 to 120] AND A1[=ACR<30 mg/g] ) vs advanced stage (G3 [=GFR 30 to 59] AND/OR A2/A3 [=ACR >=30 mg/g]) 

```{r echo = F}
DimPlot(so90_pc, reduction = "umap.harmony", group.by = "renal_function", raster = F, alpha = 0.2)
```


Group 1: Advanced stage (patients with T1D or T2D, G3 [=GFR 30 to 59] AND/OR A2/A3 [=ACR >=30 mg/g])

Group 2: Early stage (patients with T1D or T2D, [G1 and G2= ie GFR 60 to 120] AND/OR A1[=ACR<30 mg/g])

* +: higher expression in advanced stage group than early stage

* -: lower expression in advanced stage group than early stage


```{r echo = F}
FindMarkers(subset(so90_pc, features = markers), group.by = "renal_function", ident.1 = "Advanced Stage", ident.2 = "Early Stage", features = markers, 
            logfc.threshold = 0, min.pct = 0.05)
```

## Impact of drugs  of chronic hyperglycemia

### Patients with T2D only: drugs vs. no drugs

#### within those with no GLP1RA exposure : SGLt2i+ vs  SGLt2i-

```{r echo = F}
# at visit drug use
DimPlot(subset(so90_pc, group == "Type_2_Diabetes"), reduction = "umap.harmony", group.by = "noglp1_sglt2i", raster = F, alpha = 0.2)
```

Group 1: SGLT2i+ in no GLP1RA exposure with T2D

Group 2: SGLT2i- in no GLP1RA exposure with T2D


* +: higher expression in SGLT2i+ than SGLT2i-

* -: lower expression in SGLT2i+ than SGLT2i-


```{r echo = F}
FindMarkers(subset(so90_pc, features = markers), group.by = "noglp1_sglt2i", ident.1 = "SGLT2i+", ident.2 = "SGLT2i-", features = markers, 
            logfc.threshold = 0, min.pct = 0.05)
```

#### within those with no SGLt2i exposure : GLP1RA+ vs  GLP1RA-

```{r echo = F}
# at visit drug use
DimPlot(subset(so90_pc, group == "Type_2_Diabetes"), reduction = "umap.harmony", group.by = "nosglt2i_glp1", raster = F, alpha = 0.2)
```

Group 1: GLP1RA+ in no SGLT2i exposure with T2D

Group 2: GLP1RA- in no SGLT2i exposure with T2D

* +: higher expression in GLP1RA+ than GLP1RA-

* -: lower expression in GLP1RA+ than GLP1RA-


```{r echo = F}
FindMarkers(subset(so90_pc, features = markers), group.by = "nosglt2i_glp1", ident.1 = "GLP1RA+", ident.2 = "GLP1RA-", features = markers, 
            logfc.threshold = 0, min.pct = 0.05)
```

# TAL cells

```{r echo = F, include = F}
so90$celltype_tal <- ifelse(grepl("TAL-", so90$celltype_harmony),
                                 "TAL", as.character(so90$celltype_harmony))
so90_tal <- subset(so90, celltype_tal == "TAL")
so90_tal <- NormalizeData(so90_tal)
so90_tal <- ScaleData(so90_tal)
```

```{r echo = F}
FeaturePlot(so90_tal, reduction = "umap.harmony", features = markers, cols = c("lightgray", "blue"), raster = F)
FeaturePlot(so90_tal, reduction = "umap.harmony", features = "PLXNB2", cols = c("lightgray", "blue"), raster = F)

DimPlot(so90_tal, reduction = "umap.harmony", group.by = "celltype_harmony", raster = F, alpha = 0.2, label = T)
```

## Impact of chronic hyperglycemia

### Comparison of T1D/T2D vs. HC

```{r echo = F}
# T1D + T2D vs. HC
DimPlot(so90_tal, reduction = "umap.harmony", group.by = "group_diab", raster = F, alpha = 0.2)
```


Group 1: Diabetes (T1D + T2D)

Group 2: Healthy Controls (No obese controls)

* +: higher expression in diabetes group than HC
 
* -: lower expression in diabetes group than HC


```{r echo = F}
FindMarkers(subset(so90_tal, features = markers), group.by = "group_diab", ident.1 = "Diabetes", ident.2 = "HC", features = markers, 
            logfc.threshold = 0, min.pct = 0.05)
```

## Impact of renal function loss

### Patients with diabetes only (T1D + T2D): normoalbuminuric normal GFR and normal albuminuria ([G1 and G2= ie GFR 60 to 120] AND A1[=ACR<30 mg/g] ) vs advanced stage (G3 [=GFR 30 to 59] AND/OR A2/A3 [=ACR >=30 mg/g]) 

```{r echo = F}
DimPlot(so90_tal, reduction = "umap.harmony", group.by = "renal_function", raster = F, alpha = 0.2)
```


Group 1: Advanced stage (patients with T1D or T2D, G3 [=GFR 30 to 59] AND/OR A2/A3 [=ACR >=30 mg/g])

Group 2: Early stage (patients with T1D or T2D, [G1 and G2= ie GFR 60 to 120] AND/OR A1[=ACR<30 mg/g])

* +: higher expression in advanced stage group than early stage

* -: lower expression in advanced stage group than early stage


```{r echo = F}
FindMarkers(subset(so90_tal, features = markers), group.by = "renal_function", ident.1 = "Advanced Stage", ident.2 = "Early Stage", features = markers, 
            logfc.threshold = 0, min.pct = 0.05)
```

## Impact of drugs  of chronic hyperglycemia

### Patients with T2D only: drugs vs. no drugs

#### within those with no GLP1RA exposure : SGLt2i+ vs  SGLt2i-

```{r echo = F}
# at visit drug use
DimPlot(subset(so90_tal, group == "Type_2_Diabetes"), reduction = "umap.harmony", group.by = "noglp1_sglt2i", raster = F, alpha = 0.2)
```

Group 1: SGLT2i+ in no GLP1RA exposure with T2D

Group 2: SGLT2i- in no GLP1RA exposure with T2D


* +: higher expression in SGLT2i+ than SGLT2i-

* -: lower expression in SGLT2i+ than SGLT2i-


```{r echo = F}
FindMarkers(subset(so90_tal, features = markers), group.by = "noglp1_sglt2i", ident.1 = "SGLT2i+", ident.2 = "SGLT2i-", features = markers, 
            logfc.threshold = 0, min.pct = 0.05)
```

#### within those with no SGLt2i exposure : GLP1RA+ vs  GLP1RA-

```{r echo = F}
# at visit drug use
DimPlot(subset(so90_tal, group == "Type_2_Diabetes"), reduction = "umap.harmony", group.by = "nosglt2i_glp1", raster = F, alpha = 0.2)
```

Group 1: GLP1RA+ in no SGLT2i exposure with T2D

Group 2: GLP1RA- in no SGLT2i exposure with T2D

* +: higher expression in GLP1RA+ than GLP1RA-

* -: lower expression in GLP1RA+ than GLP1RA-


```{r echo = F}
FindMarkers(subset(so90_tal, features = markers), group.by = "nosglt2i_glp1", ident.1 = "GLP1RA+", ident.2 = "GLP1RA-", features = markers, 
            logfc.threshold = 0, min.pct = 0.05)
```

# PT cells

```{r echo = F, include = F}
so90$celltype_pt <- ifelse(grepl("PT-", so90$celltype_harmony),
                                 "PT", as.character(so90$celltype_harmony))
so90_pt <- subset(so90, celltype_pt == "PT")
so90_pt <- NormalizeData(so90_pt)
so90_pt <- ScaleData(so90_pt)
```

```{r echo = F}
FeaturePlot(so90_pt, reduction = "umap.harmony", features = markers, cols = c("lightgray", "blue"), raster = F)
FeaturePlot(so90_pt, reduction = "umap.harmony", features = "PLXNB2", cols = c("lightgray", "blue"), raster = F)

DimPlot(so90_pt, reduction = "umap.harmony", group.by = "celltype_harmony", raster = F, alpha = 0.2, label = T)
```

## Impact of chronic hyperglycemia

### Comparison of T1D/T2D vs. HC

```{r echo = F}
# T1D + T2D vs. HC
DimPlot(so90_pt, reduction = "umap.harmony", group.by = "group_diab", raster = F, alpha = 0.2)
```


Group 1: Diabetes (T1D + T2D)

Group 2: Healthy Controls (No obese controls)

* +: higher expression in diabetes group than HC
 
* -: lower expression in diabetes group than HC


```{r echo = F}
FindMarkers(subset(so90_pt, features = markers), group.by = "group_diab", ident.1 = "Diabetes", ident.2 = "HC", features = markers, 
            logfc.threshold = 0, min.pct = 0.05)
```

## Impact of renal function loss

### Patients with diabetes only (T1D + T2D): normoalbuminuric normal GFR and normal albuminuria ([G1 and G2= ie GFR 60 to 120] AND A1[=ACR<30 mg/g] ) vs advanced stage (G3 [=GFR 30 to 59] AND/OR A2/A3 [=ACR >=30 mg/g]) 

```{r echo = F}
DimPlot(so90_pt, reduction = "umap.harmony", group.by = "renal_function", raster = F, alpha = 0.2)
```


Group 1: Advanced stage (patients with T1D or T2D, G3 [=GFR 30 to 59] AND/OR A2/A3 [=ACR >=30 mg/g])

Group 2: Early stage (patients with T1D or T2D, [G1 and G2= ie GFR 60 to 120] AND/OR A1[=ACR<30 mg/g])

* +: higher expression in advanced stage group than early stage

* -: lower expression in advanced stage group than early stage


```{r echo = F}
FindMarkers(subset(so90_pt, features = markers), group.by = "renal_function", ident.1 = "Advanced Stage", ident.2 = "Early Stage", features = markers, 
            logfc.threshold = 0, min.pct = 0.05)
```

## Impact of drugs  of chronic hyperglycemia

### Patients with T2D only: drugs vs. no drugs

#### within those with no GLP1RA exposure : SGLt2i+ vs  SGLt2i-

```{r echo = F}
# at visit drug use
DimPlot(subset(so90_pt, group == "Type_2_Diabetes"), reduction = "umap.harmony", group.by = "noglp1_sglt2i", raster = F, alpha = 0.2)
```

Group 1: SGLT2i+ in no GLP1RA exposure with T2D

Group 2: SGLT2i- in no GLP1RA exposure with T2D


* +: higher expression in SGLT2i+ than SGLT2i-

* -: lower expression in SGLT2i+ than SGLT2i-


```{r echo = F}
FindMarkers(subset(so90_pt, features = markers), group.by = "noglp1_sglt2i", ident.1 = "SGLT2i+", ident.2 = "SGLT2i-", features = markers, 
            logfc.threshold = 0, min.pct = 0.05)
```

#### within those with no SGLt2i exposure : GLP1RA+ vs  GLP1RA-

```{r echo = F}
# at visit drug use
DimPlot(subset(so90_pt, group == "Type_2_Diabetes"), reduction = "umap.harmony", group.by = "nosglt2i_glp1", raster = F, alpha = 0.2)
```

Group 1: GLP1RA+ in no SGLT2i exposure with T2D

Group 2: GLP1RA- in no SGLT2i exposure with T2D

* +: higher expression in GLP1RA+ than GLP1RA-

* -: lower expression in GLP1RA+ than GLP1RA-


```{r echo = F}
FindMarkers(subset(so90_pt, features = markers), group.by = "nosglt2i_glp1", ident.1 = "GLP1RA+", ident.2 = "GLP1RA-", features = markers, 
            logfc.threshold = 0, min.pct = 0.05)
```


```{r echo = F, eval = F}
s3$upload_file("/mmfs1/gscratch/scrubbed/yejichoi/Bjornstad/CHCO-Code/Petter Bjornstad/Pierre Jean Saulnier/PCSK9 IDO1 PLXNB2 for Pierre.html", "test", 'PCSK9 IDO1 PLXNB2 for Pierre.html')
```
