---
title: "Prediction model of liver fat in PCOS - validation cohort"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

source('C:/Users/pylell/Documents/GitHub/Functions/01_functions.R')

library(tidyr)
library(magrittr)
library(dplyr)
library(data.table)
library(knitr)
library(pander)
library(pROC)
library(OptimalCutpoints)
library(caret)
library(rms)

set.seed(3654)

# read  data
valid <- read.csv("H:\\Endocrinology\\Green\\Carreau liver fat prediction\\Data\\Validation cohort\\validation cohort.csv")
N <- nrow(valid)

#install.packages("ElemStatLearn", dependencies=T)
library(ElemStatLearn)
library(rpart)
library(rms)

# create HS variable
valid$HS <- ifelse(valid$Liver.Fat>5.5,1,0)
# rename variables 
names(valid)[names(valid) == 'BMI.percentile'] <- 'bmi.perc'
names(valid)[names(valid) == 'Waist.circumference'] <- 'Waist_cm'

#valid_mod <- glm(HS ~ bmi.perc + Waist_cm + ALT + SHBG, data=valid,family=binomial(link="logit"))
#tabx <- mod_tablog(valid_mod)

# get predicted probability of HS
#valid$pred <- predict(valid_mod,type="response",data.frame=valid)
# using the cutoff from our previous model, if probability>0.4399669, assign as predicted to have HS
#valid$pred_HS <- ifelse(valid$pred>0.4399669,1,0)

# confusion matrix
#conf <- confusionMatrix(as.factor(valid$pred_HS),as.factor(valid$HS),positive='1')
#x <- as.matrix(conf)
#row.names(x) <- c("Predicted no HS","Predicted HS")
#colnames(x) <- c("Observed no HS","Observed HS")
#y <- as.matrix(conf,what="overall")
#z <- as.matrix(conf,what="classes")
#yz <- rbind(y,z)

# use coefficients that we got from the training cohort
# first we need to read in the test data
# read  data
test <- read.csv("H:\\Endocrinology\\Green\\Carreau liver fat prediction\\Data\\NAFLD prediction score data.csv")
# replace "." with "NA"
test[test=="."] <- NA
# remove missing observations
test <- test %>% drop_na(HS,liver.fat.perc,bmi.perc,Waist_cm,HOMA,HDL,TG,ALT,Free.Testosterone,SHBG)
# fit model to test data
fred5 <- glm(HS ~ bmi.perc + Waist_cm + ALT + SHBG, data=test,family=binomial(link="logit"))

# now fit model from test data to validation data
valid$valid_mod_from_test <- predict(fred5,valid,type="response")
# using the cutoff from our previous model, if probability>0.4399669, assign as predicted to have HS
valid$pred_HS_from_test <- ifelse(valid$valid_mod_from_test>0.4399669,1,0)

# confusion matrix
conf_from_test <- confusionMatrix(as.factor(valid$pred_HS_from_test),as.factor(valid$HS),positive='1')
x <- as.matrix(conf_from_test)
row.names(x) <- c("Predicted no HS","Predicted HS")
colnames(x) <- c("Observed no HS","Observed HS")
y <- as.matrix(conf_from_test,what="overall")
z <- as.matrix(conf_from_test,what="classes")
yz <- rbind(y,z)

```

# Background

The purpose of this analysis is to fit the predictive model of hepatic steatosis (HS) that was developed in a test set of adolescent girls with PCOS.  The predictors included in the model were BMI %ile, waist circumference, ALT, and SHBG.

# Methods

The logistic regression model was fit to the validation cohort.  The performance of the model was evaluated using a confusion matrix, and by calculating sensitivity, specificity, and other criteria.

# Results

There were `r N` subjects included in this analysis.  The results of the logistic model are shown below.

```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(x)
```
\

```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(yz)
```

