---
title: "EMERALD echo data - Alex Nguyen"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 2
    toc_float: yes
---

```{r setup, include=FALSE}
library(stringr)
library(Hmisc)
library(dplyr)
library(skimr)
library(tableone)
library(knitr)
library(tidyverse)
library(rstatix)
library(ggpubr)
library(car)
library(hablar)
library(crayon)

options(scipen=999)

knitr::opts_chunk$set(echo = FALSE,warning = FALSE)

# echo data
# need to fix variable names
cardiac <- read.csv("E:\\Nadeau\\Nadeau EMERALD\\Echo\\Alex\\Raw data\\Cardiacdata 072721.csv",na.strings = c(NA,""))
names <- colnames(cardiac)
names <- names[3:ncol(cardiac)]
rootnames <- NA
for (i in seq(1,length(names),by=2)) {
  rootnames <- c(rootnames,names[i])
  names[i+1] <- paste0(names[i],"2")
  names[i] <- paste0(names[i],"1")
}
rootnames <- rootnames[-1]
# fix column names  
numnames <- names
names <- c("Study.ID","randomization",names)
colnames(cardiac) <- names

# 4 variables have percent signs in values
# cardiac$GLS....1 cardiac$GLS....2 cardiac$GCS....1 cardiac$GCS....2
#cardiac$GLS....1 <- as.numeric(gsub("\\%","",cardiac$GLS....1))
#cardiac$GLS....2 <- as.numeric(gsub("\\%","",cardiac$GLS....2))
#cardiac$GCS....1 <- as.numeric(gsub("\\%","",cardiac$GCS....1))
#cardiac$GCS....2 <- as.numeric(gsub("\\%","",cardiac$GCS....2))

# calculate deltas
Study.ID <- as.matrix(cardiac$Study.ID)
for (x in rootnames)  {
  del.x <- paste0("delta.",x) 
  v2.x <- paste0(x,"2")
  v1.x <- paste0(x,"1")
  temp <- select(cardiac,v2.x) - select(cardiac,v1.x)
  colnames(temp) <- del.x
  Study.ID <- as.data.frame(cbind(Study.ID,temp))
}
cardiac <- merge(cardiac,Study.ID,by="Study.ID")
cardiac$group <- ifelse(cardiac$randomization==1,"Metformin","Placebo")

# merge in group sheet so we have last name and BMI percentile
# don't need treatment group since that is already in the cardiac sheet
groups <- read.csv("E:\\Nadeau\\Nadeau EMERALD\\Echo\\Alex\\Raw data\\EMERALD DATA SUMMARIZED groups.csv",na.strings = c(NA,""))
groups$Group <- NULL
cardiac <- merge(cardiac,groups,by="Study.ID",all.x = T,all.y = F)

# merge in extra views
extra <- read.csv("E:\\Nadeau\\Nadeau EMERALD\\Echo\\Alex\\Raw data\\tracedata_combined_REVISED_strain.csv",na.strings = c(NA,""))
extra_names <- colnames(extra)
extra_names <- extra_names[2:61]
extra_rootnames <- NA
for (i in seq(1,length(extra_names),by=2)) {
  extra_rootnames <- c(extra_rootnames,extra_names[i])
  extra_names[i+1] <- paste0(extra_names[i],"2")
  extra_names[i] <- paste0(extra_names[i],"1")
}
extra_rootnames <- extra_rootnames[-1]
# fix column names  
extra_numnames <- extra_names
extra_names <- c("name",extra_names)
colnames(extra) <- extra_names
# extract first and last names
extra$last_name <- sub(",.*", "", extra$name)   
extra$first_name <- sub(".*, ", "", extra$name)     
# calculate deltas
names <- as.matrix(cbind(extra$last_name,extra$first_name))
colnames(names) <- c("last_name","first_name")
for (x in extra_rootnames)  {
  del.x <- paste0("delta.",x) 
  v2.x <- paste0(x,"2")
  v1.x <- paste0(x,"1")
  temp <- select(extra,v2.x) - select(extra,v1.x)
  colnames(temp) <- del.x
  names <- as.data.frame(cbind(names,temp))
}
extra <- merge(extra,names,by=c("last_name","first_name"))
#extra[extra$last_name=="O#Dell",]$last_name <- "O'Dell"
extra[extra$last_name=="Debord",]$last_name <- "DeBord"
extra[extra$first_name=="Lucus",]$first_name <- "Lucas"

# get rid of spaces
cardiac$first_name <- gsub(" ", "", cardiac$first_name, fixed = TRUE)
cardiac$last_name <- gsub(" ", "", cardiac$last_name, fixed = TRUE)
extra$first_name <- gsub(" ", "", extra$first_name, fixed = TRUE)
extra$last_name <- gsub(" ", "", extra$last_name, fixed = TRUE)

# names don't match up
a <- cardiac[,c("first_name","last_name")]
a <- cbind(a,rep(1,nrow(a)))
colnames(a) <- c("first_name","last_name","in_cardiac_data")
b <- extra[,c("first_name","last_name")]
b <- cbind(b,rep(1,nrow(b)))
colnames(b) <- c("first_name","last_name","in_extra_views")
check <- merge(a,b,by=c("first_name","last_name"),all.x=T,all.y=T)

cardiac <- merge(cardiac,extra,by=c("first_name","last_name"),all.x = T,all.y=T)

# will not take BMI percentile from labs sheet b/c fewer missing data points on group sheet
labs <- read.csv("E:\\Nadeau\\Nadeau EMERALD\\Echo\\Alex\\Raw data\\EMERALD DATA SUMMARIZED labs.csv",na.strings = c("",".","NA"))
labs <- labs %>% convert(num(mg_kg_min_80_gluc_corrected_final,mg_kg_min_80_gluc_corrected,
                     mg_kglean__min_80_gluc_corrected,mg_kglean__min_80_gluc_corrected_final))
labs$delta_gir <- labs$mg_kg_min_80_gluc_corrected_final - labs$mg_kg_min_80_gluc_corrected
labs$delta_gir_lean <- labs$mg_kglean__min_80_gluc_corrected_final - labs$mg_kglean__min_80_gluc_corrected
labs$delta_bmi <- labs$bmi_final_clamp - labs$bmi_baseline_clamp
labs$delta_ast <- labs$ast_v5 - labs$ast
labs$delta_alt <- labs$alt_v5 - labs$alt
labs$delta_a1c <- labs$a1c_v5 - labs$a1c
labs$delta_hbg <- labs$hgb_v5 - labs$hgb
labs$delta_cre <- labs$cre_v5 - labs$cre
labs$delta_wh <- labs$w_h_final - labs$w_h
labs$delta_crp <- labs$hs_crp_final - labs$hs_crp
labs$delta_mcg_mab_mgcre <- labs$mcg_mab_mgcre_final - labs$mcg_mab_mgcre
labs$delta_tg <- labs$tg_final - labs$tg                                                                                             
labs$delta_cpep <- labs$c_peptide_final - labs$c_peptide                                                                                                
labs$delta_chol <- labs$chol_final - labs$chol                          
labs$delta_hdl <- labs$hdl_final - labs$hdl
labs$delta_ldl <- labs$ldl_final - labs$ldl
labs$delta_estradiol <- labs$estradiol_final - labs$estradiol
labs$delta_progesterone <- labs$progesterone_final - labs$progesterone
labs$delta_clamp_testosterone <- labs$clamp_total_testerone_final - labs$clamp_total_testerone
labs$delta_clamp_shbg <- labs$clamp_shbg_final - labs$clamp_shbg
labs$delta_clamp_dheas <- labs$clamp_dheas_final - labs$clamp_dheas
labs$delta_leptin <- labs$leptin_final - labs$leptin
labs$delta_adipo <- labs$adipo_baseline_final - labs$adipo_baseline
labs$delta_clamp_cystatin <- labs$clamp_cystatin_final - labs$clamp_cystatin
labs$delta_ntprobnp <- labs$ntprobnp_final - labs$ntprobnp
labs$delta_creatinine <- labs$creatinine_final - labs$creatinine
labs$delta_clamp_icam <- labs$clamp_icam_final - labs$clamp_icam
labs$delta_clamp_vcam <- labs$clamp_vcam_final - labs$clamp_vcam
labs$delta_clamp_esel <- labs$clamp_esel_final - labs$clamp_esel
labs$delta_clamp_psel <- labs$clamp_psel_final - labs$clamp_psel
labs$delta_clamp_wbc <- labs$clamp_wbc_final - labs$clamp_wbc
labs$delta_clamp_plt_ct <- labs$clamp_plt_ct_final - labs$clamp_plt_ct
labs$delta_dexa_fat_perc <- labs$dexa_fat_perc_final - labs$dexa_fat_perc
labs$delta_dexa_lean_perc <- labs$dexa_lean_perc_final - labs$dexa_lean_perc
labs$delta_vo2max_ml_kg_min <- labs$vo2max_ml_kg_min_final - labs$vo2max_ml_kg_min
labs$delta_vo2_lean_kg <- labs$vo2_lean_kg_final - labs$vo2_lean_kg
labs$lept_adipo <- labs$leptin / labs$adipo_baseline
labs$lept_adipo_final <- labs$leptin_final / labs$adipo_baseline_final
labs$delta_lept_adipo <- labs$lept_adipo_final - labs$lept_adipo
keep_labs <- labs[,c("subject_id","gender","diabetes_duration","a1c","first_name","last_name","mg_kg_min_80_gluc_corrected_final","mg_kg_min_80_gluc_corrected",
                     "delta_gir","mg_kglean__min_80_gluc_corrected","mg_kglean__min_80_gluc_corrected_final","delta_gir_lean","bmi_final_clamp","bmi_baseline_clamp",
                     "delta_bmi","delta_ast","delta_alt","delta_a1c","delta_hbg","delta_cre","delta_wh","delta_crp","delta_mcg_mab_mgcre","delta_tg",
                     "delta_cpep","delta_chol","delta_hdl","delta_ldl","delta_estradiol","delta_progesterone","delta_clamp_testosterone","delta_clamp_shbg",
                     "delta_clamp_dheas","delta_leptin","delta_adipo","delta_clamp_cystatin","delta_ntprobnp","delta_creatinine","delta_clamp_icam","delta_clamp_vcam",
                     "delta_clamp_esel","delta_clamp_psel","delta_clamp_wbc","delta_clamp_plt_ct","delta_dexa_fat_perc","delta_dexa_lean_perc","delta_vo2max_ml_kg_min",
                     "delta_vo2_lean_kg","delta_lept_adipo")]
cardiac <- merge(cardiac,keep_labs,by=c("first_name","last_name"),all.x = T, all.y = F)
ids_a <- cardiac[,c("Study.ID","randomization")]
write.csv(ids_a,"E:\\Nadeau\\Nadeau EMERALD\\Echo\\Michal\\Clean data\\comparing_ids.csv",row.names = F)

control <- read.csv("E:\\Nadeau\\Nadeau EMERALD\\Echo\\Alex\\Raw data\\EMERALD DATA SUMMARIZED control echo.csv")

# still not sure if we are using these variables
#trace <- read.csv("E:\\Nadeau\\Nadeau EMERALD\\Echo\\Alex\\Raw data\\tracedata_combined_trace")

# nonnorm <- c("delta.AV.CO","delta.AV.Diam","delta.AV.SV","delta.AV.Vmax","delta.HR","delta.LA.Diam","delta.LALd.A4c","delta.Lat.A",
#              "delta.LVEF.MOD.A2c","delta.LVLd.A4c","delta.Max.AV.CO","delta.Max.AV.meanPG","delta.Max.AV.SV","delta.Max.AV.Vmean",
#              "delta.Max.Lat.S","delta.Max.LVEDV.MOD.A2C","delta.Max.MV.Dec.T","delta.Max.Sep.A","delta.Max.Sep.E.E","delta.MV.A.Vel",
#              "delta.MV.Dec.T","delta.PA.maxPG","delta.Sep.S","delta.SupineRest.AV.CO.Dopp","delta.SupineRest.AV.maxPG","delta.SupineRest.AV.Vmax",
#              "delta.SupineRest.AV.Vmean","delta.SupineRest.LVESV.MOD.A4C","delta.SupineRest.MV.Dec.Slope","delta.SupineRest.MV.E.Vel",
#              "delta.SupineRest.Sep.A","delta.deltatime_SAX.MV_SC_TRACE")

# unadjusted group differences in delta variables
c1 <- cardiac[ , grepl( "delta" , names(cardiac))]
c1 <- c1[,!grepl("TRACE1",names(c1))]
c1 <- c1[,!grepl("TRACE2",names(c1))]
c1 <- cbind(cardiac$group,c1)
colnames(c1) = c("group",colnames(c1[,2:ncol(c1)]))

# remove variables with fewer than 3 observations 
c1 = c1[,-c(which(colMeans(is.na(c1))>=0.93))]

# make another dataframe with all of the c1 variables plus  BMI %ile, sex, diabetes duration, baseline A1c
cardiac$gender <- as.factor(cardiac$gender)
c1_adj <- cbind(c1,cardiac[,c("bmi_percentile","gender","diabetes_duration","a1c")])
# need to take out even more variables with reduced number of observations
c1_adj = c1_adj[,-c(which(colMeans(is.na(c1_adj))>=0.9069))]

# make a metformin only df
c1_adj_met <- c1_adj[c1_adj$group=="Metformin",]

t1 <- CreateTableOne(data=c1,vars=colnames(c1[,-1]),strata="group",test = T)
t1 <- print(t1,varLabels=TRUE,showAllLevels=TRUE)
t1 <- t1[,-5]

# unadjusted change within metformin group
p <- NA
for (i in 2:ncol(c1)) {
 temp <- round(t.test(c1[c1$group=="Metformin",i], mu = 0, alternative = "two.sided")$p.value,4)
 p <- c(p,temp)
}
p <- p[-1]
onesamp <- as.data.frame(cbind(colnames(c1[,2:ncol(c1)]),p))
colnames(onesamp) <- c("Variable","p-value")

# need to save the p-values from the group term in the adjusted models
# need to modify so it doesn't just print out
mod_save_temp <- NA
mod_save <- NA
for (i in 2:(ncol(c1_adj)-4)) {
  mod <- lm(data = c1_adj,c1_adj[,i]~group+bmi_percentile+diabetes_duration+a1c)
  try(ptemp <- Anova(mod)[1,4])
  mod_save_temp <- ifelse(ptemp<0.05,noquote(colnames(c1_adj)[i]) ,"")
  mod_save <- noquote(c(mod_save,mod_save_temp))
}
mod_save <- mod_save[!(mod_save %in% c(NA,""))]

# within metformin group adjusted for BMI %ile, sex, diabetes duration, baseline A1c
# these models are acting strangely - the intercept should be the mean change but is quite different from the point estimate in Table 1
# if they really want these models, will need to troublshoot
# for (i in 2:(ncol(c1_adj)-4)) {
#   mod <- lm(data = c1_adj_met,c1_adj_met[,i]~c1_adj_met[,"bmi_percentile"]+c1_adj_met[,"diabetes_duration"] + c1_adj_met[,"a1c"])
#   print(noquote(c("Adjusted model for: ", noquote(colnames(c1_adj)[i]))))
#   print(summary(mod))
# }

# if there are significant differences or changes, does improvement in SI or weight loss mediate (correlate with?) these changes?
# find variables that are significant
# need to combine unadjusted group comparisons, unadjusted within met, and adjusted group comparisons
# NOW DELTA_GIR, DELTA_GIR_LEAN, AND DELTA_BMI ARE INCLUDED IN THE MODELS ABOVE - need to REMOVE THEM
sig <- t1[as.numeric(t1[,4])<0.05,]
sig <- rownames(sig)
sig <- sig[!is.na(sig)]
sig <- noquote(sig)
sig <- sub("\\ .*", "", sig)
sig1 <- onesamp[onesamp$`p-value`<0.05,]
sig1 <- sig1[,1]
sig1 <- noquote(sig1)
final_sig <- unique(c(mod_save,sig,sig1))
final_sig <- final_sig[!(final_sig %in% c("delta_bmi","delta_gir","delta_gir_lean","delta.clamp_weight","delta.height_v5","delta.BSA..m2.","delta_creatinine",
                                          "delta_clamp_esel","delta_clamp_cystatin","delta_a1c","delta_clamp_plt_ct"))]
corrdata <- cardiac[,final_sig]
corrdata <- cbind(corrdata,cardiac[,c("delta_bmi","delta_gir","delta_gir_lean","delta_clamp_cystatin","delta_creatinine","delta_clamp_esel","delta_a1c","delta_clamp_plt_ct")])

```

# Background

The purpose of this analysis was to:

1) Test whether changes in echo variables differed in the EMERALD metformin and placebo groups, both in unadjusted models and in models adjusted for baseline BMI percentile, diabetes duration, and HbA1c
2) Test whether there were significant changes within the EMERALD metformin group, both in unadjusted models and in models adjusted for baseline BMI percentile, diabetes duration, and HbA1c
3) Examine whether any significant echo variables were associated with changes in insulin sensitivity or BMI

# Methods

Changes in echo measures were calculated as the value at visit 2 minus the value at visit 1.  Variables representing changes in the echo measures were named as "delta." followed by the name of the echo measure.

Groups were compared using t-tests for normally distributed variables and the Kruskal-Wallis test for skewed variables.  One sample t-tests were used to test whether the change within the metformin group was significantly different from zero.  Linear models were used to compare changes in the metformin and placebo groups while adjusting for baseline BMI percentile, diabetes duration, and HbA1c.

Changes in echo measures that were significant in any of the above models were next tested for association with changes in GIR (mg_kg_min_80_gluc_corrected_final - mg_kglean__min_80_gluc_corrected), changes in GIR/lean kg (mg_kglean__min_80_gluc_corrected_final - mg_kglean__min_80_gluc_corrected), and changes in BMI (bmi_final_clamp - bmi_baseline_clamp) using linear models.

# Results

```{r echo=FALSE}
# checking distributions
#for (col in 2:ncol(c1)) {
#    try(hist(c1[,col],main = names(c1[col])))
#}
```

## Differences between metformin and placebo groups in change in echo and lab measures

```{r echo=FALSE}
kable(t1,caption = "Table 1.  Differences between metformin and placebo groups in change (visit 2-visit 1) in echo and lab measures")
```
<br>

## Tests of changes in echo and lab measures within the metformin group

```{r echo=FALSE}
kable(onesamp,caption = "Table 2.  P-values for test of change  (visit 2-visit 1) within the metformin group.")
```
<br>

## Models comparing changes in echo and lab measures between the groups, adjusted for baseline BMI percentile, diabetes duration, and A1c

These are the same comparisons as in the section 3.1 above, adjusted for covariates.  The output from each model can be interpreted as follows: in the table labeled "Coefficients,"  the p-value corresponding to each term in the model is in the column labeled "Pr(>|t|)."  For example, the p-value for the effect of treatment group is in the row labeled "groupPlacebo."

```{r echo=FALSE,comment=''}
# group comparisons adjusted for BMI %ile, sex, diabetes duration, baseline A1c
# last var is delta.peakval_SAX.PM_SC_TRACE
for (i in 2:(ncol(c1_adj)-4)) {
  mod <- lm(data = c1_adj,c1_adj[,i]~group+bmi_percentile+diabetes_duration+a1c)
  print(noquote(c("Adjusted model for: ", noquote(colnames(c1_adj)[i]))))
  print(summary(mod))
}
```
<br>

## Models comparing changes in significant echo measures and changes in GIR

In these models, echo variables that were either significant in group comparisons or changed significantly within the metformin group were compared to changes in GIR.  The output from each model can be interpreted as follows: in the table labeled "Coefficients,"  the p-value corresponding to each term in the model is in the column labeled "Pr(>|t|)."  The p-value for the relationship between the change in the echo measure and the change in GIR is in the row labeled "delta_gir."

```{r echo=FALSE,comment=''}
# are variables that change significantly associated with change in GIR?
for (i in 1:length(final_sig)) {
  mod_gir <- lm(data = corrdata,corrdata[,i]~delta_gir)
  cat(noquote(c("Model for:", noquote(colnames(corrdata)[i]))))
  print(summary(mod_gir))
}
```

## Models comparing changes in significant echo measures and changes in GIR/lean kg

In these models, echo variables that were either significant in group comparisons or changed significantly within the metformin group were compared to changes in GIR/lean kg.  The output from each model can be interpreted as follows: in the table labeled "Coefficients,"  the p-value corresponding to each term in the model is in the column labeled "Pr(>|t|)."  The p-value for the relationship between the change in the echo measure and the change in GIR is in the row labeled "delta_gir_lean."

```{r echo=FALSE,comment=''}
# are variables that change significantly associated with change in GIR?
for (i in 1:length(final_sig)) {
  mod_gir <- lm(data = corrdata,corrdata[,i]~delta_gir_lean)
  print(noquote(c("Model for: ", noquote(colnames(corrdata)[i]))))
  print(summary(mod_gir))
}
```

## Models comparing changes in significant echo measures and changes in BMI

In these models, echo variables that were either significant in group comparisons or changed significantly within the metformin group were compared to changes in BMI.  The output from each model can be interpreted as follows: in the table labeled "Coefficients,"  the p-value corresponding to each term in the model is in the column labeled "Pr(>|t|)."  The p-value for the relationship between the change in the echo measure and the change in GIR is in the row labeled "delta_BMI."

```{r echo=FALSE,comment=''}
for (i in 1:length(final_sig)) {
  mod_gir <- lm(data = corrdata,corrdata[,i]~delta_bmi)
  print(noquote(c("Model for: ", noquote(colnames(corrdata)[i]))))
  print(summary(mod_gir))
}
```

## Models comparing changes in significant echo measures and changes in significant lab measures

In these models, echo variables that were either significant in group comparisons or changed significantly within the metformin group were compared to changes lab or other clinical variables that also changed significantly.  The output from each model can be interpreted as follows: in the table labeled "Coefficients,"  the p-value corresponding to each term in the model is in the column labeled "Pr(>|t|)."  The p-value for the relationship between the change in the echo measure and the change in the lab measure is in the row labeled "delta_LABNAME."

```{r echo=FALSE,comment=''}
for (i in 1:length(final_sig)) {
  mod_gir <- lm(data = corrdata,corrdata[,i]~delta_clamp_cystatin )
  print(noquote(c("Model for: ", noquote(colnames(corrdata)[i]))))
  print(summary(mod_gir))
}
for (i in 1:length(final_sig)) {
  mod_gir <- lm(data = corrdata,corrdata[,i]~delta_creatinine  )
  print(noquote(c("Model for: ", noquote(colnames(corrdata)[i]))))
  print(summary(mod_gir))
}
for (i in 1:length(final_sig)) {
  mod_gir <- lm(data = corrdata,corrdata[,i]~delta_clamp_esel   )
  print(noquote(c("Model for: ", noquote(colnames(corrdata)[i]))))
  print(summary(mod_gir))
}
for (i in 1:length(final_sig)) {
  mod_gir <- lm(data = corrdata,corrdata[,i]~delta_a1c   )
  print(noquote(c("Model for: ", noquote(colnames(corrdata)[i]))))
  print(summary(mod_gir))
}
for (i in 1:length(final_sig)) {
  mod_gir <- lm(data = corrdata,corrdata[,i]~delta_clamp_plt_ct   )
  print(noquote(c("Model for: ", noquote(colnames(corrdata)[i]))))
  print(summary(mod_gir))
}
```
