---
title: "RFA DK-25-022"
author: "Ye Ji Choi"
date: "`r lubridate::today()`"
format: 
  html:
    toc: true
    toc-depth: 5
    toc-location: left
    code-fold: true
    embed-resources: true
---

 
## Table of data of interest

* Number of cores obtained

* Average length of the cores (in cm)

* Average time between harvest and placing in LM, EM, Hypothermosol/Cryostor, Cryomold/OCT RNAlater.

* The number of participants we have also obtained snap frozen LN samples.
 
If possible, it would also be great to visualize enrollment over time to show our success in recruiting for these types of studies.

```{r echo = F, include = F}
library(dplyr)
library(arsenal)
library(ggplot2)
library(purrr)
library(REDCapR)
library(tidyr)
```

```{r echo = F, include = F}
# gather tokens for REDCap read
api_tok <- read.csv("/Volumes/Peds Endo/Petter Bjornstad/Data Harmonization/api_tokens.csv")
attempt_tok <- api_tok[api_tok$Study == "ATTEMPT",]$Token
panda_tok <- api_tok[api_tok$Study == "PANDA",]$Token
croc_tok <- api_tok[api_tok$Study == "CROCODILE",]$Token

# uri for REDCap
uri <- "https://redcap.ucdenver.edu/api/"

# read from REDCap
attempt <- redcap_read(redcap_uri = uri, token = attempt_tok)$data %>%
  dplyr::rename(record_id = "subject_id") %>%
  mutate(visit = case_when(redcap_event_name == "visit_2_arm_1" ~ "BL",
                           redcap_event_name == "visit_3_arm_1" ~ "FU")) %>% 
  group_by(record_id) %>% fill(mrn, .direction = "down") %>% ungroup()
  
attempt$record_id <- as.character(attempt$record_id)
panda <- redcap_read(redcap_uri = uri, token = panda_tok)$data %>% 
  group_by(record_id) %>% fill(mrn, .direction = "down") %>% ungroup()
croc <- redcap_read(redcap_uri = uri, token = croc_tok)$data %>%
  mutate(record_id = case_when(record_id < 10 ~ paste0("CRC-0", record_id),
                               T ~ paste0("CRC-", record_id)))
attempt$study = paste0("ATTEMPT-", attempt$visit)
attempt$study <- factor(attempt$study, levels = c("ATTEMPT-BL", "ATTEMPT-FU"))
panda$study = "PANDA"
croc$study = "CROCODILE"
```

```{r echo = F, include=F}
# biopsy stats
# subset to vars of interest
attempt_biopsy <- attempt %>%
  select(record_id, mrn, study, redcap_event_name, starts_with("bx_"), kidney_side, passes_attempt, starts_with("pass"),
         starts_with("diagnostic"), starts_with("core"), starts_with("hypo"), starts_with("oct"),
         starts_with("rna"), starts_with("cryo"))

panda_biopsy <- panda %>%
  select(record_id, mrn, study, redcap_event_name, starts_with("bx_"), kidney_side, passes_attempt, starts_with("pass"),
         starts_with("diagnostic"), starts_with("core"), starts_with("hypo"), starts_with("oct"),
         starts_with("rna"), starts_with("cryo"))

croc_biopsy <- croc %>%
  select(record_id, mrn, study, starts_with("bx_"), kidney_side, passes_attempt, starts_with("pass"),
         starts_with("diagnostic"), starts_with("core"), starts_with("hypo"), starts_with("oct"),
         starts_with("rna"), starts_with("cryo"))

# merge all three df and calculate vars of interest
t1d_biopsy_dat <- full_join(attempt_biopsy, panda_biopsy) %>%
  full_join(croc_biopsy) %>%
  filter(!is.na(bx_date)) %>%
  mutate(avg_pass_core_cm = rowMeans(select(., c("pass1_core_cm", "pass2_core_cm", "pass3_core_cm", "pass4_core_cm")), na.rm = TRUE),
         core_n = rowSums(select(., c("pass1_tissue_yn", "pass2_tissue_yn", "pass3_tissue_yn","pass4_tissue_yn")), na.rm = TRUE),
         ln2_yn = case_when(!is.na(bx_ln2_id) ~ "Yes", T ~ "No"), 
         time_to_diagnostic_formalin = case_when(diagnostic_pass == 1 ~ as.numeric(difftime(diagnostic_formalin_time, pass1_time, units = "mins")),
                                                 diagnostic_pass == 2 ~ as.numeric(difftime(diagnostic_formalin_time, pass2_time, units = "mins")),
                                                 diagnostic_pass == 3 ~ as.numeric(difftime(diagnostic_formalin_time, pass3_time, units = "mins")),
                                                 diagnostic_pass == 4 ~ as.numeric(difftime(diagnostic_formalin_time, pass4_time, units = "mins"))),
         time_to_diagnostic_glut = case_when(diagnostic_pass == 1 ~ as.numeric(difftime(diagnostic_glut_time, pass1_time, units = "mins")),
                                             diagnostic_pass == 2 ~ as.numeric(difftime(diagnostic_glut_time, pass2_time, units = "mins")),
                                             diagnostic_pass == 3 ~ as.numeric(difftime(diagnostic_glut_time, pass3_time, units = "mins")),
                                             diagnostic_pass == 4 ~ as.numeric(difftime(diagnostic_glut_time, pass4_time, units = "mins"))),
         time_to_oct = case_when(oct_pass == 1 ~ as.numeric(difftime(oct_time, pass1_time, units = "mins")),
                                 oct_pass == 2 ~ as.numeric(difftime(oct_time, pass2_time, units = "mins")),
                                 oct_pass == 3 ~ as.numeric(difftime(oct_time, pass3_time, units = "mins")),
                                 oct_pass == 4 ~ as.numeric(difftime(oct_time, pass4_time, units = "mins"))),
         time_to_hypo = case_when(cryo_hypo_pass == 1 ~ as.numeric(difftime(hypo_time, pass1_time, units = "mins")),
                                  cryo_hypo_pass == 2 ~ as.numeric(difftime(hypo_time, pass2_time, units = "mins")),
                                  cryo_hypo_pass == 3 ~ as.numeric(difftime(hypo_time, pass3_time, units = "mins")),
                                  cryo_hypo_pass == 4 ~ as.numeric(difftime(hypo_time, pass4_time, units = "mins"))),
         time_to_cryo = case_when(cryo_hypo_pass == 1 ~ as.numeric(difftime(cryo_time, pass1_time, units = "mins")),
                                  cryo_hypo_pass == 2 ~ as.numeric(difftime(cryo_time, pass2_time, units = "mins")),
                                  cryo_hypo_pass == 3 ~ as.numeric(difftime(cryo_time, pass3_time, units = "mins")),
                                  cryo_hypo_pass == 4 ~ as.numeric(difftime(cryo_time, pass4_time, units = "mins"))),
          time_to_rna = case_when(rna_pass == 1 ~ as.numeric(difftime(rna_time, pass1_time, units = "mins")),
                                  rna_pass == 2 ~ as.numeric(difftime(rna_time, pass2_time, units = "mins")),
                                  rna_pass == 3 ~ as.numeric(difftime(rna_time, pass3_time, units = "mins")),
                                  rna_pass == 4 ~ as.numeric(difftime(rna_time, pass4_time, units = "mins"))))
```
### Combined
```{r echo = F, results = 'asis'}
t1d_biopsy_dat$core_n_cat <- factor(t1d_biopsy_dat$core_n)
labs <- list(core_n = "Number of cores collected", 
             core_n_cat = "Number of cores collected",
             avg_pass_core_cm = "Avg core length (cm)",
             time_to_diagnostic_formalin = "Time to diagnostic (formalin)",
             time_to_diagnostic_glut = "Time to diagnostic (glut)", 
             time_to_oct = "Time to OCT",
             time_to_hypo = "Time to hypothermasol", 
             time_to_cryo = "Time to cryostor", 
             time_to_rna = "Time to RNALater", 
             ln2_yn = "LN2 collected")

summary(arsenal::tableby(study ~ core_n + core_n_cat + avg_pass_core_cm + time_to_diagnostic_formalin + time_to_diagnostic_glut + time_to_oct +
                           time_to_hypo + time_to_cryo + time_to_rna + ln2_yn, data = t1d_biopsy_dat), 
        labelTranslations = labs, test = F)
```
### ATTEMPT
```{r echo = F, results = 'asis'}
attempt_only_dat <- t1d_biopsy_dat %>%
  arrange(desc(study)) %>%
  distinct(mrn, .keep_all = T) %>%
  filter(grepl("ATTEMPT", study)) 

summary(arsenal::tableby(study ~ core_n + core_n_cat + avg_pass_core_cm + time_to_diagnostic_formalin + time_to_diagnostic_glut + time_to_oct +
                           time_to_hypo + time_to_cryo + time_to_rna + ln2_yn, data = subset(t1d_biopsy_dat, grepl("ATTEMPT", study))), 
        labelTranslations = labs, test = F)
```

## Visualize enrollment over time
```{r echo = F, warning = F}
attempt_screening <- attempt[!is.na(attempt$consent_date),]
croc_screening <- croc[!is.na(croc$consent_date),]
panda_screening <- panda[!is.na(panda$consent_date),]

attempt_screening_n <- nrow(attempt_screening)
croc_screening_n <- nrow(croc_screening)
panda_screening_n <- nrow(panda_screening)

consent_dates <- data.frame(study = c(rep("ATTEMPT", attempt_screening_n), rep("CROCODILE", croc_screening_n), rep("PANDA", panda_screening_n)),
                     consent_date = c(attempt_screening$consent_date, croc_screening$consent_date, panda_screening$consent_date))
```

```{r echo = F, warning = F}
ggplot(consent_dates, aes(x = consent_date, fill = study)) +
  geom_histogram(binwidth = 30, color = "white",
                 position = "stack") +
  labs(x = "Consent Date", 
       y = "Frequency", 
       title = "Enrollment of T1D over time",
       fill = "Study") +
  theme_minimal() +
  theme(panel.grid = element_blank(),
        text = element_text(size = 16)) + 
  scale_fill_manual(values = c("#1d2d44","#3e5c76", "#748cab"))
```


```{r echo = F, warning=F}
# Convert consent_date to Date type if not already
consent_dates <- consent_dates %>%
  mutate(consent_date = as.Date(consent_date))

# Calculate cumulative count by date for each study
consent_dates <- consent_dates %>%
  arrange(consent_date) %>%
  group_by(consent_date) %>%
  dplyr::summarise(count = n()) %>%
  ungroup() %>%
  mutate(cumulative_count = cumsum(count))

# Plot cumulative count over time
ggplot(consent_dates, aes(x = consent_date, y = cumulative_count)) +
  geom_line(size = 1.3, color = "#748cab") +
  labs(x = "Consent Date", 
       y = "Cumulative Frequency", 
       title = "Cumulative Enrollment of T1D over Time",
       color = "Study") +
  theme_minimal() +
  theme(panel.grid = element_blank(),
        text = element_text(size = 16))
```

For the safety data – we can pull from our safety officer report / continuous review to COMIRB (protocol deviation log). Leena can help. This will be descriptive only, and just a small table in the grant. For the MRIs, I think we primarily want to demonstrate feasibility of the multiparametric MR imaging, and report the number of participants with the sequences (R2*, ASL, ADC, T1 mapping, TKV). I am open to any way of showing this. Ian can provide feedback on OCT and how to best demonstrate those data.


## Summary of safety and tolerability data


## Kidney MRIs

showcase the kidney MRIs (i.e., number, and what time of sequences)


## Ian’s OCT data

