---
title: "TODAY Somalogic baseline analyses - elastic net Cox models - CJASN response"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
knit: (function(rmdfile, ...) { rmarkdown::render(rmdfile, output_dir='/Users/pylell/Dropbox/TODAY HTN and DKD output/') })
output: rtf_document
---

```{r setup, include=FALSE}
library(SomaDataIO)
library(limma)
library(dplyr)
library(caret)
library(purrr)
library(multtest)
library(openxlsx)
library(tableone)
library(EnhancedVolcano)
library(knitr)
library(survival)
library(broom)
library(plsRcox)
library(forestmodel)
library(survC1)
library(naniar)
library(rms)
library(readxl)
library(forester)
library(grid)
library(forestploter)

knitr::opts_chunk$set(echo = FALSE,warning = FALSE)

knitr::opts_chunk$set(echo = FALSE)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "E:/Petter Bjornstad/TODAY subaward"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/TODAY subaward/"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/PEDS/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/TODAY subaward"
}
knitr::opts_knit$set(root.dir = home_dir)

```

```{r, include=FALSE}
setwd(home_dir)
# load somalogic data, with QC samples already excluded
load("./Somalogic data raw/soma.Rdata")

# load analyte info
load("./Somalogic data raw/analytes.Rdata")

# load comorbidity data
load("./Clinical data/comorb.Rdata")

# load baseline risk factors
load("./Clinical data/TODAY/baserisk.Rdata")

# take only the baseline soma samples
# can we just take the earliest or is it possible someone would have a follow-up sample but not baseline?
# probably can take the first and then check the years to make sure compatible with TODAY
base <- soma %>% arrange(releaseid,Date.Drawn) %>% group_by(releaseid) %>% filter(row_number()==1)

# these 3 release IDs have first sample after end of recruitment, so they must be missing the baseline visit
base <- base %>% filter(!releaseid %in% c("65-85903","65-47984","65-25901"))

# merge in complication data
base <- left_join(base, comorb, by="releaseid")

# merge in baseline risk factors
base <- left_join(base, baserisk, by="releaseid")

# identify columns corresponding to proteins
#is_seq <- function(.x) grepl("^seq\\.[0-9]{4}", .x) # regex for analytes
is_seq <- function(.x) grepl("seq", .x)
seq <- is_seq(names(base))

# convert to numeric
base[,seq] <- apply(base[,seq],2,as.numeric)

# are there proteins with low variability?
no_var = caret::nearZeroVar(base[,seq])
# none

# log transform
base_log <- base %>% modify_if(is_seq(names(.)), log)

# scale by SD
base_log_scale <- base_log
predictors <- colnames(base_log[seq])
for (i in 1:length(predictors)) {
  base_log_scale[,paste0(predictors[i])] <- base_log_scale[,paste0(predictors[i])]/sd(unlist(base_log[,paste0(predictors[i])]))
}

# filter to top 100 proteins by effect size
f = "/Volumes/PEDS/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/TODAY subaward/Results/Linear and Cox models/TODAY somalogic Cox models scaled baseline adjusted.xlsx"

# MAC
mac_keep = read_excel(f,sheet = "MAC CPH")
mac_keep <- arrange(mac_keep,adj.p.value)
mac_drop <- mac_keep[101:nrow(mac_keep),"AptName"]
rownames(mac_drop) <- mac_drop$AptName
mac_drop <- as.data.frame(t(mac_drop))
mac <- base_log_scale %>% select(-colnames(mac_drop))
# MIC
mic_keep = read_excel(f,sheet = "MIC CPH")
mic_keep <- arrange(mic_keep,adj.p.value)
mic_drop <- mic_keep[101:nrow(mic_keep),"AptName"]
rownames(mic_drop) <- mic_drop$AptName
mic_drop <- as.data.frame(t(mic_drop))
mic <- base_log_scale %>% select(-colnames(mic_drop))
# MIC.OR.MIC
mic.or.mac_keep = read_excel(f,sheet = "MIC.OR.MAC CPH")
mic.or.mac_keep <- arrange(mic.or.mac_keep,adj.p.value)
mic.or.mac_drop <- mic.or.mac_keep[101:nrow(mic.or.mac_keep),"AptName"]
rownames(mic.or.mac_drop) <- mic.or.mac_drop$AptName
mic.or.mac_drop <- as.data.frame(t(mic.or.mac_drop))
mic.or.mac <- base_log_scale %>% select(-colnames(mic.or.mac_drop))
# HYPERFILTRATION
hyp_keep = read_excel(f,sheet = "HYP CPH")
hyp_keep <- arrange(hyp_keep,adj.p.value)
hyp_drop <- hyp_keep[101:nrow(hyp_keep),"AptName"]
rownames(hyp_drop) <- hyp_drop$AptName
hyp_drop <- as.data.frame(t(hyp_drop))
hyp <- base_log_scale %>% select(-colnames(hyp_drop))
# RAPID EGFR DECLINE
rapid_keep = read_excel(f,sheet = "RAPID CPH")
rapid_keep <- arrange(rapid_keep,adj.p.value)
rapid_drop <- rapid_keep[101:nrow(rapid_keep),"AptName"]
rownames(rapid_drop) <- rapid_drop$AptName
rapid_drop <- as.data.frame(t(rapid_drop))
rapid <- base_log_scale %>% select(-colnames(rapid_drop))
# HYPERTENSION
htn_keep = read_excel(f,sheet = "HTN CPH")
htn_keep <- arrange(htn_keep,adj.p.value)
htn_drop <- htn_keep[101:nrow(htn_keep),"AptName"]
rownames(htn_drop) <- htn_drop$AptName
htn_drop <- as.data.frame(t(htn_drop))
htn <- base_log_scale %>% select(-colnames(htn_drop))
# HYPERTENSION WITH SBP
htn.with.sbp_keep = read_excel(f,sheet = "HTN with SBP CPH")
htn.with.sbp_keep <- arrange(htn.with.sbp_keep,adj.p.value)
htn.with.sbp_drop <- htn.with.sbp_keep[101:nrow(htn.with.sbp_keep),"AptName"]
rownames(htn.with.sbp_drop) <- htn.with.sbp_drop$AptName
htn.with.sbp_drop <- as.data.frame(t(htn.with.sbp_drop))
htn.with.sbp <- base_log_scale %>% select(-colnames(htn.with.sbp_drop))
# HYPERTENSION WITH UACR
htn.with.uacr_keep = read_excel(f,sheet = "HTN with UACR CPH")
htn.with.uacr_keep <- arrange(htn.with.uacr_keep,adj.p.value)
htn.with.uacr_drop <- htn.with.uacr_keep[101:nrow(htn.with.uacr_keep),"AptName"]
rownames(htn.with.uacr_drop) <- htn.with.uacr_drop$AptName
htn.with.uacr_drop <- as.data.frame(t(htn.with.uacr_drop))
htn.with.uacr <- base_log_scale %>% select(-colnames(htn.with.uacr_drop))
# HYPERTENSION WITH eGFR
htn.with.egfr_keep = read_excel(f,sheet = "HTN with eGFR CPH")
htn.with.egfr_keep <- arrange(htn.with.egfr_keep,adj.p.value)
htn.with.egfr_drop <- htn.with.egfr_keep[101:nrow(htn.with.egfr_keep),"AptName"]
rownames(htn.with.egfr_drop) <- htn.with.egfr_drop$AptName
htn.with.egfr_drop <- as.data.frame(t(htn.with.egfr_drop))
htn.with.egfr <- base_log_scale %>% select(-colnames(htn.with.egfr_drop))

set.seed(3654)
source("~/Documents/Github/CHCO-Code/Petter Bjornstad/TODAY proteomics metabolomics/Proteomics analyses/easy_elasticnet.R")
```

# Albuminuria

## Elastic net with 10-fold cross-validation

```{r, include=FALSE, comment=""}
mic.or.mac_nozero <- mic.or.mac[!mic.or.mac$DAYSTOMIC.OR.MAC == 0,]
seq <- is_seq(names(mic.or.mac))

mic.or.mac_select = easy_elasticnet(data = mic.or.mac_nozero,outcome = "MIC.OR.MAC",
                           predictors = mic.or.mac_nozero[,seq], out = "min.error",
                           model_type = "cox",time = "DAYSTOMIC.OR.MAC", cv_method="kfold", folds = 10)
retained_vars = analytes[analytes$AptName %in% mic.or.mac_select, c("Target","TargetFullName")]

form = as.formula(paste0("Surv(DAYSTOMIC.OR.MAC, MIC.OR.MAC)~",paste0(mic.or.mac_select,collapse = "+")))
mod = coxph(formula = form,data = mic.or.mac_nozero)
kable(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)

# make a summary table
moddf <- as.data.frame(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)
moddf$term <- analytes[analytes$AptName %in% moddf$term, ]$Target
moddf <- cbind(moddf$term,round(moddf$estimate, digits=3),round(moddf$std.error, digits=3),round(moddf$p.value, digits=3))
colnames(moddf) <- c("Variable","Hazard Ratio","SE","P-value")
concordance_df <- as.data.frame(cbind("Albuminuria","Proteins",mod$concordance[[6]], mod$concordance[[7]]))
```

```{r, include=TRUE, comment=""}
hist((mic.or.mac_nozero[mic.or.mac_nozero$MIC.OR.MAC==1,]$DAYSTOMIC.OR.MAC)/365, main = paste("Histogram of Time to Albuminuria"), xlab = "Years")
```

```{r, include=TRUE, comment=""}
kable(moddf, row.names=F)
```

There are `r nrow(moddf)` variables in the model.

```{r, include=TRUE, comment=""}
kable(round(mod$concordance,digits=3))
```

## Model with top protein

```{r, include=FALSE, comment=""}
form = as.formula(paste0("Surv(DAYSTOMIC.OR.MAC, MIC.OR.MAC)~seq.6544.33"))
mod = coxph(formula = form,data = mic.or.mac_nozero)
kable(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)

# make a summary table
moddf <- as.data.frame(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)
moddf <- cbind(moddf$term,round(moddf$estimate, digits=3),round(moddf$std.error, digits=3),round(moddf$p.value, digits=3))
colnames(moddf) <- c("Variable","Hazard Ratio","SE","P-value")
concordance_df_temp <- as.data.frame(cbind("Albuminuria","Top protein",mod$concordance[[6]], mod$concordance[[7]]))
concordance_df <- rbind(concordance_df, concordance_df_temp)
```

```{r, include=TRUE, comment=""}
kable(moddf, row.names=F)
```

There are `r nrow(moddf)` variables in the model.

```{r, include=TRUE, comment=""}
kable(round(mod$concordance,digits=3))
```

## Model with 7 proteins nominally significant for all 3 outcomes

```{r, include=FALSE, comment=""}
form = as.formula(paste0("Surv(DAYSTOMIC.OR.MAC.x, MIC.OR.MAC.x)~seq.16915.153+seq.18339.207+seq.3538.26+seq.9348.1+seq.7871.16+seq.13990.1+seq.13710.6"))
mic.or.mac_nozero_allproteins <- left_join(mic.or.mac_nozero, base_log_scale, by="releaseid")
mod = coxph(formula = form,data = mic.or.mac_nozero_allproteins)
kable(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)

# make a summary table
moddf <- as.data.frame(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)
moddf <- cbind(moddf$term,round(moddf$estimate, digits=3),round(moddf$std.error, digits=3),round(moddf$p.value, digits=3))
colnames(moddf) <- c("Variable","Hazard Ratio","SE","P-value")
concordance_df_temp <- as.data.frame(cbind("Albuminuria","Shared proteins",mod$concordance[[6]], mod$concordance[[7]]))
concordance_df <- rbind(concordance_df, concordance_df_temp)
```

```{r, include=TRUE, comment=""}
kable(moddf, row.names=F)
```

There are `r nrow(moddf)` variables in the model.

```{r, include=TRUE, comment=""}
kable(round(mod$concordance,digits=3))
```

# Hyperfiltration

## Elastic net with 10-fold cross-validation

```{r, include=FALSE, comment=""}
hyp_nozero <- hyp[!hyp$DAYSTOHYP == 0,]
seq <- is_seq(names(hyp))

hyp_select = easy_elasticnet(data = hyp_nozero,outcome = "HYP",
                           predictors = hyp_nozero[,seq], out = "min.error",
                           model_type = "cox",time = "DAYSTOHYP", cv_method="kfold", folds = 10)
retained_vars = analytes[analytes$AptName %in% hyp_select, c("Target","TargetFullName")]

form = as.formula(paste0("Surv(DAYSTOHYP, HYP)~",paste0(hyp_select,collapse = "+")))
mod = coxph(formula = form,data = hyp_nozero)
kable(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)

# make a summary table
moddf <- as.data.frame(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)
moddf$term <- analytes[analytes$AptName %in% moddf$term, ]$Target
moddf <- cbind(moddf$term,round(moddf$estimate, digits=3),round(moddf$std.error, digits=3),round(moddf$p.value, digits=3))
colnames(moddf) <- c("Variable","Hazard Ratio","SE","P-value")
concordance_df_temp <- as.data.frame(cbind("Hyperfiltration","Proteins",mod$concordance[[6]], mod$concordance[[7]]))
concordance_df <- rbind(concordance_df, concordance_df_temp)
```

```{r, include=TRUE, comment=""}
kable(moddf, row.names=F)
```

There are `r nrow(moddf)` variables in the model.

```{r, include=TRUE, comment=""}
kable(round(mod$concordance,digits=3))
```

## Model with top protein

```{r, include=FALSE, comment=""}
form = as.formula(paste0("Surv(DAYSTOHYP, HYP)~seq.2609.59"))
mod = coxph(formula = form,data = hyp_nozero)
kable(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)

# make a summary table
moddf <- as.data.frame(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)
moddf <- cbind(moddf$term,round(moddf$estimate, digits=3),round(moddf$std.error, digits=3),round(moddf$p.value, digits=3))
colnames(moddf) <- c("Variable","Hazard Ratio","SE","P-value")
concordance_df_temp <- as.data.frame(cbind("Hyperfiltration","Top protein",mod$concordance[[6]], mod$concordance[[7]]))
concordance_df <- rbind(concordance_df, concordance_df_temp)
```

```{r, include=TRUE, comment=""}
kable(moddf, row.names=F)
```

There are `r nrow(moddf)` variables in the model.

```{r, include=TRUE, comment=""}
kable(round(mod$concordance,digits=3))
```

## Model with 7 proteins nominally significant for all 3 outcomes

```{r, include=FALSE, comment=""}
form = as.formula(paste0("Surv(DAYSTOHYP.x, HYP.x)~seq.16915.153+seq.18339.207+seq.3538.26.x+seq.9348.1.x+seq.7871.16+seq.13990.1+seq.13710.6"))
hyp_nozero_allproteins <- left_join(hyp_nozero, base_log_scale, by="releaseid")
mod = coxph(formula = form,data = hyp_nozero_allproteins)
kable(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)

# make a summary table
moddf <- as.data.frame(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)
moddf <- cbind(moddf$term,round(moddf$estimate, digits=3),round(moddf$std.error, digits=3),round(moddf$p.value, digits=3))
colnames(moddf) <- c("Variable","Hazard Ratio","SE","P-value")
concordance_df_temp <- as.data.frame(cbind("Hyperfiltration","Shared proteins",mod$concordance[[6]], mod$concordance[[7]]))
concordance_df <- rbind(concordance_df, concordance_df_temp)
```

```{r, include=TRUE, comment=""}
kable(moddf, row.names=F)
```

There are `r nrow(moddf)` variables in the model.

```{r, include=TRUE, comment=""}
kable(round(mod$concordance,digits=3))
```

# Rapid eGFR decline

## Elastic net with 10-fold cross-validation

```{r, include=FALSE, comment=""}
rapid_nozero <- rapid[!rapid$DAYSTORAPID == 0,]
seq <- is_seq(names(rapid))

rapid_select = easy_elasticnet(data = rapid_nozero,outcome = "RAPID",
                           predictors = rapid_nozero[,seq], out = "min.error",
                           model_type = "cox",time = "DAYSTORAPID", cv_method="kfold", folds = 10)
retained_vars = analytes[analytes$AptName %in% rapid_select, c("Target","TargetFullName")]

form = as.formula(paste0("Surv(DAYSTORAPID, RAPID)~",paste0(rapid_select,collapse = "+")))
mod = coxph(formula = form,data = rapid_nozero)
kable(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)

# make a summary table
moddf <- as.data.frame(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)
moddf$term <- analytes[analytes$AptName %in% moddf$term, ]$Target
moddf <- cbind(moddf$term,round(moddf$estimate, digits=3),round(moddf$std.error, digits=3),round(moddf$p.value, digits=3))
colnames(moddf) <- c("Variable","Hazard Ratio","SE","P-value")
concordance_df_temp <- as.data.frame(cbind("Rapid eGFR decline","Proteins",mod$concordance[[6]], mod$concordance[[7]]))
concordance_df <- rbind(concordance_df, concordance_df_temp)
```

```{r, include=TRUE, comment=""}
kable(moddf, row.names=F)
```

There are `r nrow(moddf)` variables in the model.

```{r, include=TRUE, comment=""}
kable(round(mod$concordance,digits=3))
```

## Model with top protein

```{r, include=FALSE, comment=""}
form = as.formula(paste0("Surv(DAYSTORAPID, RAPID)~seq.5590.11"))
mod = coxph(formula = form,data = rapid_nozero)
kable(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)

# make a summary table
moddf <- as.data.frame(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)
moddf <- cbind(moddf$term,round(moddf$estimate, digits=3),round(moddf$std.error, digits=3),round(moddf$p.value, digits=3))
colnames(moddf) <- c("Variable","Hazard Ratio","SE","P-value")
concordance_df_temp <- as.data.frame(cbind("Rapid eGFR decline","Top protein",mod$concordance[[6]], mod$concordance[[7]]))
concordance_df <- rbind(concordance_df, concordance_df_temp)
```

```{r, include=TRUE, comment=""}
kable(moddf, row.names=F)
```

There are `r nrow(moddf)` variables in the model.

```{r, include=TRUE, comment=""}
kable(round(mod$concordance,digits=3))
```

## Model with 7 proteins nominally significant for all 3 outcomes

```{r, include=FALSE, comment=""}
form = as.formula(paste0("Surv(DAYSTORAPID.x, RAPID.x)~seq.16915.153+seq.18339.207+seq.3538.26+seq.9348.1+seq.7871.16+seq.13990.1+seq.13710.6"))
rapid_nozero_allproteins <- left_join(rapid_nozero, base_log_scale, by="releaseid")
mod = coxph(formula = form,data = rapid_nozero_allproteins)
kable(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)

# make a summary table
moddf <- as.data.frame(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)
moddf <- cbind(moddf$term,round(moddf$estimate, digits=3),round(moddf$std.error, digits=3),round(moddf$p.value, digits=3))
colnames(moddf) <- c("Variable","Hazard Ratio","SE","P-value")
concordance_df_temp <- as.data.frame(cbind("Rapid eGFR decline","Shared proteins",mod$concordance[[6]], mod$concordance[[7]]))
concordance_df <- rbind(concordance_df, concordance_df_temp)
```

```{r, include=TRUE, comment=""}
kable(moddf, row.names=F)
```

There are `r nrow(moddf)` variables in the model.

```{r, include=TRUE, comment=""}
kable(round(mod$concordance,digits=3))
```

# Summary tables and plots

```{r, include=FALSE, comment=""}
colnames(concordance_df) <- c("Outcome","Model","Concordance","SE")
concordance_df$Concordance <- as.numeric(concordance_df$Concordance)
concordance_df$SE <- as.numeric(concordance_df$SE)
concordance_df$ci_low <- concordance_df$Concordance - (1.96*concordance_df$SE)
concordance_df$ci_high <- concordance_df$Concordance + (1.96*concordance_df$SE)
concordance_df[,c("Concordance","SE","ci_low", "ci_high")] <- round(concordance_df[,c("Concordance","SE","ci_low", "ci_high")],2)
colnames(concordance_df) <- c("Outcome","Model","Concordance","SE","Lower bound 95% CI","Upper bound 95% CI")
#concordance_df <- rbind(concordance_df[3,],concordance_df[1,],concordance_df[2,],
#                        concordance_df[6,],concordance_df[4,],concordance_df[5,],
#                        concordance_df[9,],concordance_df[7,],concordance_df[8,])


```

```{r, include=TRUE, comment=""}
# DKD forest plot
concordance_df_for_plot <- concordance_df
concordance_df_for_plot$` ` <- paste(rep(" ", 40), collapse = " ")
concordance_df_for_plot$`Concordance (95% CI)` <- sprintf("%.2f (%.2f to %.2f)",
                                     concordance_df_for_plot$Concordance, concordance_df_for_plot$`Lower bound 95% CI`, concordance_df_for_plot$`Upper bound 95% CI`)
p <- forest(concordance_df_for_plot[, c(1:2, 7:8)], est = concordance_df_for_plot$Concordance, lower = concordance_df_for_plot$`Lower bound 95% CI`,
       upper = concordance_df_for_plot$`Upper bound 95% CI`, ci_column = 3, xlim = c(0,1))
#p$widths <- unit(c(1,100,50,50), "mm")
png('/Volumes/PEDS/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/Proteomics and DKD/Figures/forest_plot_elastic_net_DKD_protein_subsets.png', res = 300, width = 9, height = 3, units = "in")
p
dev.off()
png('/Users/pylell/Dropbox/TODAY HTN and DKD output/forest_plot_elastic_net_DKD_protein_subsets.png', res = 300, width = 9, height = 3, units = "in")
p
dev.off()
```


