---
title: "Melanie Cree - DNL mixed models"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include=FALSE}
library(readxl)
library(nlme)
library(emmeans)
library(knitr)

knitr::opts_chunk$set(echo = FALSE,warning = FALSE)

home_dir = ifelse(.Platform$OS.type != "unix","E:/Melanie Green/DNL-GLP1/data",
                  "/Volumes/som/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects")
knitr::opts_knit$set(root.dir = home_dir)
setwd(home_dir)

```

```{r data, include=FALSE}
data_2a <- read_xlsx("./For LP, PCOS DNL without pre-baseline.xlsx",sheet="Fig 2A",skip=1)
data_2b <- read_xlsx("./For LP, PCOS DNL without pre-baseline.xlsx",sheet="Fig 2B",skip=1)
data_2c <- read_xlsx("./For LP, PCOS DNL without pre-baseline.xlsx",sheet="Fig 2C",skip=1)
data_2d <- read_xlsx("./For LP, PCOS DNL without pre-baseline.xlsx",sheet="Fig 2D",skip=1)
data_2e <- read_xlsx("./For LP, PCOS DNL without pre-baseline.xlsx",sheet="Fig 2E",skip=1)
data_2f <- read_xlsx("./For LP, PCOS DNL without pre-baseline.xlsx",sheet="Fig 2F",skip=1)
data_2g <- read_xlsx("./For LP, PCOS DNL without pre-baseline.xlsx",sheet="Fig 2G",skip=1)
data_2h <- read_xlsx("./For LP, PCOS DNL without pre-baseline.xlsx",sheet="Fig 2H",skip=1)
data_3a <- read_xlsx("./For LP, PCOS DNL without pre-baseline.xlsx",sheet="Fig 3A",skip=1)
data_3b <- read_xlsx("./For LP, PCOS DNL without pre-baseline.xlsx",sheet="Fig 3B",skip=1)
data_3c <- read_xlsx("./For LP, PCOS DNL without pre-baseline.xlsx",sheet="Fig 3C",skip=1)
data_3d <- read_xlsx("./For LP, PCOS DNL without pre-baseline.xlsx",sheet="Fig 3D",skip=1)
data_3e <- read_xlsx("./For LP, PCOS DNL without pre-baseline.xlsx",sheet="Fig 3E",skip=1)
data_3f <- read_xlsx("./For LP, PCOS DNL without pre-baseline.xlsx",sheet="Fig 3F",skip=1)
data_3g <- read_xlsx("./For LP, PCOS DNL without pre-baseline.xlsx",sheet="Fig 3G",skip=1)
data_3h <- read_xlsx("./For LP, PCOS DNL without pre-baseline.xlsx",sheet="Fig 3H",skip=1)
```

# Methods

Mixed-effects models with a random intercept for participant were used to compare groups while accounting for the correlation of repeated measures within a participant. Models with and without a group by time interaction were run for each figure panel.

# How to interpret the output

For each figure panel, the first set of output comes from a model with a group by time interaction. If this interaction is significant, it indicates that the groups differ in how they change over time. To determine if the interaction is significant, look at the p-value for "as.factor(Group):as.factor(Time)" in the table labeled "Type 3 tests of fixed effects." If the interaction is significant, continue interpreting the rest of the output from the model containing the interaction, and ignore the results from the model without the interaction. If the interaction is not significant, ignore the rest of the results from the model with the interaction and interpret the results from the model without the intearction.

For a model with a significant interaction, the following describes how to interpret the remainder of the output. The second table, labeled "Time point means," provides the estimate of the mean at each time point within each group. The third table, labeled "Pair-wise comparisons," provides p-values for the group comparison at each time point.

For a model without a significant interaction, the first table, labeled "Type 3 tests of fixed effects," provides p-values for the overall effect of group (averaged across time), and time (averaged across group). The second table, labeled "Time point means," provides the estimate of the mean in each group. The third table, labeled "Pair-wise comparisons," provides p-values for the group comparison.

# Results

## Fig 2a

### With interaction between group and time

```{r echo=FALSE,comment=''}
fig2a_mod <- lme(Value ~ as.factor(Group)*as.factor(Time),random=~1|ID,
                  data = data_2a,na.action = na.omit)
fig2a_anova <- anova.lme(fig2a_mod, type="marginal")
fig2a_means <- emmeans(fig2a_mod, ~ as.factor(Group)*as.factor(Time))
fig2a_pairs <-  pairs(fig2a_means,adjust="tukey",simple="Group")
```

```{r, echo=FALSE, message=FALSE}
kable(fig2a_anova,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig2a_means,caption="Time point means.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig2a_pairs,caption="Pair-wise comparisons.",format.args = list(scientific = FALSE))
```

### Without interaction between group and time

```{r echo=FALSE,comment=''}
fig2a_mod <- lme(Value ~ as.factor(Group) + as.factor(Time),random=~1|ID,
                  data = data_2a,na.action = na.omit)
fig2a_anova <- anova.lme(fig2a_mod, type="marginal")
fig2a_means <- emmeans(fig2a_mod, ~ as.factor(Group))
fig2a_pairs <-  pairs(fig2a_means,adjust="tukey",simple="Group")
```

```{r, echo=FALSE, message=FALSE}
kable(fig2a_anova,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig2a_means,caption="Time point means.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig2a_pairs,caption="Pair-wise comparisons.",format.args = list(scientific = FALSE))
```

## Fig 2b

### With interaction between group and time

```{r echo=FALSE,comment=''}
fig2b_mod <- lme(Value ~ as.factor(Group)*as.factor(Time),random=~1|ID,
                  data = data_2b,na.action = na.omit)
fig2b_anova <- anova.lme(fig2b_mod, type="marginal")
fig2b_means <- emmeans(fig2b_mod, ~ as.factor(Group)*as.factor(Time))
fig2b_pairs <-  pairs(fig2b_means,adjust="tukey",simple="Group")
```

```{r, echo=FALSE, message=FALSE}
kable(fig2b_anova,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig2b_means,caption="Time point means.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig2b_pairs,caption="Pair-wise comparisons.",format.args = list(scientific = FALSE))
```

### Without interaction between group and time

```{r echo=FALSE,comment=''}
fig2b_mod <- lme(Value ~ as.factor(Group) + as.factor(Time),random=~1|ID,
                  data = data_2b,na.action = na.omit)
fig2b_anova <- anova.lme(fig2b_mod, type="marginal")
fig2b_means <- emmeans(fig2b_mod, ~ as.factor(Group))
fig2b_pairs <-  pairs(fig2b_means,adjust="tukey",simple="Group")
```

```{r, echo=FALSE, message=FALSE}
kable(fig2b_anova,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig2b_means,caption="Time point means.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig2b_pairs,caption="Pair-wise comparisons.",format.args = list(scientific = FALSE))
```

## Fig 2c

### With interaction between group and time

```{r echo=FALSE,comment=''}
fig2c_mod <- lme(Value ~ as.factor(Group)*as.factor(Time),random=~1|ID,
                  data = data_2c,na.action = na.omit)
fig2c_anova <- anova.lme(fig2c_mod, type="marginal")
fig2c_means <- emmeans(fig2c_mod, ~ as.factor(Group)*as.factor(Time))
fig2c_pairs <-  pairs(fig2c_means,adjust="tukey",simple="Group")
```

```{r, echo=FALSE, message=FALSE}
kable(fig2c_anova,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig2c_means,caption="Time point means.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig2c_pairs,caption="Pair-wise comparisons.",format.args = list(scientific = FALSE))
```

### Without interaction between group and time

```{r echo=FALSE,comment=''}
fig2c_mod <- lme(Value ~ as.factor(Group) + as.factor(Time),random=~1|ID,
                  data = data_2c,na.action = na.omit)
fig2c_anova <- anova.lme(fig2c_mod, type="marginal")
fig2c_means <- emmeans(fig2c_mod, ~ as.factor(Group))
fig2c_pairs <-  pairs(fig2c_means,adjust="tukey",simple="Group")
```

```{r, echo=FALSE, message=FALSE}
kable(fig2c_anova,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig2c_means,caption="Time point means.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig2c_pairs,caption="Pair-wise comparisons.",format.args = list(scientific = FALSE))
```

## Fig 2d

### With interaction between group and time

```{r echo=FALSE,comment=''}
fig2d_mod <- lme(Value ~ as.factor(Group)*as.factor(Time),random=~1|ID,
                  data = data_2d,na.action = na.omit)
fig2d_anova <- anova.lme(fig2d_mod, type="marginal")
fig2d_means <- emmeans(fig2d_mod, ~ as.factor(Group)*as.factor(Time))
fig2d_pairs <-  pairs(fig2d_means,adjust="tukey",simple="Group")
```

```{r, echo=FALSE, message=FALSE}
kable(fig2d_anova,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig2d_means,caption="Time point means.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig2d_pairs,caption="Pair-wise comparisons.",format.args = list(scientific = FALSE))
```

### Without interaction between group and time

```{r echo=FALSE,comment=''}
fig2d_mod <- lme(Value ~ as.factor(Group) + as.factor(Time),random=~1|ID,
                  data = data_2d,na.action = na.omit)
fig2d_anova <- anova.lme(fig2d_mod, type="marginal")
fig2d_means <- emmeans(fig2d_mod, ~ as.factor(Group))
fig2d_pairs <-  pairs(fig2d_means,adjust="tukey",simple="Group")
```

```{r, echo=FALSE, message=FALSE}
kable(fig2d_anova,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig2d_means,caption="Time point means.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig2d_pairs,caption="Pair-wise comparisons.",format.args = list(scientific = FALSE))
```

## Fig 2e

### With interaction between group and time

```{r echo=FALSE,comment=''}
fig2e_mod <- lme(Value ~ as.factor(Group)*as.factor(Time),random=~1|ID,
                  data = data_2e,na.action = na.omit)
fig2e_anova <- anova.lme(fig2e_mod, type="marginal")
fig2e_means <- emmeans(fig2e_mod, ~ as.factor(Group)*as.factor(Time))
fig2e_pairs <-  pairs(fig2e_means,adjust="tukey",simple="Group")
```

```{r, echo=FALSE, message=FALSE}
kable(fig2e_anova,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig2e_means,caption="Time point means.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig2e_pairs,caption="Pair-wise comparisons.",format.args = list(scientific = FALSE))
```

### Without interaction between group and time

```{r echo=FALSE,comment=''}
fig2e_mod <- lme(Value ~ as.factor(Group) + as.factor(Time),random=~1|ID,
                  data = data_2e,na.action = na.omit)
fig2e_anova <- anova.lme(fig2e_mod, type="marginal")
fig2e_means <- emmeans(fig2e_mod, ~ as.factor(Group))
fig2e_pairs <-  pairs(fig2e_means,adjust="tukey",simple="Group")
```

```{r, echo=FALSE, message=FALSE}
kable(fig2e_anova,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig2e_means,caption="Time point means.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig2e_pairs,caption="Pair-wise comparisons.",format.args = list(scientific = FALSE))
```

## Fig 2f

### With interaction between group and time

```{r echo=FALSE,comment=''}
fig2f_mod <- lme(Value ~ as.factor(Group)*as.factor(Time),random=~1|ID,
                  data = data_2f,na.action = na.omit)
fig2f_anova <- anova.lme(fig2f_mod, type="marginal")
fig2f_means <- emmeans(fig2f_mod, ~ as.factor(Group)*as.factor(Time))
fig2f_pairs <-  pairs(fig2f_means,adjust="tukey",simple="Group")
```

```{r, echo=FALSE, message=FALSE}
kable(fig2f_anova,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig2f_means,caption="Time point means.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig2f_pairs,caption="Pair-wise comparisons.",format.args = list(scientific = FALSE))
```

### Without interaction between group and time

```{r echo=FALSE,comment=''}
fig2f_mod <- lme(Value ~ as.factor(Group) + as.factor(Time),random=~1|ID,
                  data = data_2f,na.action = na.omit)
fig2f_anova <- anova.lme(fig2f_mod, type="marginal")
fig2f_means <- emmeans(fig2f_mod, ~ as.factor(Group))
fig2f_pairs <-  pairs(fig2f_means,adjust="tukey",simple="Group")
```

```{r, echo=FALSE, message=FALSE}
kable(fig2f_anova,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig2f_means,caption="Time point means.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig2f_pairs,caption="Pair-wise comparisons.",format.args = list(scientific = FALSE))
```

## Fig 2g

### With interaction between group and time

```{r echo=FALSE,comment=''}
fig2g_mod <- lme(Value ~ as.factor(Group)*as.factor(Time),random=~1|ID,
                  data = data_2g,na.action = na.omit)
fig2g_anova <- anova.lme(fig2g_mod, type="marginal")
fig2g_means <- emmeans(fig2g_mod, ~ as.factor(Group)*as.factor(Time))
fig2g_pairs <-  pairs(fig2g_means,adjust="tukey",simple="Group")
```

```{r, echo=FALSE, message=FALSE}
kable(fig2g_anova,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig2g_means,caption="Time point means.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig2g_pairs,caption="Pair-wise comparisons.",format.args = list(scientific = FALSE))
```

### Without interaction between group and time

```{r echo=FALSE,comment=''}
fig2g_mod <- lme(Value ~ as.factor(Group) + as.factor(Time),random=~1|ID,
                  data = data_2g,na.action = na.omit)
fig2g_anova <- anova.lme(fig2g_mod, type="marginal")
fig2g_means <- emmeans(fig2g_mod, ~ as.factor(Group))
fig2g_pairs <-  pairs(fig2g_means,adjust="tukey",simple="Group")
```

```{r, echo=FALSE, message=FALSE}
kable(fig2g_anova,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig2g_means,caption="Time point means.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig2g_pairs,caption="Pair-wise comparisons.",format.args = list(scientific = FALSE))
```

## Fig 2h

### With interaction between group and time

```{r echo=FALSE,comment=''}
fig2h_mod <- lme(Value ~ as.factor(Group)*as.factor(Time),random=~1|ID,
                  data = data_2h,na.action = na.omit)
fig2h_anova <- anova.lme(fig2h_mod, type="marginal")
fig2h_means <- emmeans(fig2h_mod, ~ as.factor(Group)*as.factor(Time))
fig2h_pairs <-  pairs(fig2h_means,adjust="tukey",simple="Group")
```

```{r, echo=FALSE, message=FALSE}
kable(fig2h_anova,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig2h_means,caption="Time point means.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig2h_pairs,caption="Pair-wise comparisons.",format.args = list(scientific = FALSE))
```

### Without interaction between group and time

```{r echo=FALSE,comment=''}
fig2h_mod <- lme(Value ~ as.factor(Group) + as.factor(Time),random=~1|ID,
                  data = data_2h,na.action = na.omit)
fig2h_anova <- anova.lme(fig2h_mod, type="marginal")
fig2h_means <- emmeans(fig2h_mod, ~ as.factor(Group))
fig2h_pairs <-  pairs(fig2h_means,adjust="tukey",simple="Group")
```

```{r, echo=FALSE, message=FALSE}
kable(fig2h_anova,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig2h_means,caption="Time point means.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig2h_pairs,caption="Pair-wise comparisons.",format.args = list(scientific = FALSE))
```

## Fig 3a

### With interaction between group and time

```{r echo=FALSE,comment=''}
fig3a_mod <- lme(Value ~ as.factor(Group)*as.factor(Time),random=~1|ID,
                  data = data_3a,na.action = na.omit)
fig3a_anova <- anova.lme(fig3a_mod, type="marginal")
fig3a_means <- emmeans(fig3a_mod, ~ as.factor(Group)*as.factor(Time))
fig3a_pairs <-  pairs(fig3a_means,adjust="tukey",simple="Group")
```

```{r, echo=FALSE, message=FALSE}
kable(fig3a_anova,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig3a_means,caption="Time point means.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig3a_pairs,caption="Pair-wise comparisons.",format.args = list(scientific = FALSE))
```

### Without interaction between group and time

```{r echo=FALSE,comment=''}
fig3a_mod <- lme(Value ~ as.factor(Group) + as.factor(Time),random=~1|ID,
                  data = data_3a,na.action = na.omit)
fig3a_anova <- anova.lme(fig3a_mod, type="marginal")
fig3a_means <- emmeans(fig3a_mod, ~ as.factor(Group))
fig3a_pairs <-  pairs(fig3a_means,adjust="tukey",simple="Group")
```

```{r, echo=FALSE, message=FALSE}
kable(fig3a_anova,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig3a_means,caption="Time point means.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig3a_pairs,caption="Pair-wise comparisons.",format.args = list(scientific = FALSE))
```

## Fig 3b

```{r echo=FALSE,comment=''}
fig3b_mod <- lme(Value ~ as.factor(Group)*as.factor(Time),random=~1|ID,
                  data = data_3b,na.action = na.omit)
fig3b_anova <- anova.lme(fig3b_mod, type="marginal")
fig3b_means <- emmeans(fig3b_mod, ~ as.factor(Group)*as.factor(Time))
fig3b_pairs <-  pairs(fig3b_means,adjust="tukey",simple="Group")
```

```{r, echo=FALSE, message=FALSE}
kable(fig3b_anova,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig3b_means,caption="Time point means.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig3b_pairs,caption="Pair-wise comparisons.",format.args = list(scientific = FALSE))
```

### Without interaction between group and time

```{r echo=FALSE,comment=''}
fig3b_mod <- lme(Value ~ as.factor(Group) + as.factor(Time),random=~1|ID,
                  data = data_3b,na.action = na.omit)
fig3b_anova <- anova.lme(fig3b_mod, type="marginal")
fig3b_means <- emmeans(fig3b_mod, ~ as.factor(Group))
fig3b_pairs <-  pairs(fig3b_means,adjust="tukey",simple="Group")
```

```{r, echo=FALSE, message=FALSE}
kable(fig3b_anova,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig3b_means,caption="Time point means.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig3b_pairs,caption="Pair-wise comparisons.",format.args = list(scientific = FALSE))
```

## Fig 3c

```{r echo=FALSE,comment=''}
fig3c_mod <- lme(Value ~ as.factor(Group)*as.factor(Time),random=~1|ID,
                  data = data_3c,na.action = na.omit)
fig3c_anova <- anova.lme(fig3c_mod, type="marginal")
fig3c_means <- emmeans(fig3c_mod, ~ as.factor(Group)*as.factor(Time))
fig3c_pairs <-  pairs(fig3c_means,adjust="tukey",simple="Group")
```

```{r, echo=FALSE, message=FALSE}
kable(fig3c_anova,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig3c_means,caption="Time point means.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig3c_pairs,caption="Pair-wise comparisons.",format.args = list(scientific = FALSE))
```

### Without interaction between group and time

```{r echo=FALSE,comment=''}
fig3c_mod <- lme(Value ~ as.factor(Group) + as.factor(Time),random=~1|ID,
                  data = data_3c,na.action = na.omit)
fig3c_anova <- anova.lme(fig3c_mod, type="marginal")
fig3c_means <- emmeans(fig3c_mod, ~ as.factor(Group))
fig3c_pairs <-  pairs(fig3c_means,adjust="tukey",simple="Group")
```

```{r, echo=FALSE, message=FALSE}
kable(fig3c_anova,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig3c_means,caption="Time point means.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig3c_pairs,caption="Pair-wise comparisons.",format.args = list(scientific = FALSE))
```

## Fig 3d

```{r echo=FALSE,comment=''}
fig3d_mod <- lme(Value ~ as.factor(Group)*as.factor(Time),random=~1|ID,
                  data = data_3d,na.action = na.omit)
fig3d_anova <- anova.lme(fig3d_mod, type="marginal")
fig3d_means <- emmeans(fig3d_mod, ~ as.factor(Group)*as.factor(Time))
fig3d_pairs <-  pairs(fig3d_means,adjust="tukey",simple="Group")
```

```{r, echo=FALSE, message=FALSE}
kable(fig3d_anova,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig3d_means,caption="Time point means.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig3d_pairs,caption="Pair-wise comparisons.",format.args = list(scientific = FALSE))
```

### Without interaction between group and time

```{r echo=FALSE,comment=''}
fig3d_mod <- lme(Value ~ as.factor(Group) + as.factor(Time),random=~1|ID,
                  data = data_3d,na.action = na.omit)
fig3d_anova <- anova.lme(fig3d_mod, type="marginal")
fig3d_means <- emmeans(fig3d_mod, ~ as.factor(Group))
fig3d_pairs <-  pairs(fig3d_means,adjust="tukey",simple="Group")
```

```{r, echo=FALSE, message=FALSE}
kable(fig3d_anova,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig3d_means,caption="Time point means.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig3d_pairs,caption="Pair-wise comparisons.",format.args = list(scientific = FALSE))
```

## Fig 3e

```{r echo=FALSE,comment=''}
fig3e_mod <- lme(Value ~ as.factor(Group)*as.factor(Time),random=~1|ID,
                  data = data_3e,na.action = na.omit)
fig3e_anova <- anova.lme(fig3e_mod, type="marginal")
fig3e_means <- emmeans(fig3e_mod, ~ as.factor(Group)*as.factor(Time))
fig3e_pairs <-  pairs(fig3e_means,adjust="tukey",simple="Group")
```

```{r, echo=FALSE, message=FALSE}
kable(fig3e_anova,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig3e_means,caption="Time point means.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig3e_pairs,caption="Pair-wise comparisons.",format.args = list(scientific = FALSE))
```

### Without interaction between group and time

```{r echo=FALSE,comment=''}
fig3e_mod <- lme(Value ~ as.factor(Group) + as.factor(Time),random=~1|ID,
                  data = data_3e,na.action = na.omit)
fig3e_anova <- anova.lme(fig3e_mod, type="marginal")
fig3e_means <- emmeans(fig3e_mod, ~ as.factor(Group))
fig3e_pairs <-  pairs(fig3e_means,adjust="tukey",simple="Group")
```

```{r, echo=FALSE, message=FALSE}
kable(fig3e_anova,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig3e_means,caption="Time point means.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig3e_pairs,caption="Pair-wise comparisons.",format.args = list(scientific = FALSE))
```

## Fig 3f

```{r echo=FALSE,comment=''}
fig3f_mod <- lme(Value ~ as.factor(Group)*as.factor(Time),random=~1|ID,
                  data = data_3f,na.action = na.omit)
fig3f_anova <- anova.lme(fig3f_mod, type="marginal")
fig3f_means <- emmeans(fig3f_mod, ~ as.factor(Group)*as.factor(Time))
fig3f_pairs <-  pairs(fig3f_means,adjust="tukey",simple="Group")
```

```{r, echo=FALSE, message=FALSE}
kable(fig3f_anova,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig3f_means,caption="Time point means.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig3f_pairs,caption="Pair-wise comparisons.",format.args = list(scientific = FALSE))
```

### Without interaction between group and time

```{r echo=FALSE,comment=''}
fig3f_mod <- lme(Value ~ as.factor(Group) + as.factor(Time),random=~1|ID,
                  data = data_3f,na.action = na.omit)
fig3f_anova <- anova.lme(fig3f_mod, type="marginal")
fig3f_means <- emmeans(fig3f_mod, ~ as.factor(Group))
fig3f_pairs <-  pairs(fig3f_means,adjust="tukey",simple="Group")
```

```{r, echo=FALSE, message=FALSE}
kable(fig3f_anova,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig3f_means,caption="Time point means.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig3f_pairs,caption="Pair-wise comparisons.",format.args = list(scientific = FALSE))
```

## Fig 3g

```{r echo=FALSE,comment=''}
fig3g_mod <- lme(Value ~ as.factor(Group)*as.factor(Time),random=~1|ID,
                  data = data_3g,na.action = na.omit)
fig3g_anova <- anova.lme(fig3g_mod, type="marginal")
fig3g_means <- emmeans(fig3g_mod, ~ as.factor(Group)*as.factor(Time))
fig3g_pairs <-  pairs(fig3g_means,adjust="tukey",simple="Group")
```

```{r, echo=FALSE, message=FALSE}
kable(fig3g_anova,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig3g_means,caption="Time point means.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig3g_pairs,caption="Pair-wise comparisons.",format.args = list(scientific = FALSE))
```

### Without interaction between group and time

```{r echo=FALSE,comment=''}
fig3g_mod <- lme(Value ~ as.factor(Group) + as.factor(Time),random=~1|ID,
                  data = data_3g,na.action = na.omit)
fig3g_anova <- anova.lme(fig3g_mod, type="marginal")
fig3g_means <- emmeans(fig3g_mod, ~ as.factor(Group))
fig3g_pairs <-  pairs(fig3g_means,adjust="tukey",simple="Group")
```

```{r, echo=FALSE, message=FALSE}
kable(fig3g_anova,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig3g_means,caption="Time point means.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig3g_pairs,caption="Pair-wise comparisons.",format.args = list(scientific = FALSE))
```

## Fig 3h

```{r echo=FALSE,comment=''}
fig3h_mod <- lme(Value ~ as.factor(Group)*as.factor(Time),random=~1|ID,
                  data = data_3h,na.action = na.omit)
fig3h_anova <- anova.lme(fig3h_mod, type="marginal")
fig3h_means <- emmeans(fig3h_mod, ~ as.factor(Group)*as.factor(Time))
fig3h_pairs <-  pairs(fig3h_means,adjust="tukey",simple="Group")
```

```{r, echo=FALSE, message=FALSE}
kable(fig3h_anova,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig3h_means,caption="Time point means.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig3h_pairs,caption="Pair-wise comparisons.",format.args = list(scientific = FALSE))
```

### Without interaction between group and time

```{r echo=FALSE,comment=''}
fig3h_mod <- lme(Value ~ as.factor(Group) + as.factor(Time),random=~1|ID,
                  data = data_3h,na.action = na.omit)
fig3h_anova <- anova.lme(fig3h_mod, type="marginal")
fig3h_means <- emmeans(fig3h_mod, ~ as.factor(Group))
fig3h_pairs <-  pairs(fig3h_means,adjust="tukey",simple="Group")
```

```{r, echo=FALSE, message=FALSE}
kable(fig3h_anova,caption="Type 3 tests of fixed effects.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig3h_means,caption="Time point means.",format.args = list(scientific = FALSE))
```

```{r, echo=FALSE, message=FALSE}
kable(fig3h_pairs,caption="Pair-wise comparisons.",format.args = list(scientific = FALSE))
```