---
title: "Harmonized Data Non-Manual Kidney Volume Analyses"
author: "Callie Rountree-Jablin"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 3
    toc_float: yes
    toc_loc: left
---

```{r setup, include=FALSE}
library(tidyverse)
library(kableExtra)
library(Hmisc)
library(gtsummary)
library(magrittr)
pacman::p_load(naniar)

if(Sys.info()["sysname"] == "Windows"){
  home_dir = "B:/Petter Bjornstad"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/Peds Endo/Petter Bjornstad/"
} 
knitr::opts_knit$set(echo=FALSE, root.dir = home_dir)
```
# Overview

The aim of this report is to:

 1. Separately describe non-manual left and right kidney volume across all studies by age and disease type, using the harmonized dataset. 

 
```{r read, include=FALSE}
## read raw data and data dictionary
raw_dat <- read.csv("./Data Harmonization/Data Clean/harmonized_dataset.csv", na.strings = c(" ", "", "-9999",-9999))
dict <- read.csv("./Data Harmonization/data_dictionary_master.csv", na.strings = c(" ", "", "-9999",-9999)) %>% 
  select(variable_name, label)

## import co-enrolled IDs for IMPROVE, RH, and RH2; transpose to long format
coenroll_id <- read.csv("./Renal HERITAGE/Data_Cleaned/coenrolled_ids.csv") %>%
  pivot_longer(cols = 'improve_id':'crc_id',
               values_to = "record_id") %>% 
  select(merged_id, record_id) %>%
  filter(record_id != "")
```


```{r non-manual kv analysis data cleaning, include=FALSE}
### data cleaning for non-manual analyses

## clean main df
# add in merged IDs to data, filter down to studies and procedures where kv data was collected
# exclude PKD subjects, define variable for categorizing manual v. non-manual method
clean_dat <- left_join(raw_dat, coenroll_id, by = 'record_id') %>% 
  relocate(merged_id, .after = co_enroll_id) %>%
  filter(((study == 'CROCODILE'| study == 'PANDA') & procedure == 'bold_mri') | 
           ((study == 'PANTHER'| study == 'RENAL-HEIR'| study =='RENAL-HEIRitage') & procedure == 'clamp') | 
           (study == 'PENGUIN' & procedure == 'mri')) %>% 
  filter(participation_status != 'Removed'|is.na(participation_status)) %>% 
  filter(group != 'PKD') %>% 
  mutate(date = as.Date(date, format = '%Y-%m-%d'), # formatted date to use arrange function later on
         lkv_metric_type = case_when(!is.na(left_kidney_volume_ml) & is.na(volume_left_manual) ~ 'Non-Manual',
                                      is.na(left_kidney_volume_ml) & !is.na(volume_left_manual) ~ 'Manual',
                                     !is.na(left_kidney_volume_ml) & !is.na(volume_left_manual) ~ 'Non-Manual and Manual'),
         rkv_metric_type = case_when(!is.na(right_kidney_volume_ml) & is.na(volume_right_manual) ~ 'Non-Manual',
                                      is.na(right_kidney_volume_ml) & !is.na(volume_right_manual) ~ 'Manual',
                                     !is.na(right_kidney_volume_ml) & !is.na(volume_right_manual) ~ 'Non-Manual and Manual'),
         height_meters = height*0.01)

## create separate dfs for stratifying analyses by left and right kidney
left_clean_dat <- clean_dat %>% filter(lkv_metric_type == 'Non-Manual' | lkv_metric_type == 'Non-Manual and Manual')
right_clean_dat <- clean_dat %>% filter(rkv_metric_type == 'Non-Manual' | rkv_metric_type == 'Non-Manual and Manual')

rm(clean_dat)

## define function for collapsing subjects who are represented more than once to 1 row
# function cleans RH, RH2, CROCODILE, and PANDA co-enrolled subjects as well as PANTHER subjects w/multiple visits by
# first dropping rows with no left/right kv data at all, then keeping subject's earliest measurements for all variables 
coenroll_multivisit_cleaning <- function(dat, kv_varname){
  dat %>%
    filter(study == 'RENAL-HEIRitage' | study == 'RENAL-HEIR' | study == 'CROCODILE' | study == 'PANDA' | study == 'PANTHER') %>%
    filter(!is.na({{kv_varname}})) %>% 
    arrange(merged_id, visit, date) %>%
    mutate(date = as.character(date)) %>% 
    summarise(across(where(negate(is.numeric)), ~ ifelse(all(is.na(.x)), NA_character_, first(na.omit(.x)))),
              across(where(is.numeric), ~ ifelse(all(is.na(.x)), NA_real_, first(na.omit(.x)))), .by = c(merged_id)) %>% 
    mutate(date = as.Date(date, format = '%Y-%m-%d'))
}

## separately clean left and right data

# collapse to one row per subject (no need to merge back into main left/right dfs since PENGUIN subjects are dropped with PKD exclusion)
left_combined_clean <- coenroll_multivisit_cleaning(dat = left_clean_dat, kv_varname = left_kidney_volume_ml) %>% 
  rename('lkv_date' = 'date')

right_combined_clean <- coenroll_multivisit_cleaning(dat = right_clean_dat, kv_varname = right_kidney_volume_ml) %>% 
  rename('rkv_date' = 'date')

# save a copy for later check
og_left_combined_clean <- left_combined_clean

og_right_combined_clean <- right_combined_clean

# add in missing height data, calculate ht. adj. kv
additional_height_dat <- left_join(raw_dat, coenroll_id, by = 'record_id') %>% 
  relocate(merged_id, .after = co_enroll_id) %>%
  filter(study == 'PANTHER' | study == 'CROCODILE' | study == 'PANDA' | study == 'RENAL-HEIRitage') %>% 
  filter(participation_status != 'Removed'|is.na(participation_status)) %>%
  filter(!is.na(height)) %>%
  select(merged_id, study, visit, procedure, date, height) %>% 
  group_by(merged_id, visit, date) %>% 
  mutate(additional_height_meters = 0.01*mean(height)) %>% # take mean height if dates are exact same
  slice(1) %>% 
  ungroup() %>% 
  as.data.frame() %>% 
  select(-c(procedure, height)) %>% 
  rename('additional_height_date' = 'date')


lkv_additional_heights <- left_join(left_combined_clean, additional_height_dat, by = c('merged_id', 'study', 'visit')) %>% 
  mutate(lkv_add_height_date_diff = as.numeric(difftime(as.Date(lkv_date), as.Date(additional_height_date), units = 'days'))) %>% 
  select(merged_id, study, visit, additional_height_date, lkv_add_height_date_diff, additional_height_meters) %>%
  filter(study == 'PANTHER' | study == 'CROCODILE' | study == 'PANDA' | study == 'RENAL-HEIRitage') %>% 
  filter(lkv_add_height_date_diff >= 0) %>%  # only use heights taken on or before date of kv measurement
  group_by(merged_id) %>%
  slice_min(lkv_add_height_date_diff, n = 1) %>%  
  ungroup()

left_combined_clean <- left_join(left_combined_clean, lkv_additional_heights, by = c('merged_id', 'study', 'visit')) %>% 
  mutate(height_meters =
           case_when(!is.na(height_meters) | !is.na(additional_height_meters) ~ rowSums(select(., height_meters, additional_height_meters),
                                                                                             na.rm = TRUE))) %>%
  mutate(ht_adj_left_kv = left_kidney_volume_ml/height_meters) %>% 
  select(-additional_height_meters)

rkv_additional_heights <- left_join(right_combined_clean, additional_height_dat, by = c('merged_id', 'study', 'visit')) %>% 
  mutate(rkv_add_height_date_diff = as.numeric(difftime(as.Date(rkv_date), as.Date(additional_height_date), units = 'days'))) %>% 
  select(merged_id, study, visit, additional_height_date, rkv_add_height_date_diff, additional_height_meters) %>%
  filter(study == 'PANTHER' | study == 'CROCODILE' | study == 'PANDA' | study == 'RENAL-HEIRitage') %>% 
  filter(rkv_add_height_date_diff >= 0) %>%  # only use heights taken on or before date of kv measurement
  group_by(merged_id) %>%
  slice_min(rkv_add_height_date_diff, n = 1) %>%  
  ungroup()

right_combined_clean <- left_join(right_combined_clean, rkv_additional_heights, by = c('merged_id', 'study', 'visit')) %>% 
  mutate(height_meters =
           case_when(!is.na(height_meters) | !is.na(additional_height_meters) ~ rowSums(select(., height_meters, additional_height_meters),
                                                                                             na.rm = TRUE))) %>%
  mutate(ht_adj_right_kv = right_kidney_volume_ml/height_meters) %>% 
  select(-additional_height_meters)

rm(rkv_additional_heights, lkv_additional_heights)

## check - everything working as intended
og_left_combined_clean <- og_left_combined_clean %>% select(-height_meters)
check_left_combined <- left_combined_clean %>% select(-height_meters)
identical(og_left_combined_clean, check_left_combined[ , -c(1015:1017)])

og_right_combined_clean <- og_right_combined_clean %>% select(-height_meters)
check_right_combined <- right_combined_clean %>% select(-height_meters)
identical(og_right_combined_clean, check_right_combined[, -c(1015:1017)])

rm(og_left_combined_clean, check_left_combined, og_right_combined_clean, check_right_combined)

## labels
rownames(dict) <- dict$variable_name
dict %<>% select("label")
dict <- t(dict) %>% as.data.frame(dict)
rownames(dict) <- "label"

data_labeling <- function(dat, dictionary){
  dictionary <- dictionary[intersect(names(dat), names(dictionary))]
  dictionary[setdiff(names(dat), names(dictionary))] <- " "
  label(dat) = as.list(dictionary[match(names(dat), names(dictionary))])
  label(dat$left_kidney_volume_ml) <- "Non-Manually Measured Left Kidney Volume (mL)"
  label(dat$right_kidney_volume_ml) <- "Non-Manually Measured Right Kidney Volume (mL)"
  label(dat$lkv_metric_type) <- "Left Kidney Volume Measure Type"
  label(dat$rkv_metric_type) <- "Right Kidney Volume Measure Type"
  label(dat$height_meters) <- "Height (m)"

  return(dat)
}

left_combined_clean <- data_labeling(dat = left_combined_clean, dictionary = dict)
right_combined_clean <- data_labeling(dat = right_combined_clean, dictionary = dict)

label(left_combined_clean$ht_adj_left_kv) <- "Height Adjusted Non-Manual LKV (mL/m)"
label(right_combined_clean$ht_adj_right_kv) <- "Height Adjusted Non-Manual RKV (mL/m)"
label(left_combined_clean$lkv_add_height_date_diff) <- "Days Between LKV and Height Measures for PANTHER/CROCODILE/PANDA/RH2 Subjects "
label(right_combined_clean$rkv_add_height_date_diff) <- "Days Between RKV and Height Measures for PANTHER/CROCODILE/PANDA/RH2 Subjects"

## workspace cleanup
rm(additional_height_dat, right_clean_dat, left_clean_dat)
```


# Methods

## Data Cleaning

The primary outcome of interest for this report was non-manually calculated left or right kidney volume (KV). Results were reported for this KV variable and additionally for a height-adjusted version of this variable $\bigl(\frac{\text{KV}}{\text{height (m)}}\bigr)$. In instances where subjects had multiple KV measurements or were enrolled on multiple studies, their earliest KV measurements were used for analyses; data from the same visit the KV measurement was taken was referenced for the covariates of interest (age, disease type, etc.) unless missing in which case covariate data from the closest visit in time was used.

## Data Analysis

**All Studies (CROCODILE, PANDA, RENAL-HEIR, PANTHER, RENAL-HEIRitage):**

Unadjusted and height-adjusted KV were modeled using simple linear regression for describing associations with age and Kruskal-Wallis testing for describing KV differences by disease type. All analyses were stratified by left and right kidney; separate descriptive statistics were reported for left versus right analyses to account for subjects with only one kidney. 

R v 4.2.3 was used for all analyses.

# Left Kidney Analyses

## Results: All Studies

### Descriptive Statistics
```{r lkv tab1, echo=FALSE}
table1::table1(~ sex + age + group + diabetes_duration + height_meters + left_kidney_volume_ml + ht_adj_left_kv + 
                 lkv_metric_type + lkv_add_height_date_diff, 
               data = left_combined_clean, caption = 'Descriptive Statistics of all Participants for Left Kidney Analyses')
```

Note: Height data from the same visit KV measures were collected was not available for subjects on PANTHER, CROCODILE, PANDA, or RENAL-HEIRitage so height measurements that were closest in time to the KV measures were instead used. Only heights taken on or before the KV measurement date were used. "Days Between LKV and Height Measures for PANTHER/CROCODILE/PANDA/RH2 Subjects" summarizes the distance in days between the KV measures and height measures. 

### LKV by Age

The below results describe the association between left KV and age for subjects across all studies. Diagnostic plots for checking linear regression assumptions are also included. `left_kidney_volume_ml` is the unadjusted non-manual left KV. `ht_adj_left_kv` is the height-adjusted version of `left_kidney_volume_ml`.

```{r lkv by age, echo=FALSE}
outcome_vars <- c('left_kidney_volume_ml','ht_adj_left_kv')

for (v in outcome_vars) {
  form = as.formula(paste0(v,'~ age'))
  mod <- lm(form, data = left_combined_clean)
  print(noquote(c("Association between age and", noquote(v))))
  print(summary(mod))
  plot(x=left_combined_clean$age, y=left_combined_clean[[v]], 
       xlab='age', ylab=noquote(v), main='Y-X Scatterplot'); abline(mod, col='gray', lty=2)
  plot(mod, which=1)
  plot(ppoints(length(rstudent(mod))), sort(pnorm(rstudent(mod))), xlab='Observed Cumulative Probability', ylab='Expected Cumulative Probability',
       main='Normal Probability Plot', cex=1); abline(a=0,b=1, col='gray65', lwd=1)
  hist(rstudent(mod), xlab='Jackknife Residual', main='Histogram of Residuals', freq=F)
  cat('\n\n')
}
```

### LKV by Disease Type

The below table reports the mean and standard deviation (SD) of the non-manual KV outcomes of interest by disease type across all subjects in the harmonized dataset. Kruskal-Wallis testing was used to evaluate whether there was a significant difference in the mean ranks between groups.

```{r lkv disease type assumptions, eval=FALSE, include=FALSE}
## assumptions check
outcome_vars <- c('left_kidney_volume_ml','ht_adj_left_kv')

for (v in outcome_vars) {
  form = as.formula(paste0(v,'~ group'))
  mod <- aov(form, data = left_combined_clean)

  qqnorm(mod$residuals)
  qqline(mod$residuals)

  bartlett_result <- bartlett.test(form, data = left_combined_clean)
  print(bartlett_result)
}
```


```{r lkv by disease type, echo=FALSE}
## summary table with Kruskal-Wallis for testing
left_combined_clean %>% 
  select(merged_id, group, left_kidney_volume_ml, ht_adj_left_kv) %>% 
   tbl_summary(by = group, include = -merged_id,
              type = list(c(left_kidney_volume_ml, ht_adj_left_kv) ~ 'continuous', 
                        group ~ 'categorical'),
    statistic=list(all_continuous() ~ '{mean} ({sd})'),
    digits=everything()~c(2,2),
    missing_text='N Missing') %>% 
  add_p(test = all_continuous() ~ 'kruskal.test',
        pvalue_fun=~style_pvalue(.x, digits=3)) %>% 
  modify_header(label='**Measurement**')
```


# Right Kidney Analyses

## Results: All Studies

### Descriptive Statistics
```{r rkv tab1, echo=FALSE}
table1::table1(~ sex + age + group + diabetes_duration + height_meters + right_kidney_volume_ml + 
                 ht_adj_right_kv + rkv_metric_type + rkv_add_height_date_diff, 
               data = right_combined_clean, caption = 'Descriptive Statistics of all Participants for Right Kidney Analyses')
```

Note: Height data from the same visit KV measures were collected was not available for subjects on PANTHER, CROCODILE, PANDA, or RENAL-HEIRitage so height measurements that were closest in time to the KV measures were instead used. Only heights taken on or before the KV measurement date were used. "Days Between RKV and Height Measures for PANTHER/CROCODILE/PANDA/RH2 Subjects" summarizes the distance in days between the KV measures and height measures.

### RKV by Age

The below results describe the association between right KV and age for subjects across all studies. Diagnostic plots for checking linear regression assumptions are also included. `right_kidney_volume_ml` is the non-manual KV measurement. `ht_adj_right_kv` is the height-adjusted version of `right_kidney_volume_ml`.

```{r rkv by age, echo=FALSE}
outcome_vars <- c('right_kidney_volume_ml','ht_adj_right_kv')

for (v in outcome_vars) {
  form = as.formula(paste0(v,'~ age'))
  mod <- lm(form, data = right_combined_clean)
  print(noquote(c("Association between age and", noquote(v))))
  print(summary(mod))
  plot(x=right_combined_clean$age, y=right_combined_clean[[v]], xlab='age', ylab=noquote(v), main='Y-X Scatterplot'); abline(mod, col='gray', lty=2)
  plot(mod, which=1)
  plot(ppoints(length(rstudent(mod))), sort(pnorm(rstudent(mod))), xlab='Observed Cumulative Probability', ylab='Expected Cumulative Probability',
       main='Normal Probability Plot', cex=1); abline(a=0,b=1, col='gray65', lwd=1)
  hist(rstudent(mod), xlab='Jackknife Residual', main='Histogram of Residuals', freq=F)
  cat('\n\n')
}
```

### RKV by Disease Type

The below table reports the mean and standard deviation (SD) of the non-manual KV outcomes of interest by disease type across all subjects in the harmonized dataset. Kruskal-Wallis testing was used to evaluate whether there was a significant difference in the mean ranks between groups.

```{r rkv disease type assumptions, eval=FALSE, include=FALSE}
## assumptions check
outcome_vars <- c('right_kidney_volume_ml','ht_adj_right_kv')

for (v in outcome_vars) {
  form = as.formula(paste0(v,'~ group'))
  mod <- aov(form, data = right_combined_clean)

  qqnorm(mod$residuals)
  qqline(mod$residuals)

  bartlett_result <- bartlett.test(form, data = right_combined_clean)
  print(bartlett_result)
}
```


```{r rkv by disease type, echo=FALSE}
## summary table with Kruskal-Wallis for testing
right_combined_clean %>%
  select(record_id, group, right_kidney_volume_ml, ht_adj_right_kv) %>%
   tbl_summary(by = group, include = -record_id,
              type = list(c(right_kidney_volume_ml, ht_adj_right_kv) ~ 'continuous',
                        group ~ 'categorical'),
    statistic=list(all_continuous() ~ '{mean} ({sd})'),
    digits=everything()~c(2,2),
    missing_text='N Missing') %>%
  add_p(test = all_continuous() ~ 'kruskal.test',
        pvalue_fun=~style_pvalue(.x, digits=3)) %>%
  modify_header(label='**Measurement**')
```


