---
title: "MANGO MMTT Mixed Models <br>(Preliminary Results)"
author: "Callie Rountree-Jablin"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include=FALSE}
library(tidyverse)
library(nlme)
library(emmeans)
library(kableExtra)
knitr::opts_chunk$set(echo = FALSE,warning = FALSE)
options(scipen=999)
```


```{r read, include=FALSE}
raw_dat <- read_csv("Mango-MangoPreandPostAsses_DATA_2024-04-12_1037.csv")
```


```{r data cleaning and formatting, include=FALSE}
## create dfs for each assay
glucose <- raw_dat %>% select(pid, preop_glucose_neg10:preop_glucose_240, postop_glucose_neg10:postop_glucose_240) # checked
insulin <- raw_dat %>% select(pid, preop_i_neg10:preop_i_240, postop_i_neg10:postop_i_240)
ffa <- raw_dat %>% select(pid, preop_ffa_neg_10:preop_faa_240, postop_ffa_neg_10:postop_faa_240) %>% 
  rename(c('preop_faa_neg10' = preop_ffa_neg_10, 'postop_faa_neg10' = postop_ffa_neg_10))
glp1 <- raw_dat %>% select(pid, preop_glp_neg10:preop_glp_120, postop_glp_neg10:postop_glp_120)
glucagon <- raw_dat %>% select(pid, preop_glucagon_neg10:preop_glucagon_240, postop_glucagon_neg10:postop_glucagon_240)
triglycerides <- raw_dat %>%  select(pid, preop_trig_0:preop_trig_120, postop_trig_0:postop_trig_120)
cpep <- raw_dat %>% select(pid, preop_cpep_neg10:preop_cpep_240, postop_cpep_neg10:postop_cpep_240)

## switch to doubly long format
######################################################################################################################
# test code for converting to doubly long (used for creating pivot_doubly_long function)
#
# glucose_long <- glucose %>%  pivot_longer(cols = starts_with("preop_glucose") | starts_with("postop_glucose"), 
#                          names_to = c("study_visit", ".value", "time"), 
#                          names_sep = "_",
#                          values_to = "glucose")
#
######################################################################################################################

# define function for converting from wide to doubly long format
# .value specifies that the assay readings should be used as values in the resulting df instead of the assay name (for example, the word "glucose")
pivot_doubly_long <- function(df, assay_name){
  df %>% pivot_longer(cols = starts_with(paste0("preop_", assay_name))| starts_with(paste0("postop_", assay_name)),
                      names_to = c("study_visit", ".value", "time"),
                      names_sep = "_",
                      values_to = assay_name)
}

glucose_long <- pivot_doubly_long(glucose, "glucose")
insulin_long <- pivot_doubly_long(insulin, "i")
ffa_long <- pivot_doubly_long(ffa, "faa")
glp1_long <- pivot_doubly_long(glp1, "glp")
glucagon_long <- pivot_doubly_long(glucagon, "glucagon")
triglycerides_long <- pivot_doubly_long(triglycerides, "trig")
cpep_long <- pivot_doubly_long(cpep, "cpep")

## clean up variables
# define function for cleaning up each of the doubly long dfs
clean_DL_dfs <- function(df_name){
  df <- get(df_name)
  
  df <- df %>%
    mutate(time = recode(time, 'neg10' = '-10')) %>% 
    mutate(time = as.factor(as.numeric(time)),
           study_visit = factor(study_visit, levels = c('preop', 'postop')))
  
  assign(df_name, df, envir = .GlobalEnv)
}

dfs_to_clean <- c("glucose_long", "insulin_long", "ffa_long", "glp1_long", "glucagon_long", "triglycerides_long", "cpep_long")
invisible(lapply(dfs_to_clean, clean_DL_dfs))
```


# Methods

Mixed effects models were used to evaluate differences in MMTT labs before and after VSG bariatric surgery, while adjusting for the correlation of repeated measures within a participant. Estimated marginal means were calculated for each timepoint of the MMTT within a visit, and pairwise comparisons were used to test whether each timepoint differed across visits. 

Analyses for each of the MMTT labs were based on the following timepoints:

* Glucose, Insulin: -10, 0, 10, 20, 30, 45, 60, 90, 120, 150, 180, 210, 240 min
* FFA: -10, 0, 30, 60, 120, 180, 240 min
* GLP-1: -10,0, 10, 30, 45, 60, 120 min
* Glucagon: -10, 0, 10, 30, 45, 60, 120, 240 min
* Triglycerides: 0, 30, 60, 120 min
* C-Peptide: -10, 0, 10, 20, 30, 60, 90, 120, 180, 240 min


```{r mixed function, include=FALSE}
# define function that fits LMM then outputs a list of the ANOVA, marginal mean, and pairwise comparisons results
# inputted dat is the doubly long df for the assay of interest
mixed <- function(dat, outcome){
  form = as.formula(paste0(outcome,"~ study_visit * time"))
  mod <- lme(as.formula(form), random = ~1|pid, data = dat, na.action = na.omit,
             correlation = corCompSymm(form = ~1 | pid)) # specifies within-subject variability as exchangeable
  anova <- anova.lme(mod, type="marginal") # use ANOVA to obtain significance of interaction
  m <- emmeans(mod,c("study_visit","time")) # compute estimated marginal means
  prs <-  pairs(m,adjust="tukey",by="time") # perform pairwise comparisons by timepoint
  
  return(list(anova, m, prs))
}
```


# Results
How to interpret the output:

* ANOVA Results: This table provides the parameter estimates for each term in the model. The `study_visit:time` term tests whether the pattern of change of the lab measure throughout the duration of the MMTT differed between the pre-op and post-op visits. For example, based on the insulin results, we can see that the trajectory of insulin during the MMTT did differ between visits (p < 0.05). In instances where the `study_visit:time` term has a p-value of $\ge0.05$, the estimated marginal means and pairwise comparisons tables are omitted from the report. 

* Estimated Marginal Means: This table reports the estimated marginal means for each visit and timepoint. `lower.CL` and `upper.CL` respectively describe the lower and upper bounds of the 95% confidence interval.

* Pairwise Comparisons: This table provides p-values for comparing whether there is a notable difference in the MMTT measure between the pre-op and post-op visits at each timepoint. For example, looking at the first row of the glucose pairwise comparisons table, glucose at the -10 min timepoint was not significantly different between the two study visits (p=0.284). 

## Glucose
```{r glucose analysis, echo=FALSE}
glucose_results <- mixed(dat = glucose_long, outcome = "glucose")

glucose_results[[1]] %>% 
  kable(caption = 'ANOVA Results:', booktabs = TRUE, digits = 3) %>% 
  kable_styling(latex_options='basic', html_font='Arial') %>% 
      htmltools::HTML()
glucose_results[[2]] %>% 
  kable(caption = 'Estimated Marginal Means:', booktabs = TRUE, digits =3) %>% 
  kable_styling(latex_options='basic', html_font='Arial') %>% 
      htmltools::HTML()
glucose_results[[3]] %>%
  kable(caption = 'Pairwise Comparisons:', booktabs = TRUE, digits =3) %>% 
  kable_styling(latex_options='basic', html_font='Arial') %>% 
      htmltools::HTML()
```

## Insulin

```{r insulin analysis, echo=FALSE}
insulin_results <- mixed(dat = insulin_long, outcome = "i")

insulin_results[[1]] %>% 
  kable(caption = 'ANOVA Results:', booktabs = TRUE, digits = 3) %>% 
  kable_styling(latex_options='basic', html_font='Arial') %>% 
      htmltools::HTML()
insulin_results[[2]] %>% 
  kable(caption = 'Estimated Marginal Means:', booktabs = TRUE, digits =3) %>% 
  kable_styling(latex_options='basic', html_font='Arial') %>% 
      htmltools::HTML()
insulin_results[[3]] %>% 
  kable(caption = 'Pairwise Comparisons:', booktabs = TRUE, digits =3) %>% 
  kable_styling(latex_options='basic', html_font='Arial') %>% 
      htmltools::HTML()
```

## FFA

```{r ffa analysis, echo=FALSE}
ffa_results <- mixed(dat = ffa_long, outcome = "faa")

ffa_results[[1]] %>% 
  kable(caption = 'ANOVA Results:', booktabs = TRUE, digits = 3) %>% 
  kable_styling(latex_options='basic', html_font='Arial') %>% 
      htmltools::HTML()
ffa_results[[2]] %>% 
  kable(caption = 'Estimated Marginal Means:', booktabs = TRUE, digits =3) %>% 
  kable_styling(latex_options='basic', html_font='Arial') %>% 
      htmltools::HTML()
ffa_results[[3]] %>% 
  kable(caption = 'Pairwise Comparisons:', booktabs = TRUE, digits =3) %>% 
  kable_styling(latex_options='basic', html_font='Arial') %>% 
      htmltools::HTML()
```

## GLP-1

```{r glp1 analysis, echo=FALSE}
glp1_results <- mixed(dat = glp1_long, outcome = "glp")

glp1_results[[1]] %>% 
  kable(caption = 'ANOVA Results:', booktabs = TRUE, digits = 3) %>% 
  kable_styling(latex_options='basic', html_font='Arial') %>% 
      htmltools::HTML()
glp1_results[[2]] %>% 
  kable(caption = 'Estimated Marginal Means:', booktabs = TRUE, digits =3) %>% 
  kable_styling(latex_options='basic', html_font='Arial') %>% 
      htmltools::HTML()
glp1_results[[3]] %>% 
  kable(caption = 'Pairwise Comparisons:', booktabs = TRUE, digits =3) %>% 
  kable_styling(latex_options='basic', html_font='Arial') %>% 
      htmltools::HTML()
```

## Glucagon
<span style="color: blue;">Estimated marginal means and pairwise comparisons were not computed due to non-significant findings.</span>
```{r glucagon analysis, echo=FALSE}
glucagon_results <- mixed(dat = glucagon_long, outcome = "glucagon")

glucagon_results[[1]] %>% 
  kable(caption = 'ANOVA Results:', booktabs = TRUE, digits = 3) %>% 
  kable_styling(latex_options='basic', html_font='Arial') %>% 
      htmltools::HTML()
# glucagon_results[[2]] %>% 
#   kable(caption = 'Estimated Marginal Means:', booktabs = TRUE, digits =3) %>% 
#   kable_styling(latex_options='basic', html_font='Arial') %>% 
#       htmltools::HTML()
# glucagon_results[[3]] %>% 
#   kable(caption = 'Pairwise Comparisons:', booktabs = TRUE, digits =3) %>% 
#   kable_styling(latex_options='basic', html_font='Arial') %>% 
#       htmltools::HTML()
```

## Triglycerides
<span style="color: blue;">Estimated marginal means and pairwise comparisons were not computed due to non-significant findings.</span>
```{r triglycerides analysis, echo=FALSE}
triglycerides_results <- mixed(dat = triglycerides_long, outcome = "trig")

triglycerides_results[[1]] %>% 
  kable(caption = 'ANOVA Results:', booktabs = TRUE, digits = 3) %>% 
  kable_styling(latex_options='basic', html_font='Arial') %>% 
      htmltools::HTML()
# triglycerides_results[[2]] %>% 
#   kable(caption = 'Estimated Marginal Means:', booktabs = TRUE, digits =3) %>% 
#   kable_styling(latex_options='basic', html_font='Arial') %>% 
#       htmltools::HTML()
# triglycerides_results[[3]] %>% 
#   kable(caption = 'Pairwise Comparisons:', booktabs = TRUE, digits =3) %>% 
#   kable_styling(latex_options='basic', html_font='Arial') %>% 
#       htmltools::HTML()
```

## C-Peptide
<span style="color: blue;">Estimated marginal means and pairwise comparisons were not computed due to non-significant findings.</span>
```{r cpep analysis, echo=FALSE}
cpep_results <- mixed(dat = cpep_long, outcome = "cpep")

cpep_results[[1]] %>% 
  kable(caption = 'ANOVA Results:', booktabs = TRUE, digits = 3) %>% 
  kable_styling(latex_options='basic', html_font='Arial') %>% 
      htmltools::HTML()
# cpep_results[[2]] %>% 
#   kable(caption = 'Estimated Marginal Means:', booktabs = TRUE, digits =3) %>% 
#   kable_styling(latex_options='basic', html_font='Arial') %>% 
#       htmltools::HTML()
# cpep_results[[3]] %>% 
#   kable(caption = 'Pairwise Comparisons:', booktabs = TRUE, digits =3) %>% 
#   kable_styling(latex_options='basic', html_font='Arial') %>% 
#       htmltools::HTML()
```

# Questions
* Are the glucagon analyses based on the correct timepoints?