---
title: "MANGO Preliminary Analysis Results"
author: "Callie Rountree-Jablin"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include=FALSE}
library(tidyverse)
library(nlme)
library(lme4)
library(emmeans)
library(kableExtra)
library(gtsummary)
pacman::p_load(naniar)
options(scipen=999)

home_dir = "B:/Melanie Green/MANGO"
knitr::opts_knit$set(echo=FALSE, root.dir = home_dir)
```


```{r read, include=FALSE}
raw_dat <- read_csv("./Raw Data/Mango-MangoPreandPostAsses_DATA_2024-07-17_1932.csv")
```


```{r LMM data cleaning, include=FALSE}
## create dfs for each assay
glucose <- raw_dat %>% select(pid, preop_glucose_neg10:preop_glucose_240, postop_glucose_neg10:postop_glucose_240)
insulin <- raw_dat %>% select(pid, preop_i_neg10:preop_i_240, postop_i_neg10:postop_i_240)
ffa <- raw_dat %>% select(pid, preop_ffa_neg_10:preop_faa_240, postop_ffa_neg_10:postop_faa_240) %>% 
  rename(c(preop_faa_neg10 = preop_ffa_neg_10, postop_faa_neg10 = postop_ffa_neg_10))
glp1 <- raw_dat %>% select(pid, preop_glp_neg10:preop_glp_120, postop_glp_neg10:postop_glp_120)
glucagon <- raw_dat %>% select(pid, preop_glucagon_neg10:preop_glucagon_240, postop_glucagon_neg10:postop_glucagon_240)
triglycerides <- raw_dat %>%  select(pid, preop_trig_0:preop_trig_120, postop_trig_0:postop_trig_120)
cpep <- raw_dat %>% select(pid, preop_cpep_neg10:preop_cpep_240, postop_cpep_neg10:postop_cpep_240)

# extract pre and post weight measurements (units = kg)
weights <- raw_dat %>% select(pid, preop_mmtt_weight, postop_mmtt_weight) %>%
  rename(c(preop = preop_mmtt_weight, postop = postop_mmtt_weight)) %>% 
  pivot_longer(cols = c(preop, postop),
               names_to = "study_visit",
               values_to = "weight")

## switch to doubly long format
######################################################################################################################
# test code for converting to doubly long (used for creating pivot_doubly_long function)
#
# glucose_long <- glucose %>%  pivot_longer(cols = starts_with("preop_glucose") | starts_with("postop_glucose"), 
#                          names_to = c("study_visit", ".value", "time"), 
#                          names_sep = "_",
#                          values_to = "glucose")
#
######################################################################################################################

# define function for converting from wide to doubly long format
# .value specifies that the assay readings should be used as values in the resulting df instead of the assay name (for example, the word "glucose")
pivot_doubly_long <- function(df, assay_name){
  df %>% pivot_longer(cols = starts_with(paste0("preop_", assay_name))| starts_with(paste0("postop_", assay_name)),
                      names_to = c("study_visit", ".value", "time"),
                      names_sep = "_",
                      values_to = assay_name)
}

# convert each df to long and add in weights
glucose_long <- pivot_doubly_long(glucose, "glucose") %>% left_join(weights, by = join_by(pid, study_visit))
insulin_long <- pivot_doubly_long(insulin, "i") %>% left_join(weights, by = join_by(pid, study_visit))
ffa_long <- pivot_doubly_long(ffa, "faa") %>% left_join(weights, by = join_by(pid, study_visit))
glp1_long <- pivot_doubly_long(glp1, "glp") %>% left_join(weights, by = join_by(pid, study_visit))
glucagon_long <- pivot_doubly_long(glucagon, "glucagon") %>% left_join(weights, by = join_by(pid, study_visit))
triglycerides_long <- pivot_doubly_long(triglycerides, "trig") %>% left_join(weights, by = join_by(pid, study_visit))
cpep_long <- pivot_doubly_long(cpep, "cpep") %>% left_join(weights, by = join_by(pid, study_visit))

## clean up variables
clean_DL_dfs <- function(df_name){
  df <- get(df_name)
  
  df <- df %>%
    mutate(time = recode(time, 'neg10' = '-10')) %>% 
    mutate(time = as.factor(as.numeric(time)),
           study_visit = factor(study_visit, levels = c('preop', 'postop')))
  
  assign(df_name, df, envir = .GlobalEnv)
}

dfs_to_clean <- c("glucose_long", "insulin_long", "ffa_long", "glp1_long", "glucagon_long", "triglycerides_long", "cpep_long")
invisible(lapply(dfs_to_clean, clean_DL_dfs))
```


```{r EDA, eval=FALSE, include=FALSE}
## spaghetti plots with mean line
spaghetti_plot <- function(df, assay_name){
  df %>% 
    ggplot(aes_string(x = 'time', y = assay_name, group = 'pid')) +
    geom_line(alpha = 0.2) +
    stat_summary(fun = mean, geom = 'line', aes(group = 1), color = 'orangered') +
    stat_summary(fun = mean, geom = 'point', aes(group = 1), color = 'orangered') +
    facet_wrap( ~study_visit) +
    theme_bw()
}

spaghetti_plot(glucose_long, 'glucose')
spaghetti_plot(insulin_long, 'i')
spaghetti_plot(ffa_long, 'faa')
spaghetti_plot(glp1_long, 'glp')
spaghetti_plot(glucagon_long, 'glucagon')
spaghetti_plot(triglycerides_long, "trig")
spaghetti_plot(cpep_long, 'cpep')

## histogram of weights
weights %>% 
  ggplot(aes(x = weight)) +
  geom_histogram(binwidth = 5) +
  labs(x = 'weight (kg)') +
  facet_wrap(~study_visit) +
  theme_bw()

## missingness exploration
glucose %>% miss_var_summary()
insulin %>% miss_var_summary()
ffa %>% miss_var_summary()
glp1 %>% miss_var_summary()
glucagon %>% miss_var_summary()
triglycerides %>% miss_var_summary()
cpep %>% miss_var_summary()
```

# Methods

Mixed effects models were used to evaluate differences in MMTT labs before and after VSG bariatric surgery, while adjusting for the correlation of repeated measures within a participant. Estimated marginal means were calculated for each timepoint of the MMTT within a visit, and pairwise comparisons were used to test whether each timepoint differed across visits. Modeling with and without adjustment for changes in weight between study visits was performed; results are separately reported in sections 2 and 3 of this document. 

Analyses for each of the MMTT labs were based on the following timepoints:

* Glucose, Insulin: -10, 0, 10, 20, 30, 45, 60, 90, 120, 150, 180, 210, 240 min
* FFA: -10, 0, 30, 60, 120, 180, 240 min
* GLP-1: -10,0, 10, 30, 45, 60, 120 min
* Glucagon: -10, 0, 10, 30, 45, 60, 120, 240 min
* Triglycerides: 0, 30, 60, 120 min
* C-Peptide: -10, 0, 10, 20, 30, 60, 90, 120, 180, 240 min

Differences in non-longitudinal labs (VSG, delta metrics, etc.) before and after surgery were evaluated using paired *t*-tests or Wilcoxon signed rank testing, depending on visual inspection of each metric's pre-op and post-op histogram distributions. Metrics analyzed with *t*-tests are reported with means and standard deviations for the pre-op and post-op timepoints. In instances where Wilcoxon signed rank testing was instead used, the associated pre-op and post-op measures are reported as median [25th percentile, 75th percentile]. These results are presented in section 4.


```{r mixed functions, include=FALSE}
## define function that fits LMM then outputs a list of the ANOVA, marginal mean, and pairwise comparisons results
# inputted dat is the doubly long df for the assay of interest
mixed <- function(dat, outcome){
  form = paste0(outcome,"~ study_visit * time")
  mod <- lme(as.formula(form), random = ~1|pid, data = dat, na.action = na.omit,
             correlation = corCompSymm(form = ~1 | pid)) # specifies within-subject variability as exchangeable
  anova <- anova.lme(mod, type="marginal") # use ANOVA to get significance of interaction
  m <- emmeans(mod,c("study_visit","time")) # compute estimated marginal means to obtain effect at each timepoint
  prs <-  pairs(m,adjust="tukey",by="time") # perform pairwise comparisons by timepoint
  
  return(list(anova, m, prs))
}

## define weight adjusted version of mixed function
mixed_wt_adj <- function(dat, outcome){
  form = paste0(outcome,"~ study_visit * time + weight")
  mod <- lme(as.formula(form), random = ~1|pid, data = dat, na.action = na.omit,
             correlation = corCompSymm(form = ~1 | pid)) 
  anova <- anova.lme(mod, type="marginal") 
  m <- emmeans(mod,c("study_visit","time")) 
  prs <-  pairs(m,adjust="tukey",by="time") 
  
  return(list(anova, m, prs))
}
```


```{r convergence check, eval=FALSE, include=FALSE}
# this chunk provides code for checking whether the models of interest are converging
# the below functions print summary output for lme and lmer models
# (lmer was additionally used here b/c we were uncertain that lme provides convergence warnings)

## define function for checking non-adjusted mixed models
mixed_check <- function(dat, outcome){
  form1 = paste0(outcome,"~ study_visit * time")
  mod1 <- lme(as.formula(form1), random = ~1|pid, data = dat, na.action = na.omit,
             correlation = corCompSymm(form = ~1 | pid))
  
  form2 <- paste0(outcome,"~ study_visit * time + (1|pid)")
  mod2 <- lmer(as.formula(form2), data = dat, na.action = na.omit)
  print(summary(mod1))
  print(summary(mod2), correlation = TRUE)
}

## check models using function
mixed_check(dat = glucose_long, outcome = "glucose")
mixed_check(dat = insulin_long, outcome = "i")
mixed_check(dat = ffa_long, outcome = "faa")
mixed_check(dat = glp1_long, outcome = "glp")
mixed_check(dat = glucagon_long, outcome = "glucagon")
mixed_check(dat = triglycerides_long, outcome = "trig")
mixed_check(dat = cpep_long, outcome = "cpep")

## define function for checking weight-adjusted mixed models
mixed_wt_adj_check <- function(dat, outcome){
  form1 = paste0(outcome,"~ study_visit * time + weight")
  mod1 <- lme(as.formula(form1), random = ~1|pid, data = dat, na.action = na.omit,
             correlation = corCompSymm(form = ~1 | pid))
  
  form2 <- paste0(outcome,"~ study_visit * time + weight + (1|pid)")
  mod2 <- lmer(as.formula(form2), data = dat, na.action = na.omit)
  print(summary(mod1))
  print(summary(mod2), correlation = TRUE)
}

## check weight-adjusted models
mixed_wt_adj_check(dat = glucose_long, outcome = "glucose")
mixed_wt_adj_check(dat = insulin_long, outcome = "i")
mixed_wt_adj_check(dat = ffa_long, outcome = "faa")
mixed_check(dat = glp1_long, outcome = "glp")
mixed_wt_adj_check(dat = glucagon_long, outcome = "glucagon")
mixed_wt_adj_check(dat = triglycerides_long, outcome = "trig")
mixed_wt_adj_check(dat = cpep_long, outcome = "cpep")
```


# Mixed Model Results
How to interpret the output:

* ANOVA Results: This table provides the parameter estimates for each term in the model. The `study_visit:time` term tests whether the pattern of change of the lab measure throughout the duration of the MMTT differed between the pre-op and post-op visits. For example, based on the insulin results, we can see that the trajectory of insulin during the MMTT did differ between visits (p < 0.05). In instances where the `study_visit:time` term has a p-value of $\ge0.05$, the estimated marginal means and pairwise comparisons tables are omitted from the report. 

* Estimated Marginal Means: This table reports the estimated marginal means for each visit and timepoint. `lower.CL` and `upper.CL` respectively describe the lower and upper bounds of the 95% confidence interval.

* Pairwise Comparisons: This table provides p-values for testing whether there is a notable difference in the MMTT measure between the pre-op and post-op visits at each timepoint. For example, looking at the first row of the glucose pairwise comparisons table, there was not a significant difference in glucose between the pre-op and post-op visits at -10 min (p=0.284). 

## Glucose
```{r glucose analysis, echo=FALSE}
glucose_results <- mixed(dat = glucose_long, outcome = "glucose")

glucose_results[[1]] %>% 
  kable(caption = 'ANOVA Results:', booktabs = TRUE, digits = 3) %>% 
  kable_styling(latex_options='basic', html_font='Arial') %>% 
      htmltools::HTML()
glucose_results[[2]] %>% 
  kable(caption = 'Estimated Marginal Means:', booktabs = TRUE, digits =3) %>% 
  kable_styling(latex_options='basic', html_font='Arial') %>% 
      htmltools::HTML()
glucose_results[[3]] %>%
  kable(caption = 'Pairwise Comparisons:', booktabs = TRUE, digits =3) %>% 
  kable_styling(latex_options='basic', html_font='Arial') %>% 
      htmltools::HTML()
```

## Insulin

```{r insulin analysis, echo=FALSE}
insulin_results <- mixed(dat = insulin_long, outcome = "i")

insulin_results[[1]] %>% 
  kable(caption = 'ANOVA Results:', booktabs = TRUE, digits = 3) %>% 
  kable_styling(latex_options='basic', html_font='Arial') %>% 
      htmltools::HTML()
insulin_results[[2]] %>% 
  kable(caption = 'Estimated Marginal Means:', booktabs = TRUE, digits =3) %>% 
  kable_styling(latex_options='basic', html_font='Arial') %>% 
      htmltools::HTML()
insulin_results[[3]] %>% 
  kable(caption = 'Pairwise Comparisons:', booktabs = TRUE, digits =3) %>% 
  kable_styling(latex_options='basic', html_font='Arial') %>% 
      htmltools::HTML()
```

## FFA

```{r ffa analysis, echo=FALSE}
ffa_results <- mixed(dat = ffa_long, outcome = "faa")

ffa_results[[1]] %>% 
  kable(caption = 'ANOVA Results:', booktabs = TRUE, digits = 3) %>% 
  kable_styling(latex_options='basic', html_font='Arial') %>% 
      htmltools::HTML()
ffa_results[[2]] %>% 
  kable(caption = 'Estimated Marginal Means:', booktabs = TRUE, digits =3) %>% 
  kable_styling(latex_options='basic', html_font='Arial') %>% 
      htmltools::HTML()
ffa_results[[3]] %>% 
  kable(caption = 'Pairwise Comparisons:', booktabs = TRUE, digits =3) %>% 
  kable_styling(latex_options='basic', html_font='Arial') %>% 
      htmltools::HTML()
```

## GLP-1

```{r glp1 analysis, echo=FALSE}
glp1_results <- mixed(dat = glp1_long, outcome = "glp")

glp1_results[[1]] %>% 
  kable(caption = 'ANOVA Results:', booktabs = TRUE, digits = 3) %>% 
  kable_styling(latex_options='basic', html_font='Arial') %>% 
      htmltools::HTML()
glp1_results[[2]] %>% 
  kable(caption = 'Estimated Marginal Means:', booktabs = TRUE, digits =3) %>% 
  kable_styling(latex_options='basic', html_font='Arial') %>% 
      htmltools::HTML()
glp1_results[[3]] %>% 
  kable(caption = 'Pairwise Comparisons:', booktabs = TRUE, digits =3) %>% 
  kable_styling(latex_options='basic', html_font='Arial') %>% 
      htmltools::HTML()
```

## Glucagon
<span style="color: blue;">Estimated marginal means and pairwise comparisons were not computed due to non-significant findings.</span>
```{r glucagon analysis, echo=FALSE}
glucagon_results <- mixed(dat = glucagon_long, outcome = "glucagon")

glucagon_results[[1]] %>% 
  kable(caption = 'ANOVA Results:', booktabs = TRUE, digits = 3) %>% 
  kable_styling(latex_options='basic', html_font='Arial') %>% 
      htmltools::HTML()
# glucagon_results[[2]] %>%
#   kable(caption = 'Estimated Marginal Means:', booktabs = TRUE, digits =3) %>%
#   kable_styling(latex_options='basic', html_font='Arial') %>%
#       htmltools::HTML()
# glucagon_results[[3]] %>%
#   kable(caption = 'Pairwise Comparisons:', booktabs = TRUE, digits =3) %>%
#   kable_styling(latex_options='basic', html_font='Arial') %>%
#       htmltools::HTML()
```

## Triglycerides
<span style="color: blue;">Estimated marginal means and pairwise comparisons were not computed due to non-significant findings.</span>
```{r triglycerides analysis, echo=FALSE}
triglycerides_results <- mixed(dat = triglycerides_long, outcome = "trig")

triglycerides_results[[1]] %>% 
  kable(caption = 'ANOVA Results:', booktabs = TRUE, digits = 3) %>% 
  kable_styling(latex_options='basic', html_font='Arial') %>% 
      htmltools::HTML()
# triglycerides_results[[2]] %>% 
#   kable(caption = 'Estimated Marginal Means:', booktabs = TRUE, digits =3) %>% 
#   kable_styling(latex_options='basic', html_font='Arial') %>% 
#       htmltools::HTML()
# triglycerides_results[[3]] %>% 
#   kable(caption = 'Pairwise Comparisons:', booktabs = TRUE, digits =3) %>% 
#   kable_styling(latex_options='basic', html_font='Arial') %>% 
#       htmltools::HTML()
```

## C-Peptide
<span style="color: blue;">Estimated marginal means and pairwise comparisons were not computed due to non-significant findings.</span>
```{r cpep analysis, echo=FALSE}
cpep_results <- mixed(dat = cpep_long, outcome = "cpep")

cpep_results[[1]] %>% 
  kable(caption = 'ANOVA Results:', booktabs = TRUE, digits = 3) %>% 
  kable_styling(latex_options='basic', html_font='Arial') %>% 
      htmltools::HTML()
# cpep_results[[2]] %>% 
#   kable(caption = 'Estimated Marginal Means:', booktabs = TRUE, digits =3) %>% 
#   kable_styling(latex_options='basic', html_font='Arial') %>% 
#       htmltools::HTML()
# cpep_results[[3]] %>% 
#   kable(caption = 'Pairwise Comparisons:', booktabs = TRUE, digits =3) %>% 
#   kable_styling(latex_options='basic', html_font='Arial') %>% 
#       htmltools::HTML()
```

# Weight-Adjusted Mixed Model Results

How to interpret the output:

Output for these models can be interpreted similarly to the regular non-adjusted mixed model results with the additional consideration that all results are now adjusted for weight. 

## Glucose
```{r glucose wt adj analysis, echo=FALSE}
glucose_results <- mixed_wt_adj(dat = glucose_long, outcome = "glucose")

glucose_results[[1]] %>% 
  kable(caption = 'ANOVA Results:', booktabs = TRUE, digits = 3) %>% 
  kable_styling(latex_options='basic', html_font='Arial') %>% 
      htmltools::HTML()
glucose_results[[2]] %>% 
  kable(caption = 'Estimated Marginal Means:', booktabs = TRUE, digits =3) %>% 
  kable_styling(latex_options='basic', html_font='Arial') %>% 
      htmltools::HTML()
glucose_results[[3]] %>%
  kable(caption = 'Pairwise Comparisons:', booktabs = TRUE, digits =3) %>% 
  kable_styling(latex_options='basic', html_font='Arial') %>% 
      htmltools::HTML()
```

## Insulin

```{r insulin wt adj analysis, echo=FALSE}
insulin_results <- mixed_wt_adj(dat = insulin_long, outcome = "i")

insulin_results[[1]] %>% 
  kable(caption = 'ANOVA Results:', booktabs = TRUE, digits = 3) %>% 
  kable_styling(latex_options='basic', html_font='Arial') %>% 
      htmltools::HTML()
insulin_results[[2]] %>% 
  kable(caption = 'Estimated Marginal Means:', booktabs = TRUE, digits =3) %>% 
  kable_styling(latex_options='basic', html_font='Arial') %>% 
      htmltools::HTML()
insulin_results[[3]] %>% 
  kable(caption = 'Pairwise Comparisons:', booktabs = TRUE, digits =3) %>% 
  kable_styling(latex_options='basic', html_font='Arial') %>% 
      htmltools::HTML()
```

## FFA

```{r ffa wt adj analysis, echo=FALSE}
ffa_results <- mixed_wt_adj(dat = ffa_long, outcome = "faa")

ffa_results[[1]] %>% 
  kable(caption = 'ANOVA Results:', booktabs = TRUE, digits = 3) %>% 
  kable_styling(latex_options='basic', html_font='Arial') %>% 
      htmltools::HTML()
ffa_results[[2]] %>% 
  kable(caption = 'Estimated Marginal Means:', booktabs = TRUE, digits =3) %>% 
  kable_styling(latex_options='basic', html_font='Arial') %>% 
      htmltools::HTML()
ffa_results[[3]] %>% 
  kable(caption = 'Pairwise Comparisons:', booktabs = TRUE, digits =3) %>% 
  kable_styling(latex_options='basic', html_font='Arial') %>% 
      htmltools::HTML()
```

## GLP-1

```{r glp1 wt adj analysis, echo=FALSE}
glp1_results <- mixed_wt_adj(dat = glp1_long, outcome = "glp")

glp1_results[[1]] %>% 
  kable(caption = 'ANOVA Results:', booktabs = TRUE, digits = 3) %>% 
  kable_styling(latex_options='basic', html_font='Arial') %>% 
      htmltools::HTML()
glp1_results[[2]] %>% 
  kable(caption = 'Estimated Marginal Means:', booktabs = TRUE, digits =3) %>% 
  kable_styling(latex_options='basic', html_font='Arial') %>% 
      htmltools::HTML()
glp1_results[[3]] %>% 
  kable(caption = 'Pairwise Comparisons:', booktabs = TRUE, digits =3) %>% 
  kable_styling(latex_options='basic', html_font='Arial') %>% 
      htmltools::HTML()
```

## Glucagon
<span style="color: blue;">Estimated marginal means and pairwise comparisons were not computed due to non-significant findings.</span>
```{r glucagon wt adj analysis, echo=FALSE}
glucagon_results <- mixed_wt_adj(dat = glucagon_long, outcome = "glucagon")

glucagon_results[[1]] %>% 
  kable(caption = 'ANOVA Results:', booktabs = TRUE, digits = 3) %>% 
  kable_styling(latex_options='basic', html_font='Arial') %>% 
      htmltools::HTML()
# glucagon_results[[2]] %>%
#   kable(caption = 'Estimated Marginal Means:', booktabs = TRUE, digits =3) %>%
#   kable_styling(latex_options='basic', html_font='Arial') %>%
#       htmltools::HTML()
# glucagon_results[[3]] %>%
#   kable(caption = 'Pairwise Comparisons:', booktabs = TRUE, digits =3) %>%
#   kable_styling(latex_options='basic', html_font='Arial') %>%
#       htmltools::HTML()
```

## Triglycerides
<span style="color: blue;">Estimated marginal means and pairwise comparisons were not computed due to non-significant findings.</span>
```{r triglycerides wt adj analysis, echo=FALSE}
triglycerides_results <- mixed_wt_adj(dat = triglycerides_long, outcome = "trig")

triglycerides_results[[1]] %>% 
  kable(caption = 'ANOVA Results:', booktabs = TRUE, digits = 3) %>% 
  kable_styling(latex_options='basic', html_font='Arial') %>% 
      htmltools::HTML()
# triglycerides_results[[2]] %>% 
#   kable(caption = 'Estimated Marginal Means:', booktabs = TRUE, digits =3) %>% 
#   kable_styling(latex_options='basic', html_font='Arial') %>% 
#       htmltools::HTML()
# triglycerides_results[[3]] %>% 
#   kable(caption = 'Pairwise Comparisons:', booktabs = TRUE, digits =3) %>% 
#   kable_styling(latex_options='basic', html_font='Arial') %>% 
#       htmltools::HTML()
```

## C-Peptide
<span style="color: blue;">Estimated marginal means and pairwise comparisons were not computed due to non-significant findings.</span>
```{r cpep wt adj analysis, echo=FALSE}
cpep_results <- mixed_wt_adj(dat = cpep_long, outcome = "cpep")

cpep_results[[1]] %>% 
  kable(caption = 'ANOVA Results:', booktabs = TRUE, digits = 3) %>% 
  kable_styling(latex_options='basic', html_font='Arial') %>% 
      htmltools::HTML()
# cpep_results[[2]] %>% 
#   kable(caption = 'Estimated Marginal Means:', booktabs = TRUE, digits =3) %>% 
#   kable_styling(latex_options='basic', html_font='Arial') %>% 
#       htmltools::HTML()
# cpep_results[[3]] %>% 
#   kable(caption = 'Pairwise Comparisons:', booktabs = TRUE, digits =3) %>% 
#   kable_styling(latex_options='basic', html_font='Arial') %>% 
#       htmltools::HTML()
```


# Pre-op and Post-op Comparisons

```{r pre/post comparisons data cleaning, echo=FALSE, message=FALSE, warning=FALSE}
## create df with metrics of interest for pre/post comparison
clean_metrics <- raw_dat %>%
  # compute fasting and peak metrics
  mutate(preop_fasting_glucose = rowMeans(select(., preop_glucose_neg10, preop_glucose_0), na.rm = TRUE),
         postop_fasting_glucose = rowMeans(select(., postop_glucose_neg10, postop_glucose_0), na.rm = TRUE),
         preop_peak_glucose = do.call(pmax, c(select(., preop_glucose_neg10: preop_glucose_240), na.rm = TRUE)),
         postop_peak_glucose = do.call(pmax, c(select(., postop_glucose_neg10:postop_glucose_240), na.rm = TRUE)),
         preop_peak_insulin = do.call(pmax, c(select(., preop_i_neg10:preop_i_240), na.rm = TRUE)),
         postop_peak_insulin = do.call(pmax, c(select(., postop_i_neg10:postop_i_240), na.rm = TRUE))) %>%
  select(c(pid, preop_mmtt_bmi, postop_mmtt_bmi, preop_mmtt_w_h, postop_mmtt_w_h, 
           preop_a1c, postop_a1c, preop_alt, postop_alt, preop_ast, postop_ast, preop_ggt, postop_ggt, 
           preop_mmtt_chol, postop_mmtt_chol, preop_mmtt_hdl, postop_mmtt_hdl, preop_mmtt_tg, postop_mmtt_tg, 
           preop_mmtt_calc_ldl, postop_mmtt_calc_ldl, preop_mmtt_adiponectin, postop_mmtt_adiponectin, 
           preop_homa, postop_homa, preop_fasting_glucose, postop_fasting_glucose, preop_peak_glucose, postop_peak_glucose,
           preop_peak_insulin, postop_peak_insulin, preop_visceralfat_percent, postop_visceralfat_percent, preop_liver_fat_perc,
           postop_liver_fat_perc, preop_bodpod_fat, postop_bodpod_fat, preop_bodpod_fatfreemass, postop_bodpod_fatfreemass,
           preop_basal_fat_oxidation_mmol_min_kg, postop_basal_fat_oxidation_mmol_min_kg, preop_stage1_fat_oxidation_mmol_min_kg,
           postop_stage1_fat_oxidation_mmol_min_kg, preop_basal_carb_oxidation_mmol_min_kg,
           postop_basal_carb_oxidation_mmol_min_kg, preop_stage1_carb_oxidation_mmol_min_kg,
           postop_stage1_carb_oxidation_mmol_min_kg, basal_rq, postop_basal_rq, stage1_rq, postop_stage1_rq,
           preop_mets_3dpar, preop_mets_3dpar_v2, preop_kcals_day, preop_kcals_day_v2, preop_kcals_total,
           preop_kcals_total_v2, preop_steps_total, preop_steps_total_v2, ffq_fat_g, ffq_fat_g_postop,
           ffq_kcal, ffq_kcal_postop, ffq_pro_g, ffq_pro_g_postop, ffq_carbo_g, ffq_carbo_g_postop, ffq_totsugar_g,
           ffq_totsugar_g_postop, ffq_fructose_g, ffq_fructose_g_postop)) %>%
  # update variable names as needed to contain "pre" or "post" wording so that we can later create a study_visit column
  rename(c(preop_basal_rq = basal_rq, preop_stage1_rq = stage1_rq, preop_ffq_fat_g = ffq_fat_g,
           preop_ffq_kcal = ffq_kcal, preop_ffq_pro_g = ffq_pro_g, preop_ffq_carbo_g = ffq_carbo_g,
           preop_ffq_totsugar_g = ffq_totsugar_g, preop_ffq_fructose_g = ffq_fructose_g,
           postop_mets_3dpar = preop_mets_3dpar_v2, 
           postop_kcals_day = preop_kcals_day_v2, postop_kcals_total = preop_kcals_total_v2, 
           postop_steps_total = preop_steps_total_v2, postop_ffq_fat_g = ffq_fat_g_postop,
           postop_ffq_kcal = ffq_kcal_postop, postop_ffq_pro_g = ffq_pro_g_postop,
           postop_ffq_carbo_g = ffq_carbo_g_postop, postop_ffq_totsugar_g = ffq_totsugar_g_postop,
           postop_ffq_fructose_g = ffq_fructose_g_postop)) %>% 
  # remove percentages from visceral fat cols
  mutate(preop_visceralfat_percent = str_remove(preop_visceralfat_percent, "%") %>% as.numeric(),
         postop_visceralfat_percent = str_remove(postop_visceralfat_percent, "%") %>% as.numeric())
  
## format df to be in long format with each subject containing separate lines of data for pre and post
clean_metrics <- clean_metrics %>% 
  pivot_longer(cols = starts_with("preop_") | starts_with("postop_"),
               names_to = "combined",
               values_to = "value") %>% 
  separate(combined, into = c("study_visit", "metric1", "metric2", "metric3", "metric4", "metric5", "metric6"))
clean_metrics$metric <- apply(clean_metrics[, c("metric1", "metric2", "metric3", "metric4", "metric5", "metric6")], 
                     1, function(x) paste(na.omit(x), collapse = "_"))
clean_metrics <- clean_metrics %>%
  select(-c(metric1, metric2, metric3, metric4, metric5, metric6)) %>% 
  pivot_wider(names_from = metric, values_from = value)
```


```{r normality check, eval=FALSE, include=FALSE}
## check distribution of metrics for deciding btwn. parametric and non-parametric testing 
var_labels <- sjlabelled::get_label(clean_metrics[,-1])
ncolumn <- ncol(clean_metrics)
ncolumn <- ncolumn - 1

for (col in 2:ncolumn){
  var_name <- colnames(clean_metrics[,-1])[col]
  plot <- ggplot(clean_metrics, aes(x = .data[[var_name]])) +
    geom_histogram() +
    labs(title = var_labels[var_name]) +
    theme_bw() +
    facet_wrap(~ study_visit) 
  print(plot)
}
```

## t-test Results

How to interpret the output:

The below table reports the mean and standard deviation for each metric at pre-op and post-op. The corresponding p-values test whether there is a notable difference in means between the two timepoints.

```{r t-tests, echo=FALSE, message=FALSE, warning=FALSE}
clean_metrics %>%
  mutate(study_visit = factor(study_visit, levels = c('preop', 'postop'))) %>% 
  select(c(pid, study_visit, mmtt_adiponectin, stage1_carb_oxidation_mmol_min_kg, basal_rq)) %>% 
  tbl_summary(by = study_visit, include = -pid,
              type = list(c(mmtt_adiponectin,
                            stage1_carb_oxidation_mmol_min_kg, basal_rq) ~ 'continuous',
                          study_visit ~ 'dichotomous'),
              statistic = list(all_continuous() ~ '{mean} ({sd})'),
              digits = everything() ~ c(2,2),
              missing_text = 'N Missing') %>% 
  add_p(test = all_continuous() ~ 'paired.t.test', group = pid,
        pvalue_fun = ~ style_pvalue(.x, digits = 3)) %>% 
  modify_header(label = '**Metric**',
                stat_1 = '**Pre-op** (N={n})',
                stat_2 = '**Post-op** (N={n})') %>%
  modify_caption("t-test Results")
```

## Wilcoxon Signed Rank Results

How to interpret the output:

The table below reports the median, 25% percentile, and 75% percentile values for each metric at pre-op and post-op. Included are p-values for testing whether there is a notable difference in mean ranks between the two timepoints. In instances where p-values are not reported (ggt, mmtt_chol, fasting_glucose, stage1_rq, a1c), the models did not converge.

```{r WRS testing, echo=FALSE, message=FALSE, warning=FALSE}
clean_metrics %>%
  mutate(study_visit = factor(study_visit, levels = c('preop', 'postop'))) %>% 
  select(-c(mmtt_adiponectin, stage1_carb_oxidation_mmol_min_kg, basal_rq)) %>% 
  tbl_summary(by = study_visit, include = -pid,
              type = list(c(mmtt_bmi, mmtt_w_h, a1c, alt, ast, ggt, 
                            mmtt_chol, mmtt_hdl, mmtt_tg, mmtt_calc_ldl, homa, fasting_glucose, peak_glucose,
                            peak_insulin, visceralfat_percent, liver_fat_perc, bodpod_fat, 
                            bodpod_fatfreemass, basal_fat_oxidation_mmol_min_kg, stage1_fat_oxidation_mmol_min_kg,
                            basal_carb_oxidation_mmol_min_kg, stage1_rq, mets_3dpar, kcals_day, kcals_total, 
                            steps_total, ffq_fat_g, ffq_kcal, ffq_pro_g, ffq_carbo_g, 
                            ffq_totsugar_g, ffq_fructose_g) ~ 'continuous',
                          study_visit ~ 'dichotomous'),
              statistic = list(all_continuous() ~ '{median} [25% Pi: {p25}, 75% Pi: {p75}]'),
              digits = everything() ~ c(2,2),
              missing_text = 'N Missing') %>% 
  add_p(test = all_continuous() ~ 'paired.wilcox.test',
        group = pid,
        pvalue_fun = ~ style_pvalue(.x, digits = 3)) %>% 
  modify_header(label = '**Metric**',
                stat_1 = '**Pre-op** (N={n})',
                stat_2 = '**Post-op** (N={n})') %>%
  modify_caption("Wilcoxon Signed Rank Results") %>%
  modify_table_body(
    ~ .x %>% 
      mutate(p.value = ifelse(variable %in% c('ggt', 'mmtt_chol', 'fasting_glucose', 'stage1_rq', 'a1c'), NA, p.value))
  )

```

Pi = percentile