---
title: "Top transcripts associated with PGLO"
author: "Ye Ji Choi"
date: "today"
date-format: long
execute: 
  echo: false
  warning: false
  messages: false
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
editor: source
---

```{r libraries}
#| include: false
library(Seurat)
library(future)
library(tidyverse)
library(colorspace)
library(patchwork)
library(ggdendro)
library(cowplot)
library(ggpubr)
library(rstatix)
library(table1)
library(Biobase)
library(ReactomeGSA)
library(GSEABase)
library(msigdbr)
library(kableExtra)
library(knitr)
library(pushoverr)
set_pushover_user(user = "uv4fkwcjfp88sprou64rboojje4xvm")
set_pushover_app(token = "avi88qsjeunh4xmvz3mbgp2fzpmkey")

```

```{r scRNA data setup}
#| include: false
# Parallel processing
plan(multicore, workers = 16)
options(future.globals.maxSize=6e9)
# Import
so <- readRDS("/home/yejichoi2/Documents/seurat_data_compiled.RDS")
pushover(message = "done importing so")
```

```{r include = F}
harm_dat <- read.csv("/run/user/1023/gvfs/smb-share:server=ucdenver.pvt,share=som/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/Data Harmonization/Data Clean/harmonized_dataset.csv", na.strings = "")

dat <- harm_dat %>%
  dplyr::summarise(across(where(negate(is.numeric)), ~ ifelse(all(is.na(.x)), NA_character_, last(na.omit(.x)))),
                   across(where(is.numeric), ~ ifelse(all(is.na(.x)), NA_real_, mean(.x, na.rm = TRUE))),
                   .by = c(mrn, screen_date)) %>%
  # filter(participation_status!="Removed"|is.na(participation_status)) %>%
  dplyr::mutate(race_ethnicity_condensed = case_when(race == "White" & 
                                                ethnicity == "Not Hispanic or Latino" ~ "Not Hispanic or Latino White",
                                              race == "Black or African American" & 
                                                ethnicity == "Not Hispanic or Latino" ~ "Not Hispanic or Latino Black",
                                              ethnicity == "Hispanic or Latino" ~ "Hispanic or Latino",
                                              T ~ "Not Hispanic or Latino Other"),
         combined_tkv = coalesce(total_kidney_volume_ml, total_kidney_volume_ml_manual),
         combined_ht_tkv = coalesce(ht_adj_tkv, ht_adj_tkv_manual)) %>%
  arrange(record_id)
```

# Transcript association with PGLO

* dPT (degenerative proximal tubule), 
* dPT/DTL (degenerative proximal tubule/descending thin limb), 
* EC-AEA (afferent/efferent arterioles), 
* EC-AVR (ascending vasa recta), 
* EC-DVR (descending vasa recta), 
* EC-GC (glomerular endothelial cells), 
* MC (mesangial cells), 
* PT-S1/S2 (segments S1 and S2 of the proximal tubule, which are relevant for studying SGLT2), 
* PT-S3 (segment S3 of the proximal tubule, relevant for studying SGLT1), 
* VSMC (vascular smooth muscle cells), and 
* VSMC/P (vascular smooth muscle cells containing pericytes, which are also crucial for the vasa recta)

```{r}

```
