---
title: "TODAY Somalogic baseline analyses - adjusted Cox models"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include=FALSE}
library(SomaDataIO)
library(limma)
library(dplyr)
library(caret)
library(purrr)
library(multtest)
library(openxlsx)
library(tableone)
library(EnhancedVolcano)
library(knitr)
library(survival)
library(broom)
library(emmeans)
library(ggvenn)

knitr::opts_chunk$set(echo = FALSE,warning = FALSE)

knitr::opts_chunk$set(echo = FALSE)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "E:/Petter Bjornstad/TODAY subaward"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/TODAY subaward/"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/TODAY subaward"
}
knitr::opts_knit$set(root.dir = home_dir)
```

```{r, include=FALSE}
setwd(home_dir)
# load somalogic data, with QC samples already excluded
load("./Somalogic data raw/soma.Rdata")

# load analyte info
load("./Somalogic data raw/analytes.Rdata")

# load comorbidity data
load("./Clinical data/comorb.Rdata")

# load baseline risk factors
load("./Clinical data/TODAY/baserisk.Rdata")

# take only the baseline soma samples
# can we just take the earliest or is it possible someone would have a follow-up sample but not baseline?
# probably can take the first and then check the years to make sure compatible with TODAY
base <- soma %>% arrange(releaseid,Date.Drawn) %>% dplyr::group_by(releaseid) %>% dplyr::filter(row_number()==1)
# these 3 release IDs have first sample after end of recruitment, so they must be missing the baseline visit
base <- base %>% filter(!releaseid %in% c("65-85903","65-47984","65-25901"))

# the above section of code used to work but seems to have broken....this should accomplish the same thing?
#base <- soma %>% filter(visit==1)

# merge in complication data
base <- left_join(base, comorb, by="releaseid")
# this was previously:
# base <- merge(base, comorb, by="releaseid",all.x=T, all.y=F)

# merge in baseline risk factors
base <- left_join(base, baserisk, by="releaseid")
# this was previously:
#base <- merge(base, baserisk, by="releaseid",all.x=T, all.y=F)

# log transform UAlbCreat
base$log_UAlbCreat <- log(base$UAlbCreat + 0.0000001)

# identify columns corresponding to proteins
#is_seq <- function(.x) grepl("^seq\\.[0-9]{4}", .x) # regex for analytes
is_seq <- function(.x) grepl("seq", .x)
seq <- is_seq(names(base))

# convert to numeric
base[,seq] <- apply(base[,seq],2,as.numeric)

# are there proteins with low variability?
no_var = caret::nearZeroVar(base[,seq])
# none

# log transform
base_log <- base %>% modify_if(is_seq(names(.)), log)

# scale by SD
base_log_scale <- base_log
predictors <- colnames(base_log[seq])
for (i in 1:length(predictors)) {
  base_log_scale[,paste0(predictors[i])] <- base_log_scale[,paste0(predictors[i])]/sd(unlist(base_log[,paste0(predictors[i])]))
}

# df for descriptive statistics
t1data <- base
t1data$MIC.OR.MAC <- as.factor(t1data$MIC.OR.MAC)
t1data$MAC <- as.factor(t1data$MAC)
t1data$MAC0 <- as.factor(t1data$MAC0)
t1data$MIC <- as.factor(t1data$MIC)
t1data$MIC0 <- as.factor(t1data$MIC0)
t1data$HYP <- as.factor(t1data$HYP)
t1data$HYP0 <- as.factor(t1data$HYP0)
t1data$RAPID <- as.factor(t1data$RAPID)
t1data$RAPID0 <- as.factor(t1data$RAPID0)
t1data$NEURO <- as.factor(t1data$NEURO)
t1data$RETINO <- as.factor(t1data$RETINO)
t1data$GLYC <- as.factor(t1data$GLYC)
t1data$HTN <- as.factor(t1data$HTN)
t1data$sex <- as.factor(t1data$sex)
t1data$tg <- exp(t1data$log_trig)
t1data$tx <- as.factor(t1data$tx)
t1data$houseinc <- as.factor(t1data$houseinc)
t1data$race <- as.factor(t1data$race)
t1data$group[t1data$MIC==1 & t1data$MAC==0]<- "MIC"
t1data$group[t1data$MAC==1]<- "MAC"
t1data$group[t1data$MAC==0 & t1data$MIC==0]<- "None"
t1data$UAlbCreat <- t1data$UAlbCreat*1000
t1data$ge25yrs_baseline <- ifelse(t1data$AGEBASE >= 25, 1, 0)
# read in age at last visit
load("./Clinical data/TODAY/clinical_data_long.Rdata")
long <- long %>% group_by(releaseid) %>% filter(row_number()==1) %>% select(releaseid, fup_years, age_last_visit, ge25yrs)
t1data <- left_join(t1data, long, by = "releaseid")
t1data$ge26yrs <- ifelse(t1data$age_last_visit >= 26, 1, 0)
t1data$HTN0 <- as.factor(t1data$HTN0)
```

## Outcomes by sex

```{r echo=FALSE, include=FALSE}
t1vars <- c("MIC0","MAC0","HYP0","RAPID0","HTN0","MIC.OR.MAC","MAC","MIC","HYP","RAPID","HTN","NEURO","RETINO","GLYC",
                                          "HbA1c","trig","sbp","dbp","map","si_1_ins0","glu0min","codi","ins0min",
                                          "AGEBASE","sex","tg","dxtime",
                                          "bmi","txdesc","houseincdesc","racedesc","UAlbCreat","ge25yrs", "ge26yrs",
                                          "SerumCreat","serumcystc","EstCreatClear")

t_by_sex <- CreateTableOne(data=t1data, vars=t1vars, strata = "sex")
t_by_sex <- print(t_by_sex, nonnormal=c("UAlbCreat","dxtime","tg","HbA1c"))
```

```{r echo=FALSE, include=TRUE}
kable(t_by_sex)
```

## Interactions with sex

### Severe albuminuria and NELL1

```{r echo=FALSE, include=TRUE, comment=""}
base_log_scale$sex <- as.factor(base_log_scale$sex)
# NELL1
mod <- coxph(data = base_log_scale, formula = Surv(DAYSTOMAC, MAC)~ HbA1c+log_trig+sbp+si_1_ins0 + seq.6544.33*sex)
tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
tidymod <- tidymod %>% mutate_if(is.double, round, 4)
kable(tidymod)
```

### Severe albuminuria and CI061

```{r echo=FALSE, include=TRUE, comment=""}
# CI061
mod <- coxph(data = base_log_scale, formula = Surv(DAYSTOMAC, MAC)~ HbA1c+log_trig+sbp+si_1_ins0 + seq.5699.19*sex)
tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
tidymod <- tidymod %>% mutate_if(is.double, round, 4)
kable(tidymod)
```

### Severe albuminuria and PSB3

```{r echo=FALSE, include=TRUE, comment=""}
# PSB3
mod <- coxph(data = base_log_scale, formula = Surv(DAYSTOMAC, MAC)~ HbA1c+log_trig+sbp+si_1_ins0 + seq.18339.207*sex)
tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
tidymod <- tidymod %>% mutate_if(is.double, round, 4)
kable(tidymod)
```

### Moderate albuminuria and NELL1

```{r echo=FALSE, include=TRUE, comment=""}
# NELL1
mod <- coxph(data = base_log_scale, formula = Surv(DAYSTOMIC, MIC)~ HbA1c+log_trig+sbp+si_1_ins0 + seq.6544.33*sex)
tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
tidymod <- tidymod %>% mutate_if(is.double, round, 4)
kable(tidymod)
```

### Moderate albuminuria and CI061

```{r echo=FALSE, include=TRUE, comment=""}
# CI061
mod <- coxph(data = base_log_scale, formula = Surv(DAYSTOMIC, MIC)~ HbA1c+log_trig+sbp+si_1_ins0 + seq.5699.19*sex)
tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
tidymod <- tidymod %>% mutate_if(is.double, round, 4)
kable(tidymod)
```

### Moderate albuminuria and CTHR1

```{r echo=FALSE, include=TRUE, comment=""}
# CTHR1
mod <- coxph(data = base_log_scale, formula = Surv(DAYSTOMIC, MIC)~ HbA1c+log_trig+sbp+si_1_ins0 + seq.15467.10*sex)
tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
tidymod <- tidymod %>% mutate_if(is.double, round, 4)
kable(tidymod)
# p = 0.06!
```

### Rapid eGFR decline and ZPLD1

The interaction between sex and ZPLD1 was significant. In the model results below, the estimate for seq.5590.11 is the hazard ratio for sex = 1 (i.e., females).

```{r echo=FALSE, include=TRUE, comment=""}
# ZPLD1
mod <- coxph(data = base_log_scale, formula = Surv(DAYSTORAPID, RAPID)~ HbA1c+log_trig+sbp+si_1_ins0 + seq.5590.11*sex)
tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
tidymod <- tidymod %>% mutate_if(is.double, round, 4)
kable(tidymod)
# p = 0.03
```

In the model results below, the estimate for seq.5590.11 is the hazard ratio for sex = 2 (i.e., males).

```{r echo=FALSE, include=TRUE, comment=""}
base_log_scale$sex <- as.factor(base_log_scale$sex)
base_log_scale$sex <- relevel(base_log_scale$sex, ref="2")
# ZPLD1
mod <- coxph(data = base_log_scale, formula = Surv(DAYSTORAPID, RAPID)~ HbA1c+log_trig+sbp+si_1_ins0 + seq.5590.11*sex)
tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
tidymod <- tidymod %>% mutate_if(is.double, round, 4)
kable(tidymod)
# p = 0.03
base_log_scale$sex <- relevel(base_log_scale$sex, ref="1")
```

### Rapid eGFR decline and RELM-beta

```{r echo=FALSE, include=TRUE, comment=""}
# RELM-beta
mod <- coxph(data = base_log_scale, formula = Surv(DAYSTORAPID, RAPID)~ HbA1c+log_trig+sbp+si_1_ins0 + seq.22007.1*sex)
tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
tidymod <- tidymod %>% mutate_if(is.double, round, 4)
kable(tidymod)
```

### Rapid eGFR decline and CLU

```{r echo=FALSE, include=TRUE, comment=""}
# CLU
mod <- coxph(data = base_log_scale, formula = Surv(DAYSTORAPID, RAPID)~ HbA1c+log_trig+sbp+si_1_ins0 + seq.24941.14*sex)
tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
tidymod <- tidymod %>% mutate_if(is.double, round, 4)
kable(tidymod)
```

### Hyperfiltration and Cystatin C

```{r echo=FALSE, include=TRUE, comment=""}
# Cystatin C
mod <- coxph(data = base_log_scale, formula = Surv(DAYSTOHYP, HYP)~ HbA1c+log_trig+sbp+si_1_ins0 + seq.2609.59*sex)
tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
tidymod <- tidymod %>% mutate_if(is.double, round, 4)
kable(tidymod)
```

### Hyperfiltration and PEDF, aptamer 1

```{r echo=FALSE, include=TRUE, comment=""}
# PEDF
mod <- coxph(data = base_log_scale, formula = Surv(DAYSTOHYP, HYP)~ HbA1c+log_trig+sbp+si_1_ins0 + seq.9211.19*sex)
tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
tidymod <- tidymod %>% mutate_if(is.double, round, 4)
kable(tidymod)
```

### Hyperfiltration and PEDF, aptamer 2

```{r echo=FALSE, include=TRUE, comment=""}
# PEDF
mod <- coxph(data = base_log_scale, formula = Surv(DAYSTOHYP, HYP)~ HbA1c+log_trig+sbp+si_1_ins0 + seq.7735.17*sex)
tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
tidymod <- tidymod %>% mutate_if(is.double, round, 4)
kable(tidymod)
# p = 0.06
```

### Hyperfiltration and b2-microglobulin

```{r echo=FALSE, include=TRUE, comment=""}
# b2-Microglobulin
mod <- coxph(data = base_log_scale, formula = Surv(DAYSTOHYP, HYP)~ HbA1c+log_trig+sbp+si_1_ins0 + seq.3485.28*sex)
tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
tidymod <- tidymod %>% mutate_if(is.double, round, 4)
kable(tidymod)
```

### Hypertension and WFKN2

```{r echo=FALSE, include=TRUE, comment=""}
# WFKN2
mod <- coxph(data = base_log_scale, formula = Surv(DAYSTOHTN, HTN)~ HbA1c+log_trig+sbp+si_1_ins0 + seq.13408.23*sex)
tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
tidymod <- tidymod %>% mutate_if(is.double, round, 4)
kable(tidymod)
```

### Hypertension and SEZ6L

```{r echo=FALSE, include=TRUE, comment=""}
# SEZ6L
mod <- coxph(data = base_log_scale, formula = Surv(DAYSTOHTN, HTN)~ HbA1c+log_trig+sbp+si_1_ins0 + seq.19563.3*sex)
tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
tidymod <- tidymod %>% mutate_if(is.double, round, 4)
kable(tidymod)
```

### Hypertension and SCG3

```{r echo=FALSE, include=TRUE, comment=""}
# SCG3
mod <- coxph(data = base_log_scale, formula = Surv(DAYSTOHTN, HTN)~ HbA1c+log_trig+sbp+si_1_ins0 + seq.7957.2*sex)
tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
tidymod <- tidymod %>% mutate_if(is.double, round, 4)
kable(tidymod)
```
