---
title: "Baby study screener analysis"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include=FALSE}
library(caret)

knitr::opts_chunk$set(echo = FALSE,warning = FALSE)

#home_dir = ifelse(.Platform$OS.type != "unix","T:\\",
#                  "/Volumes/som/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects")
#knitr::opts_knit$set(root.dir = home_dir)

# read in data
data6 <- read.csv("H:\\Tartaglia\\Screener analysis\\Raw data\\Copy of 6mV_ASQ_Bayley_2.csv")
data12 <- read.csv("H:\\Tartaglia\\Screener analysis\\Raw data\\Copy of 12mV_ASQ_Bayley_2.csv")

# rename variables
data6$group <- data6$Group..XXY..1...XYY..2...XXX..3...XXYY..4.
data6$Group..XXY..1...XYY..2...XXX..3...XXYY..4. <- NULL
data12$group <- data12$Group..XXY..1...XYY..2...XXX..3...XXYY..4.
data12$Group..XXY..1...XYY..2...XXX..3...XXYY..4. <- NULL
data6$ASQ_before_Bayley <- data6$X.Y.1.N.0..ASQ.Before.Bayley
data12$ASQ_before_Bayley <- data12$X.Y.1.N.0..ASQ.Before.Bayley
data6$X.Y.1.N.0..ASQ.Before.Bayley <- NULL
data12$X.Y.1.N.0..ASQ.Before.Bayley <- NULL

# create variables indicating pass/fail
# ASQ communication
data6$ASQcomm_pass01 <- as.factor(ifelse(data6$asq6_commcat_calc %in% c(0,1),1,0))
data6$ASQcomm_pass0 <- as.factor(ifelse(data6$asq6_commcat_calc==0,1,0))
data12$ASQcomm_pass01 <- as.factor(ifelse(data12$asq12_commcat_calc %in% c(0,1),1,0))
data12$ASQcomm_pass0 <- as.factor(ifelse(data12$asq12_commcat_calc==0,1,0))
# Bayley language composite
data6$bsidlangcomp_pass3 <- as.factor(ifelse(data6$bsid_lang_qualdesc>=3,1,0))
data6$bsidlangcomp_pass4 <- as.factor(ifelse(data6$bsid_lang_qualdesc>=4,1,0))
data12$bsidlangcomp_pass3 <- as.factor(ifelse(data12$bsid_lang_qualdesc>=3,1,0))
data12$bsidlangcomp_pass4 <- as.factor(ifelse(data12$bsid_lang_qualdesc>=4,1,0))
# Bayley receptive subdomain
data6$bsidlangrecep_pass3 <- as.factor(ifelse(data6$bsid_rc_qualdesc>=3,1,0))
data6$bsidlangrecep_pass4 <- as.factor(ifelse(data6$bsid_rc_qualdesc>=4,1,0))
data12$bsidlangrecep_pass3 <- as.factor(ifelse(data12$bsid_rc_qualdesc>=3,1,0))
data12$bsidlangrecep_pass4 <- as.factor(ifelse(data12$bsid_rc_qualdesc>=4,1,0))
# Bayley expressive subdomain
data6$bsidlangexp_pass3 <- as.factor(ifelse(data6$bsid_ec_qualdesc>=3,1,0))
data6$bsidlangexp_pass4 <- as.factor(ifelse(data6$bsid_ec_qualdesc>=4,1,0))
data12$bsidlangexp_pass3 <- as.factor(ifelse(data12$bsid_ec_qualdesc>=3,1,0))
data12$bsidlangexp_pass4 <- as.factor(ifelse(data12$bsid_ec_qualdesc>=4,1,0))
# ASQ fine motor
data6$ASQfm_pass01 <- as.factor(ifelse(data6$asq6_fmotorcat_calc %in% c(0,1),1,0))
data6$ASQfm_pass0 <- as.factor(ifelse(data6$asq6_fmotorcat_calc==0,1,0))
data12$ASQfm_pass01 <- as.factor(ifelse(data12$asq12_fmotorcat_calc %in% c(0,1),1,0))
data12$ASQfm_pass0 <- as.factor(ifelse(data12$asq12_fmotorcat_calc==0,1,0))
# Bayley fine motor
data6$bsidfm_pass3 <- as.factor(ifelse(data6$bsid_gm_qualdesc>=3,1,0))
data6$bsidfm_pass4 <- as.factor(ifelse(data6$bsid_gm_qualdesc>=4,1,0))
data12$bsidfm_pass3 <- as.factor(ifelse(data12$bsid_gm_qualdesc>=3,1,0))
data12$bsidfm_pass4 <- as.factor(ifelse(data12$bsid_gm_qualdesc>=4,1,0))
# ASQ gross motor
data6$ASQgm_pass01 <- as.factor(ifelse(data6$asq6_gmotorcat_calc %in% c(0,1),1,0))
data6$ASQgm_pass0 <- as.factor(ifelse(data6$asq6_gmotorcat_calc==0,1,0))
data12$ASQgm_pass01 <- as.factor(ifelse(data12$asq12_gmotorcat_calc %in% c(0,1),1,0))
data12$ASQgm_pass0 <- as.factor(ifelse(data12$asq12_gmotorcat_calc==0,1,0))
# Bayley gross motor
data6$bsidgm_pass3 <- as.factor(ifelse(data6$bsid_gm_qualdesc>=3,1,0))
data6$bsidgm_pass4 <- as.factor(ifelse(data6$bsid_gm_qualdesc>=4,1,0))
data12$bsidgm_pass3 <- as.factor(ifelse(data12$bsid_gm_qualdesc>=3,1,0))
data12$bsidgm_pass4 <- as.factor(ifelse(data12$bsid_gm_qualdesc>=4,1,0))
# ASQ problem solving
data6$ASQps_pass01 <- as.factor(ifelse(data6$asq6_probslvcat_calc %in% c(0,1),1,0))
data6$ASQps_pass0 <- as.factor(ifelse(data6$asq6_probslvcat_calc==0,1,0))
data12$ASQps_pass01 <- as.factor(ifelse(data12$asq12_probslvcat_calc %in% c(0,1),1,0))
data12$ASQps_pass0 <- as.factor(ifelse(data12$asq12_probslvcat_calc==0,1,0))
# Bayley problem solving
data6$bsidps_pass3 <- as.factor(ifelse(data6$bsid_cog_qualdesc>=3,1,0))
data6$bsidps_pass4 <- as.factor(ifelse(data6$bsid_cog_qualdesc>=4,1,0))
data12$bsidps_pass3 <- as.factor(ifelse(data12$bsid_cog_qualdesc>=3,1,0))
data12$bsidps_pass4 <- as.factor(ifelse(data12$bsid_cog_qualdesc>=4,1,0))

# create dfs for separate analyses
# screener administered at the correct time
data6_correcttime <- data6[data6$ASQ_before_Bayley==1,]
data12_correcttime <- data12[data12$ASQ_before_Bayley==1,]
# site
data6_site1 <- data6[data6$Site..CHCO..1...Nemours..2.==1,]
data6_site2 <- data6[data6$Site..CHCO..1...Nemours..2.==2,]
data12_site1 <- data12[data12$Site..CHCO..1...Nemours..2.==1,]
data12_site2 <- data12[data12$Site..CHCO..1...Nemours..2.==2,]
# by chromosome group
# note - only 1 xxyy at 6 months so will ignore
data6xxy <- data6[data6$group==1,]
data6xyy <- data6[data6$group==2,]
data6xxx <- data6[data6$group==3,]
data12xxy <- data12[data12$group==1,]
data12xyy <- data12[data12$group==2,]
data12xxx <- data12[data12$group==3,]
data12xxyy <- data12[data12$group==4,]
```

```{r function, include=FALSE, comment=""}
#print(c("ASQ variable: ASQcomm_pass01"))
#print(c("BSID variable: bsidlangcomp_pass3"))
#print("Sensitivity:")
#sensitivity(data6$ASQcomm_pass01,data6$bsidlangcomp_pass3,positive="1")
#print("Specificity:")
#specificity(data6$ASQcomm_pass01,data6$bsidlangcomp_pass3,negative="0")
#print("PPV:")
#posPredValue(data6$ASQcomm_pass01,data6$bsidlangcomp_pass3,positive="1")
#print("NPV:")
#negPredValue(data6$ASQcomm_pass01,data6$bsidlangcomp_pass3,negative="0")

ss <- function(df,asqvar,bsidvar) {
  print(c(noquote("ASQ variable:"), noquote(eval(asqvar))))
  print(c(noquote("BSID variable:"), noquote(eval(bsidvar))))
  print(noquote("Sensitivity:"))
  print(sensitivity(df[,asqvar],df[,bsidvar],positive="1"))
  print(noquote("Specificity:"))
  print(specificity(df[,asqvar],df[,bsidvar],negative="0"))
  print(noquote("PPV:"))
  print(posPredValue(df[,asqvar],df[,bsidvar],positive="1"))
  print(noquote("NPV:"))
  print(negPredValue(df[,asqvar],df[,bsidvar],negative="0"))
}


```

# Background

# Methods

The following variables were defined:

ASQcomm_pass01: equal to 1 (pass) if asq6_commcat_calc or asq12_commcat_calc 0 or 1, otherwise equal to 0 (fail)

ASQcomm_pass0: equal to 1 (pass) if asq6_commcat_calc or asq12_commcat_calc 0, otherwise equal to 1 (fail)

bsidlangcomp_pass3: equal to 1 (pass) if bsid_lang_qualdesc >=3, otherwise equal to 0 (fail) 

bsidlangcomp_pass4: equal to 1 (pass) if bsid_lang_qualdesc >=4, otherwise equal to 0 (fail)  

bsidlangrecep_pass3: equal to 1 (pass) if bsid_rc_qualdesc >=3, otherwise equal to 0 (fail) 

bsidlangrecep_pass4: equal to 1 (pass) if bsid_rc_qualdesc >=4, otherwise equal to 0 (fail)

bsidlangexp_pass3: equal to 1 (pass) if bsid_ec_qualdesc >=3, otherwise equal to 0 (fail) 

bsidlangexp_pass4: equal to 1 (pass) if bsid_ec_qualdesc >=4, otherwise equal to 0 (fail) 

ASQfm_pass01: equal to 1 (pass) if ASQ6_fmotor_calc or ASQ12_fmotor_calc 0 or 1, otherwise equal to 0 (fail) 

ASQfm_pass0: equal to 1 (pass) if ASQ6_fmotor_calc or ASQ12_fmotor_calc 0, otherwise equal to 0 (fail) 

bsidfm_pass3: equal to 1 (pass) if bsid_fm_qualdesc >=3, otherwise equal to 0 (fail) 

bsidfm_pass4: equal to 1 (pass) if bsid_fm_qualdesc >=4, otherwise equal to 0 (fail) 

ASQgm_pass01: equal to 1 (pass) if ASQ6_gmotor_calc or ASQ12_gmotor_calc 0 or 1, otherwise equal to 0 (fail)  

ASQgm_pass0: equal to 1 (pass) if ASQ6_gmotor_calc or ASQ12_gmotor_calc 0, otherwise equal to 0 (fail) 

bsidgm_pass3: equal to 1 (pass) if bsid_gm_qualdesc >=3, otherwise equal to 0 (fail) 

bsidgm_pass4: equal to 1 (pass) if bsid_gm_qualdesc >=4, otherwise equal to 0 (fail)

ASQps_pass01: equal to 1 (pass) if ASQ6_probslvcat_calc or ASQ12_probslvcat_calc 0 or 1, otherwise equal to 0 (fail) 

ASQps_pass0: equal to 1 (pass) if ASQ6_probslvcat_calc or ASQ12_probslvcat_calc 0, otherwise equal to 0 (fail) 

bsidps_pass3: equal to 1 (pass) if bsid_cog_qualdesc >=3, otherwise equal to 0 (fail)  

bsidps_pass4: equal to 1 (pass) if bsid_cog_qualdesc >=4, otherwise equal to 0 (fail)  

# Results

## All data - 6 months

### Language

```{r echo=FALSE,comment=""}
ss(df=data6,asqvar="ASQcomm_pass01", bsidvar = "bsidlangcomp_pass3")
ss(df=data6,asqvar="ASQcomm_pass0", bsidvar = "bsidlangcomp_pass3")
ss(df=data6,asqvar="ASQcomm_pass01", bsidvar = "bsidlangcomp_pass4")
ss(df=data6,asqvar="ASQcomm_pass0", bsidvar = "bsidlangcomp_pass4")
ss(df=data6,asqvar="ASQcomm_pass01", bsidvar = "bsidlangrecep_pass3")
ss(df=data6,asqvar="ASQcomm_pass0", bsidvar = "bsidlangrecep_pass3")
ss(df=data6,asqvar="ASQcomm_pass01", bsidvar = "bsidlangrecep_pass4")
ss(df=data6,asqvar="ASQcomm_pass0", bsidvar = "bsidlangrecep_pass4")
ss(df=data6,asqvar="ASQcomm_pass01", bsidvar = "bsidlangexp_pass3")
ss(df=data6,asqvar="ASQcomm_pass0", bsidvar = "bsidlangexp_pass3")
ss(df=data6,asqvar="ASQcomm_pass01", bsidvar = "bsidlangexp_pass4")
ss(df=data6,asqvar="ASQcomm_pass0", bsidvar = "bsidlangexp_pass4")
```

### Fine motor

```{r echo=FALSE,comment=""}
ss(df=data6,asqvar="ASQfm_pass01", bsidvar = "bsidfm_pass3")
ss(df=data6,asqvar="ASQfm_pass0", bsidvar = "bsidfm_pass3")
ss(df=data6,asqvar="ASQfm_pass01", bsidvar = "bsidfm_pass4")
ss(df=data6,asqvar="ASQfm_pass0", bsidvar = "bsidfm_pass4")
```

### Gross motor

```{r echo=FALSE,comment=""}
ss(df=data6,asqvar="ASQgm_pass01", bsidvar = "bsidgm_pass3")
ss(df=data6,asqvar="ASQgm_pass0", bsidvar = "bsidgm_pass3")
ss(df=data6,asqvar="ASQgm_pass01", bsidvar = "bsidgm_pass4")
ss(df=data6,asqvar="ASQgm_pass0", bsidvar = "bsidgm_pass4")
```

### Problem solving/Cognitive

```{r echo=FALSE,comment=""}
ss(df=data6,asqvar="ASQps_pass01", bsidvar = "bsidps_pass3")
ss(df=data6,asqvar="ASQps_pass0", bsidvar = "bsidps_pass3")
ss(df=data6,asqvar="ASQps_pass01", bsidvar = "bsidps_pass4")
ss(df=data6,asqvar="ASQps_pass0", bsidvar = "bsidps_pass4")
```

## All data - 12 months

### Language

```{r echo=FALSE,comment=""}
ss(df=data12,asqvar="ASQcomm_pass01", bsidvar = "bsidlangcomp_pass3")
ss(df=data12,asqvar="ASQcomm_pass0", bsidvar = "bsidlangcomp_pass3")
ss(df=data12,asqvar="ASQcomm_pass01", bsidvar = "bsidlangcomp_pass4")
ss(df=data12,asqvar="ASQcomm_pass0", bsidvar = "bsidlangcomp_pass4")
ss(df=data12,asqvar="ASQcomm_pass01", bsidvar = "bsidlangrecep_pass3")
ss(df=data12,asqvar="ASQcomm_pass0", bsidvar = "bsidlangrecep_pass3")
ss(df=data12,asqvar="ASQcomm_pass01", bsidvar = "bsidlangrecep_pass4")
ss(df=data12,asqvar="ASQcomm_pass0", bsidvar = "bsidlangrecep_pass4")
ss(df=data12,asqvar="ASQcomm_pass01", bsidvar = "bsidlangexp_pass3")
ss(df=data12,asqvar="ASQcomm_pass0", bsidvar = "bsidlangexp_pass3")
ss(df=data12,asqvar="ASQcomm_pass01", bsidvar = "bsidlangexp_pass4")
ss(df=data12,asqvar="ASQcomm_pass0", bsidvar = "bsidlangexp_pass4")
```

### Fine motor

```{r echo=FALSE,comment=""}
ss(df=data12,asqvar="ASQfm_pass01", bsidvar = "bsidfm_pass3")
ss(df=data12,asqvar="ASQfm_pass0", bsidvar = "bsidfm_pass3")
ss(df=data12,asqvar="ASQfm_pass01", bsidvar = "bsidfm_pass4")
ss(df=data12,asqvar="ASQfm_pass0", bsidvar = "bsidfm_pass4")
```

### Gross motor

```{r echo=FALSE,comment=""}
ss(df=data12,asqvar="ASQgm_pass01", bsidvar = "bsidgm_pass3")
ss(df=data12,asqvar="ASQgm_pass0", bsidvar = "bsidgm_pass3")
ss(df=data12,asqvar="ASQgm_pass01", bsidvar = "bsidgm_pass4")
ss(df=data12,asqvar="ASQgm_pass0", bsidvar = "bsidgm_pass4")
```

### Problem solving/Cognitive

```{r echo=FALSE,comment=""}
ss(df=data12,asqvar="ASQps_pass01", bsidvar = "bsidps_pass3")
ss(df=data12,asqvar="ASQps_pass0", bsidvar = "bsidps_pass3")
ss(df=data12,asqvar="ASQps_pass01", bsidvar = "bsidps_pass4")
ss(df=data12,asqvar="ASQps_pass0", bsidvar = "bsidps_pass4")
```

## ASQ before Bayley - 6 months

### Language

```{r echo=FALSE,comment=""}
ss(df=data6_correcttime,asqvar="ASQcomm_pass01", bsidvar = "bsidlangcomp_pass3")
ss(df=data6_correcttime,asqvar="ASQcomm_pass0", bsidvar = "bsidlangcomp_pass3")
ss(df=data6_correcttime,asqvar="ASQcomm_pass01", bsidvar = "bsidlangcomp_pass4")
ss(df=data6_correcttime,asqvar="ASQcomm_pass0", bsidvar = "bsidlangcomp_pass4")
ss(df=data6_correcttime,asqvar="ASQcomm_pass01", bsidvar = "bsidlangrecep_pass3")
ss(df=data6_correcttime,asqvar="ASQcomm_pass0", bsidvar = "bsidlangrecep_pass3")
ss(df=data6_correcttime,asqvar="ASQcomm_pass01", bsidvar = "bsidlangrecep_pass4")
ss(df=data6_correcttime,asqvar="ASQcomm_pass0", bsidvar = "bsidlangrecep_pass4")
ss(df=data6_correcttime,asqvar="ASQcomm_pass01", bsidvar = "bsidlangexp_pass3")
ss(df=data6_correcttime,asqvar="ASQcomm_pass0", bsidvar = "bsidlangexp_pass3")
ss(df=data6_correcttime,asqvar="ASQcomm_pass01", bsidvar = "bsidlangexp_pass4")
ss(df=data6_correcttime,asqvar="ASQcomm_pass0", bsidvar = "bsidlangexp_pass4")
```

### Fine motor

```{r echo=FALSE,comment=""}
ss(df=data6_correcttime,asqvar="ASQfm_pass01", bsidvar = "bsidfm_pass3")
ss(df=data6_correcttime,asqvar="ASQfm_pass0", bsidvar = "bsidfm_pass3")
ss(df=data6_correcttime,asqvar="ASQfm_pass01", bsidvar = "bsidfm_pass4")
ss(df=data6_correcttime,asqvar="ASQfm_pass0", bsidvar = "bsidfm_pass4")
```

### Gross motor

```{r echo=FALSE,comment=""}
ss(df=data6_correcttime,asqvar="ASQgm_pass01", bsidvar = "bsidgm_pass3")
ss(df=data6_correcttime,asqvar="ASQgm_pass0", bsidvar = "bsidgm_pass3")
ss(df=data6_correcttime,asqvar="ASQgm_pass01", bsidvar = "bsidgm_pass4")
ss(df=data6_correcttime,asqvar="ASQgm_pass0", bsidvar = "bsidgm_pass4")
```

### Problem solving/Cognitive

```{r echo=FALSE,comment=""}
ss(df=data6_correcttime,asqvar="ASQps_pass01", bsidvar = "bsidps_pass3")
ss(df=data6_correcttime,asqvar="ASQps_pass0", bsidvar = "bsidps_pass3")
ss(df=data6_correcttime,asqvar="ASQps_pass01", bsidvar = "bsidps_pass4")
ss(df=data6_correcttime,asqvar="ASQps_pass0", bsidvar = "bsidps_pass4")
```

## ASQ before Bayley - 12 months

### Language

```{r echo=FALSE,comment=""}
ss(df=data12_correcttime,asqvar="ASQcomm_pass01", bsidvar = "bsidlangcomp_pass3")
ss(df=data12_correcttime,asqvar="ASQcomm_pass0", bsidvar = "bsidlangcomp_pass3")
ss(df=data12_correcttime,asqvar="ASQcomm_pass01", bsidvar = "bsidlangcomp_pass4")
ss(df=data12_correcttime,asqvar="ASQcomm_pass0", bsidvar = "bsidlangcomp_pass4")
ss(df=data12_correcttime,asqvar="ASQcomm_pass01", bsidvar = "bsidlangrecep_pass3")
ss(df=data12_correcttime,asqvar="ASQcomm_pass0", bsidvar = "bsidlangrecep_pass3")
ss(df=data12_correcttime,asqvar="ASQcomm_pass01", bsidvar = "bsidlangrecep_pass4")
ss(df=data12_correcttime,asqvar="ASQcomm_pass0", bsidvar = "bsidlangrecep_pass4")
ss(df=data12_correcttime,asqvar="ASQcomm_pass01", bsidvar = "bsidlangexp_pass3")
ss(df=data12_correcttime,asqvar="ASQcomm_pass0", bsidvar = "bsidlangexp_pass3")
ss(df=data12_correcttime,asqvar="ASQcomm_pass01", bsidvar = "bsidlangexp_pass4")
ss(df=data12_correcttime,asqvar="ASQcomm_pass0", bsidvar = "bsidlangexp_pass4")
```

### Fine motor

```{r echo=FALSE,comment=""}
ss(df=data12_correcttime,asqvar="ASQfm_pass01", bsidvar = "bsidfm_pass3")
ss(df=data12_correcttime,asqvar="ASQfm_pass0", bsidvar = "bsidfm_pass3")
ss(df=data12_correcttime,asqvar="ASQfm_pass01", bsidvar = "bsidfm_pass4")
ss(df=data12_correcttime,asqvar="ASQfm_pass0", bsidvar = "bsidfm_pass4")
```

### Gross motor

```{r echo=FALSE,comment=""}
ss(df=data12_correcttime,asqvar="ASQgm_pass01", bsidvar = "bsidgm_pass3")
ss(df=data12_correcttime,asqvar="ASQgm_pass0", bsidvar = "bsidgm_pass3")
ss(df=data12_correcttime,asqvar="ASQgm_pass01", bsidvar = "bsidgm_pass4")
ss(df=data12_correcttime,asqvar="ASQgm_pass0", bsidvar = "bsidgm_pass4")
```

### Problem solving/Cognitive

```{r echo=FALSE,comment=""}
ss(df=data12_correcttime,asqvar="ASQps_pass01", bsidvar = "bsidps_pass3")
ss(df=data12_correcttime,asqvar="ASQps_pass0", bsidvar = "bsidps_pass3")
ss(df=data12_correcttime,asqvar="ASQps_pass01", bsidvar = "bsidps_pass4")
ss(df=data12_correcttime,asqvar="ASQps_pass0", bsidvar = "bsidps_pass4")
```

## Site 1 - 6 months

### Language

```{r echo=FALSE,comment=""}
ss(df=data6_site1,asqvar="ASQcomm_pass01", bsidvar = "bsidlangcomp_pass3")
ss(df=data6_site1,asqvar="ASQcomm_pass0", bsidvar = "bsidlangcomp_pass3")
ss(df=data6_site1,asqvar="ASQcomm_pass01", bsidvar = "bsidlangcomp_pass4")
ss(df=data6_site1,asqvar="ASQcomm_pass0", bsidvar = "bsidlangcomp_pass4")
ss(df=data6_site1,asqvar="ASQcomm_pass01", bsidvar = "bsidlangrecep_pass3")
ss(df=data6_site1,asqvar="ASQcomm_pass0", bsidvar = "bsidlangrecep_pass3")
ss(df=data6_site1,asqvar="ASQcomm_pass01", bsidvar = "bsidlangrecep_pass4")
ss(df=data6_site1,asqvar="ASQcomm_pass0", bsidvar = "bsidlangrecep_pass4")
ss(df=data6_site1,asqvar="ASQcomm_pass01", bsidvar = "bsidlangexp_pass3")
ss(df=data6_site1,asqvar="ASQcomm_pass0", bsidvar = "bsidlangexp_pass3")
ss(df=data6_site1,asqvar="ASQcomm_pass01", bsidvar = "bsidlangexp_pass4")
ss(df=data6_site1,asqvar="ASQcomm_pass0", bsidvar = "bsidlangexp_pass4")
```

### Fine motor

```{r echo=FALSE,comment=""}
ss(df=data6_site1,asqvar="ASQfm_pass01", bsidvar = "bsidfm_pass3")
ss(df=data6_site1,asqvar="ASQfm_pass0", bsidvar = "bsidfm_pass3")
ss(df=data6_site1,asqvar="ASQfm_pass01", bsidvar = "bsidfm_pass4")
ss(df=data6_site1,asqvar="ASQfm_pass0", bsidvar = "bsidfm_pass4")
```

### Gross motor

```{r echo=FALSE,comment=""}
ss(df=data6_site1,asqvar="ASQgm_pass01", bsidvar = "bsidgm_pass3")
ss(df=data6_site1,asqvar="ASQgm_pass0", bsidvar = "bsidgm_pass3")
ss(df=data6_site1,asqvar="ASQgm_pass01", bsidvar = "bsidgm_pass4")
ss(df=data6_site1,asqvar="ASQgm_pass0", bsidvar = "bsidgm_pass4")
```

### Problem solving/Cognitive

```{r echo=FALSE,comment=""}
ss(df=data6_site1,asqvar="ASQps_pass01", bsidvar = "bsidps_pass3")
ss(df=data6_site1,asqvar="ASQps_pass0", bsidvar = "bsidps_pass3")
ss(df=data6_site1,asqvar="ASQps_pass01", bsidvar = "bsidps_pass4")
ss(df=data6_site1,asqvar="ASQps_pass0", bsidvar = "bsidps_pass4")
```

## Site 1 - 12 months

### Language

```{r echo=FALSE,comment=""}
ss(df=data12_site1,asqvar="ASQcomm_pass01", bsidvar = "bsidlangcomp_pass3")
ss(df=data12_site1,asqvar="ASQcomm_pass0", bsidvar = "bsidlangcomp_pass3")
ss(df=data12_site1,asqvar="ASQcomm_pass01", bsidvar = "bsidlangcomp_pass4")
ss(df=data12_site1,asqvar="ASQcomm_pass0", bsidvar = "bsidlangcomp_pass4")
ss(df=data12_site1,asqvar="ASQcomm_pass01", bsidvar = "bsidlangrecep_pass3")
ss(df=data12_site1,asqvar="ASQcomm_pass0", bsidvar = "bsidlangrecep_pass3")
ss(df=data12_site1,asqvar="ASQcomm_pass01", bsidvar = "bsidlangrecep_pass4")
ss(df=data12_site1,asqvar="ASQcomm_pass0", bsidvar = "bsidlangrecep_pass4")
ss(df=data12_site1,asqvar="ASQcomm_pass01", bsidvar = "bsidlangexp_pass3")
ss(df=data12_site1,asqvar="ASQcomm_pass0", bsidvar = "bsidlangexp_pass3")
ss(df=data12_site1,asqvar="ASQcomm_pass01", bsidvar = "bsidlangexp_pass4")
ss(df=data12_site1,asqvar="ASQcomm_pass0", bsidvar = "bsidlangexp_pass4")
```

### Fine motor

```{r echo=FALSE,comment=""}
ss(df=data12_site1,asqvar="ASQfm_pass01", bsidvar = "bsidfm_pass3")
ss(df=data12_site1,asqvar="ASQfm_pass0", bsidvar = "bsidfm_pass3")
ss(df=data12_site1,asqvar="ASQfm_pass01", bsidvar = "bsidfm_pass4")
ss(df=data12_site1,asqvar="ASQfm_pass0", bsidvar = "bsidfm_pass4")
```

### Gross motor

```{r echo=FALSE,comment=""}
ss(df=data12_site1,asqvar="ASQgm_pass01", bsidvar = "bsidgm_pass3")
ss(df=data12_site1,asqvar="ASQgm_pass0", bsidvar = "bsidgm_pass3")
ss(df=data12_site1,asqvar="ASQgm_pass01", bsidvar = "bsidgm_pass4")
ss(df=data12_site1,asqvar="ASQgm_pass0", bsidvar = "bsidgm_pass4")
```

### Problem solving/Cognitive

```{r echo=FALSE,comment=""}
ss(df=data12_site1,asqvar="ASQps_pass01", bsidvar = "bsidps_pass3")
ss(df=data12_site1,asqvar="ASQps_pass0", bsidvar = "bsidps_pass3")
ss(df=data12_site1,asqvar="ASQps_pass01", bsidvar = "bsidps_pass4")
ss(df=data12_site1,asqvar="ASQps_pass0", bsidvar = "bsidps_pass4")
```

## Site 2 - 6 months

### Language

```{r echo=FALSE,comment=""}
ss(df=data6_site2,asqvar="ASQcomm_pass01", bsidvar = "bsidlangcomp_pass3")
ss(df=data6_site2,asqvar="ASQcomm_pass0", bsidvar = "bsidlangcomp_pass3")
ss(df=data6_site2,asqvar="ASQcomm_pass01", bsidvar = "bsidlangcomp_pass4")
ss(df=data6_site2,asqvar="ASQcomm_pass0", bsidvar = "bsidlangcomp_pass4")
ss(df=data6_site2,asqvar="ASQcomm_pass01", bsidvar = "bsidlangrecep_pass3")
ss(df=data6_site2,asqvar="ASQcomm_pass0", bsidvar = "bsidlangrecep_pass3")
ss(df=data6_site2,asqvar="ASQcomm_pass01", bsidvar = "bsidlangrecep_pass4")
ss(df=data6_site2,asqvar="ASQcomm_pass0", bsidvar = "bsidlangrecep_pass4")
ss(df=data6_site2,asqvar="ASQcomm_pass01", bsidvar = "bsidlangexp_pass3")
ss(df=data6_site2,asqvar="ASQcomm_pass0", bsidvar = "bsidlangexp_pass3")
ss(df=data6_site2,asqvar="ASQcomm_pass01", bsidvar = "bsidlangexp_pass4")
ss(df=data6_site2,asqvar="ASQcomm_pass0", bsidvar = "bsidlangexp_pass4")
```

### Fine motor

```{r echo=FALSE,comment=""}
ss(df=data6_site2,asqvar="ASQfm_pass01", bsidvar = "bsidfm_pass3")
ss(df=data6_site2,asqvar="ASQfm_pass0", bsidvar = "bsidfm_pass3")
ss(df=data6_site2,asqvar="ASQfm_pass01", bsidvar = "bsidfm_pass4")
ss(df=data6_site2,asqvar="ASQfm_pass0", bsidvar = "bsidfm_pass4")
```

### Gross motor

```{r echo=FALSE,comment=""}
ss(df=data6_site2,asqvar="ASQgm_pass01", bsidvar = "bsidgm_pass3")
ss(df=data6_site2,asqvar="ASQgm_pass0", bsidvar = "bsidgm_pass3")
ss(df=data6_site2,asqvar="ASQgm_pass01", bsidvar = "bsidgm_pass4")
ss(df=data6_site2,asqvar="ASQgm_pass0", bsidvar = "bsidgm_pass4")
```

### Problem solving/Cognitive

```{r echo=FALSE,comment=""}
ss(df=data6_site2,asqvar="ASQps_pass01", bsidvar = "bsidps_pass3")
ss(df=data6_site2,asqvar="ASQps_pass0", bsidvar = "bsidps_pass3")
ss(df=data6_site2,asqvar="ASQps_pass01", bsidvar = "bsidps_pass4")
ss(df=data6_site2,asqvar="ASQps_pass0", bsidvar = "bsidps_pass4")
```

## Site 2 - 12 months

### Language

```{r echo=FALSE,comment=""}
ss(df=data12_site2,asqvar="ASQcomm_pass01", bsidvar = "bsidlangcomp_pass3")
ss(df=data12_site2,asqvar="ASQcomm_pass0", bsidvar = "bsidlangcomp_pass3")
ss(df=data12_site2,asqvar="ASQcomm_pass01", bsidvar = "bsidlangcomp_pass4")
ss(df=data12_site2,asqvar="ASQcomm_pass0", bsidvar = "bsidlangcomp_pass4")
ss(df=data12_site2,asqvar="ASQcomm_pass01", bsidvar = "bsidlangrecep_pass3")
ss(df=data12_site2,asqvar="ASQcomm_pass0", bsidvar = "bsidlangrecep_pass3")
ss(df=data12_site2,asqvar="ASQcomm_pass01", bsidvar = "bsidlangrecep_pass4")
ss(df=data12_site2,asqvar="ASQcomm_pass0", bsidvar = "bsidlangrecep_pass4")
ss(df=data12_site2,asqvar="ASQcomm_pass01", bsidvar = "bsidlangexp_pass3")
ss(df=data12_site2,asqvar="ASQcomm_pass0", bsidvar = "bsidlangexp_pass3")
ss(df=data12_site2,asqvar="ASQcomm_pass01", bsidvar = "bsidlangexp_pass4")
ss(df=data12_site2,asqvar="ASQcomm_pass0", bsidvar = "bsidlangexp_pass4")
```

### Fine motor

```{r echo=FALSE,comment=""}
ss(df=data12_site2,asqvar="ASQfm_pass01", bsidvar = "bsidfm_pass3")
ss(df=data12_site2,asqvar="ASQfm_pass0", bsidvar = "bsidfm_pass3")
ss(df=data12_site2,asqvar="ASQfm_pass01", bsidvar = "bsidfm_pass4")
ss(df=data12_site2,asqvar="ASQfm_pass0", bsidvar = "bsidfm_pass4")
```

### Gross motor

```{r echo=FALSE,comment=""}
ss(df=data12_site2,asqvar="ASQgm_pass01", bsidvar = "bsidgm_pass3")
ss(df=data12_site2,asqvar="ASQgm_pass0", bsidvar = "bsidgm_pass3")
ss(df=data12_site2,asqvar="ASQgm_pass01", bsidvar = "bsidgm_pass4")
ss(df=data12_site2,asqvar="ASQgm_pass0", bsidvar = "bsidgm_pass4")
```

### Problem solving/Cognitive

```{r echo=FALSE,comment=""}
ss(df=data12_site2,asqvar="ASQps_pass01", bsidvar = "bsidps_pass3")
ss(df=data12_site2,asqvar="ASQps_pass0", bsidvar = "bsidps_pass3")
ss(df=data12_site2,asqvar="ASQps_pass01", bsidvar = "bsidps_pass4")
ss(df=data12_site2,asqvar="ASQps_pass0", bsidvar = "bsidps_pass4")
```



