---
title: "Proteomics and HTN - SOMAScan correlation"
author: "Laura Pyle, Ye Ji Choi & Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
editor: source
execute:
  echo: false
  message: false
---

```{r libraries}
#| include: false
library(psych)
library(corrplot)
library(pander)
library(dplyr)
library(tableone)
library(knitr)
# Print all pander tables
panderOptions("knitr.auto.asis", FALSE)
```

```{r data import}
#| include: false
# source("~/GitHub/CHCO-Code/Petter Bjornstad/Proteomics and DKD/create_today_dkd_analysis_dataset.R")
load("/Volumes/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/Proteomics and HTN/Data_Cleaned/analysis_dataset.RData")
top_htn <- top_htn_df %>%
  filter(Target %in% c("SEZ6L","SCG3","LSAMP","T132B","Nr-CAM","IGLO5",
                       "PCD10:ECD","UNC5H4","SLIK1",
                       "Secretoglobin family 3A member 1","sICAM-5")) %>%
  slice_max(abs(log(estimate)), n = 11) %>%
  pull(AptName)
df_t2d <- df %>% filter(group=="Type 2 Diabetes")
df_lc <- df %>% filter(group=="Lean Control")
df_oc <- df %>% filter(group=="Obese Control")
```

```{r functions}
cor_and_plot_soma <- function(data, vars) {
  # Format matrix
  d <- data[, c(vars, "acr_u", "sbp", "dbp", "map", "gfr_raw_plasma", "gfr_bsa_plasma")]
  n <- c(
    analytes$EntrezGeneSymbol[match(vars, analytes$AptName)],
    "acr_u", "sbp", "dbp", "map", "gfr_raw_plasma", "gfr_bsa_plasma"
  )
  n[duplicated(n)] <- paste0(n[duplicated(n)], "(2)")
  names(d) <- n
  # Correlations
  M <- corr.test(data.matrix(d), normal = F, method = "spearman")
  corrplot(round(M$r, 3),
    p.mat = M$p, method = "circle", type = "lower", insig = "blank",
    addCoef.col = "white", diag = FALSE
  )
  # Print a nice looking table
  r <- data.frame(round(M$r, 3))
  emphasize.strong.cells(which(M$p <= 0.05, arr.ind = TRUE))
  cat("\n")
  pander(r, caption = "Coefficients")
  cat("\n")
  # print p-values
  p <- data.frame(round(M$p, 3))
  cat("\n")
  pander(p, caption = "P-values")
  cat("\n")
   # print Ns
  n <- data.frame(round(M$n, 3))
  cat("\n")
  pander(n, caption = "Number of observations")
  cat("\n")
}
```


# HTN

The proteins with the largest absolute effect size and adjusted p value $\leq 0.05$ for hypertension were: `r toString(analytes$TargetFullName[match(top_htn,analytes$AptName)])` (associated with genes `r toString(analytes$EntrezGeneSymbol[match(top_htn,analytes$AptName)])`, respectively).

## T2D

### Correlations

```{r results='asis', dpi=1200}
cor_and_plot_soma(data = df_t2d, vars = top_htn)
```

### T-tests by hypertension status

```{r include=FALSE}
t <- CreateTableOne(data=df_t2d, vars=top_htn, strata="htn")
t <- print(t)
```

```{r results='as is'}
kable(t)
```

## Lean Controls

### Correlations

```{r results='asis', dpi=1200}
cor_and_plot_soma(data = df_lc, vars = top_htn)
```

### T-tests by hypertension status

```{r include=FALSE}
t <- CreateTableOne(data=df_lc, vars=top_htn, strata="htn")
t <- print(t)
```

```{r results='as is'}
kable(t)
```

## Obese Controls

### Correlations

```{r results='asis', dpi=1200}
cor_and_plot_soma(data = df_oc, vars = top_htn)
```

### T-tests by hypertension status

```{r include=FALSE}
t <- CreateTableOne(data=df_oc, vars=top_htn, strata="htn")
t <- print(t)
```

```{r results='as is'}
kable(t)
```

```{r, include=TRUE}
file.copy("/Users/pylell/Documents/GitHub/CHCO-Code/Petter Bjornstad/Proteomics and HTN/proteomics_and_HTN_somascan.html",
          "/Users/pylell/Dropbox/TODAY HTN and DKD output/proteomics_and_HTN_somascan.html",overwrite = TRUE)
```
