---
title: "Proteomics and DKD - TODAY Pathway Analyses"
author: "Laura Pyle, Ye Ji Choi & Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
editor: source
execute:
  echo: false
  message: false
---

```{r libraries}
#| include: false
library(tidyverse)
library(ggrepel)
library(clusterProfiler)
library(ReactomePA)
library(patchwork)
```

```{r data import}
#| include: false
# source("~/GitHub/CHCO-Code/Petter Bjornstad/Proteomics and DKD/create_today_dkd_analysis_dataset.R")
load("/Volumes/Peds Endo/Petter Bjornstad/Proteomics and DKD/Data_Cleaned/analysis_dataset.RData")
```

```{r functions}
# Volcano plot
volcano <- function(data, lab = "EntrezGeneSymbol", xcol = "estimate", ycol = "p.value", 
                    top = 3,
                    xlimit = c(-1, 1), ylimit = c(0, -log10(10e-7)),
                    xlabel = "HR", pCutoff = 0.05, overlaps = 10,
                    log_t = T) {
  data <- as.data.frame(data)
  t <- data[data[, "adj.p.value"] <= pCutoff, ]
  if (log_t) {
    t <- t[order(abs(log(t[, xcol])), decreasing = T), ]
  } else {
    t <- t[order(abs(t[, xcol]), decreasing = T), ]
  }
  data$top <- data[, "AptName"] %in% t[1:top, "AptName"]
  data$logp <- -log10(data[, ycol])
  data$fc <- data[, xcol]
  data$sig <- data[, ycol] <= pCutoff
  p <- ggplot(data = data, aes(x = fc, y = logp, color = sig)) +
    geom_hline(yintercept = -log10(pCutoff), linetype = "dashed") +
    geom_point(size = 2) +
    geom_label_repel(
      data = data[data$top, ], aes(label = EntrezGeneSymbol), color = "black",
      max.overlaps = overlaps
    ) +
    scale_color_manual(values = c("grey", "#3e6dbf")) +
    xlab(xlabel) +
    ylab(bquote(~ -Log[10] ~ italic(P))) +
    theme_bw() +
    theme(legend.position = "none")
  return(p)
}
# GSEA
pathway_analysis <- function(de_genes) {
  # Log transform if necessary
  if(!any(de_genes < 0)){
    de_genes = log(de_genes)
  }
  de_genes = sort(de_genes,decreasing = T)
  de_genes = de_genes[!duplicated(names(de_genes))]
  gsea = gsePathway(de_genes)
  gsea = setReadable(gsea, "org.Hs.eg.db", "ENTREZID")
  dot = dotplot(gsea)
  # Return
  return(dot)
}
# Function for comparing two gene lists
# de_gene_list is a named list of multiple gene sets
pathway_comparison <- function(de_gene_list) {
  # Log transform if necessary
  de_gene_list = lapply(de_gene_list,function(l){
    if(!any(l < 0)){
      l = log(l)
    }
    return(l)
  })
  clust_df = lapply(names(de_gene_list),function(t){
    d = de_gene_list[[t]]
    d = d[!is.na(names(d))]
    d = d[!duplicated(names(d))]
    n = names(d)
    d = data.frame(cbind(n,d))
    colnames(d) = c("Entrez","logFC")
    d$group = t
    return(d)
  })
  clust_df = data.frame(do.call(rbind,clust_df))
  clust_df$logFC = as.numeric(clust_df$logFC)
  clust_df = clust_df[order(clust_df$logFC),]
  comp = compareCluster(Entrez~group, data=clust_df,
                        fun="groupGO", OrgDb='org.Hs.eg.db')
  # Plots
  updot <- dotplot(upck) + ggtitle("Upregulated") + 
    theme(plot.title = element_text(hjust = 0.5))
  upnet <- cnetplot(upck) + ggtitle("Upregulated") + 
    theme(plot.title = element_text(hjust = 0.5))
  downdot <- dotplot(downck) + ggtitle("Downregulated") + 
    theme(plot.title = element_text(hjust = 0.5))
  downnet <- cnetplot(downck) + ggtitle("Downregulated") + 
    theme(plot.title = element_text(hjust = 0.5))
  # Return
  return(list("updot" = updot, "upnet" = upnet,
              "downdot" = downdot, "downnet" = downnet))
}
```

# TODAY pathway analysis

## MAC

The proteins with the largest absolute effect size and adjusted p value $\leq 0.05$ for MAC were: `r toString(analytes$TargetFullName[match(top_mac,analytes$AptName)])` (associated with genes `r toString(analytes$EntrezGeneSymbol[match(top_mac,analytes$AptName)])`, respectively).

```{r message=FALSE,warning=FALSE,dpi=1200}
v_mac <- volcano(top_mac_df)
v_mac <- v_mac +
  ggtitle("Severe Albuminuria") +
  theme(plot.title = element_text(hjust = 0.5))
v_mac
```

```{r message=FALSE,warning=FALSE,dpi=1200}
mac_p <- pathway_analysis(de_genes_mac)

mac_p = mac_p + ggtitle("Severe Albuminuria") + 
  theme(plot.title = element_text(hjust = 0.5))

# ggsave(filename = "/Users/timvigers/Library/CloudStorage/Dropbox/Work/Shared/TODAY/TODAY DKD manuscript [shared]/Figures/MAC_ORA_1600x900.jpeg",plot = mac_p,
#        width = 1600,height = 900,units = "px",scale = 2.5)

mac_p_comp <- pathway_comparison(list(
  Baseline = de_genes_mac,
  Followup = de_genes_mac_10
))
mac_p_comp$updot
mac_p_comp$downdot

mac_comp_plot = mac_p_comp$updot + mac_p_comp$downdot + 
  plot_annotation(title = "Severe Albuminuria",
                  theme = theme(plot.title = element_text(hjust = 0.5)))
# ggsave(filename = "/Users/timvigers/Library/CloudStorage/Dropbox/Work/Shared/TODAY/TODAY DKD manuscript [shared]/Figures/MAC_ORA_comparison_1600x900.jpeg",
#        plot = mac_comp_plot,width = 1600,height = 900,units = "px",scale = 2.5)
```

## MIC

The proteins with the largest absolute effect size and adjusted p value $\leq 0.05$ for MIC were: `r toString(analytes$TargetFullName[match(top_mic,analytes$AptName)])` (associated with genes `r toString(analytes$EntrezGeneSymbol[match(top_mic,analytes$AptName)])`, respectively).

```{r message=FALSE,warning=FALSE,dpi=1200}
v_mic <- volcano(top_mic_df)
v_mic <- v_mic +
  ggtitle("Elevated Albuminuria") +
  theme(plot.title = element_text(hjust = 0.5))
v_mic
```

```{r message=FALSE,warning=FALSE,dpi=1200}
mic_p <- pathway_analysis(de_genes_mic)
mic_p$updot
mic_p$downdot

mic_plot = mic_p$updot + mic_p$downdot + 
  plot_annotation(title = "Elevated Albuminuria",
                  theme = theme(plot.title = element_text(hjust = 0.5)))
# ggsave(filename = "/Users/timvigers/Library/CloudStorage/Dropbox/Work/Shared/TODAY/TODAY DKD manuscript [shared]/Figures/MIC_ORA_1600x900.jpeg",plot = mic_plot,
#        width = 1600,height = 900,units = "px",scale = 2.5)

mic_p_comp <- pathway_comparison(list(
  Baseline = de_genes_mic,
  Followup = de_genes_mic_10
))
mic_p_comp$updot
mic_p_comp$downdot

mic_comp_plot = mic_p_comp$updot + mic_p_comp$downdot + 
  plot_annotation(title = "Elevated Albuminuria",
                  theme = theme(plot.title = element_text(hjust = 0.5)))
# ggsave(filename = "/Users/timvigers/Library/CloudStorage/Dropbox/Work/Shared/TODAY/TODAY DKD manuscript [shared]/Figures/MIC_ORA_comparison_1600x900.jpeg",
#        plot = mic_comp_plot,width = 1600,height = 900,units = "px",scale = 2.5)
```

## HYP

The proteins with the largest absolute effect size and adjusted p value $\leq 0.05$ for hypertension were: `r toString(analytes$TargetFullName[match(top_hyp,analytes$AptName)])` (associated with genes `r toString(analytes$EntrezGeneSymbol[match(top_hyp,analytes$AptName)])`, respectively).

```{r message=FALSE,warning=FALSE,dpi=1200}
v_hyp <- volcano(top_hyp_df)
v_hyp <- v_hyp +
  ggtitle("Hyperfiltration") +
  theme(plot.title = element_text(hjust = 0.5))
v_hyp
```

```{r message=FALSE,warning=FALSE,dpi=1200}
hyp_p <- pathway_analysis(de_genes_hyp)
hyp_p$updot
hyp_p$downdot

hyp_plot = hyp_p$updot + hyp_p$downdot + 
  plot_annotation(title = "Hyperfiltration",
                  theme = theme(plot.title = element_text(hjust = 0.5)))
# ggsave(filename = "/Users/timvigers/Library/CloudStorage/Dropbox/Work/Shared/TODAY/TODAY DKD manuscript [shared]/Figures/HYP_ORA_1600x900.jpeg",plot = hyp_plot,
#        width = 1600,height = 900,units = "px",scale = 3)

hyp_p_comp <- pathway_comparison(list(
  Baseline = de_genes_hyp,
  Followup = de_genes_hyp_10
))
hyp_p_comp$updot
hyp_p_comp$downdot

hyp_comp_plot = hyp_p_comp$updot + hyp_p_comp$downdot + 
  plot_annotation(title = "Hyperfiltration",
                  theme = theme(plot.title = element_text(hjust = 0.5)))
# ggsave(filename = "/Users/timvigers/Library/CloudStorage/Dropbox/Work/Shared/TODAY/TODAY DKD manuscript [shared]/Figures/HYP_ORA_comparison_1600x900.jpeg",
#        plot = hyp_comp_plot,width = 1600,height = 900,units = "px",scale = 2.5)
```

## RAPID

The proteins with the largest absolute effect size and adjusted p value $\leq 0.05$ for rapid eGFR decline were: `r toString(analytes$TargetFullName[match(top_rapid,analytes$AptName)])` (associated with genes `r toString(analytes$EntrezGeneSymbol[match(top_rapid,analytes$AptName)])`, respectively).

```{r message=FALSE,warning=FALSE,dpi=1200}
v_rapid <- volcano(top_rapid_df)
v_rapid <- v_rapid +
  ggtitle("Rapid eGFR Decline") +
  theme(plot.title = element_text(hjust = 0.5))
v_rapid
```

```{r message=FALSE,warning=FALSE,dpi=1200}
rapid_p <- pathway_analysis(de_genes_rapid)
rapid_p$updot
rapid_p$downdot

rapid_plot = rapid_p$updot + rapid_p$downdot + 
  plot_annotation(title = "Rapid eGFR decline",
                  theme = theme(plot.title = element_text(hjust = 0.5)))
# ggsave(filename = "/Users/timvigers/Library/CloudStorage/Dropbox/Work/Shared/TODAY/TODAY DKD manuscript [shared]/Figures/RAPID_ORA_1600x900.jpeg",plot = rapid_plot,
#        width = 1600,height = 900,units = "px",scale = 2.8)

rapid_p_comp <- pathway_comparison(list(
  Baseline = de_genes_rapid,
  Followup = de_genes_rapid_10
))
rapid_p_comp$updot
rapid_p_comp$downdot

rapid_comp_plot = rapid_p_comp$updot + rapid_p_comp$downdot + 
  plot_annotation(title = "Rapid eGFR decline",
                  theme = theme(plot.title = element_text(hjust = 0.5)))
# ggsave(filename = "/Users/timvigers/Library/CloudStorage/Dropbox/Work/Shared/TODAY/TODAY DKD manuscript [shared]/Figures/RAPID_ORA_comparison_1600x900.jpeg",
#        plot = rapid_comp_plot,width = 1600,height = 900,units = "px",scale = 2.5)
```

## HTN

The proteins with the largest absolute effect size and adjusted p value $\leq 0.05$ for hypertension were: `r toString(analytes$TargetFullName[match(top_htn,analytes$AptName)])` (associated with genes `r toString(analytes$EntrezGeneSymbol[match(top_htn,analytes$AptName)])`, respectively).

```{r message=FALSE,warning=FALSE,dpi=1200}
v_htn <- volcano(top_htn_df)
v_htn <- v_htn +
  ggtitle("Hypertension") +
  theme(plot.title = element_text(hjust = 0.5))
v_htn
```

```{r message=FALSE,warning=FALSE,dpi=1200}
htn_p <- pathway_analysis(de_genes_htn)
htn_p$updot
htn_p$downdot

htn_plot = htn_p$updot + htn_p$downdot + 
  plot_annotation(title = "Hypertension",
                  theme = theme(plot.title = element_text(hjust = 0.5)))
# ggsave(filename = "/Users/timvigers/Library/CloudStorage/Dropbox/Work/Shared/TODAY/TODAY DKD manuscript [shared]/Figures/HTN_ORA_1600x900.jpeg",plot = htn_plot,
#        width = 1600,height = 900,units = "px",scale = 2.8)

htn_p_comp <- pathway_comparison(list(
  Baseline = de_genes_htn,
  Followup = de_genes_htn_10
))
htn_p_comp$updot
htn_p_comp$downdot

htn_comp_plot = htn_p_comp$updot + htn_p_comp$downdot + 
  plot_annotation(title = "Hypertension",
                  theme = theme(plot.title = element_text(hjust = 0.5)))
# ggsave(filename = "/Users/timvigers/Library/CloudStorage/Dropbox/Work/Shared/TODAY/TODAY DKD manuscript [shared]/Figures/HTN_ORA_comparison_1600x900.jpeg",
#        plot = htn_comp_plot,width = 1600,height = 900,units = "px",scale = 2.5)
```

## HTN with SBP

The proteins with the largest absolute effect size and adjusted p value $\leq 0.05$ for hypertension (with SBP) were: `r toString(analytes$TargetFullName[match(top_htn_sbp,analytes$AptName)])` (associated with genes `r toString(analytes$EntrezGeneSymbol[match(top_htn_sbp,analytes$AptName)])`, respectively).

```{r message=FALSE,warning=FALSE,dpi=1200}
v_htn_sbp <- volcano(top_htn_sbp_df)
v_htn_sbp <- v_htn_sbp +
  ggtitle("Hypertension (with SBP)") +
  theme(plot.title = element_text(hjust = 0.5))
v_htn_sbp
```

```{r message=FALSE,warning=FALSE,dpi=1200}
htn_sbp_p <- pathway_analysis(de_genes_htn_sbp)
htn_sbp_p$updot
htn_sbp_p$downdot

htn_sbp_p_comp <- pathway_comparison(list(
  Baseline = de_genes_htn_sbp,
  Followup = de_genes_htn_sbp_10
))
htn_sbp_p_comp$updot
htn_sbp_p_comp$downdot
```

# Figure 1

```{r message=FALSE,warning=FALSE,dpi=1200}
fig1 <- (v_mic + v_mac + v_rapid) / (v_hyp + v_htn + plot_spacer())
# ggsave("/Users/timvigers/Library/CloudStorage/Dropbox/Work/Shared/TODAY/TODAY DKD manuscript [shared]/Figures/Figure_1_1600x900.jpeg",
#   units = "px",
#   width = 1600, height = 900, scale = 2, plot = fig1, device = "jpeg"
# )
```
