---
title: "Proteomics Pathways"
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
---

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
library(limma)
library(WebGestaltR)
library(knitr)
knitr::opts_chunk$set(echo = FALSE)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "E:/Petter Bjornstad/TODAY subaward"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/TODAY subaward"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/TODAY subaward"
}
knitr::opts_knit$set(root.dir = home_dir)
```

```{r}
# Map proteomics ID to sample ID
lead = read_excel("./Somalogic repository link/Omics-Petter Ancillary Samples at Colorado LEAD Center - Wash U.xlsx") %>% select(releaseid,SAMPLE_ID) %>% rename(current_label = SAMPLE_ID)
today = read_excel("./Somalogic repository link/Omics-Petter Ancillary Samples at NIDDK-TODAY - Wash U.xlsx") %>% select(releaseid,current_label)
today2 = read_excel("./Somalogic repository link/Omics-Petter Ancillary Samples at NIDDK-TODAY2 - Wash U.xlsx") %>% select(releaseid,current_label)
sample_ids = do.call(rbind,list(lead,today,today2))
colnames(sample_ids) = c("RELEASEID","SampleDescription")
# Co-morbidities?
comorb = read.csv("./Clinical data/COMORB.csv")
comorb = left_join(sample_ids,comorb,by = "RELEASEID")
# Cleanup
rm(lead,today,today2,sample_ids)
# Import proteomics
load("./Somalogic data raw/soma.Rdata")
# Just proteins (in columns) - should end up with 7596
rownames(soma) = soma$SampleDescription
soma = soma %>% select(seq.10000.28:seq.9999.1)
# Want samples in columns for leapR and log transform
soma = t(log(soma))
# Remove those missing comorb data
comorb = comorb[comorb$SampleDescription %in% colnames(soma),]
# Lookup gene names for SomaLogic proteins - some missing
load("./Somalogic data raw/analytes.Rdata")
```

# Enrichment

Enrichment was performed using the GSEA method [] as implemented in WebGestaltR [], with Reactome pathways [] as the functional database. Genes of interest (based on converting SomaLogic protein IDs to UniProt) were ranked in descending order by logFC between groups divided by the empirical Bayes estimate of standard deviation ($\frac{\beta_g}{\sigma_g}$) as calculated in the `limma` R package []. Some SomaLogic IDs are associated with more than one UniProt ID, in which case the first UniProt ID was used.

```{r}
gsea = function(limma_fit,output_name = "MAC",coef_name = "MAC",cores = 8){
  # Get genes of interest, sorting limma results by fold change (decreasing), 
  # standardized by estimated SD of the gene
  interest = data.frame(limma_fit$coefficients[,coef_name]/sqrt(limma_fit$s2.post))
  interest = interest %>% rownames_to_column(.)
  colnames(interest) = c("V1","V2")
  interest$V1 = sub("\\|.*","",analytes$UniProt[match(interest$V1,analytes$AptName)])
  interest$V2 = as.numeric(interest$V2)
  # Webgestalt
  WebGestaltR(enrichMethod="GSEA", organism="hsapiens",
              enrichDatabase="pathway_Reactome", interestGene = interest,
              interestGeneType="uniprotswissprot", sigMethod="fdr",fdrThr = 0.1,
              outputDirectory="./Results/Enrichment",fdrMethod = "BH",
              projectName = paste0(output_name,"_GSEA"),
              nThreads = cores)
}
```

## MAC vs. no MAC comparison

```{r warning=FALSE,message=FALSE}
m = model.matrix(~MAC,comorb)
m[,2] = 1 - m[,2]
fit = lmFit(soma,m)
fit <- eBayes(fit)
gsea(fit)
```


