---
title: "TODAY adenine analyses"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include=FALSE}
library(limma)
library(dplyr)
library(caret)
library(purrr)
library(multtest)
library(openxlsx)
library(tableone)
library(EnhancedVolcano)
library(knitr)
library(arsenal)
library(emmeans)
library(survival)
library(broom)
library(skimr)

knitr::opts_chunk$set(echo = FALSE,warning = FALSE)

knitr::opts_chunk$set(echo = FALSE)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "E:/Petter Bjornstad/TODAY subaward"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/TODAY subaward/"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = ""
}
knitr::opts_knit$set(root.dir = home_dir)
```

```{r, include=FALSE}
# load urine and plasma data
load("./Metabolomic data/urine.Rdata")
load("./Metabolomic data/plasma.Rdata")
urine$Date.Drawn <- as.Date(urine$Date.Drawn,format = "%m/%d/%Y")
plasma$Date.Drawn <- as.Date(plasma$Date.Drawn,format = "%m/%d/%Y")

# load comorbidity data
load("./Clinical data/comorb.Rdata")

# load baseline risk factors
load("./Clinical data/TODAY/baserisk.Rdata")

# take only the baseline samples
# can we just take the earliest or is it possible someone would have a follow-up sample but not baseline?
# probably can take the first and then check the years to make sure compatible with TODAY
urine_base <- urine %>% arrange(releaseid,Date.Drawn) %>% group_by(releaseid) %>% filter(row_number()==1)
plasma_base <- plasma %>% arrange(releaseid,Date.Drawn) %>% group_by(releaseid) %>% filter(row_number()==1)

# merge in comorb data
urine_base <- merge(urine_base, comorb, by="releaseid",all.x=T, all.y=F)
plasma_base <- merge(plasma_base, comorb, by="releaseid",all.x=T, all.y=F)

# merge in baseline risk factors
urine_base <- merge(urine_base, baserisk, by="releaseid",all.x=T, all.y=F)
plasma_base <- merge(plasma_base, baserisk, by="releaseid",all.x=T, all.y=F)

# note that plasma adenine has no variability
# for the later analyses (i.e., after adenine, definitely want to check for low variability)
# for the rest of these adenine analyses, will focus on urine results only

# create tertiles of adenine
urine_base <- urine_base %>%
  mutate(adenine_tertiles = ntile(Adenine.in.uM.mM.Creatinine, 3)) %>%
  mutate(adenine_tertiles = if_else(adenine_tertiles == 1, 'Low', if_else(adenine_tertiles == 3, 'High', "Medium"))) %>%
  arrange(Adenine.in.uM.mM.Creatinine)
urine_base$adenine_tertiles_num <- ifelse(urine_base$adenine_tertiles=="High",1,
                                          ifelse(urine_base$adenine_tertiles=="Low",0,NA))
urine_base$adenine_tertiles <- ifelse(urine_base$adenine_tertiles=="Medium",NA,urine_base$adenine_tertiles)
```

# Questions/comments

- There was no variability in plasma adenine, so these analyses focus on urine adenine results only.

# Data description

```{r, include=TRUE}
skim(urine_base)
```

# Results using adenine as a continuous variable

## Urine adenine in participants stratified by complication status

### Microalbuminuria

```{r, include=FALSE}
mic_table <- tableby(data=urine_base, MIC ~ Adenine.in.uM.mM.Creatinine,
                     numeric.test="kwt",numeric.stats=c("Nmiss","median","q1q3"))
```

```{r echo=FALSE, include=TRUE, results='asis'}
summary(mic_table)
```

### Macroalbuminuria

```{r, include=FALSE}
# urine adenine by macroalbumuniuria status
mac_table <- tableby(data=urine_base, MAC ~ Adenine.in.uM.mM.Creatinine,
                     numeric.test="kwt",numeric.stats=c("Nmiss","median","q1q3"))
```

```{r echo=FALSE, include=TRUE, results='asis'}
summary(mac_table)
```

### Hyperfiltration

```{r, include=FALSE}
# urine adenine by hyperfiltration status
hyper_table <- tableby(data=urine_base, HYP ~ Adenine.in.uM.mM.Creatinine,
                     numeric.test="kwt",numeric.stats=c("Nmiss","median","q1q3"))
```

```{r echo=FALSE, include=TRUE, results='asis'}
summary(hyper_table)
```

### Rapid eGFR decline

```{r, include=FALSE}
# urine adenine by rapid eGFR decline status
rapid_table <- tableby(data=urine_base, RAPID ~ Adenine.in.uM.mM.Creatinine,
                     numeric.test="kwt",numeric.stats=c("Nmiss","median","q1q3"))
```

```{r echo=FALSE, include=TRUE, results='asis'}
summary(rapid_table)
```

### Hypertension

```{r, include=FALSE}
# urine adenine by hypertension status
htn_table <- tableby(data=urine_base, HTN ~ Adenine.in.uM.mM.Creatinine,
                     numeric.test="kwt",numeric.stats=c("Nmiss","median","q1q3"))
```

```{r echo=FALSE, include=TRUE, results='asis'}
summary(htn_table)
```

### Neuropathy

```{r, include=FALSE}
# urine adenine by neuropathy status
neuro_table <- tableby(data=urine_base, NEURO ~ Adenine.in.uM.mM.Creatinine,
                     numeric.test="kwt",numeric.stats=c("Nmiss","median","q1q3"))
```

```{r echo=FALSE, include=TRUE, results='asis'}
summary(neuro_table)
```

### Retinopathy

```{r, include=FALSE}
# urine adenine by retinopathy status
retino_table <- tableby(data=urine_base, RETINO ~ Adenine.in.uM.mM.Creatinine,
                     numeric.test="kwt",numeric.stats=c("Nmiss","median","q1q3"))
```

```{r echo=FALSE, include=TRUE, results='asis'}
summary(retino_table)
```

## Adjusted linear models

### Microalbuminuria

```{r, include=FALSE}
MIC_mod <- lm(Adenine.in.uM.mM.Creatinine~MIC+HbA1c+log_trig+sbp+si_1_ins0, urine_base)
MIC_means <- emmeans(MIC_mod, "MIC")
```

```{r echo=FALSE, include=TRUE, comment=""}
summary(MIC_mod)

kable(MIC_means)
```

### Macroalbuminuria

```{r, include=FALSE}
MAC_mod <- lm(Adenine.in.uM.mM.Creatinine~MAC+HbA1c+log_trig+sbp+si_1_ins0, urine_base)
MAC_means <- emmeans(MAC_mod, "MAC")
```

```{r echo=FALSE, include=TRUE, comment=""}
summary(MAC_mod)

kable(MAC_means)
```

### Hyperfiltration

```{r, include=FALSE}
HYP_mod <- lm(Adenine.in.uM.mM.Creatinine~HYP+HbA1c+log_trig+sbp+si_1_ins0, urine_base)
HYP_means <- emmeans(HYP_mod, "HYP")
```

```{r echo=FALSE, include=TRUE, comment=""}
summary(HYP_mod)

kable(HYP_means)
```

### Rapid eGFR decline

```{r, include=FALSE}
rapid_mod <- lm(Adenine.in.uM.mM.Creatinine~RAPID+HbA1c+log_trig+sbp+si_1_ins0, urine_base)
rapid_means <- emmeans(rapid_mod, "RAPID")
```

```{r echo=FALSE, include=TRUE, comment=""}
summary(rapid_mod)

kable(rapid_means)
```

### Hypertension

```{r, include=FALSE}
htn_mod <- lm(Adenine.in.uM.mM.Creatinine~HTN+HbA1c+log_trig+si_1_ins0, urine_base)
htn_means <- emmeans(htn_mod, "HTN")
```

```{r echo=FALSE, include=TRUE, comment=""}
summary(htn_mod)

kable(htn_means)
```

### Neuropathy

```{r, include=FALSE}
neuro_mod <- lm(Adenine.in.uM.mM.Creatinine~NEURO+HbA1c+log_trig+sbp+si_1_ins0, urine_base)
neuro_means <- emmeans(neuro_mod, "NEURO")
```

```{r echo=FALSE, include=TRUE, comment=""}
summary(neuro_mod)

kable(neuro_means)
```

### Retinopathy

```{r, include=FALSE}
retino_mod <- lm(Adenine.in.uM.mM.Creatinine~RETINO+HbA1c+log_trig+sbp+si_1_ins0, urine_base)
retino_means <- emmeans(retino_mod, "RETINO")
```

```{r echo=FALSE, include=TRUE, comment=""}
summary(retino_mod)
kable(retino_means)
```

## Cox models of time to complications

### Microalbuminuria

```{r, include=FALSE}
mic_cox <- coxph(data = urine_base, formula = Surv(DAYSTOMIC, MIC)~Adenine.in.uM.mM.Creatinine)
mic_cox_tidy <- tidy(mic_cox, exponentiate=TRUE, conf.int = TRUE)
colnames(mic_cox_tidy) <- c("","HR","SE","Statistic","P-value","Lower bound 95% CI HR","Upper bound 95% CI HR")
```

```{r echo=FALSE, include=TRUE, comment=""}
kable(mic_cox_tidy)
```

### Macroalbuminuria

```{r, include=FALSE}
mac_cox <- coxph(data = urine_base, formula = Surv(DAYSTOMAC, MAC)~Adenine.in.uM.mM.Creatinine)
mac_cox_tidy <- tidy(mac_cox, exponentiate=TRUE, conf.int = TRUE)
colnames(mac_cox_tidy) <- c("","HR","SE","Statistic","P-value","Lower bound 95% CI HR","Upper bound 95% CI HR")
```

```{r echo=FALSE, include=TRUE, comment=""}
kable(mac_cox_tidy)
```

### Hyperfiltration

```{r, include=FALSE}
hyp_cox <- coxph(data = urine_base, formula = Surv(DAYSTOHYP, HYP)~Adenine.in.uM.mM.Creatinine)
hyp_cox_tidy <- tidy(hyp_cox, exponentiate=TRUE, conf.int = TRUE)
colnames(hyp_cox_tidy) <- c("","HR","SE","Statistic","P-value","Lower bound 95% CI HR","Upper bound 95% CI HR")
```

```{r echo=FALSE, include=TRUE, comment=""}
kable(hyp_cox_tidy)
```

### Rapid eGFR decline

```{r, include=FALSE}
rapid_cox <- coxph(data = urine_base, formula = Surv(DAYSTORAPID, RAPID)~Adenine.in.uM.mM.Creatinine)
rapid_cox_tidy <- tidy(rapid_cox, exponentiate=TRUE, conf.int = TRUE)
colnames(rapid_cox_tidy) <- c("","HR","SE","Statistic","P-value","Lower bound 95% CI HR","Upper bound 95% CI HR")
```

```{r echo=FALSE, include=TRUE, comment=""}
kable(rapid_cox_tidy)
```

### Hypertension

```{r, include=FALSE}
htn_cox <- coxph(data = urine_base, formula = Surv(DAYSTOHTN, HTN)~Adenine.in.uM.mM.Creatinine)
htn_cox_tidy <- tidy(htn_cox, exponentiate=TRUE, conf.int = TRUE)
colnames(htn_cox_tidy) <- c("","HR","SE","Statistic","P-value","Lower bound 95% CI HR","Upper bound 95% CI HR")
```

```{r echo=FALSE, include=TRUE, comment=""}
kable(htn_cox_tidy)
```

### Neuropathy

```{r, include=FALSE}
neuro_cox <- coxph(data = urine_base, formula = Surv(DAYSTONEURO, NEURO)~Adenine.in.uM.mM.Creatinine)
neuro_cox_tidy <- tidy(neuro_cox, exponentiate=TRUE, conf.int = TRUE)
colnames(neuro_cox_tidy) <- c("","HR","SE","Statistic","P-value","Lower bound 95% CI HR","Upper bound 95% CI HR")
```

```{r echo=FALSE, include=TRUE, comment=""}
kable(neuro_cox_tidy)
```

### Retinopathy

```{r, include=FALSE}
retino_cox <- coxph(data = urine_base, formula = Surv(DAYSTORETINO, RETINO)~Adenine.in.uM.mM.Creatinine)
retino_cox_tidy <- tidy(retino_cox, exponentiate=TRUE, conf.int = TRUE)
colnames(retino_cox_tidy) <- c("","HR","SE","Statistic","P-value","Lower bound 95% CI HR","Upper bound 95% CI HR")
```

```{r echo=FALSE, include=TRUE, comment=""}
kable(retino_cox_tidy)
```

## Adjusted Cox models of time to complications

### Microalbuminuria

```{r, include=FALSE}
mic_cox <- coxph(data = urine_base, formula = Surv(DAYSTOMIC, MIC)~Adenine.in.uM.mM.Creatinine+HbA1c+log_trig+sbp+si_1_ins0)
mic_cox_tidy <- tidy(mic_cox, exponentiate=TRUE, conf.int = TRUE)
colnames(mic_cox_tidy) <- c("","HR","SE","Statistic","P-value","Lower bound 95% CI HR","Upper bound 95% CI HR")
```

```{r echo=FALSE, include=TRUE, comment=""}
kable(mic_cox_tidy)
```

### Macroalbuminuria

```{r, include=FALSE}
mac_cox <- coxph(data = urine_base, formula = Surv(DAYSTOMAC, MAC)~Adenine.in.uM.mM.Creatinine+HbA1c+log_trig+sbp+si_1_ins0)
mac_cox_tidy <- tidy(mac_cox, exponentiate=TRUE, conf.int = TRUE)
colnames(mac_cox_tidy) <- c("","HR","SE","Statistic","P-value","Lower bound 95% CI HR","Upper bound 95% CI HR")
```

```{r echo=FALSE, include=TRUE, comment=""}
kable(mac_cox_tidy)
```

### Hyperfiltration

```{r, include=FALSE}
hyp_cox <- coxph(data = urine_base, formula = Surv(DAYSTOHYP, HYP)~Adenine.in.uM.mM.Creatinine+HbA1c+log_trig+sbp+si_1_ins0)
hyp_cox_tidy <- tidy(hyp_cox, exponentiate=TRUE, conf.int = TRUE)
colnames(hyp_cox_tidy) <- c("","HR","SE","Statistic","P-value","Lower bound 95% CI HR","Upper bound 95% CI HR")
```

```{r echo=FALSE, include=TRUE, comment=""}
kable(hyp_cox_tidy)
```

### Rapid eGFR decline

```{r, include=FALSE}
rapid_cox <- coxph(data = urine_base, formula = Surv(DAYSTORAPID, RAPID)~Adenine.in.uM.mM.Creatinine+HbA1c+log_trig+sbp+si_1_ins0)
rapid_cox_tidy <- tidy(rapid_cox, exponentiate=TRUE, conf.int = TRUE)
colnames(rapid_cox_tidy) <- c("","HR","SE","Statistic","P-value","Lower bound 95% CI HR","Upper bound 95% CI HR")
```

```{r echo=FALSE, include=TRUE, comment=""}
kable(rapid_cox_tidy)
```

### Hypertension

```{r, include=FALSE}
htn_cox <- coxph(data = urine_base, formula = Surv(DAYSTOHTN, HTN)~Adenine.in.uM.mM.Creatinine+HbA1c+log_trig+sbp+si_1_ins0)
htn_cox_tidy <- tidy(htn_cox, exponentiate=TRUE, conf.int = TRUE)
colnames(htn_cox_tidy) <- c("","HR","SE","Statistic","P-value","Lower bound 95% CI HR","Upper bound 95% CI HR")
```

```{r echo=FALSE, include=TRUE, comment=""}
kable(htn_cox_tidy)
```

### Neuropathy

```{r, include=FALSE}
neuro_cox <- coxph(data = urine_base, formula = Surv(DAYSTONEURO, NEURO)~Adenine.in.uM.mM.Creatinine+HbA1c+log_trig+sbp+si_1_ins0)
neuro_cox_tidy <- tidy(neuro_cox, exponentiate=TRUE, conf.int = TRUE)
colnames(neuro_cox_tidy) <- c("","HR","SE","Statistic","P-value","Lower bound 95% CI HR","Upper bound 95% CI HR")
```

```{r echo=FALSE, include=TRUE, comment=""}
kable(neuro_cox_tidy)
```

### Retinopathy

```{r, include=FALSE}
retino_cox <- coxph(data = urine_base, formula = Surv(DAYSTORETINO, RETINO)~Adenine.in.uM.mM.Creatinine+HbA1c+log_trig+sbp+si_1_ins0)
retino_cox_tidy <- tidy(retino_cox, exponentiate=TRUE, conf.int = TRUE)
colnames(retino_cox_tidy) <- c("","HR","SE","Statistic","P-value","Lower bound 95% CI HR","Upper bound 95% CI HR")
```

```{r echo=FALSE, include=TRUE, comment=""}
kable(retino_cox_tidy)
```

# Results using comparing upper and lower tertiles of adenine

## Urine adenine in participants stratified by complication status

### Microalbuminuria

```{r, include=FALSE}
mic_table <- tableby(data=urine_base, MIC ~ adenine_tertiles,
                     numeric.test="kwt",numeric.stats=c("Nmiss","median","q1q3"))
```

```{r echo=FALSE, include=TRUE, results='asis'}
summary(mic_table)
```

### Macroalbuminuria

```{r, include=FALSE}
# urine adenine by macroalbumuniuria status
mac_table <- tableby(data=urine_base, MAC ~ adenine_tertiles,
                     numeric.test="kwt",numeric.stats=c("Nmiss","median","q1q3"))
```

```{r echo=FALSE, include=TRUE, results='asis'}
summary(mac_table)
```

### Hyperfiltration

```{r, include=FALSE}
# urine adenine by hyperfiltration status
hyper_table <- tableby(data=urine_base, HYP ~ adenine_tertiles,
                     numeric.test="kwt",numeric.stats=c("Nmiss","median","q1q3"))
```

```{r echo=FALSE, include=TRUE, results='asis'}
summary(hyper_table)
```

### Rapid eGFR decline

```{r, include=FALSE}
# urine adenine by rapid eGFR decline status
rapid_table <- tableby(data=urine_base, RAPID ~ adenine_tertiles,
                     numeric.test="kwt",numeric.stats=c("Nmiss","median","q1q3"))
```

```{r echo=FALSE, include=TRUE, results='asis'}
summary(rapid_table)
```

### Hypertension

```{r, include=FALSE}
# urine adenine by hypertension status
htn_table <- tableby(data=urine_base, HTN ~ adenine_tertiles,
                     numeric.test="kwt",numeric.stats=c("Nmiss","median","q1q3"))
```

```{r echo=FALSE, include=TRUE, results='asis'}
summary(htn_table)
```

### Neuropathy

```{r, include=FALSE}
# urine adenine by neuropathy status
neuro_table <- tableby(data=urine_base, NEURO ~ adenine_tertiles,
                     numeric.test="kwt",numeric.stats=c("Nmiss","median","q1q3"))
```

```{r echo=FALSE, include=TRUE, results='asis'}
summary(neuro_table)
```

### Retinopathy

```{r, include=FALSE}
# urine adenine by retinopathy status
retino_table <- tableby(data=urine_base, RETINO ~ adenine_tertiles,
                     numeric.test="kwt",numeric.stats=c("Nmiss","median","q1q3"))
```

```{r echo=FALSE, include=TRUE, results='asis'}
summary(retino_table)
```

## Adjusted linear models

### Microalbuminuria

```{r, include=FALSE}
MIC_mod <- glm(adenine_tertiles_num~MIC+HbA1c+log_trig+sbp+si_1_ins0, urine_base, family = "binomial")
```

```{r echo=FALSE, include=TRUE, comment=""}
summary(MIC_mod)
```

### Macroalbuminuria

```{r, include=FALSE}
MAC_mod <- lm(adenine_tertiles_num~MAC+HbA1c+log_trig+sbp+si_1_ins0, urine_base)
```

```{r echo=FALSE, include=TRUE, comment=""}
summary(MAC_mod)
```

### Hyperfiltration

```{r, include=FALSE}
HYP_mod <- lm(adenine_tertiles_num~HYP+HbA1c+log_trig+sbp+si_1_ins0, urine_base)
```

```{r echo=FALSE, include=TRUE, comment=""}
summary(HYP_mod)
```

### Rapid eGFR decline

```{r, include=FALSE}
rapid_mod <- lm(adenine_tertiles_num~RAPID+HbA1c+log_trig+sbp+si_1_ins0, urine_base)
```

```{r echo=FALSE, include=TRUE, comment=""}
summary(rapid_mod)
```

### Hypertension

```{r, include=FALSE}
htn_mod <- lm(adenine_tertiles_num~HTN+HbA1c+log_trig+si_1_ins0, urine_base)
```

```{r echo=FALSE, include=TRUE, comment=""}
summary(htn_mod)
```

### Neuropathy

```{r, include=FALSE}
neuro_mod <- lm(adenine_tertiles_num~NEURO+HbA1c+log_trig+sbp+si_1_ins0, urine_base)
```

```{r echo=FALSE, include=TRUE, comment=""}
summary(neuro_mod)
```

### Retinopathy

```{r, include=FALSE}
retino_mod <- lm(adenine_tertiles_num~RETINO+HbA1c+log_trig+sbp+si_1_ins0, urine_base)
```

```{r echo=FALSE, include=TRUE, comment=""}
summary(retino_mod)
```

## Cox models of time to complications

### Microalbuminuria

```{r, include=FALSE}
mic_cox <- coxph(data = urine_base, formula = Surv(DAYSTOMIC, MIC)~adenine_tertiles)
mic_cox_tidy <- tidy(mic_cox, exponentiate=TRUE, conf.int = TRUE)
colnames(mic_cox_tidy) <- c("","HR","SE","Statistic","P-value","Lower bound 95% CI HR","Upper bound 95% CI HR")
```

```{r echo=FALSE, include=TRUE, comment=""}
kable(mic_cox_tidy)
```

### Macroalbuminuria

```{r, include=FALSE}
mac_cox <- coxph(data = urine_base, formula = Surv(DAYSTOMAC, MAC)~adenine_tertiles)
mac_cox_tidy <- tidy(mac_cox, exponentiate=TRUE, conf.int = TRUE)
colnames(mac_cox_tidy) <- c("","HR","SE","Statistic","P-value","Lower bound 95% CI HR","Upper bound 95% CI HR")
```

```{r echo=FALSE, include=TRUE, comment=""}
kable(mac_cox_tidy)
```

### Hyperfiltration

```{r, include=FALSE}
hyp_cox <- coxph(data = urine_base, formula = Surv(DAYSTOHYP, HYP)~adenine_tertiles)
hyp_cox_tidy <- tidy(hyp_cox, exponentiate=TRUE, conf.int = TRUE)
colnames(hyp_cox_tidy) <- c("","HR","SE","Statistic","P-value","Lower bound 95% CI HR","Upper bound 95% CI HR")
```

```{r echo=FALSE, include=TRUE, comment=""}
kable(hyp_cox_tidy)
```

### Rapid eGFR decline

```{r, include=FALSE}
rapid_cox <- coxph(data = urine_base, formula = Surv(DAYSTORAPID, RAPID)~adenine_tertiles)
rapid_cox_tidy <- tidy(rapid_cox, exponentiate=TRUE, conf.int = TRUE)
colnames(rapid_cox_tidy) <- c("","HR","SE","Statistic","P-value","Lower bound 95% CI HR","Upper bound 95% CI HR")
```

```{r echo=FALSE, include=TRUE, comment=""}
kable(rapid_cox_tidy)
```

### Hypertension

```{r, include=FALSE}
htn_cox <- coxph(data = urine_base, formula = Surv(DAYSTOHTN, HTN)~adenine_tertiles)
htn_cox_tidy <- tidy(htn_cox, exponentiate=TRUE, conf.int = TRUE)
colnames(htn_cox_tidy) <- c("","HR","SE","Statistic","P-value","Lower bound 95% CI HR","Upper bound 95% CI HR")
```

```{r echo=FALSE, include=TRUE, comment=""}
kable(htn_cox_tidy)
```

### Neuropathy

```{r, include=FALSE}
neuro_cox <- coxph(data = urine_base, formula = Surv(DAYSTONEURO, NEURO)~adenine_tertiles)
neuro_cox_tidy <- tidy(neuro_cox, exponentiate=TRUE, conf.int = TRUE)
colnames(neuro_cox_tidy) <- c("","HR","SE","Statistic","P-value","Lower bound 95% CI HR","Upper bound 95% CI HR")
```

```{r echo=FALSE, include=TRUE, comment=""}
kable(neuro_cox_tidy)
```

### Retinopathy

```{r, include=FALSE}
retino_cox <- coxph(data = urine_base, formula = Surv(DAYSTORETINO, RETINO)~adenine_tertiles)
retino_cox_tidy <- tidy(retino_cox, exponentiate=TRUE, conf.int = TRUE)
colnames(retino_cox_tidy) <- c("","HR","SE","Statistic","P-value","Lower bound 95% CI HR","Upper bound 95% CI HR")
```

```{r echo=FALSE, include=TRUE, comment=""}
kable(retino_cox_tidy)
```

## Adjusted Cox models of time to complications

### Microalbuminuria

```{r, include=FALSE}
mic_cox <- coxph(data = urine_base, formula = Surv(DAYSTOMIC, MIC)~adenine_tertiles+HbA1c+log_trig+sbp+si_1_ins0)
mic_cox_tidy <- tidy(mic_cox, exponentiate=TRUE, conf.int = TRUE)
colnames(mic_cox_tidy) <- c("","HR","SE","Statistic","P-value","Lower bound 95% CI HR","Upper bound 95% CI HR")
```

```{r echo=FALSE, include=TRUE, comment=""}
kable(mic_cox_tidy)
```

### Macroalbuminuria

```{r, include=FALSE}
mac_cox <- coxph(data = urine_base, formula = Surv(DAYSTOMAC, MAC)~adenine_tertiles+HbA1c+log_trig+sbp+si_1_ins0)
mac_cox_tidy <- tidy(mac_cox, exponentiate=TRUE, conf.int = TRUE)
colnames(mac_cox_tidy) <- c("","HR","SE","Statistic","P-value","Lower bound 95% CI HR","Upper bound 95% CI HR")
```

```{r echo=FALSE, include=TRUE, comment=""}
kable(mac_cox_tidy)
```

### Hyperfiltration

```{r, include=FALSE}
hyp_cox <- coxph(data = urine_base, formula = Surv(DAYSTOHYP, HYP)~adenine_tertiles+HbA1c+log_trig+sbp+si_1_ins0)
hyp_cox_tidy <- tidy(hyp_cox, exponentiate=TRUE, conf.int = TRUE)
colnames(hyp_cox_tidy) <- c("","HR","SE","Statistic","P-value","Lower bound 95% CI HR","Upper bound 95% CI HR")
```

```{r echo=FALSE, include=TRUE, comment=""}
kable(hyp_cox_tidy)
```

### Rapid eGFR decline

```{r, include=FALSE}
rapid_cox <- coxph(data = urine_base, formula = Surv(DAYSTORAPID, RAPID)~adenine_tertiles+HbA1c+log_trig+sbp+si_1_ins0)
rapid_cox_tidy <- tidy(rapid_cox, exponentiate=TRUE, conf.int = TRUE)
colnames(rapid_cox_tidy) <- c("","HR","SE","Statistic","P-value","Lower bound 95% CI HR","Upper bound 95% CI HR")
```

```{r echo=FALSE, include=TRUE, comment=""}
kable(rapid_cox_tidy)
```

### Hypertension

```{r, include=FALSE}
htn_cox <- coxph(data = urine_base, formula = Surv(DAYSTOHTN, HTN)~adenine_tertiles+HbA1c+log_trig+sbp+si_1_ins0)
htn_cox_tidy <- tidy(htn_cox, exponentiate=TRUE, conf.int = TRUE)
colnames(htn_cox_tidy) <- c("","HR","SE","Statistic","P-value","Lower bound 95% CI HR","Upper bound 95% CI HR")
```

```{r echo=FALSE, include=TRUE, comment=""}
kable(htn_cox_tidy)
```

### Neuropathy

```{r, include=FALSE}
neuro_cox <- coxph(data = urine_base, formula = Surv(DAYSTONEURO, NEURO)~adenine_tertiles+HbA1c+log_trig+sbp+si_1_ins0)
neuro_cox_tidy <- tidy(neuro_cox, exponentiate=TRUE, conf.int = TRUE)
colnames(neuro_cox_tidy) <- c("","HR","SE","Statistic","P-value","Lower bound 95% CI HR","Upper bound 95% CI HR")
```

```{r echo=FALSE, include=TRUE, comment=""}
kable(neuro_cox_tidy)
```

### Retinopathy

```{r, include=FALSE}
retino_cox <- coxph(data = urine_base, formula = Surv(DAYSTORETINO, RETINO)~adenine_tertiles+HbA1c+log_trig+sbp+si_1_ins0)
retino_cox_tidy <- tidy(retino_cox, exponentiate=TRUE, conf.int = TRUE)
colnames(retino_cox_tidy) <- c("","HR","SE","Statistic","P-value","Lower bound 95% CI HR","Upper bound 95% CI HR")
```

```{r echo=FALSE, include=TRUE, comment=""}
kable(retino_cox_tidy)
```