---
title: "TODAY Somalogic follow-up analyses"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include=FALSE}
library(tidyverse)
library(SomaDataIO)
library(parallel)
library(lmerTest)
library(emmeans)
library(DT)
knitr::opts_chunk$set(echo = FALSE,warning = FALSE)
knitr::opts_chunk$set(echo = FALSE)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "E:/Petter Bjornstad/TODAY subaward"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/TODAY subaward/"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/TODAY subaward"
}
knitr::opts_knit$set(root.dir = home_dir)
```

```{r}
# Import
load("./Somalogic data raw/soma.Rdata")
load("./Somalogic data raw/analytes.Rdata")
load("./Clinical data/comorb.Rdata")
# Merge soma and comorbidities
soma = left_join(soma,comorb,by = "releaseid")
# Filter unnecessary columns
soma = soma %>% arrange(releaseid,Date.Drawn) %>% group_by(releaseid) %>% 
  mutate(timepoint = row_number()) %>%
  select(releaseid,timepoint,Date.Drawn,MAC,MIC,HYP,RAPID,HTN,NEURO,RETINO,
         seq.10000.28:seq.9999.1) %>% ungroup()
# Create matrices for limma
Y = soma %>% select(seq.10000.28:seq.9999.1) %>% t(.)
X = soma %>% select(releaseid,timepoint,MAC:RETINO)
X = data.frame(lapply(X,as.factor))
# Log transform proteins
Y = log(Y)
```

# LMM results

```{r}
mod_fit = function(outcomes = rownames(Y),predictor,df = soma){
  # Parallel processing
  cl = makeCluster(detectCores()*0.5,type = "FORK")
  # Fit all the models, and pull out those with significant interaction term
  models = parLapply(cl,outcomes, function(p){
    # Formula and model fit
    f = as.formula(paste0("log(",p,")","~factor(",predictor,")*factor(timepoint)+(1|releaseid)"))
    mod = suppressMessages(lmer(f,data = df)) # Some singular fit warnings here, but okay to ignore per Laura
    # Calculate the difference in change over time between two groups
    contr = (c(0,0,0,1) - c(0,1,0,0)) - (c(0,0,1,0) - c(1,0,0,0))
    means_f = as.formula(paste0("~",predictor,"+timepoint"))
    emm.mod = emmeans(mod,specs = means_f,contr = list("Diff. in logFC" = contr))
    means = data.frame(emm.mod$emmeans)
    comp = data.frame(emm.mod$contrasts)
    return(comp)
  })
  stopCluster(cl)
  # Combine the fold change comparisons
  results = data.frame(do.call(rbind,models))
  rownames(results) = outcomes
  results$contrast = NULL
  colnames(results)[1] = "Diff. in logFC"
  results$p.value.adj = p.adjust(results$p.value,method = "BH")
  results = results[order(results$p.value),]
  # Plot
  p = EnhancedVolcano(results,lab = rownames(results),pCutoff = 0.05,
                      x="Diff. in logFC",y="p.value")
  # Return
  return(list("table" = round(results,3),
              "plot" = p))
}
```

Protein data were log transformed prior to analysis. For each protein, we fit a linear mixed effects model with random intercept for participant. Next, we calculated the estimated marginal means (EMMs) for each group (e.g. MAC vs. no MAC) at baseline and follow, and compared the average change over time for each group. These differences are of the form:

$$
(log(\mu_{12})-log(\mu_{11}))-(log(\mu_{02})-log(\mu_{01}))
$$

where the first subscript indicates group and the second indicates timepoint. Thus, the estimate is the difference in log fold change (log FC) over time between groups. P values for this comparison were adjusted for multiple comparisons using the Benjamini-Hochberg method, and the tables below include all proteins with a nominally significant (p $\leq$ 0.05) difference between groups.

## MAC

```{r}
mac_res = mod_fit(predictor = "MAC")
mac_res$plot
mac_res = mac_res$table
datatable(mac_res[mac_res$p.value <= 0.05,])
```

## MIC

```{r}
mic_res = mod_fit(predictor = "MIC")
datatable(mic_res[mic_res$p.value <= 0.05,])
```

## HYP

```{r}
hyp_res = mod_fit(predictor = "HYP")
datatable(hyp_res[hyp_res$p.value <= 0.05,])
```

## RAPID

```{r}
rapid_res = mod_fit(predictor = "RAPID")
datatable(rapid_res[rapid_res$p.value <= 0.05,])
```

## HTN

```{r}
htn_res = mod_fit(predictor = "HTN")
datatable(htn_res[htn_res$p.value <= 0.05,])
```

## NEURO

```{r}
neuro_res = mod_fit(predictor = "NEURO")
datatable(neuro_res[neuro_res$p.value <= 0.05,])
```

## RETINO

```{r}
retino_res = mod_fit(predictor = "RETINO")
datatable(retino_res[retino_res$p.value <= 0.05,])
```
