---
title: "TODAY Somalogic baseline sPLS-DA"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include=FALSE}
library(SomaDataIO)
library(limma)
library(dplyr)
library(caret)
library(purrr)
library(multtest)
library(openxlsx)
library(tableone)
library(EnhancedVolcano)
library(knitr)
library(survival)
library(broom)
library(mixOmics)

knitr::opts_chunk$set(echo = FALSE,warning = FALSE)

knitr::opts_chunk$set(echo = FALSE)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "E:/Petter Bjornstad/TODAY subaward"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/TODAY subaward/"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = ""
}
knitr::opts_knit$set(root.dir = home_dir)
```

```{r, include=FALSE}
# load somalogic data, with QC samples already excluded
load("./Somalogic data raw/soma.Rdata")

# load analyte info
load("./Somalogic data raw/analytes.Rdata")

# load comorbidity data
load("./Clinical data/comorb.Rdata")

# take only the baseline soma samples
# can we just take the earliest or is it possible someone would have a follow-up sample but not baseline?
# probably can take the first and then check the years to make sure compatible with TODAY
base <- soma %>% arrange(releaseid,Date.Drawn) %>% group_by(releaseid) %>% filter(row_number()==1)

# these 3 release IDs have first sample after end of recruitment, so they must be missing the baseline visit
base <- base %>% filter(!releaseid %in% c("65-85903","65-47984","65-25901"))

# merge in nephropathy data
base <- merge(base, comorb, by="releaseid",all.x=T, all.y=F)

# identify columns corresponding to proteins
#is_seq <- function(.x) grepl("^seq\\.[0-9]{4}", .x) # regex for analytes
is_seq <- function(.x) grepl("seq", .x)
seq <- is_seq(names(base))

# convert to numeric
base[,seq] <- apply(base[,seq],2,as.numeric)

# are there proteins with low variability?
no_var = caret::nearZeroVar(base[,seq])
# none

# log transform
base_log <- base %>% modify_if(is_seq(names(.)), log)

```

# Methods

Sparse partial least-squares discriminant analysis (sPLS-DA) was used to identify proteins that best discriminate between participants with and without a given comorbidity. Ten-fold cross-validation was used to determine the optimal parameters, including the number of proteins retained. Permutation testing was used to evaluate the extent of overfitting.

For each comorbidity, the output is as follows:

- An individual plot where each participant is represented by a point on the axes defined the first two sPLS-DA components, with 95% confidence intervals for the ellipse defined by the comorbidity groups.
- Next, the proteins retained by the sPLS-DA are listed along with their loadings on the first two sPLS-DA components. The first box contains the proteins retained for the first component, and the second box contains the proteins retained for the second component.
- Finally, a plot summarizes the results of the permutation testing. The red line indicates the point estimate of the AUC from the original sPLS-DA model. The histogram represents AUC values obtained by randomly permuting the class labels (i.e., comorbidity groups). If the red line is far above the rest of the histogram, this indicates potential overfitting and the models should be interpreted with caution.

# Results

## Microalbuminuria

```{r echo=FALSE, include=TRUE, cache=TRUE, comment=""}
Y <- base_log$MIC
X <- base_log[,seq]
tune.splsda <- tune.splsda(X, Y, ncomp = 5) 
choice.keepX <- tune.splsda$choice.keepX[1:2] 
## sPLS-DA function
splsda.utm <- splsda(X, Y, ncomp = 2, keepX = choice.keepX)
plotIndiv(splsda.utm,title = "sPLS-DA by microalbuminuria status",
          pch = 20,ind.names = F,ellipse = T,legend = T)

top_1 <- selectVar(splsda.utm,comp=1)
top_2 <- selectVar(splsda.utm,comp=2)
top_1d <- data.frame(top_1$value)
top_1d$AptName <- row.names(top_1d)
top_1d <- merge(top_1d,analytes,by="AptName",all.x = T, all.y = F)
top_1d <- top_1d[order(-abs(top_1d$value.var)),] 
top_2d <- data.frame(top_2$value)
top_2d$AptName <- row.names(top_2d)
top_2d <- merge(top_2d,analytes,by="AptName",all.x = T, all.y = F)
top_2d <- top_2d[order(-abs(top_2d$value.var)),] 
top_1d[,c("value.var","TargetFullName","Target")]
top_2d[,c("value.var","TargetFullName","Target")]

set.seed(3654)
plsda.perf = perf(splsda.utm,validation = "Mfold",folds = 5,auc = T)
auc_true <- as.numeric(plsda.perf$auc$comp1[1])
# Permutation testing - takes forever (even in parallel), make sure to cache
n_perm <- 1000
cl = makeForkCluster(detectCores()*0.5)
aucs = parSapply(cl,1:n_perm, function(i){
  Y <- sample(as.factor(base_log$MIC),replace = F)
  plsda_res <- plsda(X,Y,ncomp = 2)
  perf_plsda = perf(plsda_res,validation = "Mfold",folds = 5,auc = T)
  return(as.numeric(perf_plsda$auc$comp1[1]))
})
stopCluster(cl)
ggplot(as.data.frame(aucs),aes(x=aucs)) +
  geom_histogram(bins = 10) +
  geom_vline(aes(xintercept=auc_true),color="red") +
  theme_bw() + xlab("AUC")

```

## Macroalbuminuria

```{r echo=FALSE, include=TRUE, cache=TRUE, comment=""}
Y <- base_log$MAC
X <- base_log[,seq]
tune.splsda <- tune.splsda(X, Y, ncomp = 5) 
choice.keepX <- tune.splsda$choice.keepX[1:2] 
## sPLS-DA function
splsda.utm <- splsda(X, Y, ncomp = 2, keepX = choice.keepX)
plotIndiv(splsda.utm,title = "sPLS-DA by macroalbuminuria status",
          pch = 20,ind.names = F,ellipse = T,legend = T)

top_1 <- selectVar(splsda.utm,comp=1)
top_2 <- selectVar(splsda.utm,comp=2)
top_1d <- data.frame(top_1$value)
top_1d$AptName <- row.names(top_1d)
top_1d <- merge(top_1d,analytes,by="AptName",all.x = T, all.y = F)
top_1d <- top_1d[order(-abs(top_1d$value.var)),] 
top_2d <- data.frame(top_2$value)
top_2d$AptName <- row.names(top_2d)
top_2d <- merge(top_2d,analytes,by="AptName",all.x = T, all.y = F)
top_2d <- top_2d[order(-abs(top_2d$value.var)),] 
top_1d[,c("value.var","TargetFullName","Target")]
top_2d[,c("value.var","TargetFullName","Target")]

set.seed(3654)
plsda.perf = perf(splsda.utm,validation = "Mfold",folds = 5,auc = T)
auc_true <- as.numeric(plsda.perf$auc$comp1[1])
# Permutation testing - takes forever (even in parallel), make sure to cache
n_perm <- 1000
cl = makeForkCluster(detectCores()*0.5)
aucs = parSapply(cl,1:n_perm, function(i){
  Y <- sample(as.factor(base_log$MAC),replace = F)
  plsda_res <- plsda(X,Y,ncomp = 2)
  perf_plsda = perf(plsda_res,validation = "Mfold",folds = 5,auc = T)
  return(as.numeric(perf_plsda$auc$comp1[1]))
})
stopCluster(cl)
ggplot(as.data.frame(aucs),aes(x=aucs)) +
  geom_histogram(bins = 10) +
  geom_vline(aes(xintercept=auc_true),color="red") +
  theme_bw() + xlab("AUC")
```

## Hyperfiltration

```{r echo=FALSE, include=TRUE, cache=TRUE, comment=""}
Y <- base_log[!is.na(base_log$HYP),]$HYP
X <- base_log[!is.na(base_log$HYP),seq]
tune.splsda <- tune.splsda(X, Y, ncomp = 5) 
choice.keepX <- tune.splsda$choice.keepX[1:2] 
## sPLS-DA function
splsda.utm <- splsda(X, Y, ncomp = 2, keepX = choice.keepX)
plotIndiv(splsda.utm,title = "sPLS-DA by hyperfiltration status",
          pch = 20,ind.names = F,ellipse = T,legend = T)

top_1 <- selectVar(splsda.utm,comp=1)
top_2 <- selectVar(splsda.utm,comp=2)
top_1d <- data.frame(top_1$value)
top_1d$AptName <- row.names(top_1d)
top_1d <- merge(top_1d,analytes,by="AptName",all.x = T, all.y = F)
top_1d <- top_1d[order(-abs(top_1d$value.var)),] 
top_2d <- data.frame(top_2$value)
top_2d$AptName <- row.names(top_2d)
top_2d <- merge(top_2d,analytes,by="AptName",all.x = T, all.y = F)
top_2d <- top_2d[order(-abs(top_2d$value.var)),] 
top_1d[,c("value.var","TargetFullName","Target")]
top_2d[,c("value.var","TargetFullName","Target")]

set.seed(3654)
plsda.perf = perf(splsda.utm,validation = "Mfold",folds = 5,auc = T)
auc_true <- as.numeric(plsda.perf$auc$comp1[1])
# Permutation testing - takes forever (even in parallel), make sure to cache
n_perm <- 1000
cl = makeForkCluster(detectCores()*0.5)
aucs = parSapply(cl,1:n_perm, function(i){
  Y <- sample(as.factor(base_log[!is.na(base_log$HYP),]$HYP),replace = F)
  plsda_res <- plsda(X,Y,ncomp = 2)
  perf_plsda = perf(plsda_res,validation = "Mfold",folds = 5,auc = T)
  return(as.numeric(perf_plsda$auc$comp1[1]))
})
stopCluster(cl)
ggplot(as.data.frame(aucs),aes(x=aucs)) +
  geom_histogram(bins = 10) +
  geom_vline(aes(xintercept=auc_true),color="red") +
  theme_bw() + xlab("AUC")
```

## Rapid eGFR decline

```{r echo=FALSE, include=TRUE, cache=TRUE, comment=""}
Y <- base_log[!is.na(base_log$RAPID),]$RAPID
X <- base_log[!is.na(base_log$RAPID),seq]
tune.splsda <- tune.splsda(X, Y, ncomp = 5) 
choice.keepX <- tune.splsda$choice.keepX[1:2] 
## sPLS-DA function
splsda.utm <- splsda(X, Y, ncomp = 2, keepX = choice.keepX)
plotIndiv(splsda.utm,title = "sPLS-DA by rapid eGFR decline status",
          pch = 20,ind.names = F,ellipse = T,legend = T)

top_1 <- selectVar(splsda.utm,comp=1)
top_2 <- selectVar(splsda.utm,comp=2)
top_1d <- data.frame(top_1$value)
top_1d$AptName <- row.names(top_1d)
top_1d <- merge(top_1d,analytes,by="AptName",all.x = T, all.y = F)
top_1d <- top_1d[order(-abs(top_1d$value.var)),] 
top_2d <- data.frame(top_2$value)
top_2d$AptName <- row.names(top_2d)
top_2d <- merge(top_2d,analytes,by="AptName",all.x = T, all.y = F)
top_2d <- top_2d[order(-abs(top_2d$value.var)),] 
top_1d[,c("value.var","TargetFullName","Target")]
top_2d[,c("value.var","TargetFullName","Target")]

set.seed(3654)
plsda.perf = perf(splsda.utm,validation = "Mfold",folds = 5,auc = T)
auc_true <- as.numeric(plsda.perf$auc$comp1[1])
# Permutation testing - takes forever (even in parallel), make sure to cache
n_perm <- 1000
cl = makeForkCluster(detectCores()*0.5)
aucs = parSapply(cl,1:n_perm, function(i){
  Y <- sample(as.factor(base_log[!is.na(base_log$RAPID),]$RAPID),replace = F)
  plsda_res <- plsda(X,Y,ncomp = 2)
  perf_plsda = perf(plsda_res,validation = "Mfold",folds = 5,auc = T)
  return(as.numeric(perf_plsda$auc$comp1[1]))
})
stopCluster(cl)
ggplot(as.data.frame(aucs),aes(x=aucs)) +
  geom_histogram(bins = 10) +
  geom_vline(aes(xintercept=auc_true),color="red") +
  theme_bw() + xlab("AUC")
```

## Hypertension

```{r echo=FALSE, include=TRUE, cache=TRUE, comment=""}
Y <- base_log$HTN
X <- base_log[,seq]
tune.splsda <- tune.splsda(X, Y, ncomp = 5) 
choice.keepX <- tune.splsda$choice.keepX[1:2] 
## sPLS-DA function
splsda.utm <- splsda(X, Y, ncomp = 2, keepX = choice.keepX)
plotIndiv(splsda.utm,title = "sPLS-DA by hypertension status",
          pch = 20,ind.names = F,ellipse = T,legend = T)

top_1 <- selectVar(splsda.utm,comp=1)
top_2 <- selectVar(splsda.utm,comp=2)
top_1d <- data.frame(top_1$value)
top_1d$AptName <- row.names(top_1d)
top_1d <- merge(top_1d,analytes,by="AptName",all.x = T, all.y = F)
top_1d <- top_1d[order(-abs(top_1d$value.var)),] 
top_2d <- data.frame(top_2$value)
top_2d$AptName <- row.names(top_2d)
top_2d <- merge(top_2d,analytes,by="AptName",all.x = T, all.y = F)
top_2d <- top_2d[order(-abs(top_2d$value.var)),] 
top_1d[,c("value.var","TargetFullName","Target")]
top_2d[,c("value.var","TargetFullName","Target")]

set.seed(3654)
plsda.perf = perf(splsda.utm,validation = "Mfold",folds = 5,auc = T)
auc_true <- as.numeric(plsda.perf$auc$comp1[1])
# Permutation testing - takes forever (even in parallel), make sure to cache
n_perm <- 1000
cl = makeForkCluster(detectCores()*0.5)
aucs = parSapply(cl,1:n_perm, function(i){
  Y <- sample(as.factor(base_log$HTN),replace = F)
  plsda_res <- plsda(X,Y,ncomp = 2)
  perf_plsda = perf(plsda_res,validation = "Mfold",folds = 5,auc = T)
  return(as.numeric(perf_plsda$auc$comp1[1]))
})
stopCluster(cl)
ggplot(as.data.frame(aucs),aes(x=aucs)) +
  geom_histogram(bins = 10) +
  geom_vline(aes(xintercept=auc_true),color="red") +
  theme_bw() + xlab("AUC")
```

## Neuropathy

```{r echo=FALSE, include=TRUE, cache=TRUE, comment=""}
Y <- base_log$NEURO
X <- base_log[,seq]
tune.splsda <- tune.splsda(X, Y, ncomp = 5) 
choice.keepX <- tune.splsda$choice.keepX[1:2] 
## sPLS-DA function
splsda.utm <- splsda(X, Y, ncomp = 2, keepX = choice.keepX)
plotIndiv(splsda.utm,title = "sPLS-DA by neuropathy status",
          pch = 20,ind.names = F,ellipse = T,legend = T)

top_1 <- selectVar(splsda.utm,comp=1)
top_2 <- selectVar(splsda.utm,comp=2)
top_1d <- data.frame(top_1$value)
top_1d$AptName <- row.names(top_1d)
top_1d <- merge(top_1d,analytes,by="AptName",all.x = T, all.y = F)
top_1d <- top_1d[order(-abs(top_1d$value.var)),] 
top_2d <- data.frame(top_2$value)
top_2d$AptName <- row.names(top_2d)
top_2d <- merge(top_2d,analytes,by="AptName",all.x = T, all.y = F)
top_2d <- top_2d[order(-abs(top_2d$value.var)),] 
top_1d[,c("value.var","TargetFullName","Target")]
top_2d[,c("value.var","TargetFullName","Target")]

set.seed(3654)
plsda.perf = perf(splsda.utm,validation = "Mfold",folds = 5,auc = T)
auc_true <- as.numeric(plsda.perf$auc$comp1[1])
# Permutation testing - takes forever (even in parallel), make sure to cache
n_perm <- 1000
cl = makeForkCluster(detectCores()*0.5)
aucs = parSapply(cl,1:n_perm, function(i){
  Y <- sample(as.factor(base_log$NEURO),replace = F)
  plsda_res <- plsda(X,Y,ncomp = 2)
  perf_plsda = perf(plsda_res,validation = "Mfold",folds = 5,auc = T)
  return(as.numeric(perf_plsda$auc$comp1[1]))
})
stopCluster(cl)
ggplot(as.data.frame(aucs),aes(x=aucs)) +
  geom_histogram(bins = 10) +
  geom_vline(aes(xintercept=auc_true),color="red") +
  theme_bw() + xlab("AUC")
```

## Retinopathy

```{r echo=FALSE, include=TRUE, cache=TRUE, comment=""}
Y <- base_log[!is.na(base_log$RETINO),]$RETINO
X <- base_log[!is.na(base_log$RETINO),seq]
tune.splsda <- tune.splsda(X, Y, ncomp = 5) 
choice.keepX <- tune.splsda$choice.keepX[1:2] 
## sPLS-DA function
splsda.utm <- splsda(X, Y, ncomp = 2, keepX = choice.keepX)
plotIndiv(splsda.utm,title = "sPLS-DA by retinopathy status",
          pch = 20,ind.names = F,ellipse = T,legend = T)

top_1 <- selectVar(splsda.utm,comp=1)
top_2 <- selectVar(splsda.utm,comp=2)
top_1d <- data.frame(top_1$value)
top_1d$AptName <- row.names(top_1d)
top_1d <- merge(top_1d,analytes,by="AptName",all.x = T, all.y = F)
top_1d <- top_1d[order(-abs(top_1d$value.var)),] 
top_2d <- data.frame(top_2$value)
top_2d$AptName <- row.names(top_2d)
top_2d <- merge(top_2d,analytes,by="AptName",all.x = T, all.y = F)
top_2d <- top_2d[order(-abs(top_2d$value.var)),] 
top_1d[,c("value.var","TargetFullName","Target")]
top_2d[,c("value.var","TargetFullName","Target")]

set.seed(3654)
plsda.perf = perf(splsda.utm,validation = "Mfold",folds = 5,auc = T)
auc_true <- as.numeric(plsda.perf$auc$comp1[1])
# Permutation testing - takes forever (even in parallel), make sure to cache
n_perm <- 1000
cl = makeForkCluster(detectCores()*0.5)
aucs = parSapply(cl,1:n_perm, function(i){
  Y <- sample(as.factor(base_log[!is.na(base_log$RETINO),]$RETINO),replace = F)
  plsda_res <- plsda(X,Y,ncomp = 2)
  perf_plsda = perf(plsda_res,validation = "Mfold",folds = 5,auc = T)
  return(as.numeric(perf_plsda$auc$comp1[1]))
})
stopCluster(cl)
ggplot(as.data.frame(aucs),aes(x=aucs)) +
  geom_histogram(bins = 10) +
  geom_vline(aes(xintercept=auc_true),color="red") +
  theme_bw() + xlab("AUC")
```