---
title: "ERCB analysis"
author: "Laura Pyle"
date: "September 20, 2018"
output: word_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyr)
library(magrittr)
library(dplyr)
library(lubridate)
library(zoo) 
library(M3)
library(data.table)
library(Table1)
library(devtools)
library(Hmisc)
library(knitr)
#library(kableExtra)
library(psycho)
#library(tidyverse)

####################################################################################################
#Data Manipulation #
####################################################################################################

setwd("H:\\Endocrinology\\Bjornstad\\NEPTUNE\\Data")
nameFile1 <- ".\\GlomerularTranscriptome_ERCB.csv"
nameFile2 <- ".\\TItranscriptome_ERCB.csv"

# read the data
#alldata <- read.csv(nameFile)
glom <- read.csv(nameFile1,na.strings = c("N",'',' ','               '))
ti <- read.csv(nameFile2,na.strings = c("N",'',' ','               '))

# set factors and levels

# set labels
label(glom$PAT_Cohort)="Disease subtype"
label(glom$log2.SLC5A2..SGLT2.)="Log2 SLC5A2 SGLT2 GLOM"
label(glom$log2.SLC5A1..SGLT1.)="Log2 SLCA1 SGLT1 GLOM"
label(ti$PAT_Cohort)="Disease subtype"
label(ti$log2.SLC5A2..SGLT2.)="Log2 SLC5A2 SGLT2 TI"
label(ti$log2.SLC5A1..SGLT1.)="Log2 SLC5A1 SGLT1 TI"

Nglom <- nrow(distinct(glom,Sample.name))
Nti <- nrow(distinct(ti,sample.id))

####################################################################################################
# Compare expression of SGLT1 and SGLT2 mRNA across diagnosis categories #
# Stratify by age group                                                  #
####################################################################################################

# transcriptome variables by dx group in peds
tab_glom <- final_table(glom,c("log2.SLC5A2..SGLT2.","log2.SLC5A1..SGLT1."),
                    group=glom$PAT_Cohort,margin=2,single=0,ron=1)

# transcriptome variables by dx group in adults
tab_ti <- final_table(ti,c("log2.SLC5A2..SGLT2.","log2.SLC5A1..SGLT1."),
                    group=ti$PAT_Cohort,margin=2,single=0,ron=1)


```

# Background

The purpose of this analysis is to compare SGLT2 and SGLT1 mRNA expression in the kidney across GN disease subtypes in the ERCB study.

# Methods

Descriptive statistics reported include mean and standard deviations or medians and percentiles for continuous variables, and frequencies and percentages for categorical variables.  To compare characteristics in the two groups, t-tests/ANOVA or the Mann-Whitney/Wilcoxon test were used for continuous variables and the chi-square test or Fisher's exact test was used for categorical variables.  No correction was made for multiple comparisons.  All analyses were performed using R version 3.4.1.

# Results


There were `r Nglom` glomerular samples included in this analysis.  Table 1a shows the comparison of mRNA expression by GN disease subtype.

```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab_glom,caption="Table 1a.  mRNA expression in glomerular samples, by disease subtype.  Statistics reported are mean ± standard deviation.")
```

There were `r Nti` TI samples included in this analysis.  Table 1b shows the comparison of mRNA expression by GN disease subtype.

```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab_ti,caption="Table 1b.  mRNA expression in TI samples, by disease subtype.  Statistics reported are mean ± standard deviation.")
```
\
