---
title: "Colorado and Netherlands Analyses"
author: "Callie Rountree-Jablin"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 3
    toc_float: yes
    toc_loc: left
---

<!-- This file includes analyses for the merged RED Netherlands and Colorado harmonized dataset. 
Aside from adding labels, no additional data cleaning is performed in this rmd as all data cleaning was previously 
conducted in the CO_RED_data_cleaning_merge rmd file.-->

```{r setup, include=FALSE}
library(tidyverse)
library(kableExtra)
library(magrittr)
library(Hmisc)

if(Sys.info()["sysname"] == "Windows"){
  home_dir = "B:/Petter Bjornstad"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/Peds Endo/Petter Bjornstad/"
} 
knitr::opts_knit$set(echo=FALSE, root.dir = home_dir)
```


```{r read, include=FALSE}
combined_clean <- read.csv('./IHD/Clean Data/combined_CO_RED_clean.csv')
dict <- read.csv("./Data Harmonization/data_dictionary_master.csv", na.strings = c(" ", "", "-9999",-9999))

# labels
rownames(dict) <- dict$variable_name
dict %<>% select("label")
dict <- t(dict) %>% as.data.frame(dict)
rownames(dict) <- "label"
dict <- dict[intersect(names(combined_clean), names(dict))]
dict[setdiff(names(combined_clean), names(dict))] <- " "
label(combined_clean) = as.list(dict[match(names(combined_clean), names(dict))])

label(combined_clean$data_source) <- "Site"
label(combined_clean$`Body_Fat_Percentage`) <- "Body Fat Percentage"
label(combined_clean$cholesterol) <- "Cholesterol (mg/dL)"
label(combined_clean$AER_sex_adjusted) <- "Sex-Adjusted Estimate of AER"
label(combined_clean$AER_covariate_adjusted) <- "Covariate-Adjusted Estimate of AER"
label(combined_clean$NEFA_V2.1_Fasting) <- "Fasting NEFA level 100 minutes after iohexol/PAH"
label(combined_clean$NEFA_V2.7_Eu) <- "NEFA level 180 minutes after initiation of euglycemic clamp"
label(combined_clean$NEFA_delta) <- "Difference in NEFA level (fasting NEFA – NEFA during euglycemic clamp)"
label(combined_clean$INUorIOH) <- "Inulin/Iohexol Indicator for GFR Measurement"
label(combined_clean$Med_betablocker) <- "Betablocker use"
label(combined_clean$Med_calciumantagonist) <- "Calcium channel blocker use"
label(combined_clean$Med_anticoagulant) <- "Anticoagulant use"
label(combined_clean$Medication) <- "Current medication use and dose"
```


```{r lm function, include=FALSE}
# define function for outputting LM summary and diagnostic plots
lm_summary_and_diagnostics <- function(mod){
  print(summary(mod))
  plot(x = predict(mod), y = rstudent(mod), xlab = 'Predicted Value', ylab = 'Jackknife Residual', 
       main = 'Scatterplot of Residuals'); abline(h = 0, lty = 2)
  plot(ppoints(length(rstudent(mod))), sort(pnorm(rstudent(mod))), xlab='Observed Cumulative Probability', ylab='Expected Cumulative
       Probability', main='Normal Probability Plot', cex=1); abline(a=0,b=1, col='gray65', lwd=1)
  hist(rstudent(mod), xlab='Jackknife Residual', main='Histogram of Residuals', freq=F)
  }
```

This report describes preliminary analyses investigating kidney hemodynamic function and insulin sensitivity measures among youth-onset (Colorado) and adult-onset (Netherlands) Type 2 diabetes patients. 

# Methods

## Data Cleaning

Datasets from the Colorado and Netherlands sites were separately cleaned as described in sections 1.1.1 and 1.1.2 before being combined as one main dataset for analyses. Variables that were considered similar between the two sites were combined as one variable. In instances where data for a variable was only available from one site, if the variable was of interest (e.g., steady state c-peptide) it was still included in the main dataset. 

### Colorado Cleaning

The Colorado harmonized dataset was filtered to include subjects in the Type 2 Diabetes group. Subjects who were enrolled in only RENAL-HEIRitage were also excluded. In instances where subjects had multiple entries due to multiple study procedures per visit, the data was collapsed to one row where the earliest measurements were retained. Both gender-adjusted and covariate-adjusted estimates of albumin excretion rate (AER) were calculated for all subjects at the Colorado site using the derivations described in the Results section of [Younes (2010)](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2893069/).

### Netherlands Cleaning

All provided Netherlands datasets were first combined into one main dataset. One subject was dropped when the datasets were merged due to missing data on all outcomes of interest. Netherlands variables were renamed to match the Colorado notation and if needed, units were adjusted to match units used at the Colorado site. Albumin excretion rate (AER) was calculated for all Netherlands subjects based on the following formula: $$AER=\frac{\text{24 hour urine microalbuminuria quantity}}{24 \times 60}$$

<!-- ## Data Analysis (to add in once analyses more finalized) -->

# Results

## Descriptive Statistics of Participants

This section summarizes the variables available in the combined data set. 
```{r table1, echo=FALSE}

table1::table1 (~ study + group + age + diabetes_duration + sex + ethnicity + height + weight + waistcm +
                  Body_Fat_Percentage + bsa + bmi + hba1c + gfr_raw_plasma + gfr_bsa_plasma + eGFR_fas_cr_cysc + 
                  erpf_raw_plasma + erpf_bsa_plasma + ff + rvr + ra + re + glomerular_pressure + sbp + dbp + map + 
                  creatinine_s + cholesterol + hdl + ldl + acprg + acr_u + AER_sex_adjusted + AER_covariate_adjusted + 
                  airg + di + gir_190 + gir_200 + steady_state_cpeptide + steady_state_ffa + ffa_suppression + 
                  steady_state_insulin + NEFA_V2.1_Fasting + NEFA_V2.7_Eu + NEFA_delta + INUorIOH + metformin_timepoint +
                  raasi_timepoint + statin + Med_betablocker + Med_calciumantagonist + Med_anticoagulant | data_source, 
                data = combined_clean)
```

## Comparison of Measures in Youth-onset vs. Adult-onset, Adjusted for Sex

In the below output, `data_source` is the variable distinguishing between Youth-onset (Colorado) and Adult-onset (Netherlands).
```{r YT2D vs AT2D adjusted for sex, echo=FALSE}
outcome_vars <- c('steady_state_insulin', 'gir_190', 'gir_200', 'di', 'ffa_suppression')

for (v in outcome_vars) {
  form = as.formula(paste0(v,'~ data_source + sex'))
  mod <- lm(form, data = combined_clean)
  cat(noquote(c("Association between", noquote(v), "and data source, adjusted for sex\n")))
  cat("Model Form: ")
  print(form)
  lm_summary_and_diagnostics(mod)
  }
```


## Comparison of Measures in Youth-onset vs. Adult-onset, Adjusted for Sex and HbA1c

In the below output, `data_source` is the variable distinguishing between Colorado and Netherlands.
```{r YT2D vs AT2D adjusted for sex and HbA1c, echo=FALSE}
outcome_vars <- c('gfr_raw_plasma', 'gfr_bsa_plasma', 'erpf_raw_plasma', 'erpf_bsa_plasma', 
                  'glomerular_pressure', 'ra', 're', 'rvr', 'ff')

for (v in outcome_vars) {
  form = as.formula(paste0(v,'~ data_source + sex + hba1c'))
  mod <- lm(form, data = combined_clean)
  cat(noquote(c("Association between", noquote(v), "and data source, adjusted for sex and HbA1c\n")))
  cat("Model Form: ")
  print(form)
  lm_summary_and_diagnostics(mod)
  }
```


## Association Between Measures of Interest, Adjusted for Sex and HbA1c

```{r measures of interest adjusted for sex and HbA1c, echo=FALSE}
cat("Association between sex-adjusted AER and FFA suppression, adjusted for sex and HbA1c:")
AER_ffa_suppress_mod1 <- lm(AER_sex_adjusted ~ ffa_suppression + sex + hba1c, data = combined_clean)
lm_summary_and_diagnostics(AER_ffa_suppress_mod1)

cat("Association between covariate-adjusted AER and FFA suppression, adjusted for sex and HbA1c:")
AER_ffa_suppress_mod2 <- lm(AER_covariate_adjusted ~ ffa_suppression + sex + hba1c, data = combined_clean)
lm_summary_and_diagnostics(AER_ffa_suppress_mod2)

cat("Association between glomerular pressure and FFA suppression, adjusted for sex and HbA1c:")
pglo_ffa_suppress_mod <- lm(glomerular_pressure ~ ffa_suppression + sex + hba1c, data = combined_clean)
lm_summary_and_diagnostics(pglo_ffa_suppress_mod)

cat("Association between re and FFA suppression, adjusted for sex and HbA1c:")
re_ffa_suppress_mod <- lm(re ~ ffa_suppression + sex + hba1c, data = combined_clean)
lm_summary_and_diagnostics(re_ffa_suppress_mod)

cat("Association between rvr and FFA suppression, adjusted for sex and HbA1c:")
rvr_ffa_suppress_mod <- lm(rvr ~ ffa_suppression + sex + hba1c, data = combined_clean)
lm_summary_and_diagnostics(rvr_ffa_suppress_mod)
```

## Unadjusted Association Between Measures of Interest

### Raw M-value Models
Raw M-value refers to a dextrose concentration of 190 in Youth-onset (Colorado) participants.

```{r raw M mods, echo=FALSE}
raw_m_outcome_vars <- c('AER_sex_adjusted', 'AER_covariate_adjusted', 'glomerular_pressure', 'ff')

for (v in raw_m_outcome_vars) {
  form = as.formula(paste0('gir_190 ~ ', v))
  mod <- lm(form, data = combined_clean)
  cat(noquote(c("Association between raw M-value and", noquote(v))))
  cat("\nModel Form: ")
  print(form)
  lm_summary_and_diagnostics(mod)
  }
```


### GIR Models
GIR corresponds to a dextrose concentration of 200 in Youth-onset (Colorado) participants.
```{r GIR mods, echo=FALSE}
gir_outcome_vars <- c('AER_sex_adjusted', 'AER_covariate_adjusted', 'glomerular_pressure', 'ff')

for (v in gir_outcome_vars) {
  form = as.formula(paste0('gir_200 ~ ', v))
  mod <- lm(form, data = combined_clean)
  cat(noquote(c("Association between GIR and", noquote(v))))
  cat("\nModel Form: ")
  print(form)
  lm_summary_and_diagnostics(mod)
  }
```

### Steady-State Insulin Models
```{r steady state mods, echo=FALSE}
steady_state_insulin_outcome_vars <- c('AER_sex_adjusted', 'AER_covariate_adjusted', 'glomerular_pressure', 'ff')

for (v in steady_state_insulin_outcome_vars) {
  form = as.formula(paste0('steady_state_insulin ~ ', v))
  mod <- lm(form, data = combined_clean)
  cat(noquote(c("Association between steady-state insulin and", noquote(v))))
  cat("\nModel Form: ")
  print(form)
  lm_summary_and_diagnostics(mod)
  }
```

# Questions/Notes

* `gir_190` = Raw M-Value (Colorado site)
* `gir_200` = GIR (Colorado site)
* One subject from the Netherlands site was found to be missing clamp data (RED-16) - this subject was excluded from all analyses, is this appropriate?
* Variables that are missing from Netherlands dataset include: Steady-state C-peptide, AIRg, and ACPRg
* Are ACPRG and AIRg C-peptide equivalent? (both are included in the DAP) 
* Should AER analyses be based on the sex-adjusted estimate or covariate-adjusted estimate?
* Is it best to match Netherlands `Euclamp_V2_Mvalue_corr_minus_glycosuria` with Colorado `gir_190` or `gir_200`?



