---
title: "Colorado and Netherlands Analyses"
author: "Callie Rountree-Jablin"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 3
    toc_float: yes
    toc_loc: left
---

<!-- This file includes analyses for the merged RED Netherlands and Colorado harmonized dataset. 
Please reference the CO_RED_data_cleaning_merge rmd file for data cleaning used to combine the two datasets.-->

```{r setup, include=FALSE}
library(tidyverse)
library(kableExtra)
library(magrittr)
library(Hmisc)
library(emmeans)

if(Sys.info()["sysname"] == "Windows"){
  home_dir = "B:/Petter Bjornstad"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/Peds Endo/Petter Bjornstad/"
} 
knitr::opts_knit$set(echo=FALSE, root.dir = home_dir)
```


```{r read, include=FALSE}
combined_clean <- read.csv('./IHD/Clean Data/combined_CO_RED_clean.csv')
dict <- read.csv("./Data Harmonization/data_dictionary_master.csv", na.strings = c(" ", "", "-9999",-9999))

# labels
rownames(dict) <- dict$variable_name
dict %<>% select("label")
dict <- t(dict) %>% as.data.frame(dict)
rownames(dict) <- "label"
dict <- dict[intersect(names(combined_clean), names(dict))]
dict[setdiff(names(combined_clean), names(dict))] <- " "
label(combined_clean) = as.list(dict[match(names(combined_clean), names(dict))])

label(combined_clean$site) <- "Site"
label(combined_clean$`Body_Fat_Percentage`) <- "Body Fat Percentage"
label(combined_clean$cholesterol) <- "Cholesterol (mg/dL)"
label(combined_clean$AER_sex_adjusted) <- "Sex-Adjusted Estimate of AER (mg/24 hr)"
label(combined_clean$AER_covariate_adjusted) <- "Covariate-Adjusted Estimate of AER (mg/24 hr)"
label(combined_clean$NEFA_V2.1_Fasting) <- "Fasting NEFA level 100 minutes after iohexol/PAH"
label(combined_clean$NEFA_delta) <- "Difference in NEFA level (fasting NEFA – NEFA during euglycemic clamp)"
label(combined_clean$INUorIOH) <- "Inulin/Iohexol Indicator for GFR Measurement"
label(combined_clean$Med_betablocker) <- "Betablocker use"
label(combined_clean$Med_calciumantagonist) <- "Calcium channel blocker use"
label(combined_clean$Med_anticoagulant) <- "Anticoagulant use"
label(combined_clean$Medication) <- "Current medication use and dose"
label(combined_clean$bsa) <- "BSA (m^2)"

```


```{r lm function, include=FALSE}
# define function for outputting LM summary and diagnostic plots
lm_summary_and_diagnostics <- function(mod){
  print(summary(mod))
  plot(x = predict(mod), y = rstudent(mod), xlab = 'Predicted Value', ylab = 'Jackknife Residual', 
       main = 'Scatterplot of Residuals'); abline(h = 0, lty = 2)
  plot(ppoints(length(rstudent(mod))), sort(pnorm(rstudent(mod))), xlab='Observed Cumulative Probability', ylab='Expected Cumulative
       Probability', main='Normal Probability Plot', cex=1); abline(a=0,b=1, col='gray65', lwd=1)
  hist(rstudent(mod), xlab='Jackknife Residual', main='Histogram of Residuals', freq=F)
}
```

This report describes analysis results investigating kidney hemodynamic function and insulin sensitivity measures among youth-onset (Colorado) and adult-onset (Netherlands) Type 2 diabetes patients. 

# Methods

## Data Cleaning

Datasets from the Colorado and Netherlands sites were separately cleaned as described in sections 1.1.1 and 1.1.2 before being combined as one main dataset for analyses. Variables that were considered similar between the two sites were combined as one variable.

### Colorado Cleaning

The Colorado harmonized dataset was filtered to include subjects in the Type 2 Diabetes group. Subjects who were enrolled in only RENAL-HEIRitage were also excluded. In instances where subjects had multiple entries due to multiple study procedures per visit and/or co-enrollment on studies, the data was collapsed to one row where the earliest measurements were retained. BSA calculations were performed prior to collapsing the data and were based on the below formula for consistency with the Netherlands site:
$$ BSA = \sqrt {\text{weight (kg)} \times \text{height (cm)} / 3600}$$

After the data was collapsed, both gender-adjusted and covariate-adjusted estimates of albumin excretion rate (AER) were calculated for all subjects at the Colorado site using the derivations described in the Results section of [Younes (2010)](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2893069/).

### Netherlands Cleaning

All provided Netherlands datasets were first combined into one main dataset. One subject was dropped when the datasets were merged due to missing data on all outcomes of interest. Netherlands variables were renamed to match the Colorado notation and if needed, units were adjusted to match units used at the Colorado site. Albumin excretion rate (AER) was calculated for all Netherlands subjects based on the following formula: $$AER=\frac{\text{24 hour urine microalbuminuria quantity}}{24 \times 60}$$

<!-- ## Data Analysis (to add in once analyses more finalized) -->

# Results

## Descriptive Statistics of Participants

This section summarizes the variables available in the combined data set. Note that while some subjects at the Colorado site were enrolled on multiple studies, they are only represented once under `Study Name`. For example, if a subject was enrolled on RENAL-HEIR and IMPROVE, they are counted as a participant on either RENAL-HEIR or IMPROVE (whichever study they were enrolled on first). 

```{r table1, echo=FALSE}
table1::table1 (~ study + age + diabetes_duration + sex + ethnicity + height + weight + waistcm +
                  Body_Fat_Percentage + bsa + bmi + hba1c + gfr_raw_plasma + gfr_bsa_plasma + eGFR_fas_cr_cysc + 
                  erpf_raw_plasma + erpf_bsa_plasma + ff + rvr + ra + re + glomerular_pressure + sbp + dbp + map + 
                  creatinine_s + cholesterol + hdl + ldl + acr_u + AER_sex_adjusted + AER_covariate_adjusted + 
                  gir_190 + gir_200 + steady_state_ffa + ffa_suppression + 
                  steady_state_insulin + NEFA_V2.1_Fasting + NEFA_delta + metformin_timepoint +
                  raasi_timepoint + statin + Med_betablocker + Med_calciumantagonist + Med_anticoagulant | site, 
                data = combined_clean)
```

## Comparison of Measures in Youth-onset vs. Adult-onset, Adjusted for Sex

In the below output, `site` is the variable distinguishing between Youth-onset (Colorado) and Adult-onset (Netherlands).
```{r YT2D vs AT2D adjusted for sex, echo=FALSE}
outcome_vars <- c('steady_state_insulin', 'gir_190', 'gir_200', 'ffa_suppression')

for (v in outcome_vars) {
  form = as.formula(paste0(v,'~ site + sex'))
  mod <- lm(form, data = combined_clean)
  cat(noquote(c("Association between", noquote(v), "and site, adjusted for sex\n")))
  cat("Model Form: ")
  print(form)
  lm_summary_and_diagnostics(mod)
  }
```


## Comparison of Measures in Youth-onset vs. Adult-onset, Adjusted for Sex and HbA1c

In the below output, `site` is the variable distinguishing between Colorado and Netherlands.
```{r YT2D vs AT2D adjusted for sex and HbA1c, echo=FALSE}
outcome_vars <- c('gfr_raw_plasma', 'gfr_bsa_plasma', 'erpf_raw_plasma', 'erpf_bsa_plasma', 
                  'glomerular_pressure', 'ra', 're', 'rvr', 'ff')

for (v in outcome_vars) {
  form = as.formula(paste0(v,'~ site + sex + hba1c'))
  mod <- lm(form, data = combined_clean)
  cat(noquote(c("Association between", noquote(v), "and site, adjusted for sex and HbA1c\n")))
  cat("Model Form: ")
  print(form)
  lm_summary_and_diagnostics(mod)
  }
```


## Modeling with Interaction between FFA Suppression and Site, Adjusted for HbA1c

This section reports the results for models fit to evaluate whether site modifies the measures of interest (glomerular pressure, re, rvr, or AER) when FFA suppression is the predictor of interest. These models are additionally adjusted for HbA1c.

How to interpret the output:

In the R output, the `ffa_suppression:siteNetherlands` term tests whether there is an interaction effect between FFA suppression and site for the measure of interest. In instances where this effect is significant ($p < 0.05$), estimated marginal means and pairwise comparisons are provided. In the estimated marginal means output, the emmean reports an estimate of the effect of the measure of interest for each site. The pairwise comparison output provides p-values for testing whether there is a notable difference in the means between sites.


### Non-Transformed Models
```{r ffa suppression and site interaction mods, echo=FALSE}
cat("Association between glomerular pressure and FFA suppression by Site (Interaction), adjusted for HbA1c:")
pglo_ffa_suppress_mod <- lm(glomerular_pressure ~ ffa_suppression:site + ffa_suppression + site + hba1c, data = combined_clean)
lm_summary_and_diagnostics(pglo_ffa_suppress_mod)
# ffa_suppression:site not significant w this model so emmeans not computed

cat("Association between re and FFA suppression by Site (Interaction), adjusted for HbA1c:")
re_ffa_suppress_mod <- lm(re ~ ffa_suppression:site + ffa_suppression + site + hba1c, data = combined_clean)
lm_summary_and_diagnostics(re_ffa_suppress_mod)
# ffa_suppression:site not significant w this model so emmeans not computed

cat("Association between rvr and FFA suppression by Site (Interaction), adjusted for HbA1c:")
rvr_ffa_suppress_mod <- lm(rvr ~ ffa_suppression:site + ffa_suppression + site + hba1c, data = combined_clean)
lm_summary_and_diagnostics(rvr_ffa_suppress_mod)
m <- emmeans(rvr_ffa_suppress_mod, pairwise ~ site|ffa_suppression)
cat("Estimated Marginal Means for rvr:")
m$emmeans
cat("Pairwise Comparison for rvr:")
m$contrasts
```

### Log-transformed Models

Due to a notable right skew in the data, a log-transformation of the sex-adjusted AER and covariate-adjusted AER variables was performed. The below output reports results that have been back-transformed to the non-logged scale.  

$$\\[0.01in]$$

Association between sex-adjusted AER and FFA suppression by Site, adjusted for HbA1c:

```{r ffa suppression and site interaction mods w log sex adj AER, echo=FALSE}
AER_ffa_suppress_mod1 <- lm(log(AER_sex_adjusted) ~ ffa_suppression:site + ffa_suppression + site  + hba1c, data = combined_clean)

broom::tidy(AER_ffa_suppress_mod1, conf.int = F, exp = T) %>% 
  select(term, estimate, std.error, statistic, p.value) %>%
  mutate(p.value = format.pval(p.value, digits = 3, eps=0.001)) %>% 
  kable(caption = 'Model Summary:',
        col.names = c('Coefficients', 'Estimate', 'Std. Error', 't value', 'p-value'),
        booktabs = T, digits = 3,
        linesep = c('')) %>% 
  kable_styling(latex_options=c('HOLD_position', 'striped'))
# no emmeans added here because interaction not significant

plot(x = predict(AER_ffa_suppress_mod1), y = rstudent(AER_ffa_suppress_mod1), xlab = 'Predicted Value', ylab = 'Jackknife
     Residual', main = 'Scatterplot of Residuals'); abline(h = 0, lty = 2)

plot(ppoints(length(rstudent(AER_ffa_suppress_mod1))), sort(pnorm(rstudent(AER_ffa_suppress_mod1))), xlab='Observed Cumulative
Probability', ylab='Expected Cumulative Probability', main='Normal Probability Plot', cex=1); abline(a=0,b=1, col='gray65', lwd=1)

hist(rstudent(AER_ffa_suppress_mod1), xlab='Jackknife Residual', main='Histogram of Residuals', freq=F)
```

$$\\[0.01in]$$
Association between covariate-adjusted AER and FFA suppression by Site, adjusted for HbA1c:
```{r ffa suppression and site interaction mods w log cov adj AER, echo=FALSE}
AER_ffa_suppress_mod2 <- lm(log(AER_covariate_adjusted) ~ ffa_suppression:site + ffa_suppression + site + hba1c, data = combined_clean)

broom::tidy(AER_ffa_suppress_mod2, conf.int = F, exp = T) %>% 
  select(term, estimate, std.error, statistic, p.value) %>%
  mutate(p.value = format.pval(p.value, digits = 3, eps=0.001)) %>% 
  kable(caption = 'Model Summary:',
        col.names = c('Coefficients', 'Estimate', 'Std. Error', 't value', 'p-value'),
        booktabs = T, digits = 3,
        linesep = c('')) %>% 
  kable_styling(latex_options=c('HOLD_position', 'striped'))

plot(x = predict(AER_ffa_suppress_mod2), y = rstudent(AER_ffa_suppress_mod2), xlab = 'Predicted Value', ylab = 'Jackknife
     Residual', main = 'Scatterplot of Residuals'); abline(h = 0, lty = 2)

plot(ppoints(length(rstudent(AER_ffa_suppress_mod2))), sort(pnorm(rstudent(AER_ffa_suppress_mod2))), xlab='Observed Cumulative
Probability', ylab='Expected Cumulative Probability', main='Normal Probability Plot', cex=1); abline(a=0,b=1, col='gray65', lwd=1)

hist(rstudent(AER_ffa_suppress_mod2), xlab='Jackknife Residual', main='Histogram of Residuals', freq=F)

m <- emmeans(AER_ffa_suppress_mod2, c("site", "ffa_suppression"), type = "response")
prs <- pairs(m, adjust = "tukey", by = "ffa_suppression") 

m %>%
  kable(caption = 'Estimated Marginal Means for covariate-adjusted AER:', booktabs = TRUE, digits = 3,
        col.names = c('Site', 'FFA Suppression', 'emmean', 'SE', 'df', 'lower.CL', 'upper.CL')) %>%
  kable_styling(latex_options='basic', html_font='Arial') %>% 
      htmltools::HTML()

as.data.frame(prs) %>%
  select(-ffa_suppression) %>% 
  kable(caption = 'Pairwise Comparison for covariate-adjusted AER:', booktabs = TRUE, digits = 3,
        col.names = c('Contrast', 'Estimate', 'SE', 'df', 'null', 't.ratio', 'p.value')) %>% 
  kable_styling(latex_options='basic', html_font='Arial') %>% 
      htmltools::HTML()
```


# Questions/Notes

* `gir_190` = Raw M-Value (Colorado site)
* `gir_200` = GIR (Colorado site)
