---
title: "IHD analyses"
author: "Paige Dillon and Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include=FALSE}
library(dplyr)
library(tableone)
library(knitr)
library(ISLR)
library(stringr)
library(Hmisc)
library(emmeans)
library(car)

knitr::opts_chunk$set(echo = FALSE,warning = FALSE)

knitr::opts_chunk$set(echo = FALSE)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "E:/Petter Bjornstad/IHD/raw_data"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/TODAY subaward/"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = ""
}
knitr::opts_knit$set(root.dir = home_dir)
setwd(home_dir)
```

```{r, include=FALSE}
# read in raw data
data1 <- read.csv("./Final_Dataset_01.csv", na.strings = c(" ", "", "-99"))
data2 <- read.csv("./Table01data.csv", na.strings = c(" ", "", "-99"))
data3 <- read.csv("./Final_Dataset_02.csv", na.strings = c(" ", "", "-99"))
data4 <- read.csv("./Final_Dataset_03.csv", na.strings = c(" ", "", "-99"))

# merge
finaldata <- merge(data1, data2, by=("Subject_ID"), all.x = TRUE, all.y = TRUE)
finaldata2 <- merge(data3, data4, by="Subject_ID" , all.x = TRUE , all.y = TRUE)
finaldata2$X.FFA.Suppression <- NULL
merged_data <- merge(finaldata, finaldata2, by = intersect(names(finaldata), names(finaldata2))  ,  all.x = TRUE , all.y = TRUE)
merged_data$percent_FFA_suppression <- merged_data$X.FFA_Suppression..x.
merged_data$X_FFA_Supression <- NULL
merged_data$group <- ifelse(str_detect(merged_data$Subject_ID, "T"), "T2D", 
                            ifelse(str_detect(merged_data$Subject_ID, "L"), "Lean", "Obese"))
merged_data$group <- as.factor(merged_data$group)

# labels for table 01
label(merged_data$Sex)= "Sex"
label(merged_data$Age.at.time.of.Consent)= "Age at time of consent"
label(merged_data$Race..choice.American.Indian.or.Alaskan.Native.)= "Is your race American Indian or Alaskan Native?"
label(merged_data$Race..choice.Asian.)= "Is your race Asian?"
label(merged_data$Race..choice.Hawaiian.or.Pacific.Islander.)= "Is your race Hawaiian or Pacific Islander?"
label(merged_data$Race..choice.White.)= "Is your race white?"
label(merged_data$Race..choice.Unknown.)= "Race Unknown"
label(merged_data$Race..choice.Other.)= "Race Other"
label(merged_data$Race..if.Other)= "Is your race different than those previously asked?"
label(merged_data$Race..choice.Black.or.African.American.)= "Is your race Black or African American?"
label(merged_data$Ethnicity..choice.Hispanic.)= "Are you hispanic?"
label(merged_data$Ethnicity..choice.Non.Hispanic.)= "Are you non-Hispanic?"
label(merged_data$Ethnicity..choice.Unknown.Not.Reported.)= "Ethnicity Unknown or Not Reported"
label(merged_data$Screening.A1c.....)= "Screening A1C"
label(merged_data$Length.of.Diabetes.Diagnosis..months.)= "Length of diabetes diagnosis in months"
label(merged_data$Height..cm.)= "Height in Cm"
label(merged_data$Weight_kg)= "Weight in kg"
label(merged_data$BMI)= "BMI"
label(merged_data$Body.Fat....)= "Body Fat"
label(merged_data$X.Avg..Waist.Circumference..cm.)= "Average Waist Circumference"
label(merged_data$Baseline.Cholesterol..mg.dl.)= "Baseline Cholesterol (mg/dl)"
label(merged_data$Baseline.HDL..mg.dl.)= "Baseline HDL (mg/dl)"
label(merged_data$Baseline.LDL..mg.dl.)= "Baseline LDL (mg/dl)"
label(merged_data$Baseline.Triglycerides..mg.dl.)= "Baseline Triglycerides (mg/dl)"
label(merged_data$Systolic.Blood.Pressure..SBP.)= "Systolic Blood Pressure (SBP)"
label(merged_data$Diastolic.Blood.Pressure..DBP.)= "Diastolic Blood Pressure (DBP)"
label(merged_data$Mean.Arterial.Pressure..MAP.)= "Mean Arterial Pressure (MAP)"
label(merged_data$Serum.Creatinine...mg.dl.)= "Serum Creatinine (mg/dl)"
label(merged_data$Baseline.Serum.Cystatin.C..clamp.)= "Baseline Serum Cystatin"
label(merged_data$BSA.normalized.GFR)= "BSA normalized GFR"
label(merged_data$Absolute.GFR)= "Absolute GFR"
label(merged_data$Absolute.ERPF)= "Absolute ERPF"
label(merged_data$BSA.normalized.ERPF)= "BSA normalized ERPF"
label(merged_data$O2_Consumption._cortex)= "O2 consumption- cortex"
label(merged_data$O2_Consumption_medulla)= "O2 consumption medulla"
label(merged_data$O2_Consumption_Kidney)= "O2 consumption kidney"

#Labeling the other variables
label(merged_data$Subject_ID)= "Subject ID"
label(merged_data$ACR_baseline)= "ACR Baseline"
label(merged_data$GFR_BSAnormalized)= "GFR BSA Normalized"
label(merged_data$ERPF_BSAnormalized)= "ERPF BSA Normalized"
label(merged_data$Gender)= "Gender"
label(merged_data$Date.of.Diagnosis.with.Diabetes)= "Date of Diagnosis with Diabetes"
label(merged_data$Weight..kg.)= "Weight (kg)"
label(merged_data$Fat.Mass..kg.)= "Fat Mass (kg)"
label(merged_data$Notes)= "Notes"
label(merged_data$AIRg)= "AIRg"
label(merged_data$SS.insulin)= "SS. insulin"
label(merged_data$Disposition.index..M.I.xAIRg)= "Disposition Index (M.I.x AIRg)"
label(merged_data$Baseline.Serum.creatinine..clamp.)= "Baseline Serum Creatinine"
label(merged_data$Baseline.ACR)= "Baseline ACR"
label(merged_data$percent_FFA_suppression)= "Percent.FFA. Suppression"
label(merged_data$Date.of.Birth)= "Date of Birth"
label(merged_data$Baseline.Cystatin.C..mg.L.)= "Baseline Cystatin (mg/L)"
label(merged_data$ACPRg..nmol.L.)= "ACPRg (nmol/L)"
label(merged_data$SS.CP)= " SS CP"
label(merged_data$GIR)= "GIR"
label(merged_data$group)= "Group"

# make dataframe with only T2D and obese
merged_data_T2D_obese <- merged_data %>% filter(group %in% c("T2D","Obese"))
merged_data_T2D_obese$group <- as.factor(merged_data_T2D_obese$group)
merged_data_T2D_obese$group <- droplevels(merged_data_T2D_obese$group)
merged_data_T2D_obese$Sex <- as.factor(merged_data_T2D_obese$Sex)
```

# Questions for Carson

1. We did not find oxygen availability in the dataset.
2. For albuminuria, we used variable "ACR_baseline" - is that correct?

# Results

## Descriptive statistics

```{r echo=FALSE, include=FALSE}
t1 <- CreateTableOne(data=merged_data, vars=c('Sex','Age.at.time.of.Consent','Race..choice.American.Indian.or.Alaskan.Native.','Race..choice.Asian.', 'Race..choice.Black.or.African.American.','Race..choice.White.', 'Race..choice.Unknown.', 'Race..choice.Other.', 'Ethnicity..choice.Hispanic.', 'Ethnicity..choice.Non.Hispanic.','Screening.A1c.....','Height..cm.', 'Weight_kg', 'BMI', 'Body.Fat....', 'X.Avg..Waist.Circumference..cm.','Baseline.Cholesterol..mg.dl.', 'Baseline.HDL..mg.dl.', 'Baseline.LDL..mg.dl.', 'Baseline.Triglycerides..mg.dl.', 'Systolic.Blood.Pressure..SBP.', 'Diastolic.Blood.Pressure..DBP.','Mean.Arterial.Pressure..MAP.', 'Serum.Creatinine...mg.dl.', 'Baseline.Serum.Cystatin.C..clamp.','BSA.normalized.GFR', 'Absolute.GFR', 'Absolute.ERPF','BSA.normalized.ERPF', 'O2_Consumption._cortex', 'O2_Consumption_medulla','O2_Consumption_Kidney'), strata="group")
t1 <- print(t1, showAllLevels=T)
```

```{r echo=FALSE, include=TRUE}
kable(t1)
```

## Measures of insulin sensitivity, adjusted for sex

```{r echo=FALSE,comment=''}
sivars <- c("SS.CP","SS.insulin","GIR","AIRg","Disposition.index..M.I.xAIRg","percent_FFA_suppression")

# group comparisons adjusted for sex
for (v in sivars) {
  form = as.formula(paste0(v,"~group + Sex"))
  mod <- lm(form,data = merged_data_T2D_obese)
  print(noquote(c("Adjusted model for: ", noquote(v))))
  print(summary(mod))
  means <- emmeans(mod,"group")
  print(means)
}
```

## GFR and ERPF, adjusted for sex and A1c

```{r echo=FALSE,comment=''}
gfr_erpf_vars <- c("Absolute.GFR","BSA.normalized.GFR","Absolute.ERPF","BSA.normalized.ERPF")

# group comparisons adjusted for sex
for (v in gfr_erpf_vars) {
  form = as.formula(paste0(v,"~group + Sex + Screening.A1c....."))
  mod <- lm(form,data = merged_data_T2D_obese)
  print(noquote(c("Adjusted model for: ", noquote(v))))
  print(summary(mod))
  means <- emmeans(mod,"group")
  print(means)
}
```

## O2 consumption, adjusted for sex, A1c, and weight

```{r echo=FALSE,comment=''}
o2_vars <- c("O2_Consumption._cortex","O2_Consumption_medulla","O2_Consumption_Kidney")

# group comparisons adjusted for sex
for (v in o2_vars) {
  form = as.formula(paste0(v,"~group + Sex + Screening.A1c..... + Weight..kg."))
  mod <- lm(form,data = merged_data)
  print(noquote(c("Adjusted model for: ", noquote(v))))
  print(summary(mod))
  means <- emmeans(mod,"group")
  print(means)
 }
```

## FFA suppression vs. albuminuria and renal oxygen consumption

```{r echo=FALSE,comment=''}
ffa_vars <- c("O2_Consumption._cortex","O2_Consumption_medulla","O2_Consumption_Kidney","ACR_baseline")

# group comparisons adjusted for sex
for (v in ffa_vars) {
  form = as.formula(paste0(v,"~percent_FFA_suppression + Sex + Screening.A1c..... + Weight..kg."))
  mod <- lm(form,data = merged_data)
  print(noquote(c("Adjusted model for: ", noquote(v))))
  print(summary(mod))
  plot(mod)
}
```
