---
title: "Intraglomerular hemodynamic models"
author: "Cameron Severn and Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include=FALSE}
library(knitr)
library(dplyr)

knitr::opts_chunk$set(echo = FALSE,warning = FALSE)


data <- read.csv("E:\\Petter Bjornstad\\IHD\\raw_data\\Final Dataset for analysis.csv")

# there are 2 outliers
# RH-77-T has ACR>1000
# RH-42-O has O2 consumption cortex >8 and total kidney O2 consumption >9
# try removing them and fitting linear model
data_modeling <- data %>% filter(!Subject.ID %in% c("RH-77-T","RH-42-O"))

# translate and transform
data_log <- data
data_log$l.Percent.FFA.Suppression <- log(data$Percent.FFA.Suppression+11)
data_log$l.O2.Consumption...pre.post.furosemide.R2....medulla <- log(data$O2.Consumption...pre.post.furosemide.R2....medulla.+11)
data_log$l.O2.Consumption...pre.post.furosemide.R2....cortex <- log(data$O2.Consumption..pre.post.furosemide.R2....cortex.+11)
data_log$l.O2.Consumption...pre.post.furosemide.R2....kidney <- log(data$O2.Consumption...pre.post.furosemide.R2....kidney.+11)
data_log$l.Baseline.ACR <- log(data$Baseline.ACR+11)

# inverse hypberbolic sine
data_ihs <- data
data_ihs$ihs.Percent.FFA.Suppression <- asinh(data$Percent.FFA.Suppression)
data_ihs$ihs.Baseline.ACR <- asinh(data$Baseline.ACR)
data_ihs$ihs.O2.Consumption...pre.post.furosemide.R2....medulla <- asinh(data$O2.Consumption...pre.post.furosemide.R2....medulla)
data_ihs$ihs.O2.Consumption...pre.post.furosemide.R2....cortex <- asinh(data$O2.Consumption..pre.post.furosemide.R2....cortex.)
data_ihs$ihs.O2.Consumption...pre.post.furosemide.R2....kidney <- asinh(data$O2.Consumption...pre.post.furosemide.R2....kidney.)
```

# Models on untransformed data

## Baseline ACR

```{r echo=FALSE,comment=""}
mod <- lm(Baseline.ACR ~ Percent.FFA.Suppression,data=data)
summary(mod)
plot(mod)
```

## O2 consumption cortex

```{r echo=FALSE,comment=""}
mod <- lm(O2.Consumption..pre.post.furosemide.R2....cortex. ~ Percent.FFA.Suppression,data=data)
summary(mod)
plot(mod)
```

## O2 consumption medulla

```{r echo=FALSE,comment=""}
mod <- lm(O2.Consumption...pre.post.furosemide.R2....medulla. ~ Percent.FFA.Suppression,data=data)
summary(mod)
plot(mod)
```

## O2 consumption whole kidney

```{r echo=FALSE,comment=""}
mod <- lm(O2.Consumption...pre.post.furosemide.R2....kidney. ~ Percent.FFA.Suppression,data=data)
summary(mod)
plot(mod)
```

# Models on log transformed data

## Baseline ACR

```{r echo=FALSE,comment=""}
mod <- lm(l.Baseline.ACR ~ l.Percent.FFA.Suppression,data=data_log)
summary(mod)
plot(mod)
```

## O2 consumption cortex

```{r echo=FALSE,comment=""}
mod <- lm(l.O2.Consumption...pre.post.furosemide.R2....cortex ~ l.Percent.FFA.Suppression,data=data_log)
summary(mod)
plot(mod)
```

## O2 consumption medulla

```{r echo=FALSE,comment=""}
mod <- lm(l.O2.Consumption...pre.post.furosemide.R2....medulla ~ l.Percent.FFA.Suppression,data=data_log)
summary(mod)
plot(mod)
```

## O2 consumption whole kidney

```{r echo=FALSE,comment=""}
mod <- lm(l.O2.Consumption...pre.post.furosemide.R2....kidney ~ l.Percent.FFA.Suppression,data=data_log)
summary(mod)
plot(mod)
```

# Models on inverse hyperbolic sine transformed data

## Baseline ACR

```{r echo=FALSE,comment=""}
mod <- lm(ihs.Baseline.ACR ~ ihs.Percent.FFA.Suppression,data=data_ihs)
summary(mod)
plot(mod)
```

## O2 consumption cortex

```{r echo=FALSE,comment=""}
mod <- lm(ihs.O2.Consumption...pre.post.furosemide.R2....cortex ~ ihs.Percent.FFA.Suppression,data=data_ihs)
summary(mod)
plot(mod)
```

## O2 consumption medulla

```{r echo=FALSE,comment=""}
mod <- lm(ihs.O2.Consumption...pre.post.furosemide.R2....medulla ~ ihs.Percent.FFA.Suppression,data=data_ihs)
summary(mod)
plot(mod)
```

## O2 consumption whole kidney

```{r echo=FALSE,comment=""}
mod <- lm(ihs.O2.Consumption...pre.post.furosemide.R2....kidney ~ ihs.Percent.FFA.Suppression,data=data_ihs)
summary(mod)
plot(mod)
```