---
title: "Adipose insulin sensitivity in adolescents with extreme insulin resistance"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:   
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include=FALSE}
library(tableone)
library(knitr)
library(dplyr)
library(ggplot2)
library(hrbrthemes)
library(tidyr)

knitr::opts_chunk$set(echo = FALSE,warning = FALSE, message = FALSE)

data <- read.csv("E:/Melanie Green/Frontiers clamp methods review paper/Data raw/Data template for paper_MelanieLaura.csv",
                 na.strings = c("NA",".",""," ","#DIV/0!"))
data$Age.diabetes.onset <- NULL
data$BMI.category <- NULL
data$Sex <- as.factor(data$Sex)
data$bmi_cat <- ifelse(is.na(data$BMIp),NA,
                       ifelse(data$BMIp<85,"<85th %ile",
                              ifelse(data$BMIp>=85 & data$BMIp<95,"85th - <95th %ile",">=95th %ile")))
data$bmi_twocat <- ifelse(is.na(data$bmi_cat),NA,
                          ifelse(data$bmi_cat=="<85th %ile","Lean","Overweight/Obese"))

# exclude those with missing basal glucose
data <- data %>% filter(!is.na(Mean.Glucose.Basal))

# exclude 6377-31 due to bad insulin assay
data <- data %>% filter(!StudyID=="6377-31")

# fix one insulin value per Melanie
data$mean.16.insulin <- ifelse(data$StudyID=="827-110",NA,data$mean.16.insulin)

# allow only 1 time point missing, and must have 10 stage (based on insulin and glucose?)
data$nmissgluc <- apply(data[,c("Mean.Glucose.Basal","mean.10.glucose","mean.16.glucose","mean.80.glucose")], 1, function(x) sum(is.na(x)))
data$nmissins <- apply(data[,c("Basal.Insulin","mean.10.insulin","mean.16.insulin","mean.80.insulin")], 1, function(x) sum(is.na(x)))  
data <- data %>% filter(!is.na(mean.10.glucose) | !is.na(mean.10.insulin))
data <- data %>% filter(nmissgluc<=1)  
data <- data %>% filter(nmissins<=1)  

# categorize overnight glucose
data$Mean.Glucose.Basal.cat <- ifelse(is.na(data$Mean.Glucose.Basal),NA,
                                      ifelse(data$Mean.Glucose.Basal>70 & data$Mean.Glucose.Basal<90,"70 - <90 mg/dL",
                                             ifelse(data$Mean.Glucose.Basal>=90 & data$Mean.Glucose.Basal<110,"90 - <110 mg/dL",
                                                    ifelse(data$Mean.Glucose.Basal>=110 & data$Mean.Glucose.Basal<130,"110 - <130 mg/dL",
                                                           ifelse(data$Mean.Glucose.Basal>=130 & data$Mean.Glucose.Basal<150,"130 - <150 mg/dl",
                                                                  ">150 mg/dl")))))

# exclude those with baseline glucose >150
data <- data %>% filter(Mean.Glucose.Basal.cat != ">150 mg/dl")
data$Mean.Glucose.Basal.cat <- as.factor(data$Mean.Glucose.Basal.cat)
data$Mean.Glucose.Basal.cat <- droplevels(data$Mean.Glucose.Basal.cat)

data$Mean.Glucose.Basal.cat <- factor(data$Mean.Glucose.Basal.cat, 
                                      c("70 - <90 mg/dL","90 - <110 mg/dL","110 - <130 mg/dL","130 - <150 mg/dl"))

# categorize by whether glycerol Ra 10 is lower than basal glycerol Ra
data$glycerolRa_10_lt_0 <- ifelse(data$Glycerol.Ra.10<data$Basal.glycerol.RA,"Glycerol Ra 10 < Glycerol Ra 0","Glycerol Ra 10 >= Glycerol Ra 0")
data$glycerolRa_10_lt_0 <- as.factor(data$glycerolRa_10_lt_0)
# stratify by median percentage difference between max glucose Ra and glucose Ra 16
x <- data[,c("StudyID","Basal.glucose.Ra","glucose.Ra.10","glucose.Ra.16","glucose.Ra.80")]
x$max_glucra <- apply(x[,2:5],1,max)
x$glucra_diff <- x$max_glucra-x$glucose.Ra.16
x <- x %>% select(StudyID,glucra_diff)
data <- merge(data,x,by="StudyID",all.x = T,all.y=T)
median_diff <- median(x$glucra_diff,na.rm = T)
data$glucoseRa_diff_cat <- ifelse(data$glucra_diff<median_diff,"Max glucose Ra - glucose Ra 16 < median","Max glucose Ra - glucose Ra 16 >= median")
data$glucoseRa_diff_cat <- as.factor(data$glucoseRa_diff_cat)

# calculate means by BMI group
means <- data %>% group_by(bmi_twocat,Mean.Glucose.Basal.cat) %>% summarise_at(vars("Mean.Glucose.Basal","Basal.Insulin",
                                                             "Basal.glucose.Ra","Basal.FFA","Basal.glycerol",
                                                             "Basal.glycerol.RA"),mean,na.rm=T)

# need to make a long dataset
for_long <- data %>% select(-c(Sex,Age,FFA.IC50,Glycerol.IC50,Glycerol.Ra.IC50,Race,Ethnicity,Diabetes.duration_years,Approximate.total.daily.insulin,
                               BMI,BMIp,Dexa...Lean.mass,Dexa...fat.mass,Most.recent.HbA1c,Cholesterol,Triglycerides,LDL,bmi_cat,
                               bmi_twocat,Mean.Glucose.Basal.cat,M.I.80,Glucose.IC50,nmissgluc,nmissins,
                               glucoseRa_diff_cat,glycerolRa_10_lt_0,glucra_diff))
colnames(for_long) <- c("StudyID","Glucose.0","Glucose.10","Glucose.16","Glucose.80","Insulin.0","Insulin.10",
                        "Insulin.16","Insulin.80","GlucoseRa.0","GlucoseRa.10","GlucoseRa.16","GlucoseRa.80","FFA.0",
                        "FFA.10","FFA.16","FFA.80","Glycerol.0","Glycerol.10","Glycerol.16","Glycerol.80","GlycerolRa.0",
                        "GlycerolRa.10","GlycerolRa.16","GlycerolRa.80")
long <- reshape(for_long, idvar = "StudyID", varying=2:25, direction="long",sep=".")
long <- long %>% arrange(StudyID,time)
# merge in overnight glucose category
x <- data %>% select(StudyID,Mean.Glucose.Basal.cat)
long <- merge(long,x,by="StudyID",all.x = T,all.y = T)
long$time <- as.factor(long$time)

my_order <- c("70 - <90 mg/dL","90 - <110 mg/dL","110 - <130 mg/dL","130 - <150 mg/dl",">150 mg/dl")

# function to calculate predicted value at mean insulin
# returns dataframe with StudyID and y_hat
calc_predicted <- function(data, outcome, stage) {
  y_hat <- NULL
  df = data
  stage = stage
  insmean = df %>% filter(time==stage) %>% summarise(log(mean(Insulin,na.rm=T)))
  ids = as.factor(unique(df$StudyID))
  for (i in levels(ids)) {
    df_i = df[df$StudyID==i,]
    # Predictor matrix
    X = log(df_i[,"Insulin"])
    # Outcome
    Y = log(df_i[,outcome]+5)
    nonmiss = sum(!is.na(Y))
    print(nonmiss)
    if (nonmiss>2) {
      # Regression 
      mod = lm(as.numeric(Y) ~ X)
      int = coef(mod)[1]
      slope = coef(mod)[2]
      y_hat_i <- exp(int + (insmean*slope)) + 5
    } else {y_hat_i <- NA}
    y_hat <- c(y_hat,y_hat_i)
  }
  y_hat <- data.frame(cbind(unlist(unique(df$StudyID)),unlist(y_hat)))
  y_hat[,2] <- as.numeric(y_hat[,2])
  colnames(y_hat) <- c("StudyID",paste0("y_hat_",outcome))
  return(y_hat)
}

predgluc <- calc_predicted(data = long, outcome = "Glucose", stage=16)
predglucra <- calc_predicted(data = long, outcome = "GlucoseRa", stage=16)
predglyc <- calc_predicted(data = long, outcome = "Glycerol", stage=10)
predglycra <- calc_predicted(data = long, outcome = "GlycerolRa", stage=10)
predffa <- calc_predicted(data = long, outcome = "FFA", stage=10)
preds <- merge(predgluc,predglucra,by="StudyID", all.x = T, all.y = T)
preds <- merge(preds,predglyc,by="StudyID", all.x = T, all.y = T)
preds <- merge(preds,predglycra,by="StudyID", all.x = T, all.y = T)
preds <- merge(preds,predffa,by="StudyID", all.x = T, all.y = T)
data <- merge(data,preds,by="StudyID",all.x = T,all.y = T)

```

# Descriptive statistics by BMI category

```{r, include=F, warning=FALSE}
descdata <- data %>% select(-StudyID,glucra_diff )
t1 <- CreateTableOne(data=descdata,vars=colnames(descdata),strata = "bmi_twocat")
t1 <- print(t1, nonnorm=colnames(select_if(descdata, is.numeric)))
```

```{r, include=T, warning=FALSE}
kable(t1)
```

# Descriptive statistics by overnight mean glucose

```{r, include=F, warning=FALSE}
t1a <- CreateTableOne(data=descdata,vars=colnames(descdata),strata = "Mean.Glucose.Basal.cat")
t1a <- print(t1a, nonnorm=colnames(select_if(descdata, is.numeric)))
```

```{r, include=T, warning=FALSE}
kable(t1a)
```

# Descriptive statistics by glycerol Ra category

```{r, include=F, warning=FALSE}
t1b <- CreateTableOne(data=descdata,vars=colnames(descdata),strata = "glycerolRa_10_lt_0")
t1b <- print(t1b, nonnorm=colnames(select_if(descdata, is.numeric)))
```

```{r, include=T, warning=FALSE}
kable(t1b)
```

# Descriptive statistics by glucose Ra category

```{r, include=F, warning=FALSE}
t1c <- CreateTableOne(data=descdata,vars=colnames(descdata),strata = "glucoseRa_diff_cat")
t1c <- print(t1c, nonnorm=colnames(select_if(descdata, is.numeric)))
```

```{r, include=T, warning=FALSE}
kable(t1c)
```

# Figure 1 - scatterplots of fasting data

## Glucose

### Raw data

```{r, include=T, warning=FALSE}
ggplot(data, aes(x=Basal.Insulin, y=Mean.Glucose.Basal, color=bmi_twocat)) + 
      geom_point(size=6) +
      theme_ipsum()
```

### Means

```{r, include=T, warning=FALSE}
ggplot(means, aes(x=Basal.Insulin, y=Mean.Glucose.Basal, shape=bmi_twocat, color=Mean.Glucose.Basal.cat)) + 
      geom_point(size=6) +
      theme_ipsum()
```

## Glucose Ra

### Raw data

```{r, include=T, warning=FALSE}
ggplot(data, aes(x=Basal.Insulin, y=Basal.glucose.Ra, color=bmi_twocat)) + 
      geom_point(size=6) +
      theme_ipsum()
```

### Means

```{r, include=T, warning=FALSE}
ggplot(means, aes(x=Basal.Insulin, y=Basal.glucose.Ra, shape=bmi_twocat, color=Mean.Glucose.Basal.cat)) + 
      geom_point(size=6) +
      theme_ipsum()
```

## Glycerol

### Raw data

```{r, include=T, warning=FALSE}
ggplot(data, aes(x=Basal.Insulin, y=Basal.glycerol, color=bmi_twocat)) + 
      geom_point(size=6) +
      theme_ipsum()
```

### Means

```{r, include=T, warning=FALSE}
ggplot(means, aes(x=Basal.Insulin, y=Basal.glycerol, shape=bmi_twocat, color=Mean.Glucose.Basal.cat)) + 
      geom_point(size=6) +
      theme_ipsum()
```

## Glycerol Ra

### Raw data

```{r, include=T, warning=FALSE}
ggplot(data, aes(x=Basal.Insulin, y=Basal.glycerol.RA, color=bmi_twocat)) + 
      geom_point(size=6) +
      theme_ipsum()
```

### Means

```{r, include=T, warning=FALSE}
ggplot(means, aes(x=Basal.Insulin, y=Basal.glycerol.RA, shape=bmi_twocat, color=Mean.Glucose.Basal.cat)) + 
      geom_point(size=6) +
      theme_ipsum()
```

## FFA

### Raw data

```{r, include=T, warning=FALSE}
ggplot(data, aes(x=Basal.FFA, y=Basal.glycerol.RA, color=bmi_twocat)) + 
      geom_point(size=6) +
      theme_ipsum()
```

### Means

```{r, include=T, warning=FALSE}
ggplot(means, aes(x=Basal.FFA, y=Basal.glycerol.RA, shape=bmi_twocat, color=Mean.Glucose.Basal.cat)) + 
      geom_point(size=6) +
      theme_ipsum()
```

## Glycerol and FFA

### Means

```{r, include=T, warning=FALSE}
ggplot(means, aes(x=Basal.FFA, y=Basal.glycerol.RA, shape=bmi_twocat, color=Mean.Glucose.Basal.cat)) + 
      geom_point(size=6) +
      theme_ipsum()
```

# Figure 2 - boxplots by overnight glucose and stage

## Glucose

```{r, include=T, warning=FALSE}
p <- ggplot(long, aes(x=Mean.Glucose.Basal.cat, y=Glucose, fill=time)) +
      geom_boxplot() +
      scale_x_discrete(limits = function(x) my_order[my_order %in% x]) + 
      xlab("Overnight glucose") + ylab("Glucose") + theme_classic()
p
```

## Insulin

```{r, include=T, warning=FALSE}
p <- ggplot(long, aes(x=Mean.Glucose.Basal.cat, y=Insulin, fill=time)) +
      geom_boxplot() +
      scale_x_discrete(limits = function(x) my_order[my_order %in% x]) + 
      xlab("Overnight glucose") + ylab("Insulin") + theme_classic()
p
```

## Glucose Ra

```{r, include=T, warning=FALSE}
p <- ggplot(long, aes(x=Mean.Glucose.Basal.cat, y=GlucoseRa, fill=time)) +
      geom_boxplot() +
      scale_x_discrete(limits = function(x) my_order[my_order %in% x]) + 
      xlab("Overnight glucose") + ylab("Glucose Ra") + theme_classic()
p
```

## Glycerol

```{r, include=T, warning=FALSE}
p <- ggplot(long, aes(x=Mean.Glucose.Basal.cat, y=Glycerol, fill=time)) +
      geom_boxplot() +
      scale_x_discrete(limits = function(x) my_order[my_order %in% x]) + 
      xlab("Overnight glucose") + ylab("Glycerol") + theme_classic()
p
```

## Glycerol Ra

```{r, include=T, warning=FALSE}
p <- ggplot(long, aes(x=Mean.Glucose.Basal.cat, y=GlycerolRa, fill=time)) +
      geom_boxplot() +
      scale_x_discrete(limits = function(x) my_order[my_order %in% x]) + 
      xlab("Overnight glucose") + ylab("Glycerol Ra") + theme_classic()
p
```

## FFA

```{r, include=T, warning=FALSE}
p <- ggplot(long, aes(x=Mean.Glucose.Basal.cat, y=FFA, fill=time)) +
      geom_boxplot() +
      scale_x_discrete(limits = function(x) my_order[my_order %in% x]) + 
      xlab("Overnight glucose") + ylab("FFA") + theme_classic()
p
```

# Figure 3 - curves over time by overnight glucose

## Spaghetti plots

### Glucose

```{r, include=T, warning=FALSE}
p <- ggplot(long, aes(x=time, y=Glucose, group=StudyID, color=Mean.Glucose.Basal.cat)) +
      geom_line() 
p
```

### Insulin

```{r, include=T, warning=FALSE}
p <- ggplot(long, aes(x=time, y=Insulin, group=StudyID, color=Mean.Glucose.Basal.cat,)) +
      geom_line() 
p
```

### Glucose Ra

```{r, include=T, warning=FALSE}
p <- ggplot(long, aes(x=time, y=GlucoseRa, group=StudyID, color=Mean.Glucose.Basal.cat,)) +
      geom_line() 
p
```

### Glycerol

```{r, include=T, warning=FALSE}
p <- ggplot(long, aes(x=time, y=Glycerol, group=StudyID, color=Mean.Glucose.Basal.cat,)) +
      geom_line() 
p
```

### Glycerol Ra

```{r, include=T, warning=FALSE}
p <- ggplot(long, aes(x=time, y=GlycerolRa, group=StudyID, color=Mean.Glucose.Basal.cat,)) +
      geom_line() 
p
```

### FFA

```{r, include=T, warning=FALSE}
p <- ggplot(long, aes(x=time, y=FFA, group=StudyID, color=Mean.Glucose.Basal.cat,)) +
      geom_line() 
p
```

## Mean curves by time

### Glucose

```{r, include=T, warning=FALSE}
grouped <- long %>% group_by(Mean.Glucose.Basal.cat,time) %>% summarise(glucose_mean = mean(Glucose,na.rm=T))

p <- ggplot(long, aes(x=time, y=Glucose, group=StudyID, color=Mean.Glucose.Basal.cat), alpha=0.3, size=0.5) +
      geom_line() +
      stat_summary(fun=mean, aes(group=Mean.Glucose.Basal.cat), geom="line", alpha=0.8, size=3)
p
```

### Insulin

```{r, include=T, warning=FALSE}
grouped <- long %>% group_by(Mean.Glucose.Basal.cat,time) %>% summarise(insulin_mean = mean(Insulin,na.rm=T))

p <- ggplot(long, aes(x=time, y=Insulin, group=StudyID, color=Mean.Glucose.Basal.cat), alpha=0.3, size=0.5) +
      geom_line() +
      stat_summary(fun=mean, aes(group=Mean.Glucose.Basal.cat), geom="line", alpha=0.8, size=3)
p
```

### Glucose Ra

```{r, include=T, warning=FALSE}
grouped <- long %>% group_by(Mean.Glucose.Basal.cat,time) %>% summarise(glucosera_mean = mean(GlucoseRa,na.rm=T))

p <- ggplot(long, aes(x=time, y=GlucoseRa, group=StudyID, color=Mean.Glucose.Basal.cat), alpha=0.3, size=0.5) +
      geom_line() +
      stat_summary(fun=mean, aes(group=Mean.Glucose.Basal.cat), geom="line", alpha=0.8, size=3)
p
```

### Glycerol

```{r, include=T, warning=FALSE}
grouped <- long %>% group_by(Mean.Glucose.Basal.cat,time) %>% summarise(glycerol_mean = mean(Glycerol,na.rm=T))

p <- ggplot(long, aes(x=time, y=Glycerol, group=StudyID, color=Mean.Glucose.Basal.cat), alpha=0.3, size=0.5) +
      geom_line() +
      stat_summary(fun=mean, aes(group=Mean.Glucose.Basal.cat), geom="line", alpha=0.8, size=3)
p
```

### Glycerol Ra

```{r, include=T, warning=FALSE}
grouped <- long %>% group_by(Mean.Glucose.Basal.cat,time) %>% summarise(glycerolra_mean = mean(GlycerolRa,na.rm=T))

p <- ggplot(long, aes(x=time, y=GlycerolRa, group=StudyID, color=Mean.Glucose.Basal.cat), alpha=0.3, size=0.5) +
      geom_line() +
      stat_summary(fun=mean, aes(group=Mean.Glucose.Basal.cat), geom="line", alpha=0.8, size=3)
p
```

### FFA

```{r, include=T, warning=FALSE}
grouped <- long %>% group_by(Mean.Glucose.Basal.cat,time) %>% summarise(ffa_mean = mean(FFA,na.rm=T))

p <- ggplot(long, aes(x=time, y=FFA, group=StudyID, color=Mean.Glucose.Basal.cat), alpha=0.3, size=0.5) +
      geom_line() +
      stat_summary(fun=mean, aes(group=Mean.Glucose.Basal.cat), geom="line", alpha=0.8, size=3)
p
```

## Mean curves by insulin

### Glucose

```{r, include=T, warning=FALSE}
grouped_gluc <- long %>% group_by(Mean.Glucose.Basal.cat,time) %>% summarise(glucose_mean = mean(Glucose,na.rm=T))
grouped_ins <- long %>% group_by(Mean.Glucose.Basal.cat,time) %>% summarise(insulin_mean = mean(Insulin,na.rm=T))
grouped <- merge(grouped_gluc, grouped_ins, by=c("Mean.Glucose.Basal.cat","time"), all.x=T, all.y=T)

p <- ggplot(grouped, aes(x=insulin_mean, y=glucose_mean, group=Mean.Glucose.Basal.cat, color=Mean.Glucose.Basal.cat)) +
      geom_line(alpha=0.8, size=3) +
      scale_color_discrete(breaks=c("70 - <90 mg/dL","90 - <110 mg/dL","90 - <110 mg/dL","110 - <130 mg/dL","130 - <150 mg/dl",">150 mg/dl"))
p
```

### Glucose Ra

```{r, include=T, warning=FALSE}
grouped1 <- long %>% group_by(Mean.Glucose.Basal.cat,time) %>% summarise(glucosera_mean = mean(GlucoseRa,na.rm=T))
grouped2 <- long %>% group_by(Mean.Glucose.Basal.cat,time) %>% summarise(insulin_mean = mean(Insulin,na.rm=T))
grouped <- merge(grouped1, grouped2, by=c("Mean.Glucose.Basal.cat","time"), all.x=T, all.y=T)

p <- ggplot(grouped, aes(x=insulin_mean, y=glucosera_mean, group=Mean.Glucose.Basal.cat, color=Mean.Glucose.Basal.cat)) +
      geom_line(alpha=0.8, size=3) +
      scale_color_discrete(breaks=c("70 - <90 mg/dL","90 - <110 mg/dL","90 - <110 mg/dL","110 - <130 mg/dL","130 - <150 mg/dl",">150 mg/dl"))
p
```

### Glycerol

```{r, include=T, warning=FALSE}
grouped1 <- long %>% group_by(Mean.Glucose.Basal.cat,time) %>% summarise(glycerol_mean = mean(Glycerol,na.rm=T))
grouped2 <- long %>% group_by(Mean.Glucose.Basal.cat,time) %>% summarise(insulin_mean = mean(Insulin,na.rm=T))
grouped <- merge(grouped1, grouped2, by=c("Mean.Glucose.Basal.cat","time"), all.x=T, all.y=T)

p <- ggplot(grouped, aes(x=insulin_mean, y=glycerol_mean, group=Mean.Glucose.Basal.cat, color=Mean.Glucose.Basal.cat)) +
      geom_line(alpha=0.8, size=3) +
      scale_color_discrete(breaks=c("70 - <90 mg/dL","90 - <110 mg/dL","90 - <110 mg/dL","110 - <130 mg/dL","130 - <150 mg/dl",">150 mg/dl"))
p
```

### Glycerol Ra

```{r, include=T, warning=FALSE}
grouped1 <- long %>% group_by(Mean.Glucose.Basal.cat,time) %>% summarise(glycerolra_mean = mean(GlycerolRa,na.rm=T))
grouped2 <- long %>% group_by(Mean.Glucose.Basal.cat,time) %>% summarise(insulin_mean = mean(Insulin,na.rm=T))
grouped <- merge(grouped1, grouped2, by=c("Mean.Glucose.Basal.cat","time"), all.x=T, all.y=T)

p <- ggplot(grouped, aes(x=insulin_mean, y=glycerolra_mean, group=Mean.Glucose.Basal.cat, color=Mean.Glucose.Basal.cat)) +
      geom_line(alpha=0.8, size=3) +
      scale_color_discrete(breaks=c("70 - <90 mg/dL","90 - <110 mg/dL","90 - <110 mg/dL","110 - <130 mg/dL","130 - <150 mg/dl",">150 mg/dl"))
p
```

### FFA

```{r, include=T, warning=FALSE}
grouped1 <- long %>% group_by(Mean.Glucose.Basal.cat,time) %>% summarise(ffa_mean = mean(FFA,na.rm=T))
grouped2 <- long %>% group_by(Mean.Glucose.Basal.cat,time) %>% summarise(insulin_mean = mean(Insulin,na.rm=T))
grouped <- merge(grouped1, grouped2, by=c("Mean.Glucose.Basal.cat","time"), all.x=T, all.y=T)

p <- ggplot(grouped, aes(x=insulin_mean, y=ffa_mean, group=Mean.Glucose.Basal.cat, color=Mean.Glucose.Basal.cat)) +
      geom_line(alpha=0.8, size=3) +
      scale_color_discrete(breaks=c("70 - <90 mg/dL","90 - <110 mg/dL","90 - <110 mg/dL","110 - <130 mg/dL","130 - <150 mg/dl",">150 mg/dl"))
p
```

# Figure - Predicted values at mean insulin 10 or 16 by overnight glucose and BMI

## Glucose

```{r, include=T, warning=FALSE}
p <- ggplot(data, aes(x=Mean.Glucose.Basal.cat, y=y_hat_Glucose, fill=bmi_twocat)) +
      geom_boxplot() +
      scale_x_discrete(limits = function(x) my_order[my_order %in% x]) + 
      xlab("Overnight glucose") + ylab("Predicted glucose") + theme_classic()
p
```

## Glucose Ra

```{r, include=T, warning=FALSE}
p <- ggplot(data, aes(x=Mean.Glucose.Basal.cat, y=y_hat_GlucoseRa, fill=bmi_twocat)) +
      geom_boxplot() +
      scale_x_discrete(limits = function(x) my_order[my_order %in% x]) + 
      xlab("Overnight glucose") + ylab("Predicted glucose Ra") + theme_classic()
p
```

## Glycerol

```{r, include=T, warning=FALSE}
p <- ggplot(data, aes(x=Mean.Glucose.Basal.cat, y=y_hat_Glycerol, fill=bmi_twocat)) +
      geom_boxplot() +
      scale_x_discrete(limits = function(x) my_order[my_order %in% x]) + 
      xlab("Overnight glucose") + ylab("Predicted glycerol") + theme_classic()
p
```

## Glycerol Ra

```{r, include=T, warning=FALSE}
p <- ggplot(data, aes(x=Mean.Glucose.Basal.cat, y=y_hat_GlycerolRa, fill=bmi_twocat)) +
      geom_boxplot() +
      scale_x_discrete(limits = function(x) my_order[my_order %in% x]) + 
      xlab("Overnight glucose") + ylab("Predicted glycerol Ra") + theme_classic()
p
```

## FFA

```{r, include=T, warning=FALSE}
p <- ggplot(data, aes(x=Mean.Glucose.Basal.cat, y=y_hat_FFA, fill=bmi_twocat)) +
      geom_boxplot() +
      scale_x_discrete(limits = function(x) my_order[my_order %in% x]) + 
      xlab("Overnight glucose") + ylab("Predicted FFA") + theme_classic()
p
```