---
title: "Kidney scRNA for Jenny Kanter"
author: "Hailey Hampson"
date: "2024-10-09"
output: html_document
---
#1. Set Up Libraries & Directories
```{r Set up, echo=F, include = F}
# Function to install and load packages
install_and_load <- function(package_name) {
  if (!requireNamespace(package_name, quietly = TRUE)) {
    install.packages(package_name)
  }
  library(package_name, character.only = TRUE)
}

# Load required packages
library(tidyverse)
install_and_load("tidyr")
install_and_load("BiocManager")
install_and_load("arsenal")
install_and_load("dplyr")
install_and_load("ggplot2")
install_and_load("ggrepel")
install_and_load("Seurat")
install_and_load("future")
install_and_load("colorspace")
install_and_load("patchwork")
install_and_load("ggdendro")
install_and_load("cowplot")
install_and_load("ggpubr")
install_and_load("rstatix")
install_and_load("table1")
install_and_load("Biobase")
install_and_load("ReactomeGSA")
install_and_load("GSEABase")
install_and_load("msigdbr")
install_and_load("kableExtra")
install_and_load("knitr")
if (!requireNamespace('BiocManager', quietly = TRUE))
  install.packages('BiocManager')
library(table1)

# BiocManager::install('EnhancedVolcano')
library(EnhancedVolcano)
install.packages("openxlsx")
library(openxlsx)

#Increase Memory
mem.maxVSize(64000000000)

#Set up directories
dir.dat <- c("/Users/hhampson/Dropbox/Bjornstad data")
# dir.dat <- c("/Desktop/Peds/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/scRNA/data_raw")
dir.meta.dat <- c("/Volumes/Peds Endo/Petter Bjornstad/Data Harmonization/Data Clean")
# dir.dat2 <- c("/Users/hhampson/Dropbox/Bjornstad data/Clinical data")
# "smb://ucdenver.pvt/som/PEDS/RI%20Biostatistics%20Core/Shared/Shared%20Projects/Laura/Peds%20Endo/Petter%20Bjornstad/Data%20Harmonization/"
# smb://ucdenver.pvt/som/PEDS/RI%20Biostatistics%20Core/Shared/Shared%20Projects/Laura/Peds%20Endo/Petter%20Bjornstad/scRNA
dir.home <- c("/Users/hhampson/Documents/CHCO-Code/Petter Bjornstad/Kidney scRNA/Kidney scRNA")

#Load functions
# source(fs::path(dir.home,"Kidney_Functions.R"))
source("Kidney_Functions.R")
```

## Descriptive Statistics
```{r Table 1}
dat <- meta_raw

dat$sex <- as.numeric(as.factor(dat$sex))
dat$sex <- as.factor(dat$sex)
# dat$sex    <- factor(dat$sex, levels=c(0, 1))
label(dat$sex) <- "Sex"
dat$ethnicity   <- factor(dat$ethnicity, levels=c("Hispanic_Or_Latino","NonHispanic"), labels=c("Hispanic or Latino","Non-Hispanic/Non-Latino"))
label(dat$ethnicity) <- "Ethnicity"
dat$race   <- factor(dat$race, levels=c("White","Black or African American"),
                      labels=c("White","Black"))
label(dat$race) <- "Race"
label(dat$age) <- "Age (y)"
label(dat$bmi) <- "BMI (kg/m2)"
label(dat$hba1c) <- "HbA1c (mg/dL)"
label(dat$acr_u) <- "uACR (mg/g)"

dat_T2 <- dat %>% 
  filter(group=="Type 2 Diabetes")
dat_HC <- dat %>% 
  filter(group=="Lean Control")

caption1  <- "Descriptive Statistics by Sex among youth with Type 2 Diabetes"
# footnote <- "ᵃ Also known as Breslow thickness"
caption2  <- "Descriptives by Sex among HC"

#Type 2
table1(~ age + race + ethnicity + bmi + hba1c + acr_u | sex, data=dat_T2,caption = caption1)
#LC
table1(~ age + race + ethnicity + bmi + hba1c + acr_u | sex, data=dat_HC)

pvalue <- function(x, ...) {
    # Construct vectors of data y, and groups (strata) g
    y <- unlist(x)
    g <- factor(rep(1:length(x), times=sapply(x, length)))
    if (is.numeric(y)) {
        # For numeric variables, perform a standard 2-sample t-test
        p <- t.test(y ~ g)$p.value
    } else {
        # For categorical variables, perform a chi-squared test of independence
        p <- chisq.test(table(y, g))$p.value
    }
    # Format the p-value, using an HTML entity for the less-than sign.
    # The initial empty string places the output on the line below the variable label.
    c("", sub("<", "&lt;", format.pval(p, digits=3, eps=0.001)))
}
dat_T2$sex
table1(~ age + race + ethnicity + bmi + hba1c + acr_u | sex, data=dat_T2,caption = caption1)

t.test(age~sex,data=dat_T2)
t.test(bmi~sex,data=dat_T2)
t.test(hba1c~sex,data=dat_T2)
t.test(acr_u~sex,data=dat_T2)

table1(~ age + race + ethnicity + bmi + hba1c + acr_u | sex, data=dat_HC,caption = caption1)

t.test(age~sex,data=dat_T2)
t.test(bmi~sex,data=dat_T2)
t.test(hba1c~sex,data=dat_T2)
t.test(acr_u~sex,data=dat_T2)
chisq.test(table(dat_T2$race, dat_T2$sex))$p.value
chisq.test(table(dat_T2$ethnicity, dat_T2$sex))$p.value

t.test(age~sex,data=dat_HC)
t.test(bmi~sex,data=dat_HC)
t.test(hba1c~sex,data=dat_HC)
t.test(acr_u~sex,data=dat_HC)
chisq.test(table(dat_HC$race, dat_HC$sex))$p.value
chisq.test(table(dat_HC$ethnicity, dat_HC$sex))$p.value
```


#2. Load & Format scRNAseq Data 
```{r echo = F}
# Kidney scRNA data processing
so_kidney_sc <- readRDS(fs::path(dir.dat,"Kidney scRNA","PB_90samples_Harmony_rpca_Fadhl_PhilApproved_091024.RDS"))
# so_kidney_sc <- readRDS(fs::path(dir.dat,"PB_90samples_Harmony_rpca_Fadhl_PhilApproved_091024.RDS"))
# so_kidney_sc <- readRDS("/Desktop/Peds/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/scRNA/data_raw/PB_90samples_Harmony_rpca_Fadhl_PhilApproved_091024.RDS")

# Add missing covariate meta data to Seurat object
meta_raw <- read.csv(fs::path(dir.dat,"Clinical Data","renal_clearance_biopsy.csv"))
# meta_raw <- read.csv(fs::path(dir.dat2,"renal_clearance_biopsy.csv"))
# meta_raw <- read.csv("/Users/hhampson/Dropbox/Bjornstad data/Clinical data/renal_clearance_biopsy.csv")
# harm_data <- read.csv(fs::path(dir.meta.dat,"harmonized_dataset.csv"))
# harm_data2 <- harm_data %>% 
#   filter(group=="Lean Control"|group=="Obese Control"|group=="Type 2 Diabetes" |group=="Type 1 Diabetes")
# sglt2s <- harm_data2 %>% 
#   filter(procedure=="medications") %>% 
#   filter(sglt2i_ever=="Yes" | sglt2i_ever=="No" | sglti_timepoint=="No") %>% 
#   filter(visit=="baseline")
# clinical <- harm_data %>% 
#   filter(procedure=="physical_exam")
# hba1c <- harm_data %>% 
#   filter(procedure=="clamp")
# sex <- harm_data %>% 
#   filter(procedure=="screening")


# meta.vars <- c("procedure",colnames(meta_raw))
# harm_data <- harm_data %>% 
#   dplyr::select(all_of(meta.vars))
# meta_data <- harm_data[which(harm_data$sglt2i_ever!=""),]
# meta_data <- meta_data[which(harm_data$sex=="Male" | harm_data$sex=="Female"),]


#Filter out meta data participants without sc 
record_ids <- unique(so_kidney_sc@meta.data$record_id)
meta_raw <- meta_raw %>%
  filter(record_id %in% record_ids) %>%
  dplyr::select(-record_id)

# #Create study id to filter to in so
# so_kidney_sc@meta.data <- so_kidney_sc@meta.data %>%
#   mutate(study = case_when(grepl("CRC",record_id) ~ "CRC",
#                            grepl("RH",record_id) ~ "RH",
#                            grepl("IT",record_id) ~ "IT"))

#Merge new metadata with 
meta_kidney_sc <-  tidylog::left_join(so_kidney_sc@meta.data, meta_raw,by="cryostor_id")
# meta_kidney_sc <-  so_kidney_sc@meta.data

#Check there are no duplicates
length(which(duplicated(meta_kidney_sc$barcode)))

#Add rownames back to metadata to merge back into saurat object 
rownames(meta_kidney_sc) <- rownames(so_kidney_sc@meta.data)

# Add new meta data to so 
so_kidney_sc <- AddMetaData(so_kidney_sc, meta_kidney_sc)


#Filter to baseline visits only
so_kidney_sc <- subset(so_kidney_sc, visit == "baseline")
# so_kidney_sc <- subset(so_kidney_sc)

# so_kidney_sc <- subset(so_kidney_sc, record_id=="CRC" | grepl("CRC",record_id) | grepl("RH",record_id) | grepl("IT",record_id))
# meta_kidney_sc <-  meta_kidney_sc %>%
#   filter(grepl("CRC",record_id) | grepl("RH",record_id) | grepl("IT",record_id))

so_kidney_sc <- NormalizeData(so_kidney_sc)
so_kidney_sc <- ScaleData(so_kidney_sc)
# # # PCA
# so_kidney_sc <- RunPCA(so_kidney_sc, features = VariableFeatures(object = so_kidney_sc))
# ElbowPlot(so_kidney_sc)
# # Cluster cells
# so_kidney_sc <- FindNeighbors(so_kidney_sc)
# so_kidney_sc <- FindClusters(so_kidney_sc)
# # Perform UMAP and tSNE
# so_kidney_sc <- RunUMAP(so_kidney_sc, dims = 1:15)
# DimPlot(so_kidney_sc, reduction = "umap")
# 
# #UMAP by disease status
# DimPlot(so_kidney_sc, reduction = "umap", group.by = "group") 

```


#3. Examine Descriptive Statistics by Characteristics of Interest (Sex & SGLT2i Status)
## Sex (M/F)
### Descriptive table
```{r echo = F}
# Descriptive stats
kidney_meta_raw_sc <- so_kidney_sc@meta.data %>%
  filter(cryostor_id!="")
form <- paste("sex", paste(colnames(kidney_meta_raw_sc)[c(15,18:27)], collapse = " + "), sep = " ~ ")
summary(arsenal::tableby(formula = as.formula(form), data = kidney_meta_raw_sc, test = F))



#By group
kidney_meta_raw_sc <- so_HC@meta.data %>%
  filter(cryostor_id!="")
form <- paste("sex", paste(colnames(kidney_meta_raw_sc)[c(18:27)], collapse = " + "), sep = " ~ ")
summary(arsenal::tableby(formula = as.formula(form), data = kidney_meta_raw_sc, test = T))


t.test(kidney_meta_raw_sc$age ~ kidney_meta_raw_sc$sex)
t.test(kidney_meta_raw_sc$hba1c ~ kidney_meta_raw_sc$sex)

#cell numbers for the HC and T2 for MON, MAC, and GC-EC
#HC
so_HC_mon<- subset(so_HC, celltype_rpca == "MON")
ncol(so_HC_mon)
so_HC_mac<- subset(so_HC, celltype_rpca == "MAC")
ncol(so_HC_mac)
so_HC_ecgc<- subset(so_HC, celltype_rpca == "EC-GC")
ncol(so_HC_ecgc)
#T2
so_T2_mon<- subset(so_T2, celltype_rpca == "MON")
ncol(so_T2_mon)
so_T2_mon_male<- subset(so_T2_mon, sex == "Male")
ncol(so_T2_mon_male)
so_T2_mon_fem<- subset(so_T2_mon, sex == "Female")
ncol(so_T2_mon_fem)

so_T2_mac<- subset(so_T2, celltype_rpca == "MAC")
ncol(so_T2_mac)
so_T2_mac_male<- subset(so_T2_mac, sex == "Male")
ncol(so_T2_mac_male)
so_T2_mac_fem<- subset(so_T2_mac, sex == "Female")
ncol(so_T2_mac_fem)

so_T2_ecgc<- subset(so_T2, celltype_rpca == "EC-GC")
ncol(so_T2_ecgc)

ncol(so_T2)

```
## Group 
### Descriptive table
```{r echo = F}
# Descriptive stats
kidney_meta_raw_sc <- meta_kidney_sc %>%
  filter(cryostor_id!="")
form <- paste("group", paste(colnames(kidney_meta_raw_sc)[c(15,17:27)], collapse = " + "), sep = " ~ ")
summary(arsenal::tableby(formula = as.formula(form), data = kidney_meta_raw_sc, test = T))

```


## SGLT2i Y/N
### Descriptive table
```{r echo = F}
# Descriptive stats
# kidney_meta_raw_sc <- meta_raw %>%
#   filter(cryostor_id!="")
form <- paste("sglt2i_ever", paste(colnames(kidney_meta_raw_sc)[c(15,18:27)], collapse = " + "), sep = " ~ ")

summary(arsenal::tableby(formula = as.formula(form), data = kidney_meta_raw_sc, test = F))

```

#4. Visualize using UMAPs Stratified by Characteristics of Interest (Sex & SGLT2i Status)
## Stratified UMAP scRNA 
### a. Sex
```{r echo = F}
so_kidney_sc <- RunUMAP(so_kidney_sc, dims = 1:15)
DimPlot(so_kidney_sc, reduction = "umap", group.by = "sex") 

```

### b. SGLT2is
```{r echo = F}
so_kidney_sc <- RunUMAP(so_kidney_sc, dims = 1:15)
DimPlot(so_kidney_sc, reduction = "umap", group.by = "sglt2i_ever") 

```


#5. Differential Expression of Genes by Group for All Disease Types
## Top 2000 genes by Sex
```{r}
#Sex
de.markers(so_kidney_sc, NULL, "sex", id1 = "Female", id2 = "Male", NULL, "_top")
m_top <- m_top %>% head(2000)
```

#Top 2000 genes by SGLT2is
```{r}
#SGLT2is
de.markers(so_kidney_sc, NULL, "sglt2i_ever", id1 = "No", id2 = "Yes", NULL, "_top")
m_top <- m_top %>% head(2000)

```


### By Sex & SGLT2is
```{r}
# so_kidney_sc@meta.data$hepatocyte <- ifelse(grepl("Hepatocyte",meta_liver_sc$celltype),"Hepatocyte","Non-Hepatocyte")
genes <- unique(rownames(so_kidney_sc))
Idents(so_kidney_sc) <- so_kidney_sc$celltype_rpca

#Sex
de.markers(so_kidney_sc, genes, "sex", id2 = "Female", id1 = "Male", "MON", "")
m_top <- m_top %>% head(2000)
de.markers(so_kidney_sc, genes, "sex", id2 = "Female", id1 = "Male", "MAC", "")
m_top <- m_top %>% head(2000)
de.markers(so_kidney_sc, genes, "sex", id2 = "Female", id1 = "Male", "EC-GC", "")
m_top <- m_top %>% head(2000)
de.markers(so_kidney_sc, genes, "sex", id2 = "Female", id1 = "Male", "EC-LYM", "")
m_top <- m_top %>% head(2000)
de.markers(so_kidney_sc, genes, "sex", id2 = "Female", id1 = "Male", "EC-AEA", "")
m_top <- m_top %>% head(2000)
de.markers(so_kidney_sc, genes, "sex", id2 = "Female", id1 = "Male", "EC-PTC", "")
m_top <- m_top %>% head(2000)

#SGLT2is
de.markers(so_kidney_sc, genes, "sglt2i_ever", id2 = "No", id1 = "Yes", "MON", "")
m_top <- m_top %>% head(20)
de.markers(so_kidney_sc, genes, "sglt2i_ever", id2 = "No", id1 = "Yes", "MAC", "")
m_top <- m_top %>% head(20)
de.markers(so_kidney_sc, genes, "sglt2i_ever", id2 = "No", id1 = "Yes", "EC-GC", "")
m_top <- m_top %>% head(20)
de.markers(so_kidney_sc, genes, "sglt2i_ever", id2 = "No", id1 = "Yes", "EC-LYM", "")
m_top <- m_top %>% head(20)
de.markers(so_kidney_sc, genes, "sglt2i_ever", id2 = "No", id1 = "Yes", "EC-AEA", "")
m_top <- m_top %>% head(20)
de.markers(so_kidney_sc, genes, "sglt2i_ever", id2 = "No", id1 = "Yes", "EC-PTC", "")
m_top <- m_top %>% head(20)
```

## Figures
### By Sex & SGLT2is
```{r}
# Idents(so_kidney_sc) <- so_kidney_sc$celltype_rpca
# #Sex
# pt.plot_table <- dp.formatted(seurat_object = so_kidney_sc, genes = genes, group.by = "sex",
#                               celltype = NULL, m = m)
# pt.plot_table <- dp.formatted(seurat_object = so_kidney_sc, genes = genes, group.by = "sex",
#                               celltype = "MON", m = m)
# pt.plot_table <- dp.formatted(seurat_object = so_kidney_sc, genes = genes, group.by = "sex",
#                               celltype = "MAC", m = m)
# pt.plot_table <- dp.formatted(seurat_object = so_kidney_sc, genes = genes, group.by = "sex",
#                               celltype = "EC-GC", m = m)
# pt.plot_table <- dp.formatted(seurat_object = so_kidney_sc, genes = genes, group.by = "sex",
#                               celltype = "EC_LYM", m = m)
# pt.plot_table <- dp.formatted(seurat_object = so_kidney_sc, genes = genes, group.by = "sex",
#                               celltype = "EC-AEA", m = m)
# pt.plot_table <- dp.formatted(seurat_object = so_kidney_sc, genes = genes, group.by = "sex",
#                               celltype = "EC-PTC", m = m)
# 
# #SGLT2is
# pt.plot_table <- dp.formatted(seurat_object = so_kidney_sc, genes = genes, group.by = "sglt2i_ever",
#                               celltype = NULL, m = m)
# pt.plot_table <- dp.formatted(seurat_object = so_kidney_sc, genes = genes, group.by = "sglt2i_ever",
#                               celltype = "MON", m = m)
# pt.plot_table <- dp.formatted(seurat_object = so_kidney_sc, genes = genes, group.by = "sglt2i_ever",
#                               celltype = "MAC", m = m)
# pt.plot_table <- dp.formatted(seurat_object = so_kidney_sc, genes = genes, group.by = "sglt2i_ever",
#                               celltype = "EC-GC", m = m)
# pt.plot_table <- dp.formatted(seurat_object = so_kidney_sc, genes = genes, group.by = "sglt2i_ever",
#                               celltype = "EC_LYM", m = m)
# pt.plot_table <- dp.formatted(seurat_object = so_kidney_sc, genes = genes, group.by = "sglt2i_ever",
#                               celltype = "EC-AEA", m = m)
# pt.plot_table <- dp.formatted(seurat_object = so_kidney_sc, genes = genes, group.by = "sglt2i_ever",
#                               celltype = "EC-PTC", m = m)

```

#6. Differential Expression of Genes by Group Stretified by Disease Status (HC/OB/T1/T2)
##a. Healthy Controls

```{r subset by disease status}
#Filter to HC only
so_HC <- subset(so_kidney_sc, group == "Lean Control")

#Filter to OB only
so_OB <- subset(so_kidney_sc, group == "Obese Control")

#Filter to T1 only
so_T1 <- subset(so_kidney_sc, group == "Type 1 Diabetes")

#Filter to T2 only
so_T2 <- subset(so_kidney_sc, group == "Type 2 Diabetes")

length(which(meta_raw$group=="Lean Control"))
length(which(meta_raw$group=="Obese Control"))
length(which(meta_raw$group=="Type 1 Diabetes"))
length(which(meta_raw$group=="Type 2 Diabetes" & meta_raw$sex=="Female"))
length(which(meta_raw$group=="Type 2 Diabetes" & meta_raw$sex=="Male"))
```

### Sex
```{r echo = F}
#HC
so_HC <- RunUMAP(so_HC, dims = 1:15)
DimPlot(so_HC, reduction = "umap", group.by = "sex") 

#OB
so_OB <- RunUMAP(so_OB, dims = 1:15)
DimPlot(so_OB, reduction = "umap", group.by = "sex") 

#T1
so_T1 <- RunUMAP(so_T1, dims = 1:15)
DimPlot(so_T1, reduction = "umap", group.by = "sex") 

#T2
so_T2 <- RunUMAP(so_T2, dims = 1:15)
DimPlot(so_T2, reduction = "umap", group.by = "sex") 

```

### SGLT2is
```{r echo = F}
#HC
so_HC <- RunUMAP(so_HC, dims = 1:15)
DimPlot(so_HC, reduction = "umap", group.by = "sglt2i_ever") 

#OB
so_OB <- RunUMAP(so_OB, dims = 1:15)
DimPlot(so_OB, reduction = "umap", group.by = "sglt2i_ever") 

#T1
so_T1 <- RunUMAP(so_T1, dims = 1:15)
DimPlot(so_T1, reduction = "umap", group.by = "sglt2i_ever") 

#T2
so_T2 <- RunUMAP(so_T2, dims = 1:15)
DimPlot(so_T2, reduction = "umap", group.by = "sglt2i_ever") 

```


## Differential Expression by Group

## Top 2000 genes
```{r}
#Sex
#HC
de.markers(so_HC, NULL, "sex", id1 = "Female", id2 = "Male", NULL, "_top")
m_top <- m_top %>% head(20)
#OB
de.markers(so_OB, NULL, "sex", id1 = "Female", id2 = "Male", NULL, "_top")
m_top <- m_top %>% head(20)
#T1
de.markers(so_T1, NULL, "sex", id1 = "Female", id2 = "Male", NULL, "_top")
m_top <- m_top %>% head(20)
#T2
de.markers(so_T2, NULL, "sex", id1 = "Female", id2 = "Male", NULL, "_top")
m_top <- m_top %>% head(20)

#SGLT2is
#T2
de.markers(so_T2, NULL, "sglt2i_ever", id1 = "No", id2 = "Yes", NULL, "_top")
m_top <- m_top %>% head(20)

```


### By Sex & SGLT2is
###HC
```{r}
genes <- unique(rownames(so_HC))
Idents(so_HC) <- so_HC$celltype_rpca

#Sex
de.markers(so_HC, genes, "sex", id2 = "Female", id1 = "Male", "MON", "")
m_top <- m_top %>% head(20)
de.markers(so_HC, genes, "sex", id2 = "Female", id1 = "Male", "MAC", "")
m_top <- m_top %>% head(20)
de.markers(so_HC, genes, "sex", id2 = "Female", id1 = "Male", "EC-GC", "")
m_top <- m_top %>% head(20)
# de.markers(so_HC, genes, "sex", id2 = "Female", id1 = "Male", "EC-LYM", "")
de.markers(so_HC, genes, "sex", id2 = "Female", id1 = "Male", "EC-AEA", "")
m_top <- m_top %>% head(20)
de.markers(so_HC, genes, "sex", id2 = "Female", id1 = "Male", "EC-PTC", "")
m_top <- m_top %>% head(20)
```

###OB
```{r}
genes <- unique(rownames(so_OB))
Idents(so_OB) <- so_OB$celltype_rpca

#Sex
de.markers(so_OB, genes, "sex", id2 = "Female", id1 = "Male", "MON", "")
m_top <- m_top %>% head(20)
de.markers(so_OB, genes, "sex", id2 = "Female", id1 = "Male", "MAC", "")
m_top <- m_top %>% head(20)
de.markers(so_OB, genes, "sex", id2 = "Female", id1 = "Male", "EC-GC", "")
m_top <- m_top %>% head(20)
de.markers(so_OB, genes, "sex", id2 = "Female", id1 = "Male", "EC-LYM", "")
m_top <- m_top %>% head(20)
de.markers(so_OB, genes, "sex", id2 = "Female", id1 = "Male", "EC-AEA", "")
m_top <- m_top %>% head(20)
de.markers(so_OB, genes, "sex", id2 = "Female", id1 = "Male", "EC-PTC", "")
m_top <- m_top %>% head(20)
```

###T1
```{r}
genes <- unique(rownames(so_T1))
Idents(so_T1) <- so_T1$celltype_rpca

#Sex
de.markers(so_T1, genes, "sex", id2 = "Female", id1 = "Male", "MON", "")
m_top <- m_top %>% head(20)
de.markers(so_T1, genes, "sex", id2 = "Female", id1 = "Male", "MAC", "")
m_top <- m_top %>% head(20)
de.markers(so_T1, genes, "sex", id2 = "Female", id1 = "Male", "EC-GC", "")
m_top <- m_top %>% head(20)
de.markers(so_T1, genes, "sex", id2 = "Female", id1 = "Male", "EC-LYM", "")
m_top <- m_top %>% head(20)
de.markers(so_T1, genes, "sex", id2 = "Female", id1 = "Male", "EC-AEA", "")
m_top <- m_top %>% head(20)
de.markers(so_T1, genes, "sex", id2 = "Female", id1 = "Male", "EC-PTC", "")
m_top <- m_top %>% head(20)
```

###T2
```{r}
genes <- unique(rownames(so_T2))
Idents(so_T2) <- so_T2$celltype_rpca

#Sex
de.markers(so_T2, genes, "sex", id2 = "Female", id1 = "Male", "MON", "")
m_top <- m_top %>% head(20)
de.markers(so_T2, genes, "sex", id2 = "Female", id1 = "Male", "MAC", "")
m_top <- m_top %>% head(20)
de.markers(so_T2, genes, "sex", id2 = "Female", id1 = "Male", "EC-GC", "")
m_top <- m_top %>% head(20)
de.markers(so_T2, genes, "sex", id2 = "Female", id1 = "Male", "EC-LYM", "")
m_top <- m_top %>% head(20)
de.markers(so_T2, genes, "sex", id2 = "Female", id1 = "Male", "EC-AEA", "")
m_top <- m_top %>% head(20)
de.markers(so_T2, genes, "sex", id2 = "Female", id1 = "Male", "EC-PTC", "")
m_top <- m_top %>% head(20)

#SGLT2is
de.markers(so_T2, genes, "sglt2i_ever", id2 = "No", id1 = "Yes", "MON", "")
m_top <- m_top %>% head(20)
de.markers(so_T2, genes, "sglt2i_ever", id2 = "No", id1 = "Yes", "MAC", "")
m_top <- m_top %>% head(20)
de.markers(so_T2, genes, "sglt2i_ever", id2 = "No", id1 = "Yes", "EC-GC", "")
m_top <- m_top %>% head(20)
de.markers(so_T2, genes, "sglt2i_ever", id2 = "No", id1 = "Yes", "EC-LYM", "")
m_top <- m_top %>% head(20)
de.markers(so_T2, genes, "sglt2i_ever", id2 = "No", id1 = "Yes", "EC-AEA", "")
m_top <- m_top %>% head(20)
de.markers(so_T2, genes, "sglt2i_ever", id2 = "No", id1 = "Yes", "EC-PTC", "")
m_top <- m_top %>% head(20)
```
