---
title: "Kidney_Project"
author: "Hailey Hampson"
date: "2025-02-25"
output:
  word_document: default
  pdf_document: default
  html_document: default
  output_dir: "/Users/hhampson/Library/CloudStorage/OneDrive-UW/Biostatistics Core Shared Drive/Kidney scRNAseq Project/Results"
---

# 1. Set up 
## a. Libraries & Directores
```{r, include=F}
library(reprex)
# library(tidyverse)
library(BiocManager)        
library(arsenal)
library(dplyr)
library(ggplot2)
library(ggrepel)
library(Seurat)
library(future)
library(colorspace)
library(patchwork)
library(ggdendro)
library(cowplot)
library(ggpubr)
library(venn)
library(rstatix)
library(table1)
# BiocManager::install("Biobase")
library(Biobase)
# BiocManager::install("ReactomeGSA")
library(ReactomeGSA)
# BiocManager::install("GSEABase")
library(GSEABase)
# install.packages('msigdbdf', repos = 'https://igordot.r-universe.dev')
library(msigdbr)
# library(kableExtra)
library(knitr)
# BiocManager::install("SingleCellExperiment")
library(SingleCellExperiment)
# BiocManager::install("fgsea")
library(fgsea)
# BiocManager::install("EnhancedVolcano")
library(EnhancedVolcano)
library(openxlsx)
library(BiocManager)
# library(MAST)
library(ggrepel)
# library(qpcR)
library(ggpubr)
library(openxlsx)
library(ggplot2)
library(GGally)
library(GSEABase)
# library(limma)
library(reshape2)
library(data.table)
library(knitr)
# BiocManager::install("TxDb.Hsapiens.UCSC.hg19.knownGene")
library(TxDb.Hsapiens.UCSC.hg19.knownGene)
library(stringr)
# library(NMF)
library(rsvd)
library(RColorBrewer)
# BiocManager::install("devtools")
# library(devtools)
# install_github("Sun-lab/ideas",force=T)
# library(ideas)
library(foreach)
library(doRNG)
library(doParallel)
library(fs)
library(future)
# registerDoParallel(cores = 6)
library(VennDiagram)
#install.packages("survival")
library(survival)
#install.packages("lme4")  # If not already installed
library(lme4)
#install.packages("lmerTest")
library(lmerTest)
# install.packages("emmeans")
library(emmeans)
# install.packages("glmmTMB")
library(glmmTMB)
library(ggrepel)
# library(qpcR)
library(ggpubr)
#install.packages("glmmTMB")
# Reinstall glmmTMB from source
#install.packages("glmmTMB", type = "source")
# library(glmmTMB)
# Install DoubletFinder (if not already installed)
# devtools::install_github("chris-mcginnis-ucsf/DoubletFinder",force=T)
# Load the package
# Install DoubletFinder from GitHub (use devtools to install)
# if (!requireNamespace("devtools", quietly = TRUE)) install.packages("devtools")
# devtools::install_github("chris-mcginnis-ucsf/DoubletFinder",force=T)
# library(DoubletFinder)
# install.packages("emmeans")
library(emmeans)
library(ggrepel)
# library(qpcR)
# BiocManager::install("enrichplot")
# library(enrichplot)
# library(enrichR)
# dbs <- c("GO_Biological_Process_2023", 
#          "KEGG_2021_Human",
#          "Reactome_2022", 
#          "Reactome_Pathways_2024")


#Set number of cores for parallellization
#maxCores <- detectCores()
#numCores <- maxCores-1
#cl <- makeCluster(numCores)  # Create a cluster with the desired number of cores
#registerDoParallel(cl) 

#Local file path
#dir.dat <- c("/Volumes/Peds Endo/Petter Bjornstad")
#dir.dat2 <- c("/Volumes/Peds Endo/Petter Bjornstad/scRNA/data_clean")
#dir.code <- c("/Users/hhampson/Documents/CHCO-Code/Petter Bjornstad/Liver analysis/Liver scRNAseq")
#dir.results <- c("/Volumes/Peds Endo/Petter Bjornstad/Kidney Project/Results")

# #Lambda file path
# dir.dat <- c("/run/user/1026/gvfs/smb-share:server=ucdenver.pvt,share=som/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad")
# dir.code <- c("/home/Github_Repo/CHCO-Code/Petter Bjornstad/Kidney scRNA/Kidney scRNA")
# dir.results <- c(fs::path(dir.dat,"Kidney Project/Results"))

# #Mac Studio File Path
# dir.dat <- c("/Users/hhampson/Library/CloudStorage/OneDrive-UW/Biostatistics Core Shared Drive")
# dir.results <- c("/Users/hhampson/Library/CloudStorage/OneDrive-UW/Biostatistics Core Shared Drive/Kidney scRNAseq Project/Results")
# dir.ipa <- c("/Users/hhampson/Documents/IPA/Results")

#Load functions
source("Kidney_functions_sc.R")
# source("/Users/hhampson/CHCO-Code/Petter Bjornstad/Data Processing and Analysis/Standard_Functions.R")
dir.results <- c("/home/hailey/Documents/Results")

```

## b. Cyberduck 
```{r, Cyberduck setup}
# knitr::opts_knit$set(root.dir = dir.results)
# install.packages("reticulate")
# Sys.setenv(RETICULATE_PYTHON = "/home/hailey/miniconda3/envs/py310_env/bin/python")
library(reticulate)
reticulate::py_config()
# library(reticulate)
# reticulate::use_condaenv("py310_env", required = TRUE)
# reticulate::py_config()
# Sys.setenv(LD_LIBRARY_PATH = "/usr/lib/x86_64-linux-gnu")
# reticulate::use_python("/home/hailey/miniconda3/bin/python")
# reticulate::use_condaenv("base", required = TRUE)

## Load boto3 and pandas
boto3 <- reticulate::import("boto3")
pd <- reticulate::import("pandas")

## Create an S3 client
# install.packages("jsonlite")  # Install if not already installed
library(jsonlite)  # Load the package

keys <- fromJSON("/home/hailey/keys.json") # replace with your Lambda username
session <- boto3$session$Session(
  aws_access_key_id = keys$MY_ACCESS_KEY,
  aws_secret_access_key = keys$MY_SECRET_KEY
)

## Create an S3 client with the session
s3 <- session$client("s3", endpoint_url = "https://s3.kopah.uw.edu")
```

# 2. Quality Control & Preprocessing
## a. Load Kidney scRNA seq Data
```{r, include=F}
#Load PB90 with new KPMP Cell Types
#Mac Studio pathway
# so_kpmp_sc <- readRDS("/Users/hhampson/Library/CloudStorage/OneDrive-UW/Biostatistics Core Shared Drive/scRNA/data_raw/PB_90_RPCAFix_Metadata_LR_RM_kpmpV1labelled.rds")

#Lambda Pathway
invisible(gc())
bucket <- "scrna" # bucket name in Kopah
temp_file <- tempfile(fileext = ".rds") # need to create a temporary file
s3$download_file(bucket, "Kidney transcriptomics/Single cell RNA seq/PB_90_RPCAFix_Metadata_LR_RM_kpmpV1labelled.rds", temp_file)
so_kpmp_sc <- readRDS(temp_file)
invisible(gc())
```

## b. Data Cleaning, QC & Preprocessing
```{r, echo=F, warning=F, fig.width=15, fig.height=15}
#Fix Typos in kit ids in PB90
so_kpmp_sc$kit_id[which(so_kpmp_sc$kit_id=="KI-0014643")] <- "KL-0014643"
so_kpmp_sc$kit_id[which(so_kpmp_sc$kit_id=="kl-0023998")] <- "KL-0023998"

#Load harmonized data that has been filtered from 90 to the 83 participants that have baseline single cell data
# harm_meta_data <- read.csv(fs::path(dir.dat,"Kidney scRNAseq Project","Data","harmonized_data_kidney_sc_all_metadata2.csv"))
invisible(gc())
bucket <- "scrna" # bucket name in Kopah
temp_file <- tempfile(fileext = ".csv") # need to create a temporary file
s3$download_file(bucket, "data_clean/harmonized_data_kidney_sc_all_metadata2.csv", temp_file)
harm_meta_data <- read.csv(temp_file)
invisible(gc())

harm_meta_data <- harm_meta_data %>%
  dplyr::select(-X)

#Select metadata from seurat object to facilitate merge of new metadata into seurat object
meta_kidney_sc <-  so_kpmp_sc@meta.data
rownames(meta_kidney_sc) <- rownames(so_kpmp_sc@meta.data)

#Merge metadata from 83 participants at baseline into seurat object metadata
meta_kidney_sc <- meta_kidney_sc %>%
  left_join(harm_meta_data,by="kit_id")
rownames(meta_kidney_sc) <- rownames(so_kpmp_sc@meta.data)

#Pull ids from 83 participants at baseline to filter seurat object to these participants only 
ids <- harm_meta_data$kit_id

#Merge metadata back into seurat object
so_kpmp_sc <- AddMetaData(so_kpmp_sc, meta_kidney_sc)

#Check number of unique ids
length(unique(so_kpmp_sc$kit_id)) #should be 90

#Filter seurat object to only IDs that have the metadata & sc data (83 individuals at baseline)
so_kpmp_sc <- subset(so_kpmp_sc, kit_id %in% ids)

#Check number of unique ids after filtering
length(unique(so_kpmp_sc$kit_id)) #should be 83

#Remove metadatasets
rm(meta_kidney_sc,harm_meta_data)

#Remove two coenrolled IDs
so_kpmp_sc <- subset(so_kpmp_sc,kit_id!="KL-0030621")
so_kpmp_sc <- subset(so_kpmp_sc,kit_id!="KL-0029535")
#Check unique number of kit ids now (should be 81)
length(unique(so_kpmp_sc$kit_id)) #Should be 81

#Load in most up to date medication data to update medication information
# med <- read.xlsx(fs::path(dir.dat,"Kidney scRNAseq Project/Data/Biopsies_w_mrn_Oct3.xlsx"))
invisible(gc())
bucket <- "scrna" # bucket name in Kopah
temp_file <- tempfile(fileext = ".xlsx") # need to create a temporary file
s3$download_file(bucket, "data_clean/Biopsies_w_mrn_Oct3.xlsx", temp_file)
med <- read.xlsx(temp_file)
invisible(gc())
rm(s3,session,keys)

#Select Metformin, RASSI, Insulin data
med <- med %>%
  dplyr::select(all_of(c("record_id","mrn","raasi_1","insulin_1","mfm_1")))
#Pull seurat object metadata to help harmoinize in new metadata
meta_kidney_sc <-  so_kpmp_sc@meta.data
#Filter to only those with a unique identifier id in the seurat object metadata
med <- med %>%
  filter(mrn %in% as.character(meta_kidney_sc$mrn)) 
length(unique(med$mrn)) #95 total rows remain, 81 unique individual ids
#Filter to only those that have a unique record id in the seurat object
med <- med %>%
  filter(record_id %in% meta_kidney_sc$record_id) 
length(unique(med$mrn)) #81 remain
length(unique(med$record_id)) #81
rownames(meta_kidney_sc) <- rownames(so_kpmp_sc@meta.data)
med$mrn <- as.numeric(med$mrn) #Make numeric to merge
#Merge med data with seurat metadata
meta_kidney_sc <- meta_kidney_sc %>%
  left_join(med,by=c("mrn","record_id"))
rownames(meta_kidney_sc) <- rownames(so_kpmp_sc@meta.data)
length(unique(meta_kidney_sc$mrn)) #81 remain
length(unique(meta_kidney_sc$record_id)) #81

#Add Med Meta Data to Seurat object
so_kpmp_sc <- AddMetaData(so_kpmp_sc, meta_kidney_sc)
#Remove med metadatset
rm(med,meta_kidney_sc)

#Create medication & disease status groups of interest
so_kpmp_sc@meta.data <- so_kpmp_sc@meta.data %>%
  mutate(glp1_sglt2=ifelse(epic_glp1ra_1=="Yes" & epic_sglti2_1=="Yes","Yes","No")) %>%
  mutate(sglt2=ifelse(epic_sglti2_1=="Yes" & epic_glp1ra_1=="No","Yes","No")) %>%
  mutate(glp1=ifelse(epic_sglti2_1=="No" & epic_glp1ra_1=="Yes","Yes","No")) %>%
  mutate(no_med=ifelse(epic_sglti2_1=="No" & epic_glp1ra_1=="No","Yes","No"))

#Define 4 exposure groups:
#SGLT2i(+)/GLP1RA(+), SGLT2i(+)/GLP1RA(-), SGLT2i(-)/GLPRA(+), SGLT2i(-)/GLP1RA(-)
invisible(gc())
so_kpmp_sc@meta.data <- so_kpmp_sc@meta.data %>%
  mutate(medication = case_when(glp1_sglt2 == "Yes" ~ "glp1_sglt2",
                                sglt2 == "Yes" ~ "sglt2",
                                glp1 == "Yes" ~ "glp1",
                                no_med == "Yes" ~ "no_med"))
so_kpmp_sc@meta.data$medication <- factor(so_kpmp_sc@meta.data$medication, levels = c("no_med", "sglt2", "glp1","glp1_sglt2"))

#Ensure default assay in seurat object to RNA
DefaultAssay(so_kpmp_sc) <- "RNA"
invisible(gc())

#Before filtering
ncol(so_kpmp_sc)
#186125 cells
nrow(so_kpmp_sc)
#31332 genes

#YE JI's filtering code for percent expression 
#Filter out rare genes expressed in less than "gene_pct" of cells
# expr_matrix <- as.matrix(GetAssayData(so_kpmp_sc, layer = "counts"))
expr_matrix <- as.matrix(GetAssayData(so_kpmp_sc, assay = "RNA", layer = "counts"))
# expr_matrix <- so_kpmp_sc@assays$RNA@counts
# Calculate the proportion of cells expressing each gene
num_cells_per_gene <- rowSums(expr_matrix > 0)  # Count nonzero cells per gene
total_cells <- ncol(expr_matrix)  # Total number of cells
gene_proportion <- num_cells_per_gene / total_cells  # Fraction of cells expressing each gene
remove(expr_matrix)
# Keep genes expressed in at least "gene_pct" of cells
genes_to_keep <- names(gene_proportion[gene_proportion >= 0.01])
so_kpmp_sc <- subset(so_kpmp_sc, features = genes_to_keep)
#After filtering
ncol(so_kpmp_sc) #186125 nuclei
nrow(so_kpmp_sc) # 14627 genes

# Step 2: Remove mitochondrial genes (those starting with "MT")
mito_genes <- grep("^MT-", rownames(so_kpmp_sc), value = TRUE)
#keep_ids <- unique(rownames(so_kpmp_sc)[which(!rownames(so_kpmp_sc) %in% mito_genes)])
# so_kpmp_sc <- subset(so_kpmp_sc, features = setdiff(rownames(so_kpmp_sc), mito_genes))
# so_kpmp_sc <- subset(so_kpmp_sc,kit_id!="KL-0029535")
#so_kpmp_sc$Gene <- rownames(so_kpmp_sc)
#so_kpmp_sc <- subset(so_kpmp_sc, Gene %in% keep_ids)
so_kpmp_sc <- subset(so_kpmp_sc, features = setdiff(rownames(so_kpmp_sc@assays$RNA@layers$counts), mito_genes))
#so_kpmp_sc <- subset(so_kpmp_sc, features = setdiff(rownames(so_kpmp_sc@assays$RNA@counts), mito_genes))
# so_kpmp_sc <- subset(so_kpmp_sc, !rownames(so_kpmp_sc) %in% mito_genes)
# grep("^MT-", rownames(so_kpmp_sc@assays$RNA@counts), value = TRUE)
# dim(so_kpmp_sc@assays$RNA@counts) #9276 186125
# dim(so_kpmp_sc@assays$RNA@data) #9276 186125
# dim(so_kpmp_sc@assays$RNA)#9276 186125
sum(grepl("^MT-", rownames(so_kpmp_sc@assays$RNA@layers$counts))) #0
ncol(so_kpmp_sc) #186125 cells
nrow(so_kpmp_sc) #14627 genes

#Renormalize & Scale after filtering
so_kpmp_sc <- NormalizeData(so_kpmp_sc)
so_kpmp_sc <- ScaleData(so_kpmp_sc, features = VariableFeatures(so_kpmp_sc))

#Create general hepatocyte cell type variable
#Create PT and TAL pseudobulk cell type variable
so_kpmp_sc$celltype1 <- case_when(grepl("PT-",so_kpmp_sc$celltype_rpca)~"PT",
                                  grepl("TAL-",so_kpmp_sc$celltype_rpca)~"TAL",
                                  grepl("EC-",so_kpmp_sc$celltype_rpca)~"EC",
                                  grepl("POD",so_kpmp_sc$celltype_rpca)~"POD",
                                  grepl("MAC",so_kpmp_sc$celltype_rpca)~"MAC",
                                  grepl("MON",so_kpmp_sc$celltype_rpca)~"MON",
                                  grepl("PC-",so_kpmp_sc$celltype_rpca)~"PC",
                                  grepl("FIB",so_kpmp_sc$celltype_rpca)~"FIB_MC_VSMC",
                                  grepl("DTL",so_kpmp_sc$celltype_rpca)~"DTL",
                                  so_kpmp_sc$celltype_rpca=="DCT"~"DCT",
                                  so_kpmp_sc$celltype_rpca=="ATL"~"ATL",
                                  so_kpmp_sc$celltype_rpca=="B"~"B",
                                  so_kpmp_sc$celltype_rpca=="T"~"T")
so_kpmp_sc$celltype1 <- as.character(so_kpmp_sc$celltype1)

so_kpmp_sc$KPMP_celltype2 <- as.character(so_kpmp_sc$KPMP_celltype)
so_kpmp_sc$celltype2 <- ifelse(so_kpmp_sc$KPMP_celltype=="aPT" | 
                                 so_kpmp_sc$KPMP_celltype=="PT-S1/S2" | 
                                 so_kpmp_sc$KPMP_celltype == "PT-S3","PT",
                               ifelse(grepl("TAL",so_kpmp_sc$KPMP_celltype),"TAL",
                                      ifelse(grepl("EC-",so_kpmp_sc$KPMP_celltype),"EC",so_kpmp_sc$KPMP_celltype2)))
# so_kpmp_sc$celltype2 <- factor(so_kpmp_sc$celltype2)
```

## c. Visualizations 
```{r}
# PCA
so_kpmp_sc <- FindVariableFeatures(object = so_kpmp_sc)
so_kpmp_sc <- RunPCA(so_kpmp_sc, features = VariableFeatures(object = so_kpmp_sc),assay="RNA")
ElbowPlot(so_kpmp_sc)

# # Find neighbors and clusters (again using integrated data)
so_kpmp_sc <- FindNeighbors(so_kpmp_sc, assay = "RNA", dims = 1:20)
so_kpmp_sc <- FindClusters(so_kpmp_sc, resolution = 0.5)

# Perform UMAP and tSNE
# so_liver_sc <- RunUMAP(so_liver_sc, dims = 1:15)
so_kpmp_sc@reductions
DimPlot(so_kpmp_sc, reduction = "umap.harmony", raster = F)
DimPlot(so_kpmp_sc, reduction = "umap.harmony", group.by = "KPMP_celltype", raster = FALSE)
DimPlot(so_kpmp_sc, reduction = "umap", group.by = "celltype",raster = FALSE, label = TRUE, label.size = 4)+
  ggtitle("UMAP Plot with Cell Type Labels")

so_kpmp_sc$group <- factor(so_kpmp_sc$group)
DimPlot(so_kpmp_sc, reduction = "umap",group.by = "steatosis_cat",label=F,raster=F) +
  ggtitle(paste0("UMAP by Steatosis Category"))

so_kpmp_sc$fibrosis_cat <- factor(so_kpmp_sc$fibrosis_cat)
DimPlot(so_kpmp_sc, reduction = "umap",group.by = "fibrosis_cat",label=F,raster=F) +
  ggtitle(paste0("UMAP by Fibrosis Category"))


#Barcharts of proportions
# By PT subtypes
cellcount <- so_kpmp_sc@meta.data %>% 
  filter(celltype2 == "PT")

label(cellcount$group) <- "Disease Status"
# label(cellcount$steatosis_cat) <- "Steatosis Category"
# cellcount$group <- ifelse(cellcount$group=="0+1", "Low Steatosis (0+1)","High Steatosis (2+3)")
cellcount$group <- factor(cellcount$group,levels=c("Lean_Control","Obese_Control","Type_1_Diabetes","Type_2_Diabetes"),labels=c("Lean Controls","Obese Controls","Type 1 Diabetes","Type 2 Diabetes"))
cellcount$KPMP_celltype2 <- factor(cellcount$KPMP_celltype2,levels=c("PT-S1/S2","PT-S3","aPT"),labels=c("PT-S1/S2","PT-S3","aPT"))
cellcount <- cellcount %>% 
  filter(group!="Obese Controls")

prop_plot1 <- ggplot(data=cellcount,aes(group, fill = KPMP_celltype)) + 
  geom_bar(stat = "count", position = "fill") +
  theme_classic() +
  labs(x = NULL,
       y = "Proportion of Cells",
       fill = "Cell type") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  theme(plot.title = element_text(hjust=0.5, face="bold"),
        text = element_text(size = 20)) +
  ggtitle("Disease Status") +
  scale_fill_manual(values = c("#264653", "#2a9d8f", "#e9c46a", "#f4a261", "#e76f51","darkred"))

prop_plot1 

#Barcharts of proportions
# By PT subtypes
cellcount <- so_kpmp_sc@meta.data %>% 
  filter(celltype2 == "TAL")

label(cellcount$group) <- "Disease Status"
# label(cellcount$steatosis_cat) <- "Steatosis Category"
# cellcount$group <- ifelse(cellcount$group=="0+1", "Low Steatosis (0+1)","High Steatosis (2+3)")
cellcount$group <- factor(cellcount$group,levels=c("Lean_Control","Obese_Control","Type_1_Diabetes","Type_2_Diabetes"),labels=c("Lean Controls","Obese Controls","Type 1 Diabetes","Type 2 Diabetes"))
# cellcount$KPMP_celltype2 <- factor(cellcount$KPMP_celltype2,levels=c("PT-S1/S2","PT-S3","aPT"),labels=c("PT-S1/S2","PT-S3","aPT"))
cellcount <- cellcount %>%
  filter(group!="Obese Controls")

prop_plot1 <- ggplot(data=cellcount,aes(group, fill = KPMP_celltype)) + 
  geom_bar(stat = "count", position = "fill") +
  theme_classic() +
  labs(x = NULL,
       y = "Proportion of Cells",
       fill = "Cell type") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  theme(plot.title = element_text(hjust=0.5, face="bold"),
        text = element_text(size = 20)) +
  ggtitle("Disease Status") +
  scale_fill_manual(values = c("#264653", "#2a9d8f", "#e9c46a", "#f4a261", "#e76f51","darkred"))

prop_plot1 

#Barcharts of proportions
# By PT subtypes
cellcount <- so_kpmp_sc@meta.data %>% 
  filter(grepl("EC-",KPMP_celltype) | grepl("EC/VSMC",KPMP_celltype))

label(cellcount$group) <- "Disease Status"
# label(cellcount$steatosis_cat) <- "Steatosis Category"
# cellcount$group <- ifelse(cellcount$group=="0+1", "Low Steatosis (0+1)","High Steatosis (2+3)")
cellcount$group <- factor(cellcount$group,levels=c("Lean_Control","Obese_Control","Type_1_Diabetes","Type_2_Diabetes"),labels=c("Lean Controls","Obese Controls","Type 1 Diabetes","Type 2 Diabetes"))
# cellcount$KPMP_celltype2 <- factor(cellcount$KPMP_celltype2,levels=c("PT-S1/S2","PT-S3","aPT"),labels=c("PT-S1/S2","PT-S3","aPT"))
cellcount <- cellcount %>%
  filter(group!="Obese Controls")

prop_plot1 <- ggplot(data=cellcount,aes(group, fill = KPMP_celltype)) + 
  geom_bar(stat = "count", position = "fill") +
  theme_classic() +
  labs(x = NULL,
       y = "Proportion of Cells",
       fill = "Cell type") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  theme(plot.title = element_text(hjust=0.5, face="bold"),
        text = element_text(size = 20)) +
  ggtitle("Disease Status") +
  scale_fill_manual(values = c("#264653", "#2a9d8f", "#e9c46a", "#f4a261", "#e76f51","darkred"))

prop_plot1

```

# 3. Comparitive Data Analysis
## A. Type 2 Diabetes (No Med) vs. Healthy Controls
### a. PT Cells
#### i. Pseudobulk PT cells
```{r}
#PT
so_celltype <- subset(so_kpmp_sc,celltype1=="PT")
ncol(so_celltype) #37948 cells
nrow(so_celltype) #14627 genes
# Extract the gene expression data for all genes
gene_expression <- as.data.frame(GetAssayData(so_celltype, layer = "data"))

#Assign gene names as rownames, cell names as colnames
# rownames(gene_expression) <- rownames(so_celltype) #Gene Names
# colnames(gene_expression) <- colnames(so_celltype) #Cell Names
#Transpose gene expression dataset to merge with metadata
gene_expression <- t(gene_expression)
gene_expression <- data.frame(gene_expression) #Make a dataframe again after transposing
#Set gene list
gene_list_total <- colnames(gene_expression)
gene_expression$cellname <- rownames(gene_expression) 
rownames(gene_expression) <- NULL

# Extract the metadata
metadata <- so_celltype@meta.data
# metadata <- metadata %>%
#   mutate(across(everything(),~ifelse(.==".",NA,.)))
metadata$cellname <- rownames(metadata)
rownames(metadata) <- NULL

# Combine the gene expression data and metadata
data <- tidylog::full_join(metadata,gene_expression,by="cellname")
rm(metadata,gene_expression)

#Compare Health Controls to Type 2 Diabetes on no meds
data_subset <- data %>%
  filter(group=="Type_2_Diabetes" | group=="Lean_Control") %>%
  filter(medication=="no_med")
rm(data)
data_subset$group <- factor(data_subset$group)
data_subset$group <- relevel(data_subset$group,ref="Lean_Control")
data_subset <- data_subset %>% 
  dplyr::select(c("kit_id","group","age","sex","bmi",all_of(gene_list_total)))

#Try as batch loop
# Set batch size
batch_size <- 2000
total_cores <- 50

# Deduplicate gene list to avoid double-processing
gene_list_total <- unique(gene_list_total)

# Split the gene list into batches
gene_batches <- split(gene_list_total, ceiling(seq_along(gene_list_total) / batch_size))

# Define the function that processes a single gene
process_gene <- function(gene) {
  if (sum(data_subset[[gene]]) > 0) {
    m1 <- as.formula(paste0(gene, " ~ group + age + sex + bmi + (1 | kit_id)"))
    
    model1 <- tryCatch({
      glmmTMB(m1, data = data_subset, family = gaussian)
    }, error = function(e) {
      return(NULL)
    })
    
    if (!is.null(model1)) {
      Beta <- summary(model1)$coef$cond[2,1]
      PValue <- summary(model1)$coef$cond[2,4]
    } else {
      Beta <- NA
      PValue <- NA
    }
  } else {
    Beta <- NA
    PValue <- NA
  }
  
  return(data.frame(Gene = gene, Beta = Beta, PValue = PValue))
}

# Set number of cores
total_cores <- parallel::detectCores() - 1

# Initialize and register cluster
cl <- makeCluster(total_cores)
registerDoParallel(cl)

# Run analysis per batch
all_results <- lapply(seq_along(gene_batches), function(batch_idx) {
  batch_genes <- gene_batches[[batch_idx]]
  
  batch_results <- foreach(
    gene = batch_genes,
    .combine = rbind,
    .packages = c("glmmTMB", "lme4"),
    .export = c("process_gene", "data_subset")
  ) %dopar% {
    process_gene(gene)
  }
  
  cat("Processed batch", batch_idx, "with", length(batch_genes), "genes\n")
  return(batch_results)
})

# Stop cluster
stopCluster(cl)

# Combine all batch results into a single data frame
final_results <- do.call(rbind, all_results)

# #Make volcano plot of all gene results for group
full_results <- final_results %>%
  mutate(fdr=p.adjust(PValue,method="fdr"))  
# mutate(fdr3=p.adjust(PValue3,method="fdr"))
full_results$PValue10 <- -log10(pmax(full_results$PValue, 1e-10))  # Avoid log(0)
# full_results$PValue10_3 <- -log10(pmax(full_results$PValue3, 1e-10))  # Avoid log(0)
Nonconvergence_Rate <- paste0(round(((length(which(is.na(full_results$PValue)))+length(which(full_results$PValue=="NaN")))/length(full_results$Gene))*100,0),"%")
# view(full_results)

write.csv(full_results,fs::path(dir.results,"PT_Cells_No_Med_T2D_LC_adj_cov_lmm_no_zi.csv"))
# full_results <- read.csv(fs::path(dir.results,"Hep_5_Fibrosis_Steatosis.csv"))

full_results$color <- ifelse(full_results$fdr < 0.05 & full_results$Beta > 0, "lightcoral",
                             ifelse(full_results$fdr < 0.05 & full_results$Beta < 0, "lightblue", "gray"))

# Identify significant points (fdr < 0.05)
significant_df <- full_results[full_results$fdr < 0.05, ]
# 
# full_results$color3 <- ifelse(full_results$fdr3 < 0.2 & full_results$Beta3 > 0, "lightcoral",
#                               ifelse(full_results$fdr3 < 0.2 & full_results$Beta3 < 0, "lightblue", "gray"))
# 
# # Identify significant points (fdr < 0.05)
# significant_df3 <- full_results[full_results$fdr3 < 0.2, ]

Genes <- length(unique(full_results$Gene))
Cells <- ncol(so_celltype)

#Figure Range
max <- max(significant_df$Beta,na.rm=T)
min <- min(significant_df$Beta,na.rm=T)


# full_results$Log2FC <- 2^(full_results$Beta)
# Create the volcano plot using ggplot
volcano_plot <- ggplot(full_results, aes(x = Beta, y = PValue10, color = color)) +
  geom_point(alpha = 0.7) +  # Plot points with transparency
  scale_color_identity() +  # Use the color column directly
  theme_minimal() +  # Minimal theme
  labs(
    title = "Type 2 Diabetes vs. Lean Controls (All No Medication)",
    subtitle = "PT Cells, Adjusted for Age, Sex & BMI",
    x = "Fold Change",
    y = "-log10(P-Value)",
    color = "FC Direction Direction",
    caption = paste0("FDR < 0.05, Genes = ",Genes,", Cells = ",Cells,", Non-Convergence Rate: ",Nonconvergence_Rate)
  ) +
  xlim(min,max)+
  theme(
    plot.title = element_text(hjust = 0),
    axis.text.x = element_text(angle = 0, hjust = 1)
  )+
  # # Add labels for significant points
  geom_text(data = significant_df, aes(label = Gene),
            vjust = 1, hjust = 1, size = 3, check_overlap = TRUE, color = "black")
# Add labels for significant points with ggrepel
# geom_text_repel(data = significant_df, aes(label = Gene),
#                 size = 3, color = "black", box.padding = 0.5, max.overlaps = 20)

volcano_plot
pdf(fs::path(dir.results,"Plot_PT_Cells_T2D_LC_NoMed.pdf"),width=10,height=7)
print(volcano_plot)
dev.off()

```
#### GSEA
```{r echo = F}
#PT
full_results <- read.csv(fs::path(dir.results,"PT_Cells_No_Med_T2D_LC_adj_cov_lmm_no_zi.csv")) %>%
  dplyr::select(-X)
sig_pos <- full_results %>% 
  filter(Beta>0 & fdr<0.05)
sig_neg <- full_results %>% 
  filter(Beta<0 & fdr<0.05)

#Run postive enricher analysis 
sig_pos_en <- enrichr(as.character(sig_pos$Gene), dbs)
#Plot results
# Convert enrichment results to a dataframe
sig_pos_en_df <- as.data.frame(sig_pos_en[[1]])
# Extract the number of overlapping genes
sig_pos_en_df$Num_Genes <- as.numeric(sub("/.*", "", sig_pos_en_df$Overlap))
# Filter pathways with at least 3 overlapping genes
filtered_results <- subset(sig_pos_en_df, Num_Genes >= 3)
# Plot the filtered results
plotEnrich(filtered_results, title = "Model 1 Positive Hepatocytes - GO")
ggsave(fs::path(dir.results,"All_Cell_Types_pos_gsea_go.jpeg"),width=10,height=7)

# Convert enrichment results to a dataframe
sig_pos_en_df <- as.data.frame(sig_pos_en[[2]])
# Extract the number of overlapping genes
sig_pos_en_df$Num_Genes <- as.numeric(sub("/.*", "", sig_pos_en_df$Overlap))
# Filter pathways with at least 3 overlapping genes
filtered_results <- subset(sig_pos_en_df, Num_Genes >= 3)
# Plot the filtered results
plotEnrich(filtered_results, title = "Model 1 Positive Hepatocytes - Kegg")
# plotEnrich(as.data.frame(sig_pos_en[[2]]), title = "Model 1 Positive Hepatocytes - Kegg")
ggsave(fs::path(dir.results,"All_Cell_Types_pos_gsea_kegg.jpeg"),width=10,height=7)

sig_pos_en_df <- as.data.frame(sig_pos_en[[3]])
# Extract the number of overlapping genes
sig_pos_en_df$Num_Genes <- as.numeric(sub("/.*", "", sig_pos_en_df$Overlap))
# Filter pathways with at least 3 overlapping genes
filtered_results <- subset(sig_pos_en_df, Num_Genes >= 3)
# Plot the filtered results
plotEnrich(filtered_results, title = "Model 1 Positive Hepatocytes - Reactome '22")
# plotEnrich(as.data.frame(sig_pos_en[[3]]), title = "Model 1 Positive Hepatocytes - Reactome '22")
ggsave(fs::path(dir.results,"All_Cell_Types_pos_gsea_reactome22.jpeg"),width=10,height=7)

sig_pos_en_df <- as.data.frame(sig_pos_en[[4]])
# Extract the number of overlapping genes
sig_pos_en_df$Num_Genes <- as.numeric(sub("/.*", "", sig_pos_en_df$Overlap))
# Filter pathways with at least 3 overlapping genes
filtered_results <- subset(sig_pos_en_df, Num_Genes >= 3)
# Plot the filtered results
plotEnrich(filtered_results, title = "Model 1 Positive Hepatocytes - Reactome '24")
# plotEnrich(as.data.frame(sig_pos_en[[4]]), title = "Model 1 Positive Hepatocytes - Reactome '24")
ggsave(fs::path(dir.results,"All_Cell_Types_pos_gsea_reactome24.jpeg"),width=10,height=7)

#Run negative enricher analysis 
sig_neg_en <- enrichr(as.character(sig_neg$Gene), dbs)

# Convert enrichment results to a dataframe
sig_neg_en_df <- as.data.frame(sig_neg_en[[1]])
# Extract the number of overlapping genes
sig_neg_en_df$Num_Genes <- as.numeric(sub("/.*", "", sig_neg_en_df$Overlap))
# Filter pathways with at least 3 overlapping genes
filtered_results <- subset(sig_neg_en_df, Num_Genes >= 3)
# Plot the filtered results
plotEnrich(filtered_results, title = "Model 1 Negative Hepatocytes - GO")
ggsave(fs::path(dir.results,"All_Cell_Types_neg_gsea_go.jpeg"),width=10,height=7)

# Convert enrichment results to a dataframe
sig_neg_en_df <- as.data.frame(sig_neg_en[[2]])
# Extract the number of overlapping genes
sig_neg_en_df$Num_Genes <- as.numeric(sub("/.*", "", sig_neg_en_df$Overlap))
# Filter pathways with at least 3 overlapping genes
filtered_results <- subset(sig_neg_en_df, Num_Genes >= 3)
# Plot the filtered results
plotEnrich(filtered_results, title = "Model 1 Negative Hepatocytes - Kegg")
# plotEnrich(as.data.frame(sig_neg_en[[2]]), title = "Model 1 Negative Hepatocytes - Kegg")
ggsave(fs::path(dir.results,"All_Cell_Types_neg_gsea_kegg.jpeg"),width=10,height=7)

sig_neg_en_df <- as.data.frame(sig_neg_en[[3]])
# Extract the number of overlapping genes
sig_neg_en_df$Num_Genes <- as.numeric(sub("/.*", "", sig_neg_en_df$Overlap))
# Filter pathways with at least 3 overlapping genes
filtered_results <- subset(sig_neg_en_df, Num_Genes >= 3)
# Plot the filtered results
plotEnrich(filtered_results, title = "Model 1 Negative Hepatocytes - Reactome '22")
# plotEnrich(as.data.frame(sig_neg_en[[3]]), title = "Model 1 Negative Hepatocytes - Reactome '22")
ggsave(fs::path(dir.results,"All_Cell_Types_neg_gsea_reactome22.jpeg"),width=10,height=7)

sig_neg_en_df <- as.data.frame(sig_neg_en[[4]])
# Extract the number of overlapping genes
sig_neg_en_df$Num_Genes <- as.numeric(sub("/.*", "", sig_neg_en_df$Overlap))
# Filter pathways with at least 3 overlapping genes
filtered_results <- subset(sig_neg_en_df, Num_Genes >= 3)
# Plot the filtered results
plotEnrich(filtered_results, title = "Model 1 Negative Hepatocytes - Reactome '24")
# plotEnrich(as.data.frame(sig_neg_en[[4]]), title = "Model 1 Negative Hepatocytes - Reactome '24")
ggsave(fs::path(dir.results,"All_Cell_Types_neg_gsea_reactome24.jpeg"),width=10,height=7)
```
#### ii. PT-S1/2
```{r}
so_celltype <- subset(so_kpmp_sc,KPMP_celltype2=="PT-S1/S2")
# so_celltype <- subset(so_kpmp_sc,celltype1=="PT")
ncol(so_celltype) #37948 cells
nrow(so_celltype) #14627 genes
# Extract the gene expression data for all genes
gene_expression <- as.data.frame(GetAssayData(so_celltype, layer = "data"))

#Assign gene names as rownames, cell names as colnames
# rownames(gene_expression) <- rownames(so_celltype) #Gene Names
# colnames(gene_expression) <- colnames(so_celltype) #Cell Names
#Transpose gene expression dataset to merge with metadata
gene_expression <- t(gene_expression)
gene_expression <- data.frame(gene_expression) #Make a dataframe again after transposing
#Set gene list
gene_list_total <- colnames(gene_expression)
gene_expression$cellname <- rownames(gene_expression) 
rownames(gene_expression) <- NULL

# Extract the metadata
metadata <- so_celltype@meta.data
# metadata <- metadata %>%
#   mutate(across(everything(),~ifelse(.==".",NA,.)))
metadata$cellname <- rownames(metadata)
rownames(metadata) <- NULL

# Combine the gene expression data and metadata
data <- tidylog::full_join(metadata,gene_expression,by="cellname")
rm(metadata,gene_expression)

#Compare Health Controls to Type 2 Diabetes on no meds
data_subset <- data %>%
  filter(group=="Type_2_Diabetes" | group=="Lean_Control") %>%
  filter(medication=="no_med")
rm(data)
data_subset$group <- factor(data_subset$group)
data_subset$group <- relevel(data_subset$group,ref="Lean_Control")
data_subset <- data_subset %>% 
  dplyr::select(c("kit_id","group","age","sex","bmi",all_of(gene_list_total)))

# library(glmmTMB)
# library(lme4)
# library(doParallel)
# library(foreach)
# Set batch size
batch_size <- 2000
total_cores <- 50

# Deduplicate gene list to avoid double-processing
gene_list_total <- unique(gene_list_total)

# Split the gene list into batches
gene_batches <- split(gene_list_total, ceiling(seq_along(gene_list_total) / batch_size))

# Define the function that processes a single gene
process_gene <- function(gene) {
  if (sum(data_subset[[gene]]) > 0) {
    m1 <- as.formula(paste0(gene, " ~ group + age + sex + bmi + (1 | kit_id)"))
    
    model1 <- tryCatch({
      glmmTMB(m1, data = data_subset, family = gaussian)
    }, error = function(e) {
      return(NULL)
    })
    
    if (!is.null(model1)) {
      Beta <- summary(model1)$coef$cond[2,1]
      PValue <- summary(model1)$coef$cond[2,4]
    } else {
      Beta <- NA
      PValue <- NA
    }
  } else {
    Beta <- NA
    PValue <- NA
  }
  
  return(data.frame(Gene = gene, Beta = Beta, PValue = PValue))
}

# Set number of cores
total_cores <- parallel::detectCores() - 1

# Initialize and register cluster
cl <- makeCluster(total_cores)
registerDoParallel(cl)

# Run analysis per batch
all_results <- lapply(seq_along(gene_batches), function(batch_idx) {
  batch_genes <- gene_batches[[batch_idx]]
  
  batch_results <- foreach(
    gene = batch_genes,
    .combine = rbind,
    .packages = c("glmmTMB", "lme4"),
    .export = c("process_gene", "data_subset")
  ) %dopar% {
    process_gene(gene)
  }
  
  cat("Processed batch", batch_idx, "with", length(batch_genes), "genes\n")
  return(batch_results)
})

# Stop cluster
stopCluster(cl)

# Combine all batch results into a single data frame
final_results <- do.call(rbind, all_results)

# #Try as batch loop
# # Set batch size
# batch_size <- 2000
# total_cores <- 50  
# 
# # Get total number of genes
# total_genes <- length(gene_list_total)
# 
# # Create an empty list to store results
# all_results <- list()
# 
# # Loop over batches
# for (start_idx in seq(1, total_genes, by = batch_size)) {
#   
#   # Define end index for the batch
#   end_idx <- min(start_idx + batch_size - 1, total_genes)
#   
#   # Subset genes for this batch
#   gene_list <- gene_list_total[start_idx:end_idx]
#   
#   # Set number of cores to use
#   cl <- makeCluster(total_cores)
#   registerDoParallel(cl)
# 
#   # Define function to process a single gene
#   process_gene <- function(gene_idx) {
#     gene <- gene_list[gene_idx]  # Get the gene name based on the index
#     
#     # Ensure that the gene exists in your data subset
#     if (sum(data_subset[[gene]]) > 0) {
#       # Define model formula
#       m1 <- as.formula(paste0(gene, " ~ group + age + sex + bmi + (1 | kit_id)"))
#       
#       # Fit the zero-inflated mixed-effects model with error handling
#       model1 <- tryCatch({
#         # glmmTMB(m1, data = data_subset, family = gaussian, ziformula = ~1)
#         glmmTMB(m1, data = data_subset, family = gaussian)
#       }, error = function(e) {
#         return(NULL)
#       })
#       
#       if (!is.null(model1)) {
#         Beta <- summary(model1)$coef$cond[2,1]
#         PValue <- summary(model1)$coef$cond[2,4]
#         # zi <- summary(model1)$coef$zi[1,1]
#         # zi_pval <- summary(model1)$coef$zi[1,4]
#       } else {
#         Beta <- NA
#         PValue <- NA
#         # zi <- NA
#         # zi_pval <- NA
#       }
#       
#     } else {
#       Beta <- NA
#       PValue <- NA
#       # zi <- NA
#       # zi_pval <- NA
#     }
#     
#     # Return result
#     # return(data.frame(Gene = gene, Beta = Beta, PValue = PValue, ZI_Intercept = zi, ZI_PValue = zi_pval))
#     return(data.frame(Gene = gene, Beta = Beta, PValue = PValue))
#   }
#   
#   # Run the analysis in parallel
#   batch_results <- foreach(i = seq_along(gene_list), .combine = rbind, .packages = c("glmmTMB", "lme4")) %dopar% {
#     process_gene(i)
#   }
#   
#   # Stop the cluster
#   stopCluster(cl)
#   
#   # Store batch results
#   all_results[[length(all_results) + 1]] <- batch_results
#   
#   # Print progress
#   cat("Processed batch:", start_idx, "to", end_idx, "\n")
#   
# }
# 
# # Combine all batch results into a single data frame
# final_results <- do.call(rbind, all_results)

# #Make volcano plot of all gene results for group
full_results <- final_results %>%
  mutate(fdr=p.adjust(PValue,method="fdr"))  
# mutate(fdr3=p.adjust(PValue3,method="fdr"))
full_results$PValue10 <- -log10(pmax(full_results$PValue, 1e-10))  # Avoid log(0)
# full_results$PValue10_3 <- -log10(pmax(full_results$PValue3, 1e-10))  # Avoid log(0)
Nonconvergence_Rate <- paste0(round(((length(which(is.na(full_results$PValue)))+length(which(full_results$PValue=="NaN")))/length(full_results$Gene))*100,0),"%")
# view(full_results)

# full_results <- full_results %>%
#   mutate(FoldChange=2^Beta-1)  
# mutate(FoldChange3=2^Beta3-1)
# full_results <- full_results %>% 
#   dplyr::rename(color=color2,
#                 FoldChange=FoldChange2) %>%
#   # dplyr::rename(fdr=fdr2,
#   #               PValue10=PValue10_2) %>%
#   # dplyr::select(-c("Beta3","PValue3")) %>% 
#   dplyr::select(-c("color3","FoldChange3"))

# full_results2 <- full_results
# # write file
# bucket <- "scrna" # bucket name in Kopah
# temp_file <- tempfile(fileext = ".csv") # need to create a temporary file
# write.csv(full_results, row.names = FALSE) # save the temp file locally (will delete after R session)
# s3$upload_file(temp_file, bucket, "Kidney transcriptomics/Results/EC_AEA_No_Med_T2D_LC_adj_cov.csv")
# write.csv(full_results,"test.csv")

write.csv(full_results,fs::path(dir.results,"PT_S1S2_Cells_No_Med_T2D_LC_adj_cov_lmm_no_zi.csv"))
# full_results <- read.csv(fs::path(dir.results,"Hep_5_Fibrosis_Steatosis.csv"))

full_results$color <- ifelse(full_results$fdr < 0.05 & full_results$Beta > 0, "lightcoral",
                             ifelse(full_results$fdr < 0.05 & full_results$Beta < 0, "lightblue", "gray"))

# Identify significant points (fdr < 0.05)
significant_df <- full_results[full_results$fdr < 0.05, ]
# 
# full_results$color3 <- ifelse(full_results$fdr3 < 0.2 & full_results$Beta3 > 0, "lightcoral",
#                               ifelse(full_results$fdr3 < 0.2 & full_results$Beta3 < 0, "lightblue", "gray"))
# 
# # Identify significant points (fdr < 0.05)
# significant_df3 <- full_results[full_results$fdr3 < 0.2, ]

Genes <- length(unique(full_results$Gene))
Cells <- ncol(so_celltype)

#Figure Range
# max <- max(significant_df$Beta,na.rm=T)
# min <- min(significant_df$Beta,na.rm=T)

max <- max(full_results$Beta,na.rm=T)
min <- min(full_results$Beta,na.rm=T)

# Create the volcano plot using ggplot
volcano_plot <- ggplot(full_results, aes(x = Beta, y = PValue10, color = color)) +
  geom_point(alpha = 0.7) +  # Plot points with transparency
  scale_color_identity() +  # Use the color column directly
  theme_minimal() +  # Minimal theme
  labs(
    title = "Type 2 Diabetes vs. Lean Controls (All No Medication)",
    subtitle = "PT-S1/S2, Adjusted for Age, Sex & BMI",
    x = "Fold Change",
    y = "-log10(P-Value)",
    color = "FC Direction Direction",
    caption = paste0("FDR < 0.05, Genes = ",Genes,", Cells = ",Cells,", Non-Convergence Rate: ",Nonconvergence_Rate)
  ) +
  xlim(min,max)+
  theme(
    plot.title = element_text(hjust = 0),
    axis.text.x = element_text(angle = 0, hjust = 1)
  )+
  # # Add labels for significant points
  geom_text(data = significant_df, aes(label = Gene),
            vjust = 1, hjust = 1, size = 3, check_overlap = TRUE, color = "black")
# Add labels for significant points with ggrepel
# geom_text_repel(data = significant_df, aes(label = Gene),
#                 size = 3, color = "black", box.padding = 0.5, max.overlaps = 20)

volcano_plot
pdf(fs::path(dir.results,"Plot_PT_S1S2_T2D_LC_NoMed_lmm_no_zi.pdf"),width=10,height=7)
print(volcano_plot)
dev.off()

```
#### iii. PT-S3
```{r}
so_celltype <- subset(so_kpmp_sc,KPMP_celltype2=="PT-S3")
# so_celltype <- subset(so_kpmp_sc,celltype1=="PT")
ncol(so_celltype) #37948 cells
nrow(so_celltype) #14627 genes
# Extract the gene expression data for all genes
gene_expression <- as.data.frame(GetAssayData(so_celltype, layer = "data"))

#Assign gene names as rownames, cell names as colnames
# rownames(gene_expression) <- rownames(so_celltype) #Gene Names
# colnames(gene_expression) <- colnames(so_celltype) #Cell Names
#Transpose gene expression dataset to merge with metadata
gene_expression <- t(gene_expression)
gene_expression <- data.frame(gene_expression) #Make a dataframe again after transposing
#Set gene list
gene_list_total <- colnames(gene_expression)
gene_expression$cellname <- rownames(gene_expression) 
rownames(gene_expression) <- NULL

# Extract the metadata
metadata <- so_celltype@meta.data
# metadata <- metadata %>%
#   mutate(across(everything(),~ifelse(.==".",NA,.)))
metadata$cellname <- rownames(metadata)
rownames(metadata) <- NULL

# Combine the gene expression data and metadata
data <- tidylog::full_join(metadata,gene_expression,by="cellname")
rm(metadata,gene_expression)

#Compare Health Controls to Type 2 Diabetes on no meds
data_subset <- data %>%
  filter(group=="Type_2_Diabetes" | group=="Lean_Control") %>%
  filter(medication=="no_med")
rm(data)
data_subset$group <- factor(data_subset$group)
data_subset$group <- relevel(data_subset$group,ref="Lean_Control")
data_subset <- data_subset %>% 
  dplyr::select(c("kit_id","group","age","sex","bmi",all_of(gene_list_total)))

# Set batch size
batch_size <- 2000
total_cores <- 50

# Deduplicate gene list to avoid double-processing
gene_list_total <- unique(gene_list_total)

# Split the gene list into batches
gene_batches <- split(gene_list_total, ceiling(seq_along(gene_list_total) / batch_size))

# Define the function that processes a single gene
process_gene <- function(gene) {
  if (sum(data_subset[[gene]]) > 0) {
    m1 <- as.formula(paste0(gene, " ~ group + age + sex + bmi + (1 | kit_id)"))
    
    model1 <- tryCatch({
      glmmTMB(m1, data = data_subset, family = gaussian)
    }, error = function(e) {
      return(NULL)
    })
    
    if (!is.null(model1)) {
      Beta <- summary(model1)$coef$cond[2,1]
      PValue <- summary(model1)$coef$cond[2,4]
    } else {
      Beta <- NA
      PValue <- NA
    }
  } else {
    Beta <- NA
    PValue <- NA
  }
  
  return(data.frame(Gene = gene, Beta = Beta, PValue = PValue))
}

# Set number of cores
total_cores <- parallel::detectCores() - 1

# Initialize and register cluster
cl <- makeCluster(total_cores)
registerDoParallel(cl)

# Run analysis per batch
all_results <- lapply(seq_along(gene_batches), function(batch_idx) {
  batch_genes <- gene_batches[[batch_idx]]
  
  batch_results <- foreach(
    gene = batch_genes,
    .combine = rbind,
    .packages = c("glmmTMB", "lme4"),
    .export = c("process_gene", "data_subset")
  ) %dopar% {
    process_gene(gene)
  }
  
  cat("Processed batch", batch_idx, "with", length(batch_genes), "genes\n")
  return(batch_results)
})

# Stop cluster
stopCluster(cl)

# Combine all batch results into a single data frame
final_results <- do.call(rbind, all_results)

# #Make volcano plot of all gene results for group
full_results <- final_results %>%
  mutate(fdr=p.adjust(PValue,method="fdr"))  
# mutate(fdr3=p.adjust(PValue3,method="fdr"))
full_results$PValue10 <- -log10(pmax(full_results$PValue, 1e-10))  # Avoid log(0)
# full_results$PValue10_3 <- -log10(pmax(full_results$PValue3, 1e-10))  # Avoid log(0)
Nonconvergence_Rate <- paste0(round(((length(which(is.na(full_results$PValue)))+length(which(full_results$PValue=="NaN")))/length(full_results$Gene))*100,0),"%")
# view(full_results)

# full_results <- full_results %>%
#   mutate(FoldChange=2^Beta-1)  
# mutate(FoldChange3=2^Beta3-1)
# full_results <- full_results %>% 
#   dplyr::rename(color=color2,
#                 FoldChange=FoldChange2) %>%
#   # dplyr::rename(fdr=fdr2,
#   #               PValue10=PValue10_2) %>%
#   # dplyr::select(-c("Beta3","PValue3")) %>% 
#   dplyr::select(-c("color3","FoldChange3"))

# full_results2 <- full_results
# # write file
# bucket <- "scrna" # bucket name in Kopah
# temp_file <- tempfile(fileext = ".csv") # need to create a temporary file
# write.csv(full_results, row.names = FALSE) # save the temp file locally (will delete after R session)
# s3$upload_file(temp_file, bucket, "Kidney transcriptomics/Results/EC_AEA_No_Med_T2D_LC_adj_cov.csv")
# write.csv(full_results,"test.csv")

write.csv(full_results,fs::path(dir.results,"PT_S3_Cells_No_Med_T2D_LC_adj_cov_lmm_no_zi.csv"))
# full_results <- read.csv(fs::path(dir.results,"Hep_5_Fibrosis_Steatosis.csv"))

full_results$color <- ifelse(full_results$fdr < 0.05 & full_results$Beta > 0, "lightcoral",
                             ifelse(full_results$fdr < 0.05 & full_results$Beta < 0, "lightblue", "gray"))

# Identify significant points (fdr < 0.05)
significant_df <- full_results[full_results$fdr < 0.05, ]
# 
# full_results$color3 <- ifelse(full_results$fdr3 < 0.2 & full_results$Beta3 > 0, "lightcoral",
#                               ifelse(full_results$fdr3 < 0.2 & full_results$Beta3 < 0, "lightblue", "gray"))
# 
# # Identify significant points (fdr < 0.05)
# significant_df3 <- full_results[full_results$fdr3 < 0.2, ]

Genes <- length(unique(full_results$Gene))
Cells <- ncol(so_celltype)

#Figure Range
# max <- max(significant_df$Beta,na.rm=T)
# min <- min(significant_df$Beta,na.rm=T)

max <- max(full_results$Beta,na.rm=T)
min <- min(full_results$Beta,na.rm=T)

# Create the volcano plot using ggplot
volcano_plot <- ggplot(full_results, aes(x = Beta, y = PValue10, color = color)) +
  geom_point(alpha = 0.7) +  # Plot points with transparency
  scale_color_identity() +  # Use the color column directly
  theme_minimal() +  # Minimal theme
  labs(
    title = "Type 2 Diabetes vs. Lean Controls (All No Medication)",
    subtitle = "PT-S3, Adjusted for Age, Sex & BMI",
    x = "Fold Change",
    y = "-log10(P-Value)",
    color = "FC Direction Direction",
    caption = paste0("FDR < 0.05, Genes = ",Genes,", Cells = ",Cells,", Non-Convergence Rate: ",Nonconvergence_Rate)
  ) +
  xlim(min,max)+
  theme(
    plot.title = element_text(hjust = 0),
    axis.text.x = element_text(angle = 0, hjust = 1)
  )+
  # # Add labels for significant points
  geom_text(data = significant_df, aes(label = Gene),
            vjust = 1, hjust = 1, size = 3, check_overlap = TRUE, color = "black")
# Add labels for significant points with ggrepel
# geom_text_repel(data = significant_df, aes(label = Gene),
#                 size = 3, color = "black", box.padding = 0.5, max.overlaps = 20)

volcano_plot
pdf(fs::path(dir.results,"Plot_PT_S3_T2D_LC_NoMed_lmm_no_zi.pdf"),width=10,height=7)
print(volcano_plot)
dev.off()

```
#### iv. aPT
```{r}
so_celltype <- subset(so_kpmp_sc,KPMP_celltype2=="aPT")
# so_celltype <- subset(so_kpmp_sc,celltype1=="PT")
ncol(so_celltype) #37948 cells
nrow(so_celltype) #14627 genes
# Extract the gene expression data for all genes
gene_expression <- as.data.frame(GetAssayData(so_celltype, layer = "data"))

#Assign gene names as rownames, cell names as colnames
# rownames(gene_expression) <- rownames(so_celltype) #Gene Names
# colnames(gene_expression) <- colnames(so_celltype) #Cell Names
#Transpose gene expression dataset to merge with metadata
gene_expression <- t(gene_expression)
gene_expression <- data.frame(gene_expression) #Make a dataframe again after transposing
#Set gene list
gene_list_total <- colnames(gene_expression)
gene_expression$cellname <- rownames(gene_expression) 
rownames(gene_expression) <- NULL

# Extract the metadata
metadata <- so_celltype@meta.data
# metadata <- metadata %>%
#   mutate(across(everything(),~ifelse(.==".",NA,.)))
metadata$cellname <- rownames(metadata)
rownames(metadata) <- NULL

# Combine the gene expression data and metadata
data <- tidylog::full_join(metadata,gene_expression,by="cellname")
rm(metadata,gene_expression)

#Compare Health Controls to Type 2 Diabetes on no meds
data_subset <- data %>%
  filter(group=="Type_2_Diabetes" | group=="Lean_Control") %>%
  filter(medication=="no_med")
rm(data)
data_subset$group <- factor(data_subset$group)
data_subset$group <- relevel(data_subset$group,ref="Lean_Control")
data_subset <- data_subset %>% 
  dplyr::select(c("kit_id","group","age","sex","bmi",all_of(gene_list_total)))

# Set batch size
batch_size <- 2000
total_cores <- 50

# Deduplicate gene list to avoid double-processing
gene_list_total <- unique(gene_list_total)

# Split the gene list into batches
gene_batches <- split(gene_list_total, ceiling(seq_along(gene_list_total) / batch_size))

# Define the function that processes a single gene
process_gene <- function(gene) {
  if (sum(data_subset[[gene]]) > 0) {
    m1 <- as.formula(paste0(gene, " ~ group + age + sex + bmi + (1 | kit_id)"))
    
    model1 <- tryCatch({
      glmmTMB(m1, data = data_subset, family = gaussian)
    }, error = function(e) {
      return(NULL)
    })
    
    if (!is.null(model1)) {
      Beta <- summary(model1)$coef$cond[2,1]
      PValue <- summary(model1)$coef$cond[2,4]
    } else {
      Beta <- NA
      PValue <- NA
    }
  } else {
    Beta <- NA
    PValue <- NA
  }
  
  return(data.frame(Gene = gene, Beta = Beta, PValue = PValue))
}

# Set number of cores
total_cores <- parallel::detectCores() - 1

# Initialize and register cluster
cl <- makeCluster(total_cores)
registerDoParallel(cl)


rm(all_results, final_results,full_results,significant_df,vol)
# Run analysis per batch
all_results <- lapply(seq_along(gene_batches), function(batch_idx) {
  batch_genes <- gene_batches[[batch_idx]]
  
  batch_results <- foreach(
    gene = batch_genes,
    .combine = rbind,
    .packages = c("glmmTMB", "lme4"),
    .export = c("process_gene", "data_subset")
  ) %dopar% {
    process_gene(gene)
  }
  
  cat("Processed batch", batch_idx, "with", length(batch_genes), "genes\n")
  return(batch_results)
})

# Stop cluster
stopCluster(cl)

# Combine all batch results into a single data frame
final_results <- do.call(rbind, all_results)

# #Make volcano plot of all gene results for group
full_results <- final_results %>%
  mutate(fdr=p.adjust(PValue,method="fdr"))  
# mutate(fdr3=p.adjust(PValue3,method="fdr"))
full_results$PValue10 <- -log10(pmax(full_results$PValue, 1e-10))  # Avoid log(0)
# full_results$PValue10_3 <- -log10(pmax(full_results$PValue3, 1e-10))  # Avoid log(0)
Nonconvergence_Rate <- paste0(round(((length(which(is.na(full_results$PValue)))+length(which(full_results$PValue=="NaN")))/length(full_results$Gene))*100,0),"%")
# view(full_results)

write.csv(full_results,fs::path(dir.results,"aPT_Cells_No_Med_T2D_LC_adj_cov_lmm_no_zi.csv"))
# full_results <- read.csv(fs::path(dir.results,"Hep_5_Fibrosis_Steatosis.csv"))

full_results$color <- ifelse(full_results$fdr < 0.05 & full_results$Beta > 0, "lightcoral",
                             ifelse(full_results$fdr < 0.05 & full_results$Beta < 0, "lightblue", "gray"))

# Identify significant points (fdr < 0.05)
significant_df <- full_results[full_results$fdr < 0.05, ]
# 
# full_results$color3 <- ifelse(full_results$fdr3 < 0.2 & full_results$Beta3 > 0, "lightcoral",
#                               ifelse(full_results$fdr3 < 0.2 & full_results$Beta3 < 0, "lightblue", "gray"))
# 
# # Identify significant points (fdr < 0.05)
# significant_df3 <- full_results[full_results$fdr3 < 0.2, ]

Genes <- length(unique(full_results$Gene))
Cells <- ncol(so_celltype)

#Figure Range
# max <- max(significant_df$Beta,na.rm=T)
# min <- min(significant_df$Beta,na.rm=T)

max <- max(full_results$Beta,na.rm=T)
min <- min(full_results$Beta,na.rm=T)

# Create the volcano plot using ggplot
volcano_plot <- ggplot(full_results, aes(x = Beta, y = PValue10, color = color)) +
  geom_point(alpha = 0.7) +  # Plot points with transparency
  scale_color_identity() +  # Use the color column directly
  theme_minimal() +  # Minimal theme
  labs(
    title = "Type 2 Diabetes vs. Lean Controls (All No Medication)",
    subtitle = "aPT, Adjusted for Age, Sex & BMI",
    x = "Fold Change",
    y = "-log10(P-Value)",
    color = "FC Direction Direction",
    caption = paste0("FDR < 0.05, Genes = ",Genes,", Cells = ",Cells,", Non-Convergence Rate: ",Nonconvergence_Rate)
  ) +
  xlim(min,max)+
  theme(
    plot.title = element_text(hjust = 0),
    axis.text.x = element_text(angle = 0, hjust = 1)
  )+
  # # Add labels for significant points
  geom_text(data = significant_df, aes(label = Gene),
            vjust = 1, hjust = 1, size = 3, check_overlap = TRUE, color = "black")
# Add labels for significant points with ggrepel
# geom_text_repel(data = significant_df, aes(label = Gene),
#                 size = 3, color = "black", box.padding = 0.5, max.overlaps = 20)

volcano_plot
pdf(fs::path(dir.results,"Plot_aPT_T2D_LC_NoMed_lmm_no_zi.pdf"),width=10,height=7)
print(volcano_plot)
dev.off()

```

### b. EC Cells
#### i Pseudobulk EC
```{r}
#EC
so_celltype <- subset(so_kpmp_sc,celltype1=="EC")
ncol(so_celltype) #37948 cells
nrow(so_celltype) #14627 genes
# Extract the gene expression data for all genes
gene_expression <- as.data.frame(GetAssayData(so_celltype, layer = "data"))

#Assign gene names as rownames, cell names as colnames
# rownames(gene_expression) <- rownames(so_celltype) #Gene Names
# colnames(gene_expression) <- colnames(so_celltype) #Cell Names
#Transpose gene expression dataset to merge with metadata
gene_expression <- t(gene_expression)
gene_expression <- data.frame(gene_expression) #Make a dataframe again after transposing
#Set gene list
gene_list_total <- colnames(gene_expression)
gene_expression$cellname <- rownames(gene_expression) 
rownames(gene_expression) <- NULL

# Extract the metadata
metadata <- so_celltype@meta.data
# metadata <- metadata %>%
#   mutate(across(everything(),~ifelse(.==".",NA,.)))
metadata$cellname <- rownames(metadata)
rownames(metadata) <- NULL

# Combine the gene expression data and metadata
data <- tidylog::full_join(metadata,gene_expression,by="cellname")
rm(metadata,gene_expression)

#Compare Health Controls to Type 2 Diabetes on no meds
data_subset <- data %>%
  filter(group=="Type_2_Diabetes" | group=="Lean_Control") %>%
  filter(medication=="no_med")
rm(data)
data_subset$group <- factor(data_subset$group)
data_subset$group <- relevel(data_subset$group,ref="Lean_Control")
data_subset <- data_subset %>% 
  dplyr::select(c("kit_id","group","age","sex","bmi",all_of(gene_list_total)))

#Try as batch loop
# Set batch size
batch_size <- 2000
total_cores <- 50

# Deduplicate gene list to avoid double-processing
gene_list_total <- unique(gene_list_total)

# Split the gene list into batches
gene_batches <- split(gene_list_total, ceiling(seq_along(gene_list_total) / batch_size))

# Define the function that processes a single gene
process_gene <- function(gene) {
  if (sum(data_subset[[gene]]) > 0) {
    m1 <- as.formula(paste0(gene, " ~ group + age + sex + bmi + (1 | kit_id)"))
    
    model1 <- tryCatch({
      glmmTMB(m1, data = data_subset, family = gaussian)
    }, error = function(e) {
      return(NULL)
    })
    
    if (!is.null(model1)) {
      Beta <- summary(model1)$coef$cond[2,1]
      PValue <- summary(model1)$coef$cond[2,4]
    } else {
      Beta <- NA
      PValue <- NA
    }
  } else {
    Beta <- NA
    PValue <- NA
  }
  
  return(data.frame(Gene = gene, Beta = Beta, PValue = PValue))
}

# Set number of cores
total_cores <- parallel::detectCores() - 1

# Initialize and register cluster
cl <- makeCluster(total_cores)
registerDoParallel(cl)

# Run analysis per batch
all_results <- lapply(seq_along(gene_batches), function(batch_idx) {
  batch_genes <- gene_batches[[batch_idx]]
  
  batch_results <- foreach(
    gene = batch_genes,
    .combine = rbind,
    .packages = c("glmmTMB", "lme4"),
    .export = c("process_gene", "data_subset")
  ) %dopar% {
    process_gene(gene)
  }
  
  cat("Processed batch", batch_idx, "with", length(batch_genes), "genes\n")
  return(batch_results)
})

# Stop cluster
stopCluster(cl)

# Combine all batch results into a single data frame
final_results <- do.call(rbind, all_results)

# #Make volcano plot of all gene results for group
full_results <- final_results %>%
  mutate(fdr=p.adjust(PValue,method="fdr"))  
# mutate(fdr3=p.adjust(PValue3,method="fdr"))
full_results$PValue10 <- -log10(pmax(full_results$PValue, 1e-10))  # Avoid log(0)
# full_results$PValue10_3 <- -log10(pmax(full_results$PValue3, 1e-10))  # Avoid log(0)
Nonconvergence_Rate <- paste0(round(((length(which(is.na(full_results$PValue)))+length(which(full_results$PValue=="NaN")))/length(full_results$Gene))*100,0),"%")
# view(full_results)

# write.csv(full_results,"test.csv")

write.csv(full_results,fs::path(dir.results,"EC_Cells_No_Med_T2D_LC_adj_cov_lmm_no_zi.csv"))
# full_results <- read.csv(fs::path(dir.results,"Hep_5_Fibrosis_Steatosis.csv"))

full_results$color <- ifelse(full_results$fdr < 0.05 & full_results$Beta > 0, "lightcoral",
                             ifelse(full_results$fdr < 0.05 & full_results$Beta < 0, "lightblue", "gray"))

# Identify significant points (fdr < 0.05)
significant_df <- full_results[full_results$fdr < 0.05, ]
# 
# full_results$color3 <- ifelse(full_results$fdr3 < 0.2 & full_results$Beta3 > 0, "lightcoral",
#                               ifelse(full_results$fdr3 < 0.2 & full_results$Beta3 < 0, "lightblue", "gray"))
# 
# # Identify significant points (fdr < 0.05)
# significant_df3 <- full_results[full_results$fdr3 < 0.2, ]

Genes <- length(unique(full_results$Gene))
Cells <- ncol(so_celltype)

#Figure Range
max <- max(significant_df$Beta,na.rm=T)
min <- min(significant_df$Beta,na.rm=T)


# full_results$Log2FC <- 2^(full_results$Beta)
# Create the volcano plot using ggplot
volcano_plot <- ggplot(full_results, aes(x = Beta, y = PValue10, color = color)) +
  geom_point(alpha = 0.7) +  # Plot points with transparency
  scale_color_identity() +  # Use the color column directly
  theme_minimal() +  # Minimal theme
  labs(
    title = "Type 2 Diabetes vs. Lean Controls (All No Medication)",
    subtitle = "EC Cells, Adjusted for Age, Sex & BMI",
    x = "Fold Change",
    y = "-log10(P-Value)",
    color = "FC Direction Direction",
    caption = paste0("FDR < 0.05, Genes = ",Genes,", Cells = ",Cells,", Non-Convergence Rate: ",Nonconvergence_Rate)
  ) +
  xlim(min,max)+
  theme(
    plot.title = element_text(hjust = 0),
    axis.text.x = element_text(angle = 0, hjust = 1)
  )+
  # # Add labels for significant points
  geom_text(data = significant_df, aes(label = Gene),
            vjust = 1, hjust = 1, size = 3, check_overlap = TRUE, color = "black")
# Add labels for significant points with ggrepel
# geom_text_repel(data = significant_df, aes(label = Gene),
#                 size = 3, color = "black", box.padding = 0.5, max.overlaps = 20)

volcano_plot
pdf(fs::path(dir.results,"Plot_EC_Cells_T2D_LC_NoMed_lmm_no_zi.pdf"),width=10,height=7)
print(volcano_plot)
dev.off()

```
#### ii. EC-AVR
```{r}
so_celltype <- subset(so_kpmp_sc,KPMP_celltype2=="EC-AVR")
# so_celltype <- subset(so_kpmp_sc,celltype1=="PT")
ncol(so_celltype) #37948 cells
nrow(so_celltype) #14627 genes
# Extract the gene expression data for all genes
gene_expression <- as.data.frame(GetAssayData(so_celltype, layer = "data"))

#Assign gene names as rownames, cell names as colnames
# rownames(gene_expression) <- rownames(so_celltype) #Gene Names
# colnames(gene_expression) <- colnames(so_celltype) #Cell Names
#Transpose gene expression dataset to merge with metadata
gene_expression <- t(gene_expression)
gene_expression <- data.frame(gene_expression) #Make a dataframe again after transposing
#Set gene list
gene_list_total <- colnames(gene_expression)
gene_expression$cellname <- rownames(gene_expression) 
rownames(gene_expression) <- NULL

# Extract the metadata
metadata <- so_celltype@meta.data
# metadata <- metadata %>%
#   mutate(across(everything(),~ifelse(.==".",NA,.)))
metadata$cellname <- rownames(metadata)
rownames(metadata) <- NULL

# Combine the gene expression data and metadata
data <- tidylog::full_join(metadata,gene_expression,by="cellname")
rm(metadata,gene_expression)

#Compare Health Controls to Type 2 Diabetes on no meds
data_subset <- data %>%
  filter(group=="Type_2_Diabetes" | group=="Lean_Control") %>%
  filter(medication=="no_med")
rm(data)
data_subset$group <- factor(data_subset$group)
data_subset$group <- relevel(data_subset$group,ref="Lean_Control")
data_subset <- data_subset %>% 
  dplyr::select(c("kit_id","group","age","sex","bmi",all_of(gene_list_total)))

# library(glmmTMB)
# library(lme4)
# library(doParallel)
# library(foreach)
# Set batch size
batch_size <- 2000
total_cores <- 50

# Deduplicate gene list to avoid double-processing
gene_list_total <- unique(gene_list_total)

# Split the gene list into batches
gene_batches <- split(gene_list_total, ceiling(seq_along(gene_list_total) / batch_size))

# Define the function that processes a single gene
process_gene <- function(gene) {
  if (sum(data_subset[[gene]]) > 0) {
    m1 <- as.formula(paste0(gene, " ~ group + age + sex + bmi + (1 | kit_id)"))
    
    model1 <- tryCatch({
      glmmTMB(m1, data = data_subset, family = gaussian)
    }, error = function(e) {
      return(NULL)
    })
    
    if (!is.null(model1)) {
      Beta <- summary(model1)$coef$cond[2,1]
      PValue <- summary(model1)$coef$cond[2,4]
    } else {
      Beta <- NA
      PValue <- NA
    }
  } else {
    Beta <- NA
    PValue <- NA
  }
  
  return(data.frame(Gene = gene, Beta = Beta, PValue = PValue))
}

# Set number of cores
total_cores <- parallel::detectCores() - 1

# Initialize and register cluster
cl <- makeCluster(total_cores)
registerDoParallel(cl)

# Run analysis per batch
all_results <- lapply(seq_along(gene_batches), function(batch_idx) {
  batch_genes <- gene_batches[[batch_idx]]
  
  batch_results <- foreach(
    gene = batch_genes,
    .combine = rbind,
    .packages = c("glmmTMB", "lme4"),
    .export = c("process_gene", "data_subset")
  ) %dopar% {
    process_gene(gene)
  }
  
  cat("Processed batch", batch_idx, "with", length(batch_genes), "genes\n")
  return(batch_results)
})

# Stop cluster
stopCluster(cl)

# Combine all batch results into a single data frame
final_results <- do.call(rbind, all_results)

# #Make volcano plot of all gene results for group
full_results <- final_results %>%
  mutate(fdr=p.adjust(PValue,method="fdr"))  
# mutate(fdr3=p.adjust(PValue3,method="fdr"))
full_results$PValue10 <- -log10(pmax(full_results$PValue, 1e-10))  # Avoid log(0)
# full_results$PValue10_3 <- -log10(pmax(full_results$PValue3, 1e-10))  # Avoid log(0)
Nonconvergence_Rate <- paste0(round(((length(which(is.na(full_results$PValue)))+length(which(full_results$PValue=="NaN")))/length(full_results$Gene))*100,0),"%")
# view(full_results)

write.csv(full_results,fs::path(dir.results,"EC_AVR_Cells_No_Med_T2D_LC_adj_cov_lmm_no_zi.csv"))
# full_results <- read.csv(fs::path(dir.results,"Hep_5_Fibrosis_Steatosis.csv"))

full_results$color <- ifelse(full_results$fdr < 0.05 & full_results$Beta > 0, "lightcoral",
                             ifelse(full_results$fdr < 0.05 & full_results$Beta < 0, "lightblue", "gray"))

# Identify significant points (fdr < 0.05)
significant_df <- full_results[full_results$fdr < 0.05, ]
# 
# full_results$color3 <- ifelse(full_results$fdr3 < 0.2 & full_results$Beta3 > 0, "lightcoral",
#                               ifelse(full_results$fdr3 < 0.2 & full_results$Beta3 < 0, "lightblue", "gray"))
# 
# # Identify significant points (fdr < 0.05)
# significant_df3 <- full_results[full_results$fdr3 < 0.2, ]

Genes <- length(unique(full_results$Gene))
Cells <- ncol(so_celltype)

#Figure Range
# max <- max(significant_df$Beta,na.rm=T)
# min <- min(significant_df$Beta,na.rm=T)

max <- max(full_results$Beta,na.rm=T)
min <- min(full_results$Beta,na.rm=T)

# Create the volcano plot using ggplot
volcano_plot <- ggplot(full_results, aes(x = Beta, y = PValue10, color = color)) +
  geom_point(alpha = 0.7) +  # Plot points with transparency
  scale_color_identity() +  # Use the color column directly
  theme_minimal() +  # Minimal theme
  labs(
    title = "Type 2 Diabetes vs. Lean Controls (All No Medication)",
    subtitle = "EC-AVR, Adjusted for Age, Sex & BMI",
    x = "Fold Change",
    y = "-log10(P-Value)",
    color = "FC Direction Direction",
    caption = paste0("FDR < 0.05, Genes = ",Genes,", Cells = ",Cells,", Non-Convergence Rate: ",Nonconvergence_Rate)
  ) +
  xlim(min,max)+
  theme(
    plot.title = element_text(hjust = 0),
    axis.text.x = element_text(angle = 0, hjust = 1)
  )+
  # # Add labels for significant points
  geom_text(data = significant_df, aes(label = Gene),
            vjust = 1, hjust = 1, size = 3, check_overlap = TRUE, color = "black")
# Add labels for significant points with ggrepel
# geom_text_repel(data = significant_df, aes(label = Gene),
#                 size = 3, color = "black", box.padding = 0.5, max.overlaps = 20)

volcano_plot
pdf(fs::path(dir.results,"Plot_EC_AVR_T2D_LC_NoMed_lmm_no_zi.pdf"),width=10,height=7)
print(volcano_plot)
dev.off()

```
#### iii. EC-GC
```{r}
so_celltype <- subset(so_kpmp_sc,KPMP_celltype2=="EC-GC")
# so_celltype <- subset(so_kpmp_sc,celltype1=="PT")
ncol(so_celltype) #37948 cells
nrow(so_celltype) #14627 genes
# Extract the gene expression data for all genes
gene_expression <- as.data.frame(GetAssayData(so_celltype, layer = "data"))

#Assign gene names as rownames, cell names as colnames
# rownames(gene_expression) <- rownames(so_celltype) #Gene Names
# colnames(gene_expression) <- colnames(so_celltype) #Cell Names
#Transpose gene expression dataset to merge with metadata
gene_expression <- t(gene_expression)
gene_expression <- data.frame(gene_expression) #Make a dataframe again after transposing
#Set gene list
gene_list_total <- colnames(gene_expression)
gene_expression$cellname <- rownames(gene_expression) 
rownames(gene_expression) <- NULL

# Extract the metadata
metadata <- so_celltype@meta.data
# metadata <- metadata %>%
#   mutate(across(everything(),~ifelse(.==".",NA,.)))
metadata$cellname <- rownames(metadata)
rownames(metadata) <- NULL

# Combine the gene expression data and metadata
data <- tidylog::full_join(metadata,gene_expression,by="cellname")
rm(metadata,gene_expression)

#Compare Health Controls to Type 2 Diabetes on no meds
data_subset <- data %>%
  filter(group=="Type_2_Diabetes" | group=="Lean_Control") %>%
  filter(medication=="no_med")
rm(data)
data_subset$group <- factor(data_subset$group)
data_subset$group <- relevel(data_subset$group,ref="Lean_Control")
data_subset <- data_subset %>% 
  dplyr::select(c("kit_id","group","age","sex","bmi",all_of(gene_list_total)))

# library(glmmTMB)
# library(lme4)
# library(doParallel)
# library(foreach)
# Set batch size
batch_size <- 2000
total_cores <- 50

# Deduplicate gene list to avoid double-processing
gene_list_total <- unique(gene_list_total)

# Split the gene list into batches
gene_batches <- split(gene_list_total, ceiling(seq_along(gene_list_total) / batch_size))

# Define the function that processes a single gene
process_gene <- function(gene) {
  if (sum(data_subset[[gene]]) > 0) {
    m1 <- as.formula(paste0(gene, " ~ group + age + sex + bmi + (1 | kit_id)"))
    
    model1 <- tryCatch({
      glmmTMB(m1, data = data_subset, family = gaussian)
    }, error = function(e) {
      return(NULL)
    })
    
    if (!is.null(model1)) {
      Beta <- summary(model1)$coef$cond[2,1]
      PValue <- summary(model1)$coef$cond[2,4]
    } else {
      Beta <- NA
      PValue <- NA
    }
  } else {
    Beta <- NA
    PValue <- NA
  }
  
  return(data.frame(Gene = gene, Beta = Beta, PValue = PValue))
}

# Set number of cores
total_cores <- parallel::detectCores() - 1

# Initialize and register cluster
cl <- makeCluster(total_cores)
registerDoParallel(cl)

# Run analysis per batch
all_results <- lapply(seq_along(gene_batches), function(batch_idx) {
  batch_genes <- gene_batches[[batch_idx]]
  
  batch_results <- foreach(
    gene = batch_genes,
    .combine = rbind,
    .packages = c("glmmTMB", "lme4"),
    .export = c("process_gene", "data_subset")
  ) %dopar% {
    process_gene(gene)
  }
  
  cat("Processed batch", batch_idx, "with", length(batch_genes), "genes\n")
  return(batch_results)
})

# Stop cluster
stopCluster(cl)

# Combine all batch results into a single data frame
final_results <- do.call(rbind, all_results)

# #Make volcano plot of all gene results for group
full_results <- final_results %>%
  mutate(fdr=p.adjust(PValue,method="fdr"))  
# mutate(fdr3=p.adjust(PValue3,method="fdr"))
full_results$PValue10 <- -log10(pmax(full_results$PValue, 1e-10))  # Avoid log(0)
# full_results$PValue10_3 <- -log10(pmax(full_results$PValue3, 1e-10))  # Avoid log(0)
Nonconvergence_Rate <- paste0(round(((length(which(is.na(full_results$PValue)))+length(which(full_results$PValue=="NaN")))/length(full_results$Gene))*100,0),"%")
# view(full_results)

write.csv(full_results,fs::path(dir.results,"EC_GC_Cells_No_Med_T2D_LC_adj_cov_lmm_no_zi.csv"))
# full_results <- read.csv(fs::path(dir.results,"Hep_5_Fibrosis_Steatosis.csv"))

full_results$color <- ifelse(full_results$fdr < 0.05 & full_results$Beta > 0, "lightcoral",
                             ifelse(full_results$fdr < 0.05 & full_results$Beta < 0, "lightblue", "gray"))

# Identify significant points (fdr < 0.05)
significant_df <- full_results[full_results$fdr < 0.05, ]
# 
# full_results$color3 <- ifelse(full_results$fdr3 < 0.2 & full_results$Beta3 > 0, "lightcoral",
#                               ifelse(full_results$fdr3 < 0.2 & full_results$Beta3 < 0, "lightblue", "gray"))
# 
# # Identify significant points (fdr < 0.05)
# significant_df3 <- full_results[full_results$fdr3 < 0.2, ]

Genes <- length(unique(full_results$Gene))
Cells <- ncol(so_celltype)

#Figure Range
# max <- max(significant_df$Beta,na.rm=T)
# min <- min(significant_df$Beta,na.rm=T)

max <- max(full_results$Beta,na.rm=T)
min <- min(full_results$Beta,na.rm=T)

# Create the volcano plot using ggplot
volcano_plot <- ggplot(full_results, aes(x = Beta, y = PValue10, color = color)) +
  geom_point(alpha = 0.7) +  # Plot points with transparency
  scale_color_identity() +  # Use the color column directly
  theme_minimal() +  # Minimal theme
  labs(
    title = "Type 2 Diabetes vs. Lean Controls (All No Medication)",
    subtitle = "EC-GC, Adjusted for Age, Sex & BMI",
    x = "Fold Change",
    y = "-log10(P-Value)",
    color = "FC Direction Direction",
    caption = paste0("FDR < 0.05, Genes = ",Genes,", Cells = ",Cells,", Non-Convergence Rate: ",Nonconvergence_Rate)
  ) +
  xlim(min,max)+
  theme(
    plot.title = element_text(hjust = 0),
    axis.text.x = element_text(angle = 0, hjust = 1)
  )+
  # # Add labels for significant points
  geom_text(data = significant_df, aes(label = Gene),
            vjust = 1, hjust = 1, size = 3, check_overlap = TRUE, color = "black")
# Add labels for significant points with ggrepel
# geom_text_repel(data = significant_df, aes(label = Gene),
#                 size = 3, color = "black", box.padding = 0.5, max.overlaps = 20)

volcano_plot
pdf(fs::path(dir.results,"Plot_EC_GC_T2D_LC_NoMed_lmm_no_zi.pdf"),width=10,height=7)
print(volcano_plot)
dev.off()

```
#### iv. EC-PTC
```{r}
so_celltype <- subset(so_kpmp_sc,KPMP_celltype2=="EC-PTC")
# so_celltype <- subset(so_kpmp_sc,celltype1=="PT")
ncol(so_celltype) #37948 cells
nrow(so_celltype) #14627 genes
# Extract the gene expression data for all genes
gene_expression <- as.data.frame(GetAssayData(so_celltype, layer = "data"))

#Assign gene names as rownames, cell names as colnames
# rownames(gene_expression) <- rownames(so_celltype) #Gene Names
# colnames(gene_expression) <- colnames(so_celltype) #Cell Names
#Transpose gene expression dataset to merge with metadata
gene_expression <- t(gene_expression)
gene_expression <- data.frame(gene_expression) #Make a dataframe again after transposing
#Set gene list
gene_list_total <- colnames(gene_expression)
gene_expression$cellname <- rownames(gene_expression) 
rownames(gene_expression) <- NULL

# Extract the metadata
metadata <- so_celltype@meta.data
# metadata <- metadata %>%
#   mutate(across(everything(),~ifelse(.==".",NA,.)))
metadata$cellname <- rownames(metadata)
rownames(metadata) <- NULL

# Combine the gene expression data and metadata
data <- tidylog::full_join(metadata,gene_expression,by="cellname")
rm(metadata,gene_expression)

#Compare Health Controls to Type 2 Diabetes on no meds
data_subset <- data %>%
  filter(group=="Type_2_Diabetes" | group=="Lean_Control") %>%
  filter(medication=="no_med")
rm(data)
data_subset$group <- factor(data_subset$group)
data_subset$group <- relevel(data_subset$group,ref="Lean_Control")
data_subset <- data_subset %>% 
  dplyr::select(c("kit_id","group","age","sex","bmi",all_of(gene_list_total)))

# library(glmmTMB)
# library(lme4)
# library(doParallel)
# library(foreach)
# Set batch size
batch_size <- 2000
total_cores <- 50

# Deduplicate gene list to avoid double-processing
gene_list_total <- unique(gene_list_total)

# Split the gene list into batches
gene_batches <- split(gene_list_total, ceiling(seq_along(gene_list_total) / batch_size))

# Define the function that processes a single gene
process_gene <- function(gene) {
  if (sum(data_subset[[gene]]) > 0) {
    m1 <- as.formula(paste0(gene, " ~ group + age + sex + bmi + (1 | kit_id)"))
    
    model1 <- tryCatch({
      glmmTMB(m1, data = data_subset, family = gaussian)
    }, error = function(e) {
      return(NULL)
    })
    
    if (!is.null(model1)) {
      Beta <- summary(model1)$coef$cond[2,1]
      PValue <- summary(model1)$coef$cond[2,4]
    } else {
      Beta <- NA
      PValue <- NA
    }
  } else {
    Beta <- NA
    PValue <- NA
  }
  
  return(data.frame(Gene = gene, Beta = Beta, PValue = PValue))
}

# Set number of cores
total_cores <- parallel::detectCores() - 1

# Initialize and register cluster
cl <- makeCluster(total_cores)
registerDoParallel(cl)

# Run analysis per batch
all_results <- lapply(seq_along(gene_batches), function(batch_idx) {
  batch_genes <- gene_batches[[batch_idx]]
  
  batch_results <- foreach(
    gene = batch_genes,
    .combine = rbind,
    .packages = c("glmmTMB", "lme4"),
    .export = c("process_gene", "data_subset")
  ) %dopar% {
    process_gene(gene)
  }
  
  cat("Processed batch", batch_idx, "with", length(batch_genes), "genes\n")
  return(batch_results)
})

# Stop cluster
stopCluster(cl)

# Combine all batch results into a single data frame
final_results <- do.call(rbind, all_results)

# #Make volcano plot of all gene results for group
full_results <- final_results %>%
  mutate(fdr=p.adjust(PValue,method="fdr"))  
# mutate(fdr3=p.adjust(PValue3,method="fdr"))
full_results$PValue10 <- -log10(pmax(full_results$PValue, 1e-10))  # Avoid log(0)
# full_results$PValue10_3 <- -log10(pmax(full_results$PValue3, 1e-10))  # Avoid log(0)
Nonconvergence_Rate <- paste0(round(((length(which(is.na(full_results$PValue)))+length(which(full_results$PValue=="NaN")))/length(full_results$Gene))*100,0),"%")
# view(full_results)

write.csv(full_results,fs::path(dir.results,"EC_PTC_Cells_No_Med_T2D_LC_adj_cov_lmm_no_zi.csv"))
# full_results <- read.csv(fs::path(dir.results,"Hep_5_Fibrosis_Steatosis.csv"))

full_results$color <- ifelse(full_results$fdr < 0.05 & full_results$Beta > 0, "lightcoral",
                             ifelse(full_results$fdr < 0.05 & full_results$Beta < 0, "lightblue", "gray"))

# Identify significant points (fdr < 0.05)
significant_df <- full_results[full_results$fdr < 0.05, ]
# 
# full_results$color3 <- ifelse(full_results$fdr3 < 0.2 & full_results$Beta3 > 0, "lightcoral",
#                               ifelse(full_results$fdr3 < 0.2 & full_results$Beta3 < 0, "lightblue", "gray"))
# 
# # Identify significant points (fdr < 0.05)
# significant_df3 <- full_results[full_results$fdr3 < 0.2, ]

Genes <- length(unique(full_results$Gene))
Cells <- ncol(so_celltype)

#Figure Range
# max <- max(significant_df$Beta,na.rm=T)
# min <- min(significant_df$Beta,na.rm=T)

max <- max(full_results$Beta,na.rm=T)
min <- min(full_results$Beta,na.rm=T)

# Create the volcano plot using ggplot
volcano_plot <- ggplot(full_results, aes(x = Beta, y = PValue10, color = color)) +
  geom_point(alpha = 0.7) +  # Plot points with transparency
  scale_color_identity() +  # Use the color column directly
  theme_minimal() +  # Minimal theme
  labs(
    title = "Type 2 Diabetes vs. Lean Controls (All No Medication)",
    subtitle = "EC-PTC, Adjusted for Age, Sex & BMI",
    x = "Fold Change",
    y = "-log10(P-Value)",
    color = "FC Direction Direction",
    caption = paste0("FDR < 0.05, Genes = ",Genes,", Cells = ",Cells,", Non-Convergence Rate: ",Nonconvergence_Rate)
  ) +
  xlim(min,max)+
  theme(
    plot.title = element_text(hjust = 0),
    axis.text.x = element_text(angle = 0, hjust = 1)
  )+
  # # Add labels for significant points
  geom_text(data = significant_df, aes(label = Gene),
            vjust = 1, hjust = 1, size = 3, check_overlap = TRUE, color = "black")
# Add labels for significant points with ggrepel
# geom_text_repel(data = significant_df, aes(label = Gene),
#                 size = 3, color = "black", box.padding = 0.5, max.overlaps = 20)

volcano_plot
pdf(fs::path(dir.results,"Plot_EC_PTC_T2D_LC_NoMed_lmm_no_zi.pdf"),width=10,height=7)
print(volcano_plot)
dev.off()

```
#### v. EC-AEA
```{r}
so_celltype <- subset(so_kpmp_sc,KPMP_celltype2=="EC-AEA")
# so_celltype <- subset(so_kpmp_sc,celltype1=="PT")
ncol(so_celltype) #37948 cells
nrow(so_celltype) #14627 genes
# Extract the gene expression data for all genes
gene_expression <- as.data.frame(GetAssayData(so_celltype, layer = "data"))

#Assign gene names as rownames, cell names as colnames
# rownames(gene_expression) <- rownames(so_celltype) #Gene Names
# colnames(gene_expression) <- colnames(so_celltype) #Cell Names
#Transpose gene expression dataset to merge with metadata
gene_expression <- t(gene_expression)
gene_expression <- data.frame(gene_expression) #Make a dataframe again after transposing
#Set gene list
gene_list_total <- colnames(gene_expression)
gene_expression$cellname <- rownames(gene_expression) 
rownames(gene_expression) <- NULL

# Extract the metadata
metadata <- so_celltype@meta.data
# metadata <- metadata %>%
#   mutate(across(everything(),~ifelse(.==".",NA,.)))
metadata$cellname <- rownames(metadata)
rownames(metadata) <- NULL

# Combine the gene expression data and metadata
data <- tidylog::full_join(metadata,gene_expression,by="cellname")
rm(metadata,gene_expression)

#Compare Health Controls to Type 2 Diabetes on no meds
data_subset <- data %>%
  filter(group=="Type_2_Diabetes" | group=="Lean_Control") %>%
  filter(medication=="no_med")
rm(data)
data_subset$group <- factor(data_subset$group)
data_subset$group <- relevel(data_subset$group,ref="Lean_Control")
data_subset <- data_subset %>% 
  dplyr::select(c("kit_id","group","age","sex","bmi",all_of(gene_list_total)))

# library(glmmTMB)
# library(lme4)
# library(doParallel)
# library(foreach)
# Set batch size
batch_size <- 2000
total_cores <- 50

# Deduplicate gene list to avoid double-processing
gene_list_total <- unique(gene_list_total)

# Split the gene list into batches
gene_batches <- split(gene_list_total, ceiling(seq_along(gene_list_total) / batch_size))

# Define the function that processes a single gene
process_gene <- function(gene) {
  if (sum(data_subset[[gene]]) > 0) {
    m1 <- as.formula(paste0(gene, " ~ group + age + sex + bmi + (1 | kit_id)"))
    
    model1 <- tryCatch({
      glmmTMB(m1, data = data_subset, family = gaussian)
    }, error = function(e) {
      return(NULL)
    })
    
    if (!is.null(model1)) {
      Beta <- summary(model1)$coef$cond[2,1]
      PValue <- summary(model1)$coef$cond[2,4]
    } else {
      Beta <- NA
      PValue <- NA
    }
  } else {
    Beta <- NA
    PValue <- NA
  }
  
  return(data.frame(Gene = gene, Beta = Beta, PValue = PValue))
}

# Set number of cores
total_cores <- parallel::detectCores() - 1

# Initialize and register cluster
cl <- makeCluster(total_cores)
registerDoParallel(cl)

# Run analysis per batch
all_results <- lapply(seq_along(gene_batches), function(batch_idx) {
  batch_genes <- gene_batches[[batch_idx]]
  
  batch_results <- foreach(
    gene = batch_genes,
    .combine = rbind,
    .packages = c("glmmTMB", "lme4"),
    .export = c("process_gene", "data_subset")
  ) %dopar% {
    process_gene(gene)
  }
  
  cat("Processed batch", batch_idx, "with", length(batch_genes), "genes\n")
  return(batch_results)
})

# Stop cluster
stopCluster(cl)

# Combine all batch results into a single data frame
final_results <- do.call(rbind, all_results)

# #Make volcano plot of all gene results for group
full_results <- final_results %>%
  mutate(fdr=p.adjust(PValue,method="fdr"))  
# mutate(fdr3=p.adjust(PValue3,method="fdr"))
full_results$PValue10 <- -log10(pmax(full_results$PValue, 1e-10))  # Avoid log(0)
# full_results$PValue10_3 <- -log10(pmax(full_results$PValue3, 1e-10))  # Avoid log(0)
Nonconvergence_Rate <- paste0(round(((length(which(is.na(full_results$PValue)))+length(which(full_results$PValue=="NaN")))/length(full_results$Gene))*100,0),"%")
# view(full_results)

write.csv(full_results,fs::path(dir.results,"EC_AEA_Cells_No_Med_T2D_LC_adj_cov_lmm_no_zi.csv"))
# full_results <- read.csv(fs::path(dir.results,"Hep_5_Fibrosis_Steatosis.csv"))

full_results$color <- ifelse(full_results$fdr < 0.05 & full_results$Beta > 0, "lightcoral",
                             ifelse(full_results$fdr < 0.05 & full_results$Beta < 0, "lightblue", "gray"))

# Identify significant points (fdr < 0.05)
significant_df <- full_results[full_results$fdr < 0.05, ]
# 
# full_results$color3 <- ifelse(full_results$fdr3 < 0.2 & full_results$Beta3 > 0, "lightcoral",
#                               ifelse(full_results$fdr3 < 0.2 & full_results$Beta3 < 0, "lightblue", "gray"))
# 
# # Identify significant points (fdr < 0.05)
# significant_df3 <- full_results[full_results$fdr3 < 0.2, ]

Genes <- length(unique(full_results$Gene))
Cells <- ncol(so_celltype)

#Figure Range
# max <- max(significant_df$Beta,na.rm=T)
# min <- min(significant_df$Beta,na.rm=T)

max <- max(full_results$Beta,na.rm=T)
min <- min(full_results$Beta,na.rm=T)

# Create the volcano plot using ggplot
volcano_plot <- ggplot(full_results, aes(x = Beta, y = PValue10, color = color)) +
  geom_point(alpha = 0.7) +  # Plot points with transparency
  scale_color_identity() +  # Use the color column directly
  theme_minimal() +  # Minimal theme
  labs(
    title = "Type 2 Diabetes vs. Lean Controls (All No Medication)",
    subtitle = "EC-AEA, Adjusted for Age, Sex & BMI",
    x = "Fold Change",
    y = "-log10(P-Value)",
    color = "FC Direction Direction",
    caption = paste0("FDR < 0.05, Genes = ",Genes,", Cells = ",Cells,", Non-Convergence Rate: ",Nonconvergence_Rate)
  ) +
  xlim(min,max)+
  theme(
    plot.title = element_text(hjust = 0),
    axis.text.x = element_text(angle = 0, hjust = 1)
  )+
  # # Add labels for significant points
  geom_text(data = significant_df, aes(label = Gene),
            vjust = 1, hjust = 1, size = 3, check_overlap = TRUE, color = "black")
# Add labels for significant points with ggrepel
# geom_text_repel(data = significant_df, aes(label = Gene),
#                 size = 3, color = "black", box.padding = 0.5, max.overlaps = 20)

volcano_plot
pdf(fs::path(dir.results,"Plot_EC_AEA_T2D_LC_NoMed_lmm_no_zi.pdf"),width=10,height=7)
print(volcano_plot)
dev.off()

```

### c. TAL
```{r}
#TAL
so_celltype <- subset(so_kpmp_sc,celltype1=="TAL")
ncol(so_celltype) #37948 cells
nrow(so_celltype) #14627 genes
# Extract the gene expression data for all genes
gene_expression <- as.data.frame(GetAssayData(so_celltype, layer = "data"))

#Assign gene names as rownames, cell names as colnames
# rownames(gene_expression) <- rownames(so_celltype) #Gene Names
# colnames(gene_expression) <- colnames(so_celltype) #Cell Names
#Transpose gene expression dataset to merge with metadata
gene_expression <- t(gene_expression)
gene_expression <- data.frame(gene_expression) #Make a dataframe again after transposing
#Set gene list
gene_list_total <- colnames(gene_expression)
gene_expression$cellname <- rownames(gene_expression) 
rownames(gene_expression) <- NULL

# Extract the metadata
metadata <- so_celltype@meta.data
# metadata <- metadata %>%
#   mutate(across(everything(),~ifelse(.==".",NA,.)))
metadata$cellname <- rownames(metadata)
rownames(metadata) <- NULL

# Combine the gene expression data and metadata
data <- tidylog::full_join(metadata,gene_expression,by="cellname")
rm(metadata,gene_expression)

#Compare Health Controls to Type 2 Diabetes on no meds
data_subset <- data %>%
  filter(group=="Type_2_Diabetes" | group=="Lean_Control") %>%
  filter(medication=="no_med")
rm(data)
data_subset$group <- factor(data_subset$group)
data_subset$group <- relevel(data_subset$group,ref="Lean_Control")
data_subset <- data_subset %>% 
  dplyr::select(c("kit_id","group","age","sex","bmi",all_of(gene_list_total)))

#Try as batch loop
# Set batch size
batch_size <- 2000
total_cores <- 50

# Deduplicate gene list to avoid double-processing
gene_list_total <- unique(gene_list_total)

# Split the gene list into batches
gene_batches <- split(gene_list_total, ceiling(seq_along(gene_list_total) / batch_size))

# Define the function that processes a single gene
process_gene <- function(gene) {
  if (sum(data_subset[[gene]]) > 0) {
    m1 <- as.formula(paste0(gene, " ~ group + age + sex + bmi + (1 | kit_id)"))
    
    model1 <- tryCatch({
      glmmTMB(m1, data = data_subset, family = gaussian)
    }, error = function(e) {
      return(NULL)
    })
    
    if (!is.null(model1)) {
      Beta <- summary(model1)$coef$cond[2,1]
      PValue <- summary(model1)$coef$cond[2,4]
    } else {
      Beta <- NA
      PValue <- NA
    }
  } else {
    Beta <- NA
    PValue <- NA
  }
  
  return(data.frame(Gene = gene, Beta = Beta, PValue = PValue))
}

# Set number of cores
total_cores <- parallel::detectCores() - 1

# Initialize and register cluster
cl <- makeCluster(total_cores)
registerDoParallel(cl)

# Run analysis per batch
all_results <- lapply(seq_along(gene_batches), function(batch_idx) {
  batch_genes <- gene_batches[[batch_idx]]
  
  batch_results <- foreach(
    gene = batch_genes,
    .combine = rbind,
    .packages = c("glmmTMB", "lme4"),
    .export = c("process_gene", "data_subset")
  ) %dopar% {
    process_gene(gene)
  }
  
  cat("Processed batch", batch_idx, "with", length(batch_genes), "genes\n")
  return(batch_results)
})

# Stop cluster
stopCluster(cl)

# Combine all batch results into a single data frame
final_results <- do.call(rbind, all_results)

# #Make volcano plot of all gene results for group
full_results <- final_results %>%
  mutate(fdr=p.adjust(PValue,method="fdr"))  
# mutate(fdr3=p.adjust(PValue3,method="fdr"))
full_results$PValue10 <- -log10(pmax(full_results$PValue, 1e-10))  # Avoid log(0)
# full_results$PValue10_3 <- -log10(pmax(full_results$PValue3, 1e-10))  # Avoid log(0)
Nonconvergence_Rate <- paste0(round(((length(which(is.na(full_results$PValue)))+length(which(full_results$PValue=="NaN")))/length(full_results$Gene))*100,0),"%")
# view(full_results)

# write.csv(full_results,"test.csv")

write.csv(full_results,fs::path(dir.results,"TAL_Cells_No_Med_T2D_LC_adj_cov_lmm_no_zi.csv"))
# full_results <- read.csv(fs::path(dir.results,"Hep_5_Fibrosis_Steatosis.csv"))

full_results$color <- ifelse(full_results$fdr < 0.05 & full_results$Beta > 0, "lightcoral",
                             ifelse(full_results$fdr < 0.05 & full_results$Beta < 0, "lightblue", "gray"))

# Identify significant points (fdr < 0.05)
significant_df <- full_results[full_results$fdr < 0.05, ]
# 
# full_results$color3 <- ifelse(full_results$fdr3 < 0.2 & full_results$Beta3 > 0, "lightcoral",
#                               ifelse(full_results$fdr3 < 0.2 & full_results$Beta3 < 0, "lightblue", "gray"))
# 
# # Identify significant points (fdr < 0.05)
# significant_df3 <- full_results[full_results$fdr3 < 0.2, ]

Genes <- length(unique(full_results$Gene))
Cells <- ncol(so_celltype)

#Figure Range
# max <- max(significant_df$Beta,na.rm=T)
# min <- min(significant_df$Beta,na.rm=T)
max <- max(full_results$Beta,na.rm=T)
min <- min(full_results$Beta,na.rm=T)

# full_results$Log2FC <- 2^(full_results$Beta)
# Create the volcano plot using ggplot
volcano_plot <- ggplot(full_results, aes(x = Beta, y = PValue10, color = color)) +
  geom_point(alpha = 0.7) +  # Plot points with transparency
  scale_color_identity() +  # Use the color column directly
  theme_minimal() +  # Minimal theme
  labs(
    title = "Type 2 Diabetes vs. Lean Controls (All No Medication)",
    subtitle = "TAL Cells, Adjusted for Age, Sex & BMI",
    x = "Fold Change",
    y = "-log10(P-Value)",
    color = "FC Direction Direction",
    caption = paste0("FDR < 0.05, Genes = ",Genes,", Cells = ",Cells,", Non-Convergence Rate: ",Nonconvergence_Rate)
  ) +
  xlim(min,max)+
  theme(
    plot.title = element_text(hjust = 0),
    axis.text.x = element_text(angle = 0, hjust = 1)
  )+
  # # Add labels for significant points
  geom_text(data = significant_df, aes(label = Gene),
            vjust = 1, hjust = 1, size = 3, check_overlap = TRUE, color = "black")
# Add labels for significant points with ggrepel
# geom_text_repel(data = significant_df, aes(label = Gene),
#                 size = 3, color = "black", box.padding = 0.5, max.overlaps = 20)

volcano_plot
pdf(fs::path(dir.results,"Plot_TAL_Cells_T2D_LC_NoMed_lmm_no_zi.pdf"),width=10,height=7)
print(volcano_plot)
dev.off()

```
#### i. C-TAL-1 
```{r}
so_celltype <- subset(so_kpmp_sc,KPMP_celltype2=="C-TAL-1")
# so_celltype <- subset(so_kpmp_sc,celltype1=="PT")
ncol(so_celltype) #37948 cells
nrow(so_celltype) #14627 genes
# Extract the gene expression data for all genes
gene_expression <- as.data.frame(GetAssayData(so_celltype, layer = "data"))

#Assign gene names as rownames, cell names as colnames
# rownames(gene_expression) <- rownames(so_celltype) #Gene Names
# colnames(gene_expression) <- colnames(so_celltype) #Cell Names
#Transpose gene expression dataset to merge with metadata
gene_expression <- t(gene_expression)
gene_expression <- data.frame(gene_expression) #Make a dataframe again after transposing
#Set gene list
gene_list_total <- colnames(gene_expression)
gene_expression$cellname <- rownames(gene_expression) 
rownames(gene_expression) <- NULL

# Extract the metadata
metadata <- so_celltype@meta.data
# metadata <- metadata %>%
#   mutate(across(everything(),~ifelse(.==".",NA,.)))
metadata$cellname <- rownames(metadata)
rownames(metadata) <- NULL

# Combine the gene expression data and metadata
data <- tidylog::full_join(metadata,gene_expression,by="cellname")
rm(metadata,gene_expression)

#Compare Health Controls to Type 2 Diabetes on no meds
data_subset <- data %>%
  filter(group=="Type_2_Diabetes" | group=="Lean_Control") %>%
  filter(medication=="no_med")
rm(data)
data_subset$group <- factor(data_subset$group)
data_subset$group <- relevel(data_subset$group,ref="Lean_Control")
data_subset <- data_subset %>% 
  dplyr::select(c("kit_id","group","age","sex","bmi",all_of(gene_list_total)))

# library(glmmTMB)
# library(lme4)
# library(doParallel)
# library(foreach)
# Set batch size
batch_size <- 2000
total_cores <- 50

# Deduplicate gene list to avoid double-processing
gene_list_total <- unique(gene_list_total)

# Split the gene list into batches
gene_batches <- split(gene_list_total, ceiling(seq_along(gene_list_total) / batch_size))

# Define the function that processes a single gene
process_gene <- function(gene) {
  if (sum(data_subset[[gene]]) > 0) {
    m1 <- as.formula(paste0(gene, " ~ group + age + sex + bmi + (1 | kit_id)"))
    
    model1 <- tryCatch({
      glmmTMB(m1, data = data_subset, family = gaussian)
    }, error = function(e) {
      return(NULL)
    })
    
    if (!is.null(model1)) {
      Beta <- summary(model1)$coef$cond[2,1]
      PValue <- summary(model1)$coef$cond[2,4]
    } else {
      Beta <- NA
      PValue <- NA
    }
  } else {
    Beta <- NA
    PValue <- NA
  }
  
  return(data.frame(Gene = gene, Beta = Beta, PValue = PValue))
}

# Set number of cores
total_cores <- parallel::detectCores() - 1

# Initialize and register cluster
cl <- makeCluster(total_cores)
registerDoParallel(cl)

# Run analysis per batch
all_results <- lapply(seq_along(gene_batches), function(batch_idx) {
  batch_genes <- gene_batches[[batch_idx]]
  
  batch_results <- foreach(
    gene = batch_genes,
    .combine = rbind,
    .packages = c("glmmTMB", "lme4"),
    .export = c("process_gene", "data_subset")
  ) %dopar% {
    process_gene(gene)
  }
  
  cat("Processed batch", batch_idx, "with", length(batch_genes), "genes\n")
  return(batch_results)
})

# Stop cluster
stopCluster(cl)

# Combine all batch results into a single data frame
final_results <- do.call(rbind, all_results)

# #Make volcano plot of all gene results for group
full_results <- final_results %>%
  mutate(fdr=p.adjust(PValue,method="fdr"))  
# mutate(fdr3=p.adjust(PValue3,method="fdr"))
full_results$PValue10 <- -log10(pmax(full_results$PValue, 1e-10))  # Avoid log(0)
# full_results$PValue10_3 <- -log10(pmax(full_results$PValue3, 1e-10))  # Avoid log(0)
Nonconvergence_Rate <- paste0(round(((length(which(is.na(full_results$PValue)))+length(which(full_results$PValue=="NaN")))/length(full_results$Gene))*100,0),"%")
# view(full_results)

write.csv(full_results,fs::path(dir.results,"C_TAL_1_Cells_No_Med_T2D_LC_adj_cov_lmm_no_zi.csv"))
# full_results <- read.csv(fs::path(dir.results,"Hep_5_Fibrosis_Steatosis.csv"))

full_results$color <- ifelse(full_results$fdr < 0.05 & full_results$Beta > 0, "lightcoral",
                             ifelse(full_results$fdr < 0.05 & full_results$Beta < 0, "lightblue", "gray"))

# Identify significant points (fdr < 0.05)
significant_df <- full_results[full_results$fdr < 0.05, ]
# 
# full_results$color3 <- ifelse(full_results$fdr3 < 0.2 & full_results$Beta3 > 0, "lightcoral",
#                               ifelse(full_results$fdr3 < 0.2 & full_results$Beta3 < 0, "lightblue", "gray"))
# 
# # Identify significant points (fdr < 0.05)
# significant_df3 <- full_results[full_results$fdr3 < 0.2, ]

Genes <- length(unique(full_results$Gene))
Cells <- ncol(so_celltype)

#Figure Range
# max <- max(significant_df$Beta,na.rm=T)
# min <- min(significant_df$Beta,na.rm=T)

max <- max(full_results$Beta,na.rm=T)
min <- min(full_results$Beta,na.rm=T)

# Create the volcano plot using ggplot
volcano_plot <- ggplot(full_results, aes(x = Beta, y = PValue10, color = color)) +
  geom_point(alpha = 0.7) +  # Plot points with transparency
  scale_color_identity() +  # Use the color column directly
  theme_minimal() +  # Minimal theme
  labs(
    title = "Type 2 Diabetes vs. Lean Controls (All No Medication)",
    subtitle = "C-TAL-1, Adjusted for Age, Sex & BMI",
    x = "Fold Change",
    y = "-log10(P-Value)",
    color = "FC Direction Direction",
    caption = paste0("FDR < 0.05, Genes = ",Genes,", Cells = ",Cells,", Non-Convergence Rate: ",Nonconvergence_Rate)
  ) +
  xlim(min,max)+
  theme(
    plot.title = element_text(hjust = 0),
    axis.text.x = element_text(angle = 0, hjust = 1)
  )+
  # # Add labels for significant points
  geom_text(data = significant_df, aes(label = Gene),
            vjust = 1, hjust = 1, size = 3, check_overlap = TRUE, color = "black")
# Add labels for significant points with ggrepel
# geom_text_repel(data = significant_df, aes(label = Gene),
#                 size = 3, color = "black", box.padding = 0.5, max.overlaps = 20)

volcano_plot
pdf(fs::path(dir.results,"Plot_C_TAL_1_T2D_LC_NoMed_lmm_no_zi.pdf"),width=10,height=7)
print(volcano_plot)
dev.off()

```
#### ii. C-TAL-2
```{r}
so_celltype <- subset(so_kpmp_sc,KPMP_celltype2=="C-TAL-2")
# so_celltype <- subset(so_kpmp_sc,celltype1=="PT")
ncol(so_celltype) #37948 cells
nrow(so_celltype) #14627 genes
# Extract the gene expression data for all genes
gene_expression <- as.data.frame(GetAssayData(so_celltype, layer = "data"))

#Assign gene names as rownames, cell names as colnames
# rownames(gene_expression) <- rownames(so_celltype) #Gene Names
# colnames(gene_expression) <- colnames(so_celltype) #Cell Names
#Transpose gene expression dataset to merge with metadata
gene_expression <- t(gene_expression)
gene_expression <- data.frame(gene_expression) #Make a dataframe again after transposing
#Set gene list
gene_list_total <- colnames(gene_expression)
gene_expression$cellname <- rownames(gene_expression) 
rownames(gene_expression) <- NULL

# Extract the metadata
metadata <- so_celltype@meta.data
# metadata <- metadata %>%
#   mutate(across(everything(),~ifelse(.==".",NA,.)))
metadata$cellname <- rownames(metadata)
rownames(metadata) <- NULL

# Combine the gene expression data and metadata
data <- tidylog::full_join(metadata,gene_expression,by="cellname")
rm(metadata,gene_expression)

#Compare Health Controls to Type 2 Diabetes on no meds
data_subset <- data %>%
  filter(group=="Type_2_Diabetes" | group=="Lean_Control") %>%
  filter(medication=="no_med")
rm(data)
data_subset$group <- factor(data_subset$group)
data_subset$group <- relevel(data_subset$group,ref="Lean_Control")
data_subset <- data_subset %>% 
  dplyr::select(c("kit_id","group","age","sex","bmi",all_of(gene_list_total)))

# library(glmmTMB)
# library(lme4)
# library(doParallel)
# library(foreach)
# Set batch size
batch_size <- 2000
total_cores <- 50

# Deduplicate gene list to avoid double-processing
gene_list_total <- unique(gene_list_total)

# Split the gene list into batches
gene_batches <- split(gene_list_total, ceiling(seq_along(gene_list_total) / batch_size))

# Define the function that processes a single gene
process_gene <- function(gene) {
  if (sum(data_subset[[gene]]) > 0) {
    m1 <- as.formula(paste0(gene, " ~ group + age + sex + bmi + (1 | kit_id)"))
    
    model1 <- tryCatch({
      glmmTMB(m1, data = data_subset, family = gaussian)
    }, error = function(e) {
      return(NULL)
    })
    
    if (!is.null(model1)) {
      Beta <- summary(model1)$coef$cond[2,1]
      PValue <- summary(model1)$coef$cond[2,4]
    } else {
      Beta <- NA
      PValue <- NA
    }
  } else {
    Beta <- NA
    PValue <- NA
  }
  
  return(data.frame(Gene = gene, Beta = Beta, PValue = PValue))
}

# Set number of cores
total_cores <- parallel::detectCores() - 1

# Initialize and register cluster
cl <- makeCluster(total_cores)
registerDoParallel(cl)

# Run analysis per batch
all_results <- lapply(seq_along(gene_batches), function(batch_idx) {
  batch_genes <- gene_batches[[batch_idx]]
  
  batch_results <- foreach(
    gene = batch_genes,
    .combine = rbind,
    .packages = c("glmmTMB", "lme4"),
    .export = c("process_gene", "data_subset")
  ) %dopar% {
    process_gene(gene)
  }
  
  cat("Processed batch", batch_idx, "with", length(batch_genes), "genes\n")
  return(batch_results)
})

# Stop cluster
stopCluster(cl)

# Combine all batch results into a single data frame
final_results <- do.call(rbind, all_results)

# #Make volcano plot of all gene results for group
full_results <- final_results %>%
  mutate(fdr=p.adjust(PValue,method="fdr"))  
# mutate(fdr3=p.adjust(PValue3,method="fdr"))
full_results$PValue10 <- -log10(pmax(full_results$PValue, 1e-10))  # Avoid log(0)
# full_results$PValue10_3 <- -log10(pmax(full_results$PValue3, 1e-10))  # Avoid log(0)
Nonconvergence_Rate <- paste0(round(((length(which(is.na(full_results$PValue)))+length(which(full_results$PValue=="NaN")))/length(full_results$Gene))*100,0),"%")
# view(full_results)

write.csv(full_results,fs::path(dir.results,"C_TAL_2_Cells_No_Med_T2D_LC_adj_cov_lmm_no_zi.csv"))
# full_results <- read.csv(fs::path(dir.results,"Hep_5_Fibrosis_Steatosis.csv"))

full_results$color <- ifelse(full_results$fdr < 0.05 & full_results$Beta > 0, "lightcoral",
                             ifelse(full_results$fdr < 0.05 & full_results$Beta < 0, "lightblue", "gray"))

# Identify significant points (fdr < 0.05)
significant_df <- full_results[full_results$fdr < 0.05, ]
# 
# full_results$color3 <- ifelse(full_results$fdr3 < 0.2 & full_results$Beta3 > 0, "lightcoral",
#                               ifelse(full_results$fdr3 < 0.2 & full_results$Beta3 < 0, "lightblue", "gray"))
# 
# # Identify significant points (fdr < 0.05)
# significant_df3 <- full_results[full_results$fdr3 < 0.2, ]

Genes <- length(unique(full_results$Gene))
Cells <- ncol(so_celltype)

#Figure Range
# max <- max(significant_df$Beta,na.rm=T)
# min <- min(significant_df$Beta,na.rm=T)

max <- max(full_results$Beta,na.rm=T)
min <- min(full_results$Beta,na.rm=T)

# Create the volcano plot using ggplot
volcano_plot <- ggplot(full_results, aes(x = Beta, y = PValue10, color = color)) +
  geom_point(alpha = 0.7) +  # Plot points with transparency
  scale_color_identity() +  # Use the color column directly
  theme_minimal() +  # Minimal theme
  labs(
    title = "Type 2 Diabetes vs. Lean Controls (All No Medication)",
    subtitle = "C-TAL-2, Adjusted for Age, Sex & BMI",
    x = "Fold Change",
    y = "-log10(P-Value)",
    color = "FC Direction Direction",
    caption = paste0("FDR < 0.05, Genes = ",Genes,", Cells = ",Cells,", Non-Convergence Rate: ",Nonconvergence_Rate)
  ) +
  xlim(min,max)+
  theme(
    plot.title = element_text(hjust = 0),
    axis.text.x = element_text(angle = 0, hjust = 1)
  )+
  # # Add labels for significant points
  geom_text(data = significant_df, aes(label = Gene),
            vjust = 1, hjust = 1, size = 3, check_overlap = TRUE, color = "black")
# Add labels for significant points with ggrepel
# geom_text_repel(data = significant_df, aes(label = Gene),
#                 size = 3, color = "black", box.padding = 0.5, max.overlaps = 20)

volcano_plot
pdf(fs::path(dir.results,"Plot_C_TAL_2_T2D_LC_NoMed_lmm_no_zi.pdf"),width=10,height=7)
print(volcano_plot)
dev.off()

```
#### iii. dTAL
```{r}
so_celltype <- subset(so_kpmp_sc,KPMP_celltype2=="dTAL")
# so_celltype <- subset(so_kpmp_sc,celltype1=="PT")
ncol(so_celltype) #37948 cells
nrow(so_celltype) #14627 genes
# Extract the gene expression data for all genes
gene_expression <- as.data.frame(GetAssayData(so_celltype, layer = "data"))

#Assign gene names as rownames, cell names as colnames
# rownames(gene_expression) <- rownames(so_celltype) #Gene Names
# colnames(gene_expression) <- colnames(so_celltype) #Cell Names
#Transpose gene expression dataset to merge with metadata
gene_expression <- t(gene_expression)
gene_expression <- data.frame(gene_expression) #Make a dataframe again after transposing
#Set gene list
gene_list_total <- colnames(gene_expression)
gene_expression$cellname <- rownames(gene_expression) 
rownames(gene_expression) <- NULL

# Extract the metadata
metadata <- so_celltype@meta.data
# metadata <- metadata %>%
#   mutate(across(everything(),~ifelse(.==".",NA,.)))
metadata$cellname <- rownames(metadata)
rownames(metadata) <- NULL

# Combine the gene expression data and metadata
data <- tidylog::full_join(metadata,gene_expression,by="cellname")
rm(metadata,gene_expression)

#Compare Health Controls to Type 2 Diabetes on no meds
data_subset <- data %>%
  filter(group=="Type_2_Diabetes" | group=="Lean_Control") %>%
  filter(medication=="no_med")
rm(data)
data_subset$group <- factor(data_subset$group)
data_subset$group <- relevel(data_subset$group,ref="Lean_Control")
data_subset <- data_subset %>% 
  dplyr::select(c("kit_id","group","age","sex","bmi",all_of(gene_list_total)))

# library(glmmTMB)
# library(lme4)
# library(doParallel)
# library(foreach)
# Set batch size
batch_size <- 2000
total_cores <- 50

# Deduplicate gene list to avoid double-processing
gene_list_total <- unique(gene_list_total)

# Split the gene list into batches
gene_batches <- split(gene_list_total, ceiling(seq_along(gene_list_total) / batch_size))

# Define the function that processes a single gene
process_gene <- function(gene) {
  if (sum(data_subset[[gene]]) > 0) {
    m1 <- as.formula(paste0(gene, " ~ group + age + sex + bmi + (1 | kit_id)"))
    
    model1 <- tryCatch({
      glmmTMB(m1, data = data_subset, family = gaussian)
    }, error = function(e) {
      return(NULL)
    })
    
    if (!is.null(model1)) {
      Beta <- summary(model1)$coef$cond[2,1]
      PValue <- summary(model1)$coef$cond[2,4]
    } else {
      Beta <- NA
      PValue <- NA
    }
  } else {
    Beta <- NA
    PValue <- NA
  }
  
  return(data.frame(Gene = gene, Beta = Beta, PValue = PValue))
}

# Set number of cores
total_cores <- parallel::detectCores() - 1

# Initialize and register cluster
cl <- makeCluster(total_cores)
registerDoParallel(cl)

# Run analysis per batch
all_results <- lapply(seq_along(gene_batches), function(batch_idx) {
  batch_genes <- gene_batches[[batch_idx]]
  
  batch_results <- foreach(
    gene = batch_genes,
    .combine = rbind,
    .packages = c("glmmTMB", "lme4"),
    .export = c("process_gene", "data_subset")
  ) %dopar% {
    process_gene(gene)
  }
  
  cat("Processed batch", batch_idx, "with", length(batch_genes), "genes\n")
  return(batch_results)
})

# Stop cluster
stopCluster(cl)

# Combine all batch results into a single data frame
final_results <- do.call(rbind, all_results)

# #Make volcano plot of all gene results for group
full_results <- final_results %>%
  mutate(fdr=p.adjust(PValue,method="fdr"))  
# mutate(fdr3=p.adjust(PValue3,method="fdr"))
full_results$PValue10 <- -log10(pmax(full_results$PValue, 1e-10))  # Avoid log(0)
# full_results$PValue10_3 <- -log10(pmax(full_results$PValue3, 1e-10))  # Avoid log(0)
Nonconvergence_Rate <- paste0(round(((length(which(is.na(full_results$PValue)))+length(which(full_results$PValue=="NaN")))/length(full_results$Gene))*100,0),"%")
# view(full_results)

write.csv(full_results,fs::path(dir.results,"dTAL_Cells_No_Med_T2D_LC_adj_cov_lmm_no_zi.csv"))
# full_results <- read.csv(fs::path(dir.results,"Hep_5_Fibrosis_Steatosis.csv"))

full_results$color <- ifelse(full_results$fdr < 0.05 & full_results$Beta > 0, "lightcoral",
                             ifelse(full_results$fdr < 0.05 & full_results$Beta < 0, "lightblue", "gray"))

# Identify significant points (fdr < 0.05)
significant_df <- full_results[full_results$fdr < 0.05, ]
# 
# full_results$color3 <- ifelse(full_results$fdr3 < 0.2 & full_results$Beta3 > 0, "lightcoral",
#                               ifelse(full_results$fdr3 < 0.2 & full_results$Beta3 < 0, "lightblue", "gray"))
# 
# # Identify significant points (fdr < 0.05)
# significant_df3 <- full_results[full_results$fdr3 < 0.2, ]

Genes <- length(unique(full_results$Gene))
Cells <- ncol(so_celltype)

#Figure Range
# max <- max(significant_df$Beta,na.rm=T)
# min <- min(significant_df$Beta,na.rm=T)

max <- max(full_results$Beta,na.rm=T)
min <- min(full_results$Beta,na.rm=T)

# Create the volcano plot using ggplot
volcano_plot <- ggplot(full_results, aes(x = Beta, y = PValue10, color = color)) +
  geom_point(alpha = 0.7) +  # Plot points with transparency
  scale_color_identity() +  # Use the color column directly
  theme_minimal() +  # Minimal theme
  labs(
    title = "Type 2 Diabetes vs. Lean Controls (All No Medication)",
    subtitle = "dTAL, Adjusted for Age, Sex & BMI",
    x = "Fold Change",
    y = "-log10(P-Value)",
    color = "FC Direction Direction",
    caption = paste0("FDR < 0.05, Genes = ",Genes,", Cells = ",Cells,", Non-Convergence Rate: ",Nonconvergence_Rate)
  ) +
  xlim(min,max)+
  theme(
    plot.title = element_text(hjust = 0),
    axis.text.x = element_text(angle = 0, hjust = 1)
  )+
  # # Add labels for significant points
  geom_text(data = significant_df, aes(label = Gene),
            vjust = 1, hjust = 1, size = 3, check_overlap = TRUE, color = "black")
# Add labels for significant points with ggrepel
# geom_text_repel(data = significant_df, aes(label = Gene),
#                 size = 3, color = "black", box.padding = 0.5, max.overlaps = 20)

volcano_plot
pdf(fs::path(dir.results,"Plot_dTAL_T2D_LC_NoMed_lmm_no_zi.pdf"),width=10,height=7)
print(volcano_plot)
dev.off()

```
#### iv. aTAL
```{r}
so_celltype <- subset(so_kpmp_sc,KPMP_celltype2=="aTAL")
# so_celltype <- subset(so_kpmp_sc,celltype1=="PT")
ncol(so_celltype) #37948 cells
nrow(so_celltype) #14627 genes
# Extract the gene expression data for all genes
gene_expression <- as.data.frame(GetAssayData(so_celltype, layer = "data"))

#Assign gene names as rownames, cell names as colnames
# rownames(gene_expression) <- rownames(so_celltype) #Gene Names
# colnames(gene_expression) <- colnames(so_celltype) #Cell Names
#Transpose gene expression dataset to merge with metadata
gene_expression <- t(gene_expression)
gene_expression <- data.frame(gene_expression) #Make a dataframe again after transposing
#Set gene list
gene_list_total <- colnames(gene_expression)
gene_expression$cellname <- rownames(gene_expression) 
rownames(gene_expression) <- NULL

# Extract the metadata
metadata <- so_celltype@meta.data
# metadata <- metadata %>%
#   mutate(across(everything(),~ifelse(.==".",NA,.)))
metadata$cellname <- rownames(metadata)
rownames(metadata) <- NULL

# Combine the gene expression data and metadata
data <- tidylog::full_join(metadata,gene_expression,by="cellname")
rm(metadata,gene_expression)

#Compare Health Controls to Type 2 Diabetes on no meds
data_subset <- data %>%
  filter(group=="Type_2_Diabetes" | group=="Lean_Control") %>%
  filter(medication=="no_med")
rm(data)
data_subset$group <- factor(data_subset$group)
data_subset$group <- relevel(data_subset$group,ref="Lean_Control")
data_subset <- data_subset %>% 
  dplyr::select(c("kit_id","group","age","sex","bmi",all_of(gene_list_total)))

# Parameters
batch_size <- 2000
# total_cores <- 50

# Deduplicate genes
gene_list_total <- unique(gene_list_total)

# Split into batches
gene_batches <- split(gene_list_total, ceiling(seq_along(gene_list_total) / batch_size))

# Function to process a single gene
process_gene <- function(gene, data_subset) {
  tryCatch({
    formula <- as.formula(paste0(gene, " ~ medication + age + sex + bmi + (1 | kit_id)"))
    model <- suppressWarnings(lmer(formula, data = data_subset))
    
    emm <- emmeans(model, ~ medication, lmerTest.limit = 19037, pbkrtest.limit = 19037)
    
    contrast_results <- contrast(emm, list(
      "GLP1 vs No Med"        = c(0, 1, 0, -1),
      "SGT2 vs No Med"        = c(0, 0, 1, -1),
      "GLP1+SGT2 vs No Med"   = c(0, 0, 0, 1),
      "GLP1+SGT2 vs GLP1"     = c(0, -1, 1, 0),
      "GLP1+SGT2 vs SGT2"     = c(0, 0, -1, 1)
    ))
    
    contrast_df <- as.data.frame(confint(contrast_results))
    contrast_df$PValue <- summary(contrast_results)$p.value
    contrast_df$Gene <- gene
    
    return(contrast_df)
  }, error = function(e) {
    return(NULL)  # Skip genes that cause errors
  })
}

# Create a cluster
cl <- makeCluster(total_cores)

# Export necessary objects and packages to each worker
clusterExport(cl, varlist = c("data_subset", "process_gene"), envir = environment())
clusterEvalQ(cl, {
  library(lme4)
  library(emmeans)
})

# Flatten results from each batch
all_results <- list()

for (i in seq_along(gene_batches)) {
  cat("Processing batch", i, "of", length(gene_batches), "\n")
  
  # Apply function in parallel across the current batch
  batch_genes <- gene_batches[[i]]
  batch_results <- parLapply(cl, batch_genes, function(g) process_gene(g, data_subset))
  
  # Remove NULLs and bind into a single dataframe
  batch_df <- do.call(rbind, batch_results[!sapply(batch_results, is.null)])
  
  all_results[[i]] <- batch_df
}

# Stop the cluster
stopCluster(cl)

# Combine all batch results
final_results <- do.call(rbind, all_results)

# Preview
head(final_results)

# #Make volcano plot of all gene results for group
full_results <- final_results %>%
  mutate(fdr=p.adjust(PValue,method="fdr"))  
# mutate(fdr3=p.adjust(PValue3,method="fdr"))
full_results$PValue10 <- -log10(pmax(full_results$PValue, 1e-10))  # Avoid log(0)
# full_results$PValue10_3 <- -log10(pmax(full_results$PValue3, 1e-10))  # Avoid log(0)
Nonconvergence_Rate <- paste0(round(((length(which(is.na(full_results$PValue)))+length(which(full_results$PValue=="NaN")))/length(full_results$Gene))*100,0),"%")
# view(full_results)

write.csv(full_results,fs::path(dir.results,"aTAL_Cells_No_Med_T2D_LC_adj_cov_lmm_no_zi.csv"))
# full_results <- read.csv(fs::path(dir.results,"Hep_5_Fibrosis_Steatosis.csv"))

full_results$color <- ifelse(full_results$fdr < 0.05 & full_results$Beta > 0, "lightcoral",
                             ifelse(full_results$fdr < 0.05 & full_results$Beta < 0, "lightblue", "gray"))

# Identify significant points (fdr < 0.05)
significant_df <- full_results[full_results$fdr < 0.05, ]
# 
# full_results$color3 <- ifelse(full_results$fdr3 < 0.2 & full_results$Beta3 > 0, "lightcoral",
#                               ifelse(full_results$fdr3 < 0.2 & full_results$Beta3 < 0, "lightblue", "gray"))
# 
# # Identify significant points (fdr < 0.05)
# significant_df3 <- full_results[full_results$fdr3 < 0.2, ]

Genes <- length(unique(full_results$Gene))
Cells <- ncol(so_celltype)

#Figure Range
# max <- max(significant_df$Beta,na.rm=T)
# min <- min(significant_df$Beta,na.rm=T)

max <- max(full_results$Beta,na.rm=T)
min <- min(full_results$Beta,na.rm=T)

# Create the volcano plot using ggplot
volcano_plot <- ggplot(full_results, aes(x = Beta, y = PValue10, color = color)) +
  geom_point(alpha = 0.7) +  # Plot points with transparency
  scale_color_identity() +  # Use the color column directly
  theme_minimal() +  # Minimal theme
  labs(
    title = "Type 2 Diabetes vs. Lean Controls (All No Medication)",
    subtitle = "aTAL, Adjusted for Age, Sex & BMI",
    x = "Fold Change",
    y = "-log10(P-Value)",
    color = "FC Direction Direction",
    caption = paste0("FDR < 0.05, Genes = ",Genes,", Cells = ",Cells,", Non-Convergence Rate: ",Nonconvergence_Rate)
  ) +
  xlim(min,max)+
  theme(
    plot.title = element_text(hjust = 0),
    axis.text.x = element_text(angle = 0, hjust = 1)
  )+
  # # Add labels for significant points
  geom_text(data = significant_df, aes(label = Gene),
            vjust = 1, hjust = 1, size = 3, check_overlap = TRUE, color = "black")
# Add labels for significant points with ggrepel
# geom_text_repel(data = significant_df, aes(label = Gene),
#                 size = 3, color = "black", box.padding = 0.5, max.overlaps = 20)

volcano_plot
pdf(fs::path(dir.results,"Plot_aTAL_T2D_LC_NoMed_lmm_no_zi.pdf"),width=10,height=7)
print(volcano_plot)
dev.off()

# 
```

### d. Podocytes
```{r}
so_celltype <- subset(so_kpmp_sc,KPMP_celltype2=="POD")
# so_celltype <- subset(so_kpmp_sc,celltype1=="PT")
ncol(so_celltype) #37948 cells
nrow(so_celltype) #14627 genes
# Extract the gene expression data for all genes
gene_expression <- as.data.frame(GetAssayData(so_celltype, layer = "data"))

#Assign gene names as rownames, cell names as colnames
# rownames(gene_expression) <- rownames(so_celltype) #Gene Names
# colnames(gene_expression) <- colnames(so_celltype) #Cell Names
#Transpose gene expression dataset to merge with metadata
gene_expression <- t(gene_expression)
gene_expression <- data.frame(gene_expression) #Make a dataframe again after transposing
#Set gene list
gene_list_total <- colnames(gene_expression)
gene_expression$cellname <- rownames(gene_expression) 
rownames(gene_expression) <- NULL

# Extract the metadata
metadata <- so_celltype@meta.data
# metadata <- metadata %>%
#   mutate(across(everything(),~ifelse(.==".",NA,.)))
metadata$cellname <- rownames(metadata)
rownames(metadata) <- NULL

# Combine the gene expression data and metadata
data <- tidylog::full_join(metadata,gene_expression,by="cellname")
rm(metadata,gene_expression)

#Compare Health Controls to Type 2 Diabetes on no meds
data_subset <- data %>%
  filter(group=="Type_2_Diabetes" | group=="Lean_Control") %>%
  filter(medication=="no_med")
rm(data)
data_subset$group <- factor(data_subset$group)
data_subset$group <- relevel(data_subset$group,ref="Lean_Control")
data_subset <- data_subset %>% 
  dplyr::select(c("kit_id","group","age","sex","bmi",all_of(gene_list_total)))

# library(glmmTMB)
# library(lme4)
# library(doParallel)
# library(foreach)
# Set batch size
batch_size <- 2000
total_cores <- 50

# Deduplicate gene list to avoid double-processing
gene_list_total <- unique(gene_list_total)

# Split the gene list into batches
gene_batches <- split(gene_list_total, ceiling(seq_along(gene_list_total) / batch_size))

# Define the function that processes a single gene
process_gene <- function(gene) {
  if (sum(data_subset[[gene]]) > 0) {
    m1 <- as.formula(paste0(gene, " ~ group + age + sex + bmi + (1 | kit_id)"))
    
    model1 <- tryCatch({
      glmmTMB(m1, data = data_subset, family = gaussian)
    }, error = function(e) {
      return(NULL)
    })
    
    if (!is.null(model1)) {
      Beta <- summary(model1)$coef$cond[2,1]
      PValue <- summary(model1)$coef$cond[2,4]
    } else {
      Beta <- NA
      PValue <- NA
    }
  } else {
    Beta <- NA
    PValue <- NA
  }
  
  return(data.frame(Gene = gene, Beta = Beta, PValue = PValue))
}

# Set number of cores
total_cores <- parallel::detectCores() - 1

# Initialize and register cluster
cl <- makeCluster(total_cores)
registerDoParallel(cl)

# Run analysis per batch
all_results <- lapply(seq_along(gene_batches), function(batch_idx) {
  batch_genes <- gene_batches[[batch_idx]]
  
  batch_results <- foreach(
    gene = batch_genes,
    .combine = rbind,
    .packages = c("glmmTMB", "lme4"),
    .export = c("process_gene", "data_subset")
  ) %dopar% {
    process_gene(gene)
  }
  
  cat("Processed batch", batch_idx, "with", length(batch_genes), "genes\n")
  return(batch_results)
})

# Stop cluster
stopCluster(cl)

# Combine all batch results into a single data frame
final_results <- do.call(rbind, all_results)

# #Make volcano plot of all gene results for group
full_results <- final_results %>%
  mutate(fdr=p.adjust(PValue,method="fdr"))  
# mutate(fdr3=p.adjust(PValue3,method="fdr"))
full_results$PValue10 <- -log10(pmax(full_results$PValue, 1e-10))  # Avoid log(0)
# full_results$PValue10_3 <- -log10(pmax(full_results$PValue3, 1e-10))  # Avoid log(0)
Nonconvergence_Rate <- paste0(round(((length(which(is.na(full_results$PValue)))+length(which(full_results$PValue=="NaN")))/length(full_results$Gene))*100,0),"%")
# view(full_results)

write.csv(full_results,fs::path(dir.results,"POD_Cells_No_Med_T2D_LC_adj_cov_lmm_no_zi.csv"))
# full_results <- read.csv(fs::path(dir.results,"Hep_5_Fibrosis_Steatosis.csv"))

full_results$color <- ifelse(full_results$fdr < 0.05 & full_results$Beta > 0, "lightcoral",
                             ifelse(full_results$fdr < 0.05 & full_results$Beta < 0, "lightblue", "gray"))

# Identify significant points (fdr < 0.05)
significant_df <- full_results[full_results$fdr < 0.05, ]
# 
# full_results$color3 <- ifelse(full_results$fdr3 < 0.2 & full_results$Beta3 > 0, "lightcoral",
#                               ifelse(full_results$fdr3 < 0.2 & full_results$Beta3 < 0, "lightblue", "gray"))
# 
# # Identify significant points (fdr < 0.05)
# significant_df3 <- full_results[full_results$fdr3 < 0.2, ]

Genes <- length(unique(full_results$Gene))
Cells <- ncol(so_celltype)

#Figure Range
# max <- max(significant_df$Beta,na.rm=T)
# min <- min(significant_df$Beta,na.rm=T)

max <- max(full_results$Beta,na.rm=T)
min <- min(full_results$Beta,na.rm=T)

# Create the volcano plot using ggplot
volcano_plot <- ggplot(full_results, aes(x = Beta, y = PValue10, color = color)) +
  geom_point(alpha = 0.7) +  # Plot points with transparency
  scale_color_identity() +  # Use the color column directly
  theme_minimal() +  # Minimal theme
  labs(
    title = "Type 2 Diabetes vs. Lean Controls (All No Medication)",
    subtitle = "POD, Adjusted for Age, Sex & BMI",
    x = "Fold Change",
    y = "-log10(P-Value)",
    color = "FC Direction Direction",
    caption = paste0("FDR < 0.05, Genes = ",Genes,", Cells = ",Cells,", Non-Convergence Rate: ",Nonconvergence_Rate)
  ) +
  xlim(min,max)+
  theme(
    plot.title = element_text(hjust = 0),
    axis.text.x = element_text(angle = 0, hjust = 1)
  )+
  # # Add labels for significant points
  geom_text(data = significant_df, aes(label = Gene),
            vjust = 1, hjust = 1, size = 3, check_overlap = TRUE, color = "black")
# Add labels for significant points with ggrepel
# geom_text_repel(data = significant_df, aes(label = Gene),
#                 size = 3, color = "black", box.padding = 0.5, max.overlaps = 20)

volcano_plot
pdf(fs::path(dir.results,"Plot_POD_T2D_LC_NoMed_lmm_no_zi.pdf"),width=10,height=7)
print(volcano_plot)
dev.off()

```

## B. Medication Status - Type 2 Diabetes Only
### i. Pseudobulk PT Cells
```{r}
so_celltype <- subset(so_kpmp_sc,KPMP_celltype2=="aPT")
ncol(so_celltype) #37948 cells
nrow(so_celltype) #14627 genes
# Extract the gene expression data for all genes
gene_expression <- as.data.frame(GetAssayData(so_celltype, layer = "data"))

#Assign gene names as rownames, cell names as colnames
# rownames(gene_expression) <- rownames(so_celltype) #Gene Names
# colnames(gene_expression) <- colnames(so_celltype) #Cell Names
#Transpose gene expression dataset to merge with metadata
gene_expression <- t(gene_expression)
gene_expression <- data.frame(gene_expression) #Make a dataframe again after transposing
#Set gene list
gene_list_total <- colnames(gene_expression)
gene_expression$cellname <- rownames(gene_expression) 
rownames(gene_expression) <- NULL

# Extract the metadata
metadata <- so_celltype@meta.data
# metadata <- metadata %>%
#   mutate(across(everything(),~ifelse(.==".",NA,.)))
metadata$cellname <- rownames(metadata)
rownames(metadata) <- NULL

# Combine the gene expression data and metadata
data <- tidylog::full_join(metadata,gene_expression,by="cellname")
rm(metadata,gene_expression)

#Compare Health Controls to Type 2 Diabetes on no meds
data_subset <- data %>%
  filter(group=="Type_2_Diabetes") 
# filter(medication=="no_med")

data_subset$medication <- factor(data_subset$medication)
data_subset <- data_subset %>% 
  dplyr::select(c("kit_id","medication","group","age","sex","bmi",all_of(gene_list_total)))
rm(data)
gc()
rm(so_celltype)

# batch_size <- 2000
# total_cores <- 50
# 
# # Deduplicate gene list to avoid double-processing
# gene_list_total <- unique(gene_list_total)
# 
# # Split the gene list into batches
# gene_batches <- split(gene_list_total, ceiling(seq_along(gene_list_total) / batch_size))
# 
# # Define the function that processes a single gene
# process_gene <- function(gene) {
#     # tryCatch({
#   # if (sum(data_subset[[gene]]) > 0) {
#     formula <- as.formula(paste0(gene, " ~ medication + age + sex + bmi + (1 | kit_id)"))
#     model <- suppressWarnings(lmer(formula, data = data_subset))
# 
#     emm <- emmeans(model, ~ medication, lmerTest.limit = 19037, pbkrtest.limit = 19037)
# 
#     contrast_results <- contrast(emm, list(
#       "GLP1 vs No Med"        = c(0, 1, 0, -1),
#       "SGT2 vs No Med"        = c(0, 0, 1, -1),
#       "GLP1+SGT2 vs No Med"   = c(0, 0, 0, 1),
#       "GLP1+SGT2 vs GLP1"     = c(0, -1, 1, 0),
#       "GLP1+SGT2 vs SGT2"     = c(0, 0, -1, 1)
#     ))
# 
#     contrast_df <- as.data.frame(confint(contrast_results))
#     contrast_df$PValue <- summary(contrast_results)$p.value
#     contrast_df$Gene <- gene
# 
#     return(contrast_df)
#   #   } else {
#   #     return(NULL)
#   #   }
#   # }, error = function(e) {
#   #   return(NULL)
#   # })
# }
# 
# # Set number of cores
# total_cores <- parallel::detectCores() - 1
# 
# # Initialize and register cluster
# cl <- makeCluster(total_cores)
# registerDoParallel(cl)
# 
# # Run analysis per batch
# all_results <- lapply(seq_along(gene_batches), function(batch_idx) {
#   batch_genes <- gene_batches[[batch_idx]]
# 
#   batch_results <- foreach(
#     gene = batch_genes,
#     .combine = rbind,
#     .packages = c("glmmTMB", "lme4","emmeans"),
#     .export = c("process_gene", "data_subset")
#   ) %dopar% {
#     process_gene(gene)
#   }
# 
#   cat("Processed batch", batch_idx, "with", length(batch_genes), "genes\n")
#   return(batch_results)
# })
# 
# # Stop cluster
# stopCluster(cl)
# 
# # Combine all batch results into a single data frame
# final_results <- do.call(rbind, all_results)

# library(parallel)
# library(doParallel)
# library(foreach)
# library(lme4)
# library(emmeans)


# Load necessary libraries
library(parallel)
library(lme4)
library(emmeans)

# Define the number of cores to use
total_cores <- 50
batch_size <- 1

# Deduplicate gene list to avoid double-processing
gene_list_total <- unique(gene_list_total)[1:10]

# Split the gene list into batches
gene_batches <- split(gene_list_total, ceiling(seq_along(gene_list_total) / batch_size))

# Define the function that processes a single gene
process_gene <- function(gene) {
    # Define the formula for the linear mixed-effects model
    formula <- as.formula(paste0(gene, " ~ medication + age + sex + bmi + (1 | kit_id)"))
    
    # Fit the model
    model <- suppressWarnings(lmer(formula, data = data_subset))
    
    # Compute emmeans
    emm <- emmeans(model, ~ medication, lmerTest.limit = 19037, pbkrtest.limit = 19037)

    # Define the contrasts to test
    contrast_results <- contrast(emm, list(
      "GLP1 vs No Med"        = c(0, 1, 0, -1),
      "SGT2 vs No Med"        = c(0, 0, 1, -1),
      "GLP1+SGT2 vs No Med"   = c(0, 0, 0, 1),
      "GLP1+SGT2 vs GLP1"     = c(0, -1, 1, 0),
      "GLP1+SGT2 vs SGT2"     = c(0, 0, -1, 1)
    ))

    # Convert the contrast results to a data frame and add PValue and Gene columns
    contrast_df <- as.data.frame(confint(contrast_results))
    contrast_df$PValue <- summary(contrast_results)$p.value
    contrast_df$Gene <- gene

    return(contrast_df)
}

# Set up the parallel execution
# Use mclapply to process genes in parallel across multiple cores
result_list <- mclapply(gene_list_total, process_gene, mc.cores = total_cores)

# Combine the results into a single data frame
final_results <- do.call(rbind, result_list)

# View or save the final results
head(final_results)


# # #Make volcano plot of all gene results for group
full_results <- full_results %>%
  mutate(fdr=p.adjust(PValue,method="fdr"))
full_results$PValue10 <- -log10(pmax(full_results$PValue, 1e-10))  # Avoid log(0)


# write.csv(full_results,fs::path(dir.results,"PT_S1_S2_T2D_HC_No_Med.csv"))
full_results <- read.csv(fs::path(dir.results,"PT_S1_S2_T2D_HC_No_Med.csv"))


# Add a column to determine color based on the sign of the beta
# Create a new color column based on beta sign and FDR threshold
full_results$color <- ifelse(full_results$fdr < 0.2 & full_results$Beta > 0, "lightcoral",
                             ifelse(full_results$fdr < 0.2 & full_results$Beta < 0, "lightblue", "gray"))

# Identify significant points (fdr < 0.05)
significant_df <- full_results[full_results$fdr < 0.2, ]
# full_results$color <- ifelse(full_results$PValue < 0.05 & full_results$Beta > 0, "red", 
#                              ifelse(full_results$PValue < 0.05 & full_results$Beta < 0, "blue", "gray"))
# 
# # Identify significant points (PValue < 0.05)
# significant_df <- full_results[full_results$PValue < 0.05, ]

# Create the volcano plot using ggplot
volcano_plot <- ggplot(full_results, aes(x = Beta, y = PValue10, color = color)) +
  geom_point(alpha = 0.7) +  # Plot points with transparency
  scale_color_identity() +  # Use the color column directly
  theme_minimal() +  # Minimal theme
  labs(
    title = "PT-S1/S2 Cells: Lean Controls vs. Type 2 Diabetes (No Medication)",
    subtitle = "Adjusted for Age, Sex, & BMI",
    x = "Log2FC",
    y = "-log10(P-Value)",
    color = "Beta Direction",
    caption = "Positive Beta = Upregulation in T2D vs. LC, Negative Beta = Downregulation in T2D vs. LC \n Signficance threshold: FDR < 0.2"
  ) +
  # Add a horizontal line at -log10(0.05)  1.3 (FDR threshold)
  # geom_hline(yintercept = -log10(0.05), linetype = "dashed", color = "red", linewidth = 1) +
  theme(
    plot.title = element_text(hjust = 0),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )+
  # # Add labels for significant points
  # geom_text(data = significant_df, aes(label = gene), 
  #           vjust = 1, hjust = 1, size = 3, check_overlap = TRUE, color = "black")
  # Add labels for significant points with ggrepel
  geom_text_repel(data = significant_df, aes(label = Gene),
                  size = 3, color = "black", box.padding = 0.5, max.overlaps = 20)

pdf(fs::path(dir.results,"Plot_PT_S1_S2_T2D_HC_No_Med.pdf"),width=10,height=7)
print(volcano_plot)
dev.off()
```
#### ii. PT-S1/2
```{r}
#PT-S1/2
so_celltype <- subset(so_filtered,KPMP_celltype=="PT-S1/S2")
# Extract the gene expression data for all genes
gene_expression <- as.data.frame(GetAssayData(so_celltype, layer = "data"))

#Assign gene names as rownames, cell names as colnames
# rownames(gene_expression) <- rownames(so_celltype) #Gene Names
# colnames(gene_expression) <- colnames(so_celltype) #Cell Names
#Transpose gene expression dataset to merge with metadata
gene_expression <- t(gene_expression)
gene_expression <- data.frame(gene_expression) #Make a dataframe again after transposing
#Set gene list
gene_list <- colnames(gene_expression)
gene_expression$cellname <- rownames(gene_expression) 
rownames(gene_expression) <- NULL

# Extract the metadata
metadata <- so_celltype@meta.data
# metadata <- metadata %>%
#   mutate(across(everything(),~ifelse(.==".",NA,.)))
metadata$cellname <- rownames(metadata)
rownames(metadata) <- NULL

# Combine the gene expression data and metadata
data <- tidylog::full_join(metadata,gene_expression,by="cellname")
rm(metadata,gene_expression)

#Compare Health Controls to Type 2 Diabetes on no meds
data_subset <- data %>%
  filter(group=="Type_2_Diabetes") 
# filter(medication=="no_med")

data_subset$group <- factor(data_subset$group)

#Log transform gene expression
data_subset <- data_subset %>% 
  mutate(across(all_of(gene_list),~log2(.+1)))

full_results <- data.frame()

# Total number of genes
total_genes <- length(gene_list)

# Calculate the batch size
batch_size <- round(total_genes / 5)

# Simulate a vector of genes (replace this with your actual gene data)
genes <- 1:total_genes

# Loop through the genes in batches
for (i in seq(1, total_genes, by = batch_size)) {
  
  # Get the current batch (subsetting the gene vector)
  batch <- genes[i:min(i + batch_size - 1, total_genes)]
  for (gene in gene_list[batch]) { #tester genes
    # m0 <- as.formula(paste0(gene," ~ group + (1 | kit_id)"))
    # model <- lmer(m0,data=data_subset)
    #Adjust for key covariates
    m1 <- as.formula(paste0(gene," ~ medication + age + sex + bmi + (1 | kit_id)"))
    model1 <- lmer(m1,data=data_subset)
    
    # Compute estimated marginal means
    emm <- emmeans(model1, ~ medication)
    
    # Define and test the contrasts
    contrast(emm, list(
      "GLP1 vs No Med" = c(0, 1, 0, -1),  # GLP1 vs No Med
      "SGT2 vs No Med" = c(0, 0, 1, -1),  # SGT2 vs No Med
      "GLP1+SGT2 vs No Med" = c(0, 0, 0, 1),  # GLP1+SGT2 vs No Med
      "GLP1+SGT2 vs GLP1" = c(0, -1, 1, 0),  # GLP1+SGT2 vs GLP1
      "GLP1+SGT2 vs SGT2" = c(0, 0, -1, 1)  # GLP1+SGT2 vs SGT2
    ))
    
    
    # beta <- round(summary(model)$coef[2,1],3)
    # pval <-round(summary(model)$coef[2,5],5)
    beta1 <- round(summary(model1)$coef[2,1],3)
    pval1 <-round(summary(model1)$coef[2,5],5)
    
    result <- data.frame(Gene=gene,Beta=beta1,PValue=pval1)
    full_results <- rbind(full_results,result)
  }
}
# #Make volcano plot of all gene results for group
full_results <- full_results %>%
  mutate(fdr=p.adjust(PValue,method="fdr"))
full_results$PValue10 <- -log10(pmax(full_results$PValue, 1e-10))  # Avoid log(0)


# write.csv(full_results,fs::path(dir.results,"PT_S1_S2_T2D_HC_No_Med.csv"))
full_results <- read.csv(fs::path(dir.results,"PT_S1_S2_T2D_HC_No_Med.csv"))

# }
# Add a column to determine color based on the sign of the beta
# Create a new color column based on beta sign and FDR threshold
full_results$color <- ifelse(full_results$fdr < 0.2 & full_results$Beta > 0, "lightcoral",
                             ifelse(full_results$fdr < 0.2 & full_results$Beta < 0, "lightblue", "gray"))

# Identify significant points (fdr < 0.05)
significant_df <- full_results[full_results$fdr < 0.2, ]
# full_results$color <- ifelse(full_results$PValue < 0.05 & full_results$Beta > 0, "red", 
#                              ifelse(full_results$PValue < 0.05 & full_results$Beta < 0, "blue", "gray"))
# 
# # Identify significant points (PValue < 0.05)
# significant_df <- full_results[full_results$PValue < 0.05, ]

# Create the volcano plot using ggplot
volcano_plot <- ggplot(full_results, aes(x = Beta, y = PValue10, color = color)) +
  geom_point(alpha = 0.7) +  # Plot points with transparency
  scale_color_identity() +  # Use the color column directly
  theme_minimal() +  # Minimal theme
  labs(
    title = "PT-S1/S2 Cells: Lean Controls vs. Type 2 Diabetes (No Medication)",
    subtitle = "Adjusted for Age, Sex, & BMI",
    x = "Log2FC",
    y = "-log10(P-Value)",
    color = "Beta Direction",
    caption = "Positive Beta = Upregulation in T2D vs. LC, Negative Beta = Downregulation in T2D vs. LC \n Signficance threshold: FDR < 0.2"
  ) +
  # Add a horizontal line at -log10(0.05)  1.3 (FDR threshold)
  # geom_hline(yintercept = -log10(0.05), linetype = "dashed", color = "red", linewidth = 1) +
  theme(
    plot.title = element_text(hjust = 0),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )+
  # # Add labels for significant points
  # geom_text(data = significant_df, aes(label = gene), 
  #           vjust = 1, hjust = 1, size = 3, check_overlap = TRUE, color = "black")
  # Add labels for significant points with ggrepel
  geom_text_repel(data = significant_df, aes(label = Gene),
                  size = 3, color = "black", box.padding = 0.5, max.overlaps = 20)

pdf(fs::path(dir.results,"Plot_PT_S1_S2_T2D_HC_No_Med.pdf"),width=10,height=7)
print(volcano_plot)
dev.off()
```
#### iii. PT-S3
```{r}
#PT-S3
so_celltype <- subset(so_filtered,KPMP_celltype=="PT-S3")
# Extract the gene expression data for all genes
gene_expression <- as.data.frame(GetAssayData(so_celltype, layer = "data"))

#Assign gene names as rownames, cell names as colnames
# rownames(gene_expression) <- rownames(so_celltype) #Gene Names
# colnames(gene_expression) <- colnames(so_celltype) #Cell Names
#Transpose gene expression dataset to merge with metadata
gene_expression <- t(gene_expression)
gene_expression <- data.frame(gene_expression) #Make a dataframe again after transposing
#Set gene list
gene_list <- colnames(gene_expression)
gene_expression$cellname <- rownames(gene_expression) 
rownames(gene_expression) <- NULL

# Extract the metadata
metadata <- so_celltype@meta.data
# metadata <- metadata %>%
#   mutate(across(everything(),~ifelse(.==".",NA,.)))
metadata$cellname <- rownames(metadata)
rownames(metadata) <- NULL

# Combine the gene expression data and metadata
data <- tidylog::full_join(metadata,gene_expression,by="cellname")
rm(metadata,gene_expression)

#Compare Health Controls to Type 2 Diabetes on no meds
data_subset <- data %>%
  filter(group=="Type_2_Diabetes" | group=="Lean_Control") %>%
  filter(medication=="no_med")

data_subset$group <- factor(data_subset$group)

#Log transform gene expression
data_subset <- data_subset %>% 
  mutate(across(all_of(gene_list),~log2(.+1)))

full_results <- data.frame()

# Total number of genes
total_genes <- length(gene_list)

# Calculate the batch size
batch_size <- round(total_genes / 5)

# Simulate a vector of genes (replace this with your actual gene data)
genes <- 1:total_genes

# Loop through the genes in batches
for (i in seq(1, total_genes, by = batch_size)) {
  
  # Get the current batch (subsetting the gene vector)
  batch <- genes[i:min(i + batch_size - 1, total_genes)]
  for (gene in gene_list[batch]) { #tester genes
    # m0 <- as.formula(paste0(gene," ~ group + (1 | kit_id)"))
    # model <- lmer(m0,data=data_subset)
    #Adjust for key covariates
    if (sum(data_subset[[gene]])>0) {
      m1 <- as.formula(paste0(gene," ~ group + age + sex + bmi + (1 | kit_id)"))
      model1 <- lmer(m1,data=data_subset)
      
      # # Plot residuals vs fitted values
      # residuals <- residuals(model1)
      # fitted_values <- fitted(model1)
      # 
      # plot(fitted_values, residuals)
      # abline(h = 0, col = "red")
      # 
      # # Histogram of residuals
      # hist(residuals, main = "Histogram of Residuals", xlab = "Residuals", col = "lightblue", border = "black")
      # 
      # # Q-Q plot for normality
      # qqnorm(residuals)
      # qqline(residuals, col = "red")
      
      # beta <- round(summary(model)$coef[2,1],3)
      # pval <-round(summary(model)$coef[2,5],5)
      beta1 <- round(summary(model1)$coef[2,1],3)
      pval1 <-round(summary(model1)$coef[2,5],5)
      
    } else {
      beta1 <- NA
      pval1 <-NA
    }
    result <- data.frame(Gene=gene,Beta=beta1,PValue=pval1) 
    full_results <- rbind(full_results,result)
  }
}
# #Make volcano plot of all gene results for group
full_results <- full_results %>%
  mutate(fdr=p.adjust(PValue,method="fdr"))
# # Identify significant points (PValue < 0.05)
# significant_df <- full_results[full_results$PValue < 0.05, ]
full_results$PValue10 <- -log10(pmax(full_results$PValue, 1e-10))  # Avoid log(0)

# write.csv(full_results,fs::path(dir.results,"PT_S3_T2D_HC_No_Med.csv"))
full_results <- read.csv(fs::path(dir.results,"PT_S3_T2D_HC_No_Med.csv"))

# }
# Add a column to determine color based on the sign of the beta
# Create a new color column based on beta sign and FDR threshold
full_results$color <- ifelse(full_results$fdr < 0.05 & full_results$Beta > 0, "red",
                             ifelse(full_results$fdr < 0.05 & full_results$Beta < 0, "blue", "gray"))

# Identify significant points (fdr < 0.05)
significant_df <- full_results[full_results$fdr < 0.05, ]
# full_results$color <- ifelse(full_results$PValue < 0.05 & full_results$Beta > 0, "red", 
#                              ifelse(full_results$PValue < 0.05 & full_results$Beta < 0, "blue", "gray"))
#
# Create the volcano plot using ggplot
volcano_plot <- ggplot(full_results, aes(x = Beta, y = PValue10, color = color)) +
  geom_point(alpha = 0.7) +  # Plot points with transparency
  scale_color_identity() +  # Use the color column directly
  theme_minimal() +  # Minimal theme
  labs(
    title = "PT-S3 Cells: Lean Controls vs. Type 2 Diabetes (No Medication)",
    subtitle = "Adjusted for Age, Sex, & BMI",
    x = "Log2FC",
    y = "-log10(P-Value)",
    color = "Beta Direction",
    caption = "Positive Beta = Upregulation in T2D vs. LC, Negative Beta = Downregulation in T2D vs. LC \n Signficance threshold: FDR < 0.2"
  ) +
  theme(
    plot.title = element_text(hjust = 0),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )+
  ylim(0, 40) +
  # # Add labels for significant points
  # geom_text(data = significant_df, aes(label = gene), 
  #           vjust = 1, hjust = 1, size = 3, check_overlap = TRUE, color = "black")
  # Add labels for significant points with ggrepel
  # geom_text_repel(data = significant_df, aes(label = Gene),
  #                 size = 3, color = "black", box.padding = 0.5, max.overlaps = 20)
  # Add labels for significant points with ggrepel
  geom_text_repel(data = significant_df, aes(label = Gene),
                  size = 3, color = "black", box.padding = 0.5, 
                  max.overlaps = 20, min.segment.length = 0) 

pdf(fs::path(dir.results,"Plot2_PT_S3_T2D_HC_No_Med.pdf"),width=10,height=7)
print(volcano_plot)
dev.off()
```
#### iv. aPT
```{r}
# library(lme4)
# library(emmeans)
# library(parallel)
# library(doParallel)
# library(dplyr)
so_celltype <- subset(so_kpmp_sc,KPMP_celltype2=="aPT")
ncol(so_celltype) #37948 cells
nrow(so_celltype) #14627 genes
# Extract the gene expression data for all genes
gene_expression <- as.data.frame(GetAssayData(so_celltype, layer = "data"))

#Assign gene names as rownames, cell names as colnames
# rownames(gene_expression) <- rownames(so_celltype) #Gene Names
# colnames(gene_expression) <- colnames(so_celltype) #Cell Names
#Transpose gene expression dataset to merge with metadata
gene_expression <- t(gene_expression)
gene_expression <- data.frame(gene_expression) #Make a dataframe again after transposing
#Set gene list
gene_list_total <- colnames(gene_expression)
gene_expression$cellname <- rownames(gene_expression) 
rownames(gene_expression) <- NULL

# Extract the metadata
metadata <- so_celltype@meta.data
# metadata <- metadata %>%
#   mutate(across(everything(),~ifelse(.==".",NA,.)))
metadata$cellname <- rownames(metadata)
rownames(metadata) <- NULL

# Combine the gene expression data and metadata
data <- tidylog::full_join(metadata,gene_expression,by="cellname")
rm(metadata,gene_expression)

#Compare Health Controls to Type 2 Diabetes on no meds
data_subset <- data %>%
  filter(group=="Type_2_Diabetes") 
# filter(medication=="no_med")

data_subset$medication <- factor(data_subset$medication)
data_subset <- data_subset %>% 
  dplyr::select(c("kit_id","medication","group","age","sex","bmi",all_of(gene_list_total)))
rm(data)
gc()
rm(so_celltype)
gene_hold <- gene_list_total
gene_list_total <- gene_hold[1001:5000]
# Define your function for processing a gene
process_gene <- function(gene) {
    # Define the formula for the linear mixed-effects model
    formula <- as.formula(paste0(gene, " ~ medication + age + sex + bmi + (1 | kit_id)"))
    
    # Fit the model
    model <- suppressWarnings(lmer(formula, data = data_subset))
    
    # Compute emmeans
    emm <- emmeans(model, ~ medication, lmerTest.limit = 19037, pbkrtest.limit = 19037)

    # Define the contrasts to test
    contrast_results <- contrast(emm, list(
      "GLP1 vs No Med"        = c(0, 1, 0, -1),
      "SGT2 vs No Med"        = c(0, 0, 1, -1),
      "GLP1+SGT2 vs No Med"   = c(0, 0, 0, 1),
      "GLP1+SGT2 vs GLP1"     = c(0, -1, 1, 0),
      "GLP1+SGT2 vs SGT2"     = c(0, 0, -1, 1)
    ))

    # Convert the contrast results to a data frame and add PValue and Gene columns
    contrast_df <- as.data.frame(confint(contrast_results))
    contrast_df$PValue <- summary(contrast_results)$p.value
    contrast_df$Gene <- gene
    contrast_df <- as.data.drame(contrast_df)

    return(contrast_df)
}

# Set up parallel backend (use maximum available cores)
num_cores <- detectCores() - 1  # Reserve one core for the system, adjust as needed
cl <- makeCluster(num_cores)
registerDoParallel(cl)

# List of genes to process (replace with actual gene names)
# genes <- c("Gene1", "Gene2", "Gene3", ...)  # your list of genes
genes <- gene_list_total
# Run the analysis for each gene in parallel
results <- foreach(gene = genes, .combine = rbind, .packages = c("lme4", "emmeans", "dplyr")) %dopar% {
    process_gene(gene)
}

# Stop the parallel cluster
stopCluster(cl)

# View the combined results
head(results)
# results2 <- results
rm(results)
# full_results <- rbind(results,results2)
nc_rate <- full_results %>%
  group_by(contrast, Gene) %>%
  summarize(
    total_na = sum(is.na(estimate)),
    total_nan = sum(is.nan(estimate)),
    total_non_na_non_nan = sum(!is.na(estimate) & !is.nan(estimate)),
    .groups = "drop"
  ) %>% 
  ungroup() 
nc_rate2 <- full_results %>%
  group_by(contrast, Gene) %>%
  summarize(
    total_na = sum(is.na(PValue)),
    total_nan = sum(is.nan(PValue)),
    total_non_na_non_nan = sum(!is.na(PValue) & !is.nan(PValue)),
    .groups = "drop"
  ) %>% 
  ungroup() 
# Nonconvergence_Rate <- sum(nc_rate$total_na,nc_rate$total_nan)/sum(nc_rate$total_na,nc_rate$total_nan+total)
# Nonconvergence_Rate <- paste0(round(((length(which(is.na(full_results$PValue)))+length(which(full_results$PValue=="NaN")))/length(full_results$Gene))*100,0),"%")
```

