---
title: "Liver Analysis"
author: "Hailey Hampson"
date: "2024-09-24"
output: html_document
--- 

#1. Set up Libraries & Directores
```{r libraries, echo=F, include = F}
library(reprex)
library(tidyverse)
library(BiocManager)        
library(arsenal)
library(dplyr)
library(ggplot2)
library(ggrepel)
library(Seurat)
library(future)
library(colorspace)
library(patchwork)
library(ggdendro)
library(cowplot)
library(ggpubr)
library(venn)
library(rstatix)
library(table1)
library(Biobase)
library(ReactomeGSA)
library(GSEABase)
library(msigdbr)
library(kableExtra)
library(knitr)
library(SingleCellExperiment)
library(fgsea)
library(EnhancedVolcano)
library(openxlsx)
library(BiocManager)
library(MAST)
library(ggrepel)
# library(qpcR)
library(ggpubr)
library(openxlsx)
library(ggplot2)
library(GGally)
library(GSEABase)
library(limma)
library(reshape2)
library(data.table)
library(knitr)
library(TxDb.Hsapiens.UCSC.hg19.knownGene)
library(stringr)
library(NMF)
library(rsvd)
library(RColorBrewer)
library(MAST)
library(devtools)
# install_github("Sun-lab/ideas",force=T)
#library(ideas)
library(foreach)
library(doRNG)
library(doParallel)
library(fs)
registerDoParallel(cores = 6)
library(VennDiagram)


#Local file path
dir.dat <- c("/Volumes/Peds Endo/Petter Bjornstad")
dir.dat2 <- c("/Volumes/Peds Endo/Petter Bjornstad/scRNA/data_clean")
dir.code <- c("/Users/hhampson/Documents/CHCO-Code/Petter Bjornstad/Liver analysis/Liver scRNAseq")
dir.results <- c("/Volumes/Peds Endo/Petter Bjornstad/Kidney Project/Results")

#Lambda file path
# dir.dat <- c("/run/user/1026/gvfs/smb-share:server=ucdenver.pvt,share=som/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad")
# dir.code <- c("/home/Github_Repo/CHCO-Code/Petter Bjornstad/Kidney scRNA/Kidney scRNA")
# dir.results <- c(fs::path(dir.dat,"Kidney Project/Results"))

#Load functions
source("Kidney_functions_sc.R")

```

#2. Load Full Data & Clean: scRNA & MetaData
```{r}
#Increase Memory
mem.maxVSize(64000000000)
#Local
# dir.dat <- c("/Users/hhampson/Dropbox/Bjornstad data")
# so_kidney_sc <- readRDS(fs::path(dir.dat,"Kidney scRNA","PB_90samples_Harmony_rpca_Fadhl_PhilApproved_091024.RDS"))
# load("/Volumes/Peds Endo/Petter Bjornstad/Data Harmonization/Data Exports/PB90_clinical_metadata.RData")
# so_kidney_sc <- AddMetaData(so_kidney_sc, so_meta_combined)
# rm(so_meta_combined)
# gc()

#Lambda
so_kidney_sc <- readRDS(fs::path(dir.dat,"scRNA","data_raw","PB_90samples_Harmony_rpca_Fadhl_PhilApproved_091024.RDS"))
load(fs::path(dir.dat,"Data Harmonization","Data Exports","PB90_clinical_metadata_12_10.RData"))
so_kidney_sc <- AddMetaData(so_kidney_sc, so_meta_combined)
rm(so_meta_combined)
gc()

#Define 4 exposure groups: 
#SGLT2i(+)/GLP1RA(+), SGLT2i(+)/GLP1RA(-), SGLT2i(-)/GLPRA(+), SGLT2i(-)/GLP1RA(-)
so_kidney_sc@meta.data <- so_kidney_sc@meta.data %>% 
  mutate(both=ifelse(epic_glp1ra_1=="Yes" & epic_sglti2_1=="Yes","Yes","No")) %>% 
  mutate(exclusive_sglt2=ifelse(epic_sglti2_1=="Yes" & epic_glp1ra_1=="No","Yes","No")) %>% 
  mutate(exclusive_glp1=ifelse(epic_sglti2_1=="No" & epic_glp1ra_1=="Yes","Yes","No")) %>% 
  mutate(neither=ifelse(epic_sglti2_1=="No" & epic_glp1ra_1=="No","Yes","No"))


so_kidney_sc@meta.data <- so_kidney_sc@meta.data %>% 
  mutate(group2 = case_when(both == "Yes" ~ "both",
                            exclusive_sglt2 == "Yes" ~ "exclusive_sglt2",
                            exclusive_glp1 == "Yes" ~ "exclusive_glp1",
                            neither == "Yes" ~ "neither"))

so_kidney_sc@meta.data$group2 <- factor(so_kidney_sc@meta.data$group2, levels = c("neither", "both", "exclusive_sglt2", "exclusive_glp1"))

#so_diab <- subset(so_kidney_sc,group=="Type 2 Diabetes")
#rm(so_kidney_sc)
gc()

#Create senesence so in all cell types
#sens_genes <- c(sens_genes,"CDKN1A")
#sens_gene_in_data <- intersect(sens_genes, rownames(so_kidney_sc))
# Subset the Seurat object to include only genes in sens_gene_in_data
#so_sens_all <- subset(so_kidney_sc, features = sens_gene_in_data)
#rm(so_kidney_sc)
# # saveRDS(so_sens_all,fs::path(dir.dat,"so_sens_genes_all_cells_KIDNEY.RDS"))
# # rm(so_sens_all)
# # 
# # #Load sens all cells
# # so_sens <- readRDS(fs::path(dir.dat,"so_sens_genes_all_cells_KIDNEY.RDS"))
# # dat <- so_sens@meta.data
# # so_diab <- subset(so_sens, diagnosis_of_diabetes == "Yes")
# # 
# #PT Cells
# so_sens_all@meta.data$PT <- ifelse(grepl("PT-",so_sens_all@meta.data$celltype_rpca),"PT","Non-PT")
# Idents(so_sens_all) <- so_sens_all$PT
# so_sens <- subset(so_sens_all, PT=="PT")
# # saveRDS(so_sens,fs::path(dir.dat,"so_sens_genes_PT_Cells_KIDNEY.RDS"))
# rm(so_sens_all,so_sens)
# so_sens_PT <- readRDS(fs::path(dir.dat,"Liver Project","so_sens_genes_PT_Cells_KIDNEY.RDS"))
# so_sens_PT@meta.data$group1 <- ifelse(so_sens_PT@meta.data$group=="Type 2 Diabetes" | so_sens_PT@meta.data$group=="Lean Control" ,"Yes","No")
# so_sens_PT1 <- subset(so_sens_PT, group1=="Yes")
# so_sens_PT@meta.data$group2 <- ifelse(so_sens_PT@meta.data$group=="Type 2 Diabetes","Yes","No")
# so_sens_PT2 <- subset(so_sens_PT, group2=="Yes")
# so_sens_PT3 <- subset(so_sens_PT, group=="Type 2 Diabetes")
# 
# 
# 
# #TAL Cells
# so_sens_all@meta.data$TAL <- ifelse(grepl("TAL",so_sens_all@meta.data$celltype_rpca),"TAL","Non-TAL")
# Idents(so_sens_all) <- so_sens_all$TAL
# so_sens <- subset(so_sens_all, TAL=="TAL")
# # saveRDS(so_sens,fs::path(dir.dat,"so_sens_genes_TAL_Cells_KIDNEY.RDS"))
# rm(so_sens_all,so_sens,so_sens_TAL,so_sens_TAL1)
# so_sens_TAL <- readRDS(fs::path(dir.dat,"Liver Project","so_sens_genes_TAL_Cells_KIDNEY.RDS"))
# so_sens_TAL@meta.data$group1 <- ifelse(so_sens_TAL@meta.data$group=="Type 2 Diabetes"| so_sens_TAL@meta.data$group=="Lean Control","Yes","No")
# so_sens_TAL1 <- subset(so_sens_TAL, group1=="Yes")
# so_sens_TAL@meta.data$group2 <- ifelse(so_sens_TAL@meta.data$group=="Type 2 Diabetes","Yes","No")
# so_sens_TAL2 <- subset(so_sens_TAL, group2=="Yes")
# so_sens_TAL3 <- subset(so_sens_TAL, group=="Type 2 Diabetes")
```
#3. Descriptive Statistics
```{r echo = F}
dat <- readRDS("/run/user/1026/gvfs/smb-share:server=ucdenver.pvt,share=som/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/Data Harmonization/Data Exports/PB90_clinical_metadata_only_12_10.rds")
unique(so_kidney_sc$kit_id)
dat <- dat %>% 
  filter(kit_id %in% unique(so_kidney_sc$kit_id)) %>%
  mutate(both=ifelse(epic_glp1ra_1=="Yes" & epic_sglti2_1=="Yes","Yes","No")) %>% 
  mutate(exclusive_sglt2=ifelse(epic_sglti2_1=="Yes" & epic_glp1ra_1=="No","Yes","No")) %>% 
  mutate(exclusive_glp1=ifelse(epic_sglti2_1=="No" & epic_glp1ra_1=="Yes","Yes","No")) %>% 
  mutate(neither=ifelse(epic_sglti2_1=="No" & epic_glp1ra_1=="No","Yes","No"))
dat <- dat %>% 
  filter(group=="Type 2 Diabetes")

dat$record_id %in% unique(so_kidney_sc$record_id)
unique(so_kidney_sc$record_id)
table(dat$exclusive_sglt2) #-/+ #27 no, 10 yes
table(dat$exclusive_glp1) #+/- #29 no, 8 yes
table(dat$both) #+/+ #33 no, 4 yes
table(dat$neither) #-/- #22 no, 15 yes




# dat <- dat %>% 
#   filter(diagnosis_of_diabetes=="Yes")
# cor(dat$ast,dat$alt,method="pearson")
# cor(dat$ast,dat$alt,method="spearman")
# cor(dat$alt,dat$tg,method="spearman")

# dat$diagnosis_of_diabetes <- factor(dat$diagnosis_of_diabetes, levels=c("Yes","No"), labels=c("Type 2 Diabetes", "Obese Controls"))
# dat$nih_sex     <- factor(dat$nih_sex, levels=c("Male", "Female"), labels=c("Male", "Female"))
# dat$nih_ethnicity   <- factor(dat$nih_ethnicity, levels=c("Hispanic_Or_Latino","NonHispanic"), labels=c("Hispanic or Latino","Non-Hispanic/Non-Latino"))
# dat$nih_race   <- factor(dat$nih_race, levels=c("White","BlackAfAm","Multiracial","Other"),
#                          labels=c("White","Black","Multirace","Other"))
# dat$steatosis_grade <- as.factor(dat$steatosis_grade)
# dat$fibrosis_stage <- as.factor(dat$fibrosis_stage)
# dat$lobular_inflammation_percent <- as.factor(dat$lobular_inflammation_percent)
# 
# label(dat$age)      <- "Age (y)"
# label(dat$nih_sex)      <- "Sex"
# label(dat$nih_race)    <- "Race"
# label(dat$nih_ethnicity)    <- "Ethnicity"
# label(dat$diagnosis_of_diabetes)  <- "Diabetes Status"
# label(dat$bmi)   <- "Body Mass Index (kg/m2)"
# label(dat$diagnosis_of_MASLD)  <- "MASLD Status"
# label(dat$sbp)     <- "Systolic Blood Pressure (mmHg)"
# label(dat$dbp) <- "Diastolic Blood Pressure (mmHg)"
# label(dat$tg) <-  "Triglycerides (mg/dL)"
# label(dat$creatinine) <-  "Creatinine (mg/dL)"
# label(dat$steatosis_percent) <- "Steatosis Percent (%)"
# label(dat$fibrosis_stage) <- "Fibrosis Stage"
# label(dat$lobular_inflammation_percent) <- "Lobular Inflammation Percent (%)"
dat2 <- dat %>% 
  filter(group=="Type 2 Diabetes" | group=="Obese Control" | group=="Lean Control")
dat3 <- dat %>% 
  filter(group=="Type 2 Diabetes")
table1(~ age + sex + race + exclusive_glp1 + exclusive_sglt2 + both + acr_u + albuminuria_cat| group, data=dat2)
table1(~ age + sex + race + epic_sglti2_1 + acr_u + albuminuria_cat| epic_glp1ra_1 , data=dat2)

table1(~ age + nih_sex + nih_race + diagnosis_of_MASLD + bmi+ tg + ast + alt + ggt + steatosis_percent + lobular_inflammation_percent + fibrosis_stage | diagnosis_of_diabetes, data = dat)

table1(~ age + nih_sex + nih_race + diagnosis_of_MASLD + diagnosis_of_diabetes + a1c  + bmi+ tg + ast + alt + ggt + steatosis_percent + lobular_inflammation_percent + fibrosis_stage | glp1agonist, data = dat)


pvalue <- function(x,...) {
  # Construct vectors of data y, and groups (strata) g
  y <- unlist(x)
  g <- factor(rep(1:length(x), times=sapply(x, length)))
  if (is.numeric(y)) {
    # For numeric variables, perform a standard 2-sample t-test
    p <- t.test(y ~ g)$p.value
  } else {
    # For categorical variables, perform a chi-squared test of independence
    p <- chisq.test(table(y, g))$p.value
  }
  # Format the p-value, using an HTML entity for the less-than sign.
  # The initial empty string places the output on the line below the variable label.
  c("", sub("<", "&lt;", format.pval(p, digits=3, eps=0.001)))
}

mean_sd <- function(x) {
  # Handle missing values explicitly
  s <- stats::sd(x, na.rm=TRUE)
  m <- mean(x, na.rm=TRUE)
  sprintf("%.2f (%.2f)", m, s)
}

table1(~ age + nih_sex + nih_race + bmi + diagnosis_of_MASLD + sbp + dbp + tg | diagnosis_of_diabetes, data=dat,overall=F, extra.col=list(`P-value`=pvalue), render.continuous = mean_sd)

table1(~ age + nih_sex + nih_race + bmi + diagnosis_of_MASLD + sbp + dbp + tg + diagnosis_of_diabetes | glp1agonist, data=dat,overall=F, extra.col=list(`P-value`=pvalue), render.continuous = mean_sd)

table1(~ age + nih_sex + nih_race + bmi + diagnosis_of_MASLD + sbp + dbp + tg | diagnosis_of_diabetes, data=dat,render.continuous = render.continuous.custom)

table1(~ age + nih_sex + nih_race + diagnosis_of_MASLD + bmi+ tg + ast + alt + ggt + steatosis_percent + lobular_inflammation_percent + fibrosis_stage | diagnosis_of_diabetes, data = dat,render.continuous = render.continuous.custom)

# Generate the table
table1(~ age + nih_sex + nih_race + diagnosis_of_MASLD + bmi+ tg + ast + alt + ggt + steatosis_percent + lobular_inflammation_percent + fibrosis_stage | diagnosis_of_diabetes, 
       data = dat, 
       overall = , 
       extra.col = list(`P-value` = pvalue), 
       render.continuous = render.continuous.custom)
#bmi,masld,TGs,AST/ALT (Liver enzymes),ggt, steatosis grade and or percentages,lobular inflam grade, fibrosis stage

# Descriptive stats
liver_meta_raw_sc <- meta_liver_raw %>%
  filter(Cryostor_ID!="")
form <- paste("diagnosis_of_diabetes", paste(colnames(liver_meta_raw_sc)[6:36], collapse = " + "), sep = " ~ ")

summary(arsenal::tableby(formula = as.formula(form), data = liver_meta_raw_sc, test = F))

## MASLD Y/N
liver_meta_raw_sc <- meta_liver_raw %>%
  filter(Cryostor_ID!="")
form <- paste("diagnosis_of_MASLD", paste(colnames(liver_meta_raw_sc)[6:36], collapse = " + "), sep = " ~ ")

summary(arsenal::tableby(formula = as.formula(form), data = liver_meta_raw_sc, test = F))

## GLP-1RA Y/N
liver_meta_raw_sc <- liver_meta_raw %>%
  filter(Cryostor_ID!="")
form <- paste("glp1agonist", paste(colnames(liver_meta_raw_sc)[6:36], collapse = " + "), sep = " ~ ")

summary(arsenal::tableby(formula = as.formula(form), data = liver_meta_raw_sc, test = F))

```

#4. DEGS & GSEA
##a. All Cell Types
```{r}
#Among Type 2: No meds vs. GLP-1 Exclusive
##SGLT2i only -/+ #5 No, 0 yes
#GLP-1 only +/- # 2 Yes, 3 No
#Both +/+ #1 Both, 4 not both
#Neither-/- #3 neither, 2 on someting

all_genes <- rownames(so_kidney_sc)
#Diabetes Groups
#Type 2 vs. Lean Control
so_subgroup <- subset(so_kidney_sc,group=="Type 2 Diabetes" | group=="Lean Control")
so_subgroup$group <- factor(so_subgroup$group)
so_subgroup$Type_2_Diabetes <- so_subgroup$group 
degs_fxn(so=so_subgroup,cell=NULL,gene_set=all_genes,exposure="Type_2_Diabetes",exp_group="Type 2 Diabetes",ref_group="Lean Control",enrichment="Yes",top_gsea = 30)

#Create diab so
so_diab <- subset(so_kidney_sc,group=="Type 2 Diabetes")
#GLP-1 vs. None
so_sub <- subset(so_diab,group2=="exclusive_glp1" | group2=="neither")
so_sub$group2 <- factor(so_sub$group2)
#Define genes
all_genes <- rownames(so_sub)
#Rename variable
so_sub$Medication <- so_sub$group2
degs_fxn(so=so_sub,cell=NULL,gene_set=all_genes,exposure="Medication",exp_group="exclusive_glp1",ref_group="neither",enrichment="Yes",top_gsea = 30)
#GLP-1 vs. SGLT2
so_sub <- subset(so_diab,group2=="exclusive_glp1" | group2=="exclusive_sglt2")
so_sub$group2 <- factor(so_sub$group2)
#Define genes
all_genes <- rownames(so_sub)
#Rename variable
so_sub$Medication <- so_sub$group2
degs_fxn(so=so_sub,cell=NULL,gene_set=all_genes,exposure="Medication",exp_group="exclusive_glp1",ref_group="exclusive_sglt2",enrichment="Yes",top_gsea = 30)

#GLP-1 vs. Both
so_sub <- subset(so_diab,group2=="exclusive_glp1" | group2=="both")
so_sub$group2 <- factor(so_sub$group2)
#Define genes
all_genes <- rownames(so_sub)
#Rename variable
so_sub$Medication <- so_sub$group2
degs_fxn(so=so_sub,cell=NULL,gene_set=all_genes,exposure="Medication",exp_group="both",ref_group="exclusive_glp1",enrichment="Yes",top_gsea = 30)

#SGLT2 vs. None 
so_sub <- subset(so_diab,group2=="exclusive_sglt2" | group2=="neither")
so_sub$group2 <- factor(so_sub$group2)
#Define genes
all_genes <- rownames(so_sub)
#Rename variable
so_sub$Medication <- so_sub$group2
degs_fxn(so=so_sub,cell=NULL,gene_set=all_genes,exposure="Medication",exp_group="exclusive_sglt2",ref_group="neither",enrichment="Yes",top_gsea = 30)

#SGLT2 vs. Both
so_sub <- subset(so_diab,group2=="exclusive_sglt2" | group2=="both")
so_sub$group2 <- factor(so_sub$group2)
#Define genes
all_genes <- rownames(so_sub)
#Rename variable
so_sub$Medication <- so_sub$group2
degs_fxn(so=so_sub,cell=NULL,gene_set=all_genes,exposure="Medication",exp_group="both",ref_group="exclusive_sglt2",enrichment="Yes",top_gsea = 30)

#Both vs. None
so_sub <- subset(so_diab,group2=="neither" | group2=="both")
so_sub$group2 <- factor(so_sub$group2)
#Define genes
all_genes <- rownames(so_sub)
#Rename variable
so_sub$Medication <- so_sub$group2
degs_fxn(so=so_sub,cell=NULL,gene_set=all_genes,exposure="Medication",exp_group="both",ref_group="neither",enrichment="Yes",top_gsea = 30)

```

##b. Specific Cell Types
```{r}
#Major Cell Types
#unique(so_kidney_sc$celltype_rpca)
so_kidney_sc$celltype_rpca <- as.character(so_kidney_sc$celltype_rpca)
so_kidney_sc$celltype2 <- ifelse(grepl("PT-",so_kidney_sc$celltype_rpca),"PT",
                                 ifelse(grepl("TAL-",so_kidney_sc$celltype_rpca),"TAL",
                                        ifelse(grepl("PC-",so_kidney_sc$celltype_rpca),"PC",
                                               ifelse(grepl("IC-",so_kidney_sc$celltype_rpca) & !grepl("lowQuality",so_kidney_sc$celltype_rpca),"IC",
                                                      ifelse(grepl("DTL-",so_kidney_sc$celltype_rpca),"DTL",so_kidney_sc$celltype_rpca)))))
unique(so_kidney_sc$celltype2)
all_cell_types <- unique(so_kidney_sc$celltype2)
all_genes <- rownames(so_kidney_sc)

#Create diab so 
so_diab <- subset(so_kidney_sc,group=="Type 2 Diabetes")

#GLP-1 vs. None
so_sub <- subset(so_diab,group2=="exclusive_glp1" | group2 =="neither")
so_sub$group2 <- factor(so_sub$group2)
#Define genes
all_genes <- rownames(so_sub)
so_sub$Medication <- so_sub$group2
all_cell_types <- unique(so_sub$celltype2)
for (cell_type in all_cell_types) {
  degs_fxn(so=so_sub,cell=cell_type,gene_set=all_genes,exposure="Medication",exp_group="exclusive_glp1",ref_group="neither",enrichment="Yes",top_gsea = 30)
}
#GLP-1 vs. Both
so_sub <- subset(so_diab,group2=="exclusive_glp1" | group2 =="both")
so_sub$group2 <- factor(so_sub$group2)
#Define genes
all_genes <- rownames(so_sub)
so_sub$Medication <- so_sub$group2
all_cell_types <- unique(so_sub$celltype2)[-21]
for (cell_type in all_cell_types) {
  degs_fxn(so=so_sub,cell=cell_type,gene_set=all_genes,exposure="Medication",exp_group="both",ref_group="exclusive_glp1",enrichment="Yes",top_gsea = 30)
}

#GLP-1 vs. SGLT2
so_sub <- subset(so_diab,group2=="exclusive_glp1" | group2 =="exclusive_sglt2")
so_sub$group2 <- factor(so_sub$group2)
#Define genes
all_genes <- rownames(so_sub)
so_sub$Medication <- so_sub$group2
all_cell_types <- unique(so_sub$celltype2)[-23]
for (cell_type in all_cell_types) {
  degs_fxn(so=so_sub,cell=cell_type,gene_set=all_genes,exposure="Medication",exp_group="exclusive_glp1",ref_group="exclusive_sglt2",enrichment="Yes",top_gsea = 30)
}

#SGLT2 vs. None
so_sub <- subset(so_diab,group2=="exclusive_sglt2" | group2 =="neither")
so_sub$group2 <- factor(so_sub$group2)
#Define genes
all_genes <- rownames(so_sub)
so_sub$Medication <- so_sub$group2
all_cell_types <- unique(so_sub$celltype2)
for (cell_type in all_cell_types) {
  degs_fxn(so=so_sub,cell=cell_type,gene_set=all_genes,exposure="Medication",exp_group="exclusive_sglt2",ref_group="neither",enrichment="Yes",top_gsea = 30)
}

#Both vs. SGLT2
so_sub <- subset(so_diab,group2=="both" | group2 =="exclusive_sglt2")
so_sub$group2 <- factor(so_sub$group2)
#Define genes
all_genes <- rownames(so_sub)
so_sub$Medication <- so_sub$group2
all_cell_types <- unique(so_sub$celltype2)
for (cell_type in all_cell_types) {
  degs_fxn(so=so_sub,cell=cell_type,gene_set=all_genes,exposure="Medication",exp_group="both",ref_group="exclusive_sglt2",enrichment="Yes",top_gsea = 30)
}

#Both vs. None
so_sub <- subset(so_diab,group2=="both" | group2 =="neither")
so_sub$group2 <- factor(so_sub$group2)
#Define genes
all_genes <- rownames(so_sub)
so_sub$Medication <- so_sub$group2
all_cell_types <- unique(so_sub$celltype2)
for (cell_type in all_cell_types) {
  degs_fxn(so=so_sub,cell=cell_type,gene_set=all_genes,exposure="Medication",exp_group="both",ref_group="neither",enrichment="Yes",top_gsea = 30)
}
```
##c. Senescence Marker Genes 
###i. All Cells
```{r}
#Among Type 2: No meds vs. GLP-1 Exclusive
##SGLT2i only -/+ #5 No, 0 yes
#GLP-1 only +/- # 2 Yes, 3 No
#Both +/+ #1 Both, 4 not both
#Neither-/- #3 neither, 2 on someting

all_genes <- rownames(so_sens_all)
#Diabetes Groups
#Type 2 vs. Lean Control
so_subgroup <- subset(so_sens_all,group=="Type 2 Diabetes" | group=="Lean Control")
so_subgroup$group <- factor(so_subgroup$group)
so_subgroup$Type_2_Diabetes <- so_subgroup$group 
degs_fxn_sens(so=so_subgroup,cell=NULL,gene_set=all_genes,exposure="Type_2_Diabetes",exp_group="Type 2 Diabetes",ref_group="Lean Control",enrichment="Yes",top_gsea = 30)

#Create diab so
so_diab <- subset(so_sens_all,group=="Type 2 Diabetes")
#GLP-1 vs. None
so_sub <- subset(so_diab,group2=="exclusive_glp1" | group2=="neither")
so_sub$group2 <- factor(so_sub$group2)
#Define genes
all_genes <- rownames(so_sub)
#Rename variable
so_sub$Medication <- so_sub$group2
degs_fxn_sens(so=so_sub,cell=NULL,gene_set=all_genes,exposure="Medication",exp_group="exclusive_glp1",ref_group="neither",enrichment="Yes",top_gsea = 30)
#GLP-1 vs. SGLT2
so_sub <- subset(so_diab,group2=="exclusive_glp1" | group2=="exclusive_sglt2")
so_sub$group2 <- factor(so_sub$group2)
#Define genes
all_genes <- rownames(so_sub)
#Rename variable
so_sub$Medication <- so_sub$group2
degs_fxn_sens(so=so_sub,cell=NULL,gene_set=all_genes,exposure="Medication",exp_group="exclusive_glp1",ref_group="exclusive_sglt2",enrichment="Yes",top_gsea = 30)

#GLP-1 vs. Both
so_sub <- subset(so_diab,group2=="exclusive_glp1" | group2=="both")
so_sub$group2 <- factor(so_sub$group2)
#Define genes
all_genes <- rownames(so_sub)
#Rename variable
so_sub$Medication <- so_sub$group2
degs_fxn_sens(so=so_sub,cell=NULL,gene_set=all_genes,exposure="Medication",exp_group="both",ref_group="exclusive_glp1",enrichment="Yes",top_gsea = 30)

#SGLT2 vs. None 
so_sub <- subset(so_diab,group2=="exclusive_sglt2" | group2=="neither")
so_sub$group2 <- factor(so_sub$group2)
#Define genes
all_genes <- rownames(so_sub)
#Rename variable
so_sub$Medication <- so_sub$group2
degs_fxn_sens(so=so_sub,cell=NULL,gene_set=all_genes,exposure="Medication",exp_group="exclusive_sglt2",ref_group="neither",enrichment="Yes",top_gsea = 30)

#SGLT2 vs. Both
so_sub <- subset(so_diab,group2=="exclusive_sglt2" | group2=="both")
so_sub$group2 <- factor(so_sub$group2)
#Define genes
all_genes <- rownames(so_sub)
#Rename variable
so_sub$Medication <- so_sub$group2
degs_fxn_sens(so=so_sub,cell=NULL,gene_set=all_genes,exposure="Medication",exp_group="both",ref_group="exclusive_sglt2",enrichment="Yes",top_gsea = 30)

#Both vs. None
so_sub <- subset(so_diab,group2=="neither" | group2=="both")
so_sub$group2 <- factor(so_sub$group2)
#Define genes
all_genes <- rownames(so_sub)
#Rename variable
so_sub$Medication <- so_sub$group2
degs_fxn_sens(so=so_sub,cell=NULL,gene_set=all_genes,exposure="Medication",exp_group="both",ref_group="neither",enrichment="Yes",top_gsea = 30)

```
###ii. Specific Cells
```{r}
#Major Cell Types
#unique(so_sens_all$celltype_rpca)
so_sens_all$celltype_rpca <- as.character(so_sens_all$celltype_rpca)
so_sens_all$celltype2 <- ifelse(grepl("PT-",so_sens_all$celltype_rpca),"PT",
                                ifelse(grepl("TAL-",so_sens_all$celltype_rpca),"TAL",
                                       ifelse(grepl("PC-",so_sens_all$celltype_rpca),"PC",
                                              ifelse(grepl("IC-",so_sens_all$celltype_rpca) & !grepl("lowQuality",so_sens_all$celltype_rpca),"IC",
                                                     ifelse(grepl("DTL-",so_sens_all$celltype_rpca),"DTL",so_sens_all$celltype_rpca)))))
unique(so_sens_all$celltype2)
all_cell_types <- unique(so_sens_all$celltype2)
all_genes <- rownames(so_sens_all)

#Create diab so 
so_diab <- subset(so_sens_all,group=="Type 2 Diabetes")

#GLP-1 vs. None
so_sub <- subset(so_diab,group2=="exclusive_glp1" | group2 =="neither")
so_sub$group2 <- factor(so_sub$group2)
#Define genes
all_genes <- rownames(so_sub)
so_sub$Medication <- so_sub$group2
all_cell_types <- unique(so_sub$celltype2)[-20]
for (cell_type in all_cell_types) {
degs_fxn_sens(so=so_sub,cell=cell_type,gene_set=all_genes,exposure="Medication",exp_group="exclusive_glp1",ref_group="neither",enrichment="Yes",top_gsea = 30)
}
#GLP-1 vs. Both
so_sub <- subset(so_diab,group2=="exclusive_glp1" | group2 =="both")
so_sub$group2 <- factor(so_sub$group2)
#Define genes
all_genes <- rownames(so_sub)
so_sub$Medication <- so_sub$group2
all_cell_types <- unique(so_sub$celltype2)[-21]
for (cell_type in all_cell_types) {
degs_fxn_sens(so=so_sub,cell=cell_type,gene_set=all_genes,exposure="Medication",exp_group="both",ref_group="exclusive_glp1",enrichment="Yes",top_gsea = 30)
}

#GLP-1 vs. SGLT2
so_sub <- subset(so_diab,group2=="exclusive_glp1" | group2 =="exclusive_sglt2")
so_sub$group2 <- factor(so_sub$group2)
#Define genes
all_genes <- rownames(so_sub)
so_sub$Medication <- so_sub$group2
all_cell_types <- unique(so_sub$celltype2)[-23]
for (cell_type in all_cell_types) {
degs_fxn_sens(so=so_sub,cell=cell_type,gene_set=all_genes,exposure="Medication",exp_group="exclusive_glp1",ref_group="exclusive_sglt2",enrichment="Yes",top_gsea = 30)
}

#SGLT2 vs. None
so_sub <- subset(so_diab,group2=="exclusive_sglt2" | group2 =="neither")
so_sub$group2 <- factor(so_sub$group2)
#Define genes
all_genes <- rownames(so_sub)
so_sub$Medication <- so_sub$group2
all_cell_types <- unique(so_sub$celltype2)
for (cell_type in all_cell_types) {
degs_fxn_sens(so=so_sub,cell=cell_type,gene_set=all_genes,exposure="Medication",exp_group="exclusive_sglt2",ref_group="neither",enrichment="Yes",top_gsea = 30)
}

#Both vs. SGLT2
so_sub <- subset(so_diab,group2=="both" | group2 =="exclusive_sglt2")
so_sub$group2 <- factor(so_sub$group2)
#Define genes
all_genes <- rownames(so_sub)
so_sub$Medication <- so_sub$group2
all_cell_types <- unique(so_sub$celltype2)
for (cell_type in all_cell_types) {
degs_fxn_sens(so=so_sub,cell=cell_type,gene_set=all_genes,exposure="Medication",exp_group="both",ref_group="exclusive_sglt2",enrichment="Yes",top_gsea = 30)
}

#Both vs. None
so_sub <- subset(so_diab,group2=="both" | group2 =="neither")
so_sub$group2 <- factor(so_sub$group2)
#Define genes
all_genes <- rownames(so_sub)
so_sub$Medication <- so_sub$group2
all_cell_types <- unique(so_sub$celltype2)
for (cell_type in all_cell_types) {
degs_fxn_sens(so=so_sub,cell=cell_type,gene_set=all_genes,exposure="Medication",exp_group="both",ref_group="neither",enrichment="Yes",top_gsea = 30)
}
```

#5. MAST Associations
##a. All Cell Types
```{r}
<<<<<<< Updated upstream
exposure_vars <- c("ast","alt")
for (exp in exposure_vars){
  mast_fxn(so=so_liver_sn,cell=NULL,exposure=exp,covariate="diagnosis_of_diabetes",gene_set=rownames(so_liver_sn),batch_size=100,exp_group=NULL,ref_group=NULL)
}
=======
#Type 2 vs. LC
so_subset <- subset(so_kidney_sc,group=="Type 2 Diabetes" | group=="Lean Control")
so_subset$Diabetes <- so_subset$group
so_subset$Diabetes <- factor(so_subset$Diabetes)
mast_fxn(so=so_subset,cell=NULL,exposure="Diabetes",gene_set=rownames(so_subset),batch_size=1000,exp_group="Type 2 Diabetes",ref_group="Lean Control")

#T2 + OC vs. LC
so_subset <- subset(so_kidney_sc,group=="Type 2 Diabetes" | group=="Lean Control" | group=="Obese Control")
so_subset$Diabetes <- ifelse(so_subset$group=="Type 2 Diabetes" | so_subset$group=="Obese Control","T2D + OC",so_subset$group)
so_subset$Diabetes <- factor(so_subset$Diabetes)
mast_fxn(so=so_subset,cell=NULL,exposure=exp,covariate="Diabetes",gene_set=rownames(so_subset),batch_size=1000,exp_group="T2D + OC",ref_group="Lean Control")

#Medication Groups in Type 2
so_diab <- subset(so_kidney_sc,group=="Type 2 Diabetes")

#GLP-1 vs. None
so_subset <- subset(so_diab, group2=="exclusive_glp1" | group2=="neither")
so_subset$Medication <- so_subset$group2
so_subset$Medication <- factor(so_subset$Medication)
mast_fxn(so=so_subset,cell=NULL,exposure="Medication",gene_set=rownames(so_subset),batch_size=1000,exp_group="exclusive_glp1",ref_group="neither")

#GLP-1 vs. SLGT2
so_subset <- subset(so_diab, group2=="exclusive_glp1" | group2=="exclusive_sglt2")
so_subset$Medication <- so_subset$group2
so_subset$Medication <- factor(so_subset$Medication)
mast_fxn(so=so_subset,cell=NULL,exposure="Medication",gene_set=rownames(so_subset),batch_size=1000,exp_group="exclusive_glp1",ref_group="exclusive_sglt2")

#GLP-1 vs. Both
so_subset <- subset(so_diab, group2=="exclusive_glp1" | group2=="both")
so_subset$Medication <- so_subset$group2
so_subset$Medication <- factor(so_subset$Medication)
mast_fxn(so=so_subset,cell=NULL,exposure="Medication",gene_set=rownames(so_subset),batch_size=1000,exp_group="exclusive_glp1",ref_group="both")

#SLGT2 vs. None
so_subset <- subset(so_diab, group2=="exclusive_sglt2" | group2=="neither")
so_subset$Medication <- so_subset$group2
so_subset$Medication <- factor(so_subset$Medication)
mast_fxn(so=so_subset,cell=NULL,exposure="Medication",gene_set=rownames(so_subset),batch_size=1000,exp_group="exclusive_sglt2",ref_group="neither")

#SGLT2 vs. Both
so_subset <- subset(so_diab, group2=="exclusive_sglt2" | group2=="both")
so_subset$Medication <- so_subset$group2
so_subset$Medication <- factor(so_subset$Medication)
mast_fxn(so=so_subset,cell=NULL,exposure="Medication",gene_set=rownames(so_subset),batch_size=1000,exp_group="exclusive_sglt2",ref_group="both")

#Both vs. None
so_subset <- subset(so_diab, group2=="exclusive_glp1" | group2=="both")
so_subset$Medication <- so_subset$group2
so_subset$Medication <- factor(so_subset$Medication)
mast_fxn(so=so_subset,cell=NULL,exposure="Medication",gene_set=rownames(so_subset),batch_size=1000,exp_group="exclusive_glp1",ref_group="both")

>>>>>>> Stashed changes
```

##b. Specific Cell Types
```{r}
Idents(so_liver_sn) <- so_liver_sn$celltype
de.markers(so_liver_sn, genes, "diagnosis_of_diabetes", id2 = "No", id1 = "Yes", NULL, "_top")

#Gene set enrichment analysis
sce_sn_hep <- as.SingleCellExperiment(so_liver_sn_hep)
rm(so_liver_sn_hep)
## C2 category is according to canonical pathways: https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4707969/pdf/nihms-743907.pdf
geneSets <- msigdbr(species = "Homo sapiens", category = "C2", subcategory = "CP:KEGG")
### filter background to only include genes that we assessed
geneSets$gene_symbol <- toupper(geneSets$gene_symbol)
geneSets <- geneSets[geneSets$gene_symbol %in% names(sce_sn_hep),]
m_list <- geneSets %>% split(x = .$gene_symbol, f = .$gs_name)
stats <- m$p_val_adj
names(stats) <- rownames(m)
eaRes <- fgsea(pathways = m_list, stats = na.omit(stats))
#ooEA <- order(eaRes$pval, decreasing = FALSE)
#kable(head(eaRes[ooEA, 1:7], n = 20))
# Convert the leadingEdge column to comma-separated strings
eaRes$leadingEdge <- sapply(eaRes$leadingEdge, function(x) paste(x, collapse = ", "))
gc()
# Save to CSV
write.csv(eaRes, fs::path(dir.results,"GSEA_Diabetes_Hepatocytes.csv"))
gc()

#IPA Pathways 
test_ipa <- readxl::read_xls("/Users/choiyej/Dropbox/PANTHER/IPA results/tanner_stage_12_345_rh_de_table_export.xls", skip = 1)
colnames(test_ipa) <- c("pathway", "neglog_p", "ratio", "zscore", "molecules")

# Volcano plot
sig_cutoff = -log(0.05, base = 10)

test_ipa <- test_ipa %>%
  dplyr::mutate(direction = case_when(zscore < 0 & neglog_p > sig_cutoff ~ "Negative", 
                                      zscore > 0 & neglog_p > sig_cutoff ~ "Positive",
                                      zscore == 0 & neglog_p > sig_cutoff ~ "Unknown",
                                      T ~ "NS")) %>%
  group_by(direction) %>%
  dplyr::mutate(count = row_number(),
                label = case_when(direction != "NS" ~ paste0(str_sub(direction, 1,1), count))) %>%
  ungroup()

test_ipa$direction <- factor(test_ipa$direction, levels = c("Negative", "Positive", "Unknown", "NS"))

ipa.plot <- test_ipa %>% filter(neglog_p > 0) %>%
  ggplot(aes(x = zscore, y = neglog_p)) + 
  geom_point(aes(color = direction), alpha = 0.5, size = 2) +
  geom_hline(aes(yintercept = sig_cutoff), color = "red", linetype = "dashed") +
  labs(x = "Z-Score",
       y = "-log(p-value)",
       color = "Direction") + 
  scale_color_manual(values = c("#bf0603", "#457b9d",
                                "#f6bd60", "#dad7cd")) +
  ggrepel::geom_label_repel(aes(label = label, color = direction),
                            label.size = 0.15,
                            max.overlaps = 100, 
                            min.segment.length = 0.01) +
  theme_bw()

# Table
ipa.table.neg <- test_ipa %>%
  dplyr::filter(direction == "Negative") %>%
  dplyr::select(label, pathway) 
ipa.table.pos <- test_ipa %>%
  dplyr::filter(direction == "Positive") %>%
  dplyr::select(label, pathway) 
ipa.table.unk <- test_ipa %>%
  dplyr::filter(direction == "Unknown") %>%
  dplyr::select(label, pathway) 

cbind.fill <- function(...) {
  nm <- list(...) 
  nm <- lapply(nm, as.matrix)
  n <- max(sapply(nm, nrow)) 
  do.call(cbind, lapply(nm, function(x) {
    filled_matrix <- ifelse(is.na(x), "", x)
    rbind(filled_matrix, matrix("", n - nrow(filled_matrix), ncol(filled_matrix)))
  }))
}

ipa.table.comb <- cbind.fill(ipa.table.neg,ipa.table.pos,ipa.table.unk)
gg.ipa.table <- ggtexttable(ipa.table.comb, rows = NULL,
                            cols = c("Label", "Pathway", "Label", "Pathway","Label", "Pathway"),
                            theme = ttheme("blank",
                                           tbody.style = tbody_style(fill = "white", 
                                                                     size = 8.5, 
                                                                     hjust = 0,
                                                                     x = 0.1))) %>%
  tab_add_hline(at.row = 1:2, row.side = "top", linewidth = 0.8)



ggarrange(ipa.plot, gg.ipa.table,
          ncol = 2, heights = c(1,1), widths = c(0.5,1),
          legend = "top")


```


#6. Visualize Results
##a. Specific Cell Types
```{r}
so_kidney_sc$celltype_rpca <- as.character(so_kidney_sc$celltype_rpca)
so_kidney_sc$celltype2 <- ifelse(grepl("PT-",so_kidney_sc$celltype_rpca),"PT",
                                 ifelse(grepl("TAL-",so_kidney_sc$celltype_rpca),"TAL",
                                        ifelse(grepl("PC-",so_kidney_sc$celltype_rpca),"PC",
                                               ifelse(grepl("IC-",so_kidney_sc$celltype_rpca) & !grepl("lowQuality",so_kidney_sc$celltype_rpca),"IC",
                                                      ifelse(grepl("DTL-",so_kidney_sc$celltype_rpca),"DTL",so_kidney_sc$celltype_rpca)))))
unique(so_kidney_sc$celltype2)
cell_type <- c("IC","TAL","PC","MON","VSMC/MC/FIB","MAC","T","DTL","PT","IC-A_lowQuality","PEC","DCT","EC-GC","CNT","EC-PTC","TAL_highUMI","NKT/NKC","ATL","B" ,"POD","EC-AEA" ,"EC-LYM")
#"PT_lowQuality"
cell_type <- str_replace_all(cell_type,"/","_")
exposure <- "Medication"
for (cell in cell_type){
  pdf(fs::path(dir.results,paste0("venn_results_",cell,"_",exposure,".pdf")),width=20,height=15)
  visualize_function(exposure,cell)
  dev.off()
}
```

##b. All Cell Types
```{r}
exposure = "Medication"
exp_group = "exclusive_glp1"
ref_group = "neither"

cell_types <- c("TAL","PT","PC","DTL","IC","ALT","B","CNT","DCT","EC-GC","EC-PTC","MAC",
                "MON","NKT_NKC","PEC","TAL_highUMI","T","VSMC_MC_FIB","POD")

# Function to extract cell type from file name using a list of cell types
extract_cell_type <- function(filename, cell_types) {
  for (cell_type in cell_types) {
    if (str_detect(filename, cell_type)) {
      return(cell_type)
    }
  }
  return(NA)  # Return NA if no cell type is found
}

visualize_cell <- function(exposure,exp_group,ref_group) {
  
  # List all files in the results directory
  all_files <- dir_ls(path = dir.results, glob = "*.xlsx")  # List only .xlsx files
  
  # Filter the files that contain both the exposure and the word "Bulk" in the file name
  matching_files <- all_files[str_detect(all_files, exposure) & str_detect(all_files, exp_group) & str_detect(all_files, ref_group) & !str_detect(all_files,"Bulk") & !str_detect(all_files,"Senescence")]
  
  # Check if matching files exist for the current cell type and exposure
  if (length(matching_files) > 0) {
    # Loop through each matching file
    for (file in matching_files) {
      # Extract cell type from the filename
      cell_type <- extract_cell_type(basename(file), cell_types)
      
      # Dynamically assign the data to a variable named after the cell type
      if (!is.na(cell_type)) {
        assign(make.names(cell_type), read.xlsx(file), envir = .GlobalEnv)
      }
    }
  }
  
  # Initialize a list to store data
  all_data <- list()
  # Loop through each file and process
  for (file in matching_files) {
    # Read the file (assuming CSV format, modify if needed)
    data <- read.xlsx(file)
    
    # Extract cell type from the filename
    cell_type <- extract_cell_type(basename(file), cell_types)
    
    # Add the cell type as a column
    data$Cell_Type <- cell_type
    
    # Append to the list
    all_data[[length(all_data) + 1]] <- data
  }
  
  # Combine all data into a single dataframe
  combined_data <- bind_rows(all_data)
  combined_data$sig <- ifelse(combined_data$p_val_adj<0.05,"Significant","Non-Significant")
  combined_data$sig <- factor(combined_data$sig)
  
# Filter significant data
significant_data <- combined_data %>%
  filter(sig == "Significant")

# Get top 3 positive and negative genes for each Cell_Type
top_positive_per_cell_type <- significant_data %>%
  filter(avg_log2FC > 0) %>%  # Only positive values
  group_by(Cell_Type) %>%
  arrange(desc(avg_log2FC)) %>%
  slice_head(n = 3) %>%
  mutate(pos_or_neg = "Positive")

top_negative_per_cell_type <- significant_data %>%
  filter(avg_log2FC < 0) %>%  # Only negative values
  group_by(Cell_Type) %>%
  arrange(avg_log2FC) %>%  # Sort in ascending order for negative values
  slice_head(n = 3) %>%
  mutate(pos_or_neg = "Negative")

# Combine both positive and negative data
top_genes_per_cell_type <- bind_rows(top_positive_per_cell_type, top_negative_per_cell_type)

# Create the plot
p <- ggplot(combined_data, aes(x = Cell_Type, y = avg_log2FC)) +
  
  # Dot plot with jittered dots for individual data points, colored by significance
  geom_jitter(aes(color = sig), width = 0.3, alpha = 1, size = 0.5) +
  
  # Set custom colors for the dots based on significance
  scale_color_manual(values = c("Significant" = "red", "Non-Significant" = "lightblue1")) +
  
  # Add labels for the top positive and negative genes for each Cell_Type
  geom_text(data = top_genes_per_cell_type, aes(label = Gene), 
            color = "black", size = 3, 
            vjust = -0.5, hjust = 0.5, angle = 45) +  # Adjust positioning of labels
  
  # Adjust theme and labels
  theme_minimal() +
  labs(
    title = paste0("Top Genes. by ",exposure," Group (",exp_group," vs. ",ref_group,") by Cell Type"),
    x = "Cell Type",
    y = "Log2 Fold Change"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

pdf(fs::path(dir.results,paste0(exposure,"_",exp_group,"_",ref_group,".pdf")),width=20,height=20)
plot(p)
dev.off()
}

# Filter significant data
significant_data <- combined_data %>%
  filter(sig == "Significant")

# Get top 3 positive and negative genes for each Cell_Type
top_positive_per_cell_type <- significant_data %>%
  filter(avg_log2FC > 0) %>%  # Only positive values
  group_by(Cell_Type) %>%
  arrange(desc(avg_log2FC)) %>%
  slice_head(n = 3) %>%
  mutate(pos_or_neg = "Positive")

top_negative_per_cell_type <- significant_data %>%
  filter(avg_log2FC < 0) %>%  # Only negative values
  group_by(Cell_Type) %>%
  arrange(avg_log2FC) %>%  # Sort in ascending order for negative values
  slice_head(n = 3) %>%
  mutate(pos_or_neg = "Negative")

# Combine both positive and negative data
top_genes_per_cell_type <- bind_rows(top_positive_per_cell_type, top_negative_per_cell_type)

# Create the plot
ggplot(combined_data, aes(x = Cell_Type, y = avg_log2FC)) +
  
  # Dot plot with jittered dots for individual data points, colored by significance
  geom_jitter(aes(color = sig), width = 0.3, alpha = 1, size = 0.5) +
  
  # Set custom colors for the dots based on significance
  scale_color_manual(values = c("Significant" = "red", "Non-Significant" = "lightblue1")) +
  
  # Add labels for the top positive and negative genes for each Cell_Type
  geom_text(data = top_genes_per_cell_type, aes(label = Gene), 
            color = "black", size = 3, 
            vjust = -0.5, hjust = 0.5, angle = 45) +  # Adjust positioning of labels
  
  # Adjust theme and labels
  theme_minimal() +
  labs(
    title = "Top 3 Positive and Negative Log2 Fold Change by Cell Type",
    x = "Cell Type",
    y = "Log2 Fold Change"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

# Create the plot
ggplot(combined_data, aes(x = Cell_Type, y = avg_log2FC)) +
  
  # Dot plot with jittered dots for individual data points, colored by significance
  geom_jitter(aes(color = sig), width = 0.3, alpha = 1, size = 0.5) +
  
  # Set custom colors for the dots based on significance
  scale_color_manual(values = c("Significant" = "red", "Non-Significant" = "gray")) +
  
  # Add lines connecting the labels to the points
  geom_segment(data = top_genes_per_cell_type, 
               aes(x = Cell_Type, xend = Cell_Type, 
                   y = avg_log2FC, yend = avg_log2FC), 
               color = "black", size = 0.5, alpha = 0.7) +
  
  # Add labels with smaller text and adjust their positions
  geom_text(data = top_genes_per_cell_type, 
            aes(label = Gene), 
            color = "black", size = 2.5, 
            vjust = -0.5, hjust = 0.5, angle = 45) +  # Adjust positioning and size of text
  
  # Adjust theme and labels
  theme_minimal() +
  labs(
    title = paste0("Top","by Cell Type"),
    x = "Cell Type",
    y = "Log2 Fold Change"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```


#7. TODAY 
```{r}
#Load in AST and ALT in today 
today <- read.csv(fs::path(dir.dat,"TODAY subaward","Clinical data","TODAY","CBL.csv")) # length(unique(today$releaseid)) - 699 ids
today_M0 <- today %>%
  filter(mvisit=="M00")
#dplyr::select(c("releaseid","ALT","AST"))
#Load in Proteomics in Today
prot <- readRDS(fs::path(dir.dat,"Liver project","TODAY_proteomics.rds"))
prot <- prot[c(7326,which(grepl("seq.",colnames(prot))))] #length(unique(prot$releaseid)) - 376 unique ids
proteins <- colnames(prot)[which(grepl("seq.",colnames(prot)))]

#rm(today,today_M0,prot)

#Determine which proteins correlate with genes associated with alt and ast
#AST
ast_genes <- read.xlsx(fs::path(dir.dat,"Liver project","Results","Bulk_Results_for_Ast Di (High vs. Low).xlsx"))
ast_genes_pos <- ast_genes %>%
  filter(p_val_adj<0.05) %>%
  filter(avg_log2FC>0) %>%
  arrange(desc(avg_log2FC)) %>%
  dplyr::slice(1:200) 
ast_genes_pos <- ast_genes_pos$Gene
ast_genes_neg <- ast_genes %>%
  filter(p_val_adj<0.05) %>%
  filter(avg_log2FC<0) %>%
  arrange(avg_log2FC) %>%
  dplyr::slice(1:200) 
ast_genes_neg <- ast_genes_neg$Gene
ast_genes <- c(ast_genes_pos,ast_genes_neg)

#Key Protein Gene
key <- readRDS(fs::path(dir.dat,"Liver project","gene_prot.RDS"))
sig_proteins_ast <- key$AptName[which(ast_genes %in% key$EntrezGeneSymbol)]
prot_ast <- prot[c("releaseid",sig_proteins_ast)]
dat_ast <- tidylog::right_join(today_M0,prot_ast,by="releaseid")

#Create loop for linear regression between proteins & liver enzymes
results <- data.frame()
for (x in sig_proteins_ast) {
  M0 <- as.formula(paste0("AST ~ ",x))
  M1 <- lm(M0,data=dat_ast)
  beta <- summary(M1)$coef[2,1]
  pval <- summary(M1)$coef[2,4]
  results_full <- data.frame(Protein=x,Estimate=beta,Pvalue=pval)
  results <- rbind(results,results_full)
}
results_fdr <- results %>%
  mutate(fdr = p.adjust(Pvalue,method="fdr"))
results_sig <- results_fdr %>%
  filter(fdr<0.05)

#ALT
alt_genes <- read.xlsx(fs::path(dir.dat,"Liver project","Results","Bulk_Results_for_Alt Di (High vs. Low).xlsx"))
alt_genes_pos <- alt_genes %>%
  filter(p_val_adj<0.05) %>%
  filter(avg_log2FC>0) %>%
  arrange(desc(avg_log2FC)) %>%
  dplyr::slice(1:200) 
alt_genes_pos <- alt_genes_pos$Gene
alt_genes_neg <- alt_genes %>%
  filter(p_val_adj<0.05) %>%
  filter(avg_log2FC<0) %>%
  arrange(avg_log2FC) %>%
  dplyr::slice(1:200) 
alt_genes_neg <- alt_genes_neg$Gene
alt_genes <- c(alt_genes_pos,alt_genes_neg)
sig_proteins_alt <- key$AptName[which(alt_genes %in% key$EntrezGeneSymbol)]
prot_alt <- prot[c("releaseid",sig_proteins_ast)]
dat_alt <- tidylog::right_join(today_M0,prot_alt,by="releaseid")

#Create loop for linear regression between proteins & liver enzymes
results2 <- data.frame()
for (x in sig_proteins_alt) {
  M0 <- as.formula(paste0("ALT ~ ",x))
  M1 <- lm(M0,data=dat_alt)
  beta <- summary(M1)$coef[2,1]
  pval <- summary(M1)$coef[2,4]
  results_full2 <- data.frame(Protein=x,Estimate=beta,Pvalue=pval)
  results2 <- rbind(results2,results_full2)
}
results_fdr2 <- results2 %>%
  mutate(fdr = p.adjust(Pvalue,method="fdr"))
results_sig2 <- results_fdr2 %>%
  filter(fdr<0.05)

results_sig2$Gene <- key$EntrezGeneSymbol[which(results_sig2$Protein %in% key$AptName)]
results_sig2$Gene %in% alt_genes

```

#8. TEEN-LABs
```{r}
#Proteomics teen labs 
load(fs::path(dir.dat,"Teen Labs","Data_Cleaned","analysis_dataset.RData"))

```

#9. Pseudotime Analysis - Hepatocytes
```{r}


```

#10. Targeted DEGs 
```{r}


```


