---
title: "TODAY Somalogic baseline analyses - elastic net Cox models"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include=FALSE}
library(SomaDataIO)
library(limma)
library(dplyr)
library(caret)
library(purrr)
library(multtest)
library(openxlsx)
library(tableone)
library(EnhancedVolcano)
library(knitr)
library(survival)
library(broom)
library(plsRcox)
library(forestmodel)
library(survC1)
library(naniar)
library(rms)
library(readxl)

knitr::opts_chunk$set(echo = FALSE,warning = FALSE)

knitr::opts_chunk$set(echo = FALSE)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "E:/Petter Bjornstad/TODAY subaward"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/TODAY subaward/"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/TODAY subaward"
}
knitr::opts_knit$set(root.dir = home_dir)

```

```{r, include=FALSE}
setwd(home_dir)
# load somalogic data, with QC samples already excluded
load("./Somalogic data raw/soma.Rdata")

# load analyte info
load("./Somalogic data raw/analytes.Rdata")

# load comorbidity data
load("./Clinical data/comorb.Rdata")

# load baseline risk factors
load("./Clinical data/TODAY/baserisk.Rdata")

# take only the baseline soma samples
# can we just take the earliest or is it possible someone would have a follow-up sample but not baseline?
# probably can take the first and then check the years to make sure compatible with TODAY
base <- soma %>% arrange(releaseid,Date.Drawn) %>% group_by(releaseid) %>% filter(row_number()==1)

# these 3 release IDs have first sample after end of recruitment, so they must be missing the baseline visit
base <- base %>% filter(!releaseid %in% c("65-85903","65-47984","65-25901"))

# merge in complication data
base <- merge(base, comorb, by="releaseid",all.x=T, all.y=F)

# merge in baseline risk factors
base <- merge(base, baserisk, by="releaseid",all.x=T, all.y=F)

# identify columns corresponding to proteins
#is_seq <- function(.x) grepl("^seq\\.[0-9]{4}", .x) # regex for analytes
is_seq <- function(.x) grepl("seq", .x)
seq <- is_seq(names(base))

# convert to numeric
base[,seq] <- apply(base[,seq],2,as.numeric)

# are there proteins with low variability?
no_var = caret::nearZeroVar(base[,seq])
# none

# log transform
base_log <- base %>% modify_if(is_seq(names(.)), log)

# scale by SD
base_log_scale <- base_log
predictors <- colnames(base_log[seq])
for (i in 1:length(predictors)) {
  base_log_scale[predictors[i]] <- base_log_scale[predictors[i]]/sd(base_log[,paste0(predictors[i])])
}

# filter to top 100 proteins by effect size
f = "/Volumes/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/TODAY subaward/Results/Linear and Cox models/TODAY somalogic Cox models scaled baseline adjusted.xlsx"

# MAC
mac_keep = read_excel(f,sheet = "MAC CPH")
mac_keep <- arrange(mac_keep,adj.p.value)
mac_drop <- mac_keep[101:nrow(mac_keep),"AptName"]
rownames(mac_drop) <- mac_drop$AptName
mac_drop <- as.data.frame(t(mac_drop))
mac <- base_log_scale %>% select(-colnames(mac_drop))
# MIC
mic_keep = read_excel(f,sheet = "MIC CPH")
mic_keep <- arrange(mic_keep,adj.p.value)
mic_drop <- mic_keep[101:nrow(mic_keep),"AptName"]
rownames(mic_drop) <- mic_drop$AptName
mic_drop <- as.data.frame(t(mic_drop))
mic <- base_log_scale %>% select(-colnames(mic_drop))
# HYPERFILTRATION
hyp_keep = read_excel(f,sheet = "HYP CPH")
hyp_keep <- arrange(hyp_keep,adj.p.value)
hyp_drop <- hyp_keep[101:nrow(hyp_keep),"AptName"]
rownames(hyp_drop) <- hyp_drop$AptName
hyp_drop <- as.data.frame(t(hyp_drop))
hyp <- base_log_scale %>% select(-colnames(hyp_drop))
# RAPID EGFR DECLINE
rapid_keep = read_excel(f,sheet = "RAPID CPH")
rapid_keep <- arrange(rapid_keep,adj.p.value)
rapid_drop <- rapid_keep[101:nrow(rapid_keep),"AptName"]
rownames(rapid_drop) <- rapid_drop$AptName
rapid_drop <- as.data.frame(t(rapid_drop))
rapid <- base_log_scale %>% select(-colnames(rapid_drop))
# HYPERTENSION
htn_keep = read_excel(f,sheet = "HTN CPH")
htn_keep <- arrange(htn_keep,adj.p.value)
htn_drop <- htn_keep[101:nrow(htn_keep),"AptName"]
rownames(htn_drop) <- htn_drop$AptName
htn_drop <- as.data.frame(t(htn_drop))
htn <- base_log_scale %>% select(-colnames(htn_drop))

set.seed(3654)
source("~/Documents/Github/CHCO-Code/Petter Bjornstad/TODAY proteomics metabolomics/Proteomics analyses/easy_elasticnet.R")
```

# Severe albuminuria 

## Elastic net with 10-fold cross-validation

```{r, include=FALSE, comment=""}
mac_nozero <- mac[!mac$DAYSTOMAC == 0,]
seq <- is_seq(names(mac))

mac_select = easy_elasticnet(data = mac_nozero,outcome = "MAC",
                           predictors = mac_nozero[,seq], out = "min.error",
                           model_type = "cox",time = "DAYSTOMAC", cv_method="kfold", folds = 10)
retained_vars = analytes[analytes$AptName %in% mac_select, c("Target","TargetFullName")]

form = as.formula(paste0("Surv(DAYSTOMAC, MAC)~",paste0(mac_select,collapse = "+")))
mod = coxph(formula = form,data = mac_nozero)
kable(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)
```

```{r, include=TRUE, comment=""}
kable(retained_vars)

summary(mod)
```

## Model with proteins selected by elastic net and covariates

```{r, include=FALSE, comment=""}
form = as.formula(paste0("Surv(DAYSTOMAC, MAC)~",paste0(mac_select,collapse = "+"),paste0("+HbA1c+log_trig+sbp+si_1_ins0")))
mod = coxph(formula = form,data = mac_nozero)
kable(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)
```

```{r, include=TRUE, comment=""}
summary(mod)
```

## Model with only covariates

```{r, include=FALSE, comment=""}
form = as.formula(paste0("Surv(DAYSTOMAC, MAC)~HbA1c+log_trig+sbp+si_1_ins0"))
mod = coxph(formula = form,data = mac_nozero)
kable(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)
```

```{r, include=TRUE, comment=""}
summary(mod)
```

### Model validation  - compare the 3 models

```{r, include=TRUE, comment=""}
#val1 <- validate(cph(formula = form,data = mac_nozero, x=TRUE, y=TRUE, surv = TRUE), 
#                 method="crossvalidation", B = 10, dxy=TRUE)

# Summarize the results
#val1
```

# Elevated albuminuria

## 10-fold cross-validation

```{r, include=FALSE, comment=""}
mic_nozero <- mic[!mic$DAYSTOMIC == 0,]
seq <- is_seq(names(mic))

mic_select = easy_elasticnet(data = mic_nozero,outcome = "MIC",
                           predictors = mic_nozero[,seq], out = "min.error",
                           model_type = "cox",time = "DAYSTOMIC", cv_method="kfold", folds = 10)
retained_vars = analytes[analytes$AptName %in% mic_select, c("Target","TargetFullName")]

form = as.formula(paste0("Surv(DAYSTOMIC, MIC)~",paste0(mic_select,collapse = "+")))
mod = coxph(formula = form,data = mic_nozero)
kable(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)
```

```{r, include=TRUE, comment=""}
kable(retained_vars)

summary(mod)
```

### Model validation

```{r, include=TRUE, comment=""}
val1 <- validate(cph(formula = form,data = mic_nozero, x=TRUE, y=TRUE, surv = TRUE), 
                 method="crossvalidation", B = 10, dxy=TRUE)

# Summarize the results
val1
```

# Hyperfiltration

## 10-fold cross-validation

```{r, include=FALSE, comment=""}
hyp_nozero <- hyp[!hyp$DAYSTOHYP == 0,]
seq <- is_seq(names(hyp))

hyp_select = easy_elasticnet(data = hyp_nozero,outcome = "HYP",
                           predictors = hyp_nozero[,seq], out = "min.error",
                           model_type = "cox",time = "DAYSTOHYP", cv_method="kfold", folds = 10)
retained_vars = analytes[analytes$AptName %in% hyp_select, c("Target","TargetFullName")]

form = as.formula(paste0("Surv(DAYSTOHYP, HYP)~",paste0(hyp_select,collapse = "+")))
mod = coxph(formula = form,data = hyp_nozero)
kable(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)
```

```{r, include=TRUE, comment=""}
kable(retained_vars)

summary(mod)
```

### Model validation

```{r, include=TRUE, comment=""}
val1 <- validate(cph(formula = form,data = hyp_nozero, x=TRUE, y=TRUE, surv = TRUE), 
                 method="crossvalidation", B = 10, dxy=TRUE)

# Summarize the results
val1
```

# Rapid eGFR decline

## 10-fold cross-validation

```{r, include=FALSE, comment=""}
rapid_nozero <- rapid[!rapid$DAYSTORAPID == 0,]
seq <- is_seq(names(rapid))

rapid_select = easy_elasticnet(data = rapid_nozero,outcome = "RAPID",
                           predictors = rapid_nozero[,seq], out = "min.error",
                           model_type = "cox",time = "DAYSTORAPID", cv_method="kfold", folds = 10)
retained_vars = analytes[analytes$AptName %in% rapid_select, c("Target","TargetFullName")]

form = as.formula(paste0("Surv(DAYSTORAPID, RAPID)~",paste0(rapid_select,collapse = "+")))
mod = coxph(formula = form,data = rapid_nozero)
kable(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)
```

```{r, include=TRUE, comment=""}
kable(retained_vars)

summary(mod)
```

### Model validation

```{r, include=TRUE, comment=""}
val1 <- validate(cph(formula = form,data = rapid_nozero, x=TRUE, y=TRUE, surv = TRUE), 
                 method="crossvalidation", B = 10, dxy=TRUE)

# Summarize the results
val1
```

# Hypertension

## 10-fold cross-validation

```{r, include=FALSE, comment=""}
htn_nozero <- htn[!htn$DAYSTOHTN == 0,]
seq <- is_seq(names(htn))

htn_select = easy_elasticnet(data = htn_nozero,outcome = "HTN",
                           predictors = htn_nozero[,seq], out = "min.error",
                           model_type = "cox",time = "DAYSTOHTN", cv_method="kfold", folds = 10)
retained_vars = analytes[analytes$AptName %in% htn_select, c("Target","TargetFullName")]

form = as.formula(paste0("Surv(DAYSTOHTN, HTN)~",paste0(htn_select,collapse = "+")))
mod = coxph(formula = form,data = htn_nozero)
kable(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)
```

```{r, include=TRUE, comment=""}
kable(retained_vars)

summary(mod)
```

### Model validation

```{r, include=TRUE, comment=""}
val1 <- validate(cph(formula = form,data = htn_nozero, x=TRUE, y=TRUE, surv = TRUE), 
                 method="crossvalidation", B = 10, dxy=TRUE)

# Summarize the results
val1
```



