---
title: "TODAY study - proteomics, hypertension, and OSA"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include=FALSE}
library(SomaDataIO)
library(limma)
library(dplyr)
library(caret)
library(purrr)
library(multtest)
library(openxlsx)
library(tableone)
library(EnhancedVolcano)
library(knitr)
library(survival)
library(broom)
library(psych)
library(corrplot)
library(pander)
library(tableone)
library(polycor)
library(piercer)

knitr::opts_chunk$set(echo = FALSE,warning = FALSE)

knitr::opts_chunk$set(echo = FALSE)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "E:/Petter Bjornstad/TODAY subaward"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/TODAY subaward/"
} else if (Sys.info()["sysname"] == "Darwin"){
 home_dir = "/Volumes/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/TODAY subaward/"
 #home_dir = "/Users/pylell/Documents/Downtime/TODAY"
  }
knitr::opts_knit$set(root.dir = home_dir)

```

```{r, include=FALSE}
setwd(home_dir)
# load somalogic data, with QC samples already excluded
load("./Somalogic data raw/soma.Rdata")

# load analyte info
load("./Somalogic data raw/analytes.Rdata")

# load comorbidity data
load("./Clinical data/comorb.Rdata")

# load baseline risk factors
load("./Clinical data/TODAY/baserisk.Rdata")

# take only the baseline soma samples
# can we just take the earliest or is it possible someone would have a follow-up sample but not baseline?
# probably can take the first and then check the years to make sure compatible with TODAY
base <- soma %>% arrange(releaseid,Date.Drawn) %>% group_by(releaseid) %>% filter(row_number()==1)

# these 3 release IDs have first sample after end of recruitment, so they must be missing the baseline visit
base <- base %>% filter(!releaseid %in% c("65-85903","65-47984","65-25901"))

# merge in complication data
base <- merge(base, comorb, by="releaseid",all.x=T, all.y=F)

# merge in baseline risk factors
base <- merge(base, baserisk, by="releaseid",all.x=T, all.y=F)

# identify columns corresponding to proteins
#is_seq <- function(.x) grepl("^seq\\.[0-9]{4}", .x) # regex for analytes
is_seq <- function(.x) grepl("seq", .x)
seq <- is_seq(names(base))

# convert to numeric
base[,seq] <- apply(base[,seq],2,as.numeric)

# are there proteins with low variability?
#no_var = caret::nearZeroVar(base[,seq])
# none

# log transform
base_log <- base %>% modify_if(is_seq(names(.)), log)

# scale by SD
base_log_scale <- base_log
predictors <- colnames(base_log[seq])
for (i in 1:length(predictors)) {
  base_log_scale[predictors[i]] <- base_log_scale[predictors[i]]/sd(base_log[,paste0(predictors[i])])
}

# TODAY2 TME
tme <- read.csv("./Clinical data/TODAY2/TME.csv")
tme <- tme %>% select(RELEASEID,TMETYPE,DAYSTOTME)
colnames(tme) <- c("releaseid","TMETYPE","DAYSTOTME")
tme <- tme %>% filter(TMETYPE==4)

# TODAY2 PSG
psg <- read.csv("./Clinical data/TODAY2/PSG.csv")
psg$releaseid <- psg$RELEASEID
psg$RELEASEID <- NULL
psg <- psg %>% select(releaseid,CENTRAL_AHI, OSA_SEVERITY_LEVEL)

# merge base_log (not scaled data) with PSG data
analytes_keep <- analytes %>% filter(AptName %in% c("seq.13408.23","seq.19563.3","seq.7957.2","seq.2999.6","seq.18896.23","seq.5109.24",
                                                    "seq.9211.19","seq.18339.207","seq.18376.19","seq.18935.14","seq.24948.79","seq.4330.4"))
key <- analytes_keep[,c("AptName","UniProt","Target","TargetFullName")]
base_log_keep <- base_log %>% select(c("releaseid",all_of(analytes_keep$AptName)))
base_log_keep <- merge(base_log_keep,psg,by="releaseid",all.x = T, all.y = F)
base_log_keep <- merge(base_log_keep,tme,by="releaseid",all.x = T, all.y = F)
# base_log_keep now has ALL participants, OSA variables, and proteins related to hypertension
base_log_keep$osa <- NA
base_log_keep$osa <- ifelse(base_log_keep$TMETYPE==4, 1, base_log_keep$osa)
base_log_keep$osa <- ifelse(base_log_keep$OSA_SEVERITY_LEVEL %in%  c("1","2","3"), 1, base_log_keep$osa)
base_log_keep$osa <- ifelse(is.na(base_log_keep$osa), 0, base_log_keep$osa)
base_log_keep$osa <- as.factor(base_log_keep$osa)
base_log_keep$OSA_SEVERITY_LEVEL <- as.factor(base_log_keep$OSA_SEVERITY_LEVEL)

# what dfs do I need
# analysis with all participants, OSA variables, and proteins, but need to create new var to indicate OSA Y/N combining TME and PSG
# need to have descriptive statistics on the df above (all ppts)
# df with only proteins and AHI for correlations
# need a smaller df for descriptive statistics for only the participants who have AHI

# df for descriptive statistics
corr <- base_log_keep[!is.na(base_log_keep$CENTRAL_AHI),]
corr_demo <- corr
corr <- corr %>% select(-releaseid)

t_osa <- corr

corr <- corr %>% select(-c(OSA_SEVERITY_LEVEL,osa,TMETYPE,DAYSTOTME))


```

# Results

## Key to protein names

```{r, include=TRUE}
kable(key)
```

## Participants in the PSG substudy

### Summary statistics 

Note: severity 1="Mild," 2="Moderate," 3="Severe"

```{r, include=FALSE}
t <- CreateTableOne(vars=c("OSA_SEVERITY_LEVEL","CENTRAL_AHI"),data=corr_demo)
t <- print(t, nonnorm="CENTRAL_AHI", minMax=T)
```

```{r, include=TRUE}
kable(t)
```

### Correlation between AHI and proteins

```{r, include=TRUE}
 # Correlations
  M <- corr.test(data.matrix(corr), normal = F, method = "spearman")
  r <- data.frame(round(M$r, 3))
#  r <- r %>% filter(rownames(r)=="CENTRAL_AHI")
#  r <- r %>% select(-releaseid)
#  p <- data.frame(round(M$p, 3))
#  p <- p %>% filter(rownames(p)=="CENTRAL_AHI")
#  p <- p %>% select(-releaseid)
  emphasize.strong.cells(which(M$p <= 0.05, arr.ind = TRUE))
  cat("\n")
  pander(r, caption = "Coefficients")
  cat("\n")
```

### Levels of proteins associated with hypertension, by OSA status

```{r, include=FALSE}
t <- CreateTableOne(vars=c("seq.13408.23","seq.19563.3","seq.7957.2","seq.2999.6","seq.18896.23","seq.5109.24",
                                                    "seq.9211.19","seq.18339.207","seq.18376.19","seq.18935.14","seq.24948.79","seq.4330.4"),data=t_osa, strata="OSA_SEVERITY_LEVEL")
t <- print(t, minMax=T, nonnorm=c("seq.13408.23","seq.19563.3","seq.7957.2","seq.2999.6","seq.18896.23","seq.5109.24",
                                                    "seq.9211.19","seq.18339.207","seq.18376.19","seq.18935.14","seq.24948.79","seq.4330.4"))
t <- print(t)
```

```{r, include=TRUE}
kable(t)
```

### Levels of proteins associated with hypertension, by OSA status, with all OSA groups combined

```{r, include=FALSE}
t <- CreateTableOne(vars=c("seq.13408.23","seq.19563.3","seq.7957.2","seq.2999.6","seq.18896.23","seq.5109.24",
                                                    "seq.9211.19","seq.18339.207","seq.18376.19","seq.18935.14","seq.24948.79","seq.4330.4"),data=t_osa, strata="osa")
t <- print(t, minMax=T, nonnorm=c("seq.13408.23","seq.19563.3","seq.7957.2","seq.2999.6","seq.18896.23","seq.5109.24",
                                                    "seq.9211.19","seq.18339.207","seq.18376.19","seq.18935.14","seq.24948.79","seq.4330.4"))
t <- print(t)
```

```{r, include=TRUE}
kable(t)
```

### Polyserial correlations of OSA severity and proteins

```{r, include=F}
HC <- hetcor(t_osa$seq.13408.23,t_osa$OSA_SEVERITY_LEVEL, ML = TRUE)
a <- ci.rpc(r = HC$correlations[2,1], se = HC$std.errors[2,1], rn = "seq.13408.23 & OSA_SEVERITY_LEVEL")
```

```{r, include=T}
kable(a[,1:6])
```

## All participants

### Summary statistics 

```{r, include=FALSE}
t <- CreateTableOne(vars="osa",data=base_log_keep)
t <- print(t, minMax=T)
```

```{r, include=TRUE}
kable(t)
```

### Levels of proteins associated with hypertension, by OSA status

```{r, include=FALSE}
t <- CreateTableOne(vars=c("seq.13408.23","seq.19563.3","seq.7957.2","seq.2999.6","seq.18896.23","seq.5109.24",
                                                    "seq.9211.19","seq.18339.207","seq.18376.19","seq.18935.14","seq.24948.79","seq.4330.4"),data=base_log_keep, strata="osa")
t <- print(t, minMax=T, nonnorm=c("seq.13408.23","seq.19563.3","seq.7957.2","seq.2999.6","seq.18896.23","seq.5109.24",
                                                    "seq.9211.19","seq.18339.207","seq.18376.19","seq.18935.14","seq.24948.79","seq.4330.4"))
```

```{r, include=TRUE}
kable(t)
```
