---
title: "TODAY study - proteomics, hypertension, and OSA"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include=FALSE}
library(SomaDataIO)
library(limma)
library(dplyr)
library(caret)
library(purrr)
library(multtest)
library(openxlsx)
library(tableone)
library(EnhancedVolcano)
library(knitr)
library(survival)
library(broom)
library(psych)
library(corrplot)
library(pander)
library(tableone)
library(polycor)
library(piercer)

knitr::opts_chunk$set(echo = FALSE,warning = FALSE)

knitr::opts_chunk$set(echo = FALSE)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "E:/Petter Bjornstad/TODAY subaward"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/TODAY subaward/"
} else if (Sys.info()["sysname"] == "Darwin"){
 home_dir = "/Volumes/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/TODAY subaward/"
 #home_dir = "/Users/pylell/Documents/Downtime/TODAY"
  }
knitr::opts_knit$set(root.dir = home_dir)

```

```{r, include=FALSE}
setwd(home_dir)
# load somalogic data, with QC samples already excluded
load("./Somalogic data raw/soma.Rdata")

# load analyte info
load("./Somalogic data raw/analytes.Rdata")

# load comorbidity data
load("./Clinical data/comorb.Rdata")

# load baseline risk factors
load("./Clinical data/TODAY/baserisk.Rdata")

# take only the baseline soma samples
# can we just take the earliest or is it possible someone would have a follow-up sample but not baseline?
# probably can take the first and then check the years to make sure compatible with TODAY
base <- soma %>% arrange(releaseid,Date.Drawn) %>% group_by(releaseid) %>% filter(row_number()==1)

# these 3 release IDs have first sample after end of recruitment, so they must be missing the baseline visit
base <- base %>% filter(!releaseid %in% c("65-85903","65-47984","65-25901"))

# merge in complication data
base <- merge(base, comorb, by="releaseid",all.x=T, all.y=F)

# merge in baseline risk factors
base <- merge(base, baserisk, by="releaseid",all.x=T, all.y=F)

# identify columns corresponding to proteins
#is_seq <- function(.x) grepl("^seq\\.[0-9]{4}", .x) # regex for analytes
is_seq <- function(.x) grepl("seq", .x)
seq <- is_seq(names(base))

# convert to numeric
base[,seq] <- apply(base[,seq],2,as.numeric)

# are there proteins with low variability?
#no_var = caret::nearZeroVar(base[,seq])
# none

# log transform
base_log <- base %>% modify_if(is_seq(names(.)), log)

# scale by SD
base_log_scale <- base_log
predictors <- colnames(base_log[seq])
for (i in 1:length(predictors)) {
  base_log_scale[predictors[i]] <- base_log_scale[predictors[i]]/sd(base_log[,paste0(predictors[i])])
}

# TODAY ECHO
today_echo <- read.csv("./Clinical data/TODAY/echo.csv")
# TODAY2 ECHO
today2_echo <- read.csv("./Clinical data/TODAY2/echo.csv")
# combine echo datasets - need to calculate deltas, but there are some variables that exist in timepoint 2 but not 1
echo <- bind_rows(today2_echo, today_echo)
echo <- echo %>% arrange(echo,days)
# make a wide dataset and calculate deltas
echonames <- colnames(echo)
pwvnames <- pwvnames[-1]
pwvnames <- pwvnames[-1]
echo_wide <- reshape(data=echo, direction="wide", idvar = "releaseid", timevar = "TIMEPOINT")

# TODAY2 PWV
pwv <- read.csv("./Clinical data/TODAY2/pwv.csv")
pwv$releaseid <- pwv$RELEASEID
pwv$RELEASEID <- NULL
# make a wide dataset and calculate deltas
pwvnames <- colnames(pwv)
pwvnames <- pwvnames[-1]
pwvnames <- pwvnames[-1]
pwv_wide <- reshape(data=pwv, direction="wide", idvar = "releaseid", timevar = "TIMEPOINT")
pwv_wide$delta_days <- pwv_wide$DAYS.PWV2 - pwv_wide$DAYS.PWV1
pwv_wide$delta_sdnn <- pwv_wide$SDNN.PWV2 - pwv_wide$SDNN.PWV1
pwv_wide$delta_pnn50 <- pwv_wide$PNN50.PWV2 - pwv_wide$PNN50.PWV1
pwv_wide$delta_rmssd <- pwv_wide$RMSSD.PWV2 - pwv_wide$RMSSD.PWV1
pwv_wide$delta_sdann <- pwv_wide$SDANN.PWV2 - pwv_wide$SDANN.PWV1
pwv_wide$delta_sdnn_index <- pwv_wide$SDNN_INDEX.PWV2 - pwv_wide$SDNN_INDEX.PWV1
pwv_wide$delta_lf_power_normalized <- pwv_wide$LF_POWER_NORMALISED.PWV2 - pwv_wide$LF_POWER_NORMALISED.PWV1
pwv_wide$delta_hf_power_normalized <- pwv_wide$HF_POWER_NORMALISED.PWV2 - pwv_wide$HF_POWER_NORMALISED.PWV1
pwv_wide$delta_lf_hf_ratio <- pwv_wide$LF_HF_RATIO.PWV2 - pwv_wide$LF_HF_RATIO.PWV1
pwv_wide$delta_pwvf <- pwv_wide$PWVF.PWV2 - pwv_wide$PWVF.PWV1
pwv_wide$delta_pwvr <- pwv_wide$PWVR.PWV2 - pwv_wide$PWVR.PWV1
pwv_wide$delta_pwvfft <- pwv_wide$PWVFFT.PWV2 - pwv_wide$PWVFFT.PWV1
pwv_wide$delta_pwvd <- pwv_wide$PWVD.PWV2 - pwv_wide$PWVD.PWV1
pwv_wide$delta_aix <- pwv_wide$AIX.PWV2 - pwv_wide$AIX.PWV1
pwv_wide$delta_c_sbp <- pwv_wide$C_SBP.PWV2 - pwv_wide$C_SBP.PWV1
pwv_wide$delta_c_dbp <- pwv_wide$C_DBP.PWV2 - pwv_wide$C_DBP.PWV1
pwv_wide$delta_brachd <- pwv_wide$BRACHD.PWV2 - pwv_wide$BRACHD.PWV1
pwv_wide$delta_map_pm <- pwv_wide$MAP_PM.PWV2 - pwv_wide$MAP_PM.PWV1
pwv_wide$delta_hr_pm <- pwv_wide$HR_PM.PWV2 - pwv_wide$HR_PM.PWV1
pwv_wide$delta_ci <- pwv_wide$CI.PWV2 - pwv_wide$CI.PWV1
pwv_wide$delta_lvdpdt <- pwv_wide$LVDPDT.PWV2 - pwv_wide$LVDPDT.PWV1
pwv_wide$delta_lv_ctx <- pwv_wide$LV_CTX.PWV2 - pwv_wide$LV_CTX.PWV1
pwv_wide$delta_svc <- pwv_wide$SVC.PWV2 - pwv_wide$SVC.PWV1
pwv_wide$delta_svr <- pwv_wide$SVR.PWV2 - pwv_wide$SVR.PWV1

# merge in the proteins (not scaled data)
analytes_keep <- analytes %>% filter(AptName %in% c("seq.13408.23","seq.19563.3","seq.7957.2","seq.2999.6","seq.18896.23","seq.5109.24",
                                                    "seq.9211.19","seq.18339.207","seq.18376.19","seq.18935.14","seq.24948.79","seq.4330.4"))
key <- analytes_keep[,c("AptName","UniProt","Target","TargetFullName")]
base_log_keep <- base_log %>% select(c("releaseid",all_of(analytes_keep$AptName)))

# with echo data - TODAY
today_echo <- merge(base_log_keep,today_echo,by="releaseid",all.x = F, all.y = F)
# remove markers with >=80% missing 
today_echo <- today_echo %>% dplyr::select(where(~mean(is.na(.)) < 0.8))
# with echo data - TODAY2
today2_echo <- merge(base_log_keep,today2_echo,by="releaseid",all.x = F, all.y = F)
# remove markers with >=80% missing 
today2_echo <- today2_echo %>% dplyr::select(where(~mean(is.na(.)) < 0.8))

# with pwv data - time point 1
pwv_v1 <- pwv %>% filter(TIMEPOINT=="PWV1")
# remove markers with >=80% missing 
pwv_v1 <- pwv_v1 %>% dplyr::select(where(~mean(is.na(.)) < 0.8))
pwv_v1 <- merge(base_log_keep,pwv_v1,by="releaseid",all.x = F, all.y = F)

# with pwv data - time point 2
pwv_v2 <- pwv %>% filter(TIMEPOINT=="PWV2")
pwv_v2 <- pwv_v2 %>% dplyr::select(where(~mean(is.na(.)) < 0.8))
pwv_v2 <- merge(base_log_keep,pwv_v2,by="releaseid",all.x = F, all.y = F)

# with pwv data - deltas
pwv_deltas <- pwv_wide %>% select(starts_with("delta"))
pwv_deltas <- cbind(pwv_wide$releaseid, pwv_deltas)
pwv_deltas <- pwv_deltas %>% filter(!is.na(delta_days))
names(pwv_deltas)[1] <- "releaseid"
pwv_deltas <- pwv_deltas %>% dplyr::select(where(~mean(is.na(.)) < 0.8))
pwv_deltas <- merge(base_log_keep,pwv_deltas,by="releaseid",all.x = F, all.y = F)


```

# Results

## Key to protein names

```{r, include=TRUE}
kable(key)
```

## Echocardiogram

### Summary statistics - TODAY

```{r, include=FALSE}
echonames <- colnames(today_echo)
echonames <- echonames[-1]
t <- CreateTableOne(vars=echonames,data=today_echo)
t <- print(t, nonnormal = echonames, missing=T)
```

```{r, include=TRUE, comment=""}
kable(t)
```

### Summary statistics - TODAY2

```{r, include=FALSE}
echonames <- colnames(today2_echo)
echonames <- echonames[-1]
t <- CreateTableOne(vars=echonames,data=today2_echo)
t <- print(t, nonnormal = echonames, missing=T)
```

```{r, include=TRUE, comment=""}
kable(t)
```

### Correlation between TODAY echo and proteins

```{r, include=TRUE}
 # Correlations
today_echo_corr <- today_echo %>% select(-c(releaseid, days))
  M <- corr.test(data.matrix(today_echo_corr), normal = F, method = "spearman")
  r <- data.frame(round(M$r, 3))
#  r <- r %>% filter(rownames(r)=="CENTRAL_AHI")
#  r <- r %>% select(-releaseid)
#  p <- data.frame(round(M$p, 3))
#  p <- p %>% filter(rownames(p)=="CENTRAL_AHI")
#  p <- p %>% select(-releaseid)
  emphasize.strong.cells(which(M$p <= 0.05, arr.ind = TRUE))
  cat("\n")
  pander(r, caption = "Coefficients")
  cat("\n")
```

### Correlation between TODAY2 echo and proteins

```{r, include=TRUE}
 # Correlations
today2_echo_corr <- today2_echo %>% select(-c(releaseid, days))
  M <- corr.test(data.matrix(today2_echo_corr), normal = F, method = "spearman")
  r <- data.frame(round(M$r, 3))
#  r <- r %>% filter(rownames(r)=="CENTRAL_AHI")
#  r <- r %>% select(-releaseid)
#  p <- data.frame(round(M$p, 3))
#  p <- p %>% filter(rownames(p)=="CENTRAL_AHI")
#  p <- p %>% select(-releaseid)
  emphasize.strong.cells(which(M$p <= 0.05, arr.ind = TRUE))
  cat("\n")
  pander(r, caption = "Coefficients")
  cat("\n")
```

### Correlation between change in echo and proteins

## PWV

### Summary statistics 

```{r, include=FALSE}
pwvnames <- colnames(pwv)
pwvnames <- pwvnames[!(pwvnames %in% c("releaseid","TIMEPOINT"))]
t <- CreateTableOne(vars=pwvnames,data=pwv,strata = "TIMEPOINT")
t <- print(t, nonnormal = pwvnames, missing=T)
```

```{r, include=TRUE, comment=""}
kable(t)
```

### Correlation between PWV time 1 and proteins

```{r, include=TRUE}
 # Correlations
pwv_v1_corr <- pwv_v1 %>% select(-c(releaseid,TIMEPOINT))
  M <- corr.test(data.matrix(pwv_v1_corr), normal = F, method = "spearman")
  r <- data.frame(round(M$r, 3))
#  r <- r %>% filter(rownames(r)=="CENTRAL_AHI")
#  r <- r %>% select(-releaseid)
#  p <- data.frame(round(M$p, 3))
#  p <- p %>% filter(rownames(p)=="CENTRAL_AHI")
#  p <- p %>% select(-releaseid)
  emphasize.strong.cells(which(M$p <= 0.05, arr.ind = TRUE))
  cat("\n")
  pander(r, caption = "Coefficients")
  cat("\n")
```

### Correlation between PWV time 2 and proteins

```{r, include=TRUE}
 # Correlations
pwv_v2_corr <- pwv_v2 %>% select(-c(releaseid,TIMEPOINT))
  M <- corr.test(data.matrix(pwv_v2_corr), normal = F, method = "spearman")
  r <- data.frame(round(M$r, 3))
#  r <- r %>% filter(rownames(r)=="CENTRAL_AHI")
#  r <- r %>% select(-releaseid)
#  p <- data.frame(round(M$p, 3))
#  p <- p %>% filter(rownames(p)=="CENTRAL_AHI")
#  p <- p %>% select(-releaseid)
  emphasize.strong.cells(which(M$p <= 0.05, arr.ind = TRUE))
  cat("\n")
  pander(r, caption = "Coefficients")
  cat("\n")
```

### Correlation between change in PWV and proteins


```{r, include=TRUE}
 # Correlations
pwv_deltas_corr <- pwv_deltas %>% select(-c(releaseid))
  M <- corr.test(data.matrix(pwv_deltas_corr), normal = F, method = "spearman")
  r <- data.frame(round(M$r, 3))
#  r <- r %>% filter(rownames(r)=="CENTRAL_AHI")
#  r <- r %>% select(-releaseid)
#  p <- data.frame(round(M$p, 3))
#  p <- p %>% filter(rownames(p)=="CENTRAL_AHI")
#  p <- p %>% select(-releaseid)
  emphasize.strong.cells(which(M$p <= 0.05, arr.ind = TRUE))
  cat("\n")
  pander(r, caption = "Coefficients")
  cat("\n")
```
