---
title: "TODAY study - proteomics, hypertension, and OSA"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
knit: (function(rmdfile, ...) { rmarkdown::render(rmdfile, output_dir='/Users/pylell/Dropbox/TODAY HTN and DKD output/') })
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include=FALSE}
library(SomaDataIO)
library(limma)
library(dplyr)
library(caret)
library(purrr)
library(multtest)
library(openxlsx)
library(tableone)
library(EnhancedVolcano)
library(knitr)
library(survival)
library(broom)
library(psych)
library(corrplot)
library(pander)
library(tableone)
library(polycor)
library(piercer)

knitr::opts_chunk$set(echo = FALSE,warning = FALSE)

knitr::opts_chunk$set(echo = FALSE)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "E:/Petter Bjornstad/TODAY subaward"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/TODAY subaward/"
} else if (Sys.info()["sysname"] == "Darwin"){
 home_dir = "/Volumes/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/TODAY subaward/"
 #home_dir = "/Users/pylell/Documents/Downtime/TODAY"
  }
knitr::opts_knit$set(root.dir = home_dir)

```

```{r, include=FALSE}
setwd(home_dir)
# load somalogic data, with QC samples already excluded
load("./Somalogic data raw/soma.Rdata")

# load analyte info
load("./Somalogic data raw/analytes.Rdata")

# load comorbidity data
load("./Clinical data/comorb.Rdata")

# load baseline risk factors
load("./Clinical data/TODAY/baserisk.Rdata")

# take only the baseline soma samples
# can we just take the earliest or is it possible someone would have a follow-up sample but not baseline?
# probably can take the first and then check the years to make sure compatible with TODAY
base <- soma %>% arrange(releaseid,Date.Drawn) %>% group_by(releaseid) %>% filter(row_number()==1)

# these 3 release IDs have first sample after end of recruitment, so they must be missing the baseline visit
base <- base %>% filter(!releaseid %in% c("65-85903","65-47984","65-25901"))

# merge in complication data
base <- left_join(base, comorb, by="releaseid")

# merge in baseline risk factors
base <- left_join(base, baserisk, by="releaseid")

# identify columns corresponding to proteins
#is_seq <- function(.x) grepl("^seq\\.[0-9]{4}", .x) # regex for analytes
is_seq <- function(.x) grepl("seq", .x)
seq <- is_seq(names(base))

# convert to numeric
base[,seq] <- apply(base[,seq],2,as.numeric)

# are there proteins with low variability?
#no_var = caret::nearZeroVar(base[,seq])
# none

# log transform
base_log <- base %>% modify_if(is_seq(names(.)), log)

# scale by SD
base_log_scale <- base_log
predictors <- colnames(base_log[seq])
for (i in 1:length(predictors)) {
  base_log_scale[,paste0(predictors[i])] <- base_log_scale[,paste0(predictors[i])]/sd(unlist(base_log[,paste0(predictors[i])]))
}

# TODAY2 TME
tme <- read.csv("./Clinical data/TODAY2/TME.csv")
tme <- tme %>% select(RELEASEID,TMETYPE,DAYSTOTME)
colnames(tme) <- c("releaseid","TMETYPE","DAYSTOTME")
tme <- tme %>% filter(TMETYPE==4)

# TODAY2 PSG
psg <- read.csv("./Clinical data/TODAY2/PSG.csv")
psg$releaseid <- psg$RELEASEID
psg$RELEASEID <- NULL
psg <- psg %>% select(releaseid,CENTRAL_AHI, OSA_SEVERITY_LEVEL)

# merge base_log (not scaled data) with PSG data
analytes_keep <- analytes %>% filter(AptName %in% c("seq.13408.23","seq.19563.3","seq.7957.2","seq.4330.4",
                                                    "seq.2999.6","seq.18896.23","seq.3235.50","seq.8890.9",
                                                    "seq.5109.24","seq.7735.17","seq.9211.19","seq.6478.2",
                                                    "seq.18339.207","seq.18376.19","seq.9018.38","seq.5140.56",
                                                    "seq.18935.14","seq.15539.15","seq.24948.79","seq.16307.22","seq.23569.53","seq.6252.62","seq.8245.27"))
key <- analytes_keep[,c("AptName","UniProt","Target","TargetFullName")]
base_log_keep <- base_log %>% select(c("releaseid",all_of(analytes_keep$AptName)))
base_log_keep <- merge(base_log_keep,psg,by="releaseid",all.x = T, all.y = F)
base_log_keep <- merge(base_log_keep,tme,by="releaseid",all.x = T, all.y = F)
# base_log_keep now has ALL participants, OSA variables, and proteins related to hypertension
base_log_keep$osa <- NA
base_log_keep$osa <- ifelse(base_log_keep$TMETYPE==4, 1, base_log_keep$osa)
base_log_keep$osa <- ifelse(base_log_keep$OSA_SEVERITY_LEVEL %in%  c("1","2","3"), 1, base_log_keep$osa)
base_log_keep$osa <- ifelse(is.na(base_log_keep$osa), 0, base_log_keep$osa)
base_log_keep$osa <- as.factor(base_log_keep$osa)
base_log_keep$OSA_SEVERITY_LEVEL <- as.factor(base_log_keep$OSA_SEVERITY_LEVEL)

# what dfs do I need
# analysis with all participants, OSA variables, and proteins, but need to create new var to indicate OSA Y/N combining TME and PSG
# need to have descriptive statistics on the df above (all ppts)
# df with only proteins and AHI for correlations
# need a smaller df for descriptive statistics for only the participants who have AHI

# df for descriptive statistics
corr <- base_log_keep[!is.na(base_log_keep$CENTRAL_AHI),]
corr_demo <- corr
corr <- corr %>% select(-releaseid)

t_osa <- corr

corr <- corr %>% select(-c(OSA_SEVERITY_LEVEL,osa,TMETYPE,DAYSTOTME))


```


```{r, include=FALSE}
corr_plot_modified_nodict <- function(data, X, Y, cor_method = "pearson", adj_var = NULL, 
                               method = "color", insig = "pch", coef_col = NULL,
                               pch = 4, pch.col = "black", pch.cex = 0) {
  
  n_cols = length(Y)
  M <- cor(y = subset(data, select = Y),
           x = subset(data, select = X),
           use = "pairwise.complete.obs",
           method = cor_method)
  
  if (!is.na(adj_var)){
    x_vars <- rep(X, times = length(Y))
    x_coord <- rep(seq(1,length(Y)), each = length(X))
    y_vars <- rep(Y, each = length(X))
    y_coord <- rep(seq(length(X),1), times = length(Y))
    lm_extracted <- data.frame(yName = character(0),
                               xName = character(0),
                               x = numeric(0),
                               y = numeric(0),
                               adj_var = character(0),
                               adj_x_coef = numeric(0),
                               adj_x_pval = numeric(0))
    for (i in 1:length(x_vars)) {
      if(x_vars[i] != y_vars[i]) {
        lm_formula <- as.formula(paste0(y_vars[i], "~", x_vars[i], "+", adj_var))
        lm_model <- lm(lm_formula, data = data)
        lm_summary <- summary(lm_model)
        lm_coef <- lm_summary$coefficient[x_vars[i], "Estimate"]
        lm_p_val <- lm_summary$coefficient[x_vars[i], "Pr(>|t|)"]
        lm_extracted <- rbind(lm_extracted,
                              data.frame(
                                yName = y_vars[i],
                                xName = x_vars[i],
                                x = x_coord[i],
                                y = y_coord[i],
                                adj_var = adj_var,
                                adj_x_coef = lm_coef,
                                adj_x_pval = lm_p_val))
      }
    }
  }
  
  correlation_p_value <- correlation_p_value_matrix(data, relevant_vars = c(Y, X), n_cols = n_cols, cor_method = cor_method)
  
  corrplot(M,
           p.mat = correlation_p_value,
           method = method,
           tl.col = "black",
           tl.srt = 45,
           tl.cex = 1.2,
           insig = insig,
           addCoef.col = coef_col,
           addgrid.col = 'lightgray',
           number.digits = 2,
           pch.col = pch.col, 
           pch = pch,
           pch.cex = pch.cex)$corrPos -> p1
  p1_sub <- subset(p1, p.value <= 0.05)
  p1_sub2 <- subset(p1, p.value <= 0.05 & abs(corr) >= 0.7)
  
  if (nrow(p1) > 0) {
    graphics::text(p1$x, p1$y, sprintf("%.2f", p1$corr), adj = c(0.5, 0), cex=1.2)
    if (nrow(p1_sub > 0)) {
      graphics::text(p1_sub$x, p1_sub$y, stars.pval(p1_sub$p.value), adj = c(0.5, 2), cex=1.2)
    }
  }
  if (nrow(p1_sub2) > 0) {
    graphics::text(p1_sub2$x, p1_sub2$y, sprintf("%.2f", p1_sub2$corr), col = "white", adj = c(0.5, 0), cex=1.2)
    graphics::text(p1_sub2$x, p1_sub2$y, stars.pval(p1_sub2$p.value), col = "white", adj = c(0.5, 2), cex=1.2)
  }
  if (!is.na(adj_var)) {
    lm_extracted <- subset(lm_extracted, adj_x_pval <= 0.05)
    if (nrow(lm_extracted) > 0){
      graphics::rect(xleft = lm_extracted$x-.45, 
                     ybottom = lm_extracted$y-.45,
                     xright = lm_extracted$x+.45,
                     ytop = lm_extracted$y+.45)
    }
  }
}


correlation_p_value_matrix <- function(data, relevant_vars, n_cols, cor_method = "pearson") {
  # Filter relevant variables for the correlation matrix
  dat_correlation <- data %>%
    dplyr::select(all_of(relevant_vars))
  
  # Compute p-values 
  res2 <- Hmisc::rcorr(as.matrix(dat_correlation),
                       type = cor_method)
  corr_pval <- as.data.frame(res2$P) %>%
    dplyr::select(all_of(relevant_vars[1:n_cols]))
  corr_pval = corr_pval[(n_cols + 1):nrow(corr_pval),]
  return(as.matrix(corr_pval))
}
```

# Results

## Key to protein names

```{r, include=TRUE}
kable(key)
```

## Participants in the PSG substudy

### Summary statistics 

Note: severity 1="Mild," 2="Moderate," 3="Severe"

```{r, include=FALSE}
t <- CreateTableOne(vars=c("OSA_SEVERITY_LEVEL","CENTRAL_AHI"),data=corr_demo)
t <- print(t, nonnorm="CENTRAL_AHI", minMax=T)
```

```{r, include=TRUE}
kable(t)
```

### Correlation between AHI and proteins

```{r, include=TRUE}
 # Correlations
  M <- corr.test(data.matrix(corr), normal = F, method = "spearman")
  r <- data.frame(round(M$r, 3))
#  r <- r %>% filter(rownames(r)=="CENTRAL_AHI")
#  r <- r %>% select(-releaseid)
#  p <- data.frame(round(M$p, 3))
#  p <- p %>% filter(rownames(p)=="CENTRAL_AHI")
#  p <- p %>% select(-releaseid)
  emphasize.strong.cells(which(M$p <= 0.05, arr.ind = TRUE))
  cat("\n")
  pander(r, caption = "Coefficients")
  cat("\n")
```

### Levels of proteins associated with hypertension, by OSA status

```{r, include=FALSE}
t <- CreateTableOne(vars=c("seq.13408.23","seq.19563.3","seq.7957.2","seq.4330.4",
                                                    "seq.2999.6","seq.18896.23","seq.3235.50","seq.8890.9",
                                                    "seq.5109.24","seq.7735.17","seq.9211.19","seq.6478.2",
                                                    "seq.18339.207","seq.18376.19","seq.9018.38","seq.5140.56",
                                                    "seq.18935.14","seq.15539.15","seq.24948.79","seq.16307.22","seq.23569.53","seq.6252.62","seq.8245.27"),data=t_osa, strata="OSA_SEVERITY_LEVEL")
t <- print(t, minMax=T, nonnorm=c("seq.13408.23","seq.19563.3","seq.7957.2","seq.4330.4",
                                                    "seq.2999.6","seq.18896.23","seq.3235.50","seq.8890.9",
                                                    "seq.5109.24","seq.7735.17","seq.9211.19","seq.6478.2",
                                                    "seq.18339.207","seq.18376.19","seq.9018.38","seq.5140.56",
                                                    "seq.18935.14","seq.15539.15","seq.24948.79","seq.16307.22","seq.23569.53","seq.6252.62","seq.8245.27"))
t <- print(t)
```

```{r, include=TRUE}
kable(t)
```

### Levels of proteins associated with hypertension, by OSA status, with all OSA groups combined

```{r, include=FALSE}
t <- CreateTableOne(vars=c("seq.13408.23","seq.19563.3","seq.7957.2","seq.4330.4",
                                                    "seq.2999.6","seq.18896.23","seq.3235.50","seq.8890.9",
                                                    "seq.5109.24","seq.7735.17","seq.9211.19","seq.6478.2",
                                                    "seq.18339.207","seq.18376.19","seq.9018.38","seq.5140.56",
                                                    "seq.18935.14","seq.15539.15","seq.24948.79","seq.16307.22","seq.23569.53","seq.6252.62","seq.8245.27"),data=t_osa, strata="osa")
t <- print(t, minMax=T, nonnorm=c("seq.13408.23","seq.19563.3","seq.7957.2","seq.4330.4",
                                                    "seq.2999.6","seq.18896.23","seq.3235.50","seq.8890.9",
                                                    "seq.5109.24","seq.7735.17","seq.9211.19","seq.6478.2",
                                                    "seq.18339.207","seq.18376.19","seq.9018.38","seq.5140.56",
                                                    "seq.18935.14","seq.15539.15","seq.24948.79","seq.16307.22","seq.23569.53","seq.6252.62","seq.8245.27"))
t <- print(t)
```

```{r, include=TRUE}
kable(t)
```

### Polyserial correlations of OSA severity and proteins

```{r, include=F}
HC <- hetcor(t_osa$seq.13408.23,t_osa$OSA_SEVERITY_LEVEL, ML = TRUE)
a <- ci.rpc(r = HC$correlations[2,1], se = HC$std.errors[2,1], rn = "seq.13408.23 & OSA_SEVERITY_LEVEL")
```

```{r, include=T}
kable(a[,1:6])
```

```{r, include=F}
HC <- hetcor(t_osa$seq.19563.3,t_osa$OSA_SEVERITY_LEVEL, ML = TRUE)
a <- ci.rpc(r = HC$correlations[2,1], se = HC$std.errors[2,1], rn = "seq.19563.3 & OSA_SEVERITY_LEVEL")
```

```{r, include=T}
kable(a[,1:6])
```

```{r, include=F}
HC <- hetcor(t_osa$seq.7957.2,t_osa$OSA_SEVERITY_LEVEL, ML = TRUE)
a <- ci.rpc(r = HC$correlations[2,1], se = HC$std.errors[2,1], rn = "seq.7957.2 & OSA_SEVERITY_LEVEL")
```

```{r, include=T}
kable(a[,1:6])
```

```{r, include=F}
HC <- hetcor(t_osa$seq.4330.4,t_osa$OSA_SEVERITY_LEVEL, ML = TRUE)
a <- ci.rpc(r = HC$correlations[2,1], se = HC$std.errors[2,1], rn = "seq.4330.4 & OSA_SEVERITY_LEVEL")
```

```{r, include=T}
kable(a[,1:6])
```

```{r, include=F}
HC <- hetcor(t_osa$seq.2999.6,t_osa$OSA_SEVERITY_LEVEL, ML = TRUE)
a <- ci.rpc(r = HC$correlations[2,1], se = HC$std.errors[2,1], rn = "seq.2999.6 & OSA_SEVERITY_LEVEL")
```

```{r, include=T}
kable(a[,1:6])
```

```{r, include=F}
HC <- hetcor(t_osa$seq.18896.23,t_osa$OSA_SEVERITY_LEVEL, ML = TRUE)
a <- ci.rpc(r = HC$correlations[2,1], se = HC$std.errors[2,1], rn = "seq.18896.23 & OSA_SEVERITY_LEVEL")
```

```{r, include=T}
kable(a[,1:6])
```

```{r, include=F}
HC <- hetcor(t_osa$seq.3235.50,t_osa$OSA_SEVERITY_LEVEL, ML = TRUE)
a <- ci.rpc(r = HC$correlations[2,1], se = HC$std.errors[2,1], rn = "seq.3235.50 & OSA_SEVERITY_LEVEL")
```

```{r, include=T}
kable(a[,1:6])
```

```{r, include=F}
HC <- hetcor(t_osa$seq.8890.9,t_osa$OSA_SEVERITY_LEVEL, ML = TRUE)
a <- ci.rpc(r = HC$correlations[2,1], se = HC$std.errors[2,1], rn = "seq.8890.9 & OSA_SEVERITY_LEVEL")
```

```{r, include=T}
kable(a[,1:6])
```

```{r, include=F}
HC <- hetcor(t_osa$seq.5109.24,t_osa$OSA_SEVERITY_LEVEL, ML = TRUE)
a <- ci.rpc(r = HC$correlations[2,1], se = HC$std.errors[2,1], rn = "seq.5109.24 & OSA_SEVERITY_LEVEL")
```

```{r, include=T}
kable(a[,1:6])
```

```{r, include=F}
HC <- hetcor(t_osa$seq.7735.17,t_osa$OSA_SEVERITY_LEVEL, ML = TRUE)
a <- ci.rpc(r = HC$correlations[2,1], se = HC$std.errors[2,1], rn = "seq.7735.17 & OSA_SEVERITY_LEVEL")
```

```{r, include=T}
kable(a[,1:6])
```

```{r, include=F}
HC <- hetcor(t_osa$seq.9211.19,t_osa$OSA_SEVERITY_LEVEL, ML = TRUE)
a <- ci.rpc(r = HC$correlations[2,1], se = HC$std.errors[2,1], rn = "seq.9211.19 & OSA_SEVERITY_LEVEL")
```

```{r, include=T}
kable(a[,1:6])
```

```{r, include=F}
HC <- hetcor(t_osa$seq.6478.2,t_osa$OSA_SEVERITY_LEVEL, ML = TRUE)
a <- ci.rpc(r = HC$correlations[2,1], se = HC$std.errors[2,1], rn = "seq.6478.2 & OSA_SEVERITY_LEVEL")
```

```{r, include=T}
kable(a[,1:6])
```

```{r, include=F}
HC <- hetcor(t_osa$seq.18339.207,t_osa$OSA_SEVERITY_LEVEL, ML = TRUE)
a <- ci.rpc(r = HC$correlations[2,1], se = HC$std.errors[2,1], rn = "seq.18339.207 & OSA_SEVERITY_LEVEL")
```

```{r, include=T}
kable(a[,1:6])
```

```{r, include=F}
HC <- hetcor(t_osa$seq.18376.19,t_osa$OSA_SEVERITY_LEVEL, ML = TRUE)
a <- ci.rpc(r = HC$correlations[2,1], se = HC$std.errors[2,1], rn = "seq.18376.19 & OSA_SEVERITY_LEVEL")
```

```{r, include=T}
kable(a[,1:6])
```

```{r, include=F}
HC <- hetcor(t_osa$seq.9018.38,t_osa$OSA_SEVERITY_LEVEL, ML = TRUE)
a <- ci.rpc(r = HC$correlations[2,1], se = HC$std.errors[2,1], rn = "seq.9018.38 & OSA_SEVERITY_LEVEL")
```

```{r, include=T}
kable(a[,1:6])
```

```{r, include=F}
HC <- hetcor(t_osa$seq.5140.56,t_osa$OSA_SEVERITY_LEVEL, ML = TRUE)
a <- ci.rpc(r = HC$correlations[2,1], se = HC$std.errors[2,1], rn = "seq.5140.56 & OSA_SEVERITY_LEVEL")
```

```{r, include=T}
kable(a[,1:6])
```

```{r, include=F}
HC <- hetcor(t_osa$seq.18935.14,t_osa$OSA_SEVERITY_LEVEL, ML = TRUE)
a <- ci.rpc(r = HC$correlations[2,1], se = HC$std.errors[2,1], rn = "seq.18935.14 & OSA_SEVERITY_LEVEL")
```

```{r, include=T}
kable(a[,1:6])
```

```{r, include=F}
HC <- hetcor(t_osa$seq.15539.15,t_osa$OSA_SEVERITY_LEVEL, ML = TRUE)
a <- ci.rpc(r = HC$correlations[2,1], se = HC$std.errors[2,1], rn = "seq.15539.15 & OSA_SEVERITY_LEVEL")
```

```{r, include=T}
kable(a[,1:6])
```

```{r, include=F}
HC <- hetcor(t_osa$seq.24948.79,t_osa$OSA_SEVERITY_LEVEL, ML = TRUE)
a <- ci.rpc(r = HC$correlations[2,1], se = HC$std.errors[2,1], rn = "seq.24948.79 & OSA_SEVERITY_LEVEL")
```

```{r, include=T}
kable(a[,1:6])
```

```{r, include=F}
HC <- hetcor(t_osa$seq.16307.22,t_osa$OSA_SEVERITY_LEVEL, ML = TRUE)
a <- ci.rpc(r = HC$correlations[2,1], se = HC$std.errors[2,1], rn = "seq.16307.22 & OSA_SEVERITY_LEVEL")
```

```{r, include=T}
kable(a[,1:6])
```
```{r, include=F}
HC <- hetcor(t_osa$seq.23569.53,t_osa$OSA_SEVERITY_LEVEL, ML = TRUE)
a <- ci.rpc(r = HC$correlations[2,1], se = HC$std.errors[2,1], rn = "seq.23569.53 & OSA_SEVERITY_LEVEL")
```

```{r, include=T}
kable(a[,1:6])
```
```{r, include=F}
HC <- hetcor(t_osa$seq.6252.62,t_osa$OSA_SEVERITY_LEVEL, ML = TRUE)
a <- ci.rpc(r = HC$correlations[2,1], se = HC$std.errors[2,1], rn = "seq.6252.62 & OSA_SEVERITY_LEVEL")
```

```{r, include=T}
kable(a[,1:6])
```
```{r, include=F}
HC <- hetcor(t_osa$seq.8245.27,t_osa$OSA_SEVERITY_LEVEL, ML = TRUE)
a <- ci.rpc(r = HC$correlations[2,1], se = HC$std.errors[2,1], rn = "seq.8245.27 & OSA_SEVERITY_LEVEL")
```

```{r, include=T}
kable(a[,1:6])
```


## All participants

### Summary statistics 

```{r, include=FALSE}
t <- CreateTableOne(vars="osa",data=base_log_keep)
t <- print(t, minMax=T)
```

```{r, include=TRUE}
kable(t)
```

### Levels of proteins associated with hypertension, by OSA status

```{r, include=FALSE}
t <- CreateTableOne(vars=c("seq.13408.23","seq.19563.3","seq.7957.2","seq.4330.4",
                                                    "seq.2999.6","seq.18896.23","seq.3235.50","seq.8890.9",
                                                    "seq.5109.24","seq.7735.17","seq.9211.19","seq.6478.2",
                                                    "seq.18339.207","seq.18376.19","seq.9018.38","seq.5140.56",                                       "seq.18935.14","seq.15539.15","seq.24948.79","seq.16307.22","seq.23569.53","seq.6252.62","seq.8245.27"),data=base_log_keep, strata="osa")
t <- print(t, minMax=T, nonnorm=c("seq.13408.23","seq.19563.3","seq.7957.2","seq.4330.4",
                                                    "seq.2999.6","seq.18896.23","seq.3235.50","seq.8890.9",
                                                    "seq.5109.24","seq.7735.17","seq.9211.19","seq.6478.2",
                                                    "seq.18339.207","seq.18376.19","seq.9018.38","seq.5140.56",
                                                    "seq.18935.14","seq.15539.15","seq.24948.79","seq.16307.22","seq.23569.53","seq.6252.62","seq.8245.27"))
```

```{r, include=TRUE}
kable(t)
```
