---
title: "TODAY Somalogic baseline analyses - adjusted Cox models"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include=FALSE}
library(SomaDataIO)
library(limma)
library(dplyr)
library(caret)
library(purrr)
library(multtest)
library(openxlsx)
library(tableone)
library(EnhancedVolcano)
library(knitr)
library(survival)
library(broom)
library(emmeans)
library(ggvenn)
library(ggpubr)

knitr::opts_chunk$set(echo = FALSE,warning = FALSE)

knitr::opts_chunk$set(echo = FALSE)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "E:/Petter Bjornstad/TODAY subaward"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/TODAY subaward/"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/TODAY subaward"
}
knitr::opts_knit$set(root.dir = home_dir)
```

```{r, include=FALSE}
setwd(home_dir)
# load somalogic data, with QC samples already excluded
load("./Somalogic data raw/soma.Rdata")

# load analyte info
load("./Somalogic data raw/analytes.Rdata")

# load comorbidity data
load("./Clinical data/comorb.Rdata")

# load baseline risk factors
load("./Clinical data/TODAY/baserisk.Rdata")

# take only the baseline soma samples
# can we just take the earliest or is it possible someone would have a follow-up sample but not baseline?
# probably can take the first and then check the years to make sure compatible with TODAY
base <- soma %>% arrange(releaseid,Date.Drawn) %>% dplyr::group_by(releaseid) %>% dplyr::filter(row_number()==1)
# these 3 release IDs have first sample after end of recruitment, so they must be missing the baseline visit
base <- base %>% filter(!releaseid %in% c("65-85903","65-47984","65-25901"))

# the above section of code used to work but seems to have broken....this should accomplish the same thing?
#base <- soma %>% filter(visit==1)

# merge in complication data
base <- left_join(base, comorb, by="releaseid")
# this was previously:
# base <- merge(base, comorb, by="releaseid",all.x=T, all.y=F)

# merge in baseline risk factors
base <- left_join(base, baserisk, by="releaseid")
# this was previously:
#base <- merge(base, baserisk, by="releaseid",all.x=T, all.y=F)

# log transform UAlbCreat
base$log_UAlbCreat <- log(base$UAlbCreat + 0.0000001)

# identify columns corresponding to proteins
#is_seq <- function(.x) grepl("^seq\\.[0-9]{4}", .x) # regex for analytes
is_seq <- function(.x) grepl("seq", .x)
seq <- is_seq(names(base))

# convert to numeric
base[,seq] <- apply(base[,seq],2,as.numeric)

# are there proteins with low variability?
no_var = caret::nearZeroVar(base[,seq])
# none

# log transform
base_log <- base %>% modify_if(is_seq(names(.)), log)

# scale by SD
base_log_scale <- base_log
predictors <- colnames(base_log[seq])
for (i in 1:length(predictors)) {
  base_log_scale[,paste0(predictors[i])] <- base_log_scale[,paste0(predictors[i])]/sd(unlist(base_log[,paste0(predictors[i])]))
}

# create separate df for each race/ethnicity
base_log_hispanic <- base_log %>% filter(racedesc == "Hispanic")
base_log_scale_hispanic <- base_log_scale %>% filter(racedesc == "Hispanic")
base_log_nhb <- base_log %>% filter(racedesc == "Non-Hispanic Black")
base_log_scale_nhb <- base_log_scale %>% filter(racedesc == "Non-Hispanic Black")
base_log_nhw <- base_log %>% filter(racedesc == "Non-Hispanic White")
base_log_scale_nhw <- base_log_scale %>% filter(racedesc == "Non-Hispanic White")

```

```{r, include=FALSE}
# HISPANIC

# Macroalbuminuria
result_MAC <- NULL
predictors <- colnames(base_log_hispanic[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOMAC, MAC)~",predictors[i],"+HbA1c+log_trig+sbp+si_1_ins0")
  mod <- coxph(data = base_log_scale_hispanic, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_MAC <- rbind(result_MAC,tidymod)
}
result_MAC$adj.p.value <- p.adjust(result_MAC$p.value, "fdr")
result_MAC$AptName <- result_MAC$term
result_MAC$term <- NULL
result_MAC <- merge(result_MAC,analytes,by="AptName",all.x = T, all.y = F)
result_MAC <- result_MAC[order(result_MAC$p.value),] 

# Microalbuminuria
result_MIC <- NULL
predictors <- colnames(base_log_hispanic[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOMIC, MIC)~",predictors[i],"+HbA1c+log_trig+sbp+si_1_ins0")
  mod <- coxph(data = base_log_scale_hispanic, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_MIC <- rbind(result_MIC,tidymod)
}
result_MIC$adj.p.value <- p.adjust(result_MIC$p.value, "fdr")
result_MIC$AptName <- result_MIC$term
result_MIC$term <- NULL
result_MIC <- merge(result_MIC,analytes,by="AptName",all.x = T, all.y = F)
result_MIC <- result_MIC[order(result_MIC$p.value),] 

# Microalbuminuria or macroalbuminuria - whichever is first
result_MIC.OR.MAC <- NULL
predictors <- colnames(base_log_hispanic[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOMIC.OR.MAC, MIC.OR.MAC)~",predictors[i],"+HbA1c+log_trig+sbp+si_1_ins0")
  mod <- coxph(data = base_log_scale_hispanic, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_MIC.OR.MAC <- rbind(result_MIC.OR.MAC,tidymod)
}
result_MIC.OR.MAC$adj.p.value <- p.adjust(result_MIC.OR.MAC$p.value, "fdr")
result_MIC.OR.MAC$AptName <- result_MIC.OR.MAC$term
result_MIC.OR.MAC$term <- NULL
result_MIC.OR.MAC <- merge(result_MIC.OR.MAC,analytes,by="AptName",all.x = T, all.y = F)
result_MIC.OR.MAC <- result_MIC.OR.MAC[order(result_MIC.OR.MAC$p.value),] 
result_MIC.OR.MAC_hispanic <- result_MIC.OR.MAC

# hyperfiltration
result_HYP <- NULL
predictors <- colnames(base_log_hispanic[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOHYP, HYP)~",predictors[i],"+HbA1c+log_trig+sbp+si_1_ins0")
  mod <- coxph(data = base_log_scale_hispanic, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_HYP <- rbind(result_HYP,tidymod)
}
result_HYP$adj.p.value <- p.adjust(result_HYP$p.value, "fdr")
result_HYP$AptName <- result_HYP$term
result_HYP$term <- NULL
result_HYP <- merge(result_HYP,analytes,by="AptName",all.x = T, all.y = F)
result_HYP <- result_HYP[order(result_HYP$p.value),] 
result_HYP_hispanic <- result_HYP

# rapid eGFR decline
result_RAPID <- NULL
predictors <- colnames(base_log_hispanic[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTORAPID, RAPID)~",predictors[i],"+HbA1c+log_trig+sbp+si_1_ins0")
  mod <- coxph(data = base_log_scale_hispanic, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_RAPID <- rbind(result_RAPID,tidymod)
}
result_RAPID$adj.p.value <- p.adjust(result_RAPID$p.value, "fdr")
result_RAPID$AptName <- result_RAPID$term
result_RAPID$term <- NULL
result_RAPID <- merge(result_RAPID,analytes,by="AptName",all.x = T, all.y = F)
result_RAPID <- result_RAPID[order(result_RAPID$p.value),] 
result_RAPID_hispanic <- result_RAPID

# hypertension
# omit SBP
result_HTN <- NULL
predictors <- colnames(base_log_hispanic[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOHTN, HTN)~",predictors[i],"+HbA1c+log_trig+si_1_ins0")
  mod <- coxph(data = base_log_scale_hispanic, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_HTN <- rbind(result_HTN,tidymod)
}
result_HTN$adj.p.value <- p.adjust(result_HTN$p.value, "fdr")
result_HTN$AptName <- result_HTN$term
result_HTN$term <- NULL
result_HTN <- merge(result_HTN,analytes,by="AptName",all.x = T, all.y = F)
result_HTN <- result_HTN[order(result_HTN$p.value),] 

# hypertension with SBP
results_htn_bp <- NULL
predictors <- colnames(base_log_hispanic[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOHTN, HTN)~",predictors[i],"+HbA1c+log_trig+sbp+si_1_ins0")
  mod <- coxph(data = base_log_scale_hispanic, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  results_htn_bp <- rbind(results_htn_bp,tidymod)
}
results_htn_bp$adj.p.value <- p.adjust(results_htn_bp$p.value, "fdr")
results_htn_bp$AptName <- results_htn_bp$term
results_htn_bp$term <- NULL
results_htn_bp <- merge(results_htn_bp,analytes,by="AptName",all.x = T, all.y = F)
results_htn_bp <- results_htn_bp[order(results_htn_bp$p.value),] 

# hypertension with UACR
results_htn_uacr <- NULL
predictors <- colnames(base_log_hispanic[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOHTN, HTN)~",predictors[i],"+HbA1c+log_trig+log_UAlbCreat+si_1_ins0")
  mod <- coxph(data = base_log_scale_hispanic, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  results_htn_uacr <- rbind(results_htn_uacr,tidymod)
}
results_htn_uacr$adj.p.value <- p.adjust(results_htn_uacr$p.value, "fdr")
results_htn_uacr$AptName <- results_htn_uacr$term
results_htn_uacr$term <- NULL
results_htn_uacr <- merge(results_htn_uacr,analytes,by="AptName",all.x = T, all.y = F)
results_htn_uacr <- results_htn_uacr[order(results_htn_uacr$p.value),] 

# hypertension with eGFR
results_htn_egfr <- NULL
predictors <- colnames(base_log_hispanic[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOHTN, HTN)~",predictors[i],"+HbA1c+log_trig+EstCreatClear+si_1_ins0")
  mod <- coxph(data = base_log_scale_hispanic, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  results_htn_egfr <- rbind(results_htn_egfr,tidymod)
}
results_htn_egfr$adj.p.value <- p.adjust(results_htn_egfr$p.value, "fdr")
results_htn_egfr$AptName <- results_htn_egfr$term
results_htn_egfr$term <- NULL
results_htn_egfr <- merge(results_htn_egfr,analytes,by="AptName",all.x = T, all.y = F)
results_htn_egfr <- results_htn_egfr[order(results_htn_egfr$p.value),] 

# neuropathy
result_NEURO <- NULL
predictors <- colnames(base_log_hispanic[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTONEURO, NEURO)~",predictors[i],"+HbA1c+log_trig+sbp+si_1_ins0")
  mod <- coxph(data = base_log_scale_hispanic, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_NEURO <- rbind(result_NEURO,tidymod)
}
result_NEURO$adj.p.value <- p.adjust(result_NEURO$p.value, "fdr")
result_NEURO$AptName <- result_NEURO$term
result_NEURO$term <- NULL
result_NEURO <- merge(result_NEURO,analytes,by="AptName",all.x = T, all.y = F)
result_NEURO <- result_NEURO[order(result_NEURO$p.value),] 

# retinopathy
result_RETINO <- NULL
predictors <- colnames(base_log_hispanic[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTORETINO, RETINO)~",predictors[i],"+HbA1c+log_trig+sbp+si_1_ins0")
  mod <- coxph(data = base_log_scale_hispanic, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_RETINO <- rbind(result_RETINO,tidymod)
}
result_RETINO$adj.p.value <- p.adjust(result_RETINO$p.value, "fdr")
result_RETINO$AptName <- result_RETINO$term
result_RETINO$term <- NULL
result_RETINO <- merge(result_RETINO,analytes,by="AptName",all.x = T, all.y = F)
result_RETINO <- result_RETINO[order(result_RETINO$p.value),] 

# glycemic failure
# omit A1c
result_GLYC <- NULL
predictors <- colnames(base_log_hispanic[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOGLYC, GLYC)~",predictors[i],"+bmi+si_1_ins0+AGEBASE+sex")
  mod <- coxph(data = base_log_scale_hispanic, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_GLYC <- rbind(result_GLYC,tidymod)
}
result_GLYC$adj.p.value <- p.adjust(result_GLYC$p.value, "fdr")
result_GLYC$AptName <- result_GLYC$term
result_GLYC$term <- NULL
result_GLYC <- merge(result_GLYC,analytes,by="AptName",all.x = T, all.y = F)
result_GLYC <- result_GLYC[order(result_GLYC$p.value),] 

# glycemic failure with A1c
results_glyc_a1c <- NULL
predictors <- colnames(base_log_hispanic[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOGLYC, GLYC)~",predictors[i],"+HbA1c+bmi+si_1_ins0+AGEBASE+sex")
  mod <- coxph(data = base_log_scale_hispanic, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  results_glyc_a1c <- rbind(results_glyc_a1c,tidymod)
}
results_glyc_a1c$adj.p.value <- p.adjust(results_glyc_a1c$p.value, "fdr")
results_glyc_a1c$AptName <- results_glyc_a1c$term
results_glyc_a1c$term <- NULL
results_glyc_a1c <- merge(results_glyc_a1c,analytes,by="AptName",all.x = T, all.y = F)
results_glyc_a1c <- results_glyc_a1c[order(results_glyc_a1c$p.value),] 

# write all results to file
wb <- createWorkbook()
addWorksheet(wb,"MAC CPH")
writeData(wb,"MAC CPH",result_MAC,rowNames = F)
addWorksheet(wb,"MIC CPH")
writeData(wb,"MIC CPH",result_MIC,rowNames = F)
addWorksheet(wb,"MIC.OR.MAC CPH")
writeData(wb,"MIC.OR.MAC CPH",result_MIC.OR.MAC,rowNames = F)
addWorksheet(wb,"HYP CPH")
writeData(wb,"HYP CPH",result_HYP,rowNames = F)
addWorksheet(wb,"RAPID CPH")
writeData(wb,"RAPID CPH",result_RAPID,rowNames = F)
addWorksheet(wb,"HTN with SBP CPH")
writeData(wb,"HTN with SBP CPH",results_htn_bp,rowNames = F)
addWorksheet(wb,"HTN CPH")
writeData(wb,"HTN CPH",result_HTN,rowNames = F)
addWorksheet(wb,"HTN with UACR CPH")
writeData(wb,"HTN with UACR CPH",results_htn_uacr,rowNames = F)
addWorksheet(wb,"HTN with eGFR CPH")
writeData(wb,"HTN with eGFR CPH",results_htn_egfr,rowNames = F)
addWorksheet(wb,"NEURO CPH")
writeData(wb,"NEURO CPH",result_NEURO,rowNames = F)
addWorksheet(wb,"RETINO CPH")
writeData(wb,"RETINO CPH",result_RETINO,rowNames = F)
addWorksheet(wb,"GLYC CPH")
writeData(wb,"GLYC CPH",result_GLYC,rowNames = F)
addWorksheet(wb,"GLYC with A1c CPH")
writeData(wb,"GLYC with A1c CPH",results_glyc_a1c,rowNames = F)
saveWorkbook(wb,"./Results/Linear and Cox models/TODAY somalogic Cox models scaled baseline adjusted Hispanic.xlsx",overwrite = TRUE)
saveWorkbook(wb,"/Users/pylell/Dropbox/TODAY HTN and DKD output/TODAY somalogic Cox models scaled baseline adjusted Hispanic.xlsx",overwrite = TRUE)
saveWorkbook(wb,"/Users/pylell/Dropbox/TODAY glycemic manuscript [shared]/Analysis output/TODAY somalogic Cox models scaled baseline adjusted Hispanic.xlsx",overwrite = TRUE)
```

```{r, include=FALSE}
# NON-HISPANIC WHITE

# Macroalbuminuria
result_MAC <- NULL
predictors <- colnames(base_log_nhw[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOMAC, MAC)~",predictors[i],"+HbA1c+log_trig+sbp+si_1_ins0")
  mod <- coxph(data = base_log_scale_nhw, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_MAC <- rbind(result_MAC,tidymod)
}
result_MAC$adj.p.value <- p.adjust(result_MAC$p.value, "fdr")
result_MAC$AptName <- result_MAC$term
result_MAC$term <- NULL
result_MAC <- merge(result_MAC,analytes,by="AptName",all.x = T, all.y = F)
result_MAC <- result_MAC[order(result_MAC$p.value),] 

# Microalbuminuria
result_MIC <- NULL
predictors <- colnames(base_log_nhw[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOMIC, MIC)~",predictors[i],"+HbA1c+log_trig+sbp+si_1_ins0")
  mod <- coxph(data = base_log_scale_nhw, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_MIC <- rbind(result_MIC,tidymod)
}
result_MIC$adj.p.value <- p.adjust(result_MIC$p.value, "fdr")
result_MIC$AptName <- result_MIC$term
result_MIC$term <- NULL
result_MIC <- merge(result_MIC,analytes,by="AptName",all.x = T, all.y = F)
result_MIC <- result_MIC[order(result_MIC$p.value),] 

# Microalbuminuria or macroalbuminuria - whichever is first
result_MIC.OR.MAC <- NULL
predictors <- colnames(base_log_nhw[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOMIC.OR.MAC, MIC.OR.MAC)~",predictors[i],"+HbA1c+log_trig+sbp+si_1_ins0")
  mod <- coxph(data = base_log_scale_nhw, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_MIC.OR.MAC <- rbind(result_MIC.OR.MAC,tidymod)
}
result_MIC.OR.MAC$adj.p.value <- p.adjust(result_MIC.OR.MAC$p.value, "fdr")
result_MIC.OR.MAC$AptName <- result_MIC.OR.MAC$term
result_MIC.OR.MAC$term <- NULL
result_MIC.OR.MAC <- merge(result_MIC.OR.MAC,analytes,by="AptName",all.x = T, all.y = F)
result_MIC.OR.MAC <- result_MIC.OR.MAC[order(result_MIC.OR.MAC$p.value),] 
result_MIC.OR.MAC_nhw <- result_MIC.OR.MAC

# hyperfiltration
result_HYP <- NULL
predictors <- colnames(base_log_nhw[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOHYP, HYP)~",predictors[i],"+HbA1c+log_trig+sbp+si_1_ins0")
  mod <- coxph(data = base_log_scale_nhw, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_HYP <- rbind(result_HYP,tidymod)
}
result_HYP$adj.p.value <- p.adjust(result_HYP$p.value, "fdr")
result_HYP$AptName <- result_HYP$term
result_HYP$term <- NULL
result_HYP <- merge(result_HYP,analytes,by="AptName",all.x = T, all.y = F)
result_HYP <- result_HYP[order(result_HYP$p.value),] 
result_HYP_nhw <- result_HYP

# rapid eGFR decline
result_RAPID <- NULL
predictors <- colnames(base_log_nhw[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTORAPID, RAPID)~",predictors[i],"+HbA1c+log_trig+sbp+si_1_ins0")
  mod <- coxph(data = base_log_scale_nhw, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_RAPID <- rbind(result_RAPID,tidymod)
}
result_RAPID$adj.p.value <- p.adjust(result_RAPID$p.value, "fdr")
result_RAPID$AptName <- result_RAPID$term
result_RAPID$term <- NULL
result_RAPID <- merge(result_RAPID,analytes,by="AptName",all.x = T, all.y = F)
result_RAPID <- result_RAPID[order(result_RAPID$p.value),] 
result_RAPID_nhw <- result_RAPID

# hypertension
# omit SBP
result_HTN <- NULL
predictors <- colnames(base_log_nhw[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOHTN, HTN)~",predictors[i],"+HbA1c+log_trig+si_1_ins0")
  mod <- coxph(data = base_log_scale_nhw, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_HTN <- rbind(result_HTN,tidymod)
}
result_HTN$adj.p.value <- p.adjust(result_HTN$p.value, "fdr")
result_HTN$AptName <- result_HTN$term
result_HTN$term <- NULL
result_HTN <- merge(result_HTN,analytes,by="AptName",all.x = T, all.y = F)
result_HTN <- result_HTN[order(result_HTN$p.value),] 

# hypertension with SBP
results_htn_bp <- NULL
predictors <- colnames(base_log_nhw[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOHTN, HTN)~",predictors[i],"+HbA1c+log_trig+sbp+si_1_ins0")
  mod <- coxph(data = base_log_scale_nhw, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  results_htn_bp <- rbind(results_htn_bp,tidymod)
}
results_htn_bp$adj.p.value <- p.adjust(results_htn_bp$p.value, "fdr")
results_htn_bp$AptName <- results_htn_bp$term
results_htn_bp$term <- NULL
results_htn_bp <- merge(results_htn_bp,analytes,by="AptName",all.x = T, all.y = F)
results_htn_bp <- results_htn_bp[order(results_htn_bp$p.value),] 

# hypertension with UACR
results_htn_uacr <- NULL
predictors <- colnames(base_log_nhw[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOHTN, HTN)~",predictors[i],"+HbA1c+log_trig+log_UAlbCreat+si_1_ins0")
  mod <- coxph(data = base_log_scale_nhw, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  results_htn_uacr <- rbind(results_htn_uacr,tidymod)
}
results_htn_uacr$adj.p.value <- p.adjust(results_htn_uacr$p.value, "fdr")
results_htn_uacr$AptName <- results_htn_uacr$term
results_htn_uacr$term <- NULL
results_htn_uacr <- merge(results_htn_uacr,analytes,by="AptName",all.x = T, all.y = F)
results_htn_uacr <- results_htn_uacr[order(results_htn_uacr$p.value),] 

# hypertension with eGFR
results_htn_egfr <- NULL
predictors <- colnames(base_log_nhw[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOHTN, HTN)~",predictors[i],"+HbA1c+log_trig+EstCreatClear+si_1_ins0")
  mod <- coxph(data = base_log_scale_nhw, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  results_htn_egfr <- rbind(results_htn_egfr,tidymod)
}
results_htn_egfr$adj.p.value <- p.adjust(results_htn_egfr$p.value, "fdr")
results_htn_egfr$AptName <- results_htn_egfr$term
results_htn_egfr$term <- NULL
results_htn_egfr <- merge(results_htn_egfr,analytes,by="AptName",all.x = T, all.y = F)
results_htn_egfr <- results_htn_egfr[order(results_htn_egfr$p.value),] 

# neuropathy
result_NEURO <- NULL
predictors <- colnames(base_log_nhw[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTONEURO, NEURO)~",predictors[i],"+HbA1c+log_trig+sbp+si_1_ins0")
  mod <- coxph(data = base_log_scale_nhw, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_NEURO <- rbind(result_NEURO,tidymod)
}
result_NEURO$adj.p.value <- p.adjust(result_NEURO$p.value, "fdr")
result_NEURO$AptName <- result_NEURO$term
result_NEURO$term <- NULL
result_NEURO <- merge(result_NEURO,analytes,by="AptName",all.x = T, all.y = F)
result_NEURO <- result_NEURO[order(result_NEURO$p.value),] 

# retinopathy
result_RETINO <- NULL
predictors <- colnames(base_log_nhw[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTORETINO, RETINO)~",predictors[i],"+HbA1c+log_trig+sbp+si_1_ins0")
  mod <- coxph(data = base_log_scale_nhw, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_RETINO <- rbind(result_RETINO,tidymod)
}
result_RETINO$adj.p.value <- p.adjust(result_RETINO$p.value, "fdr")
result_RETINO$AptName <- result_RETINO$term
result_RETINO$term <- NULL
result_RETINO <- merge(result_RETINO,analytes,by="AptName",all.x = T, all.y = F)
result_RETINO <- result_RETINO[order(result_RETINO$p.value),] 

# glycemic failure
# omit A1c
result_GLYC <- NULL
predictors <- colnames(base_log_nhw[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOGLYC, GLYC)~",predictors[i],"+bmi+si_1_ins0+AGEBASE+sex")
  mod <- coxph(data = base_log_scale_nhw, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_GLYC <- rbind(result_GLYC,tidymod)
}
result_GLYC$adj.p.value <- p.adjust(result_GLYC$p.value, "fdr")
result_GLYC$AptName <- result_GLYC$term
result_GLYC$term <- NULL
result_GLYC <- merge(result_GLYC,analytes,by="AptName",all.x = T, all.y = F)
result_GLYC <- result_GLYC[order(result_GLYC$p.value),] 

# glycemic failure with A1c
results_glyc_a1c <- NULL
predictors <- colnames(base_log_nhw[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOGLYC, GLYC)~",predictors[i],"+HbA1c+bmi+si_1_ins0+AGEBASE+sex")
  mod <- coxph(data = base_log_scale_nhw, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  results_glyc_a1c <- rbind(results_glyc_a1c,tidymod)
}
results_glyc_a1c$adj.p.value <- p.adjust(results_glyc_a1c$p.value, "fdr")
results_glyc_a1c$AptName <- results_glyc_a1c$term
results_glyc_a1c$term <- NULL
results_glyc_a1c <- merge(results_glyc_a1c,analytes,by="AptName",all.x = T, all.y = F)
results_glyc_a1c <- results_glyc_a1c[order(results_glyc_a1c$p.value),] 

# write all results to file
wb <- createWorkbook()
addWorksheet(wb,"MAC CPH")
writeData(wb,"MAC CPH",result_MAC,rowNames = F)
addWorksheet(wb,"MIC CPH")
writeData(wb,"MIC CPH",result_MIC,rowNames = F)
addWorksheet(wb,"MIC.OR.MAC CPH")
writeData(wb,"MIC.OR.MAC CPH",result_MIC.OR.MAC,rowNames = F)
addWorksheet(wb,"HYP CPH")
writeData(wb,"HYP CPH",result_HYP,rowNames = F)
addWorksheet(wb,"RAPID CPH")
writeData(wb,"RAPID CPH",result_RAPID,rowNames = F)
addWorksheet(wb,"HTN with SBP CPH")
writeData(wb,"HTN with SBP CPH",results_htn_bp,rowNames = F)
addWorksheet(wb,"HTN CPH")
writeData(wb,"HTN CPH",result_HTN,rowNames = F)
addWorksheet(wb,"HTN with UACR CPH")
writeData(wb,"HTN with UACR CPH",results_htn_uacr,rowNames = F)
addWorksheet(wb,"HTN with eGFR CPH")
writeData(wb,"HTN with eGFR CPH",results_htn_egfr,rowNames = F)
addWorksheet(wb,"NEURO CPH")
writeData(wb,"NEURO CPH",result_NEURO,rowNames = F)
addWorksheet(wb,"RETINO CPH")
writeData(wb,"RETINO CPH",result_RETINO,rowNames = F)
addWorksheet(wb,"GLYC CPH")
writeData(wb,"GLYC CPH",result_GLYC,rowNames = F)
addWorksheet(wb,"GLYC with A1c CPH")
writeData(wb,"GLYC with A1c CPH",results_glyc_a1c,rowNames = F)
saveWorkbook(wb,"./Results/Linear and Cox models/TODAY somalogic Cox models scaled baseline adjusted NHW.xlsx",overwrite = TRUE)
saveWorkbook(wb,"/Users/pylell/Dropbox/TODAY HTN and DKD output/TODAY somalogic Cox models scaled baseline adjusted NHW.xlsx",overwrite = TRUE)
saveWorkbook(wb,"/Users/pylell/Dropbox/TODAY glycemic manuscript [shared]/Analysis output/TODAY somalogic Cox models scaled baseline adjusted NHW.xlsx",overwrite = TRUE)
```

```{r, include=FALSE}
# NHB

# Macroalbuminuria
result_MAC <- NULL
predictors <- colnames(base_log_nhb[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOMAC, MAC)~",predictors[i],"+HbA1c+log_trig+sbp+si_1_ins0")
  mod <- coxph(data = base_log_scale_nhb, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_MAC <- rbind(result_MAC,tidymod)
}
result_MAC$adj.p.value <- p.adjust(result_MAC$p.value, "fdr")
result_MAC$AptName <- result_MAC$term
result_MAC$term <- NULL
result_MAC <- merge(result_MAC,analytes,by="AptName",all.x = T, all.y = F)
result_MAC <- result_MAC[order(result_MAC$p.value),] 

# Microalbuminuria
result_MIC <- NULL
predictors <- colnames(base_log_nhb[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOMIC, MIC)~",predictors[i],"+HbA1c+log_trig+sbp+si_1_ins0")
  mod <- coxph(data = base_log_scale_nhb, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_MIC <- rbind(result_MIC,tidymod)
}
result_MIC$adj.p.value <- p.adjust(result_MIC$p.value, "fdr")
result_MIC$AptName <- result_MIC$term
result_MIC$term <- NULL
result_MIC <- merge(result_MIC,analytes,by="AptName",all.x = T, all.y = F)
result_MIC <- result_MIC[order(result_MIC$p.value),] 

# Microalbuminuria or macroalbuminuria - whichever is first
result_MIC.OR.MAC <- NULL
predictors <- colnames(base_log_nhb[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOMIC.OR.MAC, MIC.OR.MAC)~",predictors[i],"+HbA1c+log_trig+sbp+si_1_ins0")
  mod <- coxph(data = base_log_scale_nhb, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_MIC.OR.MAC <- rbind(result_MIC.OR.MAC,tidymod)
}
result_MIC.OR.MAC$adj.p.value <- p.adjust(result_MIC.OR.MAC$p.value, "fdr")
result_MIC.OR.MAC$AptName <- result_MIC.OR.MAC$term
result_MIC.OR.MAC$term <- NULL
result_MIC.OR.MAC <- merge(result_MIC.OR.MAC,analytes,by="AptName",all.x = T, all.y = F)
result_MIC.OR.MAC <- result_MIC.OR.MAC[order(result_MIC.OR.MAC$p.value),] 
result_MIC.OR.MAC_nhb <- result_MIC.OR.MAC

# hyperfiltration
result_HYP <- NULL
predictors <- colnames(base_log_nhb[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOHYP, HYP)~",predictors[i],"+HbA1c+log_trig+sbp+si_1_ins0")
  mod <- coxph(data = base_log_scale_nhb, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_HYP <- rbind(result_HYP,tidymod)
}
result_HYP$adj.p.value <- p.adjust(result_HYP$p.value, "fdr")
result_HYP$AptName <- result_HYP$term
result_HYP$term <- NULL
result_HYP <- merge(result_HYP,analytes,by="AptName",all.x = T, all.y = F)
result_HYP <- result_HYP[order(result_HYP$p.value),] 
result_HYP_nhb <- result_HYP

# rapid eGFR decline
result_RAPID <- NULL
predictors <- colnames(base_log_nhb[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTORAPID, RAPID)~",predictors[i],"+HbA1c+log_trig+sbp+si_1_ins0")
  mod <- coxph(data = base_log_scale_nhb, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_RAPID <- rbind(result_RAPID,tidymod)
}
result_RAPID$adj.p.value <- p.adjust(result_RAPID$p.value, "fdr")
result_RAPID$AptName <- result_RAPID$term
result_RAPID$term <- NULL
result_RAPID <- merge(result_RAPID,analytes,by="AptName",all.x = T, all.y = F)
result_RAPID <- result_RAPID[order(result_RAPID$p.value),] 
result_RAPID_nhb <- result_RAPID

# hypertension
# omit SBP
result_HTN <- NULL
predictors <- colnames(base_log_nhb[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOHTN, HTN)~",predictors[i],"+HbA1c+log_trig+si_1_ins0")
  mod <- coxph(data = base_log_scale_nhb, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_HTN <- rbind(result_HTN,tidymod)
}
result_HTN$adj.p.value <- p.adjust(result_HTN$p.value, "fdr")
result_HTN$AptName <- result_HTN$term
result_HTN$term <- NULL
result_HTN <- merge(result_HTN,analytes,by="AptName",all.x = T, all.y = F)
result_HTN <- result_HTN[order(result_HTN$p.value),] 

# hypertension with SBP
results_htn_bp <- NULL
predictors <- colnames(base_log_nhb[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOHTN, HTN)~",predictors[i],"+HbA1c+log_trig+sbp+si_1_ins0")
  mod <- coxph(data = base_log_scale_nhb, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  results_htn_bp <- rbind(results_htn_bp,tidymod)
}
results_htn_bp$adj.p.value <- p.adjust(results_htn_bp$p.value, "fdr")
results_htn_bp$AptName <- results_htn_bp$term
results_htn_bp$term <- NULL
results_htn_bp <- merge(results_htn_bp,analytes,by="AptName",all.x = T, all.y = F)
results_htn_bp <- results_htn_bp[order(results_htn_bp$p.value),] 

# hypertension with UACR
results_htn_uacr <- NULL
predictors <- colnames(base_log_nhb[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOHTN, HTN)~",predictors[i],"+HbA1c+log_trig+log_UAlbCreat+si_1_ins0")
  mod <- coxph(data = base_log_scale_nhb, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  results_htn_uacr <- rbind(results_htn_uacr,tidymod)
}
results_htn_uacr$adj.p.value <- p.adjust(results_htn_uacr$p.value, "fdr")
results_htn_uacr$AptName <- results_htn_uacr$term
results_htn_uacr$term <- NULL
results_htn_uacr <- merge(results_htn_uacr,analytes,by="AptName",all.x = T, all.y = F)
results_htn_uacr <- results_htn_uacr[order(results_htn_uacr$p.value),] 

# hypertension with eGFR
results_htn_egfr <- NULL
predictors <- colnames(base_log_nhb[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOHTN, HTN)~",predictors[i],"+HbA1c+log_trig+EstCreatClear+si_1_ins0")
  mod <- coxph(data = base_log_scale_nhb, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  results_htn_egfr <- rbind(results_htn_egfr,tidymod)
}
results_htn_egfr$adj.p.value <- p.adjust(results_htn_egfr$p.value, "fdr")
results_htn_egfr$AptName <- results_htn_egfr$term
results_htn_egfr$term <- NULL
results_htn_egfr <- merge(results_htn_egfr,analytes,by="AptName",all.x = T, all.y = F)
results_htn_egfr <- results_htn_egfr[order(results_htn_egfr$p.value),] 

# neuropathy
result_NEURO <- NULL
predictors <- colnames(base_log_nhb[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTONEURO, NEURO)~",predictors[i],"+HbA1c+log_trig+sbp+si_1_ins0")
  mod <- coxph(data = base_log_scale_nhb, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_NEURO <- rbind(result_NEURO,tidymod)
}
result_NEURO$adj.p.value <- p.adjust(result_NEURO$p.value, "fdr")
result_NEURO$AptName <- result_NEURO$term
result_NEURO$term <- NULL
result_NEURO <- merge(result_NEURO,analytes,by="AptName",all.x = T, all.y = F)
result_NEURO <- result_NEURO[order(result_NEURO$p.value),] 

# retinopathy
result_RETINO <- NULL
predictors <- colnames(base_log_nhb[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTORETINO, RETINO)~",predictors[i],"+HbA1c+log_trig+sbp+si_1_ins0")
  mod <- coxph(data = base_log_scale_nhb, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_RETINO <- rbind(result_RETINO,tidymod)
}
result_RETINO$adj.p.value <- p.adjust(result_RETINO$p.value, "fdr")
result_RETINO$AptName <- result_RETINO$term
result_RETINO$term <- NULL
result_RETINO <- merge(result_RETINO,analytes,by="AptName",all.x = T, all.y = F)
result_RETINO <- result_RETINO[order(result_RETINO$p.value),] 

# glycemic failure
# omit A1c
result_GLYC <- NULL
predictors <- colnames(base_log_nhb[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOGLYC, GLYC)~",predictors[i],"+bmi+si_1_ins0+AGEBASE+sex")
  mod <- coxph(data = base_log_scale_nhb, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_GLYC <- rbind(result_GLYC,tidymod)
}
result_GLYC$adj.p.value <- p.adjust(result_GLYC$p.value, "fdr")
result_GLYC$AptName <- result_GLYC$term
result_GLYC$term <- NULL
result_GLYC <- merge(result_GLYC,analytes,by="AptName",all.x = T, all.y = F)
result_GLYC <- result_GLYC[order(result_GLYC$p.value),] 

# glycemic failure with A1c
results_glyc_a1c <- NULL
predictors <- colnames(base_log_nhb[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOGLYC, GLYC)~",predictors[i],"+HbA1c+bmi+si_1_ins0+AGEBASE+sex")
  mod <- coxph(data = base_log_scale_nhb, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  results_glyc_a1c <- rbind(results_glyc_a1c,tidymod)
}
results_glyc_a1c$adj.p.value <- p.adjust(results_glyc_a1c$p.value, "fdr")
results_glyc_a1c$AptName <- results_glyc_a1c$term
results_glyc_a1c$term <- NULL
results_glyc_a1c <- merge(results_glyc_a1c,analytes,by="AptName",all.x = T, all.y = F)
results_glyc_a1c <- results_glyc_a1c[order(results_glyc_a1c$p.value),] 

# write all results to file
wb <- createWorkbook()
addWorksheet(wb,"MAC CPH")
writeData(wb,"MAC CPH",result_MAC,rowNames = F)
addWorksheet(wb,"MIC CPH")
writeData(wb,"MIC CPH",result_MIC,rowNames = F)
addWorksheet(wb,"MIC.OR.MAC CPH")
writeData(wb,"MIC.OR.MAC CPH",result_MIC.OR.MAC,rowNames = F)
addWorksheet(wb,"HYP CPH")
writeData(wb,"HYP CPH",result_HYP,rowNames = F)
addWorksheet(wb,"RAPID CPH")
writeData(wb,"RAPID CPH",result_RAPID,rowNames = F)
addWorksheet(wb,"HTN with SBP CPH")
writeData(wb,"HTN with SBP CPH",results_htn_bp,rowNames = F)
addWorksheet(wb,"HTN CPH")
writeData(wb,"HTN CPH",result_HTN,rowNames = F)
addWorksheet(wb,"HTN with UACR CPH")
writeData(wb,"HTN with UACR CPH",results_htn_uacr,rowNames = F)
addWorksheet(wb,"HTN with eGFR CPH")
writeData(wb,"HTN with eGFR CPH",results_htn_egfr,rowNames = F)
addWorksheet(wb,"NEURO CPH")
writeData(wb,"NEURO CPH",result_NEURO,rowNames = F)
addWorksheet(wb,"RETINO CPH")
writeData(wb,"RETINO CPH",result_RETINO,rowNames = F)
addWorksheet(wb,"GLYC CPH")
writeData(wb,"GLYC CPH",result_GLYC,rowNames = F)
addWorksheet(wb,"GLYC with A1c CPH")
writeData(wb,"GLYC with A1c CPH",results_glyc_a1c,rowNames = F)
saveWorkbook(wb,"./Results/Linear and Cox models/TODAY somalogic Cox models scaled baseline adjusted NHB.xlsx",overwrite = TRUE)
saveWorkbook(wb,"/Users/pylell/Dropbox/TODAY HTN and DKD output/TODAY somalogic Cox models scaled baseline adjusted NHB.xlsx",overwrite = TRUE)
saveWorkbook(wb,"/Users/pylell/Dropbox/TODAY glycemic manuscript [shared]/Analysis output/TODAY somalogic Cox models scaled baseline adjusted NHB.xlsx",overwrite = TRUE)
```


```{r, include=FALSE}
# VENN DIAGRAMS
# Overlap of proteins by race for each of the 3 outcomes

# First MIC.OR.MAC
result_MIC.OR.MAC_nom_sig_hispanic <- result_MIC.OR.MAC_hispanic %>% arrange(AptName)
result_MIC.OR.MAC_nom_sig_hispanic <- result_MIC.OR.MAC_nom_sig_hispanic$p.value < 0.05
result_MIC.OR.MAC_nom_sig_nhw <- result_MIC.OR.MAC_nhw %>% arrange(AptName)
result_MIC.OR.MAC_nom_sig_nhw <- result_MIC.OR.MAC_nom_sig_nhw$p.value < 0.05
result_MIC.OR.MAC_nom_sig_nhb <- result_MIC.OR.MAC_nhb %>% arrange(AptName)
result_MIC.OR.MAC_nom_sig_nhb <- result_MIC.OR.MAC_nom_sig_nhb$p.value < 0.05
v_MIC.OR.MAC_nom <- as.data.frame(cbind(result_MIC.OR.MAC_nom_sig_hispanic, result_MIC.OR.MAC_nom_sig_nhw, result_MIC.OR.MAC_nom_sig_nhb))
colnames(v_MIC.OR.MAC_nom) <- c("Hispanic", "Non-Hispanic White", "Non-Hispanic Black")

result_MIC.OR.MAC_fdr_sig_hispanic <- result_MIC.OR.MAC_hispanic %>% arrange(AptName)
result_MIC.OR.MAC_fdr_sig_hispanic <- result_MIC.OR.MAC_fdr_sig_hispanic$p.value < 0.05
result_MIC.OR.MAC_fdr_sig_nhw <- result_MIC.OR.MAC_nhw %>% arrange(AptName)
result_MIC.OR.MAC_fdr_sig_nhw <- result_MIC.OR.MAC_fdr_sig_nhw$p.value < 0.05
result_MIC.OR.MAC_fdr_sig_nhb <- result_MIC.OR.MAC_nhb %>% arrange(AptName)
result_MIC.OR.MAC_fdr_sig_nhb <- result_MIC.OR.MAC_fdr_sig_nhb$p.value < 0.05
v_MIC.OR.MAC_fdr <- as.data.frame(cbind(result_MIC.OR.MAC_fdr_sig_hispanic, result_MIC.OR.MAC_fdr_sig_nhw, result_MIC.OR.MAC_fdr_sig_nhb))
colnames(v_MIC.OR.MAC_fdr) <- c("Hispanic", "Non-Hispanic White", "Non-Hispanic Black")

# Hyperfiltration
result_HYP_nom_sig_hispanic <- result_HYP_hispanic %>% arrange(AptName)
result_HYP_nom_sig_hispanic <- result_HYP_nom_sig_hispanic$p.value < 0.05
result_HYP_nom_sig_nhw <- result_HYP_nhw %>% arrange(AptName)
result_HYP_nom_sig_nhw <- result_HYP_nom_sig_nhw$p.value < 0.05
result_HYP_nom_sig_nhb <- result_HYP_nhb %>% arrange(AptName)
result_HYP_nom_sig_nhb <- result_HYP_nom_sig_nhb$p.value < 0.05
v_HYP_nom <- as.data.frame(cbind(result_HYP_nom_sig_hispanic, result_HYP_nom_sig_nhw, result_HYP_nom_sig_nhb))
colnames(v_HYP_nom) <- c("Hispanic", "Non-Hispanic White", "Non-Hispanic Black")

result_HYP_fdr_sig_hispanic <- result_HYP_hispanic %>% arrange(AptName)
result_HYP_fdr_sig_hispanic <- result_HYP_fdr_sig_hispanic$p.value < 0.05
result_HYP_fdr_sig_nhw <- result_HYP_nhw %>% arrange(AptName)
result_HYP_fdr_sig_nhw <- result_HYP_fdr_sig_nhw$p.value < 0.05
result_HYP_fdr_sig_nhb <- result_HYP_nhb %>% arrange(AptName)
result_HYP_fdr_sig_nhb <- result_HYP_fdr_sig_nhb$p.value < 0.05
v_HYP_fdr <- as.data.frame(cbind(result_HYP_fdr_sig_hispanic, result_HYP_fdr_sig_nhw, result_HYP_fdr_sig_nhb))
colnames(v_HYP_fdr) <- c("Hispanic", "Non-Hispanic White", "Non-Hispanic Black")

# Rapid eGFR decline
result_RAPID_nom_sig_hispanic <- result_RAPID_hispanic %>% arrange(AptName)
result_RAPID_nom_sig_hispanic <- result_RAPID_nom_sig_hispanic$p.value < 0.05
result_RAPID_nom_sig_nhw <- result_RAPID_nhw %>% arrange(AptName)
result_RAPID_nom_sig_nhw <- result_RAPID_nom_sig_nhw$p.value < 0.05
result_RAPID_nom_sig_nhb <- result_RAPID_nhb %>% arrange(AptName)
result_RAPID_nom_sig_nhb <- result_RAPID_nom_sig_nhb$p.value < 0.05
v_RAPID_nom <- as.data.frame(cbind(result_RAPID_nom_sig_hispanic, result_RAPID_nom_sig_nhw, result_RAPID_nom_sig_nhb))
colnames(v_RAPID_nom) <- c("Hispanic", "Non-Hispanic White", "Non-Hispanic Black")

result_RAPID_fdr_sig_hispanic <- result_RAPID_hispanic %>% arrange(AptName)
result_RAPID_fdr_sig_hispanic <- result_RAPID_fdr_sig_hispanic$p.value < 0.05
result_RAPID_fdr_sig_nhw <- result_RAPID_nhw %>% arrange(AptName)
result_RAPID_fdr_sig_nhw <- result_RAPID_fdr_sig_nhw$p.value < 0.05
result_RAPID_fdr_sig_nhb <- result_RAPID_nhb %>% arrange(AptName)
result_RAPID_fdr_sig_nhb <- result_RAPID_fdr_sig_nhb$p.value < 0.05
v_RAPID_fdr <- as.data.frame(cbind(result_RAPID_fdr_sig_hispanic, result_RAPID_fdr_sig_nhw, result_RAPID_fdr_sig_nhb))
colnames(v_RAPID_fdr) <- c("Hispanic", "Non-Hispanic White", "Non-Hispanic Black")
```

```{r, include=FALSE}
# VENN DIAGRAM FOR MIC.OR.MAC

# get list of proteins significant for each race
a <- result_MIC.OR.MAC_hispanic[result_MIC.OR.MAC_hispanic$p.value < 0.05,]
a$hispanic.sig <- TRUE
a <- a %>% select(AptName, hispanic.sig, Target)
b <- result_MIC.OR.MAC_nhw[result_MIC.OR.MAC_nhw$p.value < 0.05,]
b$nhw.sig <- TRUE
b <- b %>% select(AptName, nhw.sig)
c <- result_MIC.OR.MAC_nhb[result_MIC.OR.MAC_nhb$p.value < 0.05,]
c$nhb.sig <- TRUE
c <- c %>% select(AptName, nhb.sig)
ab <- left_join(a, b, by="AptName")
abc <- left_join(ab, c, by="AptName")

# pull out results for proteins significant for all race/ethnicities
abc_sig <- abc %>% filter(hispanic.sig & nhw.sig & nhb.sig)
result_hispanic_intersect <- result_MIC.OR.MAC_hispanic %>% filter(AptName %in% abc_sig$AptName)
result_hispanic_intersect$outcome <- "Hispanic"
result_nhw_intersect <- result_MIC.OR.MAC_nhw %>% filter(AptName %in% abc_sig$AptName)
result_nhw_intersect$outcome <- "Non-Hispanic White"
result_nhb_intersect <- result_MIC.OR.MAC_nhb %>% filter(AptName %in% abc_sig$AptName)
result_nhb_intersect$outcome <- "Non-Hispanic Black"
result_intersect_temp <- rbind(result_hispanic_intersect, result_nhw_intersect, result_nhb_intersect)
result_intersect <- result_intersect_temp %>% select(outcome, AptName, Target, TargetFullName, estimate,
                                                     std.error, p.value, adj.p.value)
#write.csv(result_intersect, "/Volumes/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/TODAY subaward//Results/Linear and Cox models/TODAY somalogic Cox models scaled baseline adjusted intersection.csv",
#          row.names = F)

# plots including all proteins
vennp_MIC.OR.MAC_nom <- ggvenn(v_MIC.OR.MAC_nom, 
  c("Hispanic", "Non-Hispanic White", "Non-Hispanic Black"),
  fill_color = c("#c2dfe3", "#fff9ec","#fcb1a6"),
  stroke_size = 0.5, set_name_size = 6, digits = 2, text_size = 3
)
vennp_MIC.OR.MAC_nom + theme(plot.margin = margin(1,1,1,1, 'cm'))
png(file = "/Users/pylell/Dropbox/TODAY HTN and DKD output/venn_nominally_significant_albuminuria_by_ethnicity.png", res = 300,
    height = 2000, width = 2000)
vennp_MIC.OR.MAC_nom
dev.off()
vennp_MIC.OR.MAC_fdr <- ggvenn(v_MIC.OR.MAC_fdr, 
  c("Hispanic", "Non-Hispanic White", "Non-Hispanic Black"),
  fill_color = c("#c2dfe3", "#fff9ec","#fcb1a6"),
  stroke_size = 0.5, set_name_size = 6
)
```


```{r, include=FALSE}
# VENN DIAGRAM FOR HYPERFILTRATION

# get list of proteins significant for each race
a <- result_HYP_hispanic[result_HYP_hispanic$p.value < 0.05,]
a$hispanic.sig <- TRUE
a <- a %>% select(AptName, hispanic.sig, Target)
b <- result_HYP_nhw[result_HYP_nhw$p.value < 0.05,]
b$nhw.sig <- TRUE
b <- b %>% select(AptName, nhw.sig)
c <- result_HYP_nhb[result_HYP_nhb$p.value < 0.05,]
c$nhb.sig <- TRUE
c <- c %>% select(AptName, nhb.sig)
ab <- left_join(a, b, by="AptName")
abc <- left_join(ab, c, by="AptName")

# pull out results for proteins significant for all race/ethnicities
abc_sig <- abc %>% filter(hispanic.sig & nhw.sig & nhb.sig)
result_hispanic_intersect <- result_MIC.OR.MAC_hispanic %>% filter(AptName %in% abc_sig$AptName)
result_hispanic_intersect$outcome <- "Hispanic"
result_nhw_intersect <- result_MIC.OR.MAC_nhw %>% filter(AptName %in% abc_sig$AptName)
result_nhw_intersect$outcome <- "Non-Hispanic White"
result_nhb_intersect <- result_MIC.OR.MAC_nhb %>% filter(AptName %in% abc_sig$AptName)
result_nhb_intersect$outcome <- "Non-Hispanic Black"
result_intersect_temp <- rbind(result_hispanic_intersect, result_nhw_intersect, result_nhb_intersect)
result_intersect <- result_intersect_temp %>% select(outcome, AptName, Target, TargetFullName, estimate,
                                                     std.error, p.value, adj.p.value)
#write.csv(result_intersect, "/Volumes/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/TODAY subaward//Results/Linear and Cox models/TODAY somalogic Cox models scaled baseline adjusted intersection.csv",
#          row.names = F)

# plots including all proteins
vennp_HYP_nom <- ggvenn(v_HYP_nom, 
  c("Hispanic", "Non-Hispanic White", "Non-Hispanic Black"),
  fill_color = c("#c2dfe3", "#fff9ec","#fcb1a6"),
  stroke_size = 0.5, set_name_size = 6, digits = 2, text_size = 3
)
vennp_HYP_nom + theme(plot.margin = margin(1,1,1,1, 'cm'))
png(file = "/Users/pylell/Dropbox/TODAY HTN and DKD output/venn_nominally_significant_hyperfiltration_by_ethnicity.png", res = 300,
    height = 2000, width = 2000)
vennp_HYP_nom
dev.off()
vennp_HYP_fdr <- ggvenn(v_HYP_fdr, 
  c("Hispanic", "Non-Hispanic White", "Non-Hispanic Black"),
  fill_color = c("#c2dfe3", "#fff9ec","#fcb1a6"),
  stroke_size = 0.5, set_name_size = 6
)
```

```{r, include=FALSE}
# VENN DIAGRAM FOR RAPID EGFR DECLINE

# get list of proteins significant for each race
a <- result_RAPID_hispanic[result_RAPID_hispanic$p.value < 0.05,]
a$hispanic.sig <- TRUE
a <- a %>% select(AptName, hispanic.sig, Target)
b <- result_RAPID_nhw[result_RAPID_nhw$p.value < 0.05,]
b$nhw.sig <- TRUE
b <- b %>% select(AptName, nhw.sig)
c <- result_RAPID_nhb[result_RAPID_nhb$p.value < 0.05,]
c$nhb.sig <- TRUE
c <- c %>% select(AptName, nhb.sig)
ab <- left_join(a, b, by="AptName")
abc <- left_join(ab, c, by="AptName")

# pull out results for proteins significant for all race/ethnicities
abc_sig <- abc %>% filter(hispanic.sig & nhw.sig & nhb.sig)
result_hispanic_intersect <- result_MIC.OR.MAC_hispanic %>% filter(AptName %in% abc_sig$AptName)
result_hispanic_intersect$outcome <- "Hispanic"
result_nhw_intersect <- result_MIC.OR.MAC_nhw %>% filter(AptName %in% abc_sig$AptName)
result_nhw_intersect$outcome <- "Non-Hispanic White"
result_nhb_intersect <- result_MIC.OR.MAC_nhb %>% filter(AptName %in% abc_sig$AptName)
result_nhb_intersect$outcome <- "Non-Hispanic Black"
result_intersect_temp <- rbind(result_hispanic_intersect, result_nhw_intersect, result_nhb_intersect)
result_intersect <- result_intersect_temp %>% select(outcome, AptName, Target, TargetFullName, estimate,
                                                     std.error, p.value, adj.p.value)
#write.csv(result_intersect, "/Volumes/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/TODAY subaward//Results/Linear and Cox models/TODAY somalogic Cox models scaled baseline adjusted intersection.csv",
#          row.names = F)

# plots including all proteins
vennp_RAPID_nom <- ggvenn(v_RAPID_nom, 
  c("Hispanic", "Non-Hispanic White", "Non-Hispanic Black"),
  fill_color = c("#c2dfe3", "#fff9ec","#fcb1a6"),
  stroke_size = 0.5, set_name_size = 6, digits = 2, text_size = 3
)
vennp_RAPID_nom + theme(plot.margin = margin(2,2,2,2, 'cm'))
png(file = "/Users/pylell/Dropbox/TODAY HTN and DKD output/venn_nominally_significant_egfr_decline_by_ethnicity.png", res = 300,
    height = 2000, width = 2000)
vennp_RAPID_nom
dev.off()
vennp_RAPID_fdr <- ggvenn(v_RAPID_fdr, 
  c("Hispanic", "Non-Hispanic White", "Non-Hispanic Black"),
  fill_color = c("#c2dfe3", "#fff9ec","#fcb1a6"),
  stroke_size = 0.5, set_name_size = 6
)

p <- ggarrange(vennp_MIC.OR.MAC_nom, vennp_HYP_nom, vennp_RAPID_nom, ncol = 1, nrow = 3, align = "hv", labels = c("A", "B", "C"))

png('/Volumes/PEDS/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/Proteomics and DKD/Figures/Venn_diagram_Cox_models_by_ethnicity.png', res = 600, width = 6, height = 18, units = "in")
p
dev.off()
png('/Users/pylell/Dropbox/TODAY HTN and DKD output/Venn_diagram_Cox_models_by_ethnicity.png', res = 600, width = 6, height = 18, units = "in")
p
dev.off()
```

# Results

## Interactions with sex

### Severe albuminuria and NELL1

```{r echo=FALSE, include=TRUE, comment=""}
base_log_scale$sex <- as.factor(base_log_scale$sex)
# NELL1
mod <- coxph(data = base_log_scale, formula = Surv(DAYSTOMAC, MAC)~ HbA1c+log_trig+sbp+si_1_ins0 + seq.6544.33*sex)
tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
tidymod <- tidymod %>% mutate_if(is.double, round, 4)
kable(tidymod)
```

### Severe albuminuria and CI061

```{r echo=FALSE, include=TRUE, comment=""}
# CI061
mod <- coxph(data = base_log_scale, formula = Surv(DAYSTOMAC, MAC)~ HbA1c+log_trig+sbp+si_1_ins0 + seq.5699.19*sex)
tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
tidymod <- tidymod %>% mutate_if(is.double, round, 4)
kable(tidymod)
```

### Severe albuminuria and PSB3

```{r echo=FALSE, include=TRUE, comment=""}
# PSB3
mod <- coxph(data = base_log_scale, formula = Surv(DAYSTOMAC, MAC)~ HbA1c+log_trig+sbp+si_1_ins0 + seq.18339.207*sex)
tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
tidymod <- tidymod %>% mutate_if(is.double, round, 4)
kable(tidymod)
```

### Moderate albuminuria and NELL1

```{r echo=FALSE, include=TRUE, comment=""}
# NELL1
mod <- coxph(data = base_log_scale, formula = Surv(DAYSTOMIC, MIC)~ HbA1c+log_trig+sbp+si_1_ins0 + seq.6544.33*sex)
tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
tidymod <- tidymod %>% mutate_if(is.double, round, 4)
kable(tidymod)
```

### Moderate albuminuria and CI061

```{r echo=FALSE, include=TRUE, comment=""}
# CI061
mod <- coxph(data = base_log_scale, formula = Surv(DAYSTOMIC, MIC)~ HbA1c+log_trig+sbp+si_1_ins0 + seq.5699.19*sex)
tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
tidymod <- tidymod %>% mutate_if(is.double, round, 4)
kable(tidymod)
```

### Moderate albuminuria and CTHR1

```{r echo=FALSE, include=TRUE, comment=""}
# CTHR1
mod <- coxph(data = base_log_scale, formula = Surv(DAYSTOMIC, MIC)~ HbA1c+log_trig+sbp+si_1_ins0 + seq.15467.10*sex)
tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
tidymod <- tidymod %>% mutate_if(is.double, round, 4)
kable(tidymod)
# p = 0.06!
```

### Rapid eGFR decline and ZPLD1

The interaction between sex and ZPLD1 was significant. In the model results below, the estimate for seq.5590.11 is the hazard ratio for sex = 1 (i.e., females).

```{r echo=FALSE, include=TRUE, comment=""}
# ZPLD1
mod <- coxph(data = base_log_scale, formula = Surv(DAYSTORAPID, RAPID)~ HbA1c+log_trig+sbp+si_1_ins0 + seq.5590.11*sex)
tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
tidymod <- tidymod %>% mutate_if(is.double, round, 4)
kable(tidymod)
# p = 0.03
```

In the model results below, the estimate for seq.5590.11 is the hazard ratio for sex = 2 (i.e., males).

```{r echo=FALSE, include=TRUE, comment=""}
base_log_scale$sex <- as.factor(base_log_scale$sex)
base_log_scale$sex <- relevel(base_log_scale$sex, ref="2")
# ZPLD1
mod <- coxph(data = base_log_scale, formula = Surv(DAYSTORAPID, RAPID)~ HbA1c+log_trig+sbp+si_1_ins0 + seq.5590.11*sex)
tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
tidymod <- tidymod %>% mutate_if(is.double, round, 4)
kable(tidymod)
# p = 0.03
base_log_scale$sex <- relevel(base_log_scale$sex, ref="1")
```

### Rapid eGFR decline and RELM-beta

```{r echo=FALSE, include=TRUE, comment=""}
# RELM-beta
mod <- coxph(data = base_log_scale, formula = Surv(DAYSTORAPID, RAPID)~ HbA1c+log_trig+sbp+si_1_ins0 + seq.22007.1*sex)
tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
tidymod <- tidymod %>% mutate_if(is.double, round, 4)
kable(tidymod)
```

### Rapid eGFR decline and CLU

```{r echo=FALSE, include=TRUE, comment=""}
# CLU
mod <- coxph(data = base_log_scale, formula = Surv(DAYSTORAPID, RAPID)~ HbA1c+log_trig+sbp+si_1_ins0 + seq.24941.14*sex)
tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
tidymod <- tidymod %>% mutate_if(is.double, round, 4)
kable(tidymod)
```

### Hyperfiltration and Cystatin C

```{r echo=FALSE, include=TRUE, comment=""}
# Cystatin C
mod <- coxph(data = base_log_scale, formula = Surv(DAYSTOHYP, HYP)~ HbA1c+log_trig+sbp+si_1_ins0 + seq.2609.59*sex)
tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
tidymod <- tidymod %>% mutate_if(is.double, round, 4)
kable(tidymod)
```

### Hyperfiltration and PEDF, aptamer 1

```{r echo=FALSE, include=TRUE, comment=""}
# PEDF
mod <- coxph(data = base_log_scale, formula = Surv(DAYSTOHYP, HYP)~ HbA1c+log_trig+sbp+si_1_ins0 + seq.9211.19*sex)
tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
tidymod <- tidymod %>% mutate_if(is.double, round, 4)
kable(tidymod)
```

### Hyperfiltration and PEDF, aptamer 2

```{r echo=FALSE, include=TRUE, comment=""}
# PEDF
mod <- coxph(data = base_log_scale, formula = Surv(DAYSTOHYP, HYP)~ HbA1c+log_trig+sbp+si_1_ins0 + seq.7735.17*sex)
tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
tidymod <- tidymod %>% mutate_if(is.double, round, 4)
kable(tidymod)
# p = 0.06
```

### Hyperfiltration and b2-microglobulin

```{r echo=FALSE, include=TRUE, comment=""}
# b2-Microglobulin
mod <- coxph(data = base_log_scale, formula = Surv(DAYSTOHYP, HYP)~ HbA1c+log_trig+sbp+si_1_ins0 + seq.3485.28*sex)
tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
tidymod <- tidymod %>% mutate_if(is.double, round, 4)
kable(tidymod)
```

### Hypertension and WFKN2

```{r echo=FALSE, include=TRUE, comment=""}
# WFKN2
mod <- coxph(data = base_log_scale, formula = Surv(DAYSTOHTN, HTN)~ HbA1c+log_trig+sbp+si_1_ins0 + seq.13408.23*sex)
tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
tidymod <- tidymod %>% mutate_if(is.double, round, 4)
kable(tidymod)
```

### Hypertension and SEZ6L

```{r echo=FALSE, include=TRUE, comment=""}
# SEZ6L
mod <- coxph(data = base_log_scale, formula = Surv(DAYSTOHTN, HTN)~ HbA1c+log_trig+sbp+si_1_ins0 + seq.19563.3*sex)
tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
tidymod <- tidymod %>% mutate_if(is.double, round, 4)
kable(tidymod)
```

### Hypertension and SCG3

```{r echo=FALSE, include=TRUE, comment=""}
# SCG3
mod <- coxph(data = base_log_scale, formula = Surv(DAYSTOHTN, HTN)~ HbA1c+log_trig+sbp+si_1_ins0 + seq.7957.2*sex)
tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
tidymod <- tidymod %>% mutate_if(is.double, round, 4)
kable(tidymod)
```

```{r, include=FALSE}
# REPEAT COX MODELS, ADDING TX GROUP
# Macroalbuminuria
result_MAC <- NULL
predictors <- colnames(base_log[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOMAC, MAC)~",predictors[i],"+HbA1c+log_trig+sbp+si_1_ins0+tx")
  mod <- coxph(data = base_log_scale, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_MAC <- rbind(result_MAC,tidymod)
}
result_MAC$adj.p.value <- p.adjust(result_MAC$p.value, "fdr")
result_MAC$AptName <- result_MAC$term
result_MAC$term <- NULL
result_MAC <- merge(result_MAC,analytes,by="AptName",all.x = T, all.y = F)
result_MAC <- result_MAC[order(result_MAC$p.value),] 

# Microalbuminuria
result_MIC <- NULL
predictors <- colnames(base_log[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOMIC, MIC)~",predictors[i],"+HbA1c+log_trig+sbp+si_1_ins0+tx")
  mod <- coxph(data = base_log_scale, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_MIC <- rbind(result_MIC,tidymod)
}
result_MIC$adj.p.value <- p.adjust(result_MIC$p.value, "fdr")
result_MIC$AptName <- result_MIC$term
result_MIC$term <- NULL
result_MIC <- merge(result_MIC,analytes,by="AptName",all.x = T, all.y = F)
result_MIC <- result_MIC[order(result_MIC$p.value),] 

# Microalbuminuria or macroalbuminuria - whichever is first
result_MIC.OR.MAC <- NULL
predictors <- colnames(base_log[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOMIC.OR.MAC, MIC.OR.MAC)~",predictors[i],"+HbA1c+log_trig+sbp+si_1_ins0+tx")
  mod <- coxph(data = base_log_scale, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_MIC.OR.MAC <- rbind(result_MIC.OR.MAC,tidymod)
}
result_MIC.OR.MAC$adj.p.value <- p.adjust(result_MIC.OR.MAC$p.value, "fdr")
result_MIC.OR.MAC$AptName <- result_MIC.OR.MAC$term
result_MIC.OR.MAC$term <- NULL
result_MIC.OR.MAC <- merge(result_MIC.OR.MAC,analytes,by="AptName",all.x = T, all.y = F)
result_MIC.OR.MAC <- result_MIC.OR.MAC[order(result_MIC.OR.MAC$p.value),] 

# hyperfiltration
result_HYP <- NULL
predictors <- colnames(base_log[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOHYP, HYP)~",predictors[i],"+HbA1c+log_trig+sbp+si_1_ins0+tx")
  mod <- coxph(data = base_log_scale, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_HYP <- rbind(result_HYP,tidymod)
}
result_HYP$adj.p.value <- p.adjust(result_HYP$p.value, "fdr")
result_HYP$AptName <- result_HYP$term
result_HYP$term <- NULL
result_HYP <- merge(result_HYP,analytes,by="AptName",all.x = T, all.y = F)
result_HYP <- result_HYP[order(result_HYP$p.value),] 

# rapid eGFR decline
result_RAPID <- NULL
predictors <- colnames(base_log[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTORAPID, RAPID)~",predictors[i],"+HbA1c+log_trig+sbp+si_1_ins0+tx")
  mod <- coxph(data = base_log_scale, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_RAPID <- rbind(result_RAPID,tidymod)
}
result_RAPID$adj.p.value <- p.adjust(result_RAPID$p.value, "fdr")
result_RAPID$AptName <- result_RAPID$term
result_RAPID$term <- NULL
result_RAPID <- merge(result_RAPID,analytes,by="AptName",all.x = T, all.y = F)
result_RAPID <- result_RAPID[order(result_RAPID$p.value),] 

# hypertension
# omit SBP
result_HTN <- NULL
predictors <- colnames(base_log[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOHTN, HTN)~",predictors[i],"+HbA1c+log_trig+si_1_ins0+tx")
  mod <- coxph(data = base_log_scale, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_HTN <- rbind(result_HTN,tidymod)
}
result_HTN$adj.p.value <- p.adjust(result_HTN$p.value, "fdr")
result_HTN$AptName <- result_HTN$term
result_HTN$term <- NULL
result_HTN <- merge(result_HTN,analytes,by="AptName",all.x = T, all.y = F)
result_HTN <- result_HTN[order(result_HTN$p.value),] 

# hypertension with SBP
results_htn_bp <- NULL
predictors <- colnames(base_log[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOHTN, HTN)~",predictors[i],"+HbA1c+log_trig+sbp+si_1_ins0+tx")
  mod <- coxph(data = base_log_scale, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  results_htn_bp <- rbind(results_htn_bp,tidymod)
}
results_htn_bp$adj.p.value <- p.adjust(results_htn_bp$p.value, "fdr")
results_htn_bp$AptName <- results_htn_bp$term
results_htn_bp$term <- NULL
results_htn_bp <- merge(results_htn_bp,analytes,by="AptName",all.x = T, all.y = F)
results_htn_bp <- results_htn_bp[order(results_htn_bp$p.value),] 

# hypertension with UACR
results_htn_uacr <- NULL
predictors <- colnames(base_log[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOHTN, HTN)~",predictors[i],"+HbA1c+log_trig+log_UAlbCreat+si_1_ins0+tx")
  mod <- coxph(data = base_log_scale, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  results_htn_uacr <- rbind(results_htn_uacr,tidymod)
}
results_htn_uacr$adj.p.value <- p.adjust(results_htn_uacr$p.value, "fdr")
results_htn_uacr$AptName <- results_htn_uacr$term
results_htn_uacr$term <- NULL
results_htn_uacr <- merge(results_htn_uacr,analytes,by="AptName",all.x = T, all.y = F)
results_htn_uacr <- results_htn_uacr[order(results_htn_uacr$p.value),] 

# hypertension with eGFR
results_htn_egfr <- NULL
predictors <- colnames(base_log[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOHTN, HTN)~",predictors[i],"+HbA1c+log_trig+EstCreatClear+si_1_ins0+tx")
  mod <- coxph(data = base_log_scale, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  results_htn_egfr <- rbind(results_htn_egfr,tidymod)
}
results_htn_egfr$adj.p.value <- p.adjust(results_htn_egfr$p.value, "fdr")
results_htn_egfr$AptName <- results_htn_egfr$term
results_htn_egfr$term <- NULL
results_htn_egfr <- merge(results_htn_egfr,analytes,by="AptName",all.x = T, all.y = F)
results_htn_egfr <- results_htn_egfr[order(results_htn_egfr$p.value),] 

# neuropathy
result_NEURO <- NULL
predictors <- colnames(base_log[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTONEURO, NEURO)~",predictors[i],"+HbA1c+log_trig+sbp+si_1_ins0+tx")
  mod <- coxph(data = base_log_scale, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_NEURO <- rbind(result_NEURO,tidymod)
}
result_NEURO$adj.p.value <- p.adjust(result_NEURO$p.value, "fdr")
result_NEURO$AptName <- result_NEURO$term
result_NEURO$term <- NULL
result_NEURO <- merge(result_NEURO,analytes,by="AptName",all.x = T, all.y = F)
result_NEURO <- result_NEURO[order(result_NEURO$p.value),] 

# retinopathy
result_RETINO <- NULL
predictors <- colnames(base_log[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTORETINO, RETINO)~",predictors[i],"+HbA1c+log_trig+sbp+si_1_ins0+tx")
  mod <- coxph(data = base_log_scale, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_RETINO <- rbind(result_RETINO,tidymod)
}
result_RETINO$adj.p.value <- p.adjust(result_RETINO$p.value, "fdr")
result_RETINO$AptName <- result_RETINO$term
result_RETINO$term <- NULL
result_RETINO <- merge(result_RETINO,analytes,by="AptName",all.x = T, all.y = F)
result_RETINO <- result_RETINO[order(result_RETINO$p.value),] 

# glycemic failure
# omit A1c
result_GLYC <- NULL
predictors <- colnames(base_log[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOGLYC, GLYC)~",predictors[i],"+bmi+si_1_ins0+tx")
  mod <- coxph(data = base_log_scale, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_GLYC <- rbind(result_GLYC,tidymod)
}
result_GLYC$adj.p.value <- p.adjust(result_GLYC$p.value, "fdr")
result_GLYC$AptName <- result_GLYC$term
result_GLYC$term <- NULL
result_GLYC <- merge(result_GLYC,analytes,by="AptName",all.x = T, all.y = F)
result_GLYC <- result_GLYC[order(result_GLYC$p.value),] 

# glycemic failure with A1c
results_glyc_a1c <- NULL
predictors <- colnames(base_log[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOGLYC, GLYC)~",predictors[i],"+HbA1c+bmi+si_1_ins0+tx")
  mod <- coxph(data = base_log_scale, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  results_glyc_a1c <- rbind(results_glyc_a1c,tidymod)
}
results_glyc_a1c$adj.p.value <- p.adjust(results_glyc_a1c$p.value, "fdr")
results_glyc_a1c$AptName <- results_glyc_a1c$term
results_glyc_a1c$term <- NULL
results_glyc_a1c <- merge(results_glyc_a1c,analytes,by="AptName",all.x = T, all.y = F)
results_glyc_a1c <- results_glyc_a1c[order(results_glyc_a1c$p.value),] 

# write all results to file
wb <- createWorkbook()
addWorksheet(wb,"MAC CPH")
writeData(wb,"MAC CPH",result_MAC,rowNames = F)
addWorksheet(wb,"MIC CPH")
writeData(wb,"MIC CPH",result_MIC,rowNames = F)
addWorksheet(wb,"MIC.OR.MAC CPH")
writeData(wb,"MIC.OR.MAC CPH",result_MIC.OR.MAC,rowNames = F)
addWorksheet(wb,"HYP CPH")
writeData(wb,"HYP CPH",result_HYP,rowNames = F)
addWorksheet(wb,"RAPID CPH")
writeData(wb,"RAPID CPH",result_RAPID,rowNames = F)
addWorksheet(wb,"HTN with SBP CPH")
writeData(wb,"HTN with SBP CPH",results_htn_bp,rowNames = F)
addWorksheet(wb,"HTN CPH")
writeData(wb,"HTN CPH",result_HTN,rowNames = F)
addWorksheet(wb,"HTN with UACR CPH")
writeData(wb,"HTN with UACR CPH",results_htn_uacr,rowNames = F)
addWorksheet(wb,"HTN with eGFR CPH")
writeData(wb,"HTN with eGFR CPH",results_htn_egfr,rowNames = F)
addWorksheet(wb,"NEURO CPH")
writeData(wb,"NEURO CPH",result_NEURO,rowNames = F)
addWorksheet(wb,"RETINO CPH")
writeData(wb,"RETINO CPH",result_RETINO,rowNames = F)
addWorksheet(wb,"GLYC CPH")
writeData(wb,"GLYC CPH",result_GLYC,rowNames = F)
addWorksheet(wb,"GLYC with A1c CPH")
writeData(wb,"GLYC with A1c CPH",results_glyc_a1c,rowNames = F)
saveWorkbook(wb,"/Volumes/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/TODAY subaward/Results/Linear and Cox models/TODAY somalogic Cox models scaled baseline adjusted TX GROUP.xlsx",overwrite = TRUE)
saveWorkbook(wb,"/Users/pylell/Dropbox/TODAY HTN and DKD output/TODAY somalogic Cox models scaled baseline adjusted TX GROUP.xlsx",overwrite = TRUE)


```