---
title: "TODAY Somalogic baseline analyses - elastic net Cox models for glycemia"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
knit: (function(rmdfile, ...) { rmarkdown::render(rmdfile, output_dir='/Users/pylell/Dropbox/TODAY glycemic manuscript [shared]/') })
output: rtf_document
---

```{r setup, include=FALSE}
library(SomaDataIO)
library(limma)
library(dplyr)
library(caret)
library(purrr)
library(multtest)
library(openxlsx)
library(tableone)
library(EnhancedVolcano)
library(knitr)
library(survival)
library(broom)
library(plsRcox)
library(forestmodel)
library(survC1)
library(naniar)
library(rms)
library(readxl)
library(forester)
library(grid)
library(forestploter)
library(doParallel)
library(foreach)
library(readr)
library(nestedcv)

knitr::opts_chunk$set(echo = FALSE,warning = FALSE)

knitr::opts_chunk$set(echo = FALSE)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "E:/Petter Bjornstad/TODAY subaward"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/TODAY subaward/"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/TODAY subaward"
}
knitr::opts_knit$set(root.dir = home_dir)

```

```{r, include=FALSE}
setwd(home_dir)
# load somalogic data, with QC samples already excluded
load("./Somalogic data raw/soma.Rdata")

# load analyte info
load("./Somalogic data raw/analytes.Rdata")

# load comorbidity data
load("./Clinical data/comorb.Rdata")

# load baseline risk factors
load("./Clinical data/TODAY/baserisk.Rdata")

# take only the baseline soma samples
# can we just take the earliest or is it possible someone would have a follow-up sample but not baseline?
# probably can take the first and then check the years to make sure compatible with TODAY
base <- soma %>% arrange(releaseid,Date.Drawn) %>% group_by(releaseid) %>% filter(row_number()==1)

# these 3 release IDs have first sample after end of recruitment, so they must be missing the baseline visit
base <- base %>% filter(!releaseid %in% c("65-85903","65-47984","65-25901"))

# merge in complication data
base <- left_join(base, comorb, by="releaseid")

# merge in baseline risk factors
base <- left_join(base, baserisk, by="releaseid")

# identify columns corresponding to proteins
#is_seq <- function(.x) grepl("^seq\\.[0-9]{4}", .x) # regex for analytes
is_seq <- function(.x) grepl("seq", .x)
seq <- is_seq(names(base))

# convert to numeric
base[,seq] <- apply(base[,seq],2,as.numeric)
seq_names <- colnames(base[,seq])

# are there proteins with low variability?
no_var = caret::nearZeroVar(base[,seq])
# none

# log transform
base_log <- base %>% modify_if(is_seq(names(.)), log)

# scale by SD
base_log_scale <- base_log
predictors <- colnames(base_log[seq])
for (i in 1:length(predictors)) {
  base_log_scale[,paste0(predictors[i])] <- base_log_scale[,paste0(predictors[i])]/sd(unlist(base_log[,paste0(predictors[i])]))
}

set.seed(3654)
source("~/Documents/Github/CHCO-Code/Petter Bjornstad/TODAY proteomics metabolomics/Proteomics analyses/easy_elasticnet.R")
```

# Loss of glycemic control

## Elastic net with 10-fold cross-validation

```{r nestedcv,echo=TRUE }
file.remove("/Users/pylell/Dropbox/TODAY glycemic manuscript [shared]/Analysis output/glycemia_feature_selection.csv")

data_model <- base_log_scale %>% 
    select(releaseid, GLYC, DAYSTOGLYC, seq_names)  %>% ungroup()

y <- data_model %>% select(GLYC, DAYSTOGLYC)
x <- data_model %>% select(all_of(seq_names))

# try to write a screening function for univariable cox models
# should take x and y and return character vector of names of filtered predictors
cox_screen <- function(x, y) {
  data <- cbind(y, x)
  data <- cbind(data, base_log_scale$bmi, base_log_scale$si_1_ins0, base_log_scale$AGEBASE, base_log_scale$sex)
  predictors <- seq_names

  result_GLYC <- NULL
  for (i in 1:length(predictors)) {
    form <- paste0("Surv(DAYSTOGLYC, GLYC)~",predictors[i],"+bmi+si_1_ins0+AGEBASE+sex")
    mod <- coxph(data = base_log_scale, formula = as.formula(form))
    tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
    tidymod <- tidymod[1,]
    result_GLYC <- rbind(result_GLYC,tidymod)
  }
  result_GLYC$adj.p.value <- p.adjust(result_GLYC$p.value, "fdr")
  result_GLYC$AptName <- result_GLYC$term
  result_GLYC <- merge(result_GLYC,analytes,by="AptName",all.x = T, all.y = F)
  result_GLYC <- result_GLYC[order(result_GLYC$p.value),] 
  result_GLYC <- result_GLYC[1:100,]
  result_GLYC_vec <- result_GLYC$AptName
  return(result_GLYC_vec)
}

result_GLYC <- cox_screen(x, y)
```

```{r checking feature selection stability,echo=TRUE }
library(doSNOW)
getDoParWorkers()
getDoParName()
registerDoSNOW(makeCluster(16, type = "SOCK"))
getDoParWorkers()
getDoParName()

file.remove("/Users/pylell/Dropbox/TODAY glycemic manuscript [shared]/Analysis output/glycemia_feature_selection.csv")

data_model <- base_log_scale %>% 
    select(releaseid, GLYC, DAYSTOGLYC, seq_names)  %>% ungroup()
  
foreach(j = 1:5, .packages = c("survival", "broom", "readr")) %dopar% {
  print(j)
 
  # Getting strata training and test data.
  split <- initial_split(solar_data_model, prob = 0.7,
                         strata ="diabete_change_di")

  train <- training(split) %>% mutate(id = "train")

  y_train <- train %>% dplyr::select(all_of(outcomes)) %>%
    as.matrix()

  x_train <- train %>%
    dplyr::select(all_of(features),
                  all_of(covars1[-4]),
                  all_of(covars2)
                  ) %>%
    fastDummies::dummy_cols(remove_selected_columns = TRUE,
                          remove_first_dummy = TRUE) %>%
    as.matrix()
  
  # univariable Cox models
  result_GLYC <- NULL
  predictors <- seq_names
  for (i in 1:length(predictors)) {
    form <- paste0("Surv(DAYSTOGLYC, GLYC)~",predictors[i],"+bmi+si_1_ins0+AGEBASE+sex")
    mod <- coxph(data = base_log_scale, formula = as.formula(form))
    tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
    tidymod <- tidymod[1,]
    result_GLYC <- rbind(result_GLYC,tidymod)
  }
  result_GLYC$adj.p.value <- p.adjust(result_GLYC$p.value, "fdr")
  result_GLYC$AptName <- result_GLYC$term
  result_GLYC$term <- NULL
  result_GLYC <- merge(result_GLYC,analytes,by="AptName",all.x = T, all.y = F)
  result_GLYC <- result_GLYC[order(result_GLYC$p.value),] 
  result_GLYC <- cbind(result_GLYC, rep(j, nrow(result_GLYC)))

write_csv(result_GLYC, "/Users/pylell/Dropbox/TODAY glycemic manuscript [shared]/Analysis output/glycemia_feature_selection.csv",
           append = TRUE, col_names = F)  
  
#   y_train <- data_model %>% dplyr::select(GLYC, DAYSTOGLYC) %>%
#     as.matrix()
# 
#   x_train <- data_model %>%
#     dplyr::select(all_of(seq_names)) %>%
#     as.matrix()
# 
#   ctrl <- trainControl(
#     method = "cv",
#     number  = 10,
#     classProbs = TRUE,
#     summaryFunction = twoClassSummary
#   )
# 
#   fit <- train(
#     x = x_train,
#     y = y_train,
#     method = "glmnet",
#     trControl = ctrl,
#     metric = "ROC",
#     tuneGrid = expand.grid(alpha = 1, lambda = seq(0.001, 0.1, by = 0.001)),
#     penalty.factor = c(rep(1,length(features)), rep(0, 7))
#   )
# 
#   result <- as.data.frame(as.matrix(coef(fit$finalModel, fit$bestTune$lambda))) %>%
#     rownames_to_column() %>%
#     rename(coef = s1,
#            feature = rowname) %>%
#     mutate(outcome = "diabete_change_di")
# 
#   result_selected_feature <- result %>% filter(coef != 0)
# 
# write_csv(result_selected_feature,
#           fs::path(dir_results,
#                  "prediction",
#                  "4_feature_selection_diabete_5000_1122.csv"),
#           append = TRUE)
}
stopCluster(cl)
```

## Selected features
**Features selected in at least 85% of the 5,000 iterations were included in the prediction models**
```{r selecting features, echo=TRUE}
## Load selected feature file
feature <- read_csv(
            fs::path(dir_results, 
                   "prediction",
                   "4_feature_selection_diabete_5000_1122.csv")
,col_names = FALSE) %>%
    dplyr::rename(feature = X1,
           coefficient = X2,
           outcome = X3)

## Selected features that appear at least 4250 times of the replications
feature_n <- feature %>% 
  group_by(feature) %>% 
  dplyr::summarize(count =n()) %>%
  arrange(desc(count)) %>%
  mutate(percentage = count/5000)

# write_csv(feature_n,fs::path(dir_results, 
#                    "prediction",
#                    "4_feature_selection_percentage.csv"))
# feature_n %>% filter(!feature %in% c(covars1, covars2, "sex_1_Male") & feature != "(Intercept)"& percentage > 0.25) %>% ggplot(aes(x=percentage)) + 
#   geom_vline(aes(xintercept = 0.85), color = "blue", linetype = "dashed")+
#   # geom_vline(aes(xintercept = 0.9), color = "blue", linetype = "dashed")+
#   geom_histogram(fill = "white", color = "black", bins = 100) +
#   labs(title = "Distribution of Percentage", x = "Percentage", y = "Count") +
#   scale_x_continuous(limits = c(0.2, 1), breaks = seq(0.2,1,0.1)) +
#   theme_classic()
# 
# ggsave(filename = fs::path(dir_reports, "4_feature_selection_distribution.png"), 
#        width = 10, height = 6, dpi=300)

feature_80 <- feature_n %>% filter(count > 0.85*5000)

selected_feature_name <- feature_80$feature[-c(1:8)] # 3
```

```{r old code, include=FALSE, comment=""}

# filter to top 100 proteins by effect size
f = "/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/TODAY subaward/Results/Linear and Cox models/TODAY somalogic Cox models scaled baseline adjusted.xlsx"

# GLYCEMIA
glyc_keep = read_excel(f,sheet = "GLYC CPH")
glyc_keep <- arrange(glyc_keep,adj.p.value)
glyc_drop <- glyc_keep[101:nrow(glyc_keep),"AptName"]
rownames(glyc_drop) <- glyc_drop$AptName
glyc_drop <- as.data.frame(t(glyc_drop))
glyc <- base_log_scale %>% dplyr::select(-colnames(glyc_drop))

concordance_df_glyc <- NULL

glyc_nozero <- glyc[!glyc$DAYSTOGLYC == 0,]
seq <- is_seq(names(glyc))

glyc_select = easy_elasticnet(data = glyc_nozero,outcome = "GLYC",
                           predictors = glyc_nozero[,seq], out = "min.error",
                           model_type = "cox",time = "DAYSTOGLYC", cv_method="kfold", folds = 10)
retained_vars = analytes[analytes$AptName %in% glyc_select, c("Target","TargetFullName")]

form = as.formula(paste0("Surv(DAYSTOGLYC, GLYC)~",paste0(glyc_select,collapse = "+")))
mod = coxph(formula = form,data = glyc_nozero)
kable(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)

# make a summary table
moddf <- as.data.frame(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)
moddf$Target <- analytes[analytes$AptName %in% moddf$term, ]$Target
moddf <- cbind(moddf$Target,moddf$term,round(moddf$estimate, digits=3),round(moddf$std.error, digits=3),round(moddf$p.value, digits=3))
colnames(moddf) <- c("Target","Apt Name","Hazard Ratio","SE","P-value")
concordance_df_glyc <- as.data.frame(cbind("Loss of glycemic control","Proteins",mod$concordance[[6]], mod$concordance[[7]]))
```

```{r, include=TRUE, comment=""}
kable(moddf, row.names=F)
```

There are `r nrow(moddf)` variables in the model.

```{r, include=TRUE, comment=""}
kable(round(mod$concordance,digits=3))
```

```{r, include=TRUE, comment=""}
# make forest plot summarizing HRs for selected proteins
elastic_net_for_plot <- as.data.frame(moddf)
elastic_net_for_plot[,3:5] <- apply(elastic_net_for_plot[,3:5], 2, as.numeric)
elastic_net_for_plot$Target <- ifelse(elastic_net_for_plot$Target == "aldolase C", "Aldolase C", elastic_net_for_plot$Target)
elastic_net_for_plot$`Lower bound 95% CI` <- elastic_net_for_plot$`Hazard Ratio` - (1.96 * elastic_net_for_plot$SE)
elastic_net_for_plot$`Upper bound 95% CI` <- elastic_net_for_plot$`Hazard Ratio` + (1.96 * elastic_net_for_plot$SE)
elastic_net_for_plot <- elastic_net_for_plot %>% arrange(`P-value`)
elastic_net_for_plot$Target <- paste0(elastic_net_for_plot$Target, " (", elastic_net_for_plot$`Apt Name`, ")")
elastic_net_for_plot$` ` <- paste(rep(" ", 30), collapse = " ")
elastic_net_for_plot$`HR (95% CI)` <- sprintf("%.2f (%.2f to %.2f)",
                                     elastic_net_for_plot$`Hazard Ratio`, elastic_net_for_plot$`Lower bound 95% CI`, elastic_net_for_plot$`Upper bound 95% CI`)
elastic_net_for_plot_p <- forest(elastic_net_for_plot[,c("Target"," ","HR (95% CI)")], est = elastic_net_for_plot$`Hazard Ratio`, lower = elastic_net_for_plot$`Lower bound 95% CI`, upper = elastic_net_for_plot$`Upper bound 95% CI`, ci_column = 2, xlim = c(0, 2), ref_line = 1)
png('/Users/pylell/Dropbox/TODAY glycemic manuscript [shared]/Analysis output/forest_plot_elastic_net_proteins.png', res = 300, width = 10, height = 15, units = "in")
elastic_net_for_plot_p
dev.off()
```

## Model with proteins selected by elastic net and covariates

```{r, include=FALSE, comment=""}
form = as.formula(paste0("Surv(DAYSTOGLYC, GLYC)~",paste0(glyc_select,collapse = "+"),paste0("+bmi+si_1_ins0+AGEBASE+sex")))
mod = coxph(formula = form,data = glyc_nozero)
kable(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)

# make a summary table
moddf <- as.data.frame(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)
moddf$Target <- NA
moddf[1:(nrow(moddf)-4),]$Target <- analytes[analytes$AptName %in% moddf[1:(nrow(moddf)-4),]$term, ]$Target
moddf$Target <- paste0(moddf$Target, " (", moddf$`term`, ")")
moddf$Target <- ifelse(moddf$Target == "NA (bmi)", "BMI", moddf$Target)
moddf$Target <- ifelse(moddf$Target == "NA (si_1_ins0)", "Insulin sensitivity", moddf$Target)
moddf$Target <- ifelse(moddf$Target == "NA (AGEBASE)", "Age", moddf$Target)
moddf$Target <- ifelse(moddf$Target == "NA (sex)", "Sex", moddf$Target)
moddf <- cbind(moddf$Target,moddf$term,round(moddf$estimate, digits=3),round(moddf$std.error, digits=3),round(moddf$p.value, digits=3))
colnames(moddf) <- c("Target","Apt Name","Hazard Ratio","SE","P-value")
concordance_df_temp <- as.data.frame(cbind("Loss of glycemic control","Proteins + covariates (BMI, Si, age, sex)",mod$concordance[[6]], mod$concordance[[7]]))
concordance_df_glyc <- rbind(concordance_df_glyc, concordance_df_temp)
```

```{r, include=TRUE, comment=""}
kable(moddf, row.names=F)
```

There are `r nrow(moddf)` variables in the model.

```{r, include=TRUE, comment=""}
kable(round(mod$concordance,digits=3))
```

```{r, include=TRUE, comment=""}
# make forest plot summarizing HRs for selected proteins and covariates
elastic_net_for_plot <- as.data.frame(moddf)
elastic_net_for_plot[,3:5] <- apply(elastic_net_for_plot[,3:5], 2, as.numeric)
elastic_net_for_plot$Target <- ifelse(elastic_net_for_plot$Target == "aldolase C (seq.9876.20)", "Aldolase C (seq.9876.20)", elastic_net_for_plot$Target)
elastic_net_for_plot$`Lower bound 95% CI` <- elastic_net_for_plot$`Hazard Ratio` - (1.96 * elastic_net_for_plot$SE)
elastic_net_for_plot$`Upper bound 95% CI` <- elastic_net_for_plot$`Hazard Ratio` + (1.96 * elastic_net_for_plot$SE)
elastic_net_for_plot <- elastic_net_for_plot %>% arrange(`P-value`)
elastic_net_for_plot$` ` <- paste(rep(" ", 30), collapse = " ")
elastic_net_for_plot$`HR (95% CI)` <- sprintf("%.2f (%.2f to %.2f)",
                                     elastic_net_for_plot$`Hazard Ratio`, elastic_net_for_plot$`Lower bound 95% CI`, elastic_net_for_plot$`Upper bound 95% CI`)
elastic_net_for_plot_p <- forest(elastic_net_for_plot[,c("Target"," ","HR (95% CI)")], est = elastic_net_for_plot$`Hazard Ratio`, lower = elastic_net_for_plot$`Lower bound 95% CI`, upper = elastic_net_for_plot$`Upper bound 95% CI`, ci_column = 2, xlim = c(0, 26), ref_line = 1)
png('/Users/pylell/Dropbox/TODAY glycemic manuscript [shared]/Analysis output/forest_plot_elastic_net_proteins_covariates.png', res = 300, width = 10, height = 15, units = "in")
elastic_net_for_plot_p
dev.off()
```

## Model with only covariates

```{r, include=FALSE, comment=""}
form = as.formula(paste0("Surv(DAYSTOGLYC, GLYC)~bmi+si_1_ins0+AGEBASE+sex"))
mod = coxph(formula = form,data = glyc_nozero)
kable(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)

# make a summary table
moddf <- as.data.frame(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)
moddf <- cbind(moddf$term,round(moddf$estimate, digits=3),round(moddf$std.error, digits=3),round(moddf$p.value, digits=3))
colnames(moddf) <- c("Variable","Hazard Ratio","SE","P-value")
concordance_df_temp <- as.data.frame(cbind("Loss of glycemic control","Covariates (BMI, Si, age, sex)",mod$concordance[[6]], mod$concordance[[7]]))
concordance_df_glyc <- rbind(concordance_df_glyc, concordance_df_temp)
```

```{r, include=TRUE, comment=""}
kable(moddf, row.names=F)
```

There are `r nrow(moddf)` variables in the model.

```{r, include=TRUE, comment=""}
kable(round(mod$concordance,digits=3))
```

## Summary tables and plots

```{r, include=FALSE, comment=""}
colnames(concordance_df_glyc) <- c("Outcome","Model","Concordance","SE")
concordance_df_glyc$Concordance <- as.numeric(concordance_df_glyc$Concordance)
concordance_df_glyc$SE <- as.numeric(concordance_df_glyc$SE)
concordance_df_glyc$ci_low <- concordance_df_glyc$Concordance - (1.96*concordance_df_glyc$SE)
concordance_df_glyc$ci_high <- concordance_df_glyc$Concordance + (1.96*concordance_df_glyc$SE)
concordance_df_glyc[,c("Concordance","SE","ci_low", "ci_high")] <- round(concordance_df_glyc[,c("Concordance","SE","ci_low", "ci_high")],2)
colnames(concordance_df_glyc) <- c("Outcome","Model","Concordance","SE","Lower bound 95% CI","Upper bound 95% CI")
concordance_df_glyc <- rbind(concordance_df_glyc[3,],concordance_df_glyc[1,],concordance_df_glyc[2,])

```

```{r, include=TRUE, comment=""}
# glycemia forest plot summarizing concordance
concordance_df_for_plot <- concordance_df_glyc
concordance_df_for_plot$` ` <- paste(rep(" ", 40), collapse = " ")
concordance_df_for_plot$`Concordance (95% CI)` <- sprintf("%.2f (%.2f to %.2f)",
                                     concordance_df_for_plot$Concordance, concordance_df_for_plot$`Lower bound 95% CI`, concordance_df_for_plot$`Upper bound 95% CI`)
p <- forest(concordance_df_for_plot[, c(1:2, 7:8)], est = concordance_df_for_plot$Concordance, lower = concordance_df_for_plot$`Lower bound 95% CI`,
       upper = concordance_df_for_plot$`Upper bound 95% CI`, ci_column = 3, xlim = c(0,1))
#p$widths <- unit(c(1,100,50,50), "mm")
png('/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/Proteomics and glycemia/Figures/forest_plot_elastic_net_glycemia_concordance.png', res = 300, width = 9, height = 3, units = "in")
p
dev.off()
png('/Users/pylell/Dropbox/TODAY glycemic manuscript [shared]/Analysis output/forest_plot_elastic_net_glycemia_concordance.png', res = 300, width = 11, height = 3, units = "in")
p
dev.off()
```

# Forest plots for univariate results
```{r, include=FALSE}
# forest plot of overall results
f = "/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/TODAY subaward/Results/Linear and Cox models/TODAY somalogic Cox models scaled baseline adjusted.xlsx"
glyc_forest = read_excel(f,sheet = "GLYC CPH")
glyc_forest <- glyc_forest %>% arrange(adj.p.value)
glyc_forest <- glyc_forest[1:20,]
glyc_forest$Target <- ifelse(glyc_forest$Target == "carboxylesterase, liver", "Liver carboxylesterase", glyc_forest$Target)
glyc_forest$Target <- ifelse(glyc_forest$Target == "aldolase C", "Aldolase C", glyc_forest$Target)
glyc_forest$Target <- paste0(glyc_forest$Target, " (", glyc_forest$AptName, ")")
glyc_forest$` ` <- paste(rep(" ", 30), collapse = " ")
glyc_forest$`HR (95% CI)` <- sprintf("%.2f (%.2f to %.2f)",
                                     glyc_forest$estimate, glyc_forest$conf.low, glyc_forest$conf.high)
glyc_forest_p <- forest(glyc_forest[,c("Target"," ","HR (95% CI)")], est = glyc_forest$estimate, lower = glyc_forest$conf.low, upper = glyc_forest$conf.high, ci_column = 2, xlim = c(0, 3.5), ref_line = 1)
png('/Users/pylell/Dropbox/TODAY glycemic manuscript [shared]/Analysis output/forest_plot_univariate_CPH_overall.png', res = 300, width = 10, height = 10, units = "in")
glyc_forest_p
dev.off()

# forest plot by ethnicity
# Hispanic
f = "/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/TODAY subaward/Results/Linear and Cox models/TODAY somalogic Cox models scaled baseline adjusted Hispanic.xlsx"
glyc_forest = read_excel(f,sheet = "GLYC CPH")
glyc_forest <- glyc_forest %>% arrange(adj.p.value)
glyc_forest <- glyc_forest[1:20,]
glyc_forest$Target <- ifelse(glyc_forest$Target == "carboxylesterase, liver", "Liver carboxylesterase", glyc_forest$Target)
glyc_forest$Target <- ifelse(glyc_forest$Target == "aldolase C", "Aldolase C", glyc_forest$Target)
glyc_forest$Target <- paste0(glyc_forest$Target, " (", glyc_forest$AptName, ")")
glyc_forest$` ` <- paste(rep(" ", 30), collapse = " ")
glyc_forest$`HR (95% CI)` <- sprintf("%.2f (%.2f to %.2f)",
                                     glyc_forest$estimate, glyc_forest$conf.low, glyc_forest$conf.high)
glyc_forest_p <- forest(glyc_forest[,c("Target"," ","HR (95% CI)")], est = glyc_forest$estimate, lower = glyc_forest$conf.low, upper = glyc_forest$conf.high, ci_column = 2, xlim = c(0, 3.5), ref_line = 1)
png('/Users/pylell/Dropbox/TODAY glycemic manuscript [shared]/Analysis output/forest_plot_univariate_CPH_Hispanic.png', res = 300, width = 10, height = 10, units = "in")
glyc_forest_p
dev.off()

# NHB
f = "/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/TODAY subaward/Results/Linear and Cox models/TODAY somalogic Cox models scaled baseline adjusted NHB.xlsx"
glyc_forest = read_excel(f,sheet = "GLYC CPH")
glyc_forest <- glyc_forest %>% arrange(adj.p.value)
glyc_forest <- glyc_forest[1:20,]
glyc_forest$Target <- ifelse(glyc_forest$Target == "carboxylesterase, liver", "Liver carboxylesterase", glyc_forest$Target)
glyc_forest$Target <- ifelse(glyc_forest$Target == "aldolase C", "Aldolase C", glyc_forest$Target)
glyc_forest$Target <- paste0(glyc_forest$Target, " (", glyc_forest$AptName, ")")
glyc_forest$` ` <- paste(rep(" ", 30), collapse = " ")
glyc_forest$`HR (95% CI)` <- sprintf("%.2f (%.2f to %.2f)",
                                     glyc_forest$estimate, glyc_forest$conf.low, glyc_forest$conf.high)
glyc_forest_p <- forest(glyc_forest[,c("Target"," ","HR (95% CI)")], est = glyc_forest$estimate, lower = glyc_forest$conf.low, upper = glyc_forest$conf.high, ci_column = 2, xlim = c(0, 3.5), ref_line = 1)
png('/Users/pylell/Dropbox/TODAY glycemic manuscript [shared]/Analysis output/forest_plot_univariate_CPH_NHB.png', res = 300, width = 10, height = 10, units = "in")
glyc_forest_p
dev.off()

# NHW
f = "/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/TODAY subaward/Results/Linear and Cox models/TODAY somalogic Cox models scaled baseline adjusted NHW.xlsx"
glyc_forest = read_excel(f,sheet = "GLYC CPH")
glyc_forest <- glyc_forest %>% arrange(adj.p.value)
glyc_forest <- glyc_forest[1:20,]
glyc_forest$Target <- ifelse(glyc_forest$Target == "carboxylesterase, liver", "Liver carboxylesterase", glyc_forest$Target)
glyc_forest$Target <- ifelse(glyc_forest$Target == "aldolase C", "Aldolase C", glyc_forest$Target)
glyc_forest$Target <- paste0(glyc_forest$Target, " (", glyc_forest$AptName, ")")
glyc_forest$` ` <- paste(rep(" ", 30), collapse = " ")
glyc_forest$`HR (95% CI)` <- sprintf("%.2f (%.2f to %.2f)",
                                     glyc_forest$estimate, glyc_forest$conf.low, glyc_forest$conf.high)
glyc_forest_p <- forest(glyc_forest[,c("Target"," ","HR (95% CI)")], est = glyc_forest$estimate, lower = glyc_forest$conf.low, upper = glyc_forest$conf.high, ci_column = 2, xlim = c(0, 3.5), ref_line = 1)
png('/Users/pylell/Dropbox/TODAY glycemic manuscript [shared]/Analysis output/forest_plot_univariate_CPH_NHW.png', res = 300, width = 10, height = 10, units = "in")
glyc_forest_p
dev.off()
```


