---
title: "ULTRA Analysis"
author: "Callie Rountree-Jablin"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 3
    toc_float: yes
    toc_loc: left
---

```{r setup, include=FALSE}
library(tidyverse)
library(gtsummary)
library(kableExtra)
library(ggplot2)
library(Hmisc)
library(patchwork)
pacman::p_load(naniar)

if(Sys.info()["sysname"] == "Windows"){
  home_dir = "B:/Petter Bjornstad"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/Peds Endo/Petter Bjornstad/"
} 
knitr::opts_knit$set(echo=FALSE, root.dir = home_dir)
```

```{r read, include=FALSE}
# read raw data
raw_dat <- read_csv("./ULTRA/raw_data/ULTRAT2D_DATA_2023-10-19_1426.csv")
raw_dat <- raw_dat %>% 
  select(-c(name_last, name_first, mrn, dob, phone))
renal_flow_dat <- readxl::read_xlsx("./ULTRA/raw_data/renal_flow_dat.xlsx")
```

```{r data cleaning, include=FALSE}
## subset full raw df into separate screening and visit dfs for analysis

# screening cleaning
screening_dat <- raw_dat %>%
  filter(redcap_event_name=='screening_visit_arm_1') %>%
  mutate(Race=case_when(race___1==1~'American Indian or Alaskan Native',
                        race___2==1~'Asian',
                        race___3==1~'Hawaiian or Pacific Islander',
                        race___4==1~'Black or African American',
                        race___5==1~'White',
                        race___6==1~'Unknown',
                        race___7==1~'Other',
                        .default=NA),
         Ethnicity=case_when(ethnicity___1==1~'Hispanic',
                             ethnicity___2==1~'Non-Hispanic',
                             ethnicity___3==1~'Unknown/Not Reported',
                             .default=NA),
         t2d_less_21=factor(t2d_less_21, levels=c(0,1), labels=c('No','Yes')),
         diabetes_meds___1=factor(diabetes_meds___1, levels=c(0,1), labels=c('No','Yes')),
         diabetes_meds___2=factor(diabetes_meds___2, levels=c(0,1), labels=c('No','Yes')),
         diabetes_meds___3=factor(diabetes_meds___3, levels=c(0,1), labels=c('No','Yes')),
         diabetes_meds___4=factor(diabetes_meds___4, levels=c(0,1), labels=c('No','Yes')),
         diabetes_meds___5=factor(diabetes_meds___5, levels=c(0,1), labels=c('No','Yes')),
         diabetes_meds___6=factor(diabetes_meds___6, levels=c(0,1), labels=c('No','Yes')),
         diabetes_meds___7=factor(diabetes_meds___7, levels=c(0,1), labels=c('No','Yes')),
         insulin_type=case_when(insulin_type___1==1~'Long acting',
                                insulin_type___2==1~'Short acting',
                                insulin_type___3==1~'Other',
                                .default=NA),
         med_hx_hypertension=factor(med_hx_hypertension, levels=c(0,1), labels=c('No','Yes')),
         hypertension_meds=factor(hypertension_meds, levels=c(0,1), labels=c('No','Yes')),
         hypertension_med_type___1=factor(hypertension_med_type___1, levels=c(0,1), labels=c('No','Yes')),
         hypertension_med_type___2=factor(hypertension_med_type___2, levels=c(0,1), labels=c('No','Yes')),
         hypertension_med_type___3=factor(hypertension_med_type___3, levels=c(0,1), labels=c('No','Yes')),
         hypertension_med_type___4=factor(hypertension_med_type___4, levels=c(0,1), labels=c('No','Yes')),
         hypertension_med_type___5=factor(hypertension_med_type___5, levels=c(0,1), labels=c('No','Yes')),
         hypertension_med_type___6=factor(hypertension_med_type___6, levels=c(0,1), labels=c('No','Yes')),
         hypertension_med_type___7=factor(hypertension_med_type___7, levels=c(0,1), labels=c('No','Yes')),
         medhx_cvd=factor(medhx_cvd, levels=c(0,1), labels=c('No','Yes')),
         medhx_met=factor(medhx_met, levels=c(0,1), labels=c('No','Yes')),
         met_hx___1=factor(met_hx___1, levels=c(0,1), labels=c('No','Yes')),
         met_hx___3=factor(met_hx___3, levels=c(0,1), labels=c('No','Yes')),
         met_hx___4=factor(met_hx___4, levels=c(0,1), labels=c('No','Yes')),
         met_hx___5=factor(met_hx___5, levels=c(0,1), labels=c('No','Yes')),
         metdx_meds=factor(metdx_meds, levels=c(0,1), labels=c('No','Yes')),
         g6pd=factor(g6pd, levels=c(2,1), labels=c('Negative','Positive')),
         diabetes_yrs=(interval(as.Date(diabetes_diag, format='%Y-%m-%d'),
                                as.Date(consent_date, format='%Y-%m-%d'))/years(1))) %>% 
  select(-c(redcap_event_name:redcap_repeat_instance, 
            race___1:ethnicity___3,
            studyvisit_type:imaging_complete,
            insulin_type___1:insulin_other)) %>% 
  relocate(insulin_type, .before=insulin_tdd) %>% 
  relocate(c(Race,Ethnicity), .after=age) %>% 
  relocate(diabetes_yrs, .after=diabetes_months) %>%
  filter(record_id!='UT2D-02') %>% 
  filter(record_id!='UT2D-04')

# add labels
source('B:/Petter Bjornstad/ULTRA/raw_data/RedCap_labels_screeningdat.r')

# study visit cleaning
visit_dat <-raw_dat %>% 
  filter(redcap_event_name=='study_visit_arm_1') %>% 
  filter(record_id!='UT2D-02') %>% 
  mutate(studyvisit_type=factor(studyvisit_type, levels=c(1,2), labels=c('Baseline','1 Week'))) %>% 
  select(-c(redcap_repeat_instrument:screening_labs_complete)) %>%
  # HOMA-IR calculation (fasting insulin (mcIU/mL) * fasting glucose (mg/dl)/405)
  mutate(HOMA_IR = (labs_insulin * labs_bg)/405)

# drop peak strain values of 0 per Sungho
visit_dat$long_peak[visit_dat$long_peak == 0] <- NA
visit_dat$radial_peak[visit_dat$radial_peak == 0] <- NA
visit_dat$circum_peak[visit_dat$circum_peak == 0] <- NA

# add labels
source('B:/Petter Bjornstad/ULTRA/raw_data/RedCap_labels_visitdat.r')
Hmisc::label(visit_dat$HOMA_IR) <- "HOMA-IR" 

```

<!-- paper referenced for HOMA-IR calculation: https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5018567/#:~:text=HOMA%2DIR%20and%20HOMA%2DB,diabetes%20risk%20across%20ethnic%20groups.&text=The%20formulas%20are%20as%20follows,%2Fdl)%20%E2%80%93%2063%5D. -->

```{r all subjects visit missingness summary, include=FALSE}
## look into missingness patterns 

visit_dat %>%
  select(c(record_id, studyvisit_type, vitals_weight, vitals_sbp, vitals_dbp, vitals_map,
         vitals_pulse, labs_bg, labs_s_na, labs_s_k, labs_s_cl, labs_s_hc03, labs_glucose, 
         labs_bun, labs_s_ca, labs_tot_prot, labs_albumin, labs_bili, labs_alk_phos, labs_got_ast, 
         labs_gpt_alt, labs_sua, labs_u_na, labs_u_glucose, labs_uua, labs_cystatinc, 
         labs_s_creatinine, labs_insulin, labs_cholesterol, labs_hdl, labs_ldl,
         labs_triglycerides, labs_u_creatinine, labs_u_microalbumin, labs_acr,
         pilabs_s_osmo, pilabs_u_osmo, pilabs_copeptin, egfr, iohexol_abs_gfr,
         iohexol_bsa_gfr, iohexol_ecv, iohexol_ecv_gfr, iohexol_gfr_ecv_std,
         edv_ml, esv_ml, sv_ml, ef_percent, cardiac_output, myo_mass_dias,
         myo_mass_syst, radial_peak, circum_peak, long_peak, HOMA_IR)) %>% 
  miss_var_summary() %>% 
  kable()

vars <- c('studyvisit_type', 'vitals_weight', 'vitals_sbp', 'vitals_dbp', 'vitals_map',
'vitals_pulse', 'labs_bg', 'labs_s_na', 'labs_s_k', 'labs_s_cl', 'labs_s_hc03', 'labs_glucose', 
'labs_bun', 'labs_s_ca', 'labs_tot_prot', 'labs_albumin', 'labs_bili', 'labs_alk_phos', 'labs_got_ast', 
'labs_gpt_alt', 'labs_sua', 'labs_u_na', 'labs_u_glucose', 'labs_uua', 'labs_cystatinc', 'labs_s_creatinine', 
'labs_insulin', 'labs_cholesterol', 'labs_hdl', 'labs_ldl', 'labs_triglycerides', 'labs_u_creatinine', 
'labs_u_microalbumin', 'labs_acr', 'pilabs_s_osmo', 'pilabs_u_osmo', 'pilabs_copeptin', 'egfr', 'iohexol_abs_gfr', 
'iohexol_bsa_gfr', 'iohexol_ecv', 'iohexol_ecv_gfr', 'iohexol_gfr_ecv_std', 'edv_ml', 'esv_ml', 'sv_ml', 'ef_percent', 
'cardiac_output', 'myo_mass_dias', 'myo_mass_syst', 'radial_peak', 'circum_peak', 'long_peak', 'HOMA_IR')

miss_plot <- expss::use_labels(visit_dat[,vars],
                              {gg_miss_var(..data, facet=studyvisit_type)})
```

```{r wide format, include=FALSE}
## create df with wide format for use with later plots/analyses

wide_visit <- visit_dat %>% 
  pivot_wider(id_cols=record_id,
              names_from=studyvisit_type,
              names_prefix='Study_Visit_',
              values_from=c(vitals_sbp, vitals_dbp, vitals_map, labs_sua, labs_uua, labs_bili, iohexol_abs_gfr,
                            iohexol_bsa_gfr, edv_ml, long_peak, HOMA_IR))
```


```{r responders dat, include=FALSE}
## create responders only dfs for sensitivity analyses

# subset data to include responders only (Delta SUA >= 3 mg/dL)
wide_visit <- wide_visit %>% 
  group_by(record_id) %>% 
  mutate(abs_delta_SUA=abs(`labs_sua_Study_Visit_1 Week`-labs_sua_Study_Visit_Baseline)) %>%
  ungroup() %>% 
  relocate(abs_delta_SUA, .after=`labs_sua_Study_Visit_1 Week`)

label(wide_visit$abs_delta_SUA) <- 'Week 1 - Baseline SUA (Abs Value)'

SUA_status <- wide_visit %>% 
  select(record_id, abs_delta_SUA)

screening_dat <- left_join(screening_dat, SUA_status, by='record_id')
responders_screening_dat <- screening_dat %>% 
  filter(abs_delta_SUA >=3)

visit_dat <- left_join(visit_dat, SUA_status, by='record_id')
responders_visit_dat <- visit_dat %>% 
  filter(abs_delta_SUA >= 3)

rm(SUA_status)
```

```{r responders visit missingness summary, message=FALSE, include=FALSE}
## look into missingness patterns 

responders_visit_dat %>%
  select(c(record_id, studyvisit_type, vitals_weight, vitals_sbp, vitals_dbp, vitals_map,
         vitals_pulse, labs_bg, labs_s_na, labs_s_k, labs_s_cl, labs_s_hc03, labs_glucose, 
         labs_bun, labs_s_ca, labs_tot_prot, labs_albumin, labs_bili, labs_alk_phos, labs_got_ast, 
         labs_gpt_alt, labs_sua, labs_u_na, labs_u_glucose, labs_uua, labs_cystatinc, 
         labs_s_creatinine, labs_insulin, labs_cholesterol, labs_hdl, labs_ldl,
         labs_triglycerides, labs_u_creatinine, labs_u_microalbumin, labs_acr,
         pilabs_s_osmo, pilabs_u_osmo, pilabs_copeptin, egfr, iohexol_abs_gfr,
         iohexol_bsa_gfr, iohexol_ecv, iohexol_ecv_gfr, iohexol_gfr_ecv_std,
         edv_ml, esv_ml, sv_ml, ef_percent, cardiac_output, myo_mass_dias,
         myo_mass_syst, radial_peak, circum_peak, long_peak, HOMA_IR)) %>% 
  miss_var_summary() %>% 
  kable()

miss_plot2 <- expss::use_labels(responders_visit_dat[,vars],
                              {gg_miss_var(..data, facet=studyvisit_type)})
```

# Results: All Subjects

## Descriptive Statistics of Participants

The below table summarizes the data on participants who completed both Visit 1 (Baseline) and Visit 2 (1 Week).

```{r table 1, echo=FALSE}
## descriptive stats table 
table1::table1(~ age + Race + Ethnicity + diabetes_yrs + diabetes_age + t2d_less_21 + diabetes_meds___1 +
                 diabetes_meds___2 + diabetes_meds___3 + diabetes_meds___4 + diabetes_meds___5 + diabetes_meds___6 +
                 diabetes_meds___7 + insulin_type + insulin_tdd + med_hx_hypertension + hypertension_meds + 
                 hypertension_med_type___1 + hypertension_med_type___2 + hypertension_med_type___3 + 
                 hypertension_med_type___4 + hypertension_med_type___5 + hypertension_med_type___6 + 
                 hypertension_med_type___7 + medhx_cvd + medhx_met + met_hx___1 + met_hx___3 + met_hx___4 +
                 met_hx___5 + metdx_meds + pe_bmi + pe_bmi_percent + pe_waist + pe_hip + pe_sbp + pe_dbp +
                 pe_map + pe_pulse + a1c + sua + g6pd, data=screening_dat)
```

## Nonparametric: Baseline vs. 1-Week Visit Measures

The below table provides a comparison of baseline versus 1-week variables of interest; measures for these variables at each timepoint are summarized as a median with corresponding minimum, maximum, 25th percentile, and 75th percentile values. P-values are calculated using a Wilcoxon signed rank test. As no data is available for urine osmolality, serum osmolality, eGFR, or copeptin, these measures are not summarized.

```{r table 2, echo=FALSE, message=FALSE, warning=FALSE}
## WRS analysis results table

visit_dat %>%
  select(record_id, studyvisit_type, vitals_weight, vitals_sbp, vitals_dbp, vitals_map,
         vitals_pulse, labs_bg, labs_s_na, labs_s_k, labs_s_cl, labs_s_hc03, labs_glucose, 
         labs_bun, labs_s_ca, labs_tot_prot, labs_albumin, labs_bili, labs_alk_phos, labs_got_ast, 
         labs_gpt_alt, labs_sua, labs_u_na, labs_u_glucose, labs_uua, labs_cystatinc, 
         labs_s_creatinine, labs_insulin, labs_cholesterol, labs_hdl, labs_ldl,
         labs_triglycerides, labs_u_creatinine, labs_u_microalbumin, labs_acr, iohexol_abs_gfr,
         iohexol_bsa_gfr, iohexol_ecv, iohexol_ecv_gfr, iohexol_gfr_ecv_std,
         edv_ml, esv_ml, sv_ml, ef_percent, cardiac_output, myo_mass_dias,
         myo_mass_syst, radial_peak, circum_peak, long_peak, HOMA_IR) %>% 
  tbl_summary(by=studyvisit_type, include=-record_id,
              type=list(c(vitals_weight, vitals_sbp, vitals_dbp, vitals_map,
                          vitals_pulse, labs_bg, labs_s_na, labs_s_k, labs_s_cl, labs_s_hc03, labs_glucose, 
                          labs_bun, labs_s_ca, labs_tot_prot, labs_albumin, labs_bili, labs_alk_phos, labs_got_ast, 
                          labs_gpt_alt, labs_sua, labs_u_na, labs_u_glucose, labs_uua, labs_cystatinc, 
                          labs_s_creatinine, labs_insulin, labs_cholesterol, labs_hdl, labs_ldl,
                          labs_triglycerides, labs_u_creatinine, labs_u_microalbumin, labs_acr, iohexol_abs_gfr,
                          iohexol_bsa_gfr, iohexol_ecv, iohexol_ecv_gfr, iohexol_gfr_ecv_std,
                          edv_ml, esv_ml, sv_ml, ef_percent, cardiac_output, myo_mass_dias,
                          myo_mass_syst, radial_peak, circum_peak, long_peak, HOMA_IR) ~ 'continuous', 
                        studyvisit_type~'dichotomous'),
    statistic=list(all_continuous() ~ '{median} (Min: {min}, Max: {max}; 25% Pi: {p25}, 75% Pi: {p75}; )'),
    digits=everything()~c(2,2,2,2,2),
    missing_text = 'N Missing') %>% 
  add_p(test = all_continuous() ~ 'paired.wilcox.test', group=record_id,
        pvalue_fun=~style_pvalue(.x, digits=3)) %>% 
  modify_header(label='**Measurement**')

```

## Parametric: Baseline vs. 1-Week Visit Measures

The below table provides a comparison of baseline versus 1-week variables of interest; measures for these variables at each timepoint are summarized as a mean with standard deviation. P-values are calculated using a paired t-test. As no data is available for urine osmolality, serum osmolality, eGFR, or copeptin, these measures are not summarized.

```{r table 3, echo=FALSE, message=FALSE, warning=FALSE}
## paired t-test analysis results table

visit_dat %>%
  select(record_id, studyvisit_type, vitals_weight, vitals_sbp, vitals_dbp, vitals_map,
         vitals_pulse, labs_bg, labs_s_na, labs_s_k, labs_s_cl, labs_s_hc03, labs_glucose, 
         labs_bun, labs_s_ca, labs_tot_prot, labs_albumin, labs_bili, labs_alk_phos, labs_got_ast, 
         labs_gpt_alt, labs_sua, labs_u_na, labs_u_glucose, labs_uua, labs_cystatinc, 
         labs_s_creatinine, labs_insulin, labs_cholesterol, labs_hdl, labs_ldl,
         labs_triglycerides, labs_u_creatinine, labs_u_microalbumin, labs_acr, iohexol_abs_gfr,
         iohexol_bsa_gfr, iohexol_ecv, iohexol_ecv_gfr, iohexol_gfr_ecv_std,
         edv_ml, esv_ml, sv_ml, ef_percent, cardiac_output, myo_mass_dias,
         myo_mass_syst, radial_peak, circum_peak, long_peak, HOMA_IR) %>% 
  tbl_summary(by=studyvisit_type, include=-record_id,
              type=list(c(vitals_weight, vitals_sbp, vitals_dbp, vitals_map,
                          vitals_pulse, labs_bg, labs_s_na, labs_s_k, labs_s_cl, labs_s_hc03, labs_glucose, 
                          labs_bun, labs_s_ca, labs_tot_prot, labs_albumin, labs_bili, labs_alk_phos, labs_got_ast, 
                          labs_gpt_alt, labs_sua, labs_u_na, labs_u_glucose, labs_uua, labs_cystatinc, 
                          labs_s_creatinine, labs_insulin, labs_cholesterol, labs_hdl, labs_ldl,
                          labs_triglycerides, labs_u_creatinine, labs_u_microalbumin, labs_acr, iohexol_abs_gfr,
                          iohexol_bsa_gfr, iohexol_ecv, iohexol_ecv_gfr, iohexol_gfr_ecv_std,
                          edv_ml, esv_ml, sv_ml, ef_percent, cardiac_output, myo_mass_dias,
                          myo_mass_syst, radial_peak, circum_peak, long_peak, HOMA_IR) ~ 'continuous', 
                        studyvisit_type~'dichotomous'),
    statistic=list(all_continuous() ~ '{mean} ({sd})'),
    digits=everything()~c(2,2),
    missing_text = 'N Missing') %>% 
  add_p(test = all_continuous() ~ 'paired.t.test', group=record_id,
        pvalue_fun=~style_pvalue(.x, digits=3)) %>% 
  modify_header(label='**Measurement**')

```

## Missingness

The below plot visualizes the number of missing observations for each variable of interest by visit type (baseline vs. 1 week).

```{r echo=FALSE}
miss_plot
```

## Peak Longitudinal Strain Histograms

As the analysis results indicated that the difference in peak longitudinal strain between baseline and 1 week is significant with the parametric test (p<0.05) but not significant with the nonparametric test (p>0.05), the plots of the distribution of observations for peak longitudinal strain at each timepoint are presented below. 

```{r distributions, echo=FALSE, message=FALSE, warning=FALSE}
## peak longitudinal strain histograms

base_hist <- wide_visit %>% 
  ggplot(aes(x=long_peak_Study_Visit_Baseline)) +
  geom_histogram() +
  labs(title='Baseline',
       x='Peak Longitudinal Strain') +
  theme_bw()

v1_hist <- wide_visit %>% 
  ggplot(aes(x=`long_peak_Study_Visit_1 Week`)) +
  geom_histogram() +
  labs(title='1 Week',
       x='Peak Longitudinal Strain') +
  theme_bw()

base_hist + v1_hist
```

## Individual-Level Plots

The figures below show pre- and post (baseline vs. 1 week) changes in the variables of interest. 

```{r spaghetti plot, echo=FALSE, warning=FALSE}
## spaghetti plots for variables of interest
# non-responders in a different color
visit_dat <- visit_dat %>% 
  mutate(Status=case_when(abs_delta_SUA >= 3 ~ 'Responder', .default='Non-responder'))
label(visit_dat$Status) <- 'Responder Status'

# visit_dat %>% 
#   ggplot(aes(studyvisit_type, vitals_sbp, group=record_id)) +
#   geom_point(alpha=0.5, size=0.5, aes(color=Status)) +
#   geom_line(aes(color=Status), alpha=0.4) +
#   labs(x='timepoint', y='Systolic Blood Pressure', title='SBP Across Timepoints') +
#   scale_color_manual(values=c('red','black')) +
#   theme_bw()
# 
# visit_dat %>% 
#   ggplot(aes(studyvisit_type, vitals_dbp, group=record_id)) +
#   geom_point(alpha=0.5, size=0.5, aes(color=Status)) +
#   geom_line(aes(color=Status), alpha=0.4) +
#   labs(x='timepoint', y='Diastolic Blood Pressure', title='DBP Across Timepoints') +
#   scale_color_manual(values=c('red','black')) +
#   theme_bw()
# 
# visit_dat %>% 
#   ggplot(aes(studyvisit_type, vitals_map, group=record_id)) +
#   geom_point(alpha=0.5, size=0.5, aes(color=Status)) +
#   geom_line(aes(color=Status), alpha=0.4) +
#   labs(x='timepoint', y='Mean Arterial Pressure', title='MAP Across Timepoints') +
#   scale_color_manual(values=c('red','black')) +
#   theme_bw()
# 
# visit_dat %>% 
#   ggplot(aes(studyvisit_type, labs_sua, group=record_id)) +
#   geom_point(alpha=0.5, size=0.5, aes(color=Status)) +
#   geom_line(aes(color=Status), alpha=0.4) +
#   labs(x='timepoint', y='Serum Uric Acid (mg/dl)', title='SUA Across Timepoints') +
#   scale_color_manual(values=c('red','black')) +
#   theme_bw()
# 
# visit_dat %>% 
#   ggplot(aes(studyvisit_type, labs_uua, group=record_id)) +
#   geom_point(alpha=0.5, size=0.5, aes(color=Status)) +
#   geom_line(aes(color=Status), alpha=0.4) +
#   labs(x='timepoint', y='Urine Uric Acid (mg/dl)', title='UUA Across Timepoints') +
#   scale_color_manual(values=c('red','black')) +
#   theme_bw()
# 
# visit_dat %>% 
#   ggplot(aes(studyvisit_type, labs_bili, group=record_id)) +
#   geom_point(alpha=0.5, size=0.5, aes(color=Status)) +
#   geom_line(aes(color=Status), alpha=0.4) +
#   labs(x='timepoint', y='Serum Bilirubin (mg/dl)', title='Bilirubin Across Timepoints') +
#   scale_color_manual(values=c('red','black')) +
#   theme_bw()

visit_dat %>%
  filter(record_id!='UT2D-06') %>% # filter for complete pairs
  ggplot(aes(studyvisit_type, iohexol_abs_gfr, group=record_id)) +
  geom_point(alpha=0.8, size=1.7, aes(color=Status)) +
  geom_line(aes(color=Status), alpha=0.8, size=1) +
  labs(x='Timepoint', y='Absolute GFR (ml/min)') + #title='Absolute GFR Across Timepoints',
       #caption='Note: 1 subject is not represented due to missing data') +
  scale_color_manual(values=c('red','black')) +
  theme_bw() +
  theme(axis.text = element_text(size=12),
        axis.title = element_text(size=14),
        legend.position = 'none')

visit_dat %>%
  filter(record_id!='UT2D-06') %>% # filter for complete pairs
  ggplot(aes(studyvisit_type, iohexol_bsa_gfr, group=record_id)) +
  geom_point(alpha=0.8, size=1.7, aes(color=Status)) +
  geom_line(aes(color=Status), alpha=0.8, size=1) +
  labs(x='Timepoint', y='GFR BSA Standardized (ml/min/1.73m2)') + #, title='BSA-Adjusted GFR Across Timepoints',
       #caption='Note: 1 subject is not represented due to missing data') +
  scale_color_manual(values=c('red','black')) +
  theme_bw() +
  theme(axis.text = element_text(size=12),
        axis.title = element_text(size=14),
        legend.position = 'none')

# visit_dat %>%
#   filter(record_id!= 'UT2D-06', record_id!= 'UT2D-07') %>% # filter for complete pairs
#   ggplot(aes(studyvisit_type, edv_ml, group=record_id)) +
#   geom_point(alpha=0.5, size=0.5, aes(color=Status)) +
#   geom_line(aes(color=Status), alpha=0.4) +
#   labs(x='timepoint', y='End diastolic volume (mL)', title='End Diastolic Volume Across Timepoints',
#        caption='Note: 2 subjects are not represented due to missing data') +
#   scale_color_manual(values=c('red','black')) +
#   theme_bw() +
#   theme(plot.caption = element_text(hjust=0))
# 
# visit_dat %>% 
#   ggplot(aes(studyvisit_type, long_peak, group=record_id)) +
#   geom_point(alpha=0.5, size=0.5, aes(color=Status)) +
#   geom_line(aes(color=Status), alpha=0.3) +
#   labs(x='timepoint', y='Peak longitudinal strain', title='Longitudinal Strain Across Timepoints') +
#   scale_color_manual(values=c('red','black')) +
#   theme_bw()
```

## Change in SUA vs. Change in Variables of Interest

This section includes model output testing the correlation between change in serum uric acid (SUA) and change in the variables of interest. Note that change refers to the difference between 1 week and baseline values and corresponds to "delta" in the results. Diagnostic plots for checking linear regression assumptions are also included.

```{r regression mods, echo=FALSE}
# calculate difference in week 1 and baseline measures of interest
wide_visit <- wide_visit %>% 
  group_by(record_id) %>% 
  mutate(delta_SUA=`labs_sua_Study_Visit_1 Week`-labs_sua_Study_Visit_Baseline,
         delta_SBP=`vitals_sbp_Study_Visit_1 Week`- vitals_sbp_Study_Visit_Baseline,
         delta_DBP=`vitals_dbp_Study_Visit_1 Week`- vitals_dbp_Study_Visit_Baseline,
         delta_MAP=`vitals_map_Study_Visit_1 Week` - vitals_map_Study_Visit_Baseline,
         delta_UUA=`labs_uua_Study_Visit_1 Week`- labs_uua_Study_Visit_Baseline,
         delta_bili=`labs_bili_Study_Visit_1 Week`- labs_bili_Study_Visit_Baseline,
         delta_abs_gfr= `iohexol_abs_gfr_Study_Visit_1 Week`- iohexol_abs_gfr_Study_Visit_Baseline,
         delta_bsa_gfr= `iohexol_bsa_gfr_Study_Visit_1 Week`- iohexol_bsa_gfr_Study_Visit_Baseline,
         delta_edv=`edv_ml_Study_Visit_1 Week`- edv_ml_Study_Visit_Baseline,
         delta_long_peak=`long_peak_Study_Visit_1 Week`- long_peak_Study_Visit_Baseline) %>%
  ungroup()

label(wide_visit$delta_SUA) <- ''

# fit regression models with delta_SUA as outcome
vars <- c('delta_SBP','delta_DBP','delta_MAP','delta_UUA','delta_bili','delta_abs_gfr','delta_bsa_gfr','delta_edv','delta_long_peak')

for (v in vars) {
  form = as.formula(paste0('delta_SUA','~', v))
  mod <- lm(form, data = wide_visit)
  print(noquote(c("Correlation between delta SUA and", noquote(v))))
  print(summary(mod))
  plot(x=wide_visit[[v]], y=wide_visit$delta_SUA, ylab='delta_SUA', xlab=noquote(v), main='Y-X Scatterplot'); abline(mod, col='gray', lty=2)
  plot(mod, which=1)
  plot(ppoints(length(rstudent(mod))), sort(pnorm(rstudent(mod))), xlab='Observed Cumulative Probability', ylab='Expected Cumulative Probability',
       main='Normal Probability Plot', cex=1); abline(a=0,b=1, col='gray65', lwd=1)
  hist(rstudent(mod), xlab='Jackknife Residual', main='Histogram of Residuals', freq=F)
  cat('\n\n')
}
```

## Correlation

### Pearson 

Delta SUA and delta absolute GFR:

```{r echo=FALSE}
cor.test(wide_visit$delta_SUA, wide_visit$delta_abs_gfr, method = 'pearson')
```

Delta SUA and delta GFR bsa standardized:
```{r echo=FALSE}
cor.test(wide_visit$delta_SUA, wide_visit$delta_bsa_gfr, method = 'pearson')
```

Delta SUA and delta SBP:
```{r echo=FALSE}
cor.test(wide_visit$delta_SUA, wide_visit$delta_SBP, method = 'pearson')
```

### Spearman 

Delta SUA and delta absolute GFR:
```{r echo=FALSE}
cor.test(wide_visit$delta_SUA, wide_visit$delta_abs_gfr, method = 'spearman')
```

Delta SUA and delta GFR bsa standardized:
```{r echo=FALSE}
cor.test(wide_visit$delta_SUA, wide_visit$delta_bsa_gfr, method = 'spearman')
```

Delta SUA and delta SBP:
```{r echo=FALSE}
cor.test(wide_visit$delta_SUA, wide_visit$delta_SBP, method = 'spearman', exact = FALSE)
```

## Distributions
```{r echo=FALSE}
hist(wide_visit$delta_SUA, main = 'Distribution of Delta SUA')
hist(wide_visit$delta_abs_gfr, main = 'Distribution of Delta Absolute GFR')
hist(wide_visit$delta_bsa_gfr, main = 'Distribution of Delta GFR BSA Standardized')
hist(wide_visit$delta_SBP, main = 'Distribution of Delta SBP')
```

# Results: Responders Only

## Descriptive Statistics of Responders

The below table summarizes the data on participants who completed both Visit 1 (Baseline) and Visit 2 (1 Week) and had an optimal serum uric acid lowering response to pegloticase. Out of the 9 subjects used for the main analysis, 7 subjects were considered responders (Delta SUA >= 3 mg/dL).

```{r responders table 1, echo=FALSE}
# descriptive stats table 
table1::table1(~ age + Race + Ethnicity + diabetes_yrs + diabetes_age + t2d_less_21 + diabetes_meds___1 +
                 diabetes_meds___2 + diabetes_meds___3 + diabetes_meds___4 + diabetes_meds___5 + diabetes_meds___6 +
                 diabetes_meds___7 + insulin_type + insulin_tdd + med_hx_hypertension + hypertension_meds + 
                 hypertension_med_type___1 + hypertension_med_type___2 + hypertension_med_type___3 + 
                 hypertension_med_type___4 + hypertension_med_type___5 + hypertension_med_type___6 + 
                 hypertension_med_type___7 + medhx_cvd + medhx_met + met_hx___1 + met_hx___3 + met_hx___4 +
                 met_hx___5 + metdx_meds + pe_bmi + pe_bmi_percent + pe_waist + pe_hip + pe_sbp + pe_dbp +
                 pe_map + pe_pulse + a1c + sua + g6pd, data=responders_screening_dat)
```

## Nonparametric: Baseline vs. 1-Week Visit Measures

The below table provides a comparison of baseline versus 1-week variables of interest for responders; measures for these variables at each timepoint are summarized as a median with corresponding minimum, maximum, 25th percentile, and 75th percentile values. P-values are calculated using a Wilcoxon signed rank test. As no data is available for urine osmolality, serum osmolality, eGFR, or copeptin, these measures are not summarized.

```{r responders table 2, echo=FALSE, message=FALSE, warning=FALSE}
# WRS analysis results table

responders_visit_dat %>%
  select(record_id, studyvisit_type, vitals_weight, vitals_sbp, vitals_dbp, vitals_map,
         vitals_pulse, labs_bg, labs_s_na, labs_s_k, labs_s_cl, labs_s_hc03, labs_glucose, 
         labs_bun, labs_s_ca, labs_tot_prot, labs_albumin, labs_bili, labs_alk_phos, labs_got_ast, 
         labs_gpt_alt, labs_sua, labs_u_na, labs_u_glucose, labs_uua, labs_cystatinc, 
         labs_s_creatinine, labs_insulin, labs_cholesterol, labs_hdl, labs_ldl,
         labs_triglycerides, labs_u_creatinine, labs_u_microalbumin, labs_acr, iohexol_abs_gfr,
         iohexol_bsa_gfr, iohexol_ecv, iohexol_ecv_gfr, iohexol_gfr_ecv_std,
         edv_ml, esv_ml, sv_ml, ef_percent, cardiac_output, myo_mass_dias,
         myo_mass_syst, radial_peak, circum_peak, long_peak, HOMA_IR) %>% 
  tbl_summary(by=studyvisit_type, include=-record_id,
              type=list(c(vitals_weight, vitals_sbp, vitals_dbp, vitals_map,
                          vitals_pulse, labs_bg, labs_s_na, labs_s_k, labs_s_cl, labs_s_hc03, labs_glucose, 
                          labs_bun, labs_s_ca, labs_tot_prot, labs_albumin, labs_bili, labs_alk_phos, labs_got_ast, 
                          labs_gpt_alt, labs_sua, labs_u_na, labs_u_glucose, labs_uua, labs_cystatinc, 
                          labs_s_creatinine, labs_insulin, labs_cholesterol, labs_hdl, labs_ldl,
                          labs_triglycerides, labs_u_creatinine, labs_u_microalbumin, labs_acr, iohexol_abs_gfr,
                          iohexol_bsa_gfr, iohexol_ecv, iohexol_ecv_gfr, iohexol_gfr_ecv_std,
                          edv_ml, esv_ml, sv_ml, ef_percent, cardiac_output, myo_mass_dias,
                          myo_mass_syst, radial_peak, circum_peak, long_peak, HOMA_IR) ~ 'continuous', 
                        studyvisit_type~'dichotomous'),
    statistic=list(all_continuous() ~ '{median} (Min: {min}, Max: {max}; 25% Pi: {p25}, 75% Pi: {p75}; )'),
    digits=everything()~c(2,2,2,2,2),
    missing_text = 'N Missing') %>% 
  add_p(test = all_continuous() ~ 'paired.wilcox.test', group=record_id,
        pvalue_fun=~style_pvalue(.x, digits=3)) %>% 
  modify_header(label='**Measurement**')

```

## Parametric: Baseline vs. 1-Week Visit Measures

The below table provides a comparison of baseline versus 1-week variables of interest for responders; measures for these variables at each timepoint are summarized as a mean with standard deviation. P-values are calculated using a paired t-test. As no data is available for urine osmolality, serum osmolality, eGFR, or copeptin, these measures are not summarized.

```{r responders table 3, echo=FALSE, message=FALSE, warning=FALSE}
# paired t-test analysis results table

responders_visit_dat %>%
  select(record_id, studyvisit_type, vitals_weight, vitals_sbp, vitals_dbp, vitals_map,
         vitals_pulse, labs_bg, labs_s_na, labs_s_k, labs_s_cl, labs_s_hc03, labs_glucose, 
         labs_bun, labs_s_ca, labs_tot_prot, labs_albumin, labs_bili, labs_alk_phos, labs_got_ast, 
         labs_gpt_alt, labs_sua, labs_u_na, labs_u_glucose, labs_uua, labs_cystatinc, 
         labs_s_creatinine, labs_insulin, labs_cholesterol, labs_hdl, labs_ldl,
         labs_triglycerides, labs_u_creatinine, labs_u_microalbumin, labs_acr, iohexol_abs_gfr,
         iohexol_bsa_gfr, iohexol_ecv, iohexol_ecv_gfr, iohexol_gfr_ecv_std,
         edv_ml, esv_ml, sv_ml, ef_percent, cardiac_output, myo_mass_dias,
         myo_mass_syst, radial_peak, circum_peak, long_peak, HOMA_IR) %>% 
  tbl_summary(by=studyvisit_type, include=-record_id,
              type=list(c(vitals_weight, vitals_sbp, vitals_dbp, vitals_map,
                          vitals_pulse, labs_bg, labs_s_na, labs_s_k, labs_s_cl, labs_s_hc03, labs_glucose, 
                          labs_bun, labs_s_ca, labs_tot_prot, labs_albumin, labs_bili, labs_alk_phos, labs_got_ast, 
                          labs_gpt_alt, labs_sua, labs_u_na, labs_u_glucose, labs_uua, labs_cystatinc, 
                          labs_s_creatinine, labs_insulin, labs_cholesterol, labs_hdl, labs_ldl,
                          labs_triglycerides, labs_u_creatinine, labs_u_microalbumin, labs_acr, iohexol_abs_gfr,
                          iohexol_bsa_gfr, iohexol_ecv, iohexol_ecv_gfr, iohexol_gfr_ecv_std,
                          edv_ml, esv_ml, sv_ml, ef_percent, cardiac_output, myo_mass_dias,
                          myo_mass_syst, radial_peak, circum_peak, long_peak, HOMA_IR) ~ 'continuous', 
                        studyvisit_type~'dichotomous'),
    statistic=list(all_continuous() ~ '{mean} ({sd})'),
    digits=everything()~c(2,2),
    missing_text = 'N Missing') %>% 
  add_p(test = all_continuous() ~ 'paired.t.test', group=record_id,
        pvalue_fun=~style_pvalue(.x, digits=3)) %>% 
  modify_header(label='**Measurement**')
```

## Missingness

The below plot visualizes the number of missing observations for each variable of interest by visit type (baseline vs. 1 week) for those categorized as responders.
```{r echo=FALSE}
miss_plot2
```

## Correlation

```{r responders wide format, include=FALSE}
responders_wide_visit <- wide_visit %>% 
  filter(abs_delta_SUA >= 3)

# add in renal flow data
responders_wide_visit <- left_join(responders_wide_visit, renal_flow_dat, by = "record_id") %>% 
  select(-delta_gfr) # delta gfr is same as abs delta gfr in responders_wide_visit
```

### Pearson 

Delta SUA and delta absolute GFR:

```{r echo=FALSE}
cor.test(responders_wide_visit$delta_SUA, responders_wide_visit$delta_abs_gfr, method = 'pearson')
```

Delta SUA and delta GFR bsa standardized:
```{r echo=FALSE}
cor.test(responders_wide_visit$delta_SUA, responders_wide_visit$delta_bsa_gfr, method = 'pearson')
```

Delta SUA and delta SBP:
```{r echo=FALSE}
cor.test(responders_wide_visit$delta_SUA, responders_wide_visit$delta_SBP, method = 'pearson')
```

Delta RV 2D PWV and absolute GFR:
```{r echo=FALSE}
cor.test(responders_wide_visit$delta_RV2D_PWV, responders_wide_visit$delta_abs_gfr, method = 'pearson')
```

Delta RA 2D PWV and absolute GFR:
```{r echo=FALSE}
cor.test(responders_wide_visit$delta_RA2DPWV, responders_wide_visit$delta_abs_gfr, method = 'pearson')
```

### Spearman 

Delta SUA and delta absolute GFR:
```{r echo=FALSE}
cor.test(responders_wide_visit$delta_SUA, responders_wide_visit$delta_abs_gfr, method = 'spearman')
```

Delta SUA and delta GFR bsa standardized:
```{r echo=FALSE}
cor.test(responders_wide_visit$delta_SUA, responders_wide_visit$delta_bsa_gfr, method = 'spearman')
```

Delta SUA and delta SBP:
```{r echo=FALSE}
cor.test(responders_wide_visit$delta_SUA, responders_wide_visit$delta_SBP, method = 'spearman', exact = FALSE)
```

Delta RV 2D PWV and absolute GFR:
```{r echo=FALSE}
cor.test(responders_wide_visit$delta_RV2D_PWV, responders_wide_visit$delta_abs_gfr, method = 'spearman')
```

Delta RA 2D PWV and absolute GFR:
```{r echo=FALSE}
cor.test(responders_wide_visit$delta_RA2DPWV, responders_wide_visit$delta_abs_gfr, method = 'spearman')
```

## Distributions
```{r echo=FALSE}
hist(responders_wide_visit$delta_SUA, main = 'Distribution of Delta SUA')
hist(responders_wide_visit$delta_abs_gfr, main = 'Distribution of Delta Absolute GFR')
hist(responders_wide_visit$delta_bsa_gfr, main = 'Distribution of Delta GFR BSA Standardized')
hist(responders_wide_visit$delta_SBP, main = 'Distribution of Delta SBP')
hist(responders_wide_visit$delta_RV2D_PWV, main = 'Distribution of Delta RV 2D PWV')
hist(responders_wide_visit$delta_RA2DPWV, main = 'Distribution of Delta RA 2D PWV')
```


```{r UACR check, eval=FALSE, include=FALSE}
check_30 <- visit_dat %>% 
  filter(labs_acr >= 30) %>% 
  select(record_id, studyvisit_type, sv_date, labs_acr)

check_responders_30 <- responders_visit_dat %>% 
  filter(labs_acr >= 30) %>% 
  select(record_id, studyvisit_type, sv_date, labs_acr)

check_300 <- visit_dat %>% 
  filter(labs_acr >= 300) %>% 
  select(record_id, studyvisit_type, sv_date, labs_acr)

check_responders_300 <- responders_visit_dat %>% 
  filter(labs_acr >= 300) %>% 
  select(record_id, studyvisit_type, sv_date, labs_acr)
```


