---
title: "RH/RH2/ROCKIES"
author: "Ye Ji Choi"
date: "today"
date-format: long
execute: 
  echo: false
  warning: false
  messages: false
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
editor: source
---

scRNA data file: "PB_90samples_Harmony_rpca_Fadhl_PhilApproved_091024.RDS"


**Comments from PB on 10/30/24:**
DEGs

* Q1. Does this dataset include RH, RH2, IT2D, CRC?

This includes RH, RH2, and CRC. No IT2D. 

* Q2. Albuminuria analysis should be limited to those with T2D or OB; if not, this apparent difference b/w albuminuria is likely driven by T2D/OB vs. LC.

* Q3. SGLT2i (ever) in all? If so, same concept above applies.

This has been added.

* Q4. Albuminuria missing

This has been added.

* Q5. SGLT2i (timepoint) data are missing 

This has been added.

“subset to IDs in old dataset”

* Q6. Does this refer to dataset at JCI paper?

No, this dataset refers to PB68.

* Q7. If so, it should be T2D(-) vs. HC

The dataset refers to PB68 but the analyses for T2Di- vs. HC is included.


```{r libraries, message = F}
#| include: false
library(Seurat)
library(future)
library(tidyverse)
library(colorspace)
library(patchwork)
library(ggdendro)
library(cowplot)
library(ggpubr)
library(rstatix)
library(table1)
library(Biobase)
library(ReactomeGSA)
library(GSEABase)
library(msigdbr)
library(kableExtra)
library(knitr)
library(slingshot); library(SingleCellExperiment)
```



```{r data-setup, echo = F, message=F}
# Parallel processing
plan(multicore, workers = 16)
options(future.globals.maxSize=2e9)
so <- readRDS("/run/user/1023/gvfs/smb-share:server=ucdenver.pvt,share=som/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/scRNA/data_raw/PB_90samples_Harmony_rpca_Fadhl_PhilApproved_091024.RDS")
source("/home/yejichoi2/Documents/CHCO-Code/Petter Bjornstad/CROCODILE/crocodile_functions.R")
source("/home/yejichoi2/Documents/CHCO-Code/Petter Bjornstad/Resources/YC/R Functions/label_harmonized_function.R")
dict <- read.csv("/run/user/1023/gvfs/smb-share:server=ucdenver.pvt,share=som/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/Data Harmonization/data_dictionary_master.csv") %>%
  dplyr::select(variable_name, label)

dict <- setNames(data.frame(t(dict[ , - 1])), dict[ , 1])
dict <- dict %>%
  mutate(sce_PT.slingPseudotime_1 = "Pseudotime",
         gbm_thick_art_mean = "GBM thickness",
         p2_raw_m = "M-value",
         avg_c_k2 = "Cortical K2",
         bmi = "BMI",
         diabetes_duration = "Diabetes duration", 
         hba1c = "HbA1c")

dat <- read.csv("/run/user/1023/gvfs/smb-share:server=ucdenver.pvt,share=som/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/Data Harmonization/Data Clean/harmonized_dataset.csv") %>%
  filter(study %in% c("RENAL-HEIR", "RENAL-HEIRitage") & group != "Lean Control" | study == "CROCODILE" & group == "Lean Control") %>%
  dplyr::mutate(across(where(is.character), ~ na_if(., ""))) %>%
  dplyr::summarise(across(where(negate(is.numeric)), ~ ifelse(all(is.na(.x)), NA_character_, last(na.omit(.x)))),
                   across(where(is.numeric), ~ ifelse(all(is.na(.x)), NA_real_, mean(.x, na.rm = TRUE))),
                   .by = record_id) %>%
  mutate(albuminuria_yn = case_when(acr_u < 30 ~ "No", acr_u >= 30 ~ "Yes"),
         avg_c_f = rowMeans(dplyr::select(., lc_f, rc_f), na.rm = TRUE),
         avg_c_k1 = rowMeans(dplyr::select(., lc_k1, rc_k1), na.rm = TRUE),
         avg_c_k2 = rowMeans(dplyr::select(., lc_k2, rc_k2), na.rm = TRUE),
         avg_c_vb = rowMeans(dplyr::select(., lc_vb, rc_vb), na.rm = TRUE),
         avg_m_f = rowMeans(dplyr::select(., lm_f, rm_f), na.rm = TRUE),
         avg_m_k1 = rowMeans(dplyr::select(., lm_k1, rm_k1), na.rm = TRUE),
         avg_m_k2 = rowMeans(dplyr::select(., lm_k2, rm_k2), na.rm = TRUE),
         avg_m_vb = rowMeans(dplyr::select(., lm_vb, rm_vb), na.rm = TRUE),
         avg_pcascl = rowMeans(dplyr::select(., pcasl3d_left, pcasl3d_right), na.rm = TRUE),
         avg_k_r2_bl = rowMeans(dplyr::select(., bold_l_bl_kidney, bold_r_bl_kidney), na.rm = TRUE),
         avg_c_r2_bl = rowMeans(dplyr::select(., bold_l_bl_cortex, bold_r_bl_cortex), na.rm = TRUE),
         avg_m_r2_bl = rowMeans(dplyr::select(., bold_l_bl_medulla, bold_r_bl_medulla), na.rm = TRUE),
         avg_k_r2_pf = rowMeans(dplyr::select(., bold_l_pf_kidney, bold_r_pf_kidney), na.rm = TRUE),
         avg_c_r2_pf = rowMeans(dplyr::select(., bold_l_pf_cortex, bold_r_pf_cortex), na.rm = TRUE),
         avg_m_r2_pf = rowMeans(dplyr::select(., bold_l_pf_medulla, bold_r_pf_medulla), na.rm = TRUE),
         avg_k_fsoc = rowMeans(dplyr::select(., fsoc_l_kidney, fsoc_r_kidney), na.rm = TRUE),
         avg_c_fsoc = rowMeans(dplyr::select(., fsoc_l_cortex, fsoc_r_cortex), na.rm = TRUE),
         avg_m_fsoc = rowMeans(dplyr::select(., fsoc_l_medulla, fsoc_r_medulla), na.rm = TRUE),
         avg_c_adc = rowMeans(dplyr::select(., adc_left, adc_right), na.rm = TRUE),
         avg_c_t1 = rowMeans(dplyr::select(., t1_l_cortex, t1_r_cortex), na.rm = TRUE),
         avg_k_t1 = rowMeans(dplyr::select(., t1_l_kidney, t1_r_kidney), na.rm = TRUE),
         avg_c_k1_wo_cyst_vw = rowMeans(dplyr::select(., lc_k1_wo_cyst_vw, rc_k1_wo_cyst_vw), na.rm = TRUE),
         avg_m_k1_wo_cyst_vw = rowMeans(dplyr::select(., lm_k1_wo_cyst_vw, lm_k1_wo_cyst_vw), na.rm = TRUE),
         avg_c_k2_wo_cyst_vw = rowMeans(dplyr::select(., lc_k2_wo_cyst_vw, rc_k2_wo_cyst_vw), na.rm = TRUE),
         avg_m_k2_wo_cyst_vw = rowMeans(dplyr::select(., lm_k2_wo_cyst_vw, rm_k2_wo_cyst_vw), na.rm = TRUE),
         avg_tot_k1_wo_cyst = rowMeans(dplyr::select(., ltot_k1_wo_cyst, rtot_k1_wo_cyst), na.rm = TRUE),
         avg_tot_k2_wo_cyst = rowMeans(dplyr::select(., ltot_k2_wo_cyst, rtot_k2_wo_cyst), na.rm = TRUE)) %>%
  dplyr::select(-c(kit_id, cryostor_id))

so_meta <- so@meta.data
so_meta <- left_join(so_meta, dat) 
rownames(so_meta) <- so_meta$barcode
so <- AddMetaData(so, so_meta)
so$group_t2ob <- ifelse(so$group == "Lean Control", "LC", "T2D/OB")
so$group_t2d <- ifelse(so$group == "Type 2 Diabetes", "T2D", "Non-T2D")
so <- subset(so, group %in% c("Lean Control", "Obese Control", "Type 2 Diabetes"))
```

```{r echo = F, message = F, include = F}
options(future.globals.maxSize = 3000 * 1024^3)
# UMAP of single cell RNA sequencing (RH/RH2, CRC)
so <- NormalizeData(so)
so <- ScaleData(so)
ElbowPlot(so)
so <- RunPCA(so, ncomponents = 10, features = VariableFeatures(object = so))
so <- FindNeighbors(so)
so <- FindClusters(so)
so <- RunUMAP(so, dims = 1:30)
# so <- RunTSNE(so, dimred = "UMAP")
```

```{r echo = F, message = F}
DimPlot(object = so, reduction = "umap", group.by = "celltype_rpca", raster = F)
DimPlot(object = so, reduction = "umap", group.by = "celltype_harmony", raster = F)
DimPlot(object = so, reduction = "umap", group.by = "group", raster = F)
```

```{r echo = F}
# Bar plot with key transcripts of ox phos and TCA cycle stratified by T2D status, albuminuria and SGLT2i (RH/RH2, CRC)
so$celltype_pt <- ifelse(grepl("PT-", so$celltype_harmony),
                         "PT",
                         as.character(so$celltype_harmony))
Idents(so) <- so$celltype_pt
so_group_t2ob <- subset(so, group_t2ob == "T2D/OB")

source("/home/yejichoi2/Documents/CHCO-Code/Petter Bjornstad/CROCODILE/crocodile_functions.R")

plot_de_markers <- function(so, genes, group_column, id1, id2, subset = NULL, suffix = "_plot") {
  # Perform differential expression analysis
  de_results <- de.markers(so, genes, group_column, id1 = id1, id2 = id2, subset, suffix)
  de_results_name <- paste0("m", suffix)
  de_results <- get(de_results_name)
  
  # Add direction column to the results
  de_results <- add_direction(de_results)
  
  # Plot the results
  de_plot <- de_results %>%
    ggplot(aes(x = rownames(.), y = avg_log2FC, fill = direction)) + 
    geom_col(color = "white") +
    theme_bw() + 
    theme(panel.grid = element_blank(),
          text = element_text(size = 14),
          axis.text.x = element_text(angle = 70, hjust = 1)) +
    labs(x = "Transcripts", 
         y = "Avg Log2FC",
         fill = "Direction") +
    scale_fill_manual(values = c("NS" = "grey", "Upregulated" = "#f28482", "Downregulated" = "#457b9d"))
  
  return(de_plot)
}
```

# Individual expression levels of genes
```{r echo = F}
# source("~/Documents/CHCO-Code/Petter Bjornstad/CROCODILE/crocodile_functions.R")
# FetchData(so)
```

# DEG in PT
## Key transcripts of TCA cycle
### Combined (T2D + OB + LC)
#### By T2D status (T2D vs. OB + LC)

Group 1 = "T2D"

Group 2 = "Non-T2D"

**Negative values** indicate that the transcript has a lower expression T2D group compared to non-T2D group.

```{r echo = F}
plot_de_markers(so, genes, "group_t2d", "T2D", "Non-T2D", "PT","_tca_t2d")
kable(m_tca_t2d, digits = 3)
```

#### By T2D status (T2D + OB vs. LC)
Group 1 = "T2D/OB"

Group 2 = "LC"

**Negative values** indicate that the transcript has a lower expression T2D/OB group compared to LC group.

```{r echo = F}
plot_de_markers(so, genes, "group_t2ob", id1 = "T2D/OB", id2 = "LC", "PT","_tca_t2ob")
kable(m_tca_t2ob, digits = 3)
```


### T2D & OB only (T2D + OB)
#### By T2D status (T2D vs. OB)

Group 1 = "T2D"

Group 2 = "Obese Control"

**Negative values** indicate that the transcript has a lower expression T2D group compared to Obese group.

```{r echo = F}
plot_de_markers(so_group_t2ob, genes, "group", id1 = "Type 2 Diabetes", id2 = "Obese Control", "PT","_tca_t2d_t2dob")
kable(m_tca_t2d_t2dob, digits = 3)
```

#### By Albuminuria

Group 1 = "Albuminuria"

Group 2 = "Non-Albuminuria"

**Negative values** indicate that the transcript has a lower expression Albuminuria group compared to Non-Albuminuria group.

```{r echo = F}
plot_de_markers(so_group_t2ob, genes, "elevated_albuminuria", id1 = "Yes", id2 = "No", "PT","_tca_alb_t2dob")
kable(m_tca_alb_t2dob, digits = 3)
```

#### By SGLT2i (ever)

Group 1 = "SGLT2i"

Group 2 = "Non-SGLT2i"

**Negative values** indicate that the transcript has a lower expression SGLT2i group compared to Non-SGLT2i group.

```{r echo = F}
plot_de_markers(so_group_t2ob, genes, "epic_ever_sglt2i_1", id1 = "Yes", id2 = "No", "PT","_tca_sglt2i_ever_t2dob")
kable(m_tca_sglt2i_ever_t2dob, digits = 3)
```

#### By SGLT2i (timepoint)

Group 1 = "SGLT2i"

Group 2 = "Non-SGLT2i"

**Negative values** indicate that the transcript has a lower expression SGLT2i group compared to Non-SGLT2i group.

```{r echo = F}
plot_de_markers(so_group_t2ob, genes, "epic_sglti2_1", id1 = "Yes", id2 = "No", "PT","_tca_sglt2i_timpoint_t2dob")
kable(m_tca_sglt2i_timpoint_t2dob, digits = 3)
```


## Key transcripts of Oxidative Phosphorylation
### Combined (T2D + OB + LC)
#### By T2D status (T2D vs. OB + LC)

Group 1 = "T2D"

Group 2 = "Non-T2D"

**Negative values** indicate that the transcript has a lower expression T2D group compared to non-T2D group.

```{r echo = F}
plot_de_markers(so, oxy_phos_genes, "group_t2d", "T2D", "Non-T2D", "PT","_oxphos_t2d")
kable(m_oxphos_t2d, digits = 3)
```

#### By T2D status (T2D + OB vs. LC)
Group 1 = "T2D/OB"

Group 2 = "LC"

**Negative values** indicate that the transcript has a lower expression T2D/OB group compared to LC group.

```{r echo = F}
plot_de_markers(so, oxy_phos_genes, "group_t2ob", id1 = "T2D/OB", id2 = "LC", "PT","_oxphos_t2ob")
kable(m_oxphos_t2ob, digits = 3)
```



### T2D & OB only (T2D + OB)
#### By T2D status (T2D vs. OB)

Group 1 = "T2D"

Group 2 = "Obese Control"

**Negative values** indicate that the transcript has a lower expression T2D group compared to Obese group.

```{r echo = F}
so_group_t2ob <- subset(so, group_t2ob == "T2D/OB")
plot_de_markers(so_group_t2ob, oxy_phos_genes, "group", id1 = "Type 2 Diabetes", id2 = "Obese Control", "PT","_oxphos_t2d_t2dob")
kable(m_oxphos_t2d_t2dob, digits = 3)
```

#### By Albuminuria

Group 1 = "Albuminuria"

Group 2 = "Non-Albuminuria"

**Negative values** indicate that the transcript has a lower expression Albuminuria group compared to Non-Albuminuria group.

```{r echo = F}
plot_de_markers(so_group_t2ob, oxy_phos_genes, "elevated_albuminuria", id1 = "Yes", id2 = "No", "PT","_oxphos_alb_t2dob")
kable(m_oxphos_alb_t2dob, digits = 3)
```

#### By SGLT2i (ever)

Group 1 = "SGLT2i"

Group 2 = "Non-SGLT2i"

**Negative values** indicate that the transcript has a lower expression SGLT2i group compared to Non-SGLT2i group.

```{r echo = F}
plot_de_markers(so_group_t2ob, oxy_phos_genes, "epic_ever_sglt2i_1", id1 = "Yes", id2 = "No", "PT","_oxphos_sglt2i_ever_t2dob")
kable(m_oxphos_sglt2i_ever_t2dob, digits = 3)
```

#### By SGLT2i (timepoint)

Group 1 = "SGLT2i"

Group 2 = "Non-SGLT2i"

**Negative values** indicate that the transcript has a lower expression SGLT2i group compared to Non-SGLT2i group.

```{r echo = F}
plot_de_markers(so_group_t2ob, oxy_phos_genes, "epic_sglti2_1", id1 = "Yes", id2 = "No", "PT","_oxphos_sglt2i_timpoint_t2dob")
kable(m_oxphos_sglt2i_timpoint_t2dob, digits = 3)
```

## Key transcripts of Glycolysis

```{r echo = F}
so$t2d_sglt2i_neg <- ifelse(so$group == "Type 2 Diabetes" & so$epic_ever_sglt2i_1 == "No", "T2Di(-)", so$group)
so$t2d_sglt2i <- ifelse(so$group == "Type 2 Diabetes" & so$epic_ever_sglt2i_1 == "Yes", "T2Di(+)", so$t2d_sglt2i_neg)

glycolysis_genes <- c("PKLR", "PFKFB3", "PFKL", "ALDOC", "HK2", "ENO2", "PGK1", "PGAM1", "TPI1", "GAPDH")
```

### JCI (T2Di(-) vs. HC)
#### By T2D status (T2Di(-) vs HC)

Group 1 = "T2Di(-)"

Group 2 = "HC"

**Negative values** indicate that the transcript has a lower expression T2Di(-) group compared to HC group.

```{r echo = F}
plot_de_markers(so, glycolysis_genes, "t2d_sglt2i_neg", id1 = "T2Di(-)", id2 = "Lean Control", "PT","_glyc_t2di_hc")
kable(m_glyc_t2di_hc, digits = 3)
```

### JCI (T2Di(+) vs. T2Di(-))
#### By T2D status (T2Di(+) vs T2Di(-))

Group 1 = "T2Di(+)"

Group 2 = "T2Di(-)"

**Negative values** indicate that the transcript has a lower expression T2Di(+) group compared to T2Di(-) group.

```{r echo = F}
plot_de_markers(so, glycolysis_genes, "t2d_sglt2i", id1 = "T2Di(+)", id2 = "T2Di(-)", "PT","_glyc_t2di")
kable(m_glyc_t2di, digits = 3)
```

## Key transcripts of Gluconeogensesis

```{r echo = F}
gluconeogenesis_genes <- c("SLC25A10", "GOT2", "GOT1", "FBP1", "SLC25A11", "PCK1", "MDH1")
```

### JCI (T2Di(-) vs. HC)
#### By T2D status (T2Di(-) vs HC)

Group 1 = "T2Di(-)"

Group 2 = "HC"

**Negative values** indicate that the transcript has a lower expression T2Di(-) group compared to HC group.

```{r echo = F}
plot_de_markers(so, gluconeogenesis_genes, "t2d_sglt2i_neg", id1 = "T2Di(-)", id2 = "Lean Control", "PT","_glucogen_t2di_hc")
kable(m_glucogen_t2di_hc, digits = 3)
```

### JCI (T2Di(+) vs. T2Di(-))
#### By T2D status (T2Di(+) vs T2Di(-))

Group 1 = "T2Di(+)"

Group 2 = "T2Di(-)"

**Negative values** indicate that the transcript has a lower expression T2Di(+) group compared to T2Di(-) group.

```{r echo = F}
plot_de_markers(so, gluconeogenesis_genes, "t2d_sglt2i", id1 = "T2Di(+)", id2 = "T2Di(-)", "PT","_glucogen_t2di")
kable(m_glucogen_t2di, digits = 3)
```

## Key transcripts of Pyruvate metabolism and TCA cycle

```{r echo = F}
pyruvate_tca <- c("SDHB", "SUCLG1", "PDK2", "ACO2", "IDH3G", "SUCLA2", "HAGH", "PDHB", "LDHA")
```

### JCI (T2Di(-) vs. HC)
#### By T2D status (T2Di(-) vs HC)

Group 1 = "T2Di(-)"

Group 2 = "HC"

**Negative values** indicate that the transcript has a lower expression T2Di(-) group compared to HC group.

```{r echo = F}
plot_de_markers(so, pyruvate_tca, "t2d_sglt2i_neg", id1 = "T2Di(-)", id2 = "Lean Control", "PT","_pyruvate_tca_t2di_hc")
kable(m_pyruvate_tca_t2di_hc, digits = 3)
```

### JCI (T2Di(+) vs. T2Di(-))
#### By T2D status (T2Di(+) vs T2Di(-))

Group 1 = "T2Di(+)"

Group 2 = "T2Di(-)"

**Negative values** indicate that the transcript has a lower expression T2Di(+) group compared to T2Di(-) group.

```{r echo = F}
plot_de_markers(so, pyruvate_tca, "t2d_sglt2i", id1 = "T2Di(+)", id2 = "T2Di(-)", "PT","_pyruvate_tca_t2di")
kable(m_pyruvate_tca_t2di, digits = 3)
```

## Key transcripts of Glutathione conjugation

```{r echo = F}
glut_conj <- c("CNDP2", "GSTM4", "GSTT2B", "GSTO1", "GGCT", "GSTM3", "AKR1A1")
```

### JCI (T2Di(-) vs. HC)
#### By T2D status (T2Di(-) vs HC)

Group 1 = "T2Di(-)"

Group 2 = "HC"

**Negative values** indicate that the transcript has a lower expression T2Di(-) group compared to HC group.

```{r echo = F}
plot_de_markers(so, glut_conj, "t2d_sglt2i_neg", id1 = "T2Di(-)", id2 = "Lean Control", "PT","_glut_conj_t2di_hc")
kable(m_glut_conj_t2di_hc, digits = 3)
```

### JCI (T2Di(+) vs. T2Di(-))
#### By T2D status (T2Di(+) vs T2Di(-))

Group 1 = "T2Di(+)"

Group 2 = "T2Di(-)"

**Negative values** indicate that the transcript has a lower expression T2Di(+) group compared to T2Di(-) group.

```{r echo = F}
plot_de_markers(so, glut_conj, "t2d_sglt2i", id1 = "T2Di(+)", id2 = "T2Di(-)", "PT","_glut_conj_t2di")
kable(m_glut_conj_t2di, digits = 3)
```
## Key transcripts of Metallothioneins bind metals

```{r echo = F}
metal_genes <- c("MT1G", "MT1X", "MT1H", "MT2A")
```

### JCI (T2Di(-) vs. HC)
#### By T2D status (T2Di(-) vs HC)

Group 1 = "T2Di(-)"

Group 2 = "HC"

**Negative values** indicate that the transcript has a lower expression T2Di(-) group compared to HC group.

```{r echo = F}
plot_de_markers(so, metal_genes, "t2d_sglt2i_neg", id1 = "T2Di(-)", id2 = "Lean Control", "PT","_metal_t2di_hc")
kable(m_metal_t2di_hc, digits = 3)
```

### JCI (T2Di(+) vs. T2Di(-))
#### By T2D status (T2Di(+) vs T2Di(-))

Group 1 = "T2Di(+)"

Group 2 = "T2Di(-)"

**Negative values** indicate that the transcript has a lower expression T2Di(+) group compared to T2Di(-) group.

```{r echo = F}
plot_de_markers(so, metal_genes, "t2d_sglt2i", id1 = "T2Di(+)", id2 = "T2Di(-)", "PT","_metal_t2di")
kable(m_metal_t2di, digits = 3)
```

## Key transcripts of lipid synthesis
### Combined (T2D + OB + LC)
#### By T2D status (T2D vs. OB + LC)

Group 1 = "T2D"

Group 2 = "Non-T2D"

**Negative values** indicate that the transcript has a lower expression T2D group compared to non-T2D group.

```{r echo = F}
lipid_genes <- c("FAS", "ACACB", "ACACA")
```

```{r echo = F}
plot_de_markers(so, lipid_genes, "group_t2d", "T2D", "Non-T2D", "PT","_tca_t2d")
kable(m_tca_t2d, digits = 3)
```

#### By T2D status (T2D + OB vs. LC)
Group 1 = "T2D/OB"

Group 2 = "LC"

**Negative values** indicate that the transcript has a lower expression T2D/OB group compared to LC group.

```{r echo = F}
plot_de_markers(so, lipid_genes, "group_t2ob", id1 = "T2D/OB", id2 = "LC", "PT","_tca_t2ob")
kable(m_tca_t2ob, digits = 3)
```




### T2D & OB only (T2D + OB)
#### By T2D status (T2D vs. OB)

Group 1 = "T2D"

Group 2 = "Obese Control"

**Negative values** indicate that the transcript has a lower expression T2D group compared to Obese group.

```{r echo = F}
plot_de_markers(so_group_t2ob, lipid_genes, "group", id1 = "Type 2 Diabetes", id2 = "Obese Control", "PT","_tca_t2d_t2dob")
kable(m_tca_t2d_t2dob, digits = 3)
```

#### By Albuminuria

Group 1 = "Albuminuria"

Group 2 = "Non-Albuminuria"

**Negative values** indicate that the transcript has a lower expression Albuminuria group compared to Non-Albuminuria group.

```{r echo = F}
plot_de_markers(so_group_t2ob, lipid_genes, "elevated_albuminuria", id1 = "Yes", id2 = "No", "PT","_tca_alb_t2dob")
kable(m_tca_alb_t2dob, digits = 3)
```

#### By SGLT2i (ever)

Group 1 = "SGLT2i"

Group 2 = "Non-SGLT2i"

**Negative values** indicate that the transcript has a lower expression SGLT2i group compared to Non-SGLT2i group.

```{r echo = F}
plot_de_markers(so_group_t2ob, lipid_genes, "epic_ever_sglt2i_1", id1 = "Yes", id2 = "No", "PT","_tca_sglt2i_ever_t2dob")
kable(m_tca_sglt2i_ever_t2dob, digits = 3)
```

#### By SGLT2i (timepoint)

Group 1 = "SGLT2i"

Group 2 = "Non-SGLT2i"

**Negative values** indicate that the transcript has a lower expression SGLT2i group compared to Non-SGLT2i group.

```{r echo = F}
plot_de_markers(so_group_t2ob, lipid_genes, "epic_sglti2_1", id1 = "Yes", id2 = "No", "PT","_tca_sglt2i_timpoint_t2dob")
kable(m_tca_sglt2i_timpoint_t2dob, digits = 3)
```

# DEG in PT (Subset to JCI IDs from old dataset PB68)

scRNA data file: "seurat_data_no_computations.RDS"

```{r echo = F}
so_old <- readRDS("/run/user/1023/gvfs/smb-share:server=ucdenver.pvt,share=som/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/scRNA/data_clean/seurat_data_no_computations.RDS")
jci_ids <- read.csv("/run/user/1023/gvfs/smb-share:server=ucdenver.pvt,share=som/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/T2D scRNA SGLT2i/Data raw/clinical_from_RDS.csv") %>%
  mutate(orig.ident = gsub("-PB","", orig.ident))
dat <- read.csv("/run/user/1023/gvfs/smb-share:server=ucdenver.pvt,share=som/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/Data Harmonization/Data Clean/harmonized_dataset.csv") %>%
  dplyr::mutate(across(where(is.character), ~ na_if(., ""))) %>%
  group_by(record_id, visit) %>%
  fill(kit_id, .direction = "updown") %>% ungroup() %>%
  dplyr::summarise(across(where(negate(is.numeric)), ~ ifelse(all(is.na(.x)), NA_character_, last(na.omit(.x)))),
                   across(where(is.numeric), ~ ifelse(all(is.na(.x)), NA_real_, mean(.x, na.rm = TRUE))),
                   .by = c(kit_id)) %>% filter(!is.na(kit_id)) %>%
  dplyr::mutate(record_id = paste0(record_id, visit)) %>%
  dplyr::select(c(record_id, group, acr_u, eGFR_CKD_epi, kit_id,
                  elevated_albuminuria, epic_ever_sglt2i_1,epic_sglti2_1))

so_old_meta <- so_old@meta.data %>%
  dplyr::select(1:13) %>%
  plyr::rename(c("KL.ID..kit.number.contains.path.sc.sample." = "kit_id")) 

so_old_meta <- left_join(so_old_meta, jci_ids) %>%
  left_join(dat)
rownames(so_old_meta) <- rownames(so_old@meta.data)
so_old <- AddMetaData(so_old, so_old_meta)

# IDs included in JCI
so_oldid_subset <- subset(so_old, Age != "NA")
unique(so_oldid_subset$michigan_id)

Idents(so_oldid_subset) <- so_oldid_subset$LR_clusters
```

# Table of descriptives
```{r echo = F, results='asis'}
so_old_meta_sub <- so_oldid_subset@meta.data %>%
  distinct(michigan_id, .keep_all = T) %>%
  mutate(Age = as.numeric(Age),
         diabetes_duration = as.numeric(diabetes_duration_months)/12,
         sex = factor(sex, labels = c("Female", "Male")))
summary(arsenal::tableby(group ~ Age + sex + diabetes_duration + eGFR_CKD_epi + kwt(acr_u, "medianq1q3", "range"),
                 data = so_old_meta_sub, test = F, total = F))
```

## Key transcripts of TCA cycle
### Combined (T2D + LC)
#### By T2D status (T2D vs. LC)

Group 1 = "T2D"

Group 2 = "LC"

**Negative values** indicate that the transcript has a lower expression T2D group compared to LC group.

```{r echo = F}
plot_de_markers(so_oldid_subset, genes, "group", "Type 2 Diabetes", "Lean Control", "PT","_tca_t2d")
kable(m_tca_t2d, digits = 3)
```

### T2D only

#### By Albuminuria

Group 1 = "Albuminuria"

Group 2 = "Non-Albuminuria"

**Negative values** indicate that the transcript has a lower expression Albuminuria group compared to Non-Albuminuria group.

```{r echo = F}
so_oldid_subset_group <- subset(so_oldid_subset, group == "Type 2 Diabetes")

plot_de_markers(so_oldid_subset_group, genes, "elevated_albuminuria", id1 = "Yes", id2 = "No", "PT","_tca_alb_t2dob")
kable(m_tca_alb_t2dob, digits = 3)
```

#### By SGLT2i (ever)

Group 1 = "SGLT2i"

Group 2 = "Non-SGLT2i"

**Negative values** indicate that the transcript has a lower expression SGLT2i group compared to Non-SGLT2i group.

```{r echo = F}
plot_de_markers(so_oldid_subset_group, genes, "epic_ever_sglt2i_1", id1 = "Yes", id2 = "No", "PT","_tca_sglt2i_ever_t2dob")
kable(m_tca_sglt2i_ever_t2dob, digits = 3)
```

#### By SGLT2i (timepoint)

Group 1 = "SGLT2i"

Group 2 = "Non-SGLT2i"

**Negative values** indicate that the transcript has a lower expression SGLT2i group compared to Non-SGLT2i group.

```{r echo = F}
plot_de_markers(so_oldid_subset_group, genes, "epic_sglti2_1", id1 = "Yes", id2 = "No", "PT","_tca_sglt2i_timpoint_t2dob")
kable(m_tca_sglt2i_timpoint_t2dob, digits = 3)
```


## Key transcripts of Oxidative Phosphorylation
### Combined (T2D + LC)
#### By T2D status (T2D vs.  LC)

Group 1 = "T2D"

Group 2 = "LC"

**Negative values** indicate that the transcript has a lower expression T2D group compared to LC group.

```{r echo = F}
plot_de_markers(so_oldid_subset, oxy_phos_genes, "group", "Type 2 Diabetes", "Lean Control", "PT","_oxphos_t2d")
kable(m_oxphos_t2d, digits = 3)
```

### T2D only

#### By Albuminuria

Group 1 = "Albuminuria"

Group 2 = "Non-Albuminuria"

**Negative values** indicate that the transcript has a lower expression Albuminuria group compared to Non-Albuminuria group.

```{r echo = F}
plot_de_markers(so_oldid_subset_group, oxy_phos_genes, "elevated_albuminuria", id1 = "Yes", id2 = "No", "PT","_oxphos_alb_t2dob")
kable(m_oxphos_alb_t2dob, digits = 3)
```

#### By SGLT2i (ever)

Group 1 = "SGLT2i"

Group 2 = "Non-SGLT2i"

**Negative values** indicate that the transcript has a lower expression SGLT2i group compared to Non-SGLT2i group.

```{r echo = F}
plot_de_markers(so_oldid_subset_group, oxy_phos_genes, "epic_ever_sglt2i_1", id1 = "Yes", id2 = "No", "PT","_oxphos_sglt2i_ever_t2dob")
kable(m_oxphos_sglt2i_ever_t2dob, digits = 3)
```

#### By SGLT2i (timepoint)

Group 1 = "SGLT2i"

Group 2 = "Non-SGLT2i"

**Negative values** indicate that the transcript has a lower expression SGLT2i group compared to Non-SGLT2i group.

```{r echo = F}
plot_de_markers(so_oldid_subset_group, oxy_phos_genes, "epic_sglti2_1", id1 = "Yes", id2 = "No", "PT","_oxphos_sglt2i_timpoint_t2dob")
kable(m_oxphos_sglt2i_timpoint_t2dob, digits = 3)
```

## Key transcripts of Glycolysis

```{r echo = F}
so_oldid_subset$t2d_sglt2i_neg <- ifelse(so_oldid_subset$group == "Type 2 Diabetes" & so_oldid_subset$epic_ever_sglt2i_1 == "No", "T2Di(-)", so_oldid_subset$group)
so_oldid_subset$t2d_sglt2i <- ifelse(so_oldid_subset$group == "Type 2 Diabetes" & so_oldid_subset$epic_ever_sglt2i_1 == "Yes", "T2Di(+)", so_oldid_subset$t2d_sglt2i_neg)

glycolysis_genes <- c("PKLR", "PFKFB3", "PFKL", "ALDOC", "HK2", "ENO2", "PGK1", "PGAM1", "TPI1", "GAPDH")
```

### JCI (T2Di(-) vs. HC)
#### By T2D status (T2Di(-) vs HC)

Group 1 = "T2Di(-)"

Group 2 = "HC"

**Negative values** indicate that the transcript has a lower expression T2Di(-) group compared to HC group.

```{r echo = F}
plot_de_markers(so_oldid_subset, glycolysis_genes, "t2d_sglt2i_neg", id1 = "T2Di(-)", id2 = "Lean Control", "PT","_glyc_t2di_hc")
kable(m_glyc_t2di_hc, digits = 3)
```

### JCI (T2Di(+) vs. T2Di(-))
#### By T2D status (T2Di(+) vs T2Di(-))

Group 1 = "T2Di(+)"

Group 2 = "T2Di(-)"

**Negative values** indicate that the transcript has a lower expression T2Di(+) group compared to T2Di(-) group.

```{r echo = F}
plot_de_markers(so_oldid_subset, glycolysis_genes, "t2d_sglt2i", id1 = "T2Di(+)", id2 = "T2Di(-)", "PT","_glyc_t2di")
kable(m_glyc_t2di, digits = 3)
```

## Key transcripts of Gluconeogensesis

```{r echo = F}
gluconeogenesis_genes <- c("SLC25A10", "GOT2", "GOT1", "FBP1", "SLC25A11", "PCK1", "MDH1")
```

### JCI (T2Di(-) vs. HC)
#### By T2D status (T2Di(-) vs HC)

Group 1 = "T2Di(-)"

Group 2 = "HC"

**Negative values** indicate that the transcript has a lower expression T2Di(-) group compared to HC group.

```{r echo = F}
plot_de_markers(so_oldid_subset, gluconeogenesis_genes, "t2d_sglt2i_neg", id1 = "T2Di(-)", id2 = "Lean Control", "PT","_glucogen_t2di_hc")
kable(m_glucogen_t2di_hc, digits = 3)
```

### JCI (T2Di(+) vs. T2Di(-))
#### By T2D status (T2Di(+) vs T2Di(-))

Group 1 = "T2Di(+)"

Group 2 = "T2Di(-)"

**Negative values** indicate that the transcript has a lower expression T2Di(+) group compared to T2Di(-) group.

```{r echo = F}
plot_de_markers(so_oldid_subset, gluconeogenesis_genes, "t2d_sglt2i", id1 = "T2Di(+)", id2 = "T2Di(-)", "PT","_glucogen_t2di")
kable(m_glucogen_t2di, digits = 3)
```
## Key transcripts of Pyruvate metabolism and TCA cycle

```{r echo = F}
pyruvate_tca <- c("SDHB", "SUCLG1", "PDK2", "ACO2", "IDH3G", "SUCLA2", "HAGH", "PDHB", "LDHA")
```

### JCI (T2Di(-) vs. HC)
#### By T2D status (T2Di(-) vs HC)

Group 1 = "T2Di(-)"

Group 2 = "HC"

**Negative values** indicate that the transcript has a lower expression T2Di(-) group compared to HC group.

```{r echo = F}
plot_de_markers(so_oldid_subset, pyruvate_tca, "t2d_sglt2i_neg", id1 = "T2Di(-)", id2 = "Lean Control", "PT","_pyruvate_tca_t2di_hc")
kable(m_pyruvate_tca_t2di_hc, digits = 3)
```

### JCI (T2Di(+) vs. T2Di(-))
#### By T2D status (T2Di(+) vs T2Di(-))

Group 1 = "T2Di(+)"

Group 2 = "T2Di(-)"

**Negative values** indicate that the transcript has a lower expression T2Di(+) group compared to T2Di(-) group.

```{r echo = F}
plot_de_markers(so_oldid_subset, pyruvate_tca, "t2d_sglt2i", id1 = "T2Di(+)", id2 = "T2Di(-)", "PT","_pyruvate_tca_t2di")
kable(m_pyruvate_tca_t2di, digits = 3)
```

## Key transcripts of Glutathione conjugation

```{r echo = F}
glut_conj <- c("CNDP2", "GSTM4", "GSTT2B", "GSTO1", "GGCT", "GSTM3", "AKR1A1")
```

### JCI (T2Di(-) vs. HC)
#### By T2D status (T2Di(-) vs HC)

Group 1 = "T2Di(-)"

Group 2 = "HC"

**Negative values** indicate that the transcript has a lower expression T2Di(-) group compared to HC group.

```{r echo = F}
plot_de_markers(so_oldid_subset, glut_conj, "t2d_sglt2i_neg", id1 = "T2Di(-)", id2 = "Lean Control", "PT","_glut_conj_t2di_hc")
kable(m_glut_conj_t2di_hc, digits = 3)
```

### JCI (T2Di(+) vs. T2Di(-))
#### By T2D status (T2Di(+) vs T2Di(-))

Group 1 = "T2Di(+)"

Group 2 = "T2Di(-)"

**Negative values** indicate that the transcript has a lower expression T2Di(+) group compared to T2Di(-) group.

```{r echo = F}
plot_de_markers(so_oldid_subset, glut_conj, "t2d_sglt2i", id1 = "T2Di(+)", id2 = "T2Di(-)", "PT","_glut_conj_t2di")
kable(m_glut_conj_t2di, digits = 3)
```
## Key transcripts of Metallothioneins bind metals

```{r echo = F}
metal_genes <- c("MT1G", "MT1X", "MT1H", "MT2A")
```

### JCI (T2Di(-) vs. HC)
#### By T2D status (T2Di(-) vs HC)

Group 1 = "T2Di(-)"

Group 2 = "HC"

**Negative values** indicate that the transcript has a lower expression T2Di(-) group compared to HC group.

```{r echo = F}
plot_de_markers(so_oldid_subset, metal_genes, "t2d_sglt2i_neg", id1 = "T2Di(-)", id2 = "Lean Control", "PT","_metal_t2di_hc")
kable(m_metal_t2di_hc, digits = 3)
```

### JCI (T2Di(+) vs. T2Di(-))
#### By T2D status (T2Di(+) vs T2Di(-))

Group 1 = "T2Di(+)"

Group 2 = "T2Di(-)"

**Negative values** indicate that the transcript has a lower expression T2Di(+) group compared to T2Di(-) group.

```{r echo = F}
plot_de_markers(so_oldid_subset, metal_genes, "t2d_sglt2i", id1 = "T2Di(+)", id2 = "T2Di(-)", "PT","_metal_t2di")
kable(m_metal_t2di, digits = 3)
```

```{r echo = F, include = F}
# str_sort(unique(so$record_id))
# str_sort(unique(so_oldid_subset$record_id))
unique(so$record_id)[unique(so$record_id) %nin% unique(so_oldid_subset$record_id)]

unique(so_old$michigan_id)
unique(so_oldid_subset$record_id)

nrow(so_oldid_subset)

```


# Correlations

```{r echo = F}
# Correlation / heat map showing association between PET data and transcripts (and consider estimates of insulin sensitivity) (RH/RH2, CRC)
# Extract individual level expression of scRNA
# meta_features <- so[["RNA"]]@meta.features
# a <- rownames(meta_features)
# var_list <- c("sample", "ident", "michigan_id", genes_subset)
# pt_md <- FetchData(so, vars = var_list) %>%
# filter(ident == "PT")
```

## Combined (T2D + OB + LC)

```{r echo = F}

```

## T2D & OB only (T2D + OB)


```{r echo = F}
# Pseudotime trajectories according to T2D status, albuminuria and SGLT2i as well as proportion of PT subpopulations across these groups (RH/RH2, CRC)

```

