---
title: "Salud Aim 2 Analysis"
author: "Casey Sakamoto"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE, echo=FALSE}
library(haven) # import spss dataset
library(knitr)
library(nlme) # mixed models
library(emmeans) # contrast statement
library(tidyverse) # data manipulation
library(Hmisc)

knitr::opts_chunk$set(echo = FALSE)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "S:/Laura/Peds Endo/Lauren Shomaker/Salud"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Lauren Shomaker/Salud"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Lauren Shomaker/Salud"
}
knitr::opts_knit$set(root.dir = home_dir)

# import data
## make sure connected to vpn
HLP_MASTER_DATA <- read_sav("S:/Laura/Peds Endo/Lauren Shomaker/Salud/Data Analysis - Pyle 2020/HLP_MASTER DATA 7.15.20 LP Trim.sav")

```

# 1. Background
Aim 2 Hypothesis: We estimate that the MBI + Lifestyle (SSB) will be more effective than Lifestyle Only (HeLP) for reducing depression symptoms, insulin resistance, and stress.Depression, insulin, and stress measures were measured at baseline, 6 week, and 6 month visits.

# 2. Outstanding Data Questions/ Remarks
- We have 12 subjects with 3 time points (BL, 6wk, 6mo), 1 subj with 2 (BL, 6wk : subj 117 (has some missing data)), 5 with just BL

- Cortisol was measured before and after over 6 sessions. the change in cortisol was derived, and the 6 sessions were used as the time points rather than the 3 time points used for the other measures

# 3. Methods
CES-D scores were used to measure teen depression. HOMA-IR, fasting insulin, fasting glucose were used to measure insulin resistance. HBA1C, ALT/AST, triglycerides, total cholesterol, and HDL/LDL were measured for other metabolic outcomes. The Perceived Stress Scale (PSS), Everyday Discrimination Scale (EDS), and Psychological Stress Experience (PSE) were used to measure Stress. The change in cortisol levels before and after sessions (6 total) were also used to measure stress.

Groups were compared at each time point by fitting linear mixed models with time, treatment, an interaction term, and random intercepts. T-tests with the Tukey p-value adjustment were performed to evaluate the differences between groups at each time point. Spaghetti plots were added to visualize the outcomes at a subject level.

Model fits and contrast tests are included in the code appendix.

```{r data exploration and clean, include=FALSE}
###### light exploration #################
#look at what we are working with
unique(HLP_MASTER_DATA$subjectidno)
length(unique(HLP_MASTER_DATA$subjectidno)) # 18 total subj

# double check homairlg is a log transform of homair
a  = HLP_MASTER_DATA %>% select(subjectidno,homair, homairlg) 
a$testhomair = log(a$homair) # base e doesnt match
a$testhomair = log10(a$homair) # base 10 matches
##########################################

#####
# for some reason no more t-score in the new data set- will pull from the older data set
HLP_MASTER_DATA_7_14_20_LP_ = read_sav("S:/Laura/Peds Endo/Lauren Shomaker/Salud/Data Analysis - Pyle 2020/Older/HLP_MASTER DATA 7.14.20 LP .sav")

HLP_MASTER_DATA_7_14_20_LP_ = HLP_MASTER_DATA_7_14_20_LP_ %>% select(subjectidno, interval, pserawsumT)
####

###### cortisol derivation #####
hlp_cortisol = HLP_MASTER_DATA %>% mutate(cortisol_change1 = Cortisol_nmoll_session1_after - Cortisol_nmoll_session1_before,
                                          cortisol_change2 = Cortisol_nmoll_session2_after - Cortisol_nmoll_session2_before,
                                          cortisol_change3 = Cortisol_nmoll_session3_after - Cortisol_nmoll_session3_before,
                                          cortisol_change4 = Cortisol_nmoll_session4_after - Cortisol_nmoll_session4_before,
                                          cortisol_change5 = Cortisol_nmoll_session5_after - Cortisol_nmoll_session5_before,
                                          cortisol_change6 = Cortisol_nmoll_session6_after - Cortisol_nmoll_session6_before) %>%
  select(subjectidno, interval, cortisol_change1, cortisol_change2, cortisol_change3, cortisol_change4, cortisol_change5, cortisol_change6)

###### get already cleaned variables from data #################
analysis = HLP_MASTER_DATA %>% select(subjectidno, interval, groupno, sex, SV2age_calculated,z_score,
                                      cesdtot, homair, insulin_mcIUmL, glucose_mgdL,
                                      hba1c, ALT_UL, AST_UL, triglycerides_mgdL, cholesteroltot_mgdL, HDLC_mgdL, LDLC_mgdL,
                                      psssumitem, EDSrawscore) 

analysis = full_join(analysis, HLP_MASTER_DATA_7_14_20_LP_)
analysis = full_join(analysis, hlp_cortisol)

# order and factor variables
analysis = analysis[order(analysis$subjectidno, analysis$interval),]
analysis$groupno = factor(analysis$groupno, levels = c(0,1), labels = c("HeLP","SSB"))
analysis$interval = factor(analysis$interval, levels = c(1,2,3), labels = c("Baseline", "6 Week", "6 Month"))
analysis$sex = factor(analysis$sex, levels = c(0,1), labels = c("Male", "Female"))

# clean up workspace
rm(hlp_cortisol); rm(HLP_MASTER_DATA); rm(HLP_MASTER_DATA_7_14_20_LP_); rm(a)
###########################################
```

```{r mixed model, include = FALSE}
###### Teen Depression (cesdtot) #########

# simple models
cesdtot_mod = lme(cesdtot ~ interval+groupno+interval:groupno,
                   random=~1|subjectidno,
                   data = analysis)

cesdtot_mod_anova = anova.lme(cesdtot_mod, type="marginal")
# use emmeans 
cesdtot_mod_means = emmeans(cesdtot_mod,specs=pairwise ~ factor(interval):groupno, adjust="tukey")

#adjusted for age, bmi, sex
cesdtot_mod2 = lme(cesdtot ~ interval*groupno +sex+z_score,
                   random=~1|subjectidno,
                   data = analysis, na.action = na.omit)

cesdtot_mod_anova2 = anova.lme(cesdtot_mod2, type="marginal")
# use emmeans 
cesdtot_mod_means2 = emmeans(cesdtot_mod2,specs=pairwise ~ factor(interval):groupno, adjust="tukey")
########################################

###### Insulin Resistance (homair) #########

# simple models
homair_mod = lme(homair ~ interval+groupno+interval:groupno,
                   random=~1|subjectidno,
                   data = analysis)

homair_mod_anova = anova.lme(homair_mod, type="marginal")
# use emmeans 
homair_mod_means = emmeans(homair_mod,specs=pairwise ~ factor(interval):groupno, adjust="tukey")

#adjusted for age, bmi, sex
homair_mod2 = lme(homair ~ interval*groupno +sex+z_score,
                   random=~1|subjectidno,
                   data = analysis, na.action = na.omit)

homair_mod_anova2 = anova.lme(homair_mod2, type="marginal")
# use emmeans 
homair_mod_means2 = emmeans(homair_mod2,specs=pairwise ~ factor(interval):groupno, adjust="tukey")
########################################

###### Insulin Resistance (fasting insulin) #########
# simple models
insulin_mod = lme(insulin_mcIUmL ~ interval+groupno+interval:groupno,
                   random=~1|subjectidno,
                   data = analysis)

insulin_mod_anova = anova.lme(insulin_mod, type="marginal")
# use emmeans 
insulin_mod_means = emmeans(insulin_mod,specs=pairwise ~ factor(interval):groupno, adjust="tukey")

#adjusted for age, bmi, sex
insulin_mod2 = lme(insulin_mcIUmL ~ interval*groupno +sex+z_score,
                   random=~1|subjectidno,
                   data = analysis, na.action = na.omit)

insulin_mod_anova2 = anova.lme(insulin_mod2, type="marginal")
# use emmeans 
insulin_mod_means2 = emmeans(insulin_mod2,specs=pairwise ~ factor(interval):groupno, adjust="tukey")
########################################

###### Insulin Resistance (glucose) #########
# simple models
glucose_mod = lme(glucose_mgdL ~ interval+groupno+interval:groupno,
                   random=~1|subjectidno,
                   data = analysis)

glucose_mod_anova = anova.lme(glucose_mod, type="marginal")
# use emmeans 
glucose_mod_means = emmeans(glucose_mod,specs=pairwise ~ factor(interval):groupno, adjust="tukey")

#adjusted for age, bmi, sex
glucose_mod2 = lme(glucose_mgdL ~ interval*groupno +sex+z_score,
                   random=~1|subjectidno,
                   data = analysis, na.action = na.omit)

glucose_mod_anova2 = anova.lme(glucose_mod2, type="marginal")
# use emmeans 
glucose_mod_means2 = emmeans(glucose_mod2,specs=pairwise ~ factor(interval):groupno, adjust="tukey")
########################################

###### metabolic outcomes (hba1c) #########
# simple models
hba1c_mod = lme(hba1c ~ interval+groupno+interval:groupno,
                   random=~1|subjectidno,
                   data = analysis)

hba1c_mod_anova = anova.lme(hba1c_mod, type="marginal")
# use emmeans 
hba1c_mod_means = emmeans(hba1c_mod,specs=pairwise ~ factor(interval):groupno, adjust="tukey")

#adjusted for age, bmi, sex
hba1c_mod2 = lme(hba1c ~ interval*groupno +sex+z_score,
                   random=~1|subjectidno,
                   data = analysis, na.action = na.omit)

hba1c_mod_anova2 = anova.lme(hba1c_mod2, type="marginal")
# use emmeans 
hba1c_mod_means2 = emmeans(hba1c_mod2,specs=pairwise ~ factor(interval):groupno, adjust="tukey")
########################################

###### metabolic outcomes (AST/ALT) #########
# simple models
alt_mod = lme(ALT_UL ~ interval+groupno+interval:groupno,
                   random=~1|subjectidno,
                   data = analysis)

alt_mod_anova = anova.lme(alt_mod, type="marginal")

ast_mod = lme(AST_UL ~ interval+groupno+interval:groupno,
                   random=~1|subjectidno,
                   data = analysis)
ast_mod_anova = anova.lme(ast_mod, type="marginal")
# use emmeans 
alt_mod_means = emmeans(alt_mod,specs=pairwise ~ factor(interval):groupno, adjust="tukey")
ast_mod_means = emmeans(ast_mod,specs=pairwise ~ factor(interval):groupno, adjust="tukey")
#adjusted for age, bmi, sex
alt_mod2 = lme(ALT_UL ~ interval*groupno +sex+z_score,
                   random=~1|subjectidno,
                   data = analysis, na.action = na.omit)
ast_mod2 = lme(AST_UL ~ interval*groupno +sex+z_score,
                   random=~1|subjectidno,
                   data = analysis, na.action = na.omit)
ast_mod_anova2 = anova.lme(ast_mod2, type="marginal")
alt_mod_anova2 = anova.lme(alt_mod2, type="marginal")
# use emmeans 
alt_mod_means2 = emmeans(alt_mod2,specs=pairwise ~ factor(interval):groupno, adjust="tukey")
ast_mod_means2 = emmeans(ast_mod2,specs=pairwise ~ factor(interval):groupno, adjust="tukey")

########################################

###### metabolic outcomes lipids #########
# simple models
lipid_mod = lme(triglycerides_mgdL ~ interval+groupno+interval:groupno,
                   random=~1|subjectidno,
                   data = analysis, na.action = na.omit)

lipid_mod_anova = anova.lme(lipid_mod, type="marginal")
# use emmeans 
lipid_mod_means = emmeans(lipid_mod,specs=pairwise ~ factor(interval):groupno, adjust="tukey")

#adjusted for age, bmi, sex
lipid_mod2 = lme(triglycerides_mgdL ~ interval*groupno +sex+z_score,
                   random=~1|subjectidno,
                   data = analysis, na.action = na.omit)

lipid_mod_anova2 = anova.lme(lipid_mod2, type="marginal")
# use emmeans 
lipid_mod_means2 = emmeans(lipid_mod2,specs=pairwise ~ factor(interval):groupno, adjust="tukey")
########################################

###### metabolic outcomes cholesterol total #########
# simple models
chol_mod = lme(cholesteroltot_mgdL ~ interval+groupno+interval:groupno,
                   random=~1|subjectidno,
                   data = analysis, na.action = na.omit)

chol_mod_anova = anova.lme(chol_mod, type="marginal")
# use emmeans 
chol_mod_means = emmeans(chol_mod,specs=pairwise ~ factor(interval):groupno, adjust="tukey")

#adjusted for age, bmi, sex
chol_mod2 = lme(cholesteroltot_mgdL~ interval*groupno +sex+z_score,
                   random=~1|subjectidno,
                   data = analysis, na.action = na.omit)

chol_mod_anova2 = anova.lme(chol_mod2, type="marginal")
# use emmeans 
chol_mod_means2 = emmeans(chol_mod2,specs=pairwise ~ factor(interval):groupno, adjust="tukey")
########################################

###### metabolic outcomes (LDL/HDL) #########
# simple models
analysis$LDLC_mgdL
ldl_mod = lme(LDLC_mgdL ~ interval+groupno+interval:groupno,
                   random=~1|subjectidno,
                   data = analysis, na.action = na.omit)

ldl_mod_anova = anova.lme(ldl_mod, type="marginal")

hdl_mod = lme(HDLC_mgdL ~ interval+groupno+interval:groupno,
                   random=~1|subjectidno,
                   data = analysis, na.action = na.omit)
hdl_mod_anova = anova.lme(hdl_mod, type="marginal")
# use emmeans 
ldl_mod_means = emmeans(ldl_mod,specs=pairwise ~ factor(interval):groupno, adjust="tukey")
hdl_mod_means = emmeans(hdl_mod,specs=pairwise ~ factor(interval):groupno, adjust="tukey")
#adjusted for age, bmi, sex
ldl_mod2 = lme(LDLC_mgdL ~ interval*groupno +sex+z_score,
                   random=~1|subjectidno,
                   data = analysis, na.action = na.omit)
hdl_mod2 = lme(HDLC_mgdL ~ interval*groupno +sex+z_score,
                   random=~1|subjectidno,
                   data = analysis, na.action = na.omit)
ldl_mod_anova2 = anova.lme(ldl_mod2, type="marginal")
hdl_mod_anova2 = anova.lme(hdl_mod2, type="marginal")
# use emmeans 
ldl_mod_means2 = emmeans(ldl_mod2,specs=pairwise ~ factor(interval):groupno, adjust="tukey")
hdl_mod_means2 = emmeans(hdl_mod2,specs=pairwise ~ factor(interval):groupno, adjust="tukey")

########################################

###### stress outcomes pss #########
# simple models
analysis$psssumitem
pss_mod = lme(psssumitem ~ interval+groupno+interval:groupno,
                   random=~1|subjectidno,
                   data = analysis, na.action = na.omit)

pss_mod_anova = anova.lme(pss_mod, type="marginal")
# use emmeans 
pss_mod_means = emmeans(pss_mod,specs=pairwise ~ factor(interval):groupno, adjust="tukey")

#adjusted for age, bmi, sex
pss_mod2 = lme(psssumitem~ interval*groupno +sex+z_score,
                   random=~1|subjectidno,
                   data = analysis, na.action = na.omit)

pss_mod_anova2 = anova.lme(pss_mod2, type="marginal")
# use emmeans 
pss_mod_means2 = emmeans(pss_mod2,specs=pairwise ~ factor(interval):groupno, adjust="tukey")
########################################

###### stress outcomes eds #########
# simple models
analysis$EDSrawscore
eds_mod = lme(EDSrawscore ~ interval+groupno+interval:groupno,
                   random=~1|subjectidno,
                   data = analysis, na.action = na.omit)

eds_mod_anova = anova.lme(eds_mod, type="marginal")
# use emmeans 
eds_mod_means = emmeans(eds_mod,specs=pairwise ~ factor(interval):groupno, adjust="tukey")

#adjusted for age, bmi, sex
eds_mod2 = lme(EDSrawscore ~ interval*groupno +sex+z_score,
                   random=~1|subjectidno,
                   data = analysis, na.action = na.omit)

eds_mod_anova2 = anova.lme(eds_mod2, type="marginal")
# use emmeans 
eds_mod_means2 = emmeans(eds_mod2,specs=pairwise ~ factor(interval):groupno, adjust="tukey")
########################################

###### stress outcomes pse #########
# simple models
analysis$pserawsumT
pse_mod = lme(pserawsumT ~ interval+groupno+interval:groupno,
                   random=~1|subjectidno,
                   data = analysis, na.action = na.omit)

pse_mod_anova = anova.lme(pse_mod, type="marginal")
# use emmeans 
pse_mod_means = emmeans(pse_mod,specs=pairwise ~ factor(interval):groupno, adjust="tukey")

#adjusted for age, bmi, sex
pse_mod2 = lme(pserawsumT ~ interval*groupno +sex+z_score,
                   random=~1|subjectidno,
                   data = analysis, na.action = na.omit)

pse_mod_anova2 = anova.lme(pse_mod2, type="marginal")
# use emmeans 
pse_mod_means2 = emmeans(pse_mod2,specs=pairwise ~ factor(interval):groupno, adjust="tukey")
########################################

###### stress outcomes pse #########
# simple models
analysis$pserawsumT
pse_mod = lme(pserawsumT ~ interval+groupno+interval:groupno,
                   random=~1|subjectidno,
                   data = analysis, na.action = na.omit)

pse_mod_anova = anova.lme(pse_mod, type="marginal")
# use emmeans 
pse_mod_means = emmeans(pse_mod,specs=pairwise ~ factor(interval):groupno, adjust="tukey")

#adjusted for age, bmi, sex
pse_mod2 = lme(pserawsumT ~ interval*groupno +sex+z_score,
                   random=~1|subjectidno,
                   data = analysis, na.action = na.omit)

pse_mod_anova2 = anova.lme(pse_mod2, type="marginal")
# use emmeans 
pse_mod_means2 = emmeans(pse_mod2,specs=pairwise ~ factor(interval):groupno, adjust="tukey")
########################################

###### stress outcomes cortisol change #########
# grab the cortisol session data
# note only 13 subjects of the 18 have cortisol data
analysis_cort = analysis %>% filter(interval=="6 Week") %>% 
  select(subjectidno, groupno, cortisol_change1:cortisol_change6)%>%
  pivot_longer(cortisol_change1:cortisol_change6,
               names_to = "session",
               names_prefix = "cortisol_change",
               names_transform = list(session = as.double),
               values_to = "cortisol")
  

cortisol_mod = lme(cortisol ~ as.factor(session)*groupno,
                   random=~1|subjectidno,
                   data = analysis_cort, na.action = na.omit)

cortisol_mod_anova = anova.lme(cortisol_mod, type="marginal")
# use emmeans 
cortisol_mod_means = emmeans(cortisol_mod,specs=pairwise ~ factor(session):groupno, adjust="tukey")

# #adjusted for age, bmi, sex
# pse_mod2 = lme(pserawsumT ~ interval*groupno +sex+z_score,
#                    random=~1|subjectidno,
#                    data = analysis, na.action = na.omit)
# 
# pse_mod_anova2 = anova.lme(pse_mod2, type="marginal")
# # use emmeans 
# pse_mod_means2 = emmeans(pse_mod2,specs=pairwise ~ factor(interval):groupno, adjust="tukey")
########################################
```

# 4. Results
## Depression Symptoms
```{r cesd plot}
ggplot(analysis, aes(x = interval, y = cesdtot, color = groupno, group=subjectidno))+
    geom_line(size = 1) + labs(y = "CES-D Total")
```

No difference in groups at each time point (p = .82, .84, 1). Similar results when adjusting for covariates.

## Insulin Resistance/ Other Metabolic Outcomes
### homa-ir
```{r homair plot}
ggplot(analysis, aes(x = interval, y = homair, color = groupno, group=subjectidno))+
    geom_line(size = 1) + labs(y = "HOMA-IR")
```

No difference in groups at each time point (p = .91, 1, .69). Similar results when adjusting for covariates.

### Fasting Insulin
```{r insulin plot}
ggplot(analysis, aes(x = interval, y = insulin_mcIUmL, color = groupno, group=subjectidno))+
    geom_line(size = 1) + labs(y = "Fasting Insulin mcIU/mL")
```

No difference in groups at each time point (p = .94, 1, .64). Similar results when adjusting for covariates.

### Fasting Glucose
```{r gluc plot}
ggplot(analysis, aes(x = interval, y = glucose_mgdL, color = groupno, group=subjectidno))+
    geom_line(size = 1) + labs(y = "Glucose mg/dL")
```

No difference in groups at each time point (p = .99, .58, .99). Similar results when adjusting for covariates.

### HBA1C
```{r hba1c plot}
ggplot(analysis, aes(x = interval, y = hba1c, color = groupno, group=subjectidno))+
    geom_line(size = 1) + labs(y = "HBA1C")
```

No difference in groups at each time point (p = 1, 1, .13). Similar results when adjusting for covariates.

### ALT/AST
```{r altast plot}
ggplot(analysis, aes(x = interval, y = ALT_UL, color = groupno, group=subjectidno))+
    geom_line(size = 1) + labs(y = "ALT UL")
ggplot(analysis, aes(x = interval, y = AST_UL, color = groupno, group=subjectidno))+
    geom_line(size = 1) + labs(y = "AST UL")
```

No difference in groups for ALT at each time point (p = .93, .98, 1). Similar results when adjusting for covariates
No difference in groups for AST at each time point (p = .94, .62, 1). Similar results when adjusting for covariates

### Triglycerides
```{r trig plot}
ggplot(analysis, aes(x = interval, y = triglycerides_mgdL, color = groupno, group=subjectidno))+
    geom_line(size = 1) + labs(y = "Triglycerides mg/dL")
```

No difference in groups at each time point (p = .96, 1, .86). Similar results when adjusting for covariates.

### Cholesterol
```{r chol plot}
ggplot(analysis, aes(x = interval, y = cholesteroltot_mgdL, color = groupno, group=subjectidno))+
    geom_line(size = 1) + labs(y = "Total Cholesterol mg/dL")
```

No difference in groups at each time point (p = .99, .99, .83). Similar results when adjusting for covariates.

### LDL/HDL
```{r ldlhdl plot}
ggplot(analysis, aes(x = interval, y = LDLC_mgdL, color = groupno, group=subjectidno))+
    geom_line(size = 1) + labs(y = "LDLC mg/dL")
ggplot(analysis, aes(x = interval, y = HDLC_mgdL, color = groupno, group=subjectidno))+
    geom_line(size = 1) + labs(y = "HDLC mg/dL")
```

No difference in groups for LDL at each time point (p = .96, .83, .91). Similar results when adjusting for covariates.
No difference in groups for HDL at each time point (p = .46, .26, .48). Similar results when adjusting for covariates.

## Stress
### PSS
```{r pss plot}
ggplot(analysis, aes(x = interval, y = psssumitem, color = groupno, group=subjectidno))+
    geom_line(size = 1) + labs(y = "Perceived Stress Scale Total")
```

No difference in groups at each time point (p = .74, .78, 1). Similar results when adjusting for covariates.

### EDS
```{r eds plot}
ggplot(analysis, aes(x = interval, y = EDSrawscore, color = groupno, group=subjectidno))+
    geom_line(size = 1) + labs(y = "Everyday Discrimination Scale Raw Score")
```

No difference in groups at each time point (p = 1, 1, 1). Similar results when adjusting for covariates.

### PSE
```{r pse plot}
ggplot(analysis, aes(x = interval, y = pserawsumT, color = groupno, group=subjectidno))+
    geom_line(size = 1) + labs(y = "Psychological Stress Experience T-Score")
```

No difference in groups at each time point (p = 1, .76, 1). Similar results when adjusting for covariates.

## Stress
```{r cort plot}
ggplot(analysis_cort, aes(x = session, y = cortisol, color = groupno, group=subjectidno))+
    geom_line(size = 1)
```

No difference in groups at each time point (p = .95, 1, .43, 1, 1, .97). Similar results when adjusting for covariates.

# 5. Code Appendix
```{r cesdtot results}
kable(cesdtot_mod_anova,caption="CESDtot anova",format.args = list(scientific = FALSE))
kable(cesdtot_mod_means$contrasts[c(3,8,12),],caption="Time point means.",format.args = list(scientific = FALSE))
```

```{r homair results}
kable(homair_mod_anova,caption="HOMA-IR anova",format.args = list(scientific = FALSE))
kable(homair_mod_means$contrasts[c(3,8,12),],caption="Time point means.",format.args = list(scientific = FALSE))
```

```{r insulin results}
kable(insulin_mod_anova,caption="Insulin anova",format.args = list(scientific = FALSE))
kable(insulin_mod_means$contrasts[c(3,8,12),],caption="Time point means.",format.args = list(scientific = FALSE))
```

```{r glucose results}
kable(glucose_mod_anova,caption="Glucose anova",format.args = list(scientific = FALSE))
kable(glucose_mod_means$contrasts[c(3,8,12),],caption="Time point means.",format.args = list(scientific = FALSE))
```

```{r hba1c results}
kable(hba1c_mod_anova,caption="HBA1C anova",format.args = list(scientific = FALSE))
kable(hba1c_mod_means$contrasts[c(3,8,12),],caption="Time point means.",format.args = list(scientific = FALSE))
```

```{r AST/ALT results}
kable(alt_mod_anova,caption="ALT anova",format.args = list(scientific = FALSE))
kable(alt_mod_means$contrasts[c(3,8,12),],caption="Time point means.",format.args = list(scientific = FALSE))

kable(ast_mod_anova,caption="ALT anova",format.args = list(scientific = FALSE))
kable(ast_mod_means$contrasts[c(3,8,12),],caption="Time point means.",format.args = list(scientific = FALSE))
```

```{r Lipids results}
kable(lipid_mod_anova,caption="Lipids (triglycerides) anova",format.args = list(scientific = FALSE))
kable(lipid_mod_means$contrasts[c(3,8,12),],caption="Time point means.",format.args = list(scientific = FALSE))
```

```{r Cholesterol results}
kable(chol_mod_anova,caption="Total Cholesterol anova",format.args = list(scientific = FALSE))
kable(chol_mod_means$contrasts[c(3,8,12),],caption="Time point means.",format.args = list(scientific = FALSE))
```

```{r LDL/HDL results}
kable(ldl_mod_anova,caption="LDL anova",format.args = list(scientific = FALSE))
kable(ldl_mod_means$contrasts[c(3,8,12),],caption="Time point means.",format.args = list(scientific = FALSE))

kable(hdl_mod_anova,caption="HDL anova",format.args = list(scientific = FALSE))
kable(hdl_mod_means$contrasts[c(3,8,12),],caption="Time point means.",format.args = list(scientific = FALSE))
```

```{r PSS results}
kable(pss_mod_anova,caption="Perceived Stress Scale anova",format.args = list(scientific = FALSE))
kable(pss_mod_means$contrasts[c(3,8,12),],caption="Time point means.",format.args = list(scientific = FALSE))
```

```{r EDS results}
kable(eds_mod_anova,caption="Everyday Discrimination Scale anova",format.args = list(scientific = FALSE))
kable(eds_mod_means$contrasts[c(3,8,12),],caption="Time point means.",format.args = list(scientific = FALSE))
```

```{r PSE results}
kable(pse_mod_anova,caption="Psychological Stress Experience anova",format.args = list(scientific = FALSE))
kable(pse_mod_means$contrasts[c(3,8,12),],caption="Time point means.",format.args = list(scientific = FALSE))
```

```{r cortisol results}
kable(cortisol_mod_anova,caption="Cortisol anova",format.args = list(scientific = FALSE))
kable(cortisol_mod_means$contrasts[c(6,17,27, 36, 44, 51),],caption="Time point means.",format.args = list(scientific = FALSE))
```


