---
title: "Salud Data Exploration/Analysis"
author: "Casey Sakamoto"
date:"`r format(Sys.time(), '%d %B %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
library(haven) # import spss dataset
library(lme4) # mixed models
library(multcomp) # contrast statement
library(tidyverse) # data manipulation

knitr::opts_chunk$set(echo = FALSE)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "S:/Laura/Peds Endo/Lauren Shomaker/Salud"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Lauren Shomaker/Salud"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Lauren Shomaker/Salud"
}
knitr::opts_knit$set(root.dir = home_dir)

# import data
## make sure connected to vpn
HLP_MASTER_DATA <- read_sav("S:/Laura/Peds Endo/Lauren Shomaker/Salud/Data Analysis - Pyle 2020/HLP_MASTER DATA 7.15.20 LP Trim.sav")

```
# Aim 1
we can do

# Aim 2
Hypothesis: We estimate that the MBI + Lifestyle (SSB) will be more effective than Lifestyle Only (HeLP) for reducing depression symptoms, insulin resistance, and stress.

## Key variables

### Teen Depression
1.	CES-D survey 
– continuous (cesdtot)
- categorical as elevated/non-elevated (cesd_elevated)

### Insulin Resistance
1. HOMA-IR (homair, homair_clean- caps at 10.3 ish for 3subj, homairlg - whats this one?)
2. Fasting insulin (insulin_mclUml, insulin_mclUml_clean - capped at 49.15 for 2 subj)
3. Fasting Glucose (glucosemgdl)
4. Other Metabolic outcomes -hba1c (hba1c, hba1c_clean - capped 1 subj 5.7), ALt/AST (AST_UL_clean - capped 2 subj 35, ALT_UL_clean capped 44 for 7 subj), Lipids(tryglyc(tryglycerides_mgdl_clean - missing 1 subj), cholesterol total(cholesteroltot_mgdl), hdl/ldl (hdlc_mgdl_clean - capped at 64.75 2 subj, missing 1, ldlc_mgdl)

### Stress
1. Perceived Stress Scale (psssumitem - total score, psselevated - 1 low 2 med 3 high)
2. Everyday discrimination scale - (edsrawscore - total, sumavg_eds - avg) - both fine
3.	Psychological Stress Experience - (pserawsum, pserawmean, pserawsumT) - which to use
4.	Cortisol, changes before to after session each week-6 sessions (cortisol_nmoll_sessionN_before, cortisol_nmoll_sessionN_after)
a.	Did we see a pattern/trajectory toward more declining cortisol in SSB (vs. HELP)?
b.	Note cortisol was measured end-of-day, after exercise portion of the intervention -during evenings (~6pm and again at 6:45/7pm), so imperfect we are aware but was consistently measured this way 

# TO ASK ABOUT: 
## Variables
1. homairlg - whats this,
2. ALT clean capped for 7 subj,
3. lipids and hdlc missingfor subj 117 6wk,
4. no advice on what to use for PSE,
5. change in cortisol need to derive the difference for each session,

## Data overall
1. we have 12 subjects with 3 time points (BL, 6wk, 6mo), 1 subj with 2 (BL, 6wk , subj 117 (has some missing data)), 5 with just bl - which to potentially include/exclude or imputation methods etc?


```{r data exploration}
#look at what we are working with
unique(HLP_MASTER_DATA$subjectidno)
length(unique(HLP_MASTER_DATA$subjectidno)) # 18 total subj

# double check homairlg is a log transform of homair
a  = HLP_MASTER_DATA %>% select(subjectidno,homair, homairlg) 
a$testhomair = log(a$homair) # base e doesnt match
a$testhomair = log10(a$homair) # base 10 matches
```
Question answers: 
1. Homairlg is log10 of homair 

```{r mixed model}
# test that mixed models work on this dataset as intended and contrasts

```

MISC

# Aim 3
Hypothesis: We anticipate that MBI + Lifestyle (SSB) will be more effective than Lifestyle Only (HELP) in improving mindfulness, eating behavior, sleep quality, physical activity, cortisol and perceived discrimination.

## Key Variables

### Mindfulness
1.	Mindfulness Attention Awareness Scale (higher scores=better) - (maastotal) 
2.	Group mood forms of state mindlessness/state anxiety (State-Trait Anxiety Inventory for Children & MAAS; higher scores for both=worse) - (maasstateavg.N.00 - goes to 12 (odd nums pre sessions, evens post)), correspond with cortisol

###	Sleep
1.	Pittsburgh Sleep Quality Index - (psqi_...) lot of these, which to use?

### Physical Activity
1.	Actigraph (typically MVPA and total time used)- (totalMVPA, AverageMVPAPerday)
2.	Physical Activity Questionnaire (physactivityrawsum)

### Stress-Related Eating
1.	Emotional Eating Scale - (EEStot, EESavg) - either fine
2.	QEWP - (qewpobe qewpsbe) (0 = no 1 = yes O = objective S = subjective) - which to use?
3.	Diet Recall (totalkcals) - also measured pct proteins, carbs, fats - use any of this


TO ASK ABOUT: 
6. lots of sleep index variables which to use,
7. QEWP - which to use, subjective or objective
8. Diet recall - also have pcts of carbs, fats proteins etc, use any of this in conjunction w total kcals? 

NOTE: Mainly looking at Aim 2 (questions 1-5)