---
title: "Salud Group Comparisons"
author: "Casey Sakamoto"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  html_document:
    df_print: paged
    toc: yes
    toc_depth: 4
    toc_float: yes
---

```{r setup, include=FALSE}
library(haven) # import spss dataset
library(knitr)
library(nlme) # mixed models
library(emmeans) # contrast statement
library(tidyverse) # data manipulation
library(Hmisc)
library(table1)
library(labelled)

knitr::opts_chunk$set(echo = FALSE)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "S:/Laura/Peds Endo/Lauren Shomaker/Salud"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Lauren Shomaker/Salud"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Lauren Shomaker/Salud"
}
knitr::opts_knit$set(root.dir = home_dir)

# import data
## make sure connected to vpn
HLP_MASTER_DATA <- read_sav("S:/Laura/Peds Endo/Lauren Shomaker/Salud/Data Analysis - Pyle 2020/HLP_MASTER DATA 7.15.20 LP Trim.sav")

# more data in updated data set, so will use same subj id list as before
HLP_MASTER_DATA2 <- read_sav("S:/Laura/Peds Endo/Lauren Shomaker/Salud/Data Analysis - Pyle 2020/HLP_MASTER DATA 6.20.22.sav")

HLP_MASTER_DATA = left_join(HLP_MASTER_DATA,HLP_MASTER_DATA2)
rm(HLP_MASTER_DATA2)
```

```{r data exploration and clean, include=FALSE}
###### light exploration #################
#look at what we are working with
unique(HLP_MASTER_DATA$subjectidno)
length(unique(HLP_MASTER_DATA$subjectidno)) # 18 total subj
##########################################

#####
# for some reason no more t-score in the new data set- will pull from the older data set
HLP_MASTER_DATA_7_14_20_LP_ = read_sav("S:/Laura/Peds Endo/Lauren Shomaker/Salud/Data Analysis - Pyle 2020/Older/HLP_MASTER DATA 7.14.20 LP .sav")

HLP_MASTER_DATA_7_14_20_LP_ = HLP_MASTER_DATA_7_14_20_LP_ %>% select(subjectidno, interval, pserawsumT)
####

###### cortisol derivation #####
hlp_cortisol = HLP_MASTER_DATA %>% mutate(cortisol_change1 = Cortisol_nmoll_session1_after - Cortisol_nmoll_session1_before,
                                          cortisol_change2 = Cortisol_nmoll_session2_after - Cortisol_nmoll_session2_before,
                                          cortisol_change3 = Cortisol_nmoll_session3_after - Cortisol_nmoll_session3_before,
                                          cortisol_change4 = Cortisol_nmoll_session4_after - Cortisol_nmoll_session4_before,
                                          cortisol_change5 = Cortisol_nmoll_session5_after - Cortisol_nmoll_session5_before,
                                          cortisol_change6 = Cortisol_nmoll_session6_after - Cortisol_nmoll_session6_before,
                                          staic_change1 = staicstatetot.2.00 - staicstatetot.1.00,
                                          staic_change2 = staicstatetot.4.00 - staicstatetot.3.00,
                                          staic_change3 = staicstatetot.6.00 - staicstatetot.5.00,
                                          staic_change4 = staicstatetot.8.00 - staicstatetot.7.00,
                                          staic_change5 = staicstatetot.10.00 - staicstatetot.9.00,
                                          staic_change6 = staicstatetot.12.00 - staicstatetot.11.00,
                                          maas_change1 = maastatetot.2.00 - maastatetot.1.00,
                                          maas_change2 = maastatetot.4.00 - maastatetot.3.00,
                                          maas_change3 = maastatetot.6.00 - maastatetot.5.00,
                                          maas_change4 = maastatetot.8.00 - maastatetot.7.00,
                                          maas_change5 = maastatetot.10.00 - maastatetot.9.00,
                                          maas_change6 = maastatetot.12.00 - maastatetot.11.00,                                          ) %>%
  select(subjectidno, interval, cortisol_change1, cortisol_change2, cortisol_change3, cortisol_change4, cortisol_change5, cortisol_change6, staic_change1, staic_change2, staic_change3, staic_change4, staic_change5, staic_change6,maas_change1,maas_change2,maas_change3,maas_change4,maas_change5,maas_change6,                                      Cortisol_nmoll_session1_after,Cortisol_nmoll_session1_before,Cortisol_nmoll_session2_after,Cortisol_nmoll_session2_before,
                                      Cortisol_nmoll_session3_after,Cortisol_nmoll_session3_before,
                                      Cortisol_nmoll_session4_after,Cortisol_nmoll_session4_before,
                                      Cortisol_nmoll_session5_after,Cortisol_nmoll_session5_before,
                                      Cortisol_nmoll_session6_after,Cortisol_nmoll_session6_before)

###### get already cleaned variables from data #################
analysis = HLP_MASTER_DATA %>% select(subjectidno, interval, groupno, sex, SV1age_calculated,z_score,
                                      cesdtot, homair, insulin_mcIUmL, glucose_mgdL,
                                      hba1c, ALT_UL, AST_UL, triglycerides_mgdL, cholesteroltot_mgdL, HDLC_mgdL, LDLC_mgdL,
                                      psssumitem, EDSrawscore, psqi_globalscores, maastotal, TotalMVPA,
                                      inMVPA, physactivityrawsum, EEStot, qewpOBE, qewpSBE, totalkcals:percentfatinkcals,
                                      Sedentary, pss_totalscoreP, maasrawscoreP, cesd_elevated
                                      ) 

analysis = full_join(analysis, HLP_MASTER_DATA_7_14_20_LP_)
analysis = full_join(analysis, hlp_cortisol)

# order and factor variables
analysis = analysis[order(analysis$subjectidno, analysis$interval),]
analysis$groupno = factor(analysis$groupno, levels = c(0,1), labels = c("HeLP","SSB"))
analysis$interval = factor(analysis$interval, levels = c(1,2,3), labels = c("Baseline", "6 Week", "6 Month"))
analysis$sex = factor(analysis$sex, levels = c(0,1), labels = c("Male", "Female"))
analysis$qewpOBE = factor(analysis$qewpOBE, levels = c(0,1), labels = c("No", "Yes"))
analysis$qewpSBE = factor(analysis$qewpSBE, levels = c(0,1), labels = c("No", "Yes"))

# make separate data frame for the 6 session variables

# clean up workspace
rm(hlp_cortisol); rm(HLP_MASTER_DATA); rm(HLP_MASTER_DATA_7_14_20_LP_); rm(a)

# analysis bl for 6sessions/demogs
analysis_BL = analysis %>% filter(interval == "Baseline")
analysis_6wk = analysis %>% filter(interval == "6 Week") %>% select(subjectidno,cortisol_change1:Cortisol_nmoll_session6_before, groupno)

# cesd elevated
analysis_cesd = analysis %>% filter(cesd_elevated == 1)
###########################################
```

```{r functions n tables, include = FALSE}
# function to grab bl-6wk and bl-6mo values for each group
diffgrp = function(outcome_str){
  # make df for each group, then long to wide
  h_df = analysis %>% filter(groupno == "HeLP") %>% select(subjectidno, interval, outcome_str )%>% 
    spread(interval, outcome_str) %>%
    mutate(WK6_BL = `6 Week` - Baseline,
           Mo6_BL = `6 Month` - Baseline)
  
  s_df = analysis %>% filter(groupno == "SSB") %>% select(subjectidno, interval, outcome_str )%>% 
    spread(interval, outcome_str) %>%
    mutate(WK6_BL = `6 Week` - Baseline,
           Mo6_BL = `6 Month` - Baseline)  
  
  # pull mean and sd out
  HeLP1 = c(outcome_str,"Week 6 - BL Mean(SD):",paste(round(mean(h_df$WK6_BL,na.rm = T),2),
                                        "(",round(sd(h_df$WK6_BL, na.rm = T),2),")"),
            paste("n=",sum(!is.na(h_df$WK6_BL))))
  HeLP2 = c(outcome_str,"Month 6 - BL Mean(SD):",paste(round(mean(h_df$Mo6_BL,na.rm = T),2),
                                        "(",round(sd(h_df$Mo6_BL, na.rm = T),2),")"),
            paste("n=",sum(!is.na(h_df$Mo6_BL)))) 
  SSB1 = c(outcome_str,"Week 6 - BL Mean(SD):",paste(round(mean(s_df$WK6_BL,na.rm = T),2),
                                        "(",round(sd(s_df$WK6_BL, na.rm = T),2),")"),
            paste("n=",sum(!is.na(s_df$WK6_BL)))) 
  SSB2 = c(outcome_str,"Month 6 - BL Mean(SD):",paste(round(mean(s_df$Mo6_BL,na.rm = T),2),
                                        "(",round(sd(s_df$Mo6_BL, na.rm = T),2),")"),
            paste("n=",sum(!is.na(s_df$Mo6_BL))))     
  a = rbind(HeLP1,SSB1)
  b = rbind(HeLP2,SSB2)
  c = rbind(a,b)
  return(kable(c))
}

# table 1 by time and group 
# demogs (age, sex)
demog_table1 = table1(~ sex + SV1age_calculated|groupno,
                       render.continuous=c(.="Mean (SD)","Missing"="NMISS"), data=analysis_BL, overall= F)
#teen depression
teendep_table1 = table1(~  cesdtot |interval*groupno,
                         render.continuous=c(.="Median [Q1, Q3]","Missing"="NMISS"), data=analysis, overall= F)
#insulin res
insres_t1 = table1(~ homair+insulin_mcIUmL+glucose_mgdL+hba1c+ALT_UL+AST_UL+triglycerides_mgdL+cholesteroltot_mgdL+HDLC_mgdL+LDLC_mgdL
                   |interval*groupno,render.continuous=c(.="Median [Q1, Q3]"), data=analysis, overall= F)

# stress
analysis_6wk = remove_var_label(analysis_6wk)

stress_t1 = table1(~   psssumitem+EDSrawscore+pserawsumT + pss_totalscoreP|interval*groupno,
                         render.continuous=c(.="Median [Q1, Q3]"), data=analysis, overall= F)
cortisol_t1 = table1(~  cortisol_change1+cortisol_change2+cortisol_change3+cortisol_change4+cortisol_change5+cortisol_change6|groupno,
                         render.continuous=c(.="Median [Q1, Q3]"), data=analysis_6wk, overall= F)

cortisol_t2 = table1(~  Cortisol_nmoll_session1_before + Cortisol_nmoll_session1_after +
                        Cortisol_nmoll_session2_before + Cortisol_nmoll_session2_after +
                        Cortisol_nmoll_session3_before + Cortisol_nmoll_session3_after +
                        Cortisol_nmoll_session4_before + Cortisol_nmoll_session4_after +
                        Cortisol_nmoll_session5_before + Cortisol_nmoll_session5_after +
                        Cortisol_nmoll_session6_before + Cortisol_nmoll_session6_after |groupno,
                         render.continuous=c(.="Median [Q1, Q3]"), data=analysis_6wk, overall= F)
# mindfulness
mind_t1 =  table1(~ maastotal + maasrawscoreP |interval*groupno,
                         render.continuous=c(.="Median [Q1, Q3]"), data=analysis, overall= F)
staic_t1 = table1(~ staic_change1+staic_change2+staic_change3+staic_change4+staic_change5+staic_change6|groupno,
                         render.continuous=c(.="Median [Q1, Q3]"), data=analysis_6wk, overall= F)
maas_t1 = table1(~ maas_change1+maas_change2+maas_change3+maas_change4+maas_change5+maas_change6|groupno,
                         render.continuous=c(.="Median [Q1, Q3]"), data=analysis_6wk, overall= F)

#sleep
sleep_t1 = table1(~ psqi_globalscores |interval*groupno,
                         render.continuous=c(.="Median [Q1, Q3]"), data=analysis, overall= F)
# phys act
phys_t1 = table1(~ Sedentary+TotalMVPA+inMVPA+physactivityrawsum |interval*groupno,
                         render.continuous=c(.="Median [Q1, Q3]"), data=analysis, overall= F)

# eating
eat_t1 = table1(~EEStot+factor(qewpOBE)+factor(qewpSBE)+totalkcals+percentproteinkcals+percentcarbsinkcal+percentfatinkcals  |interval*groupno,
                         render.continuous=c(.="Median [Q1, Q3]"), data=analysis, overall= F)

```
### added in 8/9/2022
#### Cortisol visits
```{r Cort}
cortisol_t2
```



### Demographics

```{r demg t1}
demog_table1
```

### Teen Depression (CESD Survey)

For CESD elevated, n = 12 (7 bl visits, 2 6wk visits, 3 6mo visits)
```{r teen depression}
teendep_table1

ggplot(data=analysis, aes(x=interval, y=cesdtot, fill=groupno)) +
    geom_bar(stat="summary",fun = "mean", position=position_dodge(), colour="black") + ylab("Mean CESD-tot") + xlab("") + theme_classic()

ggplot(data=analysis_cesd, aes(x=interval, y=cesdtot, fill=groupno)) +
    geom_bar(stat="summary",fun = "mean", position=position_dodge(), colour="black") + ylab("Mean CESD-tot") + xlab("") + theme_classic() + ggtitle("CESD Elevated")
```

#### Within Group Differences
```{r td grp}
diffgrp("cesdtot")
```

### Insulin Resistance

```{r ins res}
insres_t1

# ins res bar charts
ggplot(data=analysis, aes(x=interval, y=homair, fill=groupno)) +
    geom_bar(stat="summary",fun = "mean", position=position_dodge(), colour="black") + ylab("Mean HomaIR") + xlab("") + theme_classic()

ggplot(data=analysis, aes(x=interval, y=hba1c, fill=groupno)) +
    geom_bar(stat="summary",fun = "mean", position=position_dodge(), colour="black") + ylab("Mean HBA1C") + xlab("") + theme_classic()

ggplot(data=analysis, aes(x=interval, y=insulin_mcIUmL, fill=groupno)) +
    geom_bar(stat="summary",fun = "mean", position=position_dodge(), colour="black") + ylab("Mean Insulin mcl UmL") + xlab("") + theme_classic()

ggplot(data=analysis, aes(x=interval, y=glucose_mgdL, fill=groupno)) +
    geom_bar(stat="summary",fun = "mean", position=position_dodge(), colour="black") + ylab("Mean Glucose mgdL") + xlab("") + theme_classic()

ggplot(data=analysis, aes(x=interval, y=ALT_UL, fill=groupno)) +
    geom_bar(stat="summary",fun = "mean", position=position_dodge(), colour="black") + ylab("Mean ALT UL") + xlab("") + theme_classic()

ggplot(data=analysis, aes(x=interval, y=AST_UL, fill=groupno)) +
    geom_bar(stat="summary",fun = "mean", position=position_dodge(), colour="black") + ylab("Mean AST UL") + xlab("") + theme_classic()

ggplot(data=analysis, aes(x=interval, y=triglycerides_mgdL, fill=groupno)) +
    geom_bar(stat="summary",fun = "mean", position=position_dodge(), colour="black") + ylab("Mean Triglycerides mgdL") + xlab("") + theme_classic()

ggplot(data=analysis, aes(x=interval, y=cholesteroltot_mgdL, fill=groupno)) +
    geom_bar(stat="summary",fun = "mean", position=position_dodge(), colour="black") + ylab("Mean Cholesterol mgdL") + xlab("") + theme_classic()

ggplot(data=analysis, aes(x=interval, y=HDLC_mgdL, fill=groupno)) +
    geom_bar(stat="summary",fun = "mean", position=position_dodge(), colour="black") + ylab("Mean HDLC mgdL") + xlab("") + theme_classic()

ggplot(data=analysis, aes(x=interval, y=LDLC_mgdL, fill=groupno)) +
    geom_bar(stat="summary",fun = "mean", position=position_dodge(), colour="black") + ylab("Mean LDLC mgdL") + xlab("") + theme_classic()
```

#### Within Group Differences
```{r insres grp}
diffgrp("homair")
diffgrp("insulin_mcIUmL")
diffgrp("glucose_mgdL")
diffgrp("hba1c")
diffgrp("ALT_UL")
diffgrp("AST_UL")
diffgrp("triglycerides_mgdL")
diffgrp("cholesteroltot_mgdL")
diffgrp("HDLC_mgdL")
diffgrp("LDLC_mgdL")
```


### Stress and Cortisol Changes
#### Stress
```{r Stress}
stress_t1
```

#### Within Group Differences

```{r stress grp}
diffgrp("psssumitem")
diffgrp("EDSrawscore")
diffgrp("pserawsumT")
diffgrp("pss_totalscoreP")
```

#### Cortisol
calculated After - Before
```{r COrt}
cortisol_t1
```



### Mindfulness (MAAS and STAIC Changes)
```{r mindful}
mind_t1
```

#### Within Group Differences

```{r maas grp}
diffgrp("maastotal")
diffgrp("maasrawscoreP")
```

#### STAIC
calculated After - Before
```{r staic}
staic_t1
```

#### MAAS STATE
calculated After - Before
```{r maas}
maas_t1
```

### Sleep
```{r psqi}
sleep_t1
```
#### Within Group Differences

```{r psqi grp}
diffgrp("psqi_globalscores")
```

### Physical Activity
```{r phys act}
phys_t1
```
#### Within Group Differences
```{r phys grp}
diffgrp("Sedentary")
diffgrp("TotalMVPA")
diffgrp("inMVPA")
diffgrp("physactivityrawsum")
```

### Stress Related Eating
```{r stress eat}
eat_t1
```
#### Within Group Differences
```{r eat grp}
diffgrp("EEStot")
diffgrp("totalkcals")
diffgrp("percentproteinkcals")
diffgrp("percentcarbsinkcal")
diffgrp("percentfatinkcals")
```
