---
title: "Bariatric surgery and insulin sensitivity"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include=FALSE}
library(tableone)
library(dplyr)
library(knitr)
library(ggpubr)
library(nlme)
library(emmeans)

knitr::opts_chunk$set(echo = FALSE,warning = FALSE)

home_dir = ifelse(.Platform$OS.type != "unix","E:/Nadeau/IMPROVE-T2D/Data raw","")
knitr::opts_knit$set(root.dir = home_dir)

```

```{r data, include=FALSE}
setwd(home_dir)
alldata <- read.csv("./IMPROVET2D-MMTTVariables_DATA_2022-07-12_1018.csv")

# select the participants
# IT2D-2,3,5,6,8,11-16 
data <- alldata %>% filter(subject_id %in% c("IT_02","IT_03","IT_05","IT_06","IT_08","IT_11","IT_12","IT_13","IT_14","IT_15","IT_16"))

# separate variables into MMTT and screening variables
screen <- data %>% filter(redcap_event_name=="screening_arm_1") %>% 
  select(subject_id,age_current,gender,screen_bmi,screen_bmi_percentile,screen_weight,hba1c,a1c_pre)
mmtt <- data %>% filter(redcap_event_name=="study_visits_arm_1") %>% select(subject_id,study_visit_mttt,mmtt_wt,mmtt_insulin_0,mmtt_insulin_10,
                          mmtt_insulin_20,mmtt_insulin_30,mmtt_insulin_45,mmtt_insulin_60,mmtt_insulin_90,mmtt_insulin_120,
                          mmtt_insulin_150,mmtt_insulin_180,mmtt_insulin_210,mmtt_insulin_240,mmtt_bg_0,mmtt_bg_10,
                          mmtt_bg_20,mmtt_bg_30,mmtt_bg_45,mmtt_bg_60,mmtt_bg_90,mmtt_bg_120,mmtt_bg_150,mmtt_bg_180,
                          mmtt_bg_210,mmtt_bg_240,mmtt_hba1c_base,mmtt_wt)
mmtt <- mmtt %>% filter(study_visit_mttt %in% c(1,2))

# calculate glucose AUC
source("C:/Users/pylell/Documents/GitHub/General-code/iauc.r")
mmtt$auc_glucose <- iAUC_insulin(data=mmtt,thresh=0,type="linear",times=c(0,10,20,30,45,60,90,120,150,180,210,240),
                          cols=c("mmtt_bg_0","mmtt_bg_10","mmtt_bg_20","mmtt_bg_30","mmtt_bg_45","mmtt_bg_60",
                                 "mmtt_bg_90","mmtt_bg_120","mmtt_bg_150","mmtt_bg_180","mmtt_bg_210",
                                 "mmtt_bg_240"))

# calculate insulin AUC
mmtt$auc_insulin <- iAUC_insulin(data=mmtt,thresh=0,type="linear",times=c(0,10,20,30,45,60,90,120,150,180,210,240),
                          cols=c("mmtt_insulin_0","mmtt_insulin_10","mmtt_insulin_20","mmtt_insulin_30","mmtt_insulin_45","mmtt_insulin_60",
                                 "mmtt_insulin_90","mmtt_insulin_120","mmtt_insulin_150","mmtt_insulin_180","mmtt_insulin_210",
                                 "mmtt_insulin_240"))

# calculate HOMA-IR
mmtt$homa_ir <- (mmtt$mmtt_bg_0*mmtt$mmtt_insulin_0)/405

# calculate HOMA-%B
mmtt$homa_b <- (360*mmtt$mmtt_insulin_0) / (mmtt$mmtt_bg_0 - 63)

# calculate Matsuda Index
mmtt$matsuda <- 10000/(sqrt(mmtt$mmtt_bg_0*mmtt$mmtt_insulin_0*
                  (((15*mmtt$mmtt_insulin_0)+(30*mmtt$mmtt_insulin_30)+(30*mmtt$mmtt_insulin_60)+(30*mmtt$mmtt_insulin_90)+(15*mmtt$mmtt_insulin_120))/120)*
                  (((15*mmtt$mmtt_bg_0)+(30*mmtt$mmtt_bg_30)+(30*mmtt$mmtt_bg_60)+(30*mmtt$mmtt_bg_90)+(15*mmtt$mmtt_bg_120))/120)))

# calculate Disposition Index
mmtt$isi <- (mmtt$mmtt_insulin_30 - mmtt$mmtt_insulin_0) / (mmtt$mmtt_bg_30-mmtt$mmtt_bg_0)
mmtt$odi <- (1/mmtt$mmtt_insulin_0) * mmtt$isi

# calculate change in Si variables
keep <- mmtt %>% select(subject_id,study_visit_mttt,auc_glucose,auc_insulin,homa_ir,homa_b,odi,mmtt_insulin_0,mmtt_bg_0,mmtt_hba1c_base,mmtt_wt)
wide <- reshape(keep,timevar = "study_visit_mttt", direction = "wide", idvar = "subject_id")
wide$auc_glucose_delta <- wide$auc_glucose.2 - wide$auc_glucose.1
wide$auc_insulin_delta <- wide$auc_insulin.2 - wide$auc_insulin.1
wide$mmtt_insulin_0_delta <- wide$mmtt_insulin_0.2 - wide$mmtt_insulin_0.1
wide$mmtt_bg_0_delta <- wide$mmtt_bg_0.2 - wide$mmtt_bg_0.1
wide$homa_ir_delta <- wide$homa_ir.2 - wide$homa_ir.1
wide$homa_b_delta <- wide$homa_b.2 - wide$homa_b.1
wide$odi_delta <- wide$odi.2 - wide$odi.1
wide$hba1c_delta <- wide$mmtt_hba1c_base.2 - wide$mmtt_hba1c_base.1
wide$wt_delta <- wide$mmtt_wt.2-wide$mmtt_wt.1

# merge change in weight back to mmtt
wt <- wide[,c("subject_id","wt_delta")]
mmtt <- merge(mmtt,wt,by="subject_id",all.x = T, all.y = T)
```

# Formulae used

HOMA-IR: mmtt_bg_0*mmtt_insulin_0/405

HOMA-B: (360*mmtt_insulin_0) / (mmtt_bg_0 - 63)

Matsuda Index: 10000/(sqrt(mmtt_bg_0 * mmtt_insulin_0 * 
                  (((15 * mmtt_insulin_0) + (30 * mmtt_insulin_30) + (30 * mmtt_insulin_60) + (30 * mmtt_insulin_90) + (15 * mmtt_insulin_120))/120) *
                  (((15 * mmtt_bg_0) + (30 * mmtt_bg_30) + (30 * mmtt_bg_60) + (30 * mmtt_bg_90) + (15 * mmtt_bg_120))/120)))

ODI: (1/mmtt_insulin_0) *  (mmtt_insulin_30 - mmtt_insulin_0) / (mmtt_bg_30-mmtt_bg_0)

# Methods

Unadjusted changes in outcomes were tested using the Wilcoxon test. Mixed-effects models were used to test changes in outcomes while adjusting for changes in weight.

# Results

## Table of unadjusted changes

```{r echo=FALSE, include=FALSE}
t1 <- CreateTableOne(data=wide, vars = c("auc_insulin_delta","auc_glucose_delta","mmtt_insulin_0_delta","mmtt_bg_0_delta",
                                         "homa_ir_delta","homa_b_delta","odi_delta"), test = TRUE)
t1 <- print(t1, nonnorm=c("auc_insulin_delta","auc_glucose_delta","mmtt_insulin_0_delta","mmtt_bg_0_delta",
                                         "homa_ir_delta","homa_b_delta","odi_delta","hba1c_delta"))

p <- NULL
t <- wilcox.test(wide$auc_insulin_delta, mu = 0, alternative = "two.sided")$p.value
p <- c(p,t)
t <- wilcox.test(wide$auc_glucose_delta, mu = 0, alternative = "two.sided")$p.value
p <- c(p,t)
t <- wilcox.test(wide$mmtt_insulin_0_delta, mu = 0, alternative = "two.sided")$p.value
p <- c(p,t)
t <- wilcox.test(wide$mmtt_bg_0_delta, mu = 0, alternative = "two.sided")$p.value
p <- c(p,t)
t <- wilcox.test(wide$homa_ir_delta, mu = 0, alternative = "two.sided")$p.value
p <- c(p,t)
t <- wilcox.test(wide$homa_b_delta, mu = 0, alternative = "two.sided")$p.value
p <- c(p,t)
t <- wilcox.test(wide$odi_delta, mu = 0, alternative = "two.sided")$p.value
p <- c(p,t)
t <- wilcox.test(wide$hba1c_delta, mu = 0, alternative = "two.sided")$p.value
p <- c(p,t)
p <- round(p,3)
p <- c("",p)

t1 <- cbind(t1,p)

```

```{r echo=FALSE}
kable(t1)
```

## Adjusted models

How to interpret the output: the first table is the ANOVA table, which gives the p-value for the effect of time (pre vs. post) and the least-squares means (i.e., adjusted means) at each time point.

### AUC insulin

```{r echo=FALSE, include=FALSE}
mod <- lme(auc_insulin ~ study_visit_mttt + wt_delta,random=~1|subject_id,data = mmtt,na.action = na.omit)
anova <- anova.lme(mod, type="marginal")
means <- emmeans(mod,"study_visit_mttt")
```

```{r echo=FALSE}
kable(anova)
kable(means)
```

### AUC glucose

```{r echo=FALSE, include=FALSE}
mod <- lme(auc_glucose ~ study_visit_mttt + wt_delta,random=~1|subject_id,data = mmtt,na.action = na.omit)
anova <- anova.lme(mod, type="marginal")
means <- emmeans(mod,"study_visit_mttt")
```

```{r echo=FALSE}
kable(anova)
kable(means)
```

### Fasting insulin

```{r echo=FALSE, include=FALSE}
mod <- lme(mmtt_insulin_0 ~ study_visit_mttt + wt_delta,random=~1|subject_id,data = mmtt,na.action = na.omit)
anova <- anova.lme(mod, type="marginal")
means <- emmeans(mod,"study_visit_mttt")
```

```{r echo=FALSE}
kable(anova)
kable(means)
```

### Fasting glucose

```{r echo=FALSE, include=FALSE}
mod <- lme(mmtt_bg_0 ~ study_visit_mttt + wt_delta,random=~1|subject_id,data = mmtt,na.action = na.omit)
anova <- anova.lme(mod, type="marginal")
means <- emmeans(mod,"study_visit_mttt")
```

```{r echo=FALSE}
kable(anova)
kable(means)
```

### HOMA-IR

```{r echo=FALSE, include=FALSE}
mod <- lme(homa_ir ~ study_visit_mttt + wt_delta,random=~1|subject_id,data = mmtt,na.action = na.omit)
anova <- anova.lme(mod, type="marginal")
means <- emmeans(mod,"study_visit_mttt")
```

```{r echo=FALSE}
kable(anova)
kable(means)
```

### HOMA-B

```{r echo=FALSE, include=FALSE}
mod <- lme(homa_b ~ study_visit_mttt + wt_delta,random=~1|subject_id,data = mmtt,na.action = na.omit)
anova <- anova.lme(mod, type="marginal")
means <- emmeans(mod,"study_visit_mttt")
```

```{r echo=FALSE}
kable(anova)
kable(means)
```

### oDI

```{r echo=FALSE, include=FALSE}
mod <- lme(odi ~ study_visit_mttt + wt_delta,random=~1|subject_id,data = mmtt,na.action = na.omit)
anova <- anova.lme(mod, type="marginal")
means <- emmeans(mod,"study_visit_mttt")
```

```{r echo=FALSE}
kable(anova)
kable(means)
```

### HbA1c

```{r echo=FALSE, include=FALSE}
mod <- lme(mmtt_hba1c_base ~ study_visit_mttt + wt_delta,random=~1|subject_id,data = mmtt,na.action = na.omit)
anova <- anova.lme(mod, type="marginal")
means <- emmeans(mod,"study_visit_mttt")
```

```{r echo=FALSE}
kable(anova)
kable(means)
```