---
title: "IMPROVE-T2D MMTT mixed models"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include=FALSE}
library(tableone)
library(dplyr)
library(knitr)
library(ggpubr)
library(lme4)
library(nlme)
library(emmeans)
library(sjPlot)
library(marginaleffects)
library(effects)

knitr::opts_chunk$set(echo = FALSE,warning = FALSE)

options(scipen=999)

#home_dir = ifelse(.Platform$OS.type != "unix","/Volumes/Shared/Shared Projects/Laura/Peds Endo/Nadeau/IMPROVE-T2D/Data raw","")
#knitr::opts_knit$set(root.dir = home_dir)

```

```{r data, include=FALSE}
#setwd(home_dir)
alldata <- read.csv("/Volumes/PEDS/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Nadeau/IMPROVE-T2D/Data raw/IMPROVET2D-MMTTVariables_DATA_2024-04-29_1548.csv")
alldata <- alldata %>% filter(!is.na(redcap_repeat_instance))
alldata$visit <- as.factor(alldata$redcap_repeat_instance) 

# select the participants
# IT2D-2,3,5,6,8,11-16, 18-20
data <- alldata %>% filter(subject_id %in% c("IT_02","IT_03","IT_05","IT_06","IT_08","IT_11","IT_12","IT_13","IT_14","IT_15","IT_16","IT_18","IT_19","IT_20"))

# calculate weight loss
wtloss <- data %>% select(subject_id, visit, mmtt_wt)
wtloss_wide <- reshape(wtloss, direction = "wide", idvar = "subject_id", timevar = "visit")
wtloss_wide$wtloss <- wtloss_wide$mmtt_wt.1 - wtloss_wide$mmtt_wt.2
wtloss_wide$per_wtloss <- (wtloss_wide$wtloss / wtloss_wide$mmtt_wt.1) * 100
wtloss_wide <- wtloss_wide %>% select(subject_id, wtloss, per_wtloss)
data <- left_join(data, wtloss_wide, by="subject_id")

# separate variables into MMTT and screening variables
screen <- data %>% filter(redcap_event_name=="screening_arm_1") %>% 
  select(subject_id,age_current,gender,screen_bmi,screen_bmi_percentile,screen_weight,hba1c,a1c_pre)
mmtt <- data %>% filter(redcap_event_name=="study_visits_arm_1") %>% select(subject_id,study_visit_mttt,mmtt_wt,mmtt_insulin_0,mmtt_insulin_10,
                          mmtt_insulin_20,mmtt_insulin_30,mmtt_insulin_45,mmtt_insulin_60,mmtt_insulin_90,mmtt_insulin_120,
                          mmtt_insulin_150,mmtt_insulin_180,mmtt_insulin_210,mmtt_insulin_240,
                          mmtt_bg_0,mmtt_bg_10,
                          mmtt_bg_20,mmtt_bg_30,mmtt_bg_45,mmtt_bg_60,mmtt_bg_90,mmtt_bg_120,mmtt_bg_150,mmtt_bg_180,
                          mmtt_bg_210,mmtt_bg_240,mmtt_hba1c_base,
                          mmtt_cpep_0,mmtt_cpep_10,
                          mmtt_cpep_20,mmtt_cpep_30,
                          #mmtt_cpep_45,
                          mmtt_cpep_60,mmtt_cpep_90,mmtt_cpep_120,
                          #mmtt_cpep_150,
                          mmtt_cpep_180,
                          #mmtt_cpep_210,
                          mmtt_cpep_240,
                          mmtt_yy_0,mmtt_yy_10,
                          #mmtt_yy_20,
                          mmtt_yy_30,mmtt_yy_45,mmtt_yy_60,
                          #mmtt_yy_90,
                          mmtt_yy_120,
                          #mmtt_yy_150,
                          #mmtt_yy_180,
                          #mmtt_yy_210,
                          mmtt_yy_240,
                          glucagon_0,glucagon_10,
                          #glucagon_20,
                          glucagon_30,glucagon_45,glucagon_60,
                          #glucagon_90,
                          glucagon_120,
                          #glucagon_150,
                          #glucagon_180,
                          #glucagon_210,
                          glucagon_240,
                          glp1_0,glp1_10,glp1_30,glp1_45,glp1_60,glp1_120,glp1_240,
                          ffa_0,
                          #ffa_10,
                          ffa_30,
                          #ffa_45,
                          ffa_60,ffa_120,ffa_240,
                          per_wtloss)
# Keep pre-surgery and 3 month visits
mmtt <- mmtt %>% filter(study_visit_mttt %in% c(1,2))
# create separate dataframes for each assay - might make reshaping a bit easier since things were measured at different time points
insulin <- mmtt %>% select(subject_id,study_visit_mttt,mmtt_insulin_0,mmtt_insulin_10,
                          mmtt_insulin_20,mmtt_insulin_30,mmtt_insulin_45,mmtt_insulin_60,mmtt_insulin_90,mmtt_insulin_120,
                          mmtt_insulin_150,mmtt_insulin_180,mmtt_insulin_210,mmtt_insulin_240)
glucose <- mmtt %>% select(subject_id,study_visit_mttt,mmtt_bg_0,mmtt_bg_10,
                          mmtt_bg_20,mmtt_bg_30,mmtt_bg_45,mmtt_bg_60,mmtt_bg_90,mmtt_bg_120,mmtt_bg_150,mmtt_bg_180,
                          mmtt_bg_210,mmtt_bg_240)
cpep <- mmtt %>% select(subject_id,study_visit_mttt,mmtt_cpep_0,mmtt_cpep_10,
                          mmtt_cpep_20,mmtt_cpep_30,mmtt_cpep_60,mmtt_cpep_90,mmtt_cpep_120,
                          mmtt_cpep_180,mmtt_cpep_240)
pyy <- mmtt %>% select(subject_id,study_visit_mttt,mmtt_yy_0,mmtt_yy_10,
                          mmtt_yy_30,mmtt_yy_45,mmtt_yy_60,mmtt_yy_120,mmtt_yy_240)
glucagon <- mmtt %>% select(subject_id,study_visit_mttt,glucagon_0,glucagon_10,glucagon_30,glucagon_45,glucagon_60,
                          glucagon_120,glucagon_240)
glp1 <- mmtt %>% select(subject_id,study_visit_mttt,glp1_0,glp1_10,glp1_30,glp1_45,glp1_60,glp1_120,glp1_240)
ffa <- mmtt %>% select(subject_id,study_visit_mttt,ffa_0,ffa_30,ffa_60,ffa_120,ffa_240)

# create "doubly long" dataset
insulin_long <- reshape(insulin, direction = "long", varying = c("mmtt_insulin_0","mmtt_insulin_10",
                          "mmtt_insulin_20","mmtt_insulin_30","mmtt_insulin_45","mmtt_insulin_60","mmtt_insulin_90","mmtt_insulin_120",
                          "mmtt_insulin_150","mmtt_insulin_180","mmtt_insulin_210","mmtt_insulin_240"), 
                          v.names = "mmtt_insulin",times = c(0,10,20,30,45,60,90,120,150,180,210,240),
                          idvar = c("subject_id","study_visit_mttt"))
insulin_long <- insulin_long %>% arrange(subject_id,study_visit_mttt,time)
insulin_long <- left_join(insulin_long, wtloss_wide, by = "subject_id")

glucose_long <- reshape(glucose, direction = "long", varying = c("mmtt_bg_0","mmtt_bg_10",
                          "mmtt_bg_20","mmtt_bg_30","mmtt_bg_45","mmtt_bg_60","mmtt_bg_90","mmtt_bg_120","mmtt_bg_150","mmtt_bg_180",
                          "mmtt_bg_210","mmtt_bg_240"), 
                          v.names = "mmtt_bg",times = c(0,10,20,30,45,60,90,120,150,180,210,240),
                          idvar = c("subject_id","study_visit_mttt"))
glucose_long <- glucose_long %>% arrange(subject_id,study_visit_mttt,time)
glucose_long <- left_join(glucose_long, wtloss_wide, by = "subject_id")

cpep_long <- reshape(cpep, direction = "long", varying = c("mmtt_cpep_0","mmtt_cpep_10",
                          "mmtt_cpep_20","mmtt_cpep_30","mmtt_cpep_60","mmtt_cpep_90","mmtt_cpep_120",
                          "mmtt_cpep_180","mmtt_cpep_240"), 
                          v.names = "mmtt_cpep",times = c(0,10,20,30,60,90,120,180,240),
                          idvar = c("subject_id","study_visit_mttt"))
cpep_long <- cpep_long %>% arrange(subject_id,study_visit_mttt,time)
cpep_long <- left_join(cpep_long, wtloss_wide, by = "subject_id")

pyy_long <- reshape(pyy, direction = "long", varying = c("mmtt_yy_0","mmtt_yy_10",
                          "mmtt_yy_30","mmtt_yy_45","mmtt_yy_60","mmtt_yy_120","mmtt_yy_240"), 
                          v.names = "mmtt_yy",times = c(0,10,30,45,60,120,240),
                          idvar = c("subject_id","study_visit_mttt"))
pyy_long <- pyy_long %>% arrange(subject_id,study_visit_mttt,time)
pyy_long <- left_join(pyy_long, wtloss_wide, by = "subject_id")

glucagon_long <- reshape(glucagon, direction = "long", varying = c("glucagon_0","glucagon_10","glucagon_30","glucagon_45","glucagon_60",
                          "glucagon_120","glucagon_240"), 
                          v.names = "glucagon",times = c(0,10,30,45,60,120,240),
                          idvar = c("subject_id","study_visit_mttt"))
glucagon_long <- glucagon_long %>% arrange(subject_id,study_visit_mttt,time)
glucagon_long <- left_join(glucagon_long, wtloss_wide, by = "subject_id")

glp1_long <- reshape(glp1, direction = "long", varying = c("glp1_0","glp1_10","glp1_30","glp1_45","glp1_60","glp1_120","glp1_240"), 
                          v.names = "glp1",times = c(0,10,30,45,60,120,240),
                          idvar = c("subject_id","study_visit_mttt"))
glp1_long <- glp1_long %>% arrange(subject_id,study_visit_mttt,time)
glp1_long <- left_join(glp1_long, wtloss_wide, by = "subject_id")

ffa_long <- reshape(ffa, direction = "long", varying = c("ffa_0","ffa_30","ffa_60","ffa_120","ffa_240"), 
                          v.names = "ffa",times = c(0,30,60,120,240),
                          idvar = c("subject_id","study_visit_mttt"))
ffa_long <- ffa_long %>% arrange(subject_id,study_visit_mttt,time)
ffa_long <- left_join(ffa_long, wtloss_wide, by = "subject_id")
```

# Methods

Mixed effects models were used to test the difference between MMTT labs before and after bariatric surgery, while accounting for the correlation of repeated measures within a participant. Estimated marginal means were calculated for each timepoint of the MMTT within a visit, and pairwise comparisons were used to test whether each timepoint differed across visits.

# Results

```{r echo=FALSE, include=FALSE}
mixed <- function(data, outcome){
  data <- as.data.frame(data)
  data$time <- as.factor(data$time)
  form = as.formula(paste0(outcome,"~ study_visit_mttt*time"))
  mod <- lme(as.formula(form),random=~1|subject_id,data = data,na.action = na.omit, correlation = corCompSymm(form = ~1|subject_id))
  anova <- anova.lme(mod, type="marginal")
  m <- emmeans(mod,c("study_visit_mttt","time"))
  prs <-  pairs(m,adjust="tukey",by="time")
  #pred1 <- data.frame(subject_id = NA, 
  #                                       study_visit_mttt = 1,
  #                                       time = unique(data$time))
  #pred2 <- data.frame(subject_id = NA, 
  #                                       study_visit_mttt = 2,
  #                                       time = unique(data$time))
  #newdata <- rbind(pred1, pred2) 
  #pred <- predictions(mod,
  #                    newdata = newdata,
  #                    re.form = NA
  #                    )
  
  ef <- effect(mod = mod, term = "study_visit_mttt:time", 
         xlevels = list(time = unique(data$time), study_visit_mmtt = as.factor(c(1,2))))
  ef <- as.data.frame(ef)
  ef <- ef %>% filter(study_visit_mttt %in% c(1,2))
  ef$study_visit_mttt <- as.factor(ef$study_visit_mttt)
  ef <- ef %>% arrange(study_visit_mttt, time)
  p <- ggplot(ef, aes(x = time, y = fit, color = study_visit_mttt)) + 
       geom_point() + geom_line(linetype = "study_visit_mmtt")
       geom_line(aes(x = time, y = fit, color = "study_visit_mmtt", group = "study_visit_mmtt")) +
       geom_ribbon(aes(ymin=fit-se, ymax=fit+se, fill=study_visit_mttt),alpha=0.3) 
  print(p)
  #p <- plot_model(mod,data = data, type="int", terms=c("time","study_visit_mttt", ci.lvl=0.95))
  #p = ggplot(data,aes_string(x = "time", y = outcome, group = "study_visit_mttt")) +
  #  geom_line(alpha=0.1) +
  #  geom_line(data=mod[[2]],aes(x="time",y=as.numeric(emmean),group=1),
  #          size = 1,inherit.aes = F) +
  #  theme_bw() +xlab("Timepoint") + ylab(ylab) 
  #e <- effect("study_visit_mttt*time", mod)
  #plot(e)
  #p <- plot_model(mod, type = "emm", terms = c("time", "study_visit_mttt"), title = "") + 
  #  geom_line() + 
  #  labs(x = "Weeks", y = "Weight (kg)", colour = "") 
  #print(p)
  return(list(anova,m,prs))
}

mixed_adj <- function(data, outcome){
  data <- as.data.frame(data)
  data$time <- as.factor(data$time)
  form = as.formula(paste0(outcome,"~ study_visit_mttt*time + per_wtloss"))
  mod <- lme(as.formula(form),random=~1|subject_id,data = data,na.action = na.omit, correlation = corCompSymm(form = ~1|subject_id))
  anova <- anova.lme(mod, type="marginal")
  m <- emmeans(mod,c("study_visit_mttt","time"))
  prs <-  pairs(m,adjust="tukey",by="time")
  plot_dat <- data$outcome
  #p <- interaction.plot(x.factor = as.factor(data$time), trace.factor = data$study_visit_mttt, response = eval(quote(data[,outcome])))
  #p <- plot_model(mod,data = insulin_long, type="int", terms=c("time","study_visit_mttt", ci.lvl=0.95))
  # p = ggplot(data,aes_string(x = "time", y = outcome, group = "study_visit_mttt")) +
  #  geom_line(alpha=0.1) +
  #  geom_line(data=mod[[2]],aes(x="time",y=as.numeric(emmean),group=1),
  #          size = 1,inherit.aes = F) +
  #  theme_bw() +xlab("Timepoint") + ylab(ylab) 
  #e <- effect("study_visit_mttt*time", mod)
  #plot(e)
  return(list(anova,m,prs))
}

```
How to interpret the output: the first table is the ANOVA table, which gives parameter estimates for each term in the model. The term labeled "study_visit_mttt:as.factor(time)" tests whether the trajectories of the lab measure during the MMTT differed across visits. The second table provides estimated marginal means for each visit and timepoint. The third table provides p-values for the comparison of each timepoint across visits (e.g., is insulin at 10 minutes different at the visits before and after surgery). 

## Insulin

### Unadjusted model

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = insulin_long, outcome = "mmtt_insulin")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
kable(mod[[3]])
#mod[[4]]
```

### Adjusted model

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = insulin_long, outcome = "mmtt_insulin")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
kable(mod[[3]])
#mod[[4]]
```

## Glucose

### Unadjusted model

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = glucose_long, outcome = "mmtt_bg")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
kable(mod[[3]])
#mod[[4]]
```

### Adjusted model

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = glucose_long, outcome = "mmtt_bg")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
kable(mod[[3]])
#mod[[4]]
```

## C-peptide

### Unadjusted model

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = cpep_long, outcome = "mmtt_cpep")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
kable(mod[[3]])
#mod[[4]]
```

### Adjusted model

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = cpep_long, outcome = "mmtt_cpep")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
kable(mod[[3]])
#mod[[4]]
```

## PYY

### Unadjusted model

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = pyy_long, outcome = "mmtt_yy")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
kable(mod[[3]])
#mod[[4]]
```

### Adjusted model

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = pyy_long, outcome = "mmtt_yy")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
kable(mod[[3]])
#mod[[4]]
```

## GLP1

### Unadjusted model

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = glp1_long, outcome = "glp1")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
kable(mod[[3]])
#mod[[4]]
```

### Adjusted model

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = glp1_long, outcome = "glp1")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
kable(mod[[3]])
#mod[[4]]
```

## FFA

### Unadjusted model

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = ffa_long, outcome = "ffa")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
kable(mod[[3]])
#mod[[4]]
```

### Adjusted model

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = ffa_long, outcome = "ffa")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
kable(mod[[3]])
#mod[[4]]
```

## Glucagon

### Unadjusted model

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = glucagon_long, outcome = "glucagon")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
kable(mod[[3]])
#mod[[4]]
```

### Adjusted model

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = glucagon_long, outcome = "glucagon")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
kable(mod[[3]])
#mod[[4]]
```