---
title: "IMPROVE-T2D 3 month analysis"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include=FALSE}
library(arsenal)
library(dplyr)
library(knitr)
library(ggpubr)
library(nlme)
library(emmeans)

knitr::opts_chunk$set(echo = FALSE,warning = FALSE)

options(scipen=999)
```

```{r functions, include=FALSE}
mixed <- function(data, outcome){
  data <- as.data.frame(data)
  form = as.formula(paste0(outcome,"~ visit"))
  mod <- lme(as.formula(form),random=~1|subject_id,data = data,na.action = na.omit)
  anova <- anova.lme(mod, type="marginal")
  m <- emmeans(mod,"visit")
  plot_dat <- data$outcome
  #p <- interaction.plot(x.factor = as.factor(data$time), trace.factor = data$study_visit_mttt, response = eval(quote(data[,outcome])))
  #p <- plot_model(mod,data = insulin_long, type="int", terms=c("time","study_visit_mttt", ci.lvl=0.95))
  # p = ggplot(data,aes_string(x = "time", y = outcome, group = "study_visit_mttt")) +
  #  geom_line(alpha=0.1) +
  #  geom_line(data=mod[[2]],aes(x="time",y=as.numeric(emmean),group=1),
  #          size = 1,inherit.aes = F) +
  #  theme_bw() +xlab("Timepoint") + ylab(ylab) 
  #e <- effect("study_visit_mttt*time", mod)
  #plot(e)
  return(list(anova,m))
}

mixed_adj <- function(data, outcome){
  data <- as.data.frame(data)
  form = as.formula(paste0(outcome,"~ visit + wtloss"))
  mod <- lme(as.formula(form),random=~1|subject_id,data = data,na.action = na.omit)
  anova <- anova.lme(mod, type="marginal")
  m <- emmeans(mod,"visit")
  plot_dat <- data$outcome
  #p <- interaction.plot(x.factor = as.factor(data$time), trace.factor = data$study_visit_mttt, response = eval(quote(data[,outcome])))
  #p <- plot_model(mod,data = insulin_long, type="int", terms=c("time","study_visit_mttt", ci.lvl=0.95))
  # p = ggplot(data,aes_string(x = "time", y = outcome, group = "study_visit_mttt")) +
  #  geom_line(alpha=0.1) +
  #  geom_line(data=mod[[2]],aes(x="time",y=as.numeric(emmean),group=1),
  #          size = 1,inherit.aes = F) +
  #  theme_bw() +xlab("Timepoint") + ylab(ylab) 
  #e <- effect("study_visit_mttt*time", mod)
  #plot(e)
  return(list(anova,m))
}
```

```{r data, include=FALSE}
# read in screening data
source("/Volumes/Shared/Shared Projects/Laura/Peds Endo/Nadeau/IMPROVE-T2D/Data raw/IMPROVET2D-ScreeningData_R_2024-02-27_1322.r")
screen <- data
data <- NULL
# select the participants
# IT2D-2,3,5,6,8,11-16, 18-20
screen <- screen %>% filter(subject_id %in% c("IT_02","IT_03","IT_05","IT_06","IT_08","IT_11","IT_12","IT_13","IT_14","IT_15","IT_16","IT_18","IT_19","IT_20"))

# read in repeating visit data
source("/Volumes/Shared/Shared Projects/Laura/Peds Endo/Nadeau/IMPROVE-T2D/Data raw/IMPROVET2D-VisitData_R_2024-02-27_1323.r")
visit <- data
data <- NULL
# select the participants
# IT2D-2,3,5,6,8,11-16, 18-20
visit <- visit %>% filter(subject_id %in% c("IT_02","IT_03","IT_05","IT_06","IT_08","IT_11","IT_12","IT_13","IT_14","IT_15","IT_16","IT_18","IT_19","IT_20"))
# FOR THIS ANALYSIS, USE ONLY BASELINE AND MONTH 3 DATA
visit <- visit %>% filter(redcap_repeat_instance %in% c(1,2))
visit$visit <- as.factor(visit$redcap_repeat_instance)

# calculate glucose AUC
source("/Users/pylell/Documents/GitHub/General-code/iauc.r")
visit$auc_glucose <- iAUC_insulin(data=visit,thresh=0,type="linear",times=c(0,10,20,30,45,60,90,120,150,180,210,240),
                          cols=c("mmtt_bg_0","mmtt_bg_10","mmtt_bg_20","mmtt_bg_30","mmtt_bg_45","mmtt_bg_60",
                                 "mmtt_bg_90","mmtt_bg_120","mmtt_bg_150","mmtt_bg_180","mmtt_bg_210",
                                 "mmtt_bg_240"))

# calculate insulin AUC
for (i in 1:nrow(visit)){
  visit$auc_insulin[i] <- ifelse (!is.na(visit[i, "mmtt_insulin_0"]), 
                                   iAUC_insulin(data = visit[i,], thresh=0,type="linear",times=c(0,10,20,30,45,60,90,120,150,180,210,240),
                                    cols=c("mmtt_insulin_0","mmtt_insulin_10","mmtt_insulin_20","mmtt_insulin_30","mmtt_insulin_45","mmtt_insulin_60",
                                    "mmtt_insulin_90","mmtt_insulin_120","mmtt_insulin_150","mmtt_insulin_180","mmtt_insulin_210",
                                    "mmtt_insulin_240")), 
                                   NA) 
}

# calculate HOMA-IR
visit$homa_ir <- (visit$mmtt_bg_0*visit$mmtt_insulin_0)/405

# calculate HOMA-%B
visit$homa_b <- (360*visit$mmtt_insulin_0) / (visit$mmtt_bg_0 - 63)

# calculate Matsuda Index
visit$matsuda <- 10000/(sqrt(visit$mmtt_bg_0*visit$mmtt_insulin_0*
                  (((15*visit$mmtt_insulin_0)+(30*visit$mmtt_insulin_30)+(30*visit$mmtt_insulin_60)+(30*visit$mmtt_insulin_90)+(15*visit$mmtt_insulin_120))/120)*
                  (((15*visit$mmtt_bg_0)+(30*visit$mmtt_bg_30)+(30*visit$mmtt_bg_60)+(30*visit$mmtt_bg_90)+(15*visit$mmtt_bg_120))/120)))

# calculate Disposition Index
visit$isi <- (visit$mmtt_insulin_30 - visit$mmtt_insulin_0) / (visit$mmtt_bg_30-visit$mmtt_bg_0)
visit$odi <- (1/visit$mmtt_insulin_0) * visit$isi

# calculate BMI
visit$mmtt_bmi <- visit$mmtt_wt / ((visit$mmtt_ht/100)^2)

# calculate 60 - 0 min RQ
visit$rmr_rq_60_minus_0 <- visit$srmr_rq - visit$brmr_rq

# calculate weight loss
wtloss <- visit %>% select(subject_id, visit, mmtt_wt)
wtloss_wide <- reshape(wtloss, direction = "wide", idvar = "subject_id", timevar = "visit")
wtloss_wide$wtloss <- wtloss_wide$mmtt_wt.1 - wtloss_wide$mmtt_wt.2
wtloss_wide <- wtloss_wide %>% select(subject_id, wtloss)
visit <- left_join(visit, wtloss_wide, by="subject_id")
```

# Results

## Descriptive statistics

```{r create_t1, include=FALSE}
#t1 <- tableby( ~ age_current +  gender.factor + race + ethnicity.factor, data = screen)
```

```{r t1, results='asis'}
#summary(t1)
```

## Unadjusted models

### Weight

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mmtt_wt")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### BMI

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mmtt_bmi")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Waist circumference

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mmtt_waist")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Hip circumference

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mmtt_hip")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Pulse

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mmtt_hr")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### SBP

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mmtt_sys_bp")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### DBP

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mmtt_dia_bp")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### MAP

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mmtt_map")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### ALT

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mmtt_gptalt_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### AST

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mmtt_gotast_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### BUN

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mmtt_bun_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Creatinine

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mmtt_creat_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Absolute GFR 

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "gfr")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### BSA normalized GFR 

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "gfr_bsa")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Total cholesterol

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mmtt_chol_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### HDL

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mmtt_hdl_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### LDL

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mmtt_ldl_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Triglycerides

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mmtt_tri_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### HbA1c

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mmtt_hba1c_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 0 minute VO2 (ml/kg.min)

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "brmr_vo2_ml")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 0 minute VO2 (L/min)

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "brmr_vo2_l")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 0 minute VCO2 (L/min)

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "brmr_vco2")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 0 minute RQ

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "brmr_rq")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 0 minute REE

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "brmr_ree")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 0 minute Fat oxidation

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "brmr_fo")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 0 minute Carb oxidation

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "brmr_co")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 60 minute VO2 (ml/kg.min)

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "srmr_vo2_ml")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 60 minute VO2 (L/min)

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "srmr_vo2_l")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 60 minute VCO2 (L/min)

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "srmr_co2")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 60 minute RQ

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "srmr_rq")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 60 minute REE

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "srmr_ree")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 60 minute Fat oxidation

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "srmr_fo")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 60 minute Carb oxidation

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "srmr_co")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Time 60-0 RQ

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "rmr_rq_60_minus_0")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Percent body fat

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "dxa_body_fat")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Percent lean mass

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "dxa_lean_mass")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Fat mass

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "dxa_fat_kg")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Lean mass

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "dxa_lean_kg")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### HOMA-IR

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "homa_ir")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### HOMA-%B

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "homa_b")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Matsuda Index

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "matsuda")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Disposition Index

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "odi")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

## Models adjusted for weight loss

### Waist circumference

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "mmtt_waist")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Hip circumference

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "mmtt_hip")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Pulse

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "mmtt_hr")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### SBP

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "mmtt_sys_bp")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### DBP

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "mmtt_dia_bp")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### MAP

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "mmtt_map")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### ALT

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "mmtt_gptalt_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### AST

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "mmtt_gotast_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### BUN

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "mmtt_bun_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Creatinine

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "mmtt_creat_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Absolute GFR 

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "gfr")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### BSA normalized GFR 

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "gfr_bsa")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Total cholesterol

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "mmtt_chol_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### HDL

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "mmtt_hdl_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### LDL

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "mmtt_ldl_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Triglycerides

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "mmtt_tri_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### HbA1c

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "mmtt_hba1c_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 0 minute VO2 (ml/kg.min)

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "brmr_vo2_ml")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 0 minute VO2 (L/min)

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "brmr_vo2_l")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 0 minute VCO2 (L/min)

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "brmr_vco2")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 0 minute RQ

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "brmr_rq")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 0 minute REE

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "brmr_ree")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 0 minute Fat oxidation

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "brmr_fo")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 0 minute Carb oxidation

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "brmr_co")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 60 minute VO2 (ml/kg.min)

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "srmr_vo2_ml")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 60 minute VO2 (L/min)

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "srmr_vo2_l")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 60 minute VCO2 (L/min)

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "srmr_co2")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 60 minute RQ

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "srmr_rq")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 60 minute REE

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "srmr_ree")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 60 minute Fat oxidation

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "srmr_fo")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 60 minute Carb oxidation

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "srmr_co")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Time 60-0 RQ

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "rmr_rq_60_minus_0")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Percent body fat

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "dxa_body_fat")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Percent lean mass

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "dxa_lean_mass")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Fat mass

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "dxa_fat_kg")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Lean mass

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "dxa_lean_kg")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### HOMA-IR

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "homa_ir")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### HOMA-%B

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "homa_b")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Matsuda Index

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "matsuda")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Disposition Index

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "odi")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

## Readmissions
