---
title: "IMPROVE-T2D 3 month analysis"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include=FALSE}
library(arsenal)
library(dplyr)
library(knitr)
library(ggpubr)
library(nlme)
library(emmeans)

knitr::opts_chunk$set(echo = FALSE,warning = FALSE)

options(scipen=999)
```

```{r functions, include=FALSE}
mixed <- function(data, outcome){
  data <- as.data.frame(data)
  form = as.formula(paste0(outcome,"~ visit"))
  mod <- lme(as.formula(form),random=~1|subject_id,data = data,na.action = na.omit, correlation = corCompSymm(form = ~1|subject_id))
  anova <- anova.lme(mod, type="marginal")
  m <- emmeans(mod,"visit")
  plot_dat <- data$outcome
 return(list(anova,m))
}

mixed_adj <- function(data, outcome){
  data <- as.data.frame(data)
  form = as.formula(paste0(outcome,"~ visit + per_wtloss"))
  mod <- lme(as.formula(form),random=~1|subject_id,data = data,na.action = na.omit, correlation = corCompSymm(form = ~1|subject_id))
  anova <- anova.lme(mod, type="marginal")
  m <- emmeans(mod,"visit")
  plot_dat <- data$outcome
  return(list(anova,m))
}
```

```{r data, include=FALSE}
# read in screening data
source("/Volumes/PEDS/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Nadeau/IMPROVE-T2D/Data raw/IMPROVET2D-ScreeningData_R_2024-02-27_1322.r")

screen <- data
data <- NULL
# select the participants
# IT2D-2,3,5,6,8,11-16, 18-20
screen <- screen %>% filter(subject_id %in% c("IT_02","IT_03","IT_05","IT_06","IT_08","IT_11","IT_12","IT_13","IT_14","IT_15","IT_16","IT_18","IT_19","IT_20"))
# code race ethnicity
screen$race <-  
  case_when(
  screen$race___1 == 1 ~ "American Indian or Alaskan Native",
  screen$race___2 == 1 ~ "Asian",
  screen$race___3 == 1 ~ "Hawaiian or Pacific Islander",
  screen$race___4 == 1 ~ "Black or African American",
  screen$race___5 == 1 ~ "White",
  screen$race___6 == 1 ~ "Unknown",
  screen$race___7 == 1 ~ "Other",
  .default = NA
)
screen$ethnicity <-  
  case_when(
  screen$ethnicity___1 == 1 ~ "Hispanic",
  screen$ethnicity___2 == 1 ~ "Non-Hispanic",
  screen$ethnicity___3 == 1 ~ "Unknown/Not reported",
  .default = NA
)
screen$race_eth <-  
  case_when(
  screen$race == "Black or African American" & screen$ethnicity == "Non-Hispanic" ~ "Black Non-Hispanic",
  screen$race == "Other" & screen$ethnicity == "Hispanic" ~ "Hispanic",
  screen$race == "White" & screen$ethnicity == "Hispanic" ~ "Hispanic",
  screen$race == "White" & screen$ethnicity == "Non-Hispanic" ~ "White Non-Hispanic",
  .default = NA
)

# read in repeating visit data
source("/Volumes/PEDS/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Nadeau/IMPROVE-T2D/Data raw/IMPROVET2D-VisitData_R_2024-02-27_1323.r")
visit <- data
data <- NULL
# select the participants
# IT2D-2,3,5,6,8,11-16, 18-20
visit <- visit %>% filter(subject_id %in% c("IT_02","IT_03","IT_05","IT_06","IT_08","IT_11","IT_12","IT_13","IT_14","IT_15","IT_16","IT_18","IT_19","IT_20"))
# FOR THIS ANALYSIS, USE ONLY BASELINE AND MONTH 3 DATA
visit <- visit %>% filter(redcap_repeat_instance %in% c(1,2))
visit$visit <- as.factor(visit$redcap_repeat_instance)

# calculate glucose AUC
source("/Users/pylell/Documents/GitHub/General-code/iauc.r")
visit$auc_glucose <- iAUC_insulin(data=visit,thresh=0,type="linear",times=c(0,10,20,30,45,60,90,120,150,180,210,240),
                          cols=c("mmtt_bg_0","mmtt_bg_10","mmtt_bg_20","mmtt_bg_30","mmtt_bg_45","mmtt_bg_60",
                                 "mmtt_bg_90","mmtt_bg_120","mmtt_bg_150","mmtt_bg_180","mmtt_bg_210",
                                 "mmtt_bg_240"))

# calculate insulin AUC
for (i in 1:nrow(visit)){
  visit$auc_insulin[i] <- ifelse (!is.na(visit[i, "mmtt_insulin_0"]), 
                                   iAUC_insulin(data = visit[i,], thresh=0,type="linear",times=c(0,10,20,30,45,60,90,120,150,180,210,240),
                                    cols=c("mmtt_insulin_0","mmtt_insulin_10","mmtt_insulin_20","mmtt_insulin_30","mmtt_insulin_45","mmtt_insulin_60",
                                    "mmtt_insulin_90","mmtt_insulin_120","mmtt_insulin_150","mmtt_insulin_180","mmtt_insulin_210",
                                    "mmtt_insulin_240")), 
                                   NA) 
}

# calculate HOMA-IR
visit$homa_ir <- (visit$mmtt_bg_0*visit$mmtt_insulin_0)/405

# calculate HOMA-%B
visit$homa_b <- (360*visit$mmtt_insulin_0) / (visit$mmtt_bg_0 - 63)

# calculate Matsuda Index
visit$matsuda <- 10000/(sqrt(visit$mmtt_bg_0*visit$mmtt_insulin_0*
                  (((15*visit$mmtt_insulin_0)+(30*visit$mmtt_insulin_30)+(30*visit$mmtt_insulin_60)+(30*visit$mmtt_insulin_90)+(15*visit$mmtt_insulin_120))/120)*
                  (((15*visit$mmtt_bg_0)+(30*visit$mmtt_bg_30)+(30*visit$mmtt_bg_60)+(30*visit$mmtt_bg_90)+(15*visit$mmtt_bg_120))/120)))

# calculate Disposition Index
visit$isi <- (visit$mmtt_insulin_30 - visit$mmtt_insulin_0) / (visit$mmtt_bg_30-visit$mmtt_bg_0)
visit$odi <- (1/visit$mmtt_insulin_0) * visit$isi
visit$odi <- ifelse(visit$odi < 0, NA, visit$odi)

# calculate BMI
visit$mmtt_bmi <- visit$mmtt_wt / ((visit$mmtt_ht/100)^2)

# calculate 60 - 0 min RQ
visit$rmr_rq_60_minus_0 <- visit$srmr_rq - visit$brmr_rq

# calculate mean kidney volume
visit$volume_mean <- rowMeans(visit[,c("volume_right","volume_left")])

# calculate mean ACR during clamp
visit$clamp_acr_mean <- rowMeans(visit[,c("clamp_acr_baseline","clamp_acr_250")])

# calculate weight loss
wtloss <- visit %>% select(subject_id, visit, mmtt_wt) 
wtloss_wide <- reshape(wtloss, direction = "wide", idvar = "subject_id", timevar = "visit")
wtloss_wide$wtloss <- wtloss_wide$mmtt_wt.1 - wtloss_wide$mmtt_wt.2
wtloss_wide$per_wtloss <- (wtloss_wide$wtloss / wtloss_wide$mmtt_wt.1) * 100
wtloss_wide <- wtloss_wide %>% select(subject_id, wtloss, per_wtloss)
visit <- left_join(visit, wtloss_wide, by="subject_id")
```

# Background

This analysis examines changes in outcomes from pre-surgery to the 3-month post-surgery visit in the IMPROVE study.

# Methods

Repeated measures mixed-effects models with a compound-symmetric correlations structure were used to examine changes from baseline to month 3. Unadjusted models and models adjusted for weight loss were considered.

# Results

## Descriptive statistics

```{r create_t1, include=FALSE}
t1 <- tableby( ~ age_current +  gender.factor + race_eth + hba1c + sys_bp + dys_bp + as.factor(sglt2i) + screen_height + 
                 screen_weight + screen_bmi + waist_circumference, data = screen)
```

```{r t1, results='asis'}
summary(t1)
```

## Unadjusted models

### Weight

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mmtt_wt")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### BMI

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mmtt_bmi")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Waist circumference

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mmtt_waist")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Hip circumference

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mmtt_hip")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Pulse

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mmtt_hr")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### SBP

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mmtt_sys_bp")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### DBP

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mmtt_dia_bp")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### MAP

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mmtt_map")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### ALT

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mmtt_gptalt_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### AST

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mmtt_gotast_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### BUN

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mmtt_bun_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Creatinine

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mmtt_creat_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Total cholesterol

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mmtt_chol_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### HDL

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mmtt_hdl_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### LDL

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mmtt_ldl_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Triglycerides

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mmtt_tri_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### HbA1c

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mmtt_hba1c_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 0 minute VO2 (ml/kg.min)

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "brmr_vo2_ml")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 0 minute VO2 (L/min)

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "brmr_vo2_l")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 0 minute VCO2 (L/min)

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "brmr_vco2")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 0 minute RQ

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "brmr_rq")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 0 minute REE

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "brmr_ree")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 0 minute Fat oxidation

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "brmr_fo")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 0 minute Carb oxidation

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "brmr_co")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 60 minute VO2 (ml/kg.min)

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "srmr_vo2_ml")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 60 minute VO2 (L/min)

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "srmr_vo2_l")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 60 minute VCO2 (L/min)

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "srmr_co2")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 60 minute RQ

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "srmr_rq")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 60 minute REE

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "srmr_ree")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 60 minute Fat oxidation

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "srmr_fo")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 60 minute Carb oxidation

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "srmr_co")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Time 60-0 RQ

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "rmr_rq_60_minus_0")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Percent body fat

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "dxa_body_fat")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Percent lean mass

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "dxa_lean_mass")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Fat mass

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "dxa_fat_kg")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Lean mass

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "dxa_lean_kg")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### HOMA-IR

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "homa_ir")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### HOMA-%B

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "homa_b")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Matsuda Index

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "matsuda")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Disposition Index

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "odi")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

## Models adjusted for percent weight loss

### Waist circumference

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "mmtt_waist")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Hip circumference

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "mmtt_hip")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Pulse

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "mmtt_hr")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### SBP

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "mmtt_sys_bp")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### DBP

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "mmtt_dia_bp")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### MAP

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "mmtt_map")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### ALT

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "mmtt_gptalt_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### AST

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "mmtt_gotast_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### BUN

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "mmtt_bun_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Creatinine

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "mmtt_creat_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Total cholesterol

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "mmtt_chol_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### HDL

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "mmtt_hdl_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### LDL

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "mmtt_ldl_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Triglycerides

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "mmtt_tri_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### HbA1c

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "mmtt_hba1c_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 0 minute VO2 (ml/kg.min)

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "brmr_vo2_ml")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 0 minute VO2 (L/min)

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "brmr_vo2_l")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 0 minute VCO2 (L/min)

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "brmr_vco2")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 0 minute RQ

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "brmr_rq")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 0 minute REE

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "brmr_ree")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 0 minute Fat oxidation

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "brmr_fo")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 0 minute Carb oxidation

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "brmr_co")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 60 minute VO2 (ml/kg.min)

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "srmr_vo2_ml")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 60 minute VO2 (L/min)

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "srmr_vo2_l")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 60 minute VCO2 (L/min)

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "srmr_co2")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 60 minute RQ

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "srmr_rq")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 60 minute REE

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "srmr_ree")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 60 minute Fat oxidation

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "srmr_fo")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 60 minute Carb oxidation

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "srmr_co")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Time 60-0 RQ

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "rmr_rq_60_minus_0")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Percent body fat

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "dxa_body_fat")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Percent lean mass

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "dxa_lean_mass")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Fat mass

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "dxa_fat_kg")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Lean mass

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "dxa_lean_kg")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### HOMA-IR

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "homa_ir")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### HOMA-%B

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "homa_b")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Matsuda Index

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "matsuda")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Disposition Index

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "odi")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### Height

### Serum sodium

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "serum_sodium")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
```

### 
