---
title: "Bromocriptine arterial health manuscript - response to reviewers"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 2
    toc_float: yes
---

```{r setup, include=FALSE}
library(readxl)
library(nlme)

knitr::opts_chunk$set(echo = FALSE,warning = FALSE)

home_dir = ifelse(.Platform$OS.type != "unix","E:\\Nadeau\\Nadeau Bromocriptine T1D DSMB\\Dissemination",
                  "/Volumes/som/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects")
knitr::opts_knit$set(root.dir = home_dir)


```

```{r, include=FALSE}
setwd(home_dir)
data <- read.csv("./BromocriptineData.csv")

# make a wide dataset
data_wide <- reshape(data,idvar="PTID",timevar = "TX",direction = "wide")

# calculate deltas (BCQR - PLA)
deltas <- data_wide[,4:28]- data_wide[,31:55]
c <- colnames(deltas)
c <- gsub("BCQR","delta",c)
colnames(deltas) <- c
data_wide <- cbind(data_wide, deltas)
```

# Results

## Models adjusted for BP

### Pressure hemodynamics

Models for pressure hemodynamic variables were adjusted for the baseline (i.e., placebo) value of the variable. These models can be interpreted as follows: the estimate of the intercept is the estimated change (BCQR - placebo) in the outcome, while adjusting for baseline. The p-value for the intercept reflects a test of whether the change in the outcome is different from zero.

```{r echo=FALSE, comment=""}
# for BP variables, just adjust for baseline (i.e., Placebo)
mod <- lm(SBP.delta ~ SBP.Placebo, data=data_wide)
print(summary(mod))
```

```{r echo=FALSE, comment=""}
mod <- lm(DBP.delta ~ DBP.Placebo, data=data_wide)
print(summary(mod))
```

```{r echo=FALSE, comment=""}
mod <- lm(PP.delta ~ PP.Placebo, data=data_wide)
print(summary(mod))
```

```{r echo=FALSE, comment=""}
mod <- lm(MAP.delta ~ MAP.Placebo, data=data_wide)
print(summary(mod))
```

### Other variables

For PWC, distensibility, and RAC, models were adjusted for the change in MAP. These models can be interpreted as follows: the estimate of the intercept is the estimated change (BCQR - placebo) in the outcome, while adjusting for the change in MAP. The p-value for the intercept reflects a test of whether the change in the outcome is different from zero.


```{r echo=FALSE, comment=""}
vars <- c[5:14]

for (v in vars) {
  form = as.formula(paste0(v,"~MAP.delta"))
  mod <- lm(form,data = data_wide)
  print(noquote(c("Adjusted model for: ", noquote(v))))
  print(summary(mod))
}
```

## Models with treatment interaction with BMI group

Mixed models were used to test the interaction of treatment with BMI group. In the ANOVA tables that follow, the p-value on the row labeled "TX:as.factor(BMI.Category)" is the test of whether this interaction is significant.

```{r echo=FALSE, comment=""}
vars_long <- colnames(data[,5:18])

#mod <- lme(SBP ~ TX*BMI.Category,random=~1|PTID,data = data,na.action = na.omit)
#mod_anova <- anova.lme(mod, type="marginal")
#print(mod_anova)

for (v in vars_long) {
  form = as.formula(paste0(v,"~TX*as.factor(BMI.Category)"))
  mod <- lme(form,random=~1|PTID,data = data,na.action = na.omit)
  mod_anova <- anova.lme(mod, type="marginal")
  print(noquote(c("Adjusted model for: ", noquote(v))))
  print(mod_anova)
}
```
