---
title: "CF Registry Data Clean"
author: "Casey Sakamoto"
date: "2/4/2022"
output: pdf_document
---

```{r setup, include=FALSE}
library(arsenal)
library(readxl)
library(readr)
library(Hmisc)
library(tidyverse)
library(performance)
library(knitr)
library(lubridate)
library(stringr)
knitr::opts_chunk$set(echo = FALSE)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "S:/Laura/Peds Endo/Christine Chan/CF registry data"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data"
}
knitr::opts_knit$set(root.dir = home_dir)
```



```{r data}
CFF19_DemogCFDiag_Del1 <- read_csv("S:/Laura/Peds Endo/Christine Chan/CF registry data/CFF19_DemogCFDiag_Del1.csv")
CFF19_Annualized_Del1 <- read_csv("S:/Laura/Peds Endo/Christine Chan/CF registry data/CFF19_Annualized_Del1.csv")
CFF19_CareEpisodes_Del1 <- read_csv("S:/Laura/Peds Endo/Christine Chan/CF registry data/CFF19_CareEpisodes_Del1.csv")
CFF19_encountersMerged_Del1 <- read_csv("S:/Laura/Peds Endo/Christine Chan/CF registry data/CFF19_encountersMerged_Del1.csv")


describe(CFF19_Annualized_Del1$cfrd_status_annual) # mostly 1's, 68587 missing (~16%)
describe(CFF19_Annualized_Del1$ReviewYear) # 2003 - 2018

describe(CFF19_encountersMerged_Del1$cfrd_status) # mostly 3's, 1784520 missing (~80%)

```

Aim 1:

- estimate prevalence of CFRD at each year as the proportion of patients with CFRD: cfrd_status in encounters_merged and annualized (it seems like this one)

- estimate incidence of CFRD per year (number of incident CFRD diagnoses per 1000 person-yrs), excluding prevalent cases

t-tests/Chisq tests for associations of indiv variables (sex, race/ethm, number of prednisone prescriptions) with CFRD within a given year

mixed effects logreg model with interactino effect for year and cfrd status to test whether these univariate relationships differ between years -- significance of interaction effect from the model assessed with anova will be used to detect a difference in effect between years.

mixed effects logreg model with variables selected by LASSO (least absolute shrinkage selection operator) or elastic -- which variables are associated with CFRD -- OR and 95%CIs to interpret results


CFRD_status variable per the codebook
1: ????
2	Impaired Glucose Tolerance (FBG < 126, 2-h PG 140-199)
3	CFRD with or without fasting hyperglycemia
4	Type 1 Diabetes
5	Type 2 Diabetes


Questions:

data cleaning/merging depending on which variables we need? do we need all forms if demog and annual/enc has the necessary info?

two cfrd status variables: one in annualized and one in encounters: difference in data sets? which to use?
  within both it looks like there are 5 levels : 
  
1	Normal Glucose Metabolism (includes normal, random, fasting, or OGTT)
2	Impaired Glucose Tolerance (FBG < 126, 2-h PG 140-199)
3	CFRD with or without fasting hyperglycemia ( 2-h PG >= 200)
4	Type 1 Diabetes
5	Type 2 Diabetes

and a lot of NA's; does normal glucose metabolism mean no cfrd? or does NA mean no cfrd?


DEMOG:
what is number of prednisone prescriptions; cant seem to find a variable, how to tell? couple variables have prednisone in it: corticosteroids1 & 2 (cant tell the diff for this either)

```{r AIM 1}

##### VARIABLES LIST ##########
# CFRD Status/ year : annualized data: cfrd_status_annual, reviewyear
#                     encounter data: cfrd_status, review year

# sex, race, ethn: demog 
# prednisone prescriptions 
describe(CFF19_encountersMerged_Del1$corticosteroids1) # i think this means oral (eg prednisone)
describe(CFF19_encountersMerged_Del1$corticosteroids2) # i think this means inhaled
                                                       # maybe there should be a 3 for inhaled in combination with bronchodilator (per the codesheet)


describe(CFF19_Annualized_Del1$A_corticosteroids1) # similar as above
describe(CFF19_Annualized_Del1$A_corticosteroids2)
describe(CFF19_Annualized_Del1$A_corticosteroids3)

CFRD_annualtest = CFF19_Annualized_Del1 %>% select(eDWID, ReviewYear ,cfrd_status_annual)
CFRD_encountertest = CFF19_encountersMerged_Del1 %>% select(eDWID, reviewyear, cfrd_status)
###############################
##### PREVALENCE/ INCIDENCE ##########

######################################

```



