---
title: "Pima"
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,dpi = 600,cache = TRUE)
knitr::opts_knit$set(root.dir = "/home/tim/Dropbox/Work/Petter Bjornstad/Pima/Youth vs. adult onset kidney disease")
library(mixOmics)
library(readxl)
library(limma)
library(knitr)
library(tidyverse)
set.seed(1017)
```

# Clinical variables from Petter

## Morphometric parameters (tissue integrity):

### Most important
- GBM
- VVMES
- VVINT
- GV
- P_SCLER

### Less important
- SV
- VPC
- VVPC
- NVPC
- P_FPW
- P_FEN
- VVAT_NS
- VVATTT_NS
- VVPT_NS
- VVDT_NS
- VVTT_NS
- GV_GLOMS
- P_SCLER

## Functional parameters (kidney function):
- GFR
- RPF
- SBP
- DBP

```{r include=FALSE}
# Clinical data
clinical_data = read.csv("./Data_raw/pima_yo.csv",stringsAsFactors=F,na.strings = "")
clinical_data$AZ = NULL
# Import and format UM serum and urine
um_serum = read.csv("./Data_raw/UM_SS_Serum_yop.csv")
um_serum = um_serum[,c(1,4:ncol(um_serum))]
um_urine = read.csv("./Data_raw/um_ss_urine_yop.csv")
um_urine = um_urine[,c(1,4:ncol(um_urine))]
colnames(um_serum)[2:ncol(um_serum)] = 
  paste0(colnames(um_serum)[2:ncol(um_serum)],"_serum")
colnames(um_urine)[2:ncol(um_urine)] = 
  paste0(colnames(um_urine)[2:ncol(um_urine)],"_urine")
# Merge
df = left_join(clinical_data,um_serum,by="NIH")
df = left_join(df,um_urine,by="NIH")
# Format JOS data
jos_data = read.csv("./Data_raw/jos_ss_yop.csv")
jos_data$DATE = lubridate::dmy(jos_data$DATE)
jos_data = jos_data %>% arrange(NIH,DATE)
jos_data = jos_data[duplicated(jos_data$NIH) == F,]
jos_data$DATE = NULL
# Merge
df = left_join(df,jos_data,by="NIH")
# Delete visit and media columns
df$VISIT = NULL
df$MEDIA = NULL
```

# Onset at Age 21

## PLS-DA

### PLS-DA Performance

```{r}
# Numeric df
X = as.data.frame(lapply(df[,5:ncol(df)],function(c){as.numeric(gsub(",","",c))}))
# Remove columns with SD = 0
nzv = nearZeroVar(X)
X = X[,-c(nzv$Position)]
# Group as outcome
Y <- as.factor(df$youth21)
# PLS-DA function
plsda_true <- plsda(X,Y,ncomp = 20)
# Check error and look at optimal number of components
perf_plsda_true <- perf(plsda_true,progressBar=FALSE, auc=TRUE,validation = "loo")
plot(perf_plsda_true,col = color.mixo(5:7), sd = TRUE, legend.position = "horizontal")
```

Based on LOO CV, 12 PLS-DA components results in the best (lowest) overall classification error rate.

```{r}
# Print AUCs
aucs = as.data.frame(perf_plsda_true$auc[1:12])
aucs = aucs[1,]
kable(t(aucs))
```

```{r cache=TRUE,warning=FALSE}
# Bootstrap AUC CIs
plsda_fun = function(data,i){
  df = data[i, ]
  pls = plsda(df[,-1],df[,"Y"],ncomp = 2)
  perf = perf(pls,progressBar=FALSE, auc=TRUE,validation = "loo")
  auc = cbind(perf$auc$comp1["AUC.mean"],perf$auc$comp2["AUC.mean"])
  colnames(auc) = c("comp1","comp2")
  auc
}
df_boot = cbind(Y,X)
# First column in b$t is component 1, second is component 2
b = boot(df_boot,plsda_fun,1000,parallel = "multicore",ncpus = 8)
# CIs
bci1 = boot.ci(b,index = 1,type = "perc")
bci2 = boot.ci(b,index = 2,type = "perc")
```

```{r}
# Compare to Cameron's elasticnet results
enet_res = read.csv("./Reports/pima_jos_enet.csv")
```