---
title: "Early Glargine in DKA"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyr)
library(magrittr)
library(dplyr)
library(lubridate)
library(zoo) 
library(M3)
library(data.table)
#library(Table1)
library(devtools)
library(Hmisc)
library(knitr)
library(ggplot2)
library(gridExtra)
library(easyGgplot2)

# read in data
final <- read.csv("/Volumes/PEDS/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Ohman/Ohman early glargine retrospective/Data/final data from sas.csv")

# use Claire's functions
source(file='C:\\Users\\pylell\\Documents\\GitHub\\General-code\\temp_table1_nofishers.r')


####################################################################################################

####################################################################################################
#Descriptive Tables #
####################################################################################################

# Set factors and levels
final$rebound_hyper <- as.factor(final$rebound_hyper)
final$rebound_hyper_250 <- as.factor(final$rebound_hyper_250)
final$rebound_hyper_300 <- as.factor(final$rebound_hyper_300)
final$group <- as.factor(final$group)
final$cerebral_edema <- as.factor(final$cerebral_edema)
levels(final$rebound_hyper)=c('No','Yes')
levels(final$rebound_hyper_250)=c('No','Yes')
levels(final$rebound_hyper_300)=c('No','Yes')
levels(final$Sex)=c('Female','Male')
levels(final$Ethnicity)=c('Black or African American','Hispanic','Multiracial','Other','White')
levels(final$Insurance_Type)=c('Unknown','Private/Military','Public','Self Pay')
final$hypo_12hr <- as.factor(final$hypo_12hr)
final$hypo_12hr54 <- as.factor(final$hypo_12hr54)
final$hypo_iv <- as.factor(final$hypo_iv)
final$hypo_iv54 <- as.factor(final$hypo_iv54)
final$ketosis_either <- as.factor(final$ketosis_either)
levels(final$hypo_12hr)=c('No','Yes')
levels(final$hypo_12hr54)=c('No','Yes')
levels(final$hypo_iv)=c('No','Yes')
levels(final$hypo_iv54)=c('No','Yes')
levels(final$ketosis_either)=c('No','Yes')
levels(final$cerebral_edema)=c('No','Yes')
final$hypokal_12hr <- as.factor(final$hypokal_12hr)
levels(final$hypokal_12hr) =c('No','Yes')
final$hypokal_iv <- as.factor(final$hypokal_iv)
levels(final$hypokal_iv)=c('No','Yes')

# set non new-onsets
levels(final$New_Onset_Diabetic)=c("Y","N")
final$New_Onset_Diabetic[final$New_Onset_Diabetic != "Y"] <- "N"

# convert to numeric
final$HbA1c_Closest_to_Admisson <- as.numeric(final$HbA1c_Closest_to_Admisson)

# calculate glargine dose per kg
# first recode 0 weights to missing
final[!is.na(final$Weight) & final$Weight==0,]$Weight <- NA
final$Glargine_per_kg <- final$Glargine/final$Weight


# Set Labels
label(final$rebound_hyper)='Rebound hyperglycemia 180 mg/dl'
label(final$rebound_hyper_250)='Rebound hyperglycemia 250 mg/dl'
label(final$rebound_hyper_300)='Rebound hyperglycemia 300 mg/dl'
label(final$group)="Group"
label(final$Sex)='Sex'
label(final$Ethnicity)='Ethnicity'
label(final$Insurance_Type)='Insurance'
label(final$Age)='Age'
label(final$Height)='Height'
label(final$Weight)='Weight'
label(final$HbA1c_Closest_to_Admisson)='HbA1c closest to admission'
label(final$Median_Income)='Median income'
label(final$median_glucose)='Median glucose'
label(final$hypo_12hr)='Hypoglycemia (<70 mg/dL) within 12 hours after d/c IV insulin'
label(final$hypo_iv)='Hypoglycemia (<70 mg/dl) during IV insulin'
label(final$hypo_12hr54)='Hypoglycemia (<54 mg/dL) within 12 hours after d/c IV insulin'
label(final$hypo_iv54)='Hypoglycemia (<54 mg/dl) during IV insulin'
label(final$ketosis_either)='Ketosis w/in 12 hrs after d/c insulin, by either BHB or urine ketones'
label(final$cerebral_edema)='Cerebral Edema'
label(final$hypokal_12hr)='Hypokalemia within 12 hours after d/c IV insulin'
label(final$hypokal_iv)='Hypokalemia during IV insulin'
label(final$dur_iv_insulin)='Length of IV insulin administration'
label(final$slopegluc)='Rate of glucose decrease (mg/dL per minute)'
label(final$Glargine_per_kg)="Glargine dose/kg"
label(final$New_Onset_Diabetic)="New onset?"

cpg <- final
cpg <- cpg[cpg$October_1_2014_Flag %in% c("After","Before"),]
cpg$October_1_2014_Flag <- droplevels(cpg$October_1_2014_Flag)
Ncpg <- nrow(cpg)

Norig <- length(final$group)
final <- final[final$group %in% c("Early","Late"),]
final$group <- droplevels(final$group)
N <- sum(!is.na(final$group))
Nexcl <- Norig - N

# calculate minutes from initiation of IV insulin to glargine
final$min_IV_glargine <- final$min_admit_glargine - final$min_admit_insstart

tab1 <- final_table(final,c('Sex','Ethnicity',"Insurance_Type","New_Onset_Diabetic",
    "Age","Height","Weight","HbA1c_Closest_to_Admisson","Median_Income","Glargine_per_kg"),
    group=final$group,margin=2,single=0,ron=2)

tab1a <- final_table(cpg,c('Sex','Ethnicity',"Insurance_Type","New_Onset_Diabetic",
    "Age","Height","Weight","HbA1c_Closest_to_Admisson","Median_Income","Glargine_per_kg"),
    group=cpg$October_1_2014_Flag,margin=2,single=0,ron=2)

# convert duration of IV insulin to hours
final$dur_iv_insulin_hours <- final$dur_iv_insulin/60


####################################################################################################

####################################################################################################
#Primary Outcome #
####################################################################################################

# early vs. late glargine variable is October.1.2014.Flag
#tbl <- table(rebound_hyper$rebound_hyper,rebound_hyper$October.1.2014.Flag)
#rslt <- chisq.test(tbl)
tab_primary <- final_table(final,c('rebound_hyper','rebound_hyper_250','rebound_hyper_300',
                                   'median_glucose','hypo_12hr','hypo_iv',
                                   'hypo_12hr54','hypo_iv54',
                                   'ketosis_either','cerebral_edema',
                                   'hypokal_12hr','hypokal_iv','dur_iv_insulin','slopegluc'),
                           group=final$group,margin=2,single=0,ron=2)

tab_primarya <- final_table(cpg,c('rebound_hyper','rebound_hyper_250','rebound_hyper_300',
                                   'median_glucose','hypo_12hr','hypo_iv',
                                   'hypo_12hr54','hypo_iv54',
                                   'ketosis_either','cerebral_edema',
                                   'hypokal_12hr','hypokal_iv','dur_iv_insulin','slopegluc'),
    group=cpg$October_1_2014_Flag,margin=2,single=0,ron=2)

# test interaction
mod1 <- glm(rebound_hyper ~ group*New_Onset_Diabetic, data=final,family="binomial")
mod2 <- glm(rebound_hyper_250 ~ group*New_Onset_Diabetic, data=final,family="binomial")
mod3 <- glm(rebound_hyper_300 ~ group*New_Onset_Diabetic, data=final,family="binomial")
mod4 <- glm(hypo_12hr ~ group*New_Onset_Diabetic, data=final,family="binomial")
mod5 <- glm(hypo_iv ~ group*New_Onset_Diabetic, data=final,family="binomial")
mod6 <- glm(hypo_12hr54 ~ group*New_Onset_Diabetic, data=final,family="binomial")
mod7 <- glm(hypo_iv54 ~ group*New_Onset_Diabetic, data=final,family="binomial")
mod8 <- glm(ketosis_either ~ group*New_Onset_Diabetic, data=final,family="binomial")
mod9 <- glm(cerebral_edema ~ group*New_Onset_Diabetic, data=final,family="binomial")
mod10 <- glm(hypokal_12hr ~ group*New_Onset_Diabetic, data=final,family="binomial")
mod11 <- glm(hypokal_iv ~ group*New_Onset_Diabetic, data=final,family="binomial")
mod12 <- lm(median_glucose ~ group*New_Onset_Diabetic, data=final)
mod13 <- lm(dur_iv_insulin ~ group*New_Onset_Diabetic, data=final)
mod14 <- lm(slopegluc ~ group*New_Onset_Diabetic, data=final)

mod15 <- glm(rebound_hyper ~ October_1_2014_Flag*New_Onset_Diabetic, data=final,family="binomial")
mod16 <- glm(rebound_hyper_250 ~ October_1_2014_Flag*New_Onset_Diabetic, data=final,family="binomial")
mod17 <- glm(rebound_hyper_300 ~ October_1_2014_Flag*New_Onset_Diabetic, data=final,family="binomial")
mod18 <- glm(hypo_12hr ~ October_1_2014_Flag*New_Onset_Diabetic, data=final,family="binomial")
mod19 <- glm(hypo_iv ~ October_1_2014_Flag*New_Onset_Diabetic, data=final,family="binomial")
mod20 <- glm(hypo_12hr54 ~ October_1_2014_Flag*New_Onset_Diabetic, data=final,family="binomial")
mod21 <- glm(hypo_iv54 ~ October_1_2014_Flag*New_Onset_Diabetic, data=final,family="binomial")
mod22 <- glm(ketosis_either ~ October_1_2014_Flag*New_Onset_Diabetic, data=final,family="binomial")
mod23 <- glm(cerebral_edema ~ October_1_2014_Flag*New_Onset_Diabetic, data=final,family="binomial")
mod24 <- glm(hypokal_12hr ~ October_1_2014_Flag*New_Onset_Diabetic, data=final,family="binomial")
mod25 <- glm(hypokal_iv ~ October_1_2014_Flag*New_Onset_Diabetic, data=final,family="binomial")
mod26 <- lm(median_glucose ~ October_1_2014_Flag*New_Onset_Diabetic, data=final)
mod27 <- lm(dur_iv_insulin ~ October_1_2014_Flag*New_Onset_Diabetic, data=final)
mod28 <- lm(slopegluc ~ October_1_2014_Flag*New_Onset_Diabetic, data=final)


# box plot IV insulin duration in early vs. late
boxplot(dur_iv_insulin_hours~group,data=final,xlab="",ylab="Duration of IV insulin (hours)")

# bar graph of % rebound hyper (>180) in early vs. late
table <- prop.table(table(final$rebound_hyper,final$group),margin=2)*100
table <- table[2,]
barplot(table,ylab = "% Rebound Hyperglycemia (>180 mg/dL)")

# bar graph of % rebound hyper (>250) in early vs. late
table_250 <- prop.table(table(final$rebound_hyper_250,final$group),margin=2)*100
table_250 <- table_250[2,]
barplot(table_250,ylab = "% Rebound Hyperglycemia (>250 mg/dL)")

# bar graph of % rebound hyper (>300) in early vs. late
table_300 <- prop.table(table(final$rebound_hyper_300,final$group),margin=2)*100
table_300 <- table_300[2,]
barplot(table_300,ylab = "% Rebound Hyperglycemia (>300 mg/dL)")

# group all the hyper data together
labels <- c(">180 mg/dL",">250 mg/dL",">300 mg/dL")
table_hyper <- as.data.frame(rbind(table, table_250, table_300))
table_hyper <- cbind(table_hyper,labels)
# make a long dataset
table_hyper_long <- gather(table_hyper,Group,value,Early:Late)

# bar graph of % severe hypo in early vs. late
table_sh <- prop.table(table(final$hypo_12hr54,final$group),margin=2)*100
table_sh <- table_sh[2,]
barplot(table_sh,ylab = "% Severe Hypoglycemia (<54 mg/dL)")

# bar graph of hypo <70 in early vs. late during IV insulin
table_hypo70_iv <- prop.table(table(final$hypo_iv,final$group),margin=2)*100
table_hypo70_iv <- table_hypo70_iv[2,]
barplot(table_hypo70_iv,ylab = "% Hypoglycemia (<70 mg/dL) during IV insulin")

# bar graph of hypo <70 in early vs. late within 12 hours of IV insulin
table_hypo70_12 <- prop.table(table(final$hypo_12hr,final$group),margin=2)*100
table_hypo70_12 <- table_hypo70_12[2,]
barplot(table_hypo70_12,ylab = "% Hypoglycemia (<70 mg/dL) within 12 h of IV insulin")

# group all the hypo<70 data together
labels <- c("During IV insulin","Within 12 h")
table_hypo70 <- as.data.frame(rbind(table_hypo70_iv, table_hypo70_12))
table_hypo70 <- cbind(table_hypo70,labels)
# make a long dataset
table_hypo70_long <- gather(table_hypo70,Group,value,Early:Late)

# bar graph of hypo <54 in early vs. late during IV insulin
table_hypo54_iv <- prop.table(table(final$hypo_iv54,final$group),margin=2)*100
table_hypo54_iv <- table_hypo54_iv[2,]
barplot(table_hypo54_iv,ylab = "% Hypoglycemia (<54 mg/dL) during IV insulin")

# bar graph of hypo <54 in early vs. late within 12 hours of IV insulin
table_hypo54_12 <- prop.table(table(final$hypo_12hr54,final$group),margin=2)*100
table_hypo54_12 <- table_hypo54_12[2,]
barplot(table_hypo54_12,ylab = "% Hypoglycemia (<54 mg/dL) within 12 h of IV insulin")

# group all the hypo<54 data together
labels <- c("During IV insulin","Within 12 h")
table_hypo54 <- as.data.frame(rbind(table_hypo54_iv, table_hypo54_12))
table_hypo54 <- cbind(table_hypo54,labels)
# make a long dataset
table_hypo54_long <- gather(table_hypo54,Group,value,Early:Late)


```

# Background

The purpose of this study was to determine if early glargine administration in children with DKA prevents rebound hyperglycemia and recurrent ketogenesis without increasing the rate of hypoglycemia.   

# Methods

Early glargine administration was defined as either (1) glargine administered before the start of IV insulin administration or (2) glargine administered after the start of IV insulin administration with >4 hours of overlap with IV insulin.  Late glargine administration was defined as glargine administered with <2 hour overlap with IV insulin.  There were `r Nexcl` participants with between 2-4 hours of overlap with IV insulin who were excluded from this analysis.
 
The primary outcome of this study is rebound hyperglycemia, defined as serum glucose >180 mg/dL within 12 hours of discontinuation of IV insulin. Secondary analyses examined the outcome of rebound hyperglycemia using cutoffs of 250 and 300 mg/dl. 

The following data cleaning and manipulation steps were performed in order to determine whether each participant met the definition of rebound hyperglycemia:

1)  Study IDs on the main data sheet ending in "_1" or something similar, to indicate distinct encounters, were stripped of the suffix starting with the underscore, so they would match the IDs on the glucose data sheet.
2)  Missing glucose values or those with comments such as ("Already Documented","See Comment","Task Duplication", and "TNP") deleted from the glucose results.
3)  Encounters that were missing insulin administration dates and/or times were deleted.
4)  For each encounter, the last date and time of IV insulin administration was determined.
5)  The main data sheet was merged with glucose results by a combination of study ID and age, since a participant could have more than one encounter.
6)  If a participant had a glucose value >180 mg/dL (or 250 or 300 mg/dl) within 12 hours of IV insulin discontinuation, s/he was classified as having rebound hyperglycemia.
7)  Subjects 5336, 5770, 6073, 6074, 6080, 6121 were excluded from analysis because they had a diagnosis of type 2 diabetes.

Additional data cleaning and manipulation steps included:

1)  Military and private insurance were combined into one category.
2)  American Indian and Asian ethnicities were combined with Other ethnicity.

Secondary outcomes were defined as follows:

1) If a participant had a glucose value <70 within 12 hours of IV insulin discontinuation, s/he was classified as having hypoglycemia within 12 hours of insulin discontinuation.  Secondary analyses used a cutoff of 54 mg/dl.
2) If a participant had a glucose value <70 during IV insulin administration, s/he was classified as having hypoglycemia during IV insulin.  Secondary analyses used a cutoff of 54 mg/dl.
3) If a participant had a beta-hydroxybutyrate value >= 1.5 mmol/L or moderate or large urine ketones within 12 hours of IV insulin discontinuation, s/he was classifed as having ketosis within 12 hours of insulin discontinuation.
4) If a participant had a potassium level <3.5 mmol/L either during IV insulin administration or within 12 hours of IV insulin discontinuation, s/he was classified as having hypokalemia during that time period.

Descriptive statistics reported include mean and standard deviations for continuous variables, and frequencies and percentages for categorical variables.  To compare characteristics in the two groups, t-tests were used for continuous variables and the chi-square test was used for categorical variables.

These analyses were repeated using the initiation date of the clinical practice guideline (10/1/2014) as the definition of early and late glargine.

In order to test whether the effect of the change in CPG differed for early onset vs. established patients, linear models with an interaction between group (i.e., early vs. late or before vs. after CPG) and early onset.  None of the interaction terms were significant, so there is no evidence of a differential effect in early onset vs. established patients.

# Results

There were `r N` subjects included in the analysis using the definition of early and late glargine.  Descriptive statistics are shown in Table 1.

```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab1,caption="Table 1.  Descriptive statistics.  Statistics given are N(%), mean?SD, or median (25th %ile, 75th %ile).  * denotes that the Mann-Whitney test was used, and ** denotes that Fisher's exact test was used.")
```


There were `r Ncpg` subjects included in the analysis using the date of CPG initiation.  Descriptive statistics for before and after the initiation of the CPG are shown in Table 1a.

```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab1a,caption="Table 1a.  Descriptive statistics.  Statistics given are N(%), mean?SD, or median (25th %ile, 75th %ile).  * denotes that the Mann-Whitney test was used, and ** denotes that Fisher's exact test was used.")
```



Table 2 shows the results of the comparisons of study outcomes using the timing definition of early and late glargine.

```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab_primary,caption="Table 2.  Comparison of study outcomes.  Statistics given are N(%), mean?SD, or median (25th %ile, 75th %ile).  * denotes that the Mann-Whitney test was used, and ** denotes that Fisher's exact test was used.")
```

Table 2a shows the results of comparison of study outcomes, before and after the initiation of the CPG.


```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab_primarya,caption="Table 2a.  Comparison of study outcomes.  Statistics given are N(%), mean?SD, or median (25th %ile, 75th %ile).  * denotes that the Mann-Whitney test was used, and ** denotes that Fisher's exact test was used.")
```

```{r, results='asis',tidy=TRUE, echo=FALSE}
boxplot(dur_iv_insulin_hours~group,data=final,xlab="",ylab="Duration of IV insulin (hours)")

ggplot2.barplot(data=table_hyper_long, xName="labels", yName="value", groupName="Group",
                position=position_dodge(), ytitle="% Rebound Hyperglycemia",xtitle="",
                groupColors = c("#3d3c3b","#aba9a7"))

barplot(table_sh,ylab = "% Severe Hypoglycemia (<54 mg/dL)")

ggplot2.barplot(data=table_hypo70_long, xName="labels", yName="value", groupName="Group",
                position=position_dodge(), ytitle="% Hypoglycemia (<70 mg/dL)",xtitle="",
                groupColors = c("#3d3c3b","#aba9a7"),xtickLabelFont=c("6","bold","black"))

ggplot2.barplot(data=table_hypo54_long, xName="labels", yName="value", groupName="Group",
                position=position_dodge(), ytitle="% Hypoglycemia (<54 mg/dL)",xtitle="",
                groupColors = c("#3d3c3b","#aba9a7"),xtickLabelFont=c("6","bold","black"))

```