---
title: "Early Glargine in DKA"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

setwd("H:\\Endocrinology\\Ohman\\Ohman early glargine\\Data")

#install_github('palmercl/Table1')

library(tidyr)
library(magrittr)
library(dplyr)
library(lubridate)
library(zoo) 
library(M3)
library(data.table)
library(Table1)
library(devtools)
library(Hmisc)
library(knitr)

####################################################################################################
#Data Manipulation #
####################################################################################################

completeFun <- function(data, desiredCols) {
  completeVec <- complete.cases(data[, desiredCols])
  return(data[completeVec, ])
}

# use Claire's functions
source(file='H:\\R\\Markdown sample\\01_functions.r')

# read main data sheet
all <- read.csv("H:\\Endocrinology\\Ohman\\Ohman early glargine\\Data\\CMC_DKA deidentified_1.30.2018 main sheet.csv",na.strings = c("","NA"," "))
head(all)
N.MRFin <- dim(distinct(all,MRFin))
N.StudyID <- dim(distinct(all,StudyID))
# create a new study ID for merging with glucose data (in combination with age because that's the only way to tell in glucose data)
all$StudyID.glucose <- sub('\\_.*','',all$StudyID)

# read glucose data sheet
gluc <- read.csv("H:\\Endocrinology\\Ohman\\Ohman early glargine\\Data\\CMC_DKA deidentified_1.30.2018 glucose.csv")
# change the name of the glucose ID variable so it can be matched to the main datasheet
setnames(gluc,"StudyID","StudyID.glucose")
# convert blank glucose and weird entries to NA
gluc$Glucose[gluc$Glucose==" "]  <- NA
gluc$Glucose[gluc$Glucose=="Already Documented"]  <- NA
gluc$Glucose[gluc$Glucose=="See Comment"]  <- NA
gluc$Glucose[gluc$Glucose=="Task Duplication"]  <- NA
gluc$Glucose[gluc$Glucose=="TNP"]  <- NA
# get rid of missing glucose values
gluc <- completeFun(gluc,"Glucose.from.Admit..Minutes.")
gluc <- completeFun(gluc,"Glucose")
gluc$Glucose <- as.numeric(as.character(gluc$Glucose))

# keep needed variables for primary analysis from main data sheet
# keep insulin end time, not glargine
myvars <- c("StudyID","StudyID.glucose","MRFin","Age","Sex","Ethnicity","Height","Weight","HbA1c.Closest.to.Admisson","HbA1c.Most.recent.prior.to.Admission","Insurance.Type",
            "Median.Income","October.1.2014.Flag","Admit_mod_Date","Admit_Time","Discharge_mod_Date","Discharge_Time",
            "Last.Insulin.Administration_mod_Date","Last.Insulin.Administration_Time","Glargine_Time","Glargine_mod_Date",
            "First.Insulin.Start_Time","First.Insulin.Start_mod_Date")
hosp_er <- all[myvars]
time_vars <- c("StudyID","StudyID.glucose","Last.Insulin.Administration_mod_Date","Last.Insulin.Administration_Time")

# get rid of missing insulin dates and times
hosp_er <- completeFun(hosp_er,"Last.Insulin.Administration_mod_Date")

# in the hosp_er dataset, study IDs are unique to an admission (e.g., 1151_1 and 1151_2)
# but there can be multiple records for the same admission, if the person moved from ER to inpatient
# keep the one with the latest date and time of IV insulin administration
sortds <- hosp_er[order(hosp_er$StudyID,desc(hosp_er$Last.Insulin.Administration_mod_Date),desc(hosp_er$Last.Insulin.Administration_Time)),]
lastinsulin <- sortds %>% group_by(StudyID) %>% filter(row_number(StudyID) == 1)

# merge last date and time of insulin administration to the glucose data
# keep only those ID/Age combinations that are in both dataframes
merged <- merge(lastinsulin,gluc,by=c("StudyID.glucose","Age"))

# calculate minutes between admit and insulin discontinuation
merged$instime <- as_datetime(hms(merged$Last.Insulin.Administration_Time))
merged$admittime <- as_datetime(hms(merged$Admit_Time))
# elapsed is just the difference between insulin d/c and admission but does not account for whether those are on different days - that's fixed later
merged$elapsed <- difftime(merged$instime,merged$admittime,units="mins") 
merged$min_admit_insdc <- (as.duration(merged$elapsed) + ddays(merged$Last.Insulin.Administration_mod_Date-merged$Admit_mod_Date))/dminutes(1)
merged$min_insdc_gluc <- ifelse(merged$Glucose.from.Admit..Minutes.>merged$min_admit_insdc,merged$Glucose.from.Admit..Minutes.-merged$min_admit_insdc,NA)
check <- c("StudyID","Admit_Time","admittime","Last.Insulin.Administration_Time","instime","elapsed","Last.Insulin.Administration_mod_Date","Admit_mod_Date","min_admit_insdc")

# create new group variable for overlap between IV insulin and glargine
# if glargine is before IV insulin start, then they are in the early group
# if glargine is after IV insulin start, then check overlap.  If overlap >4 hours, then early.  If overlap <2 hours, then late.
# Glargine time is not always the same day as admission
merged$glarginetime <- as_datetime(hms(merged$Glargine_Time))
merged$temp <- difftime(merged$glarginetime,merged$admittime,units="mins")
merged$min_admit_glargine <- (as.duration(merged$temp) +   ddays(merged$Glargine_mod_Date-merged$Admit_mod_Date))/dminutes(1)
merged$min_insdc_glargine <- merged$min_admit_glargine - merged$min_admit_insdc
merged$temp1 <- merged$min_admit_insdc-240
merged$temp2 <- merged$min_admit_insdc-120
merged$group[merged$min_admit_glargine<(merged$min_admit_insdc-240)] <- "Early"
merged$group[merged$min_admit_glargine>(merged$min_admit_insdc-120)] <- "Late"
check <- c("StudyID","Admit_mod_Date","admittime","Last.Insulin.Administration_mod_Date","instime","min_admit_insdc","Glargine_mod_Date","glarginetime","min_admit_glargine","min_insdc_glargine","temp1","temp2","group")
count <- distinct(merged,StudyID.glucose, .keep_all=TRUE)
table(count$group,useNA="always")
table(count$group,count$October.1.2014.Flag)
summary(merged$min_insdc_glargine)

# create new dataframe with glucose values that are within 12 hours of insulin discontinuation, for calculating median
gluc_med <- merged[!is.na(merged$min_insdc_gluc) & merged$min_insdc_gluc<720,]
gluc_median <- aggregate(gluc_med$Glucose,list(gluc_med$StudyID.glucose),median)
setnames(gluc_median,"Group.1","StudyID.glucose")
setnames(gluc_median,"x","median_glucose")

# determine rebound hyperglycemia - defined as glucose >180 within 12 hours of insulin discontinuation
# compare minutes after admit that glucose was measured to minutes after admit that insulin was last administered to see if w/in window
merged$rebound_hyper <- ifelse((!is.na(merged$min_insdc_gluc)) & (merged$min_insdc_gluc<720) & (merged$Glucose>180),1,0)
merged$test <- ifelse(merged$min_insdc_gluc<720,1,0)

# create a new variable for hypoglycemia (<70) within 12 hours of insulin discontinuation
merged$hypo_12hr <- ifelse((!is.na(merged$min_insdc_gluc)) & (merged$min_insdc_gluc<720) & (merged$Glucose<70),1,0)

# create a new variable for hypoglycemia (<70) during IV insulin administration
#merged$hypo_IVins <- 

# need to keep highest value of rebound_hyper
merged <- merged[order(merged$StudyID,desc(merged$rebound_hyper)),]
# having both of the next 2 statements should not work/be necessary
rebound_hyper <- merged %>% group_by(StudyID) %>% filter(row_number(StudyID) == 1)
#rebound_hyper <- merged %>% group_by(StudyID.glucose) %>% filter(row_number(StudyID.glucose) == 1)
#rebound_hyper <- merged %>% group_by(StudyID.glucose) %>% top_n(1,StudyID.glucose)
rebound_hyper[rebound_hyper$rebound_hyper==0,
             c("StudyID.glucose","min_admit_insdc","Glucose.from.Admit..Minutes.","min_insdc_gluc","Glucose","rebound_hyper","group")]

# create new dataframe to keep highest value of hypoglycemia within 12 hours
hypo12 <- merged[order(merged$StudyID,desc(merged$hypo_12hr)),]
hypo12 <- hypo12 %>% group_by(StudyID) %>% filter(row_number(StudyID) == 1)

# combine Military and Private insurance
rebound_hyper$Insnew <-as.character(rebound_hyper$Insurance.Type)
rebound_hyper$Insnew[rebound_hyper$Insurance.Type=='Military'|rebound_hyper$Insurance.Type=='Private']<-'Private/Military'
rebound_hyper$Insnew[rebound_hyper$Insurance.Type=='NaN']<-NA
rebound_hyper$Insnew <-as.factor(rebound_hyper$Insnew)

# merge back in median glucose values
rebound_hyper <- merge(rebound_hyper,gluc_median,by.x="StudyID.glucose",by.y="StudyID.glucose")


####################################################################################################

####################################################################################################
#Descriptive Tables #
####################################################################################################

# Set factors and levels
rebound_hyper$rebound_hyper <- as.factor(rebound_hyper$rebound_hyper)
rebound_hyper$group <- as.factor(rebound_hyper$group)
levels(rebound_hyper$rebound_hyper)=c('No','Yes')
#levels(rebound_hyper$October.1.2014.Flag)=c('After','Before')
levels(rebound_hyper$Sex)=c('Female','Male')
levels(rebound_hyper$Ethnicity)=c('American Indian or Alaska Native','Asian','Black or African American','Hispanic','Multiracial'
                                  ,'Native Hawaiian or Other Pacific Islander','Other','White')
levels(rebound_hyper$Insnew)=c('Private/Military','Public','Self Pay')

# convert to numeric
rebound_hyper$HbA1c.Closest.to.Admisson <- as.numeric(rebound_hyper$HbA1c.Closest.to.Admisson)

# Set Labels
label(rebound_hyper$rebound_hyper)='Rebound hyperglycemia'
label(rebound_hyper$group)
label(rebound_hyper$Sex)='Sex'
label(rebound_hyper$Ethnicity)='Ethnicity'
label(rebound_hyper$Insnew)='Insurance'
label(rebound_hyper$Age)='Age'
label(rebound_hyper$Height)='Height'
label(rebound_hyper$Weight)='Weight'
label(rebound_hyper$HbA1c.Closest.to.Admisson)='HbA1c closest to admission'
label(rebound_hyper$Median.Income)='Median income'
label(rebound_hyper$median_glucose)='Median glucose'

N <- sum(!is.na(rebound_hyper$group))
Nexcl <- sum(is.na(rebound_hyper$group))

tab1 <- final_table(rebound_hyper,c('Sex','Ethnicity',"Insnew","Age","Height","Weight","HbA1c.Closest.to.Admisson","Median.Income"),
                    group=rebound_hyper$group,margin=2,single=0,ron=2)

####################################################################################################

####################################################################################################
#Primary Outcome #
####################################################################################################

# early vs. late glargine variable is October.1.2014.Flag
#tbl <- table(rebound_hyper$rebound_hyper,rebound_hyper$October.1.2014.Flag)
#rslt <- chisq.test(tbl)
tab_primary <- final_table(rebound_hyper,'rebound_hyper',group=rebound_hyper$group,margin=2,single=0,ron=2)

####################################################################################################
#Median glucose #
####################################################################################################
tab_medgluc <- final_table(rebound_hyper,'median_glucose',group=rebound_hyper$group,margin=2,single=0,ron=2)

####################################################################################################
#Hypoglycemia within 12 hours after insulin discontinuation #
####################################################################################################
tab_hypo12 <- final_table(rebound_hyper,'rebound_hyper',group=rebound_hyper$group,margin=2,single=0,ron=2)

```

# Background

The purpose of this study was to determine if early glargine administration in children with DKA prevents rebound hyperglycemia and recurrent ketogenesis without increasing the rate of hypoglycemia.   

# Methods

Early glargine administration was defined as either (1) glargine administered before the start of IV insulin administration or (2) glargine administered after the start of IV insulin administration with >4 hours of overlap with IV insulin.  Late glargine administration was defined as glargine administered with <2 hour overlap with IV insulin.  There were `r Nexcl` participants with between 2-4 hours of overlap with IV insulin who were excluded from this analysis.
 
The primary outcome of this study is rebound hyperglycemia, defined as serum glucose >180 mg/dL within 12 hours of discontinuation of IV insulin.  

The following data cleaning and manipulation steps were performed in order to determine whether each participant met the definition of rebound hyperglycemia:

1)  Study IDs on the main data sheet ending in "_1" or something similar, to indicate distinct encounters, were stripped of the suffix starting with the underscore, so they would match the IDs on the glucose data sheet.
2)  Missing glucose values or those with comments such as ("Already Documented","See Comment","Task Duplication", and "TNP") deleted from the glucose results.
3)  Encounters that were missing insulin administration dates and/or times were deleted.
4)  For each encounter, the last date and time of IV insulin administration was determined.
5)  The main data sheet was merged with glucose results by a combination of study ID and age, since a participant could have more than one encounter.
6)  If a participant had a glucose value >180 mg/dL within 12 hours of IV insulin discontinuation, s/he was classified as having rebound hyperglycemia.

Additional data cleaning and manipulation steps included:

1)  Military and private insurance were combined into one category.

Descriptive statistics reported include mean and standard deviations for continuous variables, and frequencies and percentages for categorical variables.  To compare characteristics in the two groups, t-tests were used for continuous variables and the chi-square test or Fisher's exact test was used for categorical variables.

# Results

There were `r N` subjects included in this analysis.  Descriptive statistics are shown in Table 1.

```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab1,caption="Table 1.  Descriptive statistics.  Asterisks indicate that Fisher's exact test was used instead of the chi-square test due to small expected cell value.")
```
\

There was a significant difference in the rate of rebound hyperglycemia in the two groups (Table 2).

```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab_primary,caption="Table 2.  Comparison of rates of rebound hyperglycemia.")
```

\

There was no difference in the median glucose values during the 12 hours after insulin discontinuation (Table 3).

```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab_medgluc,caption="Table 3.  Comparison of median glucose values in the 12 hours after insulin discontinuation.")
```


