---
title: "Early Glargine Prospective"
author: "Casey Sakamoto"
date: "2/8/2022"
output: pdf_document
---

```{r setup, include=FALSE}
library(arsenal)
library(readxl)
library(readr)
library(Hmisc)
library(tidyverse)
library(performance)
library(knitr)
library(lubridate)
library(stringr)
library(table1)
knitr::opts_chunk$set(echo = FALSE)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "S://Laura/Peds Endo/Ohman/Early glargine prospective"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Ohman/Early glargine prospective"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Ohman/Early glargine prospective"
}
knitr::opts_knit$set(root.dir = home_dir)
```


```{r data, include=FALSE}
# raw dataset
EarlyGlargine = read_csv("S:/Laura/Peds Endo/Ohman/Early glargine prospective/EarlyGlargineInDKA_DATA_12.2021.csv")

# check screenfailure/wd
# screenfailures only have 1 line: demog
# 4 sf: 1, 3, 23,  110
# 7 wd (completion_status = 1) 24, 25, 29, 31, 33, 34, 115; keep these in as its a rct and itt analysis
EarlyGlargine_SF = EarlyGlargine %>% filter(screen_failure == 1 ) %>% select(`Patient ID`)

# no screen failures
EarlyGlargine_NSFW = EarlyGlargine %>% filter(!`Patient ID` %in% EarlyGlargine_SF$`Patient ID`)

########################## PATIENT CHARACTERISTICS TABLE ################################

# demographic data (1 line/subj)
EarlyGlargine_demog = EarlyGlargine_NSFW %>% filter(redcap_repeat_instrument == "demographics") %>%
  select(`Patient ID`, study_group, nih_sex, nih_race, nih_ethnicity, age_at_consent)

# clinical data (1 line/subj)
EarlyGlargine_clinicalinfo =EarlyGlargine_NSFW %>% filter(redcap_repeat_instrument == "clinical_information") %>% select(`Patient ID`,diabetes_diagnosis, `duration_of_t1d (years)`, glargine_dose_per_kg, home_regimen___1, home_regimen___2, home_regimen___3, home_regimen___4, home_regimen___5 )

# lab data (1 line/ hba1c, mult for ph and bicarbonate (we want admission, instance 1?))
EarlyGlargine_lab = EarlyGlargine_NSFW %>% filter(redcap_repeat_instrument == "lab_results_dka" & redcap_repeat_instance == 1) %>%select(`Patient ID` ,ph, bicarbonate, hba1c)

# combine these datasets
EG_patient_char = full_join(EarlyGlargine_demog, EarlyGlargine_clinicalinfo)
EG_patient_char = full_join(EG_patient_char, EarlyGlargine_lab)

rm(EarlyGlargine_demog); rm(EarlyGlargine_clinicalinfo); rm(EarlyGlargine_lab); rm(EarlyGlargine_SF)

# make a single home regimen variable
EG_patient_char = EG_patient_char %>% mutate(home_regimen = case_when(home_regimen___1 == 1 ~ 1,
                                                                                   home_regimen___2 == 1 ~ 2,
                                                                                   home_regimen___3 == 1 ~ 3,
                                                                                   home_regimen___4 == 1 ~ 4,
                                                                                   home_regimen___5 == 1 ~ 5))

EG_patient_char = EG_patient_char %>% select( -c(home_regimen___1:home_regimen___5))
  

###### ADDED 2/23 IMPUTING MISSING VALUES
# ASK BECKY TO DOUBLE CHECK REDCAP SO I DONT HAVE TO DO THIS

# Age
EG_patient_char$age_at_consent[EG_patient_char$`Patient ID`=="16.1965_07"] = 12

# t1dm 2= est 1 = new
EG_patient_char$diabetes_diagnosis[EG_patient_char$`Patient ID`=="16.1965_08"] = 2
EG_patient_char$diabetes_diagnosis[EG_patient_char$`Patient ID`=="16.1965_24"] = 2
EG_patient_char$diabetes_diagnosis[EG_patient_char$`Patient ID`=="16.1965_29"] = 2
EG_patient_char$diabetes_diagnosis[EG_patient_char$`Patient ID`=="16.1965_31"] = 2
EG_patient_char$diabetes_diagnosis[EG_patient_char$`Patient ID`=="16.1965_33"] = 2
EG_patient_char$diabetes_diagnosis[EG_patient_char$`Patient ID`=="16.1965_34"] = 2

EG_patient_char$diabetes_diagnosis[EG_patient_char$`Patient ID`=="16.9165_110"] = 1
EG_patient_char$diabetes_diagnosis[EG_patient_char$`Patient ID`=="16.1965_115"] = 1
EG_patient_char$diabetes_diagnosis[EG_patient_char$`Patient ID`=="16.1965_117"] = 1

# hba1c
EG_patient_char$hba1c[EG_patient_char$`Patient ID`=="16.1965_04"] = 10.4
EG_patient_char$hba1c[EG_patient_char$`Patient ID`=="16.1965_07"] = 8
EG_patient_char$hba1c[EG_patient_char$`Patient ID`=="16.1965_10"] = 8.1
EG_patient_char$hba1c[EG_patient_char$`Patient ID`=="16.1965_101"] = 14
EG_patient_char$hba1c[EG_patient_char$`Patient ID`=="16.1965_103"] = 12.6
EG_patient_char$hba1c[EG_patient_char$`Patient ID`=="16.1965_11"] = 15
EG_patient_char$hba1c[EG_patient_char$`Patient ID`=="16.9165_110"] = 13.5
EG_patient_char$hba1c[EG_patient_char$`Patient ID`=="16.1965_115"] = 13.4
EG_patient_char$hba1c[EG_patient_char$`Patient ID`=="16.1965_116"] = 10.7
EG_patient_char$hba1c[EG_patient_char$`Patient ID`=="16.1965_118"] = 15
EG_patient_char$hba1c[EG_patient_char$`Patient ID`=="16.1965_12"] = 14
EG_patient_char$hba1c[EG_patient_char$`Patient ID`=="16.1965_16"] = 8.4
EG_patient_char$hba1c[EG_patient_char$`Patient ID`=="16.1965_24"] = 13
EG_patient_char$hba1c[EG_patient_char$`Patient ID`=="16.1965_29"] = 14
EG_patient_char$hba1c[EG_patient_char$`Patient ID`=="16.1965_30"] = 6.6
EG_patient_char$hba1c[EG_patient_char$`Patient ID`=="16.1965_31"] = 10.1
EG_patient_char$hba1c[EG_patient_char$`Patient ID`=="16.1965_32"] = 11.5
EG_patient_char$hba1c[EG_patient_char$`Patient ID`=="16.1965_33"] = 11.1
EG_patient_char$hba1c[EG_patient_char$`Patient ID`=="16.1965_34"] = 9.0

# make Vars pretty for the table
EG_patient_char$study_group = factor(EG_patient_char$study_group, levels = c(1,2), labels = c("Early Glargine", # intervention
                                                                                              "Late Glargine"))


EG_patient_char$nih_sex = factor(EG_patient_char$nih_sex, levels = c(1,2), labels = c("Female",
                                                                                      "Male"))
label(EG_patient_char$nih_sex) = "Sex"

EG_patient_char$nih_race = factor(EG_patient_char$nih_race, levels = c(1,2,3,4,5,6,7), labels = c("American Indian or Alaska Native",
                                                                                                  "Asian",
                                                                                                  "Black or African-American",
                                                                                                  "Native Hawaiian or Other Pacific Islander",
                                                                                                  "White",
                                                                                                  "More than one race",
                                                                                                  "Unknown or not reported"))
label(EG_patient_char$nih_race) = "Race"

EG_patient_char$nih_ethnicity = factor(EG_patient_char$nih_ethnicity, levels = c(1,2), labels = c("Hispanic or Latino",
                                                                                                  "Not Hispanic or Latino"))
label(EG_patient_char$nih_ethnicity) = "Ethnicity"


EG_patient_char$diabetes_diagnosis = factor(EG_patient_char$diabetes_diagnosis, levels = c(1,2), labels = c("New onset T1D",
                                                                                                            "Established T1D"))

label(EG_patient_char$diabetes_diagnosis) = "Diabetes Diagnosis"

EG_patient_char$home_regimen = factor(EG_patient_char$home_regimen, levels = c(1,2,3,4,5), labels = c("Insulin Pump",
                                                                                                      "MDI",
                                                                                                      "SSI",
                                                                                                      "N/A (New Onset)",
                                                                                                      "Other"))

label(EG_patient_char$home_regimen) = "Home Regimen"


label(EG_patient_char$age_at_consent) = "Age at Consent"
label(EG_patient_char$hba1c) = "HbA1c"
label(EG_patient_char$ph) = "pH at admission"
label(EG_patient_char$bicarbonate) = "Bicarbonate at admission"
label(EG_patient_char$glargine_dose_per_kg) = "Glargine dose/kg"
label(EG_patient_char$`duration_of_t1d (years)`) = "Duration of T1D (years)"

########################################################################################



######################        PRIMARY OUTCOME         ##################################

# look at our main variables:
# group
EG_primary_group = EarlyGlargine_NSFW %>% filter(redcap_repeat_instrument == "demographics") %>%
  select(`Patient ID`, study_group)
# stop time
EG_primary_stoptime = EarlyGlargine_NSFW %>% filter(redcap_repeat_instrument == "clinical_information") %>%
  select(`Patient ID`, iv_insulin_stop)
# outcomes
EG_primary_lab = EarlyGlargine_NSFW %>% filter(redcap_repeat_instrument == "lab_results_dka") %>%
  select(`Patient ID`,redcap_repeat_instance,time_of_lab, glucometer)

# combine 
EG_primary_outcome = full_join(EG_primary_group, EG_primary_stoptime)
EG_primary_outcome = full_join(EG_primary_outcome, EG_primary_lab)

# make date times out of char variables
EG_primary_outcome = EG_primary_outcome %>% mutate(timediff = difftime( mdy_hm(time_of_lab), mdy_hm(iv_insulin_stop), units = "hours"))
########################################################################################
```

```{r save data, eval=FALSE}
saveRDS(EG_primary_outcome,"S:/Laura/Peds Endo/Ohman/Early glargine prospective/EG_primary_outcome.rds" )
saveRDS(EG_patient_char,"S:/Laura/Peds Endo/Ohman/Early glargine prospective/EG_patient_char.rds" )
```

```{r table1 data check, eval=FALSE}
# Check Numeric Variables Distributions
hist(EG_patient_char$age_at_consent)
hist(EG_patient_char$`duration_of_t1d (years)`)
hist(EG_patient_char$glargine_dose_per_kg)
hist(EG_patient_char$ph)
hist(EG_patient_char$bicarbonate)
hist(EG_patient_char$hba1c)
```

Analysis plan:
Chi-squared tests will be employed to evaluate whether the difference in rebound hyperglycemia between the two study groups is statistically significant.  We will also use logistic regression to adjust for potential confounders (BMI, sex, age, established T1D versus newly-diagnosed T1D).   Using a blocked and stratified randomization plan we expect to have similar numbers of newly-diagnosed T1D and known T1D patients in each group.

The data collected by CGM will be summarized for each patient and day into variables describing glycemia, such as mean glucose, area under the curve, area under the curve >180 mg/dL, and percent of time spent above or below particular glucose thresholds (e.g., >180, <60, and <70 mg/dL). 

Trends over time in these summary statistics will be evaluated graphically to give insight into blood glucose response during treatment of DKA and immediately after resolution of DKA.  For variables without an excess of zero values, we will use linear mixed-effects models to compare trajectories over time in the two groups. 

We will also calculate the mean value of each of these statistics for each patient and compare the groups using t-tests or nonparametric equivalent.  Bland-Altman plots will be used to evaluate the agreement between CGM glucose variables (e.g., mean glucose) and POC glucose. 

To interpret the clinical relevance relevance of any discrepancies between the two sets of measures, we will use Clarke error grid analysis, which classifies pairs of blood glucose measurements made by two methods into categories based on the impact the discrepancy would have on treatment decisions.

