---
title: "Early Glargine in DKA analysis"
author: "Casey Sakamoto"
date: "2/23/2022"
output:
  html_document:
    df_print: paged
  word_document: default
---

```{r setup, include=FALSE}
library(arsenal)
library(readxl)
library(readr)
library(Hmisc)
library(tidyverse)
library(performance)
library(knitr)
library(lubridate)
library(stringr)
library(table1)
knitr::opts_chunk$set(echo = FALSE)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "S:/Laura/Peds Endo/Ohman/Early glargine prospective"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Ohman/Early glargine prospective"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Ohman/Early glargine prospective"
}
knitr::opts_knit$set(root.dir = home_dir)

#  table1 function for p-values
pvalue <- function(x, ...) {
    # Construct vectors of data y, and groups (strata) g
    y <- unlist(x)
    g <- factor(rep(1:length(x), times=sapply(x, length)))
    if (is.numeric(y)) {
        # For numeric variables, perform a standard 2-sample t-test
        p <- t.test(y ~ g)$p.value
    } else {
        # For categorical variables, perform a fisher-exact test of independence
        p <- fisher.test(table(y, g))$p.value
    }
    # Format the p-value, using an HTML entity for the less-than sign.
    # The initial empty string places the output on the line below the variable label.
    c("", sub("<", "&lt;", format.pval(p, digits=3, eps=0.001)))
}

# data
EG_primary_outcome = readRDS(file = "S:/Laura/Peds Endo/Ohman/Early glargine prospective/EG_primary_outcome.rds")
EG_patient_char = readRDS(file = "S:/Laura/Peds Endo/Ohman/Early glargine prospective/EG_patient_char.rds")

# new onset duration set from na to 0 per Becky 2/18
EG_patient_char = EG_patient_char %>% mutate(`duration_of_t1d (years)` = case_when(diabetes_diagnosis == "New onset T1D" & is.na(`duration_of_t1d (years)`) == TRUE ~ 0,
                                                                                   TRUE ~ `duration_of_t1d (years)`))
label(EG_patient_char$`duration_of_t1d (years)`) = "Duration of T1D (years)"
```

# 1 Background

The purpose of this analysis is to summarize patient characteristics by treatment groups and T1D onset, and to examine the difference in rebound hyperglycemia between the two study groups.

# 2 Outstanding Data Questions/ Remarks

### updated 2/23 to include some missing data


# 3 Methods

Subjects with screen failures were not included in analysis.

Tables generated using the *table1* package in R, with t-tests and Fisher's exact tests used to compare variables across either treatment group or diabetes diagnosis.

The primary outcome is the rate of rebound hyperglycemia within 12 hours of IV insulin discontinuation. A Fisher's exact test was used to evaluate the difference in rebound hyperglycemia between the two study groups.

# 4 Results
```{r Primary Outcome data wrangle, include=FALSE}
# need to derive the outcome variable: first filter out values that are before iv stop, or after 12 hr post stop
EG_primary_outcome = EG_primary_outcome %>% filter(timediff > 0 & timediff < 12)

# look at glucometer values within the 12hr window (note: a few 180's so i opted for 180 as the cutoff rather than 179)
EG_primary_outcome = EG_primary_outcome %>% mutate(hg_12hr = case_when(glucometer >=  180 ~ 1,
                                                                             TRUE ~ 0),
                                                   hg_12hr180_250 = case_when(glucometer >= 180 & glucometer <=  250 ~ 1,
                                                                             TRUE ~ 0),
                                                   
                                                   hg_12hr250_300 = case_when(glucometer >= 250 & glucometer <=  300 ~ 1,
                                                                             TRUE ~ 0),
                                                   
                                                   hg_12hr300 = case_when(glucometer >=  300 ~ 1,
                                                                             TRUE ~ 0))

# hyperglycemia: did they have glucose > 180 ANY time within 12 hr of iv stop
EG_primary_outcome = EG_primary_outcome %>% group_by(`Patient ID`) %>% mutate(hyperglycemia = max(hg_12hr),
                                                                              hyperglycemia180250 = max(hg_12hr180_250),
                                                                              hyperglycemia250300 = max(hg_12hr250_300),
                                                                              hyperglycemia300 = max(hg_12hr300))


##################### using 180 threshold #############################
# check proportion of those who got Hyperglycemia
EG_prop = EG_primary_outcome %>% group_by(`Patient ID`, study_group) %>% summarise("hyperglycemia" = max(hyperglycemia))

# add this into the patient characteristics data frame
EG_prop2 = EG_prop %>% select(`Patient ID`, hyperglycemia)
EG_patient_char = full_join(EG_patient_char, EG_prop2)


#######################################################################

##################### using 180-250 threshold #############################
EG_prop180250 = EG_primary_outcome %>% group_by(`Patient ID`) %>% summarise("hyperglycemia180250" = max(hyperglycemia180250))

# add this into the patient characteristics data frame
EG_patient_char = full_join(EG_patient_char, EG_prop180250)
#######################################################################

##################### using 250-300 threshold #############################
EG_prop180250 = EG_primary_outcome %>% group_by(`Patient ID`) %>% summarise("hyperglycemia250300" = max(hyperglycemia250300))

# add this into the patient characteristics data frame
EG_patient_char = full_join(EG_patient_char, EG_prop180250)
#######################################################################

######################## 300 threshold ##################################
EG_prop300 = EG_primary_outcome %>% group_by(`Patient ID`) %>% summarise("hyperglycemia300" = max(hyperglycemia300))


# add this into the patient characteristics data frame
EG_patient_char = full_join(EG_patient_char, EG_prop300)

##################################################################

################## TABLE THE OUTCOMES ############################
EG_prop180250_df = EG_primary_outcome %>% group_by(`Patient ID`, study_group) %>% summarise("hyperglycemia180250" = max(hyperglycemia180250))
EG_prop250300_df = EG_primary_outcome %>% group_by(`Patient ID`, study_group) %>% summarise("hyperglycemia250300" = max(hyperglycemia250300))
EG_prop300_df = EG_primary_outcome %>% group_by(`Patient ID`, study_group) %>% summarise("hyperglycemia300" = max(hyperglycemia300))

# want a table of proportion having the outcome in each group along with a p-value
hyperglycemia_prop = full_join(EG_prop, EG_prop180250_df)
hyperglycemia_prop = full_join(hyperglycemia_prop, EG_prop250300_df)
hyperglycemia_prop = full_join(hyperglycemia_prop, EG_prop300_df)

# label variables
hyperglycemia_prop$study_group = factor(hyperglycemia_prop$study_group, levels = c(1,2), labels = c("Early Glargine", # intervention
                                                                                              "Late Glargine"))

hyperglycemia_prop$hyperglycemia = factor(hyperglycemia_prop$hyperglycemia, levels = c(0,1), labels = c("No Hyperglycemia", 
                                                                                              "Hyperglycemia"))

hyperglycemia_prop$hyperglycemia180250 = factor(hyperglycemia_prop$hyperglycemia180250, levels = c(0,1), labels = c("Outside Range", 
                                                                                              "Within Range"))

hyperglycemia_prop$hyperglycemia250300 = factor(hyperglycemia_prop$hyperglycemia250300, levels = c(0,1), labels = c("Outside Range", 
                                                                                              "Within Range"))

hyperglycemia_prop$hyperglycemia300 = factor(hyperglycemia_prop$hyperglycemia300, levels = c(0,1), labels = c("Below Threshold", 
                                                                                              "Above Threshold"))
label(hyperglycemia_prop$study_group) = "Study Group"
label(hyperglycemia_prop$hyperglycemia) = ">180mg/dL Threshold"
label(hyperglycemia_prop$hyperglycemia180250) = "180 - 250 mg/dL Threshold"
label(hyperglycemia_prop$hyperglycemia250300) = "250 - 300 mg/dL Threshold"
label(hyperglycemia_prop$hyperglycemia300) = ">300mg/dL Threshold"


# proportion having outcome

hyper_prop_table =  table1( ~ hyperglycemia + hyperglycemia180250+ hyperglycemia250300 + hyperglycemia300 | study_group, data = hyperglycemia_prop, topclass="Rtable1-zebra", overall=F, extra.col=list(`P-value`=pvalue))
##################################################################
```

```{r primary outcome, echo=FALSE}
hyper_prop_table

```

Using a cutoff of >180mg/dL, 76.0% of subjects experienced hyperglycemia. When stratified by treatment group, 76.9% of subjects in the early group and 75.0% of subjects in the late group experienced hyperglycemia. There is not enough evidence to suggest that IV rebound rate and study group are related (p = 1)

Using a BG range of 180-250mg/dL, 52.0% of subjects experienced BG levels within the range. When stratified by treatment group, 50.0% of subjects in the early group and 54.2% of subjects in the late group experienced hyperglycemia. There is not enough evidence to suggest that IV rebound rate and study group are related (p = 0.785)

Using a BG range of 250-300mg/dL, 34.0% of subjects experienced BG levels within the range. When stratified by treatment group, 30.8% of subjects in the early group and 37.5% of subjects in the late group experienced hyperglycemia. There is not enough evidence to suggest that IV rebound rate and study group are related (p = 0.767)

Using a cutoff of >300mg/dL, 26.0% of subjects experienced BG levels above the threshold. When stratified by treatment group, 23.1% of subjects in the early group and 29.2% of subjects in the late group experienced BG levels above the threshold. There is not enough evidence to suggest that IV rebound rate and study group are related (p = 0.751)

Note: 7 subjects withdrew prior to obtaining lab results

# 4.1 Patient Characteristics Table

```{r Patient Characteristics, echo=FALSE}
# tables
EG_demog_table1 =  table1( ~ nih_sex + age_at_consent + nih_race + nih_ethnicity + diabetes_diagnosis + hba1c + ph + bicarbonate + glargine_dose_per_kg
                        | study_group, data = EG_patient_char, topclass="Rtable1-zebra",
                         render.continuous=c( "Median [IQR]"="Median [Q1, Q3]"), overall=F, extra.col=list(`P-value`=pvalue))

T1D_lab_table1 =  table1( ~ `duration_of_t1d (years)` + home_regimen + hba1c + ph + bicarbonate + glargine_dose_per_kg 
                       | diabetes_diagnosis, data = EG_patient_char, topclass="Rtable1-zebra",
                        render.continuous=c( "Median [IQR]"="Median [Q1, Q3]"), overall=F, extra.col=list(`P-value`=pvalue))


EG_demog_table1 # between study groups, most variables look very similar; hba1c, bicarbonate may be a bit different
T1D_lab_table1 # between t1d status, hba1c, bicarbonate, glargine dose may be a bit different
```

# 5 Code Appendix
```{r code appendix}
# 180 threshold #############
# check proportion of those who got Hyperglycemia
mean(EG_prop$hyperglycemia) #.76 got hg
# by study group
EG_prop %>% group_by(study_group) %>% summarise("hyperglycemia" = mean(hyperglycemia))
# could also adjust for covariates using logistic regression
eg_logreg = glm(as.factor(hyperglycemia) ~ study_group + age_at_consent + nih_sex + diabetes_diagnosis, family = "binomial", data = EG_patient_char)
summary(eg_logreg)

# fisher ex test for difference in treatment groups
outcome180 = fisher.test(EG_prop$hyperglycemia, EG_prop$study_group)


###########################
# 180-250 threshold ########### 
EG_prop180250 = EG_patient_char %>% group_by(`Patient ID`) %>% summarise("hyperglycemia180250" = max(hyperglycemia180250))
mean(EG_prop180250$hyperglycemia180250, na.rm = TRUE) #.52 got hg at 250 criteria
EG_patient_char %>% group_by(study_group) %>% summarise("hyperglycemia180250" = mean(hyperglycemia180250, na.rm = TRUE)) # .50 early vs .54 late
# fisher exact test
outcome180250 = fisher.test(EG_patient_char$hyperglycemia180250, EG_patient_char$study_group) # p = 0.785
###########################

# 250-300 threshold ########### 
EG_prop250300 = EG_patient_char %>% group_by(`Patient ID`) %>% summarise("hyperglycemia250300" = max(hyperglycemia250300))
mean(EG_prop250300$hyperglycemia250300, na.rm = TRUE) #.34 got hg at 250 criteria
EG_patient_char %>% group_by(study_group) %>% summarise("hyperglycemia250300" = mean(hyperglycemia250300, na.rm = TRUE)) # .31 early vs .38 late
# fisher exact test
outcome180250 = fisher.test(EG_patient_char$hyperglycemia180250, EG_patient_char$study_group) # p = 0.767
###########################

# 300 Threshold ################
mean(EG_prop300$hyperglycemia300, na.rm = TRUE) #.26 got hg at 300 criteria
EG_patient_char %>% group_by(study_group) %>% summarise("hyperglycemia300" = mean(hyperglycemia300, na.rm = TRUE)) # .23 early vs .29 late

# chi sq test 
fisher.test(EG_patient_char$hyperglycemia300, EG_patient_char$study_group) # p = .751
###########################
```

