---
title: "Long-Acting Reversible Contraception (LARC) Analyses for BCH"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    fig-width: 12
    fig-height: 9
    page-layout: full
    theme:
      light: flatly
      dark: darkly
bibliography: /home/timvigers/Documents/Miscellaneous/zotero.bib
csl: /home/timvigers/Documents/GitHub/styles/american-medical-association.csl
editor: source
execute:
  warning: false
---

```{r setup}
#| include: false
library(Hmisc)
library(tidyverse)
library(arsenal)
library(lmerTest)
library(performance)
library(emmeans)
library(gt)
library(patchwork)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/BDC/Janet Snell-Bergeon/CALICO",
  "Windows" = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/BDC/Janet Snell-Bergeon/CALICO",
  "Linux" = "/home/timvigers/OneDrive/Vigers/BDC/Janet Snell-Bergeon/CALICO"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data import}
#| include: false
# Load data
load("./Data_Clean/analysis_data.RData")
# Combine factor levels
levels(df$insur_type) <- c(
  "Public", "Private", "Military", "Other/None", "Other/None"
)
levels(df$combined_race) <- c(
  "African American", "Other", "Asian", "Caucasian", "More than one", "Other",
  "Other"
)
# Ever on IUD and implant
ever_implant <- na.omit(unique(df$record_number[df$cv_medications___10 == "Checked"]))
df$implant_ever <- factor(df$record_number %in% ever_implant,
  levels = c(F, T), labels = c("No", "Yes")
)
ever_iud <- na.omit(unique(df$record_number[df$cv_medications___11 == "Checked"]))
df$iud_ever <- factor(df$record_number %in% ever_iud,
  levels = c(F, T), labels = c("No", "Yes")
)
label(df$implant_ever) <- "Progesterone Implant Ever Used?"
label(df$iud_ever) <- "Progesterone IUD Ever Used?"
# Remove crazy values
df$cv_dbp[df$cv_dbp >= 200] <- NA
df$cv_bmi[df$cv_bmi >= 200] <- NA
df$cv_weight[df$cv_weight >= 1000] <- NA
df$cv_monthssincepcosdx[df$cv_monthssincepcosdx >= 1000] <- NA
# Get time since last visit on LARC
df <- df %>%
  group_by(record_number) %>%
  mutate(
    last_larc = ifelse(larc == "Yes", cv_monthssincepcosdx, NA)
  ) %>%
  fill(last_larc, .direction = "down") %>%
  mutate(
    months_since_last_larc =
      cv_monthssincepcosdx - last_larc
  )
# For those who go off larc, if the visit is within 6 months of last LARC,
# include that value in the post-LARC analysis.
df$larc[df$larc == "No" & df$months_since_last_larc < 6] <- "Yes"
# Remove visits where they are off LARC after first starting
df <- df %>%
  group_by(record_number) %>%
  mutate(
    pre_post_larc =
      first(cv_monthssincepcosdx[larc == "No" & !is.na(last_larc)], na_rm = T)
  )
df$pre_post_larc[is.na(df$pre_post_larc)] <- Inf
df <- df %>% filter(cv_monthssincepcosdx < pre_post_larc | larc_ever == "No")
# Set table options
mycontrols <-
  tableby.control(numeric.stats = c("Nmiss2", "meansd", "medianq1q3", "range"))
# First visit only
cv1 <- df %>%
  group_by(record_number) %>%
  slice_min(redcap_repeat_instance) %>%
  ungroup()
```

# Data cleaning

- Participants were assumed to be on LARC if `cv_medications___10` ("Progesterone implant") or `cv_medications___11` ("Progesterone IUD") were checked at a given visit.

- There are 10 participants who were started on either implant or IUD at a clinic visit, but who don't have the medication listed at the next visit. For the purposes of these analysis, I assumed that the new medications fields should be ignored and that the current medications field better reflects who was actually on LARC.

- Months since PCOS diagnosis was used as the time variable in longitudinal analyses, and was treated as a continuous variable.

# Demographics

Univariate comparisons were performed using t-tests for continuous variables and Pearson’s Chi-squared test for categorical variables.

```{r results='asis'}
#| label: tbl-table-1
#| tbl-cap: Participant demographics by LARC use ever
f <- as.formula(paste0(
  "larc_ever ~ ",
  paste0(c(
    demo_vars, "combined_race",
    "age_first_larc", "time_pcos_to_first_larc", "implant_ever", "iud_ever",
    "time_menarche_to_first_larc", "cv_sexuallyactive", "cv_genderid"
  ), collapse = "+")
))
t1 <- tableby(f, data = cv1, control = mycontrols)
summary(t1, labelTranslations = labels(df), pfootnote = T)
```

# Mixed models

For each continuous outcome, we fit a linear mixed effects model with random effects for participant and time (months from PCOS diagnosis), and an interaction effect between time and LARC status. To evaluate the effect of starting LARC on each outcome, we also fit a linear spline model with years from LARC start as the time variable and a single knot at 0 (LARC start). If models did not converge or produced serious warnings/errors, only the random effect for participant was included.

All mixed models were fit using the `lme4` R package in version `r paste0(R.version$major,".",R.version$minor)`. @rcoreteamLanguageEnvironmentStatistical2024

```{r}
# Make a dataset for spline models
larc_only <- df %>%
  filter(larc_ever == "Yes") %>%
  mutate(
    time_from_first_larc = cv_age - age_first_larc,
    x = (time_from_first_larc > 0) * time_from_first_larc
  ) %>%
  dplyr::select(
    record_number, redcap_repeat_instance, cv_weight, cv_bmi, cv_hirsutism_num,
    cv_hgb, cv_ldl, cv_hdl, cv_sbp, cv_dbp, larc, cv_age,
    cv_monthssincepcosdx, age_first_larc, time_from_first_larc, x
  )
label(larc_only$time_from_first_larc) <- "Time From First LARC"
label(larc_only$x) <- "Slope Change After LARC"
```

## Weight (kg)

### Spaghetti plot

```{r}
#| warning: false
df %>%
  filter(!is.na(larc)) %>%
  ggplot(aes(
    x = cv_monthssincepcosdx, y = cv_weight, group = record_number,
    color = larc
  )) +
  geom_line(alpha = 0.3) +
  labs(color = "On LARC?") +
  theme_bw()
```

### Model results

#### LARC users vs. non-users

```{r}
# Fit mixed model
mod <- lmer(
  cv_weight ~ cv_monthssincepcosdx * larc +
    (1 + cv_monthssincepcosdx | record_number),
  data = df
)
# Model plot
mod_df <- mod@frame
mod_df$ind_pred <- predict(mod)
mod_df$pop_pred <- predict(mod, re.form = NA)
p1 <- ggplot(mod_df, aes(x = cv_monthssincepcosdx)) +
  geom_point(aes(y = cv_weight, group = record_number),
    alpha = 0.05, show.legend = F
  ) +
  geom_line(aes(y = ind_pred, group = record_number),
    alpha = 0.05, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = larc, color = larc)) +
  xlab("Months From PCOS Dx") +
  ylab("Weight (kg)") +
  labs(color = "On LARC?") +
  theme_classic()+
  scale_color_manual(values = c("#E41A1C", "#377EB8"))
p1
# Model summary
mod %>% gtsummary::tbl_regression()
# EMMs
emt <- emtrends(mod, "larc", var = "cv_monthssincepcosdx")
gt(as.data.frame(emt)) %>% fmt_number(decimals = 3)
gt(as.data.frame(pairs(emt))) %>% fmt_number(decimals = 3)
```

#### Change pre- vs. post-LARC (LARC users only)

```{r}
# Fit mixed model with change point
change_mod <- lmer(cv_weight ~ time_from_first_larc + x +
  (1 + time_from_first_larc | record_number), data = larc_only)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
p2 <- ggplot(mod_df, aes(x = time_from_first_larc * 12)) +
  geom_point(aes(y = cv_weight, group = record_number),
    alpha = 0.05, show.legend = F
  ) +
  geom_line(aes(y = ind_pred, group = record_number),
    alpha = 0.05, show.legend = F
  ) +
  geom_line(aes(y = pop_pred)) +
  xlab("Months From LARC Start") +
  ylab("Weight (kg)") +
  theme_classic()
p2
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

#### Combined plots

```{r}
p3=p1 / p2 + plot_layout(axis_titles = "collect")
p3
p4=p1 + p2 + plot_layout(axis_titles = "collect")
p4

ggsave("./Dissemination/Figures/NASPAG_LARC_horizontal.png",
  plot = p3,
  width = 8, height = 6, units = "in"
)
ggsave("./Dissemination/Figures/NASPAG_LARC_vertical.png",
  plot = p4,
  width = 8, height = 6, units = "in"
)
```

## BMI

### Spaghetti plot

```{r}
#| warning: false
df %>%
  filter(!is.na(larc)) %>%
  ggplot(aes(
    x = cv_monthssincepcosdx, y = cv_bmi, group = record_number,
    color = larc
  )) +
  geom_line(alpha = 0.3) +
  labs(color = "On LARC?") +
  theme_bw()
```

### Model results

#### LARC users vs. non-users

```{r}
# Fit mixed model
mod <- lmer(
  cv_bmi ~ cv_monthssincepcosdx * larc +
    (1 + cv_monthssincepcosdx | record_number),
  data = df
)
# Model plot
mod_df <- mod@frame
mod_df$ind_pred <- predict(mod)
mod_df$pop_pred <- predict(mod, re.form = NA)
ggplot(mod_df, aes(x = cv_monthssincepcosdx)) +
  geom_point(aes(y = cv_bmi, group = record_number),
    alpha = 0.05, show.legend = F
  ) +
  geom_line(aes(y = ind_pred, group = record_number),
    alpha = 0.05, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = larc, color = larc)) +
  xlab("Months from PCOS Dx") +
  labs(color = "On LARC?") +
  theme_classic()
# Model summary
mod %>% gtsummary::tbl_regression()
# EMMs
emt <- emtrends(mod, "larc", var = "cv_monthssincepcosdx")
gt(as.data.frame(emt)) %>% fmt_number(decimals = 3)
gt(as.data.frame(pairs(emt))) %>% fmt_number(decimals = 3)
```

#### Change pre- vs. post-LARC (LARC users only)

```{r}
# Fit mixed model
change_mod <- lmer(cv_bmi ~ time_from_first_larc + x +
  (1 + time_from_first_larc | record_number), data = larc_only)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = time_from_first_larc)) +
  geom_point(aes(y = cv_bmi, group = record_number),
    alpha = 0.05, show.legend = F
  ) +
  geom_line(aes(y = ind_pred, group = record_number),
    alpha = 0.05, show.legend = F
  ) +
  geom_line(aes(y = pop_pred)) +
  xlab("Years From LARC Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

## Numerical FGS score

### Spaghetti plot

```{r}
#| warning: false
df %>%
  filter(!is.na(larc)) %>%
  ggplot(aes(
    x = cv_monthssincepcosdx, y = cv_hirsutism_num, group = record_number,
    color = larc
  )) +
  geom_line(alpha = 0.3) +
  labs(color = "On LARC?") +
  theme_bw()
```

### Model results

#### LARC users vs. non-users

```{r}
# Fit mixed model
mod <- lmer(
  cv_hirsutism_num ~ cv_monthssincepcosdx * larc + (1 | record_number),
  data = df
)
# Model plot
mod_df <- mod@frame
mod_df$ind_pred <- predict(mod)
mod_df$pop_pred <- predict(mod, re.form = NA)
ggplot(mod_df, aes(x = cv_monthssincepcosdx)) +
  geom_point(aes(y = cv_hirsutism_num, group = record_number),
    alpha = 0.05, show.legend = F
  ) +
  geom_line(aes(y = ind_pred, group = record_number),
    alpha = 0.05, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = larc, color = larc)) +
  xlab("Months from PCOS Dx") +
  labs(color = "On LARC?") +
  theme_classic()
# Model summary
mod %>% gtsummary::tbl_regression()
# EMMs
emt <- emtrends(mod, "larc", var = "cv_monthssincepcosdx")
gt(as.data.frame(emt)) %>% fmt_number(decimals = 3)
gt(as.data.frame(pairs(emt))) %>% fmt_number(decimals = 3)
```

#### Change pre- vs. post-LARC (LARC users only)

```{r}
# Fit mixed model
change_mod <- lmer(cv_hirsutism_num ~ time_from_first_larc + x +
  (1 + time_from_first_larc | record_number), data = larc_only)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = time_from_first_larc)) +
  geom_point(aes(y = cv_hirsutism_num, group = record_number),
    alpha = 0.05, show.legend = F
  ) +
  geom_line(aes(y = ind_pred, group = record_number),
    alpha = 0.05, show.legend = F
  ) +
  geom_line(aes(y = pop_pred)) +
  xlab("Years From LARC Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

## Hemoglobin (g/dL)

### Spaghetti plot

```{r}
#| warning: false
df %>%
  filter(!is.na(larc)) %>%
  ggplot(aes(
    x = cv_monthssincepcosdx, y = cv_hgb, group = record_number,
    color = larc
  )) +
  geom_line(alpha = 0.3) +
  labs(color = "On LARC?") +
  theme_bw()
```

### Model results

#### LARC users vs. non-users

```{r}
# Fit mixed model
mod <- lmer(cv_hgb ~ cv_monthssincepcosdx * larc + (1 | record_number),
  data = df
)
# Model plot
mod_df <- mod@frame
mod_df$ind_pred <- predict(mod)
mod_df$pop_pred <- predict(mod, re.form = NA)
ggplot(mod_df, aes(x = cv_monthssincepcosdx)) +
  geom_point(aes(y = cv_hgb, group = record_number),
    alpha = 0.05, show.legend = F
  ) +
  geom_line(aes(y = ind_pred, group = record_number),
    alpha = 0.05, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = larc, color = larc)) +
  xlab("Months from PCOS Dx") +
  labs(color = "On LARC?") +
  theme_classic()
# Model summary
mod %>% gtsummary::tbl_regression()
# EMMs
emt <- emtrends(mod, "larc", var = "cv_monthssincepcosdx")
gt(as.data.frame(emt)) %>% fmt_number(decimals = 3)
gt(as.data.frame(pairs(emt))) %>% fmt_number(decimals = 3)
```

#### Change pre- vs. post-LARC (LARC users only)

```{r}
# Fit mixed model
change_mod <- lmer(cv_hgb ~ time_from_first_larc + x + (1 | record_number),
  data = larc_only
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = time_from_first_larc)) +
  geom_point(aes(y = cv_hgb, group = record_number),
    alpha = 0.05, show.legend = F
  ) +
  geom_line(aes(y = ind_pred, group = record_number),
    alpha = 0.05, show.legend = F
  ) +
  geom_line(aes(y = pop_pred)) +
  xlab("Years From LARC Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

## LDL (mg/dL)

### Spaghetti plot

```{r}
#| warning: false
df %>%
  filter(!is.na(larc)) %>%
  ggplot(aes(
    x = cv_monthssincepcosdx, y = cv_ldl, group = record_number,
    color = larc
  )) +
  geom_line(alpha = 0.3) +
  labs(color = "On LARC?") +
  theme_bw()
```

### Model results

#### LARC users vs. non-users

```{r}
# Fit mixed model
mod <- lmer(cv_ldl ~ cv_monthssincepcosdx * larc + (1 | record_number),
  data = df
)
# Model plot
mod_df <- mod@frame
mod_df$ind_pred <- predict(mod)
mod_df$pop_pred <- predict(mod, re.form = NA)
ggplot(mod_df, aes(x = cv_monthssincepcosdx)) +
  geom_point(aes(y = cv_ldl, group = record_number),
    alpha = 0.05, show.legend = F
  ) +
  geom_line(aes(y = ind_pred, group = record_number),
    alpha = 0.05, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = larc, color = larc)) +
  xlab("Months from PCOS Dx") +
  labs(color = "On LARC?") +
  theme_classic()
# Model summary
mod %>% gtsummary::tbl_regression()
# EMMs
emt <- emtrends(mod, "larc", var = "cv_monthssincepcosdx")
gt(as.data.frame(emt)) %>% fmt_number(decimals = 3)
gt(as.data.frame(pairs(emt))) %>% fmt_number(decimals = 3)
```

#### Change pre- vs. post-LARC (LARC users only)

```{r}
# Fit mixed model
change_mod <- lmer(cv_ldl ~ time_from_first_larc + x +
  (1 + time_from_first_larc | record_number), data = larc_only)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = time_from_first_larc)) +
  geom_point(aes(y = cv_ldl, group = record_number),
    alpha = 0.05, show.legend = F
  ) +
  geom_line(aes(y = ind_pred, group = record_number),
    alpha = 0.05, show.legend = F
  ) +
  geom_line(aes(y = pop_pred)) +
  xlab("Years From LARC Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
``` 
           
## HDL (mg/dL)

### Spaghetti plot

```{r}
#| warning: false
df %>%
  filter(!is.na(larc)) %>%
  ggplot(aes(
    x = cv_monthssincepcosdx, y = cv_hdl, group = record_number,
    color = larc
  )) +
  geom_line(alpha = 0.3) +
  labs(color = "On LARC?") +
  theme_bw()
```

### Model results

#### LARC users vs. non-users

```{r}
# Fit mixed model
mod <- lmer(cv_hdl ~ cv_monthssincepcosdx * larc + (1 | record_number),
  data = df
)
# Model plot
mod_df <- mod@frame
mod_df$ind_pred <- predict(mod)
mod_df$pop_pred <- predict(mod, re.form = NA)
ggplot(mod_df, aes(x = cv_monthssincepcosdx)) +
  geom_point(aes(y = cv_hdl, group = record_number),
    alpha = 0.05, show.legend = F
  ) +
  geom_line(aes(y = ind_pred, group = record_number),
    alpha = 0.05, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = larc, color = larc)) +
  xlab("Months from PCOS Dx") +
  labs(color = "On LARC?") +
  theme_classic()
# Model summary
mod %>% gtsummary::tbl_regression()
# EMMs
emt <- emtrends(mod, "larc", var = "cv_monthssincepcosdx")
gt(as.data.frame(emt)) %>% fmt_number(decimals = 3)
gt(as.data.frame(pairs(emt))) %>% fmt_number(decimals = 3)
```

#### Change pre- vs. post-LARC (LARC users only)

```{r}
# Fit mixed model
change_mod <- lmer(cv_hdl ~ time_from_first_larc + x +
  (1 | record_number), data = larc_only)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = time_from_first_larc)) +
  geom_point(aes(y = cv_hdl, group = record_number),
    alpha = 0.05, show.legend = F
  ) +
  geom_line(aes(y = ind_pred, group = record_number),
    alpha = 0.05, show.legend = F
  ) +
  geom_line(aes(y = pop_pred)) +
  xlab("Years From LARC Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
``` 
             
## Systolic BP (mmHg)

### Spaghetti plot

```{r}
#| warning: false
df %>%
  filter(!is.na(larc)) %>%
  ggplot(aes(
    x = cv_monthssincepcosdx, y = cv_sbp, group = record_number,
    color = larc
  )) +
  geom_line(alpha = 0.3) +
  labs(color = "On LARC?") +
  theme_bw()
```

### Model results

#### LARC users vs. non-users

```{r}
# Fit mixed model
mod <- lmer(cv_sbp ~ cv_monthssincepcosdx * larc + (1 | record_number),
  data = df
)
# Model plot
mod_df <- mod@frame
mod_df$ind_pred <- predict(mod)
mod_df$pop_pred <- predict(mod, re.form = NA)
ggplot(mod_df, aes(x = cv_monthssincepcosdx)) +
  geom_point(aes(y = cv_sbp, group = record_number),
    alpha = 0.05, show.legend = F
  ) +
  geom_line(aes(y = ind_pred, group = record_number),
    alpha = 0.05, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = larc, color = larc)) +
  xlab("Months from PCOS Dx") +
  labs(color = "On LARC?") +
  theme_classic()
# Model summary
mod %>% gtsummary::tbl_regression()
# EMMs
emt <- emtrends(mod, "larc", var = "cv_monthssincepcosdx")
gt(as.data.frame(emt)) %>% fmt_number(decimals = 3)
gt(as.data.frame(pairs(emt))) %>% fmt_number(decimals = 3)
```

#### Change pre- vs. post-LARC (LARC users only)

```{r}
# Fit mixed model
change_mod <- lmer(cv_sbp ~ time_from_first_larc + x +
  (1 | record_number), data = larc_only)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = time_from_first_larc)) +
  geom_point(aes(y = cv_sbp, group = record_number),
    alpha = 0.05, show.legend = F
  ) +
  geom_line(aes(y = ind_pred, group = record_number),
    alpha = 0.05, show.legend = F
  ) +
  geom_line(aes(y = pop_pred)) +
  xlab("Years From LARC Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```      
     
## Diastolic BP (mmHg)

### Spaghetti plot

```{r}
#| warning: false
df %>%
  filter(!is.na(larc)) %>%
  ggplot(aes(
    x = cv_monthssincepcosdx, y = cv_dbp, group = record_number,
    color = larc
  )) +
  geom_line(alpha = 0.3) +
  labs(color = "On LARC?") +
  theme_bw()
```

### Model results

#### LARC users vs. non-users

```{r}
# Fit mixed model
mod <- lmer(cv_dbp ~ cv_monthssincepcosdx * larc + (1 | record_number),
  data = df
)
# Model plot
mod_df <- mod@frame
mod_df$ind_pred <- predict(mod)
mod_df$pop_pred <- predict(mod, re.form = NA)
ggplot(mod_df, aes(x = cv_monthssincepcosdx)) +
  geom_point(aes(y = cv_dbp, group = record_number),
    alpha = 0.05, show.legend = F
  ) +
  geom_line(aes(y = ind_pred, group = record_number),
    alpha = 0.05, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = larc, color = larc)) +
  xlab("Months from PCOS Dx") +
  labs(color = "On LARC?") +
  theme_classic()
# Model summary
mod %>% gtsummary::tbl_regression()
# EMMs
emt <- emtrends(mod, "larc", var = "cv_monthssincepcosdx")
gt(as.data.frame(emt)) %>% fmt_number(decimals = 3)
gt(as.data.frame(pairs(emt))) %>% fmt_number(decimals = 3)
```

#### Change pre- vs. post-LARC (LARC users only)

```{r}
# Fit mixed model
change_mod <- lmer(cv_dbp ~ time_from_first_larc + x +
  (1 + time_from_first_larc | record_number), data = larc_only)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = time_from_first_larc)) +
  geom_point(aes(y = cv_dbp, group = record_number),
    alpha = 0.05, show.legend = F
  ) +
  geom_line(aes(y = ind_pred, group = record_number),
    alpha = 0.05, show.legend = F
  ) +
  geom_line(aes(y = pop_pred)) +
  xlab("Years From LARC Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```
