---
title: "Guideline Utilization in CALICO"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    fig-width: 12
    fig-height: 9
    page-layout: full
    grid:
      body-width: 2500px
    html-table-processing: none
editor: source
---

```{r setup}
#| include: false
library(Hmisc)
library(tidyverse)
library(gtsummary)
library(gt)
library(plotly)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/BDC/Janet Snell-Bergeon/CALICO",
  "Linux" = "/home/timvigers/OneDrive/Vigers/BDC/Janet Snell-Bergeon/CALICO"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data import}
#| include: false
# Load data
load("./Data_Clean/analysis_data.RData")
# Provider specialty levels
levels(df$pcosdx_specialty) <- c(
  "Endocrinology", "Primary Care (Pediatrics or Family Medicine)", "Gynecology",
  "Other", "Adolescent Medicine", "Other"
)
# Insurance as public, private, none/other
levels(df$insur_type) <- c("Public", "Private", "Military/None/Other", "Military/None/Other", "Military/None/Other")
# Recode specialty
df$pcosdx_specialty_combined <- fct_recode(df$pcosdx_specialty,
  "Primary Care (Pediatrics or Family Medicine)/Other" =
    "Primary Care (Pediatrics or Family Medicine)",
  "Primary Care (Pediatrics or Family Medicine)/Other" = "Other",
  "Gynecology/Adolescent Medicine" = "Gynecology",
  "Gynecology/Adolescent Medicine" = "Adolescent Medicine"
)
# Year of diagnosis
df$dx_year <- year(df$pcosdx_date)
# Lists of variables
diagnostic <- c(
  "cv_17ohp", "cv_ft", "cv_tt", "cv_tsh", "cv_hdl", "cv_ldl", "cv_ast",
  "cv_alt", "cv_liverimaging_results", "cv_hirsutism_num", "cv_hirsutism_cat",
  "cv_acneface", "cv_alopecia", "cv_osa_sx"
)
diagnostic_check <- c(
  "labs_this_visit___4", "labs_this_visit___5", "labs_this_visit___10",
  "labs_this_visit___11", "labs_this_visit___12", "labs_this_visit___33",
  "cv_acneother___0", "cv_acneother___1", "cv_acneother___2",
  paste0("cv_referrals___", c(1:9, 11, 12)),
  paste0("cv_newmeds___", 1:31)
)
demo_vars <- c(
  "site", "pcosdx_age", "combined_race", "ethnicity", "insur_type",
  "weight_cat", "pcosdx_specialty", "pcosdx_specialty_combined"
)
interact_tests <- c("cv_newmeds___1")
# Make 1 year dataset
one_year <- df %>%
  group_by(record_number) %>%
  filter(cv_monthssincepcosdx <= 12) %>%
  summarise(
    across(all_of(demo_vars), ~ first(.x, na_rm = F)),
    across(all_of(diagnostic_check), ~ any(.x == "Checked")),
    across(all_of(diagnostic), ~ any(!is.na(.x))),
    cv_mood =
      any(cv_mood %in% c("Yes", "Yes, asked by clinician, but no scale used.")),
    `Any glycemia measure` =
      any(!is.na(cv_a1c) | !is.na(cv_fbg) | !is.na(cv_2hrglucoseogtt)),
    cv_a1c = any(!is.na(cv_a1c)),
    cv_fbg = any(!is.na(cv_fbg)),
    cv_2hrglucoseogtt = any(!is.na(cv_2hrglucoseogtt)),
    cv_dietarycounseling =
      any(cv_dietarycounseling %in% c("Yes", "No, but referred to a dietitian")),
    cv_exerciseplan =
      any(cv_exerciseplan %in%
        c("Yes", "No, but referred to a exercise specialist")),
    EC = any(cv_medications___5 == "Checked") |
      any(cv_medications___6 == "Checked") |
      any(cv_medications___7 == "Checked"),
    cv_medications___8 = any(cv_medications___8 == "Checked"),
    `Any other progesterone` =
      any(cv_medications___9 == "Checked") |
        any(cv_medications___10 == "Checked") |
        any(cv_medications___11 == "Checked") |
        any(cv_medications___12 == "Checked"),
    `Weight loss medication` =
      any(cv_medications___25 == "Checked") |
        any(cv_medications___26 == "Checked") |
        any(cv_medications___27 == "Checked") |
        any(cv_medications___28 == "Checked") |
        any(cv_medications___29 == "Checked"),
    `Acne medication` =
      any(cv_medications___13 == "Checked") |
        any(cv_medications___14 == "Checked") |
        any(cv_medications___15 == "Checked"),
    cv_medications___16 = any(cv_medications___16 == "Checked"),
    `Mood medication` =
      any(cv_medications___18 == "Checked") |
        any(cv_medications___19 == "Checked"),
    cv_medications___19 = any(cv_medications___19 == "Checked")
  )
# Ever vs. never
ever_never <- df %>%
  group_by(record_number) %>%
  summarise(
    across(all_of(demo_vars), ~ first(.x, na_rm = F)),
    across(all_of(diagnostic_check), ~ any(.x == "Checked")),
    across(all_of(diagnostic), ~ any(!is.na(.x))),
    cv_mood =
      any(cv_mood %in% c("Yes", "Yes, asked by clinician, but no scale used.")),
    `Any glycemia measure` =
      any(!is.na(cv_a1c) | !is.na(cv_fbg) | !is.na(cv_2hrglucoseogtt)),
    cv_a1c = any(!is.na(cv_a1c)),
    cv_fbg = any(!is.na(cv_fbg)),
    cv_2hrglucoseogtt = any(!is.na(cv_2hrglucoseogtt)),
    cv_dietarycounseling =
      any(cv_dietarycounseling %in% c("Yes", "No, but referred to a dietitian")),
    cv_exerciseplan =
      any(cv_exerciseplan %in%
        c("Yes", "No, but referred to a exercise specialist")),
    EC = any(cv_medications___5 == "Checked") |
      any(cv_medications___6 == "Checked") |
      any(cv_medications___7 == "Checked"),
    cv_medications___8 = any(cv_medications___8 == "Checked"),
    `Any other progesterone` =
      any(cv_medications___9 == "Checked") |
        any(cv_medications___10 == "Checked") |
        any(cv_medications___11 == "Checked") |
        any(cv_medications___12 == "Checked"),
    `Weight loss medication` =
      any(cv_medications___25 == "Checked") |
        any(cv_medications___26 == "Checked") |
        any(cv_medications___27 == "Checked") |
        any(cv_medications___28 == "Checked") |
        any(cv_medications___29 == "Checked"),
    `Acne medication` =
      any(cv_medications___13 == "Checked") |
        any(cv_medications___14 == "Checked") |
        any(cv_medications___15 == "Checked"),
    cv_medications___16 = any(cv_medications___16 == "Checked"),
    `Mood medication` =
      any(cv_medications___18 == "Checked") |
        any(cv_medications___19 == "Checked"),
    cv_medications___19 = any(cv_medications___19 == "Checked")
  )
# Diagnosis year (2 year increments)
dx_2008_2009 <- df %>%
  group_by(record_number) %>%
  filter(dx_year %in% 2008:2009, cv_monthssincepcosdx == 0) %>%
  summarise(
    across(all_of(demo_vars), ~ first(.x, na_rm = F)),
    across(all_of(diagnostic_check), ~ any(.x == "Checked")),
    across(all_of(diagnostic), ~ any(!is.na(.x))),
    cv_mood =
      any(cv_mood %in% c("Yes", "Yes, asked by clinician, but no scale used.")),
    `Any glycemia measure` =
      any(!is.na(cv_a1c) | !is.na(cv_fbg) | !is.na(cv_2hrglucoseogtt)),
    cv_a1c = any(!is.na(cv_a1c)),
    cv_fbg = any(!is.na(cv_fbg)),
    cv_2hrglucoseogtt = any(!is.na(cv_2hrglucoseogtt)),
    cv_dietarycounseling =
      any(cv_dietarycounseling %in% c("Yes", "No, but referred to a dietitian")),
    cv_exerciseplan =
      any(cv_exerciseplan %in%
        c("Yes", "No, but referred to a exercise specialist")),
    EC = any(cv_medications___5 == "Checked") |
      any(cv_medications___6 == "Checked") |
      any(cv_medications___7 == "Checked"),
    cv_medications___8 = any(cv_medications___8 == "Checked"),
    `Any other progesterone` =
      any(cv_medications___9 == "Checked") |
        any(cv_medications___10 == "Checked") |
        any(cv_medications___11 == "Checked") |
        any(cv_medications___12 == "Checked"),
    `Weight loss medication` =
      any(cv_medications___25 == "Checked") |
        any(cv_medications___26 == "Checked") |
        any(cv_medications___27 == "Checked") |
        any(cv_medications___28 == "Checked") |
        any(cv_medications___29 == "Checked"),
    `Acne medication` =
      any(cv_medications___13 == "Checked") |
        any(cv_medications___14 == "Checked") |
        any(cv_medications___15 == "Checked"),
    cv_medications___16 = any(cv_medications___16 == "Checked"),
    `Mood medication` =
      any(cv_medications___18 == "Checked") |
        any(cv_medications___19 == "Checked"),
    cv_medications___19 = any(cv_medications___19 == "Checked")
  )
dx_2010_2011 <- df %>%
  group_by(record_number) %>%
  filter(dx_year %in% 2010:2011, cv_monthssincepcosdx == 0) %>%
  summarise(
    across(all_of(demo_vars), ~ first(.x, na_rm = F)),
    across(all_of(diagnostic_check), ~ any(.x == "Checked")),
    across(all_of(diagnostic), ~ any(!is.na(.x))),
    cv_mood =
      any(cv_mood %in% c("Yes", "Yes, asked by clinician, but no scale used.")),
    `Any glycemia measure` =
      any(!is.na(cv_a1c) | !is.na(cv_fbg) | !is.na(cv_2hrglucoseogtt)),
    cv_a1c = any(!is.na(cv_a1c)),
    cv_fbg = any(!is.na(cv_fbg)),
    cv_2hrglucoseogtt = any(!is.na(cv_2hrglucoseogtt)),
    cv_dietarycounseling =
      any(cv_dietarycounseling %in% c("Yes", "No, but referred to a dietitian")),
    cv_exerciseplan =
      any(cv_exerciseplan %in%
        c("Yes", "No, but referred to a exercise specialist")),
    EC = any(cv_medications___5 == "Checked") |
      any(cv_medications___6 == "Checked") |
      any(cv_medications___7 == "Checked"),
    cv_medications___8 = any(cv_medications___8 == "Checked"),
    `Any other progesterone` =
      any(cv_medications___9 == "Checked") |
        any(cv_medications___10 == "Checked") |
        any(cv_medications___11 == "Checked") |
        any(cv_medications___12 == "Checked"),
    `Weight loss medication` =
      any(cv_medications___25 == "Checked") |
        any(cv_medications___26 == "Checked") |
        any(cv_medications___27 == "Checked") |
        any(cv_medications___28 == "Checked") |
        any(cv_medications___29 == "Checked"),
    `Acne medication` =
      any(cv_medications___13 == "Checked") |
        any(cv_medications___14 == "Checked") |
        any(cv_medications___15 == "Checked"),
    cv_medications___16 = any(cv_medications___16 == "Checked"),
    `Mood medication` =
      any(cv_medications___18 == "Checked") |
        any(cv_medications___19 == "Checked"),
    cv_medications___19 = any(cv_medications___19 == "Checked")
  )
dx_2012_2013 <- df %>%
  group_by(record_number) %>%
  filter(dx_year %in% 2012:2013, cv_monthssincepcosdx == 0) %>%
  summarise(
    across(all_of(demo_vars), ~ first(.x, na_rm = F)),
    across(all_of(diagnostic_check), ~ any(.x == "Checked")),
    across(all_of(diagnostic), ~ any(!is.na(.x))),
    cv_mood =
      any(cv_mood %in% c("Yes", "Yes, asked by clinician, but no scale used.")),
    `Any glycemia measure` =
      any(!is.na(cv_a1c) | !is.na(cv_fbg) | !is.na(cv_2hrglucoseogtt)),
    cv_a1c = any(!is.na(cv_a1c)),
    cv_fbg = any(!is.na(cv_fbg)),
    cv_2hrglucoseogtt = any(!is.na(cv_2hrglucoseogtt)),
    cv_dietarycounseling =
      any(cv_dietarycounseling %in% c("Yes", "No, but referred to a dietitian")),
    cv_exerciseplan =
      any(cv_exerciseplan %in%
        c("Yes", "No, but referred to a exercise specialist")),
    EC = any(cv_medications___5 == "Checked") |
      any(cv_medications___6 == "Checked") |
      any(cv_medications___7 == "Checked"),
    cv_medications___8 = any(cv_medications___8 == "Checked"),
    `Any other progesterone` =
      any(cv_medications___9 == "Checked") |
        any(cv_medications___10 == "Checked") |
        any(cv_medications___11 == "Checked") |
        any(cv_medications___12 == "Checked"),
    `Weight loss medication` =
      any(cv_medications___25 == "Checked") |
        any(cv_medications___26 == "Checked") |
        any(cv_medications___27 == "Checked") |
        any(cv_medications___28 == "Checked") |
        any(cv_medications___29 == "Checked"),
    `Acne medication` =
      any(cv_medications___13 == "Checked") |
        any(cv_medications___14 == "Checked") |
        any(cv_medications___15 == "Checked"),
    cv_medications___16 = any(cv_medications___16 == "Checked"),
    `Mood medication` =
      any(cv_medications___18 == "Checked") |
        any(cv_medications___19 == "Checked"),
    cv_medications___19 = any(cv_medications___19 == "Checked")
  )
```

# Within 1 year of diagnosis

## Data cleaning and methods

For the following analysis, we filtered the dataset to visits where `cv_monthssincepcosdx <= 12`. Then, we grouped the data by participant ID and checked whether any of the labs were performed at any visit (for checkbox variables we looked for any visit marked "Checked" and for continuous variables we looked for any non-missing values). Participants who had the test performed were marked "TRUE" and others marked "FALSE."

`r sum(is.na(one_year$pcosdx_specialty))` participants were missing the `pcosdx_specialty` field.

`r sum(is.na(one_year$weight_cat))` participants were missing weight category.

All q values are FDR-adjusted for multiple comparisons.

## By specialty type at diagnosis

```{r}
one_year %>%
  select(-record_number, -site, -weight_cat) %>%
  tbl_summary(by = pcosdx_specialty, label = as.list(label(df))) %>%
  add_overall() %>%
  add_p(
    test.args = all_tests("fisher.test") ~ list(simulate.p.value = TRUE)
  ) %>%
  separate_p_footnotes() %>%
  add_q() %>%
  modify_footnote(everything() ~ NA)
```

```{r}
p <- one_year %>%
  group_by(pcosdx_specialty) %>%
  summarise(across(all_of(c(diagnostic, diagnostic_check)), ~ mean(.x, na.rm = T))) %>%
  filter(!is.na(pcosdx_specialty)) %>%
  pivot_longer(all_of(c(diagnostic, diagnostic_check))) %>%
  ggplot(., aes(x = reorder(name, -value), y = value)) +
  geom_bar(stat = "identity") +
  facet_wrap(~pcosdx_specialty) +
  theme_bw() +
  theme(
    axis.title.x = element_blank(), axis.text.x = element_blank(),
    axis.title.y = element_blank()
  )
ggplotly(p)
```

## By weight category at diagnosis

```{r}
one_year %>%
  select(-record_number, -site, -pcosdx_specialty) %>%
  tbl_summary(by = weight_cat, label = as.list(label(df))) %>%
  add_overall() %>%
  add_p(
    test.args = all_tests("fisher.test") ~ list(simulate.p.value = TRUE)
  ) %>%
  separate_p_footnotes() %>%
  add_q() %>%
  modify_footnote(everything() ~ NA)
```

## Interaction between specialty type and weight category at diagnosis

```{r results='asis'}
f <- cv_newmeds___1~pcosdx_specialty_combined * weight_cat
m <- glm(f, data = one_year, family = "binomial")
m_check = m$model
m_check$pred = predict(m,type = "response")

table(df$pcosdx_specialty_combined, df$weight_cat, df$cv_newmeds___1)
```

The `cv_newmeds___1~pcosdx_specialty_combined * weight_cat` model is producing numeric probabilities that are basically 0 because no one in the Gynecology/Adolescent Medicine and Overweight group were on metformin.

## By insurance type at diagnosis

```{r}
one_year %>%
  select(-record_number, -site, -pcosdx_specialty) %>%
  tbl_summary(by = insur_type, label = as.list(label(df))) %>%
  add_overall() %>%
  add_p(
    test.args = all_tests("fisher.test") ~ list(simulate.p.value = TRUE)
  ) %>%
  separate_p_footnotes() %>%
  add_q() %>%
  modify_footnote(everything() ~ NA)
```

<!-- # Ever vs. never -->

<!-- ## Data cleaning and methods -->

<!-- For the following analysis, we performed similar data cleaning to the above analyses, but did not limit the dataset to `cv_monthssincepcosdx <= 12`. -->

<!-- All q values are FDR-adjusted for multiple comparisons. -->

<!-- ## By specialty type at diagnosis -->

<!-- ```{r} -->
<!-- ever_never %>% -->
<!--   select(-record_number, -site, -weight_cat) %>% -->
<!--   tbl_summary(by = pcosdx_specialty, label = as.list(label(df))) %>% -->
<!--   add_overall() %>% -->
<!--   add_p( -->
<!--     test.args = all_tests("fisher.test") ~ list(simulate.p.value = TRUE) -->
<!--   ) %>% -->
<!--   separate_p_footnotes() %>% -->
<!--   add_q() %>% -->
<!--   modify_footnote(everything() ~ NA) %>% -->
<!--   as_gt() %>% -->
<!--   opt_interactive() -->
<!-- ``` -->

<!-- ## By weight category at diagnosis -->

<!-- ```{r} -->
<!-- ever_never %>% -->
<!--   select(-record_number, -site, -pcosdx_specialty) %>% -->
<!--   tbl_summary(by = weight_cat, label = as.list(label(df))) %>% -->
<!--   add_overall() %>% -->
<!--   add_p( -->
<!--     test.args = all_tests("fisher.test") ~ list(simulate.p.value = TRUE) -->
<!--   ) %>% -->
<!--   separate_p_footnotes() %>% -->
<!--   add_q() %>% -->
<!--   modify_footnote(everything() ~ NA) %>% -->
<!--   as_gt() %>% -->
<!--   opt_interactive() -->
<!-- ``` -->

<!-- ## By specialty type and weight category at diagnosis -->

<!-- ```{r} -->
<!-- ever_never %>% -->
<!--   mutate(pcosdx_specialty = fct_recode(one_year$pcosdx_specialty, -->
<!--     "Primary Care (Pediatrics or Family Medicine)/Other" = -->
<!--       "Primary Care (Pediatrics or Family Medicine)", -->
<!--     "Primary Care (Pediatrics or Family Medicine)/Other" = "Other", -->
<!--     "Gynecology/Adolescent Medicine" = "Gynecology", -->
<!--     "Gynecology/Adolescent Medicine" = "Adolescent Medicine" -->
<!--   )) %>% -->
<!--   select(-record_number, -site) %>% -->
<!--   filter(!is.na(pcosdx_specialty)) %>% -->
<!--   tbl_strata( -->
<!--     strata = pcosdx_specialty, -->
<!--     .tbl_fun = -->
<!--       ~ .x |> -->
<!--         tbl_summary(by = weight_cat, label = as.list(label(df))) %>% -->
<!--         add_overall() %>% -->
<!--         add_p(test.args = all_tests("fisher.test") ~ -->
<!--           list(simulate.p.value = TRUE)) %>% -->
<!--         separate_p_footnotes() %>% -->
<!--         add_q() %>% -->
<!--         modify_footnote(everything() ~ NA), -->
<!--     .header = "**{strata}**, N = {n}" -->
<!--   ) %>% -->
<!--   as_gt() %>% -->
<!--   opt_interactive() -->
<!-- ``` -->

<!-- ## By insurance type at diagnosis -->

<!-- ```{r} -->
<!-- ever_never %>% -->
<!--   select(-record_number, -site, -pcosdx_specialty) %>% -->
<!--   tbl_summary(by = insur_type, label = as.list(label(df))) %>% -->
<!--   add_overall() %>% -->
<!--   add_p( -->
<!--     test.args = all_tests("fisher.test") ~ list(simulate.p.value = TRUE) -->
<!--   ) %>% -->
<!--   separate_p_footnotes() %>% -->
<!--   add_q() %>% -->
<!--   modify_footnote(everything() ~ NA) %>% -->
<!--   as_gt() %>% -->
<!--   opt_interactive() -->
<!-- ``` -->

<!-- # By year of diagnosis -->

<!-- ## 2008 or 2009 -->

<!-- ### By specialty type at diagnosis -->

<!-- ```{r} -->
<!-- dx_2008_2009 %>% -->
<!--   select(-record_number, -site, -weight_cat) %>% -->
<!--   tbl_summary(by = pcosdx_specialty, label = as.list(label(df))) %>% -->
<!--   add_overall() %>% -->
<!--   add_p( -->
<!--     test.args = all_tests("fisher.test") ~ list(simulate.p.value = TRUE) -->
<!--   ) %>% -->
<!--   separate_p_footnotes() %>% -->
<!--   add_q() %>% -->
<!--   modify_footnote(everything() ~ NA) %>% -->
<!--   as_gt() %>% -->
<!--   opt_interactive() -->
<!-- ``` -->

<!-- ### By weight category at diagnosis -->

<!-- ```{r} -->
<!-- dx_2008_2009 %>% -->
<!--   select(-record_number, -site, -pcosdx_specialty) %>% -->
<!--   tbl_summary(by = weight_cat, label = as.list(label(df))) %>% -->
<!--   add_overall() %>% -->
<!--   add_p( -->
<!--     test.args = all_tests("fisher.test") ~ list(simulate.p.value = TRUE) -->
<!--   ) %>% -->
<!--   separate_p_footnotes() %>% -->
<!--   add_q() %>% -->
<!--   modify_footnote(everything() ~ NA) %>% -->
<!--   as_gt() %>% -->
<!--   opt_interactive() -->
<!-- ``` -->

<!-- ### By specialty type and weight category at diagnosis -->

<!-- ```{r} -->
<!-- dx_2008_2009 %>% -->
<!--   mutate(pcosdx_specialty = fct_recode(one_year$pcosdx_specialty, -->
<!--     "Primary Care (Pediatrics or Family Medicine)/Other" = -->
<!--       "Primary Care (Pediatrics or Family Medicine)", -->
<!--     "Primary Care (Pediatrics or Family Medicine)/Other" = "Other", -->
<!--     "Gynecology/Adolescent Medicine" = "Gynecology", -->
<!--     "Gynecology/Adolescent Medicine" = "Adolescent Medicine" -->
<!--   )) %>% -->
<!--   select(-record_number, -site) %>% -->
<!--   filter(!is.na(pcosdx_specialty)) %>% -->
<!--   tbl_strata( -->
<!--     strata = pcosdx_specialty, -->
<!--     .tbl_fun = -->
<!--       ~ .x |> -->
<!--         tbl_summary(by = weight_cat, label = as.list(label(df))) %>% -->
<!--         add_overall() %>% -->
<!--         add_p(test.args = all_tests("fisher.test") ~ -->
<!--           list(simulate.p.value = TRUE)) %>% -->
<!--         separate_p_footnotes() %>% -->
<!--         add_q() %>% -->
<!--         modify_footnote(everything() ~ NA), -->
<!--     .header = "**{strata}**, N = {n}" -->
<!--   ) %>% -->
<!--   as_gt() %>% -->
<!--   opt_interactive() -->
<!-- ``` -->

<!-- ### By insurance type at diagnosis -->

<!-- ```{r} -->
<!-- dx_2008_2009 %>% -->
<!--   select(-record_number, -site, -pcosdx_specialty) %>% -->
<!--   tbl_summary(by = insur_type, label = as.list(label(df))) %>% -->
<!--   add_overall() %>% -->
<!--   add_p( -->
<!--     test.args = all_tests("fisher.test") ~ list(simulate.p.value = TRUE) -->
<!--   ) %>% -->
<!--   separate_p_footnotes() %>% -->
<!--   add_q() %>% -->
<!--   modify_footnote(everything() ~ NA) %>% -->
<!--   as_gt() %>% -->
<!--   opt_interactive() -->
<!-- ``` -->
