---
title: "Guideline Utilization in CALICO"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    fig-width: 12
    fig-height: 9
    page-layout: full
    theme:
      light: flatly
      dark: darkly
bibliography: /home/timvigers/Documents/Miscellaneous/zotero.bib
csl: /home/timvigers/Documents/GitHub/styles/american-medical-association.csl
editor: source
execute:
  warning: false
---

```{r setup}
#| include: false
library(Hmisc)
library(tidyverse)
library(gtsummary)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/BDC/Janet Snell-Bergeon/CALICO",
  "Linux" = "/home/timvigers/OneDrive/Vigers/BDC/Janet Snell-Bergeon/CALICO"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data import}
#| include: false
# Load data
load("./Data_Clean/analysis_data.RData")
# Lists of variables
diagnostic <- c(
  "cv_17ohp", "cv_ft", "cv_tt", "cv_tsh", "cv_hdl", "cv_ldl", "cv_ast",
  "cv_alt", "cv_liverimaging_results", "cv_hirsutism_num", "cv_hirsutism_cat",
  "cv_acneface", "cv_alopecia"
)
diagnostic_check <- c(
  "labs_this_visit___4", "labs_this_visit___5", "labs_this_visit___10",
  "labs_this_visit___11", "labs_this_visit___12", "labs_this_visit___33",
  "cv_acneother___0", "cv_acneother___1", "cv_acneother___2",
  "cv_osa_sx", paste0("cv_referrals___", c(1:9, 11, 12)),
  paste0("cv_newmeds___", 1:31)
)
demo_vars <- c(demo_vars, "weight_cat", "pcosdx_specialty")
# Provider specialty levels
levels(df$pcosdx_specialty) <- c(
  "Endocrinology", "Primary Care (Pediatrics or Family Medicine)", "Gynecology",
  "Other", "Adolescent Medicine", "Other"
)
# Make 1 year dataset
one_year <- df %>%
  group_by(record_number) %>%
  filter(cv_monthssincepcosdx <= 12) %>%
  summarise(
    across(all_of(demo_vars), ~ first(.x, na_rm = F)),
    across(all_of(diagnostic_check), ~ any(.x == "Checked")),
    across(all_of(diagnostic), ~ any(!is.na(.x))),
    cv_mood =
      any(cv_mood %in% c("Yes", "Yes, asked by clinician, but no scale used.")),
    `Any glycemia measure` =
      any(!is.na(cv_a1c) | !is.na(cv_fbg) | !is.na(cv_2hrglucoseogtt)),
    cv_a1c = any(!is.na(cv_a1c)),
    cv_fbg = any(!is.na(cv_fbg)),
    cv_2hrglucoseogtt = any(!is.na(cv_2hrglucoseogtt)),
    cv_dietarycounseling =
      any(cv_dietarycounseling %in% c("Yes", "No, but referred to a dietitian")),
    cv_exerciseplan =
      any(cv_exerciseplan %in%
        c("Yes", "No, but referred to a exercise specialist")),
    EC = any(cv_medications___5 == "Checked") |
      any(cv_medications___6 == "Checked") |
      any(cv_medications___7 == "Checked"),
    cv_medications___8 = any(cv_medications___8 == "Checked"),
    `Any other progesterone` =
      any(cv_medications___9 == "Checked") |
        any(cv_medications___10 == "Checked") |
        any(cv_medications___11 == "Checked") |
        any(cv_medications___12 == "Checked"),
    `Weight loss medication` =
      any(cv_medications___25 == "Checked") |
        any(cv_medications___26 == "Checked") |
        any(cv_medications___27 == "Checked") |
        any(cv_medications___28 == "Checked") |
        any(cv_medications___29 == "Checked"),
    `Acne medication` =
      any(cv_medications___13 == "Checked") |
        any(cv_medications___14 == "Checked") |
        any(cv_medications___15 == "Checked"),
    cv_medications___16 = any(cv_medications___16 == "Checked"),
    `Mood medication` =
      any(cv_medications___18 == "Checked") |
        any(cv_medications___19 == "Checked"),
    cv_medications___19 = any(cv_medications___19 == "Checked")
  )
```

# Data cleaning and methods

For the following analysis, we filtered the dataset to visits where `cv_monthssincepcosdx <= 12`. Then, we grouped the data by participant ID and checked whether any of the labs were performed at any visit (for checkbox variables we looked for any visit marked "Checked" and for continuous variables we looked for any non-missing values). Participants who had the test performed were marked "TRUE" and others marked "FALSE."

`r sum(is.na(one_year$pcosdx_specialty))` participants were missing the `pcosdx_specialty` field.

`r sum(is.na(one_year$weight_cat))` participants were missing weight category.

All q values are FDR-adjusted for multiple comparisons.

# Within 1 year of diagnosis

## By specialty type

```{r}
one_year %>%
  select(-record_number, -site, -weight_cat) %>%
  tbl_summary(by = pcosdx_specialty, label = as.list(label(df))) %>%
  add_overall() %>%
  add_p(
    test.args = all_tests("fisher.test") ~ list(simulate.p.value = TRUE)
  ) %>%
  separate_p_footnotes() %>%
  add_q() %>%
  modify_footnote(everything() ~ NA)
```

## By weight category

```{r}
one_year %>%
  select(-record_number, -site, -pcosdx_specialty) %>%
  tbl_summary(by = weight_cat, label = as.list(label(df))) %>%
  add_overall() %>%
  add_p(
    test.args = all_tests("fisher.test") ~ list(simulate.p.value = TRUE)
  ) %>%
  separate_p_footnotes() %>%
  add_q() %>%
  modify_footnote(everything() ~ NA)
```

## By specialty type and weight category

```{r}
one_year %>%
  mutate(pcosdx_specialty = fct_recode(one_year$pcosdx_specialty,
    "Primary Care (Pediatrics or Family Medicine)/Other" =
      "Primary Care (Pediatrics or Family Medicine)",
    "Primary Care (Pediatrics or Family Medicine)/Other" = "Other",
    "Gynecology/Adolescent Medicine" = "Gynecology",
    "Gynecology/Adolescent Medicine" = "Adolescent Medicine"
  )) %>%
  select(-record_number, -site) %>%
  filter(!is.na(pcosdx_specialty)) %>%
  tbl_strata(
    strata = pcosdx_specialty,
    .tbl_fun =
      ~ .x |>
        tbl_summary(by = weight_cat, label = as.list(label(df))) %>%
        add_overall() %>%
        add_p(test.args = all_tests("fisher.test") ~
          list(simulate.p.value = TRUE)) %>%
        separate_p_footnotes() %>%
        add_q() %>%
        modify_footnote(everything() ~ NA),
    .header = "**{strata}**, N = {n}"
  )
```

# To Do

- Compare fasting and non-fasting lipids
- Add medication combinations
- Add ever vs. never once variables are correct for first year analysis
