---
title: "Association of Clinical and Biochemical Markers with Mental Health Disorders"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    fig-width: 12
    fig-height: 9
    page-layout: full
    theme:
      light: flatly
      dark: darkly
editor: source
---

```{r setup}
#| include: false
library(tidyverse)
library(arsenal)
knitr::opts_knit$set(root.dir = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/BDC/Janet Snell-Bergeon/CALICO")
```

```{r data import}
#| include: false
# Load data
load("./Data_Clean/analysis_data.RData")
# Check for mental health diagnoses 1 year from PCOS diagnosis
new_dx <- df %>%
  group_by(record_number) %>%
  filter(cv_monthssincepcosdx <= 1 | is.na(cv_monthssincepcosdx)) %>%
  summarise(
    anxiety = factor(any(anxiety == "Yes"),
      levels = c(F, T), labels = c("No", "Yes")
    ),
    depression = factor(any(depression == "Yes"),
      levels = c(F, T), labels = c("No", "Yes")
    ),
    red = factor(any(red == "Yes"), levels = c(F, T), labels = c("No", "Yes")),
    bed = factor(any(bed == "Yes"), levels = c(F, T), labels = c("No", "Yes"))
  )
# Basic formula for tables
f <- ~ cv_bmi_z + weight_perc + overweight_perc + weight_raw + overweight_raw +
  pcosdx_pmh___7 + cv_a1c + cv_ft_perc + cv_tt_perc + cv_shbg +
  cv_medications___17 + cv_medications___18 + cv_medications___19 +
  mental_health_screening
# Set table options
mycontrols <-
  tableby.control(
    numeric.stats = c("N", "Nmiss", "meansd", "medianq1q3", "range"),
    cat.stats = c("N", "Nmiss", "countpct")
  )
# First visit only
cv1 <- df %>%
  group_by(record_number) %>%
  slice_min(redcap_repeat_instance) %>%
  ungroup()
cv1[, c("anxiety", "depression", "red", "bed")] <- NULL
cv1 <- left_join(cv1, new_dx, by = join_by(record_number))
```

Presence or absence of a mental health disorder was defined as past medical history of the disorder within 12 months of PCOS diagnosis (e.g., `pcosdx_pmh___15` == "Checked") or a new diagnosis within 12 months of PCOS diagnosis (e.g., `cv_newdx___15` == "Checked"). Depression was defined similarly, but also included anyone with an elevated PHQ2, PHQ8, PHQ9 or CESD20.

## Anxiety

```{r results='asis'}
tbl <- tableby(update(f, anxiety ~ .), data = cv1, control = mycontrols)
summary(tbl, labelTranslations = labels(df), pfootnote = T)
```

## Depression

```{r results='asis'}
tbl <- tableby(update(f, depression ~ .), data = cv1, control = mycontrols)
summary(tbl, labelTranslations = labels(df), pfootnote = T)
```

## ADHD

```{r results='asis'}
tbl <- tableby(update(f, adhd ~ .), data = cv1, control = mycontrols)
summary(tbl, labelTranslations = labels(df), pfootnote = T)
```

## Binge eating disorder (including bulemia)

```{r results='asis'}
tbl <- tableby(update(f, bed ~ .), data = cv1, control = mycontrols)
summary(tbl, labelTranslations = labels(df), pfootnote = T)
```

## Restricting eating disorder (including anorexia)

```{r results='asis'}
tbl <- tableby(update(f, red ~ .), data = cv1, control = mycontrols)
summary(tbl, labelTranslations = labels(df), pfootnote = T)
```
