---
title: "CALICO NCH and CHOP Project"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    fig-width: 12
    fig-height: 9
    page-layout: full
    theme:
      light: flatly
      dark: darkly
bibliography: C:/Users/timvigers/Dropbox/Miscellaneous/zotero.bib
csl: C:/Users/timvigers/Documents/GitHub/styles/american-medical-association.csl
editor: source
---

```{r setup}
#| include: false
library(tidyverse)
library(arsenal)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/BDC/Janet Snell-Bergeon/CALICO",
  "Windows" = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/BDC/Janet Snell-Bergeon/CALICO",
  "Linux" = "/home/timvigers/OneDrive/Vigers/BDC/Janet Snell-Bergeon/CALICO"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data import}
#| include: false
# Load data
load("./Data_Clean/analysis_data.RData")
# Basic formula for tables
f <- as.formula(paste(
  "~",
  paste(c(aim1_vars, "mental_health_screening", "Region"), collapse = "+")
))
f <- update(f, . ~ . - pcosdx_obesitydx_age - pcosdx_obesitydx_earliestage)
# Set table options
mycontrols <-
  tableby.control(numeric.stats = c("Nmiss2", "meansd", "medianq1q3", "range"))
# First visit only
cv1 <- df %>%
  group_by(record_number) %>%
  slice_min(redcap_repeat_instance) %>%
  ungroup()
# Filter per updated "analysis plan"
cv1 <- cv1 %>%
  filter(
    !is.na(pcosdx_irregular_menses) &
      pcosdx_irregular_menses != "Unknown/Not recorded",
    pcosdx_hyperandrogenism != "Unknown/Not recorded"
  )
```

# Demographics

```{r results='asis'}
#| label: tbl-table-1
#| tbl-cap: Participant demographics
f1 <-
  as.formula(paste0("~", paste0(c(demo_vars, "combined_race"), collapse = "+")))
t1 <- tableby(f1, data = cv1, control = mycontrols)
summary(t1, labelTranslations = labels(df), pfootnote = T)
```

# By age

```{r results='asis'}
#| label: tbl-table-1-age
#| tbl-cap: Stratified by age group at diagnosis
t1_age <- tableby(update(f, age_group ~ .), data = cv1, control = mycontrols)
summary(t1_age, labelTranslations = labels(df), pfootnote = T)
```

# By obesity

## Based on percentile and raw value

```{r results='asis'}
#| label: tbl-table-1-obese-perc
#| tbl-cap: Stratified by obesity (percentile and raw value)
t1_bmi <- tableby(update(f, Obesity_perc ~ .), data = cv1, control = mycontrols)
summary(t1_bmi, labelTranslations = labels(df), pfootnote = T)
```

For those with a BMI percentile included in the dataset, participants with a BMI percentile $\geq$ 85 were considered have obesity. For those with a missing value for BMI percentile, those with a raw BMI value $\geq$ 25 were considered to have obesity. 

## Based on raw value only

```{r results='asis'}
#| label: tbl-table-1-obese-raw
#| tbl-cap: Stratified by obesity (raw value only)
t1_bmi <- tableby(update(f, Obesity_raw ~ .), data = cv1, control = mycontrols)
summary(t1_bmi, labelTranslations = labels(df), pfootnote = T)
```

# By race

```{r results='asis'}
#| label: tbl-table-1-race
#| tbl-cap: Stratified by race
t1_race <- tableby(update(f, combined_race ~ .),
  data = cv1, control = mycontrols
)
summary(t1_race, labelTranslations = labels(df), pfootnote = T)
```

# By ethnicity

```{r results='asis'}
#| label: tbl-table-1-ethnicity
#| tbl-cap: Stratified by ethnicity
t1_eth <- tableby(update(f, ethnicity ~ .), data = cv1, control = mycontrols)
summary(t1_eth, labelTranslations = labels(df), pfootnote = T)
```

# By specialty

```{r results='asis'}
#| label: tbl-table-1-specialty
#| tbl-cap: Stratified by specialty at diagnosis
t1_spec <- tableby(update(f, pcosdx_specialty ~ .),
  data = cv1, control = mycontrols
)
summary(t1_spec, labelTranslations = labels(df), pfootnote = T)
```

# Age when obesity started by mental health diagnoses

```{r}
f <- as.formula("~ pcosdx_obesitydx_age_combined")
```

Presence or absence of a given mental health diagnosis was based on the past medical history and new diagnosis fields in REDCap. For example, a participant was assumed to have depression if they checked `pcosdx_pmh___16` or `cv_newdx___16` at any visit.

## Depression

```{r results='asis'}
t1_dep <- tableby(update(f, depression ~ .), data = cv1, control = mycontrols)
summary(t1_dep, labelTranslations = labels(df), pfootnote = T)
```

## Anxiety

```{r results='asis'}
t1_anx <- tableby(update(f, anxiety ~ .), data = cv1, control = mycontrols)
summary(t1_anx, labelTranslations = labels(df), pfootnote = T)
```

## Binge eating disorder

```{r results='asis'}
t1_bed <- tableby(update(f, bed ~ .), data = cv1, control = mycontrols)
summary(t1_bed, labelTranslations = labels(df), pfootnote = T)
```

## Restrictive eating disorder

```{r results='asis'}
t1_red <- tableby(update(f, red ~ .), data = cv1, control = mycontrols)
summary(t1_red, labelTranslations = labels(df), pfootnote = T)
```

## ADHD

```{r results='asis'}
t1_adhd <- tableby(update(f, adhd ~ .), data = cv1, control = mycontrols)
summary(t1_adhd, labelTranslations = labels(df), pfootnote = T)
```
