---
title: "CALICO Health Disparities Analyses"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
editor: source
---

```{r setup}
#| include: false
library(Hmisc)
library(tidyverse)
library(arsenal)
library(performance)
library(forestmodel)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/BDC/Janet Snell-Bergeon/CALICO",
  "Linux" = "/home/timvigers/OneDrive/Vigers/BDC/Janet Snell-Bergeon/CALICO"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data import and cleaning}
#| include: false
# Load data
load("./Data_Clean/analysis_data_2024_11_22.RData")
# Select columns, remove "Other" and "Military" insurance,
# combine races
df$combined_race[df$combined_race == "Pacific Islander"] <- "Other"
df$combined_race[df$combined_race == "American Indian or Alaska Native"] <-
  "Other"
df$combined_race <- droplevels(df$combined_race)
df$combined_race <- relevel(df$combined_race, ref = "Caucasian")
df$insur_type <- factor(df$insur_type,
  levels = c(
    "Public", "Private", "Military", "None", "Other {insur_other}"
  ),
  labels = c(
    "Public/None/Military/Other", "Private", "Public/None/Military/Other",
    "Public/None/Military/Other", "Public/None/Military/Other"
  )
)
df$insur_type <- relevel(df$insur_type, ref = "Private")
# Labels
label(df$insur_type) <- "Insurance Type"
label(df$combined_race) <- "Race"
# Basic formula for tables
f <- as.formula(paste("~", paste(ses_vars, collapse = "+")))
# Set table options
mycontrols <-
  tableby.control(
    numeric.stats = c("N", "Nmiss", "meansd", "medianq1q3", "range"),
    cat.stats = c("N", "Nmiss", "countpct")
  )
# First visit only
cv1 <- df %>%
  group_by(record_number) %>%
  slice_min(redcap_repeat_instance) %>%
  ungroup()
label(cv1$pcosdx_any_famhx___27) <- "Family Hx PCOS"
label(cv1$pcosdx_any_famhx___7) <- "Family Hx T2D"
label(cv1$pcosdx_any_famhx___5) <- "Family Hx Overweight"
label(cv1$pcosdx_age) = "Age at PCOS Dx"
levels(cv1$insur_type) <- c("Private", "Public/Other")
```

# At time of diagnosis/first clinic visit

## Overall

```{r results='asis'}
t1 <- tableby(f, data = cv1, control = mycontrols)
summary(t1, labelTranslations = labels(df), pfootnote = T)
```

## By insurance type

```{r results='asis'}
t2 <- tableby(update(f, insur_type ~ .), data = cv1, control = mycontrols)
summary(t2, labelTranslations = labels(df), pfootnote = T)
```

## By race

```{r results='asis'}
t3 <- tableby(update(f, combined_race ~ .), data = cv1, control = mycontrols)
summary(t3, labelTranslations = labels(df), pfootnote = T)
```

## By ethnicity

```{r results='asis'}
t4 <- tableby(update(f, ethnicity ~ .), data = cv1, control = mycontrols)
summary(t4, labelTranslations = labels(df), pfootnote = T)
```

## By specialty

```{r results='asis'}
t5 <- tableby(update(f, pcosdx_specialty ~ .), data = cv1, control = mycontrols)
summary(t5, labelTranslations = labels(df), pfootnote = T)
```

## Missing data by insurance type

```{r results='asis'}
cv1_miss <- as.data.frame(lapply(cv1, function(c) {
  factor(is.na(c), levels = c(F, T), labels = c("Not Missing", "Missing"))
}))
cv1_miss$insur_type <- cv1$insur_type
insur_miss <- tableby(update(f, insur_type ~ .),
  data = cv1_miss,
  control = list(cat.stats = c("N", "countpct"))
)
summary(insur_miss, labelTranslations = labels(df), pfootnote = T)
```

# Baseline models

The aim of these model is to "determine differences in utilization of weight management, behavioral health and pharmacological treatment options in adolescents with PCOS based on insurance status and whether it is independent of personal medical history factors, family history, race, ethnicity, provider preference (specialty) and geographic location."

Based on this, I assumed that the outcome (dependent) variables are the following newly prescribed medications at visit 1: 

1. `cv_newmeds___18` (Antidepressant or anti-anxiety)
2. `cv_newmeds___19` (ADHD medication)
3. `cv_newmeds___17` (Atypical antipsychotic)
4. `cv_newmeds___25` (Topiramate for weight loss)
5. `cv_newmeds___26` (Phentermine for weight)
6. `cv_newmeds___27` (Topirimate/phentermine for weight loss)
7. `cv_newmeds___28` (Liraglutiade (saxcenda) for weight loss)
8. `cv_newmeds___29` (Semaglutide (Wygovy/Ozempic) for weight loss)
9. `cv_newmeds___1` (Metformin)

The following comorbidities at the time of diagnosis:

1. `pcosdx_pmh___6` (Pre-diabetes)
2. `pcosdx_pmh___7` (Type 2 diabetes)

The following mental health treatments at time of diagnosis, which were combined into a single variable:

1. `pcosdx_mentalhealthcounseling___1` (Group therapy)
2. `pcosdx_mentalhealthcounseling___2` (Individual therapy)
3. `pcosdx_mentalhealthcounseling___3` (Inpatient therapy)

I also assumed that the predictor (independent) variables to be included as covariates are: 

1. `pcosdx_age` (Age at time of PCOS diagnosis)
2. `cv_tt_perc` (Total testosterone (ng/dL) value for upper limit of normal for assay)
3. `cv_ft_perc` (Free Testosterone (pg/mL) value for upper limit of normal for sex, age/tanner stage  for assay)
4. `cv_bmi` (BMI)
5. `pcosdx_birthweight_calc` (Birthweight)
6. `pcosdx_any_famhx___27` (Any family history of PCOS - I created this variable)
7. `pcosdx_any_famhx___7` (Any family history of T2D - I created this variable)
8. `pcosdx_any_famhx___5` (Any family history of Overweight/Obesity - I created this variable)
11. `insur_type` (Insurance Type)
12. `pcosdx_specialty` (What medical specialty made the diagnosis of PCOS)
13. `combined_race` (Race - I created this variable)
14. `ethnicity` (Ethnicity)

We were also interested in examining the relationship between free testosterone and the other independent variables (total testosterone excluded).

```{r}
# List of outcomes and predictors
outcomes <- c(
  "cv_newmeds___18", "cv_newmeds___19", "cv_newmeds___17", "cv_newmeds___25",
  "cv_newmeds___26", "cv_newmeds___27", "cv_newmeds___28", "cv_newmeds___29",
  "cv_newmeds___1", "pcosdx_pmh___6", "pcosdx_pmh___7",
  "pcosdx_any_mentalhealthcounseling"
)
predictors <- c(
  "pcosdx_age", "cv_tt_perc", "cv_ft_perc", "cv_bmi", "pcosdx_birthweight_calc",
  "cv_acneother___1", "cv_acneother___2", "pcosdx_any_famhx___27",
  "pcosdx_any_famhx___7", "pcosdx_any_famhx___5", "insur_type",
  "pcosdx_specialty", "combined_race", "ethnicity"
)
# Base formula
f <- as.formula(paste0("~", paste0(predictors, collapse = "+")))
```

## `cv_newmeds___18`

There weren't a huge number of people with new antidepressant or anti-anxiety medications at clinic visit 1 (n=38). `cv_newmeds___18` was not associated with insurance type.

```{r}
m <- glm(cv_newmeds___18 ~ insur_type, data = cv1, family = "binomial")
forest_model(m)
```

## `cv_newmeds___19`

There were very few people with new ADHD medication at clinic visit 1 (n=8). `cv_newmeds___19` was not associated with insurance type.

```{r}
m <- glm(cv_newmeds___19 ~ insur_type, data = cv1, family = "binomial")
forest_model(m)
```

## `cv_newmeds___17`

There were only 2 people newly prescribed atypical antipsychotic medication at clinic visit 1 (both on public insurance), so we are unable to fit any models.

## `cv_newmeds___25`

There were only 2 people were newly prescribed Topiramate for weight loss.

```{r}
m <- glm(cv_newmeds___25 ~ insur_type, data = cv1, family = "binomial")
forest_model(m)
```

## `cv_newmeds___26`

Only one person was newly prescribed Phentermine for weight medication at clinic visit 1 (on public insurance), so we are unable to fit any models.

## `cv_newmeds___27`

No one was newly prescribed Topirimate/phentermine for weight loss at clinic visit 1.

## `cv_newmeds___28`

Only one person was newly prescribed Liraglutiade (saxcenda) for weight loss medication at clinic visit 1 (on public insurance), so we are unable to fit any models.

## `cv_newmeds___29`

There were only 4 people with a new Semaglutide (Wygovy/Ozempic) for weight loss prescription at clinic visit 1. `cv_newmeds___29` was not associated with insurance type.

```{r}
m <- glm(cv_newmeds___29 ~ insur_type, data = cv1, family = "binomial")
forest_model(m)
```

## `cv_newmeds___1`

There were a decent number of people with a new Metformin prescription at clinic visit 1 (n=299), but `cv_newmeds___1` was not associated with insurance type.

```{r}
m <- glm(cv_newmeds___1 ~ insur_type, data = cv1, family = "binomial")
forest_model(m)
```

## `pcosdx_pmh___6`

There were 69 people with pre-diabetes at PCOS diagnosis (n=69). Model adjusted for race and raw BMI value.

```{r}
m <- glm(pcosdx_pmh___6 ~ insur_type + cv_bmi + combined_race,
  data = cv1, family = "binomial"
)
forest_model(m,exponentiate = T)
```

## `pcosdx_pmh___7`

There weren't a huge number of people with type 2 diabetes at PCOS diagnosis (n=10). `pcosdx_pmh___7` was not associated with insurance type. Due to the small number of "events," we were not able to adjust this model for race.

```{r}
m <- glm(pcosdx_pmh___7 ~ insur_type + cv_bmi, data = cv1, family = "binomial")
forest_model(m,exponentiate = T)
```

## Any mental health counseling prior to diagnosis

630 participants were engaged in some kind of mental health counseling, but this was not associated with insurance type.

```{r}
m <- glm(pcosdx_any_mentalhealthcounseling ~ insur_type,
  data = cv1,
  family = "binomial"
)
forest_model(m)
```

## `cv_ft_perc`

Take out acne, specialty, etc.

```{r}
m <- lm(
  update(f, log(cv_ft_perc) ~ .
    - cv_tt_perc - cv_ft_perc - pcosdx_birthweight_calc - cv_acneother___1 -
    cv_acneother___2 - pcosdx_specialty),
  data = cv1
)
forest_model(m,exponentiate = T)
```

The model looks slightly better after log transforming the outcome, so these coefficients will be interpreted on the natural log scale. 
