---
title: "CAMP study analysis"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = FALSE)

library(tableone)
library(knitr)
library(dplyr)
library(arsenal)
library(childsds)


```


```{r data, include=FALSE}
data <- read.csv("/Volumes/Peds Endo/Lauren Shomaker/CAMP/export.csv")

# factors
data$mini_meetscriteria___1.38 <- as.factor(data$mini_meetscriteria___1.38)
data$mini_meetscriteria___2.38 <- as.factor(data$mini_meetscriteria___2.38)
data$mini_depressiveepisode_type___1.0 <- as.factor(data$mini_depressiveepisode_type___1.0)
data$mini_depressiveepisode_type___2.0 <- as.factor(data$mini_depressiveepisode_type___2.0)
data$race_corr_table <- as.factor(data$race_corr_table)
data$race_ethnicity <- as.factor(data$race_ethnicity)
data$ede3monthsloc.38 <- as.factor(data$ede3monthsloc.38)
data$edeLOCpresence.0 <- as.factor(data$edeLOCpresence.0)
data$weightstatus.0 <- as.factor(data$weightstatus.0)

# calculate change in weight, BMI Z-score, and BMI percentile from SV to EoTx
data$delta_bmi <- data$bmi_kgm2.38 - data$firsttri_bmi_kgm2.0
data$delta_bmiz <- data$zscore.38 - data$firsttri_bmi_zscore.0
data$delta_bmip <- data$bmi_percentile.38 - data$firsttri_percentile.0

# calculate additional GWG variables
data$experimental_gwg <- data$measuredwt_kg.38-data$measuredwt_kg.0
data$total_gwg <- data$predelivery_campwt_kg.52-data$prepregwt_kg.0

# calculate BMI and BMI percentile at baseline
data$prepregbmi <- data$prepregwt_kg.0 / ((data$prepreght_cm.0/100)^2)
data$sex <- "Female"
data$prepregbmi_p <- sds(data$prepregbmi,
              age = data$age.0,
              sex = data$sex, male = "Male", female =  "Female",
              ref = cdc.ref,
              item = "bmi",
              type = "perc")
data$prepregbmi_p <- data$prepregbmi_p*100

# calculate whether GWG was low, appropriate, or excessive, separately for experimental and total GWG
# all participants are under 20 years old, so use CDC criteria for BMI %ile
data$prepreg_bmi_cat <- ifelse(data$prepregbmi_p<5, "Underweight",
                               ifelse(data$prepregbmi_p>=5 & data$prepregbmi_p<85, "Normal weight",
                               ifelse(data$prepregbmi_p>=85 & data$prepregbmi_p<95, "Overweight","Obese")))


data$experimental_gwg_cat <- ifelse(is.na(data$experimental_gwg), NA, 
                                    ifelse(data$prepreg_bmi_cat=="Underweight" & data$experimental_gwg<12.5, "Low",
                                    ifelse(data$prepreg_bmi_cat=="Underweight" & data$experimental_gwg>=12.5  & data$experimental_gwg<=18, "Appropriate",
                                    ifelse(data$prepreg_bmi_cat=="Underweight" & data$experimental_gwg>18, "Excessive",
                                    ifelse(data$prepreg_bmi_cat=="Normal weight" & data$experimental_gwg<11.5, "Low",
                                    ifelse(data$prepreg_bmi_cat=="Normal weight" & data$experimental_gwg>=11.5  & data$experimental_gwg<=16, "Appropriate",
                                    ifelse(data$prepreg_bmi_cat=="Normal weight" & data$experimental_gwg>16, "Excessive",
                                    ifelse(data$prepreg_bmi_cat=="Overweight" & data$experimental_gwg<7, "Low",
                                    ifelse(data$prepreg_bmi_cat=="Overweight" & data$experimental_gwg>=7  & data$experimental_gwg<=11.5, "Appropriate",
                                    ifelse(data$prepreg_bmi_cat=="Overweight" & data$experimental_gwg>11.5, "Excessive",
                                    ifelse(data$prepreg_bmi_cat=="Obese" & data$experimental_gwg<5, "Low",
                                    ifelse(data$prepreg_bmi_cat=="Obese" & data$experimental_gwg>=5  & data$experimental_gwg<=9, "Appropriate","Excessive"))))))))))))
data$total_gwg_cat <- ifelse(is.na(data$total_gwg), NA, 
                                    ifelse(data$prepreg_bmi_cat=="Underweight" & data$total_gwg<12.5, "Low",
                                    ifelse(data$prepreg_bmi_cat=="Underweight" & data$total_gwg>=12.5  & data$total_gwg<=18, "Appropriate",
                                    ifelse(data$prepreg_bmi_cat=="Underweight" & data$total_gwg>18, "Excessive",
                                    ifelse(data$prepreg_bmi_cat=="Normal weight" & data$total_gwg<11.5, "Low",
                                    ifelse(data$prepreg_bmi_cat=="Normal weight" & data$total_gwg>=11.5  & data$total_gwg<=16, "Appropriate",
                                    ifelse(data$prepreg_bmi_cat=="Normal weight" & data$total_gwg>16, "Excessive",
                                    ifelse(data$prepreg_bmi_cat=="Overweight" & data$total_gwg<7, "Low",
                                    ifelse(data$prepreg_bmi_cat=="Overweight" & data$total_gwg>=7  & data$total_gwg<=11.5, "Appropriate",
                                    ifelse(data$prepreg_bmi_cat=="Overweight" & data$total_gwg>11.5, "Excessive",
                                    ifelse(data$prepreg_bmi_cat=="Obese" & data$total_gwg<5, "Low",
                                    ifelse(data$prepreg_bmi_cat=="Obese" & data$total_gwg>=5  & data$total_gwg<=9, "Appropriate","Excessive"))))))))))))

# change in social functioning
data$delta_sastot <- data$sastot.38-data$sastot.0
data$delta_fatherfigure_negtot <- data$nri_ff_negtot.38-data$nri_ff_negtot.0
data$delta_fatherfigure_suptot <- data$nri_ff_suptot.38-data$nri_ff_suptot.0
data$delta_motherfigure_negtot <- data$nri_mf_negtot.38-data$nri_mf_negtot.0
data$delta_motherfigure_suptot <- data$nri_mf_suptot.38-data$nri_mf_suptot.0
data$delta_romanticfigure_negtot <- data$nri_rf_negtot.38-data$nri_rf_negtot.0
data$delta_romanticfigure_suptot <- data$nri_rf_suptot.38-data$nri_rf_suptot.0
data$delta_samesex_negtot <- data$nri_ssf_negtot.38-data$nri_ssf_negtot.0
data$delta_samesex_suptot <- data$nri_ssf_suptot.38-data$nri_ssf_suptot.0
data$delta_dating_tot <- data$sas_dating_tot.38-data$sas_dating_tot.0
data$delta_familybehvr_tot <- data$sas_familybehvr_tot.38-data$sas_familybehvr_tot.0
data$delta_sas_frnds_sparetme_tot <- data$sas_frnds_sparetme_tot.38-data$sas_frnds_sparetme_tot.0
data$delta_sas_schoolbeh_tot <- data$sas_schoolbeh_tot.38-data$sas_schoolbeh_tot.0

# change in perceived stress
data$delta_psstot <- data$pss10_total.38-data$pss10_total.0

# change in depression
data$delta_cesdtot <- data$cesdtot.38-data$cesdtot.0
data$delta_edinburghtot <- data$edinburgh_total.38 - data$edinburgh_total.0

# change in overeating
data$delta_eestot <- data$ees_tot.38-data$ees_tot.0

# change in physical activity
data$delta_light <- data$actigraph_light.38-data$actigraph_light.0
data$delta_sedentary <- data$actigraph_sedentary.38-data$actigraph_sedentary.0
data$delta_mvpa <- data$average_mvpa_per_day.38-data$average_mvpa_per_day.0
# for those who don't have sufficient valid days at either baseline or follow-up, set delta to missing
data$delta_light <- ifelse(data$actigraph_valid_days.0==1 & data$actigraph_valid_days.38==1, data$delta_light, NA)
data$delta_sedentary <- ifelse(data$actigraph_valid_days.0==1 & data$actigraph_valid_days.38==1, data$delta_sedentary, NA)
data$delta_mvpa <- ifelse(data$actigraph_valid_days.0==1 & data$actigraph_valid_days.38==1, data$delta_mvpa, NA)

# change in sleep
data$delta_totalsleep <- data$total_sleep_time.38-data$total_sleep_time.0
data$delta_waso <- data$wake_after_sleep_onset.38-data$wake_after_sleep_onset.0
data$delta_sleepeff <- data$sleep_efficiency.38-data$sleep_efficiency.0

# create new variable for group for descriptive tables
data$group_condition_char <- ifelse(data$group_condition==1, "IPT + UC", "UC")

# did ppt attend 5 of 6 sessions?
data$SessionAtt80percent <- as.factor(ifelse(is.na(data$SessionAtt),NA,
                           ifelse(data$SessionAtt>=5,">=80% attendance","<80% attendance")))

# binary variable indicating whether the ppt scored 4 or higher on the acceptability questions
data$SessionAtt_4 <- as.factor(ifelse(is.na(data$SessionAtt),NA,
                            ifelse(data$SessionAtt>=4,1,0)))
data$participate_again_4 <- as.factor(ifelse(is.na(data$participate_again),NA,
                            ifelse(data$participate_again>=4,1,0)))
data$enjoy_participating_4 <- as.factor(ifelse(is.na(data$enjoy_participating),NA,
                            ifelse(data$enjoy_participating>=4,1,0)))
data$recommend_study_4 <- as.factor(ifelse(is.na(data$recommend_study),NA,
                            ifelse(data$recommend_study>=4,1,0)))
data$end_interest_4 <- as.factor(ifelse(is.na(data$end_interest),NA,
                            ifelse(data$end_interest>=4,1,0)))

```

# Methods

Because of the relatively small sample size, descriptive statistics presented for continuous variables are median (25th, 75th percentile). Groups were compared using the Kruskal-Wallis test for continuous variables and the chi-square test for categorical variables.

# Results

## Baseline descriptive statistics - by arm

```{r echo=FALSE, include=FALSE}

desc_vars <- c("age.0","gestational weeks_fractional.0","race_corr_table","race_ethnicity","prepreght_cm.0","prepregwt_kg.0",
               "prepregbmi_p","prepreg_weightstatus.0","prepreght_cm.0","prepregwt_kg.0","prepreg_bmi_percentile.0","prepregbmi",
               "prepreg_weightstatus.0","measuredht_cm.0","measuredwt_kg.0","firsttri_bmi_kgm2.0","firsttri_bmi_zscore.0",
               "firsttri_percentile.0","weightstatus.0","sas_dating_tot.0","sas_familybehvr_tot.0","sas_frnds_sparetme_tot.0",
               "sas_schoolbeh_tot.0","nri_ff_suptot.0","nri_mf_suptot.0","nri_rf_suptot.0","nri_ssf_suptot.0",
               "nri_ff_negtot.0","nri_mf_negtot.0","nri_rf_negtot.0","nri_ssf_negtot.0","pss10_total.0",
               "cesdtot.0","edinburgh_total.0","minimeetscriteria_1.0","minimeetscrtieria_2.0",
               "mini_depressiveepisode_type___1.0","mini_depressiveepisode_type___2.0",
               "ede_LOC.0","ees_tot.0","total_sleep_time.0","wake_after_sleep_onset.0","sleep_efficiency.0",
               "prepreg_bmi.0","prepreg_bmi_percentile.0")

desc_table <- CreateTableOne(data=data,vars = desc_vars,strata = "group_condition_char")
desc_table <- print(desc_table)
```

```{r echo=FALSE}
kable(desc_table)
```

## Baseline descriptive statistics - by arm - physical activity for those with Actigraph_valid_days.0 = 1

```{r echo=FALSE, include=FALSE}
data_valid <- data %>% filter(actigraph_valid_days.0==1)
desc_vars <- c("average_mvpa_per_day.0","actigraph_light.0","actigraph_sedentary.0")

desc_table <- CreateTableOne(data=data_valid,vars = desc_vars,strata = "group_condition_char")
desc_table <- print(desc_table)
```

```{r echo=FALSE}
kable(desc_table)
```

## Baseline descriptive statistics - by study completion status

```{r echo=FALSE, include=FALSE, eval=FALSE}
# can't find the variable to indicate study completion status so will not evaluate these chunks
desc_table <- CreateTableOne(data=data,vars = desc_vars,strata = "group_condition_char")
desc_table <- print(desc_table)
```

```{r echo=FALSE, eval=FALSE}
kable(desc_table)
```

## Baseline descriptive statistics - by study completion status - physical activity for those with Actigraph_valid_days.0 = 1

```{r echo=FALSE, include=FALSE, eval=FALSE}
data_valid <- data %>% filter(actigraph_valid_days.0==1)
desc_vars <- c("average_mvpa_per_day.0","actigraph_light.0","actigraph_sedentary.0")

desc_table <- CreateTableOne(data=data_valid,vars = desc_vars,strata = "group_condition_char")
desc_table <- print(desc_table)
```

```{r echo=FALSE, eval=FALSE}
kable(desc_table)
```

## Unadjusted comparisons

### Acceptability

```{r, include=FALSE}
t1 <- tableby(data=data, group_condition_char ~ SessionAtt+SessionAtt80percent+participate_again+enjoy_participating+recommend_study+end_interest+
                SessionAtt_4+participate_again_4+enjoy_participating_4+recommend_study_4+end_interest_4,
                     numeric.test="kwt",numeric.stats=c("Nmiss2","median","q1q3"))

```

```{r echo=FALSE, include=TRUE, results='asis'}
summary(t1)
```

### Gestational weight gain

```{r, include=FALSE}
t2 <- tableby(data=data, group_condition_char ~ delta_bmi+delta_bmiz+delta_bmip+experimental_gwg+total_gwg+experimental_gwg_cat+total_gwg_cat,
                     numeric.test="kwt",numeric.stats=c("Nmiss2","median","q1q3"))
```

```{r echo=FALSE, include=TRUE, results='asis'}
summary(t2)
```

### Social functioning

```{r, include=FALSE}
t3 <- tableby(data=data, group_condition_char ~ delta_sastot+delta_fatherfigure_negtot+delta_fatherfigure_suptot+delta_motherfigure_negtot+delta_motherfigure_suptot+           delta_romanticfigure_negtot+delta_romanticfigure_suptot+delta_samesex_negtot+delta_samesex_suptot+delta_dating_tot+
                delta_familybehvr_tot+delta_sas_frnds_sparetme_tot+delta_sas_schoolbeh_tot,
                     numeric.test="kwt",numeric.stats=c("Nmiss2","median","q1q3"))
```

```{r echo=FALSE, include=TRUE, results='asis'}
summary(t3)
```

### Perceived stress

```{r, include=FALSE}
t4 <- tableby(data=data, group_condition_char ~ delta_psstot,
                     numeric.test="kwt",numeric.stats=c("Nmiss2","median","q1q3"))
```

```{r echo=FALSE, include=TRUE, results='asis'}
summary(t4)
```

### Depression

```{r, include=FALSE}
t5 <- tableby(data=data, group_condition_char ~ delta_cesdtot+delta_edinburghtot+mini_meetscriteria___1.38+mini_meetscriteria___2.38
              +mini_depressiveepisode_type___1.0+mini_depressiveepisode_type___2.0,
                     numeric.test="kwt",numeric.stats=c("Nmiss2","median","q1q3"))
```

```{r t5=FALSE, include=TRUE, results='asis'}
summary(t5)
```

### Overeating

```{r, include=FALSE}
t6 <- tableby(data=data, group_condition_char ~ delta_eestot+edeLOCpresence.0+ede3monthsloc.38,
                     numeric.test="kwt",numeric.stats=c("Nmiss2","median","q1q3"))
```

```{r t5=FALSE, include=TRUE, results='asis'}
summary(t6)
```

### Physical activity

```{r, include=FALSE}
t7 <- tableby(data=data, group_condition_char ~ delta_light+delta_sedentary+delta_mvpa,
                     numeric.test="kwt",numeric.stats=c("Nmiss2","median","q1q3"))
```

```{r t5=FALSE, include=TRUE, results='asis'}
summary(t7)
```

### Sleep

```{r, include=FALSE}
t8 <- tableby(data=data, group_condition_char ~ delta_totalsleep+delta_waso+delta_sleepeff,
                     numeric.test="kwt",numeric.stats=c("Nmiss2","median","q1q3"))
```

```{r t5=FALSE, include=TRUE, results='asis'}
summary(t8)
```

## 80% confidence intervals for difference between groups

The idea behind the Cocks and Torgesen approach is that the pilot/feasibility study is powered so that a one-sided 80% confidence interval will exclude the clinically important difference if our observed estimate is zero. However, the CAMP study was not powered according to the Cocks and Torgesen approach so these results would not be easily interpreted.

```{r, eval=FALSE, comment=""}
t.test(participate_again ~ as.factor(group_condition_char), data=data, conf.level=.80)

t.test(enjoy_participating ~ as.factor(group_condition_char), data=data, conf.level=.80)

t.test(recommend_study ~ as.factor(group_condition_char), data=data, conf.level=.80)

t.test(end_interest ~ as.factor(group_condition_char), data=data, conf.level=.80)

delta_bmi+delta_bmiz+delta_bmip

delta_sastot+delta_fatherfigure_negtot+delta_fatherfigure_suptot+delta_motherfigure_negtot+delta_motherfigure_suptot+           delta_romanticfigure_negtot+delta_romanticfigure_suptot+delta_samesex_negtot+delta_samesex_suptot

delta_psstot

delta_cesdtot+delta_edinburghtot+mini_meetscriteria___1.38+mini_meetscriteria___2.38

delta_eestot+edeLOCpresence.0+ede3monthsloc.38

delta_light+delta_sedentary+delta_mvpa

delta_totalsleep+delta_waso+delta_sleepeff
```
