---
title: "EMERALD CV response to reviewers"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: word_document
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(car)
library(knitr)
library(Hmisc)

options(scipen=999)

# read in data for obese T1D vs. T2D comparison
obese <- read.csv("E:\\Nadeau\\Nadeau EMERALD\\Cardiovascular MS\\obese T1D vs. T2D.csv", 
                  na.strings = c(""," ","."))
obese$HbA1C <- as.numeric(as.character(obese$HbA1C))
obese$HR <- as.numeric(as.character(obese$HR))
obese$Serum_creatinine <- as.numeric(as.character(obese$Serum_creatinine))
obese$mcg.MAB_mgCRE <- as.numeric(as.character(obese$mcg.MAB_mgCRE))
obese$ACR <- as.numeric(as.character(obese$ACR))
obese$hs_CRP <- as.numeric(as.character(obese$hs_CRP))
obese$Leptin <- as.numeric(as.character(obese$Leptin))
obese$Adiponectin.baseline <- as.numeric(as.character(obese$Adiponectin.baseline))
obese$eGFR.Zappitelli <- as.numeric(as.character(obese$eGFR.Zappitelli))
obese$eGFR.FAS <- as.numeric(as.character(obese$eGFR.FAS))

# read in all T1D data
t1d <- read.csv("E:\\Nadeau\\Nadeau EMERALD\\Cardiovascular MS\\all T1D.csv", 
                  na.strings = c(""," ","."))

# correlation of BMI with all other variables in all T1D

# ++++++++++++++++++++++++++++
# flattenCorrMatrix
# ++++++++++++++++++++++++++++
# cormat : matrix of the correlation coefficients
# pmat : matrix of the correlation p-values
flattenCorrMatrix <- function(cormat, pmat) {
  ut <- upper.tri(cormat)
  data.frame(
    row = rownames(cormat)[row(cormat)[ut]],
    column = rownames(cormat)[col(cormat)[ut]],
    cor  =(cormat)[ut],
    p = pmat[ut]
    )
}
no <- c("Participant.ID","DOB","Age","Disease","BMI.category","Sex","Ethnicity","DM_Duration","Tanner")
yes <- colnames(t1d)
yes <- yes[!yes %in% no]
corrout <- rcorr(as.matrix(t1d[,yes]), type = "spearman")
corrout <- flattenCorrMatrix(corrout$r, corrout$P)
corrout <- corrout[corrout$row =="BMI",]
corrout$cor <- round(corrout$cor,4)
corrout$p <- round(corrout$p,4)

```
# Comparisons of obese T1D to T2D

The p-value for "as.factor(Category)" is the test of whether obese T1D are different from T2D, after adjusting for sex, diabetes duration, and BMI.

Outcome=HbA1c
```{r, results='asis',tidy=TRUE, echo=FALSE}
mod <- Anova(lm(HbA1C ~ as.factor(Category) + as.factor(Sex) + Diabetes_Duration + BMI, data=obese))
kable(mod)
```

Outcome=HR
```{r, results='asis',tidy=TRUE, echo=FALSE}
mod <- Anova(lm(HR ~ as.factor(Category) + as.factor(Sex) + Diabetes_Duration + BMI, data=obese))
kable(mod)
```

Outcome=SBP
```{r, results='asis',tidy=TRUE, echo=FALSE}
mod <- Anova(lm(SBP ~ as.factor(Category) + as.factor(Sex) + Diabetes_Duration + BMI, data=obese))
kable(mod)
```

\newpage

Outcome=DBP
```{r, results='asis',tidy=TRUE, echo=FALSE}
mod <- Anova(lm(DBP ~ as.factor(Category) + as.factor(Sex) + Diabetes_Duration + BMI, data=obese))
kable(mod)
```

Outcome=SBP %ile
```{r, results='asis',tidy=TRUE, echo=FALSE}
mod <- Anova(lm(SBP_Percentile ~ as.factor(Category) + as.factor(Sex) + Diabetes_Duration + BMI, data=obese))
kable(mod)
```

Outcome=DBP %ile
```{r, results='asis',tidy=TRUE, echo=FALSE}
mod <- Anova(lm(DBP_Percentile ~ as.factor(Category) + as.factor(Sex) + Diabetes_Duration + BMI, data=obese))
kable(mod)
```

Outcome=Pulse Pressure
```{r, results='asis',tidy=TRUE, echo=FALSE}
mod <- Anova(lm(Pulse_Pressure ~ as.factor(Category) + as.factor(Sex) + Diabetes_Duration + BMI, data=obese))
kable(mod)
```

\newpage

Outcome=MAP
```{r, results='asis',tidy=TRUE, echo=FALSE}
mod <- Anova(lm(Mean_Arterial_Pressure ~ as.factor(Category) + as.factor(Sex) + Diabetes_Duration + BMI, data=obese))
kable(mod)
```

Outcome=Serum creatinine
```{r, results='asis',tidy=TRUE, echo=FALSE}
mod <- Anova(lm(Serum_creatinine ~ as.factor(Category) + as.factor(Sex) + Diabetes_Duration + BMI, data=obese))
kable(mod)
```

Outcome=mcg.MAB_mgCRE
```{r, results='asis',tidy=TRUE, echo=FALSE}
mod <- Anova(lm(mcg.MAB_mgCRE ~ as.factor(Category) + as.factor(Sex) + Diabetes_Duration + BMI, data=obese))
kable(mod)
```

Outcome=ACR
```{r, results='asis',tidy=TRUE, echo=FALSE}
mod <- Anova(lm(ACR ~ as.factor(Category) + as.factor(Sex) + Diabetes_Duration + BMI, data=obese))
kable(mod)
```

\newpage

Outcome=hs_CRP
```{r, results='asis',tidy=TRUE, echo=FALSE}
mod <- Anova(lm(hs_CRP ~ as.factor(Category) + as.factor(Sex) + Diabetes_Duration + BMI, data=obese))
kable(mod)
```

Outcome=Leptin
```{r, results='asis',tidy=TRUE, echo=FALSE}
mod <- Anova(lm(Leptin ~ as.factor(Category) + as.factor(Sex) + Diabetes_Duration + BMI, data=obese))
kable(mod)
```

Outcome=Adiponectin
```{r, results='asis',tidy=TRUE, echo=FALSE}
mod <- Anova(lm(Adiponectin.baseline ~ as.factor(Category) + as.factor(Sex) + Diabetes_Duration + BMI, data=obese))
kable(mod)
```

\newpage

Outcome=eGFR Zappatelli
```{r, results='asis',tidy=TRUE, echo=FALSE}
mod <- Anova(lm(eGFR.Zappitelli ~ as.factor(Category) + as.factor(Sex) + Diabetes_Duration + BMI, data=obese))
kable(mod)
```

Outcome=eGFR FAS
```{r, results='asis',tidy=TRUE, echo=FALSE}
mod <- Anova(lm(eGFR.FAS ~ as.factor(Category) + as.factor(Sex) + Diabetes_Duration + BMI, data=obese))
kable(mod)
```

Outcome=Hypertensive
```{r, results='asis',tidy=TRUE, echo=FALSE}
mod <- Anova(glm(Hypertensive. ~ as.factor(Category) + as.factor(Sex) + Diabetes_Duration + BMI, data=obese))
kable(mod)
```

Outcome=Hyperfiltration
```{r, results='asis',tidy=TRUE, echo=FALSE}
mod <- Anova(glm(Hyperfiltration. ~ as.factor(Category) + as.factor(Sex) + Diabetes_Duration + BMI, data=obese))
kable(mod)
```

\newpage

Outcome=Microalbuminuria
```{r, results='asis',tidy=TRUE, echo=FALSE}
mod <- Anova(glm(Microalbuminuria. ~ as.factor(Category) + as.factor(Sex) + Diabetes_Duration + BMI, data=obese))
kable(mod)
```

\newpage

# Correlations of variables with BMI in all participants with T1D

```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(corrout)
```
