---
title: "CGM in CF - Beta cell function outcomes"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r echo=FALSE,include=FALSE}
library(Hmisc)
library(knitr)
library(rspiro)
library(tableone)
library(broom)
library(pROC)
library(dplyr)
library(sjlabelled)
#source('~/Documents/GitHub/Tim-and-Laura/tim_R_functions.R')
#source('C:\\Users\\pylell\\Documents\\GitHub\\Tim-and-Laura\\tim_R_functions.R')
#pathstart <- os.check()
```

```{r echo=FALSE, include=FALSE}
# read in OGTT data
#source("/Volumes/som/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura Tim projects/Chan 2019 ADA abstracts/Data_Cleaned/ada_2019.R")
source("S:\\Shared Projects\\Laura\\Laura Tim projects\\Christine Chan\\2019 ADA Abstracts\\Data_Cleaned\\ada_2019.R")

# read in Laura's code
#source("/Users/timvigers/Documents/GitHub/General-code/temp_table1.r")
#source("/Users/timvigers/Documents/GitHub/General-code/01_functions.r")
#source("/Users/timvigers/Documents/GitHub/General-code/laura_functions.r")
#source("/Users/timvigers/Documents/GitHub/General-code/iAUC.r")

source("C:\\Users\\pylell\\Documents\\GitHub\\General-code\\temp_table1.r")
source("C:\\Users\\pylell\\Documents\\GitHub\\General-code\\01_functions.r")
source("C:\\Users\\pylell\\Documents\\GitHub\\General-code\\laura_functions.r")
source("C:\\Users\\pylell\\Documents\\GitHub\\General-code\\iAUC.r")

# tanner stage
femtanner <- c("tanner_stage_female_breast___1","tanner_stage_female_breast___2","tanner_stage_female_breast___3","tanner_stage_female_breast___4","tanner_stage_female_breast___5" )
maletanner <- c("tanner_stage_male_testicul___1","tanner_stage_male_testicul___2","tanner_stage_male_testicul___3","tanner_stage_male_testicul___4","tanner_stage_male_testicul___5")
data$tanner[which(data$gender.factor == "Female")] <- 
  as.numeric(apply(data[which(data$gender.factor == "Female"),femtanner],1,which.max))
data$tanner[which(data$gender.factor == "Male")] <- 
  as.numeric(apply(data[which(data$gender.factor == "Male"),maletanner],1,which.max))
data$tanner <- as.factor(data$tanner)
# lung function - no Asians in this data set
# Ask Christine if height at the lung function visit should be added to REDCap
data$race_ethnicity <- as.numeric(data$race_ethnicity)
data$lung_race <- ifelse(data$race_ethnicity %in% 3:6,5,data$race_ethnicity)
predGLI <- pred_GLI(data$age_at_visit_1,(data$average_height/100),data$gender,data$lung_race,param=c("FEV1","FVC"))
data$fev_pred <- predGLI$pred.FEV1
data$fvc_pred <- predGLI$pred.FVC
data$fev_perc <- (data$raw_fev1/data$fev_pred)*100
data$fvc_perc <- (data$raw_fvc/data$fvc_pred)*100
# calculate beta cell function outcomes
data$isi <- (data$min_30_insulin - data$min_0_insulin) / (data$min_30_glucose-data$lab_ogtt_fasting)
data$isi_cpep <- (data$min_30_c_peptide - data$min_0_c_peptide) / (data$min_30_glucose-data$lab_ogtt_fasting)
data$odi <- (1/data$min_0_insulin) * data$isi
data$odi_cpep <- (1/data$min_0_c_peptide) * data$isi_cpep
glucvars <- c("lab_ogtt_fasting","min_10_glucose","min_20_glucose","min_30_glucose","min_90_glucose",
                    "min_150_glucose","min_180_glucose")
data$glucmean <- rowMeans(data[,glucvars])
insvars <- c("min_0_insulin","min_10_insulin","min_20_insulin","min_30_insulin","min_90_insulin",
             "min_150_insulin","min_180_insulin")
data$insmean <- rowMeans(data[,insvars])
data$matsuda <- 10000/(sqrt(data$lab_ogtt_fasting*data$min_0_insulin*data$glucmean*data$insmean))
data$homa_ir <- (data$lab_ogtt_fasting*data$min_0_insulin)/22.5

# also need the following
# iAUCg, iAUCins/iAUCg, iAUCcpep/iAUCg, iAUCglucagon
#glucose
cols_gluc = c("min_minus_10_glucose","lab_ogtt_fasting",
         "min_10_glucose","min_20_glucose","min_30_glucose",
         "lab_ogtt_1_hour_glucose","min_90_glucose",
         "lab_ogtt_2_hour_glucose","min_150_glucose",
         "min_180_glucose")
ogttdat <- data[,c(cols_gluc,"subject_id")]
ogttdat <- ogttdat[which(rowSums(!is.na(ogttdat[,cols_gluc]))>=4),]
ogttdat$iaucbase_glucose <- iAUC_glucose(ogttdat,ogttdat$lab_ogtt_fasting,type = "linear")
ogttdat <- ogttdat[,c("subject_id","iaucbase_glucose")]
data <- merge(data,ogttdat,by="subject_id",all.x=TRUE,all.y=TRUE)

#insulin
cols_ins = c("min_minus_10_insulin","min_0_insulin","min_10_insulin","min_20_insulin","min_30_insulin","min_60_insulin",
             "min_90_insulin","min_120_insulin","min_150_insulin","min_180_insulin")
ogttdat <- data[,c(cols_ins,"subject_id")]
ogttdat <- ogttdat[which(rowSums(!is.na(ogttdat[,cols_ins]))>=4),]
ogttdat$iaucbase_insulin <- iAUC_insulin(ogttdat,ogttdat$min_0_insulin,type = "linear")
ogttdat <- ogttdat[,c("subject_id","iaucbase_insulin")]
data <- merge(data,ogttdat,by="subject_id",all.x=TRUE,all.y=TRUE)

#C-peptide
cols_cpep = c("min_minus_10_c_peptide","min_0_c_peptide","min_10_c_peptide","min_20_c_peptide","min_30_c_peptide","min_60_c_peptide",
             "min_90_c_peptide","min_120_c_peptide","min_150_c_peptide","min_180_c_peptide")
ogttdat <- data[,c(cols_cpep,"subject_id")]
ogttdat <- ogttdat[which(rowSums(!is.na(ogttdat[,cols_cpep]))>=4),]
ogttdat$iaucbase_cpep <- iAUC_cpep(ogttdat,ogttdat$min_0_c_peptide,type = "linear")
ogttdat <- ogttdat[,c("subject_id","iaucbase_cpep")]
data <- merge(data,ogttdat,by="subject_id",all.x=TRUE,all.y=TRUE)

# glucagon
cols_glucagon = c("min_minus_10_glucagon","min_0_glucagon","min_10_glucagon","min_20_glucagon","min_30_glucagon","min_60_glucagon",
              "min_90_glucagon","min_120_glucagon","min_150_glucagon","min_180_glucagon")
ogttdat <- data[,c(cols_glucagon,"subject_id")]
ogttdat <- ogttdat[which(rowSums(!is.na(ogttdat[,cols_glucagon]))>=4),]
ogttdat <- ogttdat[which(!is.na(ogttdat[,"min_0_glucagon"])),]
ogttdat$iaucbase_glucagon <- iAUC_glucagon(ogttdat,ogttdat$min_0_glucagon,type = "linear")
ogttdat <- ogttdat[,c("subject_id","iaucbase_glucagon")]
data <- merge(data,ogttdat,by="subject_id",all.x=TRUE,all.y=TRUE)

data$iAUCins_iAUC_gluc <- data$iaucbase_insulin/data$iaucbase_glucose
data$iAUCcpep_iAUCgluc <- data$iaucbase_cpep/data$iaucbase_glucose

# add in CGM variables that were left out of the original dataset
extra <- read.csv("S:\\Shared Projects\\Laura\\Laura Tim projects\\Christine Chan\\2019 ADA Abstracts\\Data_Raw\\CGMInHealthyControls-CGMVariables_DATA_2020-04-30_1347.csv")
extra <- extra[,c("subject_id","percent_time_over_180","percent_time_over_200","cov")]
data <- merge(data,extra,by="subject_id",all.x=TRUE,all.y=FALSE)
# calculate delta BG
data$delta_bg <- data$max_sensor-data$min_sensor

# keep only those who have non-missing odi
alldata <- data[which(!is.na(data$odi) & data$odi != Inf),]

ogttvars <- c("isi","isi_cpep","odi","odi_cpep","matsuda","homa_ir","iaucbase_glucose","iaucbase_insulin","iaucbase_cpep",
              "iaucbase_glucagon","iAUCins_iAUC_gluc","iAUCcpep_iAUCgluc","Calprotectin","C.Reactive.Protein")
cgmvars <- c("min_sensor","max_sensor","average_sensor","average_auc_per_day","excursions_over_200","avg_excur_over_140_per_day",
             "avg_excur_over_200_per_day","excursions_over_140","percent_time_under_60","percent_time_under_70",
             "percent_time_over_140","percent_time_over_180","percent_time_over_200",
             "min_spent_under_60","min_spent_under_70","min_spent_over_200","min_spent_over_140",
             "total_auc","standard_deviation","mean_amp_glycemic_ex")
tab1vars <- c("age_at_visit_1", "race_ethnicity.factor","gender","poc_a1c","fev_perc","fvc_perc","g_tube_feeds","cf_pancreatic","tanner","bmi_z_score")
#tab1vars <- c("age", "race_ethnicity","gender","poc_a1c","g_tube_feeds","fev1_percent","fvc","cf_pancreatic","tanner","first_bmiz")

# make factors
alldata$gender <- as.factor(alldata$gender)
alldata$g_tube_feeds <- as.factor(alldata$g_tube_feeds)
alldata$cf_pancreatic <- as.factor(alldata$cf_pancreatic)

# labels
label(alldata$age_at_visit_1)="Age"
label(alldata$race_ethnicity.factor)="Race/ethnicity"
label(alldata$gender)="Gender"
label(alldata$poc_a1c)="HbA1c"
label(alldata$g_tube_feeds)="G-tube feeds"
label(alldata$fev_perc)="FEV-1 percent"
label(alldata$fvc_perc)="FVC"
label(alldata$cf_pancreatic)="Pancreatic insufficiency"
label(alldata$tanner)="Tanner stage"
label(alldata$bmi_z_score)="First BMI-Z"
label(alldata$isi)="ISI"
label(alldata$isi_cpep)="ISI C-peptide"
label(alldata$odi)="oDI"
label(alldata$odi_cpep)="oDI C-peptide"
label(alldata$min_sensor)="Min sensor glucose"
label(alldata$max_sensor)="Max sensor glucose"
label(alldata$average_sensor)="Average sensor glucose"
label(alldata$average_auc_per_day)="Average AUC"
label(alldata$excursions_over_200)="Excursions>200"
label(alldata$avg_excur_over_140_per_day)="Excursions>140/day"
label(alldata$avg_excur_over_200_per_day)="Excursions>200/day"
label(alldata$excursions_over_140)="Excursions>140"
label(alldata$percent_time_under_60)="% time <60"
label(alldata$percent_time_under_70)="% time <70"
label(alldata$percent_time_over_140)="% time >140"
label(alldata$min_spent_under_60)="Time spent <60"
label(alldata$min_spent_under_70)="Time spent <70"
label(alldata$min_spent_over_200)="Time spent >200"
label(alldata$min_spent_over_140)="Time spent >140"
label(alldata$percent_time_over_180)="% time >180"
label(alldata$percent_time_over_200)="% time >200"
label(alldata$total_auc)="Total AUC"
label(alldata$standard_deviation)="SD"
label(alldata$mean_amp_glycemic_ex)="MAGE"
label(alldata$r_mage)="R-MAGE"
label(alldata$iaucbase_glucose)="iAUC glucose"
label(alldata$iaucbase_insulin)="iAUC insulin"
label(alldata$iaucbase_cpep)="iAUC C-peptide"
label(alldata$iaucbase_glucagon)="iAUC glucagon"
label(alldata$iAUCins_iAUC_gluc)="iAUC insulin/iAUC glucose"
label(alldata$iAUCcpep_iAUCgluc)="iAUC C-peptide/iAUC glucose"

# create new variable for diabetes vs. no diabetes
alldata$diabetes <- ifelse(alldata$group_category.factor=="Cystic Fibrosis Related Diabetes",1,0)

# read in calprotectin and CRP
crp <- read.csv("S:\\Shared Projects\\Laura\\Laura Tim projects\\Christine Chan\\2019 ADA Abstracts\\Data_Raw\\CGMInHealthyControls-CalprotectinAndCRP_DATA_LABELS_2020-04-30_1855.csv")
crp$subject_id <- crp$Subject.ID
crp$Subject.ID <- NULL
alldata <- merge(alldata,crp,by="subject_id",all.x=TRUE,all.y=FALSE)
alldata$Calprotectin <- as.numeric(as.character(alldata$Calprotectin))
set_label(alldata$Calprotectin) <- "Calprotectin"
set_label(alldata$C.Reactive.Protein) <- "CRP"

# create tertiles of CV, MAGE, and delta BG
alldata$cov_tertile <- ntile(alldata$cov,3)
alldata$mage_tertile <- ntile(alldata$r_mage,3)
alldata$delta_bg_tertile <- ntile(alldata$delta_bg,3)

npar <- nrow(alldata)

# ANALYSIS

# table 1
tab1 <- final_table(data=alldata,variables=tab1vars,group=alldata$group_category.factor,margin=2,single=0,ron=2)

# pairwise comparisons for table 1 - HbA1c
a1 <- aov(alldata$poc_a1c~alldata$group_category.factor)
a1t <- TukeyHSD(a1)
a1t <- tidy(a1t)
a1t <- a1t[,-1]

# compare groups
tab2 <- final_table(data=alldata,variables=c(ogttvars,cgmvars),group=alldata$group_category.factor,margin=2,single=0,ron=2)

# pairwise comparisons for table 2 - ISI
a2 <- aov(alldata$isi~alldata$group_category.factor)
a2t <- TukeyHSD(a2)
a2t <- tidy(a2t)
a2t <- a2t[,-1]

# pairwise comparisons for table 2 - ISI C-peptide
a3 <- aov(alldata$isi_cpep~alldata$group_category.factor)
a3t <- tidy(TukeyHSD(a3))
a3t <- a3t[,-1]

# pairwise comparisons for table 2 - oDI
a4 <- aov(alldata$odi~alldata$group_category.factor)
a4t <- tidy(TukeyHSD(a4))
a4t <- a4t[,-1]

# pairwise comparisons for table 2 - oDI C-pep
a5 <- aov(alldata$odi~alldata$group_category.factor)
a5t <- tidy(TukeyHSD(a5))
a5t <- a5t[,-1]

# pairwise comparisons for table 2 - iAUC glucose
a6 <- aov(alldata$iaucbase_glucose~alldata$group_category.factor)
a6t <- tidy(TukeyHSD(a6))
a6t <- a6t[,-1]

# pairwise comparisons for table 2 - iAUC insulin/iAUC glucose
a7 <- aov(alldata$iAUCins_iAUC_gluc~alldata$group_category.factor)
a7t <- tidy(TukeyHSD(a7))
a7t <- a7t[,-1]

# pairwise comparisons for table 2 - iAUC C-peptide/iAUC glucose
a8 <- aov(alldata$iAUCcpep_iAUCgluc~alldata$group_category.factor)
a8t <- tidy(TukeyHSD(a8))
a8t <- a8t[,-1]

# pairwise comparisons for table 2 - max sensor glucose
a9 <- aov(alldata$max_sensor~alldata$group_category.factor)
a9t <- tidy(TukeyHSD(a9))
a9t <- a9t[,-1]

# pairwise comparisons for table 2 - excursions >140/day
a10 <- aov(alldata$excursions_over_140~alldata$group_category.factor)
a10t <- tidy(TukeyHSD(a10))
a10t <- a10t[,-1]

# pairwise comparisons for table 2 - excursions >140
a11 <- aov(alldata$excursions_over_140~alldata$group_category.factor)
a11t <- tidy(TukeyHSD(a11))
a11t <- a11t[,-1]

# pairwise comparisons for table 2 - time spent >140
a12 <- aov(alldata$min_spent_over_140~alldata$group_category.factor)
a12t <- tidy(TukeyHSD(a12))
a12t <- a12t[,-1]

# pairwise comparisons for table 2 - SD
a13 <- aov(alldata$standard_deviation ~alldata$group_category.factor)
a13t <- tidy(TukeyHSD(a13))
a13t <- a13t[,-1]

# pairwise comparisons for table 2 - MAGE
a14 <- aov(alldata$r_mage ~alldata$group_category.factor)
a14t <- tidy(TukeyHSD(a14))
a14t <- a14t[,-1]

# pairwise comparisons for table 2 - % time > 180
a15 <- aov(alldata$percent_time_over_180 ~alldata$group_category.factor)
a15t <- tidy(TukeyHSD(a15))
a15t <- a15t[,-1]

# pairwise comparisons for table 2 - % time > 200
a16 <- aov(alldata$percent_time_over_200 ~alldata$group_category.factor)
a16t <- tidy(TukeyHSD(a16))
a16t <- a16t[,-1]

# correlations
corr <- rcorr(as.matrix(alldata[, c(ogttvars,cgmvars)]), type = "pearson")
corr <- flattenCorrMatrix(corr$r, corr$P)
corr <- corr[corr$row %in% ogttvars & corr$column %in% cgmvars,]
```

```{r echo=FALSE}
# Pairwise comparisons
sigvars <- c("poc_a1c","isi","isi_cpep","odi","odi_cpep","iaucbase_glucose","iaucbase_cpep","iAUCcpep_iAUCgluc","iAUCins_iAUC_gluc","average_sensor","standard_deviation","max_sensor","average_auc_per_day","excursions_over_200","avg_excur_over_200_per_day","min_spent_over_200","excursions_over_140","avg_excur_over_140_per_day","percent_time_over_140","min_spent_over_140","mean_amp_glycemic_ex","percent_time_over_180","percent_time_over_200")
l <- list()
for (var in sigvars) {
  table <- TukeyHSD(aov(alldata[,var] ~ alldata$group_category.factor))
  table <- as.data.frame(table$`alldata$group_category.factor`)
  table <- round(table,3)
  l[[var]] <- table
}
```

```{r echo=FALSE}
# ROC analysis for diabetes
# use these variables: percent_time_over_140 percent_time_over_180 percent_time_over_200 excursions_over_200 average_sensor
# max_sensor standard_deviation cv r_mage
pct140_roc <- roc(diabetes~percent_time_over_140,data=alldata,quiet=T)
pct140_res <- coords(pct140_roc,"best",transpose=F,best.method = "closest.topleft")
pct140_res$auc <- as.numeric(pct140_roc$auc)
pct140_res <- apply(pct140_res,2,function(x){round(x,2)})

pct180_roc <- roc(diabetes~percent_time_over_180,data=alldata,quiet=T)
pct180_res <- coords(pct180_roc,"best",transpose=F,best.method = "closest.topleft")
pct180_res$auc <- as.numeric(pct180_roc$auc)
pct180_res <- apply(pct180_res,2,function(x){round(x,2)})

pct200_roc <- roc(diabetes~percent_time_over_200,data=alldata,quiet=T)
pct200_res <- coords(pct200_roc,"best",transpose=F,best.method = "closest.topleft")
pct200_res$auc <- as.numeric(pct200_roc$auc)
pct200_res <- apply(pct200_res,2,function(x){round(x,2)})

ex200_roc <- roc(diabetes~excursions_over_200,data=alldata,quiet=T)
ex200_res <- coords(ex200_roc,"best",transpose=F,best.method = "closest.topleft")
ex200_res$auc <- as.numeric(ex200_roc$auc)
ex200_res <- apply(ex200_res,2,function(x){round(x,2)})

average_roc <- roc(diabetes~average_sensor,data=alldata,quiet=T)
average_res <- coords(average_roc,"best",transpose=F,best.method = "closest.topleft")
average_res$auc <- as.numeric(average_roc$auc)
average_res <- apply(average_res,2,function(x){round(x,2)})

max_roc <- roc(diabetes~max_sensor,data=alldata,quiet=T)
max_res <- coords(max_roc,"best",transpose=F,best.method = "closest.topleft")
max_res$auc <- as.numeric(max_roc$auc)
max_res <- apply(max_res,2,function(x){round(x,2)})

sd_roc <- roc(diabetes~standard_deviation,data=alldata,quiet=T)
sd_res <- coords(sd_roc,"best",transpose=F,best.method = "closest.topleft")
sd_res$auc <- as.numeric(sd_roc$auc)
sd_res <- apply(sd_res,2,function(x){round(x,2)})

cv_roc <- roc(diabetes~cov,data=alldata,quiet=T)
cv_res <- coords(cv_roc,"best",transpose=F,best.method = "closest.topleft")
cv_res$auc <- as.numeric(cv_roc$auc)
cv_res <- apply(cv_res,2,function(x){round(x,2)})

mage_roc <- roc(diabetes~r_mage,data=alldata,quiet=T)
mage_res <- coords(mage_roc,"best",transpose=F,best.method = "closest.topleft")
mage_res$auc <- as.numeric(mage_roc$auc)
mage_res <- apply(mage_res,2,function(x){round(x,2)})

# compare calprotectin and CRP in upper and lower tertiles
tab_cal_crp_cov <-  CreateTableOne(vars = c("Calprotectin","C.Reactive.Protein"),strata="cov_tertile", 
                               data = alldata[alldata$cov_tertile==1 | alldata$cov_tertile==3,])
tab_cal_crp_cov <- print(tab_cal_crp_cov,printToggle = F)
tab_cal_crp_mage <-  CreateTableOne(vars = c("Calprotectin","C.Reactive.Protein"),strata="mage_tertile", 
                               data = alldata[alldata$mage_tertile==1 | alldata$mage_tertile==3,])
tab_cal_crp_mage <- print(tab_cal_crp_mage,printToggle = F)
tab_cal_crp_delta <-  CreateTableOne(vars = c("Calprotectin","C.Reactive.Protein"),strata="delta_bg_tertile", 
                               data = alldata[alldata$delta_bg_tertile==1 | alldata$delta_bg_tertile==3,])
tab_cal_crp_delta <- print(tab_cal_crp_delta,printToggle = F)

```

# Background

The purpose of this analysis is to examine associations between beta cell function and CGM variables in healthy controls and participants with cystic fibrosis with normal glucose tolerance, impaired glucose tolerance, and CFRD.

# Methods

Participants who had both CGM data and OGTT data with 0, 30, and 120 minute timepoints available were included in this analysis. Beta cell function outcomes were calculated as follows:

1) $ISI = (Insulin_{30} - Insulin_{0}) / (Glucose_{30} - Glucose_{0})$
2) $ISI_{Cpep} = (Cpeptide_{30} - Cpeptide_{0}) / (Glucose_{30} - Glucose_{0})$
3) $oDI = (1/Insulin_{0}) * ISI$
4) $oDI_{Cpep} = (1/Cpeptide_{0}) * ISI_{Cpep}$

Integrated AUC (iAUC) for glucose, insulin, C-peptide, and glucagon were calculated as the AUC with respect to the fasting value, using the trapezoidal method.

The distributions of all variables were examined prior to analysis.  Groups were compared using ANOVA or the Wilcoxon test for continuous variables, and the chi-square or Fisher's exact test for categorical variables.  Pearson's correlation coefficient was used to test the association between beta cell function outcomes and CGM outcomes.

ROC analysis was used to examine CGM variables associated with diabetes and to determine the cutoff that best discriminates between participants with and without CFRD.  Optimal cutoffs for each CGM metric determined by "closest to top left" algorithm:  $min((1 - sensitivities)^2 + (1 - specificities)^2)$.

# Results

There were `r npar` participants included in this analysis.  Table 1 shows baseline descriptive statistics by treatment group.

```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab1,caption="Table 1.  Descriptive statistics by treatment group.  Statistics given are N(%), mean±SD, or median (25th %ile, 75th %ile).  * denotes that the Wilcoxon test was used, and ** denotes that Fisher's exact test was used.") 
```
<br>


Descriptive statistics for beta cell function outcomes and CGM outcomes are shown in Table 2.
```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab2,caption="Table 2.  Descriptive statistics for beta cell function and CGM outcomes by treatment group.  Statistics given are N(%), mean±SD, or median (25th %ile, 75th %ile).  * denotes that the Wilcoxon test was used, and ** denotes that Fisher's exact test was used.")
```
<br>

Pairwise Comparisons for Table 2 using Tukey's Honestly Significant Difference
```{r echo=FALSE}
# Kable does not appear to work in a for loop for some reason.
kable(l[[1]],caption = sigvars[1])
kable(l[[2]],caption = sigvars[2])
kable(l[[3]],caption = sigvars[3])
kable(l[[4]],caption = sigvars[4])
kable(l[[5]],caption = sigvars[5])
kable(l[[6]],caption = sigvars[6])
kable(l[[7]],caption = sigvars[7])
kable(l[[8]],caption = sigvars[8])
kable(l[[9]],caption = sigvars[9])
kable(l[[10]],caption = sigvars[10])
kable(l[[11]],caption = sigvars[11])
kable(l[[12]],caption = sigvars[12])
kable(l[[13]],caption = sigvars[13])
kable(l[[14]],caption = sigvars[14])
kable(l[[15]],caption = sigvars[15])
kable(l[[16]],caption = sigvars[16])
kable(l[[17]],caption = sigvars[17])
kable(l[[18]],caption = sigvars[18])
kable(l[[19]],caption = sigvars[19])
kable(l[[20]],caption = sigvars[20])
kable(l[[21]],caption = sigvars[21])
kable(l[[22]],caption = sigvars[22])
kable(l[[23]],caption = sigvars[23])
```

Pearson correlation coefficients between beta cell function outcomes and CGM outcomes are shown in Table 3.
```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(corr,caption="Table 3.  Correlation between beta cell function outcomes and CGM outcomes.")
```
<br>

```{r, results='asis',tidy=TRUE, echo=FALSE}
plot(alldata$odi_cpep,alldata$max_sensor,xlab="oDI Cpep",ylab="Max sensor (mg/dL)")
abline(lm(alldata$max_sensor~alldata$odi_cpep))

plot(alldata$odi_cpep,alldata$percent_time_over_140,xlab="oDI Cpep",ylab="% time >140 mg/dL")
abline(lm(alldata$percent_time_over_140~alldata$odi_cpep))

plot(alldata$odi_cpep,alldata$r_mage,xlab="oDI Cpep",ylab="MAGE (mg/dL)")
abline(lm(alldata$r_mage~alldata$odi_cpep))

plot(alldata$Calprotectin,alldata$r_mage,xlab="Calprotectin",ylab="MAGE (mg/dL)")
abline(lm(alldata$r_mage~alldata$Calprotectin))

plot(alldata$Calprotectin,alldata$cov,xlab="Calprotectin",ylab="CV",
     xlim = c(0,max(alldata$Calprotectin,na.rm = T)),
     ylim = c(0,max(alldata$cov,na.rm = T)))
abline(lm(alldata$cov~alldata$Calprotectin))

plot(alldata$Calprotectin,alldata$delta_bg,xlab="Calprotectin",ylab="Delta BG")
abline(lm(alldata$delta_bg~alldata$Calprotectin))

plot(alldata$C.Reactive.Protein,alldata$r_mage,xlab="CRP",ylab="MAGE (mg/dL)")
abline(lm(alldata$r_mage~alldata$C.Reactive.Protein))

plot(alldata$C.Reactive.Protein,alldata$cv,xlab="CRP",ylab="CV")
abline(lm(alldata$cov~alldata$C.Reactive.Protein))

plot(alldata$C.Reactive.Protein,alldata$delta_bg,xlab="CRP",ylab="Delta BG")
abline(lm(alldata$delta_bg~alldata$C.Reactive.Protein))

plot(alldata$homa_ir,alldata$max_sensor,xlab="HOMA-IR",ylab="Max sensor")
abline(lm(alldata$max_sensor~alldata$homa_ir))

plot(alldata$homa_ir,alldata$average_sensor,xlab="HOMA-IR",ylab="Mean sensor")
abline(lm(alldata$average_sensor~alldata$homa_ir))

plot(alldata$homa_ir,alldata$avg_excur_over_200_per_day,xlab="HOMA-IR",ylab="Average excursions >200 per day")
abline(lm(alldata$avg_excur_over_200_per_day~alldata$homa_ir))

plot(alldata$homa_ir,alldata$percent_time_over_140,xlab="HOMA-IR",ylab="% time >140")
abline(lm(alldata$percent_time_over_140~alldata$homa_ir))

plot(alldata$homa_ir,alldata$percent_time_over_180,xlab="HOMA-IR",ylab="% time >180")
abline(lm(alldata$percent_time_over_180~alldata$homa_ir))

plot(alldata$homa_ir,alldata$percent_time_over_200,xlab="HOMA-IR",ylab="% time >200")
abline(lm(alldata$percent_time_over_200~alldata$homa_ir))

plot(alldata$homa_ir,alldata$standard_deviation,xlab="HOMA-IR",ylab="SD")
abline(lm(alldata$standard_deviation~alldata$homa_ir))

plot(alldata$homa_ir,alldata$r_mage,xlab="HOMA-IR",ylab="MAGE")
abline(lm(alldata$r_mage~alldata$homa_ir))

```

ROC analysis for % time >140 vs. CFRD is shown below.
```{r, results='asis',tidy=TRUE, echo=FALSE}
plot(pct140_roc)
kable(pct140_res) 

```
<br>

ROC analysis for % time >180 vs. CFRD  is shown below.
```{r, results='asis',tidy=TRUE, echo=FALSE}
plot(pct180_roc)
kable(pct180_res)

```
<br>

ROC analysis for % time >200 vs. CFRD  is shown below.

```{r, results='asis',tidy=TRUE, echo=FALSE}
plot(pct200_roc)
kable(pct200_res)

```
<br>

ROC analysis for excursions >200  vs. CFRD is shown below.

```{r, results='asis',tidy=TRUE, echo=FALSE}
plot(ex200_roc)
kable(ex200_res)

```
<br>

ROC analysis for average sensor glucose vs. CFRD  is shown below.

```{r, results='asis',tidy=TRUE, echo=FALSE}
plot(average_roc)
kable(average_res)

```
<br>

ROC analysis for maximum sensor glucose  vs. CFRD is shown below.

```{r, results='asis',tidy=TRUE, echo=FALSE}
plot(max_roc)
kable(max_res)

```
<br>

ROC analysis for SD vs. CFRD is shown below.

```{r, results='asis',tidy=TRUE, echo=FALSE}
plot(sd_roc)
kable(sd_res)

```
<br>

ROC analysis for CV vs. CFRD  is shown below.

```{r, results='asis',tidy=TRUE, echo=FALSE}
plot(cv_roc)
kable(cv_res)

```
<br>

ROC analysis for MAGE  vs. CFRD is shown below.

```{r, results='asis',tidy=TRUE, echo=FALSE}
plot(mage_roc)
kable(mage_res)

```
<br>

The tables below compare calprotectin and CRP in the lowest and highest tertiles of CV, MAGE, and delta BG.

```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab_cal_crp_cov,caption = "Calprotectin and CRP in the lowest (1) and highest (3) tertiles of CV.")
```
<br>

```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab_cal_crp_mage,caption = "Calprotectin and CRP in the lowest (1) and highest (3) tertiles of MAGE.")
```
<br>

```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab_cal_crp_delta,caption = "Calprotectin and CRP in the lowest (1) and highest (3) tertiles of delta BG.")
```
<br>


