---
title: "CGM in CF - Beta cell function outcomes"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r echo=FALSE,include=FALSE}
library(Hmisc)
library(knitr)
library(rspiro)
library(tableone)
library(broom)
library(pROC)
library(dplyr)
library(sjlabelled)
library(tableone)
library(glmnet)
#library(officedown)
```

```{r echo=FALSE, include=FALSE}
# read in OGTT data
source("S:\\Shared Projects\\Laura\\Laura Tim projects\\Christine Chan\\2019 ADA Abstracts\\Data_Cleaned\\ada_2019.R")

source("C:\\Users\\pylell\\Documents\\GitHub\\General-code\\temp_table1.r")
source("C:\\Users\\pylell\\Documents\\GitHub\\General-code\\01_functions.r")
source("C:\\Users\\pylell\\Documents\\GitHub\\General-code\\laura_functions.r")
source("C:\\Users\\pylell\\Documents\\GitHub\\General-code\\iAUC.r")

# handling duplicates - per Christine
# for 004CD, use second visit
data <- data[!data$subject_id=='004CD--1',]
# for 023, use second
data <- data[!data$subject_id=='023HC--1',]
# 059 use second
data <- data[!data$subject_id=='059--1',]
# 063 use second
data <- data[!data$subject_id=='063--1',]
# 080 use second
data <- data[!data$subject_id=='080--1',]
# 040 keep first
data <- data[!data$subject_id=='040.1--1',]

# find visits on insulin and exclude
insulin <- read.csv("S:\\Shared Projects\\Laura\\Laura Tim projects\\Christine Chan\\2019 ADA Abstracts\\Data_Raw\\insulin.csv")
data <- merge(data,insulin,by="subject_id",all.x = T,all.y = F)
data <- data[is.na(data$cf_insulin_use),]

# hardcode missing data per Christine
data[data$subject_id=="067--1",]$lab_ogtt_2_hour_glucose <- 131

# tanner stage
femtanner <- c("tanner_stage_female_breast___1","tanner_stage_female_breast___2","tanner_stage_female_breast___3","tanner_stage_female_breast___4","tanner_stage_female_breast___5" )
maletanner <- c("tanner_stage_male_testicul___1","tanner_stage_male_testicul___2","tanner_stage_male_testicul___3","tanner_stage_male_testicul___4","tanner_stage_male_testicul___5")
data$tanner[which(data$gender.factor == "Female")] <- 
  as.numeric(apply(data[which(data$gender.factor == "Female"),femtanner],1,which.max))
data$tanner[which(data$gender.factor == "Male")] <- 
  as.numeric(apply(data[which(data$gender.factor == "Male"),maletanner],1,which.max))
data$tanner <- as.factor(data$tanner)
# lung function - no Asians in this data set
# Ask Christine if height at the lung function visit should be added to REDCap
data$race_ethnicity <- as.numeric(data$race_ethnicity)
data$lung_race <- ifelse(data$race_ethnicity %in% 3:6,5,data$race_ethnicity)
predGLI <- pred_GLI(data$age_at_visit_1,(data$average_height/100),data$gender,data$lung_race,param=c("FEV1","FVC"))
data$fev_pred <- predGLI$pred.FEV1
data$fvc_pred <- predGLI$pred.FVC
data$fev_perc <- (data$raw_fev1/data$fev_pred)*100
data$fvc_perc <- (data$raw_fvc/data$fvc_pred)*100
# calculate beta cell function outcomes
data$isi <- (data$min_30_insulin - data$min_0_insulin) / (data$min_30_glucose-data$lab_ogtt_fasting)
data$isi_cpep <- (data$min_30_c_peptide - data$min_0_c_peptide) / (data$min_30_glucose-data$lab_ogtt_fasting)
data$odi <- (1/data$min_0_insulin) * data$isi
data$odi_cpep <- (1/data$min_0_c_peptide) * data$isi_cpep
glucvars <- c("lab_ogtt_fasting","min_10_glucose","min_20_glucose","min_30_glucose","min_90_glucose",
                    "min_150_glucose","min_180_glucose")
data$glucmean <- rowMeans(data[,glucvars])
insvars <- c("min_0_insulin","min_10_insulin","min_20_insulin","min_30_insulin","min_90_insulin",
             "min_150_insulin","min_180_insulin")
data$insmean <- rowMeans(data[,insvars])
data$matsuda <- 10000/(sqrt(data$lab_ogtt_fasting*data$min_0_insulin*data$glucmean*data$insmean))
data$homa_ir <- (data$lab_ogtt_fasting*data$min_0_insulin)/22.5

# also need the following
# iAUCg, iAUCins/iAUCg, iAUCcpep/iAUCg, iAUCglucagon
#glucose
cols_gluc = c("min_minus_10_glucose","lab_ogtt_fasting",
         "min_10_glucose","min_20_glucose","min_30_glucose",
         "lab_ogtt_1_hour_glucose","min_90_glucose",
         "lab_ogtt_2_hour_glucose","min_150_glucose",
         "min_180_glucose")
ogttdat <- data[,c(cols_gluc,"subject_id")]
ogttdat <- ogttdat[which(rowSums(!is.na(ogttdat[,cols_gluc]))>=4),]
ogttdat$iaucbase_glucose <- iAUC_glucose(ogttdat,ogttdat$lab_ogtt_fasting,type = "linear")
ogttdat <- ogttdat[,c("subject_id","iaucbase_glucose")]
data <- merge(data,ogttdat,by="subject_id",all.x=TRUE,all.y=TRUE)

#insulin
cols_ins = c("min_minus_10_insulin","min_0_insulin","min_10_insulin","min_20_insulin","min_30_insulin","min_60_insulin",
             "min_90_insulin","min_120_insulin","min_150_insulin","min_180_insulin")
ogttdat <- data[,c(cols_ins,"subject_id")]
ogttdat <- ogttdat[which(rowSums(!is.na(ogttdat[,cols_ins]))>=4),]
ogttdat$iaucbase_insulin <- iAUC_insulin(ogttdat,ogttdat$min_0_insulin,type = "linear")
ogttdat <- ogttdat[,c("subject_id","iaucbase_insulin")]
data <- merge(data,ogttdat,by="subject_id",all.x=TRUE,all.y=TRUE)

#C-peptide
cols_cpep = c("min_minus_10_c_peptide","min_0_c_peptide","min_10_c_peptide","min_20_c_peptide","min_30_c_peptide","min_60_c_peptide",
             "min_90_c_peptide","min_120_c_peptide","min_150_c_peptide","min_180_c_peptide")
ogttdat <- data[,c(cols_cpep,"subject_id")]
ogttdat <- ogttdat[which(rowSums(!is.na(ogttdat[,cols_cpep]))>=4),]
ogttdat$iaucbase_cpep <- iAUC_cpep(ogttdat,ogttdat$min_0_c_peptide,type = "linear")
ogttdat <- ogttdat[,c("subject_id","iaucbase_cpep")]
data <- merge(data,ogttdat,by="subject_id",all.x=TRUE,all.y=TRUE)

# glucagon
cols_glucagon = c("min_minus_10_glucagon","min_0_glucagon","min_10_glucagon","min_20_glucagon","min_30_glucagon","min_60_glucagon",
              "min_90_glucagon","min_120_glucagon","min_150_glucagon","min_180_glucagon")
ogttdat <- data[,c(cols_glucagon,"subject_id")]
ogttdat <- ogttdat[which(rowSums(!is.na(ogttdat[,cols_glucagon]))>=4),]
ogttdat <- ogttdat[which(!is.na(ogttdat[,"min_0_glucagon"])),]
ogttdat$iaucbase_glucagon <- iAUC_glucagon(ogttdat,ogttdat$min_0_glucagon,type = "linear")
ogttdat <- ogttdat[,c("subject_id","iaucbase_glucagon")]
data <- merge(data,ogttdat,by="subject_id",all.x=TRUE,all.y=TRUE)

data$iAUCins_iAUC_gluc <- data$iaucbase_insulin/data$iaucbase_glucose
data$iAUCcpep_iAUCgluc <- data$iaucbase_cpep/data$iaucbase_glucose

# add in CGM variables that were left out of the original dataset
extra <- read.csv("S:\\Shared Projects\\Laura\\Laura Tim projects\\Christine Chan\\2019 ADA Abstracts\\Data_Raw\\CGMInHealthyControls-CGMVariables_DATA_2020-04-30_1347.csv")
extra <- extra[,c("subject_id","percent_time_over_180","percent_time_over_200","cov")]
data <- merge(data,extra,by="subject_id",all.x=TRUE,all.y=FALSE)
# calculate delta BG
data$delta_bg <- data$max_sensor-data$min_sensor

# keep only those who have non-missing odi
alldata <- data[which(!is.na(data$odi) & data$odi != Inf),]
alldata <- data

# make factors
alldata$gender <- as.factor(alldata$gender)
alldata$g_tube_feeds <- as.factor(alldata$g_tube_feeds)
alldata$cf_pancreatic <- as.factor(alldata$cf_pancreatic)

# labels
label(alldata$age_at_visit_1)="Age"
label(alldata$race_ethnicity.factor)="Race/ethnicity"
label(alldata$gender)="Gender"
label(alldata$poc_a1c)="HbA1c"
label(alldata$g_tube_feeds)="G-tube feeds"
label(alldata$fev_perc)="FEV-1 percent"
label(alldata$fvc_perc)="FVC"
label(alldata$cf_pancreatic)="Pancreatic insufficiency"
label(alldata$tanner)="Tanner stage"
label(alldata$bmi_z_score)="First BMI-Z"
label(alldata$isi)="ISI"
label(alldata$isi_cpep)="ISI C-peptide"
label(alldata$odi)="oDI"
label(alldata$odi_cpep)="oDI C-peptide"
label(alldata$min_sensor)="Min sensor glucose"
label(alldata$max_sensor)="Max sensor glucose"
label(alldata$average_sensor)="Average sensor glucose"
label(alldata$average_auc_per_day)="Average AUC"
label(alldata$excursions_over_200)="Excursions>200"
label(alldata$avg_excur_over_140_per_day)="Excursions>140/day"
label(alldata$avg_excur_over_200_per_day)="Excursions>200/day"
label(alldata$excursions_over_140)="Excursions>140"
label(alldata$percent_time_under_60)="% time <60"
label(alldata$percent_time_under_70)="% time <70"
label(alldata$percent_time_over_140)="% time >140"
label(alldata$min_spent_under_60)="Time spent <60"
label(alldata$min_spent_under_70)="Time spent <70"
label(alldata$min_spent_over_200)="Time spent >200"
label(alldata$min_spent_over_140)="Time spent >140"
label(alldata$percent_time_over_180)="% time >180"
label(alldata$percent_time_over_200)="% time >200"
label(alldata$total_auc)="Total AUC"
label(alldata$standard_deviation)="SD"
label(alldata$mean_amp_glycemic_ex)="MAGE"
label(alldata$r_mage)="R-MAGE"
label(alldata$iaucbase_glucose)="iAUC glucose"
label(alldata$iaucbase_insulin)="iAUC insulin"
label(alldata$iaucbase_cpep)="iAUC C-peptide"
label(alldata$iaucbase_glucagon)="iAUC glucagon"
label(alldata$iAUCins_iAUC_gluc)="iAUC insulin/iAUC glucose"
label(alldata$iAUCcpep_iAUCgluc)="iAUC C-peptide/iAUC glucose"

# create new variable for diabetes vs. no diabetes
alldata$diabetes <- ifelse(alldata$group_category.factor=="Cystic Fibrosis Related Diabetes",1,0)

# find people missing 1 hr or 2 hr glucose but not both
keep <- alldata[(is.na(alldata$lab_ogtt_1_hour_glucose) | is.na(alldata$lab_ogtt_2_hour_glucose)),
                c("subject_id","lab_ogtt_1_hour_glucose","lab_ogtt_2_hour_glucose")]
keep <- keep[!(is.na(alldata$lab_ogtt_1_hour_glucose && is.na(alldata$lab_ogtt_2_hour_glucose))),]
write.csv(keep,"C:\\temp\\missing 1 hr or 2 hr.csv",na = "")
# per Christine, all of these people are NGT

# create new variable for prediabetes - 1 hr >=200 or 2 hr >= 140
alldata$prediabetes <- ifelse(alldata$lab_ogtt_1_hour_glucose>=200 | alldata$lab_ogtt_2_hour_glucose>=140,1,0)
alldata$prediabetes <- ifelse(alldata$prediabetes==1 & alldata$diabetes==1,0,
                              ifelse(alldata$prediabetes==1 & alldata$diabetes==0,1,0))
alldata[alldata$subject_id %in% keep$subject_id,]$prediabetes <- 0
alldata$prediabetes_or_diabetes <- ifelse(alldata$prediabetes==1 | alldata$diabetes==1,1,0)
  
  
# read in calprotectin and CRP
crp <- read.csv("S:\\Shared Projects\\Laura\\Laura Tim projects\\Christine Chan\\2019 ADA Abstracts\\Data_Raw\\CGMInHealthyControls-CalprotectinAndCRP_DATA_LABELS_2020-04-30_1855.csv")
crp$subject_id <- crp$Subject.ID
crp$Subject.ID <- NULL
alldata <- merge(alldata,crp,by="subject_id",all.x=TRUE,all.y=FALSE)
alldata$Calprotectin <- as.numeric(as.character(alldata$Calprotectin))
set_label(alldata$Calprotectin) <- "Calprotectin"
set_label(alldata$C.Reactive.Protein) <- "CRP"

# create tertiles of CV, MAGE, and delta BG
alldata$cov_tertile <- ntile(alldata$cov,3)
alldata$mage_tertile <- ntile(alldata$r_mage,3)
alldata$delta_bg_tertile <- ntile(alldata$delta_bg,3)

# fix infinite values
alldata[alldata$subject_id=="105--1",]$isi <- NA
alldata[alldata$subject_id=="105--1",]$isi_cpep <- NA
alldata[alldata$subject_id=="105--1",]$odi <- NA
alldata[alldata$subject_id=="105--1",]$odi_cpep <- NA

# also need to make separate dataset with those with fsOGTT/oDI
has_odi <- alldata[!is.na(alldata$odi),]

npar <- nrow(alldata)
npar_odi <- nrow(has_odi)

# export final data for Christine
write.csv(alldata,"S:\\Shared Projects\\Laura\\Laura Tim projects\\Christine Chan\\2019 ADA Abstracts\\Data_Cleaned\\export.csv")

# ANALYSIS

# table 1
tab1vars <- c("age_at_visit_1", "race_ethnicity.factor","gender.factor","poc_a1c","fev_perc","fvc_perc","g_tube_feeds.factor",
              "tanner","cf_pancreatic.factor","bmi_z_score")
#tab1vars <- c("age", "race_ethnicity","gender","poc_a1c","g_tube_feeds","fev1_percent","fvc","cf_pancreatic","tanner","first_bmiz")
tab1_all <- final_table(data=alldata,variables=tab1vars,group=alldata$group_category.factor,margin=2,single=0,ron=2)
tab1_odi <- final_table(data=has_odi,variables=tab1vars,group=has_odi$group_category.factor,margin=2,single=0,ron=2)

# pairwise comparisons for table 1 - HbA1c
a1 <- aov(alldata$poc_a1c~alldata$group_category.factor)
a1t <- TukeyHSD(a1)
a1t <- tidy(a1t)
a1t <- a1t[,-1]

# compare groups
ogttvars <- c("isi","isi_cpep","odi","odi_cpep","matsuda","homa_ir","iaucbase_glucose","iaucbase_insulin","iaucbase_cpep",
              "iaucbase_glucagon","iAUCins_iAUC_gluc","iAUCcpep_iAUCgluc")
has_odi[,ogttvars] <- apply(has_odi[,ogttvars],2,as.numeric)
cgmvars <- c("min_sensor","max_sensor","average_sensor","average_auc_per_day","excursions_over_200","avg_excur_over_140_per_day",
             "avg_excur_over_200_per_day","excursions_over_140","percent_time_under_60","percent_time_under_70",
             "percent_time_over_140","percent_time_over_180","percent_time_over_200",
             "min_spent_under_60","min_spent_under_70","min_spent_over_200","min_spent_over_140",
             "total_auc","standard_deviation","mean_amp_glycemic_ex","Calprotectin","C.Reactive.Protein")
tab2a <- final_table(data=alldata,variables=cgmvars,group=alldata$group_category.factor,margin=2,single=0,ron=2)
# not sure why this isn't running
#tab2b <- final_table(data=has_odi,variables=ogttvars,group=has_odi$group_category.factor,margin=2,single=0,ron=2)
tab2b <- CreateTableOne(data=has_odi,ogttvars,strata="group_category.factor")
tab2b <- print(tab2b,printToggle = F,varLabels=TRUE,missing=T)

# pairwise comparisons for table 2 - ISI
a2 <- aov(has_odi$isi~has_odi$group_category.factor)
a2t <- TukeyHSD(a2)
a2t <- tidy(a2t)
a2t <- a2t[,-1]

# pairwise comparisons for table 2 - ISI C-peptide
a3 <- aov(has_odi$isi_cpep~has_odi$group_category.factor)
a3t <- tidy(TukeyHSD(a3))
a3t <- a3t[,-1]

# pairwise comparisons for table 2 - oDI
a4 <- aov(has_odi$odi~has_odi$group_category.factor)
a4t <- tidy(TukeyHSD(a4))
a4t <- a4t[,-1]

# pairwise comparisons for table 2 - oDI C-pep
a5 <- aov(has_odi$odi~has_odi$group_category.factor)
a5t <- tidy(TukeyHSD(a5))
a5t <- a5t[,-1]

# pairwise comparisons for table 2 - iAUC glucose
a6 <- aov(has_odi$iaucbase_glucose~has_odi$group_category.factor)
a6t <- tidy(TukeyHSD(a6))
a6t <- a6t[,-1]

# pairwise comparisons for table 2 - iAUC insulin/iAUC glucose
a7 <- aov(has_odi$iAUCins_iAUC_gluc~has_odi$group_category.factor)
a7t <- tidy(TukeyHSD(a7))
a7t <- a7t[,-1]

# pairwise comparisons for table 2 - iAUC C-peptide/iAUC glucose
a8 <- aov(has_odi$iAUCcpep_iAUCgluc~has_odi$group_category.factor)
a8t <- tidy(TukeyHSD(a8))
a8t <- a8t[,-1]

# pairwise comparisons for table 2 - max sensor glucose
a9 <- aov(alldata$max_sensor~alldata$group_category.factor)
a9t <- tidy(TukeyHSD(a9))
a9t <- a9t[,-1]

# pairwise comparisons for table 2 - excursions >140/day
a10 <- aov(alldata$excursions_over_140~alldata$group_category.factor)
a10t <- tidy(TukeyHSD(a10))
a10t <- a10t[,-1]

# pairwise comparisons for table 2 - excursions >140
a11 <- aov(alldata$excursions_over_140~alldata$group_category.factor)
a11t <- tidy(TukeyHSD(a11))
a11t <- a11t[,-1]

# pairwise comparisons for table 2 - time spent >140
a12 <- aov(alldata$min_spent_over_140~alldata$group_category.factor)
a12t <- tidy(TukeyHSD(a12))
a12t <- a12t[,-1]

# pairwise comparisons for table 2 - SD
a13 <- aov(alldata$standard_deviation ~alldata$group_category.factor)
a13t <- tidy(TukeyHSD(a13))
a13t <- a13t[,-1]

# pairwise comparisons for table 2 - MAGE
a14 <- aov(alldata$r_mage ~alldata$group_category.factor)
a14t <- tidy(TukeyHSD(a14))
a14t <- a14t[,-1]

# pairwise comparisons for table 2 - % time > 180
a15 <- aov(alldata$percent_time_over_180 ~alldata$group_category.factor)
a15t <- tidy(TukeyHSD(a15))
a15t <- a15t[,-1]

# pairwise comparisons for table 2 - % time > 200
a16 <- aov(alldata$percent_time_over_200 ~alldata$group_category.factor)
a16t <- tidy(TukeyHSD(a16))
a16t <- a16t[,-1]

# correlations
corr <- rcorr(as.matrix(alldata[, c(ogttvars,cgmvars)]), type = "pearson")
corr <- flattenCorrMatrix(corr$r, corr$P)
corr <- corr[corr$row %in% ogttvars & corr$column %in% cgmvars,]
```

```{r echo=FALSE}
# Pairwise comparisons
sigvars <- c("poc_a1c","isi","isi_cpep","odi","odi_cpep","iaucbase_glucose","iaucbase_cpep","iAUCcpep_iAUCgluc","iAUCins_iAUC_gluc","average_sensor","standard_deviation","max_sensor","average_auc_per_day","excursions_over_200","avg_excur_over_200_per_day","min_spent_over_200","excursions_over_140","avg_excur_over_140_per_day","percent_time_over_140","min_spent_over_140","mean_amp_glycemic_ex","percent_time_over_180","percent_time_over_200")
l <- list()
for (var in sigvars) {
  table <- TukeyHSD(aov(alldata[,var] ~ alldata$group_category.factor))
  table <- as.data.frame(table$`alldata$group_category.factor`)
  table <- round(table,3)
  l[[var]] <- table
}
```

```{r echo=FALSE}
# ROC analysis for diabetes
# use these variables: percent_time_over_140 percent_time_over_180 percent_time_over_200 excursions_over_200 average_sensor
# max_sensor standard_deviation cv r_mage
pct140_roc <- roc(diabetes~percent_time_over_140,data=alldata,quiet=T)
pct140_res <- coords(pct140_roc,"best",transpose=F,best.method = "closest.topleft")
pct140_res$auc <- as.numeric(pct140_roc$auc)
pct140_res <- apply(pct140_res,2,function(x){round(x,2)})

pct180_roc <- roc(diabetes~percent_time_over_180,data=alldata,quiet=T)
pct180_res <- coords(pct180_roc,"best",transpose=F,best.method = "closest.topleft")
pct180_res$auc <- as.numeric(pct180_roc$auc)
pct180_res <- apply(pct180_res,2,function(x){round(x,2)})

pct200_roc <- roc(diabetes~percent_time_over_200,data=alldata,quiet=T)
pct200_res <- coords(pct200_roc,"best",transpose=F,best.method = "closest.topleft")
pct200_res$auc <- as.numeric(pct200_roc$auc)
pct200_res <- apply(pct200_res,2,function(x){round(x,2)})

ex200_roc <- roc(diabetes~excursions_over_200,data=alldata,quiet=T)
ex200_res <- coords(ex200_roc,"best",transpose=F,best.method = "closest.topleft")
ex200_res$auc <- as.numeric(ex200_roc$auc)
ex200_res <- apply(ex200_res,2,function(x){round(x,2)})

average_roc <- roc(diabetes~average_sensor,data=alldata,quiet=T)
average_res <- coords(average_roc,"best",transpose=F,best.method = "closest.topleft")
average_res$auc <- as.numeric(average_roc$auc)
average_res <- apply(average_res,2,function(x){round(x,2)})

max_roc <- roc(diabetes~max_sensor,data=alldata,quiet=T)
max_res <- coords(max_roc,"best",transpose=F,best.method = "closest.topleft")
max_res$auc <- as.numeric(max_roc$auc)
max_res <- apply(max_res,2,function(x){round(x,2)})

sd_roc <- roc(diabetes~standard_deviation,data=alldata,quiet=T)
sd_res <- coords(sd_roc,"best",transpose=F,best.method = "closest.topleft")
sd_res$auc <- as.numeric(sd_roc$auc)
sd_res <- apply(sd_res,2,function(x){round(x,2)})

cv_roc <- roc(diabetes~cov,data=alldata,quiet=T)
cv_res <- coords(cv_roc,"best",transpose=F,best.method = "closest.topleft")
cv_res$auc <- as.numeric(cv_roc$auc)
cv_res <- apply(cv_res,2,function(x){round(x,2)})

mage_roc <- roc(diabetes~r_mage,data=alldata,quiet=T)
mage_res <- coords(mage_roc,"best",transpose=F,best.method = "closest.topleft")
mage_res$auc <- as.numeric(mage_roc$auc)
mage_res <- apply(mage_res,2,function(x){round(x,2)})

# ROC analysis for prediabetes
# use these variables: percent_time_over_140 percent_time_over_180 percent_time_over_200 excursions_over_200 average_sensor
# max_sensor standard_deviation cv r_mage
pct140_roc_pre <- roc(prediabetes~percent_time_over_140,data=alldata,quiet=T)
pct140_res_pre <- coords(pct140_roc_pre,"best",transpose=F,best.method = "closest.topleft")
pct140_res_pre$auc <- as.numeric(pct140_roc_pre$auc)
pct140_res_pre <- apply(pct140_res_pre,2,function(x){round(x,2)})

pct180_roc_pre <- roc(prediabetes~percent_time_over_180,data=alldata,quiet=T)
pct180_res_pre <- coords(pct180_roc_pre,"best",transpose=F,best.method = "closest.topleft")
pct180_res_pre$auc <- as.numeric(pct180_roc_pre$auc)
pct180_res_pre <- apply(pct180_res_pre,2,function(x){round(x,2)})

pct200_roc_pre <- roc(prediabetes~percent_time_over_200,data=alldata,quiet=T)
pct200_res_pre <- coords(pct200_roc_pre,"best",transpose=F,best.method = "closest.topleft")
pct200_res_pre$auc <- as.numeric(pct200_roc_pre$auc)
pct200_res_pre <- apply(pct200_res_pre,2,function(x){round(x,2)})

ex200_roc_pre <- roc(prediabetes~excursions_over_200,data=alldata,quiet=T)
ex200_res_pre <- coords(ex200_roc_pre,"best",transpose=F,best.method = "closest.topleft")
ex200_res_pre$auc <- as.numeric(ex200_roc_pre$auc)
ex200_res_pre <- apply(ex200_res_pre,2,function(x){round(x,2)})

average_roc_pre <- roc(prediabetes~average_sensor,data=alldata,quiet=T)
average_res_pre <- coords(average_roc_pre,"best",transpose=F,best.method = "closest.topleft")
average_res_pre$auc <- as.numeric(average_roc_pre$auc)
average_res_pre <- apply(average_res_pre,2,function(x){round(x,2)})

max_roc_pre <- roc(prediabetes~max_sensor,data=alldata,quiet=T)
max_res_pre <- coords(max_roc_pre,"best",transpose=F,best.method = "closest.topleft")
max_res_pre$auc <- as.numeric(max_roc_pre$auc)
max_res_pre <- apply(max_res_pre,2,function(x){round(x,2)})

sd_roc_pre <- roc(prediabetes~standard_deviation,data=alldata,quiet=T)
sd_res_pre <- coords(sd_roc_pre,"best",transpose=F,best.method = "closest.topleft")
sd_res_pre$auc <- as.numeric(sd_roc_pre$auc)
sd_res_pre <- apply(sd_res_pre,2,function(x){round(x,2)})

cv_roc_pre <- roc(prediabetes~cov,data=alldata,quiet=T)
cv_res_pre <- coords(cv_roc_pre,"best",transpose=F,best.method = "closest.topleft")
cv_res_pre$auc <- as.numeric(cv_roc_pre$auc)
cv_res_pre <- apply(cv_res_pre,2,function(x){round(x,2)})

mage_roc_pre <- roc(prediabetes~r_mage,data=alldata,quiet=T)
mage_res_pre <- coords(mage_roc_pre,"best",transpose=F,best.method = "closest.topleft")
mage_res_pre$auc <- as.numeric(mage_roc_pre$auc)
mage_res_pre <- apply(mage_res_pre,2,function(x){round(x,2)})

# ROC analysis for prediabetes or diabetes
# use these variables: percent_time_over_140 percent_time_over_180 percent_time_over_200 excursions_over_200 average_sensor
# max_sensor standard_deviation cv r_mage
pct140_roc_pd <- roc(prediabetes_or_diabetes~percent_time_over_140,data=alldata,quiet=T)
pct140_res_pd <- coords(pct140_roc_pd,"best",transpose=F,best.method = "closest.topleft")
pct140_res_pd$auc <- as.numeric(pct140_roc_pd$auc)
pct140_res_pd <- apply(pct140_res_pd,2,function(x){round(x,2)})

pct180_roc_pd <- roc(prediabetes_or_diabetes~percent_time_over_180,data=alldata,quiet=T)
pct180_res_pd <- coords(pct180_roc_pd,"best",transpose=F,best.method = "closest.topleft")
pct180_res_pd$auc <- as.numeric(pct180_roc_pd$auc)
pct180_res_pd <- apply(pct180_res_pd,2,function(x){round(x,2)})

pct200_roc_pd <- roc(prediabetes_or_diabetes~percent_time_over_200,data=alldata,quiet=T)
pct200_res_pd <- coords(pct200_roc_pd,"best",transpose=F,best.method = "closest.topleft")
pct200_res_pd$auc <- as.numeric(pct200_roc_pd$auc)
pct200_res_pd <- apply(pct200_res_pd,2,function(x){round(x,2)})

ex200_roc_pd <- roc(prediabetes_or_diabetes~excursions_over_200,data=alldata,quiet=T)
ex200_res_pd <- coords(ex200_roc_pd,"best",transpose=F,best.method = "closest.topleft")
ex200_res_pd$auc <- as.numeric(ex200_roc_pd$auc)
ex200_res_pd <- apply(ex200_res_pd,2,function(x){round(x,2)})

average_roc_pd <- roc(prediabetes_or_diabetes~average_sensor,data=alldata,quiet=T)
average_res_pd <- coords(average_roc_pd,"best",transpose=F,best.method = "closest.topleft")
average_res_pd$auc <- as.numeric(average_roc_pd$auc)
average_res_pd <- apply(average_res_pd,2,function(x){round(x,2)})

max_roc_pd <- roc(prediabetes_or_diabetes~max_sensor,data=alldata,quiet=T)
max_res_pd <- coords(max_roc_pd,"best",transpose=F,best.method = "closest.topleft")
max_res_pd$auc <- as.numeric(max_roc_pd$auc)
max_res_pd <- apply(max_res_pd,2,function(x){round(x,2)})

sd_roc_pd <- roc(prediabetes_or_diabetes~standard_deviation,data=alldata,quiet=T)
sd_res_pd <- coords(sd_roc_pd,"best",transpose=F,best.method = "closest.topleft")
sd_res_pd$auc <- as.numeric(sd_roc_pd$auc)
sd_res_pd <- apply(sd_res_pd,2,function(x){round(x,2)})

cv_roc_pd <- roc(prediabetes_or_diabetes~cov,data=alldata,quiet=T)
cv_res_pd <- coords(cv_roc_pd,"best",transpose=F,best.method = "closest.topleft")
cv_res_pd$auc <- as.numeric(cv_roc_pd$auc)
cv_res_pd <- apply(cv_res_pd,2,function(x){round(x,2)})

mage_roc_pd <- roc(prediabetes_or_diabetes~r_mage,data=alldata,quiet=T)
mage_res_pd <- coords(mage_roc_pd,"best",transpose=F,best.method = "closest.topleft")
mage_res_pd$auc <- as.numeric(mage_roc_pd$auc)
mage_res_pd <- apply(mage_res_pd,2,function(x){round(x,2)})

# LASSO regression to predict prediabetes or diabetes based on CGM variables
# lassodf <- alldata[,c("subject_id","prediabetes_or_diabetes","percent_time_over_140", "percent_time_over_180", "percent_time_over_200","excursions_over_200", "average_sensor", "max_sensor", "standard_deviation", "cov", "r_mage")]
# lassodf <- as.data.frame(apply(lassodf[,-1],2,as.numeric))
# remove NAs
# lassodf <- na.omit(lassodf)
# X <- model.matrix(~ percent_time_over_140 + percent_time_over_180 + percent_time_over_200+excursions_over_200+average_sensor+max_sensor+standard_deviation+cov+r_mage,data=lassodf)

# find best model using cross validation
# set.seed(3654)
# cv.lasso <- cv.glmnet(X,lassodf$prediabetes_or_diabetes,alpha=1, family="binomial")
# fit the final model
# lassomod <- glmnet(X,lassodf$prediabetes_or_diabetes,alpha = 1,family = "binomial",lambda=cv.lasso$lambda.min)
# refit using regular regression
# logitmod <- glm(prediabetes_or_diabetes ~ percent_time_over_140 + average_sensor + max_sensor + 
#                   r_mage,data=alldata,family="binomial")
# logitroc <- roc(prediabetes_or_diabetes ~ percent_time_over_140 + average_sensor + max_sensor + 
#                   r_mage,data=alldata,quiet=T)
# logit_res <- coords(logitroc,"best",transpose=F,best.method = "closest.topleft")
# logit_res$auc <- as.numeric(logit_res$auc)
# logit_res <- apply(logit_res,2,function(x){round(x,2)})

# compare calprotectin and CRP in upper and lower tertiles
tab_cal_crp_cov <-  CreateTableOne(vars = c("Calprotectin","C.Reactive.Protein"),strata="cov_tertile", 
                               data = alldata[alldata$cov_tertile==1 | alldata$cov_tertile==3,])
tab_cal_crp_cov <- print(tab_cal_crp_cov,printToggle = F)
tab_cal_crp_mage <-  CreateTableOne(vars = c("Calprotectin","C.Reactive.Protein"),strata="mage_tertile", 
                               data = alldata[alldata$mage_tertile==1 | alldata$mage_tertile==3,])
tab_cal_crp_mage <- print(tab_cal_crp_mage,printToggle = F)
tab_cal_crp_delta <-  CreateTableOne(vars = c("Calprotectin","C.Reactive.Protein"),strata="delta_bg_tertile", 
                               data = alldata[alldata$delta_bg_tertile==1 | alldata$delta_bg_tertile==3,])
tab_cal_crp_delta <- print(tab_cal_crp_delta,printToggle = F)

```

# Background

The purpose of this analysis is to examine associations between beta cell function and CGM variables in healthy controls and participants with cystic fibrosis with normal glucose tolerance, impaired glucose tolerance, and CFRD.

In this version of the report, participants are excluded if they were on insulin at the time of the visit.

# Methods

Participants who had both CGM data and OGTT data with 0, 30, and 120 minute timepoints available were included in this analysis. Beta cell function outcomes were calculated as follows:

1) $ISI = (Insulin_{30} - Insulin_{0}) / (Glucose_{30} - Glucose_{0})$
2) $ISI_{Cpep} = (Cpeptide_{30} - Cpeptide_{0}) / (Glucose_{30} - Glucose_{0})$
3) $oDI = (1/Insulin_{0}) * ISI$
4) $oDI_{Cpep} = (1/Cpeptide_{0}) * ISI_{Cpep}$

Integrated AUC (iAUC) for glucose, insulin, C-peptide, and glucagon were calculated as the AUC with respect to the fasting value, using the trapezoidal method.

The distributions of all variables were examined prior to analysis.  Groups were compared using ANOVA or the Wilcoxon test for continuous variables, and the chi-square or Fisher's exact test for categorical variables.  Pearson's correlation coefficient was used to test the association between beta cell function outcomes and CGM outcomes.

ROC analysis was used to examine CGM variables associated with diabetes and to determine the cutoff that best discriminates between participants with and without CFRD.  Optimal cutoffs for each CGM metric determined by "closest to top left" algorithm:  $min((1 - sensitivities)^2 + (1 - specificities)^2)$.

# Results

There were `r npar` participants included in this analysis and `r npar_odi` participants with an fsOGTT.  Table 1a shows baseline descriptive statistics by group, and Table 1b shows baseline descriptive statistics by group for those with fsOGTTs.

```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab1_all,caption="Table 1a.  Descriptive statistics by group.  Statistics given are N(%), mean±SD, or median (25th %ile, 75th %ile).  * denotes that the Wilcoxon test was used, and ** denotes that Fisher's exact test was used.") 
```

\newpage

```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab1_odi,caption="Table 1b.  Descriptive statistics by group, for participants with fsOGTT.  Statistics given are N(%), mean±SD, or median (25th %ile, 75th %ile).  * denotes that the Wilcoxon test was used, and ** denotes that Fisher's exact test was used.") 
```
<br>

\newpage

Descriptive statistics for beta cell function outcomes and CGM outcomes are shown in Tables 2a and 2b.
```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab2a,caption="Table 2a.  Descriptive statistics for CGM outcomes by group.  Statistics given are N(%), mean±SD, or median (25th %ile, 75th %ile).  * denotes that the Wilcoxon test was used, and ** denotes that Fisher's exact test was used.")
```
<br>

\newpage

```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab2b,caption="Table 2b.  Descriptive statistics for beta cell function outcomes by group, for participants with fsOGTT.  Statistics given are N(%), mean±SD, or median (25th %ile, 75th %ile).  * denotes that the Wilcoxon test was used, and ** denotes that Fisher's exact test was used.")
```
<br>

\newpage

Pairwise Comparisons for Table 2 using Tukey's Honestly Significant Difference
```{r echo=FALSE}
# Kable does not appear to work in a for loop for some reason.
kable(l[[1]],caption = sigvars[1])
kable(l[[2]],caption = sigvars[2])
kable(l[[3]],caption = sigvars[3])
kable(l[[4]],caption = sigvars[4])
kable(l[[5]],caption = sigvars[5])
kable(l[[6]],caption = sigvars[6])
kable(l[[7]],caption = sigvars[7])
kable(l[[8]],caption = sigvars[8])
kable(l[[9]],caption = sigvars[9])
kable(l[[10]],caption = sigvars[10])
kable(l[[11]],caption = sigvars[11])
kable(l[[12]],caption = sigvars[12])
kable(l[[13]],caption = sigvars[13])
kable(l[[14]],caption = sigvars[14])
kable(l[[15]],caption = sigvars[15])
kable(l[[16]],caption = sigvars[16])
kable(l[[17]],caption = sigvars[17])
kable(l[[18]],caption = sigvars[18])
kable(l[[19]],caption = sigvars[19])
kable(l[[20]],caption = sigvars[20])
kable(l[[21]],caption = sigvars[21])
kable(l[[22]],caption = sigvars[22])
kable(l[[23]],caption = sigvars[23])
```

\newpage

Pearson correlation coefficients between beta cell function outcomes and CGM outcomes are shown in Table 3.
```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(corr,caption="Table 3.  Correlation between beta cell function outcomes and CGM outcomes.")
```
<br>

\newpage

```{r, results='asis',tidy=TRUE, echo=FALSE}
plot(alldata$odi_cpep,alldata$max_sensor,xlab="oDI Cpep",ylab="Max sensor (mg/dL)")
abline(lm(alldata$max_sensor~alldata$odi_cpep))

plot(alldata$odi_cpep,alldata$percent_time_over_140,xlab="oDI Cpep",ylab="% time >140 mg/dL")
abline(lm(alldata$percent_time_over_140~alldata$odi_cpep))

plot(alldata$odi_cpep,alldata$r_mage,xlab="oDI Cpep",ylab="MAGE (mg/dL)")
abline(lm(alldata$r_mage~alldata$odi_cpep))

plot(alldata$Calprotectin,alldata$r_mage,xlab="Calprotectin",ylab="MAGE (mg/dL)")
abline(lm(alldata$r_mage~alldata$Calprotectin))

plot(alldata$Calprotectin,alldata$cov,xlab="Calprotectin",ylab="CV",
     xlim = c(0,max(alldata$Calprotectin,na.rm = T)),
     ylim = c(0,max(alldata$cov,na.rm = T)))
abline(lm(alldata$cov~alldata$Calprotectin))

plot(alldata$Calprotectin,alldata$delta_bg,xlab="Calprotectin",ylab="Delta BG")
abline(lm(alldata$delta_bg~alldata$Calprotectin))

plot(alldata$C.Reactive.Protein,alldata$r_mage,xlab="CRP",ylab="MAGE (mg/dL)")
abline(lm(alldata$r_mage~alldata$C.Reactive.Protein))

plot(alldata$C.Reactive.Protein,alldata$cv,xlab="CRP",ylab="CV")
abline(lm(alldata$cov~alldata$C.Reactive.Protein))

plot(alldata$C.Reactive.Protein,alldata$delta_bg,xlab="CRP",ylab="Delta BG")
abline(lm(alldata$delta_bg~alldata$C.Reactive.Protein))

plot(alldata$homa_ir,alldata$max_sensor,xlab="HOMA-IR",ylab="Max sensor")
abline(lm(alldata$max_sensor~alldata$homa_ir))

plot(alldata$homa_ir,alldata$average_sensor,xlab="HOMA-IR",ylab="Mean sensor")
abline(lm(alldata$average_sensor~alldata$homa_ir))

plot(alldata$homa_ir,alldata$avg_excur_over_200_per_day,xlab="HOMA-IR",ylab="Average excursions >200 per day")
abline(lm(alldata$avg_excur_over_200_per_day~alldata$homa_ir))

plot(alldata$homa_ir,alldata$percent_time_over_140,xlab="HOMA-IR",ylab="% time >140")
abline(lm(alldata$percent_time_over_140~alldata$homa_ir))

plot(alldata$homa_ir,alldata$percent_time_over_180,xlab="HOMA-IR",ylab="% time >180")
abline(lm(alldata$percent_time_over_180~alldata$homa_ir))

plot(alldata$homa_ir,alldata$percent_time_over_200,xlab="HOMA-IR",ylab="% time >200")
abline(lm(alldata$percent_time_over_200~alldata$homa_ir))

plot(alldata$homa_ir,alldata$standard_deviation,xlab="HOMA-IR",ylab="SD")
abline(lm(alldata$standard_deviation~alldata$homa_ir))

plot(alldata$homa_ir,alldata$r_mage,xlab="HOMA-IR",ylab="MAGE")
abline(lm(alldata$r_mage~alldata$homa_ir))

```

\newpage

ROC analysis for % time >140 vs. CFRD is shown below.
```{r, results='asis',tidy=TRUE, echo=FALSE}
plot(pct140_roc)
kable(pct140_res) 

```
<br>

\newpage

ROC analysis for % time >180 vs. CFRD  is shown below.
```{r, results='asis',tidy=TRUE, echo=FALSE}
plot(pct180_roc)
kable(pct180_res)

```
<br>

\newpage

ROC analysis for % time >200 vs. CFRD  is shown below.

```{r, results='asis',tidy=TRUE, echo=FALSE}
plot(pct200_roc)
kable(pct200_res)

```
<br>

\newpage

ROC analysis for excursions >200  vs. CFRD is shown below.

```{r, results='asis',tidy=TRUE, echo=FALSE}
plot(ex200_roc)
kable(ex200_res)

```
<br>

\newpage

ROC analysis for average sensor glucose vs. CFRD  is shown below.

```{r, results='asis',tidy=TRUE, echo=FALSE}
plot(average_roc)
kable(average_res)

```
<br>

\newpage

ROC analysis for maximum sensor glucose  vs. CFRD is shown below.

```{r, results='asis',tidy=TRUE, echo=FALSE}
plot(max_roc)
kable(max_res)

```
<br>

\newpage

ROC analysis for SD vs. CFRD is shown below.

```{r, results='asis',tidy=TRUE, echo=FALSE}
plot(sd_roc)
kable(sd_res)

```
<br>

\newpage

ROC analysis for CV vs. CFRD  is shown below.

```{r, results='asis',tidy=TRUE, echo=FALSE}
plot(cv_roc)
kable(cv_res)

```
<br>

\newpage

ROC analysis for MAGE  vs. CFRD is shown below.

```{r, results='asis',tidy=TRUE, echo=FALSE}
plot(mage_roc)
kable(mage_res)

```
<br>

\newpage

ROC analysis for % time >140 vs. prediabetes is shown below.
```{r, results='asis',tidy=TRUE, echo=FALSE}
plot(pct140_roc_pre)
kable(pct140_res_pre) 

```
<br>

\newpage

ROC analysis for % time >180 vs. prediabetes  is shown below.
```{r, results='asis',tidy=TRUE, echo=FALSE}
plot(pct180_roc_pre)
kable(pct180_res_pre)

```
<br>

\newpage

ROC analysis for % time >200 vs. prediabetes  is shown below.

```{r, results='asis',tidy=TRUE, echo=FALSE}
plot(pct200_roc_pre)
kable(pct200_res_pre)

```
<br>

\newpage

ROC analysis for excursions >200  vs. prediabetes is shown below.

```{r, results='asis',tidy=TRUE, echo=FALSE}
plot(ex200_roc_pre)
kable(ex200_res_pre)

```
<br>

\newpage

ROC analysis for average sensor glucose vs. prediabetes is shown below.

```{r, results='asis',tidy=TRUE, echo=FALSE}
plot(average_roc_pre)
kable(average_res_pre)

```
<br>

\newpage

ROC analysis for maximum sensor glucose  vs. prediabetes is shown below.

```{r, results='asis',tidy=TRUE, echo=FALSE}
plot(max_roc_pre)
kable(max_res_pre)

```
<br>

\newpage

ROC analysis for SD vs. prediabetes is shown below.

```{r, results='asis',tidy=TRUE, echo=FALSE}
plot(sd_roc_pre)
kable(sd_res_pre)

```
<br>

\newpage

ROC analysis for CV vs. prediabetes  is shown below.

```{r, results='asis',tidy=TRUE, echo=FALSE}
plot(cv_roc_pre)
kable(cv_res_pre)

```
<br>

\newpage

ROC analysis for MAGE  vs. prediabetes is shown below.

```{r, results='asis',tidy=TRUE, echo=FALSE}
plot(mage_roc_pre)
kable(mage_res_pre)

```
<br>

\newpage

ROC analysis for % time >140 vs. prediabetes or diabetes is shown below.
```{r, results='asis',tidy=TRUE, echo=FALSE}
plot(pct140_roc_pd)
kable(pct140_res_pd) 

```
<br>

\newpage

ROC analysis for % time >180 vs. prediabetes or diabetes  is shown below.
```{r, results='asis',tidy=TRUE, echo=FALSE}
plot(pct180_roc_pd)
kable(pct180_res_pd)

```
<br>

\newpage

ROC analysis for % time >200 vs. prediabetes or diabetes  is shown below.

```{r, results='asis',tidy=TRUE, echo=FALSE}
plot(pct200_roc_pd)
kable(pct200_res_pd)

```
<br>

\newpage

ROC analysis for excursions >200  vs. prediabetes or diabetes is shown below.

```{r, results='asis',tidy=TRUE, echo=FALSE}
plot(ex200_roc_pd)
kable(ex200_res_pd)

```
<br>

\newpage

ROC analysis for average sensor glucose vs. prediabetes or diabetes is shown below.

```{r, results='asis',tidy=TRUE, echo=FALSE}
plot(average_roc_pd)
kable(average_res_pd)

```
<br>

\newpage

ROC analysis for maximum sensor glucose  vs. prediabetes or diabetes is shown below.

```{r, results='asis',tidy=TRUE, echo=FALSE}
plot(max_roc_pd)
kable(max_res_pd)

```
<br>

\newpage

ROC analysis for SD vs. prediabetes or diabetes is shown below.

```{r, results='asis',tidy=TRUE, echo=FALSE}
plot(sd_roc_pd)
kable(sd_res_pd)

```
<br>

\newpage

ROC analysis for CV vs. prediabetes or diabetes  is shown below.

```{r, results='asis',tidy=TRUE, echo=FALSE}
plot(cv_roc_pd)
kable(cv_res_pd)

```
<br>

\newpage

ROC analysis for MAGE  vs. prediabetes or diabetes is shown below.

```{r, results='asis',tidy=TRUE, echo=FALSE}
plot(mage_roc_pd)
kable(mage_res_pd)

```
<br>

\newpage


The tables below compare calprotectin and CRP in the lowest and highest tertiles of CV, MAGE, and delta BG.

```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab_cal_crp_cov,caption = "Calprotectin and CRP in the lowest (1) and highest (3) tertiles of CV.")
```
<br>

```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab_cal_crp_mage,caption = "Calprotectin and CRP in the lowest (1) and highest (3) tertiles of MAGE.")
```
<br>

```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab_cal_crp_delta,caption = "Calprotectin and CRP in the lowest (1) and highest (3) tertiles of delta BG.")
```
<br>


