---
title: "Aim 2 HIP study"
author: "Laura Pyle"
date: "October 24, 2018"
output: word_document
---


```{r setup, include=FALSE,results='hide',echo=FALSE,message=FALSE,warning=FALSE}
knitr::opts_chunk$set(echo = FALSE)


library(tidyr)
library(magrittr)
library(dplyr)
library(lubridate)
library(zoo) 
library(M3)
library(data.table)
library(Table1)
library(devtools)
library(Hmisc)
library(knitr)
library(lsmeans)
library(broom)

####################################################################################################
#Data Manipulation #
####################################################################################################

# read in code for creating dataset
#source(file='H:\\Endocrinology\\Kelsey\\Kesley HIP study ISS puberty\\Ergui poster\\Data for models\\makedata.r')
source(file='C:\\Users\\pylell\\Documents\\GitHub\\Kelsey-HIP\\makedata aim 2.r')


# use Claire's functions
#source(file='H:\\R\\Markdown sample\\01_functions.r')
source(file='C:\\Users\\pylell\\Documents\\GitHub\\Functions\\01_functions.r')


# first comparison will be randomized kids only
rand <- wide[!is.na(wide$Randomization.Group),]
rand$Randomization.Group <- as.factor(rand$Randomization.Group)

# set reference levels
rand$race_eth <- relevel(as.factor(rand$race_eth),ref='Hispanic')
rand$Randomization.Group <- relevel(as.factor(rand$Randomization.Group),ref='2')

# calculate deltas for secondary outcomes
rand$delta_lept <- as.numeric(rand$lept.3 - rand$lept.1)
rand$delta_adipo <- as.numeric(rand$adiponect_lv.3 - rand$adiponect_lv.1)
rand$delta_tchol <- as.numeric(rand$tcholes_lv.3 - rand$tcholes_lv.1)
rand$delta_hdl <- as.numeric(rand$hdl_lv.3 - rand$hdl_lv.1)
rand$delta_ldlmeth <- as.numeric(rand$ldl_meth_lv.3 - rand$ldl_meth_lv.1)
rand$delta_ldl <- as.numeric(rand$ldl_lv.3 - rand$ldl_lv.1)
rand$delta_tg <- as.numeric(rand$tg_lv.3 - rand$tg_lv.1)
rand$delta_crp <- as.numeric(rand$crp_lv.3 - rand$crp_lv.1)
rand$delta_fat <- as.numeric(rand$fat_percentage_dexa.3 - rand$fat_percentage_dexa.1)
rand$delta_z <- as.numeric(rand$zscore_lv.3 - rand$zscore_lv.1)
rand$delta_alt <- as.numeric(rand$alt.3 - rand$alt.1)
rand$delta_ast <- as.numeric(rand$ast.3 - rand$ast.1)
rand$delta_dhea_s <- as.numeric(rand$dhea_s.3 - rand$dhea_s.1)
rand$delta_wc <- as.numeric(rand$wc_avg_lv.3 - rand$wc_avg_lv.1)
rand$delta_igf <- as.numeric(rand$igf_lv.3 - rand$igf_lv.1)
rand$delta_sbp <- as.numeric(rand$bp_systol_exam.3 - rand$bp_systol_exam.1)
rand$delta_dbp <- as.numeric(rand$bp_diastol_exam.3 - rand$bp_diastol_exam.1)

label(rand$delta_lept)="Change in leptin"
label(rand$delta_adipo)="Change in adiponectin"
label(rand$delta_tchol)="Change in total cholesterol"
label(rand$delta_hdl)="Change in HDL"
label(rand$delta_ldl)="Change in LDL"
label(rand$delta_tg)="Change in TG"
label(rand$delta_crp)="Change in CRP"
label(rand$delta_fat)="Change in % fat by DEXA"
label(rand$delta_z)="Change in BMI Z-score"
label(rand$delta_alt)="Change in ALT"
label(rand$delta_ast)="Change in AST"
label(rand$delta_dhea_s)="Change in DHEA_S"
label(rand$delta_wc)="Change in waist circumference"
label(rand$delta_igf)="Change in IGF"
label(rand$delta_sbp)="Change in SBP"
label(rand$delta_dbp)="Change in DBP"
label(rand$tanner)='Baseline Tanner stage'
label(rand$famhxt2d)="Family Hx T2D"
label(rand$base_bmi)="Baseline BMI"
label(rand$race_eth)="Race/ethnicity"
label(rand$age_base)="Age at baseline"

N <- nrow(rand)

####################################################################################################
#Table 1 #
####################################################################################################

tab1 <- final_table(rand,c("sex","race_eth","tanner","famhxt2d","zscore_lv.1","fat_percentage_dexa.1","base_bmi","age_base"),
                    group=rand$Randomization.Group,margin=2,single=F,ron=2)


####################################################################################################
#Unadjusted models #
####################################################################################################

tab_unadjusted <- final_table(rand,c('delta_si','delta_inssec','delta_di'),group=rand$Randomization.Group,margin=2,single=F,ron=2)
names(tab_unadjusted) <- c("Characteristic","All (n=23)","Placebo (n=12)","Metformin (n=11)","P Value") 

####################################################################################################
# Adjust for sex, Tanner, race #
####################################################################################################


adj2 <-lm(delta_si~Randomization.Group + as.factor(sex) + as.factor(race_eth) + as.factor(tanner), rand)
tab2.lsmeans.group <- lsmeans(adj2,"Randomization.Group")
tab2.lsmeans.group <- tidy(tab2.lsmeans.group)
tab2.lsmeans.group$Randomization.Group <- c("Placebo","Metformin")
tab2 <- mod_tab(adj2)
tab2$Predictor <- c("Group - Met vs. Placebo","Male vs. Female", "Black vs. Hispanic","Native American vs. Hispanic","Other race/ethnicity vs. Hispanic","Tanner stage 3 vs. stage 2")

adj3 <-lm(delta_inssec~Randomization.Group + as.factor(sex) + as.factor(race_eth) + as.factor(tanner), rand)
tab3.lsmeans.group <- lsmeans(adj3,"Randomization.Group")
tab3.lsmeans.group <- tidy(tab3.lsmeans.group)
tab3.lsmeans.group$Randomization.Group <- c("Placebo","Metformin")
tab3 <- mod_tab(adj3)
tab3$Predictor <- c("Group - Met vs. Placebo","Male vs. Female", "Black vs. Hispanic","Native American vs. Hispanic","Other race/ethnicity vs. Hispanic","Tanner stage 3 vs. stage 2")

adj4 <-lm(delta_di~Randomization.Group + as.factor(sex) + as.factor(race_eth) + as.factor(tanner), rand)
tab4.lsmeans.group <- lsmeans(adj4,"Randomization.Group")
tab4.lsmeans.group <- tidy(tab4.lsmeans.group)
tab4.lsmeans.group$Randomization.Group <- c("Placebo","Metformin")
tab4 <- mod_tab(adj4)
tab4$Predictor <- c("Group - Met vs. Placebo","Male vs. Female", "Black vs. Hispanic","Native American vs. Hispanic","Other race/ethnicity vs. Hispanic","Tanner stage 3 vs. stage 2")
 
####################################################################################################
# Adjust for sex, Tanner, race, change in percent body fat #
####################################################################################################

adj5 <-lm(delta_si~Randomization.Group + as.factor(sex) + as.factor(race_eth) + as.factor(tanner) + delta_bodyfat, wide)
tab5.lsmeans.group <- lsmeans(adj5,"Randomization.Group")
tab5 <- mod_tab(adj5)
tab5$Predictor <- c("Group - Met vs. Placebo","Male vs. Female", "Black vs. Hispanic","Native American vs. Hispanic","Other race/ethnicity vs. Hispanic","Tanner stage 3 vs. stage 2","Change in % body fat")

adj6 <-lm(delta_inssec~Randomization.Group + as.factor(sex) + as.factor(race_eth) + as.factor(tanner) + delta_bodyfat, wide)
tab6.lsmeans.group <- lsmeans(adj6,"Randomization.Group")
tab6 <- mod_tab(adj6)
tab6$Predictor <- c("Group - Met vs. Placebo","Male vs. Female", "Black vs. Hispanic","Native American vs. Hispanic","Other race/ethnicity vs. Hispanic","Tanner stage 3 vs. stage 2","Change in % body fat")

adj7 <-lm(delta_di~Randomization.Group + as.factor(sex) + as.factor(race_eth) + as.factor(tanner) + delta_bodyfat, wide)
tab7.lsmeans.group <- lsmeans(adj7,"Randomization.Group")
tab7 <- mod_tab(adj7)
tab7$Predictor <- c("Group - Met vs. Placebo","Male vs. Female", "Black vs. Hispanic","Native American vs. Hispanic","Other race/ethnicity vs. Hispanic","Tanner stage 3 vs. stage 2","Change in % body fat")

#################################################
# unadjusted comparisons for secondary outcomes
#################################################

tab_unadjusted_sec <- final_table(rand,c("delta_lept","delta_adipo","delta_tchol","delta_hdl",
                                         "delta_ldl","delta_tg",
                                         "delta_crp","delta_fat","delta_z","delta_alt",
                                         "delta_ast","delta_dhea_s","delta_wc","delta_igf",
                                          "delta_sbp","delta_dbp"),group=rand$Randomization.Group,margin=2,single=F,ron=2)
names(tab_unadjusted_sec) <- c("Characteristic","All (n=23)","Placebo (n=12)","Metformin (n=11)","P Value") 



```

# Background

This report presents the results of the analysis of aim 2 of the HIP study.  Changes in insulin sensitivity, secretion, and disposition index (DI) were compared in lean and obese participants, in unadjusted analyses, as well as in analyses adjusted for sex, baseline Tanner stage, race/ethnicity, and change in percent body fat by DXA.

Changes in secondary outcomes were also compared between the groups.

# Methods

Two participants (078-T and 140-T) who did not get randomized were excluded from analysis.  Visits with missing data for insulin sensitivity, insulin secretion, and DI were deleted.  

Changes in variables were calculated as the value at visit 3 minus the value at the first visit.  Any participants with missing values for change in insulin sensitivity, insulin secretion, or DI were excluded from analysis.

The distributions of all variables were examined prior to analysis.  T-tests and chi-square tests were used for unadjusted analyses, and linear models were used for adjusted analyses.

# Results

There were `r N` subjects included in this analysis.  Descriptive statistics are shown in Table 1.

```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab1,caption="Table 1.  Descriptive statistics by study group.  Variables with a single asterisk were compared using a non-parametric test; variables with two asterisks were compared using Fisher's exact test.")
```

There were no differences between the groups in the changes in the primary outcomes, in unadjusted analyses (Table 2).

```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab_unadjusted,caption="Table 2a.  Unadjusted changes in primary outcomes.  Variables with a single asterisk were compared using a non-parametric test; variables with two asterisks were compared using Fisher's exact test.")
```

There were also no differences between the groups in the changes in the primary outcomes, in analyses adjusted for sex, race/ethnicity, and baseline Tanner stage (Tables 3-5). \

```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab2,caption="Table 3a.  Change in insulin sensitivity, adjusting for sex, race/ethnicity, and baseline Tanner stage.")
```
\
```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab2.lsmeans.group,caption="Table 3b. Least squares means of the change in insulin sensitivity (visit 3-visit 1) by treatment group, in models adjusted for sex, race/ethnicity, and baseline Tanner stage.")
```
\
```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab3,caption="Table 4a.  Change in insulin secretion, adjusting for sex, race/ethnicity, and baseline Tanner stage.")
```
\
```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab3.lsmeans.group,caption="Table 4b. Least squares means of the change in insulin secretion (visit 3-visit 1) by treatment group, in models adjusted for sex, race/ethnicity, and baseline Tanner stage.")
```
\
```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab4,caption="Table 5a.  Change in DI, adjusting for sex, race/ethnicity, and baseline Tanner stage.")
```
\
```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab4.lsmeans.group,caption="Table 5b. Least squares means of the change in DI (visit 3-visit 1) by treatment group, in models adjusted for sex, race/ethnicity, and baseline Tanner stage.")
```

Table 6 presents the results of the comparisons of changes in secondary outcomes.
```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab_unadjusted_sec,caption="Table 6.  Changes in secondary outcomes.  Variables with a single asterisk were compared using a non-parametric test; variables with two asterisks were compared using Fisher's exact test.")
```
