---
title: "Aim 1 HIP study"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: word_document
---


```{r setup, include=FALSE,results='hide',echo=FALSE,message=FALSE,warning=FALSE}
knitr::opts_chunk$set(echo = FALSE)


setwd("H:\\Endocrinology\\Kelsey\\Kesley HIP study ISS puberty\\Ergui poster")

library(tidyr)
library(magrittr)
library(dplyr)
library(lubridate)
library(zoo) 
library(M3)
library(data.table)
library(Table1)
library(devtools)
library(Hmisc)
library(knitr)


####################################################################################################
#Data Manipulation #
####################################################################################################

# read in code for creating dataset
#source(file='H:\\Endocrinology\\Kelsey\\Kesley HIP study ISS puberty\\Ergui poster\\Data for models\\makedata.r')
source(file='C:\\Users\\pylell\\Documents\\GitHub\\Kelsey-HIP\\makedata aim 1.r')

# use Claire's functions
#source(file='H:\\R\\Markdown sample\\01_functions.r')
source(file='C:\\Users\\pylell\\Documents\\GitHub\\Functions\\01_functions.r')

N <- nrow(distinct(wide,hip_id_screen))

source("C:\\Users\\pylell\\Documents\\GitHub\\General-code\\temp_table1.r")

####################################################################################################
#Unadjusted models #
####################################################################################################

tab_unadjusted <- final_table(wide,c('delta_si','delta_inssec','delta_di'),group=wide$bmi_cat_final.factor,margin=2,single=0,ron=2)

####################################################################################################
# Adjust for sex, Tanner, race #
####################################################################################################

# need to fix reference group for race
wide$race_eth <- relevel(as.factor(wide$race_eth),ref='NHW')
adj2 <-lm(delta_si~bmi_cat_final.factor + sex.factor + as.factor(race_eth) + as.factor(tanner), wide)
tab2 <- mod_tab(adj2)
tab2$Predictor <- c("Obese vs. Lean","Male vs. Female", "Asian vs. NHW","Black vs. NHW","Hispanic vs. NHW","Other race/ethnicity vs. NHW",
                    "Tanner stage 3 vs. stage 2")

adj3 <-lm(delta_inssec~bmi_cat_final.factor + sex.factor + as.factor(race_eth) + as.factor(tanner), wide)
tab3 <- mod_tab(adj3)
tab3$Predictor <- c("Obese vs. Lean","Male vs. Female", "Asian vs. NHW","Black vs. NHW","Hispanic vs. NHW","Other race/ethnicity vs. NHW",
                    "Tanner stage 3 vs. stage 2")

adj4 <-lm(delta_di~bmi_cat_final.factor + sex.factor + as.factor(race_eth) + as.factor(tanner), wide)
tab4 <- mod_tab(adj4)
tab4$Predictor <- c("Obese vs. Lean","Male vs. Female", "Asian vs. NHW","Black vs. NHW","Hispanic vs. NHW","Other race/ethnicity vs. NHW",
                    "Tanner stage 3 vs. stage 2")

####################################################################################################
# Adjust for sex, Tanner, race, change in percent body fat #
####################################################################################################


adj5 <-lm(delta_si~bmi_cat_final.factor + sex.factor + as.factor(race_eth) + as.factor(tanner) + delta_bodyfat, wide)
tab5 <- mod_tab(adj5)
tab5$Predictor <- c("Obese vs. Lean","Male vs. Female", "Asian vs. NHW","Black vs. NHW","Hispanic vs. NHW","Other race/ethnicity vs. NHW",
                    "Tanner stage 3 vs. stage 2","Change in % body fat")

adj6 <-lm(delta_inssec~bmi_cat_final.factor + sex.factor + as.factor(race_eth) + as.factor(tanner) + delta_bodyfat, wide)
tab6 <- mod_tab(adj6)
tab6$Predictor <- c("Obese vs. Lean","Male vs. Female", "Asian vs. NHW","Black vs. NHW","Hispanic vs. NHW","Other race/ethnicity vs. NHW",
                    "Tanner stage 3 vs. stage 2","Change in % body fat")

adj7 <-lm(delta_di~bmi_cat_final.factor + sex.factor + as.factor(race_eth) + as.factor(tanner) + delta_bodyfat, wide)
tab7 <- mod_tab(adj7)
tab7$Predictor <- c("Obese vs. Lean","Male vs. Female", "Asian vs. NHW","Black vs. NHW","Hispanic vs. NHW","Other race/ethnicity vs. NHW",
                    "Tanner stage 3 vs. stage 2","Change in % body fat")

####################################################################################################
#Cross-sectional comparisons #
####################################################################################################

# Set Labels
label(wide$bmi_cat_final.factor)='BMI group'
label(wide$insulin_sensitivity.1)='Insulin sensitivity, visit 1'
label(wide$insulin_secretion_mm.1)='Insulin secretion, visit 1'
label(wide$disposition_index.1)='Disposition index, visit 1'
label(wide$insulin_sensitivity.3)='Insulin sensitivity, visit 3'
label(wide$insulin_secretion_mm.3)='Insulin secretion, visit 3'
label(wide$disposition_index.3)='Disposition index, visit 3'



tab8<- final_table(wide,c("insulin_sensitivity.1","insulin_sensitivity.3","insulin_secretion_mm.1",
                           "insulin_secretion_mm.3","disposition_index.1","disposition_index.3"),
                    group=wide$bmi_cat_final.factor,margin=2,single=0,ron=2)

####################################################################################################
#Predictors of change in insulin sensitivity #
####################################################################################################

tab9 <- NULL
mod1<-lm(delta_si~ relevel(as.factor(race_eth), ref="NHW"), wide)
F.Pval <- round(anova(mod1)$"Pr(>F)"[1],digits=4)
mod1 <-mod_tab(mod1)
mod1$PValue <- c(F.Pval,"","","")
tab9 <- rbind(tab9,mod1)
mod1<-lm(delta_si~ sex.factor, wide)
mod1 <-mod_tab(mod1)
tab9 <- rbind(tab9,mod1)
mod1<-lm(delta_si~as.numeric(lept0), wide)
mod1 <-mod_tab(mod1)
tab9 <- rbind(tab9,mod1)
mod1<-lm(delta_si~adipo0, wide)
mod1 <-mod_tab(mod1)
tab9 <- rbind(tab9,mod1)
mod1<-lm(delta_si~fat_perc0, wide)
mod1 <-mod_tab(mod1)
tab9 <- rbind(tab9,mod1)
mod1<-lm(delta_si~crp0, wide)
mod1 <-mod_tab(mod1)
tab9 <- rbind(tab9,mod1)
mod1<-lm(delta_si~tg0, wide)
mod1 <-mod_tab(mod1)
tab9 <- rbind(tab9,mod1)
mod1<-lm(delta_si~hdl0, wide)
mod1 <-mod_tab(mod1)
tab9 <- rbind(tab9,mod1)
mod1<-lm(delta_si~tg_hdl0, wide)
mod1 <-mod_tab(mod1)
tab9 <- rbind(tab9,mod1)
mod1<-lm(delta_si~mets0, wide)
mod1 <-mod_tab(mod1)
tab9 <- rbind(tab9,mod1)
mod1<-lm(delta_si~dheas0, wide)
mod1 <-mod_tab(mod1)
tab9 <- rbind(tab9,mod1)
mod1<-lm(delta_si~igf0, wide)
mod1 <-mod_tab(mod1)
tab9 <- rbind(tab9,mod1)
mod1<-lm(delta_si~gluc00, wide)
mod1 <-mod_tab(mod1)
tab9 <- rbind(tab9,mod1)
mod1<-lm(delta_si~hba1c0, wide)
mod1 <-mod_tab(mod1)
tab9 <- rbind(tab9,mod1)
mod1<-lm(delta_si~famhxt2d0, wide)
mod1 <-mod_tab(mod1)
tab9 <- rbind(tab9,mod1)
tab9$Predictor<-c("Asian","Black","Hispanic","Other race","Male","Baseline leptin","Baseline adiponectin","Baseline % fat","Baseline CRP","Baseline TG",
                  "Baseline HDL","Baseline TG:HDL","Baseline mets","Baseline DHEAS","Baseline IGF","Baseline fasting glucose","Baseline HbA1c","Family hx T2D")

####################################################################################################
# Are changes over time significant #
####################################################################################################

# Set Labels
label(wide$delta_si)='Change in Si'
label(wide$delta_inssec)='Change in insulin secretion'
label(wide$delta_di)='Change in DI'
wide$all <- "Total"

tab10 <- final_table(wide,c("delta_si","delta_inssec","delta_di"),margin=2,single=0,ron=2,group=wide$all)
t1 <- t.test(wide$delta_si)
t2 <- t.test(wide$delta_inssec)
t3 <- t.test(wide$delta_di)
PValue <- round(c(t1$p.value,t2$p.value,t3$p.value),4)
tab10 <- cbind(tab10,PValue)

####################################################################################################
#Predictors of change in insulin secretion #
####################################################################################################

tab11 <- NULL
mod1<-lm(delta_inssec~ relevel(as.factor(race_eth), ref="NHW"), wide)
F.Pval <- round(anova(mod1)$"Pr(>F)"[1],digits=4)
mod1 <-mod_tab(mod1)
mod1$PValue <- c(F.Pval,"","","")
tab11 <- rbind(tab11,mod1)
mod1<-lm(delta_inssec~ sex.factor, wide)
mod1 <-mod_tab(mod1)
tab11 <- rbind(tab11,mod1)
mod1<-lm(delta_inssec~as.numeric(lept0), wide)
mod1 <-mod_tab(mod1)
tab11 <- rbind(tab11,mod1)
mod1<-lm(delta_inssec~adipo0, wide)
mod1 <-mod_tab(mod1)
tab11 <- rbind(tab11,mod1)
mod1<-lm(delta_inssec~fat_perc0, wide)
mod1 <-mod_tab(mod1)
tab11 <- rbind(tab11,mod1)
mod1<-lm(delta_inssec~crp0, wide)
mod1 <-mod_tab(mod1)
tab11 <- rbind(tab11,mod1)
mod1<-lm(delta_inssec~tg0, wide)
mod1 <-mod_tab(mod1)
tab11 <- rbind(tab11,mod1)
mod1<-lm(delta_inssec~hdl0, wide)
mod1 <-mod_tab(mod1)
tab11 <- rbind(tab11,mod1)
mod1<-lm(delta_inssec~tg_hdl0, wide)
mod1 <-mod_tab(mod1)
tab11 <- rbind(tab11,mod1)
mod1<-lm(delta_inssec~mets0, wide)
mod1 <-mod_tab(mod1)
tab11 <- rbind(tab11,mod1)
mod1<-lm(delta_inssec~dheas0, wide)
mod1 <-mod_tab(mod1)
tab11 <- rbind(tab11,mod1)
mod1<-lm(delta_inssec~igf0, wide)
mod1 <-mod_tab(mod1)
tab11 <- rbind(tab11,mod1)
mod1<-lm(delta_inssec~gluc00, wide)
mod1 <-mod_tab(mod1)
tab11 <- rbind(tab11,mod1)
mod1<-lm(delta_inssec~hba1c0, wide)
mod1 <-mod_tab(mod1)
tab11 <- rbind(tab11,mod1)
mod1<-lm(delta_inssec~famhxt2d0, wide)
mod1 <-mod_tab(mod1)
tab11 <- rbind(tab11,mod1)
tab11$Predictor<-c("Asian","Black","Hispanic","Other race","Male","Baseline leptin","Baseline adiponectin","Baseline % fat","Baseline CRP","Baseline TG",
                  "Baseline HDL","Baseline TG:HDL","Baseline mets","Baseline DHEAS","Baseline IGF","Baseline fasting glucose","Baseline HbA1c","Family hx T2D")


####################################################################################################
#Predictors of change in odi #
####################################################################################################

tab12 <- NULL
mod1<-lm(delta_di~ relevel(as.factor(race_eth), ref="NHW"), wide)
F.Pval <- round(anova(mod1)$"Pr(>F)"[1],digits=4)
mod1 <-mod_tab(mod1)
mod1$PValue <- c(F.Pval,"","","")
tab12 <- rbind(tab12,mod1)
mod1<-lm(delta_di~ sex.factor, wide)
mod1 <-mod_tab(mod1)
tab12 <- rbind(tab12,mod1)
mod1<-lm(delta_di~as.numeric(lept0), wide)
mod1 <-mod_tab(mod1)
tab12 <- rbind(tab12,mod1)
mod1<-lm(delta_di~adipo0, wide)
mod1 <-mod_tab(mod1)
tab12 <- rbind(tab12,mod1)
mod1<-lm(delta_di~fat_perc0, wide)
mod1 <-mod_tab(mod1)
tab12 <- rbind(tab12,mod1)
mod1<-lm(delta_di~crp0, wide)
mod1 <-mod_tab(mod1)
tab12 <- rbind(tab12,mod1)
mod1<-lm(delta_di~tg0, wide)
mod1 <-mod_tab(mod1)
tab12 <- rbind(tab12,mod1)
mod1<-lm(delta_di~hdl0, wide)
mod1 <-mod_tab(mod1)
tab12 <- rbind(tab12,mod1)
mod1<-lm(delta_di~tg_hdl0, wide)
mod1 <-mod_tab(mod1)
tab12 <- rbind(tab12,mod1)
mod1<-lm(delta_di~mets0, wide)
mod1 <-mod_tab(mod1)
tab12 <- rbind(tab12,mod1)
mod1<-lm(delta_di~dheas0, wide)
mod1 <-mod_tab(mod1)
tab12 <- rbind(tab12,mod1)
mod1<-lm(delta_di~igf0, wide)
mod1 <-mod_tab(mod1)
tab12 <- rbind(tab12,mod1)
mod1<-lm(delta_di~gluc00, wide)
mod1 <-mod_tab(mod1)
tab12 <- rbind(tab12,mod1)
mod1<-lm(delta_di~hba1c0, wide)
mod1 <-mod_tab(mod1)
tab12 <- rbind(tab12,mod1)
mod1<-lm(delta_di~famhxt2d0, wide)
mod1 <-mod_tab(mod1)
tab12 <- rbind(tab12,mod1)
tab12$Predictor<-c("Asian","Black","Hispanic","Other race","Male","Baseline leptin","Baseline adiponectin","Baseline % fat","Baseline CRP","Baseline TG",
                  "Baseline HDL","Baseline TG:HDL","Baseline mets","Baseline DHEAS","Baseline IGF","Baseline fasting glucose","Baseline HbA1c","Family hx T2D")

```

# Background

This report presents the results of the analysis of aim 1 of the HIP study.  Changes in insulin sensitivity, secretion, and disposition index (DI) were compared in lean and obese participants, in unadjusted analyses, as well as in analyses adjusted for sex, baseline Tanner stage, race/ethnicity, and change in percent body fat by DXA.

# Methods

Participants randomized to metformin were excluded from this analysis.  Two participants (078-T and 140-T) who did not get randomized were excluded from analysis.  Visits with missing data for insulin sensitivity, insulin secretion, and DI were deleted.  

Changes in variables were calculated as the value at visit 3 minus the value at the first visit.  Any participants without visit 1 and visit 3 were therefore excluded from analysis.

The distributions of all variables were examined prior to analysis.  T-tests were used for unadjusted analyses, and linear models were used for adjusted analyses.

# Results

There were `r N` subjects included in this analysis.

There were no differences between the groups in the changes in the primary outcomes, in unadjusted analyses (Table 1).

```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab_unadjusted,caption="Table 1.  Unadjusted changes in primary outcomes.")
```

There were also no differences between the groups in the changes in the primary outcomes, in analyses adjusted for sex, race/ethnicity, and baseline Tanner stage (Tables 2-4). \

```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab2,caption="Table 2.  Change in insulin sensitivity, adjusting for sex, race/ethnicity, and baseline Tanner stage.")
```
\
```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab3,caption="Table 3.  Change in insulin secretion, adjusting for sex, race/ethnicity, and baseline Tanner stage.")
```
\
```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab4,caption="Table 4.  Change in DI, adjusting for sex, race/ethnicity, and baseline Tanner stage.")
```

There were also no differences between the groups in the changes in the primary outcomes, in analyses adjusted for sex, race/ethnicity, baseline Tanner stage and change in percent body fat (Tables 5-7).

```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab5,caption="Table 5.  Change in insulin sensitivity, adjusting for sex, race/ethnicity, and baseline Tanner stage.")
```
\
```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab6,caption="Table 6.  Change in insulin secretion, adjusting for sex, race/ethnicity, and baseline Tanner stage.")
```
\
```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab7,caption="Table 7.  Change in DI, adjusting for sex, race/ethnicity, and baseline Tanner stage.")
```

Table 8 presents results of cross-sectional group comparisons.

```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab8,caption="Table 8.  Cross-sectional group comparisons at visit 1 and visit 3.")
```

Table 9 presents the associations between baseline predictors and change in insulin sensitivity.

```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab9,caption="Table 9.  Associations between baseline predictors and change in insulin sensitivity.")
```

Table 10 gives summary statistics for changes in the primary outcomes.
```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab10,caption="Table 10.  Changes in insulin sensitivity, insulin secretion, and disposition index.")
```

Table 11 presents the associations between baseline predictors and change in insulin secretion.

```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab11,caption="Table 11.  Associations between baseline predictors and change in insulin secretion.")
```

Table 12 presents the associations between baseline predictors and change in DI.

```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab11,caption="Table 12.  Associations between baseline predictors and change in DI.")
```
