---
title: "HbA1c and fat depot analysis"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
library(tidyverse)
library(nlme)
library(knitr)
library(tidyr)
library(tableone)
```

```{r}
load("E:/Megan Kelsey/Kelsey HIP/Kelsey_HIP_R_Code/data/dat_analysis_1.RData")
```

```{r}
data <- dat_analysis_1 %>% dplyr::select(hip_id_screen,
                                 study_visit_number_svl,
                                 liver_fat_perc,
                                 lept,
                                 hba,
                                 delta_hba,
                                 tg_lv,
                                 igf_lv,
                                 hdl_lv,
                                 fasting_glucose,
                                 fasting_insulin,
                                 insulin_sensitivity,
                                 insulin_secretion_mm,
                                 disposition_index,
                                 visceral,
                                 per_visceral,
                                 subq_fat,
                                 adiponect_lv,
                                 crp_lv,
                                 alt,
                                 ast,
                                 delta_liver_fat_perc,
                                 study_phase,
                                 fat_mass_dexa) %>% 
  filter(!is.na(liver_fat_perc)) %>%  filter(!is.na(delta_liver_fat_perc))

# calculate leptin/adipo ratio and test association with A1c
data$lept_adipo <- data$lept/data$adiponect_lv

# fix study phase
data$study_phase <- droplevels(data$study_phase)
```

# Descriptive statistics

```{r, results='asis',tidy=TRUE, echo=FALSE, message=FALSE}
t1 <- CreateTableOne(vars=c("adiponect_lv","lept","lept_adipo"),data = data,strata = "study_phase")
t1 <- print(t1,printToggle = F,varLabels=FALSE,nonnorm=c("adiponect_lv","lept","lept_adipo"))
kable(t1,caption = "Descriptive statistics by group.  P-values are from the Mann-Whitney test.")
```

# HbA1c and leptin:adiponectin ratio

The linear mixed model below tests the association between HbA1c and adiponectin, using all visits.  

```{r, results='asis',tidy=TRUE, echo=FALSE, message=FALSE}
lm1 <- lme(hba~lept_adipo,random = ~1|hip_id_screen,data = data,na.action = na.omit)
kable(summary(lm1)$tTable)

ggplot(data = data, 
                      aes_string(y = "hba",
                                 x = "lept_adipo")) +
    geom_point() +
    geom_smooth(method = "lm")
```


```{r, results='asis',tidy=TRUE, echo=FALSE, message=FALSE}
# regress A1c at visit 3 on baseline leptin, and then sequentially adjust for fat depot variables at baseline
# need to restructure dataset
baseline <- data[data$study_visit_number_svl=="Tanner 2/3",]
fup <- data[data$study_visit_number_svl=="Tanner 5",]
baseline_keep <- baseline[,c("hip_id_screen","study_visit_number_svl","lept","visceral","liver_fat_perc",
                             "per_visceral","subq_fat","lept_adipo","fat_mass_dexa")]
colnames(baseline_keep) <- c("hip_id_screen","study_visit_number_svl","lept_base","visceral_base","liver_fat_perc_base",                             "per_visceral_base","subq_fat_base","lept_adipo_base","fat_mass_dexa_base")
baseline_keep$study_visit_number_svl <- NULL
fup_keep <- fup[,c("hip_id_screen","hba")]
colnames(fup_keep) <- c("hip_id_screen","hba_tanner5")

wide <- merge(baseline_keep,fup_keep,by="hip_id_screen",all.x=T, all.y = T)

lm2 <- lm(hba_tanner5 ~ lept_base, data=wide)
lm3 <- lm(hba_tanner5 ~ lept_adipo_base, data=wide)
lm4 <- lm(hba_tanner5 ~ lept_adipo_base + visceral_base, data=wide)
lm5 <- lm(hba_tanner5 ~ lept_adipo_base + liver_fat_perc_base, data=wide)
lm6 <- lm(hba_tanner5 ~ lept_adipo_base + per_visceral_base, data=wide)
lm7 <- lm(hba_tanner5 ~ lept_adipo_base + subq_fat_base, data=wide)
lm8 <- lm(hba_tanner5 ~ lept_adipo_base + fat_mass_dexa_base, data=wide)

```

# Tanner 5 HbA1c vs. baseline leptin and leptin:adiponectin ratio, adjusting for baseline fat depot

## HbA1c vs. leptin

```{r, results='asis',tidy=TRUE, echo=FALSE, message=FALSE}
kable(summary(lm2)$coefficients)
```

## HbA1c vs. leptin:adiponectin ratio

```{r, results='asis',tidy=TRUE, echo=FALSE, message=FALSE}
kable(summary(lm3)$coefficients)
```

## HbA1c vs. leptin:adiponectin ratio, adjusting for visceral fat

```{r, results='asis',tidy=TRUE, echo=FALSE, message=FALSE}
kable(summary(lm4)$coefficients)
```

## HbA1c vs. leptin:adiponectin ratio, adjusting for liver fat percentage

```{r, results='asis',tidy=TRUE, echo=FALSE, message=FALSE}
kable(summary(lm5)$coefficients)
```

## HbA1c vs. leptin:adiponectin ratio, adjusting for percent visceral fat

```{r, results='asis',tidy=TRUE, echo=FALSE, message=FALSE}
kable(summary(lm6)$coefficients)
```

## HbA1c vs. leptin:adiponectin ratio, adjusting for subcutaneous fat

```{r, results='asis',tidy=TRUE, echo=FALSE, message=FALSE}
kable(summary(lm7)$coefficients)
```

## HbA1c vs. leptin:adiponectin ratio, adjusting for total fat mass by DXA

```{r, results='asis',tidy=TRUE, echo=FALSE, message=FALSE}
kable(summary(lm8)$coefficients)
```
