---
title: "Pancreatic Fat Analysis"
author: "Cameron Severn & Laura Pyle"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: yes
    toc_float: yes
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
library(tidyverse)
library(nlme)
library(emmeans)
library(knitr)
library(arsenal)
library(Hmisc)
library(car)
library(ggpubr)
library(doParallel)
library(caret)
library(glmnet)
library(simputation)
```

# Summary

The following models explore the effect of puberty and Obesity on pancreatic fat. 


```{r}
setwd("E:/Megan Kelsey/Kelsey HIP/Kelsey_HIP_R_Code/data")

load("E:/Megan Kelsey/Kelsey HIP/Kelsey_HIP_R_Code/data/dat_analysis_1.RData")
load("E:/Megan Kelsey/Kelsey HIP/Kelsey_HIP_R_Code/data/dat_analysis_2.RData")
```

```{r}
dat_1 <- dat_analysis_1 
#dat_2 <- dat_analysis_2 %>% filter(!is.na(liver_fat_perc))
#dat <- merge(dat_1, dat_2, all = T)

pancfat <- read.csv("E:\\Megan Kelsey\\Kelsey HIP\\Data raw\\Final_PancreaticFat.csv")
pancfat$hip_id_screen <- word(pancfat$HIP.ID,1,sep="\\.")
pancfat$HIP.ID <- NULL
pancfat$pancreatic_fat_AVG <- pancfat$AVG
pancfat$AVG <- NULL
pancfat <- pancfat %>% filter(!is.na(pancreatic_fat_AVG))

dat <- merge(dat_analysis_1, pancfat, 
                        by=c("hip_id_screen","study_visit_number_svl"),all.x = T, all.y = T)
dat <- dat %>% filter(!is.na(pancreatic_fat_AVG))

#dat <- dat_2 %>% filter(study_phase != "Normal Weight")

dat$study_phase <- as.character(dat$study_phase)

#dat$study_phase[dat$study_phase == "Obese" | dat$study_phase == "Placebo"] <- "Obese/Placebo"
#dat$study_phase[dat$study_phase == "Metformin" | dat$study_phase == "Placebo"] <- "Obese"

dat$visc_subq_ratio <- dat$visceral / dat$subq_fat
label(dat$visc_subq_ratio)="Visc:Subq Ratio"

dat$per_subq <- dat$subq_fat / (dat$visceral/(dat$per_visceral/100)) * 100
label(dat$per_subq)="Percent Subq Fat"

dat$study_visit_number_svl <- factor(dat$study_visit_number_svl, levels = c("Tanner 2/3","Tanner 5"))

label(dat$lept_adipo_ratio)="Leptin:Adiponectin Ratio"

dat$age_y <- dat$age/365.25

label(dat$age_y)="Age (years)"


b <-pancfat %>% filter(study_visit_number_svl=="Tanner 2/3") %>% select(hip_id_screen,pancreatic_fat_AVG)
colnames(b) <- c("hip_id_screen","baseline_pancreatic_fat_AVG")
b <- b %>% filter(!is.na(baseline_pancreatic_fat_AVG))

t5 <-pancfat %>% filter(study_visit_number_svl=="Tanner 5") %>% select(hip_id_screen,pancreatic_fat_AVG)
colnames(t5) <- c("hip_id_screen","t5_pancreatic_fat_AVG")
t5 <- t5 %>% filter(!is.na(t5_pancreatic_fat_AVG))

dat <- merge(dat,b,by="hip_id_screen", all.x = T, all.y = T)
dat <- merge(dat,t5,by="hip_id_screen", all.x = T, all.y = T)
dat$delta_pancreatic_fat_AVG <- dat$t5_pancreatic_fat_AVG - dat$baseline_pancreatic_fat_AVG

dat <- dat %>% filter(!is.na(study_phase))

```

```{r, results="asis"}
control <- tableby.control(numeric.stats = c("N","meansd", "range"))

tab <- tableby(study_phase ~ 
                 age_y + 
                 sex_exam + 
                 race_eth + 
                 bmi_z +
                 bmi_screen_v1 +
                 bmi_pct_screen_v1 +
                 wc_avg_lv + 
                 tcholes_lv +
                 tg_lv +
                 ldl_lv +
                 hdl_lv +
                 adiponect_lv + 
                 crp_lv +
                 fasting_glucose +
                 fasting_insulin +
                 insulin_sensitivity +
                 hba +
                 lept + 
                 lept_adipo_ratio + 
                 pancreatic_fat_AVG +
                 per_subq +
                 per_visceral + 
                 fat_percentage_dexa +
                 ast +
                 alt
               ,
               strata = study_visit_number_svl, data = dat, control = control)
summary(tab)
```


# Mixed Models

```{r, results = "asis", warning = F, message=F}
outcomes <- c("pancreatic_fat_AVG",
              "visceral",
              "per_visceral",
              "subq_fat",
              "visc_subq_ratio",
              "fat_percentage_dexa",
              "lept",
              "ast",
              "alt")

for (outcome in outcomes){
  model <- lme(as.formula(paste(outcome, "~ study_visit_number_svl*study_phase")),
               random = ~1|hip_id_screen, 
               data=dat,
               correlation = corCompSymm(),
               na.action = na.omit)
  tTable <- as.data.frame(summary(model)$tTable)
  
  mm <- emmeans(model,~study_phase*study_visit_number_svl)
  model_means <- as.data.frame(mm)
  
  model_contrasts <- as.data.frame(pairs(mm, by = "study_visit_number_svl", reverse = T))
  
  model_contrasts_2 <- as.data.frame(pairs(mm, by = "study_phase", reverse = T))
  
  
  p <- ggplot() +
    geom_jitter(data = dat, aes_string(y = outcome, x = "study_visit_number_svl", color = "study_phase"), width = 0.1) +
    geom_point(data = model_means, aes(y = emmean, x = study_visit_number_svl, color = study_phase)) +
    geom_line(data = model_means, aes(y = emmean, x = study_visit_number_svl, color = study_phase, group = study_phase)) +
    geom_errorbar(data = model_means, aes(ymax = lower.CL, ymin = upper.CL, x = study_visit_number_svl, color = study_phase, group = study_phase), width = 0.1) +
    labs(x = "", color = "") +
    theme_light()
  
  
  names(tTable) <- c("Estimated Effect", "SE", "DF", "t-value", "p-value")
  tTable <- format(tTable, digits = 3)
  names(model_means) <- c("Group", "Tanner Stage", "Mean", "SE", "DF","95% CI lower","95% CI upper")
  model_means <- format(model_means, digits = 3)
  names(model_contrasts) <- c("Test", "Tanner Stage","Estimated Difference", "SE", "DF", "t-ratio", "p-value")
  model_contrasts <- format(model_contrasts, digits = 3)
  names(model_contrasts_2) <- c("Test", "Group","Estimated Difference", "SE", "DF", "t-ratio", "p-value")
  model_contrasts_2 <- format(model_contrasts_2, digits = 3)
  
  cat("\\newpage")
  cat(paste("##",outcome))
  cat("\n")
  print(p)
  print(kable(tTable, 
              caption = "T-Table",
              booktabs = TRUE,
              longtable = FALSE))
  print(kable(model_means, 
              caption = "Model Means",
              booktabs = TRUE,
              longtable = FALSE))
  cat("\\newpage")
  print(kable(model_contrasts, 
              caption = "Contrasts by Group",
              booktabs = TRUE,
              longtable = FALSE))
  print(kable(model_contrasts_2, 
              caption = "Contrasts by Tanner Stage",
              booktabs = TRUE,
              longtable = FALSE))
  cat("\n")
  
}
```


```{r}
fat_measures_baseline <- dat %>% filter(study_visit_number_svl == "Tanner 2/3") %>% 
  select(insulin_secretion_mm, 
         insulin_sensitivity, 
         tg_lv, 
         crp_lv, 
         hba, 
         igf_lv, 
         sex_exam, 
         race_eth, 
         pancreatic_fat_AVG,
         visceral,
         per_visceral,
         subq_fat,
         visc_subq_ratio,
         fat_percentage_dexa,
         lept,
         lept_adipo_ratio,
         estradiol,
         total_testosterone,
         shbg,
         fasting_glucose,
         fasting_insulin,
         study_phase
  )

write.csv(fat_measures_baseline,"pancreatic_fat_measures_baseline.csv", row.names = F)
```

```{r}
fat_measures_tanner5 <- dat %>% filter(study_visit_number_svl == "Tanner 5") %>% 
  select(insulin_secretion_mm, 
         insulin_sensitivity, 
         tg_lv, 
         crp_lv, 
         hba, 
         igf_lv, 
         sex_exam, 
         race_eth, 
         pancreatic_fat_AVG,
         visceral,
         per_visceral,
         subq_fat,
         visc_subq_ratio,
         fat_percentage_dexa,
         lept,
         lept_adipo_ratio,
         estradiol,
         total_testosterone,
         shbg,
         fasting_glucose,
         fasting_insulin
  )

write.csv(fat_measures_tanner5,"pancreatic_fat_measures_tanner5.csv", row.names = F)
```

```{r}
fat_measures_delta <- dat %>% filter(study_visit_number_svl == "Tanner 5") %>% 
  select(sex_exam, 
         race_eth, 
         baseline_insulin_secretion_mm, 
         baseline_insulin_sensitivity, 
         baseline_tg_lv, 
         baseline_crp_lv, 
         baseline_hba, 
         baseline_igf_lv,
         baseline_lept_adipo_ratio,
         baseline_pancreatic_fat_AVG,
         baseline_visceral,
         baseline_per_visceral,
         baseline_subq_fat,
         baseline_visc_subq_ratio,
         baseline_pancreatic_fat_AVG,
         baseline_lept,
         baseline_estradiol,
         baseline_total_testosterone,
         baseline_shbg,
         baseline_fasting_glucose,
         baseline_fasting_insulin,
         
         delta_pancreatic_fat_AVG,
         delta_visceral,
         delta_per_visceral,
         delta_subq_fat,
         delta_visc_subq_ratio,
         delta_fat_percentage_dexa,
         delta_lept
  )

write.csv(fat_measures_delta,"pancreatic_fat_measures_delta.csv", row.names = F)
```

```{r}
fat_measures_tanner5_w_baseline <- dat %>% filter(study_visit_number_svl == "Tanner 5") %>% 
  select(sex_exam, 
         race_eth, 
         baseline_insulin_secretion_mm, 
         baseline_insulin_sensitivity, 
         baseline_tg_lv, 
         baseline_crp_lv, 
         baseline_hba, 
         baseline_igf_lv,
         baseline_lept_adipo_ratio,
         baseline_pancreatic_fat_AVG,
         baseline_visceral,
         baseline_per_visceral,
         baseline_subq_fat,
         baseline_visc_subq_ratio,
         baseline_fat_percentage_dexa,
         baseline_lept,
         baseline_estradiol,
         baseline_total_testosterone,
         baseline_shbg,
         baseline_fasting_glucose,
         baseline_fasting_insulin,
         
         pancreatic_fat_AVG,
         visceral,
         per_visceral,
         subq_fat,
         visc_subq_ratio,
         fat_percentage_dexa,
         lept
  )

write.csv(fat_measures_tanner5_w_baseline,"pancreatic_fat_measures_tanner5_w_baseline.csv", row.names = F)
```

# Delta Models - Simple Group Means

These group means models test whether the change in each variable is significantly different from 0 (whether that variable changed over puberty) for normal weight and obese separately, but within the same model. Statistically, it is a simple linear regression with the intercept removed. Use this to say if a specific group went up or down throughout puberty while accounting for all observations. 

```{r, results = "asis", warning = F, message=F}
dat_t5 <- dat %>% filter(study_visit_number_svl == "Tanner 5")


outcomes <- c("pancreatic_fat_AVG",
              "visceral",
              "per_visceral",
              "subq_fat",
              "visc_subq_ratio",
              "fat_percentage_dexa",
              "lept",
              "ast",
              "alt")

for (outcome in outcomes){
  model <- lm(as.formula(paste0("delta_",outcome, " ~ ", "study_phase -1")),
              data=dat_t5,
              na.action = na.omit)
  tTable <- as.data.frame(summary(model)$coefficients)
  
  
  p <- ggplot(data = dat_t5, aes_string(x = paste0("delta_",outcome), fill = "study_phase")) +
    geom_density(alpha = 0.5)
  
  
  
  cat("\\newpage")
  cat(paste("##",outcome))
  cat("\n")
  print(p)
  print(kable(tTable, 
              caption = "T-Table",
              booktabs = TRUE,
              longtable = FALSE))
  
  cat("\n")
  
}

```

# Delta Models - T-tests

T-tests stratified by group. 

```{r, results = "asis", warning = F, message=F}
outcomes <- c("pancreatic_fat_AVG",
              "visceral",
              "per_visceral",
              "subq_fat",
              "visc_subq_ratio",
              "fat_percentage_dexa",
              "lept",
              "ast",
              "alt")


for (outcome in outcomes){
  nw_model <- t.test(as.formula(paste0("delta_",outcome,"~1")),
                     data=dat_t5[dat_t5$study_phase == "Normal Weight",],
                     na.action = na.omit)
  ob_model <- t.test(as.formula(paste0("delta_",outcome,"~1")),
                     data=dat_t5[dat_t5$study_phase == "Obese",],
                     na.action = na.omit)
  results <- data.frame(group = c("Normal Weight","Obese"), estimate = c(nw_model$estimate,ob_model$estimate),p = c(nw_model$p.value,ob_model$p.value))
  
  cat("\n")
  cat(paste("##",outcome))
  cat("\n")
  print(kable(results, 
              caption = "T-test Results",
              booktabs = TRUE,
              longtable = FALSE))
  
  cat("\n")
  
}

```

# Multivariable Models

Models are fitted using important variables selected by Elastic Net models.

## Predict Tanner 5 Pancreatic Fat with Tanner 5 Covariates

Variables Considered:

```{r}
names(fat_measures_tanner5)
```

### With obesity

```{r}
model <- lm(pancreatic_fat_AVG ~ 
              study_phase +
              lept + 
              lept_adipo_ratio +
              hba +
              crp_lv +
              insulin_secretion_mm +
              fasting_glucose +
              race_eth +
              shbg +
              per_visceral +
              visc_subq_ratio
              , 
            data = dat_t5, 
            na.action = na.omit)

Anova(model, type = 3)
summary(model)
```

### Without obesity

```{r}
model <- lm(pancreatic_fat_AVG ~ 
              lept + 
              lept_adipo_ratio +
              hba +
              crp_lv +
              insulin_secretion_mm +
              fasting_glucose +
              race_eth +
              shbg +
              per_visceral +
              visc_subq_ratio
              , 
            data = dat_t5, 
            na.action = na.omit)

Anova(model, type = 3)
summary(model)
```



## Predict Tanner 2/3 Pancreatic Fat with Tanner 2/3 Covariates

Variables Considered:

```{r}
names(fat_measures_baseline)
```

### with obesity

```{r}
model <- lm(pancreatic_fat_AVG ~ 
              study_phase +
              lept + 
              race_eth + 
              hba + 
              insulin_secretion_mm + 
              visceral + 
              lept_adipo_ratio + 
              tg_lv +
              estradiol, 
            data = fat_measures_baseline, 
            na.action = na.omit)

Anova(model, type = 3)
summary(model)
```

### without obesity

```{r}
model <- lm(pancreatic_fat_AVG ~ 
              lept + 
              race_eth + 
              hba + 
              insulin_secretion_mm + 
              visceral + 
              lept_adipo_ratio + 
              tg_lv +
              estradiol, 
            data = fat_measures_baseline, 
            na.action = na.omit)

Anova(model, type = 3)
summary(model)
```


## Predict Tanner 5 Pancreatic Fat with Tanner 2/3 Covariates

Variables Considered:

```{r}
names(fat_measures_tanner5_w_baseline[-c(22:28)])
```

### with obesity

```{r}
model <- lm(pancreatic_fat_AVG ~ 
              study_phase +
              baseline_pancreatic_fat_AVG + 
              baseline_fasting_insulin +
              race_eth +
              baseline_shbg +
              baseline_lept_adipo_ratio +
              baseline_visc_subq_ratio +
              baseline_crp_lv +
              baseline_visceral +
              sex_exam
            , 
            data = dat_t5, 
            na.action = na.omit)

Anova(model, type = 3)
summary(model)
```

### without obesity

```{r}
model <- lm(pancreatic_fat_AVG ~ 
              baseline_pancreatic_fat_AVG + 
              baseline_fasting_insulin +
              race_eth +
              baseline_shbg +
              baseline_lept_adipo_ratio +
              baseline_visc_subq_ratio +
              baseline_crp_lv +
              baseline_visceral +
              sex_exam
            , 
            data = fat_measures_tanner5_w_baseline, 
            na.action = na.omit)

Anova(model, type = 3)
summary(model)
```

```{r, warning=F, message=F}
plot_dat <- dat %>% filter(!is.na(study_visit_number_svl)) %>%
  select(hip_id_screen,
         study_phase,
         study_visit_number_svl,
         insulin_secretion_mm, 
         insulin_sensitivity, 
         tg_lv, 
         crp_lv, 
         hba, 
         igf_lv, 
         sex_exam, 
         race_eth, 
         pancreatic_fat_AVG,
         visceral,
         per_visceral,
         subq_fat,
         visc_subq_ratio,
         fat_percentage_dexa,
         lept,
         lept_adipo_ratio,
         estradiol,
         total_testosterone,
         shbg,
         ast,
         alt
  ) %>% 
  unique()

healthy <- plot_dat %>% filter(study_phase == "Normal Weight")

healthy <- healthy %>% filter((hip_id_screen %in% healthy$hip_id_screen[healthy$study_visit_number_svl == "Tanner 5"]) & (hip_id_screen %in% healthy$hip_id_screen[healthy$study_visit_number_svl == "Tanner 2/3"]))

point_alpha <- 0.3
line_alpha <- 0.3

h_just <- 1
y_just <- 1
bound_extend <- 0.35
label_radius <- unit(0, "lines")

model1 <- t.test(pancreatic_fat_AVG ~ study_visit_number_svl, data = healthy, paired = T)

p1 <- ggplot(data = healthy, aes(y = pancreatic_fat_AVG, x = study_visit_number_svl)) +
  geom_boxplot(outlier.alpha = 0) +
  geom_point(alpha = point_alpha) +
  geom_line(alpha = line_alpha, aes(group = hip_id_screen)) +
  theme_light() +
  geom_label(aes(x=Inf,y=Inf,hjust=h_just,vjust=y_just,
                 label = paste("Δ", signif(-model1$estimate, 
                                           digits= 3),
                               "p =",
                               signif(model1$p.value,2))),
             label.r = label_radius) +
  labs(y = "Pancreatic Fat (%)",  x = "") +
  ylim(range(healthy$pancreatic_fat_AVG)[1],
       range(healthy$pancreatic_fat_AVG)[2] + 
         bound_extend*(range(healthy$pancreatic_fat_AVG)[2] - 
                         range(healthy$pancreatic_fat_AVG)[1]))

model2 <- t.test(visceral ~ study_visit_number_svl, data = healthy, paired = T)

p2 <- ggplot(data = healthy, aes(y = visceral, x = study_visit_number_svl)) +
  geom_boxplot(outlier.alpha = 0) +
  geom_point(alpha = point_alpha) +
  geom_line(alpha = line_alpha, aes(group = hip_id_screen)) +
  theme_light() +
  geom_label(aes(x=Inf,y=Inf,hjust=h_just,vjust=y_just,
                 label = paste("Δ", signif(-model2$estimate, 
                                           digits= 3),
                               "p =",
                               signif(model2$p.value,2))),
             label.r = label_radius) +
  labs(y = expression("Visceral Fat"~(cm^2)), x = "") +
  ylim(range(healthy$visceral)[1],
       range(healthy$visceral)[2] + 
         bound_extend*(range(healthy$visceral)[2] - 
                         range(healthy$visceral)[1]))

model3 <- t.test(lept ~ study_visit_number_svl, data = healthy, paired = T)

p3 <- ggplot(data = healthy, aes(y = lept, x = study_visit_number_svl)) +
  geom_boxplot(outlier.alpha = 0) +
  geom_point(alpha = point_alpha) +
  geom_line(alpha = line_alpha, aes(group = hip_id_screen)) +
  theme_light() +
  geom_label(aes(x=Inf,y=Inf,hjust=h_just,vjust=y_just,
                 label = paste("Δ", signif(-model3$estimate, 
                                           digits= 3),
                               "p =",
                               signif(model3$p.value,2))),
             label.r = label_radius) +
  labs(y = "Leptin (ng/mL)", x = "") +
  ylim(range(healthy$lept)[1],
       range(healthy$lept)[2] + 
         bound_extend*(range(healthy$lept)[2] - 
                         range(healthy$lept)[1]))

model4 <- t.test(fat_percentage_dexa ~ study_visit_number_svl, data = healthy, paired = T)

p4 <- ggplot(data = healthy, aes(y = fat_percentage_dexa, x = study_visit_number_svl)) +
  geom_boxplot(outlier.alpha = 0) +
  geom_point(alpha = point_alpha) +
  geom_line(alpha = line_alpha, aes(group = hip_id_screen)) +
  theme_light() +
  geom_label(aes(x=Inf,y=Inf,hjust=h_just,vjust=y_just,
                 label = paste("Δ", signif(-model4$estimate, 
                                           digits= 3),
                               "p =",
                               signif(model4$p.value,2))),
             label.r = label_radius) +
  labs(y = "Body Fat by DEXA (%)", x = "") +
  ylim(range(healthy$fat_percentage_dexa)[1],
       range(healthy$fat_percentage_dexa)[2] + 
         bound_extend*(range(healthy$fat_percentage_dexa)[2] - 
                         range(healthy$fat_percentage_dexa)[1]))

model5 <- t.test(per_visceral ~ study_visit_number_svl, data = healthy, paired = T)

p5 <- ggplot(data = healthy, aes(y = per_visceral, x = study_visit_number_svl)) +
  geom_boxplot(outlier.alpha = 0) +
  geom_point(alpha = point_alpha) +
  geom_line(alpha = line_alpha, aes(group = hip_id_screen)) +
  theme_light() +
  geom_label(aes(x=Inf,y=Inf,hjust=h_just,vjust=y_just,
                 label = paste("Δ", signif(-model5$estimate, 
                                           digits= 3),
                               "p =",
                               signif(model5$p.value,2))),
             label.r = label_radius) +
  labs(y = "Visceral Fat (%)", x = "") +
  ylim(range(healthy$per_visceral)[1],
       range(healthy$per_visceral)[2] + 
         bound_extend*(range(healthy$per_visceral)[2] - 
                         range(healthy$per_visceral)[1]))

model6 <- t.test(ast ~ study_visit_number_svl, data = healthy, paired = T)

p6 <- ggplot(data = healthy, aes(y = ast, x = study_visit_number_svl)) +
  geom_boxplot(outlier.alpha = 0) +
  geom_point(alpha = point_alpha) +
  geom_line(alpha = line_alpha, aes(group = hip_id_screen)) +
  theme_light() +
  geom_label(aes(x=Inf,y=Inf,hjust=h_just,vjust=y_just,
                 label = paste("Δ", signif(-model6$estimate, 
                                           digits= 3),
                               "p =",
                               signif(model6$p.value,2))),
             label.r = label_radius) +
  labs(y = "AST (U/L)", x = "") +
  ylim(range(healthy$ast)[1],
       range(healthy$ast)[2] + 
         bound_extend*(range(healthy$ast)[2] - 
                         range(healthy$ast)[1]))

model7 <- t.test(subq_fat ~ study_visit_number_svl, data = healthy, paired = T)

p7 <- ggplot(data = healthy, aes(y = subq_fat, x = study_visit_number_svl)) +
  geom_boxplot(outlier.alpha = 0) +
  geom_point(alpha = point_alpha) +
  geom_line(alpha = line_alpha, aes(group = hip_id_screen)) +
  theme_light() +
  geom_label(aes(x=Inf,y=Inf,hjust=h_just,vjust=y_just,
                 label = paste("Δ", signif(-model7$estimate, 
                                           digits= 3),
                               "p =",
                               signif(model7$p.value,2))),
             label.r = label_radius) +
  labs(y = expression("Subcutaneous Fat"~(cm^2)), x = "") +
  ylim(range(healthy$subq_fat)[1],
       range(healthy$subq_fat)[2] + 
         bound_extend*(range(healthy$subq_fat)[2] - 
                         range(healthy$subq_fat)[1]))

model8 <- t.test(visc_subq_ratio ~ study_visit_number_svl, data = healthy, paired = T)

p8 <- ggplot(data = healthy, aes(y = visc_subq_ratio, x = study_visit_number_svl)) +
  geom_boxplot(outlier.alpha = 0) +
  geom_point(alpha = point_alpha) +
  geom_line(alpha = line_alpha, aes(group = hip_id_screen)) +
  theme_light() +
  geom_label(aes(x=Inf,y=Inf,hjust=h_just,vjust=y_just,
                 label = paste("Δ", signif(-model8$estimate, 
                                           digits= 3),
                               "p =",
                               signif(model8$p.value,2))),
             label.r = label_radius) +
  labs(y = "Visc/Subc Ratio", x = "") +
  ylim(range(healthy$visc_subq_ratio)[1],
       range(healthy$visc_subq_ratio)[2] + 
         bound_extend*(range(healthy$visc_subq_ratio)[2] - 
                         range(healthy$visc_subq_ratio)[1]))

model9 <- t.test(alt ~ study_visit_number_svl, data = healthy, paired = T)

p9 <- ggplot(data = healthy, aes(y = alt, x = study_visit_number_svl)) +
  geom_boxplot(outlier.alpha = 0) +
  geom_point(alpha = point_alpha) +
  geom_line(alpha = line_alpha, aes(group = hip_id_screen)) +
  theme_light() +
  geom_label(aes(x=Inf,y=Inf,hjust=h_just,vjust=y_just,
                 label = paste("Δ", signif(-model9$estimate, 
                                           digits= 3),
                               "p =",
                               signif(model9$p.value,2))),
             label.r = label_radius) +
  labs(y = "ALT (U/L)", x = "") +
  ylim(range(healthy$alt)[1],
       range(healthy$alt)[2] + 
         bound_extend*(range(healthy$alt)[2] - 
                         range(healthy$alt)[1]))

scalar <- 0.9
png(filename = "pancreatic_healthy_measures.png", res = 300, width = scalar*3000, height = scalar*2000)
ggarrange(p1,p2,p3,p4,p5,p6,p7,p8,p9,
          nrow = 3, ncol = 3)
dev.off()

# ggarrange(p1,p2,p3,p4,p5,p6,p7,p8,p9,
#           nrow = 3, ncol = 3)

```


```{r, warning=F, message=F}
obese <- plot_dat %>% filter(study_phase == "Obese")

obese <- obese %>% filter((hip_id_screen %in% obese$hip_id_screen[obese$study_visit_number_svl == "Tanner 5"]) & (hip_id_screen %in% obese$hip_id_screen[obese$study_visit_number_svl == "Tanner 2/3"]))

model1 <- t.test(pancreatic_fat_AVG ~ study_visit_number_svl, data = obese, paired = T)

p1 <- ggplot(data = obese, aes(y = pancreatic_fat_AVG, x = study_visit_number_svl)) +
  geom_boxplot(outlier.alpha = 0) +
  geom_point(alpha = point_alpha) +
  geom_line(alpha = line_alpha, aes(group = hip_id_screen)) +
  theme_light() +
  geom_label(aes(x=Inf,y=Inf,hjust=h_just,vjust=y_just,
                 label = paste("Δ", signif(-model1$estimate, 
                                           digits= 3),
                               "p =",
                               signif(model1$p.value,2))),
             label.r = label_radius) +
  labs(y = "Pancreatic Fat (%)",  x = "") +
  ylim(range(obese$pancreatic_fat_AVG)[1],
       range(obese$pancreatic_fat_AVG)[2] + 
         bound_extend*(range(obese$pancreatic_fat_AVG)[2] - 
                         range(obese$pancreatic_fat_AVG)[1]))

model2 <- t.test(visceral ~ study_visit_number_svl, data = obese, paired = T)

p2 <- ggplot(data = obese, aes(y = visceral, x = study_visit_number_svl)) +
  geom_boxplot(outlier.alpha = 0) +
  geom_point(alpha = point_alpha) +
  geom_line(alpha = line_alpha, aes(group = hip_id_screen)) +
  theme_light() +
  geom_label(aes(x=Inf,y=Inf,hjust=h_just,vjust=y_just,
                 label = paste("Δ", signif(-model2$estimate, 
                                           digits= 3),
                               "p =",
                               signif(model2$p.value,2))),
             label.r = label_radius) +
  labs(y = expression("Visceral Fat"~(cm^2)), x = "") +
  ylim(range(obese$visceral)[1],
       range(obese$visceral)[2] + 
         bound_extend*(range(obese$visceral)[2] - 
                         range(obese$visceral)[1]))

model3 <- t.test(lept ~ study_visit_number_svl, data = obese, paired = T)

p3 <- ggplot(data = obese, aes(y = lept, x = study_visit_number_svl)) +
  geom_boxplot(outlier.alpha = 0) +
  geom_point(alpha = point_alpha) +
  geom_line(alpha = line_alpha, aes(group = hip_id_screen)) +
  theme_light() +
  geom_label(aes(x=Inf,y=Inf,hjust=h_just,vjust=y_just,
                 label = paste("Δ", signif(-model3$estimate, 
                                           digits= 3),
                               "p =",
                               signif(model3$p.value,2))),
             label.r = label_radius) +
  labs(y = "Leptin (ng/mL)", x = "") +
  ylim(range(obese$lept)[1],
       range(obese$lept)[2] + 
         bound_extend*(range(obese$lept)[2] - 
                         range(obese$lept)[1]))

model4 <- t.test(fat_percentage_dexa ~ study_visit_number_svl, data = obese[obese$hip_id_screen != "160-T",], paired = T)

p4 <- ggplot(data = obese, aes(y = fat_percentage_dexa, x = study_visit_number_svl)) +
  geom_boxplot(outlier.alpha = 0) +
  geom_point(alpha = point_alpha) +
  geom_line(alpha = line_alpha, aes(group = hip_id_screen)) +
  theme_light() +
  geom_label(aes(x=Inf,y=Inf,hjust=h_just,vjust=y_just,
                 label = paste("Δ", signif(-model4$estimate, 
                                           digits= 3),
                               "p =",
                               signif(model4$p.value,2))),
             label.r = label_radius) +
  labs(y = "Body Fat by DEXA (%)", x = "") +
  ylim(range(obese$fat_percentage_dexa, na.rm = T)[1],
       range(obese$fat_percentage_dexa, na.rm = T)[2] + 
         bound_extend*(range(obese$fat_percentage_dexa, na.rm = T)[2] - 
                         range(obese$fat_percentage_dexa, na.rm = T)[1]))

model5 <- t.test(per_visceral ~ study_visit_number_svl, data = obese, paired = T)

p5 <- ggplot(data = obese, aes(y = per_visceral, x = study_visit_number_svl)) +
  geom_boxplot(outlier.alpha = 0) +
  geom_point(alpha = point_alpha) +
  geom_line(alpha = line_alpha, aes(group = hip_id_screen)) +
  theme_light() +
  geom_label(aes(x=Inf,y=Inf,hjust=h_just,vjust=y_just,
                 label = paste("Δ", signif(-model5$estimate, 
                                           digits= 3),
                               "p =",
                               signif(model5$p.value,2))),
             label.r = label_radius) +
  labs(y = "Visceral Fat (%)", x = "") +
  ylim(range(obese$per_visceral)[1],
       range(obese$per_visceral)[2] + 
         bound_extend*(range(obese$per_visceral)[2] - 
                         range(obese$per_visceral)[1]))

model6 <- t.test(ast ~ study_visit_number_svl, data = obese, paired = T)

p6 <- ggplot(data = obese, aes(y = ast, x = study_visit_number_svl)) +
  geom_boxplot(outlier.alpha = 0) +
  geom_point(alpha = point_alpha) +
  geom_line(alpha = line_alpha, aes(group = hip_id_screen)) +
  theme_light() +
  geom_label(aes(x=Inf,y=Inf,hjust=h_just,vjust=y_just,
                 label = paste("Δ", signif(-model6$estimate, 
                                           digits= 3),
                               "p =",
                               signif(model6$p.value,2))),
             label.r = label_radius) +
  labs(y = "AST (U/L)", x = "") +
  ylim(range(obese$ast)[1],
       range(obese$ast)[2] + 
         bound_extend*(range(obese$ast)[2] - 
                         range(obese$ast)[1]))

model7 <- t.test(subq_fat ~ study_visit_number_svl, data = obese, paired = T)

p7 <- ggplot(data = obese, aes(y = subq_fat, x = study_visit_number_svl)) +
  geom_boxplot(outlier.alpha = 0) +
  geom_point(alpha = point_alpha) +
  geom_line(alpha = line_alpha, aes(group = hip_id_screen)) +
  theme_light() +
  geom_label(aes(x=Inf,y=Inf,hjust=h_just,vjust=y_just,
                 label = paste("Δ", signif(-model7$estimate, 
                                           digits= 3),
                               "p =",
                               signif(model7$p.value,2))),
             label.r = label_radius) +
  labs(y = expression("Subcutaneous Fat"~(cm^2)), x = "") +
  ylim(range(obese$subq_fat)[1],
       range(obese$subq_fat)[2] + 
         bound_extend*(range(obese$subq_fat)[2] - 
                         range(obese$subq_fat)[1]))

model8 <- t.test(visc_subq_ratio ~ study_visit_number_svl, data = obese, paired = T)

p8 <- ggplot(data = obese, aes(y = visc_subq_ratio, x = study_visit_number_svl)) +
  geom_boxplot(outlier.alpha = 0) +
  geom_point(alpha = point_alpha) +
  geom_line(alpha = line_alpha, aes(group = hip_id_screen)) +
  theme_light() +
  geom_label(aes(x=Inf,y=Inf,hjust=h_just,vjust=y_just,
                 label = paste("Δ", signif(-model8$estimate, 
                                           digits= 3),
                               "p =",
                               signif(model8$p.value,2))),
             label.r = label_radius) +
  labs(y = "Visc/Subc Ratio", x = "") +
  ylim(range(obese$visc_subq_ratio)[1],
       range(obese$visc_subq_ratio)[2] + 
         bound_extend*(range(obese$visc_subq_ratio)[2] - 
                         range(obese$visc_subq_ratio)[1]))

model9 <- t.test(alt ~ study_visit_number_svl, data = obese, paired = T)

p9 <- ggplot(data = obese, aes(y = alt, x = study_visit_number_svl)) +
  geom_boxplot(outlier.alpha = 0) +
  geom_point(alpha = point_alpha) +
  geom_line(alpha = line_alpha, aes(group = hip_id_screen)) +
  theme_light() +
  geom_label(aes(x=Inf,y=Inf,hjust=h_just,vjust=y_just,
                 label = paste("Δ", signif(-model9$estimate, 
                                           digits= 3),
                               "p =",
                               signif(model9$p.value,2))),
             label.r = label_radius) +
  labs(y = "ALT (U/L)", x = "") +
  ylim(range(obese$alt)[1],
       range(obese$alt)[2] + 
         bound_extend*(range(obese$alt)[2] - 
                         range(obese$alt)[1]))

scalar <- 0.9
png(filename = "pancreatic_obese_measures.png", res = 300, width = scalar*3000, height = scalar*2000)
ggarrange(p1,p2,p3,p4,p5,p6,p7,p8,p9,
          nrow = 3, ncol = 3)
dev.off()

# ggarrange(p1,p2,p3,p4,p5,p6,p7,p8,p9,
#           nrow = 3, ncol = 3)

```
