---
title: "CROCODILE scRNA Correlations"
author: "Ye Ji Choi & Laura Pyle"
date: "today"
date-format: long
execute: 
  echo: false
  warning: false
  messages: false
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
editor: source
---
```{r libraries}
#| include: false
library(Seurat)
library(future)
library(tidyverse)
library(colorspace)
library(patchwork)
library(ggdendro)
library(cowplot)
library(ggpubr)
library(rstatix)
library(table1)
library(Biobase)
library(ReactomeGSA)
library(msigdbr)
library(kableExtra)
library(knitr)
```

```{r}
so <- readRDS("/home/yejichoi2/Documents/Local data/seurat_data_CRC.RDS")
source("/home/yejichoi2/Documents/CHCO-Code/Petter Bjornstad/CROCODILE/crocodile_functions.R")
```

```{r}
load("/run/user/1023/gvfs/smb-share:server=ucdenver.pvt,share=som/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/CROCODILE/Data_Cleaned/croc_data.RData")
dict <- read.csv("/run/user/1023/gvfs/smb-share:server=ucdenver.pvt,share=som/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/Data Harmonization/data_dictionary_master.csv")
source("/home/yejichoi2/Documents/CHCO-Code/Petter Bjornstad/Resources/YC/R Functions/label_harmonized_function.R")
source("/home/yejichoi2/Documents/CHCO-Code/Petter Bjornstad/Resources/YC/R Functions/correlation_function.R")
# so <- readRDS("/Volumes/Peds Endo/Petter Bjornstad/scRNA/data_clean/seurat_data_CRC.RDS")
# source("/Users/choiyej/GitHub/CHCO-Code/Petter Bjornstad/CROCODILE/crocodile_functions.R")
# load("/Volumes/Peds Endo/Petter Bjornstad/CROCODILE/Data_Cleaned/croc_data.RData")
# dict <- read.csv("/Volumes/Peds Endo/Petter Bjornstad/Data Harmonization/data_dictionary_master.csv")
# source("/Users/choiyej/GitHub/CHCO-Code/Petter Bjornstad/Resources/YC/R Functions/label_harmonized_function.R")
```

```{r}
de.markers(so, genes, "Group", id2 = "HC", id1 = "T1D", "PT", "")
de.markers(so, genes, "Group", id2 = "HC", id1 = "T1D", "PT", "_ir")

dict <- label_harmonized_dict(dat, dict)
# Modify some labels
dict <- dict %>%
  mutate(avg_c_k2 = "Average cortical k2 (global)",
         avg_c_k2_wo_cyst_vw = "Average cortical k2 (voxelwise)",
         gbm_thick_artmean = "GBM thickness (nm)",
         p2_raw_m = "M-Value (mg/kg/min)", 
         m_i = "M/I")

glyco_genes <- c("PKM", "LDHA", "LDHB", "HIF1A", "HK2", "PFKP")

gene_dict <- c(genes, acetylcoa_genes, oxy_phos_genes, ir_genes, leptin_adipo_genes, sglt2_genes, glyco_genes)
gene_dict_values <- unique(gene_dict)

for (gene_value in gene_dict_values) {
  dict[[gene_value]] <- gene_value
}


# Blood metabolomics
blood_metab <- c("citrate", "two_oxoglutarate", "succinate", 
                 "fumarate", "malate", "aspartate",
                 "glutamate", "alanine", "arginine", "cystine", "histidine", 
                 "leucine_isoleucine", "lysine", "methionine", "phenylalanine", 
                 "proline", "serine", "thereonine",
                 "tyrosine", "valine", "glutamine")
blood_metab_tca <- c("citrate", "succinate", "fumarate", "malate")

# Spatial metabolomics
spatial_metab <- dat %>% dplyr::select(ends_with("_tissue")) %>% colnames()
spatial_metab_tca <- c("isocitric_citric_acid_tissue",
                   "aconitic_acid_tissue",
                   "succinic_acid_tissue",
                   "fumaric_acid_tissue",
                   "malic_acid_tissue", 
                   "alpha_ketoglutaric_acid_tissue",
                   "oxaloacetic_acid_tissue")
```

```{r}
par(mfrow = c(2, 2))  

for (variable in blood_metab_tca) {
    hist(dat[[variable]], main = variable, xlab = variable)
}
```
```{r}
par(mfrow = c(2,2))  

for (variable in spatial_metab_tca) {
    hist(dat[[variable]], main = variable, xlab = variable)
}
```

## Extract individual level expression in scRNA
```{r warning = F}
meta_features <- so[["RNA"]]@meta.features
a <- rownames(meta_features)
var_list <- c("sample", "ident", "michigan_id", genes_subset)
pt_md <- FetchData(so, vars = var_list) %>%
  filter(ident == "PT")
ec_md <- FetchData(so, vars = var_list) %>%
  filter(ident == "EC")

pt_md <- pt_md %>%
  group_by(michigan_id) %>%
  dplyr::summarise(across(where(is.numeric), ~ sum(.x > 0, na.rm = TRUE))) %>%
  dplyr::rename(record_id = michigan_id) %>%
  right_join(dat)

ec_md <- ec_md %>%
  group_by(michigan_id) %>%
  dplyr::summarise(across(where(is.numeric), ~ sum(.x > 0, na.rm = TRUE))) %>%
  dplyr::rename(record_id = michigan_id) %>%
  right_join(dat)
```
# Blood metabolites descriptive
```{r message=F, warning=F, results='asis'}
summary(arsenal::tableby(group ~ citrate + succinate + kwt(fumarate, "Nmiss", "median", "q1q3", "range") + malate, data = pt_md, total = F))
```

# Associations & Correlations
## Spatial metabolomics & sc PT transcripts
### Plots
```{r warning = F, message = F}
# Plots
metabo_sc(data = pt_md, gene = "isocitric_citric_acid_tissue", 
          transcripts = c("ACO1", "ACO2", "IDH1"),
          gene_name = "Isocitric/citric acid")
metabo_sc(data = pt_md, gene = "aconitic_acid_tissue", 
          transcripts = c("ACO1", "ACO2"),
          gene_name = "Aconitate")
metabo_sc(data = pt_md, gene = "succinic_acid_tissue", 
          transcripts = c("OGDH", "OGDHL", "SUCLG1", "SUCLG2"),
          gene_name = "Succinate")
metabo_sc(data = pt_md, gene = "malic_acid_tissue", 
          transcripts = c("ACO1", "ACO2", "IDH1"),
          gene_name = "Malate")
```

### Table
#### Pearson
```{r, warning=F}
correlation_table_minimal(data = pt_md,
                   relevant_vars = c(spatial_metab_tca, genes_subset),
                   n_cols = length(spatial_metab_tca),
                   dict = dict)
```
#### Spearman
```{r, warning=F}
correlation_table_minimal(data = pt_md,
                   relevant_vars = c(spatial_metab_tca, genes_subset),
                   n_cols = length(spatial_metab_tca),
                   cor_method = "spearman",
                   dict = dict)
```

### Heatmap
#### Pearson
```{r}
png(height = 2000, width = 1200, file = "/home/yejichoi2/Documents/lin_temp_folder/Correlation_tissue_scRNA.png", res = 170)
corr_plot_modified(pt_md,
                   X = genes_subset,
                   Y = spatial_metab_tca,
                   adj_var = NULL,
                   dict = dict)
dev.off()
```
#### Spearman
```{r}
png(height = 2000, width = 1200, file = "/home/yejichoi2/Documents/lin_temp_folder/Correlation_tissue_scRNA_spearman.png", res = 170)
corr_plot_modified(pt_md,
                   X = genes_subset,
                   Y = spatial_metab_tca,
                   adj_var = NULL,
                   cor_method = "spearman",
                   dict = dict)
dev.off()
```

## Blood metabolites & sc PT transcripts
### Plots
```{r message=F, warning=F}
metabo_sc(data = pt_md, gene = "citrate", 
          transcripts = c("ACO1", "ACO2", "IDH1"),
          gene_name = "Citrate")

metabo_sc(data = pt_md, gene = "succinate", 
          transcripts = c("OGDH", "OGDHL", "SUCLG1", "SUCLG2"),
          gene_name = "Succinate")

metabo_sc(data = pt_md, gene = "malate", 
          transcripts = c("ACO1", "ACO2", "IDH1"),
          gene_name = "Malate")
```

### Table
#### Pearson
```{r message = F, warning=F}
correlation_table_minimal(data = pt_md,
                   relevant_vars = c(blood_metab_tca, genes_subset),
                   n_cols = length(blood_metab_tca),
                   dict = dict)
```
#### Spearman
```{r message = F, warning=F}
correlation_table_minimal(data = pt_md,
                   relevant_vars = c(blood_metab_tca, genes_subset),
                   n_cols = length(blood_metab_tca),
                   cor_method = "spearman",
                   dict = dict)
```

### Heatmap
#### Pearson
```{r}
png(height = 2000, width = 1200, file = "/home/yejichoi2/Documents/lin_temp_folder/Correlation_blood_scRNA.png", res = 170)
corr_plot_modified(pt_md,
                   X = genes_subset,
                   Y = blood_metab_tca,
                   adj_var = NULL,
                   dict = dict)
dev.off()
```
#### Spearman
```{r}
png(height = 2000, width = 1200, file = "/home/yejichoi2/Documents/lin_temp_folder/Correlation_blood_scRNA_spearman.png", res = 170)
corr_plot_modified(pt_md,
                   X = genes_subset,
                   Y = blood_metab_tca,
                   adj_var = NULL,
                   cor_method = "spearman",
                   dict = dict)
dev.off()
```

## Blood metabolites & spatial metabolomics
### Plots
```{r message=F, warning=F}
citrate_plot <- subset(pt_md, 
                       apply(!is.na(pt_md[, c(spatial_metab_tca[1], blood_metab_tca[1])]), 1, all)) %>%
      ggplot(aes_string(y = spatial_metab_tca[1], x = blood_metab_tca[1], color = "group")) +
      geom_smooth(method = lm, se = F, aes(color=NULL), color = "black",
                  linetype = "dashed") +
      geom_point() +
      geom_smooth(method = lm, se = F) +
      labs(color = "Group",
           y = "Isocitric/citric acid",
           x = "Citrate (plasma)") +
      theme_bw()

succinate_plot <- subset(pt_md, 
                         apply(!is.na(pt_md[, c(spatial_metab_tca[3], blood_metab_tca[2])]), 1, all)) %>%
      ggplot(aes_string(y = spatial_metab_tca[3], x = blood_metab_tca[2], color = "group")) +
      geom_smooth(method = lm, se = F, aes(color=NULL), color = "black",
                  linetype = "dashed") +
      geom_point() +
      geom_smooth(method = lm, se = F) +
      labs(color = "Group",
           y = "Succinic acid",
           x = "Succinate (plasma)") +
      theme_bw()

fumarate_plot <- subset(pt_md, 
                        apply(!is.na(pt_md[, c(spatial_metab_tca[4], blood_metab_tca[3])]), 1, all)) %>%
      ggplot(aes_string(y = spatial_metab_tca[4], x = blood_metab_tca[3], color = "group")) +
      geom_smooth(method = lm, se = F, aes(color=NULL), color = "black",
                  linetype = "dashed") +
      geom_point() +
      geom_smooth(method = lm, se = F) +
      labs(color = "Group",
           y = "Fumaric acid",
           x = "Fumarate (plasma)") +
      theme_bw()

malate_plot <- subset(pt_md, 
                      apply(!is.na(pt_md[, c(spatial_metab_tca[5], blood_metab_tca[4])]), 1, all)) %>%
      ggplot(aes_string(y = spatial_metab_tca[5], x = blood_metab_tca[4], color = "group")) +
      geom_smooth(method = lm, se = F, aes(color=NULL), color = "black",
                  linetype = "dashed") +
      geom_point() +
      geom_smooth(method = lm, se = F) +
      labs(color = "Group",
           y = "Malic acid",
           x = "Malate (plasma)") +
      theme_bw()
ggarrange(citrate_plot, succinate_plot, fumarate_plot, malate_plot, common.legend = T)
```

### Table
#### Pearson
```{r message = F, warning = F}
correlation_table_minimal(data = pt_md,
                   relevant_vars = c(blood_metab_tca, spatial_metab_tca),
                   n_cols = length(blood_metab_tca),
                   dict = dict)
```
#### Spearman
```{r message = F, warning = F}
correlation_table_minimal(data = pt_md,
                   relevant_vars = c(blood_metab_tca, spatial_metab_tca),
                   n_cols = length(blood_metab_tca),
                   cor_method = "spearman",
                   dict = dict)
```
### Heatmap
#### Pearson
```{r}
png(height = 1200, width = 1200, file = "/home/yejichoi2/Documents/lin_temp_folder/Correlation_blood_tissue.png", res = 170)
corr_plot_modified(pt_md,
                   X = spatial_metab_tca, 
                   Y = blood_metab_tca,
                   adj_var = NULL,
                   dict = dict)
dev.off()
```
#### Spearman
```{r}
png(height = 1200, width = 1200, file = "/home/yejichoi2/Documents/lin_temp_folder/Correlation_blood_tissue_spearman.png", res = 170)
corr_plot_modified(pt_md,
                   X = spatial_metab_tca, 
                   Y = blood_metab_tca,
                   adj_var = NULL,
                   cor_method = "spearman",
                   dict = dict)
dev.off()
```

## sc PT transcripts w clinical variables
### Plots
```{r message=F, warning=F}
## Log GBM width, and group the two groups
metabo_sc(data = pt_md, gene = "gbm_thick_artmean", 
          transcripts = genes_subset,
          gene_name = "GBM thickness (nm)")

## Global k2 & sc PT transcripts
metabo_sc(data = pt_md, gene = "avg_c_k2", 
          transcripts = genes_subset,
          gene_name = "Cortex k2")

## Voxelwise k2 & sc PT transcripts
metabo_sc(data = pt_md, gene = "avg_c_k2_wo_cyst_vw", 
          transcripts = genes_subset,
          gene_name = "Voxel-wise Cortex k2")

## Insulin sensitivity & sc PT transcripts
metabo_sc(data = pt_md, gene = "p2_raw_m", 
          transcripts = genes_subset,
          gene_name = "M-value")

metabo_sc(data = pt_md, gene = "p2_raw_m", 
          transcripts = genes_subset_ir,
          gene_name = "M-value")
```

### Table
#### TCA Cycle
##### Pooled
```{r}
correlation_table_minimal(data = pt_md,
                   relevant_vars = c("gbm_thick_artmean",
                                     "avg_c_k2",
                                     "hba1c",
                                     "m_i",
                                     genes_subset),
                   n_cols = 4,
                   dict = dict)
# with sglt2 genes
correlation_table_minimal(data = pt_md,
                   relevant_vars = c(genes_subset, 
                                     "hba1c",
                                     "m_i",
                                     sglt2_genes),
                   n_cols = 13,
                   dict = dict)
```
##### T1D only
```{r}
correlation_table_minimal(data = subset(pt_md, group == "Type 1 Diabetes"),
                   relevant_vars = c("gbm_thick_artmean",
                                     "avg_c_k2",
                                     "hba1c",
                                     "m_i",
                                     genes_subset),
                   n_cols = 4,
                   dict = dict)
# with sglt2 genes
correlation_table_minimal(data = subset(pt_md, group == "Type 1 Diabetes"),
                   relevant_vars = c(genes_subset, 
                                     "hba1c",
                                     "m_i",
                                     sglt2_genes),
                   n_cols = 13,
                   dict = dict)
```
#### Oxy phos
##### Pooled
```{r}
correlation_table_minimal(data = pt_md,
                   relevant_vars = c("gbm_thick_artmean",
                                     "avg_c_k2",
                                     "hba1c",
                                     "m_i",
                                     genes_subset_oxy_phos),
                   n_cols = 4,
                   dict = dict)
# with sglt2 genes
correlation_table_minimal(data = pt_md,
                   relevant_vars = c(genes_subset_oxy_phos, 
                                     "hba1c",
                                     "m_i",
                                     sglt2_genes),
                   n_cols = 7,
                   dict = dict)
```

##### T1D only
```{r}
correlation_table_minimal(data = subset(pt_md, group == "Type 1 Diabetes"),
                   relevant_vars = c("gbm_thick_artmean",
                                     "avg_c_k2",
                                     "hba1c",
                                     "m_i",
                                     genes_subset_oxy_phos),
                   n_cols = 4,
                   dict = dict)
# with sglt2 genes
correlation_table_minimal(data = subset(pt_md, group == "Type 1 Diabetes"),
                   relevant_vars = c(genes_subset_oxy_phos, 
                                     "hba1c",
                                     "m_i",
                                     sglt2_genes),
                   n_cols = 7,
                   dict = dict)
```


#### Glycolytic markers
##### Pooled
```{r}
correlation_table_minimal(data = pt_md,
                   relevant_vars = c("hba1c", "m_i",
                                     glyco_genes),
                   n_cols = 2,
                   dict = dict)
```
##### T1D only
```{r}
correlation_table_minimal(data = subset(pt_md, group == "Type 1 Diabetes"),
                   relevant_vars = c("hba1c", "m_i",
                                     glyco_genes),
                   n_cols = 2,
                   dict = dict)
```

#### SGLT2i
##### Pooled
```{r}
correlation_table_minimal(data = pt_md,
                   relevant_vars = c(glyco_genes, "hba1c", "m_i",
                                     sglt2_genes),
                   n_cols = 8,
                   dict = dict)
```

##### T1D only
```{r}
correlation_table_minimal(data = subset(pt_md, group == "Type 1 Diabetes"),
                   relevant_vars = c(glyco_genes, "hba1c", "m_i",
                                     sglt2_genes),
                   n_cols = 8,
                   dict = dict)
```

### Heatmap
#### Pooled
##### Pearson
```{r}
png(height = 2000, width = 1200, file = "/home/yejichoi2/Documents/lin_temp_folder/Correlation_scRNA_clin_t1dlc.png", res = 170)
corr_plot_modified(pt_md,
                   X = genes_subset, 
                   Y = c("gbm_thick_artmean",
                         "avg_c_k2"),
                   adj_var = NULL,
                   dict = dict)
dev.off()
```
##### Spearman
```{r}
png(height = 2000, width = 1200, file = "/home/yejichoi2/Documents/lin_temp_folder/Correlation_scRNA_clin_t1dlc_spearman.png", res = 170)
corr_plot_modified(pt_md,
                   X = genes_subset, 
                   Y = c("gbm_thick_artmean",
                         "avg_c_k2"),
                   adj_var = NULL,
                   cor_method = "spearman",
                   dict = dict)
dev.off()
```
#### T1D only
##### Pearson
```{r}
png(height = 2000, width = 1200, file = "/home/yejichoi2/Documents/lin_temp_folder/Correlation_scRNA_clin_t1d.png", res = 170)
corr_plot_modified(data = subset(pt_md, group == "Type 1 Diabetes"),
                   X = genes_subset, 
                   Y = c("gbm_thick_artmean",
                         "avg_c_k2"),
                   adj_var = NULL,
                   dict = dict)
dev.off()
```
##### Spearman
```{r}
png(height = 2000, width = 1200, file = "/home/yejichoi2/Documents/lin_temp_folder/Correlation_scRNA_clin_t1d_spearman.png", res = 170)
corr_plot_modified(data = subset(pt_md, group == "Type 1 Diabetes"),
                   X = genes_subset, 
                   Y = c("gbm_thick_artmean",
                         "avg_c_k2"),
                   adj_var = NULL,
                   cor_method = "spearman",
                   dict = dict)
dev.off()
```

## Spatial metabolomics & clinical variables
### Plots
#### Pooled
```{r message=F, warning=F}
### GBM width & spatial metabolomics
metabo_sc(data = label_harmonized(pt_md, dict), gene = "gbm_thick_artmean", 
          transcripts = spatial_metab_tca,
          gene_name = "GBM Thickness (nm)")

### Global k2 & spatial metabolomics
metabo_sc(data = label_harmonized(pt_md, dict), gene = "avg_c_k2", 
          transcripts = spatial_metab_tca,
          gene_name = "Cortex k2")

###  Voxelwise k2 & spatial metabolomics
metabo_sc(data = label_harmonized(pt_md, dict), gene = "avg_c_k2_wo_cyst_vw", 
          transcripts = spatial_metab_tca,
          gene_name = "Voxel-wise Cortex k2")
```

#### T1D only
```{r message=F, warning=F}
### GBM width & spatial metabolomics
metabo_sc(data = subset(label_harmonized(pt_md, dict), group == "Type 1 Diabetes"), 
          gene = "gbm_thick_artmean", 
          transcripts = spatial_metab_tca,
          gene_name = "GBM Thickness (nm)")

### Global k2 & spatial metabolomics
metabo_sc(data = subset(label_harmonized(pt_md, dict), group == "Type 1 Diabetes"), 
          gene = "avg_c_k2", 
          transcripts = spatial_metab_tca,
          gene_name = "Cortex k2")

### Voxelwise k2 & spatial metabolomics
metabo_sc(data = subset(label_harmonized(pt_md, dict), group == "Type 1 Diabetes"), 
          gene = "avg_c_k2_wo_cyst_vw", 
          transcripts = spatial_metab_tca,
          gene_name = "Voxel-wise Cortex k2")
```

```{r message=F, warning=F}
### GBM width & spatial metabolomics
metabo_sc(data = subset(label_harmonized(pt_md, dict), group == "Type 1 Diabetes"), 
          gene = "gbm_thick_artmean", 
          transcripts = spatial_metab_tca,
          gene_name = "GBM Thickness (nm)")

### Global k2 & spatial metabolomics
metabo_sc(data = subset(label_harmonized(pt_md, dict), group == "Type 1 Diabetes"), 
          gene = "avg_c_k2", 
          transcripts = spatial_metab_tca,
          gene_name = "Cortex k2")

### Voxelwise k2 & spatial metabolomics
metabo_sc(data = subset(label_harmonized(pt_md, dict), group == "Type 1 Diabetes"), 
          gene = "avg_c_k2_wo_cyst_vw", 
          transcripts = spatial_metab_tca,
          gene_name = "Voxel-wise Cortex k2")
```

### Table
#### Pooled
##### Pearson
```{r}
correlation_table_minimal(data = pt_md,
                   relevant_vars = c("gbm_thick_artmean",
                                     "avg_c_k2",
                                     spatial_metab_tca),
                   n_cols = 2,
                   dict = dict)
```
##### Spearman
```{r}
correlation_table_minimal(data = pt_md,
                   relevant_vars = c("gbm_thick_artmean",
                                     "avg_c_k2",
                                     spatial_metab_tca),
                   cor_method = "spearman",
                   n_cols = 2,
                   dict = dict)
```
#### T1D only
##### Pearson
```{r}
correlation_table_minimal(data = subset(pt_md, group == "Type 1 Diabetes"),
                   relevant_vars = c("gbm_thick_artmean",
                                     "avg_c_k2",
                                     spatial_metab_tca),
                   n_cols = 2,
                   dict = dict)
```
##### Spearman
```{r}
correlation_table_minimal(data = subset(pt_md, group == "Type 1 Diabetes"),
                   relevant_vars = c("gbm_thick_artmean",
                                     "avg_c_k2",
                                     spatial_metab_tca),
                   cor_method = "spearman",
                   n_cols = 2,
                   dict = dict)
```
### Heatmap
#### Pooled
##### Pearson
```{r}
png(height = 1200, width = 1500, file = "/home/yejichoi2/Documents/lin_temp_folder/Correlation_tissue_clin_t1dlc.png", res = 170)
corr_plot_modified(data = pt_md,
                   X = spatial_metab_tca, 
                   Y = c("gbm_thick_artmean",
                         "avg_c_k2"),
                   adj_var = NULL,
                   dict = dict)
dev.off()
# Reduced version for EDNSG abstract
png(height = 1200, width = 1500, file = "/home/yejichoi2/Documents/lin_temp_folder/Correlation_tissue_clin_t1dlc_reduced.png", res = 170)
corr_plot_modified(data = pt_md,
                   X = spatial_metab_tca[-c(7)], 
                   Y = c("gbm_thick_artmean",
                         "avg_c_k2"),
                   adj_var = NULL,
                   dict = dict)
dev.off()
```
##### Spearman
```{r}
png(height = 1200, width = 1500, file = "/home/yejichoi2/Documents/lin_temp_folder/Correlation_tissue_clin_t1dlc_spearman.png", res = 170)
corr_plot_modified(data = pt_md,
                   X = spatial_metab_tca, 
                   Y = c("gbm_thick_artmean",
                         "avg_c_k2"),
                   cor_method = "spearman",
                   adj_var = NULL,
                   dict = dict)
dev.off()
# Reduced version for EDNSG abstract
# png(height = 1200, width = 1500, file = "/home/yejichoi2/Documents/lin_temp_folder/Correlation_tissue_clin_t1dlc_reduced_spearman.png", res = 170)
# corr_plot_modified(data = pt_md,
#                    X = spatial_metab_tca[-c(7)], 
#                    Y = c("gbm_thick_artmean",
#                          "avg_c_k2"),
#                    cor_method = "spearman",
#                    adj_var = NULL,
#                    dict = dict)
# dev.off()
```

#### T1D only
##### Pearson
```{r}
png(height = 1200, width = 1500, file = "/home/yejichoi2/Documents/lin_temp_folder/Correlation_tissue_clin_t1d.png", res = 170)
corr_plot_modified(data = subset(pt_md, group == "Type 1 Diabetes"),
                   X = spatial_metab_tca, 
                   Y = c("gbm_thick_artmean",
                         "avg_c_k2"),
                   adj_var = NULL,
                   dict = dict)
dev.off()
# Reduced version for EDNSG abstract
png(height = 1200, width = 1500, file = "/home/yejichoi2/Documents/lin_temp_folder/Correlation_tissue_clin_t1dlc_reduced.png", res = 170)
corr_plot_modified(data = subset(pt_md, group == "Type 1 Diabetes"),
                   X = spatial_metab_tca[-c(7)], 
                   Y = c("gbm_thick_artmean",
                         "avg_c_k2"),
                   adj_var = NULL,
                   dict = dict)
dev.off()
```
##### Spearman
```{r}
tiff(res = 600, height = 4000, width = 3000, file = "/home/yejichoi2/Documents/lin_temp_folder/Correlation_tissue_clin_t1d_spearman.tiff")
corr_plot_modified(data = subset(pt_md, group == "Type 1 Diabetes"),
                   X = spatial_metab_tca, 
                   Y = c("gbm_thick_artmean",
                         "avg_c_k2"),
                   adj_var = NULL,
                   cor_method = "spearman",
                   dict = dict)
dev.off()
# Reduced version for EDNSG abstract
# png(height = 1200, width = 1500, file = "/home/yejichoi2/Documents/lin_temp_folder/Correlation_tissue_clin_t1dlc_reduced_spearman.png", res = 170)
# corr_plot_modified(data = subset(pt_md, group == "Type 1 Diabetes"),
#                    X = spatial_metab_tca[-c(7)], 
#                    Y = c("gbm_thick_artmean",
#                          "avg_c_k2"),
#                    adj_var = NULL,
#                    cor_method = "spearman",
#                    dict = dict)
# dev.off()
```

## Blood metabolites & clinical variables
### Plots
```{r message=F, warning=F}
## Log GBM width, and group the two groups
metabo_sc(data = pt_md, gene = "gbm_thick_artmean", 
          transcripts = blood_metab_tca,
          gene_name = "GBM width")

#### Associate global k2 with blood metabolites
metabo_sc(data = pt_md, gene = "avg_c_k2", 
          transcripts = blood_metab_tca,
          gene_name = "Cortex k2")

#### Associate voxelwise k2 with blood metabolites
metabo_sc(data = pt_md, gene = "avg_c_k2_wo_cyst_vw", 
          transcripts = blood_metab_tca,
          gene_name = "Voxel-wise Cortex k2")

#### Associate insulin sensitivity with blood metabolites
metabo_sc(data = pt_md, gene = "p2_raw_m", 
          transcripts = blood_metab_tca,
          gene_name = "M-value")
```

## Leptin and adiponectin and clinical data
### Table
#### Pooled
```{r}
dict["p2_raw_m"] <- "M-value"
correlation_table_minimal(data = pt_md,
                   relevant_vars = c("p2_raw_m",
                                     leptin_adipo_genes),
                   n_cols = 1,
                   dict = dict)

```
#### T1D only
```{r}
correlation_table_minimal(data = subset(pt_md, group == "Type 1 Diabetes"),
                   relevant_vars = c("p2_raw_m",
                                     leptin_adipo_genes),
                   n_cols = 1,
                   dict = dict)

```

### Heatmap
#### Pooled
```{r}
# corr_plot_modified(data = pt_md,
#                    Y = leptin_adipo_genes, 
#                    X = c("bmi",
#                          "p2_raw_m"),
#                    adj_var = NULL,
#                    dict = dict)
```
#### T1D only
```{r}
corr_plot_modified(data = subset(pt_md, group == "Type 1 Diabetes"),
                   Y = leptin_adipo_genes, 
                   X = c("p2_raw_m"),
                   adj_var = NULL,
                   dict = dict)
```