---
title: "PEDSnet CAH growth curves"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 4
    toc_float: yes
---

```{r setup, include=FALSE}
library(readr)
library(ggplot2)
library(quantregGrowth)
library(tidyverse)
library(skimr)
library(dplyr)
library(stringr)
library(growthcleanr)
library(data.table)

knitr::opts_chunk$set(echo = FALSE,warning = FALSE)
```

```{r echo=FALSE}
ds <- read.csv("/Volumes/Shared/Shared Projects/Laura/Peds Endo/PEDSNet/CAH/Data_raw/CAH_BMI_Demo_23.09.18.csv")
ds$dob <- as.Date(ds$dob, format="%m/%d/%Y")
ds$measurement_date <- as.Date(ds$measurement_date, format = "%Y-%m-%d")
ds$age <- floor((ds$measurement_date - ds$dob)/365.25)
ds$age_days <- floor(ds$measurement_date - ds$dob)
ds$sex_num <- ifelse(ds$gender=="Male", 0, 1)
# some records have age <0 and age >18
ds <- ds %>% subset(age>=2)
ds <- ds %>% subset(age<=18)

# create a doubly long dataset with height and weight only, and keep the other variables separate
ds_weight <- ds[,c("person_id","age_days","sex_num","wgt")]
ds_weight$measurement <- ds_weight$wgt
ds_weight$wgt <- NULL
ds_weight$param <- "WEIGHTKG"
ds_height <- ds[,c("person_id","age_days","sex_num","hgt")]
ds_height$measurement <- ds_height$hgt
ds_height$hgt <- NULL
ds_height$param <- "HEIGHTCM"
ds_keep <- rbind(ds_weight,ds_height)
ds_keep <- as.data.table(ds_keep)

cleaned_data <- ds_keep[, gcr_result := cleangrowth(person_id, param, age_days, sex_num, measurement)]
cleaned_wt <- cleaned_data %>% filter(param=="WEIGHTKG" & gcr_result=="Include")
cleaned_wt$weight_kg <- cleaned_wt$measurement
cleaned_wt$measurement <- NULL
cleaned_wt$param <- NULL
cleaned_wt$gcr_result <- NULL
cleaned_ht <- cleaned_data %>% filter(param=="HEIGHTCM" & gcr_result=="Include")
cleaned_ht$height_cm <- cleaned_ht$measurement
cleaned_ht$measurement <- NULL
cleaned_ht$param <- NULL
cleaned_ht$gcr_result <- NULL
cleaned_data_wide <- merge(cleaned_wt, cleaned_ht, by=c("person_id", "age_days"), all.x = T, all.y = T)
cleaned_data_wide$bmi <- cleaned_data_wide$weight_kg / (cleaned_data_wide$height_cm / 100)
cleaned_data_wide$sex_num <- ifelse(!is.na(cleaned_data_wide$sex_num.x),cleaned_data_wide$sex_num.x,cleaned_data_wide$sex_num.y)
cleaned_data_wide$sex_num.x <- NULL
cleaned_data_wide$sex_num.y <- NULL
cleaned_data_wide$age <- floor(cleaned_data_wide$age_days/365.25)
cleaned_data_wide_males <- cleaned_data_wide[cleaned_data_wide$sex_num==0,]
cleaned_data_wide_females <- cleaned_data_wide[cleaned_data_wide$sex_num==1,]

# CAH subset
ds_subset <- read.csv("/Volumes/Shared/Shared Projects/Laura/Peds Endo/PEDSNet/CAH/Data_raw/CAH_BMI_Demo_subset_23.10.03.csv")
ds_subset$dob <- as.Date(ds_subset$dob, format="%m/%d/%Y")
ds_subset$measurement_date <- as.Date(ds_subset$measurement_date, format = "%Y-%m-%d")
ds_subset$age <- floor((ds_subset$measurement_date - ds_subset$dob)/365.25)
ds_subset$age_days <- floor(ds_subset$measurement_date - ds_subset$dob)
ds_subset$sex_num <- ifelse(ds_subset$gender=="Male", 0, 1)
# some records_subset have age <0 and age >18
ds_subset <- ds_subset %>% subset(age>=2)
ds_subset <- ds_subset %>% subset(age<=18)

# create a doubly long dataset with height and weight only, and keep the other variables separate
ds_subset_weight <- ds_subset[,c("person_id","age_days","sex_num","wgt")]
ds_subset_weight$measurement <- ds_subset_weight$wgt
ds_subset_weight$wgt <- NULL
ds_subset_weight$param <- "WEIGHTKG"
ds_subset_height <- ds_subset[,c("person_id","age_days","sex_num","hgt")]
ds_subset_height$measurement <- ds_subset_height$hgt
ds_subset_height$hgt <- NULL
ds_subset_height$param <- "HEIGHTCM"
ds_subset_keep <- rbind(ds_subset_weight,ds_subset_height)
ds_subset_keep <- as.data.table(ds_subset_keep)

cleaned_data_subset <- ds_subset_keep[, gcr_result := cleangrowth(person_id, param, age_days, sex_num, measurement)]
cleaned_wt_subset <- cleaned_data_subset %>% filter(param=="WEIGHTKG" & gcr_result=="Include")
cleaned_wt_subset$weight_kg <- cleaned_wt_subset$measurement
cleaned_wt_subset$measurement <- NULL
cleaned_wt_subset$param <- NULL
cleaned_wt_subset$gcr_result <- NULL
cleaned_ht_subset <- cleaned_data_subset %>% filter(param=="HEIGHTCM" & gcr_result=="Include")
cleaned_ht_subset$height_cm <- cleaned_ht_subset$measurement
cleaned_ht_subset$measurement <- NULL
cleaned_ht_subset$param <- NULL
cleaned_ht_subset$gcr_result <- NULL
cleaned_data_wide_subset <- merge(cleaned_wt_subset, cleaned_ht_subset, by=c("person_id", "age_days"), all.x = T, all.y = T)
cleaned_data_wide_subset$bmi <- cleaned_data_wide_subset$weight_kg / (cleaned_data_wide_subset$height_cm / 100)
cleaned_data_wide_subset$sex_num <- ifelse(!is.na(cleaned_data_wide_subset$sex_num.x),cleaned_data_wide_subset$sex_num.x,cleaned_data_wide_subset$sex_num.y)
cleaned_data_wide_subset$sex_num.x <- NULL
cleaned_data_wide_subset$sex_num.y <- NULL
cleaned_data_wide_subset$age <- floor(cleaned_data_wide_subset$age_days/365.25)
cleaned_data_wide_subset_males <- cleaned_data_wide[cleaned_data_wide$sex_num==0,]
cleaned_data_wide_subset_females <- cleaned_data_wide[cleaned_data_wide$sex_num==1,]
```

# All CAH data

## Males and females

### Height

```{r echo=FALSE}
# fit the model
set.seed(3654)
tau2=c(0.05, 0.1, 0.25, 0.5, 0.75, 0.9, 0.95)
tau3=c(0.1, 0.5, 0.9)
mform <-gcrq(data=cleaned_data_wide, height_cm ~ ps(age, monotone=1), tau=tau2)
mform_red <- gcrq(data=cleaned_data_wide, height_cm ~ ps(age, monotone=1), tau=tau3)
```

### Very basic plots

```{r echo=FALSE}
# very basic plots
plot(mform, res=FALSE, ylab="Height (cm)")
```

#### Plot as in package vignette

```{r echo=FALSE}
# plot per vignette
p <- plot(mform, legend=TRUE, overlap=15, grid=list(x=15,y=10), col=2, lty=1,ylab="Height (cm)", xlab="Age (yrs)")
```

#### Custom plot

```{r echo=FALSE}
plot(mform, lwd=2, legend=TRUE, 
     lty=1, cex=0.5,col=1,xaxt="n",yaxt="n",
     ylab="Height (cm)", xlab="Age (yrs))", xlim=c(0,18),ylim=c(0,200))
axis(1, at=0:18, lwd=0.5, tck=1, cex.axis=0.95, padj=-1.7)
axis(2, at=seq(0,200,10), lwd=0.5, tck=1, cex.axis=0.95, las=1, hadj=0.5)
axis(4, at=seq(0,200,5), lwd=0.5, tck=1, cex.axis=0.95, las=1, hadj=0.5)
# axis(3, at=3:19, lwd=0.5, tck=1, cex.axis=0.95, padj=3.65)
```

### Weight

```{r echo=FALSE}
# fit the model
set.seed(3654)
tau2=c(0.05, 0.1, 0.25, 0.5, 0.75, 0.9, 0.95)
tau3=c(0.1, 0.5, 0.9)
mform <-gcrq(data=cleaned_data_wide, weight_kg ~ ps(age, monotone=1), tau=tau2)
mform_red <- gcrq(data=cleaned_data_wide, weight_kg ~ ps(age, monotone=1), tau=tau3)
```

#### Very basic plots

```{r echo=FALSE}
# very basic plots
plot(mform, res=FALSE, ylab="Weight (kg)")
```

#### Plot as in package vignette

```{r echo=FALSE}
# plot per vignette
p <- plot(mform, legend=TRUE, overlap=15, grid=list(x=15,y=10), col=2, lty=1,ylab="Weight (kg)", xlab="Age (yrs)")
```

#### Custom plot

```{r echo=FALSE}
plot(mform, lwd=2, legend=TRUE, 
     lty=1, cex=0.5,col=1,xaxt="n",yaxt="n",
     ylab="Weight (kg)", xlab="Age (yrs))", xlim=c(0,18),ylim=c(0,140))
axis(1, at=0:18, lwd=0.5, tck=1, cex.axis=0.95, padj=-1.7)
axis(2, at=seq(0,140,10), lwd=0.5, tck=1, cex.axis=0.95, las=1, hadj=0.5)
axis(4, at=seq(0,140,5), lwd=0.5, tck=1, cex.axis=0.95, las=1, hadj=0.5)
# axis(3, at=3:19, lwd=0.5, tck=1, cex.axis=0.95, padj=3.65)
```

### BMI

```{r echo=FALSE}
# fit the model
set.seed(3654)
tau2=c(0.05, 0.1, 0.25, 0.5, 0.75, 0.9, 0.95)
tau3=c(0.1, 0.5, 0.9)
mform <-gcrq(data=cleaned_data_wide, bmi ~ ps(age, monotone=1), tau=tau2)
mform_red <- gcrq(data=cleaned_data_wide, bmi ~ ps(age, monotone=1), tau=tau3)
```

#### Very basic plots

```{r echo=FALSE}
# very basic plots
plot(mform, res=FALSE, ylab=expression(BMI ~ kg/m^"2"))
```

#### Plot as in package vignette

```{r echo=FALSE}
# plot per vignette
p <- plot(mform, legend=TRUE, overlap=15, grid=list(x=15,y=10), col=2, lty=1,ylab=expression(BMI ~ kg/m^"2"), xlab="Age (yrs)")
```

#### Custom plot

```{r echo=FALSE}
plot(mform, lwd=2, legend=TRUE, 
     lty=1, cex=0.5,col=1,xaxt="n",yaxt="n",
     ylab=expression(BMI ~ kg/m^"2"), xlab="Age (yrs))", xlim=c(0,18),ylim=c(0,70))
axis(1, at=0:18, lwd=0.5, tck=1, cex.axis=0.95, padj=-1.7)
axis(2, at=seq(0,70,10), lwd=0.5, tck=1, cex.axis=0.95, las=1, hadj=0.5)
axis(4, at=seq(0,70,5), lwd=0.5, tck=1, cex.axis=0.95, las=1, hadj=0.5)
# axis(3, at=3:19, lwd=0.5, tck=1, cex.axis=0.95, padj=3.65)
```

## Males

### Height

```{r echo=FALSE}
# fit the model
set.seed(3654)
tau2=c(0.05, 0.1, 0.25, 0.5, 0.75, 0.9, 0.95)
tau3=c(0.1, 0.5, 0.9)
mform <-gcrq(data=cleaned_data_wide_males, height_cm ~ ps(age, monotone=1), tau=tau2)
mform_red <- gcrq(data=cleaned_data_wide_males, height_cm ~ ps(age, monotone=1), tau=tau3)
```

### Very basic plots

```{r echo=FALSE}
# very basic plots
plot(mform, res=FALSE, ylab="Height (cm)")
```

#### Plot as in package vignette

```{r echo=FALSE}
# plot per vignette
p <- plot(mform, legend=TRUE, overlap=15, grid=list(x=15,y=10), col=2, lty=1,ylab="Height (cm)", xlab="Age (yrs)")
```

#### Custom plot

```{r echo=FALSE}
plot(mform, lwd=2, legend=TRUE, 
     lty=1, cex=0.5,col=1,xaxt="n",yaxt="n",
     ylab="Height (cm)", xlab="Age (yrs))", xlim=c(0,18),ylim=c(0,200))
axis(1, at=0:18, lwd=0.5, tck=1, cex.axis=0.95, padj=-1.7)
axis(2, at=seq(0,200,10), lwd=0.5, tck=1, cex.axis=0.95, las=1, hadj=0.5)
axis(4, at=seq(0,200,5), lwd=0.5, tck=1, cex.axis=0.95, las=1, hadj=0.5)
# axis(3, at=3:19, lwd=0.5, tck=1, cex.axis=0.95, padj=3.65)
```

### Weight

```{r echo=FALSE}
# fit the model
set.seed(3654)
tau2=c(0.05, 0.1, 0.25, 0.5, 0.75, 0.9, 0.95)
tau3=c(0.1, 0.5, 0.9)
mform <-gcrq(data=cleaned_data_wide_males, weight_kg ~ ps(age, monotone=1), tau=tau2)
mform_red <- gcrq(data=cleaned_data_wide_males, weight_kg ~ ps(age, monotone=1), tau=tau3)
```

#### Very basic plots

```{r echo=FALSE}
# very basic plots
plot(mform, res=FALSE, ylab="Weight (kg)")
```

#### Plot as in package vignette

```{r echo=FALSE}
# plot per vignette
p <- plot(mform, legend=TRUE, overlap=15, grid=list(x=15,y=10), col=2, lty=1,ylab="Weight (kg)", xlab="Age (yrs)")
```

#### Custom plot

```{r echo=FALSE}
plot(mform, lwd=2, legend=TRUE, 
     lty=1, cex=0.5,col=1,xaxt="n",yaxt="n",
     ylab="Weight (kg)", xlab="Age (yrs))", xlim=c(0,18),ylim=c(0,140))
axis(1, at=0:18, lwd=0.5, tck=1, cex.axis=0.95, padj=-1.7)
axis(2, at=seq(0,140,10), lwd=0.5, tck=1, cex.axis=0.95, las=1, hadj=0.5)
axis(4, at=seq(0,140,5), lwd=0.5, tck=1, cex.axis=0.95, las=1, hadj=0.5)
# axis(3, at=3:19, lwd=0.5, tck=1, cex.axis=0.95, padj=3.65)
```

### BMI

```{r echo=FALSE}
# fit the model
set.seed(3654)
tau2=c(0.05, 0.1, 0.25, 0.5, 0.75, 0.9, 0.95)
tau3=c(0.1, 0.5, 0.9)
mform <-gcrq(data=cleaned_data_wide_males, bmi ~ ps(age, monotone=1), tau=tau2)
mform_red <- gcrq(data=cleaned_data_wide_males, bmi ~ ps(age, monotone=1), tau=tau3)
```

#### Very basic plots

```{r echo=FALSE}
# very basic plots
plot(mform, res=FALSE, ylab=expression(BMI ~ kg/m^"2"))
```

#### Plot as in package vignette

```{r echo=FALSE}
# plot per vignette
p <- plot(mform, legend=TRUE, overlap=15, grid=list(x=15,y=10), col=2, lty=1,ylab=expression(BMI ~ kg/m^"2"), xlab="Age (yrs)")
```

#### Custom plot

```{r echo=FALSE}
plot(mform, lwd=2, legend=TRUE, 
     lty=1, cex=0.5,col=1,xaxt="n",yaxt="n",
     ylab=expression(BMI ~ kg/m^"2"), xlab="Age (yrs))", xlim=c(0,18),ylim=c(0,70))
axis(1, at=0:18, lwd=0.5, tck=1, cex.axis=0.95, padj=-1.7)
axis(2, at=seq(0,70,10), lwd=0.5, tck=1, cex.axis=0.95, las=1, hadj=0.5)
axis(4, at=seq(0,70,5), lwd=0.5, tck=1, cex.axis=0.95, las=1, hadj=0.5)
# axis(3, at=3:19, lwd=0.5, tck=1, cex.axis=0.95, padj=3.65)
```


## Females

# CAH subset

## Males and females

### Height

```{r echo=FALSE}
# fit the model
tau2=c(0.05, 0.1, 0.25, 0.5, 0.75, 0.9, 0.95)
tau3=c(0.1, 0.5, 0.9)
mform <-gcrq(data=cleaned_data_wide_subset, height_cm ~ ps(age, monotone=1), tau=tau2)
mform_red <- gcrq(data=cleaned_data_wide_subset, height_cm ~ ps(age, monotone=1), tau=tau3)
```

#### Very basic plots

```{r echo=FALSE}
# very basic plots
plot(mform, res=FALSE, ylab="Height")
```

#### Plot as in package vignette

```{r echo=FALSE}
# plot per vignette
p <- plot(mform, legend=TRUE, overlap=15, grid=list(x=15,y=10), col=2, lty=1,ylab="Height", xlab="Age (yrs)")
```

#### Custom plot

```{r echo=FALSE}
plot(mform, lwd=2, legend=TRUE, 
     lty=1, cex=0.5,col=1,xaxt="n",yaxt="n",
     ylab="Height", xlab="Age (yrs))", xlim=c(0,18),ylim=c(0,200))
axis(1, at=0:18, lwd=0.5, tck=1, cex.axis=0.95, padj=-1.7)
axis(2, at=seq(0,200,10), lwd=0.5, tck=1, cex.axis=0.95, las=1, hadj=0.5)
axis(4, at=seq(0,200,5), lwd=0.5, tck=1, cex.axis=0.95, las=1, hadj=0.5)
# axis(3, at=3:19, lwd=0.5, tck=1, cex.axis=0.95, padj=3.65)
```

## Weight

```{r echo=FALSE}
# fit the model
set.seed(3654)
tau2=c(0.05, 0.1, 0.25, 0.5, 0.75, 0.9, 0.95)
tau3=c(0.1, 0.5, 0.9)
mform <-gcrq(data=cleaned_data_wide_subset, weight_kg ~ ps(age, monotone=1), tau=tau2)
mform_red <- gcrq(data=cleaned_data_wide_subset, weight_kg ~ ps(age, monotone=1), tau=tau3)
```

### Very basic plots

```{r echo=FALSE}
# very basic plots
plot(mform, res=FALSE, ylab="Weight")
```

### Plot as in package vignette

```{r echo=FALSE}
# plot per vignette
p <- plot(mform, legend=TRUE, overlap=15, grid=list(x=15,y=10), col=2, lty=1,ylab="Weight", xlab="Age (yrs)")
```

### Custom plot

```{r echo=FALSE}
plot(mform, lwd=2, legend=TRUE, 
     lty=1, cex=0.5,col=1,xaxt="n",yaxt="n",
     ylab="Weight", xlab="Age (yrs))", xlim=c(0,18),ylim=c(0,140))
axis(1, at=0:18, lwd=0.5, tck=1, cex.axis=0.95, padj=-1.7)
axis(2, at=seq(0,140,10), lwd=0.5, tck=1, cex.axis=0.95, las=1, hadj=0.5)
axis(4, at=seq(0,140,5), lwd=0.5, tck=1, cex.axis=0.95, las=1, hadj=0.5)
# axis(3, at=3:19, lwd=0.5, tck=1, cex.axis=0.95, padj=3.65)
```

## BMI

```{r echo=FALSE}
# fit the model
set.seed(3654)
tau2=c(0.05, 0.1, 0.25, 0.5, 0.75, 0.9, 0.95)
tau3=c(0.1, 0.5, 0.9)
mform <-gcrq(data=cleaned_data_wide_subset, bmi ~ ps(age, monotone=1), tau=tau2)
mform_red <- gcrq(data=cleaned_data_wide_subset, bmi ~ ps(age, monotone=1), tau=tau3)
```

### Very basic plots

```{r echo=FALSE}
# very basic plots
plot(mform, res=FALSE, ylab="BMI")
```

### Plot as in package vignette

```{r echo=FALSE}
# plot per vignette
p <- plot(mform, legend=TRUE, overlap=15, grid=list(x=15,y=10), col=2, lty=1,ylab="BMI", xlab="Age (yrs)")
```

### Custom plot

```{r echo=FALSE}
plot(mform, lwd=2, legend=TRUE, 
     lty=1, cex=0.5,col=1,xaxt="n",yaxt="n",
     ylab="BMI", xlab="Age (yrs))", xlim=c(0,18),ylim=c(0,70))
axis(1, at=0:18, lwd=0.5, tck=1, cex.axis=0.95, padj=-1.7)
axis(2, at=seq(0,70,10), lwd=0.5, tck=1, cex.axis=0.95, las=1, hadj=0.5)
axis(4, at=seq(0,70,5), lwd=0.5, tck=1, cex.axis=0.95, las=1, hadj=0.5)
# axis(3, at=3:19, lwd=0.5, tck=1, cex.axis=0.95, padj=3.65)
```

## Males

## Females
