---
title: "Incidence of youth-onset type 2 diabetes during the COVID-19 pandemic"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 2
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = FALSE)

library(skimr)
library(stringr)
library(dplyr)
library(tableone)
library(knitr)
library(lubridate)
library(ggplot2)
library(tscount)
library(childsds)

#################
# READ IN DATA  #
#################

columbia <- read.csv("E:\\Megan Kelsey\\T2D and COVID\\Data raw\\T2D COVID multisite template 2018_2020_Columbia.csv")
names <- colnames(columbia)
names[1] <- "ptid"
colnames(columbia) <- names
columbia$site <- "Columbia"
columbia$X <- NULL
# exclude 2 participants with BMI <85th percentile
columbia <- columbia %>% filter(!ptid %in% c(1006140219,1009535804))
# exclude 7 year old because can't rule out MODY
columbia <- columbia %>% filter(!age_dx==7.4)
# exclude those with missing ab info
columbia <- columbia %>% filter(!(GAD65.Ab=="Not Tested" & Insulin.Ab=="Not Tested" & ICA.Ab=="Not Tested" & ZnT8.Ab=="Not Tested"))

chicago <- read.csv("E:\\Megan Kelsey\\T2D and COVID\\Data raw\\2021.8.30 T2D COVID UChicago FINAL deidentified2.csv")
chicago$site <- "Chicago"
# exclude 2 ppts with low BMI
chicago <- chicago %>% filter(!BMI %in% c(15.1,25.5))

mayo <- read.csv("E:\\Megan Kelsey\\T2D and COVID\\Data raw\\8-27-21MayoUpload.csv")
mayo$site <- "Mayo"

chla <- read.csv("E:\\Megan Kelsey\\T2D and COVID\\Data raw\\bmicheck_chla.csv")
chla$site <- "CHLA"
chla$ptid <- NA

jhu <- read.csv("E:\\Megan Kelsey\\T2D and COVID\\Data raw\\DATA ENTRY Hopkins t2d CLEAN no PHI.csv")
jhu$site <- "JHU"
jhu$X <- NULL
# remove 2 ppts per Risa
jhu <- jhu %>% filter(!BMI %in% c(20.16,20.06))
# missing ZnT8 data
jhu$ZnT8.Ab <- ifelse(jhu$ZnT8.Ab=="","Not Tested",jhu$ZnT8.Ab)

duke <- read.csv("E:\\Megan Kelsey\\T2D and COVID\\Data raw\\DUKE_T2D_DATA.csv")
duke$site <- "Duke"
duke$ptid <- NA

ou <- read.csv("E:\\Megan Kelsey\\T2D and COVID\\Data raw\\T2D and COVID Study Database 10Jul2021 DE-IDENTIFIED OU.csv")
ou$site <- "OU"
ou$ptid <- NA

mercy <- read.csv("E:\\Megan Kelsey\\T2D and COVID\\Data raw\\T2D COVID - Children's Mercy Final Deidentified.csv")
mercy$site <- "Mercy"
mercy$ptid <- NA

cchmc <- read.csv("E:\\Megan Kelsey\\T2D and COVID\\Data raw\\T2D COVID multisite template CCHMC.csv")
cchmc$site <- "CCHMC"
# exclude 2 ppts per Amy Shah
cchmc <- cchmc %>% filter(!(BMI %in% c(20.52,24.8)))

iu <- read.csv("E:\\Megan Kelsey\\T2D and COVID\\Data raw\\T2D COVID multisite IU 8.27.2021.csv")
iu$site <- "IU"
# correct BMI
iu$BMI <- ifelse(iu$BMI>300,31,iu$BMI)
# remove 3 ppts with low BMI
iu <- iu %>% filter(!(ptid %in% c(52,71,94)))

ucsd <- read.csv("E:\\Megan Kelsey\\T2D and COVID\\Data raw\\T2D COVID UCSD Rady Childrens 09012021 Deidentified.csv")
ucsd$site <- "UCSD"
ucsd$ptid <- NA
# remove 3 ppts with low BMI
ucsd <- ucsd %>% filter(!BMI %in% c(22.64,24.94,22.07))

cu <- read.csv("E:\\Megan Kelsey\\T2D and COVID\\Data raw\\T2D COVID_CUAnschutz.csv")
cu$site <- "Colorado"

brown <- read.csv("E:\\Megan Kelsey\\T2D and COVID\\Data raw\\T2D COVIDdata_Brown.Hasbro.csv")
brown$site <- "Brown/Hasbro"
# need to get rid of decimal at the end of all ages
brown$age_dx <- substr(brown$age_dx,1,nchar(brown$age_dx)-1)
brown$age_dx <- ifelse(brown$age_dx=="14. 11","14.11",brown$age_dx)

cnh <- read.csv("E:\\Megan Kelsey\\T2D and COVID\\Data raw\\Children's National_ UPLOAD - Estrada - 10-29-21.csv")
cnh$site <- "Children's National"
# remove 2 ppts with low BMI
#cnh <- cnh %>% filter(!ptid %in% c(160,238))
# somehow I missed their ab variables were incorrect
cnh$GAD65.Ab <- ifelse(cnh$GAD65.Ab=="Y","Negative","Not Tested")
cnh$Insulin.Ab <- ifelse(cnh$Insulin.Ab=="Y","Negative","Not Tested")
cnh$ICA.Ab <- ifelse(cnh$ICA.Ab=="Y","Negative","Not Tested")
cnh$ZnT8.Ab <- ifelse(cnh$ZnT8.Ab=="Y","Negative","Not Tested")

lurie <- read.csv("E:\\Megan Kelsey\\T2D and COVID\\Data raw\\9_28_21_LurieUpload.csv")
lurie$site <- "Lurie"
# recode ab data per Sean Delacey
lurie$GAD65.Ab <- ifelse(lurie$GAD65.Ab=="yes","Negative","Not Tested")
lurie$Insulin.Ab <- ifelse(lurie$Insulin.Ab=="yes","Negative","Not Tested")
lurie$ICA.Ab <- ifelse(lurie$ICA.Ab=="yes","Negative","Not Tested")
lurie$ZnT8.Ab <- ifelse(lurie$ZnT8.Ab=="yes","Negative","Not Tested")

pennington <- read.csv("E:\\Megan Kelsey\\T2D and COVID\\Data raw\\T2D COVID multisite template 10-15-2021 submitted to CO Pennington.csv")
pennington$site <- "Pennington"
# correct one BMI
pennington$BMI <- ifelse(pennington$BMI==23.8,27,pennington$BMI)
# delete one ppt
pennington <- pennington %>% filter(!BMI==25.3)

pennstate <- read.csv("E:\\Megan Kelsey\\T2D and COVID\\Data raw\\T2D COVID PennStateUniversity data 10-2021.csv")
pennstate$site <- "Penn State"
# 6 patients data not fixed - need to exclude for now
pennstate <- pennstate %>% filter(!(ptid %in% c("PSU006","PSU025","PSU026","PSU027","PSU101","PSU156")))

tch <- read.csv("E:\\Megan Kelsey\\T2D and COVID\\Data raw\\TCH Data-Uploaded 2nd version.csv",na.strings = c(""," ","N/A"))
tch$site <- "TCH"
# exclude 3 patients with BMI percentile <85
tch <- tch %>% filter(!ptid %in% c("TCH-32","TCH-35","TCH-69"))
tch$X <- NULL
tch$ZnT8.Ab <- ifelse(tch$ZnT8.Ab=="Negaative","NEGATIVE",tch$ZnT8.Ab)

ucsf <- read.csv("E:\\Megan Kelsey\\T2D and COVID\\Data raw\\T2D_COVID_UCSF_V2_10252021.csv")
ucsf$site <- "UCSF"
ucsf$ptid <- NA

um <- read.csv("E:\\Megan Kelsey\\T2D and COVID\\Data raw\\UM_T2D COVID multisite template 10.14.2021.csv")
um$site <- "Michigan"
# remove 1 ppt with low BMI
um <- um %>% filter(!BMI==23.14)

utsw <- read.csv("E:\\Megan Kelsey\\T2D and COVID\\Data raw\\UTSW COVID T2.csv")
utsw$site <- "UTSW"
# recode ab data per Olga Gupta
utsw$GAD65.Ab <- ifelse(utsw$GAD65.Ab=="Y","Negative","Not Tested")
utsw$Insulin.Ab <- ifelse(utsw$Insulin.Ab=="Y","Negative","Not Tested")
utsw$ICA.Ab <- ifelse(utsw$ICA.Ab=="Y","Negative","Not Tested")
utsw$ZnT8.Ab <- ifelse(utsw$ZnT8.Ab=="Y","Negative","Not Tested")

### TO BE FIXED

# VU is missing month and year is not coded correctly so will exclude
#vanderbilt <- read.csv("E:\\Megan Kelsey\\T2D and COVID\\Data raw\\.csv")
#vanderbilt$site <- "Vanderbilt"

# read in Peoria
peoria <- read.csv("E:\\Megan Kelsey\\T2D and COVID\\Data raw\\Peoria.Illinois.T2D COVID 19_2021-10-15_Validated_11.1.2021.1-De-identified.csv",
                   na.strings = c("NULL",""," "))
peoria$site <- "Peoria"
# remove 1 ppt with low BMI
peoria <- peoria %>% filter(!ptid==11)

# read in CHOP

# COMBINE DATA 
alldata <- rbind(columbia,chicago,mayo,chla,jhu,duke,ou,mercy,cchmc,iu,ucsd,cu,brown,cnh,lurie,pennington,pennstate,tch,ucsf,um,utsw,peoria)

#################
# DATA CLEANING #
#################

# clean up cases
alldata$GAD65.Ab <- toupper(alldata$GAD65.Ab)
alldata$Insulin.Ab <- toupper(alldata$Insulin.Ab)
alldata$ICA.Ab <- toupper(alldata$ICA.Ab)
alldata$ZnT8.Ab <- toupper(alldata$ZnT8.Ab)
alldata$sex <- toupper(alldata$sex)
alldata$insurance <- toupper(alldata$insurance)
alldata$Location.Diagnosis <- toupper(alldata$Location.Diagnosis)
alldata$COVID.19.at.diagnosis <- toupper(alldata$COVID.19.at.diagnosis)
alldata$Race <- toupper(alldata$Race)
alldata$Ethnicity <- toupper(alldata$Ethnicity)

# missing ab data
#View(alldata[alldata$GAD65.Ab=="" | alldata$Insulin.Ab=="" | alldata$ICA.Ab=="" | alldata$ZnT8.Ab=="",])

# clean up race/ethnicity
alldata$Race <- str_trim(alldata$Race)
alldata$Ethnicity <- str_trim(alldata$Ethnicity)
alldata$Race <- ifelse(alldata$Race %in% c("","Did not answer","DID NOT ANSWER","DECLINED","UNABLE TO OBTAIN","Unknown","HISPANIC"),"UNKNOWN",alldata$Race)
alldata$Race <- ifelse(alldata$Race %in% c("BLACK OR AFRICAN AMERICAN","BLACK/AFRICAN AMERICAN"),"BLACK",alldata$Race)
alldata$Race <- ifelse(alldata$Race %in% c("NATIVE AMERICAN/ESKIMO"),"OTHER",alldata$Race)
alldata$Race <- ifelse(alldata$Race %in% c("WHITE OR CAUCASIAN"),"WHITE",alldata$Race)
alldata$Ethnicity <- ifelse(alldata$Ethnicity %in% c("HISPANIC OR LATINO","HISPANIC/LATINO"),"HISPANIC",alldata$Ethnicity)
alldata$Ethnicity <- ifelse(alldata$Ethnicity %in% c("","DID NOT ANSWER","DECLINED","UNABLE TO OBTAIN"),"UNKNOWN",alldata$Ethnicity)
alldata$Ethnicity <- ifelse(alldata$Ethnicity %in% c("NON-HISPANIC","NON HISPANIC","NOT HISPANIC","NOT HISPANIC/LATINO","NOT HISPANIC/NOT LATINO","NOT HISPANIC OR LATINO"),
                            "NON-HISPANIC",alldata$Ethnicity)

# clean up insurance
alldata$insurance <- ifelse(alldata$insurance=="NONE","UNINSURED",alldata$insurance)
alldata$insurance <- ifelse(alldata$insurance %in% c("PUBIC","PUBILC","PUBLIC ","CHIP","CMC INDIGENT CARE","MEDICAID","MEDICAID MANAGED CARE",
                                                     "OON MEDICAID"),"PUBLIC",alldata$insurance)
alldata$insurance <- ifelse(alldata$insurance %in% c("","UNKNOWN"),"UNKNOWN",alldata$insurance)
alldata$insurance <- ifelse(alldata$insurance %in% c("NONE/UNINSURED","SELF-PAY"),"UNINSURED",alldata$insurance)
alldata$insurance <- ifelse(alldata$insurance %in% c("BLUE CROSS/BLUE SHIELD","COMMERCIAL","BC/BS"),"PRIVATE",alldata$insurance)
alldata$insurance <- ifelse(alldata$insurance=="TRICARE","OTHER",alldata$insurance)

# clean up BMI
alldata$BMI <- ifelse(alldata$BMI %in% c("?","Not Available","Not Tested"),NA,alldata$BMI)
alldata$BMI <- as.numeric(alldata$BMI)
alldata$BMI <- ifelse(alldata$BMI>300,NA,alldata$BMI)

# clean up A1c
alldata$HbA1c <- str_trim(alldata$HbA1c)
alldata$HbA1c <- ifelse(alldata$HbA1c %in% c("> 14",">14.0",">14"),14,alldata$HbA1c)
alldata$HbA1c <- ifelse(alldata$HbA1c==">14.5",14.5,alldata$HbA1c)
alldata$HbA1c <- ifelse(alldata$HbA1c %in% c("> 15",">15"),15,alldata$HbA1c)
alldata$HbA1c <- ifelse(alldata$HbA1c=="> 15.5",15.5,alldata$HbA1c)
alldata$HbA1c <- as.numeric(alldata$HbA1c)

# clean up glucose 
alldata$Serum.glucose..random. <- str_remove(alldata$Serum.glucose..random.," mg/dL")
alldata$Serum.glucose..random. <- str_remove(alldata$Serum.glucose..random.," mg/dl")
alldata$Serum.glucose..random. <- str_trim(alldata$Serum.glucose..random.)
alldata$Serum.glucose..random. <- ifelse(alldata$Serum.glucose..random.==">400",400,alldata$Serum.glucose..random.)
alldata$Serum.glucose..random. <- ifelse(alldata$Serum.glucose..random.==">500",500,alldata$Serum.glucose..random.)
alldata$Serum.glucose..random. <- ifelse(alldata$Serum.glucose..random.==">600",600,alldata$Serum.glucose..random.)
alldata$Serum.glucose..random. <- ifelse(alldata$Serum.glucose..random. %in% c(">700","> 700"),700,alldata$Serum.glucose..random.)
alldata$Serum.glucose..random. <- ifelse(alldata$Serum.glucose..random.==">750",750,alldata$Serum.glucose..random.)
alldata$Serum.glucose..random. <- ifelse(alldata$Serum.glucose..random.=="Not Tested",NA,alldata$Serum.glucose..random.)
alldata$Serum.glucose..random. <- ifelse(alldata$Serum.glucose..random.=="1,001",1001,alldata$Serum.glucose..random.)
alldata$Serum.glucose..random. <- ifelse(alldata$Serum.glucose..random.=="1,001",1001,alldata$Serum.glucose..random.)
alldata$Serum.glucose..random. <- ifelse(alldata$Serum.glucose..random.=="1,163",1163,alldata$Serum.glucose..random.)
alldata$Serum.glucose..random. <- ifelse(alldata$Serum.glucose..random.=="1,473",1473,alldata$Serum.glucose..random.)
alldata$Serum.glucose..random. <- ifelse(alldata$Serum.glucose..random.=="1,191",1191,alldata$Serum.glucose..random.)
alldata$Serum.glucose..random. <- ifelse(alldata$Serum.glucose..random. %in% c("","NULL"),NA,alldata$Serum.glucose..random.)
alldata$Serum.glucose..random. <- as.numeric(alldata$Serum.glucose..random.)

# clean up pH
alldata$pH <- ifelse(alldata$pH %in% c("","-"),NA,alldata$pH)
alldata$pH <- ifelse(alldata$pH=="Not Tested",NA,alldata$pH)
alldata$pH <- ifelse(alldata$pH=="7.3Unknown",7.3,alldata$pH)
alldata$pH <- ifelse(alldata$pH=="<7.01",7.01,alldata$pH)
alldata$pH <- as.numeric(alldata$pH)

# clean up bicarb
alldata$bicarbonate <- ifelse(alldata$bicarbonate %in% c("<10","< 10"),10,alldata$bicarbonate)
alldata$bicarbonate <- ifelse(alldata$bicarbonate=="<5",5,alldata$bicarbonate)
alldata$bicarbonate <- ifelse(alldata$bicarbonate=="Not Tested",NA,alldata$bicarbonate)
alldata$bicarbonate <- as.numeric(alldata$bicarbonate)

# clean up osmolality
alldata$Serum.osmolality <- ifelse(alldata$Serum.osmolality=="Not Tested",NA,alldata$Serum.osmolality)
alldata$Serum.osmolality <- as.numeric(alldata$Serum.osmolality)

# clean up age
alldata$age_dx <- str_trim(alldata$age_dx)
alldata$age_dx <- ifelse(alldata$age_dx=="12..75",12.75,alldata$age_dx)
alldata$age_dx <- as.numeric(alldata$age_dx)

# clean up sex
alldata$sex <- ifelse(alldata$sex=="M","MALE",alldata$sex)
alldata$sex <- ifelse(alldata$sex=="F","FEMALE",alldata$sex)
alldata$sex <- ifelse(alldata$sex %in% c(""," "),"UNKNOWN",alldata$sex)

# clean up COVID dx
alldata$COVID.19.at.diagnosis <- ifelse(alldata$COVID.19.at.diagnosis %in% c("N","NEGATIVE","0"),"NO",alldata$COVID.19.at.diagnosis)
alldata$COVID.19.at.diagnosis <- ifelse(alldata$COVID.19.at.diagnosis %in% c("ND","NOT TESTED"),"UNKNOWN",alldata$COVID.19.at.diagnosis)
alldata$COVID.19.at.diagnosis <- ifelse(alldata$COVID.19.at.diagnosis %in% c("P","POSITIVE","1"),"YES",alldata$COVID.19.at.diagnosis)
alldata$COVID.19.at.diagnosis <- ifelse(alldata$COVID.19.at.diagnosis %in% c(""," "),"UNKNOWN",alldata$COVID.19.at.diagnosis)

# clean up location
alldata$Location.Diagnosis <- ifelse(alldata$Location.Diagnosis %in% c(""," "),"UNKNOWN",alldata$Location.Diagnosis)
alldata$Location.Diagnosis <- ifelse(alldata$Location.Diagnosis %in% c("INPATIENT-FLOOR","INPATIENT-ICU","IP"),"INPATIENT",alldata$Location.Diagnosis)
alldata$Location.Diagnosis <- ifelse(alldata$Location.Diagnosis %in% c("OP","ED AND DISCHARGED"),"OUTPATIENT",alldata$Location.Diagnosis)

# convert to factors
alldata$mo_dx <- as.factor(as.character(alldata$mo_dx))
alldata$yr_dx <- as.factor(as.character(alldata$yr_dx))

# REMOVE AB POSITIVE PATIENTS!!!!!
alldata <- alldata %>% filter(!(alldata$site=="Columbia" & alldata$GAD65.Ab==""))
alldata <- alldata %>% filter(!(alldata$site=="CHLA" & alldata$GAD65.Ab=="POS"))

# calculate BMI percentile
alldata$bmip <- sds(alldata$BMI,
              age = alldata$age_dx,
              sex = alldata$sex, male = "MALE", female =  "FEMALE",
              ref = cdc.ref,
              item = "bmi",
              type = "perc")

# write file of BMI percentile <85th
bmicheck <- alldata %>% filter(bmip<0.85)
bmicheck$bmip <- bmicheck$bmip*100

bmicheck_cchmc <- bmicheck %>% filter(site=="CCHMC")
write.csv(bmicheck_cchmc,"E:\\Megan Kelsey\\T2D and COVID\\Data clean\\bmicheck_cchmc.csv",row.names = F)

bmicheck_chicago <- bmicheck %>% filter(site=="Chicago")
write.csv(bmicheck_chicago,"E:\\Megan Kelsey\\T2D and COVID\\Data clean\\bmicheck_chicago.csv",row.names = F)

bmicheck_national <- bmicheck %>% filter(site=="Children's National")
write.csv(bmicheck_national,"E:\\Megan Kelsey\\T2D and COVID\\Data clean\\bmicheck_national.csv",row.names = F)

bmicheck_chla <- bmicheck %>% filter(site=="CHLA")
write.csv(bmicheck_chla,"E:\\Megan Kelsey\\T2D and COVID\\Data clean\\bmicheck_chla.csv",row.names = F)

bmicheck_colorado <- bmicheck %>% filter(site=="Colorado")
write.csv(bmicheck_colorado,"E:\\Megan Kelsey\\T2D and COVID\\Data clean\\bmicheck_colorado.csv",row.names = F)

bmicheck_columbia <- bmicheck %>% filter(site=="Columbia")
write.csv(bmicheck_columbia,"E:\\Megan Kelsey\\T2D and COVID\\Data clean\\bmicheck_columbia.csv",row.names = F)

bmicheck_duke <- bmicheck %>% filter(site=="Duke")
write.csv(bmicheck_duke,"E:\\Megan Kelsey\\T2D and COVID\\Data clean\\bmicheck_duke.csv",row.names = F)

bmicheck_iu <- bmicheck %>% filter(site=="IU")
write.csv(bmicheck_iu,"E:\\Megan Kelsey\\T2D and COVID\\Data clean\\bmicheck_IU.csv",row.names = F)

bmicheck_JHU <- bmicheck %>% filter(site=="JHU")
write.csv(bmicheck_JHU,"E:\\Megan Kelsey\\T2D and COVID\\Data clean\\bmicheck_JHU.csv",row.names = F)

bmicheck_mayo <- bmicheck %>% filter(site=="Mayo")
write.csv(bmicheck_mayo,"E:\\Megan Kelsey\\T2D and COVID\\Data clean\\bmicheck_mayo.csv",row.names = F)

bmicheck_michigan <- bmicheck %>% filter(site=="Michigan")
write.csv(bmicheck_michigan,"E:\\Megan Kelsey\\T2D and COVID\\Data clean\\bmicheck_michigan.csv",row.names = F)

bmicheck_pennstate <- bmicheck %>% filter(site=="Penn State")
write.csv(bmicheck_pennstate,"E:\\Megan Kelsey\\T2D and COVID\\Data clean\\bmicheck_pennstate.csv",row.names = F)

bmicheck_pennington <- bmicheck %>% filter(site=="Pennington")
write.csv(bmicheck_pennington,"E:\\Megan Kelsey\\T2D and COVID\\Data clean\\bmicheck_pennington.csv",row.names = F)

bmicheck_tch <- bmicheck %>% filter(site=="TCH")
write.csv(bmicheck_tch,"E:\\Megan Kelsey\\T2D and COVID\\Data clean\\bmicheck_tch.csv",row.names = F)

bmicheck_ucsd <- bmicheck %>% filter(site=="UCSD")
write.csv(bmicheck_ucsd,"E:\\Megan Kelsey\\T2D and COVID\\Data clean\\bmicheck_ucsd.csv",row.names = F)

bmicheck_ucsf <- bmicheck %>% filter(site=="UCSF")
write.csv(bmicheck_ucsf,"E:\\Megan Kelsey\\T2D and COVID\\Data clean\\bmicheck_ucsf.csv",row.names = F)

bmicheck_utsw <- bmicheck %>% filter(site=="UTSW")
write.csv(bmicheck_utsw,"E:\\Megan Kelsey\\T2D and COVID\\Data clean\\bmicheck_utsw.csv",row.names = F)

bmicheck_peoria <- bmicheck %>% filter(site=="Peoria")

##################
# AIM 2 OUTCOMES #
##################

# DKA is pH<7.3 or bicarb<=15
alldata$dka <- (!is.na(alldata$pH) & alldata$pH<7.3 & is.na(alldata$bicarbonate)) | 
                (!is.na(alldata$bicarbonate) & alldata$bicarbonate<=15 & is.na(alldata$pH)) |
                  (!is.na(alldata$bicarbonate) & !is.na(alldata$pH) & (alldata$pH<7.3 | alldata$bicarbonate<=15))
#alldata$dka <- ifelse(is.na(alldata$pH) & is.na(alldata$bicarbonate),NA,alldata$dka)
# if DKA is missing, set to none
#alldata$dka <- ifelse(is.na(alldata$dka),FALSE,TRUE)

# HHS is osmolality >=330 AND glucose >600
alldata$hhs <- ifelse(alldata$Serum.glucose..random.>600 & alldata$Serum.osmolality>=330,TRUE,FALSE)
alldata$hhs <- ifelse(is.na(alldata$hhs),FALSE,alldata$hhs)

# metabolic decomp is DKA or HHS or both
alldata$met_decomp <- alldata$dka | alldata$hhs

# DKA severity
# first code severity of pH
alldata$pHsev <- ifelse(alldata$pH<7.1,"Severe",
                        ifelse(alldata$pH<7.2,"Moderate",
                               ifelse(alldata$pH<7.3,"Mild",NA)))
alldata$bicarbsev <- ifelse(alldata$bicarbonate<5,"Severe",
                        ifelse(alldata$bicarbonate<10,"Moderate",
                               ifelse(alldata$bicarbonate<15,"Mild",NA)))
alldata$dka_severity <- ifelse(alldata$dka==TRUE & (alldata$pHsev=="Severe" | alldata$bicarbsev=="Severe"),"Severe",
                               ifelse(alldata$dka==TRUE & (alldata$pHsev=="Moderate" | alldata$bicarbsev=="Moderate"),"Moderate",
                                      ifelse(alldata$dka==TRUE & alldata$pHsev=="Mild" & alldata$bicarbsev=="Mild","Mild",NA)))

#######################
# DEFINE TIME PERIODS #
#######################

alldata$day_dx <- 1
alldata$dx_date <- as.Date(with(alldata, paste(yr_dx, mo_dx, day_dx,sep="-")), "%Y-%m-%d")

# filter out patients not in the correct time period
alldata <- alldata %>% filter(!(alldata$dx_date<as.Date("2018-03-01")))
alldata <- alldata %>% filter(!(alldata$dx_date>as.Date("2021-02-28")))

alldata$period <- ifelse(alldata$dx_date>as.Date("2018-03-01") & alldata$dx_date<=as.Date("2019-02-28"),"3/1/18-2/28/19",
                         ifelse(alldata$dx_date>=as.Date("2019-03-01") & alldata$dx_date<=as.Date("2020-02-29"),"3/1/19-2/29/20","3/1/20-2/28/21"))
alldata$period <- as.factor(alldata$period)

alldata$pandemic <- as.factor(ifelse(alldata$period=="3/1/20-2/28/21",1,0))

#######################
# CREATE TIME SERIES  #
#######################

ts1 <- alldata %>% group_by(dx_date) %>% tally()
ts2 <- ts1$n
ts3 <- ts(ts2,start = c(2018,3), end=c(2021,2),frequency=12)

```

# Background

The purpose of this analysis is to examine the frequency of diagnosis of youth-onset type 2 diabetes during the COVID pandemic (defined as March 1, 2020-February 28, 2021) compared to prior to the pandemic (March 1, 2018-February 28, 2020).

# Outstanding data issues/questions

- CHOP data has not been added to this report

- We talked about including Vanderbilt data in an analysis of diagnoses by year instead of by month within year, since they were not able to provide the month of diagnosis.  We should discuss what type of analysis to perform, since we have only 2 years before and 1 year after.

# Methods

Lab values reported as either above or below the limit of detection were set to the limit of detection.  DKA was defined as either pH<7.3 or bicarbonate<=15.  HHS was defined as serum glucose >600 and osmolality >=330.  Metabolic decompensation was defined as either DKA or HHS.  DKA severity was defined as follows: 1) severe - pH <7.1 or bicarbonate <5, 2) moderate - pH <7.2 or bicarbonate <10, or 3) mild - pH <7.3 or bicarbonate <=15.

Groups were compared using either a t-test or ANOVA for normally distributed continuous variables, the Kruskal Wallis test for non-normally distributed variables, and the chi-square test for categorical variables.  

A negative binomial time series model was used to model the count of youth-onset T2D diagnoses per month.  A background linear trend of 5% increase per year was assumed. Seasonality was assessed by allowing the number of diagnoses in a particular month to depend on the number in the prior month, and the number 12 months prior; however, the amount of seasonality was minimal.  Because there appeared to be a peak in the number of diagnoses in August during the first 2 years, less so in the 3rd year, a covariate corresponding to August vs. other months was tested but was not significant.  The final model did not include any effect of seasonality.  A test of intervention was performed to determine whether the onset of the pandemic in March 1, 2020, had a significant effect on the count of diagnoses.

# Results

```{r include=FALSE}
vars <- c("site","mo_dx","yr_dx","period","num_ab_tested","GAD65.Ab","ICA.Ab","Insulin.Ab","ZnT8.Ab",
          "age_dx","Race","Ethnicity","sex","insurance","Location.Diagnosis","COVID.19.at.diagnosis","BMI","bmip","HbA1c",
          "Serum.glucose..random.","pH","bicarbonate","Serum.osmolality","dka","hhs","met_decomp","dka_severity")

t1 <- CreateTableOne(data=alldata,vars = vars)
t1 <- print(t1,nonnormal = c("BMI","bmip","HbA1c","Serum.glucose..random.","pH","bicarbonate","Serum.osmolality"),showAllLevels=T)
```

## Overall descriptive statistics

```{r echo=FALSE}
kable(t1,caption = "Overall descriptive statistics")
```
<br>

```{r include=FALSE}

t2 <- CreateTableOne(data=alldata,vars = vars,strata = "pandemic")
t2 <- print(t2,nonnormal = c("BMI","bmip","HbA1c","Serum.glucose..random.","pH","bicarbonate","Serum.osmolality"),showAllLevels=T)
```

## Descriptive statistics before and during pandemic

```{r echo=FALSE}
kable(t2)
```
<br>

## Descriptive statistics by site

```{r include=FALSE}
vars <- c("mo_dx","yr_dx","period","num_ab_tested","GAD65.Ab","ICA.Ab","Insulin.Ab","ZnT8.Ab",
          "age_dx","Race","Ethnicity","sex","insurance","Location.Diagnosis","COVID.19.at.diagnosis","BMI","bmip","HbA1c",
          "Serum.glucose..random.","pH","bicarbonate","Serum.osmolality","dka","hhs","met_decomp","dka_severity")

t3 <- CreateTableOne(data=alldata,vars = vars,strata="site",test = FALSE)
t3 <- print(t3,nonnormal = c("BMI","bmip","HbA1c","Serum.glucose..random.","pH","bicarbonate","Serum.osmolality"),showAllLevels=T)
```

```{r echo=FALSE}
kable(t3)
```
<br>

## Site by pandemic

```{r include=FALSE}
t3a <- CreateTableOne(data=alldata,vars = "site",strata="pandemic")
t3a <- print(t3a,showAllLevels=T)
```

```{r echo=FALSE}
kable(t3a)
```
<br>

## Descriptive statistics by year

```{r include=FALSE}
vars <- c("site","mo_dx","num_ab_tested","GAD65.Ab","ICA.Ab","Insulin.Ab","ZnT8.Ab",
          "age_dx","Race","Ethnicity","sex","insurance","Location.Diagnosis","COVID.19.at.diagnosis","BMI","bmip","HbA1c",
          "Serum.glucose..random.","pH","bicarbonate","Serum.osmolality","dka","hhs","met_decomp","dka_severity")

t4 <- CreateTableOne(data=alldata,vars = vars,strata="period")
t4 <- print(t4,nonnormal = c("BMI","bmip","HbA1c","Serum.glucose..random.","pH","bicarbonate","Serum.osmolality"),showAllLevels=T)
```

```{r echo=FALSE}
kable(t4)
```
<br>

## Sex by race

```{r include=FALSE}
t5 <- CreateTableOne(data=alldata,vars = "sex",strata="Race")
t5 <- print(t5,showAllLevels=T)
```

```{r echo=FALSE}
kable(t5)
```
<br>

## Sex by ethnicity

```{r include=FALSE}
t6 <- CreateTableOne(data=alldata,vars = "sex",strata="Ethnicity")
t6 <- print(t6,showAllLevels=T)
```

```{r echo=FALSE}
kable(t6)
```
<br>

## Race by pandemic

```{r include=FALSE}
t7 <- CreateTableOne(data=alldata,vars = "Race",strata="pandemic")
t7 <- print(t7,showAllLevels=T)
```

```{r echo=FALSE}
kable(t7)
```
<br>

## Ethnicity by pandemic

```{r include=FALSE}
t8 <- CreateTableOne(data=alldata,vars = "Ethnicity",strata="pandemic")
t8 <- print(t8,showAllLevels=T)
```

```{r echo=FALSE}
kable(t8)
```
<br>

## Figure

```{r echo=FALSE}
ggplot(alldata, aes(x=dx_date, fill=met_decomp)) + geom_bar()
```
<br>

## Model diagnostics

```{r echo=FALSE}
fit_poisson <- tsglm(ts3,distr = "poisson")
fit_nb <- tsglm(ts3,distr = "nbinom")

acf(residuals(fit_poisson))
acf(residuals(fit_nb))

pit(fit_poisson, main="PIT Poisson")
pit(fit_nb, main="PIT NB")

marcal(fit_poisson, main = "Marginal calibration")
  lines(marcal(fit_nb, plot = FALSE), lty = "dashed")
  legend("bottomright", legend = c("Pois", "NegBin"), lwd = 1, lty = c("solid", "dashed"))
# NB model is better
```
<br>

## Model results

```{r echo=FALSE,comment=""}
set.seed(3654)

# model with dependence on prior month and 12 months past to model seasonality
preds <- cbind(ts3,linearTrend=(seq(along=ts3)*(0.15/36)))
preds <- preds[,-1]
fit_nb1 <- tsglm(ts3,distr = "nbinom",model = list(past_obs=c(1,12)),xreg=preds)
a <- summary(fit_nb1,B=100)
# seasonality as represented by beta_12 is small

# now include a term for month to capture seasonality
# I don't think there is a good method for a factor predictor
#preds <- cbind(ts3,months=as.factor(rep(c("Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec","Jan","Feb"),3)),
#               linearTrend=(seq(along=ts3)*(0.15/36)))
#preds <- preds[,-1]
#fit_nb1 <- tsglm(ts3,distr = "nbinom",model = list(past_obs=c(1,12)),xreg=preds)
#a <- summary(fit_nb1,B=100)

# can we simplify to August vs. no August
preds <- cbind(ts3,aug=as.factor(rep(c(0,0,0,0,0,1,0,0,0,0,0,0),3)),
               linearTrend=(seq(along=ts3)*(0.15/36)))
preds <- preds[,-1]
fit_nb1 <- tsglm(ts3,distr = "nbinom",model = list(past_obs=c(1,12)),xreg=preds)
a <- summary(fit_nb1,B=100)
# effect of August does not appear to be important, although could be washed out by the last year
# could try a model with an interaction between pre/post pandemic and August
# although if we are really interested in August, need more pre pandemic data

# remove seasonality
# make sure linear trend is right
# beta is dependence on prior value, eta is dependence on covariates (I think), and sigmasq is overdispersion coefficient
preds <- cbind(ts3,linearTrend=((seq(along=ts3)-1)*(0.15/36))+rep(1,36))
preds <- preds[,-1]
fit_nb1 <- tsglm(ts3,distr = "nbinom",model = list(past_obs=1),xreg=preds)
colnames(fit_nb1$xreg) <- "linearTrend"
a <- summary(fit_nb1,B=100)
# test of intervention
interv_test(fit_nb1,tau=24,delta=1,est_interv=T)

```




