---
title: "Incidence of youth-onset type 2 diabetes during the COVID-19 pandemic"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 2
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = FALSE)

library(skimr)
library(stringr)
library(dplyr)
library(tableone)
library(knitr)
library(lubridate)
library(ggplot2)
library(tscount)
library(childsds)

#################
# READ IN DATA  #
#################

columbia <- read.csv("E:\\Megan Kelsey\\T2D and COVID\\Data raw\\T2D COVID multisite template 2018_2020_Columbia.csv")
names <- colnames(columbia)
names[1] <- "ptid"
colnames(columbia) <- names
columbia$site <- "Columbia"
columbia$X <- NULL

chicago <- read.csv("E:\\Megan Kelsey\\T2D and COVID\\Data raw\\2021.8.30 T2D COVID UChicago FINAL deidentified2.csv")
chicago$site <- "Chicago"

mayo <- read.csv("E:\\Megan Kelsey\\T2D and COVID\\Data raw\\8-27-21MayoUpload.csv")
mayo$site <- "Mayo"

chla <- read.csv("E:\\Megan Kelsey\\T2D and COVID\\Data raw\\CHLA_T2D Data.csv")
chla$site <- "CHLA"

jhu <- read.csv("E:\\Megan Kelsey\\T2D and COVID\\Data raw\\DATA ENTRY Hopkins t2d CLEAN no PHI.csv")
jhu$site <- "JHU"
jhu$X <- NULL

duke <- read.csv("E:\\Megan Kelsey\\T2D and COVID\\Data raw\\DUKE_T2D_DATA.csv")
duke$site <- "Duke"
duke$ptid <- NA

ou <- read.csv("E:\\Megan Kelsey\\T2D and COVID\\Data raw\\T2D and COVID Study Database 10Jul2021 DE-IDENTIFIED OU.csv")
ou$site <- "OU"
ou$ptid <- NA

mercy <- read.csv("E:\\Megan Kelsey\\T2D and COVID\\Data raw\\T2D COVID - Children's Mercy Final Deidentified.csv")
mercy$site <- "Mercy"
mercy$ptid <- NA

cchmc <- read.csv("E:\\Megan Kelsey\\T2D and COVID\\Data raw\\T2D COVID multisite template CCHMC.csv")
cchmc$site <- "CCHMC"

iu <- read.csv("E:\\Megan Kelsey\\T2D and COVID\\Data raw\\T2D COVID multisite IU 8.27.2021.csv")
iu$site <- "IU"

ucsd <- read.csv("E:\\Megan Kelsey\\T2D and COVID\\Data raw\\T2D COVID UCSD Rady Childrens 09012021 Deidentified.csv")
ucsd$site <- "UCSD"
ucsd$ptid <- NA

cu <- read.csv("E:\\Megan Kelsey\\T2D and COVID\\Data raw\\T2D COVID_CUAnschutz.csv")
cu$site <- "Colorado"

brown <- read.csv("E:\\Megan Kelsey\\T2D and COVID\\Data raw\\T2D COVIDdata_Brown.Hasbro.csv")
brown$site <- "Brown/Hasbro"
# need to get rid of decimal at the end of all ages
brown$age_dx <- substr(brown$age_dx,1,nchar(brown$age_dx)-1)
brown$age_dx <- ifelse(brown$age_dx=="14. 11","14.11",brown$age_dx)

cnh <- read.csv("E:\\Megan Kelsey\\T2D and COVID\\Data raw\\CNH T2D COVID Data - Estrada - sent 10-1-21 .csv")
cnh$site <- "Children's National"

lurie <- read.csv("E:\\Megan Kelsey\\T2D and COVID\\Data raw\\9_28_21_LurieUpload.csv")
lurie$site <- "Lurie"

pennington <- read.csv("E:\\Megan Kelsey\\T2D and COVID\\Data raw\\T2D COVID multisite template 10-15-2021 submitted to CO Pennington.csv")
pennington$site <- "Pennington"

pennstate <- read.csv("E:\\Megan Kelsey\\T2D and COVID\\Data raw\\T2D COVID PennStateUniversity data 10-2021.csv")
pennstate$site <- "Penn State"

tch <- read.csv("E:\\Megan Kelsey\\T2D and COVID\\Data raw\\TCH Data-Uploaded.csv",na.strings = c(""," ","N/A"))
tch$site <- "TCH"

ucsf <- read.csv("E:\\Megan Kelsey\\T2D and COVID\\Data raw\\T2D_COVID_UCSF_10052021.csv")
ucsf$site <- "UCSF"

um <- read.csv("E:\\Megan Kelsey\\T2D and COVID\\Data raw\\UM_T2D COVID multisite template 10.14.2021.csv")
um$site <- "Michigan"

utsw <- read.csv("E:\\Megan Kelsey\\T2D and COVID\\Data raw\\UTSW COVID T2.csv")
utsw$site <- "UTSW"

# VU is missing month and year is not coded correctly so will exclude
#vanderbilt <- read.csv("E:\\Megan Kelsey\\T2D and COVID\\Data raw\\.csv")
#vanderbilt$site <- "Vanderbilt"

# COMBINE DATA 
alldata <- rbind(columbia,chicago,mayo,chla,jhu,duke,ou,mercy,cchmc,iu,ucsd,cu,brown,cnh,lurie,pennington,pennstate,tch,ucsf,um,utsw)

#################
# DATA CLEANING #
#################

# clean up cases
alldata$GAD65.Ab <- toupper(alldata$GAD65.Ab)
alldata$Insulin.Ab <- toupper(alldata$Insulin.Ab)
alldata$ICA.Ab <- toupper(alldata$ICA.Ab)
alldata$ZnT8.Ab <- toupper(alldata$ZnT8.Ab)
alldata$sex <- toupper(alldata$sex)
alldata$insurance <- toupper(alldata$insurance)
alldata$Location.Diagnosis <- toupper(alldata$Location.Diagnosis)
alldata$COVID.19.at.diagnosis <- toupper(alldata$COVID.19.at.diagnosis)
alldata$Race <- toupper(alldata$Race)
alldata$Ethnicity <- toupper(alldata$Ethnicity)

# missing ab data
#View(alldata[alldata$GAD65.Ab=="" | alldata$Insulin.Ab=="" | alldata$ICA.Ab=="" | alldata$ZnT8.Ab=="",])

# clean up race/ethnicity
alldata$Race <- str_trim(alldata$Race)
alldata$Ethnicity <- str_trim(alldata$Ethnicity)
alldata$Race <- ifelse(alldata$Race %in% c("","Did not answer","DID NOT ANSWER","DECLINED","UNABLE TO OBTAIN","Unknown","HISPANIC"),"UNKNOWN",alldata$Race)
alldata$Race <- ifelse(alldata$Race %in% c("BLACK OR AFRICAN AMERICAN","BLACK/AFRICAN AMERICAN"),"BLACK",alldata$Race)
alldata$Race <- ifelse(alldata$Race %in% c("NATIVE AMERICAN/ESKIMO"),"OTHER",alldata$Race)
alldata$Race <- ifelse(alldata$Race %in% c("WHITE OR CAUCASIAN"),"WHITE",alldata$Race)
alldata$Ethnicity <- ifelse(alldata$Ethnicity %in% c("HISPANIC OR LATINO","HISPANIC/LATINO"),"HISPANIC",alldata$Ethnicity)
alldata$Ethnicity <- ifelse(alldata$Ethnicity %in% c("","DID NOT ANSWER","DECLINED","UNABLE TO OBTAIN"),"UNKNOWN",alldata$Ethnicity)
alldata$Ethnicity <- ifelse(alldata$Ethnicity %in% c("NON-HISPANIC","NON HISPANIC","NOT HISPANIC","NOT HISPANIC/LATINO","NOT HISPANIC/NOT LATINO"),
                            "NON-HISPANIC",alldata$Ethnicity)

# clean up insurance
alldata$insurance <- ifelse(alldata$insurance=="NONE","UNINSURED",alldata$insurance)
alldata$insurance <- ifelse(alldata$insurance %in% c("PUBIC","PUBILC","PUBLIC ","CHIP","CMC INDIGENT CARE","MEDICAID","MEDICAID MANAGED CARE",
                                                     "OON MEDICAID"),"PUBLIC",alldata$insurance)
alldata$insurance <- ifelse(alldata$insurance %in% c("","UNKNOWN"),"UNKNOWN",alldata$insurance)
alldata$insurance <- ifelse(alldata$insurance %in% c("NONE/UNINSURED","SELF-PAY"),"UNINSURED",alldata$insurance)
alldata$insurance <- ifelse(alldata$insurance %in% c("BLUE CROSS/BLUE SHIELD","COMMERCIAL"),"PRIVATE",alldata$insurance)
alldata$insurance <- ifelse(alldata$insurance=="TRICARE","OTHER",alldata$insurance)

# clean up BMI
alldata$BMI <- ifelse(alldata$BMI %in% c("?","Not Available","Not Tested"),NA,alldata$BMI)
alldata$BMI <- as.numeric(alldata$BMI)
alldata$BMI <- ifelse(alldata$BMI>300,NA,alldata$BMI)

# clean up A1c
alldata$HbA1c <- str_trim(alldata$HbA1c)
alldata$HbA1c <- ifelse(alldata$HbA1c %in% c("> 14",">14.0",">14"),14,alldata$HbA1c)
alldata$HbA1c <- ifelse(alldata$HbA1c==">14.5",14.5,alldata$HbA1c)
alldata$HbA1c <- ifelse(alldata$HbA1c %in% c("> 15",">15"),15,alldata$HbA1c)
alldata$HbA1c <- ifelse(alldata$HbA1c=="> 15.5",15.5,alldata$HbA1c)
alldata$HbA1c <- as.numeric(alldata$HbA1c)

# clean up glucose 
alldata$Serum.glucose..random. <- str_remove(alldata$Serum.glucose..random.," mg/dL")
alldata$Serum.glucose..random. <- str_remove(alldata$Serum.glucose..random.," mg/dl")
alldata$Serum.glucose..random. <- str_trim(alldata$Serum.glucose..random.)
alldata$Serum.glucose..random. <- ifelse(alldata$Serum.glucose..random.==">400",400,alldata$Serum.glucose..random.)
alldata$Serum.glucose..random. <- ifelse(alldata$Serum.glucose..random.==">500",500,alldata$Serum.glucose..random.)
alldata$Serum.glucose..random. <- ifelse(alldata$Serum.glucose..random.==">600",600,alldata$Serum.glucose..random.)
alldata$Serum.glucose..random. <- ifelse(alldata$Serum.glucose..random. %in% c(">700","> 700"),700,alldata$Serum.glucose..random.)
alldata$Serum.glucose..random. <- ifelse(alldata$Serum.glucose..random.==">750",750,alldata$Serum.glucose..random.)
alldata$Serum.glucose..random. <- ifelse(alldata$Serum.glucose..random.=="Not Tested",NA,alldata$Serum.glucose..random.)
alldata$Serum.glucose..random. <- ifelse(alldata$Serum.glucose..random.=="1,001",1001,alldata$Serum.glucose..random.)
alldata$Serum.glucose..random. <- ifelse(alldata$Serum.glucose..random.=="1,001",1001,alldata$Serum.glucose..random.)
alldata$Serum.glucose..random. <- ifelse(alldata$Serum.glucose..random.=="1,163",1163,alldata$Serum.glucose..random.)
alldata$Serum.glucose..random. <- ifelse(alldata$Serum.glucose..random.=="1,473",1473,alldata$Serum.glucose..random.)
alldata$Serum.glucose..random. <- ifelse(alldata$Serum.glucose..random. %in% c("","NULL"),NA,alldata$Serum.glucose..random.)
alldata$Serum.glucose..random. <- as.numeric(alldata$Serum.glucose..random.)

# clean up pH
alldata$pH <- ifelse(alldata$pH %in% c("","-"),NA,alldata$pH)
alldata$pH <- ifelse(alldata$pH=="Not Tested",NA,alldata$pH)
alldata$pH <- ifelse(alldata$pH=="7.3Unknown",7.3,alldata$pH)
alldata$pH <- ifelse(alldata$pH=="<7.01",7.01,alldata$pH)
alldata$pH <- as.numeric(alldata$pH)

# clean up bicarb
alldata$bicarbonate <- ifelse(alldata$bicarbonate %in% c("<10","< 10"),10,alldata$bicarbonate)
alldata$bicarbonate <- ifelse(alldata$bicarbonate=="<5",5,alldata$bicarbonate)
alldata$bicarbonate <- ifelse(alldata$bicarbonate=="Not Tested",NA,alldata$bicarbonate)
alldata$bicarbonate <- as.numeric(alldata$bicarbonate)

# clean up osmolality
alldata$Serum.osmolality <- ifelse(alldata$Serum.osmolality=="Not Tested",NA,alldata$Serum.osmolality)
alldata$Serum.osmolality <- as.numeric(alldata$Serum.osmolality)

# clean up age
alldata$age_dx <- str_trim(alldata$age_dx)
alldata$age_dx <- ifelse(alldata$age_dx=="12..75",12.75,alldata$age_dx)
alldata$age_dx <- as.numeric(alldata$age_dx)

# clean up sex
alldata$sex <- ifelse(alldata$sex=="M","MALE",alldata$sex)
alldata$sex <- ifelse(alldata$sex=="F","FEMALE",alldata$sex)
alldata$sex <- ifelse(alldata$sex %in% c(""," "),"UNKNOWN",alldata$sex)

# clean up COVID dx
alldata$COVID.19.at.diagnosis <- ifelse(alldata$COVID.19.at.diagnosis %in% c("N","NEGATIVE"),"NO",alldata$COVID.19.at.diagnosis)
alldata$COVID.19.at.diagnosis <- ifelse(alldata$COVID.19.at.diagnosis %in% c("ND","NOT TESTED"),"UNKNOWN",alldata$COVID.19.at.diagnosis)
alldata$COVID.19.at.diagnosis <- ifelse(alldata$COVID.19.at.diagnosis %in% c("P","POSITIVE"),"YES",alldata$COVID.19.at.diagnosis)
alldata$COVID.19.at.diagnosis <- ifelse(alldata$COVID.19.at.diagnosis %in% c(""," "),"UNKNOWN",alldata$COVID.19.at.diagnosis)

# clean up location
alldata$Location.Diagnosis <- ifelse(alldata$Location.Diagnosis %in% c(""," "),"UNKNOWN",alldata$Location.Diagnosis)
alldata$Location.Diagnosis <- ifelse(alldata$Location.Diagnosis %in% c("INPATIENT-FLOOR","INPATIENT-ICU","IP"),"INPATIENT",alldata$Location.Diagnosis)
alldata$Location.Diagnosis <- ifelse(alldata$Location.Diagnosis %in% c("OP"),"OUTPATIENT",alldata$Location.Diagnosis)

# convert to factors
alldata$mo_dx <- as.factor(as.character(alldata$mo_dx))
alldata$yr_dx <- as.factor(as.character(alldata$yr_dx))

# REMOVE AB POSITIVE PATIENTS!!!!!
alldata <- alldata %>% filter(!(alldata$site=="Columbia" & alldata$GAD65.Ab==""))
alldata <- alldata %>% filter(!(alldata$site=="CHLA" & alldata$GAD65.Ab=="POS"))

# calculate BMI percentile
alldata$bmip <- sds(alldata$BMI,
              age = alldata$age_dx,
              sex = alldata$sex, male = "MALE", female =  "FEMALE",
              ref = cdc.ref,
              item = "bmi",
              type = "perc")

# write file of BMI percentile <85th
bmicheck <- alldata %>% filter(bmip<0.85)
bmicheck$bmip <- bmicheck$bmip*100

bmicheck_cchmc <- bmicheck %>% filter(site=="CCHMC")
write.csv(bmicheck_cchmc,"E:\\Megan Kelsey\\T2D and COVID\\Data clean\\bmicheck_cchmc.csv",row.names = F)

bmicheck_chicago <- bmicheck %>% filter(site=="Chicago")
write.csv(bmicheck_chicago,"E:\\Megan Kelsey\\T2D and COVID\\Data clean\\bmicheck_chicago.csv",row.names = F)

bmicheck_national <- bmicheck %>% filter(site=="Children's National")
write.csv(bmicheck_national,"E:\\Megan Kelsey\\T2D and COVID\\Data clean\\bmicheck_national.csv",row.names = F)

bmicheck_chla <- bmicheck %>% filter(site=="CHLA")
write.csv(bmicheck_chla,"E:\\Megan Kelsey\\T2D and COVID\\Data clean\\bmicheck_chla.csv",row.names = F)

bmicheck_colorado <- bmicheck %>% filter(site=="Colorado")
write.csv(bmicheck_colorado,"E:\\Megan Kelsey\\T2D and COVID\\Data clean\\bmicheck_colorado.csv",row.names = F)

bmicheck_columbia <- bmicheck %>% filter(site=="Columbia")
write.csv(bmicheck_columbia,"E:\\Megan Kelsey\\T2D and COVID\\Data clean\\bmicheck_columbia.csv",row.names = F)

bmicheck_duke <- bmicheck %>% filter(site=="Duke")
write.csv(bmicheck_duke,"E:\\Megan Kelsey\\T2D and COVID\\Data clean\\bmicheck_duke.csv",row.names = F)

bmicheck_iu <- bmicheck %>% filter(site=="IU")
write.csv(bmicheck_iu,"E:\\Megan Kelsey\\T2D and COVID\\Data clean\\bmicheck_IU.csv",row.names = F)

bmicheck_JHU <- bmicheck %>% filter(site=="JHU")
write.csv(bmicheck_JHU,"E:\\Megan Kelsey\\T2D and COVID\\Data clean\\bmicheck_JHU.csv",row.names = F)

bmicheck_mayo <- bmicheck %>% filter(site=="Mayo")
write.csv(bmicheck_mayo,"E:\\Megan Kelsey\\T2D and COVID\\Data clean\\bmicheck_mayo.csv",row.names = F)

bmicheck_michigan <- bmicheck %>% filter(site=="Michigan")
write.csv(bmicheck_michigan,"E:\\Megan Kelsey\\T2D and COVID\\Data clean\\bmicheck_michigan.csv",row.names = F)

bmicheck_pennstate <- bmicheck %>% filter(site=="Penn State")
write.csv(bmicheck_pennstate,"E:\\Megan Kelsey\\T2D and COVID\\Data clean\\bmicheck_pennstate.csv",row.names = F)

bmicheck_pennington <- bmicheck %>% filter(site=="Pennington")
write.csv(bmicheck_pennington,"E:\\Megan Kelsey\\T2D and COVID\\Data clean\\bmicheck_pennington.csv",row.names = F)

bmicheck_tch <- bmicheck %>% filter(site=="TCH")
write.csv(bmicheck_tch,"E:\\Megan Kelsey\\T2D and COVID\\Data clean\\bmicheck_tch.csv",row.names = F)

bmicheck_ucsd <- bmicheck %>% filter(site=="UCSD")
write.csv(bmicheck_ucsd,"E:\\Megan Kelsey\\T2D and COVID\\Data clean\\bmicheck_ucsd.csv",row.names = F)

bmicheck_ucsf <- bmicheck %>% filter(site=="UCSF")
write.csv(bmicheck_ucsf,"E:\\Megan Kelsey\\T2D and COVID\\Data clean\\bmicheck_ucsf.csv",row.names = F)

bmicheck_utsw <- bmicheck %>% filter(site=="UTSW")
write.csv(bmicheck_utsw,"E:\\Megan Kelsey\\T2D and COVID\\Data clean\\bmicheck_utsw.csv",row.names = F)




##################
# AIM 2 OUTCOMES #
##################

# DKA is pH<7.3 or bicarb<=15
alldata$dka <- (!is.na(alldata$pH) & alldata$pH<7.3 & is.na(alldata$bicarbonate)) | 
                (!is.na(alldata$bicarbonate) & alldata$bicarbonate<=15 & is.na(alldata$pH)) |
                  (!is.na(alldata$bicarbonate) & !is.na(alldata$pH) & alldata$pH<7.3 & alldata$bicarbonate<=15)
alldata$dka <- ifelse(is.na(alldata$pH) & is.na(alldata$bicarbonate),NA,alldata$dka)

# HHS is osmolality >=330 AND glucose >600
alldata$hhs <- ifelse(alldata$Serum.glucose..random.>600 & alldata$Serum.osmolality>=330,TRUE,FALSE)

# metabolic decomp is DKA or HHS or both
alldata$met_decomp <- alldata$dka | alldata$hhs

#######################
# DEFINE TIME PERIODS #
#######################

alldata$day_dx <- 1
alldata$dx_date <- as.Date(with(alldata, paste(yr_dx, mo_dx, day_dx,sep="-")), "%Y-%m-%d")

# filter out patients not in the correct time period
alldata <- alldata %>% filter(!(alldata$dx_date<as.Date("2018-03-01")))
alldata <- alldata %>% filter(!(alldata$dx_date>as.Date("2021-02-28")))

alldata$period <- ifelse(alldata$dx_date>as.Date("2018-03-01") & alldata$dx_date<as.Date("2019-02-28"),1,
                         ifelse(alldata$dx_date>as.Date("2019-03-01") & alldata$dx_date<as.Date("2020-02-29"),2,3))
alldata$period <- as.factor(alldata$period)

alldata$pandemic <- as.factor(ifelse(alldata$period==3,1,0))

#######################
# CREATE TIME SERIES  #
#######################

ts1 <- alldata %>% group_by(dx_date) %>% tally()
ts2 <- ts1$n
ts3 <- ts(ts2,start = c(2018,3), end=c(2021,2),frequency=12)

```

# Outstanding data cleaning issues

- deleted ab positive patients 
- missing values for sex
- 1 record at IU with BMI>300
- Vanderbilt did not include month of diagnosis and year was coded as 0, -1, -2 so their data were excluded
- Several sites recorded "yes" for the individual antibodies, probably because the lead-in question was "Was Ab tested" even though the options in the drop-down were "Negative" and "Not Tested."  Do we clarify?
- 3 patients with diagnosis dates 11/19, 11/19, and 4/19 COVID positive
- TCH recorded location of diagnosis as "TCH".  How to code "ED and discharged"?
- if race=Hispanic recoded to missing (ethnicity was Hispanic in all these cases)
- 57 patients with BMI percentile <85 - exclude?

# Background

# Methods

lab values above or below limit of detection set to the limit value

# Results

```{r include=FALSE}
vars <- c("site","mo_dx","yr_dx","num_ab_tested","age_dx","Race","Ethnicity","sex","insurance","Location.Diagnosis","COVID.19.at.diagnosis","BMI","bmip","HbA1c",
          "Serum.glucose..random.","pH","bicarbonate","Serum.osmolality","dka","hhs","met_decomp")

t1 <- CreateTableOne(data=alldata,vars = vars)
t1 <- print(t1,nonnormal = c("BMI","bmip","HbA1c","Serum.glucose..random.","pH","bicarbonate","Serum.osmolality"))
```

## Overall descriptive statistics

```{r echo=FALSE}
kable(t1,caption = "Overall descriptive statistics")
```
<br>

```{r include=FALSE}
vars <- c("num_ab_tested","age_dx","Race","Ethnicity","sex","insurance","Location.Diagnosis","COVID.19.at.diagnosis","BMI","bmip","HbA1c",
          "Serum.glucose..random.","pH","bicarbonate","Serum.osmolality","dka","hhs","met_decomp")

t2 <- CreateTableOne(data=alldata,vars = vars,strata = "pandemic")
t2 <- print(t2,nonnormal = c("BMI","bmip","HbA1c","Serum.glucose..random.","pH","bicarbonate","Serum.osmolality"))
```

## Descriptive statistics before and during pandemic

```{r echo=FALSE}
kable(t2)
```
<br>

## Descriptive statistics by site

```{r include=FALSE}
vars <- c("num_ab_tested","age_dx","Race","Ethnicity","sex","insurance","Location.Diagnosis","COVID.19.at.diagnosis","BMI","bmip","HbA1c",
          "Serum.glucose..random.","pH","bicarbonate","Serum.osmolality","dka","hhs","met_decomp")

t3 <- CreateTableOne(data=alldata,vars = vars,strata="site",test = FALSE)
t3 <- print(t3,nonnormal = c("BMI","bmip","HbA1c","Serum.glucose..random.","pH","bicarbonate","Serum.osmolality"))
```

```{r echo=FALSE}
kable(t3)
```
<br>

## Descriptive statistics by year

```{r include=FALSE}
vars <- c("num_ab_tested","age_dx","Race","Ethnicity","sex","insurance","Location.Diagnosis","COVID.19.at.diagnosis","BMI","bmip","HbA1c",
          "Serum.glucose..random.","pH","bicarbonate","Serum.osmolality","dka","hhs","met_decomp")

t4 <- CreateTableOne(data=alldata,vars = vars,strata="yr_dx")
t4 <- print(t4,nonnormal = c("BMI","bmip","HbA1c","Serum.glucose..random.","pH","bicarbonate","Serum.osmolality"))
```

```{r echo=FALSE}
kable(t4)
```
<br>

## Figure

```{r echo=FALSE}
ggplot(alldata, aes(x=dx_date, fill=met_decomp)) + geom_bar()
```
<br>

## Model diagnostics

```{r echo=FALSE}
fit_poisson <- tsglm(ts3,distr = "poisson")
fit_nb <- tsglm(ts3,distr = "nbinom")

acf(residuals(fit_poisson))
acf(residuals(fit_nb))

pit(fit_poisson, main="PIT Poisson")
pit(fit_nb, main="PIT NB")

marcal(fit_poisson, main = "Marginal calibration")
  lines(marcal(fit_nb, plot = FALSE), lty = "dashed")
  legend("bottomright", legend = c("Pois", "NegBin"), lwd = 1, lty = c("solid", "dashed"))
# NB model is better
```
<br>

```{r echo=FALSE,comment=""}
set.seed(3654)

# ALSO NEED TO FIGURE OUT HOW TO ACCOUNT FOR BACKGROUND RATE OF CHANGE (SEE NOTES)

#fit_nb1 <- tsglm(ts3,distr = "nbinom",model = list(past_obs=1))
# AIC 363.7575
#fit_nb2 <- tsglm(ts3,distr = "nbinom",model = list(past_obs=c(1,2)))
# AIC 364.1613 

#fit_nb1_log <- tsglm(ts3,distr = "nbinom",model = list(past_obs=1),link = "log")
# AIC 370.3331 
#fit_nb2_log <- tsglm(ts3,distr = "nbinom",model = list(past_obs=c(1,2)), link="log")
# AIC 370.0166 

#x_pandemic <- c(rep(0,24),rep(1,12))
#fit_nb1p <- tsglm(ts3,distr = "nbinom",model = list(past_obs=1),xreg = x_pandemic)
# AIC 284.3584
#fit_nb2p <- tsglm(ts3,distr = "nbinom",model = list(past_obs=c(1,2)),xreg = x_pandemic)
# AIC 286.4812

#fit_nb1p_log <- tsglm(ts3,distr = "nbinom",model = list(past_obs=1),xreg = x_pandemic,link="log")
# AIC 282.9692
#fit_nb2p_log <- tsglm(ts3,distr = "nbinom",model = list(past_obs=c(1,2)),xreg = x_pandemic,link="log")
# AIC 285.102 

# not sure if it makes sense to test for an intervention effect when the intervention is already a term in the model?
# otherwise, how do I get p-values?
#interv_test(fit_nb1,tau=24,delta=1,est_interv=TRUE)

# probably need to fix time, this implies it's doubling
#time <- 1:length(ts3)

#regressors <- cbind(rep(1,36),linearTrend=(seq(along=ts3)*(0.15/36)))
#regressors <- regressors[,-1]

#fit_nb1 <- tsglm(ts3,distr = "nbinom",model = list(past_obs=1),xreg=regressors)
#summary(fit_nb1,B=100)
#interv_test(fit_nb1,tau=24,delta=1,est_interv=TRUE)

```




