---
title: "Growth attenuation therapy analysis"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include=FALSE}
library(knitr)
library(dplyr)
library(hrbrthemes)
library(nlme)
library(emmeans)
library(tableone)

knitr::opts_chunk$set(echo = FALSE,warning = FALSE)

knitr::opts_chunk$set(echo = FALSE)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "E:/Davis/Growth attenuation analysis/Data raw"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/som/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects"
}
setwd(home_dir)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data, include=FALSE}
source("./GrowthAttenuationThe-DemographicsMedicalH_R_2022-06-14_1722.r")
demo <- data
data <- NULL
# exclude participant 100
demo <- demo %>% filter(!record_id==100)

# set new AE vars to unknown if missing
# hairgat
demo$hairgat <- ifelse(is.na(demo$hairgat),999,demo$hairgat)
levels(demo$hairgat.factor) <- c("Yes","No","Unknown")
demo$hairgat.factor <- ifelse(is.na(demo$hairgat.factor),"Unknown",demo$hairgat.factor)
demo$hairgat.factor <- ifelse(demo$hairgat.factor=="1","Yes",
                              ifelse(demo$hairgat.factor=="2","No",demo$hairgat.factor))
# droolgat
demo$droolgat <- ifelse(is.na(demo$droolgat),999,demo$droolgat)
levels(demo$droolgat.factor) <- c("Yes","No","Unknown")
demo$droolgat.factor <- ifelse(is.na(demo$droolgat.factor),"Unknown",demo$droolgat.factor)
demo$droolgat.factor <- ifelse(demo$droolgat.factor=="1","Yes",
                              ifelse(demo$droolgat.factor=="2","No",demo$droolgat.factor))
# genitalgat
demo$genitalgat <- ifelse(is.na(demo$genitalgat),999,demo$genitalgat)
levels(demo$genitalgat.factor) <- c("Yes","No","Unknown")
demo$genitalgat.factor <- ifelse(is.na(demo$genitalgat.factor),"Unknown",demo$genitalgat.factor)
demo$genitalgat.factor <- ifelse(demo$genitalgat.factor=="1","Yes",
                              ifelse(demo$genitalgat.factor=="2","No",demo$genitalgat.factor))
# flushgat
demo$flushgat <- ifelse(is.na(demo$flushgat),999,demo$flushgat)
levels(demo$flushgat.factor) <- c("Yes","No","Unknown")
demo$flushgat.factor <- ifelse(is.na(demo$flushgat.factor),"Unknown",demo$flushgat.factor)
demo$flushgat.factor <- ifelse(demo$flushgat.factor=="1","Yes",
                              ifelse(demo$flushgat.factor=="2","No",demo$flushgat.factor))

# read in longitudinal height data
height <- read.csv("./GrowthAttenuationThe-Height_DATA_2022-06-14_0943.csv")
# exclude participant 100
height <- height %>% filter(!record_id==100)
height <- height[!is.na(height$redcap_repeat_instance),]
# get the most recent height from physical exam
height <- height %>% group_by(record_id) %>% summarise_all(last)
temp <- demo[,c("record_id","final_height","initial_height","treatment_duration_yr","gv_ave")]
height <- merge(height,temp,by="record_id",all.x = T,all.y = T)
# check growth velocity variable
height$gv_calc <- ifelse(!is.na(height$final_height), (height$final_heigh-height$initial_height)/height$treatment_duration_yr,
                         ifelse(!is.na(height$height),
                         (height$height-height$initial_height)/height$treatment_duration_yr,"")) 
# still have NAs in redcap repeat instance - but this comes from demo which is fine
# if gv negative, set to 0
height$gv_calc <- ifelse(height$gv_calc<0,0,height$gv_calc)
# if there is only one encounter, set gv_calc to 0
height$gv_calc <-ifelse(height$redcap_repeat_instance==1,NA,height$gv_calc)
height$gv_calc <- as.numeric(height$gv_calc)
height_save <- height[,c("record_id","gv_calc")]
demo <- merge(demo,height_save,by="record_id",all.x = T,all.y = T)

source("./GrowthAttenuationThe-BoneAge_R_2022-06-14_1929.r")
bone_age <- data
# exclude participant 100
bone_age <- bone_age %>% filter(!record_id==100)
data <- NULL
# two participants have data points well before start of therapy
# for now I am going to leave them in - for purposes of predicted adult height,
# Shanlee thought it would be OK to have multiple measures prior to start of therapy
# bone_age <- bone_age %>% filter(!(record_id==21 & redcap_repeat_instance==1))
# bone_age <- bone_age %>% filter(!(record_id==41 & redcap_repeat_instance==1))

source("./GrowthAttenuationThe-Labs_R_2022-06-13_1813.r")
labs <- data
# exclude participant 100
labs <- labs %>% filter(!record_id==100)
data <- NULL

# merge highest dose of estradiol, med_decrease and sex to bone age
d <- demo[,c("record_id","estradiol_highestdose","sex","stop_reason","med_decrease","tx_start_date")]
bone_age <- merge(bone_age,d,by="record_id",all.x = T, all.y = F)
bone_age$estradiol_highestdose <- as.factor(as.character(bone_age$estradiol_highestdose))
bone_age$sex <- as.factor(as.character(bone_age$sex))
# create new variable indicating if they were at the highest dose of estradiol
bone_age$at_estradiol_highest <- ifelse(bone_age$estradiol_highestdose==1 & bone_age$med_decrease!=1,1,0)
# delete any rows with missing ba_age and predicted adult height
bone_age <- bone_age %>% filter(!is.na(baest))

# calculate initial bone age, delta bone age, delta PAH and add to demo dataset (1 row per person)
# initial bone age
initialba <- bone_age[,c("record_id","tx_start_date","badate","baest","ba_sds","time_gat_ba")]
initialba <- initialba %>% filter(time_gat_ba<=0.25 & time_gat_ba >=-0.25)
initialba <- initialba %>% filter(!is.na(baest))
initialba <- initialba %>% arrange(record_id,badate) %>% group_by(record_id) %>% filter(row_number()==1)
initialba <- initialba %>% select(record_id,badate,baest,ba_sds)
colnames(initialba) <- c("record_id","date_initial_ba","initial_baest","initial_ba_sds")
# last bone age
lastba <- bone_age[,c("record_id","tx_start_date","badate","baest","ba_sds","time_gat_ba")]
lastba <- lastba %>% filter(!is.na(baest))
lastba <- lastba %>% arrange(record_id,desc(badate)) %>% group_by(record_id) %>% filter(row_number()==1)
lastba <- lastba %>% select(record_id,badate,baest,ba_sds)
colnames(lastba) <- c("record_id","date_last_ba","last_baest","last_ba_sds")
# calculate delta bone age but need to check if only 1 measure
deltaba <- merge(initialba,lastba,by="record_id",all.x = T,all.y=T)
deltaba$delta_baest <- deltaba$last_baest - deltaba$initial_baest
deltaba$delta_ba_sds <- deltaba$last_ba_sds - deltaba$initial_ba_sds
deltaba$delta_baest <- ifelse(deltaba$date_initial_ba==deltaba$date_last_ba,NA,deltaba$delta_baest)
deltaba$delta_ba_sds <- ifelse(deltaba$date_initial_ba==deltaba$date_last_ba,NA,deltaba$delta_ba_sds)
demo <- merge(demo,deltaba, by="record_id",all.x = T, all.y = T)
# first PAH
initialpah <- bone_age[,c("record_id","badate","ba_pah_calc")]
initialpah <- initialpah %>% filter(!is.na(ba_pah_calc))
initialpah <- initialpah %>% arrange(record_id,badate) %>% group_by(record_id) %>% filter(row_number()==1)
colnames(initialpah) <- c("record_id","date_initial_pah","initial_pah_calc")
# last PAH
lastpah <- bone_age[,c("record_id","badate","ba_pah_calc")]
lastpah <- lastpah %>% filter(!is.na(ba_pah_calc))
lastpah <- lastpah %>% arrange(record_id,desc(badate)) %>% group_by(record_id) %>% filter(row_number()==1)
colnames(lastpah) <- c("record_id","date_last_pah","last_pah_calc")
# calculate delta PAH but need to check if only 1 measure
deltapah <- merge(initialpah,lastpah,by="record_id",all.x = T,all.y=T)
deltapah$delta_pah_calc <- deltapah$last_pah_calc - deltapah$initial_pah_calc
deltapah$delta_pah_calc <- ifelse(deltapah$date_initial_pah==deltapah$date_last_pah,NA,deltapah$delta_pah_calc)
demo <- merge(demo,deltapah, by="record_id",all.x = T, all.y = T)

# check number of bone age measures prior to tx start
#bone_age$bapriortx <- ifelse(bone_age$badate<=bone_age$tx_start_date,1,0)
#check <- bone_age %>% group_by(record_id) %>% summarise(sum(bapriortx))
#bone_age <- merge(bone_age,check,by="record_id",all.x = T, all.y = T)
# check those with 2 data points prior to treatment
#two_bone_age_prior_tx <- bone_age[bone_age$`sum(bapriortx)`==2,
#                                 c("record_id","redcap_repeat_instance","tx_start_date","badate","sum(bapriortx)")]

#########################
# make final dataframes #
#########################

# df with participants having adverse effects excluded
bone_age_noae <- bone_age %>% filter(stop_reason %in% c(1,3,4,555))

# completers only
completers <- demo %>% filter(stop_reason==1)

# in progress
in_progress <- demo %>% filter(is.na(stop_reason))

# participants with AEs
ppt_ae <- demo %>% filter(stop_reason==2)
```

# Questions

1) Some data points are prior to start of therapy (as much as 1.4 years prior).  Some people have >1 value prior to start of therapy - should I use the one closest to start of therapy as "baseline", or use all the data?
2) For the model of predicted adult height vs. dose - I used the highest dose of estradiol.  Is that correct? 
3) For bone age at initiation - there are several versions of this variable. Which should I use? At the visit closest to and prior to start of therapy?
4) Calculate delta bone age Z-score and delta PAH using baseline and visit closest to 1 year?
5) One participant had an average growth velocity of 138 cm/year - is this correct?
6) We had discussed using a Poisson model with an offset for follow-up time to model the rate of AEs, but it seems like each participant is classified as having the AE or not, and there is no count of AEs per participant. If this is the case, we probably want to use a logistic regression model adjusted for follow-up time, OR a time-to-event model (although I don't know if we have the start date for the "other" AEs?).

# Methods

Linear mixed-effects models were used to model the trajectory of predicted adult height over time. Analyses were performed using R version 4.1.2.

# Baseline descriptive statistics

## All participants

```{r echo=FALSE, include=FALSE}
demovars <- c("sex.factor","race.factor","ethnicity.factor","language.factor","caregiver.factor",
              "state.factor","insurance.factor","genetics.factor","birthbraininjury.factor",
              "tbi.factor","respiratory_status.factor","feeds.factor","cp.factor","seizures.factor",
              "scoliosis.factor","gross_motor.factor","communication.factor","stop_reason.factor",
              "gat_start_age","treatment_duration_yr","gv_calc","wtgain_calc")
t1 <- CreateTableOne(vars=demovars,data=demo)
t1 <- print(t1,printToggle = F,varLabels=TRUE,missing=T,showAllLevels = T)
```

```{r echo=FALSE, warning=FALSE, comment=""}
kable(t1)
```

## Completers

```{r echo=FALSE, include=FALSE}
t2 <- CreateTableOne(vars=demovars,data=completers)
t2 <- print(t2,printToggle = F,varLabels=TRUE,missing=T,showAllLevels = T)
```

```{r echo=FALSE, warning=FALSE, comment=""}
kable(t2)
```

## Participants still on therapy

```{r echo=FALSE, include=FALSE}
t3 <- CreateTableOne(vars=demovars,data=in_progress)
t3 <- print(t3,printToggle = F,varLabels=TRUE,missing=T,showAllLevels = T)
```

```{r echo=FALSE, warning=FALSE, comment=""}
kable(t3)
```

## Participants who discontinued due to AEs

```{r echo=FALSE, include=FALSE}
t4 <- CreateTableOne(vars=demovars,data=ppt_ae)
t4 <- print(t4,printToggle = F,varLabels=TRUE,missing=T,showAllLevels = T)
```

```{r echo=FALSE, warning=FALSE, comment=""}
kable(t4)
```

# Models and plots - all data and participants included

## Spaghetti plot of predicted adult height

### By highest dose of estradiol

```{r echo=FALSE, warning=FALSE}
# Spaghetti plot of predicted adult height over time
p1 <- ggplot(data=bone_age, aes(x=time_gat_ba,y=ba_pah_calc,group=record_id, color=at_estradiol_highest)) + geom_line() +
      theme_ipsum()
p1
```

### By sex

```{r echo=FALSE, warning=FALSE}
# Spaghetti plot of predicted adult height over time
p2 <- ggplot(data=bone_age, aes(x=time_gat_ba,y=ba_pah_calc,group=record_id, color=sex)) + geom_line() +
      theme_ipsum()
p2
```

## Mixed model of predicted adult height 

### Model summary

```{r echo=FALSE, include=FALSE}
mod1 <- lme(ba_pah_calc ~ time_gat_ba,random=~1|record_id,data = bone_age,na.action = na.omit)
# estimation at mean value of time
mod1_pred <- matrix(predict(ref_grid(mod1)))
# estimation at baseline
mod1_pred.0yr <- ref_grid(mod1, at = list(time_gat_ba = 0))
# estimation at 1 year
mod1_pred.1yr <- ref_grid(mod1, at = list(time_gat_ba = 1))
# estimation at 2 years
mod1_pred.2yr <- ref_grid(mod1, at = list(time_gat_ba = 2))
```

```{r echo=FALSE, warning=FALSE, comment=""}
summary(mod1)
```

### Estimated value of ba_pah_calc at start of treatment

```{r echo=FALSE, warning=FALSE, comment=""}
summary(mod1_pred.0yr)
```

### Prediction interval of ba_pah_calc at start of treatment

```{r echo=FALSE, warning=FALSE, comment=""}
predict(mod1_pred.0yr, interval="prediction")
```

### Estimated value of ba_pah_calc at 1 year post-treatment

```{r echo=FALSE, warning=FALSE, comment=""}
summary(mod1_pred.1yr)
```

### Prediction interval of ba_pah_calc at 1 year post-treatment

```{r echo=FALSE, warning=FALSE, comment=""}
predict(mod1_pred.1yr, interval="prediction")
```

### Estimated value of ba_pah_calc at 2 years post-treatment

```{r echo=FALSE, warning=FALSE, comment=""}
summary(mod1_pred.2yr)
```

### Prediction interval of ba_pah_calc at 2 years post-treatment

```{r echo=FALSE, warning=FALSE, comment=""}
predict(mod1_pred.2yr, interval="prediction")
```

## Mixed model of predicted adult height vs. dose

```{r echo=FALSE, include=FALSE}
mod2 <- lme(ba_pah_calc ~ time_gat_ba*at_estradiol_highest,random=~1|record_id,data = bone_age,na.action = na.omit)
mod2_anova <- anova.lme(mod2, type="marginal")
mod2_means <- emmeans(mod2,"at_estradiol_highest")
mod2_pairs <-  pairs(mod2_means,adjust="tukey")
```

### Model summary

```{r echo=FALSE, warning=FALSE, comment=""}
mod2_anova
```

### Estimated mean ba_pah_calc in each dose category at mean value of time_gat_ba

```{r echo=FALSE, warning=FALSE, comment=""}
mod2_means 
mod2_pairs
```

# Models and plots - participants with adverse effects excluded (includes participants who have not completed therapy)

## Spaghetti plot of predicted adult height

### By highest dose of estradiol

```{r echo=FALSE, warning=FALSE}
# Spaghetti plot of predicted adult height over time
p1 <- ggplot(data=bone_age_noae, aes(x=time_gat_ba,y=ba_pah_calc,group=record_id, color=at_estradiol_highest)) + geom_line() +
      theme_ipsum()
p1
```

### By sex

```{r echo=FALSE, warning=FALSE}
# Spaghetti plot of predicted adult height over time
p2 <- ggplot(data=bone_age_noae, aes(x=time_gat_ba,y=ba_pah_calc,group=record_id, color=sex)) + geom_line() +
      theme_ipsum()
p2
```

## Mixed model of predicted adult height 

### Model summary

```{r echo=FALSE, include=FALSE}
mod1 <- lme(ba_pah_calc ~ time_gat_ba,random=~1|record_id,data = bone_age_noae,na.action = na.omit)
# estimation at mean value of time
mod1_pred <- matrix(predict(ref_grid(mod1)))
# estimation at baseline
mod1_pred.0yr <- ref_grid(mod1, at = list(time_gat_ba = 0))
# estimation at 1 year
mod1_pred.1yr <- ref_grid(mod1, at = list(time_gat_ba = 1))
# estimation at 2 years
mod1_pred.2yr <- ref_grid(mod1, at = list(time_gat_ba = 2))
```

```{r echo=FALSE, warning=FALSE, comment=""}
summary(mod1)
```

### Estimated value of ba_pah_calc at start of treatment

```{r echo=FALSE, warning=FALSE, comment=""}
summary(mod1_pred.0yr)
```

### Prediction interval of ba_pah_calc at start of treatment

```{r echo=FALSE, warning=FALSE, comment=""}
predict(mod1_pred.0yr, interval="prediction")
```

### Estimated value of ba_pah_calc at 1 year post-treatment

```{r echo=FALSE, warning=FALSE, comment=""}
summary(mod1_pred.1yr)
```

### Prediction interval of ba_pah_calc at 1 year post-treatment

```{r echo=FALSE, warning=FALSE, comment=""}
predict(mod1_pred.1yr, interval="prediction")
```

### Estimated value of ba_pah_calc at 2 years post-treatment

```{r echo=FALSE, warning=FALSE, comment=""}
summary(mod1_pred.2yr)
```

### Prediction interval of ba_pah_calc at 2 years post-treatment

```{r echo=FALSE, warning=FALSE, comment=""}
predict(mod1_pred.2yr, interval="prediction")
```

## Mixed model of predicted adult height vs. dose

```{r echo=FALSE, include=FALSE}
mod2 <- lme(ba_pah_calc ~ time_gat_ba*at_estradiol_highest,random=~1|record_id,data = bone_age_noae,na.action = na.omit)
mod2_anova <- anova.lme(mod2, type="marginal")
mod2_means <- emmeans(mod2,"at_estradiol_highest")
mod2_pairs <-  pairs(mod2_means,adjust="tukey")
```

### Model summary

```{r echo=FALSE, warning=FALSE, comment=""}
mod2_anova
```

### Estimated mean ba_pah_calc in each dose category at mean value of time_gat_ba

```{r echo=FALSE, warning=FALSE, comment=""}
mod2_means 
mod2_pairs
```

# Adverse effects

```{r echo=FALSE, include=FALSE}
aevars <- c("gi_ses.factor","skin_ses.factor","galactorrhea.factor","gynecomastia.factor",
            "fracturesgat.factor","seizuresgat.factor","hairgat.factor","droolgat.factor",
            "genitalgat.factor","flushgat.factor","other_ses")
ae_tab <- CreateTableOne(vars=aevars,data=demo)
ae_tab <- print(ae_tab,printToggle = F,varLabels=TRUE,missing=T,showAllLevels = T)
```

```{r echo=FALSE, warning=FALSE, comment=""}
kable(ae_tab)
```
