---
title: "General Software For Producing Trajectory Plots"
output:
  html_document: default
  html_notebook: default
params:
  digits: 2
  details: FALSE
# ------------------------------------------------------------------------------
# Parameters to the program that specify datasets, subsets, variable names and
# some plot parameters
# ------------------------------------------------------------------------------
  # Specify the dataset
  # SAS.dataset.path: 'C:/Users/bsc-default/Documents/RISE/SASdatasets/risecomplete_r.sas7bdat'
  # SAS.dataset.path: 'C:/Users/bsc-default/Documents/RISE/SASdatasets/adultpeds_full_i_r.sas7bdat'
  # SAS.dataset.path: 'C:/Users/bsc-default/Documents/RISE/SASdatasets/adultpeds_full_i_r.sas7bdat'
  SAS.dataset.path: 'R:/Analyses/Matched Pairs/subset.sas7bdat'
  # SAS.catalog.path: 'C:/Users/bsc-default/Documents/RISE/baselinefmtutf8.sas7bcat'
  # Specify subsets
  clinics: !r c('Adult')
  visits: !r c('BAS','M12', 'M15')
  timepoint.fu.labels: !r c('12', '15')
  # trtgrps: !r c('GLARGINE + METFORMIN','METFORMIN')
  # trtgrp.tags: !r c('GM','Met')
  # trtgrp.legend.labels: !r c('Glargine/Metformin','Metformin Alone')
  trtgrps: !r c('GLARGINE + METFORMIN', 'LIRAGLUTIDE + METFORMIN', 'METFORMIN', 'PLACEBO')
  trtgrp.tags: !r c('GM','LM','Met','Pla')
  trtgrp.legend.labels: !r c('Glargine/Metformin', 'Liraglutide+Metformin', 'Metformin Alone', 'Placebo')
  subset.bydiabvar: FALSE
  diabvar: "diabetes"
  diabvar.levels.keep: 0
  patidvar: "patid"
  clinicvar: "clinic_type"
  # yvar: "log_cpeptide_steady"
  # yvar.base: "log_cpeptide_steady_base"
  # yvar: "log_acprg"
  # yvar.base: "log_acprg_base"
  yvar: "log_acprmax"
  yvar.base: "log_acprmax_base"
  # yvar: "logcpgi_si"
  # yvar.base: "log_cpgi_base"
  # yvar: "logigi_si"
  # yvar.base: "log_igi_base"
  xvar: "log_mi"
  xvar.base: "log_mi_base"
  # xvar: "log_inv_ins0o_si"
  # xvar.base: "log_inv_ins0o_si_base"
  trtvar: "treatment"
  visitvar: "visit"
  # plot.ylab: "\nSteady State C-peptide (nmol/L)"
  # plot.ylab: "\nACPRg (nmol/L)"
  # plot.ylab: "\nACPRmax (nmol/L)"
  # plot.ylab: "IGI (pmol/mmol)"
  # plot.ylab: "CPI (nmol/mmol)"
  plot.ylab: " "
  plot.xlab: !r expression(paste(bold("M/I (x 10"^"-5"*" mmol/kg/min per pmol/L)")))
  # plot.xlab: !r expression(paste(bold("1/Fasting Insulin x 10"^"3"*" (1/(pmol/L))")))
  plot.xlo: 1
  plot.xhi: 5
  plot.ylo: 3.5
  plot.yhi: 9.0
  # plot.xlo: NULL
  # plot.xhi: NULL
  # plot.ylo: NULL
  # plot.yhi: NULL
  plot.legend.pos: "topright"
---
# Program Initialization
The code chunk below initializes program parameters

```{r loaddata, include=TRUE,error=TRUE,echo=TRUE,warning=FALSE}
# Load required packages
library(rio)
library(dplyr)
library(VIM)
library(StatAnalysis)

# Source functions from personal library
path = "./functions/"
for (nm in list.files(path, pattern = "[.][RrSsQq]$")) {
       # if(trace) cat(nm,":")
       source(file.path(path, nm))
       # if(trace) cat("\n")
}

# ------------------------------------------------------------------------------
# Read RISE data from
```

```{r initparms, include=TRUE,error=TRUE,echo=TRUE,warning=FALSE}
# ============================================================================
# Program Initialization 
# ============================================================================
# Analysis Flags:
# 1. pooledAnalysis - should treatment groups be pooled?
# 2. useImputedData - should imputed data for patients who failed treatment be
#                     included in the analysis?
# 3. IGTonly        - should the analysis be restricted to those participants who
#                     were IGT at baseline?
  clinics.keep   = params$clinics
  
# Specify the timepoints to be included in the analysis dataset using levels from
# the visit variable in the original dataset
  timepoints.keep <- params$visits
  
# Specify the treatment variable in the original dataset
  trtgrp.var <- as.symbol(params$trtvar)
  
# Specify the followup timepoints that will be analyzed for location and confidence
# ellipse from SUR model (this list should not include the baseline visit)
  timepoints.followup <- timepoints.keep[-1]

# Specify the timepoint labels that will be used in the plot
  timepoint.labels <- params$timepoint.fu.labels

# Specify the treatment group labels (not including baseline) for the legend.
# If null the values of the groups variable defined above will be used
  legend.names <- params$trtgrp.legend.labels 

# Specify the X and Y variables for the model
  xvar <- as.symbol(params$xvar)
  xvar.base <- as.symbol(params$xvar.base)

  yvar <- as.symbol(params$yvar)
  yvar.base <- as.symbol(params$yvar.base)

# Specify the x and y labels for the plot# X and Y axis labels
  xlabel <- expression(paste(bold("M/I (usual units)")))
  xlabel <- expression(paste(bold("1/Fasting Insulin x 10"^"3"*" (1 / (pmol/L))")))
  xlabel <- expression(paste(bold("M/I (x 10"^"-5"*" mmol/kg/min per pmol/L)")))
  xlabel <- params$plot.xlab

  ylabel <- "IGI (pmol/mmol)"
  ylabel <- "iAUC C/G (nmol/mmol)"
  ylabel <- "iAUC I/G (pmol/mmol)"
  ylabel <- "CPGI (nmol/mmol)"
  ylabel <- "AIRg (nmol/L)"
  ylabel <- "ACPRg (nmol/L)"
  ylabel <- "Steady State C-peptide (nmol/L)"
  ylabel <- "ACPRmax (nmol/L)"
  ylabel <- params$plot.ylab
  
  
# Specify and x and y limits for plot.  Set all to null if you want program to
# determine limits from the data
  xlo=params$plot.xlo
  xhi=params$plot.xhi
  ylo=params$plot.ylo
  yhi=params$plot.yhi

# Specify legend position by keyword. Options:
#   "bottom", "bottomright", "top", "topright"
  leg.pos <- params$plot.legend.pos
# =============================================================================
# END of Program Initialization
# =============================================================================
```

# Create Dataset
The code chunk below creates a dataset for analysis based on the program 
parameters initialized above.

```{r loaddata1, include=TRUE,error=TRUE,echo=TRUE,warning=FALSE}  
# Load the Analysis dataset
#  RISE <- import(params$SAS.dataset.path, format=params$SAS.catalog.path)
  RISE <- import(params$SAS.dataset.path)

# Apply SAS formats to create factor variables
  RISE <- as.data.frame(lapply(RISE, applySASfmt))
  
# Convert all variable names to lower case since R is case sensitive
  names(RISE) <- tolower(names(RISE))
  
# Exclude kid with extreme OGTT results - 5P069
#  RISE = RISE[RISE$PATID != '5P069',]
# Exclude the 2 kids who decompensated 5P033 and 7P031
#  RISE = RISE[!(RISE$PATID %in% c('5P033', '7P031')),]

# Subset to keep visits specified in params
  RISE = RISE[RISE[,params$visitvar] %in% params$visits,]
  RISE[,params$visitvar] <- droplevels(RISE[,params$visitvar])
  
# Subset to keep clinics specified in params
  RISE = RISE[RISE[,params$clinicvar] %in% params$clinics,]
  RISE[,params$clinicvar] <- droplevels(RISE[,params$clinicvar])
  
# Subset to keep treatment groups specified in params
  RISE = RISE[RISE[,params$trtvar] %in% params$trtgrps,]
  RISE[,params$trtvar] <- droplevels(RISE[,params$trtvar])
  
# Subset to keep only IGT participants
  if(params$subset.bydiabvar){
    RISE = RISE[RISE[,params$diabvar] %in% params$diabvar.levels.keep,]
    RISE[,params$diabvar] <- droplevels(RISE[,params$diabvar])
  }

# Keep only the needed variables for the analysis
  RISE <- RISE[,c(params$patidvar,
                  params$clinicvar,
                  params$visitvar, 
                  params$trtvar,
                  params$xvar.base,
                  params$xvar,
                  params$yvar.base,
                  params$yvar,
                  params$diabvar
                          )]

```

# Seemingly Unrelated Regression Models
The code chunk below runs the models based on the program initialization parameters.

```{r SUR,include=TRUE,error=TRUE,echo=FALSE,warning=FALSE,message=FALSE}
library(systemfit)
# SUR Regression
  SUR.models.byvisit <- by(RISE, 
                           RISE[,params$visitvar], 
                           function(x) {if(x$visit %in% 'BAS'){ x[,as.character(yvar.base)] = x[,as.character(yvar.base)] +1; x[,as.character(xvar.base)] = x[,as.character(xvar.base)] +1}
                                        systemfit(list(Y = eval(yvar) ~ eval(yvar.base) + 
                                                          eval(trtgrp.var), 
                                                      X = eval(xvar) ~ eval(xvar.base) + 
                                                          eval(trtgrp.var)),
                                                 method='SUR',data=x)})

  cat(paste("SUR Regression for the bivariate endpoint(", xvar,",", yvar,")\n"))
  lapply(SUR.models.byvisit, function(x){ return(summary(x)$coefficients[,c(1,2,4)])})
    
```


# Plot dataset
The code chunk below creates the plot dataset.

```{r Create Plot dataset}
# Version of egg plot where colors represent treatment groups, all plotting
# symbols are squares.  Baseline curve and point are black, non-baseline timepoints
# for glargine are green, and non-baseline timepoints for metformin are purple
library(ellipse)
library(shape)
library(RColorBrewer)

# -----------------------------------------------------------------------------  
# Calculate the baseline mean point in the plane and curve for the entire 
# study population
# -----------------------------------------------------------------------------  
# E[Y | E[X] ] at baseline
  # E[X] = mean of X
    mean.loginsSens <- RISE %>% filter(visit=="BAS") %>% summarize(mean(eval(xvar), na.rm=T)) %>% as.numeric
  # E[Y | E[X]] = predicted value of Y at x=E[X] for linear model of Y ~ X
    lm.model.base <- lm(eval(yvar) ~ eval(xvar), data=RISE[RISE$visit %in% 'BAS',])
    E.outcome    <- predict.lm(lm.model.base, data.frame(xvar=mean.loginsSens))
  
  # Variance-covariance matrix of (x,y) at baseline which will be used to 
  # to draw the confidence ellipse around the baseline point
  xy.matrix <- cbind(eval(xvar, envir = RISE[RISE[,params$visitvar]=='BAS',]), eval(yvar, envir = RISE[RISE[,params$visitvar]=='BAS',]))
  xy.cov    <- var(xy.matrix, use = 'complete.obs')
  n.points  <- sum(complete.cases(data.frame(xy.matrix)))

  # Confidence ellipse for baseline point
  confellipse.base <- ellipse::ellipse(xy.cov/n.points, centre=c(mean.loginsSens,E.outcome))
    
  # Predicted Baseline curve data
  baseline.curve <- data.frame(x=exp(lm.model.base$model$'eval(xvar)'),y=exp(predict.lm(lm.model.base))) 
  baseline.curve <- baseline.curve[order(baseline.curve$x),]

# -----------------------------------------------------------------------------
#  Specify the contrast matrices for each of the treatment groups. These
#  will be used below to determine the mean point and confidence ellipse
#  for each treatment group at the different visits.
#  
#  (R uses the alphabetical order of the levels
#  to determine to reference group, with the first levels always being the reference).
#  This can be changed by using the relevel() function to set a new reference, or using
#  my function ReorderFactorLevels() to change the order of the levels to whatever
#  the user chooses (e.g. PLACEBO, METFORMIN, GLARGINE + METFORMIN, LIRAGLUTIDE + METFORMIN)
#  
#  If doing a pooled analysis a treatment variable with 1-level ('Pooled') must
#  be defined and specified as trtvar in the parameters section of YAML header
# -----------------------------------------------------------------------------
   # 4 Treatment Groups
   Cmat.GM <- rbind(c(0,0,0,0,0,1,mean.loginsSens,0,0,0),
                    c(1,E.outcome,0,0,0,0,0,0,0,0))
  
   Cmat.LM <- rbind(c(0,0,0,0,0,1,mean.loginsSens,1,0,0),
                    c(1,E.outcome,1,0,0,0,0,0,0,0))
   
   Cmat.Met <- rbind(c(0,0,0,0,0,1,mean.loginsSens,0,1,0),
                    c(1,E.outcome,0,1,0,0,0,0,0,0))
   
   Cmat.Pla <- rbind(c(0,0,0,0,0,1,mean.loginsSens,0,0,1),
                    c(1,E.outcome,0,0,1,0,0,0,0,0))

   # 2 Treatment Groups
   # Cmat.GM <- rbind(c(0,0,0,1,mean.loginsSens,0),
   #                  c(1,E.outcome,0,0,0,0))
   # Cmat.LM <- rbind(c(0,0,0,0,0,1,mean.loginsSens,1,0,0),
   #                  c(1,E.outcome,1,0,0,0,0,0,0,0))
   # Cmat.Met <- rbind(c(0,0,0,1,mean.loginsSens,1),
   #                  c(1,E.outcome,1,0,0,0))
   # Cmat.Pla <- rbind(c(0,0,0,0,0,1,mean.loginsSens,0,0,1),
   #                   c(1,E.outcome,0,0,1,0,0,0,0,0))
  
# -----------------------------------------------------------------------------  
# Loop over visits and calculate the centroid and confidence ellipse for each treatment
# group at the visits (this excludes baseline since only the baseline curve and mean
# from the baseline curve are plotted for this timepoint)
# -----------------------------------------------------------------------------  
for(j in 1:length(timepoints.followup)){
  # Start loop over treatment groups
  for(i in 1:length(params$trtgrp.tags)){
    # Bivariate Mean and 95% Confidence Ellipse at followup timepoint j
    assign(paste0('confellipse.',timepoints.followup[j],'.',params$trtgrp.tags[i]), conf_Ellipse_SUR(SUR.models.byvisit[[match(timepoints.followup[j], names(SUR.models.byvisit))]], eval(as.symbol(paste0('Cmat.',params$trtgrp.tags[i])))))
                                                                                              
  # Closing bracket for loop over treatment groups in the case of unpooled data
  }
# Closing bracket for loop over visits
}
  
# -----------------------------------------------------------------------------  
# Organize the data into a list for plotting.  The code backtranforms to the
# original scale by applying exp() to the estimates based on the log transformed
# X and Y variables
# -----------------------------------------------------------------------------
  Estimates <- list()
  for(j in 1:length(timepoints.keep)){
    if(timepoints.keep[j] %in% 'BAS'){ 
        Baseline=Baseline=list(meanvec=exp(c(mean.loginsSens, E.outcome)),
                                    ellipse=exp(confellipse.base),
                                    curveXY=baseline.curve)
        Estimates[[j]] <- Baseline
    }
    else{
      groups.currentvisit <- list()
      for(i in 1:length(params$trtgrp.tags)){
        meanvec=exp(as.vector(eval(as.symbol(paste0('confellipse.',timepoints.keep[j], '.', params$trtgrp.tags[i])))$Mean))
        ellipse=exp(eval(as.symbol(paste0('confellipse.',timepoints.keep[j], '.', params$trtgrp.tags[i])))$Ellipse)
        
        groups.currentvisit[[i]] <- list(meanvec=meanvec,ellipse=ellipse)
      }
      names(groups.currentvisit) <- params$trtgrp.tags
      Estimates[[j]] <- groups.currentvisit
    }
  }
  
  names(Estimates) <- timepoints.keep 
  
```

# Trajectory Plot
The code chunk below creates the trajectory plot.

```{r Create Trajectory Plot}
# Open Graphics Device for output
  png(file='plot.png',width = 7, height=5, units = 'in', res=600)

  # Layout for paper
  # nf <- layout(matrix(c(1,1,2,2), 2, 2, byrow = TRUE), widths = c(lcm(10.5), lcm(10.5)),heights = c(lcm(12),lcm(12)))
  
  # Call Function to plot data
    plotData(Estimates, 
             timepoints = timepoints.followup,
             timepoint.labels = timepoint.labels,
             groups=params$trtgrp.tags,
             xaxis.label = xlabel,
             yaxis.label = ylabel,
             useBaseline = F,
             xmin = xlo,
             xmax = xhi,
             ymin = ylo,
             ymax = yhi,
             legend.names = legend.names,
             leg.pos = leg.pos,
             xpd = TRUE,
             plotbgcol="#f3eff5",
             plotannotcol="#330f42")
  
 dev.off()
```

# GLH Tests of Interest
General linear hypothesis tests of 
```{r GLH,include=FALSE, eval=FALSE}
# Linear Hypothesis of treatment differences equal to 0
  GLH.m15.GMvsMet <- linearHypothesis(SUR.models.byvisit[[3]],test = "Chisq",
# the constrasts:
  c('Y_eval(trtgrp.var)METFORMIN=0',
    'X_eval(trtgrp.var)METFORMIN=0'))

  GLH.m12.GMvsMet <- linearHypothesis(SUR.models.byvisit[[2]],test = "Chisq",
# the constrasts:
  c('Y_eval(trtgrp.var)METFORMIN=0',
    'X_eval(trtgrp.var)METFORMIN=0'))

# Linear Hypotheses of Treament group difference from baseline
  GLH.BasVm15.GM <- linearHypothesis(SUR.models.byvisit[[3]],test = "Chisq",
# the constrasts:
  c('Y_(Intercept)+Y_eval(yvar.base)=0',
    'X_(Intercept)+X_eval(xvar.base)=0'))

```


